### 2013-09-16 manuscripts/dpp-msbayes/dpp-msbayes.tex
Adding manuscripts.
Adding `manuscripts` directory and basic latex setup.

%&<latex>
\documentclass[letterpaper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{../utils/preamble.tex}
\input{../utils/macros.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\doublespacing
\raggedright
\setlength{\parindent}{0.5in}
\begin{linenumbers}

\begin{titlepage}
    \begin{flushleft}
        \sffamily

        \MakeUppercase{\large\bfseries Title}

        \vspace{12pt}
        \textbf{Running head:} \MakeUppercase{Running head}

        \vspace{12pt}
        Jamie R.\ Oaks$^{1,2,3}$ and Mark T.\ Holder$^{2}$

        \bigskip
        $^1$\emph{Biodiversity Institute\\
            $^2$Department of Ecology and Evolutionary Biology\\ 
            University of Kansas\\
            %1345 Jayhawk Blvd\\
            Lawrence, KS 66045\\
            USA}\\[.1in]
        $^3$\emph{Corresponding author} (\href{mailto:joaks1@ku.edu}{\tt
        joaks1@ku.edu})\\

    \end{flushleft}
\end{titlepage}

{\sffamily
    \noindent\textbf{ABSTRACT} \\
    \noindent Abstract here \ldots

    \vspace{12pt}
    \noindent\textbf{KEY WORDS: } 
}

\newpage
\noindent \ldots

\section*{Acknowledgments}

\bibliography{../bib/references}

%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\singlespacing

\renewcommand\listfigurename{Figure Captions}
\cftsetindents{fig}{0cm}{2.2cm}
\renewcommand\cftdotsep{\cftnodots}
\setlength\cftbeforefigskip{10pt}
\cftpagenumbersoff{fig}
\listoffigures

\end{linenumbers}

%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\singlespacing

\clearpage

%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

%:FIGURE-accuracy omega SS GLM
\begin{figure}[hbtp]
    \captionsetup{listformat=figList}
    \centering
    \includegraphics[width=1.0\textwidth]{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}
    \caption{
        Model-choice behavior \ldots
    }
    \label{fig_dpp_dpp_val_mc}
\end{figure}

\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}

\singlespacing

\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape supporting information}}

% PUT MAIN TEXT CITATION HERE
% \begin{thebibliography}{1}
% \providecommand{\natexlab}[1]{#1}
% \providecommand{\url}[1]{\texttt{#1}}
% \providecommand{\urlprefix}{URL }

% \bibitem

% \end{thebibliography}

%% SUPPL TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\setcounter{table}{0}

\end{document}

### 2013-09-16 manuscripts/reply/reply.tex
Adding manuscripts.
Adding `manuscripts` directory and basic latex setup.

--- 2013-09-16 manuscripts/dpp-msbayes/dpp-msbayes.tex
@@
         \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2,3}$ and Mark T.\ Holder$^{2}$
+        Jamie R. Oaks$^{1,2,5}$, Jeet Sukumaran$^{1,2}$, Jacob A.
+        Esselstyn$^{3}$, Charles W. Linkem$^{1,2}$, Cameron D.
+        Siler$^{4}$, Mark T. Holder$^{2}$ and Rafe M. Brown$^{1,2}$
 
@@
             USA}\\[.1in]
-        $^3$\emph{Corresponding author} (\href{mailto:joaks1@ku.edu}{\tt
+        $^3$\emph{Biology Department\\
+            McMaster University\\
+            % Life Sciences Building Rm 328\\
+            Hamilton, Ontario L8S4K1\\
+            Canada}\\[.1in]
+        $^4$\emph{Department of Biology\\
+            University of South Dakota\\
+            Vermillion, SD 57069\\
+            USA}\\[.1in]
+        $^5$\emph{Corresponding author} (\href{mailto:joaks1@ku.edu}{\tt
         joaks1@ku.edu})\\
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}
+    \includegraphics[width=1.0\textwidth]{../../saturation/saturation-plot.pdf}
     \caption{
-        Model-choice behavior \ldots
+        Saturation plot \ldots
     }
-    \label{fig_dpp_dpp_val_mc}
+    \label{fig_saturation_plot}
 \end{figure}

### 2013-09-16 manuscripts/utils/macros.tex
Adding manuscripts.
Adding `manuscripts` directory and basic latex setup.

% \newcommand{\change}[1]{{\color{blue} #1}\xspace}
\newcommand{\change}[1]{{\color{black} #1}\xspace}

\newcommand{\citationNeeded}{\textcolor{magenta}{\textbf{[CITATION NEEDED!]}}\xspace}
\newcommand{\tableNeeded}{\textcolor{magenta}{\textbf{[TABLE NEEDED!]}}\xspace}
\newcommand{\figureNeeded}{\textcolor{magenta}{\textbf{[FIGURE NEEDED!]}}\xspace}
\newcommand{\highLight}[1]{\textcolor{magenta}{\MakeUppercase{#1}}}

\newcommand{\editorialNote}[1]{\textcolor{red}{[\textit{#1}]}}
\newcommand{\ignore}[1]{}
\newcommand{\addTail}[1]{\textit{#1}.---}
\newcommand{\super}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\sub}[1]{\ensuremath{_{\textrm{#1}}}}
\newcommand{\dC}{\ensuremath{^\circ{\textrm{C}}}}

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\newcommand{\mthnote}[2]{{\color{red} #2}}

\newcommand{\ifTwoArgs}[3]{\ifthenelse{\equal{#1}{}\or\equal{#2}{}}{}{#3}\xspace}
\newcommand{\ifArg}[2]{\ifthenelse{\equal{#1}{}}{}{#2}\xspace}

\newcommand{\bayesfactor}[2]{\ensuremath{BF_{#1\protect\ifArg{#2}{,}#2}}}
\newcommand{\given}{\ensuremath{\,|\,}\xspace}
\newcommand{\msb}{\upshape\texttt{\MakeLowercase{ms\MakeUppercase{B}ayes}}\xspace}
\newcommand{\hky}{HKY85\xspace}
\newcommand{\uniformMin}[1]{\ensuremath{a_{#1}}\xspace}
\newcommand{\uniformMax}[1]{\ensuremath{b_{#1}}\xspace}
\newcommand{\locusRateHetShapeParameter}{\ensuremath{\alpha}\xspace}
\newcommand{\ancestralThetaVector}{\ensuremath{\boldsymbol{\theta_{A}}}\xspace}
\newcommand{\descendantThetaVector}[1]{\ensuremath{\boldsymbol{\theta_{D#1}}}\xspace}
\newcommand{\divtscaledvector}{\ensuremath{\mathbf{{\divtscaled{}{}}}}\xspace}
\newcommand{\divtvector}{\ensuremath{\boldsymbol{\divt{}}}\xspace}
\newcommand{\divtuniquevector}{\ensuremath{\mathbf{\divtunique{}}}\xspace}
\newcommand{\bottleTimeVector}{\ensuremath{\boldsymbol{\bottleTime{}}}\xspace}
\newcommand{\bottleTime}[1]{\ensuremath{\divt{B\ifArg{#1}{,}#1}}\xspace}
\newcommand{\bottleScalarVector}[1]{\ensuremath{\boldsymbol{\bottleScalar{#1}{}}}\xspace}
\newcommand{\bottleScalar}[2]{\ensuremath{\zeta_{D#1\protect\ifArg{#2}{,}#2}}\xspace}
\newcommand{\migrationRateVector}{\ensuremath{\mathbf{\migrationRate{}}}\xspace}
\newcommand{\geneTreeVector}{\ensuremath{\mathbf{\geneTree{}{}}}\xspace}
\newcommand{\alignmentVector}{\ensuremath{\mathbf{\alignment{}{}}}\xspace}
\newcommand{\alignment}[2]{\ensuremath{X_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}\xspace}
\newcommand{\geneTree}[2]{\ensuremath{G_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}\xspace}
\newcommand{\migrationRate}[1]{\ensuremath{m_{#1}}\xspace}
\newcommand{\recombinationRate}{\ensuremath{r}\xspace}
\newcommand{\ploidyScalar}[2]{\ensuremath{\rho_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}\xspace}
\newcommand{\ploidyScalarVector}{\ensuremath{\boldsymbol{\ploidyScalar{}{}}}\xspace}
\newcommand{\descendantRelativeThetaVector}[1]{\ensuremath{\boldsymbol{\eta_{D#1}}}\xspace}
\newcommand{\descendantRelativeTheta}[2]{\ensuremath{\eta_{D#1\protect\ifArg{#2}{,}#2}}\xspace}
\newcommand{\mutationRateScalarConstant}[2]{\ensuremath{\nu_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}\xspace}
\newcommand{\mutationRateScalarConstantVector}{\ensuremath{\boldsymbol{\mutationRateScalarConstant{}{}}}\xspace}
\newcommand{\locusMutationRateScalar}[1]{\ensuremath{\upsilon_{#1}}\xspace}
\newcommand{\locusMutationRateScalarVector}{\ensuremath{\boldsymbol{\upsilon}}\xspace}
\newcommand{\hkyModel}[2]{\ensuremath{\phi_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}\xspace}
\newcommand{\hkyModelVector}{\ensuremath{\boldsymbol{\hkyModel{}{}}}\xspace}
\newcommand{\mutationRate}{\ensuremath{\mu}\xspace}
\newcommand{\iid}{\textit{iid}\xspace}
\newcommand{\model}[1]{\ensuremath{\Theta}\xspace}
\newcommand{\npairs}[1]{\ensuremath{Y_{#1}}}
\newcommand{\nloci}[1]{\ensuremath{k_{#1}}\xspace}
\newcommand{\nlociTotal}{\ensuremath{K}\xspace}
\newcommand{\myTheta}[1]{\ensuremath{\theta_{#1}}}
\newcommand{\ancestralTheta}[1]{\ensuremath{\theta_{A\protect\ifArg{#1}{,}#1}}\xspace}
\newcommand{\descendantTheta}[2]{\ensuremath{\theta_{D#1\protect\ifArg{#2}{,}#2}}\xspace}
\newcommand{\meanDescendantTheta}[1]{\ensuremath{\descendantTheta{}{#1}}\xspace}
\newcommand{\nucdiv}[1]{\ensuremath{\pi_{#1}}}

\newcommand{\ssVector}[1]{\ensuremath{\mathbf{\alignmentSS{#1}{}}}\xspace}
\newcommand{\ssVectorObs}{\ensuremath{\ssVector{}^*}\xspace}
\newcommand{\ssSpace}{\ensuremath{\euclideanSpace{\ssVectorObs}}\xspace}
\newcommand{\ssVectorObsPLS}{\ensuremath{\ssVectorObs_{PLS}}\xspace}
\newcommand{\alignmentSS}[2]{\ensuremath{S_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}\xspace}
\newcommand{\alignmentSSObs}[2]{\ensuremath{\alignmentSS{#1}{#2}^*}\xspace}
\newcommand{\tol}{\ensuremath{\epsilon}\xspace}
\newcommand{\euclideanSpace}[1]{\ensuremath{B_{\tol}(#1)}\xspace}
\newcommand{\hpvector}[1]{\ensuremath{\Lambda_{#1}}}
\newcommand{\divtscaled}[2]{\ensuremath{t_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}}
\newcommand{\divt}[1]{\ensuremath{\tau_{#1}}}
\newcommand{\divtunique}[1]{\ensuremath{T_{#1}}}
\newcommand{\ssMatrix}{\ensuremath{\mathbb \alignmentSS{}{}}\xspace}
\newcommand{\ssMatrixRaw}[1]{\ensuremath{{\ssMatrix}_{stats#1}}\xspace}
\newcommand{\ssMatrixPLS}[1]{\ensuremath{{\ssMatrix}_{PLS#1}}\xspace}
\newcommand{\hpmatrix}[1]{\ensuremath{\mathcal{P}_{#1}}}
\newcommand{\meant}[2]{\ensuremath{E(\divt{#1})_{#2}}}
\newcommand{\meantestimate}{\ensuremath{\hat{E(\divt{})}}\xspace}
\newcommand{\vart}[2]{\ensuremath{Var(\divt{#1}{})_{#2}}}
\newcommand{\vmratio}[1]{\ensuremath{\Omega_{#1}}}
\newcommand{\numt}[1]{\ensuremath{\Psi_{#1}}}
\newcommand{\probnumt}[2]{\ensuremath{p(\numt{#1} = {#2})}}
\newcommand{\postprobnumt}[1]{\ensuremath{p(\numt{} = {#1}|\ssSpace)}}
\newcommand{\postprobnumtnot}[1]{\ensuremath{p(\numt{} \neq {#1}|\ssSpace)}}
\newcommand{\postprobomegasimult}{\ensuremath{p(\vmratio{} < 0.01 | \ssSpace)}\xspace}
\newcommand{\modelprior}[1]{\ensuremath{f(\model{})}}
\newcommand{\modelpost}[1]{\ensuremath{f(\model{}|\ssSpace)}}
\newcommand{\npriorsamples}{\ensuremath{n}\xspace}
\newcommand{\globalcoalunit}{\ensuremath{4\globalpopsize}\xspace}
\newcommand{\globalpopsize}{\ensuremath{N_C}\xspace}
\newcommand{\effectivePopSize}[1]{\ensuremath{N_e{#1}}\xspace}
\newcommand{\coalunit}{\ensuremath{4\effectivePopSize{}}\xspace}
\newcommand{\priorsample}[1]{\ensuremath{\hpmatrix{\modelprior{}}}}
\newcommand{\truncprior}[1]{\ensuremath{\hpmatrix{\tol}}\xspace}
\newcommand{\postsample}[1]{\ensuremath{\hpmatrix{\modelpost{}}}}
\newcommand{\abcllr}[1]{ABC\sub{LLR}}
\newcommand{\abcglm}[1]{ABC\sub{GLM}}
\newcommand{\integerPartition}[1]{\ensuremath{a({#1})}}
\newcommand{\uniqueModel}[2]{\ensuremath{M_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}}
\newcommand{\taxonLocusVector}[1]{\ensuremath{\{#1{1}{1},\ldots,#1{\npairs{}}{\nloci{\npairs{}}}\}}\xspace}
\newcommand{\taxonVector}[1]{\ensuremath{\{#1{1},\ldots,#1{\npairs{}}\}}\xspace}
\newcommand{\locusVector}[1]{\ensuremath{\{#1{1},\ldots,#1{\nlociTotal}\}}\xspace}

\newcommand{\simulationDescription}[2]{\change{Each plot represents #1
    simulation replicates using the same $#2$ samples from the prior}}
\newcommand{\simulationDistribution}{\ensuremath{\divt{} \sim U(0,
    \divt{max})}\xspace}
\newcommand{\estimateDescription}[2]{All estimates were obtained using #1 and #2}
\newcommand{\estimateDescriptionUncorrected}[1]{All estimates based on
    unadjusted posterior, \truncprior{}, obtained using #1}
\newcommand{\priorDescription}[4]{Prior settings were \priorSettings{#1}{#2}{#3}{#4}}
\newcommand{\priorSettings}[4]{$\divt{} \sim U(0, #1)$,
    $\meanDescendantTheta{} \sim U(#2, #3)$, and
    $\ancestralTheta{}{} \sim U(#2, #4)$}
\newcommand{\priorDescriptionBug}[4]{Prior settings were
    \priorSettingsBug{#1}{#2}{#3}{#4}}
\newcommand{\priorSettingsBug}[4]{$\divt{} \sim U(0, #1)$,
    $\meanDescendantTheta{} \sim U(#2, #3)$, and
    $\ancestralTheta{}{} \sim U(0.01, #4)$}
\newcommand{\simulationScheme}{simulations where \divt{} (in \globalcoalunit
    generations) for 22 population pairs is drawn from a series of uniform
    distributions, \simulationDistribution}
\newcommand{\captionPowerOmega}{Histograms of the estimated dispersion index
    of divergence times ($\hat{\vmratio{}}$) from \simulationScheme.
    The threshold for one divergence event \citep{Hickerson2006} is indicated
    by the dashed line, and the estimated probability of inferring one
    divergence event, $p(\hat{\vmratio{}}\le 0.01)$, is given for each
    \divt{max}}
\newcommand{\captionPowerPsiMode}{Histograms of the estimated number of
    divergence events ($\hat{\numt{}}$) from \simulationScheme.
    The estimated probability of inferring one divergence event,
    $p(\hat{\numt{}} = 1)$, is given for each \divt{max}}
\newcommand{\captionPowerPsi}{Histograms of the estimated posterior
    probability of one divergence event, \postprobnumt{1}, from
    \simulationScheme.
    The estimated probability of inferring one divergence event with a
    Bayes factor greater than 10 (dashed black line),
    $p(\bayesfactor{\numt{}=1}{\numt{} \ne 1} > 10)$, is given for each \divt{max}.
    The red line indicates $\postprobnumt{1} = 0.95$, and the estimated
    probability of inferring a posterior probability greater than 0.95 is given
    to the right of the line.}
\newcommand{\captionAccuracy}[1]{Accuracy and precision of #1 estimates from
    \simulationScheme.
    The proportion of estimates less than the true value ($p(\hat{#1}<#1)$) is
    given for each \divt{max}}
\newcommand{\samplingErrorTableNote}{An estimate of 1.0 for a posterior probability
    is an artifact of sampling error}

\newcommand{\refAccuracyALL}[1]{\labelcref{fig_acc_t_ss_llr_bug,fig_acc_t_ss_glm_bug,fig_acc_t_pls_llr_bug,fig_acc_t_pls_glm_bug,fig_acc_o_ss_llr_bug,fig_acc_o_ss_glm_bug,fig_acc_o_pls_llr_bug,fig_acc_o_pls_glm_bug}}
\newcommand{\refAccuracySS}[1]{\labelcref{fig_acc_t_ss_llr_bug,fig_acc_t_ss_glm_bug,fig_acc_o_ss_llr_bug,fig_acc_o_ss_glm_bug}}
\newcommand{\refAccuracySSfull}[1]{\labelcref{fig_acc_t_ssfull_llr_bug,fig_acc_t_ssfull_glm_bug,fig_acc_o_ssfull_llr_bug,fig_acc_o_ssfull_glm_bug}}
\newcommand{\refSSfull}[1]{\labelcref{fig_acc_t_ssfull_llr_bug,fig_acc_t_ssfull_glm_bug,fig_acc_o_ssfull_llr_bug,fig_acc_o_ssfull_glm_bug,fig_pow_o_ssfull_llr_bug,fig_pow_o_ssfull_glm_bug,fig_pow_psi_modes_ssfull_glm_bug}}
\newcommand{\refSS}[1]{\labelcref{fig_acc_t_ss_llr_bug,fig_acc_t_ss_glm_bug,fig_acc_o_ss_llr_bug,fig_acc_o_ss_glm_bug,fig_pow_o_ss_llr_bug,fig_pow_o_ss_glm_bug,fig_pow_psi_ss}}
\newcommand{\refAccuracyPLS}[1]{\labelcref{fig_acc_t_pls_llr_bug,fig_acc_t_pls_glm_bug,fig_acc_o_pls_llr_bug,fig_acc_o_pls_glm_bug}}
\newcommand{\refAccuracySScorrected}[1]{\labelcref{fig_acc_t_ss_llr_bug,fig_acc_t_ss_glm_bug,fig_acc_o_ss_llr_bug,fig_acc_o_ss_glm_bug}}
\newcommand{\refAccuracyPLScorrected}[1]{\labelcref{fig_acc_t_pls_llr_bug,fig_acc_t_pls_glm_bug,fig_acc_o_pls_llr_bug,fig_acc_o_pls_glm_bug}}
\newcommand{\refAccuracyUncorrected}[1]{\labelcref{fig_acc_t_ss_unc,fig_acc_t_pls_unc,fig_acc_o_ss_unc,fig_acc_o_pls_unc}}
\newcommand{\refAccuracyCorrected}[1]{\labelcref{fig_acc_t_ss_llr_bug,fig_acc_t_ss_glm_bug,fig_acc_t_pls_llr_bug,fig_acc_t_pls_glm_bug,fig_acc_o_ss_llr_bug,fig_acc_o_ss_glm_bug,fig_acc_o_pls_llr_bug,fig_acc_o_pls_glm_bug}}
\newcommand{\refAccuracyGLM}[1]{\labelcref{fig_acc_t_ss_glm_bug,fig_acc_t_pls_glm_bug,fig_acc_o_ss_glm_bug,fig_acc_o_pls_glm_bug}}
\newcommand{\refAccuracyLLR}[1]{\labelcref{fig_acc_t_ss_llr_bug,fig_acc_t_pls_llr_bug,fig_acc_o_ss_llr_bug,fig_acc_o_pls_llr_bug}}
\newcommand{\refAccuracyOmega}[1]{\labelcref{fig_acc_o_ss_llr_bug,fig_acc_o_ss_glm_bug,fig_acc_o_pls_llr_bug,fig_acc_o_pls_glm_bug}}
\newcommand{\refAccuracyOmegaUncorrected}[1]{\labelcref{fig_acc_o_ss_unc,fig_acc_o_pls_unc}}
\newcommand{\refAccuracyOmegaCorrected}[1]{\labelcref{fig_acc_o_ss_llr_bug,fig_acc_o_ss_glm_bug,fig_acc_o_pls_llr_bug,fig_acc_o_pls_glm_bug}}
\newcommand{\refAccuracyTime}[1]{\labelcref{fig_acc_t_ss_llr_bug,fig_acc_t_ss_glm_bug,fig_acc_t_pls_llr_bug,fig_acc_t_pls_glm_bug}}

\newcommand{\tn}{\tabularnewline}

### 2013-09-16 validation/bin/plot-old-results.py
Adding plotting script to `validation`.
Plotting script for old results.

@@ -1,0 +1,16 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes import plotting
+import project_util
+
+def main_cli():
+    info_path = os.path.join(project_util.RESULT_DIR, 'old',
+                'pymsbayes-results', 'pymsbayes-info.txt')
+    plotting.plot_validation_results(info_path)
+
+if __name__ == '__main__':
+    main_cli()
+

### 2013-09-16 validation/bin/plot-u-shaped-results.py
Adding plotting script for u-shaped `validation` results.

@@ -1,0 +1,16 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes import plotting
+import project_util
+
+def main_cli():
+    info_path = os.path.join(project_util.RESULT_DIR, 'u-shaped',
+                'pymsbayes-results', 'pymsbayes-info.txt')
+    plotting.plot_validation_results(info_path)
+
+if __name__ == '__main__':
+    main_cli()
+

### 2013-09-16 manuscripts/reply/reply.tex
Work on reply intro.

@@
 \newpage
-\noindent \ldots
+\noindent Both papers are largely in agreement.
+In fact, \citet{Hickerson2013} reiterate a lot of the discussion of
+\citet{Oaks2012} regarding the impact of broad uniform priors on Bayesian model
+choice.
+The main differences in their perspectives are centered around
+(1) the means by which broad uniform priors cause the poor behavior of \msb,
+and
+(2) how to potentially ameliorate this issue.
+
+\citet{Oaks2012} conclude the primary mechanism by which broad priors cause the
+poor behavior of the method is likely the low marginal likelihoods of
+parameter-rich models integrated over vast parameter space with low probability
+of producing the data, yet relatively high prior density \citep[this is often
+referred to as Lindley's paradox;][]{Lindley1957}.
+Note, this suggests the bias is extrinsic to \msb, and the numerical
+approximation machinery of the method could be sound.
+\citet{Hickerson2013} take a more pessimistic view of the bias, and suggest the
+bias is intrinsic to \msb, i.e., the method's rejection algorithm is
+inefficient and will be increasingly biased as the overall space of the model
+increases, either as a function of the number of taxon pairs or the width of
+the uniform priors on nuisance parameters.
 

### 2013-09-17 manuscripts/reply/reply.tex
Fleshing out reply intro.

@@
 
-        \MakeUppercase{\large\bfseries Title}
+        \MakeUppercase{\large\bfseries Why you should not fix a biased
+        model-choice method by adding an additional dimension of model choice:
+        A reply to Hickerson et al.}
 
         \vspace{12pt}
-        \textbf{Running head:} \MakeUppercase{Running head}
+        \textbf{Running head:} \MakeUppercase{Approximate Bayesian model
+        choice}
 
@@
 \newpage
-\noindent Both papers are largely in agreement.
+\noindent Recently, this journal has served as a venue for discussion of the
+potential pitfalls of approximate Bayesian methods of comparative
+phylogeographical model choice.
+\citet{Oaks2012} published their findings that the method \msb can often be
+biased towards inferring models of temporally clustered divergence times
+among taxon pairs.
+\citet{Hickerson2013} has published a response to this paper where they
+present a model-averaging approach that they conclude circumvents the poor
+behavior of the method revealved by \citet{Oaks2012}.
+Both of these papers are largely in agreement.
 In fact, \citet{Hickerson2013} reiterate a lot of the discussion of
@@
 approximation machinery of the method could be sound.
-\citet{Hickerson2013} take a more pessimistic view of the bias, and suggest the
-bias is intrinsic to \msb, i.e., the method's rejection algorithm is
+\citet{Hickerson2013} take a more pessimistic view of the bias, and suggest it
+is intrinsic to \msb, i.e., the method's rejection algorithm is
 inefficient and will be increasingly biased as the overall space of the model
@@
 the uniform priors on nuisance parameters.
+They support their position by giving a probabilistic argument that focuses on
+only one of the three prior models used by \citet{Oaks2012}.
+We show that this argument is based on dubious assumptions, and does not
+explain the bias of the method found in several analyses of \citet{Oaks2012}.
+We reiterate several of the findings of \citet{Oaks2012} as well as present
+results of additional analyses, which strongly suggest that Lindley's paradox
+is playing a larger role in the poor behavior of the method.
+
+\citet{Oaks2012} suggest that more cogent prior probability distributions on
+divergence models and nuisance parameters could mitigate the effect of
+Lindley's paradox.
+They also layout a set of simulation-based procedures for determining power,
+accuracy, and robustness of the method given a dataset, and recommend any
+application of \msb should be accompanied by such procedures, especially
+if any biological conclusions are going to drawn from the results.
+\citet{Hickerson2013} present an approximate Bayesian model-averaging approach
+for accommodating uncertainty in prior selection, and champion the method
+as a means of avoiding the pitfalls raised by \citet{Oaks2012}.
+Unfortunately, we show that fundamental errors in this approach render some of
+their results invalid and leave the remaining results difficult to interpret.
+Furthermore, we follow the advice of \citet{Oaks2012} and present
+simulation-based assessments of Hickerson et al.'s \citeyear{Hickerson2013}
+approach.
+Our results demonstrate that the method is biased and dangerous.
+The approach merely provides another means by which the model can ``escape''
+large parameter space, and the bias towards models with smaller space still
+remaines.
 

### 2013-09-17 manuscripts/reply/reply.tex
More work on reply.

@@
 \citet{Hickerson2013} present an approximate Bayesian model-averaging approach
-for accommodating uncertainty in prior selection, and champion the method
-as a means of avoiding the pitfalls raised by \citet{Oaks2012}.
+for accommodating uncertainty in selecting among empirically guided priors, and
+champion the method as a means of avoiding the pitfalls raised by
+\citet{Oaks2012}.
 Unfortunately, we show that fundamental errors in this approach render some of
@@
 large parameter space, and the bias towards models with smaller space still
-remaines.
+remains.
+This bias can be dangerous and often excludes the true parameter space due to
+the use of uniform priors.
+Furthermore we discuss the potential theoretical and practical problems of
+empirical Bayesian model choice.
+
+\section*{Their power analysis}
+\citet{Hickerson2013} present a power analysis in which they demonstrate that
+\msb has the power to infer multiple divergence events when divergence times
+are random over hundreds of thousands of generations (rather than millions
+as demonstrated by \citet{Oaks2012}.
+It is not surprising to find that within certain parameter space the method
+has increased power to detect temporal variation in divergences over narrower
+time windows than millions of generations.
+However, what is important to consider is whether those conditions are relevant
+to real world applications of the method.
+\citet{Oaks2012} explore the behavior of the method under three divergence time
+priors as narrow as 0--5 coalescent units, which is quite narrow considering
+that this expresses the prior belief that all 22 taxon pairs diverged within
+this window.
+\citet{Hickerson2013} limit their power analysis to a single prior of 0--1
+coalescent unit.
+It seems that being 100\% certain a priori that all taxa under comparison
+diverged within the last coalescent unit is not generally applicable to most
+empirical systems.
+Furthermore, even when such extensive prior information is available to only be
+able to detect multiple divergences on the scale of hundreds of thousands of
+generations does not seem very ``powerful.''
+
+Also, \citet{Hickerson2013} present information only on the extreme case of a
+single divergence event.
+As discussed in \citet{Oaks2012}, it seems dubious to consider the estimation
+of two divergence events as a success when divergence is random over hundreds
+of thousands of generations.
+Unlike the results of \citet{Oaks2012} it is impossible to assess the
+performance of \msb based on the results presented by \cite{Hickerson2013}
+other than to know whether or not it inferred the worst possible scenario of a
+single event.
+In an empirical system such as the Philippines, where island fragmentation
+has occurred at least X times over the past several millions of years, an
+estimate where 22 taxa share even a handful of divergence events would be
+of biogeographic interest.
+\citet{Hickerson2013} seem to suggest in their response that \msb should only
+be used as a binary estimator; i.e., was there one event or more?
+This is not consistent with how \msb has been presented, used, and interpreted
+in the past.
+
+\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
+We preface this section with a clarification of an issue that has
+been confused throughout the \msb literature.
+The mean (\meant{}), variance (\vart{}), and dispersion index
+($\vart{}/\meant{} = \vmratio{}$) of divergence times are \emph{not} parameters
+of the model implemented in \msb.
+Rather they are statistics that summarize parameters of the model.
+This is in contrast to \numt{}, which is a parameter of the model.
+
+\citet{Hickerson2013} strongly argue that \vmratio{} is a better estimator than
+\numt{}, but the logic behind the argument is dubious.
+They present a plot of \numt{} against \vmratio{} (Fig.~S1) which is simply a
+plot of sample size versus variance.
+After showing that \vmratio{} has essentially no information about the
+number of divergences, they conclude it is more informative and biogeographically
+relevant.
+We struggle to follow this logic.
+Certainly the maximum information is contained within the divergence time
+vector that \vmratio{} is summarizing.
+The temporal information contained in this vector coupled with its cardinality
+(\numt{}) is certainly more informative than its summary.
+The dispersion index is not a sufficient statistic for this vector.
+
+They also argue that ``\msb can estimate \vmratio{} much better than \numt{}.''
+However, the primary objective of \msb is to estimate the posterior probability
+of divergence models.
+\citet{Oaks2012} demonstrate that even when all assumptions of the model are
+met, \vmratio is a terrible model-choice estimator (see plots B, D \& F of
+Figure 4), whereas \numt{} performs better.
+Furthermore, \vmratio{} is limited to estimating the probability of only a
+single divergence model (the one divergence model), and thus its utility for
+model choice is extremely limited.
 
@@
 
-\bibliography{../bib/references}
+% \bibliography{../bib/references}
+\bibliography{references}
 
@@
 
-%:FIGURE-accuracy omega SS GLM
-\begin{figure}[hbtp]
-    \captionsetup{listformat=figList}
-    \centering
-    \includegraphics[width=1.0\textwidth]{../../saturation/saturation-plot.pdf}
-    \caption{
+%:FIGURE-saturation plot
+\mFigure{../../saturation/saturation-plot.pdf}{
     Saturation plot \ldots
-    }
-    \label{fig_saturation_plot}
-\end{figure}
+}{figSaturationPlot}
 
-\clearpage
+\mFigure{../../response-redux/results/hickerson/pymsbayes-results/mean_by_dispersion.pdf}{
+    Results of reanalysis.
+}{figJointPosterior}
+
+\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
+    Sampling error \ldots
+}{figSamplingError}
+
+%% Exclusion simulation results
+\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/num_tau_excluded.pdf}{
+    Histogram of the number of \divt{} parameters excluded.
+}{figExclusionSimTau}
+
+\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/prob_of_exclusion.pdf}{
+    Histogram of the posterior probability of excluding the truth.
+}{figExclusionSimProb}
+
+%% Power results
+\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
+    Histograms of the posterior probability that $\vmratio{} < 0.01$.
+}{figPowerAccOmegaMedian}
+
+\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
+    Histograms of the posterior probability that $\vmratio{} < 0.01$.
+}{figPowerAccOmegaModeGLM}
+
+\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_psi_mode.pdf}{
+    Histograms of \numt{} estimates.
+}{figPowerPsiMode}
+
+\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_omega_prob.pdf}{
+    Histograms of the posterior probability that $\vmratio{} < 0.01$.
+}{figPowerOmegaProb}
+
+\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_prob_exclusion.pdf}{
+    Histograms of the posterior probability of excluding the truth.
+}{figPowerProbExclusion}
+
+\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_num_excluded.pdf}{
+    Histograms of the number of true \divt{} parameters excluded.
+}{figPowerNumExcluded}
+
+%% Validation results
+\mFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
+    True versus estimated probability of a single divergence event.
+}{figValidationMCBehvior}
 

### 2013-09-17 model-choice/tau-prior/mc-tau-prior-plot.py
Adding plot of model-choice tau prior.

@@ -1,0 +1,39 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes.plotting import (HistData, ScatterPlot)
+from pymsbayes.utils.messaging import get_logger
+
+_LOG = get_logger(__name__)
+
+def main_cli():
+    taus = []
+    unique_taus = []
+    f = open('tau.txt', 'rU')
+    f.next()
+    for l in f:
+        l = l.strip().split()
+        taus.extend([float(t) for t in l])
+        unique_taus.extend([float(t) for t in set(l)])
+    hd_t = HistData(x = taus,
+            bins=10,
+            zorder = 0)
+    plot_t = ScatterPlot(
+            hist_data_list = [hd_t],
+            x_label = r'Divergence time ($\tau$)',
+            y_label = r'Density')
+    plot_t.savefig('tau_prior.pdf')
+    hd_t = HistData(x = unique_taus,
+            bins=10,
+            zorder = 0)
+    plot_t = ScatterPlot(
+            hist_data_list = [hd_t],
+            x_label = r'Divergence time ($\tau$)',
+            y_label = r'Density')
+    plot_t.savefig('unique_tau_prior.pdf')
+
+if __name__ == '__main__':
+    main_cli()
+

### 2013-09-17 manuscripts/reply/reply.tex
More work on reply.

@@
 
+\section*{EB}
+\citet{Hickerson2013} repeatedly refer the prior distributions used by
+\citet{Oaks2012} as ``poorly selected''.
+
 \section*{Their power analysis}
@@
 
+\section*{Our power analysis}
+
 \section*{Differing utilities of \numt{} and \vmratio{} in \msb}
@@
 model choice is extremely limited.
+The model-choice utility of \vmratio{} is limited to the probability that
+this continuous statistic, which can range from zero to infinity, is at its lower
+limit of zero. In theory, this point density will always be zero, thus an
+arbitrary threshold (0.01 is used throughout the \msb literature) must 
+be chosen to make the probability estimable.
+However, it is still not surprising to see that it is numerically difficult to
+obtain reliable estimate of the probability that the continuous \vmratio{}
+statistic is ``near'' its limit of zero.
+It is much easier, less subjective, and more interpretable to estimate
+the probability of the discrete parameter of the model, \numt{}, is
+at its lower limit of one.
+Thus, it is not surprising to see this estimator of model probability
+perform better.
+
+\section*{General thoughts}
+It is not too surprising that the model-averaging approach proposed by
+\citet{Hickerson2013} behaves so poorly.
+While this approach is trivial to implement, it is not a trivial change
+to the the basic \msb model.
+To better understand why this is, it might help to step back and get a
+sense of the scale of the \msb model.
+To do this, we will use the dataset of 22 vertebrate taxon pairs of
+\citet{Oaks2012} as an example.
+
+Following the model description and notation of \citet{Oaks2012}, we
+will tally up all of the free parameters in the \msb model.
+
+Under the simplest model in \msb (i.e., assuming no migration and
+no intra-locus recombination), the number of parameters for each
+taxon pair include:
+The population sizes of the ancestral and descendant populations
+(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
+The magnitude of population contraction in each of the descendant
+populations (\bottleScalar{1}{} and \bottleScalar{2}{}) and the
+timing of these contractions (\bottleTime{}), and the $N-1$ node heights
+(coalescent times) of the gene tree that gives rise to the $N$ gene
+copies sampled from both populations of the pair.
+
+Because we only have a single locus for each taxon, there are no locus-specific
+\myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
+\locusRateHetShapeParameter shape parameter for the gamma prior distribution on
+\locusMutationRateScalar{}.
+Lastly, there are between one and 22 divergence time parameters \divt{} in
+the vector \divtvector.
+Thus for our Philippines data and under the simplest model in \msb, there are
+581--602 free parameters.
+Under this rich model, the method is sampling over 1002 divergence models
+\citep[i.e., the number of integer partitions of $Y=22$][]{Oaks2012}.
+
+This is a very tall order, and the method only uses four summary statistics
+calculated from the sequence alignment of each taxon pair:
+$\pi$ \citep{Tajima1983}, $\theta_W$
+\citep{Watterson1975}, $\pi_{net}$ \citep{Takahata1985}, and
+$SD(\pi-\theta_W)$ \citep{Tajima1989}.
+That gives us a total of 88 summarys statistics (four for each of the 22 taxon
+pairs), which contain minimal information about many of the $\approx 600$
+parameters in the model.
+More summary statistics can be used in \msb, but most are highly correlated
+with these four statistics (which are even highly correlated among themselves),
+and thus contribute little additional information about the data.
+
+This very large number of parameters and divergence models relative to 
+summary statistics is undoubtedly the reason the method is so sensitive
+to the prior distributions.
+It also likely contributes to the methods lack of robustness.
+We use robustness here as a measure of a methods insensitivity to model
+violations.
+Robustness is an extremely important characteristic of a method to gauge its
+applicability to empirical data, because we know the model and priors will be
+wrong to some degree.
+
+The approach of \citet{Hickerson2013} adds an additional dimension of model
+choice to the model. They expand the model to sample over eight prior models.
+This extends the original model to having 582-603 free parameters and, more
+importantly, sampling over 8016 unique model states across both divergence
+models and prior models.
+This certainly is a non-trivial extension of the model, and given the methods
+very large number of parameters and models relative to the information
+used from the data, likely plays a major role in the poor behavior of this
+approach.
+
+In theory, the approach of \citet{Hickerson2013} is very appealing.  It sums
+over multiple candidate prior models to produce a posterior estimate
+marginalized over the uncertainty in prior choice.
+In general, Bayesian model-averaging is a powerful approach that leverages a
+great strengths of Bayesian statistical procedures, namely the ability to
+obtain marginalized estimates that incorporate uncertainty in nuisance
+parameters.
+However, given the basic \msb model is already struggling to estimate
+a huge number of parameters and model probabilities with scant information
+from the data, it is not surprising that adding an additional layer of
+model choice to the method causes problems.
+
+The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
+and biases of \msb are actually similar to those of \citet{Hickerson2013},
+but avoid the need for imposing additional model choice.
+\citet{Oaks2012} suggest that uniform priors are inappropriate for parameters,
+and recommend the use of more cogent probability distributions from the
+exponential family.
+If we look at the prior distribution on the divergence time parameter \divt{}
+imposed by the model-averaging prior approach of \citet{Hickerson2013} we see
+it is a mixture of overlapping uniforms with lower limits of zero
+(Figure~\ref{figMCTauPrior}).
+This looks very much like an exponential distribution.
+Thus would be much more natural and easier to simply place a gamma prior (the
+exponential being a special case of the gamma) on divergence times.
+This would allow the investigator to much better capture their prior
+uncertainty in model parameters while avoiding awkward broad uniform
+distributions, and without the need of costly model-averaging.
+This seems like a much more natural solution to the problems of Lindley's
+paradox and sampling error, which are both certainly exacerbated by the fact
+that uniform distributions are inappropriately the only choice of prior for
+many of the parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
+\descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
+\bottleScalar{1}{}, \bottleScalar{2}{}, \locusMutationRateScalar,
+\migrationRate{}, \recombinationRate;][]{Oaks2012}.
+There is no cost whatsoever to this strategy.
+Given that ABC methods merely have to draw random values from prior
+distributions, hence there are no difficult proposal probability ratios to
+calculate, etc., there is no excuse for not using the most appropriate
+distributions on parameters with conjugacy in mind.
+Also, this approach reduces the ``need'' for empirically guided priors.
+The investigator can place the majority of the prior density in regions
+of parameters space the believe, a priori, are most plausible, but still
+capture uncertainty in the tails of distributions with low density.
+As discussed in \citet{Oaks2012} the use of inflexible and unforgiving uniform
+priors necessitated the need to use broad priors to avoid excluding the truth a
+priori.
+Thus, to avoid the   behavior of the method under these priors, cheating
+and using the data was a very tempting alternative.
 
@@
 
+%% Model-choice tau prior
+\mFigure{../../model-choice/tau-prior/tau_prior.pdf}{
+    The prior distribution on divergence times imposed by the model-averaging prior.
+}{figMCTauPrior}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-09-17 manuscripts/reply/reply.tex
More work on reply.
Beginning work on EB section.

@@
 \citet{Oaks2012} as ``poorly selected''.
+However, this is bit misleading, as \citet{Oaks2012} discuss in detail how the
+selected their priors (see their section ``Specifying and simulating the joint
+prior'').
+They detail how the being limited to uniform prior distributions necessitates
+awkward broad priors to avoid excluding the truth, a priori.
+What \citet{Hickerson2013} really mean to say is that \citet{Oaks2012} did not
+use an empirical Bayesian approach, which is what they do and advocate.
+
+Given all of this, it is quite clear why one should use caution before
+overly criticizing ones choice of priors after having seen the posterior.
+This really is a strawman argument.
+They were not there before the data were seen.
 

### 2013-09-18 manuscripts/reply/reply.tex
More work on reply.

@@
 \citet{Oaks2012} as ``poorly selected''.
-However, this is bit misleading, as \citet{Oaks2012} discuss in detail how the
+However, this is bit misleading, as \citet{Oaks2012} discuss in detail how they
 selected their priors (see their section ``Specifying and simulating the joint
@@
 awkward broad priors to avoid excluding the truth, a priori.
-What \citet{Hickerson2013} really mean to say is that \citet{Oaks2012} did not
-use an empirical Bayesian approach, which is what they do and advocate.
+What \citet{Hickerson2013} really mean by ``poorly selected'' is that
+\citet{Oaks2012} did not use their data to inform their initial prior
+distributions (i.e., they did not use an empirical Bayesian approach).
+\citet{Oaks2012} did also use empirically informed priors when assessing
+both the prior sensitivity of the method and determining the power of \msb
+under the ``best'' possible real-world conditions.
+\citet{Oaks2012} discuss the potential dangers of taking an empirical Bayesian
+model choice (see last paragraph of ``Assessing prior sensitivity of \msb''),
+and we expand on this here.
+
+\subsection*{Theoretical implications of empirical priors for Bayesian model
+choice}
+\begin{linenomath}
+Bayesian inference is a rational method of inductive learning in which Bayes'
+rule is used to update our beliefs about a model as new information becomes
+available.
+If we let \allParameterValues represent the set of all possible parameter
+values, we can define a prior distribution for all $\theta \subset
+\allParameterValues$ such that $p(\theta)$ describes our belief that any given
+\myTheta{} is the true value of the parameter.
+If we let \allDatasets represent all possible datasets then we can 
+define a sampling model for all $\theta \subset
+\allParameterValues$ and $\alignment{}{} \subset \allDatasets$ such that
+$p(\alignment{}{} | \theta)$ measures our belief that any dataset \alignment{}{}
+will be generated by any model state \myTheta{}.
+After collecting a new dataset \alignment{i}{}, we can use Bayes' rule to
+calculate the posterior distribution
+\begin{equation}
+    p(\myTheta{} \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given 
+        \myTheta{})p(\myTheta{})}{p(\alignment{i}{}},
+    \label{eq:bayesrule}
+\end{equation}
+where $p(\alignment{i}{}) = \int{\myTheta{}} p(\alignment{i}{} \given
+\myTheta{})p(\myTheta{}) d\myTheta{}$.
+The posterior distribution is a measure of our beliefs after seeing the new
+information.
+\end{linenomath}
+
+This is an elegant method of updating our beliefs as data are accumulated.
+However, this all hinges on the fact that the prior ($p(\myTheta{})$) is
+defined for all possible parameters independently of the new data being
+analyzed.
+Any other datasets or external information can safely be used to inform our
+beliefs about $p(\myTheta{})$.
+However, if the data are used to both inform the prior and calculate the
+posterior, the prior becomes conditional on the data, and Bayes' rule breaks
+down.
+
+Now, this is not to say that empirical Bayesian approaches not useful.
+Quite the contrary.
+Empirical Bayes is a well studied branch of Bayesian statistics that
+has given rise to a many powerful inference methods.
+It is true that empirical Bayes methods have an uncertain theoretical basis and
+do not yield a valid posterior distribution from Bayes' rule \citp[e.g.,
+empirical Bayesian estimates of the posterior are often too narrow, off-center,
+and misshapen;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
+Nevertheless, empirical Bayes methods are often a powerful means of parameter
+estimation that display favorable frequentist properties.
+Furthermore, many post-hoc correction methods have been developed for
+estimating confidence-intervals from empirical Bayes estimates of posterior
+distributions that often display good frequentist coverage behavior
+\citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.
+
+Whereas empirical Bayes approaches can provide powerful methods for parameter
+estimation, a theoretical justification for empirical Bayes approaches to model
+choice is much more dubious.
+In Bayesian model choice, the goal is usually not to estimate parameters, but
+to estimate the relative probabilities of candidate models.
+Unlike parameter estimates, posterior probabilities represent a summary of the
+entire posterior distribution.
+Thus, given that empirical Bayesian posterior distributions are not accurate,
+there is no guarantee regarding the accuracy of probabilities summarized from
+them.
 

### 2013-09-18 manuscripts/images/coin_flip.r
Adding coin flip example plot.

@@ -1,0 +1,17 @@
+rm(list=ls(all=TRUE))
+
+q = seq(0, 1, 0.005)
+prior = dbeta(q, 0.5, 0.5)
+post = dbeta(q, 4.5, 1.5)
+epost = dbeta(q, 8.5, 2.5)
+
+pdf('coin_flip.pdf', width=7.087, height=4.5)
+par(yaxs="i", xaxs="i", cex.lab=1.2, cex.axis=1.0, mgp=c(2.2, 0.7, 0),
+    oma=c(0, 0, 0.2, 0.5), mar=c(3.2, 3.2, 0.5, 0.1))
+plot(x=c(), y = c(), xlab=bquote(theta), ylab='Density', xlim=c(0,1),
+     ylim=c(0, 3.5))
+lines(q, prior)
+lines(q, post, col='blue')
+lines(q, epost, col='red')
+dev.off()
+

### 2013-09-18 manuscripts/reply/reply.tex
Adding to reply.
Adding coin-flip example.

@@
     p(\myTheta{} \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given 
-        \myTheta{})p(\myTheta{})}{p(\alignment{i}{}},
+    \myTheta{})p(\myTheta{})}{p(\alignment{i}{})},
     \label{eq:bayesrule}
 \end{equation}
-where $p(\alignment{i}{}) = \int{\myTheta{}} p(\alignment{i}{} \given
-\myTheta{})p(\myTheta{}) d\myTheta{}$.
+where
+\begin{equation}
+    p(\alignment{i}{}) = \int_{\myTheta{}} p(\alignment{i}{} \given
+    \myTheta{})p(\myTheta{}) d\myTheta{}.
+\end{equation}
 The posterior distribution is a measure of our beliefs after seeing the new
@@
 It is true that empirical Bayes methods have an uncertain theoretical basis and
-do not yield a valid posterior distribution from Bayes' rule \citp[e.g.,
+do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
 empirical Bayesian estimates of the posterior are often too narrow, off-center,
@@
 them.
+
+This can be demonstrated with a simple, albeit contrived, example.
+Let us say that principal investigator Mary and her new postdoc Will are
+interested in the hypothesis that the mass of George Washington's periwig
+renders the United States' quarter dollar unfair.
+That is to say that our null hypothesis is that the probability of a US quarter
+landing heads when tossed is less than 0.5 ($\theta < 0.5$).
+They can certainly evaluate this hypothesis, they can have undergraduate
+worker, Joe, flip the coin for them while the tabulate the results.
+But being Bayesians, before they call Joe into the lab, they agree on a
+prior probability to place on the set of all possible probabilities that
+the quarter will land heads when it is flipped.
+Given their prior knowledge that Joe moonlights as a magician, and is notorious
+for performing coin and card tricks in the lab, they suspect there is a good
+chance that Joe will use his slight-of-hand skills to replace their quarter
+with one of his two-headed or two-tailed coins he has up his sleeve at all
+times.
+So, knowing that the beta distribution is the conjugate prior for a binomial
+likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
+(Figure~\ref{figCoinFlip}).
+
+Mary calls Joe into the lab, and hands him a regular-issue US quarter minted in
+1992.
+After five tosses, four of which was heads, Joe decides that academics are
+crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
+being computational biologists, are satisfied with their
+empirical dataset of $y = 4$ heads out of $N = 5$ trials.
+They know from the conjugacy of the beta prior, that the posterior distribution
+has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
+in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
+of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
+This allows them to plug these values into the beta cumulative distribution
+function to determine that the posterior probability of their hypothesis is
+$p(\theta < 0.5 | y=1, N=5) = 0.088$.
+Given their prior belief and dataset, this indeed is the correct posterior
+probability of the hypothesis, and Mary and Will should now update their
+posterior belief accordingly.
+
+However, upon reflecting on the results of their experiment
+(Figure~\ref{figCoinFlip}), Mary and Will regret not trusting Joe to refrain
+from turning their experiment into a magic trick.
+Their prior looks very ``poorly selected,'' and if they had only
+had more faith in Joe's integrity, they would have selected a much better
+prior. Clearly form their results, they should have used a prior centered
+around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
+much more appropriate.
+Rather than resort to flipping the coin five more times, Mary and Will decide
+to redo their analysis using the much better ``prior'' of $p(\theta) \sim
+beta(4.5, 1.5)$.
+This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
+probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
+Now convinced that the mass of President Washington's head renders the US
+Quarter unfair, Will begins composing an e-mail of complaint for the U.S.\
+Mint.
+
+If Joe's flips were a representative sample, Mary and Will's empirical Bayes
+estimate might very well be a better estimator for the parameter \myTheta{}.
+However, their empirical Bayes estimate of the probability of their hypothesis
+is incorrect and biased.
+This simple example shows how parameter estimation is fundamentally different
+from estimating the probability of a model.
+While empirical informed priors can be used to obtain well-behaved parameter
+estimators, using them for model choice is much less certain.
 
@@
 been confused throughout the \msb literature.
-The mean (\meant{}), variance (\vart{}), and dispersion index
-($\vart{}/\meant{} = \vmratio{}$) of divergence times are \emph{not} parameters
+The mean (\meant{}{}), variance (\vart{}{}), and dispersion index
+($\vart{}{}/\meant{}{} = \vmratio{}$) of divergence times are \emph{not} parameters
 of the model implemented in \msb.
@@
 
+\mFigure{../images/coin_flip.pdf}{
+    A plot of three beta density functions meant to represent a prior
+    (black), true posterior (blue), and empirical Bayes density (red).
+}{figCoinFlip}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-09-18 response-redux/bin/plot-hickerson-et-al-results.py
Adding plotting script and plot of hickerson et al posterior.

@@ -1,0 +1,56 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes import plotting
+from pymsbayes.utils.parsing import spreadsheet_iter
+import project_util
+
+def main_cli():
+    result_dir = os.path.join(project_util.PROJECT_DIR,
+            'hickerson-et-al-posterior')
+    post_sample_path = os.path.join(result_dir, 'posterior-from-mike.txt')
+    model_i = range(1, 9)
+    mean_tau = dict(zip([i for i in model_i],[[] for i in model_i]))
+    omega = dict(zip([i for i in model_i],[[] for i in model_i]))
+    for d in spreadsheet_iter([post_sample_path]):
+        model_index = (int(d['PRI.model']))
+        mean_tau[model_index].append(float(d['PRI.E.t']))
+        omega[model_index].append(float(d['PRI.omega']))
+
+    scatter_data = {}
+    xmin, xmax = 0., 0.
+    ymin, ymax = 0., 0.
+    for i in model_i:
+        markeredgecolor = '0.5'
+        if i in [5, 6]:
+            markeredgecolor = '0.05'
+        x = omega[i]
+        y = mean_tau[i]
+        sd = plotting.ScatterData(x = x, y = y,
+                markeredgecolor = markeredgecolor)
+        scatter_data[i] = sd
+        xmin = min([xmin] + x)
+        ymin = min([ymin] + y)
+        xmax = max([xmax] + x)
+        ymax = max([ymax] + y)
+    xbuff = (xmax - xmin) * 0.04
+    ybuff = (ymax - ymin) * 0.04
+    xlim = (xmin - xbuff, xmax + xbuff)
+    ylim = (ymin - ybuff, ymax + ybuff)
+    
+    sp = plotting.ScatterPlot(
+            scatter_data_list = scatter_data.values(),
+            x_label = r'$Var(\tau)/E(\tau)$ ($\Omega$)',
+            y_label = r'$E(\tau)$',
+            xlim = xlim,
+            ylim = ylim)
+    rect = [0, 0, 1, 1]
+    sp.fig.tight_layout(pad = 0.25, rect = rect)
+    sp.reset_plot()
+    sp.savefig(os.path.join(result_dir, 'mean_by_dispersion.pdf'))
+
+if __name__ == '__main__':
+    main_cli()
+

### 2013-09-18 manuscripts/reply/reply.tex
More work on reply.
Adding methods for our re-analysis of the Philippines data using
the model-averaging approach of Hickerson et al. Also adding
table of results.

@@
 If we let \allParameterValues represent the set of all possible parameter
-values, we can define a prior distribution for all $\theta \subset
+values, we can define a prior distribution for all $\theta \in
 \allParameterValues$ such that $p(\theta)$ describes our belief that any given
@@
 If we let \allDatasets represent all possible datasets then we can 
-define a sampling model for all $\theta \subset
-\allParameterValues$ and $\alignment{}{} \subset \allDatasets$ such that
+define a sampling model for all $\theta \in
+\allParameterValues$ and $\alignment{}{} \in \allDatasets$ such that
 $p(\alignment{}{} | \theta)$ measures our belief that any dataset \alignment{}{}
@@
 estimators, using them for model choice is much less certain.
+
+\subsection*{Practical dangers of narrow uniform priors for Bayesian model
+choice}
+Given that \msb can be biased toward models with smaller parameter space, and
+Hickerson et al.'s \citeyear{Hickerson2013} use of narrow, empirically informed
+uniform priors on divergence times, we sought to determine whether such an
+approach will often exclude the truth (i.e., the model-averaged posterior will
+be dominated by models that excude the true values of parameters of the model).
+We explored this possiblity in two ways.
+First, we reanalyzed the Philippines dataset using the model-averaging approach
+of \citet{Hickerson2013}, but set one of prior models with a uniform prior on
+divergence times that is unrealistically narrow, and almost certainly excludes
+the truth.
+If Lindley's paradox causes the method to prefer models with less parameter
+space, we expect \msb will preferentially sample from this incorrect model
+yielding a marginal posterior that is invalid (i.e., the model-averaged
+posterior will be dominated by an incorrect model that excludes the truth).
+Second, we generated simulated datasets for which the divergence times are
+drawn form an exponential distribution and applied the approach of
+\citet{Hickerson2013} to each of them to see how often the method excludes the
+truth.
+
+For our re-analysis of the Philippines dataset we used the model-averaging
+approach of \citet{Hickerson2013}, but with different prior models to avoid
+mixing time units. We used five priors, all of which had priors on population
+sizes of $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and
+$\ancestralTheta{} \sim U(0.0001, 0.05)$.
+Following \citet{Hickerson2013}, each of these priors had the following
+priors on divergence time parameters:
+$M_1$, $\divt{} \sim U(0, 0.1)$;
+$M_2$, $\divt{} \sim U(0, 1)$;
+$M_3$, $\divt{} \sim U(0, 5)$;
+$M_4$, $\divt{} \sim U(0, 10)$; and
+$M_5$, $\divt{} \sim U(0, 20)$.
+For our prior, we simulated $1\e6$ random samples from each of the models
+for a total sample of $5\e6$.
+For each model, we retained the 10,000 samples with the smallest Euclidean
+distance from the observed summary statistics, standardizing the statitics
+using the prior means and standard deviations of the given model.
+We then retained the 10,000 samples with the smallest Euclidean distance from
+the observed summary statistics, this time standardizing the statistics using
+the prior means and standard deviations across all five models.
+We then repeated this analysis twice, replacing the $M_1$ with
+$M_1A$ and $M_1B$, which differ only by having priors on divergence
+times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
+respectively.
+While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
+\citet{Hickerson2013} likely excludes the true divergence times of
+at least some of the 22 taxa, we are nearly certain that these
+narrower priors are incorrect and exclude the truth.
 
@@
 
+%:TABLE-model-choice-empirical
+\begin{table}[htbp]
+    \sffamily
+    \footnotesize
+    \addtolength{\tabcolsep}{-0.08cm}
+    \rowcolors{2}{}{myGray}
+    %\captionsetup{font=footnotesize}
+    \caption{Results of the model-averaging approach of \citet{Hickerson2013}
+        applied to the Philippines dataset of \citet{Oaks2012} using three sets
+        of prior models. All models used priors on population size of
+        $\meanDescendantTheta{} \sim U(0.0001,0.1)$ and $\ancestralTheta{} \sim
+        U(0.0001, 0.05)$, and differ only in their prior on divergence time
+        (\divt{}) parameters.  Each set of five models differ only in the
+        divergence time prior used for the model with the narrowest prior:
+        $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_{1A}$ ($\divt{} \sim U(0, 0.01)$),
+        or $M_{1B}$ ($\divt{} \sim U(0, 0.001)$). The approximate posterior
+        probability of each model ($p(M_i \given \ssSpace)$) is given for each
+        of the three analyses.  The posterior estimates are based on 10,000
+        samples retained from $1\e6$ prior samples
+    from each model.}
+    \centering
+    \begin{tabular}{ l l l l l }
+        \toprule
+        & & \multicolumn{3}{c}{$p(M_i \given \ssSpace)$} \\
+        \cmidrule(){3-5}
+        Model & \divt{} prior & $M_{*}=M_1$ & $M_{*}=M_{1A}$ & $M_{*}=M_{1B}$ \\
+        \midrule
+        $M_*$ & --        & 0.899 & 0.821 & 0.673 \\
+        $M_2$ & $U(0,1)$  & 0.079 & 0.136 & 0.251 \\
+        $M_3$ & $U(0,5)$  & 0.013 & 0.026 & 0.044 \\
+        $M_4$ & $U(0,10)$ & 0.006 & 0.012 & 0.022 \\
+        $M_5$ & $U(0,20)$ & 0.003 & 0.005 & 0.010 \\
+        \bottomrule
+    \end{tabular}
+    \label{tabModelChoiceEmpirical}
+\end{table}
+
 \clearpage

### 2013-09-18 manuscripts/reply/reply.tex
More work on reply.
Adding simulation-based exclusion assessment methods and results.

@@
 narrower priors are incorrect and exclude the truth.
+
+Our results show that the model-averaging approach of \citet{Hickerson2013}
+does not reduce the methods bias towards models with smaller parameter space,
+but rather allow it to manifest in a very dangerous way.
+The method strongly prefers the prior model with the narrowest distribution on
+divergence times across all three of our analyses, even when this model is
+almost certainly incorrect and excludes the true divergence times of the
+Philippine taxa.
+Given that the results of \citet{Hickerson2013} strongly prefer the models with
+the narrowest prior on divergence times, it seems quite likely that their
+model-averaged results are dominated by models that exclude at least some of
+the 22 true divergence times, making their results difficult to interpret.
+
+To better quantify the propensity of Hickerson et al.'s
+\citeyear{Hickerson2013} approach to exclude the truth, we simulated
+1000 pseudo-observed datasets in which the divergence times for the
+22 population pairs are drawn randomly from an exponential distribution
+with a mean of 0.5 ($\divt{} \sim Exp(2)$). All other parameters were
+identically distributed as in the $M_1$--$M_5$ models.
+We then repeated the analysis described above, using $5\e6$ prior samples from
+$M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, and retaining 1,000 posterior samples
+for each of the 1,000 pseudo-observed datasets.
+
+Our results show that the model-averaging approach of \citet{Hickerson2013}
+excludes the true values of parameters in 97\% of the replicates (90\% with
+GLM-regression adjustment), excluding upto 21 of the 22 true divergence times
+(Figure~\ref{figExclusionSimTau}.
+Furthermore, the posterior probability of excluding at least one true parameter
+value is very high in nearly all of the replicates
+(Figure~\ref{figExclusionSimProb}).
+We used a Bayes factor (comparing models that exclude the truth to those that
+do not) calculated for each replicate of greater than 10 as strong support for
+excluding the truth.
+With that criterion, 66\% of the replicats (87\% with GLM-regression adjustment)
+strongly support the exclusion of true values (Figure~\ref{figExclusionSimProb}).
 

### 2013-09-19 model-choice/bin/generate-samples-for-pc-plot.sh
Adding PCA plots of M1 priors.

@@ -1,0 +1,49 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=16G
+#$ -l vf=16G
+#$ -q all.q
+#$ -pe orte 8
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
+
+nprocs=4
+nprior=1000
+batch_size=250
+nsums=1000
+seed=2617237416
+
+output_dir="../priors-for-pc-plot"
+if [ ! -d "$output_dir" ]
+then
+    mkdir -p $output_dir
+fi
+
+dmc.py --np $nprocs \
+    -r 1 \
+    -o ../configs/m1.cfg \
+    -p ../configs/m1*.cfg \
+    -n $nprior \
+    --prior-batch-size $batch_size \
+    --num-standardizing-samples $nsums \
+    --output-dir $output_dir \
+    --staging-dir $staging_dir \
+    --temp-dir $staging_dir \
+    --seed $seed \
+    --generate-samples-only
+
+echo "Here are the contents of the local temp directory '${staging_dir}':"
+ls -Fla $staging_dir
+echo 'Removing the local temp directory...'
+rm -r $staging_dir
+echo 'Done!'
+

### 2013-09-19 model-choice/bin/pc-plots.r

@@ -1,0 +1,62 @@
+rm(list=ls(all=TRUE))
+
+plot_pca = function(pca_obj, predict_obj, xlim, ylim) {
+    plot(pca_obj$scores[,1], pca_obj$scores[,2],
+         xlab = 'PC1',
+         ylab = 'PC2',
+         xlim = xlim,
+         ylim = ylim)
+    points(predict_obj[1], predict_obj[2],
+           col='blue',
+           lwd=3,
+           pch=8)
+}
+
+observed_path = '../../response-redux/results/sampling-error/pymsbayes-results/observed-summary-stats/observed-1.txt'
+m1_path = '../priors-for-pc-plot/pymsbayes-results/pymsbayes-output/prior-stats-summaries/m3-prior-sample.txt'
+m1a_path = '../priors-for-pc-plot/pymsbayes-results/pymsbayes-output/prior-stats-summaries/m2-prior-sample.txt'
+m1b_path = '../priors-for-pc-plot/pymsbayes-results/pymsbayes-output/prior-stats-summaries/m1-prior-sample.txt'
+plot_path = '../priors-for-pc-plot/pc-plots.pdf'
+
+stat_indices = c(94:181)
+
+observed <- read.table(observed_path, skip=1)
+m1 = read.table(m1_path, skip=1)[stat_indices]
+colnames(m1) = colnames(observed)
+m1a = read.table(m1a_path, skip=1)[stat_indices]
+colnames(m1a) = colnames(observed)
+m1b = read.table(m1b_path, skip=1)[stat_indices]
+colnames(m1b) = colnames(observed)
+
+m1_pca <- princomp(m1)
+m1a_pca <- princomp(m1a)
+m1b_pca <- princomp(m1b)
+
+observed_m1 = predict(m1_pca, observed)
+observed_m1a = predict(m1a_pca, observed)
+observed_m1b = predict(m1b_pca, observed)
+plot_letters = LETTERS[1:3]
+
+pdf(plot_path, width=7.087, height=2.4)
+    par(yaxs = "i", xaxs = "i")
+    par(cex.axis=0.7)
+    par(mgp = c(1.5,0.6,0)) # mgp adjusts space of axis labels
+    # oma is space around all plots
+    par(oma = c(0,0,0.7,0.4), mar = c(2.6,2.9,0.5,0.1))
[…]

### 2013-09-19 manuscripts/reply/reply.tex
More work on reply.

@@
 For each model, we retained the 10,000 samples with the smallest Euclidean
-distance from the observed summary statistics, standardizing the statitics
+distance from the observed summary statistics, standardizing the statistics
 using the prior means and standard deviations of the given model.
@@
 Philippine taxa.
+Unfortunately, ``checking'' the priors by plotting the summary statistics from
+1000 random samples of each prior model along the first two orthogonal axes of
+a principle component analysis, as recommended by \citet{Hickerson2013},
+provides no warning of a problem (Figure~\ref{figPCA}.
 Given that the results of \citet{Hickerson2013} strongly prefer the models with
@@
 
+\highLight{Add how we determined exclusion.}
 Our results show that the model-averaging approach of \citet{Hickerson2013}
 excludes the true values of parameters in 97\% of the replicates (90\% with
-GLM-regression adjustment), excluding upto 21 of the 22 true divergence times
+GLM-regression adjustment), excluding up to 21 of the 22 true divergence times
 (Figure~\ref{figExclusionSimTau}.
@@
 
+The results of our empirical and simulation-based analyses clearly demonstrate
+the danger of using narrow, empirically guided priors in a Bayesian
+model-averaging framework \ldots
+
+\subsection*{Final thought on EB}
 Given all of this, it is quite clear why one should use caution before
@@
 
+\mFigure{../../model-choice/priors-for-pc-plot/pc-plots.pdf}{
+    The graphical checks recommended by \citet{Hickerson2013} for three prior
+    models: (A) $M_1$ ($\divt{} \sim U(0, 0.1)$), (B) $M_{1A}$ ($\divt{} \sim
+    U(0, 0.01)$), and (C) $M_{1B}$ ($\divt{} \sim U(0, 0.001)$).
+    The plots project the summary statistics from 1000 random samples from each
+    model onto the first two orthogonal axes of a principle component analysis,
+    with the blue dot representing the observed summary statistics from the 22
+    population pairs of Philippine vertebrates.
+}{figPCA}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-09-19 manuscripts/reply/reply.tex
Fleshing out more text for reply.

@@
 empirical Bayesian model choice.
+
+\section*{Hickerson et al.'s re-analysis of the Philippines data}
 
@@
 \subsection*{Final thought on EB}
-Given all of this, it is quite clear why one should use caution before
-overly criticizing ones choice of priors after having seen the posterior.
-This really is a strawman argument.
-They were not there before the data were seen.
+Given all of the theoretical and practical issues with empirical Bayes
+approaches to Bayesian model choice discussed in the proceeding sections, it is
+quite clear why one should use caution before overly criticizing an
+investigators choice of priors after having seen the posterior.
+In a strictly Bayesian world, this is cheating.
+As discussed by \citet{Oaks2012}, prior to analysing the data, there was a large
+amount of uncertainty regarding the divergence times of the 22 population pairs
+under study.
+Two the these pairs represent distinct species, and the taxonomy of many groups
+in the Philippines has repeatedly been shown to mask deeply divergent lineages
+\citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
+Siler2011,Siler2012,RafeStuart2012}.
+\highLight{Any new stuff to add to this list of citations?}.
+\citet{Oaks2012} also discuss how the sole use of uniform priors in \msb makes
+it very difficult for investigators to express their prior uncertainty without
+putting a lot of prior density in regions of improbable parameter space.
+The alternative, as shown above, is to risk excluding the truth before even
+conducting the analysis.
+
+\section*{Our power analysis}
+\citet{Hickerson2013} did present a power analysis in their response, however,
+they did not follow the recommendations and conduct a power analysis to
+accompany their inference of clustered divergences using the Philippines data
+and their empirical Bayesian model-averaging approach.
+We present such an analysis here.
+
+Following \citet{Oaks2012}, we simulated 1000 pseudo-replicate datasets with
+\divt{} for each of the 22 population pairs randomly drawn from a uniform
+distribution, $U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6,
+0.8, 1.0, and 2.0, in \globalcoalunit generations.
+All other parameters were identically distributed as the prior.
+Following \citet{Hickerson2013} (but avoiding mixing time units), we used five
+priors, all of which had priors on population sizes of $\meanDescendantTheta{}
+\sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim U(0.0001, 0.05)$.
+Each of these priors had the following priors on divergence time parameters:
+$M_1$, $\divt{} \sim U(0, 0.1)$;
+$M_2$, $\divt{} \sim U(0, 1)$;
+$M_3$, $\divt{} \sim U(0, 5)$;
+$M_4$, $\divt{} \sim U(0, 10)$; and
+$M_5$, $\divt{} \sim U(0, 20)$.
+For our prior, we simulated $1\e6$ random samples from each of the models
+for a total sample of $5\e6$.
+For each pseudo-replicate dataset, we retained the 1000 samples from each prior
+model with the smallest Euclidean distance from the replicates' summary
+statistics, standardizing the statistics using the prior means and standard
+deviations of the given model.
+Then, for each replicate, we retained the 1000 samples with the smallest
+Euclidean distance from the pseudo-observed summary statistics, this time
+standardizing the statistics using the prior means and standard deviations
+across all five models.
+In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
+posterior samples from each of them.
+
+We find that the approach of \citet{Hickerson2013} to estimate the dispersion
+index of divergence times (\vmratio{}) across most of the \divt{max} we
+simulated, whether evaluating the unadjusted
+(Figure~\ref{figPowerAccOmegaMedian}) or GLM-adjusted
+(Figure~\ref{figPowerAccOmegaModeGLM}) posterior estimates.
+The method only estimates \vmratio{} relatively well when the simulated
+distribution of divergence times is identical to one of the prior models
+(Figures~\ref{figPowerAccOmegaMedian}E \& \ref{figPowerAccOmegaModeGLM}E).
+This is consistent with the conclusion of \citet{Oaks2012} that \msb lacks
+robustness and is highly sensitive to the prior distribution deviating from the
+true, underlying distribution of the data.
+
+Furthermore, our results demonstrate that the approach of \citet{Hickerson2013}
+consistently infers highly clustered divergences across all the \divt{max} we
+simulated (Figure~\ref{figPowerNumExcluded}).
+The method is most likely to infer the extreme case of a single divergence event
+when populations diverge randomly over the past \globalcoalunit generations.
+Even when divergences are random over the past $8\globalpopsize$ generations,
+the most likely inference is only two divergence events, and a single 
+divergence is still estimated in more than 10\% of the replicates.
+It is very interesting to note that as \divt{max} increases, but before the
+mode inference is finally pulled away from $\numt{} = 1$, the distribution of
+\numt{} estimates closely mirrors the odd U-shaped prior on divergence models
+used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
+\citeyear{Oaks2012} Figure 5B).
+This supports the conclusions of \citet{Oaks2012} that this U-shaped
+prior coupled with the poor marginal likelihoods of models with many
+\divt{} parameters, is a major cause of the method's bias toward
+clustered divergence models.
+
+Looking at our simulation results in terms of the posterior probability of the
+divergence-time dispersion index supporting the extreme case of one divergence
+event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the method
+strongly supports one divergence in greater than 27\%
+of the replicates across all the \divt{max} we simulated
+(Figure~\ref{figPowerOmegaProb}).
+Following \citet{Hickerson2013}, We use a Bayes factor of greater than 10 as
+the criterion for incorrect inference of a single divergence event.
+There is strong support for a single divergence event in more than 90\% of the
+replicates when divergences are random over the past $2.4\globalpopsize$
+generations, and more than 60\% when over the past $3.2\globalpopsize$
+generations or less (Figure~\ref{figPowerOmegaProb}).
+
+Contrary to Hickerson et al.'s \citeyear{Hickerson2013} claim that their
+model-averaging approach ``can discriminate complex multispecies histories and
+correctly reject synchronous divergence, even when discrete divergence times
+differ by much less than \ldots millions of generations,'' our results show
+that method is biased toward strongly supporting the extreme case of a single
+divergence event when populations diverged random over the last
+$8\globalpopsize$ generations.
+To roughly put that on the scale for a vertebrate mitochondrial locus, assuming
+a mutation rate of $2\e{-8}$ per site per generation, this translates to 5
+million generations.
+Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
+million generations.
+
+The results of our power analysis further demonstrate the
+propensity of Hickerson et al.'s \citeyear{Hickerson2013} approach
+to exclude true parameter values.
+Across all but one of the \divt{max} we simulated, the method excludes the
+truth in large proportion of replicates, and across many of the \divt{max} it
+will exclude a large proportion of the true divergence time values
+(Figure~\ref{figPowerNumExcluded}.
+The posterior probability of excluding at least on true divergence value is
+also quite high across many of the \divt{max}
+(Figure~\ref{figPowerProbExclusion}).
+Only when the data are identically distributed as one the prior models does
+the method avoid excluding the truth more than 5\% of the time
+(Figure~\ref{figPowerNumExcluded}E).
+
+Given our results, we conclude that not only does is the approach of
+\citet{Hickerson2013} dangerous in terms of excluding regions of parameter
+space containing the true, but still lacks power to detect random variation in
+divergence times over the scale of $8\globalpopsize$ generations or more.
+This translates to millions of generations for mitochondrial loci, and hundreds
+of millions of generations for nuclear loci.
 
@@
 
-\section*{Our power analysis}
+\citet{Hickerson2013} translate their results from units of
+\globalcoalunit generations to generations assuming a mutation rate of
+$1.92\e{-8}$ mutations per site per generation.
+If we tranlate there results to a scale more consistant with rates of mutation
+of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
+by \citet{Hickerson2013} the method can only reliably reject the extreme case
+of one divergence event when divergences are random over a period of more than
+3 million generations.
+Again, considering \citet{Hickerson2013} are assuming prior knowledge that
+all population pairs diverged within the past coalescent unit, this
+does not seem particularly powerful.
+
+\section*{Lindley's paradox versus sampling error}
+
+\subsection*{The probabilistic argument for sampling error}
 
@@
 
+\section*{Other clarifications}
+\begin{itemize}
+    \item saturation plot
+    \item different vocabulary (``simultaneous'', etc)
+    \item we did not conclude any inference of $\numt{} > 1$ is robust
+\end{itemize}
+
 \section*{Acknowledgments}
@@
 }{figJointPosterior}
+
+\mFigure{../../response-redux/hickerson-et-al-posterior/mean_by_dispersion.pdf}{
+    Hickerson et al.'s \citeyear{Hickerson2013} joint posterior.
+}{figJointPosteriorHickerson}
 

### 2013-09-19 response-redux/bin/rescale-posteriors.py
Adding `rescale-posteriors.py` to `response-redux`.

@@ -1,0 +1,170 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes import plotting
+from pymsbayes.utils.parsing import spreadsheet_iter
+from pymsbayes.fileio import process_file_arg
+from pymsbayes.utils.stats import get_freqs, freq_less_than, median, mode_list
+import project_util
+
+def rescale_posterior(in_path, out_path, scale_factor, model_indices):
+    header = None
+    out, close = process_file_arg(out_path, 'w', compresslevel=9)
+    omegas = []
+    psis = []
+    for i, d in enumerate(spreadsheet_iter([in_path])):
+        if i == 0:
+            header = d.keys()
+            out.write('{0}\n'.format('\t'.join(header)))
+            continue
+        model_index = int(d['PRI.model'])
+        if model_index in model_indices:
+            d['PRI.E.t'] = float(d['PRI.E.t']) * scale_factor
+            d['PRI.var.t'] = float(d['PRI.var.t']) * (scale_factor * 0.5)
+            d['PRI.omega'] = float(d['PRI.omega']) * scale_factor
+            omegas.append(d['PRI.omega'])
+            psis.append(int(d['PRI.Psi']))
+        out.write('{0}\n'.format('\t'.join([
+                str(d[k]) for k in d.iterkeys()])))
+    out.close()
+    return omegas, psis
+
+def get_omega_and_mean_tau(post_path, model_indices):
+    mean_tau = dict(zip([i for i in model_indices],[[] for i in model_indices]))
+    omega = dict(zip([i for i in model_indices],[[] for i in model_indices]))
+    for d in spreadsheet_iter([post_path]):
+        model_index = (int(d['PRI.model']))
+        mean_tau[model_index].append(float(d['PRI.E.t']))
+        omega[model_index].append(float(d['PRI.omega']))
+    return omega, mean_tau
+
+def get_posterior_plot(post_path, model_indices, scaled_model_indices,
+        xlim = None, ylim = None,
+        x_label = r'$Var(\tau)/E(\tau)$ ($\Omega$)',
+        y_label = r'$E(\tau)$'):
+    omega, mean_tau = get_omega_and_mean_tau(post_path, model_indices)
+    scatter_data = {}
+    xmin, xmax = 0., 0.
+    ymin, ymax = 0., 0.
+    for i in model_indices:
+        markeredgecolor = '0.5'
+        if i in scaled_model_indices:
[…]

### 2013-09-20 manuscripts/reply/reply.tex
More work on reply.
Adding work exploring odd Hickerson et al results.
Adding scripts for rescaling invalid posteriors to be on in the
same units of time. Also, adding scripts for performing rejection
and regression on Hickerson et al.'s 80,000 best samples from
the global prior.

Moving some files around and updating figure path in `reply.tex`
accordingly.

@@
 truth.
+
+The empirically guided, narrowest prior of \citet{Hickerson2013} ($\divt{} \sim
+U(0,0.1)$) likely excludes the true values for at least some of the Philippine
+taxa.
+The prior matches posterior of \emph{mean} divergence time and is much
+narrower than posterior under two of three priors.
 
@@
 
-\mFigure{../../response-redux/hickerson-et-al-posterior/mean_by_dispersion.pdf}{
+\mFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
     Hickerson et al.'s \citeyear{Hickerson2013} joint posterior.

### 2013-09-20 response-redux/bin/plot-results.py

@@ -5,54 +5,51 @@ import sys
 
 from pymsbayes import plotting
 from pymsbayes.utils.parsing import spreadsheet_iter
+from pymsbayes.utils.stats import get_freqs, freq_less_than, median, mode_list
+
 import project_util
+from rescale_posteriors import get_posterior_plot, summarize_results
+
+def plot_posterior(post_path):
+    model_indices = range(1, 9)
+    scaled_indices = [5, 6]
+    sp, xlim, ylim = get_posterior_plot(post_path, model_indices,
+                scaled_indices)
+    rect = [0, 0, 1, 1]
+    sp.fig.tight_layout(pad = 0.25, rect = rect)
+    sp.reset_plot()
+    sp.savefig(os.path.join(os.path.dirname(post_path),
+            'mean_by_dispersion.pdf'))
 
 def main_cli():
+    prior_prob_omega_less_than = 0.0887
     for d in ['hickerson', 'alt']:
         result_dir = os.path.join(project_util.RESULT_DIR, d,
                 'pymsbayes-results')
-        post_sample_path = os.path.join(result_dir, 'pymsbayes-output',
+        post_path = os.path.join(result_dir, 'pymsbayes-output',
                 'd1', 'm12345678-combined',
                 'd1-m12345678-combined-s1-25-posterior-sample.txt.gz')
-        model_i = range(1, 9)
-        mean_tau = dict(zip([i for i in model_i],[[] for i in model_i]))
-        omega = dict(zip([i for i in model_i],[[] for i in model_i]))
-        for d in spreadsheet_iter([post_sample_path]):
-            model_index = (int(d['PRI.model']))
-            mean_tau[model_index].append(float(d['PRI.E.t']))
-            omega[model_index].append(float(d['PRI.omega']))
+        plot_posterior(post_path)
+        summarize_results(post_path)
 
-        scatter_data = {}
-        xmin, xmax = 0., 0.
-        ymin, ymax = 0., 0.
-        for i in model_i:
-            markeredgecolor = '0.5'
-            if i in [5, 6]:
-                markeredgecolor = '0.05'
-            x = omega[i]
-            y = mean_tau[i]
-            sd = plotting.ScatterData(x = x, y = y,
-                    markeredgecolor = markeredgecolor)
-            scatter_data[i] = sd
-            xmin = min([xmin] + x)
-            ymin = min([ymin] + y)
[…]

### 2013-09-20 response-redux/bin/rescale_posteriors.py

@@ -9,6 +9,34 @@ from pymsbayes.fileio import process_file_arg
 from pymsbayes.utils.stats import get_freqs, freq_less_than, median, mode_list
 import project_util
 
+def parse_results(posterior_path):
+    omegas = []
+    psis = []
+    models = []
+    for i, d in enumerate(spreadsheet_iter([posterior_path])):
+        models.append(int(d['PRI.model']))
+        omegas.append(float(d['PRI.omega']))
+        psis.append(int(d['PRI.Psi']))
+    return models, psis, omegas
+
+def summarize_results(posterior_path):
+    sum_path = posterior_path + '.summary.txt'
+    models, psis, omegas = parse_results(posterior_path)
+    psi_mode = mode_list(psis)
+    omega_mode = mode_list(omegas)
+    omega_median = median(omegas)
+    psi_probs = get_freqs(psis)
+    omega_prob = freq_less_than(omegas, 0.01)
+    model_probs = get_freqs(models)
+    with open(sum_path, 'w') as out:
+        out.write('\nResults for {0!r}:\n'.format(posterior_path))
+        out.write('psi_mode = {0}\n'.format(psi_mode))
+        out.write('psi_probs = {0}\n'.format(psi_probs))
+        out.write('omega_mode = {0}\n'.format(omega_mode))
+        out.write('omega_median = {0}\n'.format(omega_median))
+        out.write('omega_prob_less_than = {0}\n'.format(omega_prob))
+        out.write('model_probs = {0}\n'.format(model_probs))
+
 def rescale_posterior(in_path, out_path, scale_factor, model_indices):
     header = None
     out, close = process_file_arg(out_path, 'w', compresslevel=9)
@@ -18,7 +46,6 @@ def rescale_posterior(in_path, out_path, scale_factor, model_indices):
         if i == 0:
             header = d.keys()
             out.write('{0}\n'.format('\t'.join(header)))
-            continue
         model_index = int(d['PRI.model'])
         if model_index in model_indices:
             d['PRI.E.t'] = float(d['PRI.E.t']) * scale_factor
@@ -105,36 +132,47 @@ def plot_posteriors(post_path, scaled_post_path, model_indices = range(1, 9),
     pg.reset_figure()
     return pg
     
-def main_cli():
+def process_posterior(post_path, scaled_post_path):
     prior_prob_omega_less_than = 0.0887
     scaled_indices = [5, 6]
[…]

### 2013-09-20 response-redux/bin/run-rejection-on-hickerson-et-al.py

@@ -1,0 +1,67 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes.workers import EuRejectWorker, ABCToolBoxRegressWorker
+from pymsbayes.utils.tempfs import TempFileSystem
+import project_util
+
+
+def get_regression_worker(temp_fs, posterior_path, observed_path):
+    w = ABCToolBoxRegressWorker(
+            temp_fs = temp_fs,
+            observed_path = observed_path,
+            posterior_path = posterior_path,
+            num_posterior_quantiles = 10000,
+            compress = True)
+    return w
+    
+def get_reject_worker(temp_fs, prior_paths, observed_path, summary_path,
+        out_dir):
+    post_path = os.path.join(out_dir, 'posterior-sample.txt')
+    reg_worker = get_regression_worker(temp_fs, post_path, observed_path,)
+    w = EuRejectWorker(
+            temp_fs = temp_fs,
+            observed_path = observed_path,
+            prior_paths = prior_paths,
+            num_posterior_samples = 10000,
+            summary_in_path = summary_path,
+            posterior_path = post_path,
+            regression_worker = reg_worker)
+    return w
+
+
+    
+def main_cli():
+    result_dir = os.path.join(project_util.PROJECT_DIR,
+            'hickerson-et-al-posterior')
+    assert os.path.isdir(result_dir)
+    per_model_post_path = os.path.join(result_dir, 'per-model-posteriors.txt')
+    assert os.path.isfile(per_model_post_path)
+    observed_path = os.path.join(project_util.RESULT_DIR, 'sampling-error',
+            'pymsbayes-results', 'observed-summary-stats', 'observed-1.txt')
+    assert os.path.isfile(observed_path)
+    summary_path = os.path.join(project_util.RESULT_DIR, 'hickerson',
+            'pymsbayes-results', 'pymsbayes-output', 'prior-stats-summaries',
+            'm12345678-combined-stat-means-and-std-devs.txt')
+    assert os.path.isfile(summary_path)
+
+    eureject_dir = os.path.join(result_dir, 'eureject-results')
+    if not os.path.exists(eureject_dir):
+        os.mkdir(eureject_dir)
+
+    temp_fs = TempFileSystem(parent = eureject_dir)
+
+    w = get_reject_worker(
+            temp_fs = temp_fs,
+            prior_paths = [per_model_post_path],
+            observed_path = observed_path,
[…]

### 2013-09-21 manuscripts/reply/reply.tex
Working on text in reply.
Working on reply.
Also moving `mean_by_dispersion.pdf` plot in the `response-redux`
results.

@@
 phylogeographical model choice.
+The discussion is centered around the approximate Bayesian model choice method,
+\msb, which estimates the temporal distribution of divergences among
+co-distributed pairs of taxa.
 \citet{Oaks2012} published their findings that the method \msb can often be
@@
 \section*{Hickerson et al.'s re-analysis of the Philippines data}
+\citet{Hickerson2013} re-analyze the dataset of \citet{Oaks2012} using a
+model-averaging approach, where they place a discrete uniform prior over eight
+different prior models (see Table 1 of \citet{Hickerson2013}).
+In our initial attempts to explore the behavior of this approach, we found
+a fundamental error in the methodology of \citet{Hickerson2013}:
+their model mixes different time units.
 
-\section*{EB}
-\citet{Hickerson2013} repeatedly refer the prior distributions used by
+Each of the eight prior models used in the re-analysis by \citet{Hickerson2013}
+has one of two priors on the mean population size of the descendant populations
+of each taxon pair:
+$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ or
+$\meanDescendantTheta{} \sim U(0.0005, 0.04)$ or
+As described in \citet{Oaks2012}, the divergence time parameters of the model
+implemented in \msb are scaled relative to a constant reference population
+size, \myTheta{C}.
+This reference population size is defined in terms of the upper limit on the
+uniform prior on the mean population size of the descendant populations,
+\meanDescendantTheta{}, such that for the prior $\meanDescendantTheta{} \sim
+U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
+$\myTheta{C} = \uniformMax{\meanDescendantTheta{}}/2$.
+Thus, the model used by \citet{Hickerson2013} mixes two different units of
+time.
+In other words, some of their prior and posterior samples are in units of
+$0.05/\mutationRate$ generations, whereas other are in units of
+$0.02/\mutationRate$ generations.
+
+The fact that their posterior samples are in different units makes the results
+of \citet{Hickerson2013} difficult to interpret, and renders their
+regression-adjusted results invalid; a fundamental assumption of regression is
+that all of the values of the response variable are in the same units.
+Thus, the results in section ``Improved Sampling Efficiency by Prior Weighting
+Supports Asynchronous and Recent Divergence for the Philippines Vertebrate
+Data'' and presented in Figure 2 of \citet{Hickerson2013} should be
+disregarded.
+The error is easily illustrated by re-plotting their results with the different
+time units indicated (Figure~\ref{figJointPosteriorHickerson}.
+
+% In addition to this error, we were not able to reproduce the results of
+% \citet{Hickerson2013}.
+% We followed their methods, generating $5\e6$ prior samples for each of the
+% eight models in their Table 1.
+% We retained the 10,000 samples from each model with the smallest Euclidean
+% distance to the observed summary statistics, and subsequently retained 10,000
+% samples from the remaining 80,000.
+
+\section*{The dangers of empirical Bayesian model choice}
+\citet{Hickerson2013} repeatedly refer to the prior distributions used by
 \citet{Oaks2012} as ``poorly selected''.
@@
 distributions (i.e., they did not use an empirical Bayesian approach).
-\citet{Oaks2012} did also use empirically informed priors when assessing
+\citet{Oaks2012} did use empirically informed priors when assessing
 both the prior sensitivity of the method and determining the power of \msb
@@
 model choice (see last paragraph of ``Assessing prior sensitivity of \msb''),
-and we expand on this here.
+and we expand on this here by exploring the theoretical and practical
+implications of using an empirical Bayes approach to model choice.
 
@@
 taxa.
-The prior matches posterior of \emph{mean} divergence time and is much
-narrower than posterior under two of three priors.
+They set this prior to match the 95\% highest posterior density (HPD) interval for
+the mean divergence time estimated under one of the priors used by
+\citet{Oaks2012} (see Table 2 and 3 of \citet{Oaks2012}).
+Given this interval estimate is for the \emph{mean} divergence time across all
+22 taxa, it seems odd to set this as the prior, as many of the taxon pairs
+likely diverged at times older than the upper limit.
+Furthermore, the prior is \emph{excluded} from the 95\% HPD interval estimates
+of the mean divergence time under the other two prior analyses (under these
+priors the 95\% HPD is approximately 0.3--0.6; see Table~6 of
+\citet{Oaks2012}).
 
@@
 $1.92\e{-8}$ mutations per site per generation.
-If we tranlate there results to a scale more consistant with rates of mutation
+If we tranlate there results to a scale more consistent with rates of mutation
 of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
@@
 \section*{Lindley's paradox versus sampling error}
+One of the points of disagreement between \citet{Hickerson2013} and
+\citet{Oaks2012} is the underlying mechanism causing the bias toward models
+with clustered divergences (i.e., models with few divergence events).
+\citet{Oaks2012} present simulations based analyses that suggest the broad
+uniform prior on divergence time parameters hinder the marginal likelihoods of
+models with more divergence time parameters, whereas \citet{Hickerson2013}
+argue that the bias is inherent to the inefficient rejection algorithm
+implemented in \msb.
 
-\subsection*{The probabilistic argument for sampling error}
+\citet{Hickerson2013} present an interesting probabilistic argument in an
+attempt to show that insufficient prior sampling is to blame for the bias.
+They argue the widest of the three priors used by \citet{Oaks2012} was very
+unlikely to produce any prior samples with high numbers of divergence times
+that were consistent with the Philippine data.
+There are a few issues with their argument.
+First, the probabilities they present assume the gene divergence times
+estimated by \citet{Oaks2012} are correct.
+The sole purpose of the estimates of ultrametric gene trees presented by
+\citet{Oaks2012} was to provide a very rough comparison of the gene divergence
+times across the 22 taxa.  These analyses assumed an arbitrary strict clock of
+$2\e{-8}$ for all taxa, and are, of course, subject to estimation error.
+Furthermore, the units of the gene trees are in millions of year, whereas the
+divergence time prior of \msb is in generations, thus the logic of
+\citet{Hickerson2013} requires the additional assumption that all 22 Philippine
+taxa have a generation time of one year.
+Thus, the argument of \citet{Hickerson2013} that divergence time estimates of
+\citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
+dangerous, especially given our findings presented above.
+
+Finally, even \emph{if} we make all of these assumptions and assume the
+gene divergence times are without error, the probabilistic argument only
+applies to one of the three different priors used by \citet{Oaks2012}.
+The narrowest prior on divergence times used by \citet{Oaks2012} closely mirrors
+the range of gene divergence time estimates, and applying Hickerson et al.'s
+\citeyear{Hickerson2013} probability equations demonstrates that the prior
+is densely populated with samples with large number of divergence parameters
+that are consistent with the gene divergence estimates
+Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
+sampling is to blame for the bias, it should be much reduced under the narrow
+prior on \divt{}.
+However, the magnitude of the bias is very similar across all three priors
+\citet{Oaks2012}.
+\citet{Hickerson2013} point out a case where the narrow prior performs
+slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
+However, it is important to note that these results suffered from a bug
+in \msb, and there are many cases after \citet{Oaks2012} corrected the 
+bug where the narrow prior performs slightly worse (see panels D--J of
+Figures 3 and S12).
+
+To disentangle whether Lindely's paradox or insufficient prior sampling is
+to blame for the biases revealed by \citet{Oaks2012}, we must look at
+the different predictions made by these two phenomena.
+One examples, as discussed by \citet{Oaks2012} is that insufficient
+prior sampling should create higher variance in posterior estimates,
+and thus the analyses should be sensitive to the number of samples
+drawn from the prior.
+\citet{Oaks2012} do not see such sensitivity when they compare prior sample
+sizes of $2\e6$, $5\e6$, and $10\e7$.
+
+To explore this prediction further, we repeat the analysis of the Philippines
+dataset under the intermediate prior used by \citet{Oaks2012} ($\divt \sim U(0,
+10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
+(0.0005, 0.02)$), using a very large prior sample size of $10\e8$.
+When we look at the trace of the estimates of the dispersion index of
+divergence times (\vmratio{}) as the prior samples accumulate
+(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or GLM
+regression adjusted estimates.
+This strongly suggests that insufficient prior sampling did not play a large
+role in the bias found by \citet{Oaks2012}.
+
+Our results presented above that demonstrate the bias of the model-averaging
+approach of \citet{Hickerson2013} both toward models with narrower \divt{}
+priors (Table~\ref{tabModelChoiceEmpirical} and
+Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
+and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
+\ref{figPowerOmegaProb}) strongly implicate Lindley's paradox as a primary
+cause.
+In all of our model-averaging analyses, all prior models have the same number
+of samples.
+Thus, it is difficult to explain the method's propensity towards the model with
+the smallest parameter space with insufficient sampling.
+While analyses that sample each model proportional to their relative parameter
+space could be explored, it seems very likely that the broad uniform priors
+are simply inhibiting the marginal likelihoods of spacious models.
+
+As discussed by \citet{Oaks2012}, a straightforward prediction of Lindley's
+paradox is that it should disappear as the model generating the data converges
+to the prior.
+\citet{Oaks2012} test this prediction by performing 100,000 simulations to
+assess the model-choice behavior of \msb when the prior model is correct.
+They find that the bias actually switches direction (at least for the
+regression-adjusted estimates) and the method tends to underestimate the
+probability of the model with one divergence event (see Figure 4 of
+\citet{Oaks2012}).
+The same prediction is not as straightforward for the insufficient-sampling
+hypothesis.
+Even when the prior is correct, due to the discrete uniform prior on
+\numt{} implemented in \msb, models with larger numbers of divergence
+events (and thus greater parameter space) will still be less densely
+sampled than those with fewer divergence events \citep{Oaks2012}.
 
@@
 \section*{Other clarifications}
+
+\subsection*{saturation plot}
+\citet{Hickerson2013} claim that our priors ``cause much of the explored
+parameter space to be beyond the threshold of saturation in most mtDNA genes.''
+To explore this possibility, we simply plotted the summary statistics used
+over divergence times to look for the point when the statistics begin to
+saturate (Figure~\ref{figSaturationPlot}).
+Clearly, the priors with upper limits on \divt{} of 5 and 10 did not
+suffer from saturation.
+Even as far out as divergence times as 20 coalescent units, while 
+there is some saturation, there is still clearly signal in the
+summary statistics used by \msb.
+
 \begin{itemize}
-    \item saturation plot
     \item different vocabulary (``simultaneous'', etc)
@@
 
+%:TABLE-re-analysis
+% \begin{table}[htbp]
+%     \sffamily
+%     \footnotesize
+%     \addtolength{\tabcolsep}{-0.08cm}
+%     \rowcolors{2}{}{myGray}
+%     %\captionsetup{font=footnotesize}
+%     \caption{Results of the model-averaging approach of \citet{Hickerson2013}
+%         applied to the Philippines dataset of \citet{Oaks2012} using eight
+%         prior models (see \citet{Hickerson2013} Table 1).  The results of (1)
+%         \citet{Hickerson2013}, (2) our full re-analysis, and of (3) repeating
+%         the final rejection step of \citet{Hickerson2013} using the prior means and
+%         standard deviations to normalize the summary statistics, are all compared.}
+%     \centering
+%     \begin{tabular}{ l l l l }
+%         \toprule
+%          & Hickerson et al. (2013) & Re-analysis & blah \\
+%         $\hat{\numt{}}$ & 2 & 1 & 1 \\
+%         $p(\numt{} = 1 | \ssSpace)$ & 0.062 & 0.206 & 0.146 \\
+%         $BF_{\numt{} = 1, \numt{} \neq 1}$ 1.39 & 5.45 & 3.59 \\
+%         $p(\vmratio{} < 0.01 | \ssSpace)$ & 0.298 & 0.416 & 0.361 \\
+%         $BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01}$ 6.65 & 11.16 & 8.85 \\
+%         $\hat{M}$ & 1 & 2 & 2 \\
+%         $p(M_1 | \ssSpace)$ & 0.513 & 0.358 & 0.376 \\
+%         $p(M_2 | \ssSpace)$ & 0.286 & 0.369 & 0.389 \\
+%         $p(M_3 | \ssSpace)$ & 0.036 & 0.086 & 0.089 \\
+%         $p(M_4 | \ssSpace)$ & 0.002 & 0.012 & 0.011 \\
+%         $p(M_5 | \ssSpace)$ & 0.142 & 0.129 & 0.107 \\
+%         $p(M_6 | \ssSpace)$ & 0.020 & 0.043 & 0.024 \\
+%         $p(M_7 | \ssSpace)$ & 0.001 & 0.003 & 0.004 \\
+%         $p(M_8 | \ssSpace)$ & 0.000 & 0.000 & 0.001 \\
+%         \bottomrule
+%     \end{tabular}
+%     \label{tabModelChoiceEmpirical}
+% \end{table}
+
 %:TABLE-model-choice-empirical
@@
     \sffamily
-    \footnotesize
+    % \footnotesize
     \addtolength{\tabcolsep}{-0.08cm}
@@
 
-\mFigure{../../response-redux/results/hickerson/pymsbayes-results/mean_by_dispersion.pdf}{
-    Results of reanalysis.
-}{figJointPosterior}
+% \mFigure{../../response-redux/results/hickerson/pymsbayes-results/pymsbayes-output/d1/m12345678-combined/mean_by_dispersion.pdf}{
+%     Results of reanalysis.
+% }{figJointPosterior}
 
@@
 \mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
-    Sampling error \ldots
+    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
+    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
+    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
+    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
+    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
+    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
 }{figSamplingError}

### 2013-09-22 manuscripts/reply/reply.tex
Updating reply.
Fleshing out sections throughout, and adding figure captions.

@@
 results of additional analyses, which strongly suggest that Lindley's paradox
-is playing a larger role in the poor behavior of the method.
+is playing a large role in the poor behavior of the method.
 
@@
 \section*{Our power analysis}
-\citet{Hickerson2013} did present a power analysis in their response, however,
-they did not follow the recommendations and conduct a power analysis to
-accompany their inference of clustered divergences using the Philippines data
-and their empirical Bayesian model-averaging approach.
-We present such an analysis here.
+\citet{Hickerson2013} did not conduct a power analysis to accompany their
+inference of clustered divergences from the Philippines data (as recommended by
+\citet{Oaks2012}).
+Thus, we perform perform simulation-based power analyses to explore the
+power of the approximate Bayesian model-averaging approach championed
+by \citet{Hickerson2013}.
 
@@
 
+\section{Validation analyses}
+
 \section*{Lindley's paradox versus sampling error}
@@
 models and prior models.
-This certainly is a non-trivial extension of the model, and given the methods
-very large number of parameters and models relative to the information
-used from the data, likely plays a major role in the poor behavior of this
-approach.
+This certainly is a non-trivial extension of the model, and given the method's
+very large number of parameters and models relative to the information used
+from the data, likely plays a major role in the poor behavior of this approach.
+Certainly, such a large modification of the model demands simulation-based
+assessments of the behavior of the new model, as we provide here.
 
@@
 
-\subsection*{saturation plot}
+\subsection*{The validity of \msb estimates}
+\citet{Hickerson2013} claim that \citet{Oaks2012} ``assume that all previous
+\msb results are invalid.''
+With the exception of \citet{Hickerson2013}, we do not claim any previous
+results of \msb are invalid.
+Rather, we conservatively recommend that the common inference of temporally
+clustered divergences
+\citep{Barber2010, Bell2012, Carnaval2009, Chan2011, Daza2010, Hickerson2006,
+    Huang2011, Lawson2010, Leache2007, Plouviez2009, Stone2012, Voje2009},
+when not accompanied with the necessary simulation-based analyses to guide the
+interpretation of such results, should be treated with caution, because the
+method has been shown to spuriously infer clustered divergences over a range of
+prior conditions.
+
+\subsection*{Saturation of summary statistics}
 \citet{Hickerson2013} claim that our priors ``cause much of the explored
 parameter space to be beyond the threshold of saturation in most mtDNA genes.''
-To explore this possibility, we simply plotted the summary statistics used
-over divergence times to look for the point when the statistics begin to
-saturate (Figure~\ref{figSaturationPlot}).
-Clearly, the priors with upper limits on \divt{} of 5 and 10 did not
-suffer from saturation.
-Even as far out as divergence times as 20 coalescent units, while 
-there is some saturation, there is still clearly signal in the
-summary statistics used by \msb.
-
-\begin{itemize}
-    \item different vocabulary (``simultaneous'', etc)
-    \item we did not conclude any inference of $\numt{} > 1$ is robust
-\end{itemize}
+To explore this possibility, we simulated datasets under prior settings that
+match two of the three priors used by \citet{Oaks2012}: $\meanDescendantTheta
+\sim U(0.0005, 0.04)$ and $\ancestralTheta \sim U(0.0005, 0.02)$.
+We draw divergence time parameters from a uniform distribution of $U(0, 20)$,
+simulated datasets, and plotted the \divt{} values against the summary
+statistics calculated from the resulting datasets
+(Figure~\ref{figSaturationPlot}).
+Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of 5
+and 10 did not suffered from little to no effect from saturation.  Even at
+divergence times of 20 coalescent units, there is still signal in the summary
+statistics used by \msb (Figure~\ref{figSaturationPlot}).
 
 \section*{Acknowledgments}
-
-% \bibliography{../bib/references}
-\bibliography{references}
+We thank the National Science Foundation for supporting this work (DEB
+1011423).
+J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
+Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
+Society, KU Department of Ecology and Evolutionary Biology, and the KU
+Biodiversity Institute.
+
+\bibliography{../bib/references}
+% \bibliography{references}
 
@@
 \mFigure{../../saturation/saturation-plot.pdf}{
-    Saturation plot \ldots
+    The summary statistics $\pi$ \citep{Tajima1983} and $\pi_{net}$
+    \citep{Takahata1985} as a function of divergence time between populations.
+    Each plot represents 1100 pairs of parameter draws and summary statistics
+    calculated from the simulated data.
+    Prior settings for the simulations were $\divt{} \sim U(0, 20)$,
+    $\meanDescendantTheta{} \sim U(0.0005, 0.04)$, and $\ancestralTheta{} \sim
+    U(0.0005, 0.02)$.
 }{figSaturationPlot}
@@
 \mFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
-    Hickerson et al.'s \citeyear{Hickerson2013} joint posterior.
+    The joint posterior of the mean (\meant{}{}) and dispersion index ($\vmratio{} = 
+    \vart{}{}/\meant{}{}$) of divergence times for 22 vertebrate taxon pairs as
+    estimated by \citet{Hickerson2013} (see Figure 2B of \citet{Hickerson2013}).
+    The posterior samples are color-coded to indicate the erroneous mixture of
+    timescales in the analysis of \citet{Hickerson2013};
+    grey = $0.05/\mutationRate$ generations and
+    black = $0.02/\mutationRate$ generations.
 }{figJointPosteriorHickerson}
@@
 \mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/num_tau_excluded.pdf}{
-    Histogram of the number of \divt{} parameters excluded.
+    The propensity of the model-averaging approach of \citet{Hickerson2013} to
+    exclude the truth.
+    The plots illustrate the number of true \divt{} parameters excluded from analyses of
+    simulated datasets where \divt{} for 22 pairs of populations is drawn from an
+    exponential distribution, $\divt{} \sim Exp(2)$.
+    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
+    simulation replicates analyzed using $5\e6$ samples from the prior.
+    The proportion of simulation replicates in which at least one true
+    parameter value is excluded from the model preferred by a Bayes factor
+    ($p(\divt{} \notin \hat{M})$) is also given.
 }{figExclusionSimTau}
@@
 \mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/prob_of_exclusion.pdf}{
-    Histogram of the posterior probability of excluding the truth.
+    The propensity of the model-averaging approach of \citet{Hickerson2013} to
+    exclude the truth.
+    The plots illustrate the estimated probability of excluding at least one
+    true \divt{} value from analyses of simulated datasets where \divt{} for 22
+    pairs of populations is drawn from an exponential distribution, $\divt{}
+    \sim Exp(2)$.
+    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
+    simulation replicates analyzed using $5\e6$ samples from the prior.
+    The proportion of simulation replicates in which there is strong support
+    for at least one true parameter value being excluded from the model
+    ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is also given.
 }{figExclusionSimProb}
@@
 \mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
-    Histograms of the posterior probability that $\vmratio{} < 0.01$.
+    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
+    estimate the dispersion index of divergenct times (\vmratio{}) from
+    analyses of simulated datasets where \divt{} for 22 pairs of populations is
+    drawn from a series of uniform distributions, $\divt{} \sim U(0,
+    \divt{max})$.
+    The proportion of estimates less than the true value of,
+    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
+    Each plot represents unadjusted median estimates from 500 simulation
+    replicates analyzed using $5\e6$ samples from the prior.
 }{figPowerAccOmegaMedian}
@@
 \mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
-    Histograms of the posterior probability that $\vmratio{} < 0.01$.
+    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
+    estimate the dispersion index of divergenct times (\vmratio{}) from
+    analyses of simulated datasets where \divt{} for 22 pairs of populations is
+    drawn from a series of uniform distributions, $\divt{} \sim U(0,
+    \divt{max})$.
+    The proportion of estimates less than the true value of,
+    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
+    Each plot represents GLM-regression-adjusted mode estimates from 500
+    simulation replicates analyzed using $5\e6$ samples from the prior.
 }{figPowerAccOmegaModeGLM}
@@
 \mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_psi_mode.pdf}{
-    Histograms of \numt{} estimates.
+    The bias of the model-averaging approach of \citet{Hickerson2013} to infer
+    clustered divergences in our simulation-based power analyses.
+    The plots illustrate the estimated number of divergence events ($\hat{\numt{}}$)
+    from analyses of simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
+    uniform distributions, $\divt{} \sim U(0, \divt{max})$.
+    The estimated probability of the method inferring on divergecne event, $p(\numt{} = 1)$,
+    is given for each \divt{max}.
+    Each plot represents 500 simulation replicates analyzed using $5\e6$
+    samples from the prior.
 }{figPowerPsiMode}
@@
 \mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_omega_prob.pdf}{
-    Histograms of the posterior probability that $\vmratio{} < 0.01$.
+    The bias of the model-averaging approach of \citet{Hickerson2013} to
+    support one divergence event in our simulation-based power analyses.
+    The plots illustrate histograms of the estimated posterior probability that
+    the dispersion index of divergence times is less than 0.01 ($p(\vmratio{} <
+    0.01 | \ssSpace)$) from analyses of simulated datasets where \divt{} for 22
+    pairs of populations is drawn from a series of uniform distributions,
+    $\divt{} \sim U(0, \divt{max})$.
+    The proportion of simulation replicates that strongly support one
+    divergence event, $p(BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01} > 10)$,
+    is given for each \divt{max}.
+    Each plot represents 500 simulation replicates analyzed using $5\e6$
+    samples from the prior.
 }{figPowerOmegaProb}
 
+\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_num_excluded.pdf}{
+    The propensity of the model-averaging approach of \citet{Hickerson2013} to exclude
+    the truth in our simulation-based power analyses.
+    The plots illustrate the number of true \divt{} parameters excluded from analyses of
+    simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
+    uniform distributions, $\divt{} \sim U(0, \divt{max})$. The proportion
+    of simulation replicates in which at least one true parameter value is excluded from
+    the model preferred by a Bayes factor ($p(\divt{} \notin \hat{M})$) is
+    given for each \divt{max}.
+    Each plot represents 500 simulation replicates analyzed using $5\e6$
+    samples from the prior.
+}{figPowerNumExcluded}
+
 \mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_prob_exclusion.pdf}{
-    Histograms of the posterior probability of excluding the truth.
+    The propensity of the model-averaging approach of \citet{Hickerson2013} to
+    exclude the truth in our simulatoin-based power analyses.  The plots
+    illustrate the estimated posterior probability of excluding at least one
+    true \divt{} value from analyses of simulated datasets where \divt{} for 22
+    pairs of populations is drawn from a series of uniform distributions,
+    $\divt{} \sim U(0, \divt{max})$. The proportion of simulation replicates in
+    which there is strong support for at least one true parameter value being
+    excluded from the model ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is
+    given for each \divt{max}.
+    Each plot represents 500 simulation replicates analyzed using $5\e6$
+    samples from the prior.
 }{figPowerProbExclusion}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_num_excluded.pdf}{
-    Histograms of the number of true \divt{} parameters excluded.
-}{figPowerNumExcluded}
 
@@
 \mFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
-    True versus estimated probability of a single divergence event.
+    An assessment of the approximate Bayesian model-averageing approach of
+    \citet{Hickerson2013} under the ideal conditions when the prior model
+    is correct (i.e., the pseudo-replicate datasets are simulated from
+    parameters drawn from the same prior distributions used in the
+    analysis).
+    The plots show the relationship between the estimated posterior and true
+    probability of (A \& C) $\numt{}=1$ and (B \& D) $\vmratio{} < 0.01$, based
+    on 50,000 simulations.
+    The results summarize the (A \& B) unadjusted and (C \& D) GLM-adjusted
+    posterior estimate from each simulation replicate.
+    The prior settings for all replicates included five prior models with
+    $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
+    U(0.0001, 0.05)$ for all five models, and
+    $M_1: \divt{} \sim U(0, 0.1)$,
+    $M_2: \divt{} \sim U(0, 1)$,
+    $M_3: \divt{} \sim U(0, 5)$,
+    $M_4: \divt{} \sim U(0, 10)$, and
+    $M_5: \divt{} \sim U(0, 20)$.
+    The number of samples from the prior was $2.5\e6$.
+    The simulated data structure was 8 population pairs, with a single 1000
+    bp locus sampled from 10 individuals from each population.  The 50,000
+    estimates of the posterior probability of one divergence event were
+    assigned to 20 bins of width 0.05.
+    The estimated posterior probability of each bin is plotted against the
+    proportion of replicates in that bin with a true value consistent with
+    one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
 }{figValidationMCBehvior}
@@
 \mFigure{../../model-choice/tau-prior/tau_prior.pdf}{
-    The prior distribution on divergence times imposed by the model-averaging prior.
+    The prior distribution on divergence times imposed by the model-averaging prior
+    comprised of five models with different uniform priors on \divt{}:
+    $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_2$ ($\divt{} \sim U(0, 1)$), $M_3$
+    ($\divt{} \sim U(0, 5)$), $M_4$ ($\divt{} \sim U(0, 10)$), $M_5$ ($\divt{}
+    \sim U(0, 20)$).
 }{figMCTauPrior}
@@
 \mFigure{../images/coin_flip.pdf}{
-    A plot of three beta density functions meant to represent a prior
-    (black), true posterior (blue), and empirical Bayes density (red).
+    A plot of three beta probability density functions that represent a prior
+    (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
+    empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
+    Bernoulli trials, four of which are successes.
 }{figCoinFlip}

### 2013-09-22 manuscripts/reply/reply.tex
More work on text of reply.

@@
 \msb, which estimates the temporal distribution of divergences among
-co-distributed pairs of taxa.
+co-distributed pairs of taxa \citep{Huang2011}.
 \citet{Oaks2012} published their findings that the method \msb can often be
-biased towards inferring models of temporally clustered divergence times
-among taxon pairs.
+biased towards inferring models of temporally clustered divergence times among
+taxon pairs.
 \citet{Hickerson2013} has published a response to this paper where they
 present a model-averaging approach that they conclude circumvents the poor
-behavior of the method revealved by \citet{Oaks2012}.
+behavior of the method revealed by \citet{Oaks2012}.
 Both of these papers are largely in agreement.
@@
 The main differences in their perspectives are centered around
-(1) the means by which broad uniform priors cause the poor behavior of \msb,
-and
+(1) the mechanism by which broad uniform priors cause the poor behavior of
+\msb, and
 (2) how to potentially ameliorate this issue.
 
-\citet{Oaks2012} conclude the primary mechanism by which broad priors cause the
-poor behavior of the method is likely the low marginal likelihoods of
-parameter-rich models integrated over vast parameter space with low probability
-of producing the data, yet relatively high prior density \citep[this is often
-referred to as Lindley's paradox;][]{Lindley1957}.
+Regarding the mechanism behind the poor behavior of the method,
+\citet{Oaks2012} suggest the primary cause is likely the low marginal
+likelihoods of parameter-rich models integrated over vast parameter space with
+low probability of producing the data, yet relatively high prior density
+\citep[this is often referred to as Lindley's paradox;][]{Lindley1957}.
 Note, this suggests the bias is extrinsic to \msb, and the numerical
@@
 \citet{Hickerson2013} take a more pessimistic view of the bias, and suggest it
-is intrinsic to \msb, i.e., the method's rejection algorithm is
-inefficient and will be increasingly biased as the overall space of the model
-increases, either as a function of the number of taxon pairs or the width of
-the uniform priors on nuisance parameters.
+is intrinsic to \msb.
+They conclude the method's rejection algorithm is inefficient and will be
+increasingly biased as the overall space of the model increases, either as a
+function of the number of taxon pairs or the width of the uniform priors on
+nuisance parameters.
 They support their position by giving a probabilistic argument that focuses on
@@
 explain the bias of the method found in several analyses of \citet{Oaks2012}.
-We reiterate several of the findings of \citet{Oaks2012} as well as present
-results of additional analyses, which strongly suggest that Lindley's paradox
-is playing a large role in the poor behavior of the method.
-
-\citet{Oaks2012} suggest that more cogent prior probability distributions on
-divergence models and nuisance parameters could mitigate the effect of
-Lindley's paradox.
+Furthermore, we reiterate several of the findings of \citet{Oaks2012} as well
+as present results of additional analyses, which strongly suggest that
+Lindley's paradox is playing a large role in the poor behavior of the method.
+
+As for how to mitigate the bias, \citet{Oaks2012} suggest that more cogent
+prior probability distributions on divergence models and nuisance parameters
+could reduce the effect of Lindley's paradox.
 They also layout a set of simulation-based procedures for determining power,
@@
 approach.
-Our results demonstrate that the method is biased and dangerous.
-The approach merely provides another means by which the model can ``escape''
-large parameter space, and the bias towards models with smaller space still
-remains.
-This bias can be dangerous and often excludes the true parameter space due to
-the use of uniform priors.
+Our results demonstrate that the method is biased and dangerous.  The approach
+merely provides another means by which the model can ``escape'' large parameter
+space, and bias towards smaller models remains.
+This manifestation of the bias can be dangerous and often excludes the true
+parameter space due to the use of uniform priors.
 Furthermore we discuss the potential theoretical and practical problems of
-empirical Bayesian model choice.
+empirical Bayesian model choice, which is encouraged by \citet{Hickerson2013}
+without any mention of its potential complications.
 
@@
 In other words, some of their prior and posterior samples are in units of
-$0.05/\mutationRate$ generations, whereas other are in units of
+$0.05/\mutationRate$ generations, whereas others are in units of
 $0.02/\mutationRate$ generations.
@@
 that all of the values of the response variable are in the same units.
-Thus, the results in section ``Improved Sampling Efficiency by Prior Weighting
-Supports Asynchronous and Recent Divergence for the Philippines Vertebrate
-Data'' and presented in Figure 2 of \citet{Hickerson2013} should be
-disregarded.
+Thus, the results in sections ``Using ABC Model Comparison to Weight
+Alternative Priors for the Philippine Vertebrate Data'' and ``Improved Sampling
+Efficiency by Prior Weighting Supports Asynchronous and Recent Divergence for
+the Philippines Vertebrate Data'' and presented in Figure 2 of
+\citet{Hickerson2013} should be disregarded.
 The error is easily illustrated by re-plotting their results with the different
@@
 prior'').
-They detail how the being limited to uniform prior distributions necessitates
-awkward broad priors to avoid excluding the truth, a priori.
+They detail how being limited to uniform prior distributions in \msb
+necessitates awkward broad priors to avoid excluding the truth, a priori.
 What \citet{Hickerson2013} really mean by ``poorly selected'' is that
@@
 both the prior sensitivity of the method and determining the power of \msb
-under the ``best'' possible real-world conditions.
+under real-world conditions with highly informative priors.
 \citet{Oaks2012} discuss the potential dangers of taking an empirical Bayesian
-model choice (see last paragraph of ``Assessing prior sensitivity of \msb''),
-and we expand on this here by exploring the theoretical and practical
-implications of using an empirical Bayes approach to model choice.
+approach to model choice (see last paragraph of ``Assessing prior sensitivity
+of \msb''), and we expand on this here by exploring both the theoretical and
+practical implications of using empirically informed priors for Bayesian
+model choice.
 
@@
 However, this all hinges on the fact that the prior ($p(\myTheta{})$) is
-defined for all possible parameters independently of the new data being
+defined for all possible parameter values independently of the new data being
 analyzed.
@@
 beliefs about $p(\myTheta{})$.
-However, if the data are used to both inform the prior and calculate the
+However, if the same data are used to both inform the prior and calculate the
 posterior, the prior becomes conditional on the data, and Bayes' rule breaks
@@
 
-Now, this is not to say that empirical Bayesian approaches not useful.
+This is not to say that empirical Bayesian approaches not useful.
 Quite the contrary.
@@
 and misshapen;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
-Nevertheless, empirical Bayes methods are often a powerful means of parameter
-estimation that display favorable frequentist properties.
+Nevertheless, empirical Bayes methods can provide a powerful means of parameter
+estimation.
 Furthermore, many post-hoc correction methods have been developed for
 estimating confidence-intervals from empirical Bayes estimates of posterior
-distributions that often display good frequentist coverage behavior
+distributions that often exhibit good frequentist properties
 \citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.
@@
 estimation, a theoretical justification for empirical Bayes approaches to model
-choice is much more dubious.
-In Bayesian model choice, the goal is usually not to estimate parameters, but
-to estimate the relative probabilities of candidate models.
+choice is much more uncertain.
+In Bayesian model choice, the goal is not to estimate parameters, but to
+estimate the relative probabilities of candidate models.
 Unlike parameter estimates, posterior probabilities represent a summary of the
@@
 interested in the hypothesis that the mass of George Washington's periwig
-renders the United States' quarter dollar unfair.
-That is to say that our null hypothesis is that the probability of a US quarter
-landing heads when tossed is less than 0.5 ($\theta < 0.5$).
+renders the quarter dollar of the United States unfair.
+That is to say that their null hypothesis is that the probability of a US
+quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
 They can certainly evaluate this hypothesis, they can have undergraduate
@@
 If Joe's flips were a representative sample, Mary and Will's empirical Bayes
-estimate might very well be a better estimator for the parameter \myTheta{}.
+estimate might very well be a better point estimate for the parameter
+\myTheta{}.
 However, their empirical Bayes estimate of the probability of their hypothesis
@@
 choice}
-Given that \msb can be biased toward models with smaller parameter space, and
-Hickerson et al.'s \citeyear{Hickerson2013} use of narrow, empirically informed
-uniform priors on divergence times, we sought to determine whether such an
-approach will often exclude the truth (i.e., the model-averaged posterior will
-be dominated by models that excude the true values of parameters of the model).
-We explored this possiblity in two ways.
-First, we reanalyzed the Philippines dataset using the model-averaging approach
+We were alarmed when we saw the results of the model-averaging results
+presented in \citet{Hickerson2013}.
+Their results strongly favored the models with the narrowest, empirically
+informed prior on divergence time, and thus their model-averaged posterior
+estimates are dominated by models $M_1$ and $M_2$ (see Table 1 of
+\citet{Hickerson2013}).
+This concerned us because the narrowest prior of \citet{Hickerson2013}
+($\divt{} \sim U(0,0.1)$) likely excludes the true divergence times for at
+least some of the Philippine taxa.
+They set this prior to match the 95\% highest posterior density (HPD) interval for
+the mean divergence time estimated under one of the priors used by
+\citet{Oaks2012} (see Table 2 and 3 of \citet{Oaks2012}).
+Given this interval estimate is for the \emph{mean} divergence time across all
+22 taxa, it seems odd to set this as the prior, because many of the taxon pairs
+are expected to have diverged at times older than the upper limit.
+Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
+of the mean divergence time under the other two priors explored by
+\citet{Oaks2012} (under these priors the 95\% HPD is approximately 0.3--0.6;
+see Table~6 of \citet{Oaks2012}).
+Thus, when we saw the results of \citet{Hickerson2013} we suspected that
+Lindley's paradox was again biasing the method toward models with smaller
+parameter space, and as a consequence was biasing the method toward models
+that exclude the truth (i.e., the model-averaged posterior will
+be dominated by models that exclude the true values of parameters of the model).
+
+We explored the propensity of the model-averaging approach of
+\citet{Hickerson2013} to exclude the truth in two ways.
+First, we re-analyzed the Philippines dataset using the model-averaging approach
 of \citet{Hickerson2013}, but set one of prior models with a uniform prior on
 divergence times that is unrealistically narrow, and almost certainly excludes
-the truth.
+most, if not all, of the true divergence times of the 22 taxon pairs.
 If Lindley's paradox causes the method to prefer models with less parameter
 space, we expect \msb will preferentially sample from this incorrect model
-yielding a marginal posterior that is invalid (i.e., the model-averaged
+yielding a marginal posterior that is incorrect (i.e., the model-averaged
 posterior will be dominated by an incorrect model that excludes the truth).
@@
 
-The empirically guided, narrowest prior of \citet{Hickerson2013} ($\divt{} \sim
-U(0,0.1)$) likely excludes the true values for at least some of the Philippine
-taxa.
-They set this prior to match the 95\% highest posterior density (HPD) interval for
-the mean divergence time estimated under one of the priors used by
-\citet{Oaks2012} (see Table 2 and 3 of \citet{Oaks2012}).
-Given this interval estimate is for the \emph{mean} divergence time across all
-22 taxa, it seems odd to set this as the prior, as many of the taxon pairs
-likely diverged at times older than the upper limit.
-Furthermore, the prior is \emph{excluded} from the 95\% HPD interval estimates
-of the mean divergence time under the other two prior analyses (under these
-priors the 95\% HPD is approximately 0.3--0.6; see Table~6 of
-\citet{Oaks2012}).
-
 For our re-analysis of the Philippines dataset we used the model-averaging
 approach of \citet{Hickerson2013}, but with different prior models to avoid
-mixing time units. We used five priors, all of which had priors on population
-sizes of $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and
-$\ancestralTheta{} \sim U(0.0001, 0.05)$.
+their error of mixing time units.
+We used five priors, all of which had priors on population sizes of
+$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
+U(0.0001, 0.05)$.
 Following \citet{Hickerson2013}, each of these priors had the following
@@
 using the prior means and standard deviations of the given model.
-We then retained the 10,000 samples with the smallest Euclidean distance from
-the observed summary statistics, this time standardizing the statistics using
-the prior means and standard deviations across all five models.
+From the remaining 50,000 samples, we then retained the 10,000 samples with the
+smallest Euclidean distance from the observed summary statistics, this time
+standardizing the statistics using the prior means and standard deviations
+across all five models.
 We then repeated this analysis twice, replacing the $M_1$ with
@@
 While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
-\citet{Hickerson2013} likely excludes the true divergence times of
-at least some of the 22 taxa, we are nearly certain that these
-narrower priors are incorrect and exclude the truth.
+\citet{Hickerson2013} likely excludes the true divergence times of at least
+some of the 22 taxa, we are nearly certain that these narrower priors are
+incorrect and exclude most, if not all, of the divergence times of the
+Philippine taxa.
 
@@
 does not reduce the methods bias towards models with smaller parameter space,
-but rather allow it to manifest in a very dangerous way.
+but rather allows it to manifest in a very dangerous way.
 The method strongly prefers the prior model with the narrowest distribution on
@@
 model-averaged results are dominated by models that exclude at least some of
-the 22 true divergence times, making their results difficult to interpret.
+the 22 true divergence times, making their results difficult to interpret
+(in addition to the error of time units).
 
@@
 with a mean of 0.5 ($\divt{} \sim Exp(2)$). All other parameters were
-identically distributed as in the $M_1$--$M_5$ models.
+identically distributed as the $M_1$--$M_5$ models.
 We then repeated the analysis described above, using $5\e6$ prior samples from
-$M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, and retaining 1,000 posterior samples
-for each of the 1,000 pseudo-observed datasets.
+$M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, and retaining 1000 posterior samples
+for each of the 1000 pseudo-observed datasets.
 
@@
 excluding the truth.
-With that criterion, 66\% of the replicats (87\% with GLM-regression adjustment)
+With this criterion, 66\% of the replicates (87\% with GLM-regression adjustment)
 strongly support the exclusion of true values (Figure~\ref{figExclusionSimProb}).
@@
 The results of our empirical and simulation-based analyses clearly demonstrate
-the danger of using narrow, empirically guided priors in a Bayesian
-model-averaging framework \ldots
-
-\subsection*{Final thought on EB}
+the danger of using narrow, empirically guided uniform priors in a Bayesian
+model-averaging framework.
+This is especially dangerous \msb, which has already been demonstrated to be
+biased away from models with more parameter space.
+The consequence of this approach is a high risk of obtaining a model-averaged
+posterior estimate that is heavily weighted toward incorrect models that
+exclude true values of model parameters.
+
+\subsection*{Final thoughts on empirical priors in Bayesian model choice}
 Given all of the theoretical and practical issues with empirical Bayes
@@
 quite clear why one should use caution before overly criticizing an
-investigators choice of priors after having seen the posterior.
+investigator's choice of priors after having seen the resulting posterior.
 In a strictly Bayesian world, this is cheating.
@@
 putting a lot of prior density in regions of improbable parameter space.
-The alternative, as shown above, is to risk excluding the truth before even
-conducting the analysis.
-
-\section*{Our power analysis}
+The alternative, as shown above, is to risk excluding the truth before
+performing the analysis.
+
+\section*{Assessing the power of the model-averaging approach of
+    \citet{Hickerson2013}}
 \citet{Hickerson2013} did not conduct a power analysis to accompany their
@@
 \citet{Oaks2012}).
-Thus, we perform perform simulation-based power analyses to explore the
-power of the approximate Bayesian model-averaging approach championed
-by \citet{Hickerson2013}.
+Thus, we perform simulation-based power analyses to explore the power of the
+approximate Bayesian model-averaging approach championed by
+\citet{Hickerson2013}.
 
@@
 For each pseudo-replicate dataset, we retained the 1000 samples from each prior
-model with the smallest Euclidean distance from the replicates' summary
+model with the smallest Euclidean distance from the pseudo-observed summary
 statistics, standardizing the statistics using the prior means and standard
@@
 In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
-posterior samples from each of them.
-
-We find that the approach of \citet{Hickerson2013} to estimate the dispersion
-index of divergence times (\vmratio{}) across most of the \divt{max} we
-simulated, whether evaluating the unadjusted
+posterior samples for each of them.
+
+We find that the approach of \citet{Hickerson2013} struggles to estimate the
+dispersion index of divergence times (\vmratio{}) across most of the \divt{max}
+we simulated, whether evaluating the unadjusted
 (Figure~\ref{figPowerAccOmegaMedian}) or GLM-adjusted
@@
 It is very interesting to note that as \divt{max} increases, but before the
-mode inference is finally pulled away from $\numt{} = 1$, the distribution of
-\numt{} estimates closely mirrors the odd U-shaped prior on divergence models
+estimates are finally pulled away from $\numt{} = 1$, the distribution of
+\numt{} estimates closely mirror the odd U-shaped prior on divergence models
 used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
@@
 Looking at our simulation results in terms of the posterior probability of the
-divergence-time dispersion index supporting the extreme case of one divergence
-event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the method
-strongly supports one divergence in greater than 27\%
-of the replicates across all the \divt{max} we simulated
-(Figure~\ref{figPowerOmegaProb}).
-Following \citet{Hickerson2013}, We use a Bayes factor of greater than 10 as
+dispersion index of divergence times supporting the extreme case of one
+divergence event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the
+method strongly supports one divergence in greater than 27\% of the replicates
+across all the \divt{max} we simulated (Figure~\ref{figPowerOmegaProb}).
+Following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
 the criterion for incorrect inference of a single divergence event.
@@
 differ by much less than \ldots millions of generations,'' our results show
-that method is biased toward strongly supporting the extreme case of a single
+that method is biased toward supporting the extreme case of a single
 divergence event when populations diverged random over the last
 $8\globalpopsize$ generations.
-To roughly put that on the scale for a vertebrate mitochondrial locus, assuming
-a mutation rate of $2\e{-8}$ per site per generation, this translates to 5
-million generations.
+To put this on the scale roughly consistent with a vertebrate mitochondrial
+locus, assuming a mutation rate of $2\e{-8}$ per site per generation, this
+translates to 5 million generations.
 Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
@@
 
-The results of our power analysis further demonstrate the
-propensity of Hickerson et al.'s \citeyear{Hickerson2013} approach
-to exclude true parameter values.
+The results of our power analysis further demonstrate the propensity of
+Hickerson et al.'s \citeyear{Hickerson2013} approach to exclude true parameter
+values.
 Across all but one of the \divt{max} we simulated, the method excludes the
@@
 will exclude a large proportion of the true divergence time values
-(Figure~\ref{figPowerNumExcluded}.
+(Figure~\ref{figPowerNumExcluded}).
 The posterior probability of excluding at least on true divergence value is
@@
 (Figure~\ref{figPowerNumExcluded}E).
-
-Given our results, we conclude that not only does is the approach of
+Again, this demonstrates the method's lack of robustness.
+
+Given our results, we conclude that not only is the approach of
 \citet{Hickerson2013} dangerous in terms of excluding regions of parameter
@@
 
-\section*{Their power analysis}
-\citet{Hickerson2013} present a power analysis in which they demonstrate that
+\section*{The power analysis of \citet{Hickerson2013}}
+\citet{Hickerson2013} present a power analysis in which they find that
 \msb has the power to infer multiple divergence events when divergence times
-are random over hundreds of thousands of generations (rather than millions
-as demonstrated by \citet{Oaks2012}.
-It is not surprising to find that within certain parameter space the method
-has increased power to detect temporal variation in divergences over narrower
-time windows than millions of generations.
+are random over hundreds of thousands of generations (rather than millions as
+demonstrated by \citet{Oaks2012}.
+It is not surprising that under limited parameter space the method has
+increased power to detect temporal variation in divergences over narrower time
+windows than millions of generations.
 However, what is important to consider is whether those conditions are relevant
@@
 \citet{Oaks2012} explore the behavior of the method under three divergence time
-priors as narrow as 0--5 coalescent units, which is quite narrow considering
+priors, as narrow as 0--5 coalescent units. This is quite narrow considering
 that this expresses the prior belief that all 22 taxon pairs diverged within
@@
 coalescent unit.
-It seems that being 100\% certain a priori that all taxa under comparison
-diverged within the last coalescent unit is not generally applicable to most
-empirical systems.
+This prior setting assumes that there is enough a priori information to
+be 100\% certain that all taxa (18 for their simulations) diverged within
+the last $\globalpopsize$ generations.
+This does not seem applicable to most empirical systems.
 Furthermore, even when such extensive prior information is available to only be
 able to detect multiple divergences on the scale of hundreds of thousands of
-generations does not seem very ``powerful.''
+generations does not seem very powerful.
 
@@
 single divergence event.
-As discussed in \citet{Oaks2012}, it seems dubious to consider the estimation
+As discussed in \citet{Oaks2012}, it seems odd to consider the estimation
 of two divergence events as a success when divergence is random over hundreds
-of thousands of generations.
-Unlike the results of \citet{Oaks2012} it is impossible to assess the
-performance of \msb based on the results presented by \cite{Hickerson2013}
-other than to know whether or not it inferred the worst possible scenario of a
-single event.
-In an empirical system such as the Philippines, where island fragmentation
-has occurred at least X times over the past several millions of years, an
-estimate where 22 taxa share even a handful of divergence events would be
-of biogeographic interest.
-\citet{Hickerson2013} seem to suggest in their response that \msb should only
-be used as a binary estimator; i.e., was there one event or more?
-This is not consistent with how \msb has been presented, used, and interpreted
-in the past.
-
-\citet{Hickerson2013} translate their results from units of
+of thousands (or millions) of generations.
+In an empirical system such as the Philippines, where island fragmentation has
+occurred at least \highLight{XX} times over the past several millions of years,
+an estimate where 22 taxa share even a handful of divergence events would be of
+biogeographic interest.
+
+Lastly, \citet{Hickerson2013} translate their results from units of
 \globalcoalunit generations to generations assuming a mutation rate of
 $1.92\e{-8}$ mutations per site per generation.
-If we tranlate there results to a scale more consistent with rates of mutation
+If we translate there results to a scale more consistent with rates of mutation
 of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
@@
 Again, considering \citet{Hickerson2013} are assuming prior knowledge that
-all population pairs diverged within the past coalescent unit, this
+all population pairs diverged within the last coalescent unit, this
 does not seem particularly powerful.
@@
 \section{Validation analyses}
-
-\section*{Lindley's paradox versus sampling error}
+\highLight{Not sure where to put this}.
+The main contribution of this analysis is to show that the unadjusted posterior
+is a better model-choice estimator than the regression-corrected estimates
+(Figure~\ref{figValidationMCBehavior}).
+This justifies our presenting the unadjusted results in this paper.
+\highLight{Is it worth including just for this purpose?}
+
+\section*{Lindley's paradox versus insufficient prior sampling}
 One of the points of disagreement between \citet{Hickerson2013} and
@@
 with clustered divergences (i.e., models with few divergence events).
-\citet{Oaks2012} present simulations based analyses that suggest the broad
+\citet{Oaks2012} present simulation-based analyses that suggest the broad
 uniform prior on divergence time parameters hinder the marginal likelihoods of
@@
 They argue the widest of the three priors used by \citet{Oaks2012} was very
-unlikely to produce any prior samples with high numbers of divergence times
+unlikely to produce any prior samples with large numbers of divergence times
 that were consistent with the Philippine data.
@@
 $2\e{-8}$ for all taxa, and are, of course, subject to estimation error.
-Furthermore, the units of the gene trees are in millions of year, whereas the
-divergence time prior of \msb is in generations, thus the logic of
-\citet{Hickerson2013} requires the additional assumption that all 22 Philippine
-taxa have a generation time of one year.
+Furthermore, the branch-length units of the gene trees are in millions of
+years, whereas the divergence time prior of \msb is in generations, thus the
+logic of \citet{Hickerson2013} requires the additional assumption that all 22
+Philippine taxa have a generation time of one year.
 Thus, the argument of \citet{Hickerson2013} that divergence time estimates of
 \citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
-dangerous, especially given our findings presented above.
-
-Finally, even \emph{if} we make all of these assumptions and assume the
-gene divergence times are without error, the probabilistic argument only
+dangerous, especially given our findings on the behavior of empirically
+informed priors presented above.
+
+Finally, even \emph{if} we make all of these assumptions and assume the gene
+divergence times are estimated without error, the probabilistic argument only
 applies to one of the three different priors used by \citet{Oaks2012}.
@@
 \citeyear{Hickerson2013} probability equations demonstrates that the prior
-is densely populated with samples with large number of divergence parameters
-that are consistent with the gene divergence estimates
+is densely populated with samples with large numbers of divergence parameters
+that are consistent with the gene divergence estimates.
 Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
@@
 However, the magnitude of the bias is very similar across all three priors
-\citet{Oaks2012}.
+explored by \citet{Oaks2012}.
 \citet{Hickerson2013} point out a case where the narrow prior performs
@@
 the different predictions made by these two phenomena.
-One examples, as discussed by \citet{Oaks2012} is that insufficient
-prior sampling should create higher variance in posterior estimates,
-and thus the analyses should be sensitive to the number of samples
-drawn from the prior.
+One example, as discussed by \citet{Oaks2012}, is that insufficient prior
+sampling should create higher variance in posterior estimates, and thus it
+should cause analyses to be sensitive to the number of samples drawn from the
+prior.
 \citet{Oaks2012} do not see such sensitivity when they compare prior sample
@@
 divergence times (\vmratio{}) as the prior samples accumulate
-(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or GLM
-regression adjusted estimates.
-This strongly suggests that insufficient prior sampling did not play a large
+(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or
+GLM-regression-adjusted estimates.
+This suggests that insufficient prior sampling did not play a large
 role in the bias found by \citet{Oaks2012}.
@@
 Thus, it is difficult to explain the method's propensity towards the model with
-the smallest parameter space with insufficient sampling.
+the smallest parameter space via insufficient prior sampling.
 While analyses that sample each model proportional to their relative parameter
-space could be explored, it seems very likely that the broad uniform priors
-are simply inhibiting the marginal likelihoods of spacious models.
+space could be explored, it seems much more likely that the broad uniform
+priors are simply inhibiting the marginal likelihoods of spacious models.
 
@@
 sampled than those with fewer divergence events \citep{Oaks2012}.
+Thus, the results of the simulations of \citet{Oaks2012} are more consistent
+with Lindley's paradox causing the bias toward models with fewer divergence
+time parameters.
+
+Taken together, the results presented here and in \citet{Oaks2012} support
+Lindley's paradox as the primary mechanism by with broad uniform priors
+cause biases of \msb.
 
 \section*{Differing utilities of \numt{} and \vmratio{} in \msb}
-We preface this section with a clarification of an issue that has
-been confused throughout the \msb literature.
-The mean (\meant{}{}), variance (\vart{}{}), and dispersion index
-($\vart{}{}/\meant{}{} = \vmratio{}$) of divergence times are \emph{not} parameters
-of the model implemented in \msb.
-Rather they are statistics that summarize parameters of the model.
-This is in contrast to \numt{}, which is a parameter of the model.
-
-\citet{Hickerson2013} strongly argue that \vmratio{} is a better estimator than
-\numt{}, but the logic behind the argument is dubious.
-They present a plot of \numt{} against \vmratio{} (Fig.~S1) which is simply a
-plot of sample size versus variance.
-After showing that \vmratio{} has essentially no information about the
-number of divergences, they conclude it is more informative and biogeographically
-relevant.
+The primary parameter of the \msb model is the vector of divergence
+times for each of the taxon pairs,
+$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
+\citep{Oaks2012}.
+\citet{Hickerson2013} argue that the dispersion index of this vector,
+\vmratio{}, is a better model-choice estimator than the vector's cardinality,
+\numt{}.
+They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
+\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
+This plot shows, not surprisingly, that \vmratio{} has essentially no
+information about the number of divergences among taxa.
+Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
+biogeographically relevant than \numt{}.
 We struggle to follow this logic.
@@
 The temporal information contained in this vector coupled with its cardinality
-(\numt{}) is certainly more informative than its summary.
-The dispersion index is not a sufficient statistic for this vector.
-
-They also argue that ``\msb can estimate \vmratio{} much better than \numt{}.''
-However, the primary objective of \msb is to estimate the posterior probability
-of divergence models.
+(\numt{}) is certainly more informative than its variance (i.e., the dispersion
+index is not a sufficient statistic for \divtvector).
+
+\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
+better than \numt{}.''
+However, \msb is a model-choice estimate, and hence the goal is to estimate the
+posterior probability of divergence models.
 \citet{Oaks2012} demonstrate that even when all assumptions of the model are
-met, \vmratio is a terrible model-choice estimator (see plots B, D \& F of
-Figure 4), whereas \numt{} performs better.
+met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
+4), whereas \numt{} performs better.
 Furthermore, \vmratio{} is limited to estimating the probability of only a
@@
 model choice is extremely limited.
+
 The model-choice utility of \vmratio{} is limited to the probability that
@@
 However, it is still not surprising to see that it is numerically difficult to
-obtain reliable estimate of the probability that the continuous \vmratio{}
+obtain reliable estimates of the probability that the continuous \vmratio{}
 statistic is ``near'' its limit of zero.
@@
 
-\section*{General thoughts}
+\section*{Some general thoughts on the model of \msb}
 It is not too surprising that the model-averaging approach proposed by
@@
 Following the model description and notation of \citet{Oaks2012}, we
-will tally up all of the free parameters in the \msb model.
-
+take the time to tally up all of the free parameters in the \msb model.
 Under the simplest model in \msb (i.e., assuming no migration and
@@
 Robustness is an extremely important characteristic of a method to gauge its
-applicability to empirical data, because we know the model and priors will be
+applicability to real-world data, because we know the model and priors will be
 wrong to some degree.
@@
 from the data, likely plays a major role in the poor behavior of this approach.
-Certainly, such a large modification of the model demands simulation-based
-assessments of the behavior of the new model, as we provide here.
+% Certainly, such a large modification of the model demands simulation-based
+% assessments of the behavior of the new model, as we provide here.
 
@@
 In general, Bayesian model-averaging is a powerful approach that leverages a
-great strengths of Bayesian statistical procedures, namely the ability to
+great strength of Bayesian statistical procedures, namely the ability to
 obtain marginalized estimates that incorporate uncertainty in nuisance
@@
 but avoid the need for imposing additional model choice.
-\citet{Oaks2012} suggest that uniform priors are inappropriate for parameters,
-and recommend the use of more cogent probability distributions from the
-exponential family.
+\citet{Oaks2012} suggest that uniform priors are inappropriate for many
+parameters of the \msb model, and recommend the use of more cogent probability
+distributions from the exponential family.
 If we look at the prior distribution on the divergence time parameter \divt{}
@@
 This looks very much like an exponential distribution.
-Thus would be much more natural and easier to simply place a gamma prior (the
-exponential being a special case of the gamma) on divergence times.
+Thus, it would be much more natural and easier to simply place a gamma prior
+(the exponential being a special case of the gamma) on divergence times.
 This would allow the investigator to much better capture their prior
@@
 distributions, and without the need of costly model-averaging.
-This seems like a much more natural solution to the problems of Lindley's
+This seems like a much more natural solution to the problem of Lindley's
 paradox and sampling error, which are both certainly exacerbated by the fact
@@
 The investigator can place the majority of the prior density in regions
-of parameters space the believe, a priori, are most plausible, but still
+of parameters space they believe, a priori, are most plausible, but still
 capture uncertainty in the tails of distributions with low density.
 As discussed in \citet{Oaks2012} the use of inflexible and unforgiving uniform
-priors necessitated the need to use broad priors to avoid excluding the truth a
+priors necessitate the need to use broad priors to avoid excluding the truth a
 priori.
 Thus, to avoid the behavior of the method under these priors, cheating
-and using the data was a very tempting alternative.
+and using the data is a very tempting alternative.
 
@@
 statistics used by \msb (Figure~\ref{figSaturationPlot}).
+
+\section*{Conclusions}
 
@@
     one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
-}{figValidationMCBehvior}
+}{figValidationMCBehavior}
 

### 2013-09-22 manuscripts/reply/reply.tex
Initial draft of reply complete.
Sending this version to co-authors.

@@
         \vspace{12pt}
-        Jamie R. Oaks$^{1,2,5}$, Jeet Sukumaran$^{1,2}$, Jacob A.
-        Esselstyn$^{3}$, Charles W. Linkem$^{1,2}$, Cameron D.
-        Siler$^{4}$, Mark T. Holder$^{2}$ and Rafe M. Brown$^{1,2}$
+        Jamie R.\ Oaks$^{1,2}$, Jeet Sukumaran$^{3}$, Jacob A.\
+        Esselstyn$^{4}$, Charles W.\ Linkem$^{5}$, Cameron D.\
+        Siler$^{6}$, Mark T.\ Holder$^{1}$ and Rafe M.\ Brown$^{1}$
 
         \bigskip
-        $^1$\emph{Biodiversity Institute\\
-            $^2$Department of Ecology and Evolutionary Biology\\ 
-            University of Kansas\\
-            %1345 Jayhawk Blvd\\
-            Lawrence, KS 66045\\
-            USA}\\[.1in]
-        $^3$\emph{Biology Department\\
-            McMaster University\\
-            % Life Sciences Building Rm 328\\
-            Hamilton, Ontario L8S4K1\\
-            Canada}\\[.1in]
-        $^4$\emph{Department of Biology\\
-            University of South Dakota\\
-            Vermillion, SD 57069\\
-            USA}\\[.1in]
-        $^5$\emph{Corresponding author} (\href{mailto:joaks1@ku.edu}{\tt
-        joaks1@ku.edu})\\
+        $^1$\emph{Department of Ecology and Evolutionary Biology,
+            Biodiversity Institute,
+            University of Kansas,
+            Lawrence, Kansas 66045}\\[.1in]
+        $^3$\emph{Department of Biology,
+            Duke University,
+            Durham, North Carolina 27708} \\[.1in]
+        $^4$\emph{Museum of Natural Science,
+            Louisiana State University,
+            119 Foster Hall,
+            Baton Rouge, Louisiana 70803}\\[.1in]
+        $^5$\emph{Department of Biology,
+            University of Washington,
+            Seattle, Washington 98195}\\[.1in]
+        $^6$\emph{Sam Noble Museum,
+            Department of Biology,
+            University of Oklahoma,
+            Norman, Oklahoma 73072}\\[.1in]
+        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
+        joaks1@gmail.com})\\
 
@@
     \noindent\textbf{ABSTRACT} \\
-    \noindent Abstract here \ldots
+    \noindent Recently, \citet{Oaks2012} presented a simulation-based
+    assessment of the approximate Bayesian phylogeographical model choice
+    method implemented in \msb, and found the method to lack robustness and
+    power to detect temporal variation in divergences that are randomly
+    distributed over relatively broad periods of evolutionary history.
+    They conclude that a likely mechanism behind the bias toward models with
+    few divergence events is the small marginal likelihoods of models with more
+    divergence time parameters due to the integration over vast parameter space
+    imposed by broad uniform priors characterized by low likelihood and high
+    prior density.
+    This phenomenon is often referred to as Lindley's paradox
+    \citep{Lindley1957}.  \citet{Hickerson2013} have responded to this work,
+    criticising the priors used by \citet{Oaks2012} and presenting an
+    approximate Bayesian model-averaging approach that uses narrow, empirically
+    informed uniform priors.
+    \citet{Hickerson2013} conclude this approach circumvents the issues
+    revealed by \citet{Oaks2012} and yields robust inference of divergence
+    models.
+    Furthermore, they argue that the biases revealed by \citet{Oaks2012} are
+    inherent to the inefficient rejection algorithm implemented in \msb, rather
+    than a result of Lindley's paradox.
+    Here we show that their methodology suffered from a fundamental error that
+    renders some of their results invalid and the remainder difficult to
+    interpret.
+    Furthermore, we demonstrate how their approach still suffers from a bias
+    toward models with lesser parameter space, which manifests in a dangerously
+    strong tendency to sample predominantly from models that exclude the true
+    values of the model's parameters.
+    Our results suggest that their model-averaged results are dominated by
+    models that likely exclude at least some of the true divergence times
+    across the 22 pairs of Philippine vertebrates, which adds to the
+    difficulty in interpreting their results.
+    We also show that in addition to this dangerous behavior, the approach of
+    \citet{Hickerson2013} is still strongly biased toward inferring
+    simultaneous divergences even when the true divergences are randomly
+    distributed over millions of generations.
+    Our results, coupled with those of \citet{Oaks2012}, strongly suggest that
+    Lindley's paradox is a primary cause of the bias of \msb toward models with
+    lesser parameter space.
+    We discuss ways to mitigate this bias without resorting to adding an
+    additional dimension of model choice to the model or using dangerous
+    empirically informed uniform priors.
 
     \vspace{12pt}
-    \noindent\textbf{KEY WORDS: } 
+    \noindent\textbf{KEY WORDS: Approximate Bayesian computation; Bayesian
+        model choice; empirical Bayes}
+
 }
@@
 \newpage
-\noindent Recently, this journal has served as a venue for discussion of the
-potential pitfalls of approximate Bayesian methods of comparative
-phylogeographical model choice.
-The discussion is centered around the approximate Bayesian model choice method,
+\noindent Recently, this journal has served as a medium for an interesting
+discussion of the potential pitfalls of approximate Bayesian methods of
+comparative phylogeographical model choice.
+The discussion has centered around the approximate Bayesian model choice method,
 \msb, which estimates the temporal distribution of divergences among
-co-distributed pairs of taxa \citep{Huang2011}.
-\citet{Oaks2012} published their findings that the method \msb can often be
-biased towards inferring models of temporally clustered divergence times among
-taxon pairs.
-\citet{Hickerson2013} has published a response to this paper where they
-present a model-averaging approach that they conclude circumvents the poor
-behavior of the method revealed by \citet{Oaks2012}.
+co-distributed pairs of taxa \citep{Huang2011} by sampling over all
+unordered models of divergence.
+\citet{Oaks2012} published results of simulation-based power analyses that
+revealed that \msb can often be biased towards inferring models of temporally
+clustered divergence times among taxon pairs.
+\citet{Hickerson2013} has published a response to this paper where they present
+an empirical Bayes model-averaging approach that they conclude circumvents the
+poor behavior of the method revealed by \citet{Oaks2012}.
 Both of these papers are largely in agreement.
@@
 only one of the three prior models used by \citet{Oaks2012}.
-We show that this argument is based on dubious assumptions, and does not
-explain the bias of the method found in several analyses of \citet{Oaks2012}.
+We show that this argument is based on an empirical Bayesian perspective
+towards model choice and questionable assumptions, and does not explain the
+bias of the method found in many of the analyses of \citet{Oaks2012}.
 Furthermore, we reiterate several of the findings of \citet{Oaks2012} as well
@@
 accuracy, and robustness of the method given a dataset, and recommend any
-application of \msb should be accompanied by such procedures, especially
-if any biological conclusions are going to drawn from the results.
+application of \msb should be accompanied by such procedures, especially if any
+biological conclusions are drawn from the results.
 \citet{Hickerson2013} present an approximate Bayesian model-averaging approach
@@
 Furthermore, we follow the advice of \citet{Oaks2012} and present
-simulation-based assessments of Hickerson et al.'s \citeyear{Hickerson2013}
+simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
 approach.
@@
 merely provides another means by which the model can ``escape'' large parameter
-space, and bias towards smaller models remains.
+space, and bias toward smaller models remains.
 This manifestation of the bias can be dangerous and often excludes the true
-parameter space due to the use of uniform priors.
+parameter space due to the use of narrow uniform priors.
 Furthermore we discuss the potential theoretical and practical problems of
-empirical Bayesian model choice, which is encouraged by \citet{Hickerson2013}
-without any mention of its potential complications.
-
-\section*{Hickerson et al.'s re-analysis of the Philippines data}
-\citet{Hickerson2013} re-analyze the dataset of \citet{Oaks2012} using a
-model-averaging approach, where they place a discrete uniform prior over eight
+empirical Bayesian model choice, which is used and encouraged by
+\citet{Hickerson2013} without any mention of its potential complications.
+
+\section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
+\citet{Hickerson2013} re-analyzed the dataset of \citet{Oaks2012} using a
+model-averaging approach, where they placed a discrete uniform prior over eight
 different prior models (see Table 1 of \citet{Hickerson2013}).
@@
 Each of the eight prior models used in the re-analysis by \citet{Hickerson2013}
-has one of two priors on the mean population size of the descendant populations
-of each taxon pair:
+has one of two priors on the mean size of the descendant populations of each
+taxon pair:
 $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ or
-$\meanDescendantTheta{} \sim U(0.0005, 0.04)$ or
+$\meanDescendantTheta{} \sim U(0.0005, 0.04)$.
 As described in \citet{Oaks2012}, the divergence time parameters of the model
@@
 size, \myTheta{C}.
-This reference population size is defined in terms of the upper limit on the
-uniform prior on the mean population size of the descendant populations,
+This reference population size is defined in terms of the upper limit of the
+uniform prior on the mean size of the descendant populations,
 \meanDescendantTheta{}, such that for the prior $\meanDescendantTheta{} \sim
@@
 The error is easily illustrated by re-plotting their results with the different
-time units indicated (Figure~\ref{figJointPosteriorHickerson}.
+time units indicated (Figure~\ref{figJointPosteriorHickerson}).
+Another important characteristic of the approach used by \citet{Hickerson2013}
+that makes their results difficult to interpret in addition to this error is
+their use of narrow, empirically guided uniform priors in an approximate
+Bayesian model choice framework.
+We discuss the dangers of this approach in the next section, and show how the
+results of \citet{Hickerson2013} likely exclude true values of the model's
+parameters.
 
@@
 distributions (i.e., they did not use an empirical Bayesian approach).
-\citet{Oaks2012} did use empirically informed priors when assessing
-both the prior sensitivity of the method and determining the power of \msb
-under real-world conditions with highly informative priors.
-\citet{Oaks2012} discuss the potential dangers of taking an empirical Bayesian
-approach to model choice (see last paragraph of ``Assessing prior sensitivity
-of \msb''), and we expand on this here by exploring both the theoretical and
-practical implications of using empirically informed priors for Bayesian
+\citet{Oaks2012} did use empirically informed priors when assessing both the
+prior sensitivity of the method and determining the power of \msb under
+real-world conditions with highly informative priors.  \citet{Oaks2012} discuss
+the potential dangers of taking an empirical Bayesian approach to model choice
+(see the last paragraph of ``Assessing prior sensitivity of \msb'' in
+\citet{Oaks2012}), and we expand on this here by exploring both the theoretical
+and practical implications of using empirically informed priors for Bayesian
 model choice.
@@
 
-This is not to say that empirical Bayesian approaches not useful.
+This is not to say that empirical Bayesian approaches are not useful.
 Quite the contrary.
@@
 Nevertheless, empirical Bayes methods can provide a powerful means of parameter
-estimation.
+estimation that often exhibit favorable frequentist properties.
 Furthermore, many post-hoc correction methods have been developed for
 estimating confidence-intervals from empirical Bayes estimates of posterior
-distributions that often exhibit good frequentist properties
+distributions that often exhibit well-behaved frequentist coverage
+probabilities
 \citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.
@@
 estimation, a theoretical justification for empirical Bayes approaches to model
-choice is much more uncertain.
+choice is much more questionable.
 In Bayesian model choice, the goal is not to estimate parameters, but to
@@
 renders the quarter dollar of the United States unfair.
-That is to say that their null hypothesis is that the probability of a US
+That is to say their null hypothesis is that the probability of a US
 quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
@@
 the quarter will land heads when it is flipped.
-Given their prior knowledge that Joe moonlights as a magician, and is notorious
-for performing coin and card tricks in the lab, they suspect there is a good
-chance that Joe will use his slight-of-hand skills to replace their quarter
-with one of his two-headed or two-tailed coins he has up his sleeve at all
-times.
+Given that neither of them have a quarter, and their prior knowledge that Joe
+moonlights as a magician, and is notorious for performing coin and card tricks
+in the lab, they suspect there is a good chance that the quarter Joe uses will
+be either two-headed or two-tailed.
 So, knowing that the beta distribution is the conjugate prior for a binomial
@@
 
-Mary calls Joe into the lab, and hands him a regular-issue US quarter minted in
-1992.
+Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
+begin flipping it.
 After five tosses, four of which was heads, Joe decides that academics are
@@
 
-However, upon reflecting on the results of their experiment
-(Figure~\ref{figCoinFlip}), Mary and Will regret not trusting Joe to refrain
-from turning their experiment into a magic trick.
-Their prior looks very ``poorly selected,'' and if they had only
-had more faith in Joe's integrity, they would have selected a much better
-prior. Clearly form their results, they should have used a prior centered
+However, reflecting upon the results of their experiment
+(Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
+Their prior looks very ``poorly selected,'' and if they had only known that the
+coin was fair before the data were collected, they would have selected a much
+better prior.
+Clearly, from their results, they should have used a prior centered
 around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
-much more appropriate.
+much ``better.''
 Rather than resort to flipping the coin five more times, Mary and Will decide
@@
 probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
-Now convinced that the mass of President Washington's head renders the US
-Quarter unfair, Will begins composing an e-mail of complaint for the U.S.\
-Mint.
+Now convinced that the US quarter is unfair, albeit not due to the mass of
+President Washington's head as they hypothesized, Will begins composing an
+e-mail of complaint addressed to the U.S.\ Mint.
 
@@
 estimators, using them for model choice is much less certain.
+In addition to these theoretical concerns there are practical dangers to
+using narrow, empirical, uniform priors for a method that has already
+been shown to be biased toward models with lesser parameter space.
 
@@
 choice}
-We were alarmed when we saw the results of the model-averaging results
-presented in \citet{Hickerson2013}.
+We were alarmed when we saw the results of the model-averaging approach
+used by \citet{Hickerson2013}.
 Their results strongly favored the models with the narrowest, empirically
@@
 \citet{Hickerson2013}).
-This concerned us because the narrowest prior of \citet{Hickerson2013}
-($\divt{} \sim U(0,0.1)$) likely excludes the true divergence times for at
-least some of the Philippine taxa.
+This concerned us because the narrowest \divt{} prior used by
+\citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
+divergence times for at least some of the Philippine taxa.
 They set this prior to match the 95\% highest posterior density (HPD) interval for
 the mean divergence time estimated under one of the priors used by
-\citet{Oaks2012} (see Table 2 and 3 of \citet{Oaks2012}).
+\citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
 Given this interval estimate is for the \emph{mean} divergence time across all
@@
 Lindley's paradox was again biasing the method toward models with smaller
-parameter space, and as a consequence was biasing the method toward models
-that exclude the truth (i.e., the model-averaged posterior will
-be dominated by models that exclude the true values of parameters of the model).
+parameter space, and as a consequence was biasing the method toward models that
+exclude the truth (i.e., toward estimating model-averaged posteriors
+dominated by models that exclude true values of parameters of the model).
 
@@
 For our re-analysis of the Philippines dataset we used the model-averaging
-approach of \citet{Hickerson2013}, but with different prior models to avoid
-their error of mixing time units.
-We used five priors, all of which had priors on population sizes of
+approach of \citet{Hickerson2013}, but with a reduced set of prior models to
+avoid their error of mixing time units.
+We used five prior models, all of which had priors on population sizes of
 $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
 U(0.0001, 0.05)$.
-Following \citet{Hickerson2013}, each of these priors had the following
+Following \citet{Hickerson2013}, each of these models had the following
 priors on divergence time parameters:
@@
 We then repeated this analysis twice, replacing the $M_1$ with
-$M_1A$ and $M_1B$, which differ only by having priors on divergence
+$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
 times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
@@
 Our results show that the model-averaging approach of \citet{Hickerson2013}
-does not reduce the methods bias towards models with smaller parameter space,
+does not reduce the method's bias toward models with lesser parameter space,
 but rather allows it to manifest in a very dangerous way.
@@
 almost certainly incorrect and excludes the true divergence times of the
-Philippine taxa.
+Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
 Unfortunately, ``checking'' the priors by plotting the summary statistics from
@@
 a principle component analysis, as recommended by \citet{Hickerson2013},
-provides no warning of a problem (Figure~\ref{figPCA}.
+provides no warning of a problem (Figure~\ref{figPCA}).
 Given that the results of \citet{Hickerson2013} strongly prefer the models with
@@
 To better quantify the propensity of Hickerson et al.'s
-\citeyear{Hickerson2013} approach to exclude the truth, we simulated
+(\citeyear{Hickerson2013}) approach to exclude the truth, we simulated
 1000 pseudo-observed datasets in which the divergence times for the
 22 population pairs are drawn randomly from an exponential distribution
-with a mean of 0.5 ($\divt{} \sim Exp(2)$). All other parameters were
-identically distributed as the $M_1$--$M_5$ models.
+with a mean of 0.5 ($\divt{} \sim Exp(2)$).
+All other parameters were identically distributed as the $M_1$--$M_5$ models
+(Table~\ref{tabModelChoiceEmpirical}).
 We then repeated the analysis described above, using $5\e6$ prior samples from
@@
 GLM-regression adjustment), excluding up to 21 of the 22 true divergence times
-(Figure~\ref{figExclusionSimTau}.
+(Figure~\ref{figExclusionSimTau}).
 Furthermore, the posterior probability of excluding at least one true parameter
@@
 model-averaging framework.
-This is especially dangerous \msb, which has already been demonstrated to be
-biased away from models with more parameter space.
+This is especially dangerous in \msb, which has already been demonstrated to be
+biased away from models with more parameter space \citep{Oaks2012}.
 The consequence of this approach is a high risk of obtaining a model-averaged
@@
     \citet{Hickerson2013}}
-\citet{Hickerson2013} did not conduct a power analysis to accompany their
-inference of clustered divergences from the Philippines data (as recommended by
-\citet{Oaks2012}).
-Thus, we perform simulation-based power analyses to explore the power of the
-approximate Bayesian model-averaging approach championed by
-\citet{Hickerson2013}.
-
+As recommended by \citet{Oaks2012}, we perform simulation-based analyses
+to explore the power of the approximate Bayesian model-averaging approach
+championed by \citet{Hickerson2013}.
 Following \citet{Oaks2012}, we simulated 1000 pseudo-replicate datasets with
@@
 Following \citet{Hickerson2013} (but avoiding mixing time units), we used five
-priors, all of which had priors on population sizes of $\meanDescendantTheta{}
-\sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim U(0.0001, 0.05)$.
+prior models, all of which had priors on population sizes of
+$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
+U(0.0001, 0.05)$.
 Each of these priors had the following priors on divergence time parameters:
@@
 for a total sample of $5\e6$.
-For each pseudo-replicate dataset, we retained the 1000 samples from each prior
+For each pseudo-replicate dataset, we retained 1000 samples from each prior
 model with the smallest Euclidean distance from the pseudo-observed summary
@@
 deviations of the given model.
-Then, for each replicate, we retained the 1000 samples with the smallest
-Euclidean distance from the pseudo-observed summary statistics, this time
-standardizing the statistics using the prior means and standard deviations
+Then, for each replicate, we retained 1000 samples with the overall
+smallest Euclidean distance from the pseudo-observed summary statistics, this
+time standardizing the statistics using the prior means and standard deviations
 across all five models.
@@
 The method is most likely to infer the extreme case of a single divergence event
-when populations diverge randomly over the past \globalcoalunit generations.
+when populations diverged randomly over the past \globalcoalunit generations.
 Even when divergences are random over the past $8\globalpopsize$ generations,
@@
 used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
-\citeyear{Oaks2012} Figure 5B).
+(\citeyear{Oaks2012}) Figure 5B).
 This supports the conclusions of \citet{Oaks2012} that this U-shaped
@@
 
-Contrary to Hickerson et al.'s \citeyear{Hickerson2013} claim that their
+Contrary to Hickerson et al.'s (\citeyear{Hickerson2013}) claim that their
 model-averaging approach ``can discriminate complex multispecies histories and
@@
 that method is biased toward supporting the extreme case of a single
-divergence event when populations diverged random over the last
+divergence event when populations diverged randomly over the last
 $8\globalpopsize$ generations.
@@
 The results of our power analysis further demonstrate the propensity of
-Hickerson et al.'s \citeyear{Hickerson2013} approach to exclude true parameter
+Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true parameter
 values.
@@
 (Figure~\ref{figPowerProbExclusion}).
-Only when the data are identically distributed as one the prior models does
+Only when the data are identically distributed as one of the prior models does
 the method avoid excluding the truth more than 5\% of the time
@@
 \citet{Hickerson2013} dangerous in terms of excluding regions of parameter
-space containing the true, but still lacks power to detect random variation in
+space containing the truth, but still lacks power to detect random variation in
 divergence times over the scale of $8\globalpopsize$ generations or more.
-This translates to millions of generations for mitochondrial loci, and hundreds
-of millions of generations for nuclear loci.
+This roughly translates to millions of generations for mitochondrial loci, and
+hundreds of millions of generations for nuclear loci.
 
@@
 are random over hundreds of thousands of generations (rather than millions as
-demonstrated by \citet{Oaks2012}.
+demonstrated by \citet{Oaks2012}).
 It is not surprising that under limited parameter space the method has
@@
 However, what is important to consider is whether those conditions are relevant
-to real world applications of the method.
+to real-world applications of the method.
 \citet{Oaks2012} explore the behavior of the method under three divergence time
@@
 be 100\% certain that all taxa (18 for their simulations) diverged within
-the last $\globalpopsize$ generations.
+the last 4$\globalpopsize$ generations.
 This does not seem applicable to most empirical systems.
-Furthermore, even when such extensive prior information is available to only be
-able to detect multiple divergences on the scale of hundreds of thousands of
+Furthermore, even when such extensive prior information is available, to only
+be able to detect multiple divergences on the scale of hundreds of thousands of
 generations does not seem very powerful.
@@
 $1.92\e{-8}$ mutations per site per generation.
-If we translate there results to a scale more consistent with rates of mutation
+If we translate their results to a scale more consistent with rates of mutation
 of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
@@
 
-\section{Validation analyses}
-\highLight{Not sure where to put this}.
-The main contribution of this analysis is to show that the unadjusted posterior
-is a better model-choice estimator than the regression-corrected estimates
-(Figure~\ref{figValidationMCBehavior}).
-This justifies our presenting the unadjusted results in this paper.
+\section*{Validation analyses}
+\highLight{Not sure where to put this section}.
+\highLight{This justifies presenting the unadjusted results in this paper.}
 \highLight{Is it worth including just for this purpose?}
+Following \citet{Oaks2012}, we characterize the model-choice behavior of the
+model-averaging approach of \citet{Hickerson2013} under the ideal conditions
+where the prior is correct (i.e., the data are generated from parameters drawn
+from the same prior distributions used in the analysis).
+We used the same prior models as above ($M_1$--$M_5$;
+Table~\ref{tabModelChoiceEmpirical}), and generated 50,000 pseudo-replicate
+datasets under this prior (10,000 from each model).
+We used a simulated data structure of eight population pairs, with a single
+1000 bp locus sampled from 10 individuals from each population.
+We then analyzed each of these replicate datasets using the same prior,
+retaining 1000 posterior samples.
+Our results are very similar to \citet{Oaks2012}, but we note that they
+are not directly comparable as our simulations contained eight population
+pairs rather than 10 (Figure~\ref{figValidationMCBehavior}).
+We find that the method does a decent job of estimating the posterior
+probability of divergence models when all assumptions of the method
+are met, the prior is correct, and the unadjusted posterior estimates
+are used.
+Similar to \cite{Oaks2012}, we find that the regression-adjusted estimates of
+the model probabilities are biased.
 
@@
 the range of gene divergence time estimates, and applying Hickerson et al.'s
-\citeyear{Hickerson2013} probability equations demonstrates that the prior
+(\citeyear{Hickerson2013}) probability equations demonstrates that the prior
 is densely populated with samples with large numbers of divergence parameters
@@
 \citet{Oaks2012} do not see such sensitivity when they compare prior sample
-sizes of $2\e6$, $5\e6$, and $10\e7$.
+sizes of $2\e6$, $5\e6$, and $10^7$.
 
 To explore this prediction further, we repeat the analysis of the Philippines
-dataset under the intermediate prior used by \citet{Oaks2012} ($\divt \sim U(0,
+dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim U(0,
 10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
-(0.0005, 0.02)$), using a very large prior sample size of $10\e8$.
+(0.0005, 0.02)$), using a very large prior sample size of $10^8$.
 When we look at the trace of the estimates of the dispersion index of
@@
 Taken together, the results presented here and in \citet{Oaks2012} support
-Lindley's paradox as the primary mechanism by with broad uniform priors
-cause biases of \msb.
+Lindley's paradox as a primary mechanism by which broad uniform priors
+cause biases in \msb.
+Nonetheless, posterior sampling error will always be present in any numerical
+Bayesian approximation method.
+Thus, insufficient sampling of the prior will contribute to the error of all
+approximate Bayesian estimates.
+However, there is no strong evidence that it is playing a large role in
+the biases revealed herein and by \citet{Oaks2012}.
 
@@
 It is not too surprising that the model-averaging approach proposed by
-\citet{Hickerson2013} behaves so poorly.
+\citet{Hickerson2013} behaves poorly.
 While this approach is trivial to implement, it is not a trivial change
-to the the basic \msb model.
+to the basic \msb model.
 To better understand why this is, it might help to step back and get a
@@
 
-Following the model description and notation of \citet{Oaks2012}, we
-take the time to tally up all of the free parameters in the \msb model.
-Under the simplest model in \msb (i.e., assuming no migration and
+Following the model description and notation of \citet{Oaks2012}, let us tally
+up all of the free parameters in the \msb model.  Under the simplest model in
+\msb (i.e., assuming no migration and
 no intra-locus recombination), the number of parameters for each
@@
 (\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
-The magnitude of population contraction in each of the descendant
+the magnitude of population contraction in each of the descendant
 populations (\bottleScalar{1}{} and \bottleScalar{2}{}) and the
@@
 copies sampled from both populations of the pair.
-
 Because we only have a single locus for each taxon, there are no locus-specific
@@
 the vector \divtvector.
-Thus for our Philippines data and under the simplest model in \msb, there are
-581--602 free parameters.
+Overall, for our Philippines data under the simplest model in \msb, there are
+581--602 free parameters (depending on the cardinality of \divtvector).
 Under this rich model, the method is sampling over 1002 divergence models
-\citep[i.e., the number of integer partitions of $Y=22$][]{Oaks2012}.
+\citep[i.e., the number of integer partitions of $Y=22$;][]{Oaks2012}.
 
@@
 $SD(\pi-\theta_W)$ \citep{Tajima1989}.
-That gives us a total of 88 summarys statistics (four for each of the 22 taxon
+That gives us a total of 88 summary statistics (four for each of the 22 taxon
 pairs), which contain minimal information about many of the $\approx 600$
@@
 This very large number of parameters and divergence models relative to 
-summary statistics is undoubtedly the reason the method is so sensitive
+summary statistics is undoubtedly a key reason the method is so sensitive
 to the prior distributions.
-It also likely contributes to the methods lack of robustness.
-We use robustness here as a measure of a methods insensitivity to model
+It also likely contributes to the method's lack of robustness.
+We use robustness here as a measure of a method's insensitivity to model
 violations.
@@
 models and prior models.
-This certainly is a non-trivial extension of the model, and given the method's
+This certainly is a non-trivial extension of the model and, given the method's
 very large number of parameters and models relative to the information used
@@
 
-The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
-and biases of \msb are actually similar to those of \citet{Hickerson2013},
-but avoid the need for imposing additional model choice.
+The recommendations of \citet{Oaks2012} for mitigating the bias and lack of
+robustness of \msb are actually similar to those of \citet{Hickerson2013}, but
+avoid the need for imposing additional model choice.
 \citet{Oaks2012} suggest that uniform priors are inappropriate for many
@@
 \migrationRate{}, \recombinationRate;][]{Oaks2012}.
-There is no cost whatsoever to this strategy.
+
+Importantly, there is no cost to this strategy.
 Given that ABC methods merely have to draw random values from prior
-distributions, hence there are no difficult proposal probability ratios to
-calculate, etc., there is no excuse for not using the most appropriate
+distributions (hence there are no difficult proposal probability ratios to
+calculate, etc.) there is no reason for not using the most appropriate
 distributions on parameters with conjugacy in mind.
-Also, this approach reduces the ``need'' for empirically guided priors.
+Also, this approach reduces the temptation of empirically guided priors.
 The investigator can place the majority of the prior density in regions
@@
 priori.
-Thus, to avoid the behavior of the method under these priors, cheating
-and using the data is a very tempting alternative.
+Thus, to avoid the behavior of the method under these uniform priors, using the
+data is a very tempting, albeit dangerous, option \citep{Hickerson2013}.
+More appropriate prior distributions would alleviate this issue.
 
 \section*{Other clarifications}
+
+\subsection*{Graphical checks of priors}
+\citet{Hickerson2013} advocate the use of what they call graphical checks of
+prior models.
+This entails generating a small number of random samples from the prior (1000)
+and plotting the resulting summary statistics in comparison to the observed
+statistics to see if they coincide (see Figure 1 of \citet{Hickerson2013}).
+As we show above, this is a dangerous strategy for choosing priors.
+It is not surprising that the resulting plots of this approach have little
+correlation with the appropriateness of priors.
+Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
+dataset analyzed by \citet{Hickerson2013}), we do not expect that 1000
+\emph{random} draws from the vast prior parameter space will yield data and
+summary statistics consistent with the observed data.
+In fact, when it does, this can be an obvious warning sign that the prior is
+over-fit, as we show above (Table~\ref{tabModelChoiceEmpirical} and
+Figure~\ref{figPCA}).
+Thus, interpreting such plots should be avoided, and the use of posterior
+predictive analyses would be much more informative about the overall fit of
+models.
 
@@
 To explore this possibility, we simulated datasets under prior settings that
-match two of the three priors used by \citet{Oaks2012}: $\meanDescendantTheta
-\sim U(0.0005, 0.04)$ and $\ancestralTheta \sim U(0.0005, 0.02)$.
-We draw divergence time parameters from a uniform distribution of $U(0, 20)$,
-simulated datasets, and plotted the \divt{} values against the summary
-statistics calculated from the resulting datasets
+match two of the three priors used by \citet{Oaks2012}: $\meanDescendantTheta{}
+\sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
+Under this prior, we draw divergence time parameters from a uniform
+distribution of $U(0, 20)$, simulate datasets, and plot the \divt{} values
+against the summary statistics calculated from the resulting datasets
 (Figure~\ref{figSaturationPlot}).
-Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of 5
-and 10 did not suffered from little to no effect from saturation.  Even at
-divergence times of 20 coalescent units, there is still signal in the summary
-statistics used by \msb (Figure~\ref{figSaturationPlot}).
+Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
+and 10 suffered from little to no effect from saturation.
+Even at divergence times of 20 coalescent units, there is still signal in the
+summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
+Thus, the assertion of \citet{Hickerson2013} does not apply to at least
+two of the priors used by \citet{Oaks2012} and, as a result, does not
+explain the bias they found.
 
 \section*{Conclusions}
+We demonstrate how the approximate Bayesian model choice method implemented in
+\msb can be strongly biased away from models with greater parameter space.
+We agree with \citet{Oaks2012} that this is likely caused by the inappropriate
+use of uniform priors on most of the model's parameters.
+These priors necessitate the use of broad priors that place high prior density
+in unlikely regions of parameter space, less the risk of excluding the truth
+\emph{a priori}.
+This likely hinders the marginal likelihoods of models with greater parameter
+space, either due to more divergence time parameters or broader prior
+distributions on those parameters.
+We show that the empirical Bayesian model-averaging approach of
+\citet{Hickerson2013} does not mitigate this bias, but rather causes it to
+manifest by sampling predominantly from models that may exclude the true values
+of the parameters.
+Exploring cogent prior probability distributions on most of the model's
+parameters, in addition to different priors over divergence models, could
+help mitigate the methods biases.
+
+The work presented herein follows the principles of Open Notebook Science.
+All aspects of the work were recorded in real-time via version-control
+software and are publicly available at
+\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
+All information necessary to reproduce our results are provided there.
 
@@
 
-%:FIGURE-saturation plot
-\mFigure{../../saturation/saturation-plot.pdf}{
-    The summary statistics $\pi$ \citep{Tajima1983} and $\pi_{net}$
-    \citep{Takahata1985} as a function of divergence time between populations.
-    Each plot represents 1100 pairs of parameter draws and summary statistics
-    calculated from the simulated data.
-    Prior settings for the simulations were $\divt{} \sim U(0, 20)$,
-    $\meanDescendantTheta{} \sim U(0.0005, 0.04)$, and $\ancestralTheta{} \sim
-    U(0.0005, 0.02)$.
-}{figSaturationPlot}
-
-% \mFigure{../../response-redux/results/hickerson/pymsbayes-results/pymsbayes-output/d1/m12345678-combined/mean_by_dispersion.pdf}{
-%     Results of reanalysis.
-% }{figJointPosterior}
-
 \mFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
@@
 
-\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
-    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
-    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
-    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
-    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
-    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
-    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
-}{figSamplingError}
+\mFigure{../images/coin_flip.pdf}{
+    A plot of three beta probability density functions that represent a prior
+    (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
+    empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
+    Bernoulli trials, four of which are successes.
+}{figCoinFlip}
+
+\mFigure{../../model-choice/priors-for-pc-plot/pc-plots.pdf}{
+    The graphical checks recommended by \citet{Hickerson2013} for three prior
+    models: (A) $M_1$ ($\divt{} \sim U(0, 0.1)$), (B) $M_{1A}$ ($\divt{} \sim
+    U(0, 0.01)$), and (C) $M_{1B}$ ($\divt{} \sim U(0, 0.001)$).
+    The plots project the summary statistics from 1000 random samples from each
+    model onto the first two orthogonal axes of a principle component analysis,
+    with the blue dot representing the observed summary statistics from the 22
+    population pairs of Philippine vertebrates.
+}{figPCA}
 
@@
 
+\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
+    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
+    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
+    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
+    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
+    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
+    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
+}{figSamplingError}
+
 %% Model-choice tau prior
@@
 
-\mFigure{../images/coin_flip.pdf}{
-    A plot of three beta probability density functions that represent a prior
-    (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
-    empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
-    Bernoulli trials, four of which are successes.
-}{figCoinFlip}
-
-\mFigure{../../model-choice/priors-for-pc-plot/pc-plots.pdf}{
-    The graphical checks recommended by \citet{Hickerson2013} for three prior
-    models: (A) $M_1$ ($\divt{} \sim U(0, 0.1)$), (B) $M_{1A}$ ($\divt{} \sim
-    U(0, 0.01)$), and (C) $M_{1B}$ ($\divt{} \sim U(0, 0.001)$).
-    The plots project the summary statistics from 1000 random samples from each
-    model onto the first two orthogonal axes of a principle component analysis,
-    with the blue dot representing the observed summary statistics from the 22
-    population pairs of Philippine vertebrates.
-}{figPCA}
+%:FIGURE-saturation plot
+\mFigure{../../saturation/saturation-plot.pdf}{
+    The summary statistics $\pi$ \citep{Tajima1983} and $\pi_{net}$
+    \citep{Takahata1985} as a function of divergence time between populations.
+    Each plot represents 1100 pairs of parameter draws and summary statistics
+    calculated from the simulated data.
+    Prior settings for the simulations were $\divt{} \sim U(0, 20)$,
+    $\meanDescendantTheta{} \sim U(0.0005, 0.04)$, and $\ancestralTheta{} \sim
+    U(0.0005, 0.02)$.
+}{figSaturationPlot}
+
+% \mFigure{../../response-redux/results/hickerson/pymsbayes-results/pymsbayes-output/d1/m12345678-combined/mean_by_dispersion.pdf}{
+%     Results of reanalysis.
+% }{figJointPosterior}
 

### 2013-09-23 manuscripts/dpp-msbayes/dpp-msbayes.tex
Starting methods of dpp ms.

@@
         \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2,3}$ and Mark T.\ Holder$^{2}$
+        Jamie R.\ Oaks$^{1,2}$ and Mark T.\ Holder$^{1}$
 
         \bigskip
-        $^1$\emph{Biodiversity Institute\\
-            $^2$Department of Ecology and Evolutionary Biology\\ 
-            University of Kansas\\
-            %1345 Jayhawk Blvd\\
-            Lawrence, KS 66045\\
-            USA}\\[.1in]
-        $^3$\emph{Corresponding author} (\href{mailto:joaks1@ku.edu}{\tt
-        joaks1@ku.edu})\\
+        $^1$\emph{Department of Ecology and Evolutionary Biology,
+            % Biodiversity Institute,
+            University of Kansas,
+            Lawrence, Kansas 66045}\\[.1in]
+        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
+        joaks1@gmail.com})\\
 
@@
 \noindent \ldots
+
+\section*{Methods}
+
+\subsection*{The model}
+We follow much of the notation of \citet{Oaks2012}, but modify it to
+accommodate the new model \ldots
+
+\begin{linenomath}
+    Let \npairs{} be the number of pairs of populations, \nloci{i} be the
+    number of loci sampled for population pair $i$,
+    and \nlociTotal be the total number of unique loci sampled across the
+    \npairs{} pairs of populations.
+    We use \alignment{i}{j} to represent the multiple sequence alignment of
+    locus $j$ for population pair $i$.
+    $\alignmentVector = \taxonLocusVector{\alignment}$ is the full dataset,
+    i.e., a vector of sequence alignments for all pairs and loci.
+    The joint posterior distribution of our model is given by
+\begin{equation}
+    f(\model{} \given \alignmentVector, \hkyModelVector, \ploidyScalarVector,
+    \mutationRateScalarConstantVector) =
+    \frac{f(\alignmentVector \given \model{}) f(\model{})}{
+        f(\alignmentVector)}
+    \label{eq:fullModelShort}
+    % \begin{split}
+    % & f(\geneTreeVector,
+    % \numt{},
+    % \divtvector, \ancestralThetaVector,
+    % \descendantThetaVector{1}, \descendantThetaVector{2},
+    % \locusRateHetShapeParameter,
+    % \locusMutationRateScalarVector,
+    % \bottleTimeVector, \bottleScalarVector{1},
+    % \bottleScalarVector{2},
+    % \migrationRateVector \given
+    % \alignmentVector, \hkyModelVector, \ploidyScalarVector,
+    % \mutationRateScalarConstantVector) \\
+    % & = \frac{1}{f(\alignmentVector)}
+    % f(\numt{}) f(\divtvector \given \numt{})
+    % f(\locusRateHetShapeParameter)
+    % \Bigg[
+    % \prod_{i=1}^{\npairs{}}
+    % f(\ancestralTheta{i})
+    % f(\descendantTheta{1}{i}, \descendantTheta{2}{i})
+    % f(\bottleTime{i})
+    % f(\bottleScalar{1}{i}) f(\bottleScalar{2}{i})
+    % f(\migrationRate{i}) \\
+    % &
+    % \prod_{j=1}^{\nloci{i}}
+    % f(\alignment{i}{j} \given
+    % \geneTree{i}{j}, \hkyModel{i}{j})
+    % f(\geneTree{i}{j} \given
+    % \divt{i},
+    % \ancestralTheta{i}, \descendantTheta{1}{i}, \descendantTheta{2}{i},
+    % \ploidyScalar{i}{j}, \mutationRateScalarConstant{i}{j},
+    % \locusMutationRateScalar{j},
+    % \bottleTime{i}, \bottleScalar{1}{i}, \bottleScalar{2}{i},
+    % \migrationRate{i})
+    % \Bigg]
+    % \Bigg[
+    % \prod_{j=1}^{\nlociTotal}
+    % f(\locusMutationRateScalar{j} \given \locusRateHetShapeParameter)
+    % \Bigg],
+    % \label{eq:fullmodel}
+    % \end{split}
+\end{equation}
+where \hkyModelVector, \ploidyScalarVector, and \mutationRateScalarConstantVector
+are vectors of fixed parameters specified by the user.
+More specifically $\hkyModelVector = \taxonLocusVector{\hkyModelVector}$ is a
+vector of fixed \hky model parameters for each alignment.
+% where $\geneTreeVector = \taxonLocusVector{\geneTree}$ are the gene trees upon
+% which each \alignment{i}{j} evolved according to the \hky substitution model
+% parameters \hkyModel{i}{j}.
+% The \hky model parameters for each alignment reside in vector 
+% $\hkyModelVector = \taxonLocusVector{\hkyModel}$,
+% and are fixed constants provided by the user.
+% \divtvector is the vector of times, in coalescent units, when the populations
+% of each pair diverged, \taxonVector{\divt}.
+% \numt{} is the hyper-parameter controlling the number of unique \divt{} within
+% \divtvector.
+% $\ancestralThetaVector = \taxonVector{\ancestralTheta}$ is the vector of
+% \myTheta{} parameters for the ancestral population of each population pair.
+% $\descendantThetaVector{1} = \taxonVector{\descendantTheta{1}}$ is the vector of
+% \myTheta{} parameters for the $1^{st}$ descendant population of each
+% pair;
+% $\descendantThetaVector{2} = \taxonVector{\descendantTheta{2}}$ is the same for
+% the $2^{nd}$ descendant population of each pair.
+% $\ploidyScalarVector = \taxonLocusVector{\ploidyScalar}$ and
+% $\mutationRateScalarConstantVector = \taxonLocusVector{\mutationRateScalarConstant}$
+% are vectors of \myTheta{}-scaling constants provided by the user
+% (Table~\ref{tab:notation}).
+% Furthermore, there are locus-specific \myTheta{}-scaling
+% parameters in the vector $\locusMutationRateScalarVector =
+% \locusVector{\locusMutationRateScalar}$.
+% \locusRateHetShapeParameter is the shape parameter of the gamma prior
+% distribution on each \locusMutationRateScalar{}.
+% $\bottleScalarVector{1} = \taxonVector{\bottleScalar{1}}$ is the vector of
+% \myTheta{}-scaling parameters that determine the magnitude of the bottleneck in
+% the $1^{st}$ descendant population of each population pair,
+% whereas $\bottleScalarVector{2} = \taxonVector{\bottleScalar{2}}$ is the same
+% for the $2^{nd}$ descendant population of each pair;
+% for both descendants of each population pair, the bottleneck begins immediately
+% after divergence in forward-time.
+% $\bottleTimeVector = \taxonVector{\bottleTime}$ is the vector of the
+% proportions of time between present and \divt{} when the bottleneck ends for
+% both populations in each pair; after which the populations grow exponentially to
+% present.
+% $\migrationRateVector = \taxonVector{\migrationRate}$ is the vector of
+% symmetric migration rates between the descendant populations of each pair.
+\end{linenomath}
 

### 2013-09-24 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on dpp ms.
Working on model description.

--- 2013-09-16 manuscripts/dpp-msbayes/dpp-msbayes.tex
@@
 
+\section*{Methods}
+
+\subsection*{The model}
+We follow much of the notation of \citet{Oaks2012}, but modify it to
+accommodate the new model \ldots
+
+We assume an investigator is interested in inferring the distribution
+of divergence times among \npairs{} pairs of populations.
+For each pair $i$, \popSampleSize{i}{} genome copies have been sampled, with
+\popSampleSize{1}{i} copies sampled from population 1, and \popSampleSize{2}{i}
+sampled from population 2.
+From these genomes, let \nloci{i} be the number of DNA sequence loci collected
+for population pair $i$, and \nlociTotal be the total number of unique loci
+sampled across the \npairs{} pairs of populations.
+We use \alignment{i}{j} to represent the multiple sequence alignment of
+locus $j$ for population pair $i$.
+$\alignmentVector = (\alignment{1}{1}, \ldots,
+    \alignment{\npairs{}}{\nloci{\npairs{}}})$
+is the full dataset,
+i.e., a vector of sequence alignments for all pairs and loci.
+We let \geneTree{i}{j} represent the gene tree upon which \alignment{i}{j}
+evolved according to fixed \hky substitution model parameters \hkyModel{i}{j}.
+The investigtor must specify the parameters of all
+$\hkyModelVector = (\hkyModel{1}{1}, \ldots,
+\hkyModel{\npairs{}}{\nloci{\npairs{}}})$
+substitution models by which the alignments evolved along the
+$\geneTreeVector = (\geneTree{1}{1}, \ldots,
+\geneTree{\npairs{}}{\nloci{\npairs{}}})$
+gene trees.
+Furthermore, the investigator must specify a vector of fixed contants
+$\ploidyScalarVector = (\ploidyScalar{1}{1}, \ldots,
+\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$
+that scale the population size parameters for known differences in ploidy among
+loci and/or differences in generation times among population pairs.
+Lastly, the investigator must also specify a vector of fixed constants
+$\mutationRateScalarConstantVector = (\mutationRateScalarConstant{1}{1},
+\ldots, \mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$
+that scale the population size parameters for known differences in
+mutation rates among loci and/or among taxa.
+
+\begin{linenomath}
+With \alignmentVector, \hkyModelVector, \ploidyScalarVector, and
+\mutationRateScalarConstantVector in hand, the joint posterior distribution
+of our model is given by
+\begin{equation}
+    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
+    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
+    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
+    \mutationRateScalarConstantVector) =
+    \frac{p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
+        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
+        \locusMutationRateScalarVector, \ploidyScalarVector,
+        \mutationRateScalarConstantVector)
+        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
+        p(\locusRateHetShapeParameter)
+        p(\divTimeMapVector)
+        p(\demographicParamVector)}{
+        p(\alignmentVector)},
+    \label{eq:fullModelCompact}
+\end{equation}
+where
+$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
+is a vector of population divergence times for each of \npairs{} pairs,
+$\demographicParamVector = (\demographicParams{1}, \ldots,
+\demographicParams{\npairs{}})$
+is a vector of the demographic parameters for each of the \npairs{} population
+pairs,
+$\locusMutationRateScalarVector = (\locusMutationRateScalar{1}, \ldots
+\locusMutationRateScalar{\nlociTotal})$
+is a vector of locus-specific mutation-rate scaling parameters for each of the
+\nlociTotal loci,
+and \locusRateHetShapeParameter is the shape parameter of a gamma-distributed
+prior on \locusMutationRateScalar{}.
+\end{linenomath}
+
+\subsubsection*{Likelihood terms of Equation \ref{eq:fullModelCompact}}
+\begin{linenomath}
+The likelihood terms of Equation \ref{eq:fullModelCompact} can
+be expanded out as a product over population pairs and loci
+\begin{equation}
+    p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
+    p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
+    \locusMutationRateScalarVector, \ploidyScalarVector,
+    \mutationRateScalarConstantVector) = 
+    \prod_{i=1}^{\npairs{}}
+    \prod_{j=1}^{\nloci{i}}
+    p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})
+    p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
+    \locusMutationRateScalar{j}, \ploidyScalar{i}{j},
+    \mutationRateScalarConstant{i}{j}).
+    \label{eq:modelLikelihoodExpanded}
+\end{equation}
+The first term,
+$p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})$,
+is the probability of the sequence alignment of locus $j$ for population pair
+$i$ given the gene tree and \hky substitution model paramters \citep[i.e., the
+``Felsenstein likelhood'';][]{Felsenstein1981}.
+The model allows for an intra-locus recombination rate \recombinationRate,
+which, for simplicity, is assumed to be zero in Equation
+\ref{eq:fullModelCompact}.
+If \recombinationRate is non-zero, this term requires an additional product
+over the columns (sites) of each sequence alignment to allow sites to have
+different genealogies.
+The second term,
+p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
+\locusMutationRateScalar{j}, \ploidyScalar{i}{j},
+\mutationRateScalarConstant{i}{j}),
+is the probability of the gene tree under a multi-population coalescent model
+(i.e., species tree) where the ancestral population of pair $i$ diverges
+and gives rise to the two sampled descendant populations.
+Each \demographicParams contains the following demographic parameters:
+The mutation-scaled sizes of the ancestral, \ancestralTheta{}, and descendant
+populations, \descendantTheta{1}{} and \descendantTheta{2}{};
+the proportion of the first, \bottleScalar{1}{}, and second population,
+\bottleScalar{2}{}, that persists during bottlenecks that begin immmediately
+after divergence in forward-time;
+the proportion of time between present and divergence when the bottlenecks
+end for both populations, \bottleTime{};
+and the symmetric migration rate between the descendant populations,
+\migrationRate{}.
+Thus, the probability of the $\popSampleSize{i}{}-1$ coalescence times (node
+heights) of gene tree \geneTree{i}{j} is given by a multi-population
+Kingman-coalescent model \citep{Kingman1982} where the ancestral population of
+size
+$\ancestralTheta{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
+diverges at time \divTimeMap{i} into two descendant populations of constant
+size
+$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{1}{i}$
+and
+$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{2}{i}$,
+which, after time $\divTimeMap{i}\bottleTime{i}$, grow exponentially to their
+present size 
+$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
+and
+$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$,
+respectively.
+Following divergence, the descendant populations of pair $i$ exchange migrants at a symmetric rate of
+\migrationRate{i}.
+\end{linenomath}
+
+\subsubsection*{Prior terms of Equation \ref{eq:fullModelCompact}}
+\begin{linenomath}
+The term $p(\locusRateHetShapeParameter)$ is the prior density
+function for the shape-parameter for the gamma-distributed prior on
+rate heterogeneity among loci.
+This prior is $\locusRateHetShapeParameter \sim U(1, 20)$.
+The prior probability of the vector of locus-specific mutation-rate scaling parameters
+given \locusRateHetShapeParameter then expands out as a product over
+the loci
+\begin{equation}
+    p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter) =
+    \prod_{j=1}^{\nlociTotal}
+    p(\locusMutationRateScalar{j} \given \locusRateHetShapeParameter),
+    \label{eq:locusRateHetPrior}
+\end{equation}
+where each \locusMutationRateScalar{} is independently and identicially
+distributed (\iid) as
+$\locusMutationRateScalar{} \sim Gamma(\locusRateHetShapeParameter,
+1/\locusRateHetShapeParameter)$.
+If the recombination rate \recombinationRate is allowed to be
+non-zero, the prior term $p(\recombinationRate)$ would be added
+to Equation \ref{eq:fullModelCompact}, and the prior would be
+$\recombinationRate \sim Gamma(\gammaShape{\recombinationRate},
+\gammaScale{\recombinationRate})$.
+\end{linenomath}
+
+\begin{linenomath}
+The prior term for the demographic parameters, $p(\demographicParamVector)$,
+expands out into its components and as a product over the \npairs{}
+pairs of populations
+\begin{equation}
+    p(\demographicParamVector) =
+    \prod_{i=1}^{\npairs{}}
+    p(\ancestralTheta{i})
+    p(\descendantTheta{1}{i})
+    p(\descendantTheta{2}{i})
+    p(\bottleScalar{1}{i})
+    p(\bottleScalar{2}{i})
+    p(\bottleTime{i})
+    p(\migrationRate{i}).
+    \label{eq:demographicPrior}
+\end{equation}
+The priors for the demographic parameters are:
+$\ancestralTheta{} \sim Gamma(\gammaShape{\ancestralTheta{}},
+\gammaScale{\ancestralTheta{}})$,
+$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+\gammaScale{\descendantTheta{}{}})$,
+$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+\gammaScale{\descendantTheta{}{}})$,
+$\bottleScalar{1}{} \sim Beta(\betaA{\bottleScalar{}{}},
+\betaB{\bottleScalar{}{}})$,
+$\bottleScalar{2}{} \sim Beta(\betaA{\bottleScalar{}{}},
+\betaB{\bottleScalar{}{}})$,
+$\bottleTime{} \sim U(0, 1)$,
+and
+$\migrationRate{} \sim Gamma(\gammaShape{\migrationRate{}},
+\gammaScale{\migrationRate{}})$.
+\end{linenomath}
+
+\subsubsection*{Priors on divergence models}
+\begin{linenomath}
+The prior term for the vector of divergence times for each of
+the \npairs{} pairs of populations can be expanded as
+\begin{equation}
+    p(\divTimeMapVector) = p(\divTimeIndexVector)p(\divTimeVector \given \divTimeIndexVector),
+    \label{eq:divModelPrior}
+\end{equation}
+where \divTimeVector is an ordered set of unique divergence time parameters
+$\{\divTime{1}, \ldots, \divTime{\divTimeNum}\}$ whose length
+\divTimeNum can range from 1 to \npairs{},
+and \divTimeIndexVector is vector of indices of divergence time parameters 
+$(\divTimeIndex{1}, \ldots, \divTimeIndex{\npairs{}})$
+that map the divergence time in \divTimeVector to each of the \npairs{}
+pairs of populations.
+Thus, \divTimeMapVector is the result of applying the mapping function
+\begin{equation}
+    f(\divTimeVector, \divTimeIndexVector, i) = \divTime{\divTimeIndex{i}}
+    \label{eq:divTimeMapFunction}
+\end{equation}
+to each population pair $i$, such that
+$\divTimeMapVector = (\divTimeMap{1} = f(\divTimeVector, \divTimeIndexVector,
+1), \ldots, \divTimeMap{\npairs{}} = f(\divTimeVector, \divTimeIndexVector,
+\npairs{}))$.
+Each \divTime{} within \divTimeVector is \iid as
+$\divTime{} \sim Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
+\end{linenomath}
+
+Biologically speaking, \divTimeVector contains the times of divergence events,
+the length of which \divTimeNum is the number of divergence events shared across
+the \npairs{} pairs of populations.
+For example, if \divTimeVector contains a single divergence time parameter
+\divTime{1}, all \npairs{} pairs of populations are constrained to diverge at
+this time (i.e., \divTimeIndexVector would contain the index 1 repeated
+\npairs{} times, and \divTimeMapVector would contain the value \divTime{1}
+repeated \npairs{} times), whereas if it contains \npairs{} divergence time
+parameters, the model is fully generalized to allow all of the pairs to diverge
+at unique times.
+
+\begin{linenomath}
+The total sample space of ordered realizations of \divTimeMapVector is all the
+possible partitions of \npairs{} elements into 1 to \npairs{} categories.
+The total number of possible partitions is a sum of the Stirling numbers of
+the second kind over all possible number of categories \divTimeNum
+\begin{equation}
+    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
+    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
+    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
+    \label{eq:bell}
+\end{equation}
+which is the Bell Number \citep{Bell1934}.
+The original \msb model only samples over the unordered realizations of
+\divTimeIndexVector, such that the sample space is reduced to all the possible
+integer partitions of \npairs{} \citep{Oaks2012,OeisPartitionNumber,
+    OeisPartitionTriangle,Malenfant2011}.
+This has the advantage of reducing the sample space of the number of
+divergence models, but has two disadvantages.
+First, additional information about the data is lost.
+By throwing away the identity of the \npairs{} pairs of populations, all
+information regarding the amount of data (e.g., the number gene copies
+collected from each of the populations [\popSampleSize{1}{} and
+\popSampleSize{2}{}], the number of loci, and the length of the loci), the
+information content in the data (i.e., the amount of sequence variation), as
+well as the taxon- and locus-specific parameters (\hkyModel{}{},
+\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
+\locusMutationRateScalar{}) for each pair is lost.
+The original descriptions of the \msb model claim that the independence
+between the sample sizes \popSampleSize{1}{}, \popSampleSize{2}{} and
+the summary statistic $\pi_b$ calculated from the alignment for each
+population pair allows exchangeability.
+Second, the identity of the pairs of populations is lost.
+\end{linenomath}
+
+We implement two prior probability distributions over the joint space of
+\divTimeVector, \divTimeIndexVector.
+One simply places a uniform prior over all possible integer paritions of
+\npairs{}.
+
 \section*{Acknowledgments}
+We thank the National Science Foundation for supporting this work (DEB
+1011423).
+J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
+Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
+Society, KU Department of Ecology and Evolutionary Biology, and the KU
+Biodiversity Institute.
 

### 2013-09-25 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on model description for dpp ms.

@@
 prior on \locusMutationRateScalar{}.
+\end{linenomath}
+
+\begin{linenomath}
+To avoid calculating the likelihood terms of Equation \ref{eq:fullModelCompact},
+we distill each sequence alignment \alignment{}{} into a vector of insufficient
+summary statistics \alignmentSS{}{}, thus replacing the full dataset
+$\alignmentVector = (\alignment{1}{1}, \ldots,
+    \alignment{\npairs{}}{\nloci{\npairs{}}})$
+with vectors of summary statistics for each alignment
+$\ssVector{} = (\alignmentSS{1}{1}, \ldots,
+    \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$.
+This allows us to estimate the approximate joint posterior distribution
+{\small
+\begin{equation}
+    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
+    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
+    \ssSpace, \hkyModelVector, \ploidyScalarVector,
+    \mutationRateScalarConstantVector) =
+    \frac{p(\ssSpace \given \geneTreeVector, \hkyModelVector)
+        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
+        \locusMutationRateScalarVector, \ploidyScalarVector,
+        \mutationRateScalarConstantVector)
+        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
+        p(\locusRateHetShapeParameter)
+        p(\divTimeMapVector)
+        p(\demographicParamVector)}{
+        p(\ssSpace)},
+    \label{eq:approxModelCompact}
+\end{equation}
+}
+where \ssSpace is the multidimensional Euclidean space around the vector of
+summary statistics, the radius of which is the tolerance \tol.
+The approximation of this posterior is caused by the insufficiency of the
+statistics and the \tol being greater than zero.
+We describe the full model in detail before delving into the numerical
+method of estimating the approximate model.
 \end{linenomath}
@@
 population pair allows exchangeability.
+This is incorrect.
+The other summary statistics, and the alignments they are based on are
+not independent of the sample sizes.
+The alignments and summary statitics are not identically distributed because of
+taxon- and locus-specific parameters \hkyModel{}{},
+\mutationRateScalarConstant{}{}, and \ploidyScalar{}{}
 Second, the identity of the pairs of populations is lost.

### 2013-09-25 manuscripts/dpp-msbayes/dpp-msbayes.tex
Adding more text to model description in dpp ms.

@@
 \npairs{}))$.
-Each \divTime{} within \divTimeVector is \iid as
-$\divTime{} \sim Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
 \end{linenomath}
@@
 at unique times.
+Regardless of the number of divergence time parameters (\divTimeNum), each
+\divTime{} within \divTimeVector is \iid as $\divTime{} \sim
+Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
 
@@
 \divTimeIndexVector, such that the sample space is reduced to all the possible
-integer partitions of \npairs{} \citep{Oaks2012,OeisPartitionNumber,
+integer partitions of \npairs{} \citep{Oaks2012,Huang2011,OeisPartitionNumber,
     OeisPartitionTriangle,Malenfant2011}.
+We denote the set of all possible integer partitions of \npairs{} as
+\integerPartitionSet{\npairs{}} and the length of that set as
+\integerPartitionNum{\npairs{}}.
+The advantages, disadvantages, and justification of ignoring the order
+of \divTimeIndexVector is discussed in detail below.
+\end{linenomath}
+
+\begin{linenomath}
+We implement two prior probability distributions over the space
+of all possible discrete divergence models (\divTimeIndexVector).
+The first simply gives all possible unordered partitions of \npairs{} elements
+equal probability
+\begin{equation}
+    p(\divTimeIndexVector) = frac{1}{\integerPartitionNum{\npairs{}}},
+    \label{eq:divModelPriorUniform}
+\end{equation}
+i.e., a discrete uniform prior over all the integer partitions of \npairs{}
+(discrete divergence models).
+We denote this prior as
+$\divTimeIndexVector \sim DU(\integerPartitionSet{\npairs{}})$.
+\end{linenomath}
+
+The second prior we implement is based on the Dirichlet-process, which is a
+stochastic process that groups elements into an unknown number of discrete
+parameter classes \citep{Ferguson1973,Antoniak1974}.
+The process is controlled by the concentration parameter \concentrationParam,
+which determines how clustered the process will be.
+We use the Dirichlet process to place a prior over all possible ordered
+partitions of \npairs{} elements, which we denote as $\divTimeIndexVector \sim
+DP(\concentrationParam)$.
+We use a hierarchical Bayesian approach and place a second-order prior
+probability distribution (i.e., hyperprior) on \concentrationParam.  More
+specifically, we use a gamma-distributed prior $\concentrationParam =
+Gamma(\gammaShape{\concentrationParam}, \gammaScale{\concentrationParam})$,
+where \gammaShape{\concentrationParam} and \gammaScale{\concentrationParam} are
+specified by the user.
+
+\begin{linenomath}
+This provides a great deal of flexibility for specifying the prior uncertainty
+regarding divergence models.
+The \concentrationParam determines the prior probability that any two
+pairs of populations $i$ and $j$ will be assigned to the same divergence time
+parameter
+\begin{equation}
+    p(\divTimeIndex{i} = \divTimeIndex{j}) = frac{1}{1 + \concentrationParam},
+    \label{eq:dppPriorSameClass}
+\end{equation}
+and also the prior probability of the number of divergence time parameters
+\begin{equation}
+    p(\divTimeNum \given \concentrationParam, \npairs{}) = 
+    \frac{\stirlingFirst{\npairs{}}{\divTimeNum} \concentrationParam^{\divTimeNum}}
+    {\prod_{i=1}^{\npairs{}}(\concentrationParam + i - 1)},
+    \label{eq:dppPriorNumClasses}
+\end{equation}
+where \stirlingFirst{\cdot,\cdot} are the unsigned Stirling numbers of the
+first kind.
+Equations \ref{eq:dppPriorSameClass} and \ref{eq:dppPriorNumClasses} show that
+smaller values of \concentrationParam will favor fewer divergence time
+parameters, and thus more clustered models of divergence, whereas larger values
+will favor more divergence time parameters, and thus less clustered models of
+divergence.
+\end{linenomath}
+
+\subsection*{Differences between our model and the original \msb model}
+\subsubsection*{The prior on divergence models}
+One of the key differences between our model and that of \msb \citep{Huang2011}
+is the prior distribution placed on divergence models.
+As discussed in \citet{Oaks2012}, in \msb the only prior used for
+\divTimeIndexVector is a combination of a discrete uniform prior over the
+possible number of divergence events \divTimeNum from 1 to \npairs{} with a
+multinomial distribution on the number of times each index of \divTimeVector
+appears in \divTimeIndexVector, with the constraint that all \divTime
+parameters are represented at least once (see Equation 2 of \citet{Oaks2012}).
+\citet{Oaks2012} discuss how placing a uniform prior over the number of
+divergence parameters (\divTimeNum here; \numt{} in \citet{Huang2011}) imposes
+an odd U-shaped prior over discrete divergence models (\divTimeIndexVector; see
+Figure 5B of \citet{Oaks2012}).
+To avoid this, we place priors directly on the sample space of divergence
+models, thus eliminating the parameter \numt{} from the model.
+We introduce two priors on divergence models:
+(1) a prior that is uniform over all unordered divergence models, and
+(2) a Dirichlet-process prior on all ordered divergence models.
+The latter provides the user with a great deal of flexibility in
+expressing their prior beliefs about models of divergence.
+
+\subsubsection*{The priors on nuisance parameters}
+Following the recommendations of \citet{Oaks2012}, we have replaced the use of
+inappropriate, continuous uniform distributions for priors on many of the
+model's parameters (\divTime{}, \ancestralTheta{}, \descendantTheta{1}{},
+\descendantTheta{2}{}, \bottleScalar{1}{}, \bottleScalar{2}{},
+\recombinationRate, \migrationRate{}) with more flexible parametric
+distributions from the exponential family.
+We introduce gamma-distributed priors for rate parameters that have a sample
+space of all positive real numbers (\divTime{}, \ancestralTheta{},
+\descendantTheta{1}{}, \descendantTheta{2}{}, \recombinationRate,
+\migrationRate{}), and beta-distributed priors for parameters that are
+proportions bounded by zero and one (\bottleScalar{1}{} and
+\bottleScalar{2}{}).
+Not only do these priors have nice conjugacy properties, but they also provide
+the user with much greater flexibility in expressing the prior uncertainty
+regarding the parameters of the model.
+
+Another distinction between our model and that of \msb, is the prior on the
+sizes of the descendant populations of each pair.
+As described by \citet{Oaks2012}, \msb uses the joint prior
+\begin{equation}
+    \descendantTheta{1}{}, \descendantTheta{2}{} \sim
+    Dirichlet(1,1) \times 2 \times U(\uniformMin{\myTheta{}},
+    \uniformMax{\descendantTheta{}{}}),
+    \label{eq:jointThetaPrior}
+\end{equation}
+such that the user-specified uniform prior on descendant population
+size is a prior on the \emph{mean} size of the two descendant
+populations of each pair.
+Under our model, the sizes of the descendant populations of each
+pair are \iid as
+$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+\gammaScale{\descendantTheta{}{}})$
+and
+$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+\gammaScale{\descendantTheta{}{}})$.
+This relaxes the assumption that the sizes of the two descendant populations
+are interdependent and negatively correlated.
+
+\subsubsection*{Flexibility in parameterizing the model}
+We also provide the user with the ability to control the richness of
+the model.
+For the \myTheta{} parameters, the model can be fully generalized to
+allow each population pair to have three parameters:
+\ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}.
+Furthermore, any model of \myTheta{} parameters nested within this
+general model can also be specified, including the most restricted model
+where the ancestral and descendant populations of each pair share
+a single \myTheta{} parameter.
+We also allow the user to eliminate the parameters associated with
+the post-divergence bottlenecks in the descendant populations of
+each pair
+(\bottleTime{}, \bottleScalar{1}{}, and \bottleScalar{2}{}),
+which constrains the descendant populations to be of
+constant size from present back to the divergence event.
+This allows the user to specify a model that has as many as
+seven parameters per population pair
+(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{},
+\bottleTime{}, \bottleScalar{1}{}, \bottleScalar{2}{}, and
+\migrationRate{})
+or as few as
+1 parameters per pair
+(\myTheta{}),
+in addition to the $\popSampleSize{i}{} - 1$ coalescence times.
+
+\subsection*{The ABC implementation of the model}
+
 This has the advantage of reducing the sample space of the number of
@@
 Second, the identity of the pairs of populations is lost.
-\end{linenomath}
-
-We implement two prior probability distributions over the joint space of
-\divTimeVector, \divTimeIndexVector.
-One simply places a uniform prior over all possible integer paritions of
-\npairs{}.
 

### 2013-09-25 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on dpp methods.

@@
 \begin{equation}
-    p(\divTimeIndexVector) = frac{1}{\integerPartitionNum{\npairs{}}},
+    p(\divTimeIndexVector) = \frac{1}{\integerPartitionNum{\npairs{}}},
     \label{eq:divModelPriorUniform}
@@
 \begin{equation}
-    p(\divTimeIndex{i} = \divTimeIndex{j}) = frac{1}{1 + \concentrationParam},
+    p(\divTimeIndex{i} = \divTimeIndex{j}) = \frac{1}{1 + \concentrationParam},
     \label{eq:dppPriorSameClass}
@@
 \end{equation}
-where \stirlingFirst{\cdot,\cdot} are the unsigned Stirling numbers of the
+where \stirlingFirst{\cdot}{\cdot} are the unsigned Stirling numbers of the
 first kind.
@@
 or as few as
-1 parameters per pair
+one parameter per pair
 (\myTheta{}),
@@
 By throwing away the identity of the \npairs{} pairs of populations, all
-information regarding the amount of data (e.g., the number gene copies
+information associating each pair with its amount of data (e.g., the number gene copies
 collected from each of the populations [\popSampleSize{1}{} and
-\popSampleSize{2}{}], the number of loci, and the length of the loci), the
+\popSampleSize{2}{}], the number of loci, and the length of the loci),
 information content in the data (i.e., the amount of sequence variation), as
-well as the taxon- and locus-specific parameters (\hkyModel{}{},
+well as its taxon- and locus-specific parameters (\hkyModel{}{},
 \mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
-\locusMutationRateScalar{}) for each pair is lost.
+\locusMutationRateScalar{}) is lost.
 The original descriptions of the \msb model claim that the independence

### 2013-09-25 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on methods of dpp ms.

@@
 \end{linenomath}
+
+\highLight{Need to discuss units of time and how they are scaled from units
+    proportional to time to units proportional to mutations.}
 
@@
 \subsection*{The ABC implementation of the model}
+To estimate the approximate posterior of Equation \ref{eq:approxModelCompact},
+we use an ABC rejection algorithm.
+The first step of this algorithm entails collecting a random sample of
+parameters values from the joint prior and their associated summary
+statistics.
+Each sample is generated by
+(1) drawing values of all the model's parameters \hpvector{} from their respective
+prior distributions,
+(2) rescaling the divergence times in \divTimeMapVector from units
+proportional to time to units proportional to the expected number
+of mutations via Equation X,
+(3) simulate a gene tree by drawing coalescent times from
+a multi-population Kingman-coalescent model,
+(4) simulating sequences along the gene tree under the
+\hky substitution parameters of \hkyModelVector,
+and (5) calculating population genetic summary statistics from the sequences
+\ssVector{} (these are the same statistics calculated from the observed
+alignments, \ssVectorObs{}).
 

### 2013-09-25 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on methods of dpp ms.
Also, adding more macros for model notation.

@@
 \nlociTotal loci,
-and \locusRateHetShapeParameter is the shape parameter of a gamma-distributed
-prior on \locusMutationRateScalar{}.
+\locusRateHetShapeParameter is the shape parameter of a gamma-distributed
+prior on \locusMutationRateScalar{}, and
+$p(\alignmentVector)$ is the probability of the data (or the marginal
+likelihood of the model).
 \end{linenomath}
@@
 
-\highLight{Need to discuss units of time and how they are scaled from units
-    proportional to time to units proportional to mutations.}
-
 \subsubsection*{Prior terms of Equation \ref{eq:fullModelCompact}}
@@
 at unique times.
-Regardless of the number of divergence time parameters (\divTimeNum), each
-\divTime{} within \divTimeVector is \iid as $\divTime{} \sim
-Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
 
 \begin{linenomath}
-The total sample space of ordered realizations of \divTimeMapVector is all the
+Each \divTime{} within \divTimeVector is \iid as $\divTime{} \sim
+Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
+Thus, given the number of unique divergence time classes in
+\divTimeIndexVector, this determines the probability of prior term
+$p(\divTimeVector \given \divTimeIndexVector)$
+The divergence time parameters are in coalescent units relative to the size of
+a constant reference population, which we denote \myTheta{C}, that is equal to
+the expectation of the prior on the size of the descendant populations
+\begin{equation}
+    \myTheta{C} = E(\descendantTheta{}{}).
+    \label{eq:thetaC}
+\end{equation}
+Given the size of the descendant populations are \iid as
+$\descendantTheta{}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+\gammaScale{\descendantTheta{}{}})$,
+this becomes
+\begin{equation}
+    \myTheta{C} = \gammaShape{\descendantTheta{}{}}\gammaScale{\descendantTheta{}{}}.
+    \label{eq:thetaCGamma}
+\end{equation}
+The \divTime parameters are then in units $\myTheta{C}/\mutationRate$
+generations, which we denote as \globalcoalunit generations.
+Thus, each \divTime within \divTimeVector is proportional to time and
+can be converted to the number of generations of the reference population,
+which we denote \divTime{G_C}, by assuming a mutation rate multiplying by
+the effective size of the reference population
+\begin{equation}
+    \divTime{G_C} = \divTime \times \frac{
+        \gammaShape{\descendantTheta{}{}}\gammaScale{\descendantTheta{}{}}}
+        {\mutationRate}.
+    \label{eq:divTimeGenerations}
+\end{equation}
+Thus, for each of the divergence times in \divTimeVector to be on the same
+scale, and thus comparable, the relative mutation rates among the pairs of
+populations are assumed to be fixed and known according the user-provided
+values in \mutationRateScalarConstantVector.
+\end{linenomath}
+
+\begin{linenomath}
+As described by \citet{Oaks2012}, to get the divergence times in units
+proportional to the expected number of mutations, \divTimeScaled{}{}, we must
+scale them by the realized population size for locus $j$ of population-pair $i$ 
+\begin{equation}
+    \divTimeScaled{i}{j} = \divTime{i} \times \frac{\myTheta{C}}{
+        \descendantThetaMean{i} \ploidyScalar{i}{j}},
+    \label{eq:divTimeScaled}
+\end{equation}
+where \descendantThetaMean{i} is the mean of \descendantTheta{1}{} and
+\descendantTheta{2}{} for pair $i$.
+This gives us the vector of scaled divergence times
+$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
+\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$.
+\end{linenomath}
+
+\begin{linenomath}
+As for the prior term $p(\divTimeMapVector)$,
+the total sample space of ordered realizations of \divTimeMapVector is all the
 possible partitions of \npairs{} elements into 1 to \npairs{} categories.
@@
 
+\subsubsection*{Time scale}
+As described above, in our model divergence times are in units of
+$\myTheta{C}/\mutationRate$ generations, where \myTheta{C} is the expectation
+of the prior on descendant-population size.
+As described by \citet{Oaks2012}, in \msb, \myTheta{C} is half of the upper
+limit of the continuous uniform prior on the mean of the descendant-population
+sizes.
+
 \subsection*{The ABC implementation of the model}

### 2013-09-26 manuscripts/dpp-msbayes/dpp-msbayes.tex
Working on ABC algorithm description in dpp ms.

@@
 with vectors of summary statistics for each alignment
-$\ssVector{} = (\alignmentSS{1}{1}, \ldots,
-    \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$.
+$\ssVectorObs = (\alignmentSSObs{1}{1}, \ldots,
+    \alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$.
+For each populationn pair, the means of the summary statistics are calculated
+across the \nloci{} loci, reducing the vector to
+$\ssVectorObs = (\alignmentSSObs{1}{}, \ldots,
+    \alignmentSSObs{\npairs{}}{})$.
 This allows us to estimate the approximate joint posterior distribution
@@
 
-\subsection*{The ABC implementation of the model}
+\subsection*{ABC estimation of the posterior}
 To estimate the approximate posterior of Equation \ref{eq:approxModelCompact},
@@
 Each sample is generated by
-(1) drawing values of all the model's parameters \hpvector{} from their respective
-prior distributions,
-(2) rescaling the divergence times in \divTimeMapVector from units
-proportional to time to units proportional to the expected number
-of mutations via Equation X,
-(3) simulate a gene tree by drawing coalescent times from
-a multi-population Kingman-coalescent model,
-(4) simulating sequences along the gene tree under the
-\hky substitution parameters of \hkyModelVector,
-and (5) calculating population genetic summary statistics from the sequences
-\ssVector{} (these are the same statistics calculated from the observed
-alignments, \ssVectorObs{}).
+(1) drawing values of all the model's parameters, which we denote \hpvector{}, from their respective
+prior distributions;
+(2) rescaling the divergence times
+$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
+from units proportional to time to units proportional to the expected number
+of mutations via Equation \ref{eq:divTimeScaled} to get 
+$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots, \divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$;
+(3) simulating gene trees $\geneTreeVector = (\geneTree{1}{1}, \ldots, 
+\geneTree{\npairs{}}{\nloci{\npairs{}}})$
+for each locus of each population pair by drawing coalescent times from
+a multi-population Kingman-coalescent model given the demographic parameters;
+(4) simulating sequences alignments 
+$\alignmentVector = (\alignment{1}{1}, \ldots, \alignment{\npairs{}}{\nloci{\npairs{}}})$
+along the gene trees under the
+\hky substitution parameters
+$\hkyModelVector = (\hkyModel{1}{1}, \ldots, \hkyModel{\npairs{}}{\nloci{\npairs{}}})$
+that have the same number of sequences and sequence lengths as the observed
+dataset;
+(5) calculating population genetic summary statistics
+$\ssVector{} = (\alignmentSS{1}{1}, \ldots, \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$
+from the simulated sequence alignments (these are the same statistics calculated from the observed
+alignments, \ssVectorObs{});
+and (6) reducing the summary statistics to the means across loci for each
+population pair to get
+$\ssVector{} = (\alignmentSS{1}{}, \ldots, \alignmentSS{\npairs{}}{})$, which is
+the same summary statstic vector estimated from the observed data \ssVectorObs.
+After repeating this procedure \numPriorSamples times, we obtain a random sample
+of parameters
+from the model prior and there associated summary statistics
+.
 

### 2013-09-26 manuscripts/reply/reply.tex
Making correction in reply.
Changing `\locusMutationRateScalar` to `\locusRateHetShapeParameter` in
the list of uniformly-distributed priors.
commit attempt 3

@@
     than a result of Lindley's paradox.
-    Here we show that their methodology suffered from a fundamental error that
+    Here we show that \cwlnote{their}{the} methodology \cwlnote{suffered}{of \citet{Hickerson2013}} suffers from a fundamental error that
     renders some of their results invalid and the remainder difficult to
     interpret.
-    Furthermore, we demonstrate how their approach still suffers from a bias
-    toward models with lesser parameter space, which manifests in a dangerously
+    Furthermore, we demonstrate how \cwlnote{their}{the} approach \cwlnote{}{of \citet{Hickerson2013}} still suffers from a bias
+    toward models with \cwlnote{lesser}{less} parameter space, which manifests in a dangerously
     strong tendency to sample predominantly from models that exclude the true
@@
 \newpage
-\noindent Recently, this journal has served as a medium for an interesting
+
+\highLight{The goal of this paper needs to be to highlight the major differences of opinion on how ABC can be used in comparative phylogeography and how these changes to msbayes can be made that will improve the inference in the future.
+We should avoid nit-picking the Hickerson paper and try to stay positive.}
+
+\noindent Recently, this journal has served as a medium for a
 discussion of the potential pitfalls of approximate Bayesian methods of
-comparative phylogeographical model choice.
+comparative phylogeographical model choice and prior selection.
 The discussion has centered around the approximate Bayesian model choice method,
@@
 clustered divergence times among taxon pairs.
-\citet{Hickerson2013} has published a response to this paper where they present
+\citet{Hickerson2013} published a response to this paper where they present
 an empirical Bayes model-averaging approach that they conclude circumvents the
 poor behavior of the method revealed by \citet{Oaks2012}.
-Both of these papers are largely in agreement.
-In fact, \citet{Hickerson2013} reiterate a lot of the discussion of
+These papers are largely in agreement with \citet{Hickerson2013} reiterating a lot of the discussion of
 \citet{Oaks2012} regarding the impact of broad uniform priors on Bayesian model
 choice.
-The main differences in their perspectives are centered around
+The main difference in perspectives are centered around
 (1) the mechanism by which broad uniform priors cause the poor behavior of
@@
 Note, this suggests the bias is extrinsic to \msb, and the numerical
-approximation machinery of the method could be sound.
-\citet{Hickerson2013} take a more pessimistic view of the bias, and suggest it
-is intrinsic to \msb.
-They conclude the method's rejection algorithm is inefficient and will be
+approximation machinery of \msb could be sound.
+\citet{Hickerson2013} suggest it is intrinsic to \msb, concluding the method's rejection algorithm is inefficient and will be
 increasingly biased as the overall space of the model increases, either as a
@@
 nuisance parameters.
-They support their position by giving a probabilistic argument that focuses on
+\citet{Hickerson2013} support their position by giving a probabilistic argument that focuses on
 only one of the three prior models used by \citet{Oaks2012}.
-We show that this argument is based on an empirical Bayesian perspective
-towards model choice and questionable assumptions, and does not explain the
+We show that the empirical Bayesian approach used by \citet{Hickerson2013} is based on questionable assumptions, and does not explain the
 bias of the method found in many of the analyses of \citet{Oaks2012}. 
-Furthermore, we reiterate several of the findings of \citet{Oaks2012} as well
-as present results of additional analyses, which strongly suggest that
-Lindley's paradox is playing a large role in the poor behavior of the method.
-
-As for how to mitigate the bias, \citet{Oaks2012} suggest that more cogent
-prior probability distributions on divergence models and nuisance parameters
-could reduce the effect of Lindley's paradox.
-They also layout a set of simulation-based procedures for determining power,
-accuracy, and robustness of the method given a dataset, and recommend any
-application of \msb should be accompanied by such procedures, especially if any
-biological conclusions are drawn from the results.
-\citet{Hickerson2013} present an approximate Bayesian model-averaging approach
-for accommodating uncertainty in selecting among empirically guided priors, and
-champion the method as a means of avoiding the pitfalls raised by
-\citet{Oaks2012}.
-Unfortunately, we show that fundamental errors in this approach render some of
-their results invalid and leave the remaining results difficult to interpret.
-Furthermore, we follow the advice of \citet{Oaks2012} and present
-simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
-approach.
-Our results demonstrate that the method is biased and dangerous.  The approach
-merely provides another means by which the model can ``escape'' large parameter
-space, and bias toward smaller models remains.
-This manifestation of the bias can be dangerous and often excludes the true
-parameter space due to the use of narrow uniform priors.
-Furthermore we discuss the potential theoretical and practical problems of
-empirical Bayesian model choice, which is used and encouraged by
-\citet{Hickerson2013} without any mention of its potential complications.
-
+Furthermore, we show that the model-averaging approach suggested by \citet{Hickerson2013} does not correct the bias towards small priors and we present results of additional analyses, which strongly suggest that
+Lindley's paradox is playing a large role in the poor behavior of \msb.
+
+%This paragraph seems too aggressive. It also seems to repeat some of the previous points. I removed it for now.
+
+%To mitigate the bias, \citet{Oaks2012} suggest that more cogent
+%prior probability distributions on divergence models and nuisance parameters
+%could reduce the effect of Lindley's paradox.
+%They also layout a set of simulation-based procedures for determining power,
+%accuracy, and robustness of the method given a dataset, and recommend any
+%application of \msb should be accompanied by such procedures, especially if any
+%biological conclusions are drawn from the results.
+%\citet{Hickerson2013} present an approximate Bayesian model-averaging approach
+%for accommodating uncertainty in selecting among empirically guided priors, and
+%champion the method as a means of avoiding the pitfalls raised by
+%\citet{Oaks2012}.
+%Unfortunately, we show that fundamental errors in this approach render some of
+%their results invalid and leave the remaining results difficult to interpret.
+%Furthermore, we follow the advice of \citet{Oaks2012} and present
+%simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
+%approach.
+%Our results demonstrate that the new method of \citet{Hickerson2013} is biased and dangerous.  The approach
+%merely provides another means by which the model can ``escape'' large parameter
+%space, and bias toward smaller models remains.
+%This manifestation of the bias can be dangerous and often excludes the true
+%parameter space due to the use of narrow uniform priors.
+%Furthermore we discuss the potential theoretical and practical problems of
+%empirical Bayesian model choice, which is used and encouraged by
+%\citet{Hickerson2013} without any mention of its potential complications.
+
+%It may be important to point out the error made in their paper, but it can also be nit-picky, it depends on how the error impacts the overall message of our response. 
+%We should strive to keep the response short and directed at the big picture items. If there is a way to still point out the error, then maybe some of this section should be used
+\iffalse
 \section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
@@
 time units indicated (Figure~\ref{figJointPosteriorHickerson}).
+
+\cwlnote{}{This part seems important, but not related to the error in units discussed above}
 Another important characteristic of the approach used by \citet{Hickerson2013}
@@
 parameters.
+\fi
 
@@
 
+\iffalse
 \section*{The dangers of empirical Bayesian model choice}
+
 \citet{Hickerson2013} repeatedly refer to the prior distributions used by
 \citet{Oaks2012} as ``poorly selected''. 
-However, this is bit misleading, as \citet{Oaks2012} discuss in detail how they
+However, this is misleading, as \citet{Oaks2012} discuss in detail how they
 selected their priors (see their section ``Specifying and simulating the joint
 prior'').
-They detail how being limited to uniform prior distributions in \msb
-necessitates awkward broad priors to avoid excluding the truth, a priori.
+They detail that the limitation in using uniform prior distributions in \msb
+necessitates awkward broad priors to avoid excluding the truth, \textit{a priori}.
 What \citet{Hickerson2013} really mean by ``poorly selected'' is that
@@
 prior sensitivity of the method and determining the power of \msb under
-real-world conditions with highly informative priors.  \citet{Oaks2012} discuss
+real-world conditions with highly informative priors.  
+\citet{Oaks2012} discuss
 the potential dangers of taking an empirical Bayesian approach to model choice
@@
 model choice.
-
-\subsection*{Theoretical implications of empirical priors for Bayesian model
+\fi
+
+\section*{Theoretical implications of empirical priors for Bayesian model
 choice}
 \begin{linenomath}
+
 Bayesian inference is a rational method of inductive learning in which Bayes'
@@
 values, we can define a prior distribution for all $\theta \in
-\allParameterValues$ such that $p(\theta)$ describes our belief that any given
+\allParameterValues$ such that $p(\theta)$ describes our belief \cwlnote{}{is belief the right term in this section? Would probability or likelihood be better?}
+ that any given
 \myTheta{} is the true value of the parameter.
@@
 This is not to say that empirical Bayesian approaches are not useful.
-Quite the contrary.
 Empirical Bayes is a well studied branch of Bayesian statistics that
-has given rise to a many powerful inference methods.
-It is true that empirical Bayes methods have an uncertain theoretical basis and
+has given rise to many powerful inference methods.
+Unfortunately, empirical Bayes methods have an uncertain theoretical basis and
 do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
@@
 estimation, a theoretical justification for empirical Bayes approaches to model
-choice is much more questionable.
+choice is unclear.
 In Bayesian model choice, the goal is not to estimate parameters, but to
@@
 
+%I like the example, but it may be too cute.
+It is also too long.
+I think a contrived example is good, but if there could be an example that is more relevant to model choice and comparison of inaccurate posterior distributions, I think that would be better.
+This is an example of empirical Bayes, but not the potential problem with the use in msbayes.
+Unless I am interpreting wrong.
+I would be careful about making it to cute as well.
+You may intend to make it easy to understand for a general audience, but it can also seem like talking down to Hickerson et al, who we are responding to.
+It just depends on how you read it.
+\iffalse
 This can be demonstrated with a simple, albeit contrived, example.
@@
 They can certainly evaluate this hypothesis, they can have undergraduate
-worker, Joe, flip the coin for them while the tabulate the results.
+worker, Joe, flip the coin for them while they tabulate the results.
 But being Bayesians, before they call Joe into the lab, they agree on a
@@
 from estimating the probability of a model.
-While empirical informed priors can be used to obtain well-behaved parameter
+While empirically informed priors can be used to obtain well-behaved parameter
 estimators, using them for model choice is much less certain.
@@
 been shown to be biased toward models with lesser parameter space.
-
-\subsection*{Practical dangers of narrow uniform priors for Bayesian model
-choice}
-We were alarmed when we saw the results of the model-averaging approach
-used by \citet{Hickerson2013}.
-Their results strongly favored the models with the narrowest, empirically
+\fi
+
+\subsection*{Dangers of narrow uniform priors for Bayesian model
+choice} \cwlnote{}{Should this be part of the empirical Bayes section, or something else?}
+
+\cwlnote{}{It seems like this section can be boiled down to ``empirical Bayes approaches for model choice when using uniform priors can exclude the $truth$ from the prior, which is a violation of Bayes theorem" or something like that.
+Picking apart what they calculated and we calculated seems like too much detail.
+I would try to simplify how this is being presented.
+This can lead into how the use of more appropriate prior distributions should be used.}
+The results of \citet{Hickerson2013} reanalysis of the Philippine dataset strongly favored models with the narrowest, empirically
 informed prior on divergence time, and thus their model-averaged posterior
@@
 \citet{Hickerson2013}).
-This concerned us because the narrowest \divt{} prior used by
+The narrowest \divt{} prior used by
 \citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
-divergence times for at least some of the Philippine taxa.
-They set this prior to match the 95\% highest posterior density (HPD) interval for
+divergence times for at least some of the Philippine taxa, a major problem when using uniform priors.
+\citet{Hickerson2013} set this prior to match the 95\% highest posterior density (HPD) interval for
 the mean divergence time estimated under one of the priors used by
@@
 Given this interval estimate is for the \emph{mean} divergence time across all
-22 taxa, it seems odd to set this as the prior, because many of the taxon pairs
+22 taxa, it may be inappropriate to set this as the prior, because many of the taxon pairs
 are expected to have diverged at times older than the upper limit.
@@
 see Table~6 of \citet{Oaks2012}).
-Thus, when we saw the results of \citet{Hickerson2013} we suspected that
-Lindley's paradox was again biasing the method toward models with smaller
-parameter space, and as a consequence was biasing the method toward models that
+Thus, the results of \citet{Hickerson2013} indicate
+Lindley's paradox is biasing the method toward models with smaller
+parameter space, and as a consequence is biasing the method toward models that
 exclude the truth (i.e., toward estimating model-averaged posteriors
@@
 
+%This should be in the section on model-averaging, not empirical Bayes limitations
 We explored the propensity of the model-averaging approach of
@@
 Second, we generated simulated datasets for which the divergence times are
-drawn form an exponential distribution and applied the approach of
+drawn from an exponential distribution and applied the approach of
 \citet{Hickerson2013} to each of them to see how often the method excludes the
@@
 Our results show that the model-averaging approach of \citet{Hickerson2013}
-does not reduce the method's bias toward models with lesser parameter space,
+does not reduce the method's bias toward models with less parameter space,
 but rather allows it to manifest in a very dangerous way.
@@
 investigator's choice of priors after having seen the resulting posterior.
-In a strictly Bayesian world, this is cheating.
-As discussed by \citet{Oaks2012}, prior to analysing the data, there was a large
+%In a strictly Bayesian world, this is cheating.
+As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a large
 amount of uncertainty regarding the divergence times of the 22 population pairs
 under study.
-Two the these pairs represent distinct species, and the taxonomy of many groups
+Two of these pairs represent distinct species, and the taxonomy of many groups
 in the Philippines has repeatedly been shown to mask deeply divergent lineages
 \citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
-Siler2011,Siler2012,RafeStuart2012}.
-\highLight{Any new stuff to add to this list of citations?}.
+Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013}.
 \citet{Oaks2012} also discuss how the sole use of uniform priors in \msb makes
@@
     \citet{Hickerson2013}}
+\cwlnote{}{I think the empirical use of the model-averaging should go here, instead of above in the empirical Bayes section}
+
 As recommended by \citet{Oaks2012}, we perform simulation-based analyses
 to explore the power of the approximate Bayesian model-averaging approach
-championed by \citet{Hickerson2013}.
+proposed by \citet{Hickerson2013}.
 Following \citet{Oaks2012}, we simulated 1000 pseudo-replicate datasets with
@@
 
-Given our results, we conclude that not only is the approach of
-\citet{Hickerson2013} dangerous in terms of excluding regions of parameter
-space containing the truth, but still lacks power to detect random variation in
+Given our results, we conclude that the approach of
+\citet{Hickerson2013} excludes regions of parameter
+space containing the $truth$, and lacks power to detect random variation in
 divergence times over the scale of $8\globalpopsize$ generations or more.
@@
 
+%I would suggest removing this section
+\iffalse
 \section*{The power analysis of \citet{Hickerson2013}}
@@
 does not seem particularly powerful.
-
+\fi
+
+%This seems to specific, not sure about including
 \section*{Validation analyses}
@@
 
-\citet{Hickerson2013} present an interesting probabilistic argument in an
-attempt to show that insufficient prior sampling is to blame for the bias.
+\citet{Hickerson2013} present an interesting probabilistic argument
+ to show that insufficient prior sampling is to blame for the bias.
 They argue the widest of the three priors used by \citet{Oaks2012} was very
@@
 
+%This seems to nit-picky. Probably not worth including it
 \section*{Other clarifications}
@@
 \msb can be strongly biased away from models with greater parameter space.
-We agree with \citet{Oaks2012} that this is likely caused by the inappropriate
+We find that this is likely caused by the inappropriate
 use of uniform priors on most of the model's parameters.
@@
 Exploring cogent prior probability distributions on most of the model's
-parameters, in addition to different priors over divergence models, could
+parameters, in addition to different priors over divergence models, should
 help mitigate the methods biases.

### 2013-09-26 manuscripts/dpp-msbayes/dpp-msbayes.tex
Finishing draft of dpp model and ABC description.
Starting on simulation methods.

@@
 
+\subsubsection*{Estimating ordered divergence models}
+As mentioned above, \msb samples over unordered divergence models
+(i.e., unordered partitions of the \npairs{} pairs of populations).
+That is, the identity of each population pair, and all the information
+associated with it, is discarded.
+In our implementation, inference can be done on either unordered or ordered
+models of divergence.
+This is discussed in more detail in the description of the ABC implementation
+below.
+
 \subsubsection*{The priors on nuisance parameters}
@@
 \subsection*{ABC estimation of the posterior}
+\subsubsection*{Sampling from the prior}
 To estimate the approximate posterior of Equation \ref{eq:approxModelCompact},
@@
 Each sample is generated by
-(1) drawing values of all the model's parameters, which we denote \hpvector{}, from their respective
-prior distributions;
+(1) drawing values of all the model's parameters, which we denote \hpvector{},
+from their respective prior distributions;
 (2) rescaling the divergence times
@@
 of mutations via Equation \ref{eq:divTimeScaled} to get 
-$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots, \divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$;
+$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
+\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$;
 (3) simulating gene trees $\geneTreeVector = (\geneTree{1}{1}, \ldots, 
@@
 $\ssVector{} = (\alignmentSS{1}{1}, \ldots, \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$
-from the simulated sequence alignments (these are the same statistics calculated from the observed
-alignments, \ssVectorObs{});
+from the simulated sequence alignments (these are the same statistics
+calculated from the observed alignments, \ssVectorObs{});
 and (6) reducing the summary statistics to the means across loci for each
@@
 $\ssVector{} = (\alignmentSS{1}{}, \ldots, \alignmentSS{\npairs{}}{})$, which is
-the same summary statstic vector estimated from the observed data \ssVectorObs.
-After repeating this procedure \numPriorSamples times, we obtain a random sample
-of parameters
-from the model prior and there associated summary statistics
-.
+the same summary statistic vector estimated from the observed data \ssVectorObs.
+After repeating this procedure \numPriorSamples times, we obtain a random
+sample of parameter vectors
+$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$
+from the model prior and there associated vectors of summary statistics
+$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$.
 
-This has the advantage of reducing the sample space of the number of
-divergence models, but has two disadvantages.
-First, additional information about the data is lost.
-By throwing away the identity of the \npairs{} pairs of populations, all
-information associating each pair with its amount of data (e.g., the number gene copies
-collected from each of the populations [\popSampleSize{1}{} and
-\popSampleSize{2}{}], the number of loci, and the length of the loci),
-information content in the data (i.e., the amount of sequence variation), as
-well as its taxon- and locus-specific parameters (\hkyModel{}{},
+\subsubsection*{Ordering of taxon-specific summary statistics}
+As eluded to in the model description, \msb does not maintain the order of the
+taxon-specific summary statistics \alignmentSS{}{} within each \ssVector{}.
+Rather, they, along with the observed summary statistics, are re-ordered by
+descending values of average pairwise differences between the descendant
+populations \citep[$\pi_b$;][]{NeiLi1979,Huang2011}.
+This has the advantage of reducing the sample space of the number of discrete
+divergence models \divTimeMapVector, but there are at least two disadvantages.
+First, additional information in the data is lost.
+By discarding the identity of the \npairs{} pairs of populations, all
+pair-specific information about the amount of data (e.g., the number gene
+copies collected from each of the populations [\popSampleSize{1}{} and
+\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
+taxon- and locus-specific parameters (\hkyModel{}{},
 \mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
 \locusMutationRateScalar{}) is lost.
-The original descriptions of the \msb model claim that the independence
-between the sample sizes \popSampleSize{1}{}, \popSampleSize{2}{} and
-the summary statistic $\pi_b$ calculated from the alignment for each
-population pair allows exchangeability.
-This is incorrect.
-The other summary statistics, and the alignments they are based on are
-not independent of the sample sizes.
-The alignments and summary statitics are not identically distributed because of
-taxon- and locus-specific parameters \hkyModel{}{},
-\mutationRateScalarConstant{}{}, and \ploidyScalar{}{}
-Second, the identity of the pairs of populations is lost.
+Second, the results are more difficult to interpret, because they can be no
+longer directly associated to the taxa under study.
+
+The original descriptions of the \msb model claim that this re-ordering is
+justified by the fact that of $\pi_b$ (and other summary statistics) are
+unrelated to the sample sizes \popSampleSize{1}{} and \popSampleSize{2}{} of
+each pair and are thus exchangeable \citet{Hickerson2006,Huang2011}.
+This is actually incorrect for two of reasons.
+First, the expectation of $\pi_b$ is not independent of samples sizes.
+If there are more than one coalescent events in the ancestor, which is expected
+to be common on phylogeographic timescales, more samples will increase the
+probability of capturing these deeper events, and thus affect the average
+pairwise differences between the descendant populations.
+Also, other statistics that estimate gross diversity (e.g., $\pi$ and
+$\theta_W$) are clearly not independent of sample size.
+Second, for variables to be exchangeable, they do not need to be independent,
+but their marginal distributions must be the same (i.e., they must be
+identically distributed).
+The simulated alignments and their summary statistics are not identically
+distributed because of differing sample sizes \emph{and} taxon- and
+locus-specific parameters \hkyModel{}{}, \mutationRateScalarConstant{}{}, and
+\ploidyScalar{}{}.
+
+Thus, the theoretical basis for this reshuffling is questionable.
+It may be justifiable as an additional coarsening of the data,
+by removing the information associated with the identity of the
+population pairs.
+However, we do not provide a formal proof that this re-ordering of the pairs
+does not introduce bias.
+One can imagine situations in which the sampling intensity (i.e., the
+number of gene copies, loci, and locus length) is highly skewed across
+pairs and/or there is large heterogeneity in mutational processes
+(e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
+In such cases it seems possible that discarding this information could
+cause bias.
+Furthermore, given that part of the motivation for re-ordering by $\pi_b$ is to
+minimize Euclidean distances between simulated datasets in which the true model
+has a single divergence event, and this reordering increases the model's
+tendency to infer a single divergence event \citep{Huang2011}, it seems
+possible that this approach could be biasing the method \citet{Oaks2012}.
+
+To maintain compatibility and comparability with \msb we maintain
+the re-ordering of taxon-specific summary statistics by $\pi_b$, but
+we also allow the order to be maintained.
+This allows the estimation of ordered divergence models.
+
+\subsubsection*{Obtaining an approximate posterior from the prior samples}
+We use a rejection algorithm to retain an approximate posterior sample of
+\paramSampleVector{} from the from the prior sample
+$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$.
+First, the observed summary statistics \ssVectorObs, and the summary statistics
+of the prior samples 
+$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$,
+are standardized using the means and standard deviations of the statistics from
+the prior sample (i.e., the prior mean is subtracted from each statistic, and then
+it is divided by the prior standard deviation).
+After all statistics are standardized, the Euclidean distance between
+the \ssVectorObs and each \ssVector{} within \ssMatrix is calculated.
+The samples that fall within a range of tolerance \tol around \ssVectorObs
+are retained.
+The range of tolerance is determined by specifying the number of desired
+posterior samples to be retained.
+Post-hoc adjustment of the posterior sample can also be performed with a number
+of regression techniques \citep{Beaumont2002,Blum2009,Leuenberger2010}.
+
+\subsection*{Simulation-based validation of the model}
 
@@
 
-\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape supporting information}}
+\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape Supporting Information}}
 

### 2013-09-26 manuscripts/reply/reply.tex
Incorporating Charles' `cwlnote`s into the reply.

@@
     than a result of Lindley's paradox.
-    Here we show that \cwlnote{their}{the} methodology \cwlnote{suffered}{of \citet{Hickerson2013}} suffers from a fundamental error that
-    renders some of their results invalid and the remainder difficult to
-    interpret.
-    Furthermore, we demonstrate how \cwlnote{their}{the} approach \cwlnote{}{of \citet{Hickerson2013}} still suffers from a bias
-    toward models with \cwlnote{lesser}{less} parameter space, which manifests in a dangerously
-    strong tendency to sample predominantly from models that exclude the true
-    values of the model's parameters.
+    Here we show that the methodology of \citet{Hickerson2013} suffers from a
+    fundamental error that renders some of their results invalid and the
+    remainder difficult to interpret.
+    Furthermore, we demonstrate how the approach of \citet{Hickerson2013} still
+    suffers from a bias toward models with less parameter space, which
+    manifests in a dangerously strong tendency to sample predominantly from
+    models that exclude the true values of the model's parameters.
     Our results suggest that their model-averaged results are dominated by
@@
 
-\highLight{The goal of this paper needs to be to highlight the major differences of opinion on how ABC can be used in comparative phylogeography and how these changes to msbayes can be made that will improve the inference in the future.
-We should avoid nit-picking the Hickerson paper and try to stay positive.}
+%CWL:
+% \highLight{The goal of this paper needs to be to highlight the major
+%     differences of opinion on how ABC can be used in comparative phylogeography
+%     and how these changes to msbayes can be made that will improve the
+%     inference in the future. We should avoid nit-picking the Hickerson paper
+%     and try to stay positive.}
 
@@
 poor behavior of the method revealed by \citet{Oaks2012}.
-These papers are largely in agreement with \citet{Hickerson2013} reiterating a lot of the discussion of
+These papers are largely in agreement with \citet{Hickerson2013} reiterating a
+lot of the discussion of
 \citet{Oaks2012} regarding the impact of broad uniform priors on Bayesian model
@@
 approximation machinery of \msb could be sound.
-\citet{Hickerson2013} suggest it is intrinsic to \msb, concluding the method's rejection algorithm is inefficient and will be
-increasingly biased as the overall space of the model increases, either as a
-function of the number of taxon pairs or the width of the uniform priors on
-nuisance parameters.
-\citet{Hickerson2013} support their position by giving a probabilistic argument that focuses on
-only one of the three prior models used by \citet{Oaks2012}.
-We show that the empirical Bayesian approach used by \citet{Hickerson2013} is based on questionable assumptions, and does not explain the
-bias of the method found in many of the analyses of \citet{Oaks2012}. 
-Furthermore, we show that the model-averaging approach suggested by \citet{Hickerson2013} does not correct the bias towards small priors and we present results of additional analyses, which strongly suggest that
-Lindley's paradox is playing a large role in the poor behavior of \msb.
+\citet{Hickerson2013} suggest the bias is intrinsic to \msb, concluding the
+method's rejection algorithm is inefficient and will be increasingly biased as
+the overall space of the model increases, either as a function of the number of
+taxon pairs or the width of the uniform priors on nuisance parameters.
+\citet{Hickerson2013} support their position by giving a probabilistic argument
+that focuses on only one of the three prior models used by \citet{Oaks2012}.
+We show that this argument is based on an empirical Bayesian approach and
+questionable assumptions, and does not explain the bias of the method found in
+many of the analyses of \citet{Oaks2012}. 
+Furthermore, we present results of additional analyses that, when combined
+with the results of \citet{Oaks2012}, strongly suggest that Lindley's paradox
+is playing a large role in the poor behavior of \msb.
 
@@
 
-%To mitigate the bias, \citet{Oaks2012} suggest that more cogent
-%prior probability distributions on divergence models and nuisance parameters
-%could reduce the effect of Lindley's paradox.
-%They also layout a set of simulation-based procedures for determining power,
-%accuracy, and robustness of the method given a dataset, and recommend any
-%application of \msb should be accompanied by such procedures, especially if any
-%biological conclusions are drawn from the results.
-%\citet{Hickerson2013} present an approximate Bayesian model-averaging approach
-%for accommodating uncertainty in selecting among empirically guided priors, and
-%champion the method as a means of avoiding the pitfalls raised by
-%\citet{Oaks2012}.
-%Unfortunately, we show that fundamental errors in this approach render some of
-%their results invalid and leave the remaining results difficult to interpret.
-%Furthermore, we follow the advice of \citet{Oaks2012} and present
-%simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
-%approach.
-%Our results demonstrate that the new method of \citet{Hickerson2013} is biased and dangerous.  The approach
-%merely provides another means by which the model can ``escape'' large parameter
-%space, and bias toward smaller models remains.
-%This manifestation of the bias can be dangerous and often excludes the true
-%parameter space due to the use of narrow uniform priors.
-%Furthermore we discuss the potential theoretical and practical problems of
-%empirical Bayesian model choice, which is used and encouraged by
-%\citet{Hickerson2013} without any mention of its potential complications.
-
-%It may be important to point out the error made in their paper, but it can also be nit-picky, it depends on how the error impacts the overall message of our response. 
-%We should strive to keep the response short and directed at the big picture items. If there is a way to still point out the error, then maybe some of this section should be used
-\iffalse
+Regarding perspectives about mitigating the bias, \citet{Oaks2012} suggest that
+alternative prior probability distributions on divergence models and nuisance
+parameters could reduce the effect of Lindley's paradox.
+They also layout simulation-based procedures for determining power, accuracy,
+and robustness of the method given a dataset, and recommend any application of
+\msb should be accompanied by such procedures, especially if any biological
+conclusions are drawn from the results.
+Alternatively, \citet{Hickerson2013} present an approximate Bayesian
+model-averaging approach for accommodating uncertainty in selecting among
+empirically guided priors, and champion the method as a means of avoiding the
+pitfalls raised by \citet{Oaks2012}.
+We show that fundamental errors in this approach render some of their results
+invalid and leave the remaining results difficult to interpret.
+Furthermore, we follow the advice of \citet{Oaks2012} and present
+simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
+approach.
+Our results demonstrate that the model-averaging approach of
+\citet{Hickerson2013} is biased and dangerous.
+The approach provides another means by which the model can ``escape'' large
+parameter space, and the method's bias toward smaller models remains.
+This manifestation of the bias can be dangerous and often excludes the true
+parameter space due to the use of narrow uniform priors.
+Furthermore we discuss the potential theoretical and practical problems of
+empirical Bayesian model choice, which is used and encouraged by
+\citet{Hickerson2013} without discussing its potential complications.
+
+%CWL:
+%It may be important to point out the error made in their paper, but it can
+%also be nit-picky, it depends on how the error impacts the overall message of
+%our response.  We should strive to keep the response short and directed at the
+%big picture items. If there is a way to still point out the error, then maybe
+%some of this section should be used
 \section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
@@
 
-\cwlnote{}{This part seems important, but not related to the error in units discussed above}
+% \cwlnote{}{This part seems important, but not related to the error in units
+% discussed above}
 Another important characteristic of the approach used by \citet{Hickerson2013}
 that makes their results difficult to interpret in addition to this error is
-their use of narrow, empirically guided uniform priors in an approximate
-Bayesian model choice framework.
+their use of narrow, empirically guided uniform priors in an
+approximate-Bayesian model-choice framework.
 We discuss the dangers of this approach in the next section, and show how the
@@
 parameters.
-\fi
 
@@
 
-\iffalse
 \section*{The dangers of empirical Bayesian model choice}
-
 \citet{Hickerson2013} repeatedly refer to the prior distributions used by
@@
 They detail that the limitation in using uniform prior distributions in \msb
-necessitates awkward broad priors to avoid excluding the truth, \textit{a priori}.
+necessitates awkward broad priors to avoid excluding the truth, \textit{a
+    priori}.
 What \citet{Hickerson2013} really mean by ``poorly selected'' is that
@@
 model choice.
-\fi
-
-\section*{Theoretical implications of empirical priors for Bayesian model
+
+\subsection*{Theoretical implications of empirical priors for Bayesian model
 choice}
 \begin{linenomath}
-
 Bayesian inference is a rational method of inductive learning in which Bayes'
@@
 values, we can define a prior distribution for all $\theta \in
-\allParameterValues$ such that $p(\theta)$ describes our belief \cwlnote{}{is belief the right term in this section? Would probability or likelihood be better?}
- that any given
+\allParameterValues$ such that $p(\theta)$ describes our belief that any given
 \myTheta{} is the true value of the parameter.
@@
 
-%I like the example, but it may be too cute.
-It is also too long.
-I think a contrived example is good, but if there could be an example that is more relevant to model choice and comparison of inaccurate posterior distributions, I think that would be better.
-This is an example of empirical Bayes, but not the potential problem with the use in msbayes.
-Unless I am interpreting wrong.
-I would be careful about making it to cute as well.
-You may intend to make it easy to understand for a general audience, but it can also seem like talking down to Hickerson et al, who we are responding to.
-It just depends on how you read it.
-\iffalse
+%CWL:
+%I like the example, but it may be too cute. It is also too long. I think a
+%contrived example is good, but if there could be an example that is more
+%relevant to model choice and comparison of inaccurate posterior distributions,
+%I think that would be better. This is an example of empirical Bayes, but not
+%the potential problem with the use in msbayes. Unless I am interpreting wrong.
+%I would be careful about making it to cute as well. You may intend to make it
+%easy to understand for a general audience, but it can also seem like talking
+%down to Hickerson et al, who we are responding to. It just depends on how you
+%read it.
 This can be demonstrated with a simple, albeit contrived, example.
@@
 been shown to be biased toward models with lesser parameter space.
-\fi
-
-\subsection*{Dangers of narrow uniform priors for Bayesian model
-choice} \cwlnote{}{Should this be part of the empirical Bayes section, or something else?}
-
-\cwlnote{}{It seems like this section can be boiled down to ``empirical Bayes approaches for model choice when using uniform priors can exclude the $truth$ from the prior, which is a violation of Bayes theorem" or something like that.
-Picking apart what they calculated and we calculated seems like too much detail.
-I would try to simplify how this is being presented.
-This can lead into how the use of more appropriate prior distributions should be used.}
-The results of \citet{Hickerson2013} reanalysis of the Philippine dataset strongly favored models with the narrowest, empirically
+
+\subsection*{Practical dangers of empirically informed uniform priors for
+    Bayesian model choice}
+% \cwlnote{}{Should this be part of the empirical Bayes section, or something
+%     else?}
+
+% \cwlnote{}{It seems like this section can be boiled down to ``empirical Bayes
+%     approaches for model choice when using uniform priors can exclude the
+%     $truth$ from the prior, which is a violation of Bayes theorem" or something
+%     like that. Picking apart what they calculated and we calculated seems like
+%     too much detail. I would try to simplify how this is being presented. This
+%     can lead into how the use of more appropriate prior distributions should be
+%     used.}
+The results of Hickerson et al.'s \citeyear{Hickerson2013} reanalysis of the
+Philippine dataset strongly favored models with the narrowest, empirically
 informed prior on divergence time, and thus their model-averaged posterior
@@
 \citet{Hickerson2013}).
-The narrowest \divt{} prior used by
+This is concerning, because the narrowest \divt{} prior used by
 \citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
-divergence times for at least some of the Philippine taxa, a major problem when using uniform priors.
-\citet{Hickerson2013} set this prior to match the 95\% highest posterior density (HPD) interval for
-the mean divergence time estimated under one of the priors used by
-\citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
+divergence times for at least some of the Philippine taxa, a major problem when
+using uniform priors.
+\citet{Hickerson2013} set this prior to match the 95\% highest posterior
+density (HPD) interval for the mean divergence time estimated under one of the
+priors used by \citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
 Given this interval estimate is for the \emph{mean} divergence time across all
-22 taxa, it may be inappropriate to set this as the prior, because many of the taxon pairs
-are expected to have diverged at times older than the upper limit.
+22 taxa, it may be inappropriate to set this as the prior, because many of the
+taxon pairs are expected to have diverged at times older than the upper limit.
 Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
@@
 see Table~6 of \citet{Oaks2012}).
-Thus, the results of \citet{Hickerson2013} indicate
-Lindley's paradox is biasing the method toward models with smaller
-parameter space, and as a consequence is biasing the method toward models that
-exclude the truth (i.e., toward estimating model-averaged posteriors
-dominated by models that exclude true values of parameters of the model).
-
-%This should be in the section on model-averaging, not empirical Bayes limitations
+Thus, the results of \citet{Hickerson2013} indicate Lindley's paradox is
+biasing the method toward models with smaller parameter space, and as a
+consequence is biasing the method toward models that exclude the truth (i.e.,
+toward estimating model-averaged posteriors dominated by models that exclude
+true values of parameters of the model).
+
+%CWL:This should be in the section on model-averaging, not empirical Bayes
+%limitations
 We explored the propensity of the model-averaging approach of
@@
     \citet{Hickerson2013}}
-\cwlnote{}{I think the empirical use of the model-averaging should go here, instead of above in the empirical Bayes section}
+% \cwlnote{}{I think the empirical use of the model-averaging should go here,
+%     instead of above in the empirical Bayes section}
 
@@
 
-Given our results, we conclude that the approach of
-\citet{Hickerson2013} excludes regions of parameter
-space containing the $truth$, and lacks power to detect random variation in
-divergence times over the scale of $8\globalpopsize$ generations or more.
+Given our results, we conclude that the approach of \citet{Hickerson2013}
+excludes regions of parameter space containing the truth, and lacks power to
+detect random variation in divergence times over the scale of $8\globalpopsize$
+generations or more.
 This roughly translates to millions of generations for mitochondrial loci, and
@@
 
-%I would suggest removing this section
-\iffalse
+%CWL: I would suggest removing this section
 \section*{The power analysis of \citet{Hickerson2013}}
@@
 does not seem particularly powerful.
-\fi
-
-%This seems to specific, not sure about including
+
+%CWL: This seems to specific, not sure about including
 \section*{Validation analyses}
@@
 \descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
-\bottleScalar{1}{}, \bottleScalar{2}{}, \locusMutationRateScalar,
+\bottleScalar{1}{}, \bottleScalar{2}{}, \locusRateHetShapeParameter,
 \migrationRate{}, \recombinationRate;][]{Oaks2012}.
@@
 
-%This seems to nit-picky. Probably not worth including it
+%CWL: This seems to nit-picky. Probably not worth including it
 \section*{Other clarifications}
@@
 \msb can be strongly biased away from models with greater parameter space.
-We find that this is likely caused by the inappropriate
+As suggested by \citet{Oaks2012}, this is likely caused by the inappropriate
 use of uniform priors on most of the model's parameters.
@@
 parameters, in addition to different priors over divergence models, should
-help mitigate the methods biases.
+help mitigate the method's biases.
 

### 2013-09-26 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on methods of dpp ms.

@@
 \begin{equation}
-    \divTimeScaled{i}{j} = \divTime{i} \times \frac{\myTheta{C}}{
+    \divTimeScaled{i}{j} = \divTimeMap{i} \times \frac{\myTheta{C}}{
         \descendantThetaMean{i} \ploidyScalar{i}{j}},
@@
 We denote this prior as
-$\divTimeIndexVector \sim DU(\integerPartitionSet{\npairs{}})$.
+$\divTimeIndexVector \sim \priorUniform$.
 \end{linenomath}
@@
 partitions of \npairs{} elements, which we denote as $\divTimeIndexVector \sim
-DP(\concentrationParam)$.
+\priorDPP{}$.
 We use a hierarchical Bayesian approach and place a second-order prior
@@
 parameters are represented at least once (see Equation 2 of \citet{Oaks2012}).
+We denote this prior used by \msb as $\divTimeIndexVector \sim \priorOld$.
 \citet{Oaks2012} discuss how placing a uniform prior over the number of
@@
 a single \myTheta{} parameter.
+
 We also allow the user to eliminate the parameters associated with
@@
 constant size from present back to the divergence event.
-This allows the user to specify a model that has as many as
-seven parameters per population pair
+Also, rather than eliminate the bottleneck parameters, the user
+can constrain \bottleScalar{1}{} and \bottleScalar{2}{} to be
+equal to remove one free parameter from the model.
+
+Overall, our implementation allows the user to specify a model that has as many
+as seven parameters per population pair
 (\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{},
@@
 \migrationRate{})
-or as few as
-one parameter per pair
+or as few as one parameter per pair
 (\myTheta{}),
-in addition to the $\popSampleSize{i}{} - 1$ coalescence times.
+in addition to the $\popSampleSize{i}{} - 1$ coalescence-time parameters.
 
@@
 We use a rejection algorithm to retain an approximate posterior sample of
-\paramSampleVector{} from the from the prior sample
+\paramSampleVector{} from the prior sample
 $\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$.
@@
 of regression techniques \citep{Beaumont2002,Blum2009,Leuenberger2010}.
+For our analyses, we use the general linear model (GLM) regression adjustment
+\citet{Leuenberger2010}, which \citet{Oaks2012} showed performed very
+similarly to weighted local-linear regression and multinomial logistic
+regression adjustments \citep{Beaumont2002} for \msb posteriors.
 
-\subsection*{Simulation-based validation of the model}
+\subsection*{Assessing model-choice behavior and robustness}
+Following the simulation-based approach of \citet{Oaks2012}, we characterize
+the behavior of several models under the ideal conditions where the
+data are generated from parameters drawn from the same prior distributions used
+for analysis (i.e., the prior is correct).
+We selected four model priors for these analyses (Table~\ref{tabPriors}):
+(1) A model to represent the original \msb model, \modelOld, with
+priors
+$\divTimeIndexVector \sim \priorOld$,
+$\divTime{} \sim U(0,10)$,
+$\ancestralTheta{} \sim U(0, 0.05)$,
+and
+$\descendantThetaMean{} \sim U(0, 0.05)$;
+(2) a Dirichlet-process prior model, \modelDPP, with priors
+$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
+$\divTime{} \sim Exp(0.3464)$,
+$\ancestralTheta{} \sim Exp(40)$,
+and
+$\descendantTheta{}{} \sim Exp(40)$;
+(3) a model with uniform prior probability over unordered divergence models,
+\modelUniform, with priors
+$\divTimeIndexVector \sim \priorUniform$,
+$\divTime{} \sim Exp(0.3464)$,
+$\ancestralTheta{} \sim Exp(40)$,
+and
+$\descendantTheta{}{} \sim Exp(40)$;
+and
+(4) a model with the \msb prior on unordered divergence models, but with
+exponential priors on nuisance parameters,
+\modelUshaped, with priors
+$\divTimeIndexVector \sim \priorOld$,
+$\divTime{} \sim Exp(0.3464)$,
+$\ancestralTheta{} \sim Exp(40)$,
+and
+$\descendantTheta{}{} \sim Exp(40)$.
+The exponential prior on divergence time used in models \modelDPP,
+\modelUniform, and \modelUshaped was selected to have the same variance as the
+uniform prior in model \modelOld.
+The exponential prior on population size used in models \modelDPP,
+\modelUniform, and \modelUniform was selected to have the same mean as the
+uniform prior in model \modelOld.
+All of the models were the same in other respects, with three free \myTheta{}
+parameters for each population pair, two uniformly distributed ($beta(1,1)$)
+\bottleScalar{}{} parameters per pair, no migration, no recombination,
+and unordered divergence models.
+For all of our simulations, we used a simulated data structure of eight
+population pairs, with a single 1000 base-pair locus sampled from 10
+individuals from each population.
+
+For each of the four models, we generated $1\e6$ samples from the prior, and
+simulated 50,000 pseudo-replicate datasets, also drawn from the prior.
+We then analyzed each of the replicate datasets using the, retaining a
+posterior sample from its respective prior.
+A GLM-regression adjusted posterior was also estimated from each of the
+posterior samples.
+To assess the robustness of each of the four models, we also analyzed the
+pseudo-replicate datasets simulated under the other three models.
+Overall, for each model, we produced 200,000 posterior estimates,
+50,000 from the replicate datasets simulated under that model,
+and 150,000 estimated from the replicate datasets simulated under the
+other three models.
+
+For each set of 50,000 simulated datasets, we used the posterior estimates
+assess each model's model-choice behavior.
+We did this by assigning the 50,000 estimates of the posterior probability
+of one divergence event to 20 bins of width 0.05, and plotted
+the estimated posterior probability of each bin against the proportion of
+replicates in that bin with a true value consistent with one divergence
+event \cite{Huelsenbeck2004,Oaks2012}.
+We did this using two criteria for the one divergence model:
+(1) the number of divergence time parameters ($\divTimeNum = 1$) and
+(2) the dispersion index of divergence times ($\divTimeDispersion < 0.01$).
+We used the one-divergence model to assess model-choice behavior, because
+it is often of biogeographic interest and is easily comparable among
+the three different priors used on divergence models.
+
+In addition to the four models above, we also assessed the behavior of a model
+with ordered divergence models (i.e., the order of the taxon-specific summary
+statistic vectors were maintained for the observed and simulated datasets).
+For this, we used a model with identical priors as \modelDPP, but that samples
+over ordered divergence models.
+We denote this model as \modelDPPOrdered.
+We generated $1\e6$ prior samples and 50,000 pseudo-replicate datasets, and
+analyzed them as above.
+We were not able to analyze the replicate datasets of the other models under
+the ordered model, because the identity of the population pairs is not
+contained in the simulations of the other models.
+
+\subsection*{Assessing power}
 
@@
 \singlespacing
+
+\begin{table}[htbp]
+    \sffamily
+    % \footnotesize
+    \addtolength{\tabcolsep}{-0.08cm}
+    \rowcolors{2}{}{myGray}
+    %\captionsetup{font=footnotesize}
+    \caption{The model priors evaluated in our simulation-based analyses.}
+    \centering
+    \begin{tabular}{ l l l l l }
+        \toprule
+        & \multicolumn{4}{c}{Priors} \\
+        \cmidrule(){2-5}
+        Model & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
+        \midrule
+            \modelOld & $\divTimeIndexVector \sim \priorOld$
+                      & $\divTime{} \sim U(0,10)$
+                      & $\ancestralTheta{} \sim U(0, 0.05)$
+                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
+            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$
+                      & $\divTime{} \sim Exp(0.3464)$
+                      & $\ancestralTheta{} \sim Exp(40)$
+                      & $\descendantTheta{}{} \sim Exp(40)$ \\
+            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
+                          & $\divTime{} \sim Exp(0.3464)$
+                          & $\ancestralTheta{} \sim Exp(40)$
+                          & $\descendantTheta{}{} \sim Exp(40)$ \\
+            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
+                          & $\divTime{} \sim Exp(0.3464)$
+                          & $\ancestralTheta{} \sim Exp(40)$
+                          & $\descendantTheta{}{} \sim Exp(40)$ \\
+        \bottomrule
+    \end{tabular}
+    \label{tabPriors}
+\end{table}
 

### 2013-09-27 manuscripts/dpp-msbayes/dpp-msbayes.tex
Finishing draft of methods for dpp ms; starting results.

@@
 $\descendantTheta{}{} \sim Exp(40)$.
-The exponential prior on divergence time used in models \modelDPP,
-\modelUniform, and \modelUshaped was selected to have the same variance as the
+We selected the exponential prior on divergence time used in models \modelDPP,
+\modelUniform, and \modelUshaped to have the same variance as the
 uniform prior in model \modelOld.
-The exponential prior on population size used in models \modelDPP,
-\modelUniform, and \modelUniform was selected to have the same mean as the
-uniform prior in model \modelOld.
+We selected the exponential prior on population size used in models \modelDPP,
+\modelUniform, and \modelUniform to have the same mean as the uniform prior in
+model \modelOld, so that all four models have the same \myTheta{C} and thus the
+same units of time.
 All of the models were the same in other respects, with three free \myTheta{}
@@
 A GLM-regression adjusted posterior was also estimated from each of the
-posterior samples.
+posterior samples \citep{Leuenberger2010}.
 To assess the robustness of each of the four models, we also analyzed the
@@
 \subsection*{Assessing power}
+We evaluated the power of the same four models (Table~\ref{tabPriors}) to
+detect random variation in divergence times using methods similar to
+\citet{Oaks2012}.
+For all of our power simulations, we used a simulated data structure identical
+to that of the empirical dataset of Philippine vertebrates analyzed by
+\citet{Oaks2012}, which consists of 22 pairs of populations.
+For each of the four models, we generated $2\e6$ samples from the prior.
+We then generated pseudo-observed datasets from three series of models.
+One series of models, which we denote \powerSeriesOld, is identically
+distributed as \modelOld except that the divergence times for each of the 22
+pairs of populations are randomly drawn from a series of uniform distributions,
+$U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and
+2.0, in \globalcoalunit generations.
+A second series of models, \powerSeriesUniform, are identically distributed as
+\modelUniform except that the 22 divergence times are randomly drawn from the
+same series of uniform priors as above.
+The third series of models, \powerSeriesUniformExp, is also identically
+distributed as \modelUniform except that the 22 divergence times are randomly
+drawn from a series of of exponential distributions: $Exp(17.3205)$,
+$Exp(8.6603)$, $Exp(5.7735)$, $Exp(4.33013)$, $Exp(3.4641)$, and $Exp(1.7321)$.
+These exponential distributions have the same variance as their uniform
+counterparts in the first two series of models.
+
+For each of the six models in each of the three series of models, we simulated
+1000 pseudo-replicate datasets (18,000 datasets in total).
+We then analyzed each replicate dataset under all four prior models
+(Table~\ref{tabPriors}), producing 72,000 posterior estimates, each with 1000
+samples.
+We also estimated a GLM-regression adjusted posterior from each of the
+posterior samples \citep{Leuenberger2010}.
+
+\section*{Results}
 

### 2013-09-27 manuscripts/utils/macros.tex
Adding figure macros to `macros.tex`.
More work on reply.
Adding `cwlnote` to macros.
Adding macros for new divergence time notation.
Adding more macros for new model notation.
Adding more combinatoric notation to latex macros.
More work on methods of dpp ms.
Also, adding more macros for model notation.
Finishing draft of dpp model and ABC description.
Starting on simulation methods.
Adding more latex macros for dpp notation.
Adding latex notation for power series model for dpp ms.
Starting to add validation accuracy plots to dpp ms.

@@
 
-\newcommand{\mthnote}[2]{{\color{red} #2}}
+\newcommand{\mthnote}[2]{{\color{red} #2}\xspace}
+\newcommand{\cwlnote}[2]{{\color{orange} #2}\xspace}
 
@@
 
+%% New notation for divergence times
+\newcommand{\divTime}[1]{\ensuremath{\tau_{#1}}\xspace}
+\newcommand{\divTimeVector}{\ensuremath{\boldsymbol{\divTime{}}}\xspace}
+\newcommand{\divTimeIndex}[1]{\ensuremath{t_{#1}}\xspace}
+\newcommand{\divTimeIndexVector}{\ensuremath{\mathbf{\divTimeIndex{}}}\xspace}
+\newcommand{\divTimeMap}[1]{\ensuremath{T_{#1}}\xspace}
+\newcommand{\divTimeMapVector}{\ensuremath{\mathbf{\divTimeMap{}}}\xspace}
+\newcommand{\divTimeScaled}[2]{\ensuremath{\mathcal{T}_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}\xspace}
+\newcommand{\divTimeScaledVector}{\ensuremath{\mathbf{\divTimeScaled{}{}}}\xspace}
+\newcommand{\divTimeMean}{\ensuremath{\bar{\divTimeMap{}}}\xspace}
+\newcommand{\divTimeVar}{\ensuremath{s^{2}_{\divTimeMap{}}}\xspace}
+\newcommand{\divTimeDispersion}{\ensuremath{D_{\divTimeMap{}}}\xspace}
+\newcommand{\divTimeNum}{\ensuremath{\lvert \divTimeVector \rvert}\xspace}
+\newcommand{\demographicParams}[1]{\ensuremath{\Theta_{#1}}\xspace}
+\newcommand{\demographicParamVector}{\ensuremath{\mathbf{\demographicParams{}}}\xspace}
+\newcommand{\popSampleSize}[2]{\ensuremath{n_{#1\protect\ifTwoArgs{#1}{#2}{,}#2}}}
+\newcommand{\gammaShape}[1]{\ensuremath{a_{#1}}\xspace}
+\newcommand{\gammaScale}[1]{\ensuremath{b_{#1}}\xspace}
+\newcommand{\betaA}[1]{\ensuremath{a_{#1}}\xspace}
+\newcommand{\betaB}[1]{\ensuremath{b_{#1}}\xspace}
+\newcommand{\integerPartitionSet}[1]{\ensuremath{a({#1})}\xspace}
+\newcommand{\integerPartitionNum}[1]{\ensuremath{\lvert \integerPartitionSet{#1} \rvert}\xspace}
+\newcommand{\concentrationParam}{\ensuremath{\chi}\xspace}
+\newcommand{\stirlingFirst}[2]{\ensuremath{c(#1, #2)}\xspace}
+\newcommand{\descendantThetaMean}[1]{\ensuremath{\bar{\theta}_{D\protect\ifArg{#1}{,}#1}}\xspace}
+\newcommand{\numPriorSamples}{\ensuremath{\mathbf{n}}\xspace}
+\newcommand{\paramSampleVector}[1]{\ensuremath{\Lambda_{#1}}\xspace}
+\newcommand{\paramSampleMatrix}{\ensuremath{\boldsymbol{\paramSampleVector{}}}\xspace}
+\newcommand{\modelDPP}{\ensuremath{M_{DPP}}\xspace}
+\newcommand{\modelDPPOrdered}{\ensuremath{M^{\circ}_{DPP}}\xspace}
+\newcommand{\modelUniform}{\ensuremath{M_{uniform}}\xspace}
+\newcommand{\modelUshaped}{\ensuremath{M_{Ushaped}}\xspace}
+\newcommand{\modelOld}{\ensuremath{M_{msBayes}}\xspace}
+\newcommand{\priorDPP}[1]{\ensuremath{DP(\concentrationParam #1)}\xspace}
+\newcommand{\priorUniform}{\ensuremath{DU\{\integerPartitionSet{\npairs{}}\}}\xspace}
+\newcommand{\priorOld}{\ensuremath{DU\{1, \ldots, \npairs{}\}}\xspace}
+\newcommand{\powerSeriesOld}{\ensuremath{\mathcal{M}_{msBayes}}\xspace}
+\newcommand{\powerSeriesUniform}{\ensuremath{\mathcal{M}_{uniform}}\xspace}
+\newcommand{\powerSeriesUniformExp}{\ensuremath{\mathcal{M}_{uniform-exp}}\xspace}
+
+\newcommand{\allDatasets}{\ensuremath{\mathcal{\alignment{}{}}}\xspace}
+\newcommand{\allParameterValues}{\ensuremath{\boldsymbol{\Theta}}\xspace}
 \newcommand{\bayesfactor}[2]{\ensuremath{BF_{#1\protect\ifArg{#2}{,}#2}}}
@@
 
+\newcommand{\validationAccuracyCaption}[2]{Estimation accuracy for model
+    #1 when analyzing data generated under #2.
+    A random sample of 5000 posterior estimates (from 50,000) are plotted,
+    including both (A, B, \& C) unadjusted and (D, E, \&
+    F) GLM-regression-adjusted estimates.
+    Normal random variates ($N(0, 0.005)$) have been added to the estimates and
+    true values of \divTimeNum (A \& D) to reduce overlap of plot symbols.
+    The root mean square error (RMSE) calculated from the 5000 estimates is
+    provided.}
 \newcommand{\simulationDescription}[2]{\change{Each plot represents #1
@@
 \newcommand{\tn}{\tabularnewline}
+
+\newcommand{\widthFigure}[5]{\begin{figure}[htbp]
+\begin{center}
+    \includegraphics[width=#1\textwidth]{#2}
+    \captionsetup{#3}
+    \caption{#4}
+    \label{#5}
+    \end{center}
+    \end{figure}}
+
+\newcommand{\heightFigure}[5]{\begin{figure}[htbp]
+\begin{center}
+    \includegraphics[height=#1]{#2}
+    \captionsetup{#3}
+    \caption{#4}
+    \label{#5}
+    \end{center}
+    \end{figure}}
+
+\newcommand{\mFigure}[3]{\widthFigure{1.0}{#1}{listformat=figList}{#2}{#3}\clearpage}
+\newcommand{\siFigure}[3]{\widthFigure{1.0}{#1}{name=Figure S, labelformat=noSpace, listformat=sFigList}{#2}{#3}\clearpage}

### 2013-09-27 validation/bin/plot-dpp-results.py
Updating validation plotting scripts.
Plots using new notation of dpp ms.

@@ -9,7 +9,11 @@ import project_util
 def main_cli():
     info_path = os.path.join(project_util.RESULT_DIR, 'dpp',
                 'pymsbayes-results', 'pymsbayes-info.txt')
-    plotting.plot_validation_results(info_path)
+    plotting.plot_validation_results(info_path,
+            omega_symbol = r'D_T',
+            psi_symbol = r'|\mathbf{\tau}|',
+            mean_time_symbol = r'\bar{T}',
+            math_font = None)
 
 if __name__ == '__main__':
     main_cli()

### 2013-09-27 validation/bin/plot-no-sort-results.py

@@ -9,7 +9,11 @@ import project_util
 def main_cli():
     info_path = os.path.join(project_util.NO_SORT_DIR, 'results',
                 'pymsbayes-results', 'pymsbayes-info.txt')
-    plotting.plot_validation_results(info_path)
+    plotting.plot_validation_results(info_path,
+            omega_symbol = r'D_T',
+            psi_symbol = r'|\mathbf{\tau}|',
+            mean_time_symbol = r'\bar{T}',
+            math_font = None)
 
 if __name__ == '__main__':
     main_cli()

### 2013-09-27 validation/bin/plot-old-results.py

@@ -9,7 +9,11 @@ import project_util
 def main_cli():
     info_path = os.path.join(project_util.RESULT_DIR, 'old',
                 'pymsbayes-results', 'pymsbayes-info.txt')
-    plotting.plot_validation_results(info_path)
+    plotting.plot_validation_results(info_path,
+            omega_symbol = r'D_T',
+            psi_symbol = r'|\mathbf{\tau}|',
+            mean_time_symbol = r'\bar{T}',
+            math_font = None)
 
 if __name__ == '__main__':
     main_cli()

### 2013-09-27 validation/bin/plot-u-shaped-results.py

@@ -9,7 +9,11 @@ import project_util
 def main_cli():
     info_path = os.path.join(project_util.RESULT_DIR, 'u-shaped',
                 'pymsbayes-results', 'pymsbayes-info.txt')
-    plotting.plot_validation_results(info_path)
+    plotting.plot_validation_results(info_path,
+            omega_symbol = r'D_T',
+            psi_symbol = r'|\mathbf{\tau}|',
+            mean_time_symbol = r'\bar{T}',
+            math_font = None)
 
 if __name__ == '__main__':
     main_cli()

### 2013-09-27 validation/bin/plot-uniform-results.py

@@ -9,7 +9,11 @@ import project_util
 def main_cli():
     info_path = os.path.join(project_util.RESULT_DIR, 'uniform',
                 'pymsbayes-results', 'pymsbayes-info.txt')
-    plotting.plot_validation_results(info_path)
+    plotting.plot_validation_results(info_path,
+            omega_symbol = r'D_T',
+            psi_symbol = r'|\mathbf{\tau}|',
+            mean_time_symbol = r'\bar{T}',
+            math_font = None)
 
 if __name__ == '__main__':
     main_cli()

### 2013-09-27 power-comparison/bin/plot-power-results.py
Updating plotting script for power results.
Adding new notation from dpp ms to plots.

@@ -134,7 +134,9 @@ def create_plots(info_path):
             psi_plot = PowerPlotGrid(
                     observed_config_to_estimates = cfg_to_psi,
                     variable = 'psi',
-                    num_columns = 2)
+                    variable_symbol = r'|\mathbf{\tau}|',
+                    num_columns = 2,
+                    margin_top = 0.975)
             fig = psi_plot.create_grid()
             fig.savefig(os.path.join(output_dir,
                     prefix + '_power_psi_mode.pdf'))
@@ -142,7 +144,9 @@ def create_plots(info_path):
             psi_plot_glm = PowerPlotGrid(
                     observed_config_to_estimates = cfg_to_psi_glm,
                     variable = 'psi',
-                    num_columns = 2)
+                    variable_symbol = r'|\mathbf{\tau}|',
+                    num_columns = 2,
+                    margin_top = 0.975)
             fig = psi_plot_glm.create_grid()
             fig.savefig(os.path.join(output_dir,
                     prefix + '_power_psi_mode_glm.pdf'))
@@ -150,6 +154,7 @@ def create_plots(info_path):
             psi_prob_plot = ProbabilityPowerPlotGrid(
                     observed_config_to_estimates = cfg_to_psi_prob,
                     variable = 'psi',
+                    variable_symbol = r'|\mathbf{\tau}|',
                     div_model_prior = div_model_prior,
                     dpp_concentration_mean = dpp_concentration_mean,
                     bayes_factor = 10,
@@ -161,6 +166,7 @@ def create_plots(info_path):
             psi_prob_plot_glm = ProbabilityPowerPlotGrid(
                     observed_config_to_estimates = cfg_to_psi_prob_glm,
                     variable = 'psi',
+                    variable_symbol = r'|\mathbf{\tau}|',
                     div_model_prior = div_model_prior,
                     dpp_concentration_mean = dpp_concentration_mean,
                     bayes_factor = 10,
@@ -172,6 +178,7 @@ def create_plots(info_path):
             omega_plot = PowerPlotGrid(
                     observed_config_to_estimates = cfg_to_omega,
                     variable = 'omega',
+                    variable_symbol = r'D_T',
                     num_columns = 2)
             fig = omega_plot.create_grid()
             fig.savefig(os.path.join(output_dir,
@@ -180,6 +187,7 @@ def create_plots(info_path):
             omega_plot_glm = PowerPlotGrid(
                     observed_config_to_estimates = cfg_to_omega_glm,
                     variable = 'omega',
[…]

### 2013-09-27 manuscripts/utils/macros.tex
Adding/updating figure caption macros.

@@
 \newcommand{\validationAccuracyCaption}[2]{Estimation accuracy for model
-    #1 when analyzing data generated under #2.
+    #2 when analyzing data generated under #1.
     A random sample of 5000 posterior estimates (from 50,000) are plotted,
@@
     provided.}
+\newcommand{\validationModelChoiceCaption}[2]{Model-choice accuracy for model
+    #2 when analyzing data generated under #1.
+    The estimated posterior probability of a single divergence event, based on
+    (A \& C) $\divTimeNum = 1$ and (B \& D) $\divTimeDispersion < 0.01$, from
+    50,000 posterior estimates are assigned to bins of width 0.05 and plotted
+    against the proportion of replicates in each bin where the truth is
+    $\divTimeNum = 1$ or $\divTimeDispersion < 0.01$.
+    Results based on the (A \& B) unadjusted and (C \& D) GLM-adjusted
+    posterior estimates are shown.}
 \newcommand{\simulationDescription}[2]{\change{Each plot represents #1

### 2013-09-27 manuscripts/dpp-msbayes/dpp-msbayes.tex
Starting to add validation accuracy plots to dpp ms.
Adding more figures to dpp ms.

@@
 
-%:FIGURE-accuracy omega SS GLM
-\begin{figure}[hbtp]
-    \captionsetup{listformat=figList}
-    \centering
-    \includegraphics[width=1.0\textwidth]{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}
-    \caption{
-        Model-choice behavior \ldots
-    }
-    \label{fig_dpp_dpp_val_mc}
-\end{figure}
+%% Validation accuracy
+\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
+    \validationAccuracyCaption{\modelDPP}{\modelDPP}
+}{figValidationAccuracyDPPDPP}
 
-\clearpage
+\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_accuracy.pdf}{
+    \validationAccuracyCaption{\modelDPP}{\modelUniform}
+}{figValidationAccuracyDPPUniform}
+
+\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_accuracy.pdf}{
+    \validationAccuracyCaption{\modelDPP}{\modelOld}
+}{figValidationAccuracyDPPOld}
+
+\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_accuracy.pdf}{
+    \validationAccuracyCaption{\modelDPP}{\modelUshaped}
+}{figValidationAccuracyDPPUshaped}
+
+\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_accuracy.pdf}{
+    \validationAccuracyCaption{\modelUniform}{\modelDPP}
+}{figValidationAccuracyUniformDPP}
+
+\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_accuracy.pdf}{
+    \validationAccuracyCaption{\modelUniform}{\modelUniform}
+}{figValidationAccuracyUniformUniform}
+
+\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_accuracy.pdf}{
+    \validationAccuracyCaption{\modelUniform}{\modelOld}
+}{figValidationAccuracyUniformOld}
+
+\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_accuracy.pdf}{
+    \validationAccuracyCaption{\modelUniform}{\modelUshaped}
+}{figValidationAccuracyUniformUshaped}
+
+\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_accuracy.pdf}{
+    \validationAccuracyCaption{\modelOld}{\modelDPP}
+}{figValidationAccuracyOldDPP}
+
+\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_accuracy.pdf}{
+    \validationAccuracyCaption{\modelOld}{\modelUniform}
+}{figValidationAccuracyOldUniform}
+
+\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_accuracy.pdf}{
+    \validationAccuracyCaption{\modelOld}{\modelOld}
+}{figValidationAccuracyOldOld}
+
+\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_accuracy.pdf}{
+    \validationAccuracyCaption{\modelOld}{\modelUshaped}
+}{figValidationAccuracyOldUshaped}
+
+\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_accuracy.pdf}{
+    \validationAccuracyCaption{\modelUshaped}{\modelDPP}
+}{figValidationAccuracyUshapedDPP}
+
+\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_accuracy.pdf}{
+    \validationAccuracyCaption{\modelUshaped}{\modelUniform}
+}{figValidationAccuracyUshapedUniform}
+
+\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_accuracy.pdf}{
+    \validationAccuracyCaption{\modelUshaped}{\modelOld}
+}{figValidationAccuracyUshapedOld}
+
+\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_accuracy.pdf}{
+    \validationAccuracyCaption{\modelUshaped}{\modelUshaped}
+}{figValidationAccuracyUshapedUshaped}
+
+%% Validation model choice
+\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelDPP}{\modelDPP}
+}{figValidationModelChoiceDPPDPP}
+
+\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelDPP}{\modelUniform}
+}{figValidationModelChoiceDPPUniform}
+
+\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelDPP}{\modelOld}
+}{figValidationModelChoiceDPPOld}
+
+\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelDPP}{\modelUshaped}
+}{figValidationModelChoiceDPPUshaped}
+
+\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelUniform}{\modelDPP}
+}{figValidationModelChoiceUniformDPP}
+
+\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelUniform}{\modelUniform}
+}{figValidationModelChoiceUniformUniform}
+
+\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelUniform}{\modelOld}
+}{figValidationModelChoiceUniformOld}
+
+\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelUniform}{\modelUshaped}
+}{figValidationModelChoiceUniformUshaped}
+
+\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelOld}{\modelDPP}
+}{figValidationModelChoiceOldDPP}
+
+\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelOld}{\modelUniform}
+}{figValidationModelChoiceOldUniform}
+
+\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelOld}{\modelOld}
+}{figValidationModelChoiceOldOld}
+
+\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelOld}{\modelUshaped}
+}{figValidationModelChoiceOldUshaped}
+
+\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelUshaped}{\modelDPP}
+}{figValidationModelChoiceUshapedDPP}
+
+\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelUshaped}{\modelUniform}
+}{figValidationModelChoiceUshapedUniform}
+
+\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelUshaped}{\modelOld}
+}{figValidationModelChoiceUshapedOld}
+
+\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelUshaped}{\modelUshaped}
+}{figValidationModelChoiceUshapedUshaped}
+
+%% Validation ordered models
+\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
+    \validationAccuracyCaption{\modelDPPOrdered}{\modelDPPOrdered}
+}{figValidationAccuracyDPPOrderedDPPOrdered}
+
+\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
+    \validationModelChoiceCaption{\modelDPPOrdered}{\modelDPPOrdered}
+}{figValidationModelChoiceDPPOrderedDPPOrdered}
 

### 2013-09-27 manuscripts/dpp-msbayes/dpp-msbayes.tex
Adding many more figures to dpp ms.

@@
 
+%% Power accuracy results
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesUniformExp}{\modelDPP}
+}{figPowerAccuracyUniformExpDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesUniformExp}{\modelUniform}
+}{figPowerAccuracyUniformExpUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesUniformExp}{\modelOld}
+}{figPowerAccuracyUniformExpOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesUniformExp}{\modelUshaped}
+}{figPowerAccuracyUniformExpUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesUniform}{\modelDPP}
+}{figPowerAccuracyUniformDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesUniform}{\modelUniform}
+}{figPowerAccuracyUniformUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesUniform}{\modelOld}
+}{figPowerAccuracyUniformOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesUniform}{\modelUshaped}
+}{figPowerAccuracyUniformUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesOld}{\modelDPP}
+}{figPowerAccuracyOldDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesOld}{\modelUniform}
+}{figPowerAccuracyOldUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesOld}{\modelOld}
+}{figPowerAccuracyOldOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_accuracy_omega_median.pdf}{
+    \powerAccuracyCaption{\powerSeriesOld}{\modelUshaped}
+}{figPowerAccuracyOldUshaped}
+
+%% Power psi results
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesUniformExp}{\modelDPP}
+}{figPowerPsiUniformExpDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesUniformExp}{\modelUniform}
+}{figPowerPsiUniformExpUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesUniformExp}{\modelOld}
+}{figPowerPsiUniformExpOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesUniformExp}{\modelUshaped}
+}{figPowerPsiUniformExpUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesUniform}{\modelDPP}
+}{figPowerPsiUniformDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesUniform}{\modelUniform}
+}{figPowerPsiUniformUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesUniform}{\modelOld}
+}{figPowerPsiUniformOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesUniform}{\modelUshaped}
+}{figPowerPsiUniformUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesOld}{\modelDPP}
+}{figPowerPsiOldDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesOld}{\modelUniform}
+}{figPowerPsiOldUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesOld}{\modelOld}
+}{figPowerPsiOldOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_mode.pdf}{
+    \powerPsiCaption{\powerSeriesOld}{\modelUshaped}
+}{figPowerPsiOldUshaped}
+
+%% Power dispersion results
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesUniformExp}{\modelDPP}
+}{figPowerDispersionUniformExpDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesUniformExp}{\modelUniform}
+}{figPowerDispersionUniformExpUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesUniformExp}{\modelOld}
+}{figPowerDispersionUniformExpOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesUniformExp}{\modelUshaped}
+}{figPowerDispersionUniformExpUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesUniform}{\modelDPP}
+}{figPowerDispersionUniformDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesUniform}{\modelUniform}
+}{figPowerDispersionUniformUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesUniform}{\modelOld}
+}{figPowerDispersionUniformOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesUniform}{\modelUshaped}
+}{figPowerDispersionUniformUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesOld}{\modelDPP}
+}{figPowerDispersionOldDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesOld}{\modelUniform}
+}{figPowerDispersionOldUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesOld}{\modelOld}
+}{figPowerDispersionOldOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_median.pdf}{
+    \powerDispersionCaption{\powerSeriesOld}{\modelUshaped}
+}{figPowerDispersionOldUshaped}
+
+%% Power psi probability results
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesUniformExp}{\modelDPP}
+}{figPowerPsiProbUniformExpDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesUniformExp}{\modelUniform}
+}{figPowerPsiProbUniformExpUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesUniformExp}{\modelOld}
+}{figPowerPsiProbUniformExpOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesUniformExp}{\modelUshaped}
+}{figPowerPsiProbUniformExpUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesUniform}{\modelDPP}
+}{figPowerPsiProbUniformDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesUniform}{\modelUniform}
+}{figPowerPsiProbUniformUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesUniform}{\modelOld}
+}{figPowerPsiProbUniformOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesUniform}{\modelUshaped}
+}{figPowerPsiProbUniformUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesOld}{\modelDPP}
+}{figPowerPsiProbOldDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesOld}{\modelUniform}
+}{figPowerPsiProbOldUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesOld}{\modelOld}
+}{figPowerPsiProbOldOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_prob.pdf}{
+    \powerPsiProbCaption{\powerSeriesOld}{\modelUshaped}
+}{figPowerPsiProbOldUshaped}
+
+%% Power dispersion prob results
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesUniformExp}{\modelDPP}
+}{figPowerDispersionProbUniformExpDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesUniformExp}{\modelUniform}
+}{figPowerDispersionProbUniformExpUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesUniformExp}{\modelOld}
+}{figPowerDispersionProbUniformExpOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesUniformExp}{\modelUshaped}
+}{figPowerDispersionProbUniformExpUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesUniform}{\modelDPP}
+}{figPowerDispersionProbUniformDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUniform}
+}{figPowerDispersionProbUniformUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesUniform}{\modelOld}
+}{figPowerDispersionProbUniformOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUshaped}
+}{figPowerDispersionProbUniformUshaped}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesOld}{\modelDPP}
+}{figPowerDispersionProbOldDPP}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesOld}{\modelUniform}
+}{figPowerDispersionProbOldUniform}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesOld}{\modelOld}
+}{figPowerDispersionProbOldOld}
+
+\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_prob.pdf}{
+    \powerDispersionProbCaption{\powerSeriesOld}{\modelUshaped}
+}{figPowerDispersionProbOldUshaped}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-09-27 manuscripts/utils/macros.tex
Adding many more figures to dpp ms.

--- 2013-09-27 manuscripts/utils/macros.tex
@@
     provided.}
+\newcommand{\validationModelChoiceCaption}[2]{Model-choice accuracy for model
+    #2 when analyzing data generated under #1.
+    The estimated posterior probability of a single divergence event, based on
+    (A \& C) $\divTimeNum = 1$ and (B \& D) $\divTimeDispersion < 0.01$, from
+    50,000 posterior estimates are assigned to bins of width 0.05 and plotted
+    against the proportion of replicates in each bin where the truth is
+    $\divTimeNum = 1$ or $\divTimeDispersion < 0.01$.
+    Results based on the (A \& B) unadjusted and (C \& D) GLM-adjusted
+    posterior estimates are shown.}
+\newcommand{\powerAccuracyCaption}[2]{Estimation accuracy for model
+    #2 when analyzing data generated under the series of models #1.
+    The true versus estimated value of the dispersion index of divergence
+    times (\divTimeDispersion) is plotted for 1000 datasets simulated
+    under each of the #1 models, and the proportion of estimates less
+    than the truth, $p(\hat{\divTimeDispersion} < \divTimeDispersion)$,
+    is shown for each data model.}
+\newcommand{\powerPsiCaption}[2]{
+    The power of model #2 to detect random variation in divergence times as
+    simulated under the series of models #1.
+    The plots illustrate the estimated number of divergence events
+    ($\hat{\divTimeNum}$) from analyses of 1000 datasets simulated under each
+    of the #1 models, with the the estimated probability of the model inferring
+    one divergence event, $p(\hat{\divTimeNum} = 1)$, given for each data
+    model.}
+\newcommand{\powerDispersionCaption}[2]{
+    The power of model #2 to detect random variation in divergence times as
+    simulated under the series of models #1.
+    The plots illustrate the estimated dispersion index of divergence times
+    ($\hat{\divTimeDispersion}$) from analyses of 1000 datasets simulated under
+    each of the #1 models, with the the estimated probability of the model
+    inferring one divergence event, $p(\hat{\divTimeDispersion} < 0.01)$, given
+    for each data model.}
+\newcommand{\powerPsiProbCaption}[2]{
+    The tendency of model #2 to support one divergence event when there is
+    random variation in divergence times as simulated under the series of
+    models #1.
+    The plots illustrate histograms of the estimated posterior probability of
+    the one divergence model, $p(\divTimeNum = 1 | \ssSpace)$, from analyses of
+    1000 datasets simulated under each of the #1 models, with the the estimated
+    probability of the model strongly supporting one divergence event,
+    $p(BF_{\divTimeNum = 1, \divTimeNum \neq 1} > 10)$, given for each data
+    model.}
+\newcommand{\powerDispersionProbCaption}[2]{
+    The tendency of model #2 to support one divergence event when there is
+    random variation in divergence times as simulated under the series of
+    models #1.
+    The plots illustrate histograms of the estimated posterior probability of
+    the one divergence model, $p(\divTimeDispersion < 0.01 | \ssSpace)$, from
+    analyses of 1000 datasets simulated under each of the #1 models, with the
+    the estimated probability of the model strongly supporting one divergence
+    event, $p(BF_{\divTimeDispersion < 0.01, \divTimeDispersion \geq 0.01} >
+    10)$, given for each data model.}
 \newcommand{\simulationDescription}[2]{\change{Each plot represents #1

### 2013-09-28 manuscripts/dpp-msbayes/dpp-msbayes.tex
Working on results of dpp ms.
Also adding table of models used to generated datasets for
power analyses.

@@
 For our analyses, we use the general linear model (GLM) regression adjustment
-\citet{Leuenberger2010}, which \citet{Oaks2012} showed performed very
+\citet{Leuenberger2010} as implemented in \abctoolbox
+\citep[v1.1;][]{ABCtoolbox}, which \citet{Oaks2012} showed performed very
 similarly to weighted local-linear regression and multinomial logistic
@@
 same series of uniform priors as above.
-The third series of models, \powerSeriesUniformExp, is also identically
+The third series of models, \powerSeriesExp, is also identically
 distributed as \modelUniform except that the 22 divergence times are randomly
@@
 \section*{Results}
+\subsection*{Validation analyses: Estimation accuracy}
+In terms of estimating summary statistics of model parameters (\divTimeNum,
+\divTimeDispersion, and \divTimeMean), the \modelDPP performs similarly when
+applied to datasets generated under all four models of Table \ref{tabPriors}
+(Figs.\
+\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP}).
+The same is true for models \modelUniform (Figs.\
+\labelcref{figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform})
+and \modelUshaped (Figs.\
+\labelcref{figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped}).
+All three of the models with exponential priors on nuisance parameters
+(\modelDPP, \modelUniform, and \modelUshaped) perform well when applied
+to the data generated under the original \msb model, \modelOld (Figs.\
+\labelcref{figValidationAccuracyOldDPP,figValidationAccuracyOldUniform,figValidationAccuracyOldUshaped}).
+The \modelOld model also performed well on its own datasets (Figure
+\ref{figValidationAccuracyOldOld}), but performed poorly when applied to the
+data generated under the other three models (Figs.\
+\labelcref{figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld}).
+Thus, the newly implemented models display greater robustness than
+the original model.
+
+Posterior adjustment with GLM regression does improve the estimation
+accuracy for the continuous variables \divTimeDispersion and \divTimeMean,
+however, this is not always true for estimating the discrete variable
+\divTimeNum (Figs.\
+\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP,figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform,figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped,figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld,figValidationAccuracyOldOld}).
+
+\subsection*{Validation analyses: Model-choice accuracy}
+The \msb model, and our modifications of it, is a model-choice method
+with the primary purpose of estimating the probabilities of models
+of divergence across taxa.
+Thus, it is critical to assess the method's ability to accurately
+estimate the posterior probabilities of divergence models.
+Our results demonstrate that the unadjusted estimates of
+divergence models are generally more accurate than the regression
+adjusted estimates, especially when using $\divTimeDispersion < 0.01$
+to assess the probability of the one divergence model (Figs.\
+\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
+Thus, we will focus our discussion of the results on the unadjusted estimates.
+
+The model-choice behavior of the \modelDPP is relatively accurate when applied
+to datasets generated under models \modelDPP, \modelUniform, and \modelOld
+(Figs.\
+\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP}).
+It is less accurate and tends to underestimate the probability of the
+one-divergence model when applied to the data generated under \modelUshaped
+(Figure \ref{figValidationModelChoiceUshapedDPP}).
+The \modelUniform model performs almost identically as \modelDPP in
+terms of model-choice accuracy (Figs.\
+\labelcref{figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}).
 
@@
 
+\begin{table}[htbp]
+    \sffamily
+    % \footnotesize
+    % \addtolength{\tabcolsep}{-0.08cm}
+    \rowcolors{2}{}{myGray}
+    %\captionsetup{font=footnotesize}
+    \caption{The models we used to simulate pseudo-replicate datasets for
+        assessing the power of the models in Table \ref{tabPriors}.}
+    \centering
+    \begin{tabular}{ l l l l l }
+        \toprule
+        & \multicolumn{4}{c}{Priors} \\
+        \cmidrule(){2-5}
+        Model series & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
+        \midrule
+            \powerSeriesOld & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,0.2)$
+                            & $\ancestralTheta{} \sim U(0, 0.05)$
+                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,0.4)$
+                            & $\ancestralTheta{} \sim U(0, 0.05)$
+                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,0.6)$
+                            & $\ancestralTheta{} \sim U(0, 0.05)$
+                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,0.8)$
+                            & $\ancestralTheta{} \sim U(0, 0.05)$
+                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,1.0)$
+                            & $\ancestralTheta{} \sim U(0, 0.05)$
+                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,2.0)$
+                            & $\ancestralTheta{} \sim U(0, 0.05)$
+                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
+        \midrule
+            \powerSeriesUniform & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,0.2)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,0.4)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,0.6)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,0.8)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,1.0)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim U(0,2.0)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+        \midrule
+            \powerSeriesExp & $\divTimeNum = 22$
+                            & $\divTime{} \sim Exp(17.3205)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim Exp(8.6603)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim Exp(5.7735)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim Exp(4.3301)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim Exp(3.4641)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTimeNum = 22$
+                            & $\divTime{} \sim Exp(1.7321)$
+                            & $\ancestralTheta{} \sim Exp(40)$
+                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+        \bottomrule
+    \end{tabular}
+    \label{tabPowerModels}
+\end{table}
+
+\clearpage
+
 %% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniformExp}{\modelDPP}
-}{figPowerAccuracyUniformExpDPP}
+    \powerAccuracyCaption{\powerSeriesExp}{\modelDPP}
+}{figPowerAccuracyExpDPP}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniformExp}{\modelUniform}
-}{figPowerAccuracyUniformExpUniform}
+    \powerAccuracyCaption{\powerSeriesExp}{\modelUniform}
+}{figPowerAccuracyExpUniform}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniformExp}{\modelOld}
-}{figPowerAccuracyUniformExpOld}
+    \powerAccuracyCaption{\powerSeriesExp}{\modelOld}
+}{figPowerAccuracyExpOld}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniformExp}{\modelUshaped}
-}{figPowerAccuracyUniformExpUshaped}
+    \powerAccuracyCaption{\powerSeriesExp}{\modelUshaped}
+}{figPowerAccuracyExpUshaped}
 
@@
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniformExp}{\modelDPP}
-}{figPowerPsiUniformExpDPP}
+    \powerPsiCaption{\powerSeriesExp}{\modelDPP}
+}{figPowerPsiExpDPP}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniformExp}{\modelUniform}
-}{figPowerPsiUniformExpUniform}
+    \powerPsiCaption{\powerSeriesExp}{\modelUniform}
+}{figPowerPsiExpUniform}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniformExp}{\modelOld}
-}{figPowerPsiUniformExpOld}
+    \powerPsiCaption{\powerSeriesExp}{\modelOld}
+}{figPowerPsiExpOld}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniformExp}{\modelUshaped}
-}{figPowerPsiUniformExpUshaped}
+    \powerPsiCaption{\powerSeriesExp}{\modelUshaped}
+}{figPowerPsiExpUshaped}
 
@@
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniformExp}{\modelDPP}
-}{figPowerDispersionUniformExpDPP}
+    \powerDispersionCaption{\powerSeriesExp}{\modelDPP}
+}{figPowerDispersionExpDPP}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniformExp}{\modelUniform}
-}{figPowerDispersionUniformExpUniform}
+    \powerDispersionCaption{\powerSeriesExp}{\modelUniform}
+}{figPowerDispersionExpUniform}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniformExp}{\modelOld}
-}{figPowerDispersionUniformExpOld}
+    \powerDispersionCaption{\powerSeriesExp}{\modelOld}
+}{figPowerDispersionExpOld}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniformExp}{\modelUshaped}
-}{figPowerDispersionUniformExpUshaped}
+    \powerDispersionCaption{\powerSeriesExp}{\modelUshaped}
+}{figPowerDispersionExpUshaped}
 
@@
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniformExp}{\modelDPP}
-}{figPowerPsiProbUniformExpDPP}
+    \powerPsiProbCaption{\powerSeriesExp}{\modelDPP}
+}{figPowerPsiProbExpDPP}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniformExp}{\modelUniform}
-}{figPowerPsiProbUniformExpUniform}
+    \powerPsiProbCaption{\powerSeriesExp}{\modelUniform}
+}{figPowerPsiProbExpUniform}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniformExp}{\modelOld}
-}{figPowerPsiProbUniformExpOld}
+    \powerPsiProbCaption{\powerSeriesExp}{\modelOld}
+}{figPowerPsiProbExpOld}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniformExp}{\modelUshaped}
-}{figPowerPsiProbUniformExpUshaped}
+    \powerPsiProbCaption{\powerSeriesExp}{\modelUshaped}
+}{figPowerPsiProbExpUshaped}
 
@@
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniformExp}{\modelDPP}
-}{figPowerDispersionProbUniformExpDPP}
+    \powerDispersionProbCaption{\powerSeriesExp}{\modelDPP}
+}{figPowerDispersionProbExpDPP}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniformExp}{\modelUniform}
-}{figPowerDispersionProbUniformExpUniform}
+    \powerDispersionProbCaption{\powerSeriesExp}{\modelUniform}
+}{figPowerDispersionProbExpUniform}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniformExp}{\modelOld}
-}{figPowerDispersionProbUniformExpOld}
+    \powerDispersionProbCaption{\powerSeriesExp}{\modelOld}
+}{figPowerDispersionProbExpOld}
 
 \mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniformExp}{\modelUshaped}
-}{figPowerDispersionProbUniformExpUshaped}
+    \powerDispersionProbCaption{\powerSeriesExp}{\modelUshaped}
+}{figPowerDispersionProbExpUshaped}
 

### 2013-09-28 manuscripts/dpp-msbayes/dpp-msbayes.tex
Working on results of dpp ms.

@@
 \labelcref{figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}).
+
+The original \msb model \modelOld performs reasonably well when applied
+to its own datasets (Figure \ref{figValidationModelChoiceOldOld}), but
+it performs poorly when applied to data generated under any of the other
+models (Figs.\
+\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
+In these cases the \modelOld model is strongly biased towards over-estimating
+the posterior probability of the one-divergence model, especially when the
+divergence models are not distributed under its U-shaped prior (Figs.\
+\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld}).
+The other model with the U-shaped prior on divergence models, but exponential
+priors on nuisance parameters (\modelUshaped), performs similarly to \modelOld
+in that it performs well when applied to its own data, but overestimates
+the probability of the one-divergence model when applied to data generated
+by the other models (Figs.\
+\labelcref{figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceUshapedUshaped}).
+However, the bias toward over-estimating the probability of the one-divergence
+model is stronger in the \modelOld model than \modelUshaped.
+
+Overall, the result suggest that the \modelDPP and \modelUniform models are
+relatively robust in terms of model-choice accuracy, and when model violations
+do cause them to be biased, they tend to under-estimate the probability of
+the biogeographically interesting model of a single divergence event.
+In contrast, the \modelOld model is very sensitive to model violations,
+and strongly over-estimates the one-divergence model whenever the model
+is misspecified.
+Furthermore, the results suggest that using exponentially distributed priors on
+nuisance parameters rather than uniform priors helps the \modelUshaped model
+perform better than \modelOld, but it is still hindered by the Ushaped prior on
+divergence models and tends to over-estimate the probability of the
+one-divergence model whenever there are violations of the mode.
+
+\subsection*{Validation analyses: Ordered divergence models}
+Our results show that the model estimating ordered models of divergence
+\modelDPPOrdered performs similarly to its counterpart sampling over
+unordered models \modelDPP when the prior is correct (Figs.\
+\labelcref{figValidationAccuracyDPPOrderedDPPOrdered,figValidationModelChoiceDPPOrderedDPPOrdered}).
 

### 2013-09-28 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on results of dpp ms.

@@
 \labelcref{figValidationAccuracyDPPOrderedDPPOrdered,figValidationModelChoiceDPPOrderedDPPOrdered}).
+
+\subsection*{Power analyses: Estimation accuracy}
+All of the models we evaluated (Table \ref{tabPriors}) struggled to estimate
+the dispersion index of divergence times \divTimeDispersion regardless of which
+of the three series of models (Table \ref{tabPowerModels}) the data were
+generated under (Figs.\
+\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform,figPowerAccuracyExpOld,figPowerAccuracyExpUshaped,figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped,figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
+When the divergence times of the 22 population pairs are randomly drawn from a
+series of exponential priors (\powerSeriesExp), the \modelDPP model is the
+best estimator of \divTimeDispersion, followed by \modelUshaped (figs.\
+\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform}
+The \modelOld model is strongly biased toward underestimating
+\divTimeDispersion, estimating values of zero for most of the replicates across
+all the data models of \powerSeriesExp (Figure \ref{figPowerAccuracyExpOld}).
+The results of the \modelUshaped model are intermediate between those of
+\modelOld and the better performing models \modelDPP and \modelUniform
+(Figure \ref{figPowerAccuracyExpUshaped}).
+
+When the true divergence times are randomly drawn from a series of uniform
+priors (\powerSeriesUniform), all of the models poorly estimate
+\divTimeDispersion (Figs.\
+\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped}).
+The \modelDPP and \modelUniform models tend to over-estimate the variance in
+divergence times (Figs.\
+\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform}), whereas
+the \modelOld model, again, underestimates \divTimeDispersion, estimating
+values of zero for most replicates across all the data models of
+\powerSeriesUniform (Figure \ref{figPowerAccuracyUniformOld}).
+Again, the performance of the \modelUshaped model is intermediate between the
+\modelOld and \modelDPP/\modelUniform models (Figure
+\ref{figPowerAccuracyUniformUshaped}).
+This pattern of results across the four models is very similar when applied
+to data simulated under the series of models of \powerSeriesOld (Figs.\
+\labelcref{figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
+
+\subsection*{Power analyses: model-choice}
 

### 2013-09-28 manuscripts/dpp-msbayes/dpp-msbayes.tex
Finishing draft of results for dpp ms.

@@
 \subsection*{Power analyses: model-choice}
+There is a large amount of heterogeneity in estimating the number of divergence
+events (\divTimeNum) both among the prior models, and across the different
+series of data models (Table \ref{tabPowerModels}) (Figs.\
+\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped}).
+The \modelDPP model tends to infer highly clustered divergence models across
+all three series of data models when the divergences are recent, but performs
+the best of the four prior models  when divergences are random over a timescale
+of about one to two coalescent units (Figs.\
+\labelcref{figPowerPsiExpDPP,figPowerPsiUniformDPP,figPowerPsiOldDPP,}).
+The \modelUniform model performs the best when divergences are random
+over the most recent timescales, but tends to only infer a tight
+cluster of \divTimeNum values even when divergences are random
+over longer periods (Figs.\
+\labelcref{figPowerPsiExpUniform,figPowerPsiUniformUniform,figPowerPsiOldUniform,}).
+The \modelOld model performs the worst of the four prior models across
+all three series of data models, inferring a single divergence event across
+most of the 18,000 simulations (Figs.\
+\labelcref{figPowerPsiExpOld,figPowerPsiUniformOld,figPowerPsiOldOld,}).
+Even when the simulated data are identically distributed as the \modelOld model
+except for the divergence times, the \modelOld is still strongly biased towards
+a single event, whereas the \modelDPP model is the least likely to infer a
+single event when applied to these data.
+Using exponential priors on nuisance parameters does increase the power
+of the \modelUshaped model compared \modelOld across all three series
+of data models, but the U-shaped prior still prevents the model
+from performing as well as the \modelDPP and \modelUniform models (Figs.\
+\labelcref{figPowerPsiExpUshaped,figPowerPsiUniformUshaped,figPowerPsiOldUshaped,}).
+
+The improved power of the new models are even more pronounced when looking at
+estimates of the dispersion index of divergence times (\divTimeDispersion)
+across the simulations (Figs.\
+\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped}).
+The \modelDPP and \modelUniform models perform similarly across all three
+series of data models, inferring values of \divTimeDispersion consistent with
+one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
+replicates across all the simulations (Figs.\
+\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionOldDPP,figPowerDispersionOldUniform}).
+The \modelOld model infers values consistent with a single divergence event in
+most of the replicates across all the simulations (Figs.\
+\labelcref{figPowerDispersionExpOld,figPowerDispersionUniformOld,figPowerDispersionOldOld}).
+Using exponential priors on nuisance parameters greatly increases the power
+of the \modelUshaped model relative to \modelOld, but it still has
+less power than the models with DPP or uniform priors across
+divergence models (Figs.\
+\labelcref{figPowerDispersionExpUshaped,figPowerDispersionUniformUshaped,figPowerDispersionOldUshaped}).
+
+When looking at the estimated posterior probability of the single-divergence
+model across the power analyses, we also see the increased power of the
+new models (Figs.\
+\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped}).
+The \modelDPP model estimates low posterior probability of
+$\divTimeNum = 1$ across all of the simulations (Figs.\
+\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbUniformDPP,figPowerPsiProbOldDPP}).
+The \modelDPP does infer strong support for one divergence as gauged by a Bayes
+factor of greater than 10.
+However, this is not that unexpected given the low prior probability of a
+single divergence under this model.
+Also, the Bayes factor calculations for the \modelDPP replicates are only
+approximate; they are calculated base on the prior mean of the concentration
+parameter \concentrationParam.
+
+The \modelUniform model infers very low posterior probabilities for the
+one-divergence model across all the simulations (Figs.\
+\labelcref{figPowerPsiProbExpUniform,figPowerPsiProbUniformUniform,figPowerPsiProbOldUniform}).
+The\modelOld model infers high posterior probabilities of a single divergence
+for most replicates across all simulations (Figs.\
+\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld}).
+Again, we see that the exponential priors on nuisance parameters greatly
+increase power and result in lower estimates of the probability of
+one divergence when comparing \modelUshaped to \modelOld (Figs.\
+\labelcref{figPowerPsiProbExpUshaped,figPowerPsiProbUniformUshaped,figPowerPsiProbOldUshaped}).
+
+Lastly, when we look at the estimated posterior probability of
+\divTimeDispersion being consistent with one divergence parameter
+($p(\divTimeDispersion < 0.01 | \ssSpace)$), we see the same pattern of model
+behavior, with \modelDPP and \modelUniform inferring low probabilities across
+all simulations, \modelOld inferring high probabilities, and \modelUshaped
+inferring intermediate values (Figs.\
+\labelcref{figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
 

### 2013-09-29 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on dpp ms.

@@
 
+\section*{Introduction}
+Being able to accurately infer the temporal distribution of divergences
+across co-distributed taxa
+
+Understanding evolutionary history is super cool.
+Adding a comparative framework extra super cool.
+Shared evolutionary history across taxa is often of interest:
+processes of diversification and community assembly,
+host-symbiont dynamics
+
 \section*{Methods}
@@
 \subsection*{The model}
+In this section, we describe our model, which is a modification
+of the model implemented in \msb \citep{Huang2011,Oaks2012}.
 We follow much of the notation of \citet{Oaks2012}, but modify it to
-accommodate the new model \ldots
+aid in our description of the new model.
+A summary of our notation can be found in Table \ref{tabNotation}.
 
@@
 $\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
-is a vector of population divergence times for each of \npairs{} pairs,
+is a vector of population divergence times for each of the \npairs{} pairs of
+populations,
 $\demographicParamVector = (\demographicParams{1}, \ldots,
@@
     \alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$.
-For each populationn pair, the means of the summary statistics are calculated
+For each population pair, the means of the summary statistics are calculated
 across the \nloci{} loci, reducing the vector to
@@
 summary statistics, the radius of which is the tolerance \tol.
-The approximation of this posterior is caused by the insufficiency of the
-statistics and the \tol being greater than zero.
+The sources of approximation are the insufficiency of the statistics and the
+\tol being greater than zero.
 We describe the full model in detail before delving into the numerical
@@
 is the probability of the sequence alignment of locus $j$ for population pair
-$i$ given the gene tree and \hky substitution model paramters \citep[i.e., the
+$i$ given the gene tree and \hky substitution model parameters \citep[i.e., the
 ``Felsenstein likelhood'';][]{Felsenstein1981}.
@@
 The term $p(\locusRateHetShapeParameter)$ is the prior density
-function for the shape-parameter for the gamma-distributed prior on
+function for the shape-parameter of the gamma-distributed prior on
 rate heterogeneity among loci.
@@
 \divTimeNum can range from 1 to \npairs{},
-and \divTimeIndexVector is vector of indices of divergence time parameters 
+and \divTimeIndexVector is the vector of indices of divergence time parameters 
 $(\divTimeIndex{1}, \ldots, \divTimeIndex{\npairs{}})$
-that map the divergence time in \divTimeVector to each of the \npairs{}
+that map the divergence times in \divTimeVector to each of the \npairs{}
 pairs of populations.
@@
 \divTimeIndexVector, this determines the probability of prior term
-$p(\divTimeVector \given \divTimeIndexVector)$
+$p(\divTimeVector \given \divTimeIndexVector)$.
 The divergence time parameters are in coalescent units relative to the size of
@@
 \end{equation}
-The \divTime parameters are then in units $\myTheta{C}/\mutationRate$
+The \divTime{} parameters are then in units $\myTheta{C}/\mutationRate$
 generations, which we denote as \globalcoalunit generations.
-Thus, each \divTime within \divTimeVector is proportional to time and
-can be converted to the number of generations of the reference population,
-which we denote \divTime{G_C}, by assuming a mutation rate multiplying by
-the effective size of the reference population
+Thus, each \divTime{} within \divTimeVector is proportional to time and can be
+converted to the number of generations of the reference population, which we
+denote \divTime{G_C}, by assuming a mutation rate and multiplying by the
+effective size of the reference population
 \begin{equation}
@@
 As described by \citet{Oaks2012}, to get the divergence times in units
-proportional to the expected number of mutations, \divTimeScaled{}{}, we must
+proportional to the expected number of mutations, we must
 scale them by the realized population size for locus $j$ of population-pair $i$ 
@@
 \begin{linenomath}
-As for the prior term $p(\divTimeMapVector)$,
-the total sample space of ordered realizations of \divTimeMapVector is all the
-possible partitions of \npairs{} elements into 1 to \npairs{} categories.
+As for the prior term $p(\divTimeIndexVector)$, the total sample space of
+ordered realizations of \divTimeIndexVector is all the possible partitions of
+\npairs{} elements into 1 to \npairs{} categories.
 The total number of possible partitions is a sum of the Stirling numbers of
-the second kind over all possible number of categories \divTimeNum
+the second kind over all possible numbers of categories \divTimeNum
 \begin{equation}
@@
 \priorDPP{}$.
-We use a hierarchical Bayesian approach and place a second-order prior
-probability distribution (i.e., hyperprior) on \concentrationParam.  More
-specifically, we use a gamma-distributed prior $\concentrationParam =
+We take a hierarchical approach and use a second-order prior probability
+distribution (i.e., hyperprior) for \concentrationParam.
+More specifically, we use a gamma-distributed prior $\concentrationParam =
 Gamma(\gammaShape{\concentrationParam}, \gammaScale{\concentrationParam})$,
@@
 One of the key differences between our model and that of \msb \citep{Huang2011}
-is the prior distribution placed on divergence models.
+is the prior distribution on divergence models.
 As discussed in \citet{Oaks2012}, in \msb the only prior used for
@@
 multinomial distribution on the number of times each index of \divTimeVector
-appears in \divTimeIndexVector, with the constraint that all \divTime
+appears in \divTimeIndexVector, with the constraint that all \divTime{}
 parameters are represented at least once (see Equation 2 of \citet{Oaks2012}).
@@
 \citet{Oaks2012} discuss how placing a uniform prior over the number of
-divergence parameters (\divTimeNum here; \numt{} in \citet{Huang2011}) imposes
-an odd U-shaped prior over discrete divergence models (\divTimeIndexVector; see
+divergence parameters (denoted \divTimeNum here, and as \numt{} in
+\citet{Huang2011}) imposes an odd U-shaped prior over discrete divergence
+models (\divTimeIndexVector; see
 Figure 5B of \citet{Oaks2012}).
@@
 Following the recommendations of \citet{Oaks2012}, we have replaced the use of
-inappropriate, continuous uniform distributions for priors on many of the
+continuous uniform distributions for priors on many of the
 model's parameters (\divTime{}, \ancestralTheta{}, \descendantTheta{1}{},
@@
 Not only do these priors have nice conjugacy properties, but they also provide
-the user with much greater flexibility in expressing the prior uncertainty
+the user with much greater flexibility in expressing prior uncertainty
 regarding the parameters of the model.
@@
 can constrain \bottleScalar{1}{} and \bottleScalar{2}{} to be
-equal to remove one free parameter from the model.
+equal to remove one free parameter from the model for each of
+the population pairs.
 
@@
 As described by \citet{Oaks2012}, in \msb, \myTheta{C} is half of the upper
-limit of the continuous uniform prior on the mean of the descendant-population
+limit of the continuous uniform prior on the mean of the descendant population
 sizes.
 
-\subsection*{ABC estimation of the posterior}
+\subsection*{ABC estimation of the posterior of the model}
 \subsubsection*{Sampling from the prior}
@@
 a multi-population Kingman-coalescent model given the demographic parameters;
-(4) simulating sequences alignments 
+(4) simulating sequence alignments 
 $\alignmentVector = (\alignment{1}{1}, \ldots, \alignment{\npairs{}}{\nloci{\npairs{}}})$
@@
 $\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$
-from the model prior and there associated vectors of summary statistics
+from the model prior and their associated vectors of summary statistics
 $\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$.
+
+For all of our analyses below, we use four summary statistics for each of
+the pairs of populations:
+$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
+\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
+Furthermore, in addition to model parameters, each sample \hpvector{}
+also contains four statistics that summarize \divTimeMapVector:
+the mean (\divTimeMean), variance (\divTimeVar), dispersion index
+($\divTimeDispersion = \divTimeVar/\divTimeMean$), and the number
+of divergence time parameters (\divTimeNum).
+These have been denoted as \meant{}{}, \vart{}{}, \vmratio{}, and \numt{},
+respectively \citep{Hickerson2006,Huang2011,Oaks2012}.
+We use the new notation to avoid confusion that the values represent prior or
+posterior expectations, and to indicate that \vmratio{} is not a parameter of
+the model and \numt{} is no longer a parameter in the new implementation.
 
@@
 Second, the results are more difficult to interpret, because they can be no
-longer directly associated to the taxa under study.
+longer be directly associated to the taxa under study.
 
@@
 unrelated to the sample sizes \popSampleSize{1}{} and \popSampleSize{2}{} of
-each pair and are thus exchangeable \citet{Hickerson2006,Huang2011}.
+each pair and are thus exchangeable \citep{Hickerson2006,Huang2011}.
 This is actually incorrect for two of reasons.
@@
 tendency to infer a single divergence event \citep{Huang2011}, it seems
-possible that this approach could be biasing the method \citet{Oaks2012}.
+possible that this approach could be biasing the method \citep{Oaks2012}.
 
 To maintain compatibility and comparability with \msb we maintain
-the re-ordering of taxon-specific summary statistics by $\pi_b$, but
-we also allow the order to be maintained.
-This allows the estimation of ordered divergence models.
+the re-ordering of taxon-specific summary statistics by $\pi_b$.
+However, we also allow the order to be maintained, and ordered
+divergence models to be estimated.
 
@@
 After all statistics are standardized, the Euclidean distance between
-the \ssVectorObs and each \ssVector{} within \ssMatrix is calculated.
+\ssVectorObs and each \ssVector{} within \ssMatrix is calculated.
 The samples that fall within a range of tolerance \tol around \ssVectorObs
@@
 For our analyses, we use the general linear model (GLM) regression adjustment
-\citet{Leuenberger2010} as implemented in \abctoolbox
+\citep{Leuenberger2010} as implemented in \abctoolbox
 \citep[v1.1;][]{ABCtoolbox}, which \citet{Oaks2012} showed performed very
@@
 We selected the exponential prior on divergence time used in models \modelDPP,
-\modelUniform, and \modelUshaped to have the same variance as the
-uniform prior in model \modelOld.
+\modelUniform, and \modelUshaped to have the same variance as the uniform prior
+in model \modelOld.
 We selected the exponential prior on population size used in models \modelDPP,
@@
 simulated 50,000 pseudo-replicate datasets, also drawn from the prior.
-We then analyzed each of the replicate datasets using the, retaining a
-posterior sample from its respective prior.
+We then analyzed each of the replicate datasets, retaining a
+posterior of 1000 samples from the respective prior.
 A GLM-regression adjusted posterior was also estimated from each of the
@@
 For each set of 50,000 simulated datasets, we used the posterior estimates
-assess each model's model-choice behavior.
+to assess the model-choice behavior of each model.
 We did this by assigning the 50,000 estimates of the posterior probability
@@
 replicates in that bin with a true value consistent with one divergence
-event \cite{Huelsenbeck2004,Oaks2012}.
+event \citep{Huelsenbeck2004,Oaks2012}.
 We did this using two criteria for the one divergence model:
 (1) the number of divergence time parameters ($\divTimeNum = 1$) and
-(2) the dispersion index of divergence times ($\divTimeDispersion < 0.01$).
+(2) the dispersion index of divergence times \citep[$\divTimeDispersion <
+0.01$][]{Hickerson2006,Huang2011}.
 We used the one-divergence model to assess model-choice behavior, because
@@
 \subsection*{Validation analyses: Model-choice accuracy}
-The \msb model, and our modifications of it, is a model-choice method
+The \msb model, and our modification of it, is a model-choice method
 with the primary purpose of estimating the probabilities of models
@@
 estimate the posterior probabilities of divergence models.
-Our results demonstrate that the unadjusted estimates of
-divergence models are generally more accurate than the regression
-adjusted estimates, especially when using $\divTimeDispersion < 0.01$
-to assess the probability of the one divergence model (Figs.\
+Our results demonstrate that the unadjusted estimates of divergence models are
+generally more accurate than the regression-adjusted estimates, especially when
+using $\divTimeDispersion < 0.01$ to assess the probability of the one
+divergence model (Figs.\
 \labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
@@
 nuisance parameters rather than uniform priors helps the \modelUshaped model
-perform better than \modelOld, but it is still hindered by the Ushaped prior on
-divergence models and tends to over-estimate the probability of the
-one-divergence model whenever there are violations of the mode.
+perform better than \modelOld, but it is still hindered by the U-shaped prior
+on divergence models and tends to over-estimate the probability of the
+one-divergence model whenever there are violations of the model.
 
@@
 \modelDPPOrdered performs similarly to its counterpart sampling over
-unordered models \modelDPP when the prior is correct (Figs.\
+unordered models (\modelDPP) when the prior is correct (Figs.\
 \labelcref{figValidationAccuracyDPPOrderedDPPOrdered,figValidationModelChoiceDPPOrderedDPPOrdered}).
@@
 \subsection*{Power analyses: Estimation accuracy}
-All of the models we evaluated (Table \ref{tabPriors}) struggled to estimate
-the dispersion index of divergence times \divTimeDispersion regardless of which
-of the three series of models (Table \ref{tabPowerModels}) the data were
-generated under (Figs.\
+All of the models we evaluated (Table \ref{tabPriors}) struggle to estimate the
+dispersion index of divergence times \divTimeDispersion regardless of which of
+the three series of models (Table \ref{tabPowerModels}) the data were generated
+under (Figs.\
 \labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform,figPowerAccuracyExpOld,figPowerAccuracyExpUshaped,figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped,figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
@@
 series of exponential priors (\powerSeriesExp), the \modelDPP model is the
-best estimator of \divTimeDispersion, followed by \modelUshaped (figs.\
-\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform}
+best estimator of \divTimeDispersion, followed by \modelUniform (figs.\
+\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform}).
 The \modelOld model is strongly biased toward underestimating
@@
 The results of the \modelUshaped model are intermediate between those of
-\modelOld and the better performing models \modelDPP and \modelUniform
+\modelOld and the new models \modelDPP and \modelUniform
 (Figure \ref{figPowerAccuracyExpUshaped}).
@@
 
-\subsection*{Power analyses: model-choice}
+\subsection*{Power analyses: Model choice}
 There is a large amount of heterogeneity in estimating the number of divergence
@@
 Even when the simulated data are identically distributed as the \modelOld model
-except for the divergence times, the \modelOld is still strongly biased towards
-a single event, whereas the \modelDPP model is the least likely to infer a
-single event when applied to these data.
+(except for the divergence times) the \modelOld is still strongly biased
+towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
+model is the least likely to infer a single event when applied to these data
+(Figure \ref{figPowerPsiOldDPP}).
 Using exponential priors on nuisance parameters does increase the power
-of the \modelUshaped model compared \modelOld across all three series
+of the \modelUshaped model compared to \modelOld across all three series
 of data models, but the U-shaped prior still prevents the model
@@
 Also, the Bayes factor calculations for the \modelDPP replicates are only
-approximate; they are calculated base on the prior mean of the concentration
+approximate; they are calculated based on the prior mean of the concentration
 parameter \concentrationParam.
@@
 one-divergence model across all the simulations (Figs.\
-\labelcref{figPowerPsiProbExpUniform,figPowerPsiProbUniformUniform,figPowerPsiProbOldUniform}).
-The\modelOld model infers high posterior probabilities of a single divergence
-for most replicates across all simulations (Figs.\
+\labelcref{figPowerPsiProbExpUniform,figPowerPsiProbUniformUniform,figPowerPsiProbOldUniform}),
+whereas the \modelOld model infers high posterior probabilities of a single
+divergence for most replicates across all simulations (Figs.\
 \labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld}).
@@
 \labelcref{figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
+
+\section*{Discussion}
+Our results demonstrate that using alternative priors on nuisance parameters
+and divergence models improved the behavior of the \msb model.
+The model-choice estimation of our implementation is more accurate and,
+perhaps more importantly, shows greater robustness to model
+violations (Figs.\
+\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
+The original model is very sensitive to model violations and, when present,
+strongly over-estimates the probability of one divergence event shared
+across all taxa (Figs.\
+\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
+When more appropriate priors are used for nuisance parameters, and either
+a Dirichlet-process or uniform prior applied across divergence models,
+the model is less sensitive to violations (Figs.\
+\labelcref{figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}),
+and, when they do cause bias, the method is conservative and tends to
+under-estimate the probability of biogeographically interesting models of
+temporally clustered divergences 
+\labelcref{figValidationModelChoiceUshapedDPP,figValidationModelChoiceUshapedUniform}).
+
+Furthermore, the modifications to the model also improve the method's power to
+detect random variation in divergence times, reducing the bias toward
+estimating clustered divergences (Figs.\
+\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped,figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped,figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped,figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
+Our results are similar to those of \citet{Oaks2012} in that we also find
+the \msb model to infer strong support for clustered divergences when
+divergences are random over quite broad timescales.
+\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld,figPowerDispersionProbExpOld,figPowerDispersionProbUniformOld,figPowerDispersionProbOldOld}).
+Our results expand on this by showing that this behavior is consistent
+across of range conditions underlying the data.
+When using the new priors, the method is less prone to spurious inference of
+clustered divergence models, and much less likely to incorrectly infer such
+models with strong support.
+
+By evaluating a model intermediate between the old and new implementation
+(\modelUshaped), we were able to determine the relative effects of the
+modifications to model.
+Across all of our analyses, the results show that using better priors on
+nuisance parameters alone does improve the performance of the method.
+The magnitude of the bias toward inferring support for the one-divergence
+model when there are model violations is reduced when the exponential
+priors are used in place of the uniform priors.
+\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped}).
+Furthermore, better priors on on nuisance parameters improves the method's
+power to detect temporally random divergences.
+Throughout our analyses, the intermediate model (\modelUshaped) performs
+better than the \msb model, but not as well as the models with alternative
+priors on divergence models.
+This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
+models of temporally clustered divergences is caused by a combination of
+(1) uniform priors on nuisance parameters hindering the marginal likelihoods of
+models with more \divTime{} parameters and
+(2) the U-shaped prior on divergence models giving low prior density to models
+with intermediate numbers of divergence parameters.
+
+While our modifications improve the behavior of the model, we urge
+caution when using the method and interpreting its results.
+The method attempts to approximate the posterior of a very parameter rich model
+using little information from the data.
+For example, when applied to the dataset of Philippine vertebrates analyzed by
+\citet{Oaks2012} the model can have as many as 604--625 free parameters
+(depending on \divTimeNum), and samples over 1002 divergence models.
+Our modifications allow this model to be simplified to 471--492 free
+parameters.
+Sampling over hundreds of parameters and more than a thousand models is asking
+a lot from a few summary statistics calculated from the sequence alignments of
+each taxon pair.
+These simple summary statistics contain little information about many
+of the free parameters in the model.
+Thus, we expect the improved method will still be sensitive to priors, and the
+power, while improved, may still be low.
+Furthermore, ABC methods in general have been shown to be biased
+\citep{Robert2011} for model choice.
+As recommended by \citet{Oaks2012}, any results from the method should be
+accompanied by assessments of prior sensitivity and power.
+We encourage investigators to view this method and its predecessor as a means
+of exploring their data rather than as rigorous means of evaluating hypotheses.
+
+Given the difficulty of this estimation problem, we anticipate that
+full-likelihood methods that can leverage all of the information present in the
+sequence data will become increasingly important in robustly estimating the
+temporal history of divergences across taxa \citep{JeetDiss}.
+With improving numerical methods for sampling over models of differing
+dimensionality \citep{Green1995,Lemey2009} and increasing efficiency of
+likelihood calculations \citep{Ayres2012}, analyzing rich comparative
+phylogeograpical models in full-likelihood Bayesian framework is becoming
+computationally practical, especially when considering that the computational
+load of simulating millions of random datasets from the prior under the ABC
+approach is nontrivial.
 

### 2013-09-29 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on dpp ms.
Adding Intro, Discussion and notation table.

@@
 
-        \MakeUppercase{\large\bfseries Title}
+        \MakeUppercase{\large\bfseries An Improved Approximate-Bayesian
+            Model-choice Method for Estimating Shared Evolutionary History}
 
@@
 \section*{Introduction}
-Being able to accurately infer the temporal distribution of divergences
-across co-distributed taxa
-
-Understanding evolutionary history is super cool.
-Adding a comparative framework extra super cool.
-Shared evolutionary history across taxa is often of interest:
-processes of diversification and community assembly,
-host-symbiont dynamics
+Understanding the processes that generate biodiversity and regulate
+community assembly is a major goal of Evolutionary Biology.
+Given the dynamic nature of our planet, a major component of this goal is
+understanding how geological and climatic processes can effect the
+evolutionary history of whole groups of co-distributed species and their
+associated microbiota, including pathogens.
+Such large-scale processes predict patterns of shared evolutionary history,
+e.g., temporally clustered divergences, across effected populations of
+organisms.
+At recent timescales, such events should leave a signature in the genetic
+variation within and among the effected lineages.
+Thus, being able to accurately estimate models of shared evolutionary
+histories across co-distributed taxa would provide an important
+tool for better understanding how regional and global biogeographic
+processes affect biodiversity.
+
+This inference problem is challenging due to the stochastic nature by which
+mutations occur in populations and how they are inherited over generations
+\citep{Hudson1990,WakeleyCoalescent}.
+Thus, a method for estimating historical patterns of divergences across taxa
+should explicitly model the stochastic mutational and ancestral processes that
+generate and filter the genetic variation we observe in present-day genetic
+data.
+An appealing approach would be a comparative, Bayesian model-choice method for
+inferring the probability of competing divergence histories while integrating
+over uncertainty in mutational and ancestral processes via models of nucleotide
+substitution and coalescence.
+
+Such a method has been developed in an approximate-Bayesian model-choice
+framework and is implemented in the software package \msb
+\citep{Hickerson2006,Huang2011}.
+Recently, however, \citet{Oaks2012} found the method can be strongly biased
+toward spuriously estimating biogeographically interesting models of highly
+clustered divergence times across taxa.
+The results of their simulation-based assessment of the method suggest that a
+combination of broad, uniformly distributed priors on nuisance parameters and a
+U-shaped prior on divergence models were causing the bias.
+The model implemented in \msb uses a discrete uniform prior over the number of
+possible divergence events, which \citet{Oaks2012} show imposes a prior on
+divergence models that puts most of the prior mass on models with either very
+few or very many divergence times shared across the taxa.
+Furthermore, the uniform priors on most of the model's nuisance parameters
+likely hinder the marginal likelihoods of parameter-rich models, i.e., models
+with many divergence time parameters \citep[this is often referred to as
+Lindley's paradox;][]{Lindley1957}.
+Thus, together, the priors on divergence models and nuisance parameters used in
+\msb could create a strong bias toward models with very few divergence time
+parameters.
+
+Following the recommendations of \citet{Oaks2012} we modify the \msb model to
+test whether alternative parameterizations and priors will improve the behavior
+of the approximate Bayesian model-choice approach.
+We implement a Dirichlet-process prior over all possible models of divergence,
+and use gamma and beta probability distributions in place of uniform priors on
+many of the model's parameters.
+Our simulation-based analyses show that the new implementation has improved
+robustness, accuracy, and power compared to the original model.
+Also, our results show that both the uniform priors on nuisance parameters and
+the U-shaped prior on divergence model contribute to the biases of \msb.
 
@@
 \singlespacing
+
+%:TABLE-notation descriptions
+\begin{table}[htbp]
+    \sffamily
+    \scriptsize
+    \rowcolors{2}{}{myGray}
+    \addtolength{\tabcolsep}{-0.1cm}
+\caption{Summary of the notation used throughout this work; modified from \citet{Oaks2012}.}
+    \centering
+    \begin{tabular}{ l >{\raggedright\hangindent=0.5cm}m{14cm} }
+        \toprule
+        \textbf{Symbol} & \textbf{Description} \tn
+        \midrule
+        \npairs{} & Number of population pairs. \tn
+        \popSampleSize{i}{} & The number of genome copies sampled from population pair $i$, with \popSampleSize{1}{i} sampled from population 1, and \popSampleSize{2}{i} from population 2. \tn
+        \nloci{i} & Number of loci sampled from population pair $i$. \tn
+        \nlociTotal & Total number of unique loci sampled. \tn
+        \alignment{i}{j} & Sequence alignment of locus $j$ sampled from population pair $i$. \tn
+        \alignmentSSObs{i}{j} & Population genetic summary statistics calculated from \alignment{i}{j}. \tn
+        \alignmentVector & Vector containing the sequence alignments of each locus from each population pair: $(\alignment{1}{1},\ldots,\alignment{\npairs{}}{\nloci{\npairs{}}})$. \tn
+        \ssVectorObs & Vector containing the summary statistics of each locus from each population pair: $(\alignmentSSObs{1}{1},\ldots,\alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$. \tn
+        \ssSpace & Multi-dimensional Euclidean space around the observed summary statistics, \ssVectorObs. \tn
+        \tol & Radius of \ssSpace, i.e., the tolerance of the ABC estimation. \tn
+        \geneTree{i}{j} & Gene tree of the sequences in \alignment{i}{j}. \tn
+        \geneTreeVector & Vector containing the gene trees of each locus from each population pair: $(\geneTree{1}{1},\ldots,\geneTree{\npairs{}}{\nloci{\npairs{}}})$. \tn
+        \divTimeNum & Number of unique population divergence times among the \npairs{} population pairs. \tn
+        \divTime{} & One of the \divTimeNum \iid divergence time paramters in \globalcoalunit generations. \tn
+        \divTimeVector & Ordered set of divergence time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
+        \divTimeIndex{i} & The index of the divergence time parameter in \divTimeVector that is mapped to population pair $i$. \tn
+        \divTimeIndexVector & Vector of divergence time parameter indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn
+        \divTimeMap{i} & Time of divergence in \globalcoalunit generations between the populations of pair $i$. \tn
+        \divTimeMapVector & Vector of divergence times for each of the population pairs: $(\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})$. \tn
+        \divTimeScaled{i}{j} & Scaled time of divergence between the populations of pair $i$ for locus $j$. \tn
+        \divTimeScaledVector & Vector containing the scaled divergence times of each locus from each population pair: $(\divTimeScaled{1}{1},\ldots,\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$. \tn
+        \descendantTheta{1}{i}, \descendantTheta{2}{i} & \myTheta{} of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
+        \ancestralTheta{i} & \myTheta{} of the population ancestral to pair $i$. \tn
+        \descendantThetaVector{1}, \descendantThetaVector{2} & Vectors $(\descendantTheta{1}{1},\ldots,\descendantTheta{1}{\npairs{}})$ and $(\descendantTheta{2}{1},\ldots,\descendantTheta{2}{\npairs{}})$, respectively.
+\tn
+        \ancestralThetaVector & Vector containing the \ancestralTheta{} parameters for each population pair: $(\ancestralTheta{1},\ldots,\ancestralTheta{\npairs{}})$. \tn
+        \locusMutationRateScalar{j} & mutation-rate scaling parameter for locus $j$. \tn
+        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate parameters: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
+        \locusRateHetShapeParameter & Hyper-parameter for the shape of the gamma prior distribution on each \locusMutationRateScalar{}. \tn
+        \bottleScalar{1}{i}, \bottleScalar{2}{i} & \myTheta{}-scaling parameters that determine the magnitude of the population bottleneck in the $1^{st}$ and $2^{nd}$ descendant population of pair $i$, respectively.
+The bottleneck in each descendant population begins immediately after divergence.
+\tn
+        \bottleScalarVector{1}, \bottleScalarVector{2} & Vectors $(\bottleScalar{1}{1},\ldots,\bottleScalar{1}{\npairs{}})$ and $(\bottleScalar{2}{1},\ldots,\bottleScalar{2}{\npairs{}})$, respectively.
+\tn
+        \bottleTime{i} & Proportion of time between present and \divt{i} when the bottleneck ends for the descendant populations of pair $i$. \tn
+        \bottleTimeVector & Vector containing the \bottleTime{} parameters for each population pair: $(\bottleTime{1},\ldots,\bottleTime{\npairs{}})$. \tn
+        \migrationRate{i} & Symmetric migration rate between the descendant populations of pair $i$. \tn
+        \migrationRateVector & Vector containing the migration rates for each population pair: $(\migrationRate{1},\ldots,\migrationRate{\npairs{}})$. \tn
+        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
+This constant is intended to allow the user to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa.
+\tn
+        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
+This constant is intended to allow the user to scale \myTheta{} for differences in mutation rates among loci or among taxa.
+\tn
+        \ploidyScalarVector & Vector of ploidy and/or generation-time scaling constants: $(\ploidyScalar{1}{1},\ldots,\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$. \tn
+        \mutationRateScalarConstantVector & Vector of mutation-rate scaling constants: $(\mutationRateScalarConstant{1}{1},\ldots,\mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$. \tn
+        \divTimeMean  & Mean of \divTimeMapVector. \tn
+        \divTimeVar  & Variance of \divTimeMapVector. \tn
+        \divTimeDispersion  & Dispersion index of \divTimeMapVector ($\divTimeVar/\divTimeMean$). \tn
+        \numPriorSamples & Number of samples from the joint prior. \tn
+        \paramSampleVector{} & Vector of parameter values drawn from the joint prior. \tn
+        \ssVector{} & Vector containing the summary statistics calculated from data simulated under parameter values drawn from the prior (\hpvector{}). \tn
+        \paramSampleMatrix{} & Random sample of $\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples}$ drawn form the prior. \tn
+        \ssMatrix & Summary statistic vectors $\ssVector{1}, \ldots, \ssVector{\numPriorSamples}$ for each $\hpvector{1}, \ldots, \hpvector{\numPriorSamples}$ drawn from the prior. \tn
+        \bottomrule
+    \end{tabular}
+    \label{tabNotation}
+\end{table}
+
+\clearpage
 

### 2013-09-29 manuscripts/dpp-msbayes/dpp-msbayes.tex
Finishing initial draft of dpp ms.

@@
         \vspace{12pt}
-        \textbf{Running head:} \MakeUppercase{Running head}
+        \textbf{Running head:} \MakeUppercase{Improved Estimation of Shared Evolutionary History}
 
@@
     \noindent\textbf{ABSTRACT} \\
-    \noindent Abstract here \ldots
+    \noindent To understand the processes that generate biodiversity, it is
+    important to account for large-scale processes that effect the evolutionary
+    history of groups of co-distributed populations of organisms.
+    Such events predict temporally clustered divergences times, a pattern
+    that can be estimated from genetic data from co-distributed species.
+    We introduce a new approximate-Bayesian model for comparative
+    phylogeographical model-choice that estimates the temporal distribution of
+    divergences across taxa from multi-locus DNA sequence data.
+    Our model is an extension of the model implemented in \msb.
+    By reparameterizing the model, and using more flexible priors on divergence
+    models and nuisance parameters, we improve the robustness, accuracy, and power
+    of the method for estimating the posterior probabilities of models
+    of divergence across taxa.
+    Our results demonstrate that the bias of the \msb model toward inferring
+    models of clustered divergences is caused by a combination of (1) uniform
+    priors on nuisance parameters hindering the marginal likelihoods of models
+    with more divergence time parameters, and (2) a prior on divergence models
+    that disfavors models with intermediate numbers of divergence time
+    parameters.
 
     \vspace{12pt}
-    \noindent\textbf{KEY WORDS: } 
+    \noindent\textbf{KEY WORDS: Dirichlet-process prior; approximate Bayesian
+    computation; model choice; phylogeography; biogeography} 
 }
@@
 \newpage
-\noindent \ldots
 
@@
 associated microbiota, including pathogens.
-Such large-scale processes predict patterns of shared evolutionary history,
-e.g., temporally clustered divergences, across effected populations of
-organisms.
+Such large-scale processes predict patterns of shared evolutionary history
+across effected populations of organisms, e.g., temporally clustered
+divergences across co-distributed species.
 At recent timescales, such events should leave a signature in the genetic
@@
 over uncertainty in mutational and ancestral processes via models of nucleotide
-substitution and coalescence.
-
-Such a method has been developed in an approximate-Bayesian model-choice
-framework and is implemented in the software package \msb
-\citep{Hickerson2006,Huang2011}.
-Recently, however, \citet{Oaks2012} found the method can be strongly biased
+substitution and lineage coalescence.
+
+The software package \msb implements such an approach in an
+approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
+Recently, however, \citet{Oaks2012} found that \msb can be strongly biased
 toward spuriously estimating biogeographically interesting models of highly
@@
 The model implemented in \msb uses a discrete uniform prior over the number of
-possible divergence events, which \citet{Oaks2012} show imposes a prior on
+possible divergence parameters, which \citet{Oaks2012} show imposes a prior on
 divergence models that puts most of the prior mass on models with either very
 few or very many divergence times shared across the taxa.
-Furthermore, the uniform priors on most of the model's nuisance parameters
+Furthermore, uniform priors used for most of the model's nuisance parameters
 likely hinder the marginal likelihoods of parameter-rich models, i.e., models
@@
 \msb could create a strong bias toward models with very few divergence time
-parameters.
+parameters \citep{Oaks2012}.
 
@@
 Also, our results show that both the uniform priors on nuisance parameters and
-the U-shaped prior on divergence model contribute to the biases of \msb.
+the U-shaped prior on divergence models contribute to the biases of \msb.
 

### 2013-09-29 manuscripts/dpp-msbayes/dpp-msbayes.tex
Modifying dpp ms.
Adding software info and links to beginning of methods.

@@
 \subsection*{The model}
-In this section, we describe our model, which is a modification
-of the model implemented in \msb \citep{Huang2011,Oaks2012}.
+In this section, we describe our model, which is a modification of the model
+implemented in \msb \citep{Huang2011,Oaks2012}.
+The code implementing the new model is freely available in the open-source
+software package \dppmsbayes
+(\href{https://github.com/joaks1/dpp-msbayes}{https://github.com/joaks1/dpp-msbayes}).
+Our analyses described below utilize the freely avaliable, open-source software
+package \pymsbayes (\href{https://github.com/joaks1/PyMsBayes}{https://github.com/joaks1/PyMsBayes}),
+which is a Python API that provides a multi-processing interface to \msb and
+\dppmsbayes.
+We performed the work described below following the principles of
+Open Notebook Science.
+Using version-control software, we make progress in all aspects of the work
+freely and publicly available in real-time at
+\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
+All information necessary to reproduce our results is provided there.
 We follow much of the notation of \citet{Oaks2012}, but modify it to

### 2013-10-01 manuscripts/reply/reply.tex
Fixing typo in figure caption of reply.
Incorporating edits to reply draft.

@@
     divergence time parameters due to the integration over vast parameter space
-    imposed by broad uniform priors characterized by low likelihood and high
-    prior density.
-    This phenomenon is often referred to as Lindley's paradox
-    \citep{Lindley1957}.  \citet{Hickerson2013} have responded to this work,
-    criticising the priors used by \citet{Oaks2012} and presenting an
-    approximate Bayesian model-averaging approach that uses narrow, empirically
-    informed uniform priors.
-    \citet{Hickerson2013} conclude this approach circumvents the issues
-    revealed by \citet{Oaks2012} and yields robust inference of divergence
-    models.
+    with low likeihood and high prior density that is imposed by broad uniform
+    priors.
+    \citet{Hickerson2013} have responded to this work, criticising the priors
+    used by \citet{Oaks2012} and presenting an approximate Bayesian
+    model-averaging approach that uses narrow, empirically informed uniform
+    priors.
+    \citet{Hickerson2013} use this method to re-analyze the empirical dataset
+    of \citet{Oaks2012} comprised of 22 pairs of vertebrate populations from
+    the Philippines, and conclude the model-averaging approach circumvents the
+    issues revealed by \citet{Oaks2012} and yields robust inference of
+    divergence models.
     Furthermore, they argue that the biases revealed by \citet{Oaks2012} are
     inherent to the inefficient rejection algorithm implemented in \msb, rather
-    than a result of Lindley's paradox.
-    Here we show that the methodology of \citet{Hickerson2013} suffers from a
-    fundamental error that renders some of their results invalid and the
-    remainder difficult to interpret.
-    Furthermore, we demonstrate how the approach of \citet{Hickerson2013} still
-    suffers from a bias toward models with less parameter space, which
-    manifests in a dangerously strong tendency to sample predominantly from
-    models that exclude the true values of the model's parameters.
+    than a result of low marginal likelihoods of parameter-rich models.
+    Here we discuss an error in the methodology of \citet{Hickerson2013} that
+    renders some of their results invalid and the remainder difficult to
+    interpret.
+    Furthermore, we demonstrate that the approach of \citet{Hickerson2013} is
+    biased toward models with less parameter space, which manifests in a strong
+    tendency to sample predominantly from models that exclude the true values
+    of the model's parameters.
     Our results suggest that their model-averaged results are dominated by
@@
     difficulty in interpreting their results.
-    We also show that in addition to this dangerous behavior, the approach of
-    \citet{Hickerson2013} is still strongly biased toward inferring
-    simultaneous divergences even when the true divergences are randomly
-    distributed over millions of generations.
+    We also show the approach of \citet{Hickerson2013} is still biased toward
+    inferring simultaneous divergences even when the true divergences are
+    randomly distributed over millions of generations.
     Our results, coupled with those of \citet{Oaks2012}, strongly suggest that
-    Lindley's paradox is a primary cause of the bias of \msb toward models with
-    lesser parameter space.
+    low marginal likelihoods of complex models is a primary cause of the bias
+    of \msb toward models with less parameter space.
     We discuss ways to mitigate this bias without resorting to adding an
-    additional dimension of model choice to the model or using dangerous
+    additional dimension of model choice to the model or using questionable
     empirically informed uniform priors.
@@
 unordered models of divergence.
-\citet{Oaks2012} published results of simulation-based power analyses that
-revealed that \msb can often be biased towards inferring models of temporally
-clustered divergence times among taxon pairs.
+\citet{Oaks2012} published results of simulation-based power analyses
+showing that \msb can often be biased toward inferring models of temporally
+clustered divergence times across taxon pairs.
 \citet{Hickerson2013} published a response to this paper where they present
@@
 poor behavior of the method revealed by \citet{Oaks2012}.
-These papers are largely in agreement with \citet{Hickerson2013} reiterating a
-lot of the discussion of
-\citet{Oaks2012} regarding the impact of broad uniform priors on Bayesian model
-choice.
+These papers are largely in agreement, with \citet{Hickerson2013} reiterating
+much of the discussion of \citet{Oaks2012} regarding the impact of broad
+uniform priors on Bayesian model choice.
 The main difference in perspectives are centered around
@@
 \msb, and
-(2) how to potentially ameliorate this issue.
-
-Regarding the mechanism behind the poor behavior of the method,
-\citet{Oaks2012} suggest the primary cause is likely the low marginal
-likelihoods of parameter-rich models integrated over vast parameter space with
-low probability of producing the data, yet relatively high prior density
-\citep[this is often referred to as Lindley's paradox;][]{Lindley1957}.
+(2) how to potentially mitigate this issue.
+
+\citet{Oaks2012} suggest the primary cause of the poor behavior of the method
+is likely the low marginal likelihoods of parameter-rich models integrated over
+vast parameter space with low probability of producing the data, yet relatively
+high prior density \citep{Lindley1957}.
 Note, this suggests the bias is extrinsic to \msb, and the numerical
@@
 that focuses on only one of the three prior models used by \citet{Oaks2012}.
-We show that this argument is based on an empirical Bayesian approach and
+We show that this argument is based on an empirical Bayesian reasoning and
 questionable assumptions, and does not explain the bias of the method found in
 many of the analyses of \citet{Oaks2012}. 
-Furthermore, we present results of additional analyses that, when combined
-with the results of \citet{Oaks2012}, strongly suggest that Lindley's paradox
-is playing a large role in the poor behavior of \msb.
+Furthermore, we present results of additional analyses that, when combined with
+the results of \citet{Oaks2012}, strongly suggest small marginal likelihoods of
+models with more parameter space is a primary mechanism by which uniform priors
+are causing the poor behavior of \msb.
 
@@
 
-Regarding perspectives about mitigating the bias, \citet{Oaks2012} suggest that
-alternative prior probability distributions on divergence models and nuisance
-parameters could reduce the effect of Lindley's paradox.
-They also layout simulation-based procedures for determining power, accuracy,
+To mitigate the bias, \citet{Oaks2012} suggest alternative prior probability
+distributions on divergence models and nuisance parameters could increase the
+likelihoods of models with more divergence time parameters.
+They also describe simulation-based procedures for determining power, accuracy,
 and robustness of the method given a dataset, and recommend any application of
-\msb should be accompanied by such procedures, especially if any biological
-conclusions are drawn from the results.
+\msb should be accompanied by such procedures.
 Alternatively, \citet{Hickerson2013} present an approximate Bayesian
@@
 pitfalls raised by \citet{Oaks2012}.
-We show that fundamental errors in this approach render some of their results
-invalid and leave the remaining results difficult to interpret.
-Furthermore, we follow the advice of \citet{Oaks2012} and present
+We discuss errors in this approach that render some of their results invalid
+and leave the remaining results difficult to interpret.
+Furthermore, we follow the recommendations of \citet{Oaks2012} and present
 simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
@@
 Our results demonstrate that the model-averaging approach of
-\citet{Hickerson2013} is biased and dangerous.
-The approach provides another means by which the model can ``escape'' large
-parameter space, and the method's bias toward smaller models remains.
-This manifestation of the bias can be dangerous and often excludes the true
-parameter space due to the use of narrow uniform priors.
+\citet{Hickerson2013} is biased toward clustered divergence models and
+predisposed to excluding true parameter space.
+The model-averaging approach provides another means by which the model can
+``escape'' large parameter space, and the method's bias toward smaller models
+remains.
+This manifestation of the bias often excludes the true parameter space due to
+the use of narrow uniform priors.
 Furthermore we discuss the potential theoretical and practical problems of
-empirical Bayesian model choice, which is used and encouraged by
-\citet{Hickerson2013} without discussing its potential complications.
+empirical Bayesian model choice.
+% which is used and encouraged by
+% \citet{Hickerson2013} without discussing its potential complications.
 
@@
 different prior models (see Table 1 of \citet{Hickerson2013}).
-In our initial attempts to explore the behavior of this approach, we found
-a fundamental error in the methodology of \citet{Hickerson2013}:
-their model mixes different time units.
+However, there is a fundamental error in the methodology of
+\citet{Hickerson2013}: their model mixes different time units.
 
@@
 of \citet{Hickerson2013} difficult to interpret, and renders their
-regression-adjusted results invalid; a fundamental assumption of regression is
-that all of the values of the response variable are in the same units.
+regression-adjusted results invalid.
+A fundamental assumption of regression is that all of the values of the
+response variable are in the same units.
 Thus, the results in sections ``Using ABC Model Comparison to Weight
@@
 Another important characteristic of the approach used by \citet{Hickerson2013}
-that makes their results difficult to interpret in addition to this error is
-their use of narrow, empirically guided uniform priors in an
-approximate-Bayesian model-choice framework.
-We discuss the dangers of this approach in the next section, and show how the
-results of \citet{Hickerson2013} likely exclude true values of the model's
-parameters.
+that makes their results difficult to interpret is their use of narrow,
+empirically guided uniform priors in an approximate-Bayesian model-choice
+framework.
+We discuss the risks associated with this approach in the next section, and
+show how the results of \citet{Hickerson2013} likely exclude true values of the
+model's parameters.
 
@@
 
-\section*{The dangers of empirical Bayesian model choice}
+\section*{The potential implications of empirical Bayesian model choice}
 \citet{Hickerson2013} repeatedly refer to the prior distributions used by
-\citet{Oaks2012} as ``poorly selected''. 
+\citet{Oaks2012} as ``poorly selected.''
 However, this is misleading, as \citet{Oaks2012} discuss in detail how they
@@
 prior'').
-They detail that the limitation in using uniform prior distributions in \msb
-necessitates awkward broad priors to avoid excluding the truth, \textit{a
-    priori}.
-What \citet{Hickerson2013} really mean by ``poorly selected'' is that
+They mention how the use of uniform prior distributions in \msb for many of the
+model's parameters necessitates awkward broad priors to avoid excluding the
+truth, \textit{a priori}.
+What \citet{Hickerson2013} mean by ``poorly selected'' is that
 \citet{Oaks2012} did not use their data to inform their initial prior
@@
 has given rise to many powerful inference methods.
-Unfortunately, empirical Bayes methods have an uncertain theoretical basis and
+However, empirical Bayes methods have an uncertain theoretical basis and
 do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
@@
 estimation, a theoretical justification for empirical Bayes approaches to model
-choice is unclear.
+choice is questionable.
 In Bayesian model choice, the goal is not to estimate parameters, but to
@@
 them.
-
-%CWL:
-%I like the example, but it may be too cute. It is also too long. I think a
-%contrived example is good, but if there could be an example that is more
-%relevant to model choice and comparison of inaccurate posterior distributions,
-%I think that would be better. This is an example of empirical Bayes, but not
-%the potential problem with the use in msbayes. Unless I am interpreting wrong.
-%I would be careful about making it to cute as well. You may intend to make it
-%easy to understand for a general audience, but it can also seem like talking
-%down to Hickerson et al, who we are responding to. It just depends on how you
-%read it.
-This can be demonstrated with a simple, albeit contrived, example.
-Let us say that principal investigator Mary and her new postdoc Will are
-interested in the hypothesis that the mass of George Washington's periwig
-renders the quarter dollar of the United States unfair.
-That is to say their null hypothesis is that the probability of a US
-quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
-They can certainly evaluate this hypothesis, they can have undergraduate
-worker, Joe, flip the coin for them while they tabulate the results.
-But being Bayesians, before they call Joe into the lab, they agree on a
-prior probability to place on the set of all possible probabilities that
-the quarter will land heads when it is flipped.
-Given that neither of them have a quarter, and their prior knowledge that Joe
-moonlights as a magician, and is notorious for performing coin and card tricks
-in the lab, they suspect there is a good chance that the quarter Joe uses will
-be either two-headed or two-tailed.
-So, knowing that the beta distribution is the conjugate prior for a binomial
-likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
-(Figure~\ref{figCoinFlip}).
-
-Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
-begin flipping it.
-After five tosses, four of which was heads, Joe decides that academics are
-crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
-being computational biologists, are satisfied with their
-empirical dataset of $y = 4$ heads out of $N = 5$ trials.
-They know from the conjugacy of the beta prior, that the posterior distribution
-has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
-in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
-of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
-This allows them to plug these values into the beta cumulative distribution
-function to determine that the posterior probability of their hypothesis is
-$p(\theta < 0.5 | y=1, N=5) = 0.088$.
-Given their prior belief and dataset, this indeed is the correct posterior
-probability of the hypothesis, and Mary and Will should now update their
-posterior belief accordingly.
-
-However, reflecting upon the results of their experiment
-(Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
-Their prior looks very ``poorly selected,'' and if they had only known that the
-coin was fair before the data were collected, they would have selected a much
-better prior.
-Clearly, from their results, they should have used a prior centered
-around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
-much ``better.''
-Rather than resort to flipping the coin five more times, Mary and Will decide
-to redo their analysis using the much better ``prior'' of $p(\theta) \sim
-beta(4.5, 1.5)$.
-This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
-probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
-Now convinced that the US quarter is unfair, albeit not due to the mass of
-President Washington's head as they hypothesized, Will begins composing an
-e-mail of complaint addressed to the U.S.\ Mint.
-
-If Joe's flips were a representative sample, Mary and Will's empirical Bayes
-estimate might very well be a better point estimate for the parameter
-\myTheta{}.
-However, their empirical Bayes estimate of the probability of their hypothesis
-is incorrect and biased.
-This simple example shows how parameter estimation is fundamentally different
-from estimating the probability of a model.
-While empirically informed priors can be used to obtain well-behaved parameter
-estimators, using them for model choice is much less certain.
-In addition to these theoretical concerns there are practical dangers to
-using narrow, empirical, uniform priors for a method that has already
-been shown to be biased toward models with lesser parameter space.
-
-\subsection*{Practical dangers of empirically informed uniform priors for
+In addition to these theoretical concerns there are practical dangers to using
+narrow, empirical, uniform priors for a method that has already been shown to
+be biased toward models with less parameter space.
+% %CWL:
+% %I like the example, but it may be too cute. It is also too long. I think a
+% %contrived example is good, but if there could be an example that is more
+% %relevant to model choice and comparison of inaccurate posterior distributions,
+% %I think that would be better. This is an example of empirical Bayes, but not
+% %the potential problem with the use in msbayes. Unless I am interpreting wrong.
+% %I would be careful about making it to cute as well. You may intend to make it
+% %easy to understand for a general audience, but it can also seem like talking
+% %down to Hickerson et al, who we are responding to. It just depends on how you
+% %read it.
+% This can be demonstrated with a simple, albeit contrived, example.
+% Let us say that principal investigator Mary and her new postdoc Will are
+% interested in the hypothesis that the mass of George Washington's periwig
+% renders the quarter dollar of the United States unfair.
+% That is to say their null hypothesis is that the probability of a US
+% quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
+% They can certainly evaluate this hypothesis, they can have undergraduate
+% worker, Joe, flip the coin for them while they tabulate the results.
+% But being Bayesians, before they call Joe into the lab, they agree on a
+% prior probability to place on the set of all possible probabilities that
+% the quarter will land heads when it is flipped.
+% Given that neither of them have a quarter, and their prior knowledge that Joe
+% moonlights as a magician, and is notorious for performing coin and card tricks
+% in the lab, they suspect there is a good chance that the quarter Joe uses will
+% be either two-headed or two-tailed.
+% So, knowing that the beta distribution is the conjugate prior for a binomial
+% likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
+% (Figure~\ref{figCoinFlip}).
+
+% Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
+% begin flipping it.
+% After five tosses, four of which was heads, Joe decides that academics are
+% crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
+% being computational biologists, are satisfied with their
+% empirical dataset of $y = 4$ heads out of $N = 5$ trials.
+% They know from the conjugacy of the beta prior, that the posterior distribution
+% has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
+% in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
+% of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
+% This allows them to plug these values into the beta cumulative distribution
+% function to determine that the posterior probability of their hypothesis is
+% $p(\theta < 0.5 | y=1, N=5) = 0.088$.
+% Given their prior belief and dataset, this indeed is the correct posterior
+% probability of the hypothesis, and Mary and Will should now update their
+% posterior belief accordingly.
+
+% However, reflecting upon the results of their experiment
+% (Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
+% Their prior looks very ``poorly selected,'' and if they had only known that the
+% coin was fair before the data were collected, they would have selected a much
+% better prior.
+% Clearly, from their results, they should have used a prior centered
+% around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
+% much ``better.''
+% Rather than resort to flipping the coin five more times, Mary and Will decide
+% to redo their analysis using the much better ``prior'' of $p(\theta) \sim
+% beta(4.5, 1.5)$.
+% This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
+% probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
+% Now convinced that the US quarter is unfair, albeit not due to the mass of
+% President Washington's head as they hypothesized, Will begins composing an
+% e-mail of complaint addressed to the U.S.\ Mint.
+
+% If Joe's flips were a representative sample, Mary and Will's empirical Bayes
+% estimate might very well be a better point estimate for the parameter
+% \myTheta{}.
+% However, their empirical Bayes estimate of the probability of their hypothesis
+% is incorrect and biased.
+% This simple example shows how parameter estimation is fundamentally different
+% from estimating the probability of a model.
+% While empirically informed priors can be used to obtain well-behaved parameter
+% estimators, using them for model choice is much less certain.
+
+\subsection*{Practical concerns about empirically informed uniform priors for
     Bayesian model choice}
@@
 Given this interval estimate is for the \emph{mean} divergence time across all
-22 taxa, it may be inappropriate to set this as the prior, because many of the
-taxon pairs are expected to have diverged at times older than the upper limit.
+22 taxa, it may be inappropriate to set this as the limit on the prior, because
+many of the taxon pairs are expected to have diverged at times older than the
+upper limit.
 Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
@@
 see Table~6 of \citet{Oaks2012}).
-Thus, the results of \citet{Hickerson2013} indicate Lindley's paradox is
-biasing the method toward models with smaller parameter space, and as a
-consequence is biasing the method toward models that exclude the truth (i.e.,
-toward estimating model-averaged posteriors dominated by models that exclude
-true values of parameters of the model).
+Thus, the results of \citet{Hickerson2013} indicate that integration over
+parameter space with low likelihood is biasing the method toward models with
+smaller parameter space, and as a consequence is biasing the method toward
+models that exclude the truth (i.e., toward estimating model-averaged
+posteriors dominated by models that exclude true values of parameters of the
+model).
 
@@
 most, if not all, of the true divergence times of the 22 taxon pairs.
-If Lindley's paradox causes the method to prefer models with less parameter
-space, we expect \msb will preferentially sample from this incorrect model
-yielding a marginal posterior that is incorrect (i.e., the model-averaged
-posterior will be dominated by an incorrect model that excludes the truth).
+If small likelihoods of large models cause the method to prefer models with
+less parameter space, we expect \msb will preferentially sample from this
+incorrect model yielding a marginal posterior that is incorrect (i.e., the
+model-averaged posterior will be dominated by an incorrect model that excludes
+the truth).
 Second, we generated simulated datasets for which the divergence times are
@@
 approach of \citet{Hickerson2013}, but with a reduced set of prior models to
-avoid their error of mixing time units.
+avoid mixing time units.
 We used five prior models, all of which had priors on population sizes of
@@
 $M_5$, $\divt{} \sim U(0, 20)$.
-For our prior, we simulated $1\e6$ random samples from each of the models
-for a total sample of $5\e6$.
+We simulated $1\e6$ random samples from each of the models for a total of
+$5\e6$ prior samples.
 For each model, we retained the 10,000 samples with the smallest Euclidean
@@
 does not reduce the method's bias toward models with less parameter space,
-but rather allows it to manifest in a very dangerous way.
+but rather allows it to manifest in a rather dangerous way.
 The method strongly prefers the prior model with the narrowest distribution on
@@
 To better quantify the propensity of Hickerson et al.'s
-(\citeyear{Hickerson2013}) approach to exclude the truth, we simulated
-1000 pseudo-observed datasets in which the divergence times for the
-22 population pairs are drawn randomly from an exponential distribution
-with a mean of 0.5 ($\divt{} \sim Exp(2)$).
+(\citeyear{Hickerson2013}) approach to exclude the truth, we simulated 1000
+datasets in which the divergence times for the 22 population pairs are drawn
+randomly from an exponential distribution with a mean of 0.5 ($\divt{} \sim
+Exp(2)$).
 All other parameters were identically distributed as the $M_1$--$M_5$ models
 (Table~\ref{tabModelChoiceEmpirical}).
-We then repeated the analysis described above, using $5\e6$ prior samples from
-$M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, and retaining 1000 posterior samples
-for each of the 1000 pseudo-observed datasets.
-
-\highLight{Add how we determined exclusion.}
+We then repeated the analysis described above using $1\e6$ random samples from
+prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, retaining 1000 posterior
+samples for each of the 1000 simulated datasets.
+
+% \highLight{Add how we determined exclusion.}
+For each simulation replicate, we estimated the Bayes factor in favor
+of excluding the truth as the ratio of the posterior to prior odds of
+excluding the true value of at least one parameter.
+Whenever the Bayes factor preferred a model excluding the truth we counted
+the number of 22 true divergence times that were excluded by the preferred
+model.
 Our results show that the model-averaging approach of \citet{Hickerson2013}
@@
 (Figure~\ref{figExclusionSimTau}).
+We also used the mode estimate of the preferred model for each replicate to
+estimate the number of true values excluded, which produced very similar
+results.
 Furthermore, the posterior probability of excluding at least one true parameter
@@
 (Figure~\ref{figExclusionSimProb}).
-We used a Bayes factor (comparing models that exclude the truth to those that
-do not) calculated for each replicate of greater than 10 as strong support for
-excluding the truth.
-With this criterion, 66\% of the replicates (87\% with GLM-regression adjustment)
-strongly support the exclusion of true values (Figure~\ref{figExclusionSimProb}).
+Using a Bayes factor of greater than 10 as a criterion for strong support, 66\%
+of the replicates (87\% with GLM-regression adjustment) strongly support the
+exclusion of true values (Figure~\ref{figExclusionSimProb}).
 
@@
 model-averaging framework.
-This is especially dangerous in \msb, which has already been demonstrated to be
-biased away from models with more parameter space \citep{Oaks2012}.
+This is especially concerning in \msb, which has already been demonstrated to
+be biased away from models with more parameter space \citep{Oaks2012}.
 The consequence of this approach is a high risk of obtaining a model-averaged
@@
 
-\subsection*{Final thoughts on empirical priors in Bayesian model choice}
+\subsection*{Additional thoughts on empirical priors in Bayesian model choice}
 Given all of the theoretical and practical issues with empirical Bayes
@@
 
+While our results above clearly demonstrate the risks inherent to the empirical
+Bayesian model-choice approach used by \citet{Hickerson2013}, one could justify
+such risk if the method does indeed increase the power of the method and thus
+decrease bias toward clustered divergence models, as claimed by
+\citet{Hickerson2013}.
+In the next section, we use simulation-based analyses to evaluate the power of
+Hickerson et al.'s (\citeyear{Hickerson2013}) approach, which show the method is
+still biased toward highly clustered models of divergence.
+
 \section*{Assessing the power of the model-averaging approach of
@@
 proposed by \citet{Hickerson2013}.
-Following \citet{Oaks2012}, we simulated 1000 pseudo-replicate datasets with
-\divt{} for each of the 22 population pairs randomly drawn from a uniform
-distribution, $U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6,
-0.8, 1.0, and 2.0, in \globalcoalunit generations.
+Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
+the 22 population pairs randomly drawn from a uniform distribution, $U(0,
+\divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and 2.0, in
+\globalcoalunit generations.
 All other parameters were identically distributed as the prior.
-Following \citet{Hickerson2013} (but avoiding mixing time units), we used five
-prior models, all of which had priors on population sizes of
-$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
-U(0.0001, 0.05)$.
-Each of these priors had the following priors on divergence time parameters:
-$M_1$, $\divt{} \sim U(0, 0.1)$;
-$M_2$, $\divt{} \sim U(0, 1)$;
-$M_3$, $\divt{} \sim U(0, 5)$;
-$M_4$, $\divt{} \sim U(0, 10)$; and
-$M_5$, $\divt{} \sim U(0, 20)$.
-For our prior, we simulated $1\e6$ random samples from each of the models
-for a total sample of $5\e6$.
-For each pseudo-replicate dataset, we retained 1000 samples from each prior
-model with the smallest Euclidean distance from the pseudo-observed summary
-statistics, standardizing the statistics using the prior means and standard
-deviations of the given model.
-Then, for each replicate, we retained 1000 samples with the overall
-smallest Euclidean distance from the pseudo-observed summary statistics, this
-time standardizing the statistics using the prior means and standard deviations
-across all five models.
+We use the same $5\e6$ samples from the same set of five prior models as above
+($M_1$, $M_2$, $M_3$, $M_4$, and $M_5$).
+As described above, for each simulated dataset, we approximated the posterior
+by retaining 1000 samples from the prior with the smallest Euclidean distance
+from the true summary statistics.
 In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
@@
 
-The results of our power analysis further demonstrate the propensity of
-Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true parameter
-values.
+The results of our power analyses further demonstrate the propensity of
+Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
+parameter values.
 Across all but one of the \divt{max} we simulated, the method excludes the
-truth in large proportion of replicates, and across many of the \divt{max} it
+truth in a large proportion of replicates, and across many of the \divt{max} it
 will exclude a large proportion of the true divergence time values
@@
 excludes regions of parameter space containing the truth, and lacks power to
-detect random variation in divergence times over the scale of $8\globalpopsize$
+detect random variation in divergence times over a scale of $8\globalpopsize$
 generations or more.
@@
 In an empirical system such as the Philippines, where island fragmentation has
-occurred at least \highLight{XX} times over the past several millions of years,
-an estimate where 22 taxa share even a handful of divergence events would be of
-biogeographic interest.
+occurred at least 10 times over the past several millions of years
+\citep{Haq1987,Rohling1998,Siddall2003,Miller2005}, an estimate where 22 taxa
+share even a handful of divergence events would be of biogeographic interest.
 
@@
 
-%CWL: This seems to specific, not sure about including
-\section*{Validation analyses}
-\highLight{Not sure where to put this section}.
-\highLight{This justifies presenting the unadjusted results in this paper.}
-\highLight{Is it worth including just for this purpose?}
-Following \citet{Oaks2012}, we characterize the model-choice behavior of the
-model-averaging approach of \citet{Hickerson2013} under the ideal conditions
-where the prior is correct (i.e., the data are generated from parameters drawn
-from the same prior distributions used in the analysis).
-We used the same prior models as above ($M_1$--$M_5$;
-Table~\ref{tabModelChoiceEmpirical}), and generated 50,000 pseudo-replicate
-datasets under this prior (10,000 from each model).
-We used a simulated data structure of eight population pairs, with a single
-1000 bp locus sampled from 10 individuals from each population.
-We then analyzed each of these replicate datasets using the same prior,
-retaining 1000 posterior samples.
-Our results are very similar to \citet{Oaks2012}, but we note that they
-are not directly comparable as our simulations contained eight population
-pairs rather than 10 (Figure~\ref{figValidationMCBehavior}).
-We find that the method does a decent job of estimating the posterior
-probability of divergence models when all assumptions of the method
-are met, the prior is correct, and the unadjusted posterior estimates
-are used.
-Similar to \cite{Oaks2012}, we find that the regression-adjusted estimates of
-the model probabilities are biased.
-
-\section*{Lindley's paradox versus insufficient prior sampling}
+\section*{Model likelihoods or insufficient prior sampling as the cause of bias
+in \msb}
 One of the points of disagreement between \citet{Hickerson2013} and
@@
 \citet{Oaks2012} present simulation-based analyses that suggest the broad
-uniform prior on divergence time parameters hinder the marginal likelihoods of
-models with more divergence time parameters, whereas \citet{Hickerson2013}
+uniform prior on divergence time parameters decrease the marginal likelihoods
+of models with more divergence time parameters, whereas \citet{Hickerson2013}
 argue that the bias is inherent to the inefficient rejection algorithm
@@
 \citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
-dangerous, especially given our findings on the behavior of empirically
+questionable, especially given our findings on the behavior of empirically
 informed priors presented above.
@@
 The narrowest prior on divergence times used by \citet{Oaks2012} closely mirrors
-the range of gene divergence time estimates, and applying Hickerson et al.'s
+the range of estimates of gene-divergence times, and applying Hickerson et al.'s
 (\citeyear{Hickerson2013}) probability equations demonstrates that the prior
@@
 
-To disentangle whether Lindely's paradox or insufficient prior sampling is
-to blame for the biases revealed by \citet{Oaks2012}, we must look at
-the different predictions made by these two phenomena.
+To disentangle whether model likelihoods or insufficient prior sampling is to
+blame for the biases revealed by \citet{Oaks2012}, we must look at the
+different predictions made by these two phenomena.
 One example, as discussed by \citet{Oaks2012}, is that insufficient prior
@@
 and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
-\ref{figPowerOmegaProb}) strongly implicate Lindley's paradox as a primary
-cause.
+\ref{figPowerOmegaProb}) strongly suggest the primary cause is the uniform
+priors reducing the marginal likelihoods of large models.
 In all of our model-averaging analyses, all prior models have the same number
 of samples.
-Thus, it is difficult to explain the method's propensity towards the model with
+Thus, it is difficult to explain the method's propensity toward the model with
 the smallest parameter space via insufficient prior sampling.
@@
 
-As discussed by \citet{Oaks2012}, a straightforward prediction of Lindley's
-paradox is that it should disappear as the model generating the data converges
-to the prior.
+As discussed by \citet{Oaks2012}, a straightforward prediction of reduced
+likelihoods of complex models is that they should disappear as the model
+generating the data converges to the prior.
 \citet{Oaks2012} test this prediction by performing 100,000 simulations to
@@
 Thus, the results of the simulations of \citet{Oaks2012} are more consistent
-with Lindley's paradox causing the bias toward models with fewer divergence
-time parameters.
+with marginal model likelihoods causing the bias toward models with fewer
+divergence time parameters.
 
 Taken together, the results presented here and in \citet{Oaks2012} support
-Lindley's paradox as a primary mechanism by which broad uniform priors
-cause biases in \msb.
+reduced marginal likelihoods of large models as a primary mechanism by which
+broad uniform priors cause biases in \msb.
 Nonetheless, posterior sampling error will always be present in any numerical
@@
 
+%CWL: This seems to specific, not sure about including
+\section*{Validation analyses}
+% \highLight{Not sure where to put this section}.
+% \highLight{This justifies presenting the unadjusted results in this paper.}
+% \highLight{Is it worth including just for this purpose?}
+Following \citet{Oaks2012}, we characterize the model-choice behavior of the
+model-averaging approach of \citet{Hickerson2013} under the ideal conditions
+where the prior is correct (i.e., the data are generated from parameters drawn
+from the same prior distributions used in the analysis).
+We used the same prior models as above ($M_1$--$M_5$;
+Table~\ref{tabModelChoiceEmpirical}), and generated 50,000 pseudo-replicate
+datasets under this prior (10,000 from each model).
+We used a simulated data structure of eight population pairs, with a single
+1000 base-pair locus sampled from 10 individuals from each population.
+We then analyzed each of these replicate datasets using the same prior,
+retaining 1000 posterior samples.
+Our results are very similar to \citet{Oaks2012}, but we note that they
+are not directly comparable as our simulations contained eight population
+pairs rather than 10 (Figure~\ref{figValidationMCBehavior}).
+We find that the approach of \citet{Hickerson2013} estimates the posterior
+probability of divergence models reasonably well when all assumptions of the
+method are met (i.e., the prior is correct) and the unadjusted posterior
+estimates are used.
+Similar to \cite{Oaks2012}, we find that the regression-adjusted estimates of
+the model probabilities are biased.
+
 \section*{Differing utilities of \numt{} and \vmratio{} in \msb}
-The primary parameter of the \msb model is the vector of divergence
+The main parameter of the \msb model is the vector of divergence
 times for each of the taxon pairs,
@@
 \citet{Hickerson2013} argue that the dispersion index of this vector,
-\vmratio{}, is a better model-choice estimator than the vector's cardinality,
+\vmratio{}, is a better model-choice estimator than the number of 
+divergence time parameters within the vector,
 \numt{}.
@@
 \citet{Hickerson2013}), which is simply a plot of sample size versus variance.
-This plot shows, not surprisingly, that \vmratio{} has essentially no
-information about the number of divergences among taxa.
+This plot shows, not surprisingly, that \vmratio{} has very little information
+about the number of divergences among taxa.
 Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
 biogeographically relevant than \numt{}.
-We struggle to follow this logic.
+% We struggle to follow this logic.
 Certainly the maximum information is contained within the divergence time
 vector that \vmratio{} is summarizing.
-The temporal information contained in this vector coupled with its cardinality
-(\numt{}) is certainly more informative than its variance (i.e., the dispersion
-index is not a sufficient statistic for \divtvector).
+Clearly, the number of divergence time parameters within the vector and their
+values is more informative than its variance (i.e., the dispersion index is not
+a sufficient statistic for \divtvector).
 
@@
 better than \numt{}.''
-However, \msb is a model-choice estimate, and hence the goal is to estimate the
+However, \msb is a model-choice estimator, and hence the goal is to estimate the
 posterior probability of divergence models.
@@
 at its lower limit of one.
-Thus, it is not surprising to see this estimator of model probability
-perform better.
+Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better
+estimator of model probability than \vmratio{}.
 
 \section*{Some general thoughts on the model of \msb}
-It is not too surprising that the model-averaging approach proposed by
-\citet{Hickerson2013} behaves poorly.
+Our results demonstrating the poor behavior of the model-averaging approach
+proposed by \citet{Hickerson2013} are not too surprising.
 While this approach is trivial to implement, it is not a trivial change
-to the basic \msb model.
-To better understand why this is, it might help to step back and get a
+to the \msb model.
+To better understand the reason for this, it might help to step back and get a
 sense of the scale of the \msb model.
@@
 timing of these contractions (\bottleTime{}), and the $N-1$ node heights
-(coalescent times) of the gene tree that gives rise to the $N$ gene
+(coalescent times) of the gene tree that gave rise to the $N$ gene
 copies sampled from both populations of the pair.
@@
 581--602 free parameters (depending on the cardinality of \divtvector).
-Under this rich model, the method is sampling over 1002 divergence models
-\citep[i.e., the number of integer partitions of $Y=22$;][]{Oaks2012}.
-
-This is a very tall order, and the method only uses four summary statistics
-calculated from the sequence alignment of each taxon pair:
-$\pi$ \citep{Tajima1983}, $\theta_W$
-\citep{Watterson1975}, $\pi_{net}$ \citep{Takahata1985}, and
-$SD(\pi-\theta_W)$ \citep{Tajima1989}.
-That gives us a total of 88 summary statistics (four for each of the 22 taxon
+Furthermore, under this rich model, the method is sampling over 1002 divergence
+models \citep[i.e., the number of integer partitions of $Y=22$;][]{Oaks2012}.
+
+This is a very difficult inference problem, and the method only uses four
+summary statistics calculated from the sequence alignment of each taxon pair:
+$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
+\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
+That gives us a total of 88 summary statistics (four from each of the 22 taxon
 pairs), which contain minimal information about many of the $\approx 600$
@@
 
-This very large number of parameters and divergence models relative to 
+The large number of parameters and divergence models relative to 
 summary statistics is undoubtedly a key reason the method is so sensitive
@@
 It also likely contributes to the method's lack of robustness.
-We use robustness here as a measure of a method's insensitivity to model
-violations.
+% We use robustness here as a measure of a method's insensitivity to model
+% violations.
 Robustness is an extremely important characteristic of a method to gauge its
@@
 models and prior models.
-This certainly is a non-trivial extension of the model and, given the method's
-very large number of parameters and models relative to the information used
+This is a non-trivial extension of the model and, given the method's very large
+number of parameters and divergence models relative to the information used
 from the data, likely plays a major role in the poor behavior of this approach.
@@
 \citet{Oaks2012} suggest that uniform priors are inappropriate for many
-parameters of the \msb model, and recommend the use of more cogent probability
+parameters of the \msb model, and recommend the use of probability
 distributions from the exponential family.
 If we look at the prior distribution on the divergence time parameter \divt{}
-imposed by the model-averaging prior approach of \citet{Hickerson2013} we see
+imposed by the model-averaging approach of \citet{Hickerson2013} we see
 it is a mixture of overlapping uniforms with lower limits of zero
 (Figure~\ref{figMCTauPrior}).
-This looks very much like an exponential distribution.
-Thus, it would be much more natural and easier to simply place a gamma prior
-(the exponential being a special case of the gamma) on divergence times.
-This would allow the investigator to much better capture their prior
-uncertainty in model parameters while avoiding awkward broad uniform
-distributions, and without the need of costly model-averaging.
-This seems like a much more natural solution to the problem of Lindley's
-paradox and sampling error, which are both certainly exacerbated by the fact
-that uniform distributions are inappropriately the only choice of prior for
-many of the parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
+This looks very much like an exponential distribution, except that each sample
+of divergence times are restricted to the hard bounds of one of the eight prior
+models.
+Thus, it seems more appropriate to place a gamma prior (the exponential being a
+special case of the gamma) on divergence times.
+This would provide investigators flexibility to represent their prior
+uncertainty in model parameters while avoiding broad uniform
+distributions and the need of costly model-averaging.
+This seems like a natural solution to the problem of skewed marginal
+likelihoods and/or sampling error, which are both likely exacerbated by the
+fact that uniform distributions are the only choice of prior for many of the
+parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
 \descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
@@
 Importantly, there is no cost to this strategy.
-Given that ABC methods merely have to draw random values from prior
-distributions (hence there are no difficult proposal probability ratios to
-calculate, etc.) there is no reason for not using the most appropriate
-distributions on parameters with conjugacy in mind.
+Given that ABC methods only need to draw random values from prior distributions
+(hence there are no difficult prior and proposal ratios to calculate, etc.) it
+is easy to use the most appropriate distributions on parameters.
 Also, this approach reduces the temptation of empirically guided priors.
-The investigator can place the majority of the prior density in regions
-of parameters space they believe, a priori, are most plausible, but still
+The investigator can place the majority of the prior density in regions of
+parameter space they believe, \emph{a priori}, are most plausible, but still
 capture uncertainty in the tails of distributions with low density.
-As discussed in \citet{Oaks2012} the use of inflexible and unforgiving uniform
-priors necessitate the need to use broad priors to avoid excluding the truth a
-priori.
+As discussed in \citet{Oaks2012} the use of uniform priors necessitate
+broad priors to avoid excluding the truth \emph{a priori}.
 Thus, to avoid the behavior of the method under these uniform priors, using the
-data is a very tempting, albeit dangerous, option \citep{Hickerson2013}.
+data is a very tempting, albeit risky, option \citep{Hickerson2013}.
 More appropriate prior distributions would alleviate this issue.
@@
 statistics to see if they coincide (see Figure 1 of \citet{Hickerson2013}).
-As we show above, this is a dangerous strategy for choosing priors.
-It is not surprising that the resulting plots of this approach have little
-correlation with the appropriateness of priors.
+As we show above, this is a risky strategy, because the resulting plots of this
+approach have little correlation with the appropriateness of priors.
 Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
@@
 summary statistics consistent with the observed data.
-In fact, when it does, this can be an obvious warning sign that the prior is
-over-fit, as we show above (Table~\ref{tabModelChoiceEmpirical} and
-Figure~\ref{figPCA}).
+In fact, when the random draws are tightly clustered around the observed
+statistics, this can be an indication that the prior is over-fit, as we show
+above (Table~\ref{tabModelChoiceEmpirical} and Figure~\ref{figPCA}).
 Thus, interpreting such plots should be avoided, and the use of posterior
@@
 \emph{a priori}.
-This likely hinders the marginal likelihoods of models with greater parameter
+This likely reduces the marginal likelihoods of models with greater parameter
 space, either due to more divergence time parameters or broader prior
@@
 of the parameters.
-Exploring cogent prior probability distributions on most of the model's
+Exploring alternative prior probability distributions on most of the model's
 parameters, in addition to different priors over divergence models, should
@@
 
-\mFigure{../images/coin_flip.pdf}{
-    A plot of three beta probability density functions that represent a prior
-    (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
-    empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
-    Bernoulli trials, four of which are successes.
-}{figCoinFlip}
+% \mFigure{../images/coin_flip.pdf}{
+%     A plot of three beta probability density functions that represent a prior
+%     (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
+%     empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
+%     Bernoulli trials, four of which are successes.
+% }{figCoinFlip}
 
@@
     The accuracy of the model-averaging approach of \citet{Hickerson2013} to
-    estimate the dispersion index of divergenct times (\vmratio{}) from
+    estimate the dispersion index of divergence times (\vmratio{}) from
     analyses of simulated datasets where \divt{} for 22 pairs of populations is
@@
     uniform distributions, $\divt{} \sim U(0, \divt{max})$.
-    The estimated probability of the method inferring on divergecne event, $p(\numt{} = 1)$,
+    The estimated probability of the method inferring one divergence event, $p(\hat{\numt{}} = 1)$,
     is given for each \divt{max}.

### 2013-10-01 manuscripts/dpp-msbayes/dpp-msbayes.tex
Working on Mark's edits of dpp ms.

@@
     \noindent To understand the processes that generate biodiversity, it is
-    important to account for large-scale processes that effect the evolutionary
+    important to account for large-scale processes that affect the evolutionary
     history of groups of co-distributed populations of organisms.
@@
 Understanding the processes that generate biodiversity and regulate
-community assembly is a major goal of Evolutionary Biology.
+community assembly is a major goal of evolutionary biology.
 Given the dynamic nature of our planet, a major component of this goal is
-understanding how geological and climatic processes can effect the
+understanding how geological and climatic processes can affect the
 evolutionary history of whole groups of co-distributed species and their
@@
 At recent timescales, such events should leave a signature in the genetic
-variation within and among the effected lineages.
+variation within and among the affected lineages.
 Thus, being able to accurately estimate models of shared evolutionary
@@
 evolved according to fixed \hky substitution model parameters \hkyModel{i}{j}.
-The investigtor must specify the parameters of all
+The investigator must specify the parameters of all
 $\hkyModelVector = (\hkyModel{1}{1}, \ldots,
@@
 $i$ given the gene tree and \hky substitution model parameters \citep[i.e., the
-``Felsenstein likelhood'';][]{Felsenstein1981}.
+``Felsenstein likelihood'';][]{Felsenstein1981}.
 The model allows for an intra-locus recombination rate \recombinationRate,
@@
 and gives rise to the two sampled descendant populations.
-Each \demographicParams contains the following demographic parameters:
+Each \demographicParams{} contains the following demographic parameters:
 The mutation-scaled sizes of the ancestral, \ancestralTheta{}, and descendant
@@
 the proportion of the first, \bottleScalar{1}{}, and second population,
-\bottleScalar{2}{}, that persists during bottlenecks that begin immmediately
+\bottleScalar{2}{}, that persist during bottlenecks that begin immediately
 after divergence in forward-time;
@@
 \begin{equation}
-    \myTheta{C} = E(\descendantTheta{}{}).
+    \myTheta{C} = \mathbb{E}(\descendantTheta{}{}).
     \label{eq:thetaC}
@@
 \end{equation}
-The \divTime{} parameters are then in units $\myTheta{C}/\mutationRate$
+The \divTime{} parameters are in units $\myTheta{C}/\mutationRate$
 generations, which we denote as \globalcoalunit generations.
@@
 Thus, for each of the divergence times in \divTimeVector to be on the same
-scale, and thus comparable, the relative mutation rates among the pairs of
-populations are assumed to be fixed and known according the user-provided
-values in \mutationRateScalarConstantVector.
+scale, the relative mutation rates among the pairs of populations are assumed
+to be fixed and known according the user-provided values in
+\mutationRateScalarConstantVector.
 \end{linenomath}
@@
 \priorDPP{}$.
-We take a hierarchical approach and use a second-order prior probability
+We take a hierarchical approach and use a prior probability
 distribution (i.e., hyperprior) for \concentrationParam.
@@
 \begin{equation}
-    p(\divTimeNum \given \concentrationParam, \npairs{}) = 
+    p(\divTimeNum \boldsymbol{\mid} \concentrationParam, \npairs{}) = 
     \frac{\stirlingFirst{\npairs{}}{\divTimeNum} \concentrationParam^{\divTimeNum}}
@@
 divergence parameters (denoted \divTimeNum here, and as \numt{} in
-\citet{Huang2011}) imposes an odd U-shaped prior over discrete divergence
+\citet{Huang2011}) imposes an U-shaped prior over discrete divergence
 models (\divTimeIndexVector; see
@@
     \descendantTheta{1}{}, \descendantTheta{2}{} \sim
-    Dirichlet(1,1) \times 2 \times U(\uniformMin{\myTheta{}},
+    Beta(1,1) \times 2 \times U(\uniformMin{\myTheta{}},
     \uniformMax{\descendantTheta{}{}}),

### 2013-10-01 manuscripts/dpp-msbayes/dpp-msbayes.tex
Incorporating more edits of dpp ms from Mark.

@@
 of our model is given by
+\begin{equation}
+    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
+    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
+    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
+    \mutationRateScalarConstantVector) =
+    \frac{p(\alignmentVector \given \geneTreeVector, \divTimeMapVector,
+        \demographicParamVector, \locusMutationRateScalarVector,
+        \locusRateHetShapeParameter, \hkyModelVector, \ploidyScalarVector,
+        \mutationRateScalarConstantVector)
+        p(\geneTreeVector, \divTimeMapVector, \demographicParamVector,
+        \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
+        \hkyModelVector, \ploidyScalarVector,
+        \mutationRateScalarConstantVector)}{
+        p(\alignmentVector)},
+    \label{eq:fullModelCompactJoint}
+\end{equation}
+which can be expanded out into components that are assumed to be independent
+to get
 \begin{equation}
@@
 \subsubsection*{Flexibility in parameterizing the model}
-We also provide the user with the ability to control the richness of
-the model.
+In our implementation, we provide the ability to control the richness of the
+model.
 For the \myTheta{} parameters, the model can be fully generalized to
@@
 
-We also allow the user to eliminate the parameters associated with
-the post-divergence bottlenecks in the descendant populations of
-each pair
+We also provide the option of eliminating the parameters associated with the
+post-divergence bottlenecks in the descendant populations of each pair
 (\bottleTime{}, \bottleScalar{1}{}, and \bottleScalar{2}{}),
@@
 constant size from present back to the divergence event.
-Also, rather than eliminate the bottleneck parameters, the user
-can constrain \bottleScalar{1}{} and \bottleScalar{2}{} to be
-equal to remove one free parameter from the model for each of
-the population pairs.
-
-Overall, our implementation allows the user to specify a model that has as many
-as seven parameters per population pair
+Also, rather than eliminate the bottleneck parameters,
+we allow \bottleScalar{1}{} and \bottleScalar{2}{} to be constrained to be
+equal, which removes one free parameter from the model for each of the
+population pairs.
+
+Overall, our implementation allows an investigator to specify a model that has
+as many as seven parameters per population pair
 (\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{},
@@
 \subsubsection*{Ordering of taxon-specific summary statistics}
-As eluded to in the model description, \msb does not maintain the order of the
+As alluded to in the model description, \msb does not maintain the order of the
 taxon-specific summary statistics \alignmentSS{}{} within each \ssVector{}.
-Rather, they, along with the observed summary statistics, are re-ordered by
-descending values of average pairwise differences between the descendant
-populations \citep[$\pi_b$;][]{NeiLi1979,Huang2011}.
+Rather, the summary statistics are re-ordered by descending values of average
+pairwise differences between the descendant populations
+\citep[$\pi_b$;][]{NeiLi1979,Huang2011}.
 This has the advantage of reducing the sample space of the number of discrete
@@
 (e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
-In such cases it seems possible that discarding this information could
+In such cases, it seems possible that discarding this information could
 cause bias.
-Furthermore, given that part of the motivation for re-ordering by $\pi_b$ is to
-minimize Euclidean distances between simulated datasets in which the true model
-has a single divergence event, and this reordering increases the model's
-tendency to infer a single divergence event \citep{Huang2011}, it seems
+Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
+is to minimize Euclidean distances between simulated datasets in which the true
+model has a single divergence event, and (2) the reordering increases the
+model's tendency to infer a single divergence event \citep{Huang2011}, it seems
 possible that this approach could be biasing the method \citep{Oaks2012}.
@@
 \subsection*{Validation analyses: Ordered divergence models}
-Our results show that the model estimating ordered models of divergence
-\modelDPPOrdered performs similarly to its counterpart sampling over
-unordered models (\modelDPP) when the prior is correct (Figs.\
+Our results show that the ordered models of divergence, \modelDPPOrdered,
+perform similarly to sampling over unordered models (\modelDPP) when the prior
+is correct (Figs.\
 \labelcref{figValidationAccuracyDPPOrderedDPPOrdered,figValidationModelChoiceDPPOrderedDPPOrdered}).
+This suggests that the method is not adversely affected by the large increase
+in possible discrete models (from 22 unordered to 4140 ordered models) when
+there are 8 pairs of populations.
 

### 2013-10-02 manuscripts/dpp-msbayes/dpp-msbayes.tex
More editing of dpp ms.

@@
 Such large-scale processes predict patterns of shared evolutionary history
-across effected populations of organisms, e.g., temporally clustered
+across affected populations of organisms, e.g., temporally clustered
 divergences across co-distributed species.
@@
 combination of broad, uniformly distributed priors on nuisance parameters and a
-U-shaped prior on divergence models were causing the bias.
+U-shaped prior on divergence models are causing the bias.
 The model implemented in \msb uses a discrete uniform prior over the number of
@@
 likely hinder the marginal likelihoods of parameter-rich models, i.e., models
-with many divergence time parameters \citep[this is often referred to as
-Lindley's paradox;][]{Lindley1957}.
+with many divergence time parameters \citep{Lindley1957}.
 Thus, together, the priors on divergence models and nuisance parameters used in
@@
 
-Following the recommendations of \citet{Oaks2012} we modify the \msb model to
+Following the recommendations of \citet{Oaks2012}, we modify the \msb model to
 test whether alternative parameterizations and priors will improve the behavior
-of the approximate Bayesian model-choice approach.
+of the approximate-Bayesian model-choice approach.
 We implement a Dirichlet-process prior over all possible models of divergence,
@@
 gene trees.
-Furthermore, the investigator must specify a vector of fixed contants
+Furthermore, the investigator must specify a vector of fixed constants
 $\ploidyScalarVector = (\ploidyScalar{1}{1}, \ldots,
 \ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$
-that scale the population size parameters for known differences in ploidy among
-loci and/or differences in generation times among population pairs.
-Lastly, the investigator must also specify a vector of fixed constants
+that scale the population-size parameters for known differences in ploidy among
+loci and/or differences in generation times among population pairs.  Lastly,
+the investigator must also specify a vector of fixed constants
 $\mutationRateScalarConstantVector = (\mutationRateScalarConstant{1}{1},
 \ldots, \mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$
-that scale the population size parameters for known differences in
+that scale the population-size parameters for known differences in
 mutation rates among loci and/or among taxa.
@@
 \end{equation}
-where each \locusMutationRateScalar{} is independently and identicially
+where each \locusMutationRateScalar{} is independently and identically
 distributed (\iid) as
@@
 \end{equation}
-The \divTime{} parameters are in units $\myTheta{C}/\mutationRate$
+The \divTime{} parameters are in units of $\myTheta{C}/\mutationRate$
 generations, which we denote as \globalcoalunit generations.
@@
 scale, the relative mutation rates among the pairs of populations are assumed
-to be fixed and known according the user-provided values in
+to be fixed and known according to the user-provided values in
 \mutationRateScalarConstantVector.
@@
 distribution (i.e., hyperprior) for \concentrationParam.
-More specifically, we use a gamma-distributed prior $\concentrationParam =
+More specifically, we use a gamma-distributed prior $\concentrationParam \sim
 Gamma(\gammaShape{\concentrationParam}, \gammaScale{\concentrationParam})$,
@@
 is the prior distribution on divergence models.
-As discussed in \citet{Oaks2012}, in \msb the only prior used for
+As discussed in \citet{Oaks2012}, in \msb the prior used for
 \divTimeIndexVector is a combination of a discrete uniform prior over the
@@
 \bottleScalar{2}{}).
-Not only do these priors have nice conjugacy properties, but they also provide
-the user with much greater flexibility in expressing prior uncertainty
-regarding the parameters of the model.
+These priors provide the user with much greater flexibility in expressing prior
+uncertainty regarding the parameters of the model.
 
@@
 The first step of this algorithm entails collecting a random sample of
-parameters values from the joint prior and their associated summary
+parameter values from the joint prior and their associated summary
 statistics.
@@
 $\ssVector{} = (\alignmentSS{1}{1}, \ldots, \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$
-from the simulated sequence alignments (these are the same statistics
-calculated from the observed alignments, \ssVectorObs{});
+from the simulated sequence alignments;
+% (these are the same statistics
+% calculated from the observed alignments, \ssVectorObs{});
 and (6) reducing the summary statistics to the means across loci for each
@@
 are standardized using the means and standard deviations of the statistics from
-the prior sample (i.e., the prior mean is subtracted from each statistic, and then
-it is divided by the prior standard deviation).
+the prior sample (i.e., the prior mean is subtracted from each statistic, and the
+difference is divided by the prior standard deviation).
 After all statistics are standardized, the Euclidean distance between
@@
 drawn from a series of of exponential distributions: $Exp(17.3205)$,
-$Exp(8.6603)$, $Exp(5.7735)$, $Exp(4.33013)$, $Exp(3.4641)$, and $Exp(1.7321)$.
+$Exp(8.6603)$, $Exp(5.7735)$, $Exp(4.3301)$, $Exp(3.4641)$, and $Exp(1.7321)$.
 These exponential distributions have the same variance as their uniform
@@
 This suggests that the method is not adversely affected by the large increase
-in possible discrete models (from 22 unordered to 4140 ordered models) when
-there are 8 pairs of populations.
+in the number of possible discrete models (from 22 unordered to 4140 ordered
+models) when there are eight pairs of populations.
 
@@
 \ref{figPowerAccuracyUniformUshaped}).
-This pattern of results across the four models is very similar when applied
-to data simulated under the series of models of \powerSeriesOld (Figs.\
+This pattern of results across the four models is very similar to their
+behavior when applied to data simulated under the series of models of
+\powerSeriesOld (Figs.\
 \labelcref{figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
@@
 When looking at the estimated posterior probability of the single-divergence
-model across the power analyses, we also see the increased power of the
-new models (Figs.\
+model across the power analyses, we also see increased power of the new models
+(Figs.\
 \labelcref{figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped}).
@@
 and divergence models improved the behavior of the \msb model.
-The model-choice estimation of our implementation is more accurate and,
-perhaps more importantly, shows greater robustness to model
-violations (Figs.\
+In our implementation, model-choice estimation is more accurate and shows
+greater robustness to model violations (Figs.\
 \labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
@@
 under-estimate the probability of biogeographically interesting models of
-temporally clustered divergences 
+temporally clustered divergences (Figs.\
 \labelcref{figValidationModelChoiceUshapedDPP,figValidationModelChoiceUshapedUniform}).
@@
 the \msb model to infer strong support for clustered divergences when
-divergences are random over quite broad timescales.
+divergences are random over quite broad timescales (Figs.\
 \labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld,figPowerDispersionProbExpOld,figPowerDispersionProbUniformOld,figPowerDispersionProbOldOld}).
 Our results expand on this by showing that this behavior is consistent
-across of range conditions underlying the data.
+across a range conditions underlying the data.
 When using the new priors, the method is less prone to spurious inference of
@@
 model when there are model violations is reduced when the exponential
-priors are used in place of the uniform priors.
+priors are used in place of the uniform priors (Figs.\
 \labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped}).
-Furthermore, better priors on on nuisance parameters improves the method's
+Furthermore, better priors on nuisance parameters improves the method's
 power to detect temporally random divergences.
@@
 The method attempts to approximate the posterior of a very parameter rich model
-using little information from the data.
+using relatively little information from the data.
 For example, when applied to the dataset of Philippine vertebrates analyzed by
@@
 parameters.
-Sampling over hundreds of parameters and more than a thousand models is asking
-a lot from a few summary statistics calculated from the sequence alignments of
-each taxon pair.
-These simple summary statistics contain little information about many
-of the free parameters in the model.
+Sampling over hundreds of parameters and more than a thousand models is a
+difficult inference problem, especially when only using a small number of
+summary statistics calculated from the sequence alignments of each taxon pair.
+The population-genetic summary statistics used by our method contain little
+information about many of the free parameters in the model.
 Thus, we expect the improved method will still be sensitive to priors, and the
@@
 We encourage investigators to view this method and its predecessor as a means
-of exploring their data rather than as rigorous means of evaluating hypotheses.
+of exploring their data rather than a rigorous means of evaluating hypotheses.
 
@@
 full-likelihood methods that can leverage all of the information present in the
-sequence data will become increasingly important in robustly estimating the
+sequence data will become increasingly important for robustly estimating the
 temporal history of divergences across taxa \citep{JeetDiss}.
@@
 likelihood calculations \citep{Ayres2012}, analyzing rich comparative
-phylogeograpical models in full-likelihood Bayesian framework is becoming
+phylogeograpical models in a full-likelihood Bayesian framework is becoming
 computationally practical, especially when considering that the computational

### 2013-10-07 manuscripts/dpp-msbayes/dpp-msbayes.tex
Fixing error in methods for dpp ms.
Our prior on the concentration parameter of the Dirichlet process was
different between the validation and power analyses, because the
simulated data structure was different (8 pairs vs 22 pairs).

@@
 \citet{Oaks2012}.
-For all of our power simulations, we used a simulated data structure identical
-to that of the empirical dataset of Philippine vertebrates analyzed by
-\citet{Oaks2012}, which consists of 22 pairs of populations.
+The only difference from the prior models used in the validation analyses above
+is the prior on the concentration parameter \concentrationParam for the
+\modelDPP model.
+Rather than the prior
+$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
+we use the prior
+$\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
+over the discrete divergence models for the model \modelDPP.
+The reason for this difference is because for all of our power simulations, we
+used a simulated data structure identical to that of the empirical dataset of
+Philippine vertebrates analyzed by \citet{Oaks2012}, which consists of 22 pairs
+of populations.
 For each of the four models, we generated $2\e6$ samples from the prior.
@@
     %\captionsetup{font=footnotesize}
-    \caption{The model priors evaluated in our simulation-based analyses.}
+    \caption{The model priors evaluated in our simulation-based analyses.  For
+        the \modelDPP model, the prior on the concentration parameter,
+        $\concentrationParam \sim Gamma(\cdot,\cdot)$, was set to $Gamma(2,2)$
+        for the validation analyses and $Gamma(1.5,18.1)$ for the power
+        analyses.}
     \centering
@@
                       & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$
+            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
                       & $\divTime{} \sim Exp(0.3464)$

### 2013-10-07 empirical-analyses/bin/dmc-negros-panay.sh
Adding `empirical-analyses` project.

@@ -1,0 +1,57 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=16G
+#$ -l vf=16G
+#$ -q all.q
+#$ -pe orte 8
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
+
+
+nprocs=8
+nprior=10000000
+batch_size=25000
+nsums=200000
+npost=10000
+nquantiles=10000
+reporting_freq=10
+sort_index=0
+seed=845225390
+
+output_dir="../results/negros-panay/dpp-ordered"
+if [ ! -d "$output_dir" ]
+then
+    mkdir -p $output_dir
+fi
+
+dmc.py --np $nprocs \
+    -o ../configs/negros-panay.cfg \
+    -p ../configs/negros-panay.cfg \
+    -n $nprior \
+    --prior-batch-size $batch_size \
+    --num-posterior-samples $npost \
+    --num-standardizing-samples $nsums \
+    -q $nquantiles \
+    --reporting-frequency $reporting_freq \
+    --sort-index $sort_index \
+    --output-dir $output_dir \
+    --staging-dir $staging_dir \
+    --temp-dir $staging_dir \
+    --compress \
+    --seed $seed
+
+echo "Here are the contents of the local temp directory '${staging_dir}':"
+ls -Fla $staging_dir
+echo 'Removing the local temp directory...'
+rm -r $staging_dir
+echo 'Done!'
+

### 2013-10-07 empirical-analyses/bin/dmc-philippines-dpp.sh

@@ -1,0 +1,57 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=16G
+#$ -l vf=16G
+#$ -q all.q
+#$ -pe orte 8
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
+
+
+nprocs=8
+nprior=10000000
+batch_size=25000
+nsums=200000
+npost=10000
+nquantiles=10000
+reporting_freq=10
+sort_index=7
+seed=238691091
+
+output_dir="../results/philippines/dpp"
+if [ ! -d "$output_dir" ]
+then
+    mkdir -p $output_dir
+fi
+
+dmc.py --np $nprocs \
+    -o ../configs/philippines-dpp.cfg \
+    -p ../configs/philippines-dpp.cfg \
+    -n $nprior \
+    --prior-batch-size $batch_size \
+    --num-posterior-samples $npost \
+    --num-standardizing-samples $nsums \
+    -q $nquantiles \
+    --reporting-frequency $reporting_freq \
+    --sort-index $sort_index \
+    --output-dir $output_dir \
+    --staging-dir $staging_dir \
+    --temp-dir $staging_dir \
+    --compress \
+    --seed $seed
+
+echo "Here are the contents of the local temp directory '${staging_dir}':"
+ls -Fla $staging_dir
+echo 'Removing the local temp directory...'
+rm -r $staging_dir
+echo 'Done!'
+

### 2013-10-07 empirical-analyses/bin/dmc-philippines-inform.sh

@@ -1,0 +1,57 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=16G
+#$ -l vf=16G
+#$ -q all.q
+#$ -pe orte 8
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
+
+
+nprocs=8
+nprior=10000000
+batch_size=25000
+nsums=200000
+npost=10000
+nquantiles=10000
+reporting_freq=10
+sort_index=7
+seed=33741111
+
+output_dir="../results/philippines/dpp-inform"
+if [ ! -d "$output_dir" ]
+then
+    mkdir -p $output_dir
+fi
+
+dmc.py --np $nprocs \
+    -o ../configs/philippines-dpp-inform.cfg \
+    -p ../configs/philippines-dpp-inform.cfg \
+    -n $nprior \
+    --prior-batch-size $batch_size \
+    --num-posterior-samples $npost \
+    --num-standardizing-samples $nsums \
+    -q $nquantiles \
+    --reporting-frequency $reporting_freq \
+    --sort-index $sort_index \
+    --output-dir $output_dir \
+    --staging-dir $staging_dir \
+    --temp-dir $staging_dir \
+    --compress \
+    --seed $seed
+
+echo "Here are the contents of the local temp directory '${staging_dir}':"
+ls -Fla $staging_dir
+echo 'Removing the local temp directory...'
+rm -r $staging_dir
+echo 'Done!'
+

### 2013-10-07 empirical-analyses/bin/dmc-philippines-simple.sh

@@ -1,0 +1,57 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=16G
+#$ -l vf=16G
+#$ -q all.q
+#$ -pe orte 8
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
+
+
+nprocs=8
+nprior=10000000
+batch_size=25000
+nsums=200000
+npost=10000
+nquantiles=10000
+reporting_freq=10
+sort_index=7
+seed=590656225
+
+output_dir="../results/philippines/dpp-simple"
+if [ ! -d "$output_dir" ]
+then
+    mkdir -p $output_dir
+fi
+
+dmc.py --np $nprocs \
+    -o ../configs/philippines-dpp-simple.cfg \
+    -p ../configs/philippines-dpp-simple.cfg \
+    -n $nprior \
+    --prior-batch-size $batch_size \
+    --num-posterior-samples $npost \
+    --num-standardizing-samples $nsums \
+    -q $nquantiles \
+    --reporting-frequency $reporting_freq \
+    --sort-index $sort_index \
+    --output-dir $output_dir \
+    --staging-dir $staging_dir \
+    --temp-dir $staging_dir \
+    --compress \
+    --seed $seed
+
+echo "Here are the contents of the local temp directory '${staging_dir}':"
+ls -Fla $staging_dir
+echo 'Removing the local temp directory...'
+rm -r $staging_dir
+echo 'Done!'
+

### 2013-10-07 empirical-analyses/bin/dmc-philippines-uniform.sh

@@ -1,0 +1,57 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=16G
+#$ -l vf=16G
+#$ -q all.q
+#$ -pe orte 8
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
+
+
+nprocs=8
+nprior=10000000
+batch_size=25000
+nsums=200000
+npost=10000
+nquantiles=10000
+reporting_freq=10
+sort_index=7
+seed=519088215
+
+output_dir="../results/philippines/uniform"
+if [ ! -d "$output_dir" ]
+then
+    mkdir -p $output_dir
+fi
+
+dmc.py --np $nprocs \
+    -o ../configs/philippines-uniform.cfg \
+    -p ../configs/philippines-uniform.cfg \
+    -n $nprior \
+    --prior-batch-size $batch_size \
+    --num-posterior-samples $npost \
+    --num-standardizing-samples $nsums \
+    -q $nquantiles \
+    --reporting-frequency $reporting_freq \
+    --sort-index $sort_index \
+    --output-dir $output_dir \
+    --staging-dir $staging_dir \
+    --temp-dir $staging_dir \
+    --compress \
+    --seed $seed
+
+echo "Here are the contents of the local temp directory '${staging_dir}':"
+ls -Fla $staging_dir
+echo 'Removing the local temp directory...'
+rm -r $staging_dir
+echo 'Done!'
+

### 2013-10-07 empirical-analyses/bin/dmc-negros-panay-unordered.sh
Adding command line script to `empirical-analyses`.
Adding script for analysis of unordered models for negros-panay
data.

@@ -1,0 +1,57 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=16G
+#$ -l vf=16G
+#$ -q all.q
+#$ -pe orte 8
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
+
+
+nprocs=8
+nprior=10000000
+batch_size=25000
+nsums=200000
+npost=10000
+nquantiles=10000
+reporting_freq=10
+sort_index=7
+seed=845225390
+
+output_dir="../results/negros-panay/dpp-ordered"
+if [ ! -d "$output_dir" ]
+then
+    mkdir -p $output_dir
+fi
+
+dmc.py --np $nprocs \
+    -o ../configs/negros-panay.cfg \
+    -p ../configs/negros-panay.cfg \
+    -n $nprior \
+    --prior-batch-size $batch_size \
+    --num-posterior-samples $npost \
+    --num-standardizing-samples $nsums \
+    -q $nquantiles \
+    --reporting-frequency $reporting_freq \
+    --sort-index $sort_index \
+    --output-dir $output_dir \
+    --staging-dir $staging_dir \
+    --temp-dir $staging_dir \
+    --compress \
+    --seed $seed
+
+echo "Here are the contents of the local temp directory '${staging_dir}':"
+ls -Fla $staging_dir
+echo 'Removing the local temp directory...'
+rm -r $staging_dir
+echo 'Done!'
+

### 2013-10-07 empirical-analyses/bin/dmc-negros-panay-unordered.sh
Updating command-line script for unordered analysis of negros-panay

@@ -27,7 +27,7 @@ reporting_freq=10
 sort_index=7
 seed=845225390
 
-output_dir="../results/negros-panay/dpp-ordered"
+output_dir="../results/negros-panay/dpp-unordered"
 if [ ! -d "$output_dir" ]
 then
     mkdir -p $output_dir

### 2013-10-08 slides/bin/marginal_plot.py
Adding script to `slides/bin`.
Adding script for creating plot to serve as an example of how broad
priors can affect marginal likelihoods.

@@ -1,0 +1,112 @@
+#! /usr/bin/env python
+
+import os
+import sys
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib
+import numpy
+from matplotlib import pyplot as plt
+from matplotlib import cm
+
+def get_bivariate_normal_and_uniform_densities(maximum = 20.0,
+        mean = (2.0, 3.0),
+        variance = (0.2, 0.2),
+        covariance = 0.0,
+        npoints = 50):
+    a = numpy.linspace(0, maximum, npoints)
+    b = numpy.linspace(0, maximum, npoints)
+    X, Y = numpy.meshgrid(a, b)
+    Z1 = get_bivariate_normal_density(X, Y,
+            mean = mean,
+            variance = variance,
+            covariance = covariance)
+    Z2 = (Z1 * 0.0) + (1.0 / (maximum ** 2))
+    return X, Y, Z1, Z2
+
+def get_bivariate_normal_density(x, y,
+        mean = (2.0, 3.0),
+        variance = (0.2, 0.2),
+        covariance = 0.0):
+    return matplotlib.mlab.bivariate_normal(x, y,
+            sigmax = variance[0],
+            sigmay = variance[1],
+            mux = mean[0],
+            muy = mean[1],
+            sigmaxy = covariance)
+
+def get_marginal_likelihood(x, y, z):
+    max_x, max_y = 0.0, 0.0
+    for i in x:
+        max_x = max([max_x] + [max(i)])
+    for i in y:
+        max_y = max([max_y] + [max(i)])
+    prior = 1.0 / (max_x * max_y)
+    l = 0.0
+    w = 0.0
+    for i in range(len(z)):
+        for j in range(len(z[0])):
+            l += (z[i][j] * prior)
+            w += prior
+    return l/w
+
+def get_marginal_likelihood_constrained(x, y, z):
+    assert len(x) == len(y)
+    max_x, max_y = 0.0, 0.0
+    for i, a in enumerate(x):
+        assert len(x[i]) == len(y[i])
+        max_x = max([max_x] + [max(a)])
+    for a in y:
+        max_y = max([max_y] + [max(a)])
+    assert max_x == max_y
+    prior = 1.0 / max_x
+    l = 0.0
+    w = 0.0
+    for i in range(len(z)):
+        l += (z[i][i] * prior)
+        w += prior
+    return l/w
+
+def get_marginal_plot(maximum = 20.0,
+        mean = (2.0, 3.0),
[…]

### 2013-10-08 slides/bin/marginal_plot.py
Working on `marginal_plot.py` script.

@@ -82,28 +82,46 @@ def get_marginal_plot(maximum = 20.0,
     sys.stdout.write('marginal likelihood of 1-parameter model: {0}\n'.format(ml_1p))
     fig = plt.figure()
     ax = fig.add_subplot(111, projection = '3d')
-    ax.plot_surface(X, Y, Z1, rstride=1, cstride=1)#, cmap=cm.YlGnBu_r)
+    ax.plot_surface(X, Y, Z1, rstride=1, cstride=1, linewidth=0.0, antialiased=False, shade=True, cmap=cm.coolwarm, zorder=200)
     a, b, c = [], [], []
     for i in range(len(X)):
         a.append(X[i][i])
         b.append(Y[i][i])
         c.append(Z1[i][i])
+    xmin, xmax = ax.get_xlim()
+    ymin, ymax = ax.get_ylim()
+    prior_d = 1.0 / (xmax * ymax)
+    prior_d *= 2.0
+    x_back_line = ax.plot([xmin, xmax], [ymax, ymax], [prior_d, prior_d])
+    x_front_line = ax.plot([xmin, xmax], [ymin, ymin], [prior_d, prior_d], zorder=200)
+    y_back_line = ax.plot([xmin, xmin], [ymin, ymax], [prior_d, prior_d], zorder=-10)
+    y_front_line = ax.plot([xmax, xmax], [ymin, ymax], [prior_d, prior_d], zorder=200)
+    plt.setp([x_back_line, y_back_line, x_front_line, y_front_line],
+            color = 'r',
+            linestyle = '--',
+            linewidth = 1.0,
+            marker = '')
     identity_line = ax.plot(a, b, c)
     plt.setp(identity_line,
-            color = 'r',
+            color = 'w',
             linestyle = '-',
-            linewidth = 1.0,
+            linewidth = 0.75,
             marker = '',
             zorder = 100)
+    ax.set_xlabel(r'$T_1$', size=14.0)
+    ax.set_ylabel(r'$T_2$', size=14.0)
+    ax.set_zlabel('Density', size=14.0)
+    rect = [-0.12, 0, 1, 1.02]
+    fig.tight_layout(pad = 0.25, rect = rect)
     return ax, fig
 
 def main_cli():
-    maximum = 20.0
+    maximum = 1.0
     ax, fig = get_marginal_plot(maximum = maximum,
-            mean = (10.0, 11.0),
-            variance = (0.3, 0.3),
+            mean = (0.15, 0.247),
[…]

### 2013-10-08 slides/bin/marginal_plot.py
Working on slides.
Updating `marginal_plot.py` script to generate series of 2D and 3D
plots, and adding those plots. Also, removing R plots they are
replacing.

Adding new plots to `slides.tex`, and removing all uses of
'Lindley's paradox.'

@@ -7,12 +7,13 @@ import matplotlib
 import numpy
 from matplotlib import pyplot as plt
 from matplotlib import cm
+from scipy.stats import gamma
 
-def get_bivariate_normal_and_uniform_densities(maximum = 20.0,
-        mean = (2.0, 3.0),
-        variance = (0.2, 0.2),
+def get_bivariate_normal_and_uniform_densities(maximum = 1.0,
+        mean = (0.15, 0.247),
+        variance = (0.039, 0.026),
         covariance = 0.0,
-        npoints = 50):
+        npoints = 100):
     a = numpy.linspace(0, maximum, npoints)
     b = numpy.linspace(0, maximum, npoints)
     X, Y = numpy.meshgrid(a, b)
@@ -66,11 +67,89 @@ def get_marginal_likelihood_constrained(x, y, z):
         w += prior
     return l/w
 
-def get_marginal_plot(maximum = 20.0,
-        mean = (2.0, 3.0),
-        variance = (0.2, 0.2),
+def get_marginal_plot_2d(maximum = 1.0,
+        likelihood_shape = 50.0,
+        likelihood_scale = 0.002,
+        prior_shape = 3.0,
+        prior_scale = 0.06,
+        npoints = 500,
+        include_uniform_prior = True,
+        include_gamma_prior = True,
+        linewidth = 2.0,
+        prior_label_x = 0.5):
+    x = numpy.linspace(0.0000001, maximum, npoints)
+    likelihood = gamma(likelihood_shape, scale = likelihood_scale)
+    y = [likelihood.pdf(i) for i in x]
+    fig = plt.figure()
+    ax = fig.add_subplot(1,1,1)
+    likelihood_line = ax.plot(x, y)
+    plt.setp(likelihood_line,
+            color = '0.3',
+            linestyle = '-',
+            linewidth = linewidth,
+            marker = '',
+            zorder = 200)
+    max_idx = y.index(max(y)) + int(round(0.01 * npoints))
+    label_target = (x[max_idx], y[max_idx])
+    label_position = (label_target[0] + (0.08 * maximum), label_target[1])
+    plt.annotate(r'$p(X \mid \, \theta)$',
+            xy = label_target,
+            arrowprops = dict(arrowstyle = '->'),
+            xytext = label_position,
+            size = 18.0)
+    prior_x = prior_label_x * maximum
+    u_density = 1.0 / maximum
+    ymax = max(ax.get_ylim())
+    prior_label_position = (prior_x, u_density + (0.1 * ymax))
[…]

### 2013-10-08 empirical-analyses/bin/dmc-negros-panay-unordered.sh
Updating commandline scripts for `empirical-analyses`.
Increasing the number of prior samples for analyses of the Philippines
dataset.

@@ -18,12 +18,12 @@ staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
 
 
 nprocs=8
-nprior=10000000
+nprior=50000000
 batch_size=25000
 nsums=200000
 npost=10000
 nquantiles=10000
-reporting_freq=10
+reporting_freq=50
 sort_index=7
 seed=845225390
 

### 2013-10-08 empirical-analyses/bin/dmc-negros-panay.sh

@@ -18,12 +18,12 @@ staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
 
 
 nprocs=8
-nprior=10000000
+nprior=50000000
 batch_size=25000
 nsums=200000
 npost=10000
 nquantiles=10000
-reporting_freq=10
+reporting_freq=50
 sort_index=0
 seed=845225390
 

### 2013-10-08 empirical-analyses/bin/dmc-philippines-dpp.sh

@@ -18,12 +18,12 @@ staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
 
 
 nprocs=8
-nprior=10000000
+nprior=20000000
 batch_size=25000
 nsums=200000
 npost=10000
 nquantiles=10000
-reporting_freq=10
+reporting_freq=20
 sort_index=7
 seed=238691091
 

### 2013-10-08 empirical-analyses/bin/dmc-philippines-inform.sh

@@ -18,12 +18,12 @@ staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
 
 
 nprocs=8
-nprior=10000000
+nprior=20000000
 batch_size=25000
 nsums=200000
 npost=10000
 nquantiles=10000
-reporting_freq=10
+reporting_freq=20
 sort_index=7
 seed=33741111
 

### 2013-10-08 empirical-analyses/bin/dmc-philippines-simple.sh

@@ -18,12 +18,12 @@ staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
 
 
 nprocs=8
-nprior=10000000
+nprior=20000000
 batch_size=25000
 nsums=200000
 npost=10000
 nquantiles=10000
-reporting_freq=10
+reporting_freq=20
 sort_index=7
 seed=590656225
 

### 2013-10-08 empirical-analyses/bin/dmc-philippines-uniform.sh

@@ -18,12 +18,12 @@ staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
 
 
 nprocs=8
-nprior=10000000
+nprior=20000000
 batch_size=25000
 nsums=200000
 npost=10000
 nquantiles=10000
-reporting_freq=10
+reporting_freq=20
 sort_index=7
 seed=519088215
 

### 2013-10-09 slides/bin/marginal_plot.py
Replacing 3D marginal PDFs with PNGs.

@@ -149,7 +149,8 @@ def get_marginal_plot_3d(maximum = 1.0,
         covariance = 0.0,
         npoints = 100,
         include_prior = True,
-        include_constrained_density = True):
+        include_constrained_density = True,
+        linewidth = 0.1):
     X, Y, Z1, Z2 = get_bivariate_normal_and_uniform_densities(maximum = maximum,
             mean = mean,
             variance = variance,
@@ -161,7 +162,7 @@ def get_marginal_plot_3d(maximum = 1.0,
     sys.stdout.write('marginal likelihood of 1-parameter model: {0}\n'.format(ml_1p))
     fig = plt.figure()
     ax = fig.add_subplot(111, projection = '3d')
-    ax.plot_surface(X, Y, Z1, rstride=1, cstride=1, linewidth=0.0, antialiased=False, shade=True, cmap=cm.coolwarm, zorder=200)
+    ax.plot_surface(X, Y, Z1, rstride=1, cstride=1, linewidth=linewidth, antialiased=False, shade=True, cmap=cm.coolwarm, zorder=200)
     if include_prior:
         xmin, xmax = ax.get_xlim()
         ymin, ymax = ax.get_ylim()
@@ -204,8 +205,9 @@ def main_cli():
             covariance=0.0,
             npoints = 100,
             include_prior = True,
-            include_constrained_density = True)
-    fig.savefig('../images/marginal-plot-3d.pdf')
+            include_constrained_density = True,
+            linewidth=0.1)
+    fig.savefig('../images/marginal-plot-3d.png', dpi=300)
 
     ax, fig = get_marginal_plot_3d(maximum = maximum,
             mean = (0.15, 0.247),
@@ -214,7 +216,7 @@ def main_cli():
             npoints = 100,
             include_prior = False,
             include_constrained_density = False)
-    fig.savefig('../images/marginal-plot-3d-bare.pdf')
+    fig.savefig('../images/marginal-plot-3d-bare.png', dpi=300)
 
     ax, fig = get_marginal_plot_3d(maximum = maximum,
             mean = (0.15, 0.247),
@@ -223,16 +225,7 @@ def main_cli():
             npoints = 100,
             include_prior = True,
             include_constrained_density = False)
-    fig.savefig('../images/marginal-plot-3d-prior.pdf')
-
-    ax, fig = get_marginal_plot_3d(maximum = maximum,
-            mean = (0.15, 0.247),
-            variance = (0.039, 0.026),
[…]

### 2013-10-10 slides/bin/integer_partition_plot.r
Adding R script for generating integer partition plot.

@@ -1,0 +1,118 @@
+#!/usr/bin/env Rscript
+
+##############################################################################
+## Functions
+
+integer_partition = function(n) {
+    part_nums = c()
+    for (k in 1:n) {
+        part_nums = c(part_nums, A(n=n, k=k))
+    }
+    return(part_nums)
+}
+
+A = function(n, k) {
+    if ((k == 1) || (n == k)) {
+        return(1)
+    }
+    if ((k > n) || (k < 1)) {
+        return(0)
+    }
+    return(A(n=(n-1), k=(k-1)) + A(n=(n-k), k=k))
+}
+
+expand_integer_vector = function(v) {
+    expanded = c()
+    for (i in 1:length(v)) {
+        expanded = c(expanded, rep(i, v[i]))
+    }
+    return(expanded)
+}
+
+get_psi_prior = function(n) {
+    integer_parts = integer_partition(n=n)
+    return(1/(n * integer_parts))
+}
+
+parse_file_ext = function(path) {
+    return(tail(strsplit(path, '\\.')[[1]], n=1))
+}
+
+plot_div_model_prior = function(n, output_path="partition_numbers.pdf") {
+    prior_probs = get_psi_prior(n=n)
+    names(prior_probs) = c(1:n)
+    partition_numbers = expand_integer_vector(integer_partition(n=n))
+    plot_letters = LETTERS[1:2]
+    format = 'pdf'
+    ext = parse_file_ext(output_path)
+    if ((tolower(ext) == 'eps') || (tolower(ext) == 'ps')) {
+        format = 'eps'
+    }
+
+    if (format == 'eps') {
+        postscript(output_path, onefile = F, paper="special", horizontal=F, width=7.087, height=3)
+    } else {
+        pdf(output_path, width=7.087, height=3)
+    }
+    par(yaxs = "i", xaxs = "i")
+    par(cex.axis=0.7)
+    par(mgp = c(1.4,0.4,0)) # mgp adjusts space of axis labels
+    # oma is space around all plots
+    par(oma = c(0,0,0.3,0.2), mar = c(2.65,2.45,0.5,0.1))
+    par(mfrow = c(1, 2))
+    hist(partition_numbers,
+            breaks= c(1:(n+1)),
+    		right=FALSE,
+    		xlab="",
+    		ylab=bquote("# of divergence models"),
+            main="",
+    		# xlim=c(0,22),
+    		# ylim=c(0,140),
[…]

### 2013-10-10 slides/bin/plot-validation-results.py
Adding plotting scripts to `slides/bin`.
Adding scripts for generating plots that are better suited for
presentation slides.

Also adding one of the resulting plots `omega_over_sampling.pdf`.

More edits to `slides.tex` as well.

@@ -1,0 +1,31 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes import plotting
+import project_util
+
+def main_cli():
+    for d in ['dpp', 'old']:
+        info_path = os.path.join(project_util.VALIDATION_RESULT_DIR, d,
+                    'pymsbayes-results', 'pymsbayes-info.txt')
+        plotting.plot_validation_results(info_path,
+                plot_dir = project_util.IMAGE_DIR,
+                omega_symbol = r'D_T',
+                psi_symbol = r'|\mathbf{\tau}|',
+                mean_time_symbol = r'\bar{T}',
+                plot_accuracy = False,
+                prob_plot_glm = False,
+                prob_plot_height = 3.5,
+                prob_plot_margin_left = 0,
+                prob_plot_margin_bottom = 0,
+                prob_plot_margin_right = 1,
+                prob_plot_margin_top = 0.975,
+                prob_plot_padding_between_horizontal = 0.5,
+                prob_plot_padding_between_vertical = 1.0,
+                math_font = None)
+
+if __name__ == '__main__':
+    main_cli()
+

### 2013-10-10 slides/bin/plot_sampling_error.py

@@ -1,0 +1,85 @@
+#! /usr/bin/env python
+
+import os
+import sys
+import re
+import matplotlib
+
+from pymsbayes import plotting
+from pymsbayes.utils.parsing import parse_posterior_summary_file
+from pymsbayes.fileio import process_file_arg
+import project_util
+
+def main_cli():
+    sum_file_pattern = re.compile(
+            r'^d1-m1-s1-(?P<iter_index>\d{1,4})-posterior-summary.txt.gz$')
+    result_dir = os.path.join(project_util.REDUX_RESULT_DIR, 'sampling-error',
+            'pymsbayes-results')
+    post_dir = os.path.join(result_dir, 'pymsbayes-output',
+            'd1', 'm1')
+    post_sum_paths  = {}
+    for f in os.listdir(post_dir):
+        m = sum_file_pattern.match(f)
+        if m:
+            post_sum_paths[int(m.group('iter_index'))] = os.path.join(post_dir,
+                    f)
+    iter_indices = sorted(post_sum_paths.iterkeys())
+    nsamples = []
+    omega_HPD_low = []
+    omega_HPD_high = []
+    omega_HPD_low_glm = []
+    omega_HPD_high_glm = []
+    for i in iter_indices:
+        nsamples.append((i + 1) * 100000)
+        f, close = process_file_arg(post_sum_paths[i], 'r')
+        post_sum = parse_posterior_summary_file(f)
+        hpd_int  = post_sum['PRI.omega']['HPD_95_interval']
+        omega_HPD_low.append(float(hpd_int[0]))
+        omega_HPD_high.append(float(hpd_int[1]))
+        hpd_int_glm  = post_sum['PRI.omega']['HPD_95_interval_glm']
+        omega_HPD_low_glm.append(float(hpd_int_glm[0]))
+        omega_HPD_high_glm.append(float(hpd_int_glm[1]))
+        f.close()
+
+    sp_list = []
+    sd_list = []
+    sd_list.append(plotting.ScatterData(x = nsamples, y = omega_HPD_low))
+    sd_list.append(plotting.ScatterData(x = nsamples, y = omega_HPD_high))
+    # matplotlib.rc('text',**{'usetex': True})
+    sp = plotting.ScatterPlot(
+            scatter_data_list = sd_list,
+            y_label = r'95% HPD $D_T$',
+            ylim = (-0.05, None),
+            center_text = r'Unadjusted')
+    sp.center_text_size = 12.0
[…]

### 2013-10-10 slides/bin/project_util.py

@@ -1,0 +1,20 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+BIN_DIR = os.path.abspath(os.path.dirname(__file__))
+SLIDE_DIR = os.path.abspath(os.path.dirname(BIN_DIR))
+IMAGE_DIR = os.path.abspath(os.path.join(SLIDE_DIR, 'images'))
+BASE_DIR = os.path.abspath(os.path.dirname(SLIDE_DIR))
+REDUX_DIR = os.path.abspath(os.path.join(BASE_DIR, 'response-redux'))
+REDUX_RESULT_DIR = os.path.abspath(os.path.join(REDUX_DIR, 'results'))
+VALIDATION_DIR = os.path.abspath(os.path.join(BASE_DIR, 'validation'))
+VALIDATION_RESULT_DIR = os.path.abspath(os.path.join(VALIDATION_DIR, 'results'))
+
+def main():
+    sys.stdout.write("%s" % PROJECT_DIR)
+
+if __name__ == '__main__':
+    main()
+

### 2013-10-13 slides/bin/plot-power-results.py
Adding plotting script for power results.
Adding `plot-power-results.py` to `slides/bin` for generating
slide-friendly plots of power results.

@@ -1,0 +1,365 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes.utils.parsing import DMCSimulationResults, spreadsheet_iter
+from pymsbayes.config import MsBayesConfig
+from pymsbayes.plotting import (PowerPlotGrid, ProbabilityPowerPlotGrid,
+        AccuracyPowerPlotGrid)
+from pymsbayes.utils.messaging import get_logger
+import project_util
+
+_LOG = get_logger(__name__)
+
+class PowerResult(object):
+    def __init__(self):
+        self.true = []
+        self.mode = []
+        self.mode_glm = []
+        self.median = []
+        self.prob = []
+        self.prob_glm = []
+
+    @classmethod
+    def parse_result_summary(cls, result_summary_path):
+        psi = cls()
+        omega = cls()
+        for d in spreadsheet_iter([result_summary_path]):
+            psi.true.append(int(d['psi_true']))
+            psi.mode.append(int(d['psi_mode']))
+            psi.mode_glm.append(int(round(float(d['psi_mode_glm']))))
+            psi.prob.append(float(d['psi_1_prob']))
+            psi.prob_glm.append(float(d['psi_1_prob_glm']))
+            omega.true.append(float(d['omega_true']))
+            omega.mode.append(float(d['omega_mode']))
+            omega.median.append(float(d['omega_median']))
+            omega.mode_glm.append(float(d['omega_mode_glm']))
+            omega.prob.append(float(d['omega_prob_less']))
+            omega.prob_glm.append(float(d['omega_prob_less_glm']))
+        return psi, omega
+
+def parse_results(dmc_sim,
+        observed_prefixes_to_include = None,
+        observed_suffixes_to_include = None,
+        prior_suffixes_to_include = None):
+    psi_results = {}
+    omega_results = {}
+    observed_index_to_name = {}
+    observed_configs = {}
+    for k, v in dmc_sim.observed_index_to_config.iteritems():
+        if observed_prefixes_to_include:
+            include = False
+            for s in observed_prefixes_to_include:
+                if os.path.basename(v).startswith(s):
+                    include = True
+                    break
+            if not include:
+                continue
+        if observed_suffixes_to_include:
+            include = False
+            for s in observed_suffixes_to_include:
+                if os.path.splitext(v)[0].endswith(s):
+                    include = True
+                    break
+            if not include:
+                continue
+        if os.path.basename(v).startswith('exp-'):
[…]

### 2013-10-13 slides/bin/plot-validation-results.py
Updating `plot-validation-results.py` in `slides/bin`.

@@ -51,6 +51,29 @@ def main_cli():
             sub_plots['dpp']['old'],
             sub_plots['dpp']['dpp'],
             ]
+    
+    pg = plotting.PlotGrid(subplots = sub_plot_list[:1],
+            num_columns = 1,
+            share_x = True,
+            share_y = True,
+            title = 'Posterior probability of one divergence',
+            title_size = 12.0,
+            title_top = False,
+            y_title = 'True probability of one divergence',
+            y_title_position = 0.001,
+            y_title_size = 12.0,
+            width = 5.2,
+            height = 4.0,
+            auto_height = False)
+    pg.label_schema = None
+    pg.auto_adjust_margins = False
+    pg.margin_bottom = 0.05
+    pg.margin_left = 0.04
+    pg.margin_top = 1
+    pg.margin_right = 1
+    pg.reset_figure()
+
+    pg.savefig('../images/validation-model-choice-old.pdf')
 
     pg = plotting.PlotGrid(subplots = sub_plot_list,
             num_columns = 2,

### 2013-10-14 slides/bin/plot-power-results.py
Updating plotting scripts for slides.
Also, updating the images used in `slides.tex`.

@@ -166,8 +166,23 @@ def create_plots(info_path,
             margin_top = 0.83
             margin_bottom = 0.085
             margin_left = 0.017
+            margin_right = 0.95
             column_label_size = 22.0
             title_size = 18.0
+            row_labels = []
+            if prior_name == 'old':
+                row_labels.append(r'$M_{msBayes}$')
+            elif prior_name == 'u-shaped':
+                row_labels.append(r'$M_{Ushaped}$')
+            elif prior_name == 'uniform':
+                row_labels.append(r'$M_{Uniform}$')
+            elif prior_name == 'dpp':
+                row_labels.append(r'$M_{DPP}$')
+            else:
+                row_labels.append('')
+            row_label_size = 28.0
+            row_label_offset = 0.08
+
 
             psi_plot = PowerPlotGrid(
                     observed_config_to_estimates = cfg_to_psi,
@@ -187,9 +202,13 @@ def create_plots(info_path,
             pg.margin_bottom = margin_bottom
             pg.margin_top = margin_top
             pg.margin_left = margin_left
+            pg.margin_right = margin_right
             pg.column_label_offset = column_label_offset
             pg.column_label_size = column_label_size
             pg.title_size = title_size
+            pg.row_labels = row_labels
+            pg.row_label_size = row_label_size
+            pg.row_label_offset = row_label_offset
             pg.reset_figure()
             pg.savefig(os.path.join(output_dir,
                     prefix + '_power_psi_mode.pdf'))
@@ -231,9 +250,13 @@ def create_plots(info_path,
             pg.margin_bottom = margin_bottom
             pg.margin_top = margin_top
             pg.margin_left = margin_left
+            pg.margin_right = margin_right
             pg.column_label_offset = column_label_offset
             pg.column_label_size = column_label_size
             pg.title_size = title_size
+            pg.row_labels = row_labels
+            pg.row_label_size = row_label_size
+            pg.row_label_offset = row_label_offset
             pg.reset_figure()
             pg.savefig(os.path.join(output_dir,
                     prefix + '_power_psi_prob.pdf'))
@@ -272,9 +295,13 @@ def create_plots(info_path,
             pg.margin_bottom = margin_bottom
             pg.margin_top = margin_top
             pg.margin_left = margin_left
+            pg.margin_right = margin_right
             pg.column_label_offset = column_label_offset
[…]

### 2013-10-14 slides/bin/plot-validation-results.py

@@ -47,8 +47,8 @@ def main_cli():
 
     sub_plot_list = [
             sub_plots['old']['old'],
-            sub_plots['old']['dpp'],
             sub_plots['dpp']['old'],
+            sub_plots['old']['dpp'],
             sub_plots['dpp']['dpp'],
             ]
     
@@ -93,8 +93,8 @@ def main_cli():
             row_label_size = 20.0,
             column_label_offset = 0.05,
             row_label_offset = 0.05,
-            super_title = r'\textbf{Analysis model}',
-            super_y_title = r'\textbf{Data model}',
+            super_title = r'\textbf{Data model}',
+            super_y_title = r'\textbf{Analysis model}',
             super_title_size = 20.0,
             super_y_title_size = 20.0,
             super_y_title_right = True)
@@ -108,42 +108,49 @@ def main_cli():
 
     pg.savefig('../images/validation-model-choice-old-dpp.pdf')
 
-    sub_plot_list.extend([
-        sub_plots['uniform']['old'],
-        sub_plots['uniform']['dpp'],
-        sub_plots['u-shaped']['old'],
-        sub_plots['u-shaped']['dpp'],
-        ])
+    sub_plot_list = [
+            sub_plots['old']['old'],
+            sub_plots['dpp']['old'],
+            sub_plots['uniform']['old'],
+            sub_plots['u-shaped']['old'],
+            sub_plots['old']['dpp'],
+            sub_plots['dpp']['dpp'],
+            sub_plots['uniform']['dpp'],
+            sub_plots['u-shaped']['dpp'],
+            ]
+    for p in sub_plot_list:
+        p.set_extra_y_label('')
 
     pg = plotting.PlotGrid(subplots = sub_plot_list,
-            num_columns = 2,
+            num_columns = 4,
             share_x = True,
             share_y = True,
             title = 'Posterior probability of one divergence',
             title_size = 12.0,
             title_top = False,
             y_title = 'True probability of one divergence',
-            y_title_position = 0.03,
+            y_title_position = 0.015,
             y_title_size = 12.0,
-            height = 10.2,
+            height = 4.8,
+            width = 11.0,
             auto_height = False,
-            column_labels = [r'$M_{msBayes}$', r'$M_{DPP}$'],
[…]

### 2013-10-15 empirical-analyses/bin/plot-results.py
More work on slides.

@@ -1,0 +1,61 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes.utils.parsing import DMCSimulationResults, spreadsheet_iter
+from pymsbayes.config import MsBayesConfig
+from pymsbayes.plotting import (Ticks, HistData, ScatterPlot, PlotGrid)
+from pymsbayes.utils.messaging import get_logger
+import project_util
+
+_LOG = get_logger(__name__)
+
+def create_plots(info_path, out_dir):
+    if not os.path.exists(out_dir):
+        os.mkdir(out_dir)
+    dmc_sim = DMCSimulationResults(info_path)
+    result_prefix = dmc_sim.get_result_path_prefix(1, 1, 1) + '99-'
+    psi_path = result_prefix + 'psi-results.txt'
+    psis = []
+    for d in spreadsheet_iter([psi_path]):
+        n = int(round(10000 * float(d['estimated_prob'])))
+        psis.extend([int(d['num_of_div_events'])] * n)
+    bins = range(1, dmc_sim.num_taxon_pairs + 2)
+    hd = HistData(x = psis,
+            normed = True,
+            bins = bins,
+            histtype = 'bar',
+            align = 'mid',
+            orientation = 'vertical',
+            zorder = 0)
+    tick_labels = []
+    for x in bins[0:-1]:
+        if x % 2:
+            tick_labels.append(str(x))
+        else:
+            tick_labels.append('')
+    xticks_obj = Ticks(ticks = bins,
+            labels = tick_labels,
+            horizontalalignment = 'left')
+    hist = ScatterPlot(hist_data_list = [hd],
+            x_label = 'Number of divergence events',
+            y_label = 'Posterior probability')
+    pg = PlotGrid(subplots = [hist],
+            num_columns = 1,
+            height = 4.0,
+            width = 6.5,
+            label_schema = None,
+            auto_height = False)
+    pg.auto_adjust_margins = False
+    pg.margin_top = 1
+    pg.reset_figure()
+    pg.savefig(os.path.join(out_dir, 'philippines-dpp-psi-posterior.pdf'))
+
+
+def main_cli():
+    create_plots(project_util.PHILIPPINES_DPP_INFO, project_util.PLOT_DIR)
+
+if __name__ == '__main__':
+    main_cli()
+

### 2013-10-15 empirical-analyses/bin/project_util.py

@@ -1,0 +1,22 @@
+#! /usr/bin/env python
+
+import os
+import sys
+import glob
+
+BIN_DIR = os.path.abspath(os.path.dirname(__file__))
+PROJECT_DIR = os.path.abspath(os.path.dirname(BIN_DIR))
+PLOT_DIR = os.path.join(PROJECT_DIR, 'plots')
+CONFIG_DIR = os.path.abspath(os.path.join(PROJECT_DIR, 'configs'))
+RESULTS_DIR = os.path.abspath(os.path.join(PROJECT_DIR, 'results'))
+PHILIPPINES_DIR = os.path.abspath(os.path.join(RESULTS_DIR, 'philippines'))
+PHILIPPINES_DPP_DIR = os.path.abspath(os.path.join(PHILIPPINES_DIR, 'dpp'))
+PHILIPPINES_DPP_INFO = os.path.join(PHILIPPINES_DPP_DIR, 'pymsbayes-results', 'pymsbayes-info.txt')
+
+
+def main():
+    sys.stdout.write("%s" % PROJECT_DIR)
+
+if __name__ == '__main__':
+    main()
+

### 2013-10-16 empirical-analyses/bin/plot-results.py
Updating plotting of `empirical-analyses`.

@@ -3,6 +3,8 @@
 import os
 import sys
 
+import matplotlib
+
 from pymsbayes.utils.parsing import DMCSimulationResults, spreadsheet_iter
 from pymsbayes.config import MsBayesConfig
 from pymsbayes.plotting import (Ticks, HistData, ScatterPlot, PlotGrid)
@@ -12,6 +14,8 @@ import project_util
 _LOG = get_logger(__name__)
 
 def create_plots(info_path, out_dir):
+    matplotlib.rc('text',**{'usetex': True})
+    old = ([1] * 992) + ([2] * 8)
     if not os.path.exists(out_dir):
         os.mkdir(out_dir)
     dmc_sim = DMCSimulationResults(info_path)
@@ -29,6 +33,13 @@ def create_plots(info_path, out_dir):
             align = 'mid',
             orientation = 'vertical',
             zorder = 0)
+    hd_old= HistData(x = old,
+            normed = True,
+            bins = bins,
+            histtype = 'bar',
+            align = 'mid',
+            orientation = 'vertical',
+            zorder = 0)
     tick_labels = []
     for x in bins[0:-1]:
         if x % 2:
@@ -40,7 +51,15 @@ def create_plots(info_path, out_dir):
             horizontalalignment = 'left')
     hist = ScatterPlot(hist_data_list = [hd],
             x_label = 'Number of divergence events',
-            y_label = 'Posterior probability')
+            y_label = 'Posterior probability',
+            xticks_obj = xticks_obj)
+    hist_old = ScatterPlot(hist_data_list = [hd_old],
+            x_label = 'Number of divergence events',
+            y_label = 'Posterior probability',
+            xticks_obj = xticks_obj)
+    hist.set_xlim(left = bins[0], right = bins[-1])
+    hist_old.set_xlim(left = bins[0], right = bins[-1])
+    hist.set_ylim(bottom = 0.0, top = 0.1)
     pg = PlotGrid(subplots = [hist],
             num_columns = 1,
             height = 4.0,
@@ -51,6 +70,25 @@ def create_plots(info_path, out_dir):
     pg.margin_top = 1
     pg.reset_figure()
     pg.savefig(os.path.join(out_dir, 'philippines-dpp-psi-posterior.pdf'))
+
+    hist.set_ylim(bottom = 0.0, top = 1.0)
+    hist.set_ylabel('')
+    hist_old.set_ylim(bottom = 0.0, top = 1.0)
[…]

### 2013-10-16 empirical-analyses/bin/plot-results.py
Updating plotting for `empirical-analyses`

@@ -14,7 +14,7 @@ import project_util
 _LOG = get_logger(__name__)
 
 def create_plots(info_path, out_dir):
-    matplotlib.rc('text',**{'usetex': True})
+    # matplotlib.rc('text',**{'usetex': True})
     old = ([1] * 992) + ([2] * 8)
     if not os.path.exists(out_dir):
         os.mkdir(out_dir)
@@ -82,7 +82,8 @@ def create_plots(info_path, out_dir):
             share_y = True,
             label_schema = None,
             auto_height = False,
-            column_labels = [r'\texttt{msBayes}', r'\texttt{dpp-msbayes}'],
+            # column_labels = [r'\texttt{msBayes}', r'\texttt{dpp-msbayes}'],
+            column_labels = [r'msBayes', r'dpp-msbayes'],
             column_label_size = 18.0)
     pg.auto_adjust_margins = False
     pg.margin_top = 0.92

### 2013-10-16 empirical-analyses/bin/plot-results.py
More work on slides.

@@ -8,10 +8,22 @@ import matplotlib
 from pymsbayes.utils.parsing import DMCSimulationResults, spreadsheet_iter
 from pymsbayes.config import MsBayesConfig
 from pymsbayes.plotting import (Ticks, HistData, ScatterPlot, PlotGrid)
+from pymsbayes.utils.stats import Partition
+from pymsbayes.utils.probability import GammaDistribution
 from pymsbayes.utils.messaging import get_logger
 import project_util
 
 _LOG = get_logger(__name__)
+
+def get_dpp_psi_values(num_elements, shape, scale, num_sims = 100000):
+    conc = GammaDistribution(shape, scale)
+    p = Partition([1] * num_elements)
+    psis = []
+    for i in range(num_sims):
+        a = conc.draw()
+        x = p.dirichlet_process_draw(a)
+        psis.append(len(set(x)))
+    return psis
 
 def create_plots(info_path, out_dir):
     # matplotlib.rc('text',**{'usetex': True})
@@ -90,6 +102,38 @@ def create_plots(info_path, out_dir):
     pg.padding_between_horizontal = 1.0
     pg.reset_figure()
     pg.savefig(os.path.join(out_dir, 'philippines-dpp-psi-posterior-old-vs-dpp.pdf'))
+
+    prior_psis = get_dpp_psi_values(dmc_sim.num_taxon_pairs, 1.5, 18.099702, num_sims = 100000)
+    prior_hd = HistData(x = prior_psis,
+            normed = True,
+            bins = bins,
+            histtype = 'bar',
+            align = 'mid',
+            orientation = 'vertical',
+            zorder = 0)
+    prior_hist = ScatterPlot(hist_data_list = [prior_hd],
+            x_label = 'Number of divergence events',
+            y_label = 'Probability',
+            xticks_obj = xticks_obj)
+    prior_hist.set_xlim(left = bins[0], right = bins[-1])
+    prior_hist.set_ylim(bottom = 0.0, top = 0.12)
+    hist.set_ylim(bottom = 0.0, top = 0.12)
+    pg = PlotGrid(subplots = [prior_hist, hist],
+            num_columns = 2,
+            height = 3.5,
+            width = 8.0,
+            share_x = True,
+            share_y = True,
+            label_schema = None,
+            auto_height = False,
+            # column_labels = [r'\texttt{msBayes}', r'\texttt{dpp-msbayes}'],
[…]

### 2013-10-18 manuscripts/reply/reply-figures.tex
Moving sections of `reply.tex` into separate files.
Moving the main text, tables, and figures of `reply.tex` to
`reply-main.tex`, `reply-tables.tex`, and `reply-figures.tex`.

--- 2013-10-01 manuscripts/reply/reply.tex
@@
-%&<latex>
-\documentclass[letterpaper,12pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{../utils/preamble.tex}
-\input{../utils/macros.tex}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\doublespacing
-\raggedright
-\setlength{\parindent}{0.5in}
-\begin{linenumbers}
-
-\begin{titlepage}
-    \begin{flushleft}
-        \sffamily
-
-        \MakeUppercase{\large\bfseries Why you should not fix a biased
-        model-choice method by adding an additional dimension of model choice:
-        A reply to Hickerson et al.}
-
-        \vspace{12pt}
-        \textbf{Running head:} \MakeUppercase{Approximate Bayesian model
-        choice}
-
-        \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2}$, Jeet Sukumaran$^{3}$, Jacob A.\
-        Esselstyn$^{4}$, Charles W.\ Linkem$^{5}$, Cameron D.\
-        Siler$^{6}$, Mark T.\ Holder$^{1}$ and Rafe M.\ Brown$^{1}$
-
-        \bigskip
-        $^1$\emph{Department of Ecology and Evolutionary Biology,
-            Biodiversity Institute,
-            University of Kansas,
-            Lawrence, Kansas 66045}\\[.1in]
-        $^3$\emph{Department of Biology,
-            Duke University,
-            Durham, North Carolina 27708} \\[.1in]
-        $^4$\emph{Museum of Natural Science,
-            Louisiana State University,
-            119 Foster Hall,
-            Baton Rouge, Louisiana 70803}\\[.1in]
-        $^5$\emph{Department of Biology,
-            University of Washington,
-            Seattle, Washington 98195}\\[.1in]
-        $^6$\emph{Sam Noble Museum,
-            Department of Biology,
-            University of Oklahoma,
-            Norman, Oklahoma 73072}\\[.1in]
-        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
-        joaks1@gmail.com})\\
-
-    \end{flushleft}
-\end{titlepage}
-
-{\sffamily
-    \noindent\textbf{ABSTRACT} \\
-    \noindent Recently, \citet{Oaks2012} presented a simulation-based
-    assessment of the approximate Bayesian phylogeographical model choice
-    method implemented in \msb, and found the method to lack robustness and
-    power to detect temporal variation in divergences that are randomly
-    distributed over relatively broad periods of evolutionary history.
-    They conclude that a likely mechanism behind the bias toward models with
-    few divergence events is the small marginal likelihoods of models with more
-    divergence time parameters due to the integration over vast parameter space
-    with low likeihood and high prior density that is imposed by broad uniform
-    priors.
-    \citet{Hickerson2013} have responded to this work, criticising the priors
-    used by \citet{Oaks2012} and presenting an approximate Bayesian
-    model-averaging approach that uses narrow, empirically informed uniform
-    priors.
-    \citet{Hickerson2013} use this method to re-analyze the empirical dataset
-    of \citet{Oaks2012} comprised of 22 pairs of vertebrate populations from
-    the Philippines, and conclude the model-averaging approach circumvents the
-    issues revealed by \citet{Oaks2012} and yields robust inference of
-    divergence models.
-    Furthermore, they argue that the biases revealed by \citet{Oaks2012} are
-    inherent to the inefficient rejection algorithm implemented in \msb, rather
-    than a result of low marginal likelihoods of parameter-rich models.
-    Here we discuss an error in the methodology of \citet{Hickerson2013} that
-    renders some of their results invalid and the remainder difficult to
-    interpret.
-    Furthermore, we demonstrate that the approach of \citet{Hickerson2013} is
-    biased toward models with less parameter space, which manifests in a strong
-    tendency to sample predominantly from models that exclude the true values
-    of the model's parameters.
-    Our results suggest that their model-averaged results are dominated by
-    models that likely exclude at least some of the true divergence times
-    across the 22 pairs of Philippine vertebrates, which adds to the
-    difficulty in interpreting their results.
-    We also show the approach of \citet{Hickerson2013} is still biased toward
-    inferring simultaneous divergences even when the true divergences are
-    randomly distributed over millions of generations.
-    Our results, coupled with those of \citet{Oaks2012}, strongly suggest that
-    low marginal likelihoods of complex models is a primary cause of the bias
-    of \msb toward models with less parameter space.
-    We discuss ways to mitigate this bias without resorting to adding an
-    additional dimension of model choice to the model or using questionable
-    empirically informed uniform priors.
-
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Approximate Bayesian computation; Bayesian
-        model choice; empirical Bayes}
-
-}
-
-\newpage
-
-%CWL:
-% \highLight{The goal of this paper needs to be to highlight the major
-%     differences of opinion on how ABC can be used in comparative phylogeography
-%     and how these changes to msbayes can be made that will improve the
-%     inference in the future. We should avoid nit-picking the Hickerson paper
-%     and try to stay positive.}
-
-\noindent Recently, this journal has served as a medium for a
-discussion of the potential pitfalls of approximate Bayesian methods of
-comparative phylogeographical model choice and prior selection.
-The discussion has centered around the approximate Bayesian model choice method,
-\msb, which estimates the temporal distribution of divergences among
-co-distributed pairs of taxa \citep{Huang2011} by sampling over all
-unordered models of divergence.
-\citet{Oaks2012} published results of simulation-based power analyses
-showing that \msb can often be biased toward inferring models of temporally
-clustered divergence times across taxon pairs.
-\citet{Hickerson2013} published a response to this paper where they present
-an empirical Bayes model-averaging approach that they conclude circumvents the
-poor behavior of the method revealed by \citet{Oaks2012}.
-These papers are largely in agreement, with \citet{Hickerson2013} reiterating
-much of the discussion of \citet{Oaks2012} regarding the impact of broad
-uniform priors on Bayesian model choice.
-The main difference in perspectives are centered around
-(1) the mechanism by which broad uniform priors cause the poor behavior of
-\msb, and
-(2) how to potentially mitigate this issue.
-
-\citet{Oaks2012} suggest the primary cause of the poor behavior of the method
-is likely the low marginal likelihoods of parameter-rich models integrated over
-vast parameter space with low probability of producing the data, yet relatively
-high prior density \citep{Lindley1957}.
-Note, this suggests the bias is extrinsic to \msb, and the numerical
-approximation machinery of \msb could be sound.
-\citet{Hickerson2013} suggest the bias is intrinsic to \msb, concluding the
-method's rejection algorithm is inefficient and will be increasingly biased as
-the overall space of the model increases, either as a function of the number of
-taxon pairs or the width of the uniform priors on nuisance parameters.
-\citet{Hickerson2013} support their position by giving a probabilistic argument
-that focuses on only one of the three prior models used by \citet{Oaks2012}.
-We show that this argument is based on an empirical Bayesian reasoning and
-questionable assumptions, and does not explain the bias of the method found in
-many of the analyses of \citet{Oaks2012}. 
-Furthermore, we present results of additional analyses that, when combined with
-the results of \citet{Oaks2012}, strongly suggest small marginal likelihoods of
-models with more parameter space is a primary mechanism by which uniform priors
-are causing the poor behavior of \msb.
-
-%This paragraph seems too aggressive. It also seems to repeat some of the previous points. I removed it for now.
-
-To mitigate the bias, \citet{Oaks2012} suggest alternative prior probability
-distributions on divergence models and nuisance parameters could increase the
-likelihoods of models with more divergence time parameters.
-They also describe simulation-based procedures for determining power, accuracy,
-and robustness of the method given a dataset, and recommend any application of
-\msb should be accompanied by such procedures.
-Alternatively, \citet{Hickerson2013} present an approximate Bayesian
-model-averaging approach for accommodating uncertainty in selecting among
-empirically guided priors, and champion the method as a means of avoiding the
-pitfalls raised by \citet{Oaks2012}.
-We discuss errors in this approach that render some of their results invalid
-and leave the remaining results difficult to interpret.
-Furthermore, we follow the recommendations of \citet{Oaks2012} and present
-simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
-approach.
-Our results demonstrate that the model-averaging approach of
-\citet{Hickerson2013} is biased toward clustered divergence models and
-predisposed to excluding true parameter space.
-The model-averaging approach provides another means by which the model can
-``escape'' large parameter space, and the method's bias toward smaller models
-remains.
-This manifestation of the bias often excludes the true parameter space due to
-the use of narrow uniform priors.
-Furthermore we discuss the potential theoretical and practical problems of
-empirical Bayesian model choice.
-% which is used and encouraged by
-% \citet{Hickerson2013} without discussing its potential complications.
-
-%CWL:
-%It may be important to point out the error made in their paper, but it can
-%also be nit-picky, it depends on how the error impacts the overall message of
-%our response.  We should strive to keep the response short and directed at the
-%big picture items. If there is a way to still point out the error, then maybe
-%some of this section should be used
-\section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
-\citet{Hickerson2013} re-analyzed the dataset of \citet{Oaks2012} using a
-model-averaging approach, where they placed a discrete uniform prior over eight
-different prior models (see Table 1 of \citet{Hickerson2013}).
-However, there is a fundamental error in the methodology of
-\citet{Hickerson2013}: their model mixes different time units.
-
-Each of the eight prior models used in the re-analysis by \citet{Hickerson2013}
-has one of two priors on the mean size of the descendant populations of each
-taxon pair:
-$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ or
-$\meanDescendantTheta{} \sim U(0.0005, 0.04)$.
-As described in \citet{Oaks2012}, the divergence time parameters of the model
-implemented in \msb are scaled relative to a constant reference population
-size, \myTheta{C}.
-This reference population size is defined in terms of the upper limit of the
-uniform prior on the mean size of the descendant populations,
-\meanDescendantTheta{}, such that for the prior $\meanDescendantTheta{} \sim
-U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
-$\myTheta{C} = \uniformMax{\meanDescendantTheta{}}/2$.
-Thus, the model used by \citet{Hickerson2013} mixes two different units of
-time.
-In other words, some of their prior and posterior samples are in units of
-$0.05/\mutationRate$ generations, whereas others are in units of
-$0.02/\mutationRate$ generations.
-
-The fact that their posterior samples are in different units makes the results
-of \citet{Hickerson2013} difficult to interpret, and renders their
-regression-adjusted results invalid.
-A fundamental assumption of regression is that all of the values of the
-response variable are in the same units.
-Thus, the results in sections ``Using ABC Model Comparison to Weight
-Alternative Priors for the Philippine Vertebrate Data'' and ``Improved Sampling
-Efficiency by Prior Weighting Supports Asynchronous and Recent Divergence for
-the Philippines Vertebrate Data'' and presented in Figure 2 of
-\citet{Hickerson2013} should be disregarded.
-The error is easily illustrated by re-plotting their results with the different
-time units indicated (Figure~\ref{figJointPosteriorHickerson}).
-
-% \cwlnote{}{This part seems important, but not related to the error in units
-% discussed above}
-Another important characteristic of the approach used by \citet{Hickerson2013}
-that makes their results difficult to interpret is their use of narrow,
-empirically guided uniform priors in an approximate-Bayesian model-choice
-framework.
-We discuss the risks associated with this approach in the next section, and
-show how the results of \citet{Hickerson2013} likely exclude true values of the
-model's parameters.
-
-% In addition to this error, we were not able to reproduce the results of
-% \citet{Hickerson2013}.
-% We followed their methods, generating $5\e6$ prior samples for each of the
-% eight models in their Table 1.
-% We retained the 10,000 samples from each model with the smallest Euclidean
-% distance to the observed summary statistics, and subsequently retained 10,000
-% samples from the remaining 80,000.
-
-\section*{The potential implications of empirical Bayesian model choice}
-\citet{Hickerson2013} repeatedly refer to the prior distributions used by
-\citet{Oaks2012} as ``poorly selected.''
-However, this is misleading, as \citet{Oaks2012} discuss in detail how they
-selected their priors (see their section ``Specifying and simulating the joint
-prior'').
-They mention how the use of uniform prior distributions in \msb for many of the
-model's parameters necessitates awkward broad priors to avoid excluding the
-truth, \textit{a priori}.
-What \citet{Hickerson2013} mean by ``poorly selected'' is that
-\citet{Oaks2012} did not use their data to inform their initial prior
-distributions (i.e., they did not use an empirical Bayesian approach).
-\citet{Oaks2012} did use empirically informed priors when assessing both the
-prior sensitivity of the method and determining the power of \msb under
-real-world conditions with highly informative priors.  
-\citet{Oaks2012} discuss
-the potential dangers of taking an empirical Bayesian approach to model choice
-(see the last paragraph of ``Assessing prior sensitivity of \msb'' in
-\citet{Oaks2012}), and we expand on this here by exploring both the theoretical
-and practical implications of using empirically informed priors for Bayesian
-model choice.
-
-\subsection*{Theoretical implications of empirical priors for Bayesian model
-choice}
-\begin{linenomath}
-Bayesian inference is a rational method of inductive learning in which Bayes'
-rule is used to update our beliefs about a model as new information becomes
-available.
-If we let \allParameterValues represent the set of all possible parameter
-values, we can define a prior distribution for all $\theta \in
-\allParameterValues$ such that $p(\theta)$ describes our belief that any given
-\myTheta{} is the true value of the parameter.
-If we let \allDatasets represent all possible datasets then we can 
-define a sampling model for all $\theta \in
-\allParameterValues$ and $\alignment{}{} \in \allDatasets$ such that
-$p(\alignment{}{} | \theta)$ measures our belief that any dataset \alignment{}{}
-will be generated by any model state \myTheta{}.
-After collecting a new dataset \alignment{i}{}, we can use Bayes' rule to
-calculate the posterior distribution
-\begin{equation}
-    p(\myTheta{} \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given 
-    \myTheta{})p(\myTheta{})}{p(\alignment{i}{})},
-    \label{eq:bayesrule}
-\end{equation}
-where
-\begin{equation}
-    p(\alignment{i}{}) = \int_{\myTheta{}} p(\alignment{i}{} \given
-    \myTheta{})p(\myTheta{}) d\myTheta{}.
-\end{equation}
-The posterior distribution is a measure of our beliefs after seeing the new
-information.
-\end{linenomath}
-
-This is an elegant method of updating our beliefs as data are accumulated.
-However, this all hinges on the fact that the prior ($p(\myTheta{})$) is
-defined for all possible parameter values independently of the new data being
-analyzed.
-Any other datasets or external information can safely be used to inform our
-beliefs about $p(\myTheta{})$.
-However, if the same data are used to both inform the prior and calculate the
-posterior, the prior becomes conditional on the data, and Bayes' rule breaks
-down.
-
-This is not to say that empirical Bayesian approaches are not useful.
-Empirical Bayes is a well studied branch of Bayesian statistics that
-has given rise to many powerful inference methods.
-However, empirical Bayes methods have an uncertain theoretical basis and
-do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
-empirical Bayesian estimates of the posterior are often too narrow, off-center,
-and misshapen;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
-Nevertheless, empirical Bayes methods can provide a powerful means of parameter
-estimation that often exhibit favorable frequentist properties.
-Furthermore, many post-hoc correction methods have been developed for
-estimating confidence-intervals from empirical Bayes estimates of posterior
-distributions that often exhibit well-behaved frequentist coverage
-probabilities
-\citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.
-
-Whereas empirical Bayes approaches can provide powerful methods for parameter
-estimation, a theoretical justification for empirical Bayes approaches to model
-choice is questionable.
-In Bayesian model choice, the goal is not to estimate parameters, but to
-estimate the relative probabilities of candidate models.
-Unlike parameter estimates, posterior probabilities represent a summary of the
-entire posterior distribution.
-Thus, given that empirical Bayesian posterior distributions are not accurate,
-there is no guarantee regarding the accuracy of probabilities summarized from
-them.
-In addition to these theoretical concerns there are practical dangers to using
-narrow, empirical, uniform priors for a method that has already been shown to
-be biased toward models with less parameter space.
-% %CWL:
-% %I like the example, but it may be too cute. It is also too long. I think a
-% %contrived example is good, but if there could be an example that is more
-% %relevant to model choice and comparison of inaccurate posterior distributions,
-% %I think that would be better. This is an example of empirical Bayes, but not
-% %the potential problem with the use in msbayes. Unless I am interpreting wrong.
-% %I would be careful about making it to cute as well. You may intend to make it
-% %easy to understand for a general audience, but it can also seem like talking
-% %down to Hickerson et al, who we are responding to. It just depends on how you
-% %read it.
-% This can be demonstrated with a simple, albeit contrived, example.
-% Let us say that principal investigator Mary and her new postdoc Will are
-% interested in the hypothesis that the mass of George Washington's periwig
-% renders the quarter dollar of the United States unfair.
-% That is to say their null hypothesis is that the probability of a US
-% quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
-% They can certainly evaluate this hypothesis, they can have undergraduate
-% worker, Joe, flip the coin for them while they tabulate the results.
-% But being Bayesians, before they call Joe into the lab, they agree on a
-% prior probability to place on the set of all possible probabilities that
-% the quarter will land heads when it is flipped.
-% Given that neither of them have a quarter, and their prior knowledge that Joe
-% moonlights as a magician, and is notorious for performing coin and card tricks
-% in the lab, they suspect there is a good chance that the quarter Joe uses will
-% be either two-headed or two-tailed.
-% So, knowing that the beta distribution is the conjugate prior for a binomial
-% likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
-% (Figure~\ref{figCoinFlip}).
-
-% Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
-% begin flipping it.
-% After five tosses, four of which was heads, Joe decides that academics are
-% crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
-% being computational biologists, are satisfied with their
-% empirical dataset of $y = 4$ heads out of $N = 5$ trials.
-% They know from the conjugacy of the beta prior, that the posterior distribution
-% has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
-% in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
-% of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
-% This allows them to plug these values into the beta cumulative distribution
-% function to determine that the posterior probability of their hypothesis is
-% $p(\theta < 0.5 | y=1, N=5) = 0.088$.
-% Given their prior belief and dataset, this indeed is the correct posterior
-% probability of the hypothesis, and Mary and Will should now update their
-% posterior belief accordingly.
-
-% However, reflecting upon the results of their experiment
-% (Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
-% Their prior looks very ``poorly selected,'' and if they had only known that the
-% coin was fair before the data were collected, they would have selected a much
-% better prior.
-% Clearly, from their results, they should have used a prior centered
-% around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
-% much ``better.''
-% Rather than resort to flipping the coin five more times, Mary and Will decide
-% to redo their analysis using the much better ``prior'' of $p(\theta) \sim
-% beta(4.5, 1.5)$.
-% This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
-% probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
-% Now convinced that the US quarter is unfair, albeit not due to the mass of
-% President Washington's head as they hypothesized, Will begins composing an
-% e-mail of complaint addressed to the U.S.\ Mint.
-
-% If Joe's flips were a representative sample, Mary and Will's empirical Bayes
-% estimate might very well be a better point estimate for the parameter
-% \myTheta{}.
-% However, their empirical Bayes estimate of the probability of their hypothesis
-% is incorrect and biased.
-% This simple example shows how parameter estimation is fundamentally different
-% from estimating the probability of a model.
-% While empirically informed priors can be used to obtain well-behaved parameter
-% estimators, using them for model choice is much less certain.
-
-\subsection*{Practical concerns about empirically informed uniform priors for
-    Bayesian model choice}
-% \cwlnote{}{Should this be part of the empirical Bayes section, or something
-%     else?}
-
-% \cwlnote{}{It seems like this section can be boiled down to ``empirical Bayes
-%     approaches for model choice when using uniform priors can exclude the
-%     $truth$ from the prior, which is a violation of Bayes theorem" or something
-%     like that. Picking apart what they calculated and we calculated seems like
-%     too much detail. I would try to simplify how this is being presented. This
-%     can lead into how the use of more appropriate prior distributions should be
-%     used.}
-The results of Hickerson et al.'s \citeyear{Hickerson2013} reanalysis of the
-Philippine dataset strongly favored models with the narrowest, empirically
-informed prior on divergence time, and thus their model-averaged posterior
-estimates are dominated by models $M_1$ and $M_2$ (see Table 1 of
-\citet{Hickerson2013}).
-This is concerning, because the narrowest \divt{} prior used by
-\citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
-divergence times for at least some of the Philippine taxa, a major problem when
-using uniform priors.
-\citet{Hickerson2013} set this prior to match the 95\% highest posterior
-density (HPD) interval for the mean divergence time estimated under one of the
-priors used by \citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
-Given this interval estimate is for the \emph{mean} divergence time across all
-22 taxa, it may be inappropriate to set this as the limit on the prior, because
-many of the taxon pairs are expected to have diverged at times older than the
-upper limit.
-Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
-of the mean divergence time under the other two priors explored by
-\citet{Oaks2012} (under these priors the 95\% HPD is approximately 0.3--0.6;
-see Table~6 of \citet{Oaks2012}).
-Thus, the results of \citet{Hickerson2013} indicate that integration over
-parameter space with low likelihood is biasing the method toward models with
-smaller parameter space, and as a consequence is biasing the method toward
-models that exclude the truth (i.e., toward estimating model-averaged
-posteriors dominated by models that exclude true values of parameters of the
-model).
-
-%CWL:This should be in the section on model-averaging, not empirical Bayes
-%limitations
-We explored the propensity of the model-averaging approach of
-\citet{Hickerson2013} to exclude the truth in two ways.
-First, we re-analyzed the Philippines dataset using the model-averaging approach
-of \citet{Hickerson2013}, but set one of prior models with a uniform prior on
-divergence times that is unrealistically narrow, and almost certainly excludes
-most, if not all, of the true divergence times of the 22 taxon pairs.
-If small likelihoods of large models cause the method to prefer models with
-less parameter space, we expect \msb will preferentially sample from this
-incorrect model yielding a marginal posterior that is incorrect (i.e., the
-model-averaged posterior will be dominated by an incorrect model that excludes
-the truth).
-Second, we generated simulated datasets for which the divergence times are
-drawn from an exponential distribution and applied the approach of
-\citet{Hickerson2013} to each of them to see how often the method excludes the
-truth.
-
-For our re-analysis of the Philippines dataset we used the model-averaging
-approach of \citet{Hickerson2013}, but with a reduced set of prior models to
-avoid mixing time units.
-We used five prior models, all of which had priors on population sizes of
-$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
-U(0.0001, 0.05)$.
-Following \citet{Hickerson2013}, each of these models had the following
-priors on divergence time parameters:
-$M_1$, $\divt{} \sim U(0, 0.1)$;
-$M_2$, $\divt{} \sim U(0, 1)$;
-$M_3$, $\divt{} \sim U(0, 5)$;
-$M_4$, $\divt{} \sim U(0, 10)$; and
-$M_5$, $\divt{} \sim U(0, 20)$.
-We simulated $1\e6$ random samples from each of the models for a total of
-$5\e6$ prior samples.
-For each model, we retained the 10,000 samples with the smallest Euclidean
-distance from the observed summary statistics, standardizing the statistics
-using the prior means and standard deviations of the given model.
-From the remaining 50,000 samples, we then retained the 10,000 samples with the
-smallest Euclidean distance from the observed summary statistics, this time
-standardizing the statistics using the prior means and standard deviations
-across all five models.
-We then repeated this analysis twice, replacing the $M_1$ with
-$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
-times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
-respectively.
-While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
-\citet{Hickerson2013} likely excludes the true divergence times of at least
-some of the 22 taxa, we are nearly certain that these narrower priors are
-incorrect and exclude most, if not all, of the divergence times of the
-Philippine taxa.
-
-Our results show that the model-averaging approach of \citet{Hickerson2013}
-does not reduce the method's bias toward models with less parameter space,
-but rather allows it to manifest in a rather dangerous way.
-The method strongly prefers the prior model with the narrowest distribution on
-divergence times across all three of our analyses, even when this model is
-almost certainly incorrect and excludes the true divergence times of the
-Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
-Unfortunately, ``checking'' the priors by plotting the summary statistics from
-1000 random samples of each prior model along the first two orthogonal axes of
-a principle component analysis, as recommended by \citet{Hickerson2013},
-provides no warning of a problem (Figure~\ref{figPCA}).
-Given that the results of \citet{Hickerson2013} strongly prefer the models with
-the narrowest prior on divergence times, it seems quite likely that their
-model-averaged results are dominated by models that exclude at least some of
-the 22 true divergence times, making their results difficult to interpret
-(in addition to the error of time units).
-
-To better quantify the propensity of Hickerson et al.'s
-(\citeyear{Hickerson2013}) approach to exclude the truth, we simulated 1000
-datasets in which the divergence times for the 22 population pairs are drawn
-randomly from an exponential distribution with a mean of 0.5 ($\divt{} \sim
-Exp(2)$).
-All other parameters were identically distributed as the $M_1$--$M_5$ models
-(Table~\ref{tabModelChoiceEmpirical}).
-We then repeated the analysis described above using $1\e6$ random samples from
-prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, retaining 1000 posterior
-samples for each of the 1000 simulated datasets.
-
-% \highLight{Add how we determined exclusion.}
-For each simulation replicate, we estimated the Bayes factor in favor
-of excluding the truth as the ratio of the posterior to prior odds of
-excluding the true value of at least one parameter.
-Whenever the Bayes factor preferred a model excluding the truth we counted
-the number of 22 true divergence times that were excluded by the preferred
-model.
-Our results show that the model-averaging approach of \citet{Hickerson2013}
-excludes the true values of parameters in 97\% of the replicates (90\% with
-GLM-regression adjustment), excluding up to 21 of the 22 true divergence times
-(Figure~\ref{figExclusionSimTau}).
-We also used the mode estimate of the preferred model for each replicate to
-estimate the number of true values excluded, which produced very similar
-results.
-Furthermore, the posterior probability of excluding at least one true parameter
-value is very high in nearly all of the replicates
-(Figure~\ref{figExclusionSimProb}).
-Using a Bayes factor of greater than 10 as a criterion for strong support, 66\%
-of the replicates (87\% with GLM-regression adjustment) strongly support the
-exclusion of true values (Figure~\ref{figExclusionSimProb}).
-
-The results of our empirical and simulation-based analyses clearly demonstrate
-the danger of using narrow, empirically guided uniform priors in a Bayesian
-model-averaging framework.
-This is especially concerning in \msb, which has already been demonstrated to
-be biased away from models with more parameter space \citep{Oaks2012}.
-The consequence of this approach is a high risk of obtaining a model-averaged
-posterior estimate that is heavily weighted toward incorrect models that
-exclude true values of model parameters.
-
-\subsection*{Additional thoughts on empirical priors in Bayesian model choice}
-Given all of the theoretical and practical issues with empirical Bayes
-approaches to Bayesian model choice discussed in the proceeding sections, it is
-quite clear why one should use caution before overly criticizing an
-investigator's choice of priors after having seen the resulting posterior.
-%In a strictly Bayesian world, this is cheating.
-As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a large
-amount of uncertainty regarding the divergence times of the 22 population pairs
-under study.
-Two of these pairs represent distinct species, and the taxonomy of many groups
-in the Philippines has repeatedly been shown to mask deeply divergent lineages
-\citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
-    Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013}.
-\citet{Oaks2012} also discuss how the sole use of uniform priors in \msb makes
-it very difficult for investigators to express their prior uncertainty without
-putting a lot of prior density in regions of improbable parameter space.
-The alternative, as shown above, is to risk excluding the truth before
-performing the analysis.
-
-While our results above clearly demonstrate the risks inherent to the empirical
-Bayesian model-choice approach used by \citet{Hickerson2013}, one could justify
-such risk if the method does indeed increase the power of the method and thus
-decrease bias toward clustered divergence models, as claimed by
-\citet{Hickerson2013}.
-In the next section, we use simulation-based analyses to evaluate the power of
-Hickerson et al.'s (\citeyear{Hickerson2013}) approach, which show the method is
-still biased toward highly clustered models of divergence.
-
-\section*{Assessing the power of the model-averaging approach of
-    \citet{Hickerson2013}}
-% \cwlnote{}{I think the empirical use of the model-averaging should go here,
-%     instead of above in the empirical Bayes section}
-
-As recommended by \citet{Oaks2012}, we perform simulation-based analyses
-to explore the power of the approximate Bayesian model-averaging approach
-proposed by \citet{Hickerson2013}.
-Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
-the 22 population pairs randomly drawn from a uniform distribution, $U(0,
-\divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and 2.0, in
-\globalcoalunit generations.
-All other parameters were identically distributed as the prior.
-We use the same $5\e6$ samples from the same set of five prior models as above
-($M_1$, $M_2$, $M_3$, $M_4$, and $M_5$).
-As described above, for each simulated dataset, we approximated the posterior
-by retaining 1000 samples from the prior with the smallest Euclidean distance
-from the true summary statistics.
-In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
-posterior samples for each of them.
-
-We find that the approach of \citet{Hickerson2013} struggles to estimate the
-dispersion index of divergence times (\vmratio{}) across most of the \divt{max}
-we simulated, whether evaluating the unadjusted
-(Figure~\ref{figPowerAccOmegaMedian}) or GLM-adjusted
-(Figure~\ref{figPowerAccOmegaModeGLM}) posterior estimates.
-The method only estimates \vmratio{} relatively well when the simulated
-distribution of divergence times is identical to one of the prior models
-(Figures~\ref{figPowerAccOmegaMedian}E \& \ref{figPowerAccOmegaModeGLM}E).
-This is consistent with the conclusion of \citet{Oaks2012} that \msb lacks
-robustness and is highly sensitive to the prior distribution deviating from the
-true, underlying distribution of the data.
-
-Furthermore, our results demonstrate that the approach of \citet{Hickerson2013}
-consistently infers highly clustered divergences across all the \divt{max} we
-simulated (Figure~\ref{figPowerNumExcluded}).
-The method is most likely to infer the extreme case of a single divergence event
-when populations diverged randomly over the past \globalcoalunit generations.
-Even when divergences are random over the past $8\globalpopsize$ generations,
-the most likely inference is only two divergence events, and a single 
-divergence is still estimated in more than 10\% of the replicates.
-It is very interesting to note that as \divt{max} increases, but before the
-estimates are finally pulled away from $\numt{} = 1$, the distribution of
-\numt{} estimates closely mirror the odd U-shaped prior on divergence models
-used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
-(\citeyear{Oaks2012}) Figure 5B).
-This supports the conclusions of \citet{Oaks2012} that this U-shaped
-prior coupled with the poor marginal likelihoods of models with many
-\divt{} parameters, is a major cause of the method's bias toward
-clustered divergence models.
-
-Looking at our simulation results in terms of the posterior probability of the
-dispersion index of divergence times supporting the extreme case of one
-divergence event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the
-method strongly supports one divergence in greater than 27\% of the replicates
-across all the \divt{max} we simulated (Figure~\ref{figPowerOmegaProb}).
-Following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
-the criterion for incorrect inference of a single divergence event.
-There is strong support for a single divergence event in more than 90\% of the
-replicates when divergences are random over the past $2.4\globalpopsize$
-generations, and more than 60\% when over the past $3.2\globalpopsize$
-generations or less (Figure~\ref{figPowerOmegaProb}).
-
-Contrary to Hickerson et al.'s (\citeyear{Hickerson2013}) claim that their
-model-averaging approach ``can discriminate complex multispecies histories and
-correctly reject synchronous divergence, even when discrete divergence times
-differ by much less than \ldots millions of generations,'' our results show
-that method is biased toward supporting the extreme case of a single
-divergence event when populations diverged randomly over the last
-$8\globalpopsize$ generations.
-To put this on the scale roughly consistent with a vertebrate mitochondrial
-locus, assuming a mutation rate of $2\e{-8}$ per site per generation, this
-translates to 5 million generations.
-Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
-million generations.
-
-The results of our power analyses further demonstrate the propensity of
-Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
-parameter values.
-Across all but one of the \divt{max} we simulated, the method excludes the
-truth in a large proportion of replicates, and across many of the \divt{max} it
-will exclude a large proportion of the true divergence time values
-(Figure~\ref{figPowerNumExcluded}).
-The posterior probability of excluding at least on true divergence value is
-also quite high across many of the \divt{max}
-(Figure~\ref{figPowerProbExclusion}).
-Only when the data are identically distributed as one of the prior models does
-the method avoid excluding the truth more than 5\% of the time
-(Figure~\ref{figPowerNumExcluded}E).
-Again, this demonstrates the method's lack of robustness.
-
-Given our results, we conclude that the approach of \citet{Hickerson2013}
-excludes regions of parameter space containing the truth, and lacks power to
-detect random variation in divergence times over a scale of $8\globalpopsize$
-generations or more.
-This roughly translates to millions of generations for mitochondrial loci, and
-hundreds of millions of generations for nuclear loci.
-
-%CWL: I would suggest removing this section
-\section*{The power analysis of \citet{Hickerson2013}}
-\citet{Hickerson2013} present a power analysis in which they find that
-\msb has the power to infer multiple divergence events when divergence times
-are random over hundreds of thousands of generations (rather than millions as
-demonstrated by \citet{Oaks2012}).
-It is not surprising that under limited parameter space the method has
-increased power to detect temporal variation in divergences over narrower time
-windows than millions of generations.
-However, what is important to consider is whether those conditions are relevant
-to real-world applications of the method.
-\citet{Oaks2012} explore the behavior of the method under three divergence time
-priors, as narrow as 0--5 coalescent units. This is quite narrow considering
-that this expresses the prior belief that all 22 taxon pairs diverged within
-this window.
-\citet{Hickerson2013} limit their power analysis to a single prior of 0--1
-coalescent unit.
-This prior setting assumes that there is enough a priori information to
-be 100\% certain that all taxa (18 for their simulations) diverged within
-the last 4$\globalpopsize$ generations.
-This does not seem applicable to most empirical systems.
-Furthermore, even when such extensive prior information is available, to only
-be able to detect multiple divergences on the scale of hundreds of thousands of
-generations does not seem very powerful.
-
-Also, \citet{Hickerson2013} present information only on the extreme case of a
-single divergence event.
-As discussed in \citet{Oaks2012}, it seems odd to consider the estimation
-of two divergence events as a success when divergence is random over hundreds
-of thousands (or millions) of generations.
-In an empirical system such as the Philippines, where island fragmentation has
-occurred at least 10 times over the past several millions of years
-\citep{Haq1987,Rohling1998,Siddall2003,Miller2005}, an estimate where 22 taxa
-share even a handful of divergence events would be of biogeographic interest.
-
-Lastly, \citet{Hickerson2013} translate their results from units of
-\globalcoalunit generations to generations assuming a mutation rate of
-$1.92\e{-8}$ mutations per site per generation.
-If we translate their results to a scale more consistent with rates of mutation
-of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
-by \citet{Hickerson2013} the method can only reliably reject the extreme case
-of one divergence event when divergences are random over a period of more than
-3 million generations.
-Again, considering \citet{Hickerson2013} are assuming prior knowledge that
-all population pairs diverged within the last coalescent unit, this
-does not seem particularly powerful.
-
-\section*{Model likelihoods or insufficient prior sampling as the cause of bias
-in \msb}
-One of the points of disagreement between \citet{Hickerson2013} and
-\citet{Oaks2012} is the underlying mechanism causing the bias toward models
-with clustered divergences (i.e., models with few divergence events).
-\citet{Oaks2012} present simulation-based analyses that suggest the broad
-uniform prior on divergence time parameters decrease the marginal likelihoods
-of models with more divergence time parameters, whereas \citet{Hickerson2013}
-argue that the bias is inherent to the inefficient rejection algorithm
-implemented in \msb.
-
-\citet{Hickerson2013} present an interesting probabilistic argument
- to show that insufficient prior sampling is to blame for the bias.
-They argue the widest of the three priors used by \citet{Oaks2012} was very
-unlikely to produce any prior samples with large numbers of divergence times
-that were consistent with the Philippine data.
-There are a few issues with their argument.
-First, the probabilities they present assume the gene divergence times
-estimated by \citet{Oaks2012} are correct.
-The sole purpose of the estimates of ultrametric gene trees presented by
-\citet{Oaks2012} was to provide a very rough comparison of the gene divergence
-times across the 22 taxa.  These analyses assumed an arbitrary strict clock of
-$2\e{-8}$ for all taxa, and are, of course, subject to estimation error.
-Furthermore, the branch-length units of the gene trees are in millions of
-years, whereas the divergence time prior of \msb is in generations, thus the
-logic of \citet{Hickerson2013} requires the additional assumption that all 22
-Philippine taxa have a generation time of one year.
-Thus, the argument of \citet{Hickerson2013} that divergence time estimates of
-\citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
-questionable, especially given our findings on the behavior of empirically
-informed priors presented above.
-
-Finally, even \emph{if} we make all of these assumptions and assume the gene
-divergence times are estimated without error, the probabilistic argument only
-applies to one of the three different priors used by \citet{Oaks2012}.
-The narrowest prior on divergence times used by \citet{Oaks2012} closely mirrors
-the range of estimates of gene-divergence times, and applying Hickerson et al.'s
-(\citeyear{Hickerson2013}) probability equations demonstrates that the prior
-is densely populated with samples with large numbers of divergence parameters
-that are consistent with the gene divergence estimates.
-Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
-sampling is to blame for the bias, it should be much reduced under the narrow
-prior on \divt{}.
-However, the magnitude of the bias is very similar across all three priors
-explored by \citet{Oaks2012}.
-\citet{Hickerson2013} point out a case where the narrow prior performs
-slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
-However, it is important to note that these results suffered from a bug
-in \msb, and there are many cases after \citet{Oaks2012} corrected the 
-bug where the narrow prior performs slightly worse (see panels D--J of
-Figures 3 and S12).
-
-To disentangle whether model likelihoods or insufficient prior sampling is to
-blame for the biases revealed by \citet{Oaks2012}, we must look at the
-different predictions made by these two phenomena.
-One example, as discussed by \citet{Oaks2012}, is that insufficient prior
-sampling should create higher variance in posterior estimates, and thus it
-should cause analyses to be sensitive to the number of samples drawn from the
-prior.
-\citet{Oaks2012} do not see such sensitivity when they compare prior sample
-sizes of $2\e6$, $5\e6$, and $10^7$.
-
-To explore this prediction further, we repeat the analysis of the Philippines
-dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim U(0,
-10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
-(0.0005, 0.02)$), using a very large prior sample size of $10^8$.
-When we look at the trace of the estimates of the dispersion index of
-divergence times (\vmratio{}) as the prior samples accumulate
-(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or
-GLM-regression-adjusted estimates.
-This suggests that insufficient prior sampling did not play a large
-role in the bias found by \citet{Oaks2012}.
-
-Our results presented above that demonstrate the bias of the model-averaging
-approach of \citet{Hickerson2013} both toward models with narrower \divt{}
-priors (Table~\ref{tabModelChoiceEmpirical} and
-Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
-and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
-\ref{figPowerOmegaProb}) strongly suggest the primary cause is the uniform
-priors reducing the marginal likelihoods of large models.
-In all of our model-averaging analyses, all prior models have the same number
-of samples.
-Thus, it is difficult to explain the method's propensity toward the model with
-the smallest parameter space via insufficient prior sampling.
-While analyses that sample each model proportional to their relative parameter
-space could be explored, it seems much more likely that the broad uniform
-priors are simply inhibiting the marginal likelihoods of spacious models.
-
-As discussed by \citet{Oaks2012}, a straightforward prediction of reduced
-likelihoods of complex models is that they should disappear as the model
-generating the data converges to the prior.
-\citet{Oaks2012} test this prediction by performing 100,000 simulations to
-assess the model-choice behavior of \msb when the prior model is correct.
-They find that the bias actually switches direction (at least for the
-regression-adjusted estimates) and the method tends to underestimate the
-probability of the model with one divergence event (see Figure 4 of
-\citet{Oaks2012}).
-The same prediction is not as straightforward for the insufficient-sampling
-hypothesis.
-Even when the prior is correct, due to the discrete uniform prior on
-\numt{} implemented in \msb, models with larger numbers of divergence
-events (and thus greater parameter space) will still be less densely
-sampled than those with fewer divergence events \citep{Oaks2012}.
-Thus, the results of the simulations of \citet{Oaks2012} are more consistent
-with marginal model likelihoods causing the bias toward models with fewer
-divergence time parameters.
-
-Taken together, the results presented here and in \citet{Oaks2012} support
-reduced marginal likelihoods of large models as a primary mechanism by which
-broad uniform priors cause biases in \msb.
-Nonetheless, posterior sampling error will always be present in any numerical
-Bayesian approximation method.
-Thus, insufficient sampling of the prior will contribute to the error of all
-approximate Bayesian estimates.
-However, there is no strong evidence that it is playing a large role in
-the biases revealed herein and by \citet{Oaks2012}.
-
-%CWL: This seems to specific, not sure about including
-\section*{Validation analyses}
-% \highLight{Not sure where to put this section}.
-% \highLight{This justifies presenting the unadjusted results in this paper.}
-% \highLight{Is it worth including just for this purpose?}
-Following \citet{Oaks2012}, we characterize the model-choice behavior of the
-model-averaging approach of \citet{Hickerson2013} under the ideal conditions
-where the prior is correct (i.e., the data are generated from parameters drawn
-from the same prior distributions used in the analysis).
-We used the same prior models as above ($M_1$--$M_5$;
-Table~\ref{tabModelChoiceEmpirical}), and generated 50,000 pseudo-replicate
-datasets under this prior (10,000 from each model).
-We used a simulated data structure of eight population pairs, with a single
-1000 base-pair locus sampled from 10 individuals from each population.
-We then analyzed each of these replicate datasets using the same prior,
-retaining 1000 posterior samples.
-Our results are very similar to \citet{Oaks2012}, but we note that they
-are not directly comparable as our simulations contained eight population
-pairs rather than 10 (Figure~\ref{figValidationMCBehavior}).
-We find that the approach of \citet{Hickerson2013} estimates the posterior
-probability of divergence models reasonably well when all assumptions of the
-method are met (i.e., the prior is correct) and the unadjusted posterior
-estimates are used.
-Similar to \cite{Oaks2012}, we find that the regression-adjusted estimates of
-the model probabilities are biased.
-
-\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
-The main parameter of the \msb model is the vector of divergence
-times for each of the taxon pairs,
-$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
-\citep{Oaks2012}.
-\citet{Hickerson2013} argue that the dispersion index of this vector,
-\vmratio{}, is a better model-choice estimator than the number of 
-divergence time parameters within the vector,
-\numt{}.
-They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
-\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
-This plot shows, not surprisingly, that \vmratio{} has very little information
-about the number of divergences among taxa.
-Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
-biogeographically relevant than \numt{}.
-% We struggle to follow this logic.
-Certainly the maximum information is contained within the divergence time
-vector that \vmratio{} is summarizing.
-Clearly, the number of divergence time parameters within the vector and their
-values is more informative than its variance (i.e., the dispersion index is not
-a sufficient statistic for \divtvector).
-
-\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
-better than \numt{}.''
-However, \msb is a model-choice estimator, and hence the goal is to estimate the
-posterior probability of divergence models.
-\citet{Oaks2012} demonstrate that even when all assumptions of the model are
-met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
-4), whereas \numt{} performs better.
-Furthermore, \vmratio{} is limited to estimating the probability of only a
-single divergence model (the one divergence model), and thus its utility for
-model choice is extremely limited.
-
-The model-choice utility of \vmratio{} is limited to the probability that
-this continuous statistic, which can range from zero to infinity, is at its lower
-limit of zero. In theory, this point density will always be zero, thus an
-arbitrary threshold (0.01 is used throughout the \msb literature) must 
-be chosen to make the probability estimable.
-However, it is still not surprising to see that it is numerically difficult to
-obtain reliable estimates of the probability that the continuous \vmratio{}
-statistic is ``near'' its limit of zero.
-It is much easier, less subjective, and more interpretable to estimate
-the probability of the discrete parameter of the model, \numt{}, is
-at its lower limit of one.
-Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better
-estimator of model probability than \vmratio{}.
-
-\section*{Some general thoughts on the model of \msb}
-Our results demonstrating the poor behavior of the model-averaging approach
-proposed by \citet{Hickerson2013} are not too surprising.
-While this approach is trivial to implement, it is not a trivial change
-to the \msb model.
-To better understand the reason for this, it might help to step back and get a
-sense of the scale of the \msb model.
-To do this, we will use the dataset of 22 vertebrate taxon pairs of
-\citet{Oaks2012} as an example.
-
-Following the model description and notation of \citet{Oaks2012}, let us tally
-up all of the free parameters in the \msb model.  Under the simplest model in
-\msb (i.e., assuming no migration and
-no intra-locus recombination), the number of parameters for each
-taxon pair include:
-The population sizes of the ancestral and descendant populations
-(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
-the magnitude of population contraction in each of the descendant
-populations (\bottleScalar{1}{} and \bottleScalar{2}{}) and the
-timing of these contractions (\bottleTime{}), and the $N-1$ node heights
-(coalescent times) of the gene tree that gave rise to the $N$ gene
-copies sampled from both populations of the pair.
-Because we only have a single locus for each taxon, there are no locus-specific
-\myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
-\locusRateHetShapeParameter shape parameter for the gamma prior distribution on
-\locusMutationRateScalar{}.
-Lastly, there are between one and 22 divergence time parameters \divt{} in
-the vector \divtvector.
-Overall, for our Philippines data under the simplest model in \msb, there are
-581--602 free parameters (depending on the cardinality of \divtvector).
-Furthermore, under this rich model, the method is sampling over 1002 divergence
-models \citep[i.e., the number of integer partitions of $Y=22$;][]{Oaks2012}.
-
-This is a very difficult inference problem, and the method only uses four
-summary statistics calculated from the sequence alignment of each taxon pair:
-$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
-\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
-That gives us a total of 88 summary statistics (four from each of the 22 taxon
-pairs), which contain minimal information about many of the $\approx 600$
-parameters in the model.
-More summary statistics can be used in \msb, but most are highly correlated
-with these four statistics (which are even highly correlated among themselves),
-and thus contribute little additional information about the data.
-
-The large number of parameters and divergence models relative to 
-summary statistics is undoubtedly a key reason the method is so sensitive
-to the prior distributions.
-It also likely contributes to the method's lack of robustness.
-% We use robustness here as a measure of a method's insensitivity to model
-% violations.
-Robustness is an extremely important characteristic of a method to gauge its
-applicability to real-world data, because we know the model and priors will be
-wrong to some degree.
-
-The approach of \citet{Hickerson2013} adds an additional dimension of model
-choice to the model. They expand the model to sample over eight prior models.
-This extends the original model to having 582-603 free parameters and, more
-importantly, sampling over 8016 unique model states across both divergence
-models and prior models.
-This is a non-trivial extension of the model and, given the method's very large
-number of parameters and divergence models relative to the information used
-from the data, likely plays a major role in the poor behavior of this approach.
-% Certainly, such a large modification of the model demands simulation-based
-% assessments of the behavior of the new model, as we provide here.
-
-In theory, the approach of \citet{Hickerson2013} is very appealing.  It sums
-over multiple candidate prior models to produce a posterior estimate
-marginalized over the uncertainty in prior choice.
-In general, Bayesian model-averaging is a powerful approach that leverages a
-great strength of Bayesian statistical procedures, namely the ability to
-obtain marginalized estimates that incorporate uncertainty in nuisance
-parameters.
-However, given the basic \msb model is already struggling to estimate
-a huge number of parameters and model probabilities with scant information
-from the data, it is not surprising that adding an additional layer of
-model choice to the method causes problems.
-
-The recommendations of \citet{Oaks2012} for mitigating the bias and lack of
-robustness of \msb are actually similar to those of \citet{Hickerson2013}, but
-avoid the need for imposing additional model choice.
-\citet{Oaks2012} suggest that uniform priors are inappropriate for many
-parameters of the \msb model, and recommend the use of probability
-distributions from the exponential family.
-If we look at the prior distribution on the divergence time parameter \divt{}
-imposed by the model-averaging approach of \citet{Hickerson2013} we see
-it is a mixture of overlapping uniforms with lower limits of zero
-(Figure~\ref{figMCTauPrior}).
-This looks very much like an exponential distribution, except that each sample
-of divergence times are restricted to the hard bounds of one of the eight prior
-models.
-Thus, it seems more appropriate to place a gamma prior (the exponential being a
-special case of the gamma) on divergence times.
-This would provide investigators flexibility to represent their prior
-uncertainty in model parameters while avoiding broad uniform
-distributions and the need of costly model-averaging.
-This seems like a natural solution to the problem of skewed marginal
-likelihoods and/or sampling error, which are both likely exacerbated by the
-fact that uniform distributions are the only choice of prior for many of the
-parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
-\descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
-\bottleScalar{1}{}, \bottleScalar{2}{}, \locusRateHetShapeParameter,
-\migrationRate{}, \recombinationRate;][]{Oaks2012}.
-
-Importantly, there is no cost to this strategy.
-Given that ABC methods only need to draw random values from prior distributions
-(hence there are no difficult prior and proposal ratios to calculate, etc.) it
-is easy to use the most appropriate distributions on parameters.
-Also, this approach reduces the temptation of empirically guided priors.
-The investigator can place the majority of the prior density in regions of
-parameter space they believe, \emph{a priori}, are most plausible, but still
-capture uncertainty in the tails of distributions with low density.
-As discussed in \citet{Oaks2012} the use of uniform priors necessitate
-broad priors to avoid excluding the truth \emph{a priori}.
-Thus, to avoid the behavior of the method under these uniform priors, using the
-data is a very tempting, albeit risky, option \citep{Hickerson2013}.
-More appropriate prior distributions would alleviate this issue.
-
-%CWL: This seems to nit-picky. Probably not worth including it
-\section*{Other clarifications}
-
-\subsection*{Graphical checks of priors}
-\citet{Hickerson2013} advocate the use of what they call graphical checks of
-prior models.
-This entails generating a small number of random samples from the prior (1000)
-and plotting the resulting summary statistics in comparison to the observed
-statistics to see if they coincide (see Figure 1 of \citet{Hickerson2013}).
-As we show above, this is a risky strategy, because the resulting plots of this
-approach have little correlation with the appropriateness of priors.
-Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
-dataset analyzed by \citet{Hickerson2013}), we do not expect that 1000
-\emph{random} draws from the vast prior parameter space will yield data and
-summary statistics consistent with the observed data.
-In fact, when the random draws are tightly clustered around the observed
-statistics, this can be an indication that the prior is over-fit, as we show
-above (Table~\ref{tabModelChoiceEmpirical} and Figure~\ref{figPCA}).
-Thus, interpreting such plots should be avoided, and the use of posterior
-predictive analyses would be much more informative about the overall fit of
-models.
-
-\subsection*{The validity of \msb estimates}
-\citet{Hickerson2013} claim that \citet{Oaks2012} ``assume that all previous
-\msb results are invalid.''
-With the exception of \citet{Hickerson2013}, we do not claim any previous
-results of \msb are invalid.
-Rather, we conservatively recommend that the common inference of temporally
-clustered divergences
-\citep{Barber2010, Bell2012, Carnaval2009, Chan2011, Daza2010, Hickerson2006,
-    Huang2011, Lawson2010, Leache2007, Plouviez2009, Stone2012, Voje2009},
-when not accompanied with the necessary simulation-based analyses to guide the
-interpretation of such results, should be treated with caution, because the
-method has been shown to spuriously infer clustered divergences over a range of
-prior conditions.
-
-\subsection*{Saturation of summary statistics}
-\citet{Hickerson2013} claim that our priors ``cause much of the explored
-parameter space to be beyond the threshold of saturation in most mtDNA genes.''
-To explore this possibility, we simulated datasets under prior settings that
-match two of the three priors used by \citet{Oaks2012}: $\meanDescendantTheta{}
-\sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
-Under this prior, we draw divergence time parameters from a uniform
-distribution of $U(0, 20)$, simulate datasets, and plot the \divt{} values
-against the summary statistics calculated from the resulting datasets
-(Figure~\ref{figSaturationPlot}).
-Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
-and 10 suffered from little to no effect from saturation.
-Even at divergence times of 20 coalescent units, there is still signal in the
-summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
-Thus, the assertion of \citet{Hickerson2013} does not apply to at least
-two of the priors used by \citet{Oaks2012} and, as a result, does not
-explain the bias they found.
-
-\section*{Conclusions}
-We demonstrate how the approximate Bayesian model choice method implemented in
-\msb can be strongly biased away from models with greater parameter space.
-As suggested by \citet{Oaks2012}, this is likely caused by the inappropriate
-use of uniform priors on most of the model's parameters.
-These priors necessitate the use of broad priors that place high prior density
-in unlikely regions of parameter space, less the risk of excluding the truth
-\emph{a priori}.
-This likely reduces the marginal likelihoods of models with greater parameter
-space, either due to more divergence time parameters or broader prior
-distributions on those parameters.
-We show that the empirical Bayesian model-averaging approach of
-\citet{Hickerson2013} does not mitigate this bias, but rather causes it to
-manifest by sampling predominantly from models that may exclude the true values
-of the parameters.
-Exploring alternative prior probability distributions on most of the model's
-parameters, in addition to different priors over divergence models, should
-help mitigate the method's biases.
-
-The work presented herein follows the principles of Open Notebook Science.
-All aspects of the work were recorded in real-time via version-control
-software and are publicly available at
-\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
-All information necessary to reproduce our results are provided there.
-
-\section*{Acknowledgments}
-We thank the National Science Foundation for supporting this work (DEB
-1011423).
-J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
-Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
-Society, KU Department of Ecology and Evolutionary Biology, and the KU
-Biodiversity Institute.
-
-\bibliography{../bib/references}
-% \bibliography{references}
-
-%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\renewcommand\listfigurename{Figure Captions}
-\cftsetindents{fig}{0cm}{2.2cm}
-\renewcommand\cftdotsep{\cftnodots}
-\setlength\cftbeforefigskip{10pt}
-\cftpagenumbersoff{fig}
-\listoffigures
-
-\end{linenumbers}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-%:TABLE-re-analysis
-% \begin{table}[htbp]
-%     \sffamily
-%     \footnotesize
-%     \addtolength{\tabcolsep}{-0.08cm}
-%     \rowcolors{2}{}{myGray}
-%     %\captionsetup{font=footnotesize}
-%     \caption{Results of the model-averaging approach of \citet{Hickerson2013}
-%         applied to the Philippines dataset of \citet{Oaks2012} using eight
-%         prior models (see \citet{Hickerson2013} Table 1).  The results of (1)
-%         \citet{Hickerson2013}, (2) our full re-analysis, and of (3) repeating
-%         the final rejection step of \citet{Hickerson2013} using the prior means and
-%         standard deviations to normalize the summary statistics, are all compared.}
-%     \centering
-%     \begin{tabular}{ l l l l }
-%         \toprule
-%          & Hickerson et al. (2013) & Re-analysis & blah \\
-%         $\hat{\numt{}}$ & 2 & 1 & 1 \\
-%         $p(\numt{} = 1 | \ssSpace)$ & 0.062 & 0.206 & 0.146 \\
-%         $BF_{\numt{} = 1, \numt{} \neq 1}$ 1.39 & 5.45 & 3.59 \\
-%         $p(\vmratio{} < 0.01 | \ssSpace)$ & 0.298 & 0.416 & 0.361 \\
-%         $BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01}$ 6.65 & 11.16 & 8.85 \\
-%         $\hat{M}$ & 1 & 2 & 2 \\
-%         $p(M_1 | \ssSpace)$ & 0.513 & 0.358 & 0.376 \\
-%         $p(M_2 | \ssSpace)$ & 0.286 & 0.369 & 0.389 \\
-%         $p(M_3 | \ssSpace)$ & 0.036 & 0.086 & 0.089 \\
-%         $p(M_4 | \ssSpace)$ & 0.002 & 0.012 & 0.011 \\
-%         $p(M_5 | \ssSpace)$ & 0.142 & 0.129 & 0.107 \\
-%         $p(M_6 | \ssSpace)$ & 0.020 & 0.043 & 0.024 \\
-%         $p(M_7 | \ssSpace)$ & 0.001 & 0.003 & 0.004 \\
-%         $p(M_8 | \ssSpace)$ & 0.000 & 0.000 & 0.001 \\
-%         \bottomrule
-%     \end{tabular}
-%     \label{tabModelChoiceEmpirical}
-% \end{table}
-
-%:TABLE-model-choice-empirical
-\begin{table}[htbp]
-    \sffamily
-    % \footnotesize
-    \addtolength{\tabcolsep}{-0.08cm}
-    \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
-    \caption{Results of the model-averaging approach of \citet{Hickerson2013}
-        applied to the Philippines dataset of \citet{Oaks2012} using three sets
-        of prior models. All models used priors on population size of
-        $\meanDescendantTheta{} \sim U(0.0001,0.1)$ and $\ancestralTheta{} \sim
-        U(0.0001, 0.05)$, and differ only in their prior on divergence time
-        (\divt{}) parameters.  Each set of five models differ only in the
-        divergence time prior used for the model with the narrowest prior:
-        $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_{1A}$ ($\divt{} \sim U(0, 0.01)$),
-        or $M_{1B}$ ($\divt{} \sim U(0, 0.001)$). The approximate posterior
-        probability of each model ($p(M_i \given \ssSpace)$) is given for each
-        of the three analyses.  The posterior estimates are based on 10,000
-        samples retained from $1\e6$ prior samples
-    from each model.}
-    \centering
-    \begin{tabular}{ l l l l l }
-        \toprule
-        & & \multicolumn{3}{c}{$p(M_i \given \ssSpace)$} \\
-        \cmidrule(){3-5}
-        Model & \divt{} prior & $M_{*}=M_1$ & $M_{*}=M_{1A}$ & $M_{*}=M_{1B}$ \\
-        \midrule
-        $M_*$ & --        & 0.899 & 0.821 & 0.673 \\
-        $M_2$ & $U(0,1)$  & 0.079 & 0.136 & 0.251 \\
-        $M_3$ & $U(0,5)$  & 0.013 & 0.026 & 0.044 \\
-        $M_4$ & $U(0,10)$ & 0.006 & 0.012 & 0.022 \\
-        $M_5$ & $U(0,20)$ & 0.003 & 0.005 & 0.010 \\
-        \bottomrule
-    \end{tabular}
-    \label{tabModelChoiceEmpirical}
-\end{table}
-
-\clearpage
-
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
 \mFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
@@
 % }{figJointPosterior}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-
-\singlespacing
-
-\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape supporting information}}
-
-% PUT MAIN TEXT CITATION HERE
-% \begin{thebibliography}{1}
-% \providecommand{\natexlab}[1]{#1}
-% \providecommand{\url}[1]{\texttt{#1}}
-% \providecommand{\urlprefix}{URL }
-
-% \bibitem
-
-% \end{thebibliography}
-
-%% SUPPL TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\setcounter{table}{0}
-
-\end{document}

### 2013-10-18 manuscripts/reply/reply-main.tex
Moving sections of `reply.tex` into separate files.
Moving the main text, tables, and figures of `reply.tex` to
`reply-main.tex`, `reply-tables.tex`, and `reply-figures.tex`.

--- 2013-10-01 manuscripts/reply/reply.tex
@@
-%&<latex>
-\documentclass[letterpaper,12pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{../utils/preamble.tex}
-\input{../utils/macros.tex}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\doublespacing
-\raggedright
-\setlength{\parindent}{0.5in}
-\begin{linenumbers}
-
-\begin{titlepage}
-    \begin{flushleft}
-        \sffamily
-
-        \MakeUppercase{\large\bfseries Why you should not fix a biased
-        model-choice method by adding an additional dimension of model choice:
-        A reply to Hickerson et al.}
-
-        \vspace{12pt}
-        \textbf{Running head:} \MakeUppercase{Approximate Bayesian model
-        choice}
-
-        \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2}$, Jeet Sukumaran$^{3}$, Jacob A.\
-        Esselstyn$^{4}$, Charles W.\ Linkem$^{5}$, Cameron D.\
-        Siler$^{6}$, Mark T.\ Holder$^{1}$ and Rafe M.\ Brown$^{1}$
-
-        \bigskip
-        $^1$\emph{Department of Ecology and Evolutionary Biology,
-            Biodiversity Institute,
-            University of Kansas,
-            Lawrence, Kansas 66045}\\[.1in]
-        $^3$\emph{Department of Biology,
-            Duke University,
-            Durham, North Carolina 27708} \\[.1in]
-        $^4$\emph{Museum of Natural Science,
-            Louisiana State University,
-            119 Foster Hall,
-            Baton Rouge, Louisiana 70803}\\[.1in]
-        $^5$\emph{Department of Biology,
-            University of Washington,
-            Seattle, Washington 98195}\\[.1in]
-        $^6$\emph{Sam Noble Museum,
-            Department of Biology,
-            University of Oklahoma,
-            Norman, Oklahoma 73072}\\[.1in]
-        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
-        joaks1@gmail.com})\\
-
-    \end{flushleft}
-\end{titlepage}
-
-{\sffamily
-    \noindent\textbf{ABSTRACT} \\
-    \noindent Recently, \citet{Oaks2012} presented a simulation-based
-    assessment of the approximate Bayesian phylogeographical model choice
-    method implemented in \msb, and found the method to lack robustness and
-    power to detect temporal variation in divergences that are randomly
-    distributed over relatively broad periods of evolutionary history.
-    They conclude that a likely mechanism behind the bias toward models with
-    few divergence events is the small marginal likelihoods of models with more
-    divergence time parameters due to the integration over vast parameter space
-    with low likeihood and high prior density that is imposed by broad uniform
-    priors.
-    \citet{Hickerson2013} have responded to this work, criticising the priors
-    used by \citet{Oaks2012} and presenting an approximate Bayesian
-    model-averaging approach that uses narrow, empirically informed uniform
-    priors.
-    \citet{Hickerson2013} use this method to re-analyze the empirical dataset
-    of \citet{Oaks2012} comprised of 22 pairs of vertebrate populations from
-    the Philippines, and conclude the model-averaging approach circumvents the
-    issues revealed by \citet{Oaks2012} and yields robust inference of
-    divergence models.
-    Furthermore, they argue that the biases revealed by \citet{Oaks2012} are
-    inherent to the inefficient rejection algorithm implemented in \msb, rather
-    than a result of low marginal likelihoods of parameter-rich models.
-    Here we discuss an error in the methodology of \citet{Hickerson2013} that
-    renders some of their results invalid and the remainder difficult to
-    interpret.
-    Furthermore, we demonstrate that the approach of \citet{Hickerson2013} is
-    biased toward models with less parameter space, which manifests in a strong
-    tendency to sample predominantly from models that exclude the true values
-    of the model's parameters.
-    Our results suggest that their model-averaged results are dominated by
-    models that likely exclude at least some of the true divergence times
-    across the 22 pairs of Philippine vertebrates, which adds to the
-    difficulty in interpreting their results.
-    We also show the approach of \citet{Hickerson2013} is still biased toward
-    inferring simultaneous divergences even when the true divergences are
-    randomly distributed over millions of generations.
-    Our results, coupled with those of \citet{Oaks2012}, strongly suggest that
-    low marginal likelihoods of complex models is a primary cause of the bias
-    of \msb toward models with less parameter space.
-    We discuss ways to mitigate this bias without resorting to adding an
-    additional dimension of model choice to the model or using questionable
-    empirically informed uniform priors.
-
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Approximate Bayesian computation; Bayesian
-        model choice; empirical Bayes}
-
-}
-
-\newpage
-
-%CWL:
-% \highLight{The goal of this paper needs to be to highlight the major
-%     differences of opinion on how ABC can be used in comparative phylogeography
-%     and how these changes to msbayes can be made that will improve the
-%     inference in the future. We should avoid nit-picking the Hickerson paper
-%     and try to stay positive.}
-
-\noindent Recently, this journal has served as a medium for a
+Recently, this journal has served as a medium for a
 discussion of the potential pitfalls of approximate Bayesian methods of
@@
 All information necessary to reproduce our results are provided there.
-
-\section*{Acknowledgments}
-We thank the National Science Foundation for supporting this work (DEB
-1011423).
-J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
-Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
-Society, KU Department of Ecology and Evolutionary Biology, and the KU
-Biodiversity Institute.
-
-\bibliography{../bib/references}
-% \bibliography{references}
-
-%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\renewcommand\listfigurename{Figure Captions}
-\cftsetindents{fig}{0cm}{2.2cm}
-\renewcommand\cftdotsep{\cftnodots}
-\setlength\cftbeforefigskip{10pt}
-\cftpagenumbersoff{fig}
-\listoffigures
-
-\end{linenumbers}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-%:TABLE-re-analysis
-% \begin{table}[htbp]
-%     \sffamily
-%     \footnotesize
-%     \addtolength{\tabcolsep}{-0.08cm}
-%     \rowcolors{2}{}{myGray}
-%     %\captionsetup{font=footnotesize}
-%     \caption{Results of the model-averaging approach of \citet{Hickerson2013}
-%         applied to the Philippines dataset of \citet{Oaks2012} using eight
-%         prior models (see \citet{Hickerson2013} Table 1).  The results of (1)
-%         \citet{Hickerson2013}, (2) our full re-analysis, and of (3) repeating
-%         the final rejection step of \citet{Hickerson2013} using the prior means and
-%         standard deviations to normalize the summary statistics, are all compared.}
-%     \centering
-%     \begin{tabular}{ l l l l }
-%         \toprule
-%          & Hickerson et al. (2013) & Re-analysis & blah \\
-%         $\hat{\numt{}}$ & 2 & 1 & 1 \\
-%         $p(\numt{} = 1 | \ssSpace)$ & 0.062 & 0.206 & 0.146 \\
-%         $BF_{\numt{} = 1, \numt{} \neq 1}$ 1.39 & 5.45 & 3.59 \\
-%         $p(\vmratio{} < 0.01 | \ssSpace)$ & 0.298 & 0.416 & 0.361 \\
-%         $BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01}$ 6.65 & 11.16 & 8.85 \\
-%         $\hat{M}$ & 1 & 2 & 2 \\
-%         $p(M_1 | \ssSpace)$ & 0.513 & 0.358 & 0.376 \\
-%         $p(M_2 | \ssSpace)$ & 0.286 & 0.369 & 0.389 \\
-%         $p(M_3 | \ssSpace)$ & 0.036 & 0.086 & 0.089 \\
-%         $p(M_4 | \ssSpace)$ & 0.002 & 0.012 & 0.011 \\
-%         $p(M_5 | \ssSpace)$ & 0.142 & 0.129 & 0.107 \\
-%         $p(M_6 | \ssSpace)$ & 0.020 & 0.043 & 0.024 \\
-%         $p(M_7 | \ssSpace)$ & 0.001 & 0.003 & 0.004 \\
-%         $p(M_8 | \ssSpace)$ & 0.000 & 0.000 & 0.001 \\
-%         \bottomrule
-%     \end{tabular}
-%     \label{tabModelChoiceEmpirical}
-% \end{table}
-
-%:TABLE-model-choice-empirical
-\begin{table}[htbp]
-    \sffamily
-    % \footnotesize
-    \addtolength{\tabcolsep}{-0.08cm}
-    \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
-    \caption{Results of the model-averaging approach of \citet{Hickerson2013}
-        applied to the Philippines dataset of \citet{Oaks2012} using three sets
-        of prior models. All models used priors on population size of
-        $\meanDescendantTheta{} \sim U(0.0001,0.1)$ and $\ancestralTheta{} \sim
-        U(0.0001, 0.05)$, and differ only in their prior on divergence time
-        (\divt{}) parameters.  Each set of five models differ only in the
-        divergence time prior used for the model with the narrowest prior:
-        $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_{1A}$ ($\divt{} \sim U(0, 0.01)$),
-        or $M_{1B}$ ($\divt{} \sim U(0, 0.001)$). The approximate posterior
-        probability of each model ($p(M_i \given \ssSpace)$) is given for each
-        of the three analyses.  The posterior estimates are based on 10,000
-        samples retained from $1\e6$ prior samples
-    from each model.}
-    \centering
-    \begin{tabular}{ l l l l l }
-        \toprule
-        & & \multicolumn{3}{c}{$p(M_i \given \ssSpace)$} \\
-        \cmidrule(){3-5}
-        Model & \divt{} prior & $M_{*}=M_1$ & $M_{*}=M_{1A}$ & $M_{*}=M_{1B}$ \\
-        \midrule
-        $M_*$ & --        & 0.899 & 0.821 & 0.673 \\
-        $M_2$ & $U(0,1)$  & 0.079 & 0.136 & 0.251 \\
-        $M_3$ & $U(0,5)$  & 0.013 & 0.026 & 0.044 \\
-        $M_4$ & $U(0,10)$ & 0.006 & 0.012 & 0.022 \\
-        $M_5$ & $U(0,20)$ & 0.003 & 0.005 & 0.010 \\
-        \bottomrule
-    \end{tabular}
-    \label{tabModelChoiceEmpirical}
-\end{table}
-
-\clearpage
-
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
-\mFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
-    The joint posterior of the mean (\meant{}{}) and dispersion index ($\vmratio{} = 
-    \vart{}{}/\meant{}{}$) of divergence times for 22 vertebrate taxon pairs as
-    estimated by \citet{Hickerson2013} (see Figure 2B of \citet{Hickerson2013}).
-    The posterior samples are color-coded to indicate the erroneous mixture of
-    timescales in the analysis of \citet{Hickerson2013};
-    grey = $0.05/\mutationRate$ generations and
-    black = $0.02/\mutationRate$ generations.
-}{figJointPosteriorHickerson}
-
-% \mFigure{../images/coin_flip.pdf}{
-%     A plot of three beta probability density functions that represent a prior
-%     (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
-%     empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
-%     Bernoulli trials, four of which are successes.
-% }{figCoinFlip}
-
-\mFigure{../../model-choice/priors-for-pc-plot/pc-plots.pdf}{
-    The graphical checks recommended by \citet{Hickerson2013} for three prior
-    models: (A) $M_1$ ($\divt{} \sim U(0, 0.1)$), (B) $M_{1A}$ ($\divt{} \sim
-    U(0, 0.01)$), and (C) $M_{1B}$ ($\divt{} \sim U(0, 0.001)$).
-    The plots project the summary statistics from 1000 random samples from each
-    model onto the first two orthogonal axes of a principle component analysis,
-    with the blue dot representing the observed summary statistics from the 22
-    population pairs of Philippine vertebrates.
-}{figPCA}
-
-%% Exclusion simulation results
-\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/num_tau_excluded.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to
-    exclude the truth.
-    The plots illustrate the number of true \divt{} parameters excluded from analyses of
-    simulated datasets where \divt{} for 22 pairs of populations is drawn from an
-    exponential distribution, $\divt{} \sim Exp(2)$.
-    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
-    simulation replicates analyzed using $5\e6$ samples from the prior.
-    The proportion of simulation replicates in which at least one true
-    parameter value is excluded from the model preferred by a Bayes factor
-    ($p(\divt{} \notin \hat{M})$) is also given.
-}{figExclusionSimTau}
-
-\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/prob_of_exclusion.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to
-    exclude the truth.
-    The plots illustrate the estimated probability of excluding at least one
-    true \divt{} value from analyses of simulated datasets where \divt{} for 22
-    pairs of populations is drawn from an exponential distribution, $\divt{}
-    \sim Exp(2)$.
-    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
-    simulation replicates analyzed using $5\e6$ samples from the prior.
-    The proportion of simulation replicates in which there is strong support
-    for at least one true parameter value being excluded from the model
-    ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is also given.
-}{figExclusionSimProb}
-
-%% Power results
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
-    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
-    estimate the dispersion index of divergence times (\vmratio{}) from
-    analyses of simulated datasets where \divt{} for 22 pairs of populations is
-    drawn from a series of uniform distributions, $\divt{} \sim U(0,
-    \divt{max})$.
-    The proportion of estimates less than the true value of,
-    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
-    Each plot represents unadjusted median estimates from 500 simulation
-    replicates analyzed using $5\e6$ samples from the prior.
-}{figPowerAccOmegaMedian}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
-    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
-    estimate the dispersion index of divergence times (\vmratio{}) from
-    analyses of simulated datasets where \divt{} for 22 pairs of populations is
-    drawn from a series of uniform distributions, $\divt{} \sim U(0,
-    \divt{max})$.
-    The proportion of estimates less than the true value of,
-    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
-    Each plot represents GLM-regression-adjusted mode estimates from 500
-    simulation replicates analyzed using $5\e6$ samples from the prior.
-}{figPowerAccOmegaModeGLM}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_psi_mode.pdf}{
-    The bias of the model-averaging approach of \citet{Hickerson2013} to infer
-    clustered divergences in our simulation-based power analyses.
-    The plots illustrate the estimated number of divergence events ($\hat{\numt{}}$)
-    from analyses of simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
-    uniform distributions, $\divt{} \sim U(0, \divt{max})$.
-    The estimated probability of the method inferring one divergence event, $p(\hat{\numt{}} = 1)$,
-    is given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerPsiMode}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_omega_prob.pdf}{
-    The bias of the model-averaging approach of \citet{Hickerson2013} to
-    support one divergence event in our simulation-based power analyses.
-    The plots illustrate histograms of the estimated posterior probability that
-    the dispersion index of divergence times is less than 0.01 ($p(\vmratio{} <
-    0.01 | \ssSpace)$) from analyses of simulated datasets where \divt{} for 22
-    pairs of populations is drawn from a series of uniform distributions,
-    $\divt{} \sim U(0, \divt{max})$.
-    The proportion of simulation replicates that strongly support one
-    divergence event, $p(BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01} > 10)$,
-    is given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerOmegaProb}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_num_excluded.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to exclude
-    the truth in our simulation-based power analyses.
-    The plots illustrate the number of true \divt{} parameters excluded from analyses of
-    simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
-    uniform distributions, $\divt{} \sim U(0, \divt{max})$. The proportion
-    of simulation replicates in which at least one true parameter value is excluded from
-    the model preferred by a Bayes factor ($p(\divt{} \notin \hat{M})$) is
-    given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerNumExcluded}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_prob_exclusion.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to
-    exclude the truth in our simulatoin-based power analyses.  The plots
-    illustrate the estimated posterior probability of excluding at least one
-    true \divt{} value from analyses of simulated datasets where \divt{} for 22
-    pairs of populations is drawn from a series of uniform distributions,
-    $\divt{} \sim U(0, \divt{max})$. The proportion of simulation replicates in
-    which there is strong support for at least one true parameter value being
-    excluded from the model ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is
-    given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerProbExclusion}
-
-%% Validation results
-\mFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
-    An assessment of the approximate Bayesian model-averageing approach of
-    \citet{Hickerson2013} under the ideal conditions when the prior model
-    is correct (i.e., the pseudo-replicate datasets are simulated from
-    parameters drawn from the same prior distributions used in the
-    analysis).
-    The plots show the relationship between the estimated posterior and true
-    probability of (A \& C) $\numt{}=1$ and (B \& D) $\vmratio{} < 0.01$, based
-    on 50,000 simulations.
-    The results summarize the (A \& B) unadjusted and (C \& D) GLM-adjusted
-    posterior estimate from each simulation replicate.
-    The prior settings for all replicates included five prior models with
-    $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
-    U(0.0001, 0.05)$ for all five models, and
-    $M_1: \divt{} \sim U(0, 0.1)$,
-    $M_2: \divt{} \sim U(0, 1)$,
-    $M_3: \divt{} \sim U(0, 5)$,
-    $M_4: \divt{} \sim U(0, 10)$, and
-    $M_5: \divt{} \sim U(0, 20)$.
-    The number of samples from the prior was $2.5\e6$.
-    The simulated data structure was 8 population pairs, with a single 1000
-    bp locus sampled from 10 individuals from each population.  The 50,000
-    estimates of the posterior probability of one divergence event were
-    assigned to 20 bins of width 0.05.
-    The estimated posterior probability of each bin is plotted against the
-    proportion of replicates in that bin with a true value consistent with
-    one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
-}{figValidationMCBehavior}
-
-\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
-    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
-    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
-    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
-    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
-    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
-    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
-}{figSamplingError}
-
-%% Model-choice tau prior
-\mFigure{../../model-choice/tau-prior/tau_prior.pdf}{
-    The prior distribution on divergence times imposed by the model-averaging prior
-    comprised of five models with different uniform priors on \divt{}:
-    $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_2$ ($\divt{} \sim U(0, 1)$), $M_3$
-    ($\divt{} \sim U(0, 5)$), $M_4$ ($\divt{} \sim U(0, 10)$), $M_5$ ($\divt{}
-    \sim U(0, 20)$).
-}{figMCTauPrior}
-
-%:FIGURE-saturation plot
-\mFigure{../../saturation/saturation-plot.pdf}{
-    The summary statistics $\pi$ \citep{Tajima1983} and $\pi_{net}$
-    \citep{Takahata1985} as a function of divergence time between populations.
-    Each plot represents 1100 pairs of parameter draws and summary statistics
-    calculated from the simulated data.
-    Prior settings for the simulations were $\divt{} \sim U(0, 20)$,
-    $\meanDescendantTheta{} \sim U(0.0005, 0.04)$, and $\ancestralTheta{} \sim
-    U(0.0005, 0.02)$.
-}{figSaturationPlot}
-
-% \mFigure{../../response-redux/results/hickerson/pymsbayes-results/pymsbayes-output/d1/m12345678-combined/mean_by_dispersion.pdf}{
-%     Results of reanalysis.
-% }{figJointPosterior}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-
-\singlespacing
-
-\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape supporting information}}
-
-% PUT MAIN TEXT CITATION HERE
-% \begin{thebibliography}{1}
-% \providecommand{\natexlab}[1]{#1}
-% \providecommand{\url}[1]{\texttt{#1}}
-% \providecommand{\urlprefix}{URL }
-
-% \bibitem
-
-% \end{thebibliography}
-
-%% SUPPL TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\setcounter{table}{0}
-
-\end{document}

### 2013-10-18 manuscripts/reply/reply.tex
Moving sections of `reply.tex` into separate files.
Moving the main text, tables, and figures of `reply.tex` to
`reply-main.tex`, `reply-tables.tex`, and `reply-figures.tex`.

@@
 %     and try to stay positive.}
-
-\noindent Recently, this journal has served as a medium for a
-discussion of the potential pitfalls of approximate Bayesian methods of
-comparative phylogeographical model choice and prior selection.
-The discussion has centered around the approximate Bayesian model choice method,
-\msb, which estimates the temporal distribution of divergences among
-co-distributed pairs of taxa \citep{Huang2011} by sampling over all
-unordered models of divergence.
-\citet{Oaks2012} published results of simulation-based power analyses
-showing that \msb can often be biased toward inferring models of temporally
-clustered divergence times across taxon pairs.
-\citet{Hickerson2013} published a response to this paper where they present
-an empirical Bayes model-averaging approach that they conclude circumvents the
-poor behavior of the method revealed by \citet{Oaks2012}.
-These papers are largely in agreement, with \citet{Hickerson2013} reiterating
-much of the discussion of \citet{Oaks2012} regarding the impact of broad
-uniform priors on Bayesian model choice.
-The main difference in perspectives are centered around
-(1) the mechanism by which broad uniform priors cause the poor behavior of
-\msb, and
-(2) how to potentially mitigate this issue.
-
-\citet{Oaks2012} suggest the primary cause of the poor behavior of the method
-is likely the low marginal likelihoods of parameter-rich models integrated over
-vast parameter space with low probability of producing the data, yet relatively
-high prior density \citep{Lindley1957}.
-Note, this suggests the bias is extrinsic to \msb, and the numerical
-approximation machinery of \msb could be sound.
-\citet{Hickerson2013} suggest the bias is intrinsic to \msb, concluding the
-method's rejection algorithm is inefficient and will be increasingly biased as
-the overall space of the model increases, either as a function of the number of
-taxon pairs or the width of the uniform priors on nuisance parameters.
-\citet{Hickerson2013} support their position by giving a probabilistic argument
-that focuses on only one of the three prior models used by \citet{Oaks2012}.
-We show that this argument is based on an empirical Bayesian reasoning and
-questionable assumptions, and does not explain the bias of the method found in
-many of the analyses of \citet{Oaks2012}. 
-Furthermore, we present results of additional analyses that, when combined with
-the results of \citet{Oaks2012}, strongly suggest small marginal likelihoods of
-models with more parameter space is a primary mechanism by which uniform priors
-are causing the poor behavior of \msb.
-
-%This paragraph seems too aggressive. It also seems to repeat some of the previous points. I removed it for now.
-
-To mitigate the bias, \citet{Oaks2012} suggest alternative prior probability
-distributions on divergence models and nuisance parameters could increase the
-likelihoods of models with more divergence time parameters.
-They also describe simulation-based procedures for determining power, accuracy,
-and robustness of the method given a dataset, and recommend any application of
-\msb should be accompanied by such procedures.
-Alternatively, \citet{Hickerson2013} present an approximate Bayesian
-model-averaging approach for accommodating uncertainty in selecting among
-empirically guided priors, and champion the method as a means of avoiding the
-pitfalls raised by \citet{Oaks2012}.
-We discuss errors in this approach that render some of their results invalid
-and leave the remaining results difficult to interpret.
-Furthermore, we follow the recommendations of \citet{Oaks2012} and present
-simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
-approach.
-Our results demonstrate that the model-averaging approach of
-\citet{Hickerson2013} is biased toward clustered divergence models and
-predisposed to excluding true parameter space.
-The model-averaging approach provides another means by which the model can
-``escape'' large parameter space, and the method's bias toward smaller models
-remains.
-This manifestation of the bias often excludes the true parameter space due to
-the use of narrow uniform priors.
-Furthermore we discuss the potential theoretical and practical problems of
-empirical Bayesian model choice.
-% which is used and encouraged by
-% \citet{Hickerson2013} without discussing its potential complications.
-
-%CWL:
-%It may be important to point out the error made in their paper, but it can
-%also be nit-picky, it depends on how the error impacts the overall message of
-%our response.  We should strive to keep the response short and directed at the
-%big picture items. If there is a way to still point out the error, then maybe
-%some of this section should be used
-\section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
-\citet{Hickerson2013} re-analyzed the dataset of \citet{Oaks2012} using a
-model-averaging approach, where they placed a discrete uniform prior over eight
-different prior models (see Table 1 of \citet{Hickerson2013}).
-However, there is a fundamental error in the methodology of
-\citet{Hickerson2013}: their model mixes different time units.
-
-Each of the eight prior models used in the re-analysis by \citet{Hickerson2013}
-has one of two priors on the mean size of the descendant populations of each
-taxon pair:
-$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ or
-$\meanDescendantTheta{} \sim U(0.0005, 0.04)$.
-As described in \citet{Oaks2012}, the divergence time parameters of the model
-implemented in \msb are scaled relative to a constant reference population
-size, \myTheta{C}.
-This reference population size is defined in terms of the upper limit of the
-uniform prior on the mean size of the descendant populations,
-\meanDescendantTheta{}, such that for the prior $\meanDescendantTheta{} \sim
-U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
-$\myTheta{C} = \uniformMax{\meanDescendantTheta{}}/2$.
-Thus, the model used by \citet{Hickerson2013} mixes two different units of
-time.
-In other words, some of their prior and posterior samples are in units of
-$0.05/\mutationRate$ generations, whereas others are in units of
-$0.02/\mutationRate$ generations.
-
-The fact that their posterior samples are in different units makes the results
-of \citet{Hickerson2013} difficult to interpret, and renders their
-regression-adjusted results invalid.
-A fundamental assumption of regression is that all of the values of the
-response variable are in the same units.
-Thus, the results in sections ``Using ABC Model Comparison to Weight
-Alternative Priors for the Philippine Vertebrate Data'' and ``Improved Sampling
-Efficiency by Prior Weighting Supports Asynchronous and Recent Divergence for
-the Philippines Vertebrate Data'' and presented in Figure 2 of
-\citet{Hickerson2013} should be disregarded.
-The error is easily illustrated by re-plotting their results with the different
-time units indicated (Figure~\ref{figJointPosteriorHickerson}).
-
-% \cwlnote{}{This part seems important, but not related to the error in units
-% discussed above}
-Another important characteristic of the approach used by \citet{Hickerson2013}
-that makes their results difficult to interpret is their use of narrow,
-empirically guided uniform priors in an approximate-Bayesian model-choice
-framework.
-We discuss the risks associated with this approach in the next section, and
-show how the results of \citet{Hickerson2013} likely exclude true values of the
-model's parameters.
-
-% In addition to this error, we were not able to reproduce the results of
-% \citet{Hickerson2013}.
-% We followed their methods, generating $5\e6$ prior samples for each of the
-% eight models in their Table 1.
-% We retained the 10,000 samples from each model with the smallest Euclidean
-% distance to the observed summary statistics, and subsequently retained 10,000
-% samples from the remaining 80,000.
-
-\section*{The potential implications of empirical Bayesian model choice}
-\citet{Hickerson2013} repeatedly refer to the prior distributions used by
-\citet{Oaks2012} as ``poorly selected.''
-However, this is misleading, as \citet{Oaks2012} discuss in detail how they
-selected their priors (see their section ``Specifying and simulating the joint
-prior'').
-They mention how the use of uniform prior distributions in \msb for many of the
-model's parameters necessitates awkward broad priors to avoid excluding the
-truth, \textit{a priori}.
-What \citet{Hickerson2013} mean by ``poorly selected'' is that
-\citet{Oaks2012} did not use their data to inform their initial prior
-distributions (i.e., they did not use an empirical Bayesian approach).
-\citet{Oaks2012} did use empirically informed priors when assessing both the
-prior sensitivity of the method and determining the power of \msb under
-real-world conditions with highly informative priors.  
-\citet{Oaks2012} discuss
-the potential dangers of taking an empirical Bayesian approach to model choice
-(see the last paragraph of ``Assessing prior sensitivity of \msb'' in
-\citet{Oaks2012}), and we expand on this here by exploring both the theoretical
-and practical implications of using empirically informed priors for Bayesian
-model choice.
-
-\subsection*{Theoretical implications of empirical priors for Bayesian model
-choice}
-\begin{linenomath}
-Bayesian inference is a rational method of inductive learning in which Bayes'
-rule is used to update our beliefs about a model as new information becomes
-available.
-If we let \allParameterValues represent the set of all possible parameter
-values, we can define a prior distribution for all $\theta \in
-\allParameterValues$ such that $p(\theta)$ describes our belief that any given
-\myTheta{} is the true value of the parameter.
-If we let \allDatasets represent all possible datasets then we can 
-define a sampling model for all $\theta \in
-\allParameterValues$ and $\alignment{}{} \in \allDatasets$ such that
-$p(\alignment{}{} | \theta)$ measures our belief that any dataset \alignment{}{}
-will be generated by any model state \myTheta{}.
-After collecting a new dataset \alignment{i}{}, we can use Bayes' rule to
-calculate the posterior distribution
-\begin{equation}
-    p(\myTheta{} \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given 
-    \myTheta{})p(\myTheta{})}{p(\alignment{i}{})},
-    \label{eq:bayesrule}
-\end{equation}
-where
-\begin{equation}
-    p(\alignment{i}{}) = \int_{\myTheta{}} p(\alignment{i}{} \given
-    \myTheta{})p(\myTheta{}) d\myTheta{}.
-\end{equation}
-The posterior distribution is a measure of our beliefs after seeing the new
-information.
-\end{linenomath}
-
-This is an elegant method of updating our beliefs as data are accumulated.
-However, this all hinges on the fact that the prior ($p(\myTheta{})$) is
-defined for all possible parameter values independently of the new data being
-analyzed.
-Any other datasets or external information can safely be used to inform our
-beliefs about $p(\myTheta{})$.
-However, if the same data are used to both inform the prior and calculate the
-posterior, the prior becomes conditional on the data, and Bayes' rule breaks
-down.
-
-This is not to say that empirical Bayesian approaches are not useful.
-Empirical Bayes is a well studied branch of Bayesian statistics that
-has given rise to many powerful inference methods.
-However, empirical Bayes methods have an uncertain theoretical basis and
-do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
-empirical Bayesian estimates of the posterior are often too narrow, off-center,
-and misshapen;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
-Nevertheless, empirical Bayes methods can provide a powerful means of parameter
-estimation that often exhibit favorable frequentist properties.
-Furthermore, many post-hoc correction methods have been developed for
-estimating confidence-intervals from empirical Bayes estimates of posterior
-distributions that often exhibit well-behaved frequentist coverage
-probabilities
-\citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.
-
-Whereas empirical Bayes approaches can provide powerful methods for parameter
-estimation, a theoretical justification for empirical Bayes approaches to model
-choice is questionable.
-In Bayesian model choice, the goal is not to estimate parameters, but to
-estimate the relative probabilities of candidate models.
-Unlike parameter estimates, posterior probabilities represent a summary of the
-entire posterior distribution.
-Thus, given that empirical Bayesian posterior distributions are not accurate,
-there is no guarantee regarding the accuracy of probabilities summarized from
-them.
-In addition to these theoretical concerns there are practical dangers to using
-narrow, empirical, uniform priors for a method that has already been shown to
-be biased toward models with less parameter space.
-% %CWL:
-% %I like the example, but it may be too cute. It is also too long. I think a
-% %contrived example is good, but if there could be an example that is more
-% %relevant to model choice and comparison of inaccurate posterior distributions,
-% %I think that would be better. This is an example of empirical Bayes, but not
-% %the potential problem with the use in msbayes. Unless I am interpreting wrong.
-% %I would be careful about making it to cute as well. You may intend to make it
-% %easy to understand for a general audience, but it can also seem like talking
-% %down to Hickerson et al, who we are responding to. It just depends on how you
-% %read it.
-% This can be demonstrated with a simple, albeit contrived, example.
-% Let us say that principal investigator Mary and her new postdoc Will are
-% interested in the hypothesis that the mass of George Washington's periwig
-% renders the quarter dollar of the United States unfair.
-% That is to say their null hypothesis is that the probability of a US
-% quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
-% They can certainly evaluate this hypothesis, they can have undergraduate
-% worker, Joe, flip the coin for them while they tabulate the results.
-% But being Bayesians, before they call Joe into the lab, they agree on a
-% prior probability to place on the set of all possible probabilities that
-% the quarter will land heads when it is flipped.
-% Given that neither of them have a quarter, and their prior knowledge that Joe
-% moonlights as a magician, and is notorious for performing coin and card tricks
-% in the lab, they suspect there is a good chance that the quarter Joe uses will
-% be either two-headed or two-tailed.
-% So, knowing that the beta distribution is the conjugate prior for a binomial
-% likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
-% (Figure~\ref{figCoinFlip}).
-
-% Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
-% begin flipping it.
-% After five tosses, four of which was heads, Joe decides that academics are
-% crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
-% being computational biologists, are satisfied with their
-% empirical dataset of $y = 4$ heads out of $N = 5$ trials.
-% They know from the conjugacy of the beta prior, that the posterior distribution
-% has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
-% in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
-% of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
-% This allows them to plug these values into the beta cumulative distribution
-% function to determine that the posterior probability of their hypothesis is
-% $p(\theta < 0.5 | y=1, N=5) = 0.088$.
-% Given their prior belief and dataset, this indeed is the correct posterior
-% probability of the hypothesis, and Mary and Will should now update their
-% posterior belief accordingly.
-
-% However, reflecting upon the results of their experiment
-% (Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
-% Their prior looks very ``poorly selected,'' and if they had only known that the
-% coin was fair before the data were collected, they would have selected a much
-% better prior.
-% Clearly, from their results, they should have used a prior centered
-% around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
-% much ``better.''
-% Rather than resort to flipping the coin five more times, Mary and Will decide
-% to redo their analysis using the much better ``prior'' of $p(\theta) \sim
-% beta(4.5, 1.5)$.
-% This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
-% probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
-% Now convinced that the US quarter is unfair, albeit not due to the mass of
-% President Washington's head as they hypothesized, Will begins composing an
-% e-mail of complaint addressed to the U.S.\ Mint.
-
-% If Joe's flips were a representative sample, Mary and Will's empirical Bayes
-% estimate might very well be a better point estimate for the parameter
-% \myTheta{}.
-% However, their empirical Bayes estimate of the probability of their hypothesis
-% is incorrect and biased.
-% This simple example shows how parameter estimation is fundamentally different
-% from estimating the probability of a model.
-% While empirically informed priors can be used to obtain well-behaved parameter
-% estimators, using them for model choice is much less certain.
-
-\subsection*{Practical concerns about empirically informed uniform priors for
-    Bayesian model choice}
-% \cwlnote{}{Should this be part of the empirical Bayes section, or something
-%     else?}
-
-% \cwlnote{}{It seems like this section can be boiled down to ``empirical Bayes
-%     approaches for model choice when using uniform priors can exclude the
-%     $truth$ from the prior, which is a violation of Bayes theorem" or something
-%     like that. Picking apart what they calculated and we calculated seems like
-%     too much detail. I would try to simplify how this is being presented. This
-%     can lead into how the use of more appropriate prior distributions should be
-%     used.}
-The results of Hickerson et al.'s \citeyear{Hickerson2013} reanalysis of the
-Philippine dataset strongly favored models with the narrowest, empirically
-informed prior on divergence time, and thus their model-averaged posterior
-estimates are dominated by models $M_1$ and $M_2$ (see Table 1 of
-\citet{Hickerson2013}).
-This is concerning, because the narrowest \divt{} prior used by
-\citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
-divergence times for at least some of the Philippine taxa, a major problem when
-using uniform priors.
-\citet{Hickerson2013} set this prior to match the 95\% highest posterior
-density (HPD) interval for the mean divergence time estimated under one of the
-priors used by \citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
-Given this interval estimate is for the \emph{mean} divergence time across all
-22 taxa, it may be inappropriate to set this as the limit on the prior, because
-many of the taxon pairs are expected to have diverged at times older than the
-upper limit.
-Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
-of the mean divergence time under the other two priors explored by
-\citet{Oaks2012} (under these priors the 95\% HPD is approximately 0.3--0.6;
-see Table~6 of \citet{Oaks2012}).
-Thus, the results of \citet{Hickerson2013} indicate that integration over
-parameter space with low likelihood is biasing the method toward models with
-smaller parameter space, and as a consequence is biasing the method toward
-models that exclude the truth (i.e., toward estimating model-averaged
-posteriors dominated by models that exclude true values of parameters of the
-model).
-
-%CWL:This should be in the section on model-averaging, not empirical Bayes
-%limitations
-We explored the propensity of the model-averaging approach of
-\citet{Hickerson2013} to exclude the truth in two ways.
-First, we re-analyzed the Philippines dataset using the model-averaging approach
-of \citet{Hickerson2013}, but set one of prior models with a uniform prior on
-divergence times that is unrealistically narrow, and almost certainly excludes
-most, if not all, of the true divergence times of the 22 taxon pairs.
-If small likelihoods of large models cause the method to prefer models with
-less parameter space, we expect \msb will preferentially sample from this
-incorrect model yielding a marginal posterior that is incorrect (i.e., the
-model-averaged posterior will be dominated by an incorrect model that excludes
-the truth).
-Second, we generated simulated datasets for which the divergence times are
-drawn from an exponential distribution and applied the approach of
-\citet{Hickerson2013} to each of them to see how often the method excludes the
-truth.
-
-For our re-analysis of the Philippines dataset we used the model-averaging
-approach of \citet{Hickerson2013}, but with a reduced set of prior models to
-avoid mixing time units.
-We used five prior models, all of which had priors on population sizes of
-$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
-U(0.0001, 0.05)$.
-Following \citet{Hickerson2013}, each of these models had the following
-priors on divergence time parameters:
-$M_1$, $\divt{} \sim U(0, 0.1)$;
-$M_2$, $\divt{} \sim U(0, 1)$;
-$M_3$, $\divt{} \sim U(0, 5)$;
-$M_4$, $\divt{} \sim U(0, 10)$; and
-$M_5$, $\divt{} \sim U(0, 20)$.
-We simulated $1\e6$ random samples from each of the models for a total of
-$5\e6$ prior samples.
-For each model, we retained the 10,000 samples with the smallest Euclidean
-distance from the observed summary statistics, standardizing the statistics
-using the prior means and standard deviations of the given model.
-From the remaining 50,000 samples, we then retained the 10,000 samples with the
-smallest Euclidean distance from the observed summary statistics, this time
-standardizing the statistics using the prior means and standard deviations
-across all five models.
-We then repeated this analysis twice, replacing the $M_1$ with
-$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
-times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
-respectively.
-While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
-\citet{Hickerson2013} likely excludes the true divergence times of at least
-some of the 22 taxa, we are nearly certain that these narrower priors are
-incorrect and exclude most, if not all, of the divergence times of the
-Philippine taxa.
-
-Our results show that the model-averaging approach of \citet{Hickerson2013}
-does not reduce the method's bias toward models with less parameter space,
-but rather allows it to manifest in a rather dangerous way.
-The method strongly prefers the prior model with the narrowest distribution on
-divergence times across all three of our analyses, even when this model is
-almost certainly incorrect and excludes the true divergence times of the
-Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
-Unfortunately, ``checking'' the priors by plotting the summary statistics from
-1000 random samples of each prior model along the first two orthogonal axes of
-a principle component analysis, as recommended by \citet{Hickerson2013},
-provides no warning of a problem (Figure~\ref{figPCA}).
-Given that the results of \citet{Hickerson2013} strongly prefer the models with
-the narrowest prior on divergence times, it seems quite likely that their
-model-averaged results are dominated by models that exclude at least some of
-the 22 true divergence times, making their results difficult to interpret
-(in addition to the error of time units).
-
-To better quantify the propensity of Hickerson et al.'s
-(\citeyear{Hickerson2013}) approach to exclude the truth, we simulated 1000
-datasets in which the divergence times for the 22 population pairs are drawn
-randomly from an exponential distribution with a mean of 0.5 ($\divt{} \sim
-Exp(2)$).
-All other parameters were identically distributed as the $M_1$--$M_5$ models
-(Table~\ref{tabModelChoiceEmpirical}).
-We then repeated the analysis described above using $1\e6$ random samples from
-prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, retaining 1000 posterior
-samples for each of the 1000 simulated datasets.
-
-% \highLight{Add how we determined exclusion.}
-For each simulation replicate, we estimated the Bayes factor in favor
-of excluding the truth as the ratio of the posterior to prior odds of
-excluding the true value of at least one parameter.
-Whenever the Bayes factor preferred a model excluding the truth we counted
-the number of 22 true divergence times that were excluded by the preferred
-model.
-Our results show that the model-averaging approach of \citet{Hickerson2013}
-excludes the true values of parameters in 97\% of the replicates (90\% with
-GLM-regression adjustment), excluding up to 21 of the 22 true divergence times
-(Figure~\ref{figExclusionSimTau}).
-We also used the mode estimate of the preferred model for each replicate to
-estimate the number of true values excluded, which produced very similar
-results.
-Furthermore, the posterior probability of excluding at least one true parameter
-value is very high in nearly all of the replicates
-(Figure~\ref{figExclusionSimProb}).
-Using a Bayes factor of greater than 10 as a criterion for strong support, 66\%
-of the replicates (87\% with GLM-regression adjustment) strongly support the
-exclusion of true values (Figure~\ref{figExclusionSimProb}).
-
-The results of our empirical and simulation-based analyses clearly demonstrate
-the danger of using narrow, empirically guided uniform priors in a Bayesian
-model-averaging framework.
-This is especially concerning in \msb, which has already been demonstrated to
-be biased away from models with more parameter space \citep{Oaks2012}.
-The consequence of this approach is a high risk of obtaining a model-averaged
-posterior estimate that is heavily weighted toward incorrect models that
-exclude true values of model parameters.
-
-\subsection*{Additional thoughts on empirical priors in Bayesian model choice}
-Given all of the theoretical and practical issues with empirical Bayes
-approaches to Bayesian model choice discussed in the proceeding sections, it is
-quite clear why one should use caution before overly criticizing an
-investigator's choice of priors after having seen the resulting posterior.
-%In a strictly Bayesian world, this is cheating.
-As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a large
-amount of uncertainty regarding the divergence times of the 22 population pairs
-under study.
-Two of these pairs represent distinct species, and the taxonomy of many groups
-in the Philippines has repeatedly been shown to mask deeply divergent lineages
-\citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
-    Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013}.
-\citet{Oaks2012} also discuss how the sole use of uniform priors in \msb makes
-it very difficult for investigators to express their prior uncertainty without
-putting a lot of prior density in regions of improbable parameter space.
-The alternative, as shown above, is to risk excluding the truth before
-performing the analysis.
-
-While our results above clearly demonstrate the risks inherent to the empirical
-Bayesian model-choice approach used by \citet{Hickerson2013}, one could justify
-such risk if the method does indeed increase the power of the method and thus
-decrease bias toward clustered divergence models, as claimed by
-\citet{Hickerson2013}.
-In the next section, we use simulation-based analyses to evaluate the power of
-Hickerson et al.'s (\citeyear{Hickerson2013}) approach, which show the method is
-still biased toward highly clustered models of divergence.
-
-\section*{Assessing the power of the model-averaging approach of
-    \citet{Hickerson2013}}
-% \cwlnote{}{I think the empirical use of the model-averaging should go here,
-%     instead of above in the empirical Bayes section}
-
-As recommended by \citet{Oaks2012}, we perform simulation-based analyses
-to explore the power of the approximate Bayesian model-averaging approach
-proposed by \citet{Hickerson2013}.
-Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
-the 22 population pairs randomly drawn from a uniform distribution, $U(0,
-\divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and 2.0, in
-\globalcoalunit generations.
-All other parameters were identically distributed as the prior.
-We use the same $5\e6$ samples from the same set of five prior models as above
-($M_1$, $M_2$, $M_3$, $M_4$, and $M_5$).
-As described above, for each simulated dataset, we approximated the posterior
-by retaining 1000 samples from the prior with the smallest Euclidean distance
-from the true summary statistics.
-In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
-posterior samples for each of them.
-
-We find that the approach of \citet{Hickerson2013} struggles to estimate the
-dispersion index of divergence times (\vmratio{}) across most of the \divt{max}
-we simulated, whether evaluating the unadjusted
-(Figure~\ref{figPowerAccOmegaMedian}) or GLM-adjusted
-(Figure~\ref{figPowerAccOmegaModeGLM}) posterior estimates.
-The method only estimates \vmratio{} relatively well when the simulated
-distribution of divergence times is identical to one of the prior models
-(Figures~\ref{figPowerAccOmegaMedian}E \& \ref{figPowerAccOmegaModeGLM}E).
-This is consistent with the conclusion of \citet{Oaks2012} that \msb lacks
-robustness and is highly sensitive to the prior distribution deviating from the
-true, underlying distribution of the data.
-
-Furthermore, our results demonstrate that the approach of \citet{Hickerson2013}
-consistently infers highly clustered divergences across all the \divt{max} we
-simulated (Figure~\ref{figPowerNumExcluded}).
-The method is most likely to infer the extreme case of a single divergence event
-when populations diverged randomly over the past \globalcoalunit generations.
-Even when divergences are random over the past $8\globalpopsize$ generations,
-the most likely inference is only two divergence events, and a single 
-divergence is still estimated in more than 10\% of the replicates.
-It is very interesting to note that as \divt{max} increases, but before the
-estimates are finally pulled away from $\numt{} = 1$, the distribution of
-\numt{} estimates closely mirror the odd U-shaped prior on divergence models
-used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
-(\citeyear{Oaks2012}) Figure 5B).
-This supports the conclusions of \citet{Oaks2012} that this U-shaped
-prior coupled with the poor marginal likelihoods of models with many
-\divt{} parameters, is a major cause of the method's bias toward
-clustered divergence models.
-
-Looking at our simulation results in terms of the posterior probability of the
-dispersion index of divergence times supporting the extreme case of one
-divergence event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the
-method strongly supports one divergence in greater than 27\% of the replicates
-across all the \divt{max} we simulated (Figure~\ref{figPowerOmegaProb}).
-Following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
-the criterion for incorrect inference of a single divergence event.
-There is strong support for a single divergence event in more than 90\% of the
-replicates when divergences are random over the past $2.4\globalpopsize$
-generations, and more than 60\% when over the past $3.2\globalpopsize$
-generations or less (Figure~\ref{figPowerOmegaProb}).
-
-Contrary to Hickerson et al.'s (\citeyear{Hickerson2013}) claim that their
-model-averaging approach ``can discriminate complex multispecies histories and
-correctly reject synchronous divergence, even when discrete divergence times
-differ by much less than \ldots millions of generations,'' our results show
-that method is biased toward supporting the extreme case of a single
-divergence event when populations diverged randomly over the last
-$8\globalpopsize$ generations.
-To put this on the scale roughly consistent with a vertebrate mitochondrial
-locus, assuming a mutation rate of $2\e{-8}$ per site per generation, this
-translates to 5 million generations.
-Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
-million generations.
-
-The results of our power analyses further demonstrate the propensity of
-Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
-parameter values.
-Across all but one of the \divt{max} we simulated, the method excludes the
-truth in a large proportion of replicates, and across many of the \divt{max} it
-will exclude a large proportion of the true divergence time values
-(Figure~\ref{figPowerNumExcluded}).
-The posterior probability of excluding at least on true divergence value is
-also quite high across many of the \divt{max}
-(Figure~\ref{figPowerProbExclusion}).
-Only when the data are identically distributed as one of the prior models does
-the method avoid excluding the truth more than 5\% of the time
-(Figure~\ref{figPowerNumExcluded}E).
-Again, this demonstrates the method's lack of robustness.
-
-Given our results, we conclude that the approach of \citet{Hickerson2013}
-excludes regions of parameter space containing the truth, and lacks power to
-detect random variation in divergence times over a scale of $8\globalpopsize$
-generations or more.
-This roughly translates to millions of generations for mitochondrial loci, and
-hundreds of millions of generations for nuclear loci.
-
-%CWL: I would suggest removing this section
-\section*{The power analysis of \citet{Hickerson2013}}
-\citet{Hickerson2013} present a power analysis in which they find that
-\msb has the power to infer multiple divergence events when divergence times
-are random over hundreds of thousands of generations (rather than millions as
-demonstrated by \citet{Oaks2012}).
-It is not surprising that under limited parameter space the method has
-increased power to detect temporal variation in divergences over narrower time
-windows than millions of generations.
-However, what is important to consider is whether those conditions are relevant
-to real-world applications of the method.
-\citet{Oaks2012} explore the behavior of the method under three divergence time
-priors, as narrow as 0--5 coalescent units. This is quite narrow considering
-that this expresses the prior belief that all 22 taxon pairs diverged within
-this window.
-\citet{Hickerson2013} limit their power analysis to a single prior of 0--1
-coalescent unit.
-This prior setting assumes that there is enough a priori information to
-be 100\% certain that all taxa (18 for their simulations) diverged within
-the last 4$\globalpopsize$ generations.
-This does not seem applicable to most empirical systems.
-Furthermore, even when such extensive prior information is available, to only
-be able to detect multiple divergences on the scale of hundreds of thousands of
-generations does not seem very powerful.
-
-Also, \citet{Hickerson2013} present information only on the extreme case of a
-single divergence event.
-As discussed in \citet{Oaks2012}, it seems odd to consider the estimation
-of two divergence events as a success when divergence is random over hundreds
-of thousands (or millions) of generations.
-In an empirical system such as the Philippines, where island fragmentation has
-occurred at least 10 times over the past several millions of years
-\citep{Haq1987,Rohling1998,Siddall2003,Miller2005}, an estimate where 22 taxa
-share even a handful of divergence events would be of biogeographic interest.
-
-Lastly, \citet{Hickerson2013} translate their results from units of
-\globalcoalunit generations to generations assuming a mutation rate of
-$1.92\e{-8}$ mutations per site per generation.
-If we translate their results to a scale more consistent with rates of mutation
-of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
-by \citet{Hickerson2013} the method can only reliably reject the extreme case
-of one divergence event when divergences are random over a period of more than
-3 million generations.
-Again, considering \citet{Hickerson2013} are assuming prior knowledge that
-all population pairs diverged within the last coalescent unit, this
-does not seem particularly powerful.
-
-\section*{Model likelihoods or insufficient prior sampling as the cause of bias
-in \msb}
-One of the points of disagreement between \citet{Hickerson2013} and
-\citet{Oaks2012} is the underlying mechanism causing the bias toward models
-with clustered divergences (i.e., models with few divergence events).
-\citet{Oaks2012} present simulation-based analyses that suggest the broad
-uniform prior on divergence time parameters decrease the marginal likelihoods
-of models with more divergence time parameters, whereas \citet{Hickerson2013}
-argue that the bias is inherent to the inefficient rejection algorithm
-implemented in \msb.
-
-\citet{Hickerson2013} present an interesting probabilistic argument
- to show that insufficient prior sampling is to blame for the bias.
-They argue the widest of the three priors used by \citet{Oaks2012} was very
-unlikely to produce any prior samples with large numbers of divergence times
-that were consistent with the Philippine data.
-There are a few issues with their argument.
-First, the probabilities they present assume the gene divergence times
-estimated by \citet{Oaks2012} are correct.
-The sole purpose of the estimates of ultrametric gene trees presented by
-\citet{Oaks2012} was to provide a very rough comparison of the gene divergence
-times across the 22 taxa.  These analyses assumed an arbitrary strict clock of
-$2\e{-8}$ for all taxa, and are, of course, subject to estimation error.
-Furthermore, the branch-length units of the gene trees are in millions of
-years, whereas the divergence time prior of \msb is in generations, thus the
-logic of \citet{Hickerson2013} requires the additional assumption that all 22
-Philippine taxa have a generation time of one year.
-Thus, the argument of \citet{Hickerson2013} that divergence time estimates of
-\citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
-questionable, especially given our findings on the behavior of empirically
-informed priors presented above.
-
-Finally, even \emph{if} we make all of these assumptions and assume the gene
-divergence times are estimated without error, the probabilistic argument only
-applies to one of the three different priors used by \citet{Oaks2012}.
-The narrowest prior on divergence times used by \citet{Oaks2012} closely mirrors
-the range of estimates of gene-divergence times, and applying Hickerson et al.'s
-(\citeyear{Hickerson2013}) probability equations demonstrates that the prior
-is densely populated with samples with large numbers of divergence parameters
-that are consistent with the gene divergence estimates.
-Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
-sampling is to blame for the bias, it should be much reduced under the narrow
-prior on \divt{}.
-However, the magnitude of the bias is very similar across all three priors
-explored by \citet{Oaks2012}.
-\citet{Hickerson2013} point out a case where the narrow prior performs
-slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
-However, it is important to note that these results suffered from a bug
-in \msb, and there are many cases after \citet{Oaks2012} corrected the 
-bug where the narrow prior performs slightly worse (see panels D--J of
-Figures 3 and S12).
-
-To disentangle whether model likelihoods or insufficient prior sampling is to
-blame for the biases revealed by \citet{Oaks2012}, we must look at the
-different predictions made by these two phenomena.
-One example, as discussed by \citet{Oaks2012}, is that insufficient prior
-sampling should create higher variance in posterior estimates, and thus it
-should cause analyses to be sensitive to the number of samples drawn from the
-prior.
-\citet{Oaks2012} do not see such sensitivity when they compare prior sample
-sizes of $2\e6$, $5\e6$, and $10^7$.
-
-To explore this prediction further, we repeat the analysis of the Philippines
-dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim U(0,
-10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
-(0.0005, 0.02)$), using a very large prior sample size of $10^8$.
-When we look at the trace of the estimates of the dispersion index of
-divergence times (\vmratio{}) as the prior samples accumulate
-(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or
-GLM-regression-adjusted estimates.
-This suggests that insufficient prior sampling did not play a large
-role in the bias found by \citet{Oaks2012}.
-
-Our results presented above that demonstrate the bias of the model-averaging
-approach of \citet{Hickerson2013} both toward models with narrower \divt{}
-priors (Table~\ref{tabModelChoiceEmpirical} and
-Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
-and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
-\ref{figPowerOmegaProb}) strongly suggest the primary cause is the uniform
-priors reducing the marginal likelihoods of large models.
-In all of our model-averaging analyses, all prior models have the same number
-of samples.
-Thus, it is difficult to explain the method's propensity toward the model with
-the smallest parameter space via insufficient prior sampling.
-While analyses that sample each model proportional to their relative parameter
-space could be explored, it seems much more likely that the broad uniform
-priors are simply inhibiting the marginal likelihoods of spacious models.
-
-As discussed by \citet{Oaks2012}, a straightforward prediction of reduced
-likelihoods of complex models is that they should disappear as the model
-generating the data converges to the prior.
-\citet{Oaks2012} test this prediction by performing 100,000 simulations to
-assess the model-choice behavior of \msb when the prior model is correct.
-They find that the bias actually switches direction (at least for the
-regression-adjusted estimates) and the method tends to underestimate the
-probability of the model with one divergence event (see Figure 4 of
-\citet{Oaks2012}).
-The same prediction is not as straightforward for the insufficient-sampling
-hypothesis.
-Even when the prior is correct, due to the discrete uniform prior on
-\numt{} implemented in \msb, models with larger numbers of divergence
-events (and thus greater parameter space) will still be less densely
-sampled than those with fewer divergence events \citep{Oaks2012}.
-Thus, the results of the simulations of \citet{Oaks2012} are more consistent
-with marginal model likelihoods causing the bias toward models with fewer
-divergence time parameters.
-
-Taken together, the results presented here and in \citet{Oaks2012} support
-reduced marginal likelihoods of large models as a primary mechanism by which
-broad uniform priors cause biases in \msb.
-Nonetheless, posterior sampling error will always be present in any numerical
-Bayesian approximation method.
-Thus, insufficient sampling of the prior will contribute to the error of all
-approximate Bayesian estimates.
-However, there is no strong evidence that it is playing a large role in
-the biases revealed herein and by \citet{Oaks2012}.
-
-%CWL: This seems to specific, not sure about including
-\section*{Validation analyses}
-% \highLight{Not sure where to put this section}.
-% \highLight{This justifies presenting the unadjusted results in this paper.}
-% \highLight{Is it worth including just for this purpose?}
-Following \citet{Oaks2012}, we characterize the model-choice behavior of the
-model-averaging approach of \citet{Hickerson2013} under the ideal conditions
-where the prior is correct (i.e., the data are generated from parameters drawn
-from the same prior distributions used in the analysis).
-We used the same prior models as above ($M_1$--$M_5$;
-Table~\ref{tabModelChoiceEmpirical}), and generated 50,000 pseudo-replicate
-datasets under this prior (10,000 from each model).
-We used a simulated data structure of eight population pairs, with a single
-1000 base-pair locus sampled from 10 individuals from each population.
-We then analyzed each of these replicate datasets using the same prior,
-retaining 1000 posterior samples.
-Our results are very similar to \citet{Oaks2012}, but we note that they
-are not directly comparable as our simulations contained eight population
-pairs rather than 10 (Figure~\ref{figValidationMCBehavior}).
-We find that the approach of \citet{Hickerson2013} estimates the posterior
-probability of divergence models reasonably well when all assumptions of the
-method are met (i.e., the prior is correct) and the unadjusted posterior
-estimates are used.
-Similar to \cite{Oaks2012}, we find that the regression-adjusted estimates of
-the model probabilities are biased.
-
-\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
-The main parameter of the \msb model is the vector of divergence
-times for each of the taxon pairs,
-$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
-\citep{Oaks2012}.
-\citet{Hickerson2013} argue that the dispersion index of this vector,
-\vmratio{}, is a better model-choice estimator than the number of 
-divergence time parameters within the vector,
-\numt{}.
-They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
-\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
-This plot shows, not surprisingly, that \vmratio{} has very little information
-about the number of divergences among taxa.
-Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
-biogeographically relevant than \numt{}.
-% We struggle to follow this logic.
-Certainly the maximum information is contained within the divergence time
-vector that \vmratio{} is summarizing.
-Clearly, the number of divergence time parameters within the vector and their
-values is more informative than its variance (i.e., the dispersion index is not
-a sufficient statistic for \divtvector).
-
-\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
-better than \numt{}.''
-However, \msb is a model-choice estimator, and hence the goal is to estimate the
-posterior probability of divergence models.
-\citet{Oaks2012} demonstrate that even when all assumptions of the model are
-met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
-4), whereas \numt{} performs better.
-Furthermore, \vmratio{} is limited to estimating the probability of only a
-single divergence model (the one divergence model), and thus its utility for
-model choice is extremely limited.
-
-The model-choice utility of \vmratio{} is limited to the probability that
-this continuous statistic, which can range from zero to infinity, is at its lower
-limit of zero. In theory, this point density will always be zero, thus an
-arbitrary threshold (0.01 is used throughout the \msb literature) must 
-be chosen to make the probability estimable.
-However, it is still not surprising to see that it is numerically difficult to
-obtain reliable estimates of the probability that the continuous \vmratio{}
-statistic is ``near'' its limit of zero.
-It is much easier, less subjective, and more interpretable to estimate
-the probability of the discrete parameter of the model, \numt{}, is
-at its lower limit of one.
-Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better
-estimator of model probability than \vmratio{}.
-
-\section*{Some general thoughts on the model of \msb}
-Our results demonstrating the poor behavior of the model-averaging approach
-proposed by \citet{Hickerson2013} are not too surprising.
-While this approach is trivial to implement, it is not a trivial change
-to the \msb model.
-To better understand the reason for this, it might help to step back and get a
-sense of the scale of the \msb model.
-To do this, we will use the dataset of 22 vertebrate taxon pairs of
-\citet{Oaks2012} as an example.
-
-Following the model description and notation of \citet{Oaks2012}, let us tally
-up all of the free parameters in the \msb model.  Under the simplest model in
-\msb (i.e., assuming no migration and
-no intra-locus recombination), the number of parameters for each
-taxon pair include:
-The population sizes of the ancestral and descendant populations
-(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
-the magnitude of population contraction in each of the descendant
-populations (\bottleScalar{1}{} and \bottleScalar{2}{}) and the
-timing of these contractions (\bottleTime{}), and the $N-1$ node heights
-(coalescent times) of the gene tree that gave rise to the $N$ gene
-copies sampled from both populations of the pair.
-Because we only have a single locus for each taxon, there are no locus-specific
-\myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
-\locusRateHetShapeParameter shape parameter for the gamma prior distribution on
-\locusMutationRateScalar{}.
-Lastly, there are between one and 22 divergence time parameters \divt{} in
-the vector \divtvector.
-Overall, for our Philippines data under the simplest model in \msb, there are
-581--602 free parameters (depending on the cardinality of \divtvector).
-Furthermore, under this rich model, the method is sampling over 1002 divergence
-models \citep[i.e., the number of integer partitions of $Y=22$;][]{Oaks2012}.
-
-This is a very difficult inference problem, and the method only uses four
-summary statistics calculated from the sequence alignment of each taxon pair:
-$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
-\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
-That gives us a total of 88 summary statistics (four from each of the 22 taxon
-pairs), which contain minimal information about many of the $\approx 600$
-parameters in the model.
-More summary statistics can be used in \msb, but most are highly correlated
-with these four statistics (which are even highly correlated among themselves),
-and thus contribute little additional information about the data.
-
-The large number of parameters and divergence models relative to 
-summary statistics is undoubtedly a key reason the method is so sensitive
-to the prior distributions.
-It also likely contributes to the method's lack of robustness.
-% We use robustness here as a measure of a method's insensitivity to model
-% violations.
-Robustness is an extremely important characteristic of a method to gauge its
-applicability to real-world data, because we know the model and priors will be
-wrong to some degree.
-
-The approach of \citet{Hickerson2013} adds an additional dimension of model
-choice to the model. They expand the model to sample over eight prior models.
-This extends the original model to having 582-603 free parameters and, more
-importantly, sampling over 8016 unique model states across both divergence
-models and prior models.
-This is a non-trivial extension of the model and, given the method's very large
-number of parameters and divergence models relative to the information used
-from the data, likely plays a major role in the poor behavior of this approach.
-% Certainly, such a large modification of the model demands simulation-based
-% assessments of the behavior of the new model, as we provide here.
-
-In theory, the approach of \citet{Hickerson2013} is very appealing.  It sums
-over multiple candidate prior models to produce a posterior estimate
-marginalized over the uncertainty in prior choice.
-In general, Bayesian model-averaging is a powerful approach that leverages a
-great strength of Bayesian statistical procedures, namely the ability to
-obtain marginalized estimates that incorporate uncertainty in nuisance
-parameters.
-However, given the basic \msb model is already struggling to estimate
-a huge number of parameters and model probabilities with scant information
-from the data, it is not surprising that adding an additional layer of
-model choice to the method causes problems.
-
-The recommendations of \citet{Oaks2012} for mitigating the bias and lack of
-robustness of \msb are actually similar to those of \citet{Hickerson2013}, but
-avoid the need for imposing additional model choice.
-\citet{Oaks2012} suggest that uniform priors are inappropriate for many
-parameters of the \msb model, and recommend the use of probability
-distributions from the exponential family.
-If we look at the prior distribution on the divergence time parameter \divt{}
-imposed by the model-averaging approach of \citet{Hickerson2013} we see
-it is a mixture of overlapping uniforms with lower limits of zero
-(Figure~\ref{figMCTauPrior}).
-This looks very much like an exponential distribution, except that each sample
-of divergence times are restricted to the hard bounds of one of the eight prior
-models.
-Thus, it seems more appropriate to place a gamma prior (the exponential being a
-special case of the gamma) on divergence times.
-This would provide investigators flexibility to represent their prior
-uncertainty in model parameters while avoiding broad uniform
-distributions and the need of costly model-averaging.
-This seems like a natural solution to the problem of skewed marginal
-likelihoods and/or sampling error, which are both likely exacerbated by the
-fact that uniform distributions are the only choice of prior for many of the
-parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
-\descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
-\bottleScalar{1}{}, \bottleScalar{2}{}, \locusRateHetShapeParameter,
-\migrationRate{}, \recombinationRate;][]{Oaks2012}.
-
-Importantly, there is no cost to this strategy.
-Given that ABC methods only need to draw random values from prior distributions
-(hence there are no difficult prior and proposal ratios to calculate, etc.) it
-is easy to use the most appropriate distributions on parameters.
-Also, this approach reduces the temptation of empirically guided priors.
-The investigator can place the majority of the prior density in regions of
-parameter space they believe, \emph{a priori}, are most plausible, but still
-capture uncertainty in the tails of distributions with low density.
-As discussed in \citet{Oaks2012} the use of uniform priors necessitate
-broad priors to avoid excluding the truth \emph{a priori}.
-Thus, to avoid the behavior of the method under these uniform priors, using the
-data is a very tempting, albeit risky, option \citep{Hickerson2013}.
-More appropriate prior distributions would alleviate this issue.
-
-%CWL: This seems to nit-picky. Probably not worth including it
-\section*{Other clarifications}
-
-\subsection*{Graphical checks of priors}
-\citet{Hickerson2013} advocate the use of what they call graphical checks of
-prior models.
-This entails generating a small number of random samples from the prior (1000)
-and plotting the resulting summary statistics in comparison to the observed
-statistics to see if they coincide (see Figure 1 of \citet{Hickerson2013}).
-As we show above, this is a risky strategy, because the resulting plots of this
-approach have little correlation with the appropriateness of priors.
-Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
-dataset analyzed by \citet{Hickerson2013}), we do not expect that 1000
-\emph{random} draws from the vast prior parameter space will yield data and
-summary statistics consistent with the observed data.
-In fact, when the random draws are tightly clustered around the observed
-statistics, this can be an indication that the prior is over-fit, as we show
-above (Table~\ref{tabModelChoiceEmpirical} and Figure~\ref{figPCA}).
-Thus, interpreting such plots should be avoided, and the use of posterior
-predictive analyses would be much more informative about the overall fit of
-models.
-
-\subsection*{The validity of \msb estimates}
-\citet{Hickerson2013} claim that \citet{Oaks2012} ``assume that all previous
-\msb results are invalid.''
-With the exception of \citet{Hickerson2013}, we do not claim any previous
-results of \msb are invalid.
-Rather, we conservatively recommend that the common inference of temporally
-clustered divergences
-\citep{Barber2010, Bell2012, Carnaval2009, Chan2011, Daza2010, Hickerson2006,
-    Huang2011, Lawson2010, Leache2007, Plouviez2009, Stone2012, Voje2009},
-when not accompanied with the necessary simulation-based analyses to guide the
-interpretation of such results, should be treated with caution, because the
-method has been shown to spuriously infer clustered divergences over a range of
-prior conditions.
-
-\subsection*{Saturation of summary statistics}
-\citet{Hickerson2013} claim that our priors ``cause much of the explored
-parameter space to be beyond the threshold of saturation in most mtDNA genes.''
-To explore this possibility, we simulated datasets under prior settings that
-match two of the three priors used by \citet{Oaks2012}: $\meanDescendantTheta{}
-\sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
-Under this prior, we draw divergence time parameters from a uniform
-distribution of $U(0, 20)$, simulate datasets, and plot the \divt{} values
-against the summary statistics calculated from the resulting datasets
-(Figure~\ref{figSaturationPlot}).
-Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
-and 10 suffered from little to no effect from saturation.
-Even at divergence times of 20 coalescent units, there is still signal in the
-summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
-Thus, the assertion of \citet{Hickerson2013} does not apply to at least
-two of the priors used by \citet{Oaks2012} and, as a result, does not
-explain the bias they found.
-
-\section*{Conclusions}
-We demonstrate how the approximate Bayesian model choice method implemented in
-\msb can be strongly biased away from models with greater parameter space.
-As suggested by \citet{Oaks2012}, this is likely caused by the inappropriate
-use of uniform priors on most of the model's parameters.
-These priors necessitate the use of broad priors that place high prior density
-in unlikely regions of parameter space, less the risk of excluding the truth
-\emph{a priori}.
-This likely reduces the marginal likelihoods of models with greater parameter
-space, either due to more divergence time parameters or broader prior
-distributions on those parameters.
-We show that the empirical Bayesian model-averaging approach of
-\citet{Hickerson2013} does not mitigate this bias, but rather causes it to
-manifest by sampling predominantly from models that may exclude the true values
-of the parameters.
-Exploring alternative prior probability distributions on most of the model's
-parameters, in addition to different priors over divergence models, should
-help mitigate the method's biases.
-
-The work presented herein follows the principles of Open Notebook Science.
-All aspects of the work were recorded in real-time via version-control
-software and are publicly available at
-\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
-All information necessary to reproduce our results are provided there.
+\input{reply-main.tex}
 
@@
 
-%:TABLE-re-analysis
-% \begin{table}[htbp]
-%     \sffamily
-%     \footnotesize
-%     \addtolength{\tabcolsep}{-0.08cm}
-%     \rowcolors{2}{}{myGray}
-%     %\captionsetup{font=footnotesize}
-%     \caption{Results of the model-averaging approach of \citet{Hickerson2013}
-%         applied to the Philippines dataset of \citet{Oaks2012} using eight
-%         prior models (see \citet{Hickerson2013} Table 1).  The results of (1)
-%         \citet{Hickerson2013}, (2) our full re-analysis, and of (3) repeating
-%         the final rejection step of \citet{Hickerson2013} using the prior means and
-%         standard deviations to normalize the summary statistics, are all compared.}
-%     \centering
-%     \begin{tabular}{ l l l l }
-%         \toprule
-%          & Hickerson et al. (2013) & Re-analysis & blah \\
-%         $\hat{\numt{}}$ & 2 & 1 & 1 \\
-%         $p(\numt{} = 1 | \ssSpace)$ & 0.062 & 0.206 & 0.146 \\
-%         $BF_{\numt{} = 1, \numt{} \neq 1}$ 1.39 & 5.45 & 3.59 \\
-%         $p(\vmratio{} < 0.01 | \ssSpace)$ & 0.298 & 0.416 & 0.361 \\
-%         $BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01}$ 6.65 & 11.16 & 8.85 \\
-%         $\hat{M}$ & 1 & 2 & 2 \\
-%         $p(M_1 | \ssSpace)$ & 0.513 & 0.358 & 0.376 \\
-%         $p(M_2 | \ssSpace)$ & 0.286 & 0.369 & 0.389 \\
-%         $p(M_3 | \ssSpace)$ & 0.036 & 0.086 & 0.089 \\
-%         $p(M_4 | \ssSpace)$ & 0.002 & 0.012 & 0.011 \\
-%         $p(M_5 | \ssSpace)$ & 0.142 & 0.129 & 0.107 \\
-%         $p(M_6 | \ssSpace)$ & 0.020 & 0.043 & 0.024 \\
-%         $p(M_7 | \ssSpace)$ & 0.001 & 0.003 & 0.004 \\
-%         $p(M_8 | \ssSpace)$ & 0.000 & 0.000 & 0.001 \\
-%         \bottomrule
-%     \end{tabular}
-%     \label{tabModelChoiceEmpirical}
-% \end{table}
-
-%:TABLE-model-choice-empirical
-\begin{table}[htbp]
-    \sffamily
-    % \footnotesize
-    \addtolength{\tabcolsep}{-0.08cm}
-    \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
-    \caption{Results of the model-averaging approach of \citet{Hickerson2013}
-        applied to the Philippines dataset of \citet{Oaks2012} using three sets
-        of prior models. All models used priors on population size of
-        $\meanDescendantTheta{} \sim U(0.0001,0.1)$ and $\ancestralTheta{} \sim
-        U(0.0001, 0.05)$, and differ only in their prior on divergence time
-        (\divt{}) parameters.  Each set of five models differ only in the
-        divergence time prior used for the model with the narrowest prior:
-        $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_{1A}$ ($\divt{} \sim U(0, 0.01)$),
-        or $M_{1B}$ ($\divt{} \sim U(0, 0.001)$). The approximate posterior
-        probability of each model ($p(M_i \given \ssSpace)$) is given for each
-        of the three analyses.  The posterior estimates are based on 10,000
-        samples retained from $1\e6$ prior samples
-    from each model.}
-    \centering
-    \begin{tabular}{ l l l l l }
-        \toprule
-        & & \multicolumn{3}{c}{$p(M_i \given \ssSpace)$} \\
-        \cmidrule(){3-5}
-        Model & \divt{} prior & $M_{*}=M_1$ & $M_{*}=M_{1A}$ & $M_{*}=M_{1B}$ \\
-        \midrule
-        $M_*$ & --        & 0.899 & 0.821 & 0.673 \\
-        $M_2$ & $U(0,1)$  & 0.079 & 0.136 & 0.251 \\
-        $M_3$ & $U(0,5)$  & 0.013 & 0.026 & 0.044 \\
-        $M_4$ & $U(0,10)$ & 0.006 & 0.012 & 0.022 \\
-        $M_5$ & $U(0,20)$ & 0.003 & 0.005 & 0.010 \\
-        \bottomrule
-    \end{tabular}
-    \label{tabModelChoiceEmpirical}
-\end{table}
+\input{reply-tables.tex}
 
@@
 
-\mFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
-    The joint posterior of the mean (\meant{}{}) and dispersion index ($\vmratio{} = 
-    \vart{}{}/\meant{}{}$) of divergence times for 22 vertebrate taxon pairs as
-    estimated by \citet{Hickerson2013} (see Figure 2B of \citet{Hickerson2013}).
-    The posterior samples are color-coded to indicate the erroneous mixture of
-    timescales in the analysis of \citet{Hickerson2013};
-    grey = $0.05/\mutationRate$ generations and
-    black = $0.02/\mutationRate$ generations.
-}{figJointPosteriorHickerson}
-
-% \mFigure{../images/coin_flip.pdf}{
-%     A plot of three beta probability density functions that represent a prior
-%     (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
-%     empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
-%     Bernoulli trials, four of which are successes.
-% }{figCoinFlip}
-
-\mFigure{../../model-choice/priors-for-pc-plot/pc-plots.pdf}{
-    The graphical checks recommended by \citet{Hickerson2013} for three prior
-    models: (A) $M_1$ ($\divt{} \sim U(0, 0.1)$), (B) $M_{1A}$ ($\divt{} \sim
-    U(0, 0.01)$), and (C) $M_{1B}$ ($\divt{} \sim U(0, 0.001)$).
-    The plots project the summary statistics from 1000 random samples from each
-    model onto the first two orthogonal axes of a principle component analysis,
-    with the blue dot representing the observed summary statistics from the 22
-    population pairs of Philippine vertebrates.
-}{figPCA}
-
-%% Exclusion simulation results
-\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/num_tau_excluded.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to
-    exclude the truth.
-    The plots illustrate the number of true \divt{} parameters excluded from analyses of
-    simulated datasets where \divt{} for 22 pairs of populations is drawn from an
-    exponential distribution, $\divt{} \sim Exp(2)$.
-    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
-    simulation replicates analyzed using $5\e6$ samples from the prior.
-    The proportion of simulation replicates in which at least one true
-    parameter value is excluded from the model preferred by a Bayes factor
-    ($p(\divt{} \notin \hat{M})$) is also given.
-}{figExclusionSimTau}
-
-\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/prob_of_exclusion.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to
-    exclude the truth.
-    The plots illustrate the estimated probability of excluding at least one
-    true \divt{} value from analyses of simulated datasets where \divt{} for 22
-    pairs of populations is drawn from an exponential distribution, $\divt{}
-    \sim Exp(2)$.
-    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
-    simulation replicates analyzed using $5\e6$ samples from the prior.
-    The proportion of simulation replicates in which there is strong support
-    for at least one true parameter value being excluded from the model
-    ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is also given.
-}{figExclusionSimProb}
-
-%% Power results
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
-    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
-    estimate the dispersion index of divergence times (\vmratio{}) from
-    analyses of simulated datasets where \divt{} for 22 pairs of populations is
-    drawn from a series of uniform distributions, $\divt{} \sim U(0,
-    \divt{max})$.
-    The proportion of estimates less than the true value of,
-    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
-    Each plot represents unadjusted median estimates from 500 simulation
-    replicates analyzed using $5\e6$ samples from the prior.
-}{figPowerAccOmegaMedian}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
-    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
-    estimate the dispersion index of divergence times (\vmratio{}) from
-    analyses of simulated datasets where \divt{} for 22 pairs of populations is
-    drawn from a series of uniform distributions, $\divt{} \sim U(0,
-    \divt{max})$.
-    The proportion of estimates less than the true value of,
-    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
-    Each plot represents GLM-regression-adjusted mode estimates from 500
-    simulation replicates analyzed using $5\e6$ samples from the prior.
-}{figPowerAccOmegaModeGLM}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_psi_mode.pdf}{
-    The bias of the model-averaging approach of \citet{Hickerson2013} to infer
-    clustered divergences in our simulation-based power analyses.
-    The plots illustrate the estimated number of divergence events ($\hat{\numt{}}$)
-    from analyses of simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
-    uniform distributions, $\divt{} \sim U(0, \divt{max})$.
-    The estimated probability of the method inferring one divergence event, $p(\hat{\numt{}} = 1)$,
-    is given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerPsiMode}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_omega_prob.pdf}{
-    The bias of the model-averaging approach of \citet{Hickerson2013} to
-    support one divergence event in our simulation-based power analyses.
-    The plots illustrate histograms of the estimated posterior probability that
-    the dispersion index of divergence times is less than 0.01 ($p(\vmratio{} <
-    0.01 | \ssSpace)$) from analyses of simulated datasets where \divt{} for 22
-    pairs of populations is drawn from a series of uniform distributions,
-    $\divt{} \sim U(0, \divt{max})$.
-    The proportion of simulation replicates that strongly support one
-    divergence event, $p(BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01} > 10)$,
-    is given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerOmegaProb}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_num_excluded.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to exclude
-    the truth in our simulation-based power analyses.
-    The plots illustrate the number of true \divt{} parameters excluded from analyses of
-    simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
-    uniform distributions, $\divt{} \sim U(0, \divt{max})$. The proportion
-    of simulation replicates in which at least one true parameter value is excluded from
-    the model preferred by a Bayes factor ($p(\divt{} \notin \hat{M})$) is
-    given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerNumExcluded}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_prob_exclusion.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to
-    exclude the truth in our simulatoin-based power analyses.  The plots
-    illustrate the estimated posterior probability of excluding at least one
-    true \divt{} value from analyses of simulated datasets where \divt{} for 22
-    pairs of populations is drawn from a series of uniform distributions,
-    $\divt{} \sim U(0, \divt{max})$. The proportion of simulation replicates in
-    which there is strong support for at least one true parameter value being
-    excluded from the model ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is
-    given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerProbExclusion}
-
-%% Validation results
-\mFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
-    An assessment of the approximate Bayesian model-averageing approach of
-    \citet{Hickerson2013} under the ideal conditions when the prior model
-    is correct (i.e., the pseudo-replicate datasets are simulated from
-    parameters drawn from the same prior distributions used in the
-    analysis).
-    The plots show the relationship between the estimated posterior and true
-    probability of (A \& C) $\numt{}=1$ and (B \& D) $\vmratio{} < 0.01$, based
-    on 50,000 simulations.
-    The results summarize the (A \& B) unadjusted and (C \& D) GLM-adjusted
-    posterior estimate from each simulation replicate.
-    The prior settings for all replicates included five prior models with
-    $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
-    U(0.0001, 0.05)$ for all five models, and
-    $M_1: \divt{} \sim U(0, 0.1)$,
-    $M_2: \divt{} \sim U(0, 1)$,
-    $M_3: \divt{} \sim U(0, 5)$,
-    $M_4: \divt{} \sim U(0, 10)$, and
-    $M_5: \divt{} \sim U(0, 20)$.
-    The number of samples from the prior was $2.5\e6$.
-    The simulated data structure was 8 population pairs, with a single 1000
-    bp locus sampled from 10 individuals from each population.  The 50,000
-    estimates of the posterior probability of one divergence event were
-    assigned to 20 bins of width 0.05.
-    The estimated posterior probability of each bin is plotted against the
-    proportion of replicates in that bin with a true value consistent with
-    one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
-}{figValidationMCBehavior}
-
-\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
-    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
-    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
-    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
-    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
-    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
-    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
-}{figSamplingError}
-
-%% Model-choice tau prior
-\mFigure{../../model-choice/tau-prior/tau_prior.pdf}{
-    The prior distribution on divergence times imposed by the model-averaging prior
-    comprised of five models with different uniform priors on \divt{}:
-    $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_2$ ($\divt{} \sim U(0, 1)$), $M_3$
-    ($\divt{} \sim U(0, 5)$), $M_4$ ($\divt{} \sim U(0, 10)$), $M_5$ ($\divt{}
-    \sim U(0, 20)$).
-}{figMCTauPrior}
-
-%:FIGURE-saturation plot
-\mFigure{../../saturation/saturation-plot.pdf}{
-    The summary statistics $\pi$ \citep{Tajima1983} and $\pi_{net}$
-    \citep{Takahata1985} as a function of divergence time between populations.
-    Each plot represents 1100 pairs of parameter draws and summary statistics
-    calculated from the simulated data.
-    Prior settings for the simulations were $\divt{} \sim U(0, 20)$,
-    $\meanDescendantTheta{} \sim U(0.0005, 0.04)$, and $\ancestralTheta{} \sim
-    U(0.0005, 0.02)$.
-}{figSaturationPlot}
-
-% \mFigure{../../response-redux/results/hickerson/pymsbayes-results/pymsbayes-output/d1/m12345678-combined/mean_by_dispersion.pdf}{
-%     Results of reanalysis.
-% }{figJointPosterior}
+\input{reply-figures.tex}
 

### 2013-10-18 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Breakin up `dpp-msbayes.tex`.
Moving the main text, tables, and figures of `dpp-msbayes.tex` to
`dpp-msbayes-main.tex`, `dpp-msbayes-tables.tex`, and
`dpp-msbayes-figures.tex`.

--- 2013-10-01 manuscripts/dpp-msbayes/dpp-msbayes.tex
@@
-%&<latex>
-\documentclass[letterpaper,12pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{../utils/preamble.tex}
-\input{../utils/macros.tex}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\doublespacing
-\raggedright
-\setlength{\parindent}{0.5in}
-\begin{linenumbers}
-
-\begin{titlepage}
-    \begin{flushleft}
-        \sffamily
-
-        \MakeUppercase{\large\bfseries An Improved Approximate-Bayesian
-            Model-choice Method for Estimating Shared Evolutionary History}
-
-        \vspace{12pt}
-        \textbf{Running head:} \MakeUppercase{Improved Estimation of Shared Evolutionary History}
-
-        \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2}$ and Mark T.\ Holder$^{1}$
-
-        \bigskip
-        $^1$\emph{Department of Ecology and Evolutionary Biology,
-            % Biodiversity Institute,
-            University of Kansas,
-            Lawrence, Kansas 66045}\\[.1in]
-        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
-        joaks1@gmail.com})\\
-
-    \end{flushleft}
-\end{titlepage}
-
-{\sffamily
-    \noindent\textbf{ABSTRACT} \\
-    \noindent To understand the processes that generate biodiversity, it is
-    important to account for large-scale processes that affect the evolutionary
-    history of groups of co-distributed populations of organisms.
-    Such events predict temporally clustered divergences times, a pattern
-    that can be estimated from genetic data from co-distributed species.
-    We introduce a new approximate-Bayesian model for comparative
-    phylogeographical model-choice that estimates the temporal distribution of
-    divergences across taxa from multi-locus DNA sequence data.
-    Our model is an extension of the model implemented in \msb.
-    By reparameterizing the model, and using more flexible priors on divergence
-    models and nuisance parameters, we improve the robustness, accuracy, and power
-    of the method for estimating the posterior probabilities of models
-    of divergence across taxa.
-    Our results demonstrate that the bias of the \msb model toward inferring
-    models of clustered divergences is caused by a combination of (1) uniform
-    priors on nuisance parameters hindering the marginal likelihoods of models
-    with more divergence time parameters, and (2) a prior on divergence models
-    that disfavors models with intermediate numbers of divergence time
-    parameters.
-
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Dirichlet-process prior; approximate Bayesian
-    computation; model choice; phylogeography; biogeography} 
-}
-
-\newpage
-
-\section*{Introduction}
-Understanding the processes that generate biodiversity and regulate
-community assembly is a major goal of evolutionary biology.
-Given the dynamic nature of our planet, a major component of this goal is
-understanding how geological and climatic processes can affect the
-evolutionary history of whole groups of co-distributed species and their
-associated microbiota, including pathogens.
-Such large-scale processes predict patterns of shared evolutionary history
-across effected populations of organisms, e.g., temporally clustered
-divergences across co-distributed species.
-At recent timescales, such events should leave a signature in the genetic
-variation within and among the affected lineages.
-Thus, being able to accurately estimate models of shared evolutionary
-histories across co-distributed taxa would provide an important
-tool for better understanding how regional and global biogeographic
-processes affect biodiversity.
-
-This inference problem is challenging due to the stochastic nature by which
-mutations occur in populations and how they are inherited over generations
-\citep{Hudson1990,WakeleyCoalescent}.
-Thus, a method for estimating historical patterns of divergences across taxa
-should explicitly model the stochastic mutational and ancestral processes that
-generate and filter the genetic variation we observe in present-day genetic
-data.
-An appealing approach would be a comparative, Bayesian model-choice method for
-inferring the probability of competing divergence histories while integrating
-over uncertainty in mutational and ancestral processes via models of nucleotide
-substitution and lineage coalescence.
-
-The software package \msb implements such an approach in an
-approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
-Recently, however, \citet{Oaks2012} found that \msb can be strongly biased
-toward spuriously estimating biogeographically interesting models of highly
-clustered divergence times across taxa.
-The results of their simulation-based assessment of the method suggest that a
-combination of broad, uniformly distributed priors on nuisance parameters and a
-U-shaped prior on divergence models were causing the bias.
-The model implemented in \msb uses a discrete uniform prior over the number of
-possible divergence parameters, which \citet{Oaks2012} show imposes a prior on
-divergence models that puts most of the prior mass on models with either very
-few or very many divergence times shared across the taxa.
-Furthermore, uniform priors used for most of the model's nuisance parameters
-likely hinder the marginal likelihoods of parameter-rich models, i.e., models
-with many divergence time parameters \citep[this is often referred to as
-Lindley's paradox;][]{Lindley1957}.
-Thus, together, the priors on divergence models and nuisance parameters used in
-\msb could create a strong bias toward models with very few divergence time
-parameters \citep{Oaks2012}.
-
-Following the recommendations of \citet{Oaks2012} we modify the \msb model to
-test whether alternative parameterizations and priors will improve the behavior
-of the approximate Bayesian model-choice approach.
-We implement a Dirichlet-process prior over all possible models of divergence,
-and use gamma and beta probability distributions in place of uniform priors on
-many of the model's parameters.
-Our simulation-based analyses show that the new implementation has improved
-robustness, accuracy, and power compared to the original model.
-Also, our results show that both the uniform priors on nuisance parameters and
-the U-shaped prior on divergence models contribute to the biases of \msb.
-
-\section*{Methods}
-
-\subsection*{The model}
-In this section, we describe our model, which is a modification of the model
-implemented in \msb \citep{Huang2011,Oaks2012}.
-The code implementing the new model is freely available in the open-source
-software package \dppmsbayes
-(\href{https://github.com/joaks1/dpp-msbayes}{https://github.com/joaks1/dpp-msbayes}).
-Our analyses described below utilize the freely avaliable, open-source software
-package \pymsbayes (\href{https://github.com/joaks1/PyMsBayes}{https://github.com/joaks1/PyMsBayes}),
-which is a Python API that provides a multi-processing interface to \msb and
-\dppmsbayes.
-We performed the work described below following the principles of
-Open Notebook Science.
-Using version-control software, we make progress in all aspects of the work
-freely and publicly available in real-time at
-\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
-All information necessary to reproduce our results is provided there.
-We follow much of the notation of \citet{Oaks2012}, but modify it to
-aid in our description of the new model.
-A summary of our notation can be found in Table \ref{tabNotation}.
-
-We assume an investigator is interested in inferring the distribution
-of divergence times among \npairs{} pairs of populations.
-For each pair $i$, \popSampleSize{i}{} genome copies have been sampled, with
-\popSampleSize{1}{i} copies sampled from population 1, and \popSampleSize{2}{i}
-sampled from population 2.
-From these genomes, let \nloci{i} be the number of DNA sequence loci collected
-for population pair $i$, and \nlociTotal be the total number of unique loci
-sampled across the \npairs{} pairs of populations.
-We use \alignment{i}{j} to represent the multiple sequence alignment of
-locus $j$ for population pair $i$.
-$\alignmentVector = (\alignment{1}{1}, \ldots,
-    \alignment{\npairs{}}{\nloci{\npairs{}}})$
-is the full dataset,
-i.e., a vector of sequence alignments for all pairs and loci.
-We let \geneTree{i}{j} represent the gene tree upon which \alignment{i}{j}
-evolved according to fixed \hky substitution model parameters \hkyModel{i}{j}.
-The investigator must specify the parameters of all
-$\hkyModelVector = (\hkyModel{1}{1}, \ldots,
-\hkyModel{\npairs{}}{\nloci{\npairs{}}})$
-substitution models by which the alignments evolved along the
-$\geneTreeVector = (\geneTree{1}{1}, \ldots,
-\geneTree{\npairs{}}{\nloci{\npairs{}}})$
-gene trees.
-Furthermore, the investigator must specify a vector of fixed contants
-$\ploidyScalarVector = (\ploidyScalar{1}{1}, \ldots,
-\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$
-that scale the population size parameters for known differences in ploidy among
-loci and/or differences in generation times among population pairs.
-Lastly, the investigator must also specify a vector of fixed constants
-$\mutationRateScalarConstantVector = (\mutationRateScalarConstant{1}{1},
-\ldots, \mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$
-that scale the population size parameters for known differences in
-mutation rates among loci and/or among taxa.
-
-\begin{linenomath}
-With \alignmentVector, \hkyModelVector, \ploidyScalarVector, and
-\mutationRateScalarConstantVector in hand, the joint posterior distribution
-of our model is given by
-\begin{equation}
-    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
-    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) =
-    \frac{p(\alignmentVector \given \geneTreeVector, \divTimeMapVector,
-        \demographicParamVector, \locusMutationRateScalarVector,
-        \locusRateHetShapeParameter, \hkyModelVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)
-        p(\geneTreeVector, \divTimeMapVector, \demographicParamVector,
-        \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-        \hkyModelVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)}{
-        p(\alignmentVector)},
-    \label{eq:fullModelCompactJoint}
-\end{equation}
-which can be expanded out into components that are assumed to be independent
-to get
-\begin{equation}
-    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
-    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) =
-    \frac{p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
-        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
-        \locusMutationRateScalarVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)
-        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
-        p(\locusRateHetShapeParameter)
-        p(\divTimeMapVector)
-        p(\demographicParamVector)}{
-        p(\alignmentVector)},
-    \label{eq:fullModelCompact}
-\end{equation}
-where
-$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
-is a vector of population divergence times for each of the \npairs{} pairs of
-populations,
-$\demographicParamVector = (\demographicParams{1}, \ldots,
-\demographicParams{\npairs{}})$
-is a vector of the demographic parameters for each of the \npairs{} population
-pairs,
-$\locusMutationRateScalarVector = (\locusMutationRateScalar{1}, \ldots
-\locusMutationRateScalar{\nlociTotal})$
-is a vector of locus-specific mutation-rate scaling parameters for each of the
-\nlociTotal loci,
-\locusRateHetShapeParameter is the shape parameter of a gamma-distributed
-prior on \locusMutationRateScalar{}, and
-$p(\alignmentVector)$ is the probability of the data (or the marginal
-likelihood of the model).
-\end{linenomath}
-
-\begin{linenomath}
-To avoid calculating the likelihood terms of Equation \ref{eq:fullModelCompact},
-we distill each sequence alignment \alignment{}{} into a vector of insufficient
-summary statistics \alignmentSS{}{}, thus replacing the full dataset
-$\alignmentVector = (\alignment{1}{1}, \ldots,
-    \alignment{\npairs{}}{\nloci{\npairs{}}})$
-with vectors of summary statistics for each alignment
-$\ssVectorObs = (\alignmentSSObs{1}{1}, \ldots,
-    \alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$.
-For each population pair, the means of the summary statistics are calculated
-across the \nloci{} loci, reducing the vector to
-$\ssVectorObs = (\alignmentSSObs{1}{}, \ldots,
-    \alignmentSSObs{\npairs{}}{})$.
-This allows us to estimate the approximate joint posterior distribution
-{\small
-\begin{equation}
-    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
-    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-    \ssSpace, \hkyModelVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) =
-    \frac{p(\ssSpace \given \geneTreeVector, \hkyModelVector)
-        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
-        \locusMutationRateScalarVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)
-        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
-        p(\locusRateHetShapeParameter)
-        p(\divTimeMapVector)
-        p(\demographicParamVector)}{
-        p(\ssSpace)},
-    \label{eq:approxModelCompact}
-\end{equation}
-}
-where \ssSpace is the multidimensional Euclidean space around the vector of
-summary statistics, the radius of which is the tolerance \tol.
-The sources of approximation are the insufficiency of the statistics and the
-\tol being greater than zero.
-We describe the full model in detail before delving into the numerical
-method of estimating the approximate model.
-\end{linenomath}
-
-\subsubsection*{Likelihood terms of Equation \ref{eq:fullModelCompact}}
-\begin{linenomath}
-The likelihood terms of Equation \ref{eq:fullModelCompact} can
-be expanded out as a product over population pairs and loci
-\begin{equation}
-    p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
-    p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
-    \locusMutationRateScalarVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) = 
-    \prod_{i=1}^{\npairs{}}
-    \prod_{j=1}^{\nloci{i}}
-    p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})
-    p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
-    \locusMutationRateScalar{j}, \ploidyScalar{i}{j},
-    \mutationRateScalarConstant{i}{j}).
-    \label{eq:modelLikelihoodExpanded}
-\end{equation}
-The first term,
-$p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})$,
-is the probability of the sequence alignment of locus $j$ for population pair
-$i$ given the gene tree and \hky substitution model parameters \citep[i.e., the
-``Felsenstein likelihood'';][]{Felsenstein1981}.
-The model allows for an intra-locus recombination rate \recombinationRate,
-which, for simplicity, is assumed to be zero in Equation
-\ref{eq:fullModelCompact}.
-If \recombinationRate is non-zero, this term requires an additional product
-over the columns (sites) of each sequence alignment to allow sites to have
-different genealogies.
-The second term,
-p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
-\locusMutationRateScalar{j}, \ploidyScalar{i}{j},
-\mutationRateScalarConstant{i}{j}),
-is the probability of the gene tree under a multi-population coalescent model
-(i.e., species tree) where the ancestral population of pair $i$ diverges
-and gives rise to the two sampled descendant populations.
-Each \demographicParams{} contains the following demographic parameters:
-The mutation-scaled sizes of the ancestral, \ancestralTheta{}, and descendant
-populations, \descendantTheta{1}{} and \descendantTheta{2}{};
-the proportion of the first, \bottleScalar{1}{}, and second population,
-\bottleScalar{2}{}, that persist during bottlenecks that begin immediately
-after divergence in forward-time;
-the proportion of time between present and divergence when the bottlenecks
-end for both populations, \bottleTime{};
-and the symmetric migration rate between the descendant populations,
-\migrationRate{}.
-Thus, the probability of the $\popSampleSize{i}{}-1$ coalescence times (node
-heights) of gene tree \geneTree{i}{j} is given by a multi-population
-Kingman-coalescent model \citep{Kingman1982} where the ancestral population of
-size
-$\ancestralTheta{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
-diverges at time \divTimeMap{i} into two descendant populations of constant
-size
-$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{1}{i}$
-and
-$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{2}{i}$,
-which, after time $\divTimeMap{i}\bottleTime{i}$, grow exponentially to their
-present size 
-$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
-and
-$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$,
-respectively.
-Following divergence, the descendant populations of pair $i$ exchange migrants at a symmetric rate of
-\migrationRate{i}.
-\end{linenomath}
-
-\subsubsection*{Prior terms of Equation \ref{eq:fullModelCompact}}
-\begin{linenomath}
-The term $p(\locusRateHetShapeParameter)$ is the prior density
-function for the shape-parameter of the gamma-distributed prior on
-rate heterogeneity among loci.
-This prior is $\locusRateHetShapeParameter \sim U(1, 20)$.
-The prior probability of the vector of locus-specific mutation-rate scaling parameters
-given \locusRateHetShapeParameter then expands out as a product over
-the loci
-\begin{equation}
-    p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter) =
-    \prod_{j=1}^{\nlociTotal}
-    p(\locusMutationRateScalar{j} \given \locusRateHetShapeParameter),
-    \label{eq:locusRateHetPrior}
-\end{equation}
-where each \locusMutationRateScalar{} is independently and identicially
-distributed (\iid) as
-$\locusMutationRateScalar{} \sim Gamma(\locusRateHetShapeParameter,
-1/\locusRateHetShapeParameter)$.
-If the recombination rate \recombinationRate is allowed to be
-non-zero, the prior term $p(\recombinationRate)$ would be added
-to Equation \ref{eq:fullModelCompact}, and the prior would be
-$\recombinationRate \sim Gamma(\gammaShape{\recombinationRate},
-\gammaScale{\recombinationRate})$.
-\end{linenomath}
-
-\begin{linenomath}
-The prior term for the demographic parameters, $p(\demographicParamVector)$,
-expands out into its components and as a product over the \npairs{}
-pairs of populations
-\begin{equation}
-    p(\demographicParamVector) =
-    \prod_{i=1}^{\npairs{}}
-    p(\ancestralTheta{i})
-    p(\descendantTheta{1}{i})
-    p(\descendantTheta{2}{i})
-    p(\bottleScalar{1}{i})
-    p(\bottleScalar{2}{i})
-    p(\bottleTime{i})
-    p(\migrationRate{i}).
-    \label{eq:demographicPrior}
-\end{equation}
-The priors for the demographic parameters are:
-$\ancestralTheta{} \sim Gamma(\gammaShape{\ancestralTheta{}},
-\gammaScale{\ancestralTheta{}})$,
-$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$,
-$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$,
-$\bottleScalar{1}{} \sim Beta(\betaA{\bottleScalar{}{}},
-\betaB{\bottleScalar{}{}})$,
-$\bottleScalar{2}{} \sim Beta(\betaA{\bottleScalar{}{}},
-\betaB{\bottleScalar{}{}})$,
-$\bottleTime{} \sim U(0, 1)$,
-and
-$\migrationRate{} \sim Gamma(\gammaShape{\migrationRate{}},
-\gammaScale{\migrationRate{}})$.
-\end{linenomath}
-
-\subsubsection*{Priors on divergence models}
-\begin{linenomath}
-The prior term for the vector of divergence times for each of
-the \npairs{} pairs of populations can be expanded as
-\begin{equation}
-    p(\divTimeMapVector) = p(\divTimeIndexVector)p(\divTimeVector \given \divTimeIndexVector),
-    \label{eq:divModelPrior}
-\end{equation}
-where \divTimeVector is an ordered set of unique divergence time parameters
-$\{\divTime{1}, \ldots, \divTime{\divTimeNum}\}$ whose length
-\divTimeNum can range from 1 to \npairs{},
-and \divTimeIndexVector is the vector of indices of divergence time parameters 
-$(\divTimeIndex{1}, \ldots, \divTimeIndex{\npairs{}})$
-that map the divergence times in \divTimeVector to each of the \npairs{}
-pairs of populations.
-Thus, \divTimeMapVector is the result of applying the mapping function
-\begin{equation}
-    f(\divTimeVector, \divTimeIndexVector, i) = \divTime{\divTimeIndex{i}}
-    \label{eq:divTimeMapFunction}
-\end{equation}
-to each population pair $i$, such that
-$\divTimeMapVector = (\divTimeMap{1} = f(\divTimeVector, \divTimeIndexVector,
-1), \ldots, \divTimeMap{\npairs{}} = f(\divTimeVector, \divTimeIndexVector,
-\npairs{}))$.
-\end{linenomath}
-
-Biologically speaking, \divTimeVector contains the times of divergence events,
-the length of which \divTimeNum is the number of divergence events shared across
-the \npairs{} pairs of populations.
-For example, if \divTimeVector contains a single divergence time parameter
-\divTime{1}, all \npairs{} pairs of populations are constrained to diverge at
-this time (i.e., \divTimeIndexVector would contain the index 1 repeated
-\npairs{} times, and \divTimeMapVector would contain the value \divTime{1}
-repeated \npairs{} times), whereas if it contains \npairs{} divergence time
-parameters, the model is fully generalized to allow all of the pairs to diverge
-at unique times.
-
-\begin{linenomath}
-Each \divTime{} within \divTimeVector is \iid as $\divTime{} \sim
-Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
-Thus, given the number of unique divergence time classes in
-\divTimeIndexVector, this determines the probability of prior term
-$p(\divTimeVector \given \divTimeIndexVector)$.
-The divergence time parameters are in coalescent units relative to the size of
-a constant reference population, which we denote \myTheta{C}, that is equal to
-the expectation of the prior on the size of the descendant populations
-\begin{equation}
-    \myTheta{C} = \mathbb{E}(\descendantTheta{}{}).
-    \label{eq:thetaC}
-\end{equation}
-Given the size of the descendant populations are \iid as
-$\descendantTheta{}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$,
-this becomes
-\begin{equation}
-    \myTheta{C} = \gammaShape{\descendantTheta{}{}}\gammaScale{\descendantTheta{}{}}.
-    \label{eq:thetaCGamma}
-\end{equation}
-The \divTime{} parameters are in units $\myTheta{C}/\mutationRate$
-generations, which we denote as \globalcoalunit generations.
-Thus, each \divTime{} within \divTimeVector is proportional to time and can be
-converted to the number of generations of the reference population, which we
-denote \divTime{G_C}, by assuming a mutation rate and multiplying by the
-effective size of the reference population
-\begin{equation}
-    \divTime{G_C} = \divTime \times \frac{
-        \gammaShape{\descendantTheta{}{}}\gammaScale{\descendantTheta{}{}}}
-        {\mutationRate}.
-    \label{eq:divTimeGenerations}
-\end{equation}
-Thus, for each of the divergence times in \divTimeVector to be on the same
-scale, the relative mutation rates among the pairs of populations are assumed
-to be fixed and known according the user-provided values in
-\mutationRateScalarConstantVector.
-\end{linenomath}
-
-\begin{linenomath}
-As described by \citet{Oaks2012}, to get the divergence times in units
-proportional to the expected number of mutations, we must
-scale them by the realized population size for locus $j$ of population-pair $i$ 
-\begin{equation}
-    \divTimeScaled{i}{j} = \divTimeMap{i} \times \frac{\myTheta{C}}{
-        \descendantThetaMean{i} \ploidyScalar{i}{j}},
-    \label{eq:divTimeScaled}
-\end{equation}
-where \descendantThetaMean{i} is the mean of \descendantTheta{1}{} and
-\descendantTheta{2}{} for pair $i$.
-This gives us the vector of scaled divergence times
-$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
-\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$.
-\end{linenomath}
-
-\begin{linenomath}
-As for the prior term $p(\divTimeIndexVector)$, the total sample space of
-ordered realizations of \divTimeIndexVector is all the possible partitions of
-\npairs{} elements into 1 to \npairs{} categories.
-The total number of possible partitions is a sum of the Stirling numbers of
-the second kind over all possible numbers of categories \divTimeNum
-\begin{equation}
-    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
-    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
-    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
-    \label{eq:bell}
-\end{equation}
-which is the Bell Number \citep{Bell1934}.
-The original \msb model only samples over the unordered realizations of
-\divTimeIndexVector, such that the sample space is reduced to all the possible
-integer partitions of \npairs{} \citep{Oaks2012,Huang2011,OeisPartitionNumber,
-    OeisPartitionTriangle,Malenfant2011}.
-We denote the set of all possible integer partitions of \npairs{} as
-\integerPartitionSet{\npairs{}} and the length of that set as
-\integerPartitionNum{\npairs{}}.
-The advantages, disadvantages, and justification of ignoring the order
-of \divTimeIndexVector is discussed in detail below.
-\end{linenomath}
-
-\begin{linenomath}
-We implement two prior probability distributions over the space
-of all possible discrete divergence models (\divTimeIndexVector).
-The first simply gives all possible unordered partitions of \npairs{} elements
-equal probability
-\begin{equation}
-    p(\divTimeIndexVector) = \frac{1}{\integerPartitionNum{\npairs{}}},
-    \label{eq:divModelPriorUniform}
-\end{equation}
-i.e., a discrete uniform prior over all the integer partitions of \npairs{}
-(discrete divergence models).
-We denote this prior as
-$\divTimeIndexVector \sim \priorUniform$.
-\end{linenomath}
-
-The second prior we implement is based on the Dirichlet-process, which is a
-stochastic process that groups elements into an unknown number of discrete
-parameter classes \citep{Ferguson1973,Antoniak1974}.
-The process is controlled by the concentration parameter \concentrationParam,
-which determines how clustered the process will be.
-We use the Dirichlet process to place a prior over all possible ordered
-partitions of \npairs{} elements, which we denote as $\divTimeIndexVector \sim
-\priorDPP{}$.
-We take a hierarchical approach and use a prior probability
-distribution (i.e., hyperprior) for \concentrationParam.
-More specifically, we use a gamma-distributed prior $\concentrationParam =
-Gamma(\gammaShape{\concentrationParam}, \gammaScale{\concentrationParam})$,
-where \gammaShape{\concentrationParam} and \gammaScale{\concentrationParam} are
-specified by the user.
-
-\begin{linenomath}
-This provides a great deal of flexibility for specifying the prior uncertainty
-regarding divergence models.
-The \concentrationParam determines the prior probability that any two
-pairs of populations $i$ and $j$ will be assigned to the same divergence time
-parameter
-\begin{equation}
-    p(\divTimeIndex{i} = \divTimeIndex{j}) = \frac{1}{1 + \concentrationParam},
-    \label{eq:dppPriorSameClass}
-\end{equation}
-and also the prior probability of the number of divergence time parameters
-\begin{equation}
-    p(\divTimeNum \boldsymbol{\mid} \concentrationParam, \npairs{}) = 
-    \frac{\stirlingFirst{\npairs{}}{\divTimeNum} \concentrationParam^{\divTimeNum}}
-    {\prod_{i=1}^{\npairs{}}(\concentrationParam + i - 1)},
-    \label{eq:dppPriorNumClasses}
-\end{equation}
-where \stirlingFirst{\cdot}{\cdot} are the unsigned Stirling numbers of the
-first kind.
-Equations \ref{eq:dppPriorSameClass} and \ref{eq:dppPriorNumClasses} show that
-smaller values of \concentrationParam will favor fewer divergence time
-parameters, and thus more clustered models of divergence, whereas larger values
-will favor more divergence time parameters, and thus less clustered models of
-divergence.
-\end{linenomath}
-
-\subsection*{Differences between our model and the original \msb model}
-\subsubsection*{The prior on divergence models}
-One of the key differences between our model and that of \msb \citep{Huang2011}
-is the prior distribution on divergence models.
-As discussed in \citet{Oaks2012}, in \msb the only prior used for
-\divTimeIndexVector is a combination of a discrete uniform prior over the
-possible number of divergence events \divTimeNum from 1 to \npairs{} with a
-multinomial distribution on the number of times each index of \divTimeVector
-appears in \divTimeIndexVector, with the constraint that all \divTime{}
-parameters are represented at least once (see Equation 2 of \citet{Oaks2012}).
-We denote this prior used by \msb as $\divTimeIndexVector \sim \priorOld$.
-\citet{Oaks2012} discuss how placing a uniform prior over the number of
-divergence parameters (denoted \divTimeNum here, and as \numt{} in
-\citet{Huang2011}) imposes an U-shaped prior over discrete divergence
-models (\divTimeIndexVector; see
-Figure 5B of \citet{Oaks2012}).
-To avoid this, we place priors directly on the sample space of divergence
-models, thus eliminating the parameter \numt{} from the model.
-We introduce two priors on divergence models:
-(1) a prior that is uniform over all unordered divergence models, and
-(2) a Dirichlet-process prior on all ordered divergence models.
-The latter provides the user with a great deal of flexibility in
-expressing their prior beliefs about models of divergence.
-
-\subsubsection*{Estimating ordered divergence models}
-As mentioned above, \msb samples over unordered divergence models
-(i.e., unordered partitions of the \npairs{} pairs of populations).
-That is, the identity of each population pair, and all the information
-associated with it, is discarded.
-In our implementation, inference can be done on either unordered or ordered
-models of divergence.
-This is discussed in more detail in the description of the ABC implementation
-below.
-
-\subsubsection*{The priors on nuisance parameters}
-Following the recommendations of \citet{Oaks2012}, we have replaced the use of
-continuous uniform distributions for priors on many of the
-model's parameters (\divTime{}, \ancestralTheta{}, \descendantTheta{1}{},
-\descendantTheta{2}{}, \bottleScalar{1}{}, \bottleScalar{2}{},
-\recombinationRate, \migrationRate{}) with more flexible parametric
-distributions from the exponential family.
-We introduce gamma-distributed priors for rate parameters that have a sample
-space of all positive real numbers (\divTime{}, \ancestralTheta{},
-\descendantTheta{1}{}, \descendantTheta{2}{}, \recombinationRate,
-\migrationRate{}), and beta-distributed priors for parameters that are
-proportions bounded by zero and one (\bottleScalar{1}{} and
-\bottleScalar{2}{}).
-Not only do these priors have nice conjugacy properties, but they also provide
-the user with much greater flexibility in expressing prior uncertainty
-regarding the parameters of the model.
-
-Another distinction between our model and that of \msb, is the prior on the
-sizes of the descendant populations of each pair.
-As described by \citet{Oaks2012}, \msb uses the joint prior
-\begin{equation}
-    \descendantTheta{1}{}, \descendantTheta{2}{} \sim
-    Beta(1,1) \times 2 \times U(\uniformMin{\myTheta{}},
-    \uniformMax{\descendantTheta{}{}}),
-    \label{eq:jointThetaPrior}
-\end{equation}
-such that the user-specified uniform prior on descendant population
-size is a prior on the \emph{mean} size of the two descendant
-populations of each pair.
-Under our model, the sizes of the descendant populations of each
-pair are \iid as
-$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$
-and
-$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$.
-This relaxes the assumption that the sizes of the two descendant populations
-are interdependent and negatively correlated.
-
-\subsubsection*{Flexibility in parameterizing the model}
-In our implementation, we provide the ability to control the richness of the
-model.
-For the \myTheta{} parameters, the model can be fully generalized to
-allow each population pair to have three parameters:
-\ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}.
-Furthermore, any model of \myTheta{} parameters nested within this
-general model can also be specified, including the most restricted model
-where the ancestral and descendant populations of each pair share
-a single \myTheta{} parameter.
-
-We also provide the option of eliminating the parameters associated with the
-post-divergence bottlenecks in the descendant populations of each pair
-(\bottleTime{}, \bottleScalar{1}{}, and \bottleScalar{2}{}),
-which constrains the descendant populations to be of
-constant size from present back to the divergence event.
-Also, rather than eliminate the bottleneck parameters,
-we allow \bottleScalar{1}{} and \bottleScalar{2}{} to be constrained to be
-equal, which removes one free parameter from the model for each of the
-population pairs.
-
-Overall, our implementation allows an investigator to specify a model that has
-as many as seven parameters per population pair
-(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{},
-\bottleTime{}, \bottleScalar{1}{}, \bottleScalar{2}{}, and
-\migrationRate{})
-or as few as one parameter per pair
-(\myTheta{}),
-in addition to the $\popSampleSize{i}{} - 1$ coalescence-time parameters.
-
-\subsubsection*{Time scale}
-As described above, in our model divergence times are in units of
-$\myTheta{C}/\mutationRate$ generations, where \myTheta{C} is the expectation
-of the prior on descendant-population size.
-As described by \citet{Oaks2012}, in \msb, \myTheta{C} is half of the upper
-limit of the continuous uniform prior on the mean of the descendant population
-sizes.
-
-\subsection*{ABC estimation of the posterior of the model}
-\subsubsection*{Sampling from the prior}
-To estimate the approximate posterior of Equation \ref{eq:approxModelCompact},
-we use an ABC rejection algorithm.
-The first step of this algorithm entails collecting a random sample of
-parameters values from the joint prior and their associated summary
-statistics.
-Each sample is generated by
-(1) drawing values of all the model's parameters, which we denote \hpvector{},
-from their respective prior distributions;
-(2) rescaling the divergence times
-$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
-from units proportional to time to units proportional to the expected number
-of mutations via Equation \ref{eq:divTimeScaled} to get 
-$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
-\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$;
-(3) simulating gene trees $\geneTreeVector = (\geneTree{1}{1}, \ldots, 
-\geneTree{\npairs{}}{\nloci{\npairs{}}})$
-for each locus of each population pair by drawing coalescent times from
-a multi-population Kingman-coalescent model given the demographic parameters;
-(4) simulating sequence alignments 
-$\alignmentVector = (\alignment{1}{1}, \ldots, \alignment{\npairs{}}{\nloci{\npairs{}}})$
-along the gene trees under the
-\hky substitution parameters
-$\hkyModelVector = (\hkyModel{1}{1}, \ldots, \hkyModel{\npairs{}}{\nloci{\npairs{}}})$
-that have the same number of sequences and sequence lengths as the observed
-dataset;
-(5) calculating population genetic summary statistics
-$\ssVector{} = (\alignmentSS{1}{1}, \ldots, \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$
-from the simulated sequence alignments (these are the same statistics
-calculated from the observed alignments, \ssVectorObs{});
-and (6) reducing the summary statistics to the means across loci for each
-population pair to get
-$\ssVector{} = (\alignmentSS{1}{}, \ldots, \alignmentSS{\npairs{}}{})$, which is
-the same summary statistic vector estimated from the observed data \ssVectorObs.
-After repeating this procedure \numPriorSamples times, we obtain a random
-sample of parameter vectors
-$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$
-from the model prior and their associated vectors of summary statistics
-$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$.
-
-For all of our analyses below, we use four summary statistics for each of
-the pairs of populations:
-$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
-\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
-Furthermore, in addition to model parameters, each sample \hpvector{}
-also contains four statistics that summarize \divTimeMapVector:
-the mean (\divTimeMean), variance (\divTimeVar), dispersion index
-($\divTimeDispersion = \divTimeVar/\divTimeMean$), and the number
-of divergence time parameters (\divTimeNum).
-These have been denoted as \meant{}{}, \vart{}{}, \vmratio{}, and \numt{},
-respectively \citep{Hickerson2006,Huang2011,Oaks2012}.
-We use the new notation to avoid confusion that the values represent prior or
-posterior expectations, and to indicate that \vmratio{} is not a parameter of
-the model and \numt{} is no longer a parameter in the new implementation.
-
-\subsubsection*{Ordering of taxon-specific summary statistics}
-As alluded to in the model description, \msb does not maintain the order of the
-taxon-specific summary statistics \alignmentSS{}{} within each \ssVector{}.
-Rather, the summary statistics are re-ordered by descending values of average
-pairwise differences between the descendant populations
-\citep[$\pi_b$;][]{NeiLi1979,Huang2011}.
-This has the advantage of reducing the sample space of the number of discrete
-divergence models \divTimeMapVector, but there are at least two disadvantages.
-First, additional information in the data is lost.
-By discarding the identity of the \npairs{} pairs of populations, all
-pair-specific information about the amount of data (e.g., the number gene
-copies collected from each of the populations [\popSampleSize{1}{} and
-\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
-taxon- and locus-specific parameters (\hkyModel{}{},
-\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
-\locusMutationRateScalar{}) is lost.
-Second, the results are more difficult to interpret, because they can be no
-longer be directly associated to the taxa under study.
-
-The original descriptions of the \msb model claim that this re-ordering is
-justified by the fact that of $\pi_b$ (and other summary statistics) are
-unrelated to the sample sizes \popSampleSize{1}{} and \popSampleSize{2}{} of
-each pair and are thus exchangeable \citep{Hickerson2006,Huang2011}.
-This is actually incorrect for two of reasons.
-First, the expectation of $\pi_b$ is not independent of samples sizes.
-If there are more than one coalescent events in the ancestor, which is expected
-to be common on phylogeographic timescales, more samples will increase the
-probability of capturing these deeper events, and thus affect the average
-pairwise differences between the descendant populations.
-Also, other statistics that estimate gross diversity (e.g., $\pi$ and
-$\theta_W$) are clearly not independent of sample size.
-Second, for variables to be exchangeable, they do not need to be independent,
-but their marginal distributions must be the same (i.e., they must be
-identically distributed).
-The simulated alignments and their summary statistics are not identically
-distributed because of differing sample sizes \emph{and} taxon- and
-locus-specific parameters \hkyModel{}{}, \mutationRateScalarConstant{}{}, and
-\ploidyScalar{}{}.
-
-Thus, the theoretical basis for this reshuffling is questionable.
-It may be justifiable as an additional coarsening of the data,
-by removing the information associated with the identity of the
-population pairs.
-However, we do not provide a formal proof that this re-ordering of the pairs
-does not introduce bias.
-One can imagine situations in which the sampling intensity (i.e., the
-number of gene copies, loci, and locus length) is highly skewed across
-pairs and/or there is large heterogeneity in mutational processes
-(e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
-In such cases, it seems possible that discarding this information could
-cause bias.
-Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
-is to minimize Euclidean distances between simulated datasets in which the true
-model has a single divergence event, and (2) the reordering increases the
-model's tendency to infer a single divergence event \citep{Huang2011}, it seems
-possible that this approach could be biasing the method \citep{Oaks2012}.
-
-To maintain compatibility and comparability with \msb we maintain
-the re-ordering of taxon-specific summary statistics by $\pi_b$.
-However, we also allow the order to be maintained, and ordered
-divergence models to be estimated.
-
-\subsubsection*{Obtaining an approximate posterior from the prior samples}
-We use a rejection algorithm to retain an approximate posterior sample of
-\paramSampleVector{} from the prior sample
-$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$.
-First, the observed summary statistics \ssVectorObs, and the summary statistics
-of the prior samples 
-$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$,
-are standardized using the means and standard deviations of the statistics from
-the prior sample (i.e., the prior mean is subtracted from each statistic, and then
-it is divided by the prior standard deviation).
-After all statistics are standardized, the Euclidean distance between
-\ssVectorObs and each \ssVector{} within \ssMatrix is calculated.
-The samples that fall within a range of tolerance \tol around \ssVectorObs
-are retained.
-The range of tolerance is determined by specifying the number of desired
-posterior samples to be retained.
-Post-hoc adjustment of the posterior sample can also be performed with a number
-of regression techniques \citep{Beaumont2002,Blum2009,Leuenberger2010}.
-For our analyses, we use the general linear model (GLM) regression adjustment
-\citep{Leuenberger2010} as implemented in \abctoolbox
-\citep[v1.1;][]{ABCtoolbox}, which \citet{Oaks2012} showed performed very
-similarly to weighted local-linear regression and multinomial logistic
-regression adjustments \citep{Beaumont2002} for \msb posteriors.
-
-\subsection*{Assessing model-choice behavior and robustness}
-Following the simulation-based approach of \citet{Oaks2012}, we characterize
-the behavior of several models under the ideal conditions where the
-data are generated from parameters drawn from the same prior distributions used
-for analysis (i.e., the prior is correct).
-We selected four model priors for these analyses (Table~\ref{tabPriors}):
-(1) A model to represent the original \msb model, \modelOld, with
-priors
-$\divTimeIndexVector \sim \priorOld$,
-$\divTime{} \sim U(0,10)$,
-$\ancestralTheta{} \sim U(0, 0.05)$,
-and
-$\descendantThetaMean{} \sim U(0, 0.05)$;
-(2) a Dirichlet-process prior model, \modelDPP, with priors
-$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
-and
-$\descendantTheta{}{} \sim Exp(40)$;
-(3) a model with uniform prior probability over unordered divergence models,
-\modelUniform, with priors
-$\divTimeIndexVector \sim \priorUniform$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
-and
-$\descendantTheta{}{} \sim Exp(40)$;
-and
-(4) a model with the \msb prior on unordered divergence models, but with
-exponential priors on nuisance parameters,
-\modelUshaped, with priors
-$\divTimeIndexVector \sim \priorOld$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
-and
-$\descendantTheta{}{} \sim Exp(40)$.
-We selected the exponential prior on divergence time used in models \modelDPP,
-\modelUniform, and \modelUshaped to have the same variance as the uniform prior
-in model \modelOld.
-We selected the exponential prior on population size used in models \modelDPP,
-\modelUniform, and \modelUniform to have the same mean as the uniform prior in
-model \modelOld, so that all four models have the same \myTheta{C} and thus the
-same units of time.
-All of the models were the same in other respects, with three free \myTheta{}
-parameters for each population pair, two uniformly distributed ($beta(1,1)$)
-\bottleScalar{}{} parameters per pair, no migration, no recombination,
-and unordered divergence models.
-For all of our simulations, we used a simulated data structure of eight
-population pairs, with a single 1000 base-pair locus sampled from 10
-individuals from each population.
-
-For each of the four models, we generated $1\e6$ samples from the prior, and
-simulated 50,000 pseudo-replicate datasets, also drawn from the prior.
-We then analyzed each of the replicate datasets, retaining a
-posterior of 1000 samples from the respective prior.
-A GLM-regression adjusted posterior was also estimated from each of the
-posterior samples \citep{Leuenberger2010}.
-To assess the robustness of each of the four models, we also analyzed the
-pseudo-replicate datasets simulated under the other three models.
-Overall, for each model, we produced 200,000 posterior estimates,
-50,000 from the replicate datasets simulated under that model,
-and 150,000 estimated from the replicate datasets simulated under the
-other three models.
-
-For each set of 50,000 simulated datasets, we used the posterior estimates
-to assess the model-choice behavior of each model.
-We did this by assigning the 50,000 estimates of the posterior probability
-of one divergence event to 20 bins of width 0.05, and plotted
-the estimated posterior probability of each bin against the proportion of
-replicates in that bin with a true value consistent with one divergence
-event \citep{Huelsenbeck2004,Oaks2012}.
-We did this using two criteria for the one divergence model:
-(1) the number of divergence time parameters ($\divTimeNum = 1$) and
-(2) the dispersion index of divergence times \citep[$\divTimeDispersion <
-0.01$][]{Hickerson2006,Huang2011}.
-We used the one-divergence model to assess model-choice behavior, because
-it is often of biogeographic interest and is easily comparable among
-the three different priors used on divergence models.
-
-In addition to the four models above, we also assessed the behavior of a model
-with ordered divergence models (i.e., the order of the taxon-specific summary
-statistic vectors were maintained for the observed and simulated datasets).
-For this, we used a model with identical priors as \modelDPP, but that samples
-over ordered divergence models.
-We denote this model as \modelDPPOrdered.
-We generated $1\e6$ prior samples and 50,000 pseudo-replicate datasets, and
-analyzed them as above.
-We were not able to analyze the replicate datasets of the other models under
-the ordered model, because the identity of the population pairs is not
-contained in the simulations of the other models.
-
-\subsection*{Assessing power}
-We evaluated the power of the same four models (Table~\ref{tabPriors}) to
-detect random variation in divergence times using methods similar to
-\citet{Oaks2012}.
-For all of our power simulations, we used a simulated data structure identical
-to that of the empirical dataset of Philippine vertebrates analyzed by
-\citet{Oaks2012}, which consists of 22 pairs of populations.
-For each of the four models, we generated $2\e6$ samples from the prior.
-We then generated pseudo-observed datasets from three series of models.
-One series of models, which we denote \powerSeriesOld, is identically
-distributed as \modelOld except that the divergence times for each of the 22
-pairs of populations are randomly drawn from a series of uniform distributions,
-$U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and
-2.0, in \globalcoalunit generations.
-A second series of models, \powerSeriesUniform, are identically distributed as
-\modelUniform except that the 22 divergence times are randomly drawn from the
-same series of uniform priors as above.
-The third series of models, \powerSeriesExp, is also identically
-distributed as \modelUniform except that the 22 divergence times are randomly
-drawn from a series of of exponential distributions: $Exp(17.3205)$,
-$Exp(8.6603)$, $Exp(5.7735)$, $Exp(4.33013)$, $Exp(3.4641)$, and $Exp(1.7321)$.
-These exponential distributions have the same variance as their uniform
-counterparts in the first two series of models.
-
-For each of the six models in each of the three series of models, we simulated
-1000 pseudo-replicate datasets (18,000 datasets in total).
-We then analyzed each replicate dataset under all four prior models
-(Table~\ref{tabPriors}), producing 72,000 posterior estimates, each with 1000
-samples.
-We also estimated a GLM-regression adjusted posterior from each of the
-posterior samples \citep{Leuenberger2010}.
-
-\section*{Results}
-\subsection*{Validation analyses: Estimation accuracy}
-In terms of estimating summary statistics of model parameters (\divTimeNum,
-\divTimeDispersion, and \divTimeMean), the \modelDPP performs similarly when
-applied to datasets generated under all four models of Table \ref{tabPriors}
-(Figs.\
-\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP}).
-The same is true for models \modelUniform (Figs.\
-\labelcref{figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform})
-and \modelUshaped (Figs.\
-\labelcref{figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped}).
-All three of the models with exponential priors on nuisance parameters
-(\modelDPP, \modelUniform, and \modelUshaped) perform well when applied
-to the data generated under the original \msb model, \modelOld (Figs.\
-\labelcref{figValidationAccuracyOldDPP,figValidationAccuracyOldUniform,figValidationAccuracyOldUshaped}).
-The \modelOld model also performed well on its own datasets (Figure
-\ref{figValidationAccuracyOldOld}), but performed poorly when applied to the
-data generated under the other three models (Figs.\
-\labelcref{figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld}).
-Thus, the newly implemented models display greater robustness than
-the original model.
-
-Posterior adjustment with GLM regression does improve the estimation
-accuracy for the continuous variables \divTimeDispersion and \divTimeMean,
-however, this is not always true for estimating the discrete variable
-\divTimeNum (Figs.\
-\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP,figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform,figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped,figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld,figValidationAccuracyOldOld}).
-
-\subsection*{Validation analyses: Model-choice accuracy}
-The \msb model, and our modification of it, is a model-choice method
-with the primary purpose of estimating the probabilities of models
-of divergence across taxa.
-Thus, it is critical to assess the method's ability to accurately
-estimate the posterior probabilities of divergence models.
-Our results demonstrate that the unadjusted estimates of divergence models are
-generally more accurate than the regression-adjusted estimates, especially when
-using $\divTimeDispersion < 0.01$ to assess the probability of the one
-divergence model (Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
-Thus, we will focus our discussion of the results on the unadjusted estimates.
-
-The model-choice behavior of the \modelDPP is relatively accurate when applied
-to datasets generated under models \modelDPP, \modelUniform, and \modelOld
-(Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP}).
-It is less accurate and tends to underestimate the probability of the
-one-divergence model when applied to the data generated under \modelUshaped
-(Figure \ref{figValidationModelChoiceUshapedDPP}).
-The \modelUniform model performs almost identically as \modelDPP in
-terms of model-choice accuracy (Figs.\
-\labelcref{figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}).
-
-The original \msb model \modelOld performs reasonably well when applied
-to its own datasets (Figure \ref{figValidationModelChoiceOldOld}), but
-it performs poorly when applied to data generated under any of the other
-models (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
-In these cases the \modelOld model is strongly biased towards over-estimating
-the posterior probability of the one-divergence model, especially when the
-divergence models are not distributed under its U-shaped prior (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld}).
-The other model with the U-shaped prior on divergence models, but exponential
-priors on nuisance parameters (\modelUshaped), performs similarly to \modelOld
-in that it performs well when applied to its own data, but overestimates
-the probability of the one-divergence model when applied to data generated
-by the other models (Figs.\
-\labelcref{figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceUshapedUshaped}).
-However, the bias toward over-estimating the probability of the one-divergence
-model is stronger in the \modelOld model than \modelUshaped.
-
-Overall, the result suggest that the \modelDPP and \modelUniform models are
-relatively robust in terms of model-choice accuracy, and when model violations
-do cause them to be biased, they tend to under-estimate the probability of
-the biogeographically interesting model of a single divergence event.
-In contrast, the \modelOld model is very sensitive to model violations,
-and strongly over-estimates the one-divergence model whenever the model
-is misspecified.
-Furthermore, the results suggest that using exponentially distributed priors on
-nuisance parameters rather than uniform priors helps the \modelUshaped model
-perform better than \modelOld, but it is still hindered by the U-shaped prior
-on divergence models and tends to over-estimate the probability of the
-one-divergence model whenever there are violations of the model.
-
-\subsection*{Validation analyses: Ordered divergence models}
-Our results show that the ordered models of divergence, \modelDPPOrdered,
-perform similarly to sampling over unordered models (\modelDPP) when the prior
-is correct (Figs.\
-\labelcref{figValidationAccuracyDPPOrderedDPPOrdered,figValidationModelChoiceDPPOrderedDPPOrdered}).
-This suggests that the method is not adversely affected by the large increase
-in possible discrete models (from 22 unordered to 4140 ordered models) when
-there are 8 pairs of populations.
-
-\subsection*{Power analyses: Estimation accuracy}
-All of the models we evaluated (Table \ref{tabPriors}) struggle to estimate the
-dispersion index of divergence times \divTimeDispersion regardless of which of
-the three series of models (Table \ref{tabPowerModels}) the data were generated
-under (Figs.\
-\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform,figPowerAccuracyExpOld,figPowerAccuracyExpUshaped,figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped,figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
-When the divergence times of the 22 population pairs are randomly drawn from a
-series of exponential priors (\powerSeriesExp), the \modelDPP model is the
-best estimator of \divTimeDispersion, followed by \modelUniform (figs.\
-\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform}).
-The \modelOld model is strongly biased toward underestimating
-\divTimeDispersion, estimating values of zero for most of the replicates across
-all the data models of \powerSeriesExp (Figure \ref{figPowerAccuracyExpOld}).
-The results of the \modelUshaped model are intermediate between those of
-\modelOld and the new models \modelDPP and \modelUniform
-(Figure \ref{figPowerAccuracyExpUshaped}).
-
-When the true divergence times are randomly drawn from a series of uniform
-priors (\powerSeriesUniform), all of the models poorly estimate
-\divTimeDispersion (Figs.\
-\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped}).
-The \modelDPP and \modelUniform models tend to over-estimate the variance in
-divergence times (Figs.\
-\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform}), whereas
-the \modelOld model, again, underestimates \divTimeDispersion, estimating
-values of zero for most replicates across all the data models of
-\powerSeriesUniform (Figure \ref{figPowerAccuracyUniformOld}).
-Again, the performance of the \modelUshaped model is intermediate between the
-\modelOld and \modelDPP/\modelUniform models (Figure
-\ref{figPowerAccuracyUniformUshaped}).
-This pattern of results across the four models is very similar when applied
-to data simulated under the series of models of \powerSeriesOld (Figs.\
-\labelcref{figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
-
-\subsection*{Power analyses: Model choice}
-There is a large amount of heterogeneity in estimating the number of divergence
-events (\divTimeNum) both among the prior models, and across the different
-series of data models (Table \ref{tabPowerModels}) (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped}).
-The \modelDPP model tends to infer highly clustered divergence models across
-all three series of data models when the divergences are recent, but performs
-the best of the four prior models  when divergences are random over a timescale
-of about one to two coalescent units (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiUniformDPP,figPowerPsiOldDPP,}).
-The \modelUniform model performs the best when divergences are random
-over the most recent timescales, but tends to only infer a tight
-cluster of \divTimeNum values even when divergences are random
-over longer periods (Figs.\
-\labelcref{figPowerPsiExpUniform,figPowerPsiUniformUniform,figPowerPsiOldUniform,}).
-The \modelOld model performs the worst of the four prior models across
-all three series of data models, inferring a single divergence event across
-most of the 18,000 simulations (Figs.\
-\labelcref{figPowerPsiExpOld,figPowerPsiUniformOld,figPowerPsiOldOld,}).
-Even when the simulated data are identically distributed as the \modelOld model
-(except for the divergence times) the \modelOld is still strongly biased
-towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
-model is the least likely to infer a single event when applied to these data
-(Figure \ref{figPowerPsiOldDPP}).
-Using exponential priors on nuisance parameters does increase the power
-of the \modelUshaped model compared to \modelOld across all three series
-of data models, but the U-shaped prior still prevents the model
-from performing as well as the \modelDPP and \modelUniform models (Figs.\
-\labelcref{figPowerPsiExpUshaped,figPowerPsiUniformUshaped,figPowerPsiOldUshaped,}).
-
-The improved power of the new models are even more pronounced when looking at
-estimates of the dispersion index of divergence times (\divTimeDispersion)
-across the simulations (Figs.\
-\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped}).
-The \modelDPP and \modelUniform models perform similarly across all three
-series of data models, inferring values of \divTimeDispersion consistent with
-one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
-replicates across all the simulations (Figs.\
-\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionOldDPP,figPowerDispersionOldUniform}).
-The \modelOld model infers values consistent with a single divergence event in
-most of the replicates across all the simulations (Figs.\
-\labelcref{figPowerDispersionExpOld,figPowerDispersionUniformOld,figPowerDispersionOldOld}).
-Using exponential priors on nuisance parameters greatly increases the power
-of the \modelUshaped model relative to \modelOld, but it still has
-less power than the models with DPP or uniform priors across
-divergence models (Figs.\
-\labelcref{figPowerDispersionExpUshaped,figPowerDispersionUniformUshaped,figPowerDispersionOldUshaped}).
-
-When looking at the estimated posterior probability of the single-divergence
-model across the power analyses, we also see the increased power of the
-new models (Figs.\
-\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped}).
-The \modelDPP model estimates low posterior probability of
-$\divTimeNum = 1$ across all of the simulations (Figs.\
-\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbUniformDPP,figPowerPsiProbOldDPP}).
-The \modelDPP does infer strong support for one divergence as gauged by a Bayes
-factor of greater than 10.
-However, this is not that unexpected given the low prior probability of a
-single divergence under this model.
-Also, the Bayes factor calculations for the \modelDPP replicates are only
-approximate; they are calculated based on the prior mean of the concentration
-parameter \concentrationParam.
-
-The \modelUniform model infers very low posterior probabilities for the
-one-divergence model across all the simulations (Figs.\
-\labelcref{figPowerPsiProbExpUniform,figPowerPsiProbUniformUniform,figPowerPsiProbOldUniform}),
-whereas the \modelOld model infers high posterior probabilities of a single
-divergence for most replicates across all simulations (Figs.\
-\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld}).
-Again, we see that the exponential priors on nuisance parameters greatly
-increase power and result in lower estimates of the probability of
-one divergence when comparing \modelUshaped to \modelOld (Figs.\
-\labelcref{figPowerPsiProbExpUshaped,figPowerPsiProbUniformUshaped,figPowerPsiProbOldUshaped}).
-
-Lastly, when we look at the estimated posterior probability of
-\divTimeDispersion being consistent with one divergence parameter
-($p(\divTimeDispersion < 0.01 | \ssSpace)$), we see the same pattern of model
-behavior, with \modelDPP and \modelUniform inferring low probabilities across
-all simulations, \modelOld inferring high probabilities, and \modelUshaped
-inferring intermediate values (Figs.\
-\labelcref{figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
-
-\section*{Discussion}
-Our results demonstrate that using alternative priors on nuisance parameters
-and divergence models improved the behavior of the \msb model.
-The model-choice estimation of our implementation is more accurate and,
-perhaps more importantly, shows greater robustness to model
-violations (Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
-The original model is very sensitive to model violations and, when present,
-strongly over-estimates the probability of one divergence event shared
-across all taxa (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
-When more appropriate priors are used for nuisance parameters, and either
-a Dirichlet-process or uniform prior applied across divergence models,
-the model is less sensitive to violations (Figs.\
-\labelcref{figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}),
-and, when they do cause bias, the method is conservative and tends to
-under-estimate the probability of biogeographically interesting models of
-temporally clustered divergences 
-\labelcref{figValidationModelChoiceUshapedDPP,figValidationModelChoiceUshapedUniform}).
-
-Furthermore, the modifications to the model also improve the method's power to
-detect random variation in divergence times, reducing the bias toward
-estimating clustered divergences (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped,figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped,figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped,figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
-Our results are similar to those of \citet{Oaks2012} in that we also find
-the \msb model to infer strong support for clustered divergences when
-divergences are random over quite broad timescales.
-\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld,figPowerDispersionProbExpOld,figPowerDispersionProbUniformOld,figPowerDispersionProbOldOld}).
-Our results expand on this by showing that this behavior is consistent
-across of range conditions underlying the data.
-When using the new priors, the method is less prone to spurious inference of
-clustered divergence models, and much less likely to incorrectly infer such
-models with strong support.
-
-By evaluating a model intermediate between the old and new implementation
-(\modelUshaped), we were able to determine the relative effects of the
-modifications to model.
-Across all of our analyses, the results show that using better priors on
-nuisance parameters alone does improve the performance of the method.
-The magnitude of the bias toward inferring support for the one-divergence
-model when there are model violations is reduced when the exponential
-priors are used in place of the uniform priors.
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped}).
-Furthermore, better priors on on nuisance parameters improves the method's
-power to detect temporally random divergences.
-Throughout our analyses, the intermediate model (\modelUshaped) performs
-better than the \msb model, but not as well as the models with alternative
-priors on divergence models.
-This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
-models of temporally clustered divergences is caused by a combination of
-(1) uniform priors on nuisance parameters hindering the marginal likelihoods of
-models with more \divTime{} parameters and
-(2) the U-shaped prior on divergence models giving low prior density to models
-with intermediate numbers of divergence parameters.
-
-While our modifications improve the behavior of the model, we urge
-caution when using the method and interpreting its results.
-The method attempts to approximate the posterior of a very parameter rich model
-using little information from the data.
-For example, when applied to the dataset of Philippine vertebrates analyzed by
-\citet{Oaks2012} the model can have as many as 604--625 free parameters
-(depending on \divTimeNum), and samples over 1002 divergence models.
-Our modifications allow this model to be simplified to 471--492 free
-parameters.
-Sampling over hundreds of parameters and more than a thousand models is asking
-a lot from a few summary statistics calculated from the sequence alignments of
-each taxon pair.
-These simple summary statistics contain little information about many
-of the free parameters in the model.
-Thus, we expect the improved method will still be sensitive to priors, and the
-power, while improved, may still be low.
-Furthermore, ABC methods in general have been shown to be biased
-\citep{Robert2011} for model choice.
-As recommended by \citet{Oaks2012}, any results from the method should be
-accompanied by assessments of prior sensitivity and power.
-We encourage investigators to view this method and its predecessor as a means
-of exploring their data rather than as rigorous means of evaluating hypotheses.
-
-Given the difficulty of this estimation problem, we anticipate that
-full-likelihood methods that can leverage all of the information present in the
-sequence data will become increasingly important in robustly estimating the
-temporal history of divergences across taxa \citep{JeetDiss}.
-With improving numerical methods for sampling over models of differing
-dimensionality \citep{Green1995,Lemey2009} and increasing efficiency of
-likelihood calculations \citep{Ayres2012}, analyzing rich comparative
-phylogeograpical models in full-likelihood Bayesian framework is becoming
-computationally practical, especially when considering that the computational
-load of simulating millions of random datasets from the prior under the ABC
-approach is nontrivial.
-
-\section*{Acknowledgments}
-We thank the National Science Foundation for supporting this work (DEB
-1011423).
-J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
-Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
-Society, KU Department of Ecology and Evolutionary Biology, and the KU
-Biodiversity Institute.
-
-\bibliography{../bib/references}
-
-%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\renewcommand\listfigurename{Figure Captions}
-\cftsetindents{fig}{0cm}{2.2cm}
-\renewcommand\cftdotsep{\cftnodots}
-\setlength\cftbeforefigskip{10pt}
-\cftpagenumbersoff{fig}
-\listoffigures
-
-\end{linenumbers}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-%:TABLE-notation descriptions
-\begin{table}[htbp]
-    \sffamily
-    \scriptsize
-    \rowcolors{2}{}{myGray}
-    \addtolength{\tabcolsep}{-0.1cm}
-\caption{Summary of the notation used throughout this work; modified from \citet{Oaks2012}.}
-    \centering
-    \begin{tabular}{ l >{\raggedright\hangindent=0.5cm}m{14cm} }
-        \toprule
-        \textbf{Symbol} & \textbf{Description} \tn
-        \midrule
-        \npairs{} & Number of population pairs. \tn
-        \popSampleSize{i}{} & The number of genome copies sampled from population pair $i$, with \popSampleSize{1}{i} sampled from population 1, and \popSampleSize{2}{i} from population 2. \tn
-        \nloci{i} & Number of loci sampled from population pair $i$. \tn
-        \nlociTotal & Total number of unique loci sampled. \tn
-        \alignment{i}{j} & Sequence alignment of locus $j$ sampled from population pair $i$. \tn
-        \alignmentSSObs{i}{j} & Population genetic summary statistics calculated from \alignment{i}{j}. \tn
-        \alignmentVector & Vector containing the sequence alignments of each locus from each population pair: $(\alignment{1}{1},\ldots,\alignment{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \ssVectorObs & Vector containing the summary statistics of each locus from each population pair: $(\alignmentSSObs{1}{1},\ldots,\alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \ssSpace & Multi-dimensional Euclidean space around the observed summary statistics, \ssVectorObs. \tn
-        \tol & Radius of \ssSpace, i.e., the tolerance of the ABC estimation. \tn
-        \geneTree{i}{j} & Gene tree of the sequences in \alignment{i}{j}. \tn
-        \geneTreeVector & Vector containing the gene trees of each locus from each population pair: $(\geneTree{1}{1},\ldots,\geneTree{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \divTimeNum & Number of unique population divergence times among the \npairs{} population pairs. \tn
-        \divTime{} & One of the \divTimeNum \iid divergence time paramters in \globalcoalunit generations. \tn
-        \divTimeVector & Ordered set of divergence time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
-        \divTimeIndex{i} & The index of the divergence time parameter in \divTimeVector that is mapped to population pair $i$. \tn
-        \divTimeIndexVector & Vector of divergence time parameter indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn
-        \divTimeMap{i} & Time of divergence in \globalcoalunit generations between the populations of pair $i$. \tn
-        \divTimeMapVector & Vector of divergence times for each of the population pairs: $(\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})$. \tn
-        \divTimeScaled{i}{j} & Scaled time of divergence between the populations of pair $i$ for locus $j$. \tn
-        \divTimeScaledVector & Vector containing the scaled divergence times of each locus from each population pair: $(\divTimeScaled{1}{1},\ldots,\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \descendantTheta{1}{i}, \descendantTheta{2}{i} & \myTheta{} of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
-        \ancestralTheta{i} & \myTheta{} of the population ancestral to pair $i$. \tn
-        \descendantThetaVector{1}, \descendantThetaVector{2} & Vectors $(\descendantTheta{1}{1},\ldots,\descendantTheta{1}{\npairs{}})$ and $(\descendantTheta{2}{1},\ldots,\descendantTheta{2}{\npairs{}})$, respectively.
-\tn
-        \ancestralThetaVector & Vector containing the \ancestralTheta{} parameters for each population pair: $(\ancestralTheta{1},\ldots,\ancestralTheta{\npairs{}})$. \tn
-        \locusMutationRateScalar{j} & mutation-rate scaling parameter for locus $j$. \tn
-        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate parameters: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
-        \locusRateHetShapeParameter & Hyper-parameter for the shape of the gamma prior distribution on each \locusMutationRateScalar{}. \tn
-        \bottleScalar{1}{i}, \bottleScalar{2}{i} & \myTheta{}-scaling parameters that determine the magnitude of the population bottleneck in the $1^{st}$ and $2^{nd}$ descendant population of pair $i$, respectively.
-The bottleneck in each descendant population begins immediately after divergence.
-\tn
-        \bottleScalarVector{1}, \bottleScalarVector{2} & Vectors $(\bottleScalar{1}{1},\ldots,\bottleScalar{1}{\npairs{}})$ and $(\bottleScalar{2}{1},\ldots,\bottleScalar{2}{\npairs{}})$, respectively.
-\tn
-        \bottleTime{i} & Proportion of time between present and \divt{i} when the bottleneck ends for the descendant populations of pair $i$. \tn
-        \bottleTimeVector & Vector containing the \bottleTime{} parameters for each population pair: $(\bottleTime{1},\ldots,\bottleTime{\npairs{}})$. \tn
-        \migrationRate{i} & Symmetric migration rate between the descendant populations of pair $i$. \tn
-        \migrationRateVector & Vector containing the migration rates for each population pair: $(\migrationRate{1},\ldots,\migrationRate{\npairs{}})$. \tn
-        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
-This constant is intended to allow the user to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa.
-\tn
-        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
-This constant is intended to allow the user to scale \myTheta{} for differences in mutation rates among loci or among taxa.
-\tn
-        \ploidyScalarVector & Vector of ploidy and/or generation-time scaling constants: $(\ploidyScalar{1}{1},\ldots,\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \mutationRateScalarConstantVector & Vector of mutation-rate scaling constants: $(\mutationRateScalarConstant{1}{1},\ldots,\mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \divTimeMean  & Mean of \divTimeMapVector. \tn
-        \divTimeVar  & Variance of \divTimeMapVector. \tn
-        \divTimeDispersion  & Dispersion index of \divTimeMapVector ($\divTimeVar/\divTimeMean$). \tn
-        \numPriorSamples & Number of samples from the joint prior. \tn
-        \paramSampleVector{} & Vector of parameter values drawn from the joint prior. \tn
-        \ssVector{} & Vector containing the summary statistics calculated from data simulated under parameter values drawn from the prior (\hpvector{}). \tn
-        \paramSampleMatrix{} & Random sample of $\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples}$ drawn form the prior. \tn
-        \ssMatrix & Summary statistic vectors $\ssVector{1}, \ldots, \ssVector{\numPriorSamples}$ for each $\hpvector{1}, \ldots, \hpvector{\numPriorSamples}$ drawn from the prior. \tn
-        \bottomrule
-    \end{tabular}
-    \label{tabNotation}
-\end{table}
-
-\clearpage
-
-\begin{table}[htbp]
-    \sffamily
-    % \footnotesize
-    \addtolength{\tabcolsep}{-0.08cm}
-    \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
-    \caption{The model priors evaluated in our simulation-based analyses.}
-    \centering
-    \begin{tabular}{ l l l l l }
-        \toprule
-        & \multicolumn{4}{c}{Priors} \\
-        \cmidrule(){2-5}
-        Model & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
-        \midrule
-            \modelOld & $\divTimeIndexVector \sim \priorOld$
-                      & $\divTime{} \sim U(0,10)$
-                      & $\ancestralTheta{} \sim U(0, 0.05)$
-                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$
-                      & $\divTime{} \sim Exp(0.3464)$
-                      & $\ancestralTheta{} \sim Exp(40)$
-                      & $\descendantTheta{}{} \sim Exp(40)$ \\
-            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
-                          & $\divTime{} \sim Exp(0.3464)$
-                          & $\ancestralTheta{} \sim Exp(40)$
-                          & $\descendantTheta{}{} \sim Exp(40)$ \\
-            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
-                          & $\divTime{} \sim Exp(0.3464)$
-                          & $\ancestralTheta{} \sim Exp(40)$
-                          & $\descendantTheta{}{} \sim Exp(40)$ \\
-        \bottomrule
-    \end{tabular}
-    \label{tabPriors}
-\end{table}
-
-\clearpage
-
-\begin{table}[htbp]
-    \sffamily
-    % \footnotesize
-    % \addtolength{\tabcolsep}{-0.08cm}
-    \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
-    \caption{The models we used to simulate pseudo-replicate datasets for
-        assessing the power of the models in Table \ref{tabPriors}.}
-    \centering
-    \begin{tabular}{ l l l l l }
-        \toprule
-        & \multicolumn{4}{c}{Priors} \\
-        \cmidrule(){2-5}
-        Model series & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
-        \midrule
-            \powerSeriesOld & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.2)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.4)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.6)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.8)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,1.0)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,2.0)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-        \midrule
-            \powerSeriesUniform & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.2)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.4)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.6)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.8)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,1.0)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,2.0)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-        \midrule
-            \powerSeriesExp & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(17.3205)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(8.6603)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(5.7735)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(4.3301)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(3.4641)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(1.7321)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-        \bottomrule
-    \end{tabular}
-    \label{tabPowerModels}
-\end{table}
-
-\clearpage
-
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
 %% Validation accuracy
@@
 }{figPowerDispersionProbOldUshaped}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-
-\singlespacing
-
-\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape Supporting Information}}
-
-% PUT MAIN TEXT CITATION HERE
-% \begin{thebibliography}{1}
-% \providecommand{\natexlab}[1]{#1}
-% \providecommand{\url}[1]{\texttt{#1}}
-% \providecommand{\urlprefix}{URL }
-
-% \bibitem
-
-% \end{thebibliography}
-
-%% SUPPL TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\setcounter{table}{0}
-
-\end{document}

### 2013-10-18 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Breakin up `dpp-msbayes.tex`.
Moving the main text, tables, and figures of `dpp-msbayes.tex` to
`dpp-msbayes-main.tex`, `dpp-msbayes-tables.tex`, and
`dpp-msbayes-figures.tex`.

--- 2013-10-07 manuscripts/dpp-msbayes/dpp-msbayes.tex
@@
-%&<latex>
-\documentclass[letterpaper,12pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{../utils/preamble.tex}
-\input{../utils/macros.tex}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\doublespacing
-\raggedright
-\setlength{\parindent}{0.5in}
-\begin{linenumbers}
-
-\begin{titlepage}
-    \begin{flushleft}
-        \sffamily
-
-        \MakeUppercase{\large\bfseries An Improved Approximate-Bayesian
-            Model-choice Method for Estimating Shared Evolutionary History}
-
-        \vspace{12pt}
-        \textbf{Running head:} \MakeUppercase{Improved Estimation of Shared Evolutionary History}
-
-        \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2}$ and Mark T.\ Holder$^{1}$
-
-        \bigskip
-        $^1$\emph{Department of Ecology and Evolutionary Biology,
-            % Biodiversity Institute,
-            University of Kansas,
-            Lawrence, Kansas 66045}\\[.1in]
-        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
-        joaks1@gmail.com})\\
-
-    \end{flushleft}
-\end{titlepage}
-
-{\sffamily
-    \noindent\textbf{ABSTRACT} \\
-    \noindent To understand the processes that generate biodiversity, it is
-    important to account for large-scale processes that affect the evolutionary
-    history of groups of co-distributed populations of organisms.
-    Such events predict temporally clustered divergences times, a pattern
-    that can be estimated from genetic data from co-distributed species.
-    We introduce a new approximate-Bayesian model for comparative
-    phylogeographical model-choice that estimates the temporal distribution of
-    divergences across taxa from multi-locus DNA sequence data.
-    Our model is an extension of the model implemented in \msb.
-    By reparameterizing the model, and using more flexible priors on divergence
-    models and nuisance parameters, we improve the robustness, accuracy, and power
-    of the method for estimating the posterior probabilities of models
-    of divergence across taxa.
-    Our results demonstrate that the bias of the \msb model toward inferring
-    models of clustered divergences is caused by a combination of (1) uniform
-    priors on nuisance parameters hindering the marginal likelihoods of models
-    with more divergence time parameters, and (2) a prior on divergence models
-    that disfavors models with intermediate numbers of divergence time
-    parameters.
-
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Dirichlet-process prior; approximate Bayesian
-    computation; model choice; phylogeography; biogeography} 
-}
-
-\newpage
-
 \section*{Introduction}
@@
 approach is nontrivial.
-
-\section*{Acknowledgments}
-We thank the National Science Foundation for supporting this work (DEB
-1011423).
-J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
-Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
-Society, KU Department of Ecology and Evolutionary Biology, and the KU
-Biodiversity Institute.
-
-\bibliography{../bib/references}
-
-%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\renewcommand\listfigurename{Figure Captions}
-\cftsetindents{fig}{0cm}{2.2cm}
-\renewcommand\cftdotsep{\cftnodots}
-\setlength\cftbeforefigskip{10pt}
-\cftpagenumbersoff{fig}
-\listoffigures
-
-\end{linenumbers}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-%:TABLE-notation descriptions
-\begin{table}[htbp]
-    \sffamily
-    \scriptsize
-    \rowcolors{2}{}{myGray}
-    \addtolength{\tabcolsep}{-0.1cm}
-\caption{Summary of the notation used throughout this work; modified from \citet{Oaks2012}.}
-    \centering
-    \begin{tabular}{ l >{\raggedright\hangindent=0.5cm}m{14cm} }
-        \toprule
-        \textbf{Symbol} & \textbf{Description} \tn
-        \midrule
-        \npairs{} & Number of population pairs. \tn
-        \popSampleSize{i}{} & The number of genome copies sampled from population pair $i$, with \popSampleSize{1}{i} sampled from population 1, and \popSampleSize{2}{i} from population 2. \tn
-        \nloci{i} & Number of loci sampled from population pair $i$. \tn
-        \nlociTotal & Total number of unique loci sampled. \tn
-        \alignment{i}{j} & Sequence alignment of locus $j$ sampled from population pair $i$. \tn
-        \alignmentSSObs{i}{j} & Population genetic summary statistics calculated from \alignment{i}{j}. \tn
-        \alignmentVector & Vector containing the sequence alignments of each locus from each population pair: $(\alignment{1}{1},\ldots,\alignment{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \ssVectorObs & Vector containing the summary statistics of each locus from each population pair: $(\alignmentSSObs{1}{1},\ldots,\alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \ssSpace & Multi-dimensional Euclidean space around the observed summary statistics, \ssVectorObs. \tn
-        \tol & Radius of \ssSpace, i.e., the tolerance of the ABC estimation. \tn
-        \geneTree{i}{j} & Gene tree of the sequences in \alignment{i}{j}. \tn
-        \geneTreeVector & Vector containing the gene trees of each locus from each population pair: $(\geneTree{1}{1},\ldots,\geneTree{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \divTimeNum & Number of unique population divergence times among the \npairs{} population pairs. \tn
-        \divTime{} & One of the \divTimeNum \iid divergence time paramters in \globalcoalunit generations. \tn
-        \divTimeVector & Ordered set of divergence time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
-        \divTimeIndex{i} & The index of the divergence time parameter in \divTimeVector that is mapped to population pair $i$. \tn
-        \divTimeIndexVector & Vector of divergence time parameter indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn
-        \divTimeMap{i} & Time of divergence in \globalcoalunit generations between the populations of pair $i$. \tn
-        \divTimeMapVector & Vector of divergence times for each of the population pairs: $(\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})$. \tn
-        \divTimeScaled{i}{j} & Scaled time of divergence between the populations of pair $i$ for locus $j$. \tn
-        \divTimeScaledVector & Vector containing the scaled divergence times of each locus from each population pair: $(\divTimeScaled{1}{1},\ldots,\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \descendantTheta{1}{i}, \descendantTheta{2}{i} & \myTheta{} of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
-        \ancestralTheta{i} & \myTheta{} of the population ancestral to pair $i$. \tn
-        \descendantThetaVector{1}, \descendantThetaVector{2} & Vectors $(\descendantTheta{1}{1},\ldots,\descendantTheta{1}{\npairs{}})$ and $(\descendantTheta{2}{1},\ldots,\descendantTheta{2}{\npairs{}})$, respectively.
-\tn
-        \ancestralThetaVector & Vector containing the \ancestralTheta{} parameters for each population pair: $(\ancestralTheta{1},\ldots,\ancestralTheta{\npairs{}})$. \tn
-        \locusMutationRateScalar{j} & mutation-rate scaling parameter for locus $j$. \tn
-        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate parameters: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
-        \locusRateHetShapeParameter & Hyper-parameter for the shape of the gamma prior distribution on each \locusMutationRateScalar{}. \tn
-        \bottleScalar{1}{i}, \bottleScalar{2}{i} & \myTheta{}-scaling parameters that determine the magnitude of the population bottleneck in the $1^{st}$ and $2^{nd}$ descendant population of pair $i$, respectively.
-The bottleneck in each descendant population begins immediately after divergence.
-\tn
-        \bottleScalarVector{1}, \bottleScalarVector{2} & Vectors $(\bottleScalar{1}{1},\ldots,\bottleScalar{1}{\npairs{}})$ and $(\bottleScalar{2}{1},\ldots,\bottleScalar{2}{\npairs{}})$, respectively.
-\tn
-        \bottleTime{i} & Proportion of time between present and \divt{i} when the bottleneck ends for the descendant populations of pair $i$. \tn
-        \bottleTimeVector & Vector containing the \bottleTime{} parameters for each population pair: $(\bottleTime{1},\ldots,\bottleTime{\npairs{}})$. \tn
-        \migrationRate{i} & Symmetric migration rate between the descendant populations of pair $i$. \tn
-        \migrationRateVector & Vector containing the migration rates for each population pair: $(\migrationRate{1},\ldots,\migrationRate{\npairs{}})$. \tn
-        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
-This constant is intended to allow the user to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa.
-\tn
-        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
-This constant is intended to allow the user to scale \myTheta{} for differences in mutation rates among loci or among taxa.
-\tn
-        \ploidyScalarVector & Vector of ploidy and/or generation-time scaling constants: $(\ploidyScalar{1}{1},\ldots,\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \mutationRateScalarConstantVector & Vector of mutation-rate scaling constants: $(\mutationRateScalarConstant{1}{1},\ldots,\mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \divTimeMean  & Mean of \divTimeMapVector. \tn
-        \divTimeVar  & Variance of \divTimeMapVector. \tn
-        \divTimeDispersion  & Dispersion index of \divTimeMapVector ($\divTimeVar/\divTimeMean$). \tn
-        \numPriorSamples & Number of samples from the joint prior. \tn
-        \paramSampleVector{} & Vector of parameter values drawn from the joint prior. \tn
-        \ssVector{} & Vector containing the summary statistics calculated from data simulated under parameter values drawn from the prior (\hpvector{}). \tn
-        \paramSampleMatrix{} & Random sample of $\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples}$ drawn form the prior. \tn
-        \ssMatrix & Summary statistic vectors $\ssVector{1}, \ldots, \ssVector{\numPriorSamples}$ for each $\hpvector{1}, \ldots, \hpvector{\numPriorSamples}$ drawn from the prior. \tn
-        \bottomrule
-    \end{tabular}
-    \label{tabNotation}
-\end{table}
-
-\clearpage
-
-\begin{table}[htbp]
-    \sffamily
-    % \footnotesize
-    \addtolength{\tabcolsep}{-0.08cm}
-    \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
-    \caption{The model priors evaluated in our simulation-based analyses.  For
-        the \modelDPP model, the prior on the concentration parameter,
-        $\concentrationParam \sim Gamma(\cdot,\cdot)$, was set to $Gamma(2,2)$
-        for the validation analyses and $Gamma(1.5,18.1)$ for the power
-        analyses.}
-    \centering
-    \begin{tabular}{ l l l l l }
-        \toprule
-        & \multicolumn{4}{c}{Priors} \\
-        \cmidrule(){2-5}
-        Model & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
-        \midrule
-            \modelOld & $\divTimeIndexVector \sim \priorOld$
-                      & $\divTime{} \sim U(0,10)$
-                      & $\ancestralTheta{} \sim U(0, 0.05)$
-                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
-                      & $\divTime{} \sim Exp(0.3464)$
-                      & $\ancestralTheta{} \sim Exp(40)$
-                      & $\descendantTheta{}{} \sim Exp(40)$ \\
-            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
-                          & $\divTime{} \sim Exp(0.3464)$
-                          & $\ancestralTheta{} \sim Exp(40)$
-                          & $\descendantTheta{}{} \sim Exp(40)$ \\
-            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
-                          & $\divTime{} \sim Exp(0.3464)$
-                          & $\ancestralTheta{} \sim Exp(40)$
-                          & $\descendantTheta{}{} \sim Exp(40)$ \\
-        \bottomrule
-    \end{tabular}
-    \label{tabPriors}
-\end{table}
-
-\clearpage
-
-\begin{table}[htbp]
-    \sffamily
-    % \footnotesize
-    % \addtolength{\tabcolsep}{-0.08cm}
-    \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
-    \caption{The models we used to simulate pseudo-replicate datasets for
-        assessing the power of the models in Table \ref{tabPriors}.}
-    \centering
-    \begin{tabular}{ l l l l l }
-        \toprule
-        & \multicolumn{4}{c}{Priors} \\
-        \cmidrule(){2-5}
-        Model series & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
-        \midrule
-            \powerSeriesOld & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.2)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.4)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.6)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.8)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,1.0)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,2.0)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-        \midrule
-            \powerSeriesUniform & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.2)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.4)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.6)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.8)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,1.0)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,2.0)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-        \midrule
-            \powerSeriesExp & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(17.3205)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(8.6603)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(5.7735)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(4.3301)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(3.4641)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(1.7321)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-        \bottomrule
-    \end{tabular}
-    \label{tabPowerModels}
-\end{table}
-
-\clearpage
-
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
-%% Validation accuracy
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelDPP}
-}{figValidationAccuracyDPPDPP}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelUniform}
-}{figValidationAccuracyDPPUniform}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelOld}
-}{figValidationAccuracyDPPOld}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelUshaped}
-}{figValidationAccuracyDPPUshaped}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelDPP}
-}{figValidationAccuracyUniformDPP}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelUniform}
-}{figValidationAccuracyUniformUniform}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelOld}
-}{figValidationAccuracyUniformOld}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelUshaped}
-}{figValidationAccuracyUniformUshaped}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelDPP}
-}{figValidationAccuracyOldDPP}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelUniform}
-}{figValidationAccuracyOldUniform}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelOld}
-}{figValidationAccuracyOldOld}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelUshaped}
-}{figValidationAccuracyOldUshaped}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelDPP}
-}{figValidationAccuracyUshapedDPP}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelUniform}
-}{figValidationAccuracyUshapedUniform}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelOld}
-}{figValidationAccuracyUshapedOld}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelUshaped}
-}{figValidationAccuracyUshapedUshaped}
-
-%% Validation model choice
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelDPP}
-}{figValidationModelChoiceDPPDPP}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelUniform}
-}{figValidationModelChoiceDPPUniform}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelOld}
-}{figValidationModelChoiceDPPOld}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelUshaped}
-}{figValidationModelChoiceDPPUshaped}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelDPP}
-}{figValidationModelChoiceUniformDPP}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelUniform}
-}{figValidationModelChoiceUniformUniform}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelOld}
-}{figValidationModelChoiceUniformOld}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelUshaped}
-}{figValidationModelChoiceUniformUshaped}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelDPP}
-}{figValidationModelChoiceOldDPP}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelUniform}
-}{figValidationModelChoiceOldUniform}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelOld}
-}{figValidationModelChoiceOldOld}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelUshaped}
-}{figValidationModelChoiceOldUshaped}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelDPP}
-}{figValidationModelChoiceUshapedDPP}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelUniform}
-}{figValidationModelChoiceUshapedUniform}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelOld}
-}{figValidationModelChoiceUshapedOld}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelUshaped}
-}{figValidationModelChoiceUshapedUshaped}
-
-%% Validation ordered models
-\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPPOrdered}{\modelDPPOrdered}
-}{figValidationAccuracyDPPOrderedDPPOrdered}
-
-\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPPOrdered}{\modelDPPOrdered}
-}{figValidationModelChoiceDPPOrderedDPPOrdered}
-
-%% Power accuracy results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerAccuracyExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerAccuracyExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelOld}
-}{figPowerAccuracyExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerAccuracyExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerAccuracyUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerAccuracyUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerAccuracyUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerAccuracyUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerAccuracyOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerAccuracyOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelOld}
-}{figPowerAccuracyOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerAccuracyOldUshaped}
-
-%% Power psi results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerPsiExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerPsiExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelOld}
-}{figPowerPsiExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerPsiExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerPsiUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerPsiUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerPsiUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerPsiUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerPsiOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerPsiOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelOld}
-}{figPowerPsiOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerPsiOldUshaped}
-
-%% Power dispersion results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerDispersionExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerDispersionExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelOld}
-}{figPowerDispersionExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerDispersionExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerDispersionUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerDispersionUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerDispersionUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerDispersionUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerDispersionOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerDispersionOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelOld}
-}{figPowerDispersionOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerDispersionOldUshaped}
-
-%% Power psi probability results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerPsiProbExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerPsiProbExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelOld}
-}{figPowerPsiProbExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerPsiProbExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerPsiProbUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerPsiProbUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerPsiProbUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerPsiProbUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerPsiProbOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerPsiProbOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelOld}
-}{figPowerPsiProbOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerPsiProbOldUshaped}
-
-%% Power dispersion prob results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerDispersionProbExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerDispersionProbExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelOld}
-}{figPowerDispersionProbExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerDispersionProbExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerDispersionProbUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerDispersionProbUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerDispersionProbUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerDispersionProbUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerDispersionProbOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerDispersionProbOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelOld}
-}{figPowerDispersionProbOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerDispersionProbOldUshaped}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-
-\singlespacing
-
-\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape Supporting Information}}
-
-% PUT MAIN TEXT CITATION HERE
-% \begin{thebibliography}{1}
-% \providecommand{\natexlab}[1]{#1}
-% \providecommand{\url}[1]{\texttt{#1}}
-% \providecommand{\urlprefix}{URL }
-
-% \bibitem
-
-% \end{thebibliography}
-
-%% SUPPL TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\setcounter{table}{0}
-
-\end{document}

### 2013-10-18 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex
Breakin up `dpp-msbayes.tex`.
Moving the main text, tables, and figures of `dpp-msbayes.tex` to
`dpp-msbayes-main.tex`, `dpp-msbayes-tables.tex`, and
`dpp-msbayes-figures.tex`.

--- 2013-10-07 manuscripts/dpp-msbayes/dpp-msbayes.tex
@@
-%&<latex>
-\documentclass[letterpaper,12pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{../utils/preamble.tex}
-\input{../utils/macros.tex}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\doublespacing
-\raggedright
-\setlength{\parindent}{0.5in}
-\begin{linenumbers}
-
-\begin{titlepage}
-    \begin{flushleft}
-        \sffamily
-
-        \MakeUppercase{\large\bfseries An Improved Approximate-Bayesian
-            Model-choice Method for Estimating Shared Evolutionary History}
-
-        \vspace{12pt}
-        \textbf{Running head:} \MakeUppercase{Improved Estimation of Shared Evolutionary History}
-
-        \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2}$ and Mark T.\ Holder$^{1}$
-
-        \bigskip
-        $^1$\emph{Department of Ecology and Evolutionary Biology,
-            % Biodiversity Institute,
-            University of Kansas,
-            Lawrence, Kansas 66045}\\[.1in]
-        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
-        joaks1@gmail.com})\\
-
-    \end{flushleft}
-\end{titlepage}
-
-{\sffamily
-    \noindent\textbf{ABSTRACT} \\
-    \noindent To understand the processes that generate biodiversity, it is
-    important to account for large-scale processes that affect the evolutionary
-    history of groups of co-distributed populations of organisms.
-    Such events predict temporally clustered divergences times, a pattern
-    that can be estimated from genetic data from co-distributed species.
-    We introduce a new approximate-Bayesian model for comparative
-    phylogeographical model-choice that estimates the temporal distribution of
-    divergences across taxa from multi-locus DNA sequence data.
-    Our model is an extension of the model implemented in \msb.
-    By reparameterizing the model, and using more flexible priors on divergence
-    models and nuisance parameters, we improve the robustness, accuracy, and power
-    of the method for estimating the posterior probabilities of models
-    of divergence across taxa.
-    Our results demonstrate that the bias of the \msb model toward inferring
-    models of clustered divergences is caused by a combination of (1) uniform
-    priors on nuisance parameters hindering the marginal likelihoods of models
-    with more divergence time parameters, and (2) a prior on divergence models
-    that disfavors models with intermediate numbers of divergence time
-    parameters.
-
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Dirichlet-process prior; approximate Bayesian
-    computation; model choice; phylogeography; biogeography} 
-}
-
-\newpage
-
-\section*{Introduction}
-Understanding the processes that generate biodiversity and regulate
-community assembly is a major goal of evolutionary biology.
-Given the dynamic nature of our planet, a major component of this goal is
-understanding how geological and climatic processes can affect the
-evolutionary history of whole groups of co-distributed species and their
-associated microbiota, including pathogens.
-Such large-scale processes predict patterns of shared evolutionary history
-across affected populations of organisms, e.g., temporally clustered
-divergences across co-distributed species.
-At recent timescales, such events should leave a signature in the genetic
-variation within and among the affected lineages.
-Thus, being able to accurately estimate models of shared evolutionary
-histories across co-distributed taxa would provide an important
-tool for better understanding how regional and global biogeographic
-processes affect biodiversity.
-
-This inference problem is challenging due to the stochastic nature by which
-mutations occur in populations and how they are inherited over generations
-\citep{Hudson1990,WakeleyCoalescent}.
-Thus, a method for estimating historical patterns of divergences across taxa
-should explicitly model the stochastic mutational and ancestral processes that
-generate and filter the genetic variation we observe in present-day genetic
-data.
-An appealing approach would be a comparative, Bayesian model-choice method for
-inferring the probability of competing divergence histories while integrating
-over uncertainty in mutational and ancestral processes via models of nucleotide
-substitution and lineage coalescence.
-
-The software package \msb implements such an approach in an
-approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
-Recently, however, \citet{Oaks2012} found that \msb can be strongly biased
-toward spuriously estimating biogeographically interesting models of highly
-clustered divergence times across taxa.
-The results of their simulation-based assessment of the method suggest that a
-combination of broad, uniformly distributed priors on nuisance parameters and a
-U-shaped prior on divergence models are causing the bias.
-The model implemented in \msb uses a discrete uniform prior over the number of
-possible divergence parameters, which \citet{Oaks2012} show imposes a prior on
-divergence models that puts most of the prior mass on models with either very
-few or very many divergence times shared across the taxa.
-Furthermore, uniform priors used for most of the model's nuisance parameters
-likely hinder the marginal likelihoods of parameter-rich models, i.e., models
-with many divergence time parameters \citep{Lindley1957}.
-Thus, together, the priors on divergence models and nuisance parameters used in
-\msb could create a strong bias toward models with very few divergence time
-parameters \citep{Oaks2012}.
-
-Following the recommendations of \citet{Oaks2012}, we modify the \msb model to
-test whether alternative parameterizations and priors will improve the behavior
-of the approximate-Bayesian model-choice approach.
-We implement a Dirichlet-process prior over all possible models of divergence,
-and use gamma and beta probability distributions in place of uniform priors on
-many of the model's parameters.
-Our simulation-based analyses show that the new implementation has improved
-robustness, accuracy, and power compared to the original model.
-Also, our results show that both the uniform priors on nuisance parameters and
-the U-shaped prior on divergence models contribute to the biases of \msb.
-
-\section*{Methods}
-
-\subsection*{The model}
-In this section, we describe our model, which is a modification of the model
-implemented in \msb \citep{Huang2011,Oaks2012}.
-The code implementing the new model is freely available in the open-source
-software package \dppmsbayes
-(\href{https://github.com/joaks1/dpp-msbayes}{https://github.com/joaks1/dpp-msbayes}).
-Our analyses described below utilize the freely avaliable, open-source software
-package \pymsbayes (\href{https://github.com/joaks1/PyMsBayes}{https://github.com/joaks1/PyMsBayes}),
-which is a Python API that provides a multi-processing interface to \msb and
-\dppmsbayes.
-We performed the work described below following the principles of
-Open Notebook Science.
-Using version-control software, we make progress in all aspects of the work
-freely and publicly available in real-time at
-\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
-All information necessary to reproduce our results is provided there.
-We follow much of the notation of \citet{Oaks2012}, but modify it to
-aid in our description of the new model.
-A summary of our notation can be found in Table \ref{tabNotation}.
-
-We assume an investigator is interested in inferring the distribution
-of divergence times among \npairs{} pairs of populations.
-For each pair $i$, \popSampleSize{i}{} genome copies have been sampled, with
-\popSampleSize{1}{i} copies sampled from population 1, and \popSampleSize{2}{i}
-sampled from population 2.
-From these genomes, let \nloci{i} be the number of DNA sequence loci collected
-for population pair $i$, and \nlociTotal be the total number of unique loci
-sampled across the \npairs{} pairs of populations.
-We use \alignment{i}{j} to represent the multiple sequence alignment of
-locus $j$ for population pair $i$.
-$\alignmentVector = (\alignment{1}{1}, \ldots,
-    \alignment{\npairs{}}{\nloci{\npairs{}}})$
-is the full dataset,
-i.e., a vector of sequence alignments for all pairs and loci.
-We let \geneTree{i}{j} represent the gene tree upon which \alignment{i}{j}
-evolved according to fixed \hky substitution model parameters \hkyModel{i}{j}.
-The investigator must specify the parameters of all
-$\hkyModelVector = (\hkyModel{1}{1}, \ldots,
-\hkyModel{\npairs{}}{\nloci{\npairs{}}})$
-substitution models by which the alignments evolved along the
-$\geneTreeVector = (\geneTree{1}{1}, \ldots,
-\geneTree{\npairs{}}{\nloci{\npairs{}}})$
-gene trees.
-Furthermore, the investigator must specify a vector of fixed constants
-$\ploidyScalarVector = (\ploidyScalar{1}{1}, \ldots,
-\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$
-that scale the population-size parameters for known differences in ploidy among
-loci and/or differences in generation times among population pairs.  Lastly,
-the investigator must also specify a vector of fixed constants
-$\mutationRateScalarConstantVector = (\mutationRateScalarConstant{1}{1},
-\ldots, \mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$
-that scale the population-size parameters for known differences in
-mutation rates among loci and/or among taxa.
-
-\begin{linenomath}
-With \alignmentVector, \hkyModelVector, \ploidyScalarVector, and
-\mutationRateScalarConstantVector in hand, the joint posterior distribution
-of our model is given by
-\begin{equation}
-    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
-    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) =
-    \frac{p(\alignmentVector \given \geneTreeVector, \divTimeMapVector,
-        \demographicParamVector, \locusMutationRateScalarVector,
-        \locusRateHetShapeParameter, \hkyModelVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)
-        p(\geneTreeVector, \divTimeMapVector, \demographicParamVector,
-        \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-        \hkyModelVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)}{
-        p(\alignmentVector)},
-    \label{eq:fullModelCompactJoint}
-\end{equation}
-which can be expanded out into components that are assumed to be independent
-to get
-\begin{equation}
-    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
-    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) =
-    \frac{p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
-        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
-        \locusMutationRateScalarVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)
-        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
-        p(\locusRateHetShapeParameter)
-        p(\divTimeMapVector)
-        p(\demographicParamVector)}{
-        p(\alignmentVector)},
-    \label{eq:fullModelCompact}
-\end{equation}
-where
-$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
-is a vector of population divergence times for each of the \npairs{} pairs of
-populations,
-$\demographicParamVector = (\demographicParams{1}, \ldots,
-\demographicParams{\npairs{}})$
-is a vector of the demographic parameters for each of the \npairs{} population
-pairs,
-$\locusMutationRateScalarVector = (\locusMutationRateScalar{1}, \ldots
-\locusMutationRateScalar{\nlociTotal})$
-is a vector of locus-specific mutation-rate scaling parameters for each of the
-\nlociTotal loci,
-\locusRateHetShapeParameter is the shape parameter of a gamma-distributed
-prior on \locusMutationRateScalar{}, and
-$p(\alignmentVector)$ is the probability of the data (or the marginal
-likelihood of the model).
-\end{linenomath}
-
-\begin{linenomath}
-To avoid calculating the likelihood terms of Equation \ref{eq:fullModelCompact},
-we distill each sequence alignment \alignment{}{} into a vector of insufficient
-summary statistics \alignmentSS{}{}, thus replacing the full dataset
-$\alignmentVector = (\alignment{1}{1}, \ldots,
-    \alignment{\npairs{}}{\nloci{\npairs{}}})$
-with vectors of summary statistics for each alignment
-$\ssVectorObs = (\alignmentSSObs{1}{1}, \ldots,
-    \alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$.
-For each population pair, the means of the summary statistics are calculated
-across the \nloci{} loci, reducing the vector to
-$\ssVectorObs = (\alignmentSSObs{1}{}, \ldots,
-    \alignmentSSObs{\npairs{}}{})$.
-This allows us to estimate the approximate joint posterior distribution
-{\small
-\begin{equation}
-    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
-    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-    \ssSpace, \hkyModelVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) =
-    \frac{p(\ssSpace \given \geneTreeVector, \hkyModelVector)
-        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
-        \locusMutationRateScalarVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)
-        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
-        p(\locusRateHetShapeParameter)
-        p(\divTimeMapVector)
-        p(\demographicParamVector)}{
-        p(\ssSpace)},
-    \label{eq:approxModelCompact}
-\end{equation}
-}
-where \ssSpace is the multidimensional Euclidean space around the vector of
-summary statistics, the radius of which is the tolerance \tol.
-The sources of approximation are the insufficiency of the statistics and the
-\tol being greater than zero.
-We describe the full model in detail before delving into the numerical
-method of estimating the approximate model.
-\end{linenomath}
-
-\subsubsection*{Likelihood terms of Equation \ref{eq:fullModelCompact}}
-\begin{linenomath}
-The likelihood terms of Equation \ref{eq:fullModelCompact} can
-be expanded out as a product over population pairs and loci
-\begin{equation}
-    p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
-    p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
-    \locusMutationRateScalarVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) = 
-    \prod_{i=1}^{\npairs{}}
-    \prod_{j=1}^{\nloci{i}}
-    p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})
-    p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
-    \locusMutationRateScalar{j}, \ploidyScalar{i}{j},
-    \mutationRateScalarConstant{i}{j}).
-    \label{eq:modelLikelihoodExpanded}
-\end{equation}
-The first term,
-$p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})$,
-is the probability of the sequence alignment of locus $j$ for population pair
-$i$ given the gene tree and \hky substitution model parameters \citep[i.e., the
-``Felsenstein likelihood'';][]{Felsenstein1981}.
-The model allows for an intra-locus recombination rate \recombinationRate,
-which, for simplicity, is assumed to be zero in Equation
-\ref{eq:fullModelCompact}.
-If \recombinationRate is non-zero, this term requires an additional product
-over the columns (sites) of each sequence alignment to allow sites to have
-different genealogies.
-The second term,
-p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
-\locusMutationRateScalar{j}, \ploidyScalar{i}{j},
-\mutationRateScalarConstant{i}{j}),
-is the probability of the gene tree under a multi-population coalescent model
-(i.e., species tree) where the ancestral population of pair $i$ diverges
-and gives rise to the two sampled descendant populations.
-Each \demographicParams{} contains the following demographic parameters:
-The mutation-scaled sizes of the ancestral, \ancestralTheta{}, and descendant
-populations, \descendantTheta{1}{} and \descendantTheta{2}{};
-the proportion of the first, \bottleScalar{1}{}, and second population,
-\bottleScalar{2}{}, that persist during bottlenecks that begin immediately
-after divergence in forward-time;
-the proportion of time between present and divergence when the bottlenecks
-end for both populations, \bottleTime{};
-and the symmetric migration rate between the descendant populations,
-\migrationRate{}.
-Thus, the probability of the $\popSampleSize{i}{}-1$ coalescence times (node
-heights) of gene tree \geneTree{i}{j} is given by a multi-population
-Kingman-coalescent model \citep{Kingman1982} where the ancestral population of
-size
-$\ancestralTheta{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
-diverges at time \divTimeMap{i} into two descendant populations of constant
-size
-$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{1}{i}$
-and
-$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{2}{i}$,
-which, after time $\divTimeMap{i}\bottleTime{i}$, grow exponentially to their
-present size 
-$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
-and
-$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$,
-respectively.
-Following divergence, the descendant populations of pair $i$ exchange migrants at a symmetric rate of
-\migrationRate{i}.
-\end{linenomath}
-
-\subsubsection*{Prior terms of Equation \ref{eq:fullModelCompact}}
-\begin{linenomath}
-The term $p(\locusRateHetShapeParameter)$ is the prior density
-function for the shape-parameter of the gamma-distributed prior on
-rate heterogeneity among loci.
-This prior is $\locusRateHetShapeParameter \sim U(1, 20)$.
-The prior probability of the vector of locus-specific mutation-rate scaling parameters
-given \locusRateHetShapeParameter then expands out as a product over
-the loci
-\begin{equation}
-    p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter) =
-    \prod_{j=1}^{\nlociTotal}
-    p(\locusMutationRateScalar{j} \given \locusRateHetShapeParameter),
-    \label{eq:locusRateHetPrior}
-\end{equation}
-where each \locusMutationRateScalar{} is independently and identically
-distributed (\iid) as
-$\locusMutationRateScalar{} \sim Gamma(\locusRateHetShapeParameter,
-1/\locusRateHetShapeParameter)$.
-If the recombination rate \recombinationRate is allowed to be
-non-zero, the prior term $p(\recombinationRate)$ would be added
-to Equation \ref{eq:fullModelCompact}, and the prior would be
-$\recombinationRate \sim Gamma(\gammaShape{\recombinationRate},
-\gammaScale{\recombinationRate})$.
-\end{linenomath}
-
-\begin{linenomath}
-The prior term for the demographic parameters, $p(\demographicParamVector)$,
-expands out into its components and as a product over the \npairs{}
-pairs of populations
-\begin{equation}
-    p(\demographicParamVector) =
-    \prod_{i=1}^{\npairs{}}
-    p(\ancestralTheta{i})
-    p(\descendantTheta{1}{i})
-    p(\descendantTheta{2}{i})
-    p(\bottleScalar{1}{i})
-    p(\bottleScalar{2}{i})
-    p(\bottleTime{i})
-    p(\migrationRate{i}).
-    \label{eq:demographicPrior}
-\end{equation}
-The priors for the demographic parameters are:
-$\ancestralTheta{} \sim Gamma(\gammaShape{\ancestralTheta{}},
-\gammaScale{\ancestralTheta{}})$,
-$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$,
-$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$,
-$\bottleScalar{1}{} \sim Beta(\betaA{\bottleScalar{}{}},
-\betaB{\bottleScalar{}{}})$,
-$\bottleScalar{2}{} \sim Beta(\betaA{\bottleScalar{}{}},
-\betaB{\bottleScalar{}{}})$,
-$\bottleTime{} \sim U(0, 1)$,
-and
-$\migrationRate{} \sim Gamma(\gammaShape{\migrationRate{}},
-\gammaScale{\migrationRate{}})$.
-\end{linenomath}
-
-\subsubsection*{Priors on divergence models}
-\begin{linenomath}
-The prior term for the vector of divergence times for each of
-the \npairs{} pairs of populations can be expanded as
-\begin{equation}
-    p(\divTimeMapVector) = p(\divTimeIndexVector)p(\divTimeVector \given \divTimeIndexVector),
-    \label{eq:divModelPrior}
-\end{equation}
-where \divTimeVector is an ordered set of unique divergence time parameters
-$\{\divTime{1}, \ldots, \divTime{\divTimeNum}\}$ whose length
-\divTimeNum can range from 1 to \npairs{},
-and \divTimeIndexVector is the vector of indices of divergence time parameters 
-$(\divTimeIndex{1}, \ldots, \divTimeIndex{\npairs{}})$
-that map the divergence times in \divTimeVector to each of the \npairs{}
-pairs of populations.
-Thus, \divTimeMapVector is the result of applying the mapping function
-\begin{equation}
-    f(\divTimeVector, \divTimeIndexVector, i) = \divTime{\divTimeIndex{i}}
-    \label{eq:divTimeMapFunction}
-\end{equation}
-to each population pair $i$, such that
-$\divTimeMapVector = (\divTimeMap{1} = f(\divTimeVector, \divTimeIndexVector,
-1), \ldots, \divTimeMap{\npairs{}} = f(\divTimeVector, \divTimeIndexVector,
-\npairs{}))$.
-\end{linenomath}
-
-Biologically speaking, \divTimeVector contains the times of divergence events,
-the length of which \divTimeNum is the number of divergence events shared across
-the \npairs{} pairs of populations.
-For example, if \divTimeVector contains a single divergence time parameter
-\divTime{1}, all \npairs{} pairs of populations are constrained to diverge at
-this time (i.e., \divTimeIndexVector would contain the index 1 repeated
-\npairs{} times, and \divTimeMapVector would contain the value \divTime{1}
-repeated \npairs{} times), whereas if it contains \npairs{} divergence time
-parameters, the model is fully generalized to allow all of the pairs to diverge
-at unique times.
-
-\begin{linenomath}
-Each \divTime{} within \divTimeVector is \iid as $\divTime{} \sim
-Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
-Thus, given the number of unique divergence time classes in
-\divTimeIndexVector, this determines the probability of prior term
-$p(\divTimeVector \given \divTimeIndexVector)$.
-The divergence time parameters are in coalescent units relative to the size of
-a constant reference population, which we denote \myTheta{C}, that is equal to
-the expectation of the prior on the size of the descendant populations
-\begin{equation}
-    \myTheta{C} = \mathbb{E}(\descendantTheta{}{}).
-    \label{eq:thetaC}
-\end{equation}
-Given the size of the descendant populations are \iid as
-$\descendantTheta{}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$,
-this becomes
-\begin{equation}
-    \myTheta{C} = \gammaShape{\descendantTheta{}{}}\gammaScale{\descendantTheta{}{}}.
-    \label{eq:thetaCGamma}
-\end{equation}
-The \divTime{} parameters are in units of $\myTheta{C}/\mutationRate$
-generations, which we denote as \globalcoalunit generations.
-Thus, each \divTime{} within \divTimeVector is proportional to time and can be
-converted to the number of generations of the reference population, which we
-denote \divTime{G_C}, by assuming a mutation rate and multiplying by the
-effective size of the reference population
-\begin{equation}
-    \divTime{G_C} = \divTime \times \frac{
-        \gammaShape{\descendantTheta{}{}}\gammaScale{\descendantTheta{}{}}}
-        {\mutationRate}.
-    \label{eq:divTimeGenerations}
-\end{equation}
-Thus, for each of the divergence times in \divTimeVector to be on the same
-scale, the relative mutation rates among the pairs of populations are assumed
-to be fixed and known according to the user-provided values in
-\mutationRateScalarConstantVector.
-\end{linenomath}
-
-\begin{linenomath}
-As described by \citet{Oaks2012}, to get the divergence times in units
-proportional to the expected number of mutations, we must
-scale them by the realized population size for locus $j$ of population-pair $i$ 
-\begin{equation}
-    \divTimeScaled{i}{j} = \divTimeMap{i} \times \frac{\myTheta{C}}{
-        \descendantThetaMean{i} \ploidyScalar{i}{j}},
-    \label{eq:divTimeScaled}
-\end{equation}
-where \descendantThetaMean{i} is the mean of \descendantTheta{1}{} and
-\descendantTheta{2}{} for pair $i$.
-This gives us the vector of scaled divergence times
-$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
-\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$.
-\end{linenomath}
-
-\begin{linenomath}
-As for the prior term $p(\divTimeIndexVector)$, the total sample space of
-ordered realizations of \divTimeIndexVector is all the possible partitions of
-\npairs{} elements into 1 to \npairs{} categories.
-The total number of possible partitions is a sum of the Stirling numbers of
-the second kind over all possible numbers of categories \divTimeNum
-\begin{equation}
-    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
-    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
-    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
-    \label{eq:bell}
-\end{equation}
-which is the Bell Number \citep{Bell1934}.
-The original \msb model only samples over the unordered realizations of
-\divTimeIndexVector, such that the sample space is reduced to all the possible
-integer partitions of \npairs{} \citep{Oaks2012,Huang2011,OeisPartitionNumber,
-    OeisPartitionTriangle,Malenfant2011}.
-We denote the set of all possible integer partitions of \npairs{} as
-\integerPartitionSet{\npairs{}} and the length of that set as
-\integerPartitionNum{\npairs{}}.
-The advantages, disadvantages, and justification of ignoring the order
-of \divTimeIndexVector is discussed in detail below.
-\end{linenomath}
-
-\begin{linenomath}
-We implement two prior probability distributions over the space
-of all possible discrete divergence models (\divTimeIndexVector).
-The first simply gives all possible unordered partitions of \npairs{} elements
-equal probability
-\begin{equation}
-    p(\divTimeIndexVector) = \frac{1}{\integerPartitionNum{\npairs{}}},
-    \label{eq:divModelPriorUniform}
-\end{equation}
-i.e., a discrete uniform prior over all the integer partitions of \npairs{}
-(discrete divergence models).
-We denote this prior as
-$\divTimeIndexVector \sim \priorUniform$.
-\end{linenomath}
-
-The second prior we implement is based on the Dirichlet-process, which is a
-stochastic process that groups elements into an unknown number of discrete
-parameter classes \citep{Ferguson1973,Antoniak1974}.
-The process is controlled by the concentration parameter \concentrationParam,
-which determines how clustered the process will be.
-We use the Dirichlet process to place a prior over all possible ordered
-partitions of \npairs{} elements, which we denote as $\divTimeIndexVector \sim
-\priorDPP{}$.
-We take a hierarchical approach and use a prior probability
-distribution (i.e., hyperprior) for \concentrationParam.
-More specifically, we use a gamma-distributed prior $\concentrationParam \sim
-Gamma(\gammaShape{\concentrationParam}, \gammaScale{\concentrationParam})$,
-where \gammaShape{\concentrationParam} and \gammaScale{\concentrationParam} are
-specified by the user.
-
-\begin{linenomath}
-This provides a great deal of flexibility for specifying the prior uncertainty
-regarding divergence models.
-The \concentrationParam determines the prior probability that any two
-pairs of populations $i$ and $j$ will be assigned to the same divergence time
-parameter
-\begin{equation}
-    p(\divTimeIndex{i} = \divTimeIndex{j}) = \frac{1}{1 + \concentrationParam},
-    \label{eq:dppPriorSameClass}
-\end{equation}
-and also the prior probability of the number of divergence time parameters
-\begin{equation}
-    p(\divTimeNum \boldsymbol{\mid} \concentrationParam, \npairs{}) = 
-    \frac{\stirlingFirst{\npairs{}}{\divTimeNum} \concentrationParam^{\divTimeNum}}
-    {\prod_{i=1}^{\npairs{}}(\concentrationParam + i - 1)},
-    \label{eq:dppPriorNumClasses}
-\end{equation}
-where \stirlingFirst{\cdot}{\cdot} are the unsigned Stirling numbers of the
-first kind.
-Equations \ref{eq:dppPriorSameClass} and \ref{eq:dppPriorNumClasses} show that
-smaller values of \concentrationParam will favor fewer divergence time
-parameters, and thus more clustered models of divergence, whereas larger values
-will favor more divergence time parameters, and thus less clustered models of
-divergence.
-\end{linenomath}
-
-\subsection*{Differences between our model and the original \msb model}
-\subsubsection*{The prior on divergence models}
-One of the key differences between our model and that of \msb \citep{Huang2011}
-is the prior distribution on divergence models.
-As discussed in \citet{Oaks2012}, in \msb the prior used for
-\divTimeIndexVector is a combination of a discrete uniform prior over the
-possible number of divergence events \divTimeNum from 1 to \npairs{} with a
-multinomial distribution on the number of times each index of \divTimeVector
-appears in \divTimeIndexVector, with the constraint that all \divTime{}
-parameters are represented at least once (see Equation 2 of \citet{Oaks2012}).
-We denote this prior used by \msb as $\divTimeIndexVector \sim \priorOld$.
-\citet{Oaks2012} discuss how placing a uniform prior over the number of
-divergence parameters (denoted \divTimeNum here, and as \numt{} in
-\citet{Huang2011}) imposes an U-shaped prior over discrete divergence
-models (\divTimeIndexVector; see
-Figure 5B of \citet{Oaks2012}).
-To avoid this, we place priors directly on the sample space of divergence
-models, thus eliminating the parameter \numt{} from the model.
-We introduce two priors on divergence models:
-(1) a prior that is uniform over all unordered divergence models, and
-(2) a Dirichlet-process prior on all ordered divergence models.
-The latter provides the user with a great deal of flexibility in
-expressing their prior beliefs about models of divergence.
-
-\subsubsection*{Estimating ordered divergence models}
-As mentioned above, \msb samples over unordered divergence models
-(i.e., unordered partitions of the \npairs{} pairs of populations).
-That is, the identity of each population pair, and all the information
-associated with it, is discarded.
-In our implementation, inference can be done on either unordered or ordered
-models of divergence.
-This is discussed in more detail in the description of the ABC implementation
-below.
-
-\subsubsection*{The priors on nuisance parameters}
-Following the recommendations of \citet{Oaks2012}, we have replaced the use of
-continuous uniform distributions for priors on many of the
-model's parameters (\divTime{}, \ancestralTheta{}, \descendantTheta{1}{},
-\descendantTheta{2}{}, \bottleScalar{1}{}, \bottleScalar{2}{},
-\recombinationRate, \migrationRate{}) with more flexible parametric
-distributions from the exponential family.
-We introduce gamma-distributed priors for rate parameters that have a sample
-space of all positive real numbers (\divTime{}, \ancestralTheta{},
-\descendantTheta{1}{}, \descendantTheta{2}{}, \recombinationRate,
-\migrationRate{}), and beta-distributed priors for parameters that are
-proportions bounded by zero and one (\bottleScalar{1}{} and
-\bottleScalar{2}{}).
-These priors provide the user with much greater flexibility in expressing prior
-uncertainty regarding the parameters of the model.
-
-Another distinction between our model and that of \msb, is the prior on the
-sizes of the descendant populations of each pair.
-As described by \citet{Oaks2012}, \msb uses the joint prior
-\begin{equation}
-    \descendantTheta{1}{}, \descendantTheta{2}{} \sim
-    Beta(1,1) \times 2 \times U(\uniformMin{\myTheta{}},
-    \uniformMax{\descendantTheta{}{}}),
-    \label{eq:jointThetaPrior}
-\end{equation}
-such that the user-specified uniform prior on descendant population
-size is a prior on the \emph{mean} size of the two descendant
-populations of each pair.
-Under our model, the sizes of the descendant populations of each
-pair are \iid as
-$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$
-and
-$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$.
-This relaxes the assumption that the sizes of the two descendant populations
-are interdependent and negatively correlated.
-
-\subsubsection*{Flexibility in parameterizing the model}
-In our implementation, we provide the ability to control the richness of the
-model.
-For the \myTheta{} parameters, the model can be fully generalized to
-allow each population pair to have three parameters:
-\ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}.
-Furthermore, any model of \myTheta{} parameters nested within this
-general model can also be specified, including the most restricted model
-where the ancestral and descendant populations of each pair share
-a single \myTheta{} parameter.
-
-We also provide the option of eliminating the parameters associated with the
-post-divergence bottlenecks in the descendant populations of each pair
-(\bottleTime{}, \bottleScalar{1}{}, and \bottleScalar{2}{}),
-which constrains the descendant populations to be of
-constant size from present back to the divergence event.
-Also, rather than eliminate the bottleneck parameters,
-we allow \bottleScalar{1}{} and \bottleScalar{2}{} to be constrained to be
-equal, which removes one free parameter from the model for each of the
-population pairs.
-
-Overall, our implementation allows an investigator to specify a model that has
-as many as seven parameters per population pair
-(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{},
-\bottleTime{}, \bottleScalar{1}{}, \bottleScalar{2}{}, and
-\migrationRate{})
-or as few as one parameter per pair
-(\myTheta{}),
-in addition to the $\popSampleSize{i}{} - 1$ coalescence-time parameters.
-
-\subsubsection*{Time scale}
-As described above, in our model divergence times are in units of
-$\myTheta{C}/\mutationRate$ generations, where \myTheta{C} is the expectation
-of the prior on descendant-population size.
-As described by \citet{Oaks2012}, in \msb, \myTheta{C} is half of the upper
-limit of the continuous uniform prior on the mean of the descendant population
-sizes.
-
-\subsection*{ABC estimation of the posterior of the model}
-\subsubsection*{Sampling from the prior}
-To estimate the approximate posterior of Equation \ref{eq:approxModelCompact},
-we use an ABC rejection algorithm.
-The first step of this algorithm entails collecting a random sample of
-parameter values from the joint prior and their associated summary
-statistics.
-Each sample is generated by
-(1) drawing values of all the model's parameters, which we denote \hpvector{},
-from their respective prior distributions;
-(2) rescaling the divergence times
-$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
-from units proportional to time to units proportional to the expected number
-of mutations via Equation \ref{eq:divTimeScaled} to get 
-$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
-\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$;
-(3) simulating gene trees $\geneTreeVector = (\geneTree{1}{1}, \ldots, 
-\geneTree{\npairs{}}{\nloci{\npairs{}}})$
-for each locus of each population pair by drawing coalescent times from
-a multi-population Kingman-coalescent model given the demographic parameters;
-(4) simulating sequence alignments 
-$\alignmentVector = (\alignment{1}{1}, \ldots, \alignment{\npairs{}}{\nloci{\npairs{}}})$
-along the gene trees under the
-\hky substitution parameters
-$\hkyModelVector = (\hkyModel{1}{1}, \ldots, \hkyModel{\npairs{}}{\nloci{\npairs{}}})$
-that have the same number of sequences and sequence lengths as the observed
-dataset;
-(5) calculating population genetic summary statistics
-$\ssVector{} = (\alignmentSS{1}{1}, \ldots, \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$
-from the simulated sequence alignments;
-% (these are the same statistics
-% calculated from the observed alignments, \ssVectorObs{});
-and (6) reducing the summary statistics to the means across loci for each
-population pair to get
-$\ssVector{} = (\alignmentSS{1}{}, \ldots, \alignmentSS{\npairs{}}{})$, which is
-the same summary statistic vector estimated from the observed data \ssVectorObs.
-After repeating this procedure \numPriorSamples times, we obtain a random
-sample of parameter vectors
-$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$
-from the model prior and their associated vectors of summary statistics
-$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$.
-
-For all of our analyses below, we use four summary statistics for each of
-the pairs of populations:
-$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
-\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
-Furthermore, in addition to model parameters, each sample \hpvector{}
-also contains four statistics that summarize \divTimeMapVector:
-the mean (\divTimeMean), variance (\divTimeVar), dispersion index
-($\divTimeDispersion = \divTimeVar/\divTimeMean$), and the number
-of divergence time parameters (\divTimeNum).
-These have been denoted as \meant{}{}, \vart{}{}, \vmratio{}, and \numt{},
-respectively \citep{Hickerson2006,Huang2011,Oaks2012}.
-We use the new notation to avoid confusion that the values represent prior or
-posterior expectations, and to indicate that \vmratio{} is not a parameter of
-the model and \numt{} is no longer a parameter in the new implementation.
-
-\subsubsection*{Ordering of taxon-specific summary statistics}
-As alluded to in the model description, \msb does not maintain the order of the
-taxon-specific summary statistics \alignmentSS{}{} within each \ssVector{}.
-Rather, the summary statistics are re-ordered by descending values of average
-pairwise differences between the descendant populations
-\citep[$\pi_b$;][]{NeiLi1979,Huang2011}.
-This has the advantage of reducing the sample space of the number of discrete
-divergence models \divTimeMapVector, but there are at least two disadvantages.
-First, additional information in the data is lost.
-By discarding the identity of the \npairs{} pairs of populations, all
-pair-specific information about the amount of data (e.g., the number gene
-copies collected from each of the populations [\popSampleSize{1}{} and
-\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
-taxon- and locus-specific parameters (\hkyModel{}{},
-\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
-\locusMutationRateScalar{}) is lost.
-Second, the results are more difficult to interpret, because they can be no
-longer be directly associated to the taxa under study.
-
-The original descriptions of the \msb model claim that this re-ordering is
-justified by the fact that of $\pi_b$ (and other summary statistics) are
-unrelated to the sample sizes \popSampleSize{1}{} and \popSampleSize{2}{} of
-each pair and are thus exchangeable \citep{Hickerson2006,Huang2011}.
-This is actually incorrect for two of reasons.
-First, the expectation of $\pi_b$ is not independent of samples sizes.
-If there are more than one coalescent events in the ancestor, which is expected
-to be common on phylogeographic timescales, more samples will increase the
-probability of capturing these deeper events, and thus affect the average
-pairwise differences between the descendant populations.
-Also, other statistics that estimate gross diversity (e.g., $\pi$ and
-$\theta_W$) are clearly not independent of sample size.
-Second, for variables to be exchangeable, they do not need to be independent,
-but their marginal distributions must be the same (i.e., they must be
-identically distributed).
-The simulated alignments and their summary statistics are not identically
-distributed because of differing sample sizes \emph{and} taxon- and
-locus-specific parameters \hkyModel{}{}, \mutationRateScalarConstant{}{}, and
-\ploidyScalar{}{}.
-
-Thus, the theoretical basis for this reshuffling is questionable.
-It may be justifiable as an additional coarsening of the data,
-by removing the information associated with the identity of the
-population pairs.
-However, we do not provide a formal proof that this re-ordering of the pairs
-does not introduce bias.
-One can imagine situations in which the sampling intensity (i.e., the
-number of gene copies, loci, and locus length) is highly skewed across
-pairs and/or there is large heterogeneity in mutational processes
-(e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
-In such cases, it seems possible that discarding this information could
-cause bias.
-Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
-is to minimize Euclidean distances between simulated datasets in which the true
-model has a single divergence event, and (2) the reordering increases the
-model's tendency to infer a single divergence event \citep{Huang2011}, it seems
-possible that this approach could be biasing the method \citep{Oaks2012}.
-
-To maintain compatibility and comparability with \msb we maintain
-the re-ordering of taxon-specific summary statistics by $\pi_b$.
-However, we also allow the order to be maintained, and ordered
-divergence models to be estimated.
-
-\subsubsection*{Obtaining an approximate posterior from the prior samples}
-We use a rejection algorithm to retain an approximate posterior sample of
-\paramSampleVector{} from the prior sample
-$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$.
-First, the observed summary statistics \ssVectorObs, and the summary statistics
-of the prior samples 
-$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$,
-are standardized using the means and standard deviations of the statistics from
-the prior sample (i.e., the prior mean is subtracted from each statistic, and the
-difference is divided by the prior standard deviation).
-After all statistics are standardized, the Euclidean distance between
-\ssVectorObs and each \ssVector{} within \ssMatrix is calculated.
-The samples that fall within a range of tolerance \tol around \ssVectorObs
-are retained.
-The range of tolerance is determined by specifying the number of desired
-posterior samples to be retained.
-Post-hoc adjustment of the posterior sample can also be performed with a number
-of regression techniques \citep{Beaumont2002,Blum2009,Leuenberger2010}.
-For our analyses, we use the general linear model (GLM) regression adjustment
-\citep{Leuenberger2010} as implemented in \abctoolbox
-\citep[v1.1;][]{ABCtoolbox}, which \citet{Oaks2012} showed performed very
-similarly to weighted local-linear regression and multinomial logistic
-regression adjustments \citep{Beaumont2002} for \msb posteriors.
-
-\subsection*{Assessing model-choice behavior and robustness}
-Following the simulation-based approach of \citet{Oaks2012}, we characterize
-the behavior of several models under the ideal conditions where the
-data are generated from parameters drawn from the same prior distributions used
-for analysis (i.e., the prior is correct).
-We selected four model priors for these analyses (Table~\ref{tabPriors}):
-(1) A model to represent the original \msb model, \modelOld, with
-priors
-$\divTimeIndexVector \sim \priorOld$,
-$\divTime{} \sim U(0,10)$,
-$\ancestralTheta{} \sim U(0, 0.05)$,
-and
-$\descendantThetaMean{} \sim U(0, 0.05)$;
-(2) a Dirichlet-process prior model, \modelDPP, with priors
-$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
-and
-$\descendantTheta{}{} \sim Exp(40)$;
-(3) a model with uniform prior probability over unordered divergence models,
-\modelUniform, with priors
-$\divTimeIndexVector \sim \priorUniform$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
-and
-$\descendantTheta{}{} \sim Exp(40)$;
-and
-(4) a model with the \msb prior on unordered divergence models, but with
-exponential priors on nuisance parameters,
-\modelUshaped, with priors
-$\divTimeIndexVector \sim \priorOld$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
-and
-$\descendantTheta{}{} \sim Exp(40)$.
-We selected the exponential prior on divergence time used in models \modelDPP,
-\modelUniform, and \modelUshaped to have the same variance as the uniform prior
-in model \modelOld.
-We selected the exponential prior on population size used in models \modelDPP,
-\modelUniform, and \modelUniform to have the same mean as the uniform prior in
-model \modelOld, so that all four models have the same \myTheta{C} and thus the
-same units of time.
-All of the models were the same in other respects, with three free \myTheta{}
-parameters for each population pair, two uniformly distributed ($beta(1,1)$)
-\bottleScalar{}{} parameters per pair, no migration, no recombination,
-and unordered divergence models.
-For all of our simulations, we used a simulated data structure of eight
-population pairs, with a single 1000 base-pair locus sampled from 10
-individuals from each population.
-
-For each of the four models, we generated $1\e6$ samples from the prior, and
-simulated 50,000 pseudo-replicate datasets, also drawn from the prior.
-We then analyzed each of the replicate datasets, retaining a
-posterior of 1000 samples from the respective prior.
-A GLM-regression adjusted posterior was also estimated from each of the
-posterior samples \citep{Leuenberger2010}.
-To assess the robustness of each of the four models, we also analyzed the
-pseudo-replicate datasets simulated under the other three models.
-Overall, for each model, we produced 200,000 posterior estimates,
-50,000 from the replicate datasets simulated under that model,
-and 150,000 estimated from the replicate datasets simulated under the
-other three models.
-
-For each set of 50,000 simulated datasets, we used the posterior estimates
-to assess the model-choice behavior of each model.
-We did this by assigning the 50,000 estimates of the posterior probability
-of one divergence event to 20 bins of width 0.05, and plotted
-the estimated posterior probability of each bin against the proportion of
-replicates in that bin with a true value consistent with one divergence
-event \citep{Huelsenbeck2004,Oaks2012}.
-We did this using two criteria for the one divergence model:
-(1) the number of divergence time parameters ($\divTimeNum = 1$) and
-(2) the dispersion index of divergence times \citep[$\divTimeDispersion <
-0.01$][]{Hickerson2006,Huang2011}.
-We used the one-divergence model to assess model-choice behavior, because
-it is often of biogeographic interest and is easily comparable among
-the three different priors used on divergence models.
-
-In addition to the four models above, we also assessed the behavior of a model
-with ordered divergence models (i.e., the order of the taxon-specific summary
-statistic vectors were maintained for the observed and simulated datasets).
-For this, we used a model with identical priors as \modelDPP, but that samples
-over ordered divergence models.
-We denote this model as \modelDPPOrdered.
-We generated $1\e6$ prior samples and 50,000 pseudo-replicate datasets, and
-analyzed them as above.
-We were not able to analyze the replicate datasets of the other models under
-the ordered model, because the identity of the population pairs is not
-contained in the simulations of the other models.
-
-\subsection*{Assessing power}
-We evaluated the power of the same four models (Table~\ref{tabPriors}) to
-detect random variation in divergence times using methods similar to
-\citet{Oaks2012}.
-The only difference from the prior models used in the validation analyses above
-is the prior on the concentration parameter \concentrationParam for the
-\modelDPP model.
-Rather than the prior
-$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
-we use the prior
-$\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
-over the discrete divergence models for the model \modelDPP.
-The reason for this difference is because for all of our power simulations, we
-used a simulated data structure identical to that of the empirical dataset of
-Philippine vertebrates analyzed by \citet{Oaks2012}, which consists of 22 pairs
-of populations.
-For each of the four models, we generated $2\e6$ samples from the prior.
-We then generated pseudo-observed datasets from three series of models.
-One series of models, which we denote \powerSeriesOld, is identically
-distributed as \modelOld except that the divergence times for each of the 22
-pairs of populations are randomly drawn from a series of uniform distributions,
-$U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and
-2.0, in \globalcoalunit generations.
-A second series of models, \powerSeriesUniform, are identically distributed as
-\modelUniform except that the 22 divergence times are randomly drawn from the
-same series of uniform priors as above.
-The third series of models, \powerSeriesExp, is also identically
-distributed as \modelUniform except that the 22 divergence times are randomly
-drawn from a series of of exponential distributions: $Exp(17.3205)$,
-$Exp(8.6603)$, $Exp(5.7735)$, $Exp(4.3301)$, $Exp(3.4641)$, and $Exp(1.7321)$.
-These exponential distributions have the same variance as their uniform
-counterparts in the first two series of models.
-
-For each of the six models in each of the three series of models, we simulated
-1000 pseudo-replicate datasets (18,000 datasets in total).
-We then analyzed each replicate dataset under all four prior models
-(Table~\ref{tabPriors}), producing 72,000 posterior estimates, each with 1000
-samples.
-We also estimated a GLM-regression adjusted posterior from each of the
-posterior samples \citep{Leuenberger2010}.
-
-\section*{Results}
-\subsection*{Validation analyses: Estimation accuracy}
-In terms of estimating summary statistics of model parameters (\divTimeNum,
-\divTimeDispersion, and \divTimeMean), the \modelDPP performs similarly when
-applied to datasets generated under all four models of Table \ref{tabPriors}
-(Figs.\
-\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP}).
-The same is true for models \modelUniform (Figs.\
-\labelcref{figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform})
-and \modelUshaped (Figs.\
-\labelcref{figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped}).
-All three of the models with exponential priors on nuisance parameters
-(\modelDPP, \modelUniform, and \modelUshaped) perform well when applied
-to the data generated under the original \msb model, \modelOld (Figs.\
-\labelcref{figValidationAccuracyOldDPP,figValidationAccuracyOldUniform,figValidationAccuracyOldUshaped}).
-The \modelOld model also performed well on its own datasets (Figure
-\ref{figValidationAccuracyOldOld}), but performed poorly when applied to the
-data generated under the other three models (Figs.\
-\labelcref{figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld}).
-Thus, the newly implemented models display greater robustness than
-the original model.
-
-Posterior adjustment with GLM regression does improve the estimation
-accuracy for the continuous variables \divTimeDispersion and \divTimeMean,
-however, this is not always true for estimating the discrete variable
-\divTimeNum (Figs.\
-\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP,figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform,figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped,figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld,figValidationAccuracyOldOld}).
-
-\subsection*{Validation analyses: Model-choice accuracy}
-The \msb model, and our modification of it, is a model-choice method
-with the primary purpose of estimating the probabilities of models
-of divergence across taxa.
-Thus, it is critical to assess the method's ability to accurately
-estimate the posterior probabilities of divergence models.
-Our results demonstrate that the unadjusted estimates of divergence models are
-generally more accurate than the regression-adjusted estimates, especially when
-using $\divTimeDispersion < 0.01$ to assess the probability of the one
-divergence model (Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
-Thus, we will focus our discussion of the results on the unadjusted estimates.
-
-The model-choice behavior of the \modelDPP is relatively accurate when applied
-to datasets generated under models \modelDPP, \modelUniform, and \modelOld
-(Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP}).
-It is less accurate and tends to underestimate the probability of the
-one-divergence model when applied to the data generated under \modelUshaped
-(Figure \ref{figValidationModelChoiceUshapedDPP}).
-The \modelUniform model performs almost identically as \modelDPP in
-terms of model-choice accuracy (Figs.\
-\labelcref{figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}).
-
-The original \msb model \modelOld performs reasonably well when applied
-to its own datasets (Figure \ref{figValidationModelChoiceOldOld}), but
-it performs poorly when applied to data generated under any of the other
-models (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
-In these cases the \modelOld model is strongly biased towards over-estimating
-the posterior probability of the one-divergence model, especially when the
-divergence models are not distributed under its U-shaped prior (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld}).
-The other model with the U-shaped prior on divergence models, but exponential
-priors on nuisance parameters (\modelUshaped), performs similarly to \modelOld
-in that it performs well when applied to its own data, but overestimates
-the probability of the one-divergence model when applied to data generated
-by the other models (Figs.\
-\labelcref{figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceUshapedUshaped}).
-However, the bias toward over-estimating the probability of the one-divergence
-model is stronger in the \modelOld model than \modelUshaped.
-
-Overall, the result suggest that the \modelDPP and \modelUniform models are
-relatively robust in terms of model-choice accuracy, and when model violations
-do cause them to be biased, they tend to under-estimate the probability of
-the biogeographically interesting model of a single divergence event.
-In contrast, the \modelOld model is very sensitive to model violations,
-and strongly over-estimates the one-divergence model whenever the model
-is misspecified.
-Furthermore, the results suggest that using exponentially distributed priors on
-nuisance parameters rather than uniform priors helps the \modelUshaped model
-perform better than \modelOld, but it is still hindered by the U-shaped prior
-on divergence models and tends to over-estimate the probability of the
-one-divergence model whenever there are violations of the model.
-
-\subsection*{Validation analyses: Ordered divergence models}
-Our results show that the ordered models of divergence, \modelDPPOrdered,
-perform similarly to sampling over unordered models (\modelDPP) when the prior
-is correct (Figs.\
-\labelcref{figValidationAccuracyDPPOrderedDPPOrdered,figValidationModelChoiceDPPOrderedDPPOrdered}).
-This suggests that the method is not adversely affected by the large increase
-in the number of possible discrete models (from 22 unordered to 4140 ordered
-models) when there are eight pairs of populations.
-
-\subsection*{Power analyses: Estimation accuracy}
-All of the models we evaluated (Table \ref{tabPriors}) struggle to estimate the
-dispersion index of divergence times \divTimeDispersion regardless of which of
-the three series of models (Table \ref{tabPowerModels}) the data were generated
-under (Figs.\
-\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform,figPowerAccuracyExpOld,figPowerAccuracyExpUshaped,figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped,figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
-When the divergence times of the 22 population pairs are randomly drawn from a
-series of exponential priors (\powerSeriesExp), the \modelDPP model is the
-best estimator of \divTimeDispersion, followed by \modelUniform (figs.\
-\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform}).
-The \modelOld model is strongly biased toward underestimating
-\divTimeDispersion, estimating values of zero for most of the replicates across
-all the data models of \powerSeriesExp (Figure \ref{figPowerAccuracyExpOld}).
-The results of the \modelUshaped model are intermediate between those of
-\modelOld and the new models \modelDPP and \modelUniform
-(Figure \ref{figPowerAccuracyExpUshaped}).
-
-When the true divergence times are randomly drawn from a series of uniform
-priors (\powerSeriesUniform), all of the models poorly estimate
-\divTimeDispersion (Figs.\
-\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped}).
-The \modelDPP and \modelUniform models tend to over-estimate the variance in
-divergence times (Figs.\
-\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform}), whereas
-the \modelOld model, again, underestimates \divTimeDispersion, estimating
-values of zero for most replicates across all the data models of
-\powerSeriesUniform (Figure \ref{figPowerAccuracyUniformOld}).
-Again, the performance of the \modelUshaped model is intermediate between the
-\modelOld and \modelDPP/\modelUniform models (Figure
-\ref{figPowerAccuracyUniformUshaped}).
-This pattern of results across the four models is very similar to their
-behavior when applied to data simulated under the series of models of
-\powerSeriesOld (Figs.\
-\labelcref{figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
-
-\subsection*{Power analyses: Model choice}
-There is a large amount of heterogeneity in estimating the number of divergence
-events (\divTimeNum) both among the prior models, and across the different
-series of data models (Table \ref{tabPowerModels}) (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped}).
-The \modelDPP model tends to infer highly clustered divergence models across
-all three series of data models when the divergences are recent, but performs
-the best of the four prior models  when divergences are random over a timescale
-of about one to two coalescent units (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiUniformDPP,figPowerPsiOldDPP,}).
-The \modelUniform model performs the best when divergences are random
-over the most recent timescales, but tends to only infer a tight
-cluster of \divTimeNum values even when divergences are random
-over longer periods (Figs.\
-\labelcref{figPowerPsiExpUniform,figPowerPsiUniformUniform,figPowerPsiOldUniform,}).
-The \modelOld model performs the worst of the four prior models across
-all three series of data models, inferring a single divergence event across
-most of the 18,000 simulations (Figs.\
-\labelcref{figPowerPsiExpOld,figPowerPsiUniformOld,figPowerPsiOldOld,}).
-Even when the simulated data are identically distributed as the \modelOld model
-(except for the divergence times) the \modelOld is still strongly biased
-towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
-model is the least likely to infer a single event when applied to these data
-(Figure \ref{figPowerPsiOldDPP}).
-Using exponential priors on nuisance parameters does increase the power
-of the \modelUshaped model compared to \modelOld across all three series
-of data models, but the U-shaped prior still prevents the model
-from performing as well as the \modelDPP and \modelUniform models (Figs.\
-\labelcref{figPowerPsiExpUshaped,figPowerPsiUniformUshaped,figPowerPsiOldUshaped,}).
-
-The improved power of the new models are even more pronounced when looking at
-estimates of the dispersion index of divergence times (\divTimeDispersion)
-across the simulations (Figs.\
-\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped}).
-The \modelDPP and \modelUniform models perform similarly across all three
-series of data models, inferring values of \divTimeDispersion consistent with
-one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
-replicates across all the simulations (Figs.\
-\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionOldDPP,figPowerDispersionOldUniform}).
-The \modelOld model infers values consistent with a single divergence event in
-most of the replicates across all the simulations (Figs.\
-\labelcref{figPowerDispersionExpOld,figPowerDispersionUniformOld,figPowerDispersionOldOld}).
-Using exponential priors on nuisance parameters greatly increases the power
-of the \modelUshaped model relative to \modelOld, but it still has
-less power than the models with DPP or uniform priors across
-divergence models (Figs.\
-\labelcref{figPowerDispersionExpUshaped,figPowerDispersionUniformUshaped,figPowerDispersionOldUshaped}).
-
-When looking at the estimated posterior probability of the single-divergence
-model across the power analyses, we also see increased power of the new models
-(Figs.\
-\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped}).
-The \modelDPP model estimates low posterior probability of
-$\divTimeNum = 1$ across all of the simulations (Figs.\
-\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbUniformDPP,figPowerPsiProbOldDPP}).
-The \modelDPP does infer strong support for one divergence as gauged by a Bayes
-factor of greater than 10.
-However, this is not that unexpected given the low prior probability of a
-single divergence under this model.
-Also, the Bayes factor calculations for the \modelDPP replicates are only
-approximate; they are calculated based on the prior mean of the concentration
-parameter \concentrationParam.
-
-The \modelUniform model infers very low posterior probabilities for the
-one-divergence model across all the simulations (Figs.\
-\labelcref{figPowerPsiProbExpUniform,figPowerPsiProbUniformUniform,figPowerPsiProbOldUniform}),
-whereas the \modelOld model infers high posterior probabilities of a single
-divergence for most replicates across all simulations (Figs.\
-\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld}).
-Again, we see that the exponential priors on nuisance parameters greatly
-increase power and result in lower estimates of the probability of
-one divergence when comparing \modelUshaped to \modelOld (Figs.\
-\labelcref{figPowerPsiProbExpUshaped,figPowerPsiProbUniformUshaped,figPowerPsiProbOldUshaped}).
-
-Lastly, when we look at the estimated posterior probability of
-\divTimeDispersion being consistent with one divergence parameter
-($p(\divTimeDispersion < 0.01 | \ssSpace)$), we see the same pattern of model
-behavior, with \modelDPP and \modelUniform inferring low probabilities across
-all simulations, \modelOld inferring high probabilities, and \modelUshaped
-inferring intermediate values (Figs.\
-\labelcref{figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
-
-\section*{Discussion}
-Our results demonstrate that using alternative priors on nuisance parameters
-and divergence models improved the behavior of the \msb model.
-In our implementation, model-choice estimation is more accurate and shows
-greater robustness to model violations (Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
-The original model is very sensitive to model violations and, when present,
-strongly over-estimates the probability of one divergence event shared
-across all taxa (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
-When more appropriate priors are used for nuisance parameters, and either
-a Dirichlet-process or uniform prior applied across divergence models,
-the model is less sensitive to violations (Figs.\
-\labelcref{figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}),
-and, when they do cause bias, the method is conservative and tends to
-under-estimate the probability of biogeographically interesting models of
-temporally clustered divergences (Figs.\
-\labelcref{figValidationModelChoiceUshapedDPP,figValidationModelChoiceUshapedUniform}).
-
-Furthermore, the modifications to the model also improve the method's power to
-detect random variation in divergence times, reducing the bias toward
-estimating clustered divergences (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped,figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped,figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped,figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
-Our results are similar to those of \citet{Oaks2012} in that we also find
-the \msb model to infer strong support for clustered divergences when
-divergences are random over quite broad timescales (Figs.\
-\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld,figPowerDispersionProbExpOld,figPowerDispersionProbUniformOld,figPowerDispersionProbOldOld}).
-Our results expand on this by showing that this behavior is consistent
-across a range conditions underlying the data.
-When using the new priors, the method is less prone to spurious inference of
-clustered divergence models, and much less likely to incorrectly infer such
-models with strong support.
-
-By evaluating a model intermediate between the old and new implementation
-(\modelUshaped), we were able to determine the relative effects of the
-modifications to model.
-Across all of our analyses, the results show that using better priors on
-nuisance parameters alone does improve the performance of the method.
-The magnitude of the bias toward inferring support for the one-divergence
-model when there are model violations is reduced when the exponential
-priors are used in place of the uniform priors (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped}).
-Furthermore, better priors on nuisance parameters improves the method's
-power to detect temporally random divergences.
-Throughout our analyses, the intermediate model (\modelUshaped) performs
-better than the \msb model, but not as well as the models with alternative
-priors on divergence models.
-This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
-models of temporally clustered divergences is caused by a combination of
-(1) uniform priors on nuisance parameters hindering the marginal likelihoods of
-models with more \divTime{} parameters and
-(2) the U-shaped prior on divergence models giving low prior density to models
-with intermediate numbers of divergence parameters.
-
-While our modifications improve the behavior of the model, we urge
-caution when using the method and interpreting its results.
-The method attempts to approximate the posterior of a very parameter rich model
-using relatively little information from the data.
-For example, when applied to the dataset of Philippine vertebrates analyzed by
-\citet{Oaks2012} the model can have as many as 604--625 free parameters
-(depending on \divTimeNum), and samples over 1002 divergence models.
-Our modifications allow this model to be simplified to 471--492 free
-parameters.
-Sampling over hundreds of parameters and more than a thousand models is a
-difficult inference problem, especially when only using a small number of
-summary statistics calculated from the sequence alignments of each taxon pair.
-The population-genetic summary statistics used by our method contain little
-information about many of the free parameters in the model.
-Thus, we expect the improved method will still be sensitive to priors, and the
-power, while improved, may still be low.
-Furthermore, ABC methods in general have been shown to be biased
-\citep{Robert2011} for model choice.
-As recommended by \citet{Oaks2012}, any results from the method should be
-accompanied by assessments of prior sensitivity and power.
-We encourage investigators to view this method and its predecessor as a means
-of exploring their data rather than a rigorous means of evaluating hypotheses.
-
-Given the difficulty of this estimation problem, we anticipate that
-full-likelihood methods that can leverage all of the information present in the
-sequence data will become increasingly important for robustly estimating the
-temporal history of divergences across taxa \citep{JeetDiss}.
-With improving numerical methods for sampling over models of differing
-dimensionality \citep{Green1995,Lemey2009} and increasing efficiency of
-likelihood calculations \citep{Ayres2012}, analyzing rich comparative
-phylogeograpical models in a full-likelihood Bayesian framework is becoming
-computationally practical, especially when considering that the computational
-load of simulating millions of random datasets from the prior under the ABC
-approach is nontrivial.
-
-\section*{Acknowledgments}
-We thank the National Science Foundation for supporting this work (DEB
-1011423).
-J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
-Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
-Society, KU Department of Ecology and Evolutionary Biology, and the KU
-Biodiversity Institute.
-
-\bibliography{../bib/references}
-
-%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\renewcommand\listfigurename{Figure Captions}
-\cftsetindents{fig}{0cm}{2.2cm}
-\renewcommand\cftdotsep{\cftnodots}
-\setlength\cftbeforefigskip{10pt}
-\cftpagenumbersoff{fig}
-\listoffigures
-
-\end{linenumbers}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
 %:TABLE-notation descriptions
@@
 \end{table}
-
-\clearpage
-
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
-%% Validation accuracy
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelDPP}
-}{figValidationAccuracyDPPDPP}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelUniform}
-}{figValidationAccuracyDPPUniform}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelOld}
-}{figValidationAccuracyDPPOld}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelUshaped}
-}{figValidationAccuracyDPPUshaped}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelDPP}
-}{figValidationAccuracyUniformDPP}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelUniform}
-}{figValidationAccuracyUniformUniform}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelOld}
-}{figValidationAccuracyUniformOld}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelUshaped}
-}{figValidationAccuracyUniformUshaped}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelDPP}
-}{figValidationAccuracyOldDPP}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelUniform}
-}{figValidationAccuracyOldUniform}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelOld}
-}{figValidationAccuracyOldOld}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelUshaped}
-}{figValidationAccuracyOldUshaped}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelDPP}
-}{figValidationAccuracyUshapedDPP}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelUniform}
-}{figValidationAccuracyUshapedUniform}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelOld}
-}{figValidationAccuracyUshapedOld}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelUshaped}
-}{figValidationAccuracyUshapedUshaped}
-
-%% Validation model choice
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelDPP}
-}{figValidationModelChoiceDPPDPP}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelUniform}
-}{figValidationModelChoiceDPPUniform}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelOld}
-}{figValidationModelChoiceDPPOld}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelUshaped}
-}{figValidationModelChoiceDPPUshaped}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelDPP}
-}{figValidationModelChoiceUniformDPP}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelUniform}
-}{figValidationModelChoiceUniformUniform}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelOld}
-}{figValidationModelChoiceUniformOld}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelUshaped}
-}{figValidationModelChoiceUniformUshaped}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelDPP}
-}{figValidationModelChoiceOldDPP}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelUniform}
-}{figValidationModelChoiceOldUniform}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelOld}
-}{figValidationModelChoiceOldOld}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelUshaped}
-}{figValidationModelChoiceOldUshaped}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelDPP}
-}{figValidationModelChoiceUshapedDPP}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelUniform}
-}{figValidationModelChoiceUshapedUniform}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelOld}
-}{figValidationModelChoiceUshapedOld}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelUshaped}
-}{figValidationModelChoiceUshapedUshaped}
-
-%% Validation ordered models
-\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPPOrdered}{\modelDPPOrdered}
-}{figValidationAccuracyDPPOrderedDPPOrdered}
-
-\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPPOrdered}{\modelDPPOrdered}
-}{figValidationModelChoiceDPPOrderedDPPOrdered}
-
-%% Power accuracy results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerAccuracyExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerAccuracyExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelOld}
-}{figPowerAccuracyExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerAccuracyExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerAccuracyUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerAccuracyUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerAccuracyUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerAccuracyUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerAccuracyOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerAccuracyOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelOld}
-}{figPowerAccuracyOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerAccuracyOldUshaped}
-
-%% Power psi results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerPsiExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerPsiExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelOld}
-}{figPowerPsiExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerPsiExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerPsiUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerPsiUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerPsiUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerPsiUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerPsiOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerPsiOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelOld}
-}{figPowerPsiOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerPsiOldUshaped}
-
-%% Power dispersion results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerDispersionExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerDispersionExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelOld}
-}{figPowerDispersionExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerDispersionExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerDispersionUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerDispersionUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerDispersionUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerDispersionUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerDispersionOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerDispersionOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelOld}
-}{figPowerDispersionOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerDispersionOldUshaped}
-
-%% Power psi probability results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerPsiProbExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerPsiProbExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelOld}
-}{figPowerPsiProbExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerPsiProbExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerPsiProbUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerPsiProbUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerPsiProbUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerPsiProbUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerPsiProbOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerPsiProbOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelOld}
-}{figPowerPsiProbOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerPsiProbOldUshaped}
-
-%% Power dispersion prob results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerDispersionProbExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerDispersionProbExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelOld}
-}{figPowerDispersionProbExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerDispersionProbExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerDispersionProbUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerDispersionProbUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerDispersionProbUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerDispersionProbUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerDispersionProbOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerDispersionProbOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelOld}
-}{figPowerDispersionProbOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerDispersionProbOldUshaped}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-
-\singlespacing
-
-\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape Supporting Information}}
-
-% PUT MAIN TEXT CITATION HERE
-% \begin{thebibliography}{1}
-% \providecommand{\natexlab}[1]{#1}
-% \providecommand{\url}[1]{\texttt{#1}}
-% \providecommand{\urlprefix}{URL }
-
-% \bibitem
-
-% \end{thebibliography}
-
-%% SUPPL TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\setcounter{table}{0}
-
-\end{document}

### 2013-10-18 manuscripts/dpp-msbayes/dpp-msbayes.tex
Breakin up `dpp-msbayes.tex`.
Moving the main text, tables, and figures of `dpp-msbayes.tex` to
`dpp-msbayes-main.tex`, `dpp-msbayes-tables.tex`, and
`dpp-msbayes-figures.tex`.

--- 2013-10-01 manuscripts/dpp-msbayes/dpp-msbayes.tex
@@
 
-\section*{Introduction}
-Understanding the processes that generate biodiversity and regulate
-community assembly is a major goal of evolutionary biology.
-Given the dynamic nature of our planet, a major component of this goal is
-understanding how geological and climatic processes can affect the
-evolutionary history of whole groups of co-distributed species and their
-associated microbiota, including pathogens.
-Such large-scale processes predict patterns of shared evolutionary history
-across effected populations of organisms, e.g., temporally clustered
-divergences across co-distributed species.
-At recent timescales, such events should leave a signature in the genetic
-variation within and among the affected lineages.
-Thus, being able to accurately estimate models of shared evolutionary
-histories across co-distributed taxa would provide an important
-tool for better understanding how regional and global biogeographic
-processes affect biodiversity.
-
-This inference problem is challenging due to the stochastic nature by which
-mutations occur in populations and how they are inherited over generations
-\citep{Hudson1990,WakeleyCoalescent}.
-Thus, a method for estimating historical patterns of divergences across taxa
-should explicitly model the stochastic mutational and ancestral processes that
-generate and filter the genetic variation we observe in present-day genetic
-data.
-An appealing approach would be a comparative, Bayesian model-choice method for
-inferring the probability of competing divergence histories while integrating
-over uncertainty in mutational and ancestral processes via models of nucleotide
-substitution and lineage coalescence.
-
-The software package \msb implements such an approach in an
-approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
-Recently, however, \citet{Oaks2012} found that \msb can be strongly biased
-toward spuriously estimating biogeographically interesting models of highly
-clustered divergence times across taxa.
-The results of their simulation-based assessment of the method suggest that a
-combination of broad, uniformly distributed priors on nuisance parameters and a
-U-shaped prior on divergence models were causing the bias.
-The model implemented in \msb uses a discrete uniform prior over the number of
-possible divergence parameters, which \citet{Oaks2012} show imposes a prior on
-divergence models that puts most of the prior mass on models with either very
-few or very many divergence times shared across the taxa.
-Furthermore, uniform priors used for most of the model's nuisance parameters
-likely hinder the marginal likelihoods of parameter-rich models, i.e., models
-with many divergence time parameters \citep[this is often referred to as
-Lindley's paradox;][]{Lindley1957}.
-Thus, together, the priors on divergence models and nuisance parameters used in
-\msb could create a strong bias toward models with very few divergence time
-parameters \citep{Oaks2012}.
-
-Following the recommendations of \citet{Oaks2012} we modify the \msb model to
-test whether alternative parameterizations and priors will improve the behavior
-of the approximate Bayesian model-choice approach.
-We implement a Dirichlet-process prior over all possible models of divergence,
-and use gamma and beta probability distributions in place of uniform priors on
-many of the model's parameters.
-Our simulation-based analyses show that the new implementation has improved
-robustness, accuracy, and power compared to the original model.
-Also, our results show that both the uniform priors on nuisance parameters and
-the U-shaped prior on divergence models contribute to the biases of \msb.
-
-\section*{Methods}
-
-\subsection*{The model}
-In this section, we describe our model, which is a modification of the model
-implemented in \msb \citep{Huang2011,Oaks2012}.
-The code implementing the new model is freely available in the open-source
-software package \dppmsbayes
-(\href{https://github.com/joaks1/dpp-msbayes}{https://github.com/joaks1/dpp-msbayes}).
-Our analyses described below utilize the freely avaliable, open-source software
-package \pymsbayes (\href{https://github.com/joaks1/PyMsBayes}{https://github.com/joaks1/PyMsBayes}),
-which is a Python API that provides a multi-processing interface to \msb and
-\dppmsbayes.
-We performed the work described below following the principles of
-Open Notebook Science.
-Using version-control software, we make progress in all aspects of the work
-freely and publicly available in real-time at
-\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
-All information necessary to reproduce our results is provided there.
-We follow much of the notation of \citet{Oaks2012}, but modify it to
-aid in our description of the new model.
-A summary of our notation can be found in Table \ref{tabNotation}.
-
-We assume an investigator is interested in inferring the distribution
-of divergence times among \npairs{} pairs of populations.
-For each pair $i$, \popSampleSize{i}{} genome copies have been sampled, with
-\popSampleSize{1}{i} copies sampled from population 1, and \popSampleSize{2}{i}
-sampled from population 2.
-From these genomes, let \nloci{i} be the number of DNA sequence loci collected
-for population pair $i$, and \nlociTotal be the total number of unique loci
-sampled across the \npairs{} pairs of populations.
-We use \alignment{i}{j} to represent the multiple sequence alignment of
-locus $j$ for population pair $i$.
-$\alignmentVector = (\alignment{1}{1}, \ldots,
-    \alignment{\npairs{}}{\nloci{\npairs{}}})$
-is the full dataset,
-i.e., a vector of sequence alignments for all pairs and loci.
-We let \geneTree{i}{j} represent the gene tree upon which \alignment{i}{j}
-evolved according to fixed \hky substitution model parameters \hkyModel{i}{j}.
-The investigator must specify the parameters of all
-$\hkyModelVector = (\hkyModel{1}{1}, \ldots,
-\hkyModel{\npairs{}}{\nloci{\npairs{}}})$
-substitution models by which the alignments evolved along the
-$\geneTreeVector = (\geneTree{1}{1}, \ldots,
-\geneTree{\npairs{}}{\nloci{\npairs{}}})$
-gene trees.
-Furthermore, the investigator must specify a vector of fixed contants
-$\ploidyScalarVector = (\ploidyScalar{1}{1}, \ldots,
-\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$
-that scale the population size parameters for known differences in ploidy among
-loci and/or differences in generation times among population pairs.
-Lastly, the investigator must also specify a vector of fixed constants
-$\mutationRateScalarConstantVector = (\mutationRateScalarConstant{1}{1},
-\ldots, \mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$
-that scale the population size parameters for known differences in
-mutation rates among loci and/or among taxa.
-
-\begin{linenomath}
-With \alignmentVector, \hkyModelVector, \ploidyScalarVector, and
-\mutationRateScalarConstantVector in hand, the joint posterior distribution
-of our model is given by
-\begin{equation}
-    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
-    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) =
-    \frac{p(\alignmentVector \given \geneTreeVector, \divTimeMapVector,
-        \demographicParamVector, \locusMutationRateScalarVector,
-        \locusRateHetShapeParameter, \hkyModelVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)
-        p(\geneTreeVector, \divTimeMapVector, \demographicParamVector,
-        \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-        \hkyModelVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)}{
-        p(\alignmentVector)},
-    \label{eq:fullModelCompactJoint}
-\end{equation}
-which can be expanded out into components that are assumed to be independent
-to get
-\begin{equation}
-    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
-    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) =
-    \frac{p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
-        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
-        \locusMutationRateScalarVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)
-        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
-        p(\locusRateHetShapeParameter)
-        p(\divTimeMapVector)
-        p(\demographicParamVector)}{
-        p(\alignmentVector)},
-    \label{eq:fullModelCompact}
-\end{equation}
-where
-$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
-is a vector of population divergence times for each of the \npairs{} pairs of
-populations,
-$\demographicParamVector = (\demographicParams{1}, \ldots,
-\demographicParams{\npairs{}})$
-is a vector of the demographic parameters for each of the \npairs{} population
-pairs,
-$\locusMutationRateScalarVector = (\locusMutationRateScalar{1}, \ldots
-\locusMutationRateScalar{\nlociTotal})$
-is a vector of locus-specific mutation-rate scaling parameters for each of the
-\nlociTotal loci,
-\locusRateHetShapeParameter is the shape parameter of a gamma-distributed
-prior on \locusMutationRateScalar{}, and
-$p(\alignmentVector)$ is the probability of the data (or the marginal
-likelihood of the model).
-\end{linenomath}
-
-\begin{linenomath}
-To avoid calculating the likelihood terms of Equation \ref{eq:fullModelCompact},
-we distill each sequence alignment \alignment{}{} into a vector of insufficient
-summary statistics \alignmentSS{}{}, thus replacing the full dataset
-$\alignmentVector = (\alignment{1}{1}, \ldots,
-    \alignment{\npairs{}}{\nloci{\npairs{}}})$
-with vectors of summary statistics for each alignment
-$\ssVectorObs = (\alignmentSSObs{1}{1}, \ldots,
-    \alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$.
-For each population pair, the means of the summary statistics are calculated
-across the \nloci{} loci, reducing the vector to
-$\ssVectorObs = (\alignmentSSObs{1}{}, \ldots,
-    \alignmentSSObs{\npairs{}}{})$.
-This allows us to estimate the approximate joint posterior distribution
-{\small
-\begin{equation}
-    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
-    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
-    \ssSpace, \hkyModelVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) =
-    \frac{p(\ssSpace \given \geneTreeVector, \hkyModelVector)
-        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
-        \locusMutationRateScalarVector, \ploidyScalarVector,
-        \mutationRateScalarConstantVector)
-        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
-        p(\locusRateHetShapeParameter)
-        p(\divTimeMapVector)
-        p(\demographicParamVector)}{
-        p(\ssSpace)},
-    \label{eq:approxModelCompact}
-\end{equation}
-}
-where \ssSpace is the multidimensional Euclidean space around the vector of
-summary statistics, the radius of which is the tolerance \tol.
-The sources of approximation are the insufficiency of the statistics and the
-\tol being greater than zero.
-We describe the full model in detail before delving into the numerical
-method of estimating the approximate model.
-\end{linenomath}
-
-\subsubsection*{Likelihood terms of Equation \ref{eq:fullModelCompact}}
-\begin{linenomath}
-The likelihood terms of Equation \ref{eq:fullModelCompact} can
-be expanded out as a product over population pairs and loci
-\begin{equation}
-    p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
-    p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
-    \locusMutationRateScalarVector, \ploidyScalarVector,
-    \mutationRateScalarConstantVector) = 
-    \prod_{i=1}^{\npairs{}}
-    \prod_{j=1}^{\nloci{i}}
-    p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})
-    p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
-    \locusMutationRateScalar{j}, \ploidyScalar{i}{j},
-    \mutationRateScalarConstant{i}{j}).
-    \label{eq:modelLikelihoodExpanded}
-\end{equation}
-The first term,
-$p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})$,
-is the probability of the sequence alignment of locus $j$ for population pair
-$i$ given the gene tree and \hky substitution model parameters \citep[i.e., the
-``Felsenstein likelihood'';][]{Felsenstein1981}.
-The model allows for an intra-locus recombination rate \recombinationRate,
-which, for simplicity, is assumed to be zero in Equation
-\ref{eq:fullModelCompact}.
-If \recombinationRate is non-zero, this term requires an additional product
-over the columns (sites) of each sequence alignment to allow sites to have
-different genealogies.
-The second term,
-p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
-\locusMutationRateScalar{j}, \ploidyScalar{i}{j},
-\mutationRateScalarConstant{i}{j}),
-is the probability of the gene tree under a multi-population coalescent model
-(i.e., species tree) where the ancestral population of pair $i$ diverges
-and gives rise to the two sampled descendant populations.
-Each \demographicParams{} contains the following demographic parameters:
-The mutation-scaled sizes of the ancestral, \ancestralTheta{}, and descendant
-populations, \descendantTheta{1}{} and \descendantTheta{2}{};
-the proportion of the first, \bottleScalar{1}{}, and second population,
-\bottleScalar{2}{}, that persist during bottlenecks that begin immediately
-after divergence in forward-time;
-the proportion of time between present and divergence when the bottlenecks
-end for both populations, \bottleTime{};
-and the symmetric migration rate between the descendant populations,
-\migrationRate{}.
-Thus, the probability of the $\popSampleSize{i}{}-1$ coalescence times (node
-heights) of gene tree \geneTree{i}{j} is given by a multi-population
-Kingman-coalescent model \citep{Kingman1982} where the ancestral population of
-size
-$\ancestralTheta{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
-diverges at time \divTimeMap{i} into two descendant populations of constant
-size
-$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{1}{i}$
-and
-$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{2}{i}$,
-which, after time $\divTimeMap{i}\bottleTime{i}$, grow exponentially to their
-present size 
-$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
-and
-$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$,
-respectively.
-Following divergence, the descendant populations of pair $i$ exchange migrants at a symmetric rate of
-\migrationRate{i}.
-\end{linenomath}
-
-\subsubsection*{Prior terms of Equation \ref{eq:fullModelCompact}}
-\begin{linenomath}
-The term $p(\locusRateHetShapeParameter)$ is the prior density
-function for the shape-parameter of the gamma-distributed prior on
-rate heterogeneity among loci.
-This prior is $\locusRateHetShapeParameter \sim U(1, 20)$.
-The prior probability of the vector of locus-specific mutation-rate scaling parameters
-given \locusRateHetShapeParameter then expands out as a product over
-the loci
-\begin{equation}
-    p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter) =
-    \prod_{j=1}^{\nlociTotal}
-    p(\locusMutationRateScalar{j} \given \locusRateHetShapeParameter),
-    \label{eq:locusRateHetPrior}
-\end{equation}
-where each \locusMutationRateScalar{} is independently and identicially
-distributed (\iid) as
-$\locusMutationRateScalar{} \sim Gamma(\locusRateHetShapeParameter,
-1/\locusRateHetShapeParameter)$.
-If the recombination rate \recombinationRate is allowed to be
-non-zero, the prior term $p(\recombinationRate)$ would be added
-to Equation \ref{eq:fullModelCompact}, and the prior would be
-$\recombinationRate \sim Gamma(\gammaShape{\recombinationRate},
-\gammaScale{\recombinationRate})$.
-\end{linenomath}
-
-\begin{linenomath}
-The prior term for the demographic parameters, $p(\demographicParamVector)$,
-expands out into its components and as a product over the \npairs{}
-pairs of populations
-\begin{equation}
-    p(\demographicParamVector) =
-    \prod_{i=1}^{\npairs{}}
-    p(\ancestralTheta{i})
-    p(\descendantTheta{1}{i})
-    p(\descendantTheta{2}{i})
-    p(\bottleScalar{1}{i})
-    p(\bottleScalar{2}{i})
-    p(\bottleTime{i})
-    p(\migrationRate{i}).
-    \label{eq:demographicPrior}
-\end{equation}
-The priors for the demographic parameters are:
-$\ancestralTheta{} \sim Gamma(\gammaShape{\ancestralTheta{}},
-\gammaScale{\ancestralTheta{}})$,
-$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$,
-$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$,
-$\bottleScalar{1}{} \sim Beta(\betaA{\bottleScalar{}{}},
-\betaB{\bottleScalar{}{}})$,
-$\bottleScalar{2}{} \sim Beta(\betaA{\bottleScalar{}{}},
-\betaB{\bottleScalar{}{}})$,
-$\bottleTime{} \sim U(0, 1)$,
-and
-$\migrationRate{} \sim Gamma(\gammaShape{\migrationRate{}},
-\gammaScale{\migrationRate{}})$.
-\end{linenomath}
-
-\subsubsection*{Priors on divergence models}
-\begin{linenomath}
-The prior term for the vector of divergence times for each of
-the \npairs{} pairs of populations can be expanded as
-\begin{equation}
-    p(\divTimeMapVector) = p(\divTimeIndexVector)p(\divTimeVector \given \divTimeIndexVector),
-    \label{eq:divModelPrior}
-\end{equation}
-where \divTimeVector is an ordered set of unique divergence time parameters
-$\{\divTime{1}, \ldots, \divTime{\divTimeNum}\}$ whose length
-\divTimeNum can range from 1 to \npairs{},
-and \divTimeIndexVector is the vector of indices of divergence time parameters 
-$(\divTimeIndex{1}, \ldots, \divTimeIndex{\npairs{}})$
-that map the divergence times in \divTimeVector to each of the \npairs{}
-pairs of populations.
-Thus, \divTimeMapVector is the result of applying the mapping function
-\begin{equation}
-    f(\divTimeVector, \divTimeIndexVector, i) = \divTime{\divTimeIndex{i}}
-    \label{eq:divTimeMapFunction}
-\end{equation}
-to each population pair $i$, such that
-$\divTimeMapVector = (\divTimeMap{1} = f(\divTimeVector, \divTimeIndexVector,
-1), \ldots, \divTimeMap{\npairs{}} = f(\divTimeVector, \divTimeIndexVector,
-\npairs{}))$.
-\end{linenomath}
-
-Biologically speaking, \divTimeVector contains the times of divergence events,
-the length of which \divTimeNum is the number of divergence events shared across
-the \npairs{} pairs of populations.
-For example, if \divTimeVector contains a single divergence time parameter
-\divTime{1}, all \npairs{} pairs of populations are constrained to diverge at
-this time (i.e., \divTimeIndexVector would contain the index 1 repeated
-\npairs{} times, and \divTimeMapVector would contain the value \divTime{1}
-repeated \npairs{} times), whereas if it contains \npairs{} divergence time
-parameters, the model is fully generalized to allow all of the pairs to diverge
-at unique times.
-
-\begin{linenomath}
-Each \divTime{} within \divTimeVector is \iid as $\divTime{} \sim
-Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
-Thus, given the number of unique divergence time classes in
-\divTimeIndexVector, this determines the probability of prior term
-$p(\divTimeVector \given \divTimeIndexVector)$.
-The divergence time parameters are in coalescent units relative to the size of
-a constant reference population, which we denote \myTheta{C}, that is equal to
-the expectation of the prior on the size of the descendant populations
-\begin{equation}
-    \myTheta{C} = \mathbb{E}(\descendantTheta{}{}).
-    \label{eq:thetaC}
-\end{equation}
-Given the size of the descendant populations are \iid as
-$\descendantTheta{}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$,
-this becomes
-\begin{equation}
-    \myTheta{C} = \gammaShape{\descendantTheta{}{}}\gammaScale{\descendantTheta{}{}}.
-    \label{eq:thetaCGamma}
-\end{equation}
-The \divTime{} parameters are in units $\myTheta{C}/\mutationRate$
-generations, which we denote as \globalcoalunit generations.
-Thus, each \divTime{} within \divTimeVector is proportional to time and can be
-converted to the number of generations of the reference population, which we
-denote \divTime{G_C}, by assuming a mutation rate and multiplying by the
-effective size of the reference population
-\begin{equation}
-    \divTime{G_C} = \divTime \times \frac{
-        \gammaShape{\descendantTheta{}{}}\gammaScale{\descendantTheta{}{}}}
-        {\mutationRate}.
-    \label{eq:divTimeGenerations}
-\end{equation}
-Thus, for each of the divergence times in \divTimeVector to be on the same
-scale, the relative mutation rates among the pairs of populations are assumed
-to be fixed and known according the user-provided values in
-\mutationRateScalarConstantVector.
-\end{linenomath}
-
-\begin{linenomath}
-As described by \citet{Oaks2012}, to get the divergence times in units
-proportional to the expected number of mutations, we must
-scale them by the realized population size for locus $j$ of population-pair $i$ 
-\begin{equation}
-    \divTimeScaled{i}{j} = \divTimeMap{i} \times \frac{\myTheta{C}}{
-        \descendantThetaMean{i} \ploidyScalar{i}{j}},
-    \label{eq:divTimeScaled}
-\end{equation}
-where \descendantThetaMean{i} is the mean of \descendantTheta{1}{} and
-\descendantTheta{2}{} for pair $i$.
-This gives us the vector of scaled divergence times
-$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
-\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$.
-\end{linenomath}
-
-\begin{linenomath}
-As for the prior term $p(\divTimeIndexVector)$, the total sample space of
-ordered realizations of \divTimeIndexVector is all the possible partitions of
-\npairs{} elements into 1 to \npairs{} categories.
-The total number of possible partitions is a sum of the Stirling numbers of
-the second kind over all possible numbers of categories \divTimeNum
-\begin{equation}
-    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
-    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
-    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
-    \label{eq:bell}
-\end{equation}
-which is the Bell Number \citep{Bell1934}.
-The original \msb model only samples over the unordered realizations of
-\divTimeIndexVector, such that the sample space is reduced to all the possible
-integer partitions of \npairs{} \citep{Oaks2012,Huang2011,OeisPartitionNumber,
-    OeisPartitionTriangle,Malenfant2011}.
-We denote the set of all possible integer partitions of \npairs{} as
-\integerPartitionSet{\npairs{}} and the length of that set as
-\integerPartitionNum{\npairs{}}.
-The advantages, disadvantages, and justification of ignoring the order
-of \divTimeIndexVector is discussed in detail below.
-\end{linenomath}
-
-\begin{linenomath}
-We implement two prior probability distributions over the space
-of all possible discrete divergence models (\divTimeIndexVector).
-The first simply gives all possible unordered partitions of \npairs{} elements
-equal probability
-\begin{equation}
-    p(\divTimeIndexVector) = \frac{1}{\integerPartitionNum{\npairs{}}},
-    \label{eq:divModelPriorUniform}
-\end{equation}
-i.e., a discrete uniform prior over all the integer partitions of \npairs{}
-(discrete divergence models).
-We denote this prior as
-$\divTimeIndexVector \sim \priorUniform$.
-\end{linenomath}
-
-The second prior we implement is based on the Dirichlet-process, which is a
-stochastic process that groups elements into an unknown number of discrete
-parameter classes \citep{Ferguson1973,Antoniak1974}.
-The process is controlled by the concentration parameter \concentrationParam,
-which determines how clustered the process will be.
-We use the Dirichlet process to place a prior over all possible ordered
-partitions of \npairs{} elements, which we denote as $\divTimeIndexVector \sim
-\priorDPP{}$.
-We take a hierarchical approach and use a prior probability
-distribution (i.e., hyperprior) for \concentrationParam.
-More specifically, we use a gamma-distributed prior $\concentrationParam =
-Gamma(\gammaShape{\concentrationParam}, \gammaScale{\concentrationParam})$,
-where \gammaShape{\concentrationParam} and \gammaScale{\concentrationParam} are
-specified by the user.
-
-\begin{linenomath}
-This provides a great deal of flexibility for specifying the prior uncertainty
-regarding divergence models.
-The \concentrationParam determines the prior probability that any two
-pairs of populations $i$ and $j$ will be assigned to the same divergence time
-parameter
-\begin{equation}
-    p(\divTimeIndex{i} = \divTimeIndex{j}) = \frac{1}{1 + \concentrationParam},
-    \label{eq:dppPriorSameClass}
-\end{equation}
-and also the prior probability of the number of divergence time parameters
-\begin{equation}
-    p(\divTimeNum \boldsymbol{\mid} \concentrationParam, \npairs{}) = 
-    \frac{\stirlingFirst{\npairs{}}{\divTimeNum} \concentrationParam^{\divTimeNum}}
-    {\prod_{i=1}^{\npairs{}}(\concentrationParam + i - 1)},
-    \label{eq:dppPriorNumClasses}
-\end{equation}
-where \stirlingFirst{\cdot}{\cdot} are the unsigned Stirling numbers of the
-first kind.
-Equations \ref{eq:dppPriorSameClass} and \ref{eq:dppPriorNumClasses} show that
-smaller values of \concentrationParam will favor fewer divergence time
-parameters, and thus more clustered models of divergence, whereas larger values
-will favor more divergence time parameters, and thus less clustered models of
-divergence.
-\end{linenomath}
-
-\subsection*{Differences between our model and the original \msb model}
-\subsubsection*{The prior on divergence models}
-One of the key differences between our model and that of \msb \citep{Huang2011}
-is the prior distribution on divergence models.
-As discussed in \citet{Oaks2012}, in \msb the only prior used for
-\divTimeIndexVector is a combination of a discrete uniform prior over the
-possible number of divergence events \divTimeNum from 1 to \npairs{} with a
-multinomial distribution on the number of times each index of \divTimeVector
-appears in \divTimeIndexVector, with the constraint that all \divTime{}
-parameters are represented at least once (see Equation 2 of \citet{Oaks2012}).
-We denote this prior used by \msb as $\divTimeIndexVector \sim \priorOld$.
-\citet{Oaks2012} discuss how placing a uniform prior over the number of
-divergence parameters (denoted \divTimeNum here, and as \numt{} in
-\citet{Huang2011}) imposes an U-shaped prior over discrete divergence
-models (\divTimeIndexVector; see
-Figure 5B of \citet{Oaks2012}).
-To avoid this, we place priors directly on the sample space of divergence
-models, thus eliminating the parameter \numt{} from the model.
-We introduce two priors on divergence models:
-(1) a prior that is uniform over all unordered divergence models, and
-(2) a Dirichlet-process prior on all ordered divergence models.
-The latter provides the user with a great deal of flexibility in
-expressing their prior beliefs about models of divergence.
-
-\subsubsection*{Estimating ordered divergence models}
-As mentioned above, \msb samples over unordered divergence models
-(i.e., unordered partitions of the \npairs{} pairs of populations).
-That is, the identity of each population pair, and all the information
-associated with it, is discarded.
-In our implementation, inference can be done on either unordered or ordered
-models of divergence.
-This is discussed in more detail in the description of the ABC implementation
-below.
-
-\subsubsection*{The priors on nuisance parameters}
-Following the recommendations of \citet{Oaks2012}, we have replaced the use of
-continuous uniform distributions for priors on many of the
-model's parameters (\divTime{}, \ancestralTheta{}, \descendantTheta{1}{},
-\descendantTheta{2}{}, \bottleScalar{1}{}, \bottleScalar{2}{},
-\recombinationRate, \migrationRate{}) with more flexible parametric
-distributions from the exponential family.
-We introduce gamma-distributed priors for rate parameters that have a sample
-space of all positive real numbers (\divTime{}, \ancestralTheta{},
-\descendantTheta{1}{}, \descendantTheta{2}{}, \recombinationRate,
-\migrationRate{}), and beta-distributed priors for parameters that are
-proportions bounded by zero and one (\bottleScalar{1}{} and
-\bottleScalar{2}{}).
-Not only do these priors have nice conjugacy properties, but they also provide
-the user with much greater flexibility in expressing prior uncertainty
-regarding the parameters of the model.
-
-Another distinction between our model and that of \msb, is the prior on the
-sizes of the descendant populations of each pair.
-As described by \citet{Oaks2012}, \msb uses the joint prior
-\begin{equation}
-    \descendantTheta{1}{}, \descendantTheta{2}{} \sim
-    Beta(1,1) \times 2 \times U(\uniformMin{\myTheta{}},
-    \uniformMax{\descendantTheta{}{}}),
-    \label{eq:jointThetaPrior}
-\end{equation}
-such that the user-specified uniform prior on descendant population
-size is a prior on the \emph{mean} size of the two descendant
-populations of each pair.
-Under our model, the sizes of the descendant populations of each
-pair are \iid as
-$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$
-and
-$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
-\gammaScale{\descendantTheta{}{}})$.
-This relaxes the assumption that the sizes of the two descendant populations
-are interdependent and negatively correlated.
-
-\subsubsection*{Flexibility in parameterizing the model}
-In our implementation, we provide the ability to control the richness of the
-model.
-For the \myTheta{} parameters, the model can be fully generalized to
-allow each population pair to have three parameters:
-\ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}.
-Furthermore, any model of \myTheta{} parameters nested within this
-general model can also be specified, including the most restricted model
-where the ancestral and descendant populations of each pair share
-a single \myTheta{} parameter.
-
-We also provide the option of eliminating the parameters associated with the
-post-divergence bottlenecks in the descendant populations of each pair
-(\bottleTime{}, \bottleScalar{1}{}, and \bottleScalar{2}{}),
-which constrains the descendant populations to be of
-constant size from present back to the divergence event.
-Also, rather than eliminate the bottleneck parameters,
-we allow \bottleScalar{1}{} and \bottleScalar{2}{} to be constrained to be
-equal, which removes one free parameter from the model for each of the
-population pairs.
-
-Overall, our implementation allows an investigator to specify a model that has
-as many as seven parameters per population pair
-(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{},
-\bottleTime{}, \bottleScalar{1}{}, \bottleScalar{2}{}, and
-\migrationRate{})
-or as few as one parameter per pair
-(\myTheta{}),
-in addition to the $\popSampleSize{i}{} - 1$ coalescence-time parameters.
-
-\subsubsection*{Time scale}
-As described above, in our model divergence times are in units of
-$\myTheta{C}/\mutationRate$ generations, where \myTheta{C} is the expectation
-of the prior on descendant-population size.
-As described by \citet{Oaks2012}, in \msb, \myTheta{C} is half of the upper
-limit of the continuous uniform prior on the mean of the descendant population
-sizes.
-
-\subsection*{ABC estimation of the posterior of the model}
-\subsubsection*{Sampling from the prior}
-To estimate the approximate posterior of Equation \ref{eq:approxModelCompact},
-we use an ABC rejection algorithm.
-The first step of this algorithm entails collecting a random sample of
-parameters values from the joint prior and their associated summary
-statistics.
-Each sample is generated by
-(1) drawing values of all the model's parameters, which we denote \hpvector{},
-from their respective prior distributions;
-(2) rescaling the divergence times
-$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
-from units proportional to time to units proportional to the expected number
-of mutations via Equation \ref{eq:divTimeScaled} to get 
-$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
-\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$;
-(3) simulating gene trees $\geneTreeVector = (\geneTree{1}{1}, \ldots, 
-\geneTree{\npairs{}}{\nloci{\npairs{}}})$
-for each locus of each population pair by drawing coalescent times from
-a multi-population Kingman-coalescent model given the demographic parameters;
-(4) simulating sequence alignments 
-$\alignmentVector = (\alignment{1}{1}, \ldots, \alignment{\npairs{}}{\nloci{\npairs{}}})$
-along the gene trees under the
-\hky substitution parameters
-$\hkyModelVector = (\hkyModel{1}{1}, \ldots, \hkyModel{\npairs{}}{\nloci{\npairs{}}})$
-that have the same number of sequences and sequence lengths as the observed
-dataset;
-(5) calculating population genetic summary statistics
-$\ssVector{} = (\alignmentSS{1}{1}, \ldots, \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$
-from the simulated sequence alignments (these are the same statistics
-calculated from the observed alignments, \ssVectorObs{});
-and (6) reducing the summary statistics to the means across loci for each
-population pair to get
-$\ssVector{} = (\alignmentSS{1}{}, \ldots, \alignmentSS{\npairs{}}{})$, which is
-the same summary statistic vector estimated from the observed data \ssVectorObs.
-After repeating this procedure \numPriorSamples times, we obtain a random
-sample of parameter vectors
-$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$
-from the model prior and their associated vectors of summary statistics
-$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$.
-
-For all of our analyses below, we use four summary statistics for each of
-the pairs of populations:
-$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
-\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
-Furthermore, in addition to model parameters, each sample \hpvector{}
-also contains four statistics that summarize \divTimeMapVector:
-the mean (\divTimeMean), variance (\divTimeVar), dispersion index
-($\divTimeDispersion = \divTimeVar/\divTimeMean$), and the number
-of divergence time parameters (\divTimeNum).
-These have been denoted as \meant{}{}, \vart{}{}, \vmratio{}, and \numt{},
-respectively \citep{Hickerson2006,Huang2011,Oaks2012}.
-We use the new notation to avoid confusion that the values represent prior or
-posterior expectations, and to indicate that \vmratio{} is not a parameter of
-the model and \numt{} is no longer a parameter in the new implementation.
-
-\subsubsection*{Ordering of taxon-specific summary statistics}
-As alluded to in the model description, \msb does not maintain the order of the
-taxon-specific summary statistics \alignmentSS{}{} within each \ssVector{}.
-Rather, the summary statistics are re-ordered by descending values of average
-pairwise differences between the descendant populations
-\citep[$\pi_b$;][]{NeiLi1979,Huang2011}.
-This has the advantage of reducing the sample space of the number of discrete
-divergence models \divTimeMapVector, but there are at least two disadvantages.
-First, additional information in the data is lost.
-By discarding the identity of the \npairs{} pairs of populations, all
-pair-specific information about the amount of data (e.g., the number gene
-copies collected from each of the populations [\popSampleSize{1}{} and
-\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
-taxon- and locus-specific parameters (\hkyModel{}{},
-\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
-\locusMutationRateScalar{}) is lost.
-Second, the results are more difficult to interpret, because they can be no
-longer be directly associated to the taxa under study.
-
-The original descriptions of the \msb model claim that this re-ordering is
-justified by the fact that of $\pi_b$ (and other summary statistics) are
-unrelated to the sample sizes \popSampleSize{1}{} and \popSampleSize{2}{} of
-each pair and are thus exchangeable \citep{Hickerson2006,Huang2011}.
-This is actually incorrect for two of reasons.
-First, the expectation of $\pi_b$ is not independent of samples sizes.
-If there are more than one coalescent events in the ancestor, which is expected
-to be common on phylogeographic timescales, more samples will increase the
-probability of capturing these deeper events, and thus affect the average
-pairwise differences between the descendant populations.
-Also, other statistics that estimate gross diversity (e.g., $\pi$ and
-$\theta_W$) are clearly not independent of sample size.
-Second, for variables to be exchangeable, they do not need to be independent,
-but their marginal distributions must be the same (i.e., they must be
-identically distributed).
-The simulated alignments and their summary statistics are not identically
-distributed because of differing sample sizes \emph{and} taxon- and
-locus-specific parameters \hkyModel{}{}, \mutationRateScalarConstant{}{}, and
-\ploidyScalar{}{}.
-
-Thus, the theoretical basis for this reshuffling is questionable.
-It may be justifiable as an additional coarsening of the data,
-by removing the information associated with the identity of the
-population pairs.
-However, we do not provide a formal proof that this re-ordering of the pairs
-does not introduce bias.
-One can imagine situations in which the sampling intensity (i.e., the
-number of gene copies, loci, and locus length) is highly skewed across
-pairs and/or there is large heterogeneity in mutational processes
-(e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
-In such cases, it seems possible that discarding this information could
-cause bias.
-Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
-is to minimize Euclidean distances between simulated datasets in which the true
-model has a single divergence event, and (2) the reordering increases the
-model's tendency to infer a single divergence event \citep{Huang2011}, it seems
-possible that this approach could be biasing the method \citep{Oaks2012}.
-
-To maintain compatibility and comparability with \msb we maintain
-the re-ordering of taxon-specific summary statistics by $\pi_b$.
-However, we also allow the order to be maintained, and ordered
-divergence models to be estimated.
-
-\subsubsection*{Obtaining an approximate posterior from the prior samples}
-We use a rejection algorithm to retain an approximate posterior sample of
-\paramSampleVector{} from the prior sample
-$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$.
-First, the observed summary statistics \ssVectorObs, and the summary statistics
-of the prior samples 
-$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$,
-are standardized using the means and standard deviations of the statistics from
-the prior sample (i.e., the prior mean is subtracted from each statistic, and then
-it is divided by the prior standard deviation).
-After all statistics are standardized, the Euclidean distance between
-\ssVectorObs and each \ssVector{} within \ssMatrix is calculated.
-The samples that fall within a range of tolerance \tol around \ssVectorObs
-are retained.
-The range of tolerance is determined by specifying the number of desired
-posterior samples to be retained.
-Post-hoc adjustment of the posterior sample can also be performed with a number
-of regression techniques \citep{Beaumont2002,Blum2009,Leuenberger2010}.
-For our analyses, we use the general linear model (GLM) regression adjustment
-\citep{Leuenberger2010} as implemented in \abctoolbox
-\citep[v1.1;][]{ABCtoolbox}, which \citet{Oaks2012} showed performed very
-similarly to weighted local-linear regression and multinomial logistic
-regression adjustments \citep{Beaumont2002} for \msb posteriors.
-
-\subsection*{Assessing model-choice behavior and robustness}
-Following the simulation-based approach of \citet{Oaks2012}, we characterize
-the behavior of several models under the ideal conditions where the
-data are generated from parameters drawn from the same prior distributions used
-for analysis (i.e., the prior is correct).
-We selected four model priors for these analyses (Table~\ref{tabPriors}):
-(1) A model to represent the original \msb model, \modelOld, with
-priors
-$\divTimeIndexVector \sim \priorOld$,
-$\divTime{} \sim U(0,10)$,
-$\ancestralTheta{} \sim U(0, 0.05)$,
-and
-$\descendantThetaMean{} \sim U(0, 0.05)$;
-(2) a Dirichlet-process prior model, \modelDPP, with priors
-$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
-and
-$\descendantTheta{}{} \sim Exp(40)$;
-(3) a model with uniform prior probability over unordered divergence models,
-\modelUniform, with priors
-$\divTimeIndexVector \sim \priorUniform$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
-and
-$\descendantTheta{}{} \sim Exp(40)$;
-and
-(4) a model with the \msb prior on unordered divergence models, but with
-exponential priors on nuisance parameters,
-\modelUshaped, with priors
-$\divTimeIndexVector \sim \priorOld$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
-and
-$\descendantTheta{}{} \sim Exp(40)$.
-We selected the exponential prior on divergence time used in models \modelDPP,
-\modelUniform, and \modelUshaped to have the same variance as the uniform prior
-in model \modelOld.
-We selected the exponential prior on population size used in models \modelDPP,
-\modelUniform, and \modelUniform to have the same mean as the uniform prior in
-model \modelOld, so that all four models have the same \myTheta{C} and thus the
-same units of time.
-All of the models were the same in other respects, with three free \myTheta{}
-parameters for each population pair, two uniformly distributed ($beta(1,1)$)
-\bottleScalar{}{} parameters per pair, no migration, no recombination,
-and unordered divergence models.
-For all of our simulations, we used a simulated data structure of eight
-population pairs, with a single 1000 base-pair locus sampled from 10
-individuals from each population.
-
-For each of the four models, we generated $1\e6$ samples from the prior, and
-simulated 50,000 pseudo-replicate datasets, also drawn from the prior.
-We then analyzed each of the replicate datasets, retaining a
-posterior of 1000 samples from the respective prior.
-A GLM-regression adjusted posterior was also estimated from each of the
-posterior samples \citep{Leuenberger2010}.
-To assess the robustness of each of the four models, we also analyzed the
-pseudo-replicate datasets simulated under the other three models.
-Overall, for each model, we produced 200,000 posterior estimates,
-50,000 from the replicate datasets simulated under that model,
-and 150,000 estimated from the replicate datasets simulated under the
-other three models.
-
-For each set of 50,000 simulated datasets, we used the posterior estimates
-to assess the model-choice behavior of each model.
-We did this by assigning the 50,000 estimates of the posterior probability
-of one divergence event to 20 bins of width 0.05, and plotted
-the estimated posterior probability of each bin against the proportion of
-replicates in that bin with a true value consistent with one divergence
-event \citep{Huelsenbeck2004,Oaks2012}.
-We did this using two criteria for the one divergence model:
-(1) the number of divergence time parameters ($\divTimeNum = 1$) and
-(2) the dispersion index of divergence times \citep[$\divTimeDispersion <
-0.01$][]{Hickerson2006,Huang2011}.
-We used the one-divergence model to assess model-choice behavior, because
-it is often of biogeographic interest and is easily comparable among
-the three different priors used on divergence models.
-
-In addition to the four models above, we also assessed the behavior of a model
-with ordered divergence models (i.e., the order of the taxon-specific summary
-statistic vectors were maintained for the observed and simulated datasets).
-For this, we used a model with identical priors as \modelDPP, but that samples
-over ordered divergence models.
-We denote this model as \modelDPPOrdered.
-We generated $1\e6$ prior samples and 50,000 pseudo-replicate datasets, and
-analyzed them as above.
-We were not able to analyze the replicate datasets of the other models under
-the ordered model, because the identity of the population pairs is not
-contained in the simulations of the other models.
-
-\subsection*{Assessing power}
-We evaluated the power of the same four models (Table~\ref{tabPriors}) to
-detect random variation in divergence times using methods similar to
-\citet{Oaks2012}.
-For all of our power simulations, we used a simulated data structure identical
-to that of the empirical dataset of Philippine vertebrates analyzed by
-\citet{Oaks2012}, which consists of 22 pairs of populations.
-For each of the four models, we generated $2\e6$ samples from the prior.
-We then generated pseudo-observed datasets from three series of models.
-One series of models, which we denote \powerSeriesOld, is identically
-distributed as \modelOld except that the divergence times for each of the 22
-pairs of populations are randomly drawn from a series of uniform distributions,
-$U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and
-2.0, in \globalcoalunit generations.
-A second series of models, \powerSeriesUniform, are identically distributed as
-\modelUniform except that the 22 divergence times are randomly drawn from the
-same series of uniform priors as above.
-The third series of models, \powerSeriesExp, is also identically
-distributed as \modelUniform except that the 22 divergence times are randomly
-drawn from a series of of exponential distributions: $Exp(17.3205)$,
-$Exp(8.6603)$, $Exp(5.7735)$, $Exp(4.33013)$, $Exp(3.4641)$, and $Exp(1.7321)$.
-These exponential distributions have the same variance as their uniform
-counterparts in the first two series of models.
-
-For each of the six models in each of the three series of models, we simulated
-1000 pseudo-replicate datasets (18,000 datasets in total).
-We then analyzed each replicate dataset under all four prior models
-(Table~\ref{tabPriors}), producing 72,000 posterior estimates, each with 1000
-samples.
-We also estimated a GLM-regression adjusted posterior from each of the
-posterior samples \citep{Leuenberger2010}.
-
-\section*{Results}
-\subsection*{Validation analyses: Estimation accuracy}
-In terms of estimating summary statistics of model parameters (\divTimeNum,
-\divTimeDispersion, and \divTimeMean), the \modelDPP performs similarly when
-applied to datasets generated under all four models of Table \ref{tabPriors}
-(Figs.\
-\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP}).
-The same is true for models \modelUniform (Figs.\
-\labelcref{figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform})
-and \modelUshaped (Figs.\
-\labelcref{figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped}).
-All three of the models with exponential priors on nuisance parameters
-(\modelDPP, \modelUniform, and \modelUshaped) perform well when applied
-to the data generated under the original \msb model, \modelOld (Figs.\
-\labelcref{figValidationAccuracyOldDPP,figValidationAccuracyOldUniform,figValidationAccuracyOldUshaped}).
-The \modelOld model also performed well on its own datasets (Figure
-\ref{figValidationAccuracyOldOld}), but performed poorly when applied to the
-data generated under the other three models (Figs.\
-\labelcref{figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld}).
-Thus, the newly implemented models display greater robustness than
-the original model.
-
-Posterior adjustment with GLM regression does improve the estimation
-accuracy for the continuous variables \divTimeDispersion and \divTimeMean,
-however, this is not always true for estimating the discrete variable
-\divTimeNum (Figs.\
-\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP,figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform,figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped,figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld,figValidationAccuracyOldOld}).
-
-\subsection*{Validation analyses: Model-choice accuracy}
-The \msb model, and our modification of it, is a model-choice method
-with the primary purpose of estimating the probabilities of models
-of divergence across taxa.
-Thus, it is critical to assess the method's ability to accurately
-estimate the posterior probabilities of divergence models.
-Our results demonstrate that the unadjusted estimates of divergence models are
-generally more accurate than the regression-adjusted estimates, especially when
-using $\divTimeDispersion < 0.01$ to assess the probability of the one
-divergence model (Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
-Thus, we will focus our discussion of the results on the unadjusted estimates.
-
-The model-choice behavior of the \modelDPP is relatively accurate when applied
-to datasets generated under models \modelDPP, \modelUniform, and \modelOld
-(Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP}).
-It is less accurate and tends to underestimate the probability of the
-one-divergence model when applied to the data generated under \modelUshaped
-(Figure \ref{figValidationModelChoiceUshapedDPP}).
-The \modelUniform model performs almost identically as \modelDPP in
-terms of model-choice accuracy (Figs.\
-\labelcref{figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}).
-
-The original \msb model \modelOld performs reasonably well when applied
-to its own datasets (Figure \ref{figValidationModelChoiceOldOld}), but
-it performs poorly when applied to data generated under any of the other
-models (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
-In these cases the \modelOld model is strongly biased towards over-estimating
-the posterior probability of the one-divergence model, especially when the
-divergence models are not distributed under its U-shaped prior (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld}).
-The other model with the U-shaped prior on divergence models, but exponential
-priors on nuisance parameters (\modelUshaped), performs similarly to \modelOld
-in that it performs well when applied to its own data, but overestimates
-the probability of the one-divergence model when applied to data generated
-by the other models (Figs.\
-\labelcref{figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceUshapedUshaped}).
-However, the bias toward over-estimating the probability of the one-divergence
-model is stronger in the \modelOld model than \modelUshaped.
-
-Overall, the result suggest that the \modelDPP and \modelUniform models are
-relatively robust in terms of model-choice accuracy, and when model violations
-do cause them to be biased, they tend to under-estimate the probability of
-the biogeographically interesting model of a single divergence event.
-In contrast, the \modelOld model is very sensitive to model violations,
-and strongly over-estimates the one-divergence model whenever the model
-is misspecified.
-Furthermore, the results suggest that using exponentially distributed priors on
-nuisance parameters rather than uniform priors helps the \modelUshaped model
-perform better than \modelOld, but it is still hindered by the U-shaped prior
-on divergence models and tends to over-estimate the probability of the
-one-divergence model whenever there are violations of the model.
-
-\subsection*{Validation analyses: Ordered divergence models}
-Our results show that the ordered models of divergence, \modelDPPOrdered,
-perform similarly to sampling over unordered models (\modelDPP) when the prior
-is correct (Figs.\
-\labelcref{figValidationAccuracyDPPOrderedDPPOrdered,figValidationModelChoiceDPPOrderedDPPOrdered}).
-This suggests that the method is not adversely affected by the large increase
-in possible discrete models (from 22 unordered to 4140 ordered models) when
-there are 8 pairs of populations.
-
-\subsection*{Power analyses: Estimation accuracy}
-All of the models we evaluated (Table \ref{tabPriors}) struggle to estimate the
-dispersion index of divergence times \divTimeDispersion regardless of which of
-the three series of models (Table \ref{tabPowerModels}) the data were generated
-under (Figs.\
-\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform,figPowerAccuracyExpOld,figPowerAccuracyExpUshaped,figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped,figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
-When the divergence times of the 22 population pairs are randomly drawn from a
-series of exponential priors (\powerSeriesExp), the \modelDPP model is the
-best estimator of \divTimeDispersion, followed by \modelUniform (figs.\
-\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform}).
-The \modelOld model is strongly biased toward underestimating
-\divTimeDispersion, estimating values of zero for most of the replicates across
-all the data models of \powerSeriesExp (Figure \ref{figPowerAccuracyExpOld}).
-The results of the \modelUshaped model are intermediate between those of
-\modelOld and the new models \modelDPP and \modelUniform
-(Figure \ref{figPowerAccuracyExpUshaped}).
-
-When the true divergence times are randomly drawn from a series of uniform
-priors (\powerSeriesUniform), all of the models poorly estimate
-\divTimeDispersion (Figs.\
-\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped}).
-The \modelDPP and \modelUniform models tend to over-estimate the variance in
-divergence times (Figs.\
-\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform}), whereas
-the \modelOld model, again, underestimates \divTimeDispersion, estimating
-values of zero for most replicates across all the data models of
-\powerSeriesUniform (Figure \ref{figPowerAccuracyUniformOld}).
-Again, the performance of the \modelUshaped model is intermediate between the
-\modelOld and \modelDPP/\modelUniform models (Figure
-\ref{figPowerAccuracyUniformUshaped}).
-This pattern of results across the four models is very similar when applied
-to data simulated under the series of models of \powerSeriesOld (Figs.\
-\labelcref{figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
-
-\subsection*{Power analyses: Model choice}
-There is a large amount of heterogeneity in estimating the number of divergence
-events (\divTimeNum) both among the prior models, and across the different
-series of data models (Table \ref{tabPowerModels}) (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped}).
-The \modelDPP model tends to infer highly clustered divergence models across
-all three series of data models when the divergences are recent, but performs
-the best of the four prior models  when divergences are random over a timescale
-of about one to two coalescent units (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiUniformDPP,figPowerPsiOldDPP,}).
-The \modelUniform model performs the best when divergences are random
-over the most recent timescales, but tends to only infer a tight
-cluster of \divTimeNum values even when divergences are random
-over longer periods (Figs.\
-\labelcref{figPowerPsiExpUniform,figPowerPsiUniformUniform,figPowerPsiOldUniform,}).
-The \modelOld model performs the worst of the four prior models across
-all three series of data models, inferring a single divergence event across
-most of the 18,000 simulations (Figs.\
-\labelcref{figPowerPsiExpOld,figPowerPsiUniformOld,figPowerPsiOldOld,}).
-Even when the simulated data are identically distributed as the \modelOld model
-(except for the divergence times) the \modelOld is still strongly biased
-towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
-model is the least likely to infer a single event when applied to these data
-(Figure \ref{figPowerPsiOldDPP}).
-Using exponential priors on nuisance parameters does increase the power
-of the \modelUshaped model compared to \modelOld across all three series
-of data models, but the U-shaped prior still prevents the model
-from performing as well as the \modelDPP and \modelUniform models (Figs.\
-\labelcref{figPowerPsiExpUshaped,figPowerPsiUniformUshaped,figPowerPsiOldUshaped,}).
-
-The improved power of the new models are even more pronounced when looking at
-estimates of the dispersion index of divergence times (\divTimeDispersion)
-across the simulations (Figs.\
-\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped}).
-The \modelDPP and \modelUniform models perform similarly across all three
-series of data models, inferring values of \divTimeDispersion consistent with
-one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
-replicates across all the simulations (Figs.\
-\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionOldDPP,figPowerDispersionOldUniform}).
-The \modelOld model infers values consistent with a single divergence event in
-most of the replicates across all the simulations (Figs.\
-\labelcref{figPowerDispersionExpOld,figPowerDispersionUniformOld,figPowerDispersionOldOld}).
-Using exponential priors on nuisance parameters greatly increases the power
-of the \modelUshaped model relative to \modelOld, but it still has
-less power than the models with DPP or uniform priors across
-divergence models (Figs.\
-\labelcref{figPowerDispersionExpUshaped,figPowerDispersionUniformUshaped,figPowerDispersionOldUshaped}).
-
-When looking at the estimated posterior probability of the single-divergence
-model across the power analyses, we also see the increased power of the
-new models (Figs.\
-\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped}).
-The \modelDPP model estimates low posterior probability of
-$\divTimeNum = 1$ across all of the simulations (Figs.\
-\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbUniformDPP,figPowerPsiProbOldDPP}).
-The \modelDPP does infer strong support for one divergence as gauged by a Bayes
-factor of greater than 10.
-However, this is not that unexpected given the low prior probability of a
-single divergence under this model.
-Also, the Bayes factor calculations for the \modelDPP replicates are only
-approximate; they are calculated based on the prior mean of the concentration
-parameter \concentrationParam.
-
-The \modelUniform model infers very low posterior probabilities for the
-one-divergence model across all the simulations (Figs.\
-\labelcref{figPowerPsiProbExpUniform,figPowerPsiProbUniformUniform,figPowerPsiProbOldUniform}),
-whereas the \modelOld model infers high posterior probabilities of a single
-divergence for most replicates across all simulations (Figs.\
-\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld}).
-Again, we see that the exponential priors on nuisance parameters greatly
-increase power and result in lower estimates of the probability of
-one divergence when comparing \modelUshaped to \modelOld (Figs.\
-\labelcref{figPowerPsiProbExpUshaped,figPowerPsiProbUniformUshaped,figPowerPsiProbOldUshaped}).
-
-Lastly, when we look at the estimated posterior probability of
-\divTimeDispersion being consistent with one divergence parameter
-($p(\divTimeDispersion < 0.01 | \ssSpace)$), we see the same pattern of model
-behavior, with \modelDPP and \modelUniform inferring low probabilities across
-all simulations, \modelOld inferring high probabilities, and \modelUshaped
-inferring intermediate values (Figs.\
-\labelcref{figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
-
-\section*{Discussion}
-Our results demonstrate that using alternative priors on nuisance parameters
-and divergence models improved the behavior of the \msb model.
-The model-choice estimation of our implementation is more accurate and,
-perhaps more importantly, shows greater robustness to model
-violations (Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
-The original model is very sensitive to model violations and, when present,
-strongly over-estimates the probability of one divergence event shared
-across all taxa (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
-When more appropriate priors are used for nuisance parameters, and either
-a Dirichlet-process or uniform prior applied across divergence models,
-the model is less sensitive to violations (Figs.\
-\labelcref{figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}),
-and, when they do cause bias, the method is conservative and tends to
-under-estimate the probability of biogeographically interesting models of
-temporally clustered divergences 
-\labelcref{figValidationModelChoiceUshapedDPP,figValidationModelChoiceUshapedUniform}).
-
-Furthermore, the modifications to the model also improve the method's power to
-detect random variation in divergence times, reducing the bias toward
-estimating clustered divergences (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped,figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped,figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped,figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
-Our results are similar to those of \citet{Oaks2012} in that we also find
-the \msb model to infer strong support for clustered divergences when
-divergences are random over quite broad timescales.
-\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld,figPowerDispersionProbExpOld,figPowerDispersionProbUniformOld,figPowerDispersionProbOldOld}).
-Our results expand on this by showing that this behavior is consistent
-across of range conditions underlying the data.
-When using the new priors, the method is less prone to spurious inference of
-clustered divergence models, and much less likely to incorrectly infer such
-models with strong support.
-
-By evaluating a model intermediate between the old and new implementation
-(\modelUshaped), we were able to determine the relative effects of the
-modifications to model.
-Across all of our analyses, the results show that using better priors on
-nuisance parameters alone does improve the performance of the method.
-The magnitude of the bias toward inferring support for the one-divergence
-model when there are model violations is reduced when the exponential
-priors are used in place of the uniform priors.
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped}).
-Furthermore, better priors on on nuisance parameters improves the method's
-power to detect temporally random divergences.
-Throughout our analyses, the intermediate model (\modelUshaped) performs
-better than the \msb model, but not as well as the models with alternative
-priors on divergence models.
-This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
-models of temporally clustered divergences is caused by a combination of
-(1) uniform priors on nuisance parameters hindering the marginal likelihoods of
-models with more \divTime{} parameters and
-(2) the U-shaped prior on divergence models giving low prior density to models
-with intermediate numbers of divergence parameters.
-
-While our modifications improve the behavior of the model, we urge
-caution when using the method and interpreting its results.
-The method attempts to approximate the posterior of a very parameter rich model
-using little information from the data.
-For example, when applied to the dataset of Philippine vertebrates analyzed by
-\citet{Oaks2012} the model can have as many as 604--625 free parameters
-(depending on \divTimeNum), and samples over 1002 divergence models.
-Our modifications allow this model to be simplified to 471--492 free
-parameters.
-Sampling over hundreds of parameters and more than a thousand models is asking
-a lot from a few summary statistics calculated from the sequence alignments of
-each taxon pair.
-These simple summary statistics contain little information about many
-of the free parameters in the model.
-Thus, we expect the improved method will still be sensitive to priors, and the
-power, while improved, may still be low.
-Furthermore, ABC methods in general have been shown to be biased
-\citep{Robert2011} for model choice.
-As recommended by \citet{Oaks2012}, any results from the method should be
-accompanied by assessments of prior sensitivity and power.
-We encourage investigators to view this method and its predecessor as a means
-of exploring their data rather than as rigorous means of evaluating hypotheses.
-
-Given the difficulty of this estimation problem, we anticipate that
-full-likelihood methods that can leverage all of the information present in the
-sequence data will become increasingly important in robustly estimating the
-temporal history of divergences across taxa \citep{JeetDiss}.
-With improving numerical methods for sampling over models of differing
-dimensionality \citep{Green1995,Lemey2009} and increasing efficiency of
-likelihood calculations \citep{Ayres2012}, analyzing rich comparative
-phylogeograpical models in full-likelihood Bayesian framework is becoming
-computationally practical, especially when considering that the computational
-load of simulating millions of random datasets from the prior under the ABC
-approach is nontrivial.
+\input{dpp-msbayes-main.tex}
 
@@
 
-%:TABLE-notation descriptions
-\begin{table}[htbp]
-    \sffamily
-    \scriptsize
-    \rowcolors{2}{}{myGray}
-    \addtolength{\tabcolsep}{-0.1cm}
-\caption{Summary of the notation used throughout this work; modified from \citet{Oaks2012}.}
-    \centering
-    \begin{tabular}{ l >{\raggedright\hangindent=0.5cm}m{14cm} }
-        \toprule
-        \textbf{Symbol} & \textbf{Description} \tn
-        \midrule
-        \npairs{} & Number of population pairs. \tn
-        \popSampleSize{i}{} & The number of genome copies sampled from population pair $i$, with \popSampleSize{1}{i} sampled from population 1, and \popSampleSize{2}{i} from population 2. \tn
-        \nloci{i} & Number of loci sampled from population pair $i$. \tn
-        \nlociTotal & Total number of unique loci sampled. \tn
-        \alignment{i}{j} & Sequence alignment of locus $j$ sampled from population pair $i$. \tn
-        \alignmentSSObs{i}{j} & Population genetic summary statistics calculated from \alignment{i}{j}. \tn
-        \alignmentVector & Vector containing the sequence alignments of each locus from each population pair: $(\alignment{1}{1},\ldots,\alignment{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \ssVectorObs & Vector containing the summary statistics of each locus from each population pair: $(\alignmentSSObs{1}{1},\ldots,\alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \ssSpace & Multi-dimensional Euclidean space around the observed summary statistics, \ssVectorObs. \tn
-        \tol & Radius of \ssSpace, i.e., the tolerance of the ABC estimation. \tn
-        \geneTree{i}{j} & Gene tree of the sequences in \alignment{i}{j}. \tn
-        \geneTreeVector & Vector containing the gene trees of each locus from each population pair: $(\geneTree{1}{1},\ldots,\geneTree{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \divTimeNum & Number of unique population divergence times among the \npairs{} population pairs. \tn
-        \divTime{} & One of the \divTimeNum \iid divergence time paramters in \globalcoalunit generations. \tn
-        \divTimeVector & Ordered set of divergence time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
-        \divTimeIndex{i} & The index of the divergence time parameter in \divTimeVector that is mapped to population pair $i$. \tn
-        \divTimeIndexVector & Vector of divergence time parameter indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn
-        \divTimeMap{i} & Time of divergence in \globalcoalunit generations between the populations of pair $i$. \tn
-        \divTimeMapVector & Vector of divergence times for each of the population pairs: $(\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})$. \tn
-        \divTimeScaled{i}{j} & Scaled time of divergence between the populations of pair $i$ for locus $j$. \tn
-        \divTimeScaledVector & Vector containing the scaled divergence times of each locus from each population pair: $(\divTimeScaled{1}{1},\ldots,\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \descendantTheta{1}{i}, \descendantTheta{2}{i} & \myTheta{} of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
-        \ancestralTheta{i} & \myTheta{} of the population ancestral to pair $i$. \tn
-        \descendantThetaVector{1}, \descendantThetaVector{2} & Vectors $(\descendantTheta{1}{1},\ldots,\descendantTheta{1}{\npairs{}})$ and $(\descendantTheta{2}{1},\ldots,\descendantTheta{2}{\npairs{}})$, respectively.
-\tn
-        \ancestralThetaVector & Vector containing the \ancestralTheta{} parameters for each population pair: $(\ancestralTheta{1},\ldots,\ancestralTheta{\npairs{}})$. \tn
-        \locusMutationRateScalar{j} & mutation-rate scaling parameter for locus $j$. \tn
-        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate parameters: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
-        \locusRateHetShapeParameter & Hyper-parameter for the shape of the gamma prior distribution on each \locusMutationRateScalar{}. \tn
-        \bottleScalar{1}{i}, \bottleScalar{2}{i} & \myTheta{}-scaling parameters that determine the magnitude of the population bottleneck in the $1^{st}$ and $2^{nd}$ descendant population of pair $i$, respectively.
-The bottleneck in each descendant population begins immediately after divergence.
-\tn
-        \bottleScalarVector{1}, \bottleScalarVector{2} & Vectors $(\bottleScalar{1}{1},\ldots,\bottleScalar{1}{\npairs{}})$ and $(\bottleScalar{2}{1},\ldots,\bottleScalar{2}{\npairs{}})$, respectively.
-\tn
-        \bottleTime{i} & Proportion of time between present and \divt{i} when the bottleneck ends for the descendant populations of pair $i$. \tn
-        \bottleTimeVector & Vector containing the \bottleTime{} parameters for each population pair: $(\bottleTime{1},\ldots,\bottleTime{\npairs{}})$. \tn
-        \migrationRate{i} & Symmetric migration rate between the descendant populations of pair $i$. \tn
-        \migrationRateVector & Vector containing the migration rates for each population pair: $(\migrationRate{1},\ldots,\migrationRate{\npairs{}})$. \tn
-        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
-This constant is intended to allow the user to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa.
-\tn
-        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
-This constant is intended to allow the user to scale \myTheta{} for differences in mutation rates among loci or among taxa.
-\tn
-        \ploidyScalarVector & Vector of ploidy and/or generation-time scaling constants: $(\ploidyScalar{1}{1},\ldots,\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \mutationRateScalarConstantVector & Vector of mutation-rate scaling constants: $(\mutationRateScalarConstant{1}{1},\ldots,\mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \divTimeMean  & Mean of \divTimeMapVector. \tn
-        \divTimeVar  & Variance of \divTimeMapVector. \tn
-        \divTimeDispersion  & Dispersion index of \divTimeMapVector ($\divTimeVar/\divTimeMean$). \tn
-        \numPriorSamples & Number of samples from the joint prior. \tn
-        \paramSampleVector{} & Vector of parameter values drawn from the joint prior. \tn
-        \ssVector{} & Vector containing the summary statistics calculated from data simulated under parameter values drawn from the prior (\hpvector{}). \tn
-        \paramSampleMatrix{} & Random sample of $\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples}$ drawn form the prior. \tn
-        \ssMatrix & Summary statistic vectors $\ssVector{1}, \ldots, \ssVector{\numPriorSamples}$ for each $\hpvector{1}, \ldots, \hpvector{\numPriorSamples}$ drawn from the prior. \tn
-        \bottomrule
-    \end{tabular}
-    \label{tabNotation}
-\end{table}
-
-\clearpage
-
-\begin{table}[htbp]
-    \sffamily
-    % \footnotesize
-    \addtolength{\tabcolsep}{-0.08cm}
-    \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
-    \caption{The model priors evaluated in our simulation-based analyses.}
-    \centering
-    \begin{tabular}{ l l l l l }
-        \toprule
-        & \multicolumn{4}{c}{Priors} \\
-        \cmidrule(){2-5}
-        Model & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
-        \midrule
-            \modelOld & $\divTimeIndexVector \sim \priorOld$
-                      & $\divTime{} \sim U(0,10)$
-                      & $\ancestralTheta{} \sim U(0, 0.05)$
-                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$
-                      & $\divTime{} \sim Exp(0.3464)$
-                      & $\ancestralTheta{} \sim Exp(40)$
-                      & $\descendantTheta{}{} \sim Exp(40)$ \\
-            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
-                          & $\divTime{} \sim Exp(0.3464)$
-                          & $\ancestralTheta{} \sim Exp(40)$
-                          & $\descendantTheta{}{} \sim Exp(40)$ \\
-            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
-                          & $\divTime{} \sim Exp(0.3464)$
-                          & $\ancestralTheta{} \sim Exp(40)$
-                          & $\descendantTheta{}{} \sim Exp(40)$ \\
-        \bottomrule
-    \end{tabular}
-    \label{tabPriors}
-\end{table}
-
-\clearpage
-
-\begin{table}[htbp]
-    \sffamily
-    % \footnotesize
-    % \addtolength{\tabcolsep}{-0.08cm}
-    \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
-    \caption{The models we used to simulate pseudo-replicate datasets for
-        assessing the power of the models in Table \ref{tabPriors}.}
-    \centering
-    \begin{tabular}{ l l l l l }
-        \toprule
-        & \multicolumn{4}{c}{Priors} \\
-        \cmidrule(){2-5}
-        Model series & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
-        \midrule
-            \powerSeriesOld & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.2)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.4)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.6)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.8)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,1.0)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,2.0)$
-                            & $\ancestralTheta{} \sim U(0, 0.05)$
-                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-        \midrule
-            \powerSeriesUniform & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.2)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.4)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.6)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.8)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,1.0)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,2.0)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-        \midrule
-            \powerSeriesExp & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(17.3205)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(8.6603)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(5.7735)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(4.3301)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(3.4641)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-                            & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(1.7321)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
-        \bottomrule
-    \end{tabular}
-    \label{tabPowerModels}
-\end{table}
+\input{dpp-msbayes-tables.tex}
 
@@
 
-%% Validation accuracy
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelDPP}
-}{figValidationAccuracyDPPDPP}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelUniform}
-}{figValidationAccuracyDPPUniform}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelOld}
-}{figValidationAccuracyDPPOld}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelUshaped}
-}{figValidationAccuracyDPPUshaped}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelDPP}
-}{figValidationAccuracyUniformDPP}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelUniform}
-}{figValidationAccuracyUniformUniform}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelOld}
-}{figValidationAccuracyUniformOld}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelUshaped}
-}{figValidationAccuracyUniformUshaped}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelDPP}
-}{figValidationAccuracyOldDPP}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelUniform}
-}{figValidationAccuracyOldUniform}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelOld}
-}{figValidationAccuracyOldOld}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelUshaped}
-}{figValidationAccuracyOldUshaped}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelDPP}
-}{figValidationAccuracyUshapedDPP}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelUniform}
-}{figValidationAccuracyUshapedUniform}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelOld}
-}{figValidationAccuracyUshapedOld}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelUshaped}
-}{figValidationAccuracyUshapedUshaped}
-
-%% Validation model choice
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelDPP}
-}{figValidationModelChoiceDPPDPP}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelUniform}
-}{figValidationModelChoiceDPPUniform}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelOld}
-}{figValidationModelChoiceDPPOld}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelUshaped}
-}{figValidationModelChoiceDPPUshaped}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelDPP}
-}{figValidationModelChoiceUniformDPP}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelUniform}
-}{figValidationModelChoiceUniformUniform}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelOld}
-}{figValidationModelChoiceUniformOld}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelUshaped}
-}{figValidationModelChoiceUniformUshaped}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelDPP}
-}{figValidationModelChoiceOldDPP}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelUniform}
-}{figValidationModelChoiceOldUniform}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelOld}
-}{figValidationModelChoiceOldOld}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelUshaped}
-}{figValidationModelChoiceOldUshaped}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelDPP}
-}{figValidationModelChoiceUshapedDPP}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelUniform}
-}{figValidationModelChoiceUshapedUniform}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelOld}
-}{figValidationModelChoiceUshapedOld}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelUshaped}
-}{figValidationModelChoiceUshapedUshaped}
-
-%% Validation ordered models
-\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPPOrdered}{\modelDPPOrdered}
-}{figValidationAccuracyDPPOrderedDPPOrdered}
-
-\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPPOrdered}{\modelDPPOrdered}
-}{figValidationModelChoiceDPPOrderedDPPOrdered}
-
-%% Power accuracy results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerAccuracyExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerAccuracyExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelOld}
-}{figPowerAccuracyExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerAccuracyExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerAccuracyUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerAccuracyUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerAccuracyUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerAccuracyUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerAccuracyOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerAccuracyOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelOld}
-}{figPowerAccuracyOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerAccuracyOldUshaped}
-
-%% Power psi results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerPsiExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerPsiExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelOld}
-}{figPowerPsiExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerPsiExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerPsiUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerPsiUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerPsiUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerPsiUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerPsiOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerPsiOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelOld}
-}{figPowerPsiOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerPsiOldUshaped}
-
-%% Power dispersion results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerDispersionExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerDispersionExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelOld}
-}{figPowerDispersionExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerDispersionExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerDispersionUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerDispersionUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerDispersionUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerDispersionUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerDispersionOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerDispersionOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelOld}
-}{figPowerDispersionOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerDispersionOldUshaped}
-
-%% Power psi probability results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerPsiProbExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerPsiProbExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelOld}
-}{figPowerPsiProbExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerPsiProbExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerPsiProbUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerPsiProbUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerPsiProbUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerPsiProbUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerPsiProbOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerPsiProbOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelOld}
-}{figPowerPsiProbOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerPsiProbOldUshaped}
-
-%% Power dispersion prob results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerDispersionProbExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerDispersionProbExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelOld}
-}{figPowerDispersionProbExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerDispersionProbExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerDispersionProbUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerDispersionProbUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerDispersionProbUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerDispersionProbUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerDispersionProbOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerDispersionProbOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelOld}
-}{figPowerDispersionProbOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerDispersionProbOldUshaped}
+\input{dpp-msbayes-figures.tex}
 

### 2013-10-18 manuscripts/reply/reply-main.tex
Incorporating comments from Mark.

@@
-Recently, this journal has served as a medium for a
-discussion of the potential pitfalls of approximate Bayesian methods of
-comparative phylogeographical model choice and prior selection.
+Recently, the journal \emph{Evolution} has served as a medium for a discussion
+of the potential pitfalls of approximate Bayesian methods of comparative
+phylogeographical model choice and prior selection.
 The discussion has centered around the approximate Bayesian model choice method,
@@
 clustered divergence times across taxon pairs.
-\citet{Hickerson2013} published a response to this paper where they present
-an empirical Bayes model-averaging approach that they conclude circumvents the
-poor behavior of the method revealed by \citet{Oaks2012}.
+\citet{Hickerson2013} published a response to \citet{Oaks2012} where they
+present an empirical Bayes model-averaging approach that they conclude
+circumvents the poor behavior of the method revealed by \citet{Oaks2012}.
 These papers are largely in agreement, with \citet{Hickerson2013} reiterating
@@
 models with more parameter space is a primary mechanism by which uniform priors
-are causing the poor behavior of \msb.
+are causing the bias of \msb.
 
@@
 approach.
-Our results demonstrate that the model-averaging approach of
+Our results demonstrate that the empirical Bayesian approach of
 \citet{Hickerson2013} is biased toward clustered divergence models and
@@
 \begin{linenomath}
-Bayesian inference is a rational method of inductive learning in which Bayes'
-rule is used to update our beliefs about a model as new information becomes
-available.
+Bayesian inference is a method of inductive learning in which Bayes' rule is
+used to update our beliefs about a model as new information becomes available.
 If we let \allParameterValues represent the set of all possible parameter
@@
 empirical Bayesian estimates of the posterior are often too narrow, off-center,
-and misshapen;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
+and incorrectly shaped;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
 Nevertheless, empirical Bayes methods can provide a powerful means of parameter
@@
 model-averaging framework.
-This is especially concerning in \msb, which has already been demonstrated to
-be biased away from models with more parameter space \citep{Oaks2012}.
+% This is especially concerning in \msb, which has already been demonstrated to
+% be biased away from models with more parameter space \citep{Oaks2012}.
 The consequence of this approach is a high risk of obtaining a model-averaged
@@
 it very difficult for investigators to express their prior uncertainty without
-putting a lot of prior density in regions of improbable parameter space.
+putting high prior density in regions of improbable parameter space.
 The alternative, as shown above, is to risk excluding the truth before
@@
 
-Also, \citet{Hickerson2013} present information only on the extreme case of a
+\citet{Hickerson2013} present information only on the extreme case of a
 single divergence event.
@@
 Again, considering \citet{Hickerson2013} are assuming prior knowledge that
-all population pairs diverged within the last coalescent unit, this
-does not seem particularly powerful.
+all population pairs diverged within the last coalescent unit, the
+method has low power to detect random variation in divergence times.
 
@@
 of samples.
-Thus, it is difficult to explain the method's propensity toward the model with
-the smallest parameter space via insufficient prior sampling.
+% Thus, it is difficult to explain the method's propensity toward the model with
+% the smallest parameter space via insufficient prior sampling.
 While analyses that sample each model proportional to their relative parameter
@@
 a huge number of parameters and model probabilities with scant information
-from the data, it is not surprising that adding an additional layer of
+from the data, it is not surprising that adding another dimension of
 model choice to the method causes problems.
@@
 
-Importantly, there is no cost to this strategy.
+% Importantly, this strategy is easy to implement.
 Given that ABC methods only need to draw random values from prior distributions

### 2013-10-18 manuscripts/reply/reply.tex
Incorporating comments from Mark.

@@
     priors.
-    \citet{Hickerson2013} have responded to this work, criticising the priors
+    \citet{Hickerson2013} have responded to this work, questioning the priors
     used by \citet{Oaks2012} and presenting an approximate Bayesian

### 2013-11-12 empirical-analyses/bin/plot-results.py
Updating plotting of empirical results.
Adding plotting of empirical results under old msbayes implementation
with priors comparable to results under new implementation. Updating
plotting script, plots, and adding new plots with letter labels.

@@ -25,18 +25,23 @@ def get_dpp_psi_values(num_elements, shape, scale, num_sims = 100000):
         psis.append(len(set(x)))
     return psis
 
-def create_plots(info_path, out_dir):
+def create_plots(dpp_info_path, old_info_path, out_dir):
     # matplotlib.rc('text',**{'usetex': True})
-    old = ([1] * 992) + ([2] * 8)
+    # old = ([1] * 992) + ([2] * 8)
     if not os.path.exists(out_dir):
         os.mkdir(out_dir)
-    dmc_sim = DMCSimulationResults(info_path)
-    result_prefix = dmc_sim.get_result_path_prefix(1, 1, 1) + '99-'
-    psi_path = result_prefix + 'psi-results.txt'
+    dmc_sim = DMCSimulationResults(dpp_info_path)
+    dmc_sim_old = DMCSimulationResults(old_info_path)
+    psi_path = dmc_sim.get_result_path_prefix(1, 1, 1) + '99-psi-results.txt'
+    psi_path_old = dmc_sim_old.get_result_path_prefix(1, 1, 1) + '99-psi-results.txt'
     psis = []
     for d in spreadsheet_iter([psi_path]):
         n = int(round(10000 * float(d['estimated_prob'])))
         psis.extend([int(d['num_of_div_events'])] * n)
+    psis_old = []
+    for d in spreadsheet_iter([psi_path_old]):
+        n = int(round(10000 * float(d['estimated_prob'])))
+        psis_old.extend([int(d['num_of_div_events'])] * n)
     bins = range(1, dmc_sim.num_taxon_pairs + 2)
     hd = HistData(x = psis,
             normed = True,
@@ -45,7 +50,8 @@ def create_plots(info_path, out_dir):
             align = 'mid',
             orientation = 'vertical',
             zorder = 0)
-    hd_old= HistData(x = old,
+    # hd_old= HistData(x = old,
+    hd_old= HistData(x = psis_old,
             normed = True,
             bins = bins,
             histtype = 'bar',
@@ -83,9 +89,11 @@ def create_plots(info_path, out_dir):
     pg.reset_figure()
     pg.savefig(os.path.join(out_dir, 'philippines-dpp-psi-posterior.pdf'))
 
-    hist.set_ylim(bottom = 0.0, top = 1.0)
[…]

### 2013-11-12 empirical-analyses/bin/project_util.py

@@ -11,7 +11,9 @@ CONFIG_DIR = os.path.abspath(os.path.join(PROJECT_DIR, 'configs'))
 RESULTS_DIR = os.path.abspath(os.path.join(PROJECT_DIR, 'results'))
 PHILIPPINES_DIR = os.path.abspath(os.path.join(RESULTS_DIR, 'philippines'))
 PHILIPPINES_DPP_DIR = os.path.abspath(os.path.join(PHILIPPINES_DIR, 'dpp'))
+PHILIPPINES_OLD_DIR = os.path.abspath(os.path.join(PHILIPPINES_DIR, 'old'))
 PHILIPPINES_DPP_INFO = os.path.join(PHILIPPINES_DPP_DIR, 'pymsbayes-results', 'pymsbayes-info.txt')
+PHILIPPINES_OLD_INFO = os.path.join(PHILIPPINES_OLD_DIR, 'pymsbayes-results', 'pymsbayes-info.txt')
 
 
 def main():

### 2013-11-13 manuscripts/reply/reply-main.tex
Working on reply manuscript.
Making miscellaneous edits throughout, starting to move some of the text
to a Supporting Info (SI) section. Adding files which will contain SI
text (`reply-si-main.tex`) and SI figures (`reply-si-figures.tex`).

@@
 However, this is misleading, as \citet{Oaks2012} discuss in detail how they
-selected their priors (see their section ``Specifying and simulating the joint
+selected their priors to reflect the large amount of uncertainty about the
+model's parameters (see their section ``Specifying and simulating the joint
 prior'').
-They mention how the use of uniform prior distributions in \msb for many of the
-model's parameters necessitates awkward broad priors to avoid excluding the
-truth, \textit{a priori}.
+\citet{Oaks2012} further discuss how the use of uniform prior distributions in
+\msb for many of the model's parameters necessitates awkward broad priors to
+avoid excluding the truth, \textit{a priori}.
 What \citet{Hickerson2013} mean by ``poorly selected'' is that
@@
 prior sensitivity of the method and determining the power of \msb under
-real-world conditions with highly informative priors.  
+real-world situations when a large amount of prior information is available. 
 \citet{Oaks2012} discuss
@@
 %limitations
-We explored the propensity of the model-averaging approach of
-\citet{Hickerson2013} to exclude the truth in two ways.
+We explored the propensity of the empirically informed model-averaging approach
+of \citet{Hickerson2013} to exclude the truth in two ways.
 First, we re-analyzed the Philippines dataset using the model-averaging approach
@@
 
-%CWL: This seems to specific, not sure about including
-\section*{Validation analyses}
-% \highLight{Not sure where to put this section}.
-% \highLight{This justifies presenting the unadjusted results in this paper.}
-% \highLight{Is it worth including just for this purpose?}
-Following \citet{Oaks2012}, we characterize the model-choice behavior of the
-model-averaging approach of \citet{Hickerson2013} under the ideal conditions
-where the prior is correct (i.e., the data are generated from parameters drawn
-from the same prior distributions used in the analysis).
-We used the same prior models as above ($M_1$--$M_5$;
-Table~\ref{tabModelChoiceEmpirical}), and generated 50,000 pseudo-replicate
-datasets under this prior (10,000 from each model).
-We used a simulated data structure of eight population pairs, with a single
-1000 base-pair locus sampled from 10 individuals from each population.
-We then analyzed each of these replicate datasets using the same prior,
-retaining 1000 posterior samples.
-Our results are very similar to \citet{Oaks2012}, but we note that they
-are not directly comparable as our simulations contained eight population
-pairs rather than 10 (Figure~\ref{figValidationMCBehavior}).
-We find that the approach of \citet{Hickerson2013} estimates the posterior
-probability of divergence models reasonably well when all assumptions of the
-method are met (i.e., the prior is correct) and the unadjusted posterior
-estimates are used.
-Similar to \cite{Oaks2012}, we find that the regression-adjusted estimates of
-the model probabilities are biased.
-
 \section*{Differing utilities of \numt{} and \vmratio{} in \msb}
@@
 
-\subsection*{Graphical checks of priors}
+\subsection*{Graphical prior comparisons}
 \citet{Hickerson2013} advocate the use of what they call graphical checks of
 prior models.
-This entails generating a small number of random samples from the prior (1000)
+This entails generating a small number (1000) of random samples from the prior
 and plotting the resulting summary statistics in comparison to the observed
 statistics to see if they coincide (see Figure 1 of \citet{Hickerson2013}).
-As we show above, this is a risky strategy, because the resulting plots of this
-approach have little correlation with the appropriateness of priors.
+As we show above, this strategy can be misleading, because the resulting plots
+of this approach have little correlation with the appropriateness of priors.
 Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
@@
 summary statistics consistent with the observed data.
-In fact, when the random draws are tightly clustered around the observed
+In fact, when such random draws are tightly clustered around the observed
 statistics, this can be an indication that the prior is over-fit, as we show
 above (Table~\ref{tabModelChoiceEmpirical} and Figure~\ref{figPCA}).
-Thus, interpreting such plots should be avoided, and the use of posterior
-predictive analyses would be much more informative about the overall fit of
-models.
-
-\subsection*{The validity of \msb estimates}
-\citet{Hickerson2013} claim that \citet{Oaks2012} ``assume that all previous
-\msb results are invalid.''
-With the exception of \citet{Hickerson2013}, we do not claim any previous
-results of \msb are invalid.
-Rather, we conservatively recommend that the common inference of temporally
-clustered divergences
-\citep{Barber2010, Bell2012, Carnaval2009, Chan2011, Daza2010, Hickerson2006,
-    Huang2011, Lawson2010, Leache2007, Plouviez2009, Stone2012, Voje2009},
-when not accompanied with the necessary simulation-based analyses to guide the
-interpretation of such results, should be treated with caution, because the
-method has been shown to spuriously infer clustered divergences over a range of
-prior conditions.
+Thus, using such plots to select priors should be avoided, and the use of
+posterior predictive analyses would be much more informative about the overall
+fit of models.
 
 \subsection*{Saturation of summary statistics}
-\citet{Hickerson2013} claim that our priors ``cause much of the explored
-parameter space to be beyond the threshold of saturation in most mtDNA genes.''
-To explore this possibility, we simulated datasets under prior settings that
-match two of the three priors used by \citet{Oaks2012}: $\meanDescendantTheta{}
-\sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
+\citet{Hickerson2013} claim the priors used by \citet{Oaks2012} ``cause much of
+the explored parameter space to be beyond the threshold of saturation in most
+mtDNA genes.'' To explore this possibility, we simulated datasets under prior
+settings that match two of the three priors used by \citet{Oaks2012}:
+$\meanDescendantTheta{} \sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim
+U(0.0005, 0.02)$.
 Under this prior, we draw divergence time parameters from a uniform
@@
 \msb can be strongly biased away from models with greater parameter space.
-As suggested by \citet{Oaks2012}, this is likely caused by the inappropriate
-use of uniform priors on most of the model's parameters.
-These priors necessitate the use of broad priors that place high prior density
-in unlikely regions of parameter space, less the risk of excluding the truth
-\emph{a priori}.
+As suggested by \citet{Oaks2012}, this is likely caused by the use of uniform
+priors on most of the model's parameters.
+Uniform distributions necessitate the use of broad priors that place high prior
+density in unlikely regions of parameter space, less the risk of excluding the
+truth \emph{a priori}.
 This likely reduces the marginal likelihoods of models with greater parameter
@@
 
+Based on our results, we discourage the use of empirically informed uniform
+priors in Bayesian model choice, especially when combined with model-averaging
+\ldots
+
+We do not ``assumed that all previous \msb results are invalid,'' as suggested
+by \citet{Hickerson2013}.
+Rather, we conservatively recommend that the common inference of temporally
+clustered divergences \citep{Barber2010, Bell2012, Carnaval2009, Chan2011,
+Daza2010, Hickerson2006, Huang2011, Lawson2010, Leache2007, Plouviez2009,
+Stone2012, Voje2009}, when not accompanied with the necessary simulation-based
+analyses to guide the interpretation of such results, should be treated with
+caution, because the method has been shown to spuriously infer clustered
+divergences over a range of prior conditions.
+
 The work presented herein follows the principles of Open Notebook Science.
@@
 \href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
-All information necessary to reproduce our results are provided there.
+All information necessary to reproduce our results is provided there.

### 2013-11-13 manuscripts/reply/reply-si-main.tex
Working on reply manuscript.
Making miscellaneous edits throughout, starting to move some of the text
to a Supporting Info (SI) section. Adding files which will contain SI
text (`reply-si-main.tex`) and SI figures (`reply-si-figures.tex`).

--- 2013-10-18 manuscripts/reply/reply-main.tex
@@
-Recently, this journal has served as a medium for a
-discussion of the potential pitfalls of approximate Bayesian methods of
-comparative phylogeographical model choice and prior selection.
-The discussion has centered around the approximate Bayesian model choice method,
-\msb, which estimates the temporal distribution of divergences among
-co-distributed pairs of taxa \citep{Huang2011} by sampling over all
-unordered models of divergence.
-\citet{Oaks2012} published results of simulation-based power analyses
-showing that \msb can often be biased toward inferring models of temporally
-clustered divergence times across taxon pairs.
-\citet{Hickerson2013} published a response to this paper where they present
-an empirical Bayes model-averaging approach that they conclude circumvents the
-poor behavior of the method revealed by \citet{Oaks2012}.
-These papers are largely in agreement, with \citet{Hickerson2013} reiterating
-much of the discussion of \citet{Oaks2012} regarding the impact of broad
-uniform priors on Bayesian model choice.
-The main difference in perspectives are centered around
-(1) the mechanism by which broad uniform priors cause the poor behavior of
-\msb, and
-(2) how to potentially mitigate this issue.
-
-\citet{Oaks2012} suggest the primary cause of the poor behavior of the method
-is likely the low marginal likelihoods of parameter-rich models integrated over
-vast parameter space with low probability of producing the data, yet relatively
-high prior density \citep{Lindley1957}.
-Note, this suggests the bias is extrinsic to \msb, and the numerical
-approximation machinery of \msb could be sound.
-\citet{Hickerson2013} suggest the bias is intrinsic to \msb, concluding the
-method's rejection algorithm is inefficient and will be increasingly biased as
-the overall space of the model increases, either as a function of the number of
-taxon pairs or the width of the uniform priors on nuisance parameters.
-\citet{Hickerson2013} support their position by giving a probabilistic argument
-that focuses on only one of the three prior models used by \citet{Oaks2012}.
-We show that this argument is based on an empirical Bayesian reasoning and
-questionable assumptions, and does not explain the bias of the method found in
-many of the analyses of \citet{Oaks2012}. 
-Furthermore, we present results of additional analyses that, when combined with
-the results of \citet{Oaks2012}, strongly suggest small marginal likelihoods of
-models with more parameter space is a primary mechanism by which uniform priors
-are causing the poor behavior of \msb.
-
-%This paragraph seems too aggressive. It also seems to repeat some of the previous points. I removed it for now.
-
-To mitigate the bias, \citet{Oaks2012} suggest alternative prior probability
-distributions on divergence models and nuisance parameters could increase the
-likelihoods of models with more divergence time parameters.
-They also describe simulation-based procedures for determining power, accuracy,
-and robustness of the method given a dataset, and recommend any application of
-\msb should be accompanied by such procedures.
-Alternatively, \citet{Hickerson2013} present an approximate Bayesian
-model-averaging approach for accommodating uncertainty in selecting among
-empirically guided priors, and champion the method as a means of avoiding the
-pitfalls raised by \citet{Oaks2012}.
-We discuss errors in this approach that render some of their results invalid
-and leave the remaining results difficult to interpret.
-Furthermore, we follow the recommendations of \citet{Oaks2012} and present
-simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
-approach.
-Our results demonstrate that the model-averaging approach of
-\citet{Hickerson2013} is biased toward clustered divergence models and
-predisposed to excluding true parameter space.
-The model-averaging approach provides another means by which the model can
-``escape'' large parameter space, and the method's bias toward smaller models
-remains.
-This manifestation of the bias often excludes the true parameter space due to
-the use of narrow uniform priors.
-Furthermore we discuss the potential theoretical and practical problems of
-empirical Bayesian model choice.
-% which is used and encouraged by
-% \citet{Hickerson2013} without discussing its potential complications.
-
 %CWL:
@@
 
-% \cwlnote{}{This part seems important, but not related to the error in units
-% discussed above}
-Another important characteristic of the approach used by \citet{Hickerson2013}
-that makes their results difficult to interpret is their use of narrow,
-empirically guided uniform priors in an approximate-Bayesian model-choice
-framework.
-We discuss the risks associated with this approach in the next section, and
-show how the results of \citet{Hickerson2013} likely exclude true values of the
-model's parameters.
-
-% In addition to this error, we were not able to reproduce the results of
-% \citet{Hickerson2013}.
-% We followed their methods, generating $5\e6$ prior samples for each of the
-% eight models in their Table 1.
-% We retained the 10,000 samples from each model with the smallest Euclidean
-% distance to the observed summary statistics, and subsequently retained 10,000
-% samples from the remaining 80,000.
-
-\section*{The potential implications of empirical Bayesian model choice}
-\citet{Hickerson2013} repeatedly refer to the prior distributions used by
-\citet{Oaks2012} as ``poorly selected.''
-However, this is misleading, as \citet{Oaks2012} discuss in detail how they
-selected their priors (see their section ``Specifying and simulating the joint
-prior'').
-They mention how the use of uniform prior distributions in \msb for many of the
-model's parameters necessitates awkward broad priors to avoid excluding the
-truth, \textit{a priori}.
-What \citet{Hickerson2013} mean by ``poorly selected'' is that
-\citet{Oaks2012} did not use their data to inform their initial prior
-distributions (i.e., they did not use an empirical Bayesian approach).
-\citet{Oaks2012} did use empirically informed priors when assessing both the
-prior sensitivity of the method and determining the power of \msb under
-real-world conditions with highly informative priors.  
-\citet{Oaks2012} discuss
-the potential dangers of taking an empirical Bayesian approach to model choice
-(see the last paragraph of ``Assessing prior sensitivity of \msb'' in
-\citet{Oaks2012}), and we expand on this here by exploring both the theoretical
-and practical implications of using empirically informed priors for Bayesian
-model choice.
-
-\subsection*{Theoretical implications of empirical priors for Bayesian model
-choice}
-\begin{linenomath}
-Bayesian inference is a rational method of inductive learning in which Bayes'
-rule is used to update our beliefs about a model as new information becomes
-available.
-If we let \allParameterValues represent the set of all possible parameter
-values, we can define a prior distribution for all $\theta \in
-\allParameterValues$ such that $p(\theta)$ describes our belief that any given
-\myTheta{} is the true value of the parameter.
-If we let \allDatasets represent all possible datasets then we can 
-define a sampling model for all $\theta \in
-\allParameterValues$ and $\alignment{}{} \in \allDatasets$ such that
-$p(\alignment{}{} | \theta)$ measures our belief that any dataset \alignment{}{}
-will be generated by any model state \myTheta{}.
-After collecting a new dataset \alignment{i}{}, we can use Bayes' rule to
-calculate the posterior distribution
-\begin{equation}
-    p(\myTheta{} \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given 
-    \myTheta{})p(\myTheta{})}{p(\alignment{i}{})},
-    \label{eq:bayesrule}
-\end{equation}
-where
-\begin{equation}
-    p(\alignment{i}{}) = \int_{\myTheta{}} p(\alignment{i}{} \given
-    \myTheta{})p(\myTheta{}) d\myTheta{}.
-\end{equation}
-The posterior distribution is a measure of our beliefs after seeing the new
-information.
-\end{linenomath}
-
-This is an elegant method of updating our beliefs as data are accumulated.
-However, this all hinges on the fact that the prior ($p(\myTheta{})$) is
-defined for all possible parameter values independently of the new data being
-analyzed.
-Any other datasets or external information can safely be used to inform our
-beliefs about $p(\myTheta{})$.
-However, if the same data are used to both inform the prior and calculate the
-posterior, the prior becomes conditional on the data, and Bayes' rule breaks
-down.
-
-This is not to say that empirical Bayesian approaches are not useful.
-Empirical Bayes is a well studied branch of Bayesian statistics that
-has given rise to many powerful inference methods.
-However, empirical Bayes methods have an uncertain theoretical basis and
-do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
-empirical Bayesian estimates of the posterior are often too narrow, off-center,
-and misshapen;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
-Nevertheless, empirical Bayes methods can provide a powerful means of parameter
-estimation that often exhibit favorable frequentist properties.
-Furthermore, many post-hoc correction methods have been developed for
-estimating confidence-intervals from empirical Bayes estimates of posterior
-distributions that often exhibit well-behaved frequentist coverage
-probabilities
-\citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.
-
-Whereas empirical Bayes approaches can provide powerful methods for parameter
-estimation, a theoretical justification for empirical Bayes approaches to model
-choice is questionable.
-In Bayesian model choice, the goal is not to estimate parameters, but to
-estimate the relative probabilities of candidate models.
-Unlike parameter estimates, posterior probabilities represent a summary of the
-entire posterior distribution.
-Thus, given that empirical Bayesian posterior distributions are not accurate,
-there is no guarantee regarding the accuracy of probabilities summarized from
-them.
-In addition to these theoretical concerns there are practical dangers to using
-narrow, empirical, uniform priors for a method that has already been shown to
-be biased toward models with less parameter space.
-% %CWL:
-% %I like the example, but it may be too cute. It is also too long. I think a
-% %contrived example is good, but if there could be an example that is more
-% %relevant to model choice and comparison of inaccurate posterior distributions,
-% %I think that would be better. This is an example of empirical Bayes, but not
-% %the potential problem with the use in msbayes. Unless I am interpreting wrong.
-% %I would be careful about making it to cute as well. You may intend to make it
-% %easy to understand for a general audience, but it can also seem like talking
-% %down to Hickerson et al, who we are responding to. It just depends on how you
-% %read it.
-% This can be demonstrated with a simple, albeit contrived, example.
-% Let us say that principal investigator Mary and her new postdoc Will are
-% interested in the hypothesis that the mass of George Washington's periwig
-% renders the quarter dollar of the United States unfair.
-% That is to say their null hypothesis is that the probability of a US
-% quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
-% They can certainly evaluate this hypothesis, they can have undergraduate
-% worker, Joe, flip the coin for them while they tabulate the results.
-% But being Bayesians, before they call Joe into the lab, they agree on a
-% prior probability to place on the set of all possible probabilities that
-% the quarter will land heads when it is flipped.
-% Given that neither of them have a quarter, and their prior knowledge that Joe
-% moonlights as a magician, and is notorious for performing coin and card tricks
-% in the lab, they suspect there is a good chance that the quarter Joe uses will
-% be either two-headed or two-tailed.
-% So, knowing that the beta distribution is the conjugate prior for a binomial
-% likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
-% (Figure~\ref{figCoinFlip}).
-
-% Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
-% begin flipping it.
-% After five tosses, four of which was heads, Joe decides that academics are
-% crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
-% being computational biologists, are satisfied with their
-% empirical dataset of $y = 4$ heads out of $N = 5$ trials.
-% They know from the conjugacy of the beta prior, that the posterior distribution
-% has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
-% in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
-% of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
-% This allows them to plug these values into the beta cumulative distribution
-% function to determine that the posterior probability of their hypothesis is
-% $p(\theta < 0.5 | y=1, N=5) = 0.088$.
-% Given their prior belief and dataset, this indeed is the correct posterior
-% probability of the hypothesis, and Mary and Will should now update their
-% posterior belief accordingly.
-
-% However, reflecting upon the results of their experiment
-% (Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
-% Their prior looks very ``poorly selected,'' and if they had only known that the
-% coin was fair before the data were collected, they would have selected a much
-% better prior.
-% Clearly, from their results, they should have used a prior centered
-% around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
-% much ``better.''
-% Rather than resort to flipping the coin five more times, Mary and Will decide
-% to redo their analysis using the much better ``prior'' of $p(\theta) \sim
-% beta(4.5, 1.5)$.
-% This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
-% probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
-% Now convinced that the US quarter is unfair, albeit not due to the mass of
-% President Washington's head as they hypothesized, Will begins composing an
-% e-mail of complaint addressed to the U.S.\ Mint.
-
-% If Joe's flips were a representative sample, Mary and Will's empirical Bayes
-% estimate might very well be a better point estimate for the parameter
-% \myTheta{}.
-% However, their empirical Bayes estimate of the probability of their hypothesis
-% is incorrect and biased.
-% This simple example shows how parameter estimation is fundamentally different
-% from estimating the probability of a model.
-% While empirically informed priors can be used to obtain well-behaved parameter
-% estimators, using them for model choice is much less certain.
-
-\subsection*{Practical concerns about empirically informed uniform priors for
-    Bayesian model choice}
-% \cwlnote{}{Should this be part of the empirical Bayes section, or something
-%     else?}
-
-% \cwlnote{}{It seems like this section can be boiled down to ``empirical Bayes
-%     approaches for model choice when using uniform priors can exclude the
-%     $truth$ from the prior, which is a violation of Bayes theorem" or something
-%     like that. Picking apart what they calculated and we calculated seems like
-%     too much detail. I would try to simplify how this is being presented. This
-%     can lead into how the use of more appropriate prior distributions should be
-%     used.}
-The results of Hickerson et al.'s \citeyear{Hickerson2013} reanalysis of the
-Philippine dataset strongly favored models with the narrowest, empirically
-informed prior on divergence time, and thus their model-averaged posterior
-estimates are dominated by models $M_1$ and $M_2$ (see Table 1 of
-\citet{Hickerson2013}).
-This is concerning, because the narrowest \divt{} prior used by
-\citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
-divergence times for at least some of the Philippine taxa, a major problem when
-using uniform priors.
-\citet{Hickerson2013} set this prior to match the 95\% highest posterior
-density (HPD) interval for the mean divergence time estimated under one of the
-priors used by \citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
-Given this interval estimate is for the \emph{mean} divergence time across all
-22 taxa, it may be inappropriate to set this as the limit on the prior, because
-many of the taxon pairs are expected to have diverged at times older than the
-upper limit.
-Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
-of the mean divergence time under the other two priors explored by
-\citet{Oaks2012} (under these priors the 95\% HPD is approximately 0.3--0.6;
-see Table~6 of \citet{Oaks2012}).
-Thus, the results of \citet{Hickerson2013} indicate that integration over
-parameter space with low likelihood is biasing the method toward models with
-smaller parameter space, and as a consequence is biasing the method toward
-models that exclude the truth (i.e., toward estimating model-averaged
-posteriors dominated by models that exclude true values of parameters of the
-model).
-
-%CWL:This should be in the section on model-averaging, not empirical Bayes
-%limitations
-We explored the propensity of the model-averaging approach of
-\citet{Hickerson2013} to exclude the truth in two ways.
-First, we re-analyzed the Philippines dataset using the model-averaging approach
-of \citet{Hickerson2013}, but set one of prior models with a uniform prior on
-divergence times that is unrealistically narrow, and almost certainly excludes
-most, if not all, of the true divergence times of the 22 taxon pairs.
-If small likelihoods of large models cause the method to prefer models with
-less parameter space, we expect \msb will preferentially sample from this
-incorrect model yielding a marginal posterior that is incorrect (i.e., the
-model-averaged posterior will be dominated by an incorrect model that excludes
-the truth).
-Second, we generated simulated datasets for which the divergence times are
-drawn from an exponential distribution and applied the approach of
-\citet{Hickerson2013} to each of them to see how often the method excludes the
-truth.
-
-For our re-analysis of the Philippines dataset we used the model-averaging
-approach of \citet{Hickerson2013}, but with a reduced set of prior models to
-avoid mixing time units.
-We used five prior models, all of which had priors on population sizes of
-$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
-U(0.0001, 0.05)$.
-Following \citet{Hickerson2013}, each of these models had the following
-priors on divergence time parameters:
-$M_1$, $\divt{} \sim U(0, 0.1)$;
-$M_2$, $\divt{} \sim U(0, 1)$;
-$M_3$, $\divt{} \sim U(0, 5)$;
-$M_4$, $\divt{} \sim U(0, 10)$; and
-$M_5$, $\divt{} \sim U(0, 20)$.
-We simulated $1\e6$ random samples from each of the models for a total of
-$5\e6$ prior samples.
-For each model, we retained the 10,000 samples with the smallest Euclidean
-distance from the observed summary statistics, standardizing the statistics
-using the prior means and standard deviations of the given model.
-From the remaining 50,000 samples, we then retained the 10,000 samples with the
-smallest Euclidean distance from the observed summary statistics, this time
-standardizing the statistics using the prior means and standard deviations
-across all five models.
-We then repeated this analysis twice, replacing the $M_1$ with
-$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
-times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
-respectively.
-While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
-\citet{Hickerson2013} likely excludes the true divergence times of at least
-some of the 22 taxa, we are nearly certain that these narrower priors are
-incorrect and exclude most, if not all, of the divergence times of the
-Philippine taxa.
-
-Our results show that the model-averaging approach of \citet{Hickerson2013}
-does not reduce the method's bias toward models with less parameter space,
-but rather allows it to manifest in a rather dangerous way.
-The method strongly prefers the prior model with the narrowest distribution on
-divergence times across all three of our analyses, even when this model is
-almost certainly incorrect and excludes the true divergence times of the
-Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
-Unfortunately, ``checking'' the priors by plotting the summary statistics from
-1000 random samples of each prior model along the first two orthogonal axes of
-a principle component analysis, as recommended by \citet{Hickerson2013},
-provides no warning of a problem (Figure~\ref{figPCA}).
-Given that the results of \citet{Hickerson2013} strongly prefer the models with
-the narrowest prior on divergence times, it seems quite likely that their
-model-averaged results are dominated by models that exclude at least some of
-the 22 true divergence times, making their results difficult to interpret
-(in addition to the error of time units).
-
-To better quantify the propensity of Hickerson et al.'s
-(\citeyear{Hickerson2013}) approach to exclude the truth, we simulated 1000
-datasets in which the divergence times for the 22 population pairs are drawn
-randomly from an exponential distribution with a mean of 0.5 ($\divt{} \sim
-Exp(2)$).
-All other parameters were identically distributed as the $M_1$--$M_5$ models
-(Table~\ref{tabModelChoiceEmpirical}).
-We then repeated the analysis described above using $1\e6$ random samples from
-prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, retaining 1000 posterior
-samples for each of the 1000 simulated datasets.
-
-% \highLight{Add how we determined exclusion.}
-For each simulation replicate, we estimated the Bayes factor in favor
-of excluding the truth as the ratio of the posterior to prior odds of
-excluding the true value of at least one parameter.
-Whenever the Bayes factor preferred a model excluding the truth we counted
-the number of 22 true divergence times that were excluded by the preferred
-model.
-Our results show that the model-averaging approach of \citet{Hickerson2013}
-excludes the true values of parameters in 97\% of the replicates (90\% with
-GLM-regression adjustment), excluding up to 21 of the 22 true divergence times
-(Figure~\ref{figExclusionSimTau}).
-We also used the mode estimate of the preferred model for each replicate to
-estimate the number of true values excluded, which produced very similar
-results.
-Furthermore, the posterior probability of excluding at least one true parameter
-value is very high in nearly all of the replicates
-(Figure~\ref{figExclusionSimProb}).
-Using a Bayes factor of greater than 10 as a criterion for strong support, 66\%
-of the replicates (87\% with GLM-regression adjustment) strongly support the
-exclusion of true values (Figure~\ref{figExclusionSimProb}).
-
-The results of our empirical and simulation-based analyses clearly demonstrate
-the danger of using narrow, empirically guided uniform priors in a Bayesian
-model-averaging framework.
-This is especially concerning in \msb, which has already been demonstrated to
-be biased away from models with more parameter space \citep{Oaks2012}.
-The consequence of this approach is a high risk of obtaining a model-averaged
-posterior estimate that is heavily weighted toward incorrect models that
-exclude true values of model parameters.
-
-\subsection*{Additional thoughts on empirical priors in Bayesian model choice}
-Given all of the theoretical and practical issues with empirical Bayes
-approaches to Bayesian model choice discussed in the proceeding sections, it is
-quite clear why one should use caution before overly criticizing an
-investigator's choice of priors after having seen the resulting posterior.
-%In a strictly Bayesian world, this is cheating.
-As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a large
-amount of uncertainty regarding the divergence times of the 22 population pairs
-under study.
-Two of these pairs represent distinct species, and the taxonomy of many groups
-in the Philippines has repeatedly been shown to mask deeply divergent lineages
-\citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
-    Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013}.
-\citet{Oaks2012} also discuss how the sole use of uniform priors in \msb makes
-it very difficult for investigators to express their prior uncertainty without
-putting a lot of prior density in regions of improbable parameter space.
-The alternative, as shown above, is to risk excluding the truth before
-performing the analysis.
-
-While our results above clearly demonstrate the risks inherent to the empirical
-Bayesian model-choice approach used by \citet{Hickerson2013}, one could justify
-such risk if the method does indeed increase the power of the method and thus
-decrease bias toward clustered divergence models, as claimed by
-\citet{Hickerson2013}.
-In the next section, we use simulation-based analyses to evaluate the power of
-Hickerson et al.'s (\citeyear{Hickerson2013}) approach, which show the method is
-still biased toward highly clustered models of divergence.
-
-\section*{Assessing the power of the model-averaging approach of
-    \citet{Hickerson2013}}
-% \cwlnote{}{I think the empirical use of the model-averaging should go here,
-%     instead of above in the empirical Bayes section}
-
-As recommended by \citet{Oaks2012}, we perform simulation-based analyses
-to explore the power of the approximate Bayesian model-averaging approach
-proposed by \citet{Hickerson2013}.
-Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
-the 22 population pairs randomly drawn from a uniform distribution, $U(0,
-\divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and 2.0, in
-\globalcoalunit generations.
-All other parameters were identically distributed as the prior.
-We use the same $5\e6$ samples from the same set of five prior models as above
-($M_1$, $M_2$, $M_3$, $M_4$, and $M_5$).
-As described above, for each simulated dataset, we approximated the posterior
-by retaining 1000 samples from the prior with the smallest Euclidean distance
-from the true summary statistics.
-In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
-posterior samples for each of them.
-
-We find that the approach of \citet{Hickerson2013} struggles to estimate the
-dispersion index of divergence times (\vmratio{}) across most of the \divt{max}
-we simulated, whether evaluating the unadjusted
-(Figure~\ref{figPowerAccOmegaMedian}) or GLM-adjusted
-(Figure~\ref{figPowerAccOmegaModeGLM}) posterior estimates.
-The method only estimates \vmratio{} relatively well when the simulated
-distribution of divergence times is identical to one of the prior models
-(Figures~\ref{figPowerAccOmegaMedian}E \& \ref{figPowerAccOmegaModeGLM}E).
-This is consistent with the conclusion of \citet{Oaks2012} that \msb lacks
-robustness and is highly sensitive to the prior distribution deviating from the
-true, underlying distribution of the data.
-
-Furthermore, our results demonstrate that the approach of \citet{Hickerson2013}
-consistently infers highly clustered divergences across all the \divt{max} we
-simulated (Figure~\ref{figPowerNumExcluded}).
-The method is most likely to infer the extreme case of a single divergence event
-when populations diverged randomly over the past \globalcoalunit generations.
-Even when divergences are random over the past $8\globalpopsize$ generations,
-the most likely inference is only two divergence events, and a single 
-divergence is still estimated in more than 10\% of the replicates.
-It is very interesting to note that as \divt{max} increases, but before the
-estimates are finally pulled away from $\numt{} = 1$, the distribution of
-\numt{} estimates closely mirror the odd U-shaped prior on divergence models
-used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
-(\citeyear{Oaks2012}) Figure 5B).
-This supports the conclusions of \citet{Oaks2012} that this U-shaped
-prior coupled with the poor marginal likelihoods of models with many
-\divt{} parameters, is a major cause of the method's bias toward
-clustered divergence models.
-
-Looking at our simulation results in terms of the posterior probability of the
-dispersion index of divergence times supporting the extreme case of one
-divergence event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the
-method strongly supports one divergence in greater than 27\% of the replicates
-across all the \divt{max} we simulated (Figure~\ref{figPowerOmegaProb}).
-Following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
-the criterion for incorrect inference of a single divergence event.
-There is strong support for a single divergence event in more than 90\% of the
-replicates when divergences are random over the past $2.4\globalpopsize$
-generations, and more than 60\% when over the past $3.2\globalpopsize$
-generations or less (Figure~\ref{figPowerOmegaProb}).
-
-Contrary to Hickerson et al.'s (\citeyear{Hickerson2013}) claim that their
-model-averaging approach ``can discriminate complex multispecies histories and
-correctly reject synchronous divergence, even when discrete divergence times
-differ by much less than \ldots millions of generations,'' our results show
-that method is biased toward supporting the extreme case of a single
-divergence event when populations diverged randomly over the last
-$8\globalpopsize$ generations.
-To put this on the scale roughly consistent with a vertebrate mitochondrial
-locus, assuming a mutation rate of $2\e{-8}$ per site per generation, this
-translates to 5 million generations.
-Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
-million generations.
-
-The results of our power analyses further demonstrate the propensity of
-Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
-parameter values.
-Across all but one of the \divt{max} we simulated, the method excludes the
-truth in a large proportion of replicates, and across many of the \divt{max} it
-will exclude a large proportion of the true divergence time values
-(Figure~\ref{figPowerNumExcluded}).
-The posterior probability of excluding at least on true divergence value is
-also quite high across many of the \divt{max}
-(Figure~\ref{figPowerProbExclusion}).
-Only when the data are identically distributed as one of the prior models does
-the method avoid excluding the truth more than 5\% of the time
-(Figure~\ref{figPowerNumExcluded}E).
-Again, this demonstrates the method's lack of robustness.
-
-Given our results, we conclude that the approach of \citet{Hickerson2013}
-excludes regions of parameter space containing the truth, and lacks power to
-detect random variation in divergence times over a scale of $8\globalpopsize$
-generations or more.
-This roughly translates to millions of generations for mitochondrial loci, and
-hundreds of millions of generations for nuclear loci.
-
-%CWL: I would suggest removing this section
-\section*{The power analysis of \citet{Hickerson2013}}
-\citet{Hickerson2013} present a power analysis in which they find that
-\msb has the power to infer multiple divergence events when divergence times
-are random over hundreds of thousands of generations (rather than millions as
-demonstrated by \citet{Oaks2012}).
-It is not surprising that under limited parameter space the method has
-increased power to detect temporal variation in divergences over narrower time
-windows than millions of generations.
-However, what is important to consider is whether those conditions are relevant
-to real-world applications of the method.
-\citet{Oaks2012} explore the behavior of the method under three divergence time
-priors, as narrow as 0--5 coalescent units. This is quite narrow considering
-that this expresses the prior belief that all 22 taxon pairs diverged within
-this window.
-\citet{Hickerson2013} limit their power analysis to a single prior of 0--1
-coalescent unit.
-This prior setting assumes that there is enough a priori information to
-be 100\% certain that all taxa (18 for their simulations) diverged within
-the last 4$\globalpopsize$ generations.
-This does not seem applicable to most empirical systems.
-Furthermore, even when such extensive prior information is available, to only
-be able to detect multiple divergences on the scale of hundreds of thousands of
-generations does not seem very powerful.
-
-Also, \citet{Hickerson2013} present information only on the extreme case of a
-single divergence event.
-As discussed in \citet{Oaks2012}, it seems odd to consider the estimation
-of two divergence events as a success when divergence is random over hundreds
-of thousands (or millions) of generations.
-In an empirical system such as the Philippines, where island fragmentation has
-occurred at least 10 times over the past several millions of years
-\citep{Haq1987,Rohling1998,Siddall2003,Miller2005}, an estimate where 22 taxa
-share even a handful of divergence events would be of biogeographic interest.
-
-Lastly, \citet{Hickerson2013} translate their results from units of
-\globalcoalunit generations to generations assuming a mutation rate of
-$1.92\e{-8}$ mutations per site per generation.
-If we translate their results to a scale more consistent with rates of mutation
-of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
-by \citet{Hickerson2013} the method can only reliably reject the extreme case
-of one divergence event when divergences are random over a period of more than
-3 million generations.
-Again, considering \citet{Hickerson2013} are assuming prior knowledge that
-all population pairs diverged within the last coalescent unit, this
-does not seem particularly powerful.
-
-\section*{Model likelihoods or insufficient prior sampling as the cause of bias
-in \msb}
-One of the points of disagreement between \citet{Hickerson2013} and
-\citet{Oaks2012} is the underlying mechanism causing the bias toward models
-with clustered divergences (i.e., models with few divergence events).
-\citet{Oaks2012} present simulation-based analyses that suggest the broad
-uniform prior on divergence time parameters decrease the marginal likelihoods
-of models with more divergence time parameters, whereas \citet{Hickerson2013}
-argue that the bias is inherent to the inefficient rejection algorithm
-implemented in \msb.
-
-\citet{Hickerson2013} present an interesting probabilistic argument
- to show that insufficient prior sampling is to blame for the bias.
-They argue the widest of the three priors used by \citet{Oaks2012} was very
-unlikely to produce any prior samples with large numbers of divergence times
-that were consistent with the Philippine data.
-There are a few issues with their argument.
-First, the probabilities they present assume the gene divergence times
-estimated by \citet{Oaks2012} are correct.
-The sole purpose of the estimates of ultrametric gene trees presented by
-\citet{Oaks2012} was to provide a very rough comparison of the gene divergence
-times across the 22 taxa.  These analyses assumed an arbitrary strict clock of
-$2\e{-8}$ for all taxa, and are, of course, subject to estimation error.
-Furthermore, the branch-length units of the gene trees are in millions of
-years, whereas the divergence time prior of \msb is in generations, thus the
-logic of \citet{Hickerson2013} requires the additional assumption that all 22
-Philippine taxa have a generation time of one year.
-Thus, the argument of \citet{Hickerson2013} that divergence time estimates of
-\citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
-questionable, especially given our findings on the behavior of empirically
-informed priors presented above.
-
-Finally, even \emph{if} we make all of these assumptions and assume the gene
-divergence times are estimated without error, the probabilistic argument only
-applies to one of the three different priors used by \citet{Oaks2012}.
-The narrowest prior on divergence times used by \citet{Oaks2012} closely mirrors
-the range of estimates of gene-divergence times, and applying Hickerson et al.'s
-(\citeyear{Hickerson2013}) probability equations demonstrates that the prior
-is densely populated with samples with large numbers of divergence parameters
-that are consistent with the gene divergence estimates.
-Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
-sampling is to blame for the bias, it should be much reduced under the narrow
-prior on \divt{}.
-However, the magnitude of the bias is very similar across all three priors
-explored by \citet{Oaks2012}.
-\citet{Hickerson2013} point out a case where the narrow prior performs
-slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
-However, it is important to note that these results suffered from a bug
-in \msb, and there are many cases after \citet{Oaks2012} corrected the 
-bug where the narrow prior performs slightly worse (see panels D--J of
-Figures 3 and S12).
-
-To disentangle whether model likelihoods or insufficient prior sampling is to
-blame for the biases revealed by \citet{Oaks2012}, we must look at the
-different predictions made by these two phenomena.
-One example, as discussed by \citet{Oaks2012}, is that insufficient prior
-sampling should create higher variance in posterior estimates, and thus it
-should cause analyses to be sensitive to the number of samples drawn from the
-prior.
-\citet{Oaks2012} do not see such sensitivity when they compare prior sample
-sizes of $2\e6$, $5\e6$, and $10^7$.
-
-To explore this prediction further, we repeat the analysis of the Philippines
-dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim U(0,
-10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
-(0.0005, 0.02)$), using a very large prior sample size of $10^8$.
-When we look at the trace of the estimates of the dispersion index of
-divergence times (\vmratio{}) as the prior samples accumulate
-(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or
-GLM-regression-adjusted estimates.
-This suggests that insufficient prior sampling did not play a large
-role in the bias found by \citet{Oaks2012}.
-
-Our results presented above that demonstrate the bias of the model-averaging
-approach of \citet{Hickerson2013} both toward models with narrower \divt{}
-priors (Table~\ref{tabModelChoiceEmpirical} and
-Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
-and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
-\ref{figPowerOmegaProb}) strongly suggest the primary cause is the uniform
-priors reducing the marginal likelihoods of large models.
-In all of our model-averaging analyses, all prior models have the same number
-of samples.
-Thus, it is difficult to explain the method's propensity toward the model with
-the smallest parameter space via insufficient prior sampling.
-While analyses that sample each model proportional to their relative parameter
-space could be explored, it seems much more likely that the broad uniform
-priors are simply inhibiting the marginal likelihoods of spacious models.
-
-As discussed by \citet{Oaks2012}, a straightforward prediction of reduced
-likelihoods of complex models is that they should disappear as the model
-generating the data converges to the prior.
-\citet{Oaks2012} test this prediction by performing 100,000 simulations to
-assess the model-choice behavior of \msb when the prior model is correct.
-They find that the bias actually switches direction (at least for the
-regression-adjusted estimates) and the method tends to underestimate the
-probability of the model with one divergence event (see Figure 4 of
-\citet{Oaks2012}).
-The same prediction is not as straightforward for the insufficient-sampling
-hypothesis.
-Even when the prior is correct, due to the discrete uniform prior on
-\numt{} implemented in \msb, models with larger numbers of divergence
-events (and thus greater parameter space) will still be less densely
-sampled than those with fewer divergence events \citep{Oaks2012}.
-Thus, the results of the simulations of \citet{Oaks2012} are more consistent
-with marginal model likelihoods causing the bias toward models with fewer
-divergence time parameters.
-
-Taken together, the results presented here and in \citet{Oaks2012} support
-reduced marginal likelihoods of large models as a primary mechanism by which
-broad uniform priors cause biases in \msb.
-Nonetheless, posterior sampling error will always be present in any numerical
-Bayesian approximation method.
-Thus, insufficient sampling of the prior will contribute to the error of all
-approximate Bayesian estimates.
-However, there is no strong evidence that it is playing a large role in
-the biases revealed herein and by \citet{Oaks2012}.
-
 %CWL: This seems to specific, not sure about including
@@
 the model probabilities are biased.
-
-\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
-The main parameter of the \msb model is the vector of divergence
-times for each of the taxon pairs,
-$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
-\citep{Oaks2012}.
-\citet{Hickerson2013} argue that the dispersion index of this vector,
-\vmratio{}, is a better model-choice estimator than the number of 
-divergence time parameters within the vector,
-\numt{}.
-They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
-\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
-This plot shows, not surprisingly, that \vmratio{} has very little information
-about the number of divergences among taxa.
-Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
-biogeographically relevant than \numt{}.
-% We struggle to follow this logic.
-Certainly the maximum information is contained within the divergence time
-vector that \vmratio{} is summarizing.
-Clearly, the number of divergence time parameters within the vector and their
-values is more informative than its variance (i.e., the dispersion index is not
-a sufficient statistic for \divtvector).
-
-\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
-better than \numt{}.''
-However, \msb is a model-choice estimator, and hence the goal is to estimate the
-posterior probability of divergence models.
-\citet{Oaks2012} demonstrate that even when all assumptions of the model are
-met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
-4), whereas \numt{} performs better.
-Furthermore, \vmratio{} is limited to estimating the probability of only a
-single divergence model (the one divergence model), and thus its utility for
-model choice is extremely limited.
-
-The model-choice utility of \vmratio{} is limited to the probability that
-this continuous statistic, which can range from zero to infinity, is at its lower
-limit of zero. In theory, this point density will always be zero, thus an
-arbitrary threshold (0.01 is used throughout the \msb literature) must 
-be chosen to make the probability estimable.
-However, it is still not surprising to see that it is numerically difficult to
-obtain reliable estimates of the probability that the continuous \vmratio{}
-statistic is ``near'' its limit of zero.
-It is much easier, less subjective, and more interpretable to estimate
-the probability of the discrete parameter of the model, \numt{}, is
-at its lower limit of one.
-Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better
-estimator of model probability than \vmratio{}.
-
-\section*{Some general thoughts on the model of \msb}
-Our results demonstrating the poor behavior of the model-averaging approach
-proposed by \citet{Hickerson2013} are not too surprising.
-While this approach is trivial to implement, it is not a trivial change
-to the \msb model.
-To better understand the reason for this, it might help to step back and get a
-sense of the scale of the \msb model.
-To do this, we will use the dataset of 22 vertebrate taxon pairs of
-\citet{Oaks2012} as an example.
-
-Following the model description and notation of \citet{Oaks2012}, let us tally
-up all of the free parameters in the \msb model.  Under the simplest model in
-\msb (i.e., assuming no migration and
-no intra-locus recombination), the number of parameters for each
-taxon pair include:
-The population sizes of the ancestral and descendant populations
-(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
-the magnitude of population contraction in each of the descendant
-populations (\bottleScalar{1}{} and \bottleScalar{2}{}) and the
-timing of these contractions (\bottleTime{}), and the $N-1$ node heights
-(coalescent times) of the gene tree that gave rise to the $N$ gene
-copies sampled from both populations of the pair.
-Because we only have a single locus for each taxon, there are no locus-specific
-\myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
-\locusRateHetShapeParameter shape parameter for the gamma prior distribution on
-\locusMutationRateScalar{}.
-Lastly, there are between one and 22 divergence time parameters \divt{} in
-the vector \divtvector.
-Overall, for our Philippines data under the simplest model in \msb, there are
-581--602 free parameters (depending on the cardinality of \divtvector).
-Furthermore, under this rich model, the method is sampling over 1002 divergence
-models \citep[i.e., the number of integer partitions of $Y=22$;][]{Oaks2012}.
-
-This is a very difficult inference problem, and the method only uses four
-summary statistics calculated from the sequence alignment of each taxon pair:
-$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
-\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
-That gives us a total of 88 summary statistics (four from each of the 22 taxon
-pairs), which contain minimal information about many of the $\approx 600$
-parameters in the model.
-More summary statistics can be used in \msb, but most are highly correlated
-with these four statistics (which are even highly correlated among themselves),
-and thus contribute little additional information about the data.
-
-The large number of parameters and divergence models relative to 
-summary statistics is undoubtedly a key reason the method is so sensitive
-to the prior distributions.
-It also likely contributes to the method's lack of robustness.
-% We use robustness here as a measure of a method's insensitivity to model
-% violations.
-Robustness is an extremely important characteristic of a method to gauge its
-applicability to real-world data, because we know the model and priors will be
-wrong to some degree.
-
-The approach of \citet{Hickerson2013} adds an additional dimension of model
-choice to the model. They expand the model to sample over eight prior models.
-This extends the original model to having 582-603 free parameters and, more
-importantly, sampling over 8016 unique model states across both divergence
-models and prior models.
-This is a non-trivial extension of the model and, given the method's very large
-number of parameters and divergence models relative to the information used
-from the data, likely plays a major role in the poor behavior of this approach.
-% Certainly, such a large modification of the model demands simulation-based
-% assessments of the behavior of the new model, as we provide here.
-
-In theory, the approach of \citet{Hickerson2013} is very appealing.  It sums
-over multiple candidate prior models to produce a posterior estimate
-marginalized over the uncertainty in prior choice.
-In general, Bayesian model-averaging is a powerful approach that leverages a
-great strength of Bayesian statistical procedures, namely the ability to
-obtain marginalized estimates that incorporate uncertainty in nuisance
-parameters.
-However, given the basic \msb model is already struggling to estimate
-a huge number of parameters and model probabilities with scant information
-from the data, it is not surprising that adding an additional layer of
-model choice to the method causes problems.
-
-The recommendations of \citet{Oaks2012} for mitigating the bias and lack of
-robustness of \msb are actually similar to those of \citet{Hickerson2013}, but
-avoid the need for imposing additional model choice.
-\citet{Oaks2012} suggest that uniform priors are inappropriate for many
-parameters of the \msb model, and recommend the use of probability
-distributions from the exponential family.
-If we look at the prior distribution on the divergence time parameter \divt{}
-imposed by the model-averaging approach of \citet{Hickerson2013} we see
-it is a mixture of overlapping uniforms with lower limits of zero
-(Figure~\ref{figMCTauPrior}).
-This looks very much like an exponential distribution, except that each sample
-of divergence times are restricted to the hard bounds of one of the eight prior
-models.
-Thus, it seems more appropriate to place a gamma prior (the exponential being a
-special case of the gamma) on divergence times.
-This would provide investigators flexibility to represent their prior
-uncertainty in model parameters while avoiding broad uniform
-distributions and the need of costly model-averaging.
-This seems like a natural solution to the problem of skewed marginal
-likelihoods and/or sampling error, which are both likely exacerbated by the
-fact that uniform distributions are the only choice of prior for many of the
-parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
-\descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
-\bottleScalar{1}{}, \bottleScalar{2}{}, \locusRateHetShapeParameter,
-\migrationRate{}, \recombinationRate;][]{Oaks2012}.
-
-Importantly, there is no cost to this strategy.
-Given that ABC methods only need to draw random values from prior distributions
-(hence there are no difficult prior and proposal ratios to calculate, etc.) it
-is easy to use the most appropriate distributions on parameters.
-Also, this approach reduces the temptation of empirically guided priors.
-The investigator can place the majority of the prior density in regions of
-parameter space they believe, \emph{a priori}, are most plausible, but still
-capture uncertainty in the tails of distributions with low density.
-As discussed in \citet{Oaks2012} the use of uniform priors necessitate
-broad priors to avoid excluding the truth \emph{a priori}.
-Thus, to avoid the behavior of the method under these uniform priors, using the
-data is a very tempting, albeit risky, option \citep{Hickerson2013}.
-More appropriate prior distributions would alleviate this issue.
-
-%CWL: This seems to nit-picky. Probably not worth including it
-\section*{Other clarifications}
-
-\subsection*{Graphical checks of priors}
-\citet{Hickerson2013} advocate the use of what they call graphical checks of
-prior models.
-This entails generating a small number of random samples from the prior (1000)
-and plotting the resulting summary statistics in comparison to the observed
-statistics to see if they coincide (see Figure 1 of \citet{Hickerson2013}).
-As we show above, this is a risky strategy, because the resulting plots of this
-approach have little correlation with the appropriateness of priors.
-Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
-dataset analyzed by \citet{Hickerson2013}), we do not expect that 1000
-\emph{random} draws from the vast prior parameter space will yield data and
-summary statistics consistent with the observed data.
-In fact, when the random draws are tightly clustered around the observed
-statistics, this can be an indication that the prior is over-fit, as we show
-above (Table~\ref{tabModelChoiceEmpirical} and Figure~\ref{figPCA}).
-Thus, interpreting such plots should be avoided, and the use of posterior
-predictive analyses would be much more informative about the overall fit of
-models.
-
-\subsection*{The validity of \msb estimates}
-\citet{Hickerson2013} claim that \citet{Oaks2012} ``assume that all previous
-\msb results are invalid.''
-With the exception of \citet{Hickerson2013}, we do not claim any previous
-results of \msb are invalid.
-Rather, we conservatively recommend that the common inference of temporally
-clustered divergences
-\citep{Barber2010, Bell2012, Carnaval2009, Chan2011, Daza2010, Hickerson2006,
-    Huang2011, Lawson2010, Leache2007, Plouviez2009, Stone2012, Voje2009},
-when not accompanied with the necessary simulation-based analyses to guide the
-interpretation of such results, should be treated with caution, because the
-method has been shown to spuriously infer clustered divergences over a range of
-prior conditions.
-
-\subsection*{Saturation of summary statistics}
-\citet{Hickerson2013} claim that our priors ``cause much of the explored
-parameter space to be beyond the threshold of saturation in most mtDNA genes.''
-To explore this possibility, we simulated datasets under prior settings that
-match two of the three priors used by \citet{Oaks2012}: $\meanDescendantTheta{}
-\sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
-Under this prior, we draw divergence time parameters from a uniform
-distribution of $U(0, 20)$, simulate datasets, and plot the \divt{} values
-against the summary statistics calculated from the resulting datasets
-(Figure~\ref{figSaturationPlot}).
-Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
-and 10 suffered from little to no effect from saturation.
-Even at divergence times of 20 coalescent units, there is still signal in the
-summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
-Thus, the assertion of \citet{Hickerson2013} does not apply to at least
-two of the priors used by \citet{Oaks2012} and, as a result, does not
-explain the bias they found.
-
-\section*{Conclusions}
-We demonstrate how the approximate Bayesian model choice method implemented in
-\msb can be strongly biased away from models with greater parameter space.
-As suggested by \citet{Oaks2012}, this is likely caused by the inappropriate
-use of uniform priors on most of the model's parameters.
-These priors necessitate the use of broad priors that place high prior density
-in unlikely regions of parameter space, less the risk of excluding the truth
-\emph{a priori}.
-This likely reduces the marginal likelihoods of models with greater parameter
-space, either due to more divergence time parameters or broader prior
-distributions on those parameters.
-We show that the empirical Bayesian model-averaging approach of
-\citet{Hickerson2013} does not mitigate this bias, but rather causes it to
-manifest by sampling predominantly from models that may exclude the true values
-of the parameters.
-Exploring alternative prior probability distributions on most of the model's
-parameters, in addition to different priors over divergence models, should
-help mitigate the method's biases.
-
-The work presented herein follows the principles of Open Notebook Science.
-All aspects of the work were recorded in real-time via version-control
-software and are publicly available at
-\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
-All information necessary to reproduce our results are provided there.

### 2013-11-13 manuscripts/reply/reply.tex
Working on reply manuscript.
Making miscellaneous edits throughout, starting to move some of the text
to a Supporting Info (SI) section. Adding files which will contain SI
text (`reply-si-main.tex`) and SI figures (`reply-si-figures.tex`).

@@
 \input{../utils/macros.tex}
+
+\newcommand{\msTitle}{Why you should not fix a biased model-choice method by
+adding an additional dimension of model choice: A reply to Hickerson et al.
+\xspace}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-        \MakeUppercase{\large\bfseries Why you should not fix a biased
-        model-choice method by adding an additional dimension of model choice:
-        A reply to Hickerson et al.}
+        \MakeUppercase{\large\bfseries \msTitle}
 
@@
         Esselstyn$^{4}$, Charles W.\ Linkem$^{5}$, Cameron D.\
-        Siler$^{6}$, Mark T.\ Holder$^{1}$ and Rafe M.\ Brown$^{1}$
+        Siler$^{6}$, Rafe M.\ Brown$^{1}$ and Mark T.\ Holder$^{1}$
 
@@
 
+% PUT MAIN TEXT CITATION HERE
+{
 \renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape supporting information}}
+\begin{thebibliography}{1}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{\texttt{#1}}
+\providecommand{\urlprefix}{URL }
 
-% PUT MAIN TEXT CITATION HERE
-% \begin{thebibliography}{1}
-% \providecommand{\natexlab}[1]{#1}
-% \providecommand{\url}[1]{\texttt{#1}}
-% \providecommand{\urlprefix}{URL }
+\bibitem[{Oaks et~al.(2014)Oaks, Sukumaran, Esselstyn, Linkem, Siler, Brown,
+  and Holder}]{Oaks2013reply}
+Oaks, J.~R., J.~Sukumaran, J.~A.\ Esselstyn, C.~W.\ Linkem, C.~D.\ Siler,
+    R.~M.\ Brown, and M.~T.\ Holder,
+  % 2013.
+\newblock \msTitle
+% \newblock Evolution 67:991--1010.
 
-% \bibitem
+\end{thebibliography}
+}
 
-% \end{thebibliography}
-
-%% SUPPL TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\setcounter{table}{0}
+\doublespacing
+\input{reply-si-main.tex}
 

### 2013-11-16 manuscripts/reply/reply-main.tex
Updating reply manuscript.
Fleshing out the theoretical implications of empirical Bayesian model
choice.

@@
 Bayesian inference is a method of inductive learning in which Bayes' rule is
-used to update our beliefs about a model as new information becomes available.
+used to update our beliefs about a model $M$ as new information becomes
+available.
 If we let \allParameterValues represent the set of all possible parameter
-values, we can define a prior distribution for all $\theta \in
-\allParameterValues$ such that $p(\theta)$ describes our belief that any given
-\myTheta{} is the true value of the parameter.
+values for model $M$, we can define a prior distribution for all $\theta \in
+\allParameterValues$ such that $p(\theta \given M)$ describes our belief that
+any given \myTheta{} is the true value of the parameter.
 If we let \allDatasets represent all possible datasets then we can 
@@
 \allParameterValues$ and $\alignment{}{} \in \allDatasets$ such that
-$p(\alignment{}{} | \theta)$ measures our belief that any dataset \alignment{}{}
-will be generated by any model state \myTheta{}.
+$p(\alignment{}{} | \theta, M)$ measures our belief that any dataset \alignment{}{}
+will be generated by any state \myTheta{} of model $M$.
 After collecting a new dataset \alignment{i}{}, we can use Bayes' rule to
@@
 \begin{equation}
-    p(\myTheta{} \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given 
-    \myTheta{})p(\myTheta{})}{p(\alignment{i}{})},
+    p(\myTheta{} \given \alignment{i}{}, M) = \frac{p(\alignment{i}{} \given
+    \myTheta{}, M)p(\myTheta{}, M)}{p(\alignment{i}{} \given M)},
     \label{eq:bayesrule}
@@
 \begin{equation}
-    p(\alignment{i}{}) = \int_{\myTheta{}} p(\alignment{i}{} \given
-    \myTheta{})p(\myTheta{}) d\myTheta{}.
+    p(\alignment{i}{} \given M) = \int_{\myTheta{}} p(\alignment{i}{} \given
+    \myTheta{}, M)p(\myTheta{} \given M) d\myTheta{}
+    \label{eq:marginallikelhiood}
 \end{equation}
+is the marginal likelihood of the model.
 The posterior distribution is a measure of our beliefs after seeing the new
@@
 This is an elegant method of updating our beliefs as data are accumulated.
-However, this all hinges on the fact that the prior ($p(\myTheta{})$) is
-defined for all possible parameter values independently of the new data being
-analyzed.
+However, this all hinges on the fact that the prior ($p(\myTheta{} \given M)$)
+is defined for all possible parameter values independently of the new data
+being analyzed.
 Any other datasets or external information can safely be used to inform our
-beliefs about $p(\myTheta{})$.
+beliefs about $p(\myTheta{} \given M)$.
 However, if the same data are used to both inform the prior and calculate the
@@
 
-This is not to say that empirical Bayesian approaches are not useful.
-Empirical Bayes is a well studied branch of Bayesian statistics that
-has given rise to many powerful inference methods.
-However, empirical Bayes methods have an uncertain theoretical basis and
+Thus, empirical Bayes methods have an uncertain theoretical basis and
 do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
@@
 and incorrectly shaped;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
-Nevertheless, empirical Bayes methods can provide a powerful means of parameter
-estimation that often exhibit favorable frequentist properties.
+This is not to say that empirical Bayesian approaches are not useful.
+Empirical Bayes is a well-studied branch of Bayesian statistics that has given
+rise to many powerful inference methods that provide powerful means of
+parameter estimation that often exhibit favorable frequentist properties.
 Furthermore, many post-hoc correction methods have been developed for
@@
 choice is questionable.
-In Bayesian model choice, the goal is not to estimate parameters, but to
-estimate the relative probabilities of candidate models.
+In Bayesian model choice, the primary goal is not to estimate parameters, but
+to estimate the relative probabilities of candidate models.
+In a simple example of Bayesian model choice where two candidate models, $M_1$
+and $M_2$, are being compared, the goal is to estimate the posterior
+probabilities of these two models.
+Again, we can use Bayes' rule to calculate this as
+\begin{equation}
+    p(M_1 \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given
+    M_1)p(M_1)}{p(\alignment{i}{} \given M_1)p(M_1) + p(\alignment{i}{} \given
+    M_2)p(M_2)}.
+    \label{eq:bayesmodelchoice}
+\end{equation}
+By comparing Equations \ref{eq:bayesrule} and \ref{eq:bayesmodelchoice}, we
+can see the fundamental differences between Bayesian parameter estimation and
+model choice.
+In Equation \ref{eq:bayesrule}, we can see that the posterior density of any
+state $\myTheta{}$ of the model, is the prior density updated by the
+probabiliity of the data given the state $\myTheta{}$ (the likelihood of
+$\myTheta$).
+The integral over the entire parameter space of the model, which is defined by
+the priors, only appears as a normalizing constant in the denominator.
+Thus, as long as the joint prior distribution contains the values of $\myTheta$
+that maximize the probability of the data (i.e., the maximum likelihood) and the
+data are strongly informative relative to the prior, the modes of the parameter
+estimates will be relatively robust to prior choice, even if the shape of the
+posterior is incorrect due to using the data to inform the priors.
+This is why empirical Bayes can work well for estimating the values of model
+parameters.
+
+However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
+Bayesian model choice it is now the \emph{marginal} likelihood of a model
+the updates the prior probability of that model to yield posterior
+probability.
+Thus the integral over the entire parameter space of the likelihood weighted by
+the prior probability density is no longer a normalizing constant, but rather
+is what informs the posterior probability of that model from the data.
+As a result, Bayesian model choice tends to be much more senstitive to priors
+than Bayesian parameter estimation.
+Another important difference of Bayesian model choice that is illustrated by
+Equations~\ref{eq:bayesmodelchoice} is that the value of interest, the
+posterior probability of a model, is not a parameter of the model, but rather a
+single value that is a summary of the entire posterior distribution.
+This further complicates the meaning of model posterior probabilities when
+Bayes' rule is violated by empirically informing model priors.
+Model posterior probabilities are inherently measures of our belief in
+the models after our prior beliefs are updated by the data being analyzed.
+Unlike model parameters, there are no true values for model probabilities,
+and there are no measures of uncertainty in the posterior estimates. 
+For empirical Bayesian parameter estimation, the hope is that by giving the
+data more weight relative to the prior, the posterior distribution will be
+peaked near the true value(s) of the model's parameter(s).
+There is no such justification for empirical Bayesian model choice, because
+there are no true model probabilities that we are trying to estimate.
+Rather using the data twice will simply mislead our posterior
+beliefs in the models being compared.
+
+These distinctions between Bayesian parameter estimation and model choice can
+be illustrated with a simple example.
+Let us say we are interested in the probability of a particular coin
+being weighted toward landing heads.
+We already have data from flipping a different coin 20 times
+that landed both heads and tails 10 times, and so we decide to use
+these data in specifying a beta prior on fairness of the new coin of
+$beta(a=10, b=10)$ (Figure~\ref{figCoinFlip}).
+
+% Let us say that principal investigator Mary and her new postdoc Will are
+% interested in the hypothesis that the mass of George Washington's periwig
+% renders the quarter dollar of the United States unfair.
+% That is to say their null hypothesis is that the probability of a US
+% quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
+% They can certainly evaluate this hypothesis, they can have undergraduate
+% worker, Joe, flip the coin for them while they tabulate the results.
+% But being Bayesians, before they call Joe into the lab, they agree on a
+% prior probability to place on the set of all possible probabilities that
+% the quarter will land heads when it is flipped.
+% Given that neither of them have a quarter, and their prior knowledge that Joe
+% moonlights as a magician, and is notorious for performing coin and card tricks
+% in the lab, they suspect there is a good chance that the quarter Joe uses will
+% be either two-headed or two-tailed.
+% So, knowing that the beta distribution is the conjugate prior for a binomial
+% likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
+% (Figure~\ref{figCoinFlip}).
+
+% Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
+% begin flipping it.
+% After five tosses, four of which was heads, Joe decides that academics are
+% crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
+% being computational biologists, are satisfied with their
+% empirical dataset of $y = 4$ heads out of $N = 5$ trials.
+% They know from the conjugacy of the beta prior, that the posterior distribution
+% has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
+% in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
+% of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
+% This allows them to plug these values into the beta cumulative distribution
+% function to determine that the posterior probability of their hypothesis is
+% $p(\theta < 0.5 | y=1, N=5) = 0.088$.
+% Given their prior belief and dataset, this indeed is the correct posterior
+% probability of the hypothesis, and Mary and Will should now update their
+% posterior belief accordingly.
+
+% However, reflecting upon the results of their experiment
+% (Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
+% Their prior looks very ``poorly selected,'' and if they had only known that the
+% coin was fair before the data were collected, they would have selected a much
+% better prior.
+% Clearly, from their results, they should have used a prior centered
+% around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
+% much ``better.''
+% Rather than resort to flipping the coin five more times, Mary and Will decide
+% to redo their analysis using the much better ``prior'' of $p(\theta) \sim
+% beta(4.5, 1.5)$.
+% This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
+% probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
+% Now convinced that the US quarter is unfair, albeit not due to the mass of
+% President Washington's head as they hypothesized, Will begins composing an
+% e-mail of complaint addressed to the U.S.\ Mint.
+
+% If Joe's flips were a representative sample, Mary and Will's empirical Bayes
+% estimate might very well be a better point estimate for the parameter
+% \myTheta{}.
+% However, their empirical Bayes estimate of the probability of their hypothesis
+% is incorrect and biased.
+% This simple example shows how parameter estimation is fundamentally different
+% from estimating the probability of a model.
+% While empirically informed priors can be used to obtain well-behaved parameter
+% estimators, using them for model choice is much less certain.
+
 Unlike parameter estimates, posterior probabilities represent a summary of the
@@
 them.
+
+Equation~\ref{eq:bayesrule} above shows Bayes rule for calculating the
+posterior distributions of the parameter(s) of some model $M$, and thus
+represents Bayesian parameter estimation given model $M$
+We can add the implicit notation for the model $M$ being used to
+the marginal likelihood in Equation~\ref{eq:marginallikelhiood} to get
+\begin{equation}
+    p(\alignment{i}{} \given M) = \int_{\myTheta{}} p(\alignment{i}{} \given
+    \myTheta{}, M)p(\myTheta{} \given M) d\myTheta{}.
+    \label{eq:marginallikelhioodexpanded}
+\end{equation}
+In a simple example of Bayesian model choice where two candidate models, $M_1$
+and $M_2$, are being compared, the posterior probability of $M_1$ is
+\begin{equation}
+    p(M_1 \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given
+    M_1)p(M_1)}{p(\alignment{i}{} \given M_1)p(M_1) + p(\alignment{i}{} \given
+    M_2)p(M_2)}.
+    \label{eq:bayesmodelchoice}
+\end{equation}
+
 In addition to these theoretical concerns there are practical dangers to using

### 2013-11-16 manuscripts/images/coin_flip.r

@@ -1,15 +1,37 @@
 rm(list=ls(all=TRUE))
 
+prior_heads = 10
+prior_tails = 10
+heads = 3
+tails = 7
+
+post_heads = prior_heads + heads
+post_tails = prior_tails + tails
+epost_heads = prior_heads + (2 * heads)
+epost_tails = prior_tails + (2 * tails)
+
 q = seq(0, 1, 0.005)
-prior = dbeta(q, 0.5, 0.5)
-post = dbeta(q, 4.5, 1.5)
-epost = dbeta(q, 8.5, 2.5)
+prior = dbeta(q, prior_heads, prior_tails)
+post = dbeta(q, post_heads, post_tails)
+epost = dbeta(q, epost_heads, epost_tails)
+ymax = max(epost) + (0.02 * max(epost))
+
+post_prob_unfair_heads = pbeta(0.5, post_heads, post_tails,
+                               lower.tail = F)
+epost_prob_unfair_heads = pbeta(0.5, epost_heads, epost_tails,
+                                lower.tail = F)
+post_mode = (post_heads - 1) / (post_heads + post_tails - 2)
+epost_mode = (epost_heads - 1) / (epost_heads + epost_tails - 2)
+cat(paste('post prob unfair heads: ', post_prob_unfair_heads, '\n'))
+cat(paste('emp post prob unfair heads: ', epost_prob_unfair_heads, '\n'))
+cat(paste('post mode: ', post_mode, '\n'))
+cat(paste('emp post mode: ', epost_mode, '\n'))
 
 pdf('coin_flip.pdf', width=7.087, height=4.5)
 par(yaxs="i", xaxs="i", cex.lab=1.2, cex.axis=1.0, mgp=c(2.2, 0.7, 0),
     oma=c(0, 0, 0.2, 0.5), mar=c(3.2, 3.2, 0.5, 0.1))
 plot(x=c(), y = c(), xlab=bquote(theta), ylab='Density', xlim=c(0,1),
-     ylim=c(0, 3.5))
+     ylim=c(0, ymax))
 lines(q, prior)
 lines(q, post, col='blue')
 lines(q, epost, col='red')

### 2013-11-19 manuscripts/reply/reply-main.tex
Working on reply.
Moving error section to SI, and working on Intro.

@@
 Recently, the journal \emph{Evolution} has served as a medium for a discussion
-of the potential pitfalls of approximate Bayesian methods of comparative
-phylogeographical model choice and prior selection.
-The discussion has centered around the approximate Bayesian model choice method,
-\msb, which estimates the temporal distribution of divergences among
-co-distributed pairs of taxa \citep{Huang2011} by sampling over all
-unordered models of divergence.
-\citet{Oaks2012} published results of simulation-based power analyses
-showing that \msb can often be biased toward inferring models of temporally
-clustered divergence times across taxon pairs.
+of potential promises and pitfalls of approximate Bayesian methods of
+comparative phylogeographical model choice and prior selection.
+The discussion has centered around the model-choice method, \msb
+\citep{Huang2011}, which uses approximate Bayesian computation (ABC) to
+estimate the temporal distribution of divergences among co-distributed pairs of
+taxa by sampling over models of divergence that range from a single
+divergence-time parameter (i.e., all pairs of taxa diverged at the same time)
+to the fully generalized model in which each pair is assigned a divergence-time
+parameter.
+A full description of the model is given in \citet{Oaks2012}.
+
+\citet{Oaks2012} published results of simulation-based power analyses showing
+that \msb is biased toward inferring models of temporally clustered divergence
+times across taxon pairs.
 \citet{Hickerson2013} published a response to \citet{Oaks2012} where they
-present an empirical Bayes model-averaging approach that they conclude
-circumvents the poor behavior of the method revealed by \citet{Oaks2012}.
+present an empirical Bayesian model-averaging approach that they conclude
+circumvents the problems revealed by \citet{Oaks2012}.
 These papers are largely in agreement, with \citet{Hickerson2013} reiterating
@@
 uniform priors on Bayesian model choice.
-The main difference in perspectives are centered around
-(1) the mechanism by which broad uniform priors cause the poor behavior of
-\msb, and
-(2) how to potentially mitigate this issue.
+Both papers agree that when broad uniform priors are used to reflect \emph{a
+prior} uncertainty about the divergence times of all pairs of taxa under study,
+the method can be biased toward inferring biogeographically interesting
+scenarios of temporally clustered divergences among taxa (i.e. models with few
+divergence-time parameters).
+Both papers also agree that this bias toward models with few divergence-time
+parameters is due to the much greater parameter space of models with more
+divergence-time parameters.
+However, the papers disgree about the primary mechanism by which the greater
+parameter space of parameter-rich models causes the bias toward clustered
+divergence models;
+\citet{Oaks2012} suggest a statistical issue is the cause, whereas
+\citep{Hickerson2013} prefer a numerical explanation.
+% The main areas of disagreement between the papers are centered around
+% (1) the mechanism by which broad uniform priors cause the bias of \msb toward
+% models with few divergence-time parameters, and
+% (2) how to potentially mitigate this issue.
 
@@
 \citet{Hickerson2013} suggest the bias is intrinsic to \msb, concluding the
-method's rejection algorithm is inefficient and will be increasingly biased as
-the overall space of the model increases, either as a function of the number of
-taxon pairs or the width of the uniform priors on nuisance parameters.
-\citet{Hickerson2013} support their position by giving a probabilistic argument
-that focuses on only one of the three prior models used by \citet{Oaks2012}.
-We show that this argument is based on an empirical Bayesian reasoning and
-questionable assumptions, and does not explain the bias of the method found in
-many of the analyses of \citet{Oaks2012}. 
-Furthermore, we present results of additional analyses that, when combined with
+method's rejection algorithm is inefficient and will be biased
+toward models with fewer divergence-time parameters, because the parameter
+space of these models are more densely sampled relative to models with more
+divergence-time parameters.
+We agree that sampling error is present in all numerical Bayesian estimation
+methods, however, we show here that it is not a major contributer to the
+biases found by \citet{Oaks2012}.
+We present results of additional analyses that, when combined with
 the results of \citet{Oaks2012}, strongly suggest small marginal likelihoods of
@@
 
-%This paragraph seems too aggressive. It also seems to repeat some of the previous points. I removed it for now.
-
-To mitigate the bias, \citet{Oaks2012} suggest alternative prior probability
-distributions on divergence models and nuisance parameters could increase the
-likelihoods of models with more divergence time parameters.
-They also describe simulation-based procedures for determining power, accuracy,
-and robustness of the method given a dataset, and recommend any application of
-\msb should be accompanied by such procedures.
-Alternatively, \citet{Hickerson2013} present an approximate Bayesian
-model-averaging approach for accommodating uncertainty in selecting among
-empirically guided priors, and champion the method as a means of avoiding the
+% We show that this argument is based on empirical Bayesian reasoning and
+% questionable assumptions, and does not explain the bias of the method found in
+% many of the analyses of \citet{Oaks2012}.
+
+There is also some disagreement between the papers also regarding potential
+solutions to the bias.
+\citet{Oaks2012} suggest alternative prior probability distributions on
+divergence-times and other nuisance parameters could increase the marginal
+likelihoods of models with more divergence-time parameters and thus reduce the
+bias.
+Alternatively, \citet{Hickerson2013} present an approach that uses empirically
+informed priors and Bayesian model-averaging in an attempt to accommodate
+uncertainty in selecting priors.
+\citet{Hickerson2013} champion this strategy as a means of avoiding the
 pitfalls raised by \citet{Oaks2012}.
-We discuss errors in this approach that render some of their results invalid
-and leave the remaining results difficult to interpret.
-Furthermore, we follow the recommendations of \citet{Oaks2012} and present
-simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
-approach.
-Our results demonstrate that the empirical Bayesian approach of
-\citet{Hickerson2013} is biased toward clustered divergence models and
-predisposed to excluding true parameter space.
-The model-averaging approach provides another means by which the model can
-``escape'' large parameter space, and the method's bias toward smaller models
-remains.
-This manifestation of the bias often excludes the true parameter space due to
-the use of narrow uniform priors.
+In general, we agree with the use of Bayesian model averaging to obtain a
+posterior estimate marginalized over uncertainty in prior choice.
+However, we question whether adding an additional dimension of model choice
+that samples over empirically informed uniform priors is a fruitful solution
+for an already biased model-choice method.
+In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
+made an error by mixing different time units, which makes the results presented
+in their response difficult to interpret (see Supporting Information for
+details).
+However, here we show that even when that error is avoided, the empirical
+Bayesian model-averaging approach of \citet{Hickerson2013} remains biased
+toward clustered divergence models, and is predisposed to excluding true
+parameter space.
+Their approach provides another means by which the model can ``escape'' large
+parameter space, and the method's bias toward smaller models remains.
+Importantly, this bias can manifest by preferring narrow, empirically informed
+priors that exclude true parameter space.
 Furthermore we discuss the potential theoretical and practical problems of
 empirical Bayesian model choice.
-% which is used and encouraged by
-% \citet{Hickerson2013} without discussing its potential complications.
-
-%CWL:
-%It may be important to point out the error made in their paper, but it can
-%also be nit-picky, it depends on how the error impacts the overall message of
-%our response.  We should strive to keep the response short and directed at the
-%big picture items. If there is a way to still point out the error, then maybe
-%some of this section should be used
-\section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
-\citet{Hickerson2013} re-analyzed the dataset of \citet{Oaks2012} using a
-model-averaging approach, where they placed a discrete uniform prior over eight
-different prior models (see Table 1 of \citet{Hickerson2013}).
-However, there is a fundamental error in the methodology of
-\citet{Hickerson2013}: their model mixes different time units.
-
-Each of the eight prior models used in the re-analysis by \citet{Hickerson2013}
-has one of two priors on the mean size of the descendant populations of each
-taxon pair:
-$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ or
-$\meanDescendantTheta{} \sim U(0.0005, 0.04)$.
-As described in \citet{Oaks2012}, the divergence time parameters of the model
-implemented in \msb are scaled relative to a constant reference population
-size, \myTheta{C}.
-This reference population size is defined in terms of the upper limit of the
-uniform prior on the mean size of the descendant populations,
-\meanDescendantTheta{}, such that for the prior $\meanDescendantTheta{} \sim
-U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
-$\myTheta{C} = \uniformMax{\meanDescendantTheta{}}/2$.
-Thus, the model used by \citet{Hickerson2013} mixes two different units of
-time.
-In other words, some of their prior and posterior samples are in units of
-$0.05/\mutationRate$ generations, whereas others are in units of
-$0.02/\mutationRate$ generations.
-
-The fact that their posterior samples are in different units makes the results
-of \citet{Hickerson2013} difficult to interpret, and renders their
-regression-adjusted results invalid.
-A fundamental assumption of regression is that all of the values of the
-response variable are in the same units.
-Thus, the results in sections ``Using ABC Model Comparison to Weight
-Alternative Priors for the Philippine Vertebrate Data'' and ``Improved Sampling
-Efficiency by Prior Weighting Supports Asynchronous and Recent Divergence for
-the Philippines Vertebrate Data'' and presented in Figure 2 of
-\citet{Hickerson2013} should be disregarded.
-The error is easily illustrated by re-plotting their results with the different
-time units indicated (Figure~\ref{figJointPosteriorHickerson}).
+
+% We discuss errors in this approach that render some of their results invalid
+% and leave the remaining results difficult to interpret.
+% Furthermore, we follow the recommendations of \citet{Oaks2012} and present
+% simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
+% approach.
 
@@
 selected their priors to reflect the large amount of uncertainty about the
-model's parameters (see their section ``Specifying and simulating the joint
-prior'').
+model's parameters for their study system (see their section ``Specifying and
+simulating the joint prior'').
 \citet{Oaks2012} further discuss how the use of uniform prior distributions in
-\msb for many of the model's parameters necessitates awkward broad priors to
-avoid excluding the truth, \textit{a priori}.
+\msb for many of the model's parameters requires investigators to select broad
+priors to avoid excluding the truth \textit{a priori}, and how these broad
+uniform priors place too much density in improbable regions of parameter space.
 What \citet{Hickerson2013} mean by ``poorly selected'' is that
@@
 \citet{Oaks2012} did use empirically informed priors when assessing both the
-prior sensitivity of the method and determining the power of \msb under
-real-world situations when a large amount of prior information is available. 
-\citet{Oaks2012} discuss
+prior sensitivity of the method and the power of \msb under real-world
+situations when a large amount of prior information is available. 
+In doing so, \citet{Oaks2012} discuss
 the potential dangers of taking an empirical Bayesian approach to model choice
@@
 probabiliity of the data given the state $\myTheta{}$ (the likelihood of
-$\myTheta$).
+$\myTheta{}$).
 The integral over the entire parameter space of the model, which is defined by
 the priors, only appears as a normalizing constant in the denominator.
-Thus, as long as the joint prior distribution contains the values of $\myTheta$
+Thus, as long as the joint prior distribution contains the values of $\myTheta{}$
 that maximize the probability of the data (i.e., the maximum likelihood) and the
@@
 is what informs the posterior probability of that model from the data.
-As a result, Bayesian model choice tends to be much more senstitive to priors
+As a result, Bayesian model choice tends to be much more sensitive to priors
 than Bayesian parameter estimation.
@@
 
-\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
-The main parameter of the \msb model is the vector of divergence
-times for each of the taxon pairs,
-$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
-\citep{Oaks2012}.
-\citet{Hickerson2013} argue that the dispersion index of this vector,
-\vmratio{}, is a better model-choice estimator than the number of 
-divergence time parameters within the vector,
-\numt{}.
-They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
-\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
-This plot shows, not surprisingly, that \vmratio{} has very little information
-about the number of divergences among taxa.
-Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
-biogeographically relevant than \numt{}.
-% We struggle to follow this logic.
-Certainly the maximum information is contained within the divergence time
-vector that \vmratio{} is summarizing.
-Clearly, the number of divergence time parameters within the vector and their
-values is more informative than its variance (i.e., the dispersion index is not
-a sufficient statistic for \divtvector).
-
-\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
-better than \numt{}.''
-However, \msb is a model-choice estimator, and hence the goal is to estimate the
-posterior probability of divergence models.
-\citet{Oaks2012} demonstrate that even when all assumptions of the model are
-met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
-4), whereas \numt{} performs better.
-Furthermore, \vmratio{} is limited to estimating the probability of only a
-single divergence model (the one divergence model), and thus its utility for
-model choice is extremely limited.
-
-The model-choice utility of \vmratio{} is limited to the probability that
-this continuous statistic, which can range from zero to infinity, is at its lower
-limit of zero. In theory, this point density will always be zero, thus an
-arbitrary threshold (0.01 is used throughout the \msb literature) must 
-be chosen to make the probability estimable.
-However, it is still not surprising to see that it is numerically difficult to
-obtain reliable estimates of the probability that the continuous \vmratio{}
-statistic is ``near'' its limit of zero.
-It is much easier, less subjective, and more interpretable to estimate
-the probability of the discrete parameter of the model, \numt{}, is
-at its lower limit of one.
-Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better
-estimator of model probability than \vmratio{}.
-
 \section*{Some general thoughts on the model of \msb}

### 2013-11-19 manuscripts/reply/reply-si-main.tex
Working on reply.
Moving error section to SI, and working on Intro.

@@
 the model probabilities are biased.
+
+\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
+The main parameter of the \msb model is the vector of divergence
+times for each of the taxon pairs,
+$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
+\citep{Oaks2012}.
+\citet{Hickerson2013} argue that the dispersion index of this vector,
+\vmratio{}, is a better model-choice estimator than the number of 
+divergence time parameters within the vector,
+\numt{}.
+They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
+\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
+This plot shows, not surprisingly, that \vmratio{} has very little information
+about the number of divergences among taxa.
+Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
+biogeographically relevant than \numt{}.
+% We struggle to follow this logic.
+Certainly the maximum information is contained within the divergence time
+vector that \vmratio{} is summarizing.
+Clearly, the number of divergence time parameters within the vector and their
+values is more informative than its variance (i.e., the dispersion index is not
+a sufficient statistic for \divtvector).
+
+\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
+better than \numt{}.''
+However, \msb is a model-choice estimator, and hence the goal is to estimate the
+posterior probability of divergence models.
+\citet{Oaks2012} demonstrate that even when all assumptions of the model are
+met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
+4), whereas \numt{} performs better.
+Furthermore, \vmratio{} is limited to estimating the probability of only a
+single divergence model (the one divergence model), and thus its utility for
+model choice is extremely limited.
+
+The model-choice utility of \vmratio{} is limited to the probability that
+this continuous statistic, which can range from zero to infinity, is at its lower
+limit of zero. In theory, this point density will always be zero, thus an
+arbitrary threshold (0.01 is used throughout the \msb literature) must 
+be chosen to make the probability estimable.
+However, it is still not surprising to see that it is numerically difficult to
+obtain reliable estimates of the probability that the continuous \vmratio{}
+statistic is ``near'' its limit of zero.
+It is much easier, less subjective, and more interpretable to estimate
+the probability of the discrete parameter of the model, \numt{}, is
+at its lower limit of one.
+Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better
+estimator of model probability than \vmratio{}.

### 2013-11-27 manuscripts/reply/reply-figures.tex
Working on reply ms.
Working on section about theory of empirical Bayesian model choice.

@@
-\mFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
-    The joint posterior of the mean (\meant{}{}) and dispersion index ($\vmratio{} = 
-    \vart{}{}/\meant{}{}$) of divergence times for 22 vertebrate taxon pairs as
-    estimated by \citet{Hickerson2013} (see Figure 2B of \citet{Hickerson2013}).
-    The posterior samples are color-coded to indicate the erroneous mixture of
-    timescales in the analysis of \citet{Hickerson2013};
-    grey = $0.05/\mutationRate$ generations and
-    black = $0.02/\mutationRate$ generations.
-}{figJointPosteriorHickerson}
-
-% \mFigure{../images/coin_flip.pdf}{
-%     A plot of three beta probability density functions that represent a prior
-%     (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
-%     empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
-%     Bernoulli trials, four of which are successes.
-% }{figCoinFlip}
+\mFigure{../images/coin_flip.pdf}{
+    A plot of three beta probability density functions that represent a prior
+    (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
+    empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
+    Bernoulli trials, four of which are successes.
+}{figCoinFlip}
 

### 2013-11-27 manuscripts/reply/reply-main.tex
Working on reply ms.
Working on section about theory of empirical Bayesian model choice.

@@
 Recently, the journal \emph{Evolution} has served as a medium for a discussion
-of potential promises and pitfalls of approximate Bayesian methods of
+of potential promises and pitfalls of approximate-Bayesian methods of
 comparative phylogeographical model choice and prior selection.
@@
 Both papers agree that when broad uniform priors are used to reflect \emph{a
-prior} uncertainty about the divergence times of all pairs of taxa under study,
+prior} uncertainty about divergence times of all pairs of taxa under study,
 the method can be biased toward inferring biogeographically interesting
@@
 
-\citet{Oaks2012} suggest the primary cause of the poor behavior of the method
-is likely the low marginal likelihoods of parameter-rich models integrated over
-vast parameter space with low probability of producing the data, yet relatively
-high prior density \citep{Lindley1957}.
+\citet{Oaks2012} suggest the primary cause of the bias of \msb is the low
+marginal likelihoods of parameter-rich models integrated over vast parameter
+space with low probability of producing the data, yet relatively high prior
+density \citep{Lindley1957} imposed by the uniformly distributed prior on
+divergence times.
 Note, this suggests the bias is extrinsic to \msb, and the numerical
@@
 
-There is also some disagreement between the papers also regarding potential
-solutions to the bias.
+There are also differences between the papers regarding potential solutions to
+the bias.
 \citet{Oaks2012} suggest alternative prior probability distributions on
@@
 uncertainty in selecting priors.
-\citet{Hickerson2013} champion this strategy as a means of avoiding the
-pitfalls raised by \citet{Oaks2012}.
+% \citet{Hickerson2013} champion this strategy as a means of avoiding the
+% pitfalls raised by \citet{Oaks2012}.
 In general, we agree with the use of Bayesian model averaging to obtain a
@@
 for an already biased model-choice method.
+
+In this paper, we discuss the potential theoretical and practical problems of
+using empirically informed priors for Bayesian model choice.
+Furthermore, we evaluate the empirical Bayesian model averaging approach of
+\citet{Hickerson2013} as a potential solution to the biases revealed by
+\citet{Oaks2012}.
 In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
@@
 parameter space.
-Their approach provides another means by which the model can ``escape'' large
-parameter space, and the method's bias toward smaller models remains.
+The approach provides another means by which the method can ``escape'' models
+with large parameter space, and the method's bias toward smaller models
+remains.
 Importantly, this bias can manifest by preferring narrow, empirically informed
 priors that exclude true parameter space.
-Furthermore we discuss the potential theoretical and practical problems of
-empirical Bayesian model choice.
 
@@
 % discussed above}
-Another important characteristic of the approach used by \citet{Hickerson2013}
-that makes their results difficult to interpret is their use of narrow,
-empirically guided uniform priors in an approximate-Bayesian model-choice
-framework.
-We discuss the risks associated with this approach in the next section, and
-show how the results of \citet{Hickerson2013} likely exclude true values of the
-model's parameters.
+% Another important characteristic of the approach used by \citet{Hickerson2013}
+% that makes their results difficult to interpret is their use of narrow,
+% empirically guided uniform priors in an approximate-Bayesian model-choice
+% framework.
+% We discuss the risks associated with this approach in the next section, and
+% show how the results of \citet{Hickerson2013} likely exclude true values of the
+% model's parameters.
 
@@
 selected their priors to reflect the large amount of uncertainty about the
-model's parameters for their study system (see their section ``Specifying and
-simulating the joint prior'').
+model's parameters for their study system (see section ``Specifying and
+simulating the joint prior'' of \citet{Oaks2012}).
 \citet{Oaks2012} further discuss how the use of uniform prior distributions in
@@
 situations when a large amount of prior information is available. 
-In doing so, \citet{Oaks2012} discuss
-the potential dangers of taking an empirical Bayesian approach to model choice
-(see the last paragraph of ``Assessing prior sensitivity of \msb'' in
-\citet{Oaks2012}), and we expand on this here by exploring both the theoretical
-and practical implications of using empirically informed priors for Bayesian
-model choice.
+In doing so, \citet{Oaks2012} discuss the potential dangers of taking an
+empirical Bayesian approach to model choice (see the last paragraph of
+``Assessing prior sensitivity of \msb'' in \citet{Oaks2012}), and we expand on
+this here by exploring both the theoretical and practical implications of using
+empirically informed priors for Bayesian model choice.
 
@@
 Empirical Bayes is a well-studied branch of Bayesian statistics that has given
-rise to many powerful inference methods that provide powerful means of
+rise to many inference methods that provide means of
 parameter estimation that often exhibit favorable frequentist properties.
@@
 to estimate the relative probabilities of candidate models.
-In a simple example of Bayesian model choice where two candidate models, $M_1$
-and $M_2$, are being compared, the goal is to estimate the posterior
-probabilities of these two models.
+In a simple example where two candidate models, $M_1$ and $M_2$, are being
+compared, the goal is to estimate the posterior probabilities of these two
+models.
 Again, we can use Bayes' rule to calculate this as
@@
 By comparing Equations \ref{eq:bayesrule} and \ref{eq:bayesmodelchoice}, we
-can see the fundamental differences between Bayesian parameter estimation and
+see fundamental differences between Bayesian parameter estimation and
 model choice.
-In Equation \ref{eq:bayesrule}, we can see that the posterior density of any
+In Equation \ref{eq:bayesrule}, we see that the posterior density of any
 state $\myTheta{}$ of the model, is the prior density updated by the
-probabiliity of the data given the state $\myTheta{}$ (the likelihood of
+probability of the data given the state $\myTheta{}$ (the likelihood of
 $\myTheta{}$).
@@
 the priors, only appears as a normalizing constant in the denominator.
-Thus, as long as the joint prior distribution contains the values of $\myTheta{}$
-that maximize the probability of the data (i.e., the maximum likelihood) and the
-data are strongly informative relative to the prior, the modes of the parameter
-estimates will be relatively robust to prior choice, even if the shape of the
-posterior is incorrect due to using the data to inform the priors.
+Thus, as long as the joint prior distribution contains the values of
+$\myTheta{}$ that maximize the probability of the data (i.e., the maximum
+likelihood) and the data are strongly informative relative to the prior, the
+modes of the estimated posterior distributions of parameters will be relatively
+robust to prior choice, even if the shape of the posterior is incorrect due to
+using the data to inform the priors.
 This is why empirical Bayes can work well for estimating the values of model
@@
 
+% Unlike parameter estimates, posterior probabilities represent a summary of the
+% entire posterior distribution.
+% Thus, given that empirical Bayesian posterior distributions are not accurate,
+% there is no guarantee regarding the accuracy of probabilities summarized from
+% them.
+
 However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
-Bayesian model choice it is now the \emph{marginal} likelihood of a model
-the updates the prior probability of that model to yield posterior
-probability.
+Bayesian model choice it is now the \emph{marginal} likelihood of a model the
+updates the prior to yield the model's posterior probability.
 Thus the integral over the entire parameter space of the likelihood weighted by
@@
 than Bayesian parameter estimation.
-Another important difference of Bayesian model choice that is illustrated by
-Equations~\ref{eq:bayesmodelchoice} is that the value of interest, the
-posterior probability of a model, is not a parameter of the model, but rather a
-single value that is a summary of the entire posterior distribution.
-This further complicates the meaning of model posterior probabilities when
-Bayes' rule is violated by empirically informing model priors.
-Model posterior probabilities are inherently measures of our belief in
-the models after our prior beliefs are updated by the data being analyzed.
-Unlike model parameters, there are no true values for model probabilities,
-and there are no measures of uncertainty in the posterior estimates. 
-For empirical Bayesian parameter estimation, the hope is that by giving the
+
+Another important difference of Bayesian model choice illustrated by
+Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the
+posterior probability of a model, is not a function of \myTheta{}, because
+it is integrated out of the marginal likelihoods of the candidate models.
+Thus, unlike parameter estimates, the estimated posterior probability of a
+model is a single value (rather than a distribution) lacking a measure of
+posterior uncertainty, and in most cases has no true value.
+In most empirical applications, we know all candidate models are incorrect, but
+our goal is to summarize the relative probabilities that they could produce the
+observed data.
+Model posterior probabilities are inherently measures of our belief in the
+models after our prior beliefs are updated by the data being analyzed.
+
+This complicates the meaning of model posterior probabilities when Bayes' rule
+is violated by empirically informing priors.
+For empirical Bayesian parameter estimation, the objective is that by giving the
 data more weight relative to the prior, the posterior distribution will be
@@
 There is no such justification for empirical Bayesian model choice, because
-there are no true model probabilities that we are trying to estimate.
-Rather using the data twice will simply mislead our posterior
-beliefs in the models being compared.
+there are no true values for the model probabilities being estimated.
+Using the data twice will simply mislead our posterior beliefs in the models
+being compared.
 
@@
 be illustrated with a simple example.
-Let us say we are interested in the probability of a particular coin
-being weighted toward landing heads.
-We already have data from flipping a different coin 20 times
-that landed both heads and tails 10 times, and so we decide to use
-these data in specifying a beta prior on fairness of the new coin of
-$beta(a=10, b=10)$ (Figure~\ref{figCoinFlip}).
+Let us say we are interested in the probability of a particular coin being
+weighted toward landing heads (i.e., $p(p(heads) > 0.5)$).
+We already have data from flipping a different coin 20 times that landed both
+heads and tails 10 times, and so we decide to use these data in specifying a
+beta prior on fairness of the new coin of $beta(a=10, b=10)$
+(Figure~\ref{figCoinFlip}).
+We collect data by flipping the coin $N=10$ times, $y=3$ of which land heads.
+Given the beta distribution is a conjugate prior for a binomial likelihood, the
+posterior distribution has the nice analytical form $\theta \given y,N \sim
+beta(a + y, b + N - y)$, which for the new dataset is simply $beta(13, 17)$;
+this is the true posterior distribution of $\theta$ given our prior information
+and data (Figure~\ref{figCoinFlip}).
+Thus, the posterior mode estimate of the probability of heads is 0.429, and the
+posterior probability of our hypothesis is then $p(p(heads) > 0.5 \given
+heads=3, N=10) = 0.23$.
+To give the data more weight relative to the prior, we could use it twice, and
+calculate an empirical Bayes estimate using a prior of $beta(13,17)$.  This
+results in a distribution of $beta(16, 24)$ (Figure~\ref{figCoinFlip}), with a
+mode of 0.395, and the $p(p(heads) > 0.5 \given heads=3, N=10) = 0.10$.
+Whereas the estimate of the parameter $p(heads)$ is a distribution, the mode of
+which changed only 8\% under the empirically informed prior, the posterior
+probability of our model of interest ($p(p(heads) > 0.5)$) is a single value,
+which changed by 56\% under the empirical prior.
+From this example, we can see how Bayesian model choice is more sensitive
+to priors, and how using the data twice can mislead the posterior
+probability of a model in light of new data.
+
+In addition to these theoretical concerns there are practical problems with
+using narrow, empirical, uniform priors for a method that has already been
+shown to be biased toward models with less parameter space.
 
@@
 
-Unlike parameter estimates, posterior probabilities represent a summary of the
-entire posterior distribution.
-Thus, given that empirical Bayesian posterior distributions are not accurate,
-there is no guarantee regarding the accuracy of probabilities summarized from
-them.
-
-Equation~\ref{eq:bayesrule} above shows Bayes rule for calculating the
-posterior distributions of the parameter(s) of some model $M$, and thus
-represents Bayesian parameter estimation given model $M$
-We can add the implicit notation for the model $M$ being used to
-the marginal likelihood in Equation~\ref{eq:marginallikelhiood} to get
-\begin{equation}
-    p(\alignment{i}{} \given M) = \int_{\myTheta{}} p(\alignment{i}{} \given
-    \myTheta{}, M)p(\myTheta{} \given M) d\myTheta{}.
-    \label{eq:marginallikelhioodexpanded}
-\end{equation}
-In a simple example of Bayesian model choice where two candidate models, $M_1$
-and $M_2$, are being compared, the posterior probability of $M_1$ is
-\begin{equation}
-    p(M_1 \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given
-    M_1)p(M_1)}{p(\alignment{i}{} \given M_1)p(M_1) + p(\alignment{i}{} \given
-    M_2)p(M_2)}.
-    \label{eq:bayesmodelchoice}
-\end{equation}
-
-In addition to these theoretical concerns there are practical dangers to using
-narrow, empirical, uniform priors for a method that has already been shown to
-be biased toward models with less parameter space.
 % %CWL:

### 2013-11-27 manuscripts/reply/reply-si-figures.tex
Working on reply ms.
Working on section about theory of empirical Bayesian model choice.

\mFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
    The joint posterior of the mean (\meant{}{}) and dispersion index ($\vmratio{} = 
    \vart{}{}/\meant{}{}$) of divergence times for 22 vertebrate taxon pairs as
    estimated by \citet{Hickerson2013} (see Figure 2B of \citet{Hickerson2013}).
    The posterior samples are color-coded to indicate the erroneous mixture of
    timescales in the analysis of \citet{Hickerson2013};
    grey = $0.05/\mutationRate$ generations and
    black = $0.02/\mutationRate$ generations.
}{figJointPosteriorHickerson}

### 2013-11-28 manuscripts/reply/reply-main.tex
Working on reply ms.
Working on section covering the theory of empirical Bayesian
model choice.

@@
     p(\myTheta{} \given \alignment{i}{}, M) = \frac{p(\alignment{i}{} \given
-    \myTheta{}, M)p(\myTheta{}, M)}{p(\alignment{i}{} \given M)},
+    \myTheta{}, M)p(\myTheta{} \given M)}{p(\alignment{i}{} \given M)},
     \label{eq:bayesrule}
@@
 However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
-Bayesian model choice it is now the \emph{marginal} likelihood of a model the
+Bayesian model choice it is now the \emph{marginal} likelihood of a model that
 updates the prior to yield the model's posterior probability.
@@
 than Bayesian parameter estimation.
+In other words, the Bayesian model choice is much more sensitive to the marginal
+likelihood of models than parameter estimation.
+Whereas parameter estimation will tend to succeed if the topography of the
+likelihood surface is captured regardless of scale, model choice must compare
+the average height of this surface between models.
 
@@
 model is a single value (rather than a distribution) lacking a measure of
-posterior uncertainty, and in most cases has no true value.
-In most empirical applications, we know all candidate models are incorrect, but
-our goal is to summarize the relative probabilities that they could produce the
-observed data.
-Model posterior probabilities are inherently measures of our belief in the
+posterior uncertainty.
+Also, unlike model parameters, the posterior probabilities of candidate
+models have no clear true values.
+In most applications, we know all candidate models are incorrect, but our goal
+is to summarize the relative probabilities that they could produce the observed
+data.
+Thus, model posterior probabilities are inherently measures of our belief in the
 models after our prior beliefs are updated by the data being analyzed.
@@
 there are no true values for the model probabilities being estimated.
-Using the data twice will simply mislead our posterior beliefs in the models
-being compared.
+By using the data twice, we fail to account for prior uncertainty and mislead
+our posterior beliefs in the models being compared; we will be over confident
+in some models and under confident in others.
 
@@
 be illustrated with a simple example.
-Let us say we are interested in the probability of a particular coin being
-weighted toward landing heads (i.e., $p(p(heads) > 0.5)$).
+Let us say we are interested in the fairness of a particular coin, and we
+denote the probability of landing heads as \myTheta{}.
+More specifically, we are interested in the probability of two models, $M_1$
+and $M_2$.
+In both models the outcomes of flipping the coin are assumed to be binomially
+distributed, but under $M_1$ the coin is weighted toward landing heads (i.e.,
+$\myTheta{} > 0.5)$), where as under $M_2$, the coin is weighted toward landing
+tails (i.e., $\myTheta{} < 0.5$).
 We already have data from flipping a different coin 20 times that landed both
@@
 posterior distribution has the nice analytical form $\theta \given y,N \sim
-beta(a + y, b + N - y)$, which for the new dataset is simply $beta(13, 17)$;
-this is the true posterior distribution of $\theta$ given our prior information
-and data (Figure~\ref{figCoinFlip}).
-Thus, the posterior mode estimate of the probability of heads is 0.429, and the
-posterior probability of our hypothesis is then $p(p(heads) > 0.5 \given
-heads=3, N=10) = 0.23$.
+beta(a + y, b + N - y)$, which for the new dataset is simply $beta(13, 17)$
+(Figure~\ref{figCoinFlip}).
+The posterior mode estimate of the probability of heads is 0.429, and following
+Equation~\ref{eq:marginallikelhiood} the marginal likelihoods of our models of
+interest are
+\begin{equation}
+    p(y=3, N=10 \given M_1) = \int_{0.5}^{1} p(y=3, N=10 \given
+    \myTheta{}, M_1)p(\myTheta{} \given M_1) d\myTheta{} \approx 0.029,
+\end{equation}
+and
+\begin{equation}
+    p(y=3, N=10 \given M_2) = \int_{0}^{0.5} p(y=3, N=10 \given
+    \myTheta{}, M_2)p(\myTheta{} \given M_2) d\myTheta{} \approx 0.097.
+\end{equation}
+Given the models have equal probability under our prior, and following
+Equation~\ref{eq:bayesmodelchoice}, we can calculate the posterior
+probability of Model 1 as
+\begin{equation}
+    p(M_1 \given y=3, N=10) = \frac{p(y=3, N=10 \given M_1)}{p(y=3, N=10 \given
+    M_1) + p(y=3, N=10 \given M_2)} \approx 0.23.
+\end{equation}
+This is the true posterior probability of Model 1 given our prior information
+and data.
+
 To give the data more weight relative to the prior, we could use it twice, and
@@
 results in a distribution of $beta(16, 24)$ (Figure~\ref{figCoinFlip}), with a
-mode of 0.395, and the $p(p(heads) > 0.5 \given heads=3, N=10) = 0.10$.
+mode of 0.395, and $p(M_1 \given y=3, N=10) = 0.10$.
 Whereas the estimate of the parameter $p(heads)$ is a distribution, the mode of
 which changed only 8\% under the empirically informed prior, the posterior
-probability of our model of interest ($p(p(heads) > 0.5)$) is a single value,
-which changed by 56\% under the empirical prior.
-From this example, we can see how Bayesian model choice is more sensitive
-to priors, and how using the data twice can mislead the posterior
-probability of a model in light of new data.
+probability of Model 1 is a single value, which changed by 56\% under the
+empirical prior.
+This demonstrates how Bayesian model choice is more sensitive to priors than
+parameter estimation.
+Also, we can see how using the empirically informed prior ignores our prior
+uncertainty, leading to an underestimate of our posterior uncertainty
+(Figure~\ref{figCoinFlip}), and ultimately misleads us toward overconfidence in
+Model 2.
+Our mode estimate of \myTheta{} under the empirical prior might be closer
+to the true value than our posterior mode.
+Let us assume the truth of the matter is that the coin under study has a broad
+flat edge, and as a result lands on its edge at a certain frequency.
+Hence, both of our models are incorrect (i.e., the outcomes of flipping the
+coin are not binomially distributed).
+From this example, we can see how Bayesian model choice is more sensitive to
+priors, and how using the data twice can mislead the posterior probability of a
+model in light of new data.
 

### 2013-11-28 manuscripts/reply/reply-figures.tex
Working on reply ms.
More work on the section about empirical Bayes model choice theory.

@@
     A plot of three beta probability density functions that represent a prior
-    (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
-    empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
-    Bernoulli trials, four of which are successes.
+    (black; $beta(10, 10)$), true posterior (blue; $beta(13, 17)$), and
+    empirical Bayes density (red; $beta(16, 24)$) for a dataset of 10 Bernoulli
+    trials, three of which are successes.
 }{figCoinFlip}

### 2013-11-28 manuscripts/reply/reply-main.tex
Working on reply ms.
More work on the section about empirical Bayes model choice theory.

@@
 the priors, only appears as a normalizing constant in the denominator.
-Thus, as long as the joint prior distribution contains the values of
-$\myTheta{}$ that maximize the probability of the data (i.e., the maximum
-likelihood) and the data are strongly informative relative to the prior, the
-modes of the estimated posterior distributions of parameters will be relatively
-robust to prior choice, even if the shape of the posterior is incorrect due to
-using the data to inform the priors.
+Thus, as long as the prior distribution contains the values of $\myTheta{}$
+that maximize the probability of the data (i.e., the maximum likelihood) and
+the data are strongly informative relative to the prior, the values of the
+parameters that maximize the posterior distribution will be relatively robust
+to prior choice, even if the posterior distribution is technically incorrect
+due to using the data to inform the priors.
 This is why empirical Bayes can work well for estimating the values of model
@@
 As a result, Bayesian model choice tends to be much more sensitive to priors
-than Bayesian parameter estimation.
-In other words, the Bayesian model choice is much more sensitive to the marginal
-likelihood of models than parameter estimation.
-Whereas parameter estimation will tend to succeed if the topography of the
-likelihood surface is captured regardless of scale, model choice must compare
-the average height of this surface between models.
+than parameter estimation.
 
@@
 models have no clear true values.
-In most applications, we know all candidate models are incorrect, but our goal
-is to summarize the relative probabilities that they could produce the observed
-data.
-Thus, model posterior probabilities are inherently measures of our belief in the
-models after our prior beliefs are updated by the data being analyzed.
+In some simple situations we might suspect that one model is correct and the
+rest are wrong, but in most applications, we know that all the models under
+consideration are oversimplifications of reality.
+Thus, model posterior probabilities are inherently measures of our belief in
+the models after our prior beliefs are updated by the data being analyzed.
 
 This complicates the meaning of model posterior probabilities when Bayes' rule
-is violated by empirically informing priors.
+is violated by informing priors with the same data to be analyzed.
 For empirical Bayesian parameter estimation, the objective is that by giving the
@@
 Let us say we are interested in the fairness of a particular coin, and we
-denote the probability of landing heads as \myTheta{}.
+denote the unknown probability of it landing heads as \myTheta{}.
 More specifically, we are interested in the probability of two models, $M_1$
@@
 distributed, but under $M_1$ the coin is weighted toward landing heads (i.e.,
-$\myTheta{} > 0.5)$), where as under $M_2$, the coin is weighted toward landing
+$\myTheta{} > 0.5)$), whereas under $M_2$, the coin is weighted toward landing
 tails (i.e., $\myTheta{} < 0.5$).
@@
 (Figure~\ref{figCoinFlip}).
-The posterior mode estimate of the probability of heads is 0.429, and following
-Equation~\ref{eq:marginallikelhiood} the marginal likelihoods of our models of
-interest are
+The maximum a posteriori (MAP) estimate of the probability of heads is 0.429,
+and following Equation~\ref{eq:marginallikelhiood} the marginal likelihoods of
+our models of interest are
 \begin{equation}
@@
 \end{equation}
-This is the true posterior probability of Model 1 given our prior information
+This is the correct posterior probability of Model 1 given our prior information
 and data.
@@
 To give the data more weight relative to the prior, we could use it twice, and
-calculate an empirical Bayes estimate using a prior of $beta(13,17)$.  This
-results in a distribution of $beta(16, 24)$ (Figure~\ref{figCoinFlip}), with a
-mode of 0.395, and $p(M_1 \given y=3, N=10) = 0.10$.
-Whereas the estimate of the parameter $p(heads)$ is a distribution, the mode of
-which changed only 8\% under the empirically informed prior, the posterior
-probability of Model 1 is a single value, which changed by 56\% under the
-empirical prior.
-This demonstrates how Bayesian model choice is more sensitive to priors than
-parameter estimation.
-Also, we can see how using the empirically informed prior ignores our prior
-uncertainty, leading to an underestimate of our posterior uncertainty
-(Figure~\ref{figCoinFlip}), and ultimately misleads us toward overconfidence in
-Model 2.
-Our mode estimate of \myTheta{} under the empirical prior might be closer
-to the true value than our posterior mode.
-Let us assume the truth of the matter is that the coin under study has a broad
-flat edge, and as a result lands on its edge at a certain frequency.
-Hence, both of our models are incorrect (i.e., the outcomes of flipping the
-coin are not binomially distributed).
-From this example, we can see how Bayesian model choice is more sensitive to
-priors, and how using the data twice can mislead the posterior probability of a
-model in light of new data.
-
-In addition to these theoretical concerns there are practical problems with
-using narrow, empirical, uniform priors for a method that has already been
-shown to be biased toward models with less parameter space.
+calculate an empirical Bayes estimate using a prior of $beta(13,17)$.
+This results in a distribution of $beta(16, 24)$ (Figure~\ref{figCoinFlip}),
+with a MAP estimate of 0.395, and $p(M_1 \given y=3, N=10) = 0.10$.
+The estimated posterior distribution of the parameter, and resulting MAP
+estimate, is similar whether or not an empirically informed prior is used.
+However, the posterior probability of Model 1 is very sensitive to the
+empirical prior, decreasing by 56\%.
+By using the empirically informed prior, we ignored prior uncertainty, leading
+to an underestimate of our posterior uncertainty (Figure~\ref{figCoinFlip}).
+While this did not greatly affect our estimate of \myTheta{}, it did mislead us
+toward overconfidence in Model 2.
+% Our mode estimate of \myTheta{} under the empirical prior might be closer
+% to the true value than our posterior mode.
+% Let us assume the truth of the matter is that the coin under study has a broad
+% flat edge, and as a result lands on its edge at a certain frequency.
+% Hence, both of our models are incorrect (i.e., the outcomes of flipping the
+% coin are not binomially distributed).
 
@@
 %     used.}
+In addition to the theoretical concerns discussed above, there are practical
+problems with using narrow, empirically informed, uniform priors for a method
+that has already been shown to be biased toward models with less parameter
+space.
 The results of Hickerson et al.'s \citeyear{Hickerson2013} reanalysis of the

### 2013-11-28 manuscripts/reply/reply.tex
Working on reply ms.
Working on section about theory of empirical Bayesian model choice.
author line fix

@@
         Esselstyn$^{4}$, Charles W.\ Linkem$^{5}$, Cameron D.\
-        Siler$^{6}$, Rafe M.\ Brown$^{1}$ and Mark T.\ Holder$^{1}$
+        Siler$^{6}$, and Rafe M.\ Brown$^{1}$
 
@@
 
+\newpage
+\singlespacing
+
+\input{reply-si-figures.tex}
+
 \end{document}

### 2013-12-01 manuscripts/reply/short-reply.tex
added content of main-reply.tex to short-reply.tex

--- 2013-11-28 manuscripts/reply/reply-main.tex
@@
+%&<latex>
+\documentclass[letterpaper,12pt]{article}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\input{../utils/preamble.tex}
+\input{../utils/macros.tex}
+
+\newcommand{\msTitle}{Why you should not fix a biased model-choice method by
+adding an additional dimension of model choice: A reply to Hickerson et al.
+\xspace}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
+\doublespacing
+\raggedright
+\setlength{\parindent}{0.5in}
+\begin{linenumbers}
+
+\begin{titlepage}
+    \begin{flushleft}
+        \sffamily
+
+        \MakeUppercase{\large\bfseries \msTitle}
+
+        \vspace{12pt}
+        \textbf{Running head:} \MakeUppercase{Approximate Bayesian model
+        choice}
+
+        \vspace{12pt}
+        Jamie R.\ Oaks$^{1,2}$, Jeet Sukumaran$^{3}$, Jacob A.\
+        Esselstyn$^{4}$, Charles W.\ Linkem$^{5}$, Cameron D.\
+        Siler$^{6}$, and Rafe M.\ Brown$^{1}$
+
+        \bigskip
+        $^1$\emph{Department of Ecology and Evolutionary Biology,
+            Biodiversity Institute,
+            University of Kansas,
+            Lawrence, Kansas 66045}\\[.1in]
+        $^3$\emph{Department of Biology,
+            Duke University,
+            Durham, North Carolina 27708} \\[.1in]
+        $^4$\emph{Museum of Natural Science,
+            Louisiana State University,
+            119 Foster Hall,
+            Baton Rouge, Louisiana 70803}\\[.1in]
+        $^5$\emph{Department of Biology,
+            University of Washington,
+            Seattle, Washington 98195}\\[.1in]
+        $^6$\emph{Sam Noble Museum,
+            Department of Biology,
+            University of Oklahoma,
+            Norman, Oklahoma 73072}\\[.1in]
+        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
+        joaks1@gmail.com})\\
+
+    \end{flushleft}
+\end{titlepage}
+
+{\sffamily
+    \noindent\textbf{ABSTRACT} \\
+    \noindent Recently, \citet{Oaks2012} presented a simulation-based
+    assessment of the approximate Bayesian phylogeographical model choice
+    method implemented in \msb, and found the method to lack robustness and
+    power to detect temporal variation in divergences that are randomly
+    distributed over relatively broad periods of evolutionary history.
+    They conclude that a likely mechanism behind the bias toward models with
+    few divergence events is the small marginal likelihoods of models with more
+    divergence time parameters due to the integration over vast parameter space
+    with low likeihood and high prior density that is imposed by broad uniform
+    priors.
+    \citet{Hickerson2013} have responded to this work, questioning the priors
+    used by \citet{Oaks2012} and presenting an approximate Bayesian
+    model-averaging approach that uses narrow, empirically informed uniform
+    priors.
+    \citet{Hickerson2013} use this method to re-analyze the empirical dataset
+    of \citet{Oaks2012} comprised of 22 pairs of vertebrate populations from
+    the Philippines, and conclude the model-averaging approach circumvents the
+    issues revealed by \citet{Oaks2012} and yields robust inference of
+    divergence models.
+    Furthermore, they argue that the biases revealed by \citet{Oaks2012} are
+    inherent to the inefficient rejection algorithm implemented in \msb, rather
+    than a result of low marginal likelihoods of parameter-rich models.
+    Here we discuss an error in the methodology of \citet{Hickerson2013} that
+    renders some of their results invalid and the remainder difficult to
+    interpret.
+    Furthermore, we demonstrate that the approach of \citet{Hickerson2013} is
+    biased toward models with less parameter space, which manifests in a strong
+    tendency to sample predominantly from models that exclude the true values
+    of the model's parameters.
+    Our results suggest that their model-averaged results are dominated by
+    models that likely exclude at least some of the true divergence times
+    across the 22 pairs of Philippine vertebrates, which adds to the
+    difficulty in interpreting their results.
+    We also show the approach of \citet{Hickerson2013} is still biased toward
+    inferring simultaneous divergences even when the true divergences are
+    randomly distributed over millions of generations.
+    Our results, coupled with those of \citet{Oaks2012}, strongly suggest that
+    low marginal likelihoods of complex models is a primary cause of the bias
+    of \msb toward models with less parameter space.
+    We discuss ways to mitigate this bias without resorting to adding an
+    additional dimension of model choice to the model or using questionable
+    empirically informed uniform priors.
+
+    \vspace{12pt}
+    \noindent\textbf{KEY WORDS: Approximate Bayesian computation; Bayesian
+        model choice; empirical Bayes}
+
+}
+
+\newpage
+
 Recently, the journal \emph{Evolution} has served as a medium for a discussion
@@
 All information necessary to reproduce our results is provided there.
+
+\section*{Acknowledgments}
+We thank the National Science Foundation for supporting this work (DEB
+1011423).
+J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
+Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
+Society, KU Department of Ecology and Evolutionary Biology, and the KU
+Biodiversity Institute.
+
+\bibliography{../bib/references}
+% \bibliography{references}
+
+%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\singlespacing
+
+\renewcommand\listfigurename{Figure Captions}
+\cftsetindents{fig}{0cm}{2.2cm}
+\renewcommand\cftdotsep{\cftnodots}
+\setlength\cftbeforefigskip{10pt}
+\cftpagenumbersoff{fig}
+\listoffigures
+
+\end{linenumbers}
+
+%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\singlespacing
+
+\input{reply-tables.tex}
+
+\clearpage
+
+%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+
+\input{reply-figures.tex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\setcounter{page}{1}
+
+\singlespacing
+
+% PUT MAIN TEXT CITATION HERE
+{
+\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape supporting information}}
+\begin{thebibliography}{1}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{\texttt{#1}}
+\providecommand{\urlprefix}{URL }
+
+\bibitem[{Oaks et~al.(2014)Oaks, Sukumaran, Esselstyn, Linkem, Siler, Brown,
+  and Holder}]{Oaks2013reply}
+Oaks, J.~R., J.~Sukumaran, J.~A.\ Esselstyn, C.~W.\ Linkem, C.~D.\ Siler,
+    R.~M.\ Brown, and M.~T.\ Holder,
+  % 2013.
+\newblock \msTitle
+% \newblock Evolution 67:991--1010.
+
+\end{thebibliography}
+}
+
+\doublespacing
+\input{reply-si-main.tex}
+
+\newpage
+\singlespacing
+
+\input{reply-si-figures.tex}
+
+\end{document}

### 2013-12-02 manuscripts/reply/short-reply.tex
abstract

@@
 
-\newcommand{\msTitle}{Why you should not fix a biased model-choice method by
-adding an additional dimension of model choice: A reply to Hickerson et al.
+\newcommand{\msTitle}{Nuh uh: A reply to Hickerson et al.
 \xspace}
@@
     \noindent\textbf{ABSTRACT} \\
-    \noindent Recently, \citet{Oaks2012} presented a simulation-based
-    assessment of the approximate Bayesian phylogeographical model choice
-    method implemented in \msb, and found the method to lack robustness and
-    power to detect temporal variation in divergences that are randomly
-    distributed over relatively broad periods of evolutionary history.
-    They conclude that a likely mechanism behind the bias toward models with
-    few divergence events is the small marginal likelihoods of models with more
-    divergence time parameters due to the integration over vast parameter space
-    with low likeihood and high prior density that is imposed by broad uniform
+    \noindent Biogeographers frequently seek to explain population and species
+    differentiation on geographical phenomena.
+    Establishing that a set of population splitting events occurred
+    at the same time can be a potentially persuasive argument that a set of taxa
+    are affected by the same geographic events.
+    Unfortunately, estimating divergence times precisely when one lacks
+    precise information about the rate of molecular evolution is a notoriously
+    difficult problem in evolutionary biology.
+    \citep{Huang2011} introduced an approximate Bayesian approach which
+    can be used to estimate the probability of a model in which multiple
+    sets of taxa diverge at the same time. 
+    This method (implemented in the software \msb) does not require
+    a common set of character data that links all of 
+    the taxa under investigation.
+    Recently, \citet{Oaks2012} used this model choice framework to 
+    study 22 pairs of vertebrate lineages which are distributed across
+    the Philippines; they also studied the behavior of the \msb approach
+    using computer simulations.
+    \citet{Oaks2012} found that the model was very sensitive to the 
+    prior that is chosen and that the inference method had low power
+    to detect variation in divergences times.
+    These results were not surprising, in light a rich statistical literature
+    that shows that the marginal likelihood of a model is quite sensitive to the
+    use of vague priors.
+    Because this sensitivity to prior assumptions affects the crucial insights 
+    that a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended
+    that users of the approach should carefully assess the robustness of their 
+    conclusions to different priors.
+    According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
+    was due to excessively broad prior on divergence times leading to 
+    inadequate numbers of simulation replicates.
+    They proposed a new model-averaging approach that uses narrow, empirically informed uniform
     priors.
-    \citet{Hickerson2013} have responded to this work, questioning the priors
-    used by \citet{Oaks2012} and presenting an approximate Bayesian
-    model-averaging approach that uses narrow, empirically informed uniform
-    priors.
-    \citet{Hickerson2013} use this method to re-analyze the empirical dataset
-    of \citet{Oaks2012} comprised of 22 pairs of vertebrate populations from
-    the Philippines, and conclude the model-averaging approach circumvents the
-    issues revealed by \citet{Oaks2012} and yields robust inference of
-    divergence models.
-    Furthermore, they argue that the biases revealed by \citet{Oaks2012} are
-    inherent to the inefficient rejection algorithm implemented in \msb, rather
-    than a result of low marginal likelihoods of parameter-rich models.
-    Here we discuss an error in the methodology of \citet{Hickerson2013} that
-    renders some of their results invalid and the remainder difficult to
-    interpret.
-    Furthermore, we demonstrate that the approach of \citet{Hickerson2013} is
-    biased toward models with less parameter space, which manifests in a strong
-    tendency to sample predominantly from models that exclude the true values
-    of the model's parameters.
-    Our results suggest that their model-averaged results are dominated by
-    models that likely exclude at least some of the true divergence times
-    across the 22 pairs of Philippine vertebrates, which adds to the
-    difficulty in interpreting their results.
-    We also show the approach of \citet{Hickerson2013} is still biased toward
-    inferring simultaneous divergences even when the true divergences are
-    randomly distributed over millions of generations.
-    Our results, coupled with those of \citet{Oaks2012}, strongly suggest that
-    low marginal likelihoods of complex models is a primary cause of the bias
-    of \msb toward models with less parameter space.
-    We discuss ways to mitigate this bias without resorting to adding an
-    additional dimension of model choice to the model or using questionable
-    empirically informed uniform priors.
-
+    Here we demonstrate that the approach of \citet{Hickerson2013} is
+    dangerous in the sense that the empirically-derived priors often
+    exclude from consideration the true values of the models' parameters.
+    On a more fundamental level, we question the value of adopting an 
+    empirical Bayesian stance for this model choice problem.
+    The robust Bayesian approach of conducting analyses under a variety
+    of priors can reveal prior sensitivity and communicate which assumptions
+    underlie model inference.
     \vspace{12pt}
@@
 \newpage
-
-Recently, the journal \emph{Evolution} has served as a medium for a discussion
-of potential promises and pitfalls of approximate-Bayesian methods of
-comparative phylogeographical model choice and prior selection.
-The discussion has centered around the model-choice method, \msb
-\citep{Huang2011}, which uses approximate Bayesian computation (ABC) to
+\citet{}
+
+\msb
+\citep{Huang2011} uses approximate Bayesian computation (ABC) to
 estimate the temporal distribution of divergences among co-distributed pairs of

### 2013-12-02 manuscripts/reply/short-reply.tex
opening paragraphs

@@
 \newpage
-\citet{}
-
 \msb
 \citep{Huang2011} uses approximate Bayesian computation (ABC) to
-estimate the temporal distribution of divergences among co-distributed pairs of
-taxa by sampling over models of divergence that range from a single
-divergence-time parameter (i.e., all pairs of taxa diverged at the same time)
-to the fully generalized model in which each pair is assigned a divergence-time
-parameter.
+estimate the distribution of divergence times among co-distributed pairs of
+taxa.
+It approximates a posterior probability over models that range from a single
+divergence-time parameter (i.e., simultaneous divergence of all pairs of taxa)
+to the fully generalized model in which each pair of taxa diverged at a
+unique time.
 A full description of the model is given in \citet{Oaks2012}.
-
-\citet{Oaks2012} published results of simulation-based power analyses showing
-that \msb is biased toward inferring models of temporally clustered divergence
-times across taxon pairs.
-\citet{Hickerson2013} published a response to \citet{Oaks2012} where they
-present an empirical Bayesian model-averaging approach that they conclude
-circumvents the problems revealed by \citet{Oaks2012}.
-These papers are largely in agreement, with \citet{Hickerson2013} reiterating
-much of the discussion of \citet{Oaks2012} regarding the impact of broad
-uniform priors on Bayesian model choice.
-Both papers agree that when broad uniform priors are used to reflect \emph{a
-prior} uncertainty about divergence times of all pairs of taxa under study,
-the method can be biased toward inferring biogeographically interesting
-scenarios of temporally clustered divergences among taxa (i.e. models with few
-divergence-time parameters).
-Both papers also agree that this bias toward models with few divergence-time
-parameters is due to the much greater parameter space of models with more
-divergence-time parameters.
-However, the papers disgree about the primary mechanism by which the greater
-parameter space of parameter-rich models causes the bias toward clustered
-divergence models;
-\citet{Oaks2012} suggest a statistical issue is the cause, whereas
-\citep{Hickerson2013} prefer a numerical explanation.
-% The main areas of disagreement between the papers are centered around
-% (1) the mechanism by which broad uniform priors cause the bias of \msb toward
-% models with few divergence-time parameters, and
-% (2) how to potentially mitigate this issue.
-
-\citet{Oaks2012} suggest the primary cause of the bias of \msb is the low
+Frequently, the exact value of the divergence times are hard 
+to determine because researchers typically lack precise knowledge 
+of the rate at which substitutions occur.
+So the exact divergence time is often viewed as a nuisance parameter, and
+researchers focus on the number of divergence events.
+The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
+on the fundamental methodological point about the model selection performed in \msb:
+\begin{itemize}
+   \item the use of vague priors on the divergence time parameters can 
+    lead to support models with few divergence time events. Thus, the 
+    primary inference enabled by the approach is very sensitive to the
+    priors on divergence times.
+\end{itemize}
+
+\citet{Oaks2012} suggest that primary cause of the bias of \msb is the low
 marginal likelihoods of parameter-rich models integrated over vast parameter

### 2013-12-02 manuscripts/reply/short-reply.tex
Working on intro of reply ms.
Editing `short-reply.tex` up to the theory section.

@@
     difficult problem in evolutionary biology.
-    \citep{Huang2011} introduced an approximate Bayesian approach which
-    can be used to estimate the probability of a model in which multiple
-    sets of taxa diverge at the same time. 
-    This method (implemented in the software \msb) does not require
-    a common set of character data that links all of 
-    the taxa under investigation.
-    Recently, \citet{Oaks2012} used this model choice framework to 
+    \citep{Huang2011} introduced an approximate Bayesian approach which can be
+    used to estimate the probability of a model in which multiple sets of taxa
+    diverge at the same time. 
+    This method (implemented in the software \msb) does not require a common
+    set of character data that links all of the taxa under investigation, but
+    does require the relative mutation rates among the taxa and loci to be
+    assumed a priori.
+    Recently, \citet{Oaks2012} used this model-choice framework to 
     study 22 pairs of vertebrate lineages which are distributed across
@@
     to detect variation in divergences times.
-    These results were not surprising, in light a rich statistical literature
+    These results were not surprising, in light of a rich statistical literature
     that shows that the marginal likelihood of a model is quite sensitive to the
@@
     According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
-    was due to excessively broad prior on divergence times leading to 
+    was due to excessively broad priors on divergence times leading to 
     inadequate numbers of simulation replicates.
@@
     On a more fundamental level, we question the value of adopting an 
-    empirical Bayesian stance for this model choice problem.
+    empirical Bayesian stance for this model-choice problem.
     The robust Bayesian approach of conducting analyses under a variety
@@
 
-\citet{Oaks2012} suggest that primary cause of the bias of \msb is the low
+\citet{Oaks2012} suggest the primary cause of the bias of \msb is the low
 marginal likelihoods of parameter-rich models integrated over vast parameter
@@
 divergence times.
-Note, this suggests the bias is extrinsic to \msb, and the numerical
+Note, this is a statistical issue extrinsic to \msb, and the numerical
 approximation machinery of \msb could be sound.
-\citet{Hickerson2013} suggest the bias is intrinsic to \msb, concluding the
-method's rejection algorithm is inefficient and will be biased
-toward models with fewer divergence-time parameters, because the parameter
-space of these models are more densely sampled relative to models with more
-divergence-time parameters.
+\citet{Hickerson2013} suggest the bias is the result of a numerical issue
+intrinsic to \msb, concluding the method's rejection algorithm is inefficient
+and will be biased toward models with fewer divergence-time parameters, because
+the parameter space of these models are more densely sampled relative to models
+with more divergence-time parameters.
 We agree that sampling error is present in all numerical Bayesian estimation
@@
 
-There are also differences between the papers regarding potential solutions to
-the bias.
+% There are also differences between the papers regarding potential solutions to
+% the bias.
 \citet{Oaks2012} suggest alternative prior probability distributions on
@@
 details).
-However, here we show that even when that error is avoided, the empirical
+However, here we show that even when this error is avoided, the empirical
 Bayesian model-averaging approach of \citet{Hickerson2013} remains biased
@@
 remains.
-Importantly, this bias can manifest by preferring narrow, empirically informed
-priors that exclude true parameter space.
+Importantly, this bias can manifest in the preference of models that exclude
+from consideration the true values of the parameters.
 
@@
 \citet{Hickerson2013} repeatedly refer to the prior distributions used by
-\citet{Oaks2012} as ``poorly selected.''
+\citet{Oaks2012} for divergence-time parameters as ``poorly selected.''
 However, this is misleading, as \citet{Oaks2012} discuss in detail how they
-selected their priors to reflect the large amount of uncertainty about the
-model's parameters for their study system (see section ``Specifying and
-simulating the joint prior'' of \citet{Oaks2012}).
+selected their prior to reflect the large amount of uncertainty about the
+timing of divergences across all 22 of the taxa in their study (see section
+``Specifying and simulating the joint prior'' of \citet{Oaks2012}).
 \citet{Oaks2012} further discuss how the use of uniform prior distributions in
@@
 uniform priors place too much density in improbable regions of parameter space.
-What \citet{Hickerson2013} mean by ``poorly selected'' is that
-\citet{Oaks2012} did not use their data to inform their initial prior
-distributions (i.e., they did not use an empirical Bayesian approach).
-\citet{Oaks2012} did use empirically informed priors when assessing both the
-prior sensitivity of the method and the power of \msb under real-world
-situations when a large amount of prior information is available. 
-In doing so, \citet{Oaks2012} discuss the potential dangers of taking an
-empirical Bayesian approach to model choice (see the last paragraph of
-``Assessing prior sensitivity of \msb'' in \citet{Oaks2012}), and we expand on
-this here by exploring both the theoretical and practical implications of using
-empirically informed priors for Bayesian model choice.
+
+\citet{Hickerson2013} suggest that \citet{Oaks2012} should have used their
+data to inform their prior on divergence times (i.e., an empirical Bayesian
+approach).
+However, \citet{Oaks2012} did use empirically informed priors to mimic
+empirical situations in which a large amount of prior information is available;
+they did so to assess the prior sensitivity of the method and to determine if
+uniform priors in such situations are narrow enough to avoid the bias toward
+models with few divergence events.
+The results of \citet{Oaks2012} and \citet{Hickerson2013} showed the method is
+highly sensitive to the prior on divergence times.
+Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
+divergences even under the informed priors.
+The main argument of \citet{Hickerson2013} is that \citet{Oaks2012} did not
+go far enough, and suggest a very narrow, highly informed uniform prior
+on divergence times is necessary to avoid the method's preference for
+models with few divergence-time parameters.
+Such an empirical Bayesian approach to model selection raises some theoretical
+and practical concerns, some of which were discussed by \citet{Oaks2012} (see
+the last paragraph of ``Assessing prior sensitivity of \msb'' in
+\citet{Oaks2012}); we expand on this here.
 

### 2013-12-02 model-choice/bin/generate-samples-for-pc-plot.sh
Modifying `generate-samples-for-pc-plot.sh`.
Modifying to generate prior samples for prior models used
by Oaks et al. 2013.

@@ -31,7 +31,7 @@ fi
 dmc.py --np $nprocs \
     -r 1 \
     -o ../configs/m1.cfg \
-    -p ../configs/m1*.cfg \
+    -p ../configs/m1*.cfg ../configs/m[345].cfg\
     -n $nprior \
     --prior-batch-size $batch_size \
     --num-standardizing-samples $nsums \

### 2013-12-02 model-choice/bin/pc-plots.r
Updating `pc-plots.r`.
Updating script for plotting "graphical checks" based on
principal components.

@@ -16,6 +16,9 @@ observed_path = '../../response-redux/results/sampling-error/pymsbayes-results/o
 m1_path = '../priors-for-pc-plot/pymsbayes-results/pymsbayes-output/prior-stats-summaries/m3-prior-sample.txt'
 m1a_path = '../priors-for-pc-plot/pymsbayes-results/pymsbayes-output/prior-stats-summaries/m2-prior-sample.txt'
 m1b_path = '../priors-for-pc-plot/pymsbayes-results/pymsbayes-output/prior-stats-summaries/m1-prior-sample.txt'
+m3_path = '../priors-for-pc-plot/pymsbayes-results/pymsbayes-output/prior-stats-summaries/m4-prior-sample.txt'
+m4_path = '../priors-for-pc-plot/pymsbayes-results/pymsbayes-output/prior-stats-summaries/m5-prior-sample.txt'
+m5_path = '../priors-for-pc-plot/pymsbayes-results/pymsbayes-output/prior-stats-summaries/m6-prior-sample.txt'
 plot_path = '../priors-for-pc-plot/pc-plots.pdf'
 
 stat_indices = c(94:181)
@@ -27,23 +30,35 @@ m1a = read.table(m1a_path, skip=1)[stat_indices]
 colnames(m1a) = colnames(observed)
 m1b = read.table(m1b_path, skip=1)[stat_indices]
 colnames(m1b) = colnames(observed)
+m3 = read.table(m3_path, skip=1)[stat_indices]
+colnames(m3) = colnames(observed)
+m4 = read.table(m4_path, skip=1)[stat_indices]
+colnames(m4) = colnames(observed)
+m5 = read.table(m5_path, skip=1)[stat_indices]
+colnames(m5) = colnames(observed)
 
 m1_pca <- princomp(m1)
 m1a_pca <- princomp(m1a)
 m1b_pca <- princomp(m1b)
+m3_pca <- princomp(m3)
+m4_pca <- princomp(m4)
+m5_pca <- princomp(m5)
 
 observed_m1 = predict(m1_pca, observed)
 observed_m1a = predict(m1a_pca, observed)
 observed_m1b = predict(m1b_pca, observed)
-plot_letters = LETTERS[1:3]
+observed_m3 = predict(m3_pca, observed)
[…]

### 2013-12-03 model-choice/bin/pc-plots.r
Updating `pc-plots.r` and resulting `pc-plots.pdf`.

@@ -1,9 +1,10 @@
 rm(list=ls(all=TRUE))
 
-plot_pca = function(pca_obj, predict_obj, xlim, ylim) {
+plot_pca = function(pca_obj, predict_obj, xlim, ylim,
+                    xlab = 'PC1', ylab = 'PC2') {
     plot(pca_obj$scores[,1], pca_obj$scores[,2],
-         xlab = 'PC1',
-         ylab = 'PC2',
+         xlab = xlab,
+         ylab = ylab,
          xlim = xlim,
          ylim = ylim)
     points(predict_obj[1], predict_obj[2],
@@ -52,7 +53,7 @@ observed_m4 = predict(m4_pca, observed)
 observed_m5 = predict(m5_pca, observed)
 plot_letters = LETTERS[1:6]
 
-pdf(plot_path, width=7.087, height=5.0)
+pdf(plot_path, width=7.087, height=4.0)
     par(yaxs = "i", xaxs = "i")
     par(cex.axis=0.7)
     par(mgp = c(1.5,0.6,0)) # mgp adjusts space of axis labels
@@ -62,31 +63,31 @@ pdf(plot_path, width=7.087, height=5.0)
 
     xlim = c(-0.4, 0.4)
     ylim = c(-0.2, 0.2)
-    plot_pca(m1_pca, observed_m1, xlim, ylim)
+    plot_pca(m1_pca, observed_m1, xlim, ylim, xlab = '')
     mtext(plot_letters[[1]], side=3, adj=0, line=0, cex=0.75, font=2)
 	mtext(bquote(italic(M)[1]), side=3, adj=0.5, line=0, cex=0.75)
 
-    plot_pca(m1a_pca, observed_m1a, xlim, ylim)
+    plot_pca(m1a_pca, observed_m1a, xlim, ylim, xlab = '', ylab = '')
     mtext(plot_letters[[2]], side=3, adj=0, line=0, cex=0.75, font=2)
 	mtext(bquote(italic(M)['1A']), side=3, adj=0.5, line=0, cex=0.75)
 
-    plot_pca(m1b_pca, observed_m1b, xlim, ylim)
+    plot_pca(m1b_pca, observed_m1b, xlim, ylim, xlab = '', ylab = '')
     mtext(plot_letters[[3]], side=3, adj=0, line=0, cex=0.75, font=2)
[…]

### 2013-12-03 manuscripts/reply/short-reply.tex
Working on empirical Bayes section of `short-reply.tex`.

@@
 divergences even under the informed priors.
-The main argument of \citet{Hickerson2013} is that \citet{Oaks2012} did not
-go far enough, and suggest a very narrow, highly informed uniform prior
-on divergence times is necessary to avoid the method's preference for
-models with few divergence-time parameters.
+
+The main argument of \citet{Hickerson2013} is that the priors used by
+\citet{Oaks2012} were not informative enough.
+They suggest a very narrow, highly informed uniform prior on divergence times
+is necessary to avoid the method's preference for models with few
+divergence-time parameters.
 Such an empirical Bayesian approach to model selection raises some theoretical
@@
 
-For our re-analysis of the Philippines dataset we used the model-averaging
+There was an error in the methods of \citet{Hickerson2013} resulting in
+an erroneous model that mixes different units of time (see SI for
+details).
+For our re-analysis of the Philippines dataset we followed the model-averaging
 approach of \citet{Hickerson2013}, but with a reduced set of prior models to
-avoid mixing time units.
+avoid the error.
 We used five prior models, all of which had priors on population sizes of
@@
 Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
-Unfortunately, ``checking'' the priors by plotting the summary statistics from
-1000 random samples of each prior model along the first two orthogonal axes of
-a principle component analysis, as recommended by \citet{Hickerson2013},
-provides no warning of a problem (Figure~\ref{figPCA}).
+
+\citet{Hickerson2013} vetted the priors used in their model-averaging approach
+via ``graphical checks,'' in which the summary statistics from 1000 random
+samples of each prior model are plotted along the first two orthogonal axes of
+a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
+To determine if such prior predictive analyses would indicate the $M_{1A}$ and
+$M_{1B}$ are are problematic, we performed these graphical checks on our prior
+models.
+Unfortunately, these prior predictive checks provide no warning that these
+priors are too narrow (Figure~\ref{figPCA}).
+Rather, they suggest these incorrect priors are ``better fit''
+(Figure~\ref{figPCA}A--C) than the valid priors used by \citet{Oaks2012}
+(Figure~\ref{figPCA}D--F).
+
 Given that the results of \citet{Hickerson2013} strongly prefer the models with

### 2013-12-03 manuscripts/reply/reply-figures.tex
Working on reply ms.
Moving principal components plots to Supporting Information. Also,
updating figure caption for the PC plots.

@@
 }{figCoinFlip}
-
-\mFigure{../../model-choice/priors-for-pc-plot/pc-plots.pdf}{
-    The graphical checks recommended by \citet{Hickerson2013} for three prior
-    models: (A) $M_1$ ($\divt{} \sim U(0, 0.1)$), (B) $M_{1A}$ ($\divt{} \sim
-    U(0, 0.01)$), and (C) $M_{1B}$ ($\divt{} \sim U(0, 0.001)$).
-    The plots project the summary statistics from 1000 random samples from each
-    model onto the first two orthogonal axes of a principle component analysis,
-    with the blue dot representing the observed summary statistics from the 22
-    population pairs of Philippine vertebrates.
-}{figPCA}
 

### 2013-12-03 manuscripts/reply/reply-si-figures.tex
Working on reply ms.
Moving principal components plots to Supporting Information. Also,
updating figure caption for the PC plots.

@@
-\mFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
+\siFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
     The joint posterior of the mean (\meant{}{}) and dispersion index ($\vmratio{} = 
@@
 }{figJointPosteriorHickerson}
+
+\siFigure{../../model-choice/priors-for-pc-plot/pc-plots.pdf}{
+    The prior predictive graphical checks recommended by \citet{Hickerson2013}
+    for six prior models:
+    (A) $M_1$ ($\divt{} \sim U(0, 0.1)$),
+    (B) $M_{1A}$ ($\divt{} \sim U(0, 0.01)$),
+    (C) $M_{1B}$ ($\divt{} \sim U(0, 0.001)$), 
+    (D) $M_3$ ($\divt{} \sim U(0, 5)$),
+    (E) $M_4$ ($\divt{} \sim U(0, 10)$), and
+    (F) $M_5$ ($\divt{} \sim U(0, 20)$).
+    The three models that likely exclude true values of some divergence times
+    of the 22 pairs of Philippine vertebrate taxa (A--C) appear to have a
+    better ``fit'' than the priors that likely cover the true divergence times
+    (D--F).
+    The plots project the summary statistics from 1000 random samples from each
+    model onto the first two orthogonal axes of a principle component analysis,
+    with the blue dot representing the observed summary statistics from the 22
+    population pairs of Philippine vertebrates.
+}{figPCA}

### 2013-12-03 manuscripts/reply/short-reply.tex
Working on reply ms.
Moving principal components plots to Supporting Information. Also,
updating figure caption for the PC plots.

@@
 Unfortunately, these prior predictive checks provide no warning that these
-priors are too narrow (Figure~\ref{figPCA}).
+priors are too narrow (Figure~S\ref{figPCA}).
 Rather, they suggest these incorrect priors are ``better fit''
-(Figure~\ref{figPCA}A--C) than the valid priors used by \citet{Oaks2012}
-(Figure~\ref{figPCA}D--F).
+(Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
+\citet{Oaks2012} (Figure~S\ref{figPCA}D--F).
 

### 2013-12-03 manuscripts/reply/reply-si-main.tex
Working on reply ms.
Adding reference for dpp-msbayes manuscript to `references.bib`, moving
sections from `short-reply.tex` to `reply-si-main.tex`, and miscallenous
edits throughout `short-reply.tex`.

@@
 estimator of model probability than \vmratio{}.
+
+\section*{Other clarifications}
+
+\subsection*{Graphical prior comparisons}
+\citet{Hickerson2013} advocate the use of what they call graphical checks of
+prior models.
+This entails generating a small number (1000) of random samples from the prior
+and plotting the resulting summary statistics in comparison to the observed
+statistics to see if they coincide (see Figure 1 of \citet{Hickerson2013}).
+As we show above, this strategy can be misleading, because the resulting plots
+of this approach have little correlation with the appropriateness of priors.
+Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
+dataset analyzed by \citet{Hickerson2013}), we do not expect that 1000
+\emph{random} draws from the vast prior parameter space will yield data and
+summary statistics consistent with the observed data.
+In fact, when such random draws are tightly clustered around the observed
+statistics, this can be an indication that the prior is over-fit, as we show
+above (Table~\ref{tabModelChoiceEmpirical} and Figure~\ref{figPCA}).
+Thus, using such plots to select priors should be avoided, and the use of
+posterior predictive analyses would be much more informative about the overall
+fit of models.
+
+\subsection*{Saturation of summary statistics}
+\citet{Hickerson2013} claim the priors used by \citet{Oaks2012} ``cause much of
+the explored parameter space to be beyond the threshold of saturation in most
+mtDNA genes.'' To explore this possibility, we simulated datasets under prior
+settings that match two of the three priors used by \citet{Oaks2012}:
+$\meanDescendantTheta{} \sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim
+U(0.0005, 0.02)$.
+Under this prior, we draw divergence time parameters from a uniform
+distribution of $U(0, 20)$, simulate datasets, and plot the \divt{} values
+against the summary statistics calculated from the resulting datasets
+(Figure~\ref{figSaturationPlot}).
+Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
+and 10 suffered from little to no effect from saturation.
+Even at divergence times of 20 coalescent units, there is still signal in the
+summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
+Thus, the assertion of \citet{Hickerson2013} does not apply to at least
+two of the priors used by \citet{Oaks2012} and, as a result, does not
+explain the bias they found.

### 2013-12-03 manuscripts/reply/short-reply.tex
Working on reply ms.
Adding reference for dpp-msbayes manuscript to `references.bib`, moving
sections from `short-reply.tex` to `reply-si-main.tex`, and miscallenous
edits throughout `short-reply.tex`.

@@
     that users of the approach should carefully assess the robustness of their 
-    conclusions to different priors.
+    conclusions to different priors and the method's ability to detect
+    temporal variation in divergences.
     According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
@@
 
-\citet{Oaks2012} suggest the primary cause of the bias of \msb is the low
-marginal likelihoods of parameter-rich models integrated over vast parameter
-space with low probability of producing the data, yet relatively high prior
-density \citep{Lindley1957} imposed by the uniformly distributed prior on
+\citet{Oaks2012} suggest the primary cause of the spurious support for models
+with few divergence parameters is the greater marginal likelihoods of these
+models under vague uniform priors relative to more parameter-rich models.
+Models with more divergence-time parameters integrate over \emph{much} greater
+parameter space with low probability of producing the data, yet relatively high
+prior density \citep{Lindley1957} imposed by the uniformly distributed prior on
 divergence times.
@@
 models with more parameter space is a primary mechanism by which uniform priors
-are causing the bias of \msb.
+are causing \msb to prefer highly clustered models of divergences.
 
@@
 Alternatively, \citet{Hickerson2013} present an approach that uses empirically
-informed priors and Bayesian model-averaging in an attempt to accommodate
+informed uniform priors and Bayesian model-averaging in an attempt to accommodate
 uncertainty in selecting priors.
@@
 that samples over empirically informed uniform priors is a fruitful solution
-for an already biased model-choice method.
+for a model-choice method that is very sensitive to priors.
 
@@
 using empirically informed priors for Bayesian model choice.
-Furthermore, we evaluate the empirical Bayesian model averaging approach of
-\citet{Hickerson2013} as a potential solution to the biases revealed by
+Furthermore, we evaluate the empirical Bayesian model-averaging approach of
+\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
 \citet{Oaks2012}.
 In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
-made an error by mixing different time units, which makes the results presented
-in their response difficult to interpret (see Supporting Information for
-details).
-However, here we show that even when this error is avoided, the empirical
-Bayesian model-averaging approach of \citet{Hickerson2013} remains biased
-toward clustered divergence models, and is predisposed to excluding true
-parameter space.
+made an error by mixing different multiple units of time, which makes the
+results presented in their response difficult to interpret (see Supporting
+Information for details).
+Even when avoiding this error, we find Hickerson et al.'s
+\citeyear{Hickerson2013} empirical Bayesian model-averaging implementation of
+\msb remains biased toward clustered divergence models, and is predisposed to
+excluding true parameter space.
 The approach provides another means by which the method can ``escape'' models
@@
 Importantly, this bias can manifest in the preference of models that exclude
-from consideration the true values of the parameters.
+from consideration the true values of the model's parameters.
+Our results, when combined with those of \citet{Oaks2012}, suggest that it is
+difficult to choose a uniformly distributed prior on divergence times that is
+broad enough to confidently contain the true values of parameters while being
+narrow enough to avoid bias toward models with less parameter space.
+However, recent work shows altnerative probability distributions without a
+hard upper bound (e.g. gamma) can accomodate prior uncertainty in divergence
+times while increasing the method's power to detect temporal variation in
+divergences
+\citet{Oaks2014dpp}.
 
@@
 uniform priors in such situations are narrow enough to avoid the bias toward
-models with few divergence events.
+models with few divergence-time parameters.
 The results of \citet{Oaks2012} and \citet{Hickerson2013} showed the method is
@@
 \end{equation}
-where
+as a measure of our beliefs after seeing the new information, where
 \begin{equation}
@@
 is the marginal likelihood of the model.
-The posterior distribution is a measure of our beliefs after seeing the new
-information.
 \end{linenomath}
@@
 Empirical Bayes is a well-studied branch of Bayesian statistics that has given
-rise to many inference methods that provide means of
-parameter estimation that often exhibit favorable frequentist properties.
+rise to many inference methods that provide means of parameter estimation that
+often exhibit favorable frequentist properties.
 Furthermore, many post-hoc correction methods have been developed for
-estimating confidence-intervals from empirical Bayes estimates of posterior
+estimating confidence-intervals from empirical Bayes estimates of
 distributions that often exhibit well-behaved frequentist coverage
@@
 models have no clear true values.
-In some simple situations we might suspect that one model is correct and the
-rest are wrong, but in most applications, we know that all the models under
-consideration are oversimplifications of reality.
-Thus, model posterior probabilities are inherently measures of our belief in
-the models after our prior beliefs are updated by the data being analyzed.
+% In some simple situations we might suspect that one model is correct and the
+% rest are wrong, but in most applications, we know that all the models under
+% consideration are oversimplifications of reality.
+Model posterior probabilities are inherently measures of our belief in the
+models after our prior beliefs are updated by the data being analyzed.
 
@@
 \end{equation}
-Given the models have equal probability under our prior, and following
-Equation~\ref{eq:bayesmodelchoice}, we can calculate the posterior
-probability of Model 1 as
+Given the models have equal probability under our prior, we can calculate the
+posterior probability of Model 1 as
 \begin{equation}
@@
 \end{equation}
-This is the correct posterior probability of Model 1 given our prior information
+This is the correct posterior probability of Model 1 given our prior
 and data.
@@
 To give the data more weight relative to the prior, we could use it twice, and
-calculate an empirical Bayes estimate using a prior of $beta(13,17)$.
-This results in a distribution of $beta(16, 24)$ (Figure~\ref{figCoinFlip}),
-with a MAP estimate of 0.395, and $p(M_1 \given y=3, N=10) = 0.10$.
+calculate an empirical Bayes estimate using a prior of $beta(13,17)$.  This
+results in a ``posterior'' distribution of $beta(16, 24)$
+(Figure~\ref{figCoinFlip}), with a MAP estimate of 0.395, and $p(M_1 \given
+y=3, N=10) = 0.10$.
 The estimated posterior distribution of the parameter, and resulting MAP
@@
 see Table~6 of \citet{Oaks2012}).
-Thus, the results of \citet{Hickerson2013} indicate that integration over
-parameter space with low likelihood is biasing the method toward models with
-smaller parameter space, and as a consequence is biasing the method toward
-models that exclude the truth (i.e., toward estimating model-averaged
-posteriors dominated by models that exclude true values of parameters of the
-model).
+The strong preference for the questionable priors on divergence times (their
+$M_1$ and $M_2$) in the model-averaged results of \citet{Hickerson2013} suggest
+their approach is biased toward models with less parameter space, and as a
+consequence, is biased toward models that likely exclude the truth.
+I.e., the approach is biased toward estimating model-averaged posteriors
+dominated by models that exclude true values of parameters of the model.
 
@@
 
-There was an error in the methods of \citet{Hickerson2013} resulting in
-an erroneous model that mixes different units of time (see SI for
-details).
 For our re-analysis of the Philippines dataset we followed the model-averaging
 approach of \citet{Hickerson2013}, but with a reduced set of prior models to
-avoid the error.
+avoid their error of mixing units of time (see SI for details).
 We used five prior models, all of which had priors on population sizes of
@@
 across all five models.
-We then repeated this analysis twice, replacing the $M_1$ with
+We then repeated this analysis twice, replacing the $M_1$ model with
 $M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
@@
 does not reduce the method's bias toward models with less parameter space,
-but rather allows it to manifest in a rather dangerous way.
+but rather allows it to manifest in a concerning way.
 The method strongly prefers the prior model with the narrowest distribution on
@@
 
-\citet{Hickerson2013} vetted the priors used in their model-averaging approach
-via ``graphical checks,'' in which the summary statistics from 1000 random
-samples of each prior model are plotted along the first two orthogonal axes of
-a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
+However, \citet{Hickerson2013} vetted the priors used in their model-averaging
+approach via ``graphical checks,'' in which the summary statistics from 1000
+random samples of each prior model are plotted along the first two orthogonal
+axes of a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
 To determine if such prior predictive analyses would indicate the $M_{1A}$ and
-$M_{1B}$ are are problematic, we performed these graphical checks on our prior
-models.
+$M_{1B}$ models are problematic, we performed these graphical checks on our
+prior models.
 Unfortunately, these prior predictive checks provide no warning that these
 priors are too narrow (Figure~S\ref{figPCA}).
-Rather, they suggest these incorrect priors are ``better fit''
+Rather, the graphs suggest these incorrect priors are ``better fit''
 (Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
@@
 model-averaged results are dominated by models that exclude at least some of
-the 22 true divergence times, making their results difficult to interpret
-(in addition to the error of time units).
+the 22 true divergence times, making their results difficult to interpret.
 
@@
 excluding the true value of at least one parameter.
-Whenever the Bayes factor preferred a model excluding the truth we counted
-the number of 22 true divergence times that were excluded by the preferred
+Whenever the Bayes factor preferred a model excluding the truth we counted the
+number of the 22 true divergence times that were excluded by the preferred
 model.
@@
 We also used the mode estimate of the preferred model for each replicate to
-estimate the number of true values excluded, which produced very similar
-results.
+estimate the number of true parameter values excluded, which produced very
+similar results.
 Furthermore, the posterior probability of excluding at least one true parameter
@@
 % be biased away from models with more parameter space \citep{Oaks2012}.
-The consequence of this approach is a high risk of obtaining a model-averaged
+The consequence of this approach is the risk of obtaining a model-averaged
 posterior estimate that is heavily weighted toward incorrect models that
@@
 \subsection*{Additional thoughts on empirical priors in Bayesian model choice}
-Given all of the theoretical and practical issues with empirical Bayes
-approaches to Bayesian model choice discussed in the proceeding sections, it is
-quite clear why one should use caution before overly criticizing an
-investigator's choice of priors after having seen the resulting posterior.
+Given the theoretical and practical issues with empirical Bayes approaches to
+Bayesian model choice discussed in the proceeding sections, it is quite clear
+why one should use caution before overly criticizing an investigator's choice
+of priors after having seen the resulting posterior.
 %In a strictly Bayesian world, this is cheating.
-As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a large
-amount of uncertainty regarding the divergence times of the 22 population pairs
-under study.
+As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a
+large amount of uncertainty regarding the divergence times of the 22 population
+pairs under study.
 Two of these pairs represent distinct species, and the taxonomy of many groups
@@
     Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013}.
-\citet{Oaks2012} also discuss how the sole use of uniform priors in \msb makes
-it very difficult for investigators to express their prior uncertainty without
-putting high prior density in regions of improbable parameter space.
-The alternative, as shown above, is to risk excluding the truth before
-performing the analysis.
-
+% \citet{Oaks2012} also discuss how uniform priors on postive real-valued parameters make
+% it difficult for investigators to express their prior uncertainty without
+% putting high prior density in regions of improbable parameter space.
+The alternative to priors that reflect prior uncertainty, as shown above, is to
+risk excluding the truth before performing the analysis.
+
+\section*{Assessing the power of the model-averaging approach of
+    \citet{Hickerson2013}}
 While our results above clearly demonstrate the risks inherent to the empirical
@@
 such risk if the method does indeed increase the power of the method and thus
-decrease bias toward clustered divergence models, as claimed by
-\citet{Hickerson2013}.
-In the next section, we use simulation-based analyses to evaluate the power of
-Hickerson et al.'s (\citeyear{Hickerson2013}) approach, which show the method is
-still biased toward highly clustered models of divergence.
-
-\section*{Assessing the power of the model-averaging approach of
-    \citet{Hickerson2013}}
-% \cwlnote{}{I think the empirical use of the model-averaging should go here,
-%     instead of above in the empirical Bayes section}
-
-As recommended by \citet{Oaks2012}, we perform simulation-based analyses
-to explore the power of the approximate Bayesian model-averaging approach
-proposed by \citet{Hickerson2013}.
+decrease bias toward clustered divergence models.
+We assess this possiblity using simulation-based analyses to evaluate the power
+of Hickerson et al.'s (\citeyear{Hickerson2013}) approach.
 Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
@@
 We find that the approach of \citet{Hickerson2013} struggles to estimate the
-dispersion index of divergence times (\vmratio{}) across most of the \divt{max}
+variance of divergence times (\vmratio{}) across most of the \divt{max}
 we simulated, whether evaluating the unadjusted
@@
 (Figures~\ref{figPowerAccOmegaMedian}E \& \ref{figPowerAccOmegaModeGLM}E).
-This is consistent with the conclusion of \citet{Oaks2012} that \msb lacks
-robustness and is highly sensitive to the prior distribution deviating from the
-true, underlying distribution of the data.
+This is consistent with the findings of \citet{Oaks2012}  and
+\citet{Hickerson2013} that \msb is highly sensitive to the prior distribution
+deviating from the true, underlying distribution of the data.
 
@@
 when populations diverged randomly over the past \globalcoalunit generations.
-Even when divergences are random over the past $8\globalpopsize$ generations,
-the most likely inference is only two divergence events, and a single 
+When divergences are random over the past $8\globalpopsize$ generations, the
+most likely inference is only two divergence-time parameters, and a single
 divergence is still estimated in more than 10\% of the replicates.
-It is very interesting to note that as \divt{max} increases, but before the
+It is interesting to note that as \divt{max} increases, but before the
 estimates are finally pulled away from $\numt{} = 1$, the distribution of
-\numt{} estimates closely mirror the odd U-shaped prior on divergence models
+\numt{} estimates closely mirror the U-shaped prior on divergence models
 used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
@@
 This supports the conclusions of \citet{Oaks2012} that this U-shaped
-prior coupled with the poor marginal likelihoods of models with many
-\divt{} parameters, is a major cause of the method's bias toward
+prior coupled with the small marginal likelihoods of models with many
+divergence parameters, is a major cause of the method's bias toward
 clustered divergence models.
@@
 
-Contrary to Hickerson et al.'s (\citeyear{Hickerson2013}) claim that their
-model-averaging approach ``can discriminate complex multispecies histories and
-correctly reject synchronous divergence, even when discrete divergence times
-differ by much less than \ldots millions of generations,'' our results show
-that method is biased toward supporting the extreme case of a single
-divergence event when populations diverged randomly over the last
-$8\globalpopsize$ generations.
+Our results show that the empirical Baysian model-averaging approach is biased
+toward supporting the extreme case of a single divergence event when
+populations diverged randomly over the last $8\globalpopsize$ generations.
 To put this on the scale roughly consistent with a vertebrate mitochondrial
@@
 (Figure~\ref{figPowerNumExcluded}E).
-Again, this demonstrates the method's lack of robustness.
+Again, this demonstrates the method's sensitiviy to priors.
 
@@
 demonstrated by \citet{Oaks2012}).
-It is not surprising that under limited parameter space the method has
+It is not surprising that under certain prior conditions the method has
 increased power to detect temporal variation in divergences over narrower time
@@
 \citet{Oaks2012} explore the behavior of the method under three divergence time
-priors, as narrow as 0--5 coalescent units. This is quite narrow considering
-that this expresses the prior belief that all 22 taxon pairs diverged within
-this window.
+priors, as narrow as 0--5 coalescent units.
+This is quite narrow considering that this expresses the prior belief that all
+22 taxon pairs diverged within this window.
 \citet{Hickerson2013} limit their power analysis to a single prior of 0--1
 coalescent unit.
-This prior setting assumes that there is enough a priori information to
-be 100\% certain that all taxa (18 for their simulations) diverged within
+This prior setting assumes that there is enough \emph{a priori} information to
+be certain that all taxa (18 for their simulations) diverged within
 the last 4$\globalpopsize$ generations.
 This does not seem applicable to most empirical systems.
-Furthermore, even when such extensive prior information is available, to only
-be able to detect multiple divergences on the scale of hundreds of thousands of
-generations does not seem very powerful.
-
-\citet{Hickerson2013} present information only on the extreme case of a
-single divergence event.
-As discussed in \citet{Oaks2012}, it seems odd to consider the estimation
-of two divergence events as a success when divergence is random over hundreds
-of thousands (or millions) of generations.
-In an empirical system such as the Philippines, where island fragmentation has
-occurred at least 10 times over the past several millions of years
-\citep{Haq1987,Rohling1998,Siddall2003,Miller2005}, an estimate where 22 taxa
-share even a handful of divergence events would be of biogeographic interest.
-
-Lastly, \citet{Hickerson2013} translate their results from units of
-\globalcoalunit generations to generations assuming a mutation rate of
-$1.92\e{-8}$ mutations per site per generation.
-If we translate their results to a scale more consistent with rates of mutation
-of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
-by \citet{Hickerson2013} the method can only reliably reject the extreme case
-of one divergence event when divergences are random over a period of more than
-3 million generations.
-Again, considering \citet{Hickerson2013} are assuming prior knowledge that
-all population pairs diverged within the last coalescent unit, the
-method has low power to detect random variation in divergence times.
+Furthermore, even when such prior information is available, the results of
+\citet{Hickerson2013} demonstrate the method is limited to detecting multiple
+divergences on the scale of hundreds of thousands of generations, assuming a
+per-site rate of $1.92\e{-8}$ mutations per generation.
+Assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), their
+results demonstrate the method can reliably reject one divergence event when
+divergences are random over a period of 3 million generations.
 
@@
 in \msb}
-One of the points of disagreement between \citet{Hickerson2013} and
-\citet{Oaks2012} is the underlying mechanism causing the bias toward models
-with clustered divergences (i.e., models with few divergence events).
+% One of the points of disagreement between \citet{Hickerson2013} and
+% \citet{Oaks2012} is the underlying mechanism causing the bias toward models
+% with clustered divergences (i.e., models with few divergence events).
 \citet{Oaks2012} present simulation-based analyses that suggest the broad
-uniform prior on divergence time parameters decrease the marginal likelihoods
-of models with more divergence time parameters, whereas \citet{Hickerson2013}
-argue that the bias is inherent to the inefficient rejection algorithm
-implemented in \msb.
+uniform prior on divergence-time parameters result in small marginal
+likelihoods of models with more divergence-time parameters, and thus a
+preference for models with few divergence parameters.
+\citet{Hickerson2013} conclude that the bias is due to the inefficient
+rejection algorithm implemented in \msb.
 
@@
 that were consistent with the Philippine data.
-There are a few issues with their argument.
-First, the probabilities they present assume the gene divergence times
+There are a few issues with this argument.
+The probabilities they present assume the gene divergence times
 estimated by \citet{Oaks2012} are correct.
@@
 \citet{Oaks2012} was to provide a very rough comparison of the gene divergence
-times across the 22 taxa.  These analyses assumed an arbitrary strict clock of
-$2\e{-8}$ for all taxa, and are, of course, subject to estimation error.
+times across the 22 taxa.
+These analyses assumed an arbitrary strict mutation rate of $2\e{-8}$ for all
+taxa, and are, of course, subject to estimation error.
 Furthermore, the branch-length units of the gene trees are in millions of
@@
 \citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
-questionable, especially given our findings on the behavior of empirically
-informed priors presented above.
-
-Finally, even \emph{if} we make all of these assumptions and assume the gene
-divergence times are estimated without error, the probabilistic argument only
-applies to one of the three different priors used by \citet{Oaks2012}.
-The narrowest prior on divergence times used by \citet{Oaks2012} closely mirrors
-the range of estimates of gene-divergence times, and applying Hickerson et al.'s
-(\citeyear{Hickerson2013}) probability equations demonstrates that the prior
-is densely populated with samples with large numbers of divergence parameters
-that are consistent with the gene divergence estimates.
+questionable, especially given our findings above on the behavior of \msb when
+empirically informed uniform priors are employed.
+
+Also, if we make all of these assumptions and assume the gene divergence times
+are estimated without error, the probabilistic argument only applies to one of
+the three different priors used by \citet{Oaks2012}.
+The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5
+mya)$) closely mirrors the range of estimates of gene-divergence times.
+Applying Hickerson et al.'s (\citeyear{Hickerson2013}) probability equations
+demonstrates this prior is densely populated with samples with large numbers of
+divergence parameters that are consistent with the gene divergence estimates.
 Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
@@
 % the smallest parameter space via insufficient prior sampling.
-While analyses that sample each model proportional to their relative parameter
-space could be explored, it seems much more likely that the broad uniform
-priors are simply inhibiting the marginal likelihoods of spacious models.
-
-As discussed by \citet{Oaks2012}, a straightforward prediction of reduced
-likelihoods of complex models is that they should disappear as the model
-generating the data converges to the prior.
-\citet{Oaks2012} test this prediction by performing 100,000 simulations to
+While analyses that sample each model proportional to their parameter space
+could be explored, it seems much more likely that the marginal likelihoods
+under broad uniform priors prefer simpler models.
+
+As discussed by \citet{Oaks2012}, a straightforward prediction of marginal
+likelihoods causing the preference for simple models is that the bias should
+disappear as the model generating the data converges to the prior.
+\citet{Oaks2012} tested this prediction by performing 100,000 simulations to
 assess the model-choice behavior of \msb when the prior model is correct.
@@
 Thus, the results of the simulations of \citet{Oaks2012} are more consistent
-with marginal model likelihoods causing the bias toward models with fewer
-divergence time parameters.
+with larger marginal likelihoods of models with fewer divergence-time
+parameters.
 
 Taken together, the results presented here and in \citet{Oaks2012} support
-reduced marginal likelihoods of large models as a primary mechanism by which
-broad uniform priors cause biases in \msb.
-Nonetheless, posterior sampling error will always be present in any numerical
+greater marginal likelihoods of simple models under broad priors as a primary
+mechanism by which \msb prefers models with clustered divergence times.
+Nonetheless, sampling error will always be present in any numerical
 Bayesian approximation method.
-Thus, insufficient sampling of the prior will contribute to the error of all
-approximate Bayesian estimates.
-However, there is no strong evidence that it is playing a large role in
-the biases revealed herein and by \citet{Oaks2012}.
 
@@
 %CWL: This seems to nit-picky. Probably not worth including it
-\section*{Other clarifications}
-
-\subsection*{Graphical prior comparisons}
-\citet{Hickerson2013} advocate the use of what they call graphical checks of
-prior models.
-This entails generating a small number (1000) of random samples from the prior
-and plotting the resulting summary statistics in comparison to the observed
-statistics to see if they coincide (see Figure 1 of \citet{Hickerson2013}).
-As we show above, this strategy can be misleading, because the resulting plots
-of this approach have little correlation with the appropriateness of priors.
-Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
-dataset analyzed by \citet{Hickerson2013}), we do not expect that 1000
-\emph{random} draws from the vast prior parameter space will yield data and
-summary statistics consistent with the observed data.
-In fact, when such random draws are tightly clustered around the observed
-statistics, this can be an indication that the prior is over-fit, as we show
-above (Table~\ref{tabModelChoiceEmpirical} and Figure~\ref{figPCA}).
-Thus, using such plots to select priors should be avoided, and the use of
-posterior predictive analyses would be much more informative about the overall
-fit of models.
-
-\subsection*{Saturation of summary statistics}
-\citet{Hickerson2013} claim the priors used by \citet{Oaks2012} ``cause much of
-the explored parameter space to be beyond the threshold of saturation in most
-mtDNA genes.'' To explore this possibility, we simulated datasets under prior
-settings that match two of the three priors used by \citet{Oaks2012}:
-$\meanDescendantTheta{} \sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim
-U(0.0005, 0.02)$.
-Under this prior, we draw divergence time parameters from a uniform
-distribution of $U(0, 20)$, simulate datasets, and plot the \divt{} values
-against the summary statistics calculated from the resulting datasets
-(Figure~\ref{figSaturationPlot}).
-Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
-and 10 suffered from little to no effect from saturation.
-Even at divergence times of 20 coalescent units, there is still signal in the
-summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
-Thus, the assertion of \citet{Hickerson2013} does not apply to at least
-two of the priors used by \citet{Oaks2012} and, as a result, does not
-explain the bias they found.
 
@@
 of the parameters.
-Exploring alternative prior probability distributions on most of the model's
-parameters, in addition to different priors over divergence models, should
-help mitigate the method's biases.
-
-Based on our results, we discourage the use of empirically informed uniform
-priors in Bayesian model choice, especially when combined with model-averaging
-\ldots
+
+Whether or not one chooses to use empirically informed priors, our results
+suggest that it is difficult to choose a uniformly distributed prior on
+divergence times that is broad enough to confidently contain the true values of
+parameters while being narrow enough to avoid bias toward models with less
+parameter space.
+Fortunately, alternative prior probability distributions allow prior
+uncertainty to be accommodated while avoiding excessive prior density in
+regions of low likelihood, which greatly improves the power of the method
+\citet{Oaks2014dpp}.
+
+% Based on our results, we discourage the use of empirically informed uniform
+% priors in Bayesian model choice, especially when combined with model-averaging
+% \ldots
 

### 2013-12-04 manuscripts/reply/reply-figures.tex
Working on reply ms.
Moving several figures to Supporting Information. Miscellaneous
edits throughout.

@@
 %% Power results
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
-    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
-    estimate the dispersion index of divergence times (\vmratio{}) from
-    analyses of simulated datasets where \divt{} for 22 pairs of populations is
-    drawn from a series of uniform distributions, $\divt{} \sim U(0,
-    \divt{max})$.
-    The proportion of estimates less than the true value of,
-    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
-    Each plot represents unadjusted median estimates from 500 simulation
-    replicates analyzed using $5\e6$ samples from the prior.
-}{figPowerAccOmegaMedian}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
-    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
-    estimate the dispersion index of divergence times (\vmratio{}) from
-    analyses of simulated datasets where \divt{} for 22 pairs of populations is
-    drawn from a series of uniform distributions, $\divt{} \sim U(0,
-    \divt{max})$.
-    The proportion of estimates less than the true value of,
-    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
-    Each plot represents GLM-regression-adjusted mode estimates from 500
-    simulation replicates analyzed using $5\e6$ samples from the prior.
-}{figPowerAccOmegaModeGLM}
 
@@
 
-%% Validation results
-\mFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
-    An assessment of the approximate Bayesian model-averageing approach of
-    \citet{Hickerson2013} under the ideal conditions when the prior model
-    is correct (i.e., the pseudo-replicate datasets are simulated from
-    parameters drawn from the same prior distributions used in the
-    analysis).
-    The plots show the relationship between the estimated posterior and true
-    probability of (A \& C) $\numt{}=1$ and (B \& D) $\vmratio{} < 0.01$, based
-    on 50,000 simulations.
-    The results summarize the (A \& B) unadjusted and (C \& D) GLM-adjusted
-    posterior estimate from each simulation replicate.
-    The prior settings for all replicates included five prior models with
-    $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
-    U(0.0001, 0.05)$ for all five models, and
-    $M_1: \divt{} \sim U(0, 0.1)$,
-    $M_2: \divt{} \sim U(0, 1)$,
-    $M_3: \divt{} \sim U(0, 5)$,
-    $M_4: \divt{} \sim U(0, 10)$, and
-    $M_5: \divt{} \sim U(0, 20)$.
-    The number of samples from the prior was $2.5\e6$.
-    The simulated data structure was 8 population pairs, with a single 1000
-    bp locus sampled from 10 individuals from each population.  The 50,000
-    estimates of the posterior probability of one divergence event were
-    assigned to 20 bins of width 0.05.
-    The estimated posterior probability of each bin is plotted against the
-    proportion of replicates in that bin with a true value consistent with
-    one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
-}{figValidationMCBehavior}
-
 \mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
@@
 
-%:FIGURE-saturation plot
-\mFigure{../../saturation/saturation-plot.pdf}{
-    The summary statistics $\pi$ \citep{Tajima1983} and $\pi_{net}$
-    \citep{Takahata1985} as a function of divergence time between populations.
-    Each plot represents 1100 pairs of parameter draws and summary statistics
-    calculated from the simulated data.
-    Prior settings for the simulations were $\divt{} \sim U(0, 20)$,
-    $\meanDescendantTheta{} \sim U(0.0005, 0.04)$, and $\ancestralTheta{} \sim
-    U(0.0005, 0.02)$.
-}{figSaturationPlot}
-
 % \mFigure{../../response-redux/results/hickerson/pymsbayes-results/pymsbayes-output/d1/m12345678-combined/mean_by_dispersion.pdf}{

### 2013-12-04 manuscripts/reply/reply-si-figures.tex
Working on reply ms.
Moving several figures to Supporting Information. Miscellaneous
edits throughout.

@@
 }{figPCA}
+
+\siFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
+    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
+    estimate the dispersion index of divergence times (\vmratio{}) from
+    analyses of simulated datasets where \divt{} for 22 pairs of populations is
+    drawn from a series of uniform distributions, $\divt{} \sim U(0,
+    \divt{max})$.
+    The proportion of estimates less than the true value of,
+    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
+    Each plot represents unadjusted median estimates from 500 simulation
+    replicates analyzed using $5\e6$ samples from the prior.
+}{figPowerAccOmegaMedian}
+
+\siFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
+    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
+    estimate the dispersion index of divergence times (\vmratio{}) from
+    analyses of simulated datasets where \divt{} for 22 pairs of populations is
+    drawn from a series of uniform distributions, $\divt{} \sim U(0,
+    \divt{max})$.
+    The proportion of estimates less than the true value of,
+    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
+    Each plot represents GLM-regression-adjusted mode estimates from 500
+    simulation replicates analyzed using $5\e6$ samples from the prior.
+}{figPowerAccOmegaModeGLM}
+
+%% Validation results
+\siFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
+    An assessment of the approximate Bayesian model-averageing approach of
+    \citet{Hickerson2013} under the ideal conditions when the prior model
+    is correct (i.e., the pseudo-replicate datasets are simulated from
+    parameters drawn from the same prior distributions used in the
+    analysis).
+    The plots show the relationship between the estimated posterior and true
+    probability of (A \& C) $\numt{}=1$ and (B \& D) $\vmratio{} < 0.01$, based
+    on 50,000 simulations.
+    The results summarize the (A \& B) unadjusted and (C \& D) GLM-adjusted
+    posterior estimate from each simulation replicate.
+    The prior settings for all replicates included five prior models with
+    $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
+    U(0.0001, 0.05)$ for all five models, and
+    $M_1: \divt{} \sim U(0, 0.1)$,
+    $M_2: \divt{} \sim U(0, 1)$,
+    $M_3: \divt{} \sim U(0, 5)$,
+    $M_4: \divt{} \sim U(0, 10)$, and
+    $M_5: \divt{} \sim U(0, 20)$.
+    The number of samples from the prior was $2.5\e6$.
+    The simulated data structure was 8 population pairs, with a single 1000
+    bp locus sampled from 10 individuals from each population.  The 50,000
+    estimates of the posterior probability of one divergence event were
+    assigned to 20 bins of width 0.05.
+    The estimated posterior probability of each bin is plotted against the
+    proportion of replicates in that bin with a true value consistent with
+    one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
+}{figValidationMCBehavior}
+
+%:FIGURE-saturation plot
+\siFigure{../../saturation/saturation-plot.pdf}{
+    The summary statistics $\pi$ \citep{Tajima1983} and $\pi_{net}$
+    \citep{Takahata1985} as a function of divergence time between populations.
+    Each plot represents 1100 pairs of parameter draws and summary statistics
+    calculated from the simulated data.
+    Prior settings for the simulations were $\divt{} \sim U(0, 20)$,
+    $\meanDescendantTheta{} \sim U(0.0005, 0.04)$, and $\ancestralTheta{} \sim
+    U(0.0005, 0.02)$.
+}{figSaturationPlot}

### 2013-12-04 manuscripts/reply/short-reply.tex
Working on reply ms.
Moving several figures to Supporting Information. Miscellaneous
edits throughout.

@@
 %CWL: I would suggest removing this section
-\section*{The power analysis of \citet{Hickerson2013}}
-\citet{Hickerson2013} present a power analysis in which they find that
-\msb has the power to infer multiple divergence events when divergence times
-are random over hundreds of thousands of generations (rather than millions as
-demonstrated by \citet{Oaks2012}).
-It is not surprising that under certain prior conditions the method has
-increased power to detect temporal variation in divergences over narrower time
-windows than millions of generations.
-However, what is important to consider is whether those conditions are relevant
-to real-world applications of the method.
-\citet{Oaks2012} explore the behavior of the method under three divergence time
-priors, as narrow as 0--5 coalescent units.
-This is quite narrow considering that this expresses the prior belief that all
-22 taxon pairs diverged within this window.
-\citet{Hickerson2013} limit their power analysis to a single prior of 0--1
-coalescent unit.
-This prior setting assumes that there is enough \emph{a priori} information to
-be certain that all taxa (18 for their simulations) diverged within
-the last 4$\globalpopsize$ generations.
-This does not seem applicable to most empirical systems.
-Furthermore, even when such prior information is available, the results of
-\citet{Hickerson2013} demonstrate the method is limited to detecting multiple
-divergences on the scale of hundreds of thousands of generations, assuming a
-per-site rate of $1.92\e{-8}$ mutations per generation.
-Assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), their
-results demonstrate the method can reliably reject one divergence event when
+\section*{The use of power analysis to guide applications of \msb}
+\citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
+divergence-time prior of 0--1 coalescent units ago.
+They found that under these prior conditions \msb has the power, assuming a
+per-site rate of $1.92\e{-8}$ mutations per generation, to detect multiple
+divergence events among 18 taxa when the true divergences were random over
+hundreds of thousands of generations or more.
+\citet{Oaks2012} performed similar power analyses under three uniform
+divergence-time priors as narrow as 0--5 coalescent units, and found the
+method was able to detect multiple events among 22 when divergences were random
+over millions of generations.
+As recommended by \citet{Oaks2012}, it is important that investigators perform
+power analyses to determine the method's power for their dataset, and decide if
+\msb has sufficient power to address their hypotheses; in the case of the
+Philippines dataset, it did not.
+When doing so, it is important to consider what prior conditions are
+relevant to their empirical system.
+A divergence-time prior of 0--5 coalescent units is quite narrow considering
+that this expresses the prior belief that all 22 taxon pairs diverged within
+this window.
+Certainly, it is a rare for there to be enough \emph{a priori} information to
+be certain that all taxa diverged within the last 4$\globalpopsize$ generations
+(i.e., 0--1 coalescent units).
+Also, it seems unlikely that when such prior information is available that
+being able to detect multiple divergences on the scale of hundreds of thousands
+of generations will add much insight about the evolutionary history of the
+taxa.
+Assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), with such
+prior information the method can reliably reject one divergence event when
 divergences are random over a period of 3 million generations.
+However, as noted by \citet{Oaks2012}, inferring more than one divergence time
+shared across all taxa does not confirm the method is working well when
+analyzing data generated under random temporal variation in divergences (e.g.,
+an inference of two divergence events could be biogeographically interesting
+yet spurious).
+Investigators should not limit their assessment of method's power to only
+differentiating inferences of one event or more (i.e. $\numt{} = 1$ versus
+$\numt{} > 1$).
+Rather, looking at the distribution of estimates, as in \citet{Oaks2012},
+provides much more information about the behavior of the method.
 
@@
 % with clustered divergences (i.e., models with few divergence events).
-\citet{Oaks2012} present simulation-based analyses that suggest the broad
-uniform prior on divergence-time parameters result in small marginal
-likelihoods of models with more divergence-time parameters, and thus a
-preference for models with few divergence parameters.
-\citet{Hickerson2013} conclude that the bias is due to the inefficient
-rejection algorithm implemented in \msb.
-
-\citet{Hickerson2013} present an interesting probabilistic argument
- to show that insufficient prior sampling is to blame for the bias.
-They argue the widest of the three priors used by \citet{Oaks2012} was very
-unlikely to produce any prior samples with large numbers of divergence times
-that were consistent with the Philippine data.
-There are a few issues with this argument.
+It is important to determine the mechanism by which broad uniform priors cause
+\msb to support clustered models of divergence.
+It is well established that vague priors can be problematic in Bayesian model
+selection.
+Models that integrate over more parameter space characterized by low
+probability of producing the data and relatively high prior density will have
+smaller marginal likelihoods \citep{Jeffries1935,Lindley1957}.
+Given the uniformly distributed priors on divergence times (and other nuisance
+parameters) employed in \msb, models with more divergence parameters will be
+forced to integrate over \emph{much} greater parameter space, all with equal
+prior density, and much of it with low likelihood.
+Thus, it is not surprising in light of this fundamental statistical issue that
+the method tends to support simple models.
+
+However, \citet{Hickerson2013} conclude that the bias is due to insufficient
+prior sampling.
+They argue the widest of the three priors used by \citet{Oaks2012} would
+infrequently produce samples with many independent young divergence times.
+However, there are some issues with their argument.
 The probabilities they present assume the gene divergence times
@@
 Furthermore, the branch-length units of the gene trees are in millions of
-years, whereas the divergence time prior of \msb is in generations, thus the
-logic of \citet{Hickerson2013} requires the additional assumption that all 22
-Philippine taxa have a generation time of one year.
-Thus, the argument of \citet{Hickerson2013} that divergence time estimates of
+years, whereas the divergence time prior of \msb is in generations, thus
+\citet{Hickerson2013} make the implicit assumption that all 22 Philippine taxa
+have a generation time of one year.
+The argument of \citet{Hickerson2013} that divergence time estimates of
 \citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
-questionable, especially given our findings above on the behavior of \msb when
-empirically informed uniform priors are employed.
-
-Also, if we make all of these assumptions and assume the gene divergence times
-are estimated without error, the probabilistic argument only applies to one of
-the three different priors used by \citet{Oaks2012}.
+questionable, especially given our findings presented above regarding the
+behavior of \msb when empirically informed uniform priors are employed.
+
+If we make all of these assumptions and assume the gene divergence times are
+estimated without error, the Hickerson et al.'s \citeyear{Hickerson2013}
+argument only applies to one of the three different priors used by
+\citet{Oaks2012}.
 The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5
@@
 demonstrates this prior is densely populated with samples with large numbers of
-divergence parameters that are consistent with the gene divergence estimates.
-Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
-sampling is to blame for the bias, it should be much reduced under the narrow
-prior on \divt{}.
+divergence parameters with values younger than the estimated gene divergence
+estimates.
+Thus, if insufficient prior sampling is to blame for the bias, it should be
+much reduced under the narrow prior on \divt{}.
 However, the magnitude of the bias is very similar across all three priors
@@
 To disentangle whether model likelihoods or insufficient prior sampling is to
-blame for the biases revealed by \citet{Oaks2012}, we must look at the
+blame for the method's spurious support for simple models, we must look at the
 different predictions made by these two phenomena.
@@
 sampling should create higher variance in posterior estimates, and thus it
-should cause analyses to be sensitive to the number of samples drawn from the
-prior.
+should cause analyses to be highly sensitive to the number of samples drawn
+from the prior.
 \citet{Oaks2012} do not see such sensitivity when they compare prior sample
@@
 GLM-regression-adjusted estimates.
-This suggests that insufficient prior sampling did not play a large
-role in the bias found by \citet{Oaks2012}.
-
-Our results presented above that demonstrate the bias of the model-averaging
-approach of \citet{Hickerson2013} both toward models with narrower \divt{}
-priors (Table~\ref{tabModelChoiceEmpirical} and
-Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
-and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
-\ref{figPowerOmegaProb}) strongly suggest the primary cause is the uniform
-priors reducing the marginal likelihoods of large models.
-In all of our model-averaging analyses, all prior models have the same number
-of samples.
-% Thus, it is difficult to explain the method's propensity toward the model with
-% the smallest parameter space via insufficient prior sampling.
-While analyses that sample each model proportional to their parameter space
-could be explored, it seems much more likely that the marginal likelihoods
-under broad uniform priors prefer simpler models.
+While sampling error is always a reality, clearly it is not playing
+a large role in the results of \citet{Oaks2012} or presented above.
 
@@
 assess the model-choice behavior of \msb when the prior model is correct.
-They find that the bias actually switches direction (at least for the
-regression-adjusted estimates) and the method tends to underestimate the
-probability of the model with one divergence event (see Figure 4 of
+The results confirm the prediction as \msb actually tends to underestimate
+the probability of the one-divergence model (see Figure 4 of
 \citet{Oaks2012}).
-The same prediction is not as straightforward for the insufficient-sampling
-hypothesis.
+These results are not clearly predicted if insufficient-sampling was causing
+the bias.
 Even when the prior is correct, due to the discrete uniform prior on
@@
 
+Furthermore, the results above demonstrate the model-averaging
+approach of \citet{Hickerson2013} is prefers models with narrower \divt{}
+priors (Table~\ref{tabModelChoiceEmpirical} and
+Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
+and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
+\ref{figPowerOmegaProb}).
+% This further suggests the primary cause is the uniform
+% priors reducing the marginal likelihoods of large models.
+In all of these analyses, each of the prior models have the same number of
+samples.
+While analyses that sample each model proportional to their parameter space
+could be explored, it clear that the marginal likelihoods under broad uniform
+priors will be greater for simpler models.
+% Thus, it is difficult to explain the method's propensity toward the model with
+% the smallest parameter space via insufficient prior sampling.
 Taken together, the results presented here and in \citet{Oaks2012} support
@@
 
-%CWL: This seems to nit-picky. Probably not worth including it
-
 \section*{Conclusions}

### 2013-12-05 manuscripts/reply/short-reply.tex
Adding references.
Adding references as examples of taxonomic uncertainty in the
Philippines.

@@
 \citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
-    Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013}.
+Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013,Rafe2013AREES,Siler2014kikuchii}.
 % \citet{Oaks2012} also discuss how uniform priors on postive real-valued parameters make
@@
 % putting high prior density in regions of improbable parameter space.
-The alternative to priors that reflect prior uncertainty, as shown above, is to
-risk excluding the truth before performing the analysis.
+When limited to uniformly distributed priors, the alternative to priors that
+reflect prior uncertainty, as shown above, is to risk excluding the truth
+before performing the analysis.
 

### 2013-12-06 manuscripts/reply/short-reply.tex
Working on `short-reply.tex`.
More edits throughout the ms.

@@
     \citep{Huang2011} introduced an approximate Bayesian approach which can be
-    used to estimate the probability of a model in which multiple sets of taxa
+    used to estimate the probabilities of models in which multiple sets of taxa
     diverge at the same time. 
@@
     does require the relative mutation rates among the taxa and loci to be
-    assumed a priori.
+    assumed \emph{a priori}.
     Recently, \citet{Oaks2012} used this model-choice framework to 
@@
     that users of the approach should carefully assess the robustness of their 
-    conclusions to different priors and the method's ability to detect
-    temporal variation in divergences.
+    conclusions to different priors.
     According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
@@
     underlie model inference.
+    Furthermore, simulations provide insight into the temporal resolution of
+    the method, which in turn helps guide appropriate interpretation of
+    results.
+
     \vspace{12pt}
@@
         model choice; empirical Bayes} 
-
 }
@@
 A full description of the model is given in \citet{Oaks2012}.
-Frequently, the exact value of the divergence times are hard 
+Frequently, the exact values of the divergence times are hard 
 to determine because researchers typically lack precise knowledge 
@@
 \begin{itemize}
-   \item the use of vague priors on the divergence time parameters can 
-    lead to support models with few divergence time events. Thus, the 
+   \item The use of vague priors on the divergence time parameters can lead to
+       support models with few divergence events shared across taxa. Thus, the
        primary inference enabled by the approach is very sensitive to the
@@
 
-\citet{Oaks2012} suggest the primary cause of the spurious support for models
-with few divergence parameters is the greater marginal likelihoods of these
-models under vague uniform priors relative to more parameter-rich models.
+This is not surprising given the rich statistical literature that shows that
+model marginal likelihoods are quite sensitive to the priors used in Bayesian
+model selection
+\citep[e.g.,][]{Jeffreys1935,Lindley1957}.
+Accordingly, \citet{Oaks2012} suggest the primary cause of the spurious support
+for models with few divergence parameters is the greater marginal likelihoods
+of these models under vague uniform priors, relative to more parameter-rich
+models.
 Models with more divergence-time parameters integrate over \emph{much} greater
 parameter space with low probability of producing the data, yet relatively high
-prior density \citep{Lindley1957} imposed by the uniformly distributed prior on
-divergence times.
-Note, this is a statistical issue extrinsic to \msb, and the numerical
+prior density imposed by the uniformly distributed prior on divergence times.
+Note, this is a statistical issue, extrinsic to \msb, and the numerical
 approximation machinery of \msb could be sound.
+
 \citet{Hickerson2013} suggest the bias is the result of a numerical issue
@@
 biases found by \citet{Oaks2012}.
-We present results of additional analyses that, when combined with
-the results of \citet{Oaks2012}, strongly suggest small marginal likelihoods of
-models with more parameter space is a primary mechanism by which uniform priors
-are causing \msb to prefer highly clustered models of divergences.
+We present results of additional analyses that, when combined with the results
+of \citet{Oaks2012}, strongly suggest that \msb prefers models with few
+divergence-time parameters, because they have greater likelihoods when
+averaged over broad uniform priors on divergences times.
 
@@
 divergence-times and other nuisance parameters could increase the marginal
-likelihoods of models with more divergence-time parameters and thus reduce the
-bias.
+likelihoods of models with more divergence-time parameters and thus reduce
+spurious support for models of temporally clustered divergences.
 Alternatively, \citet{Hickerson2013} present an approach that uses empirically
@@
 that samples over empirically informed uniform priors is a fruitful solution
-for a model-choice method that is very sensitive to priors.
-
-In this paper, we discuss the potential theoretical and practical problems of
-using empirically informed priors for Bayesian model choice.
+for a model-choice method that is highly sensitive to priors.
+
+In this paper, we discuss the potential theoretical and practical
+considerations of using empirically informed priors for Bayesian model choice.
 Furthermore, we evaluate the empirical Bayesian model-averaging approach of
@@
 In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
-made an error by mixing different multiple units of time, which makes the
-results presented in their response difficult to interpret (see Supporting
-Information for details).
-Even when avoiding this error, we find Hickerson et al.'s
-\citeyear{Hickerson2013} empirical Bayesian model-averaging implementation of
-\msb remains biased toward clustered divergence models, and is predisposed to
-excluding true parameter space.
-The approach provides another means by which the method can ``escape'' models
-with large parameter space, and the method's bias toward smaller models
-remains.
-Importantly, this bias can manifest in the preference of models that exclude
-from consideration the true values of the model's parameters.
+made an error by mixing different units of time, which makes the results
+presented in their response difficult to interpret (see Supporting Information
+for details).
+Even when avoiding this error, we find their empirical Bayesian model-averaging
+implementation of \msb remains biased toward clustered divergence models.
+Furthermore, the approach provides another means by which the method can
+``escape'' models with large parameter space, which manifests in the preference
+of models that exclude from consideration the true values of the model's
+parameters.
 Our results, when combined with those of \citet{Oaks2012}, suggest that it is
@@
 broad enough to confidently contain the true values of parameters while being
-narrow enough to avoid bias toward models with less parameter space.
-However, recent work shows altnerative probability distributions without a
-hard upper bound (e.g. gamma) can accomodate prior uncertainty in divergence
-times while increasing the method's power to detect temporal variation in
-divergences
+narrow enough to avoid spurious support of models with less parameter space.
+However, recent work shows more flexible probability distributions without a
+hard upper bound (e.g., gamma) can accommodate prior uncertainty in divergence
+times without inhibiting the marginal likelihoods of models with more
+divergence-time parameters, and as a result, increase the method's power to
+detect temporal variation in divergences
 \citet{Oaks2014dpp}.
-
-% We discuss errors in this approach that render some of their results invalid
-% and leave the remaining results difficult to interpret.
-% Furthermore, we follow the recommendations of \citet{Oaks2012} and present
-% simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
-% approach.
-
-% \cwlnote{}{This part seems important, but not related to the error in units
-% discussed above}
-% Another important characteristic of the approach used by \citet{Hickerson2013}
-% that makes their results difficult to interpret is their use of narrow,
-% empirically guided uniform priors in an approximate-Bayesian model-choice
-% framework.
-% We discuss the risks associated with this approach in the next section, and
-% show how the results of \citet{Hickerson2013} likely exclude true values of the
-% model's parameters.
-
-% In addition to this error, we were not able to reproduce the results of
-% \citet{Hickerson2013}.
-% We followed their methods, generating $5\e6$ prior samples for each of the
-% eight models in their Table 1.
-% We retained the 10,000 samples from each model with the smallest Euclidean
-% distance to the observed summary statistics, and subsequently retained 10,000
-% samples from the remaining 80,000.
 
@@
 \msb for many of the model's parameters requires investigators to select broad
-priors to avoid excluding the truth \textit{a priori}, and how these broad
-uniform priors place too much density in improbable regions of parameter space.
+priors to avoid excluding the truth \textit{a priori}, resulting in too much
+density in improbable regions of parameter space.
 
@@
 they did so to assess the prior sensitivity of the method and to determine if
-uniform priors in such situations are narrow enough to avoid the bias toward
-models with few divergence-time parameters.
-The results of \citet{Oaks2012} and \citet{Hickerson2013} showed the method is
+uniform priors in such situations are narrow enough to avoid spurious support
+of models with few divergence-time parameters.
+The results of \citet{Oaks2012} and \citet{Hickerson2013} show the method is
 highly sensitive to the prior on divergence times.
 Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
-divergences even under the informed priors.
+divergences under the informed priors.
 
@@
 to an underestimate of our posterior uncertainty (Figure~\ref{figCoinFlip}).
-While this did not greatly affect our estimate of \myTheta{}, it did mislead us
-toward overconfidence in Model 2.
+While this did not greatly affect our estimate of \myTheta{}, it misled us
+to be overconfident in Model 2.
 % Our mode estimate of \myTheta{} under the empirical prior might be closer
@@
 
-% Let us say that principal investigator Mary and her new postdoc Will are
-% interested in the hypothesis that the mass of George Washington's periwig
-% renders the quarter dollar of the United States unfair.
-% That is to say their null hypothesis is that the probability of a US
-% quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
-% They can certainly evaluate this hypothesis, they can have undergraduate
-% worker, Joe, flip the coin for them while they tabulate the results.
-% But being Bayesians, before they call Joe into the lab, they agree on a
-% prior probability to place on the set of all possible probabilities that
-% the quarter will land heads when it is flipped.
-% Given that neither of them have a quarter, and their prior knowledge that Joe
-% moonlights as a magician, and is notorious for performing coin and card tricks
-% in the lab, they suspect there is a good chance that the quarter Joe uses will
-% be either two-headed or two-tailed.
-% So, knowing that the beta distribution is the conjugate prior for a binomial
-% likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
-% (Figure~\ref{figCoinFlip}).
-
-% Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
-% begin flipping it.
-% After five tosses, four of which was heads, Joe decides that academics are
-% crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
-% being computational biologists, are satisfied with their
-% empirical dataset of $y = 4$ heads out of $N = 5$ trials.
-% They know from the conjugacy of the beta prior, that the posterior distribution
-% has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
-% in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
-% of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
-% This allows them to plug these values into the beta cumulative distribution
-% function to determine that the posterior probability of their hypothesis is
-% $p(\theta < 0.5 | y=1, N=5) = 0.088$.
-% Given their prior belief and dataset, this indeed is the correct posterior
-% probability of the hypothesis, and Mary and Will should now update their
-% posterior belief accordingly.
-
-% However, reflecting upon the results of their experiment
-% (Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
-% Their prior looks very ``poorly selected,'' and if they had only known that the
-% coin was fair before the data were collected, they would have selected a much
-% better prior.
-% Clearly, from their results, they should have used a prior centered
-% around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
-% much ``better.''
-% Rather than resort to flipping the coin five more times, Mary and Will decide
-% to redo their analysis using the much better ``prior'' of $p(\theta) \sim
-% beta(4.5, 1.5)$.
-% This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
-% probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
-% Now convinced that the US quarter is unfair, albeit not due to the mass of
-% President Washington's head as they hypothesized, Will begins composing an
-% e-mail of complaint addressed to the U.S.\ Mint.
-
-% If Joe's flips were a representative sample, Mary and Will's empirical Bayes
-% estimate might very well be a better point estimate for the parameter
-% \myTheta{}.
-% However, their empirical Bayes estimate of the probability of their hypothesis
-% is incorrect and biased.
-% This simple example shows how parameter estimation is fundamentally different
-% from estimating the probability of a model.
-% While empirically informed priors can be used to obtain well-behaved parameter
-% estimators, using them for model choice is much less certain.
-
-% %CWL:
-% %I like the example, but it may be too cute. It is also too long. I think a
-% %contrived example is good, but if there could be an example that is more
-% %relevant to model choice and comparison of inaccurate posterior distributions,
-% %I think that would be better. This is an example of empirical Bayes, but not
-% %the potential problem with the use in msbayes. Unless I am interpreting wrong.
-% %I would be careful about making it to cute as well. You may intend to make it
-% %easy to understand for a general audience, but it can also seem like talking
-% %down to Hickerson et al, who we are responding to. It just depends on how you
-% %read it.
 % This can be demonstrated with a simple, albeit contrived, example.
@@
 22 taxa, it may be inappropriate to set this as the limit on the prior, because
-many of the taxon pairs are expected to have diverged at times older than the
+some of the taxon pairs are expected to have diverged at times older than the
 upper limit.
@@
 their approach is biased toward models with less parameter space, and as a
-consequence, is biased toward models that likely exclude the truth.
-I.e., the approach is biased toward estimating model-averaged posteriors
-dominated by models that exclude true values of parameters of the model.
-
-%CWL:This should be in the section on model-averaging, not empirical Bayes
-%limitations
-We explored the propensity of the empirically informed model-averaging approach
-of \citet{Hickerson2013} to exclude the truth in two ways.
+consequence, is biased toward estimating model-averaged posteriors dominated by
+models that exclude true values of parameters of the model.
+We explored this possibility in two ways.
 First, we re-analyzed the Philippines dataset using the model-averaging approach
@@
 less parameter space, we expect \msb will preferentially sample from this
-incorrect model yielding a marginal posterior that is incorrect (i.e., the
+incorrect model yielding a posterior that is incorrect (i.e., the
 model-averaged posterior will be dominated by an incorrect model that excludes
@@
 
-For our re-analysis of the Philippines dataset we followed the model-averaging
+\subsubsection*{Re-analyses of the Philippines dataset using empirical Bayesian
+model averaging}
+
+For our re-analyses of the Philippines dataset we followed the model-averaging
 approach of \citet{Hickerson2013}, but with a reduced set of prior models to
@@
 Our results show that the model-averaging approach of \citet{Hickerson2013}
-does not reduce the method's bias toward models with less parameter space,
-but rather allows it to manifest in a concerning way.
-The method strongly prefers the prior model with the narrowest distribution on
-divergence times across all three of our analyses, even when this model is
-almost certainly incorrect and excludes the true divergence times of the
-Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
+does not reduce the method's support of models with less parameter space.
+Rather, the method strongly prefers the prior model with the narrowest
+distribution on divergence times across all three of our analyses, even when
+this model is almost certainly incorrect and excludes the true divergence times
+of the Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
 
@@
 
+\subsubsection*{Simulation-based assessment of Hickerson et al.'s
+\citeyear{Hickerson2013} model averaging over empirical priors}
+
 To better quantify the propensity of Hickerson et al.'s
 (\citeyear{Hickerson2013}) approach to exclude the truth, we simulated 1000
-datasets in which the divergence times for the 22 population pairs are drawn
+datasets in which the divergence times for the 22-population pairs are drawn
 randomly from an exponential distribution with a mean of 0.5 ($\divt{} \sim
@@
 
-% \highLight{Add how we determined exclusion.}
 For each simulation replicate, we estimated the Bayes factor in favor
@@
 Our results show that the model-averaging approach of \citet{Hickerson2013}
-excludes the true values of parameters in 97\% of the replicates (90\% with
-GLM-regression adjustment), excluding up to 21 of the 22 true divergence times
-(Figure~\ref{figExclusionSimTau}).
+favors a model that excludes the true values of parameters in 97\% of the
+replicates (90\% with GLM-regression adjustment), excluding up to 21 of the 22
+true divergence times (Figure~\ref{figExclusionSimTau}).
 We also used the mode estimate of the preferred model for each replicate to
@@
 similar results.
-Furthermore, the posterior probability of excluding at least one true parameter
+Importantly, the posterior probability of excluding at least one true parameter
 value is very high in nearly all of the replicates
@@
 
-The results of our empirical and simulation-based analyses clearly demonstrate
-the danger of using narrow, empirically guided uniform priors in a Bayesian
-model-averaging framework.
-% This is especially concerning in \msb, which has already been demonstrated to
-% be biased away from models with more parameter space \citep{Oaks2012}.
-The consequence of this approach is the risk of obtaining a model-averaged
-posterior estimate that is heavily weighted toward incorrect models that
-exclude true values of model parameters.
+The results of the preceding empirical and simulation-based analyses clearly
+demonstrate the risk of using narrow, empirically guided uniform priors in a
+Bayesian model-averaging framework.
+The consequence of this approach is obtaining a model-averaged posterior
+estimate that is heavily weighted toward incorrect models that exclude true
+values of model parameters.
 
 \subsection*{Additional thoughts on empirical priors in Bayesian model choice}
+The preceding sections are not a general critique of Bayesian model averaging.
+Rather, model averaging can provide an elegant way of incorporating
+model uncertainty in Bayesian inference.
+However, when averaging over models with narrow and broad uniform priors in
+situations where the likelihood density is not expected to be uniformly
+distributed, the posterior could be dominated by models that exclude from
+consideration the true values of parameters due to the larger marginal
+likelihoods of the models that are integrated over less space with high prior
+weight yet low likelihood.
+
 Given the theoretical and practical issues with empirical Bayes approaches to
-Bayesian model choice discussed in the proceeding sections, it is quite clear
-why one should use caution before overly criticizing an investigator's choice
-of priors after having seen the resulting posterior.
+Bayesian model choice, it is clear why one should use caution before overly
+criticizing an investigator's choice of priors after having seen the resulting
+posterior.
 %In a strictly Bayesian world, this is cheating.
@@
 When limited to uniformly distributed priors, the alternative to priors that
-reflect prior uncertainty, as shown above, is to risk excluding the truth
-before performing the analysis.
+reflect prior uncertainty, as shown above, is to risk excluding the true values
+one seeks to estimate.
+However continuous distributions perhaps more appropriate as priors for
+positive real-valued parameters have been shown to greatly reduce spurious
+support for clustered divergence models while allowing prior uncertainty to be
+accommodated \citep{Oaks2014dpp}.
 
@@
 decrease bias toward clustered divergence models.
-We assess this possiblity using simulation-based analyses to evaluate the power
+We assess this possibility using simulation-based analyses to evaluate the power
 of Hickerson et al.'s (\citeyear{Hickerson2013}) approach.
@@
 simulated (Figure~\ref{figPowerNumExcluded}).
-The method is most likely to infer the extreme case of a single divergence event
-when populations diverged randomly over the past \globalcoalunit generations.
+The method is most likely to infer the extreme case of a single divergence
+event shared across 22 taxa when populations diverged randomly over the past
+\globalcoalunit generations.
 When divergences are random over the past $8\globalpopsize$ generations, the
-most likely inference is only two divergence-time parameters, and a single
+most likely inference is only two divergence events, and a single
 divergence is still estimated in more than 10\% of the replicates.
@@
 (\citeyear{Oaks2012}) Figure 5B).
-This supports the conclusions of \citet{Oaks2012} that this U-shaped
+This suggests this U-shaped
 prior coupled with the small marginal likelihoods of models with many
-divergence parameters, is a major cause of the method's bias toward
-clustered divergence models.
+divergence parameters, is a major cause of the method's bias toward clustered
+divergence models; this is consistent with \citet{Oaks2012} and confirmed by
+\citet{Oaks2014dpp}.
 
@@
 the criterion for incorrect inference of a single divergence event.
-There is strong support for a single divergence event in more than 90\% of the
-replicates when divergences are random over the past $2.4\globalpopsize$
-generations, and more than 60\% when over the past $3.2\globalpopsize$
-generations or less (Figure~\ref{figPowerOmegaProb}).
-
-Our results show that the empirical Baysian model-averaging approach is biased
-toward supporting the extreme case of a single divergence event when
-populations diverged randomly over the last $8\globalpopsize$ generations.
+By this criterion, there is strong support for a single divergence event in
+more than 90\% of the replicates when divergences are random over the past
+$2.4\globalpopsize$ generations, and more than 60\% when over the past
+$3.2\globalpopsize$ generations or less (Figure~\ref{figPowerOmegaProb}).
+
+Our results show that the empirical Baysian model-averaging approach leads to
+spuriously strong support for the extreme case of a single divergence event
+when populations diverged randomly over the last $8\globalpopsize$ generations.
 To put this on the scale roughly consistent with a vertebrate mitochondrial
@@
 
-The results of our power analyses further demonstrate the propensity of
+Also, the results of our power analyses further demonstrate the propensity of
 Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
 parameter values.
-Across all but one of the \divt{max} we simulated, the method excludes the
-truth in a large proportion of replicates, and across many of the \divt{max} it
-will exclude a large proportion of the true divergence time values
-(Figure~\ref{figPowerNumExcluded}).
-The posterior probability of excluding at least on true divergence value is
-also quite high across many of the \divt{max}
+Across all but one of the \divt{max} we simulated, in a large proportion of
+replicates the method favors a model that excludes the truth, and across many
+of the \divt{max} the preferred model will exclude a large proportion of the
+true divergence-time values (Figure~\ref{figPowerNumExcluded}).
+Importantly, the posterior probability of excluding at least on true divergence
+value is also quite high across many of the \divt{max}
 (Figure~\ref{figPowerProbExclusion}).
@@
 (Figure~\ref{figPowerNumExcluded}E).
-Again, this demonstrates the method's sensitiviy to priors.
-
-Given our results, we conclude that the approach of \citet{Hickerson2013}
-excludes regions of parameter space containing the truth, and lacks power to
-detect random variation in divergence times over a scale of $8\globalpopsize$
-generations or more.
-This roughly translates to millions of generations for mitochondrial loci, and
-hundreds of millions of generations for nuclear loci.
+Again, this demonstrates the method's sensitivity to priors.
+
+We find the model-averaging approach employing empirically informed uniform
+priors lacks power to detect random variation in divergence times over a scale
+of $8\globalpopsize$ generations or more, which roughly translates to millions
+of generations for mitochondrial loci.
+As a result, it is difficult to justify the risk of the approach to to exclude
+regions of parameter space containing the truth.
 
@@
 divergence-time prior of 0--1 coalescent units ago.
-They found that under these prior conditions \msb has the power, assuming a
-per-site rate of $1.92\e{-8}$ mutations per generation, to detect multiple
+They found that under these prior conditions \msb can, assuming a
+per-site rate of $1.92\e{-8}$ mutations per generation, detect multiple
 divergence events among 18 taxa when the true divergences were random over
@@
 of generations will add much insight about the evolutionary history of the
-taxa.
-Assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), with such
-prior information the method can reliably reject one divergence event when
-divergences are random over a period of 3 million generations.
-However, as noted by \citet{Oaks2012}, inferring more than one divergence time
+taxa; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$),
+this translates to a temporal resolution of 3 million generations or more.
+
+As noted by \citet{Oaks2012}, inferring more than one divergence time
 shared across all taxa does not confirm the method is working well when
@@
 yet spurious).
-Investigators should not limit their assessment of method's power to only
-differentiating inferences of one event or more (i.e. $\numt{} = 1$ versus
-$\numt{} > 1$).
+Thus, it is important that investigators not limit their assessment of method's
+power to only differentiating inferences of one event or more (i.e. $\numt{} =
+1$ versus $\numt{} > 1$).
 Rather, looking at the distribution of estimates, as in \citet{Oaks2012},
@@
 
-\section*{Model likelihoods or insufficient prior sampling as the cause of bias
-in \msb}
-% One of the points of disagreement between \citet{Hickerson2013} and
-% \citet{Oaks2012} is the underlying mechanism causing the bias toward models
-% with clustered divergences (i.e., models with few divergence events).
-It is important to determine the mechanism by which broad uniform priors cause
-\msb to support clustered models of divergence.
+\section*{The causes of support for models of co-divergence}
+To determine how best to improve the behavior of \msb, it is important to
+determine the mechanism by which broad uniform priors cause spurious support
+for clustered models of divergence.
 It is well established that vague priors can be problematic in Bayesian model
@@
 probability of producing the data and relatively high prior density will have
-smaller marginal likelihoods \citep{Jeffries1935,Lindley1957}.
+smaller marginal likelihoods \citep{Jeffreys1935,Lindley1957}.
 Given the uniformly distributed priors on divergence times (and other nuisance
@@
 prior density, and much of it with low likelihood.
-Thus, it is not surprising in light of this fundamental statistical issue that
-the method tends to support simple models.
+In light of this fundamental statistical issue, it is not surprising that the
+method tends to support simple models.
 
@@
 infrequently produce samples with many independent young divergence times.
-However, there are some issues with their argument.
-The probabilities they present assume the gene divergence times
-estimated by \citet{Oaks2012} are correct.
-The sole purpose of the estimates of ultrametric gene trees presented by
-\citet{Oaks2012} was to provide a very rough comparison of the gene divergence
-times across the 22 taxa.
+However, their argument assumes the gene divergence times presented
+in \citet{Oaks2012} were estimated without error.
+These estimates were intended to only provide a very rough comparison of the
+gene divergence times across the 22 taxa.
 These analyses assumed an arbitrary strict mutation rate of $2\e{-8}$ for all
@@
 
-If we make all of these assumptions and assume the gene divergence times are
-estimated without error, the Hickerson et al.'s \citeyear{Hickerson2013}
-argument only applies to one of the three different priors used by
-\citet{Oaks2012}.
+Even if we assume the gene divergence times are estimated without error,
+Hickerson et al.'s \citeyear{Hickerson2013} argument only applies to one of the
+three different priors used by \citet{Oaks2012}.
 The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5
 mya)$) closely mirrors the range of estimates of gene-divergence times.
-Applying Hickerson et al.'s (\citeyear{Hickerson2013}) probability equations
-demonstrates this prior is densely populated with samples with large numbers of
-divergence parameters with values younger than the estimated gene divergence
-estimates.
+Applying Hickerson et al.'s (\citeyear{Hickerson2013}) probability sampling
+probability argument demonstrates this prior is densely populated with samples
+with large numbers of divergence parameters with values younger than the
+estimated gene divergence estimates.
 Thus, if insufficient prior sampling is to blame for the bias, it should be
@@
 One example, as discussed by \citet{Oaks2012}, is that insufficient prior
-sampling should create higher variance in posterior estimates, and thus it
+sampling should create large variance among posterior estimates, and thus it
 should cause analyses to be highly sensitive to the number of samples drawn
 from the prior.
+Furthermore, if the sampling error causes bias toward models with less
+parameter space, as suggested by \citet{Hickerson2013}, we expect to see
+support for these models decrease as sampling increases.
 \citet{Oaks2012} do not see such sensitivity when they compare prior sample
@@
 To explore this prediction further, we repeat the analysis of the Philippines
-dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim U(0,
-10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
-(0.0005, 0.02)$), using a very large prior sample size of $10^8$.
+dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim
+U(0, 10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{}
+\sim (0.0005, 0.02)$), using a very large prior sample size of $10^8$.
 When we look at the trace of the estimates of the dispersion index of
@@
 (Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or
-GLM-regression-adjusted estimates.
-While sampling error is always a reality, clearly it is not playing
-a large role in the results of \citet{Oaks2012} or presented above.
-
-As discussed by \citet{Oaks2012}, a straightforward prediction of marginal
-likelihoods causing the preference for simple models is that the bias should
+GLM-regression-adjusted estimates, as predicted by \citet{Hickerson2013}.
+While sampling error is always a reality, it does not appear to be playing a
+large role in the results of \citet{Oaks2012} or presented above.
+
+As discussed by \citet{Oaks2012}, a straightforward prediction if marginal
+likelihoods are causing the preference for simple models is that the bias should
 disappear as the model generating the data converges to the prior.
@@
 \citet{Oaks2012}).
-These results are not clearly predicted if insufficient-sampling was causing
+We confirmed this same behavior for the model-averaging approach used by
+\citet{Hickerson2013} (see SI text and Figure~\ref{figValidationMCBehavior}.
+These results are not clearly predicted if insufficient sampling was causing
 the bias.
@@
 Thus, the results of the simulations of \citet{Oaks2012} are more consistent
-with larger marginal likelihoods of models with fewer divergence-time
-parameters.
-
-Furthermore, the results above demonstrate the model-averaging
-approach of \citet{Hickerson2013} is prefers models with narrower \divt{}
+with the fundamental statistical issue of the sensitivity of the marginal
+likelihoods to priors.
+
+Furthermore, the results presented herein demonstrate the model-averaging
+approach of \citet{Hickerson2013} prefers models with narrower \divt{}
 priors (Table~\ref{tabModelChoiceEmpirical} and
@@
 \ref{figPowerOmegaProb}).
-% This further suggests the primary cause is the uniform
-% priors reducing the marginal likelihoods of large models.
 In all of these analyses, each of the prior models have the same number of
 samples.
+Thus, while sampling error will always be present in any numerical Bayesian
+approximation method, insufficient sampling is an unlikely explanation for
+support of prior models with less parameters space.
 While analyses that sample each model proportional to their parameter space
-could be explored, it clear that the marginal likelihoods under broad uniform
-priors will be greater for simpler models.
-% Thus, it is difficult to explain the method's propensity toward the model with
-% the smallest parameter space via insufficient prior sampling.
-Taken together, the results presented here and in \citet{Oaks2012} support
-greater marginal likelihoods of simple models under broad priors as a primary
-mechanism by which \msb prefers models with clustered divergence times.
-Nonetheless, sampling error will always be present in any numerical
-Bayesian approximation method.
-
-\section*{Some general thoughts on the model of \msb}
-Our results demonstrating the poor behavior of the model-averaging approach
-proposed by \citet{Hickerson2013} are not too surprising.
-While this approach is trivial to implement, it is not a trivial change
-to the \msb model.
-To better understand the reason for this, it might help to step back and get a
-sense of the scale of the \msb model.
-To do this, we will use the dataset of 22 vertebrate taxon pairs of
-\citet{Oaks2012} as an example.
-
-Following the model description and notation of \citet{Oaks2012}, let us tally
-up all of the free parameters in the \msb model.  Under the simplest model in
+could be explored, it is clear that the marginal likelihoods under broad
+uniform priors on divergence times will be greater for models with fewer
+divergence-time dimensions.
+
+\section*{General thoughts on the model of \msb}
+Our findings are not very surprising in light of difficult inference problem
+\msb seeks to tackle.
+To get a sense of this task, let us tally up all the free parameters of the
+\msb model as applied to the dataset of \citet{Oaks2012}.
+% While this approach is trivial to implement, it is not a trivial change
+% to the \msb model.
+% To better understand the reason for this, it might help to step back and get a
+% sense of the scale of the \msb model.
+% To do this, we will use the dataset of 22 vertebrate taxon pairs of
+% \citet{Oaks2012} as an example.
+% Following the model description and notation of \citet{Oaks2012}, let us tally
+% up all of the free parameters in the \msb model.
+Under the simplest model in
 \msb (i.e., assuming no migration and
@@
 copies sampled from both populations of the pair.
-Because we only have a single locus for each taxon, there are no locus-specific
-\myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
-\locusRateHetShapeParameter shape parameter for the gamma prior distribution on
-\locusMutationRateScalar{}.
+% Because we only have a single locus for each taxon, there are no locus-specific
+% \myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
+% \locusRateHetShapeParameter shape parameter for the gamma prior distribution on
+% \locusMutationRateScalar{}.
 Lastly, there are between one and 22 divergence time parameters \divt{} in
 the vector \divtvector.
-Overall, for our Philippines data under the simplest model in \msb, there are
-581--602 free parameters (depending on the cardinality of \divtvector).
-Furthermore, under this rich model, the method is sampling over 1002 divergence
-models \citep[i.e., the number of integer partitions of $Y=22$;][]{Oaks2012}.
-
-This is a very difficult inference problem, and the method only uses four
-summary statistics calculated from the sequence alignment of each taxon pair:
+Overall, under the simplest model in \msb, for the dataset of \citet{Oaks2012}
+with 22 taxon pairs, there are 581--602 free parameters (depending on the
+number of divergence-time parameters in \divtvector).
+Furthermore, under this rich model, the method is estimating the probability of
+1002 divergence models \citep[i.e., the number of integer partitions of
+$Y=22$;][]{Oaks2012}.
+
+This is a very difficult inference problem, especially considering that all the
+information in the sequence alignment of each taxon pair is distilled into four
+summary statistics:
 $\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
@@
 More summary statistics can be used in \msb, but most are highly correlated
-with these four statistics (which are even highly correlated among themselves),
-and thus contribute little additional information about the data.
-
-The large number of parameters and divergence models relative to 
-summary statistics is undoubtedly a key reason the method is so sensitive
-to the prior distributions.
-It also likely contributes to the method's lack of robustness.
-% We use robustness here as a measure of a method's insensitivity to model
-% violations.
-Robustness is an extremely important characteristic of a method to gauge its
+with these four statistics, and thus contribute little additional information
+about the sequence data.
+The large number of parameters and divergence models relative to summary
+statistics is undoubtedly a key reason the method lacks robustness to prior
+conditions.
+Robustness is an important characteristic of a method to gauge its
 applicability to real-world data, because we know the model and priors will be
@@
 
-The approach of \citet{Hickerson2013} adds an additional dimension of model
-choice to the model. They expand the model to sample over eight prior models.
-This extends the original model to having 582-603 free parameters and, more
-importantly, sampling over 8016 unique model states across both divergence
-models and prior models.
-This is a non-trivial extension of the model and, given the method's very large
-number of parameters and divergence models relative to the information used
-from the data, likely plays a major role in the poor behavior of this approach.
-% Certainly, such a large modification of the model demands simulation-based
-% assessments of the behavior of the new model, as we provide here.
-
-In theory, the approach of \citet{Hickerson2013} is very appealing.  It sums
-over multiple candidate prior models to produce a posterior estimate
-marginalized over the uncertainty in prior choice.
-In general, Bayesian model-averaging is a powerful approach that leverages a
-great strength of Bayesian statistical procedures, namely the ability to
-obtain marginalized estimates that incorporate uncertainty in nuisance
-parameters.
-However, given the basic \msb model is already struggling to estimate
-a huge number of parameters and model probabilities with scant information
-from the data, it is not surprising that adding another dimension of
-model choice to the method causes problems.
-
-The recommendations of \citet{Oaks2012} for mitigating the bias and lack of
-robustness of \msb are actually similar to those of \citet{Hickerson2013}, but
-avoid the need for imposing additional model choice.
-\citet{Oaks2012} suggest that uniform priors are inappropriate for many
+The model-averaging approach of \citet{Hickerson2013} adds an additional
+dimension of model choice to the model.
+They expand the model to sample over eight prior models.
+This adds an additional free parameter to the model and, more importantly,
+forces the model to sample over 8016 unique models.
+While this approach is trivial to implement, it is a non-trivial extension of
+the model.
+Given the method's very large number of free parameters and models to sample
+relative to the information used from the data, is likely a major reason the
+method is so sensitive to priors, and tends to support models with the least
+parameter space when used with uniform priors.
+In theory, the model-averaging approach of \citet{Hickerson2013} is very
+appealing.
+It leverages a great strength of Bayesian statistical procedures, namely the
+ability to obtain marginalized estimates that incorporate uncertainty in
+nuisance parameters.
+However, when averaging over both narrow-empirical and diffuse uniform priors
+for a parameter that is expected to have a very non-uniform likelihood density,
+and in the context of a model-choice method that is highly sensitive to priors,
+it is not too surprising to find undesirable behavior.
+
+The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
+of \msb are similar to those of \citet{Hickerson2013}, but avoid the
+need for imposing an additional dimension of model choice.
+\citet{Oaks2012} suggest that uniform priors may not be appropriate for many
 parameters of the \msb model, and recommend the use of probability
 distributions from the exponential family.
-If we look at the prior distribution on the divergence time parameter \divt{}
-imposed by the model-averaging approach of \citet{Hickerson2013} we see
-it is a mixture of overlapping uniforms with lower limits of zero
-(Figure~\ref{figMCTauPrior}).
-This looks very much like an exponential distribution, except that each sample
-of divergence times are restricted to the hard bounds of one of the eight prior
-models.
-Thus, it seems more appropriate to place a gamma prior (the exponential being a
-special case of the gamma) on divergence times.
-This would provide investigators flexibility to represent their prior
-uncertainty in model parameters while avoiding broad uniform
-distributions and the need of costly model-averaging.
-This seems like a natural solution to the problem of skewed marginal
-likelihoods and/or sampling error, which are both likely exacerbated by the
-fact that uniform distributions are the only choice of prior for many of the
-parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
-\descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
-\bottleScalar{1}{}, \bottleScalar{2}{}, \locusRateHetShapeParameter,
-\migrationRate{}, \recombinationRate;][]{Oaks2012}.
+If we look at the prior distribution on divergence-time parameters imposed by
+the model-averaging approach of \citet{Hickerson2013} we see it is a mixture of
+overlapping uniforms with lower limits of zero (Figure~\ref{figMCTauPrior}).
+This looks very much like an exponential distribution, except that in any state
+of the model, all the divergence-time parameters are restricted to the hard
+bounds of one of the uniform distributions.
+Thus, it seems more appropriate to simply place a gamma prior (the exponential
+being a special case of the gamma) on divergence times.
+This would capture the prior uncertainty that \citet{Hickerson2013} are
+suggesting for divergence times (Figure~\ref{figMCTauPrior}) while avoiding
+costly model-averaging and the constraint that all divergence times must fall
+within the hard bounds of the current model state.
+It also would allow an investigator to place the majority of the prior density
+in regions of parameter space they believe, \emph{a priori}, are most
+plausible, but still capture uncertainty in the tails of distributions with low
+density.
+Most importantly, we have found the use of more flexible prior distributions in
+place of uniforms improves the power of the method to detect temporal variation
+in divergences and reduces the bias toward models of clustered divergences
+\citep{Oaks2014dpp}.
+% This seems like a natural solution to the problem of skewed marginal
+% likelihoods and/or sampling error, which are both likely exacerbated by the
+% fact that uniform distributions are the only choice of prior for many of the
+% parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
+% \descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
+% \bottleScalar{1}{}, \bottleScalar{2}{}, \locusRateHetShapeParameter,
+% \migrationRate{}, \recombinationRate;][]{Oaks2012}.
 
 % Importantly, this strategy is easy to implement.
-Given that ABC methods only need to draw random values from prior distributions
-(hence there are no difficult prior and proposal ratios to calculate, etc.) it
-is easy to use the most appropriate distributions on parameters.
-Also, this approach reduces the temptation of empirically guided priors.
-The investigator can place the majority of the prior density in regions of
-parameter space they believe, \emph{a priori}, are most plausible, but still
-capture uncertainty in the tails of distributions with low density.
-As discussed in \citet{Oaks2012} the use of uniform priors necessitate
-broad priors to avoid excluding the truth \emph{a priori}.
-Thus, to avoid the behavior of the method under these uniform priors, using the
-data is a very tempting, albeit risky, option \citep{Hickerson2013}.
-More appropriate prior distributions would alleviate this issue.
+% Given that ABC methods only need to draw random values from prior distributions
+% (hence there are no difficult prior and proposal ratios to calculate, etc.) it
+% is easy to use the most appropriate distributions on parameters.
+% Also, this approach reduces the temptation of empirically guided priors.
+% The investigator can place the majority of the prior density in regions of
+% parameter space they believe, \emph{a priori}, are most plausible, but still
+% capture uncertainty in the tails of distributions with low density.
+% As discussed in \citet{Oaks2012} the use of uniform priors necessitate
+% broad priors to avoid excluding the truth \emph{a priori}.
+% Thus, to avoid the behavior of the method under these uniform priors, using the
+% data is a very tempting, albeit risky, option \citep{Hickerson2013}.
+% More appropriate prior distributions would alleviate this issue.
 
@@
 truth \emph{a priori}.
-This likely reduces the marginal likelihoods of models with greater parameter
+These broad priors reduce the marginal likelihoods of models with greater parameter
 space, either due to more divergence time parameters or broader prior
@@
 Whether or not one chooses to use empirically informed priors, our results
-suggest that it is difficult to choose a uniformly distributed prior on
+suggest that it is difficult to choose an uniformly distributed prior on
 divergence times that is broad enough to confidently contain the true values of
@@
 parameter space.
+While we do not assume ``that all previous \msb results are invalid,'' as
+suggested by \citet{Hickerson2013}, we do conservatively recommend that the
+common inference of temporally clustered divergences \citep{Barber2010,
+    Bell2012, Carnaval2009, Chan2011, Daza2010, Hickerson2006, Huang2011,
+Lawson2010, Leache2007, Plouviez2009, Stone2012, Voje2009}, when not
+accompanied with the necessary analyses to assess the robustness and temporal
+resolution of such results, should be treated with caution, because the method
+has been shown to spuriously infer clustered divergences over a range of prior
+conditions.
 Fortunately, alternative prior probability distributions allow prior
@@
 
-We do not ``assumed that all previous \msb results are invalid,'' as suggested
-by \citet{Hickerson2013}.
-Rather, we conservatively recommend that the common inference of temporally
-clustered divergences \citep{Barber2010, Bell2012, Carnaval2009, Chan2011,
-Daza2010, Hickerson2006, Huang2011, Lawson2010, Leache2007, Plouviez2009,
-Stone2012, Voje2009}, when not accompanied with the necessary simulation-based
-analyses to guide the interpretation of such results, should be treated with
-caution, because the method has been shown to spuriously infer clustered
-divergences over a range of prior conditions.
-
 The work presented herein follows the principles of Open Notebook Science.
-All aspects of the work were recorded in real-time via version-control
-software and are publicly available at
+All aspects of the work were recorded in real-time via version-control software
+and are publicly available at
 \href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
@@
 \section*{Acknowledgments}
-We thank the National Science Foundation for supporting this work (DEB
-1011423).
+J.\ Oaks thanks the National Science Foundation for supporting this work (DEB
+1011423 and DBI 1308885).
 J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate

### 2013-12-06 manuscripts/reply/short-reply.tex
Editing sections throughout `short-reply` ms.

@@
     at the same time can be a potentially persuasive argument that a set of taxa
-    are affected by the same geographic events.
+    were affected by the same geographic events.
     Unfortunately, estimating divergence times precisely when one lacks
@@
     Furthermore, simulations provide insight into the temporal resolution of
-    the method, which in turn helps guide appropriate interpretation of
-    results.
+    the method, which in turn helps guide interpretation of results.
 
@@
 \begin{itemize}
-   \item The use of vague priors on the divergence time parameters can lead to
+   \item The use of vague priors on the divergence-time parameters can lead to
        support models with few divergence events shared across taxa. Thus, the
@@
 for details).
-Even when avoiding this error, we find their empirical Bayesian model-averaging
-implementation of \msb remains biased toward clustered divergence models.
+Here, we avoid this error, but still find their approach to be biased toward
+clustered divergence models.
 Furthermore, the approach provides another means by which the method can
@@
 (Figure~\ref{figCoinFlip}).
-We collect data by flipping the coin $N=10$ times, $y=3$ of which land heads.
+We collect data by flipping the coin of interest $N=10$ times, $y=3$ of which
+land heads.
 Given the beta distribution is a conjugate prior for a binomial likelihood, the
@@
 see Table~6 of \citet{Oaks2012}).
+
 The strong preference for the questionable priors on divergence times (their
@@
 model uncertainty in Bayesian inference.
-However, when averaging over models with narrow and broad uniform priors in
-situations where the likelihood density is not expected to be uniformly
-distributed, the posterior could be dominated by models that exclude from
+However, when averaging over models with narrow and broad uniform priors on a
+parameter with a likelihood density that is not expected to be uniformly
+distributed, the posterior can be dominated by models that exclude from
 consideration the true values of parameters due to the larger marginal
@@
 decrease bias toward clustered divergence models.
-We assess this possibility using simulation-based analyses to evaluate the power
-of Hickerson et al.'s (\citeyear{Hickerson2013}) approach.
+We assess this possibility using simulation-based analyses.
 Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
@@
 \globalcoalunit generations.
-All other parameters were identically distributed as the prior.
-We use the same $5\e6$ samples from the same set of five prior models as above
-($M_1$, $M_2$, $M_3$, $M_4$, and $M_5$).
-As described above, for each simulated dataset, we approximated the posterior
+All other parameters were identically distributed as the prior models.
+As above, we generated a prior sample of $5\e6$ total samples from the five
+prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$
+(Table~\ref{tabModelChoiceEmpirical}).
+For each simulated dataset, we approximated the posterior
 by retaining 1000 samples from the prior with the smallest Euclidean distance
-from the true summary statistics.
+from the true summary statistics, as described above.
 In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
@@
 we simulated, whether evaluating the unadjusted
-(Figure~\ref{figPowerAccOmegaMedian}) or GLM-adjusted
-(Figure~\ref{figPowerAccOmegaModeGLM}) posterior estimates.
+(Figure~S\ref{figPowerAccOmegaMedian}) or GLM-adjusted
+(Figure~S\ref{figPowerAccOmegaModeGLM}) posterior estimates.
 The method only estimates \vmratio{} relatively well when the simulated
 distribution of divergence times is identical to one of the prior models
-(Figures~\ref{figPowerAccOmegaMedian}E \& \ref{figPowerAccOmegaModeGLM}E).
+(Figures~S\ref{figPowerAccOmegaMedian}E \& S\ref{figPowerAccOmegaModeGLM}E).
 This is consistent with the findings of \citet{Oaks2012}  and
@@
 (\citeyear{Oaks2012}) Figure 5B).
-This suggests this U-shaped
-prior coupled with the small marginal likelihoods of models with many
-divergence parameters, is a major cause of the method's bias toward clustered
-divergence models; this is consistent with \citet{Oaks2012} and confirmed by
-\citet{Oaks2014dpp}.
+This suggests this U-shaped prior coupled with the small marginal likelihoods
+of models with many divergence parameters, is a major cause of the method's
+bias toward clustered divergence models; this is consistent with
+\citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.
 
@@
 method strongly supports one divergence in greater than 27\% of the replicates
-across all the \divt{max} we simulated (Figure~\ref{figPowerOmegaProb}).
-Following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
+across all the \divt{max} we simulated (Figure~\ref{figPowerOmegaProb});
+following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
 the criterion for incorrect inference of a single divergence event.
-By this criterion, there is strong support for a single divergence event in
+Furthermore, there is strong support for a single divergence event in
 more than 90\% of the replicates when divergences are random over the past
 $2.4\globalpopsize$ generations, and more than 60\% when over the past
-$3.2\globalpopsize$ generations or less (Figure~\ref{figPowerOmegaProb}).
-
-Our results show that the empirical Baysian model-averaging approach leads to
+$3.2\globalpopsize$ generations (Figure~\ref{figPowerOmegaProb}).
+
+Our results show that the empirical Bayesian model-averaging approach leads to
 spuriously strong support for the extreme case of a single divergence event
@@
 million generations.
+Given the empirically Bayesian model-averaging approach of \citet{Hickerson2013}
+lacks power to detect random variation in divergence times over such
+timescales, it is difficult to justify the risk of the approach to exclude
+regions of parameter space containing the truth.
 
@@
 
-We find the model-averaging approach employing empirically informed uniform
-priors lacks power to detect random variation in divergence times over a scale
-of $8\globalpopsize$ generations or more, which roughly translates to millions
-of generations for mitochondrial loci.
-As a result, it is difficult to justify the risk of the approach to to exclude
-regions of parameter space containing the truth.
-
 %CWL: I would suggest removing this section
-\section*{The use of power analysis to guide applications of \msb}
+\section*{The importance of power analysis to guide applications of \msb}
 \citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
@@
 power analyses to determine the method's power for their dataset, and decide if
-\msb has sufficient power to address their hypotheses; in the case of the
-Philippines dataset, it did not.
+\msb has sufficient temporal resolution to address their hypotheses; in the
+case of the Philippines dataset, it did not.
 When doing so, it is important to consider what prior conditions are
@@
 this window.
-Certainly, it is a rare for there to be enough \emph{a priori} information to
+Certainly, it is  rare for there to be enough \emph{a priori} information to
 be certain that all taxa diverged within the last 4$\globalpopsize$ generations
@@
 However, \citet{Hickerson2013} conclude that the bias is due to insufficient
-prior sampling.
+sampling.
 They argue the widest of the three priors used by \citet{Oaks2012} would
-infrequently produce samples with many independent young divergence times.
-However, their argument assumes the gene divergence times presented
-in \citet{Oaks2012} were estimated without error.
-These estimates were intended to only provide a very rough comparison of the
+infrequently produce samples with many independent population divergence times
+as recent as the estimated gene divergence times presented in \citet{Oaks2012}.
+However, this argument assumes the gene divergence times presented in
+\citet{Oaks2012} were estimated without error.
+These estimates were intended to only provide a rough comparison of the
 gene divergence times across the 22 taxa.
@@
 mya)$) closely mirrors the range of estimates of gene-divergence times.
-Applying Hickerson et al.'s (\citeyear{Hickerson2013}) probability sampling
-probability argument demonstrates this prior is densely populated with samples
-with large numbers of divergence parameters with values younger than the
-estimated gene divergence estimates.
+Applying Hickerson et al.'s (\citeyear{Hickerson2013}) sampling-probability
+argument demonstrates this prior is densely populated with samples with large
+numbers of divergence parameters with values younger than the estimated gene
+divergence estimates.
 Thus, if insufficient prior sampling is to blame for the bias, it should be
@@
 from the prior.
-Furthermore, if the sampling error causes bias toward models with less
+Furthermore, if sampling error is biased toward models with less
 parameter space, as suggested by \citet{Hickerson2013}, we expect to see
@@
 divergence times (\vmratio{}) as the prior samples accumulate
-(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or
-GLM-regression-adjusted estimates, as predicted by \citet{Hickerson2013}.
+(Figure~\ref{figSamplingError}) we do not see the trend predicted by
+\citet{Hickerson2013} in either the unadjusted or GLM-regression-adjusted
+estimates.
 While sampling error is always a reality, it does not appear to be playing a
@@
 We confirmed this same behavior for the model-averaging approach used by
-\citet{Hickerson2013} (see SI text and Figure~\ref{figValidationMCBehavior}.
+\citet{Hickerson2013} (see SI text and Figure~\ref{figValidationMCBehavior}).
 These results are not clearly predicted if insufficient sampling was causing
@@
 Thus, the results of the simulations of \citet{Oaks2012} are more consistent
-with the fundamental statistical issue of the sensitivity of the marginal
-likelihoods to priors.
-
-Furthermore, the results presented herein demonstrate the model-averaging
-approach of \citet{Hickerson2013} prefers models with narrower \divt{}
-priors (Table~\ref{tabModelChoiceEmpirical} and
+with the fundamental sensitivity of marginal likelihoods to priors.
+
+This is further demonstrated by the results presented herein that show the
+model-averaging approach of \citet{Hickerson2013} prefers models with narrower
+\divt{} priors (Table~\ref{tabModelChoiceEmpirical} and
 Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
@@
 approximation method, insufficient sampling is an unlikely explanation for
-support of prior models with less parameters space.
+support of prior models with less parameter space.
 While analyses that sample each model proportional to their parameter space
@@
 \msb seeks to tackle.
-To get a sense of this task, let us tally up all the free parameters of the
-\msb model as applied to the dataset of \citet{Oaks2012}.
+To get a sense of the challenge \msb is faced with, we tally up all the free
+parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
 % While this approach is trivial to implement, it is not a trivial change
@@
 % up all of the free parameters in the \msb model.
-Under the simplest model in
-\msb (i.e., assuming no migration and
-no intra-locus recombination), the number of parameters for each
-taxon pair include:
+Under the simplest model in \msb (i.e., assuming no migration and no
+intra-locus recombination), the number of parameters for each taxon pair
+include:
 The population sizes of the ancestral and descendant populations
@@
 the vector \divtvector.
-Overall, under the simplest model in \msb, for the dataset of \citet{Oaks2012}
-with 22 taxon pairs, there are 581--602 free parameters (depending on the
-number of divergence-time parameters in \divtvector).
+Overall, when applying the simplest model in \msb to the dataset of
+\citet{Oaks2012} with 22 taxon pairs, there are 581--602 free parameters
+(depending on the number of divergence-time parameters in \divtvector).
 Furthermore, under this rich model, the method is estimating the probability of
@@
 More summary statistics can be used in \msb, but most are highly correlated
-with these four statistics, and thus contribute little additional information
-about the sequence data.
+with the four default statistics, and thus contribute little additional
+information about the parameters from the sequence data.
 The large number of parameters and divergence models relative to summary
@@
 the model.
-Given the method's very large number of free parameters and models to sample
-relative to the information used from the data, is likely a major reason the
-method is so sensitive to priors, and tends to support models with the least
-parameter space when used with uniform priors.
+% Given the method's very large number of free parameters and models to sample
+% relative to the information used from the data, is likely a major reason the
+% method is so sensitive to priors, and tends to support models with the least
+% parameter space when used with uniform priors.
 In theory, the model-averaging approach of \citet{Hickerson2013} is very
@@
 and in the context of a model-choice method that is highly sensitive to priors,
-it is not too surprising to find undesirable behavior.
+it is not too surprising to find the approach struggles.
 

### 2013-12-07 manuscripts/reply/reply-si-main.tex
Some minor editing of the reply ms.
Sending this version to co-authors.

--- 2013-11-19 manuscripts/reply/reply-si-main.tex
@@
-%CWL:
-%It may be important to point out the error made in their paper, but it can
-%also be nit-picky, it depends on how the error impacts the overall message of
-%our response.  We should strive to keep the response short and directed at the
-%big picture items. If there is a way to still point out the error, then maybe
-%some of this section should be used
 \section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
@@
 different prior models (see Table 1 of \citet{Hickerson2013}).
-However, there is a fundamental error in the methodology of
-\citet{Hickerson2013}: their model mixes different time units.
+However, there was an error in their methodology; their model mixes different
+units of time.
 
@@
 
-\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
+\section*{Additional clarifications from \citet{Hickerson2013}}
+
+\subsection*{Saturation of summary statistics}
+\citet{Hickerson2013} claim the priors used by \citet{Oaks2012} ``cause much of
+the explored parameter space to be beyond the threshold of saturation in most
+mtDNA genes.'' To explore this possibility, we simulated datasets under prior
+settings that match two of the three priors used by \citet{Oaks2012}:
+$\meanDescendantTheta{} \sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim
+U(0.0005, 0.02)$.
+Under this prior, we draw divergence time parameters from a uniform
+distribution of $U(0, 20)$ coalescent units, simulate datasets, and plot the
+\divt{} values against the summary statistics calculated from the resulting
+datasets (Figure~\ref{figSaturationPlot}).
+Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
+and 10 coalescent units suffered from little affect from saturation.
+Even at divergence times of 20 coalescent units, there is still signal in the
+summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
+Thus, the assertion of \citet{Hickerson2013} does not apply to at least
+two of the priors used by \citet{Oaks2012} and, as a result, does not
+explain the bias they found.
+
+\subsection*{Graphical prior comparisons}
+\citet{Hickerson2013} advocate the use of what they call graphical checks of
+prior models.
+This prior-predictive approach generating a small number (1000) of random
+samples from the prior and plotting the resulting summary statistics in
+comparison to the observed statistics to see if they coincide (see Figure 1 of
+\citet{Hickerson2013}).
+% As we show above, this strategy can be misleading, because the resulting plots
+% of this approach have little correlation with the appropriateness of priors.
+Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
+dataset analyzed by \citet{Hickerson2013}), we do not expect that 1000
+\emph{random} draws from the vast prior parameter space will yield data and
+summary statistics consistent with the observed data.
+In fact, when such random draws are tightly clustered around the observed
+statistics, this can be an indication that the prior is over-fit, as we show in
+the main text (Table~\ref{tabModelChoiceEmpirical} and Figure~S\ref{figPCA}).
+Thus, using such plots to select priors should be avoided, and the use of
+posterior-predictive analyses would be much more informative about the overall
+fit of models.
+
+\subsection*{Differing utilities of \numt{} and \vmratio{} in \msb}
 The main parameter of the \msb model is the vector of divergence
@@
 They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
-\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
+\citet{Hickerson2013}), which is essentially a plot of sample size versus
+variance.
 This plot shows, not surprisingly, that \vmratio{} has very little information
@@
 better than \numt{}.''
-However, \msb is a model-choice estimator, and hence the goal is to estimate the
-posterior probability of divergence models.
-\citet{Oaks2012} demonstrate that even when all assumptions of the model are
-met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
-4), whereas \numt{} performs better.
+However, \citet{Oaks2012} demonstrate that even when all assumptions of the
+model are met, \vmratio is a poor model-choice estimator (see plots B, D \& F
+of Figure 4), whereas \numt{} performs better.
 Furthermore, \vmratio{} is limited to estimating the probability of only a
-single divergence model (the one divergence model), and thus its utility for
+single model (the one-divergence model), and thus its utility for
 model choice is extremely limited.

### 2013-12-07 manuscripts/reply/short-reply.tex
Some minor editing of the reply ms.
Sending this version to co-authors.

@@
 We confirmed this same behavior for the model-averaging approach used by
-\citet{Hickerson2013} (see SI text and Figure~\ref{figValidationMCBehavior}).
+\citet{Hickerson2013} (see SI text and Figure~S\ref{figValidationMCBehavior}).
 These results are not clearly predicted if insufficient sampling was causing
@@
 density.
-Most importantly, we have found the use of more flexible prior distributions in
-place of uniforms improves the power of the method to detect temporal variation
+Most importantly, we have found the use of more flexible distributions in place
+of uniform priors improves the power of the method to detect temporal variation
 in divergences and reduces the bias toward models of clustered divergences
@@
 \section*{Conclusions}
-We demonstrate how the approximate Bayesian model choice method implemented in
+We demonstrate how the approximate Bayesian model-choice method implemented in
 \msb can be strongly biased away from models with greater parameter space.
@@
 truth \emph{a priori}.
-These broad priors reduce the marginal likelihoods of models with greater parameter
-space, either due to more divergence time parameters or broader prior
-distributions on those parameters.
+These broad priors reduce the marginal likelihoods of models with more
+divergence-time parameters or broader prior distributions on those parameters.
 We show that the empirical Bayesian model-averaging approach of

### 2013-12-10 power-comparison/bin/plot-power-results-comparison.py
Working on plotting script for power analyses.
Updating `plot-power-results-comparison.py` script. It now creates
decent looking plots of all power results.

@@ -2,6 +2,7 @@
 
 import os
 import sys
+import argparse
 import matplotlib
 
 from pymsbayes.utils.parsing import DMCSimulationResults, spreadsheet_iter
@@ -11,6 +12,7 @@ from pymsbayes.utils import probability
 from pymsbayes.plotting import (PowerPlotGrid, ProbabilityPowerPlotGrid,
         AccuracyPowerPlotGrid)
 from pymsbayes.utils.messaging import get_logger
+from pymsbayes.utils.argparse_utils import arg_is_file, arg_is_dir
 import project_util
 
 _LOG = get_logger(__name__)
@@ -107,15 +109,16 @@ def parse_results(dmc_sim,
 def create_plots(info_path,
         observed_prefixes_to_include = None,
         observed_suffixes_to_include = None,
-        prior_suffixes_to_include = None):
+        prior_suffixes_to_include = None,
+        output_dir = None):
     num_columns = len(observed_suffixes_to_include)
     matplotlib.rc('text',**{'usetex': True})
     dmc_sim = DMCSimulationResults(info_path)
     prior_configs = {}
     for k, v in dmc_sim.prior_index_to_config.iteritems():
         prior_configs[k] = MsBayesConfig(v)
-    # output_dir = os.path.join(os.path.dirname(info_path), 'plots')
-    output_dir = project_util.IMAGE_DIR
+    if not output_dir:
+        output_dir = project_util.IMAGE_DIR
     if not os.path.exists(output_dir):
         os.mkdir(output_dir)
     prior_prob_omega_less_than = {
@@ -207,8 +210,8 @@ def create_plots(info_path,
             row_label_size = 28.0
             row_label_offset = 0.08
             draw_bayes_factor_line = False
-            if num_columns > 4:
-                draw_bayes_factor_line = True
+            # if num_columns > 4:
+            #     draw_bayes_factor_line = True
 
 
             psi_plot = PowerPlotGrid(
@@ -256,8 +259,8 @@ def create_plots(info_path,
             psi_prob_sub_plot_map[prior_name] = psi_prob_plot.subplots
             for cfg, sp in psi_prob_plot.cfg_to_subplot.iteritems():
                 sp.set_left_text('')
-                if num_columns <= 4:
-                    sp.set_right_text('')
+                # if num_columns <= 4:
+                sp.set_right_text('')
 
             # psi_prob_plot_glm = ProbabilityPowerPlotGrid(
[…]

### 2013-12-10 validation/bin/plot-validation-model-choice-comparison.py
Updating validation plotting scripts.
Updating names of output plot files.

@@ -115,7 +115,7 @@ def main_cli():
         pg.reset_figure()
 
         pg.savefig(os.path.join(project_util.IMAGE_DIR,
-                'validation-model-choice-summary-' + variable + '.pdf'))
+                'validation-model-choice-' + variable + '.pdf'))
 
 if __name__ == '__main__':
     main_cli()

### 2013-12-10 validation/bin/plot-validation-parameter-comparison.py

@@ -130,7 +130,7 @@ def main_cli():
         pg.reset_figure()
 
         pg.savefig(os.path.join(project_util.IMAGE_DIR,
-                'validation-accuracy-summary-' + variable + '.pdf'))
+                'validation-accuracy-' + variable + '.pdf'))
 
 if __name__ == '__main__':
     main_cli()

### 2013-12-11 manuscripts/dpp-msbayes/dpp-msbayes-si-figures.tex
Working on dpp ms.
Incorporating new plots into SI figures (`dpp-msbayes-si-figures.tex`).
Adding some macros to `macros.tex` for the new figure captions.
Also, adding SI to the main `dpp-msbayes.tex` file.

%% Validation accuracy
\siFigure{../../validation/images/validation-accuracy-psi.pdf}{
    \validationAccuracyComparisonCaption{unadjusted}{\divTimeNum}
    \jitterComment
}{figValidationAccuracyPsi}

\siFigure{../../validation/images/validation-accuracy-psi-glm.pdf}{
    \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeNum}
    \jitterComment
}{figValidationAccuracyPsiGlm}

\siFigure{../../validation/images/validation-accuracy-omega.pdf}{
    \validationAccuracyComparisonCaption{unadjusted}{\divTimeDispersion}
}{figValidationAccuracyOmega}

\siFigure{../../validation/images/validation-accuracy-omega-glm.pdf}{
    \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeDispersion}
}{figValidationAccuracyOmegaGlm}

\siFigure{../../validation/images/validation-accuracy-time.pdf}{
    \validationAccuracyComparisonCaption{unadjusted}{\divTimeMean}
}{figValidationAccuracyTime}

\siFigure{../../validation/images/validation-accuracy-time-glm.pdf}{
    \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeMean}
}{figValidationAccuracyTimeGlm}

%% Validation model choice
\siFigure{../../validation/images/validation-model-choice-psi.pdf}{
    \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeNum = 1$}
}{figValidationModelChoicePsi}

\siFigure{../../validation/images/validation-model-choice-psi-glm.pdf}{
    \validationModelChoiceComparisonCaption{GLM-adjusted}{$\divTimeNum = 1$}
}{figValidationModelChoicePsiGlm}

\siFigure{../../validation/images/validation-model-choice-omega.pdf}{
    \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeDispersion < 0.01$}
}{figValidationModelChoiceOmega}

\siFigure{../../validation/images/validation-model-choice-omega-glm.pdf}{
    \validationModelChoiceComparisonCaption{GLM-adjusted}{$\divTimeDispersion < 0.01$}
}{figValidationModelChoiceOmegaGlm}

%% Validation ordered models
\siFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
    \validationAccuracyCaption{\modelDPPOrdered}{\modelDPPOrdered}
}{figValidationAccuracyDPPOrderedDPPOrdered}

\siFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
    \validationModelChoiceCaption{\modelDPPOrdered}{\modelDPPOrdered}
}{figValidationModelChoiceDPPOrderedDPPOrdered}

%% Power accuracy results
\siFigure{../../power-comparison/images/old-power-omega-accuracy-6.pdf}{
    \powerAccuracyComparisonCaption{\powerSeriesOld}
}{figPowerAccuracyOld}

\siFigure{../../power-comparison/images/uniform-power-omega-accuracy-6.pdf}{
    \powerAccuracyComparisonCaption{\powerSeriesUniform}
}{figPowerAccuracyUniform}

\siFigure{../../power-comparison/images/exp-power-omega-accuracy-6.pdf}{
    \powerAccuracyComparisonCaption{\powerSeriesExp}
}{figPowerAccuracyExp}

%% Power psi results
\siFigure{../../power-comparison/images/old-power-psi-6.pdf}{
    \powerComment{\powerSeriesOld}
    \powerPsiComment{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siFigure{../../power-comparison/images/uniform-power-psi-6.pdf}{
    \powerComment{\powerSeriesUniform}
    \powerPsiComment{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siFigure{../../power-comparison/images/exp-power-psi-6.pdf}{
    \powerComment{\powerSeriesExp}
    \powerPsiComment{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

%% Power dispersion results
\siFigure{../../power-comparison/images/old-power-omega-6.pdf}{
    \powerComment{\powerSeriesOld}
    \powerDispersionComment{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siFigure{../../power-comparison/images/uniform-power-omega-6.pdf}{
    \powerComment{\powerSeriesUniform}
    \powerDispersionComment{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siFigure{../../power-comparison/images/exp-power-omega-6.pdf}{
    \powerComment{\powerSeriesExp}
    \powerDispersionComment{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

%% Power psi probability results
\siFigure{../../power-comparison/images/old-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesOld}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siFigure{../../power-comparison/images/uniform-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesUniform}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siFigure{../../power-comparison/images/exp-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesExp}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

%% Power dispersion prob results
\siFigure{../../power-comparison/images/old-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesOld}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siFigure{../../power-comparison/images/uniform-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesUniform}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siFigure{../../power-comparison/images/exp-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesExp}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

### 2013-12-11 manuscripts/dpp-msbayes/dpp-msbayes.tex
Working on dpp ms.
Incorporating new plots into SI figures (`dpp-msbayes-si-figures.tex`).
Adding some macros to `macros.tex` for the new figure captions.
Also, adding SI to the main `dpp-msbayes.tex` file.

@@
 \input{../utils/macros.tex}
+
+\newcommand{\msTitle}{An Improved Approximate-Bayesian Model-choice Method for
+    Estimating Shared Evolutionary History \xspace}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-        \MakeUppercase{\large\bfseries An Improved Approximate-Bayesian
-            Model-choice Method for Estimating Shared Evolutionary History}
+        \MakeUppercase{\large\bfseries \msTitle}
 
@@
 
-\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape Supporting Information}}
+% PUT MAIN TEXT CITATION HERE
+{
+\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape supporting information}}
+\begin{thebibliography}{1}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{\texttt{#1}}
+\providecommand{\urlprefix}{URL }
 
-% PUT MAIN TEXT CITATION HERE
-% \begin{thebibliography}{1}
-% \providecommand{\natexlab}[1]{#1}
-% \providecommand{\url}[1]{\texttt{#1}}
-% \providecommand{\urlprefix}{URL }
+\bibitem[{Oaks and Holder (2014)Oaks and Holder}]{Oaks2014dpp}
+Oaks, J.~R.\ and M.~T.\ Holder,
+  % 2013.
+\newblock \msTitle
+% \newblock Evolution 67:991--1010.
 
-% \bibitem
+\end{thebibliography}
+}
 
-% \end{thebibliography}
+\newpage
+\singlespacing
 
-%% SUPPL TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\setcounter{table}{0}
+\input{dpp-msbayes-si-figures.tex}
 

### 2013-12-11 manuscripts/utils/macros.tex
Working on results of dpp ms.
Also adding table of models used to generated datasets for
power analyses.
Modifying dpp ms.
Adding software info and links to beginning of methods.
More editing of dpp ms.
Working on dpp ms.
Incorporating new plots into SI figures (`dpp-msbayes-si-figures.tex`).
Adding some macros to `macros.tex` for the new figure captions.
Also, adding SI to the main `dpp-msbayes.tex` file.

@@
 \newcommand{\modelDPPOrdered}{\ensuremath{M^{\circ}_{DPP}}\xspace}
-\newcommand{\modelUniform}{\ensuremath{M_{uniform}}\xspace}
+\newcommand{\modelUniform}{\ensuremath{M_{Uniform}}\xspace}
 \newcommand{\modelUshaped}{\ensuremath{M_{Ushaped}}\xspace}
@@
 \newcommand{\powerSeriesOld}{\ensuremath{\mathcal{M}_{msBayes}}\xspace}
-\newcommand{\powerSeriesUniform}{\ensuremath{\mathcal{M}_{uniform}}\xspace}
-\newcommand{\powerSeriesUniformExp}{\ensuremath{\mathcal{M}_{uniform-exp}}\xspace}
+\newcommand{\powerSeriesUniform}{\ensuremath{\mathcal{M}_{Uniform}}\xspace}
+\newcommand{\powerSeriesExp}{\ensuremath{\mathcal{M}_{Exp}}\xspace}
 
@@
 \newcommand{\msb}{\upshape\texttt{\MakeLowercase{ms\MakeUppercase{B}ayes}}\xspace}
+\newcommand{\abctoolbox}{\upshape\texttt{ABCtoolbox}\xspace}
+\newcommand{\dppmsbayes}{\upshape\texttt{dpp-msbayes}\xspace}
+\newcommand{\pymsbayes}{\upshape\texttt{PyMsBayes}\xspace}
 \newcommand{\hky}{HKY85\xspace}
@@
 
+\newcommand{\validationAccuracyComparisonCaption}[2]{Accuracy of
+    #1 estimates of #2 when data generated under models (A, E, I, \& M)
+    \modelOld, (B, F, J, \& N) \modelUshaped, (C, G, K, \& O) \modelUniform,
+    and (D, H, L, \& P) \modelDPP are analyzed with models (A--D) \modelOld,
+    (E--H) \modelUshaped, (I--L) \modelUniform, and (M--P) \modelDPP.  A random
+    sample of 5000 posterior estimates (from 50,000) are plotted.  The root
+mean square error (RMSE) calculated from the 5000 estimates is provided.}
+\newcommand{\jitterComment}{Random normal variates ($N(0, 0.005)$) have been
+    added to the estimates and true values of \divTimeNum to reduce overlap of
+    plot symbols.}
+\newcommand{\timeConversionComment}{The divergence times were randomly drawn as
+    indicated above each column of plots, where time is respresented as
+    millions of generations ago (MGA) according to a per-site rate of 1\e{-8}
+    mutations per generation.}
+\newcommand{\validationModelChoiceComparisonCaption}[2]{Model-choice accuracy
+    for models (A--D) \modelOld, (E--H) \modelUshaped, (I--L) \modelUniform,
+    and (M--P) \modelDPP when analyzing data generated under models (A, E, I,
+    \& M) \modelOld, (B, F, J, \& N) \modelUshaped, (C, G, K, \& O)
+    \modelUniform, and (D, H, L, \& P) \modelDPP.
+    The #1 posterior probability of a single divergence event, based on #2,
+    from 50,000 posterior estimates are assigned to bins of width 0.05 and
+    plotted against the proportion of replicates in each bin where the truth is
+    #2.}
 \newcommand{\validationAccuracyCaption}[2]{Estimation accuracy for model
@@
     posterior estimates are shown.}
+\newcommand{\powerAccuracyComparisonCaption}[1]{Accuracy of \divTimeDispersion
+    estimates for models (A--F) \modelOld, (G--L) \modelUshaped, (M--R)
+    \modelUniform, and (S--X) \modelDPP when analyzing data generated under the
+    #1 series of models.
+    The true versus estimated value of the dispersion index of divergence times
+    (\divTimeDispersion, in \globalcoalunit generations) is plotted for 1000
+    datasets simulated under each of the #1 models, and the proportion of
+    estimates less than the truth, $p(\hat{\divTimeDispersion} <
+    \divTimeDispersion)$, is shown for each data model.
+    \timeConversionComment}
+\newcommand{\powerComment}[1]{The power of models
+    (A--F) \modelOld, (G--L) \modelUshaped, (M--R) \modelUniform, and (S--X)
+    \modelDPP to detect random variation in divergence times as simulated under
+    the #1 series of models.}
+\newcommand{\powerSupportComment}[1]{
+    The tendency of models (A--F) \modelOld, (G--L) \modelUshaped, (M--R)
+    \modelUniform, and (S--X) \modelDPP to support one divergence event when
+    there is random variation in divergence times as simulated under the #1
+    series of models.}
+\newcommand{\powerPsiComment}[1]{ 
+    The plots illustrate the estimated number of
+    divergence events ($\hat{\divTimeNum}$) from analyses of 1000 datasets
+    simulated under each of the #1 models, with the the estimated probability
+    of the model inferring one divergence event, $p(\hat{\divTimeNum} = 1)$,
+    given for each combination.}
+\newcommand{\powerProbComment}[2]{ 
+    The plots illustrate histograms of the estimated posterior probability of
+    the one divergence model, #1, from analyses of 1000 datasets simulated
+    under each of the #2 models.}
+\newcommand{\powerDispersionComment}[1]{ 
+    The plots illustrate the estimated dispersion index of divergence times
+    ($\hat{\divTimeDispersion}$) from analyses of 1000 datasets simulated under
+    each of the #1 models, with the the estimated probability of the model
+    inferring one divergence event, $p(\hat{\divTimeDispersion} < 0.01)$, given
+    for each combination.}
 \newcommand{\powerAccuracyCaption}[2]{Estimation accuracy for model

### 2013-12-12 manuscripts/reply/short-reply.tex
Incorporting Jake's edits of reply ms.

@@
     using computer simulations.
-    \citet{Oaks2012} found that the model was very sensitive to the 
-    prior that is chosen and that the inference method had low power
-    to detect variation in divergences times.
+    \citet{Oaks2012} found that the model was very sensitive to the prior and
+    that the inference method had low power to detect variation in divergences
+    times.
     These results were not surprising, in light of a rich statistical literature
-    that shows that the marginal likelihood of a model is quite sensitive to the
+    showing that the marginal likelihood of a model is sensitive to the
     use of vague priors.
@@
     According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
-    was due to excessively broad priors on divergence times leading to 
+    was due to excessively broad priors on divergence times, leading to 
     inadequate numbers of simulation replicates.
-    They proposed a new model-averaging approach that uses narrow, empirically informed uniform
-    priors.
+    They proposed a new model-averaging approach that uses narrow, empirically
+    informed uniform priors.
     Here we demonstrate that the approach of \citet{Hickerson2013} is
@@
     exclude from consideration the true values of the models' parameters.
-    On a more fundamental level, we question the value of adopting an 
-    empirical Bayesian stance for this model-choice problem.
+    On a more fundamental level, we question the value of adopting an empirical
+    Bayesian stance for this model-choice problem, because it can mislead model
+    posterior probabilities, which are inherently measures of belief in the
+    models after prior knowledge is updated by the data.
     The robust Bayesian approach of conducting analyses under a variety
@@
 of the rate at which substitutions occur.
-So the exact divergence time is often viewed as a nuisance parameter, and
-researchers focus on the number of divergence events.
+The exact divergence time is therefore often viewed as a nuisance parameter,
+and researchers focus on the number of divergence events.
 The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
@@
 \citep[e.g.,][]{Jeffreys1935,Lindley1957}.
-Accordingly, \citet{Oaks2012} suggest the primary cause of the spurious support
+Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
 for models with few divergence parameters is the greater marginal likelihoods
@@
 prior density imposed by the uniformly distributed prior on divergence times.
-Note, this is a statistical issue, extrinsic to \msb, and the numerical
-approximation machinery of \msb could be sound.
+Note, this is a statistical issue, extrinsic to \msb, and does not question the
+soundness of the numerical approximation machinery of \msb.
 
@@
 biases found by \citet{Oaks2012}.
-We present results of additional analyses that, when combined with the results
+We present results here of analyses that, when combined with the results
 of \citet{Oaks2012}, strongly suggest that \msb prefers models with few
@@
 Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
-divergences under the informed priors.
+divergences under the empirically informed priors.
 
@@
 We already have data from flipping a different coin 20 times that landed both
-heads and tails 10 times, and so we decide to use these data in specifying a
-beta prior on fairness of the new coin of $beta(a=10, b=10)$
+heads and tails 10 times each, and so we decide to use these data in specifying
+a beta prior on fairness of the new coin of $beta(a=10, b=10)$
 (Figure~\ref{figCoinFlip}).
@@
 We explored this possibility in two ways.
-First, we re-analyzed the Philippines dataset using the model-averaging approach
-of \citet{Hickerson2013}, but set one of prior models with a uniform prior on
-divergence times that is unrealistically narrow, and almost certainly excludes
-most, if not all, of the true divergence times of the 22 taxon pairs.
+First, we re-analyzed the Philippines dataset using the model-averaging
+approach of \citet{Hickerson2013}, but set one of the prior models with a
+uniform prior on divergence times that is unrealistically narrow, and almost
+certainly excludes most, if not all, of the true divergence times of the 22
+taxon pairs.
 If small likelihoods of large models cause the method to prefer models with
@@
 Bayesian model choice, it is clear why one should use caution before overly
-criticizing an investigator's choice of priors after having seen the resulting
-posterior.
+criticizing an investigator's choice of priors \emph{after} having seen the
+resulting posterior.
 %In a strictly Bayesian world, this is cheating.
@@
 true divergence-time values (Figure~\ref{figPowerNumExcluded}).
-Importantly, the posterior probability of excluding at least on true divergence
-value is also quite high across many of the \divt{max}
+Importantly, the posterior probability of excluding at least one true
+divergence value is also quite high across many of the \divt{max}
 (Figure~\ref{figPowerProbExclusion}).
@@
 yet spurious).
-Thus, it is important that investigators not limit their assessment of method's
-power to only differentiating inferences of one event or more (i.e. $\numt{} =
-1$ versus $\numt{} > 1$).
+Thus, it is important that investigators not limit their assessment of the
+method's power to only differentiating inferences of one event or more (i.e.
+$\numt{} = 1$ versus $\numt{} > 1$).
 Rather, looking at the distribution of estimates, as in \citet{Oaks2012},
@@
 support for these models decrease as sampling increases.
-\citet{Oaks2012} do not see such sensitivity when they compare prior sample
+\citet{Oaks2012} did not see such sensitivity when they compared prior sample
 sizes of $2\e6$, $5\e6$, and $10^7$.
@@
 the magnitude of population contraction in each of the descendant
-populations (\bottleScalar{1}{} and \bottleScalar{2}{}) and the
+populations (\bottleScalar{1}{} and \bottleScalar{2}{}), the
 timing of these contractions (\bottleTime{}), and the $N-1$ node heights
 (coalescent times) of the gene tree that gave rise to the $N$ gene
-copies sampled from both populations of the pair.
+copies sampled from both populations.
 % Because we only have a single locus for each taxon, there are no locus-specific

### 2013-12-12 manuscripts/reply/reply-si-main.tex
Working on SI text for reply ms.

@@
 $\meanDescendantTheta{} \sim U(0.0005, 0.04)$.
-As described in \citet{Oaks2012}, the divergence time parameters of the model
+As described in \citet{Oaks2012}, the divergence-time parameters of the model
 implemented in \msb are scaled relative to a constant reference population
@@
 We used the same prior models as above ($M_1$--$M_5$;
-Table~\ref{tabModelChoiceEmpirical}), and generated 50,000 pseudo-replicate
-datasets under this prior (10,000 from each model).
+Table~\ref{tabModelChoiceEmpirical}), and simulated 50,000 datasets under this
+prior (10,000 from each model).
 We used a simulated data structure of eight population pairs, with a single
 1000 base-pair locus sampled from 10 individuals from each population.
-We then analyzed each of these replicate datasets using the same prior,
-retaining 1000 posterior samples.
+We then analyzed each of these replicate datasets using the same prior with 2.5
+million samples (500,000 from each of the five prior models), retaining 1000
+posterior samples.
 Our results are very similar to \citet{Oaks2012}, but we note that they
@@
 U(0.0005, 0.02)$.
-Under this prior, we draw divergence time parameters from a uniform
+Under this prior, we randomly sample divergence-time parameters from a uniform
 distribution of $U(0, 20)$ coalescent units, simulate datasets, and plot the
@@
 Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
-and 10 coalescent units suffered from little affect from saturation.
+and 10 coalescent units suffered little to no effect from saturation.
 Even at divergence times of 20 coalescent units, there is still signal in the
@@
 prior models.
-This prior-predictive approach generating a small number (1000) of random
-samples from the prior and plotting the resulting summary statistics in
+This prior-predictive approach entails generating a small number (1000) of
+random samples from the prior and plotting the resulting summary statistics in
 comparison to the observed statistics to see if they coincide (see Figure 1 of
@@
 \subsection*{Differing utilities of \numt{} and \vmratio{} in \msb}
-The main parameter of the \msb model is the vector of divergence
+The primary component of the \msb model is the vector of divergence
 times for each of the taxon pairs,
@@
 \vmratio{}, is a better model-choice estimator than the number of 
-divergence time parameters within the vector,
+divergence-time parameters within the vector,
 \numt{}.
@@
 about the number of divergences among taxa.
-Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
+Nonetheless, \citet{Hickerson2013} conclude \vmratio{} is more informative and
 biogeographically relevant than \numt{}.
 % We struggle to follow this logic.
-Certainly the maximum information is contained within the divergence time
-vector that \vmratio{} is summarizing.
+However, all of the information about the temporal distribution of divergences
+is contained within the divergence-time vector that \vmratio{} is summarizing.
 Clearly, the number of divergence time parameters within the vector and their
@@
 a sufficient statistic for \divtvector).
-
 \citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
@@
 However, \citet{Oaks2012} demonstrate that even when all assumptions of the
-model are met, \vmratio is a poor model-choice estimator (see plots B, D \& F
-of Figure 4), whereas \numt{} performs better.
-Furthermore, \vmratio{} is limited to estimating the probability of only a
-single model (the one-divergence model), and thus its utility for
-model choice is extremely limited.
+model are met, \vmratio{} is a poor model-choice estimator (see plots B, D \& F
+of Figure 4 in \citet{Oaks2012}), whereas \numt{} performs better.
 
-The model-choice utility of \vmratio{} is limited to the probability that
-this continuous statistic, which can range from zero to infinity, is at its lower
-limit of zero. In theory, this point density will always be zero, thus an
-arbitrary threshold (0.01 is used throughout the \msb literature) must 
-be chosen to make the probability estimable.
-However, it is still not surprising to see that it is numerically difficult to
-obtain reliable estimates of the probability that the continuous \vmratio{}
-statistic is ``near'' its limit of zero.
-It is much easier, less subjective, and more interpretable to estimate
-the probability of the discrete parameter of the model, \numt{}, is
-at its lower limit of one.
+Importantly, \vmratio{} is limited to estimating the probability of only a
+single model (the one-divergence model), and thus its utility for model-choice
+is very limited.
+I.e., it can only be informative about the probability of whether there is one
+divergence shared among the taxa ($\vmratio{} = 0.0$) or there is greater than
+one divergence ($\vmratio{} > 0.0$).
+As a result, not only is its model-choice utility limited, but it is also
+very difficult to estimate.
+\vmratio{} can range from zero to infinity, and the point density that it is
+at its lower limit of zero will always be zero.
+Thus, an arbitrary threshold (0.01 is used throughout the \msb literature) must
+be chosen to make the probability of ``simultaneous'' divergence estimable.
+Even with this arbitrary threshold, it is still not surprising to see that it
+is numerically difficult to obtain reliable estimates of the probability that
+\vmratio{} is ``near'' its lower limit of zero.
+It is easier, less subjective, and more interpretable to estimate the
+probability of the discrete parameter of the model, \numt{}, is at its lower
+limit of one.
 Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better

### 2013-12-12 model-choice/bin/plot-power-results-pretty.py
Adding plotting script for reply power results.

@@ -1,0 +1,425 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes.utils.parsing import DMCSimulationResults, spreadsheet_iter
+from pymsbayes.config import MsBayesConfig
+from pymsbayes.plotting import (PowerPlotGrid, ProbabilityPowerPlotGrid,
+        AccuracyPowerPlotGrid)
+from pymsbayes.utils.messaging import get_logger
+import project_util
+
+_LOG = get_logger(__name__)
+
+class TauExclusionResult(object):
+    def __init__(self):
+        self.num_excluded = []
+        self.num_excluded_glm = []
+        self.prob_of_exclusion = []
+        self.prob_of_exclusion_glm = []
+        self.prior_prob_of_exclusion = []
+        self.bf_of_exclusion = []
+        self.bf_of_exclusion_glm = []
+        approx_prior_exclusion = 0.39184
+        prior_odds = approx_prior_exclusion / (1.0 - approx_prior_exclusion)
+        post_odds = prior_odds * 10
+        self.bf_10_exclusion_prob = post_odds / (1.0 + post_odds)
+        self.prob_of_bf_exclusion = None
+        self.prob_of_bf_exclusion_glm = None
+
+    def calc_prob_of_bf_exclusion(self):
+        self.prob_of_bf_exclusion = (len([
+                1 for x in self.bf_of_exclusion if x > 10.0]) /
+                float(len(self.bf_of_exclusion)))
+        self.prob_of_bf_exclusion_glm = (len([
+                1 for x in self.bf_of_exclusion_glm if x > 10.0]) /
+                float(len(self.bf_of_exclusion_glm)))
+
+class PowerResult(object):
+    def __init__(self):
+        self.true = []
+        self.mode = []
+        self.mode_glm = []
+        self.median = []
+        self.prob = []
+        self.prob_glm = []
+
+    @classmethod
+    def parse_result_summary(cls, result_summary_path):
+        psi = cls()
+        omega = cls()
+        tau = TauExclusionResult()
+        for d in spreadsheet_iter([result_summary_path]):
+            psi.true.append(int(d['psi_true']))
+            psi.mode.append(int(d['psi_mode']))
+            psi.mode_glm.append(int(round(float(d['psi_mode_glm']))))
+            psi.prob.append(float(d['psi_1_prob']))
+            psi.prob_glm.append(float(d['psi_1_prob_glm']))
[…]

### 2013-12-12 power-comparison/bin/plot-power-results-comparison.py
Fixing axis labels in plotting scripts.
Fixing axis labels in plotting scripts for validation and
power-comparison projects. Also, updating plots generated by the
scripts.

@@ -487,7 +487,7 @@ def create_plots(info_path,
                 num_columns = num_columns,
                 share_x = False,
                 share_y = False,
-                title = r'Estimated variance of divergence events, $\hat{D_T}$',
+                title = r'Estimated variance of divergence times, $\hat{D_T}$',
                 title_size = 14.0,
                 title_top = False,
                 y_title = 'Density',
@@ -570,10 +570,10 @@ def create_plots(info_path,
                 num_columns = num_columns,
                 share_x = False,
                 share_y = False,
-                title = r'True variance of divergence events, $D_T$',
+                title = r'True variance of divergence times, $D_T$',
                 title_size = 14.0,
                 title_top = False,
-                y_title = r'Estimated variance of divergence events, $\hat{D_T}$',
+                y_title = r'Estimated variance of divergence times, $\hat{D_T}$',
                 y_title_position = 0.001,
                 y_title_size = 14.0,
                 height = 9.0,

### 2013-12-12 validation/bin/plot-validation-parameter-comparison.py

@@ -44,19 +44,19 @@ def main_cli():
                 r'Estimated number of divergence events, $\hat{|\mathbf{\tau}|}$ (mode)'),
             1: ('omega',
                 r'True variance of divergence times, $D_T$',
-                r'Estimated variance of divergence events, $\hat{D_T}$ (median)'),
+                r'Estimated variance of divergence times, $\hat{D_T}$ (median)'),
             2: ('time',
                 r'True mean of divergence times, $\bar{T}$',
-                r'Estimated mean of divergence events, $\hat{\bar{T}}$ (median)'),
+                r'Estimated mean of divergence times, $\hat{\bar{T}}$ (median)'),
             3: ('psi-glm',
                 r'True number of divergence events, $|\mathbf{\tau}|$',
                 r'GLM-adjusted number of divergence events, $\hat{|\mathbf{\tau}|}$ (mode)'),
             4: ('omega-glm',
                 r'True variance of divergence times, $D_T$',
-                r'GLM-adjusted variance of divergence events, $\hat{D_T}$ (median)'),
+                r'GLM-adjusted variance of divergence times, $\hat{D_T}$ (median)'),
             5: ('time-glm',
                 r'True mean of divergence times, $\bar{T}$',
-                r'GLM-adjusted mean of divergence events, $\hat{\bar{T}}$ (median)'),
+                r'GLM-adjusted mean of divergence times, $\hat{\bar{T}}$ (median)'),
                 }
     for idx, (variable, x_label, y_label) in sub_plot_map.iteritems():
         sub_plots = {}

### 2013-12-13 manuscripts/reply/short-reply.tex
Working on reply ms.
Moving Jake to acknowledgements.

@@
         \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2}$, Jeet Sukumaran$^{3}$, Jacob A.\
-        Esselstyn$^{4}$, Charles W.\ Linkem$^{5}$, Cameron D.\
-        Siler$^{6}$, and Rafe M.\ Brown$^{1}$
+        Jamie R.\ Oaks$^{1,2}$, Jeet Sukumaran$^{3}$, Charles W.\ Linkem$^{4}$,
+        Cameron D.\ Siler$^{5}$, and Rafe M.\ Brown$^{1}$
 
@@
             Durham, North Carolina 27708} \\[.1in]
-        $^4$\emph{Museum of Natural Science,
-            Louisiana State University,
-            119 Foster Hall,
-            Baton Rouge, Louisiana 70803}\\[.1in]
-        $^5$\emph{Department of Biology,
+        $^4$\emph{Department of Biology,
             University of Washington,
             Seattle, Washington 98195}\\[.1in]
-        $^6$\emph{Sam Noble Museum,
+        $^5$\emph{Sam Noble Museum,
             Department of Biology,
@@
 \section*{Acknowledgments}
+We thank Jake Esselstyn for insightful comments that greatly improved this
+work.
 J.\ Oaks thanks the National Science Foundation for supporting this work (DEB

### 2013-12-13 manuscripts/reply/reply-figures.tex
Incorporating new figures into reply ms.
Incorporating updated figures (the same results presented in a
more concise manner). Also, some edits throughout the text.

@@
     A plot of three beta probability density functions that represent a prior
-    (black; $beta(10, 10)$), true posterior (blue; $beta(13, 17)$), and
-    empirical Bayes density (red; $beta(16, 24)$) for a dataset of 10 Bernoulli
-    trials, three of which are successes.
+    (black; $beta(10, 10)$), posterior (blue; $beta(13, 17)$), and empirical
+    Bayes density (red; $beta(16, 24)$) for a dataset of 10 coin flips, three
+    of which are successes.
 }{figCoinFlip}
@@
 \mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/num_tau_excluded.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to
-    exclude the truth.
-    The plots illustrate the number of true \divt{} parameters excluded from analyses of
-    simulated datasets where \divt{} for 22 pairs of populations is drawn from an
-    exponential distribution, $\divt{} \sim Exp(2)$.
+    Histograms of the number of true divergence times excluded from the model
+    preferred by the empirically informed model-averaging approach of
+    \citet{Hickerson2013} when applied to simulated datasets in which divergence
+    times for 22 pairs of populations is drawn from an exponential
+    distribution, $\divt{} \sim Exp(2)$.
     The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
@@
     The proportion of simulation replicates in which at least one true
-    parameter value is excluded from the model preferred by a Bayes factor
-    ($p(\divt{} \notin \hat{M})$) is also given.
+    parameter value is excluded from the preferred model ($p(\divt{} \notin
+    \hat{M})$) is also given.
 }{figExclusionSimTau}
@@
 \mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/prob_of_exclusion.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to
-    exclude the truth.
-    The plots illustrate the estimated probability of excluding at least one
-    true \divt{} value from analyses of simulated datasets where \divt{} for 22
-    pairs of populations is drawn from an exponential distribution, $\divt{}
-    \sim Exp(2)$.
+    Histograms of the support (estimated posterior probabilities) for excluding
+    at least one true divergence time when the empirically informed
+    model-averaging approach of \citet{Hickerson2013} is applied to simulated
+    datasets in which divergence times for 22 pairs of populations is drawn
+    from an exponential distribution, $\divt{} \sim Exp(2)$.
     The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
@@
 
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_psi_mode.pdf}{
-    The bias of the model-averaging approach of \citet{Hickerson2013} to infer
-    clustered divergences in our simulation-based power analyses.
-    The plots illustrate the estimated number of divergence events ($\hat{\numt{}}$)
-    from analyses of simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
-    uniform distributions, $\divt{} \sim U(0, \divt{max})$.
-    The estimated probability of the method inferring one divergence event, $p(\hat{\numt{}} = 1)$,
-    is given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerPsiMode}
+\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-4.pdf}{
+    The tendency of the empirically informed model-averaging approach of
+    \citet{Hickerson2013} to (A--D) infer clustered divergences and (E--H)
+    support the extreme model of one divergence when applied to simulated
+    datasets in which the divergence times are randomly drawn from the uniform
+    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
+    column of plots (divergence-time distributions are given in units of
+    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
+    mutations per generation).
+    Four of the six \divt{max} we simulated are provided; please see
+    Figure~S\ref{figPower6} for a summary of all of the results.
+}{figPower4}
 
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_omega_prob.pdf}{
-    The bias of the model-averaging approach of \citet{Hickerson2013} to
-    support one divergence event in our simulation-based power analyses.
-    The plots illustrate histograms of the estimated posterior probability that
-    the dispersion index of divergence times is less than 0.01 ($p(\vmratio{} <
-    0.01 | \ssSpace)$) from analyses of simulated datasets where \divt{} for 22
-    pairs of populations is drawn from a series of uniform distributions,
-    $\divt{} \sim U(0, \divt{max})$.
-    The proportion of simulation replicates that strongly support one
-    divergence event, $p(BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01} > 10)$,
-    is given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerOmegaProb}
+\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-4.pdf}{
+    Histograms of the (A--D) number of true divergence-time parameters excluded
+    from the preferred model and the (E--H) posterior probability of excluding
+    at least one divergence-time parameter when the empirically informed
+    model-averaging approach of \citet{Hickerson2013} is applied to simlated
+    datasets in which divergence times are randomly drawn from the uniform
+    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
+    column of plots (divergence-time distributions are given in units of
+    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
+    mutations per generation).
+    Four of the six \divt{max} we simulated are provided; please see
+    Figure~S\ref{figPowerExclusion6} for a summary of all of the results.
+}{figPowerExclusion4}
 
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_num_excluded.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to exclude
-    the truth in our simulation-based power analyses.
-    The plots illustrate the number of true \divt{} parameters excluded from analyses of
-    simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
-    uniform distributions, $\divt{} \sim U(0, \divt{max})$. The proportion
-    of simulation replicates in which at least one true parameter value is excluded from
-    the model preferred by a Bayes factor ($p(\divt{} \notin \hat{M})$) is
-    given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerNumExcluded}
-
-\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_prob_exclusion.pdf}{
-    The propensity of the model-averaging approach of \citet{Hickerson2013} to
-    exclude the truth in our simulatoin-based power analyses.  The plots
-    illustrate the estimated posterior probability of excluding at least one
-    true \divt{} value from analyses of simulated datasets where \divt{} for 22
-    pairs of populations is drawn from a series of uniform distributions,
-    $\divt{} \sim U(0, \divt{max})$. The proportion of simulation replicates in
-    which there is strong support for at least one true parameter value being
-    excluded from the model ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is
-    given for each \divt{max}.
-    Each plot represents 500 simulation replicates analyzed using $5\e6$
-    samples from the prior.
-}{figPowerProbExclusion}
-
+%% Sampling error
 \mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
@@
 }{figMCTauPrior}
-
-% \mFigure{../../response-redux/results/hickerson/pymsbayes-results/pymsbayes-output/d1/m12345678-combined/mean_by_dispersion.pdf}{
-%     Results of reanalysis.
-% }{figJointPosterior}

### 2013-12-13 manuscripts/reply/reply-si-figures.tex
Incorporating new figures into reply ms.
Incorporating updated figures (the same results presented in a
more concise manner). Also, some edits throughout the text.

@@
 
-\siFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
-    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
-    estimate the dispersion index of divergence times (\vmratio{}) from
-    analyses of simulated datasets where \divt{} for 22 pairs of populations is
-    drawn from a series of uniform distributions, $\divt{} \sim U(0,
-    \divt{max})$.
-    The proportion of estimates less than the true value of,
-    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
-    Each plot represents unadjusted median estimates from 500 simulation
-    replicates analyzed using $5\e6$ samples from the prior.
-}{figPowerAccOmegaMedian}
+\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-accuracy-6.pdf}{
+    The accuracy of (A--F) unadjusted and (G--L) GLM-adjusted estimates of
+    dispersion index of divergence times (\vmratio{}) when the empirically
+    informed model-averaging approach of \citet{Hickerson2013} is applied to
+    simlated datasets in which divergence times are randomly drawn from the
+    uniform distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top
+    of each column of plots (divergence-time distributions are given in units
+    of millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
+    mutations per generation).
+}{figPowerAccuracy}
 
-\siFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
-    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
-    estimate the dispersion index of divergence times (\vmratio{}) from
-    analyses of simulated datasets where \divt{} for 22 pairs of populations is
-    drawn from a series of uniform distributions, $\divt{} \sim U(0,
-    \divt{max})$.
-    The proportion of estimates less than the true value of,
-    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
-    Each plot represents GLM-regression-adjusted mode estimates from 500
-    simulation replicates analyzed using $5\e6$ samples from the prior.
-}{figPowerAccOmegaModeGLM}
+\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-6.pdf}{
+    The tendency of the empirically informed model-averaging approach of
+    \citet{Hickerson2013} to (A--F) infer clustered divergences and (G--L)
+    support the extreme model of one divergence when applied to simulated
+    datasets in which the divergence times are randomly drawn from the uniform
+    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
+    column of plots (divergence-time distributions are given in units of
+    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
+    mutations per generation).
+}{figPower6}
+
+\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-6.pdf}{
+    Histograms of the (A--F) number of true divergence-time parameters excluded
+    from the preferred model and the (G--L) posterior probability of excluding
+    at least one divergence-time parameter when the empirically informed
+    model-averaging approach of \citet{Hickerson2013} is applied to simlated
+    datasets in which divergence times are randomly drawn from the uniform
+    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
+    column of plots (divergence-time distributions are given in units of
+    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
+    mutations per generation).
+}{figPowerExclusion6}
 
@@
     \citet{Hickerson2013} under the ideal conditions when the prior model
-    is correct (i.e., the pseudo-replicate datasets are simulated from
-    parameters drawn from the same prior distributions used in the
-    analysis).
+    is correct (i.e., the datasets are simulated from parameters drawn from the
+    same prior distributions used in the analysis).
     The plots show the relationship between the estimated posterior and true

### 2013-12-13 manuscripts/reply/short-reply.tex
Incorporating new figures into reply ms.
Incorporating updated figures (the same results presented in a
more concise manner). Also, some edits throughout the text.

@@
 we simulated, whether evaluating the unadjusted
-(Figure~S\ref{figPowerAccOmegaMedian}) or GLM-adjusted
-(Figure~S\ref{figPowerAccOmegaModeGLM}) posterior estimates.
+(Figure~S\ref{figPowerAccuracy}A--F) or GLM-adjusted
+(Figure~S\ref{figPowerAccuracy}G--L) posterior estimates.
 The method only estimates \vmratio{} relatively well when the simulated
 distribution of divergence times is identical to one of the prior models
-(Figures~S\ref{figPowerAccOmegaMedian}E \& S\ref{figPowerAccOmegaModeGLM}E).
+(Figures~S\ref{figPowerAccuracy}E\&K).
 This is consistent with the findings of \citet{Oaks2012}  and
@@
 consistently infers highly clustered divergences across all the \divt{max} we
-simulated (Figure~\ref{figPowerNumExcluded}).
+simulated (Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
 The method is most likely to infer the extreme case of a single divergence
@@
 \numt{} estimates closely mirror the U-shaped prior on divergence models
-used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
+used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
 (\citeyear{Oaks2012}) Figure 5B).
@@
 method strongly supports one divergence in greater than 27\% of the replicates
-across all the \divt{max} we simulated (Figure~\ref{figPowerOmegaProb});
+across all the \divt{max} we simulated (Figure~\ref{figPower4}E--H \&
+S\ref{figPower6}G--L);
 following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
@@
 $2.4\globalpopsize$ generations, and more than 60\% when over the past
-$3.2\globalpopsize$ generations (Figure~\ref{figPowerOmegaProb}).
+$3.2\globalpopsize$ generations (Figure~\ref{figPower4}E--H \&
+S\ref{figPower6}G--L).
 
@@
 of the \divt{max} the preferred model will exclude a large proportion of the
-true divergence-time values (Figure~\ref{figPowerNumExcluded}).
+true divergence-time values (Figure~\ref{figPowerExclusion4}A--D \&
+S\ref{figPowerExclusion6}A--F).
 Importantly, the posterior probability of excluding at least one true
 divergence value is also quite high across many of the \divt{max}
-(Figure~\ref{figPowerProbExclusion}).
+(Figure~\ref{figPowerExclusion4}E--H \& S\ref{figPowerExclusion6}G--L).
 Only when the data are identically distributed as one of the prior models does
 the method avoid excluding the truth more than 5\% of the time
-(Figure~\ref{figPowerNumExcluded}E).
+(Figure~\ref{figPowerExclusion4}C).
 Again, this demonstrates the method's sensitivity to priors.
@@
 \divt{} priors (Table~\ref{tabModelChoiceEmpirical} and
-Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
-and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
-\ref{figPowerOmegaProb}).
+Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerExclusion4})
+and models with fewer \divt{} parameters (Figure~\ref{figPower4}).
 In all of these analyses, each of the prior models have the same number of
@@
 \section*{General thoughts on the model of \msb}
-Our findings are not very surprising in light of difficult inference problem
-\msb seeks to tackle.
-To get a sense of the challenge \msb is faced with, we tally up all the free
+Our findings are not surprising in light of the difficult inference problem
+with which \msb is faced.
+To get a sense of the difficulty of this problem, we tally up all the free
 parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
@@
 information about the parameters from the sequence data.
-The large number of parameters and divergence models relative to summary
-statistics is undoubtedly a key reason the method lacks robustness to prior
-conditions.
+The large number of parameters and divergence models relative to the amount of
+information in the data is undoubtedly a key reason the method lacks robustness
+to prior conditions.
 Robustness is an important characteristic of a method to gauge its
@@
 and in the context of a model-choice method that is highly sensitive to priors,
-it is not too surprising to find the approach struggles.
+it is not too surprising that the approach struggles.
 

### 2013-12-13 manuscripts/reply/reply-main.tex
Updating reply ms.
Incorporating changes made in `short-repy.tex` into `reply.tex` and
`reply-main.tex`. Leaving `short-reply.tex` in repo, because Mark has
been working on this copy.

@@
-Recently, the journal \emph{Evolution} has served as a medium for a discussion
-of potential promises and pitfalls of approximate-Bayesian methods of
-comparative phylogeographical model choice and prior selection.
-The discussion has centered around the model-choice method, \msb
-\citep{Huang2011}, which uses approximate Bayesian computation (ABC) to
-estimate the temporal distribution of divergences among co-distributed pairs of
-taxa by sampling over models of divergence that range from a single
-divergence-time parameter (i.e., all pairs of taxa diverged at the same time)
-to the fully generalized model in which each pair is assigned a divergence-time
-parameter.
+\msb \citep{Huang2011} uses approximate Bayesian computation (ABC) to estimate
+the distribution of divergence times among co-distributed pairs of taxa.
+It approximates a posterior probability over models that range from a single
+divergence-time parameter (i.e., simultaneous divergence of all pairs of taxa)
+to the fully generalized model in which each pair of taxa diverged at a
+unique time.
 A full description of the model is given in \citet{Oaks2012}.
-
-\citet{Oaks2012} published results of simulation-based power analyses showing
-that \msb is biased toward inferring models of temporally clustered divergence
-times across taxon pairs.
-\citet{Hickerson2013} published a response to \citet{Oaks2012} where they
-present an empirical Bayesian model-averaging approach that they conclude
-circumvents the problems revealed by \citet{Oaks2012}.
-These papers are largely in agreement, with \citet{Hickerson2013} reiterating
-much of the discussion of \citet{Oaks2012} regarding the impact of broad
-uniform priors on Bayesian model choice.
-Both papers agree that when broad uniform priors are used to reflect \emph{a
-prior} uncertainty about divergence times of all pairs of taxa under study,
-the method can be biased toward inferring biogeographically interesting
-scenarios of temporally clustered divergences among taxa (i.e. models with few
-divergence-time parameters).
-Both papers also agree that this bias toward models with few divergence-time
-parameters is due to the much greater parameter space of models with more
-divergence-time parameters.
-However, the papers disgree about the primary mechanism by which the greater
-parameter space of parameter-rich models causes the bias toward clustered
-divergence models;
-\citet{Oaks2012} suggest a statistical issue is the cause, whereas
-\citep{Hickerson2013} prefer a numerical explanation.
-% The main areas of disagreement between the papers are centered around
-% (1) the mechanism by which broad uniform priors cause the bias of \msb toward
-% models with few divergence-time parameters, and
-% (2) how to potentially mitigate this issue.
-
-\citet{Oaks2012} suggest the primary cause of the bias of \msb is the low
-marginal likelihoods of parameter-rich models integrated over vast parameter
-space with low probability of producing the data, yet relatively high prior
-density \citep{Lindley1957} imposed by the uniformly distributed prior on
-divergence times.
-Note, this suggests the bias is extrinsic to \msb, and the numerical
-approximation machinery of \msb could be sound.
-\citet{Hickerson2013} suggest the bias is intrinsic to \msb, concluding the
-method's rejection algorithm is inefficient and will be biased
-toward models with fewer divergence-time parameters, because the parameter
-space of these models are more densely sampled relative to models with more
-divergence-time parameters.
+Frequently, the exact values of the divergence times are hard 
+to determine because researchers typically lack precise knowledge 
+of the rate at which substitutions occur.
+The exact divergence time is therefore often viewed as a nuisance parameter,
+and researchers focus on the number of divergence events.
+The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
+on the fundamental methodological point about the model selection performed in \msb:
+\begin{itemize}
+   \item The use of vague priors on the divergence-time parameters can lead to
+       support models with few divergence events shared across taxa. Thus, the
+       primary inference enabled by the approach is very sensitive to the
+       priors on divergence times.
+\end{itemize}
+
+This is not surprising given the rich statistical literature that shows that
+model marginal likelihoods are quite sensitive to the priors used in Bayesian
+model selection
+\citep[e.g.,][]{Jeffreys1935,Lindley1957}.
+Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
+for models with few divergence parameters is the greater marginal likelihoods
+of these models under vague uniform priors, relative to more parameter-rich
+models.
+Models with more divergence-time parameters integrate over \emph{much} greater
+parameter space with low probability of producing the data, yet relatively high
+prior density imposed by the uniformly distributed prior on divergence times.
+Note, this is a statistical issue, extrinsic to \msb, and does not question the
+soundness of the numerical approximation machinery of \msb.
+
+\citet{Hickerson2013} suggest the bias is the result of a numerical issue
+intrinsic to \msb, concluding the method's rejection algorithm is inefficient
+and will be biased toward models with fewer divergence-time parameters, because
+the parameter space of these models are more densely sampled relative to models
+with more divergence-time parameters.
 We agree that sampling error is present in all numerical Bayesian estimation
@@
 biases found by \citet{Oaks2012}.
-We present results of additional analyses that, when combined with
-the results of \citet{Oaks2012}, strongly suggest small marginal likelihoods of
-models with more parameter space is a primary mechanism by which uniform priors
-are causing the bias of \msb.
-
-% We show that this argument is based on empirical Bayesian reasoning and
-% questionable assumptions, and does not explain the bias of the method found in
-% many of the analyses of \citet{Oaks2012}.
-
-There are also differences between the papers regarding potential solutions to
-the bias.
+We present results here of analyses that, when combined with the results
+of \citet{Oaks2012}, strongly suggest that \msb prefers models with few
+divergence-time parameters, because they have greater likelihoods when
+averaged over broad uniform priors on divergences times.
+
 \citet{Oaks2012} suggest alternative prior probability distributions on
 divergence-times and other nuisance parameters could increase the marginal
-likelihoods of models with more divergence-time parameters and thus reduce the
-bias.
+likelihoods of models with more divergence-time parameters and thus reduce
+spurious support for models of temporally clustered divergences.
 Alternatively, \citet{Hickerson2013} present an approach that uses empirically
-informed priors and Bayesian model-averaging in an attempt to accommodate
+informed uniform priors and Bayesian model-averaging in an attempt to accommodate
 uncertainty in selecting priors.
-% \citet{Hickerson2013} champion this strategy as a means of avoiding the
-% pitfalls raised by \citet{Oaks2012}.
 In general, we agree with the use of Bayesian model averaging to obtain a
@@
 that samples over empirically informed uniform priors is a fruitful solution
-for an already biased model-choice method.
-
-In this paper, we discuss the potential theoretical and practical problems of
-using empirically informed priors for Bayesian model choice.
-Furthermore, we evaluate the empirical Bayesian model averaging approach of
-\citet{Hickerson2013} as a potential solution to the biases revealed by
+for a model-choice method that is highly sensitive to priors.
+
+In this paper, we discuss the potential theoretical and practical
+considerations of using empirically informed priors for Bayesian model choice.
+Furthermore, we evaluate the empirical Bayesian model-averaging approach of
+\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
 \citet{Oaks2012}.
 In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
-made an error by mixing different time units, which makes the results presented
-in their response difficult to interpret (see Supporting Information for
-details).
-However, here we show that even when that error is avoided, the empirical
-Bayesian model-averaging approach of \citet{Hickerson2013} remains biased
-toward clustered divergence models, and is predisposed to excluding true
-parameter space.
-The approach provides another means by which the method can ``escape'' models
-with large parameter space, and the method's bias toward smaller models
-remains.
-Importantly, this bias can manifest by preferring narrow, empirically informed
-priors that exclude true parameter space.
-
-% We discuss errors in this approach that render some of their results invalid
-% and leave the remaining results difficult to interpret.
-% Furthermore, we follow the recommendations of \citet{Oaks2012} and present
-% simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
-% approach.
-
-% \cwlnote{}{This part seems important, but not related to the error in units
-% discussed above}
-% Another important characteristic of the approach used by \citet{Hickerson2013}
-% that makes their results difficult to interpret is their use of narrow,
-% empirically guided uniform priors in an approximate-Bayesian model-choice
-% framework.
-% We discuss the risks associated with this approach in the next section, and
-% show how the results of \citet{Hickerson2013} likely exclude true values of the
-% model's parameters.
-
-% In addition to this error, we were not able to reproduce the results of
-% \citet{Hickerson2013}.
-% We followed their methods, generating $5\e6$ prior samples for each of the
-% eight models in their Table 1.
-% We retained the 10,000 samples from each model with the smallest Euclidean
-% distance to the observed summary statistics, and subsequently retained 10,000
-% samples from the remaining 80,000.
+made an error by mixing different units of time, which makes the results
+presented in their response difficult to interpret (see Supporting Information
+for details).
+Here, we avoid this error, but still find their approach to be biased toward
+clustered divergence models.
+Furthermore, the approach provides another means by which the method can
+``escape'' models with large parameter space, which manifests in the preference
+of models that exclude from consideration the true values of the model's
+parameters.
+Our results, when combined with those of \citet{Oaks2012}, suggest that it is
+difficult to choose a uniformly distributed prior on divergence times that is
+broad enough to confidently contain the true values of parameters while being
+narrow enough to avoid spurious support of models with less parameter space.
+However, recent work shows more flexible probability distributions without a
+hard upper bound (e.g., gamma) can accommodate prior uncertainty in divergence
+times without inhibiting the marginal likelihoods of models with more
+divergence-time parameters, and as a result, increase the method's power to
+detect temporal variation in divergences
+\citet{Oaks2014dpp}.
 
@@
 \citet{Hickerson2013} repeatedly refer to the prior distributions used by
-\citet{Oaks2012} as ``poorly selected.''
+\citet{Oaks2012} for divergence-time parameters as ``poorly selected.''
 However, this is misleading, as \citet{Oaks2012} discuss in detail how they
-selected their priors to reflect the large amount of uncertainty about the
-model's parameters for their study system (see section ``Specifying and
-simulating the joint prior'' of \citet{Oaks2012}).
+selected their prior to reflect the large amount of uncertainty about the
+timing of divergences across all 22 of the taxa in their study (see section
+``Specifying and simulating the joint prior'' of \citet{Oaks2012}).
 \citet{Oaks2012} further discuss how the use of uniform prior distributions in
 \msb for many of the model's parameters requires investigators to select broad
-priors to avoid excluding the truth \textit{a priori}, and how these broad
-uniform priors place too much density in improbable regions of parameter space.
-What \citet{Hickerson2013} mean by ``poorly selected'' is that
-\citet{Oaks2012} did not use their data to inform their initial prior
-distributions (i.e., they did not use an empirical Bayesian approach).
-\citet{Oaks2012} did use empirically informed priors when assessing both the
-prior sensitivity of the method and the power of \msb under real-world
-situations when a large amount of prior information is available. 
-In doing so, \citet{Oaks2012} discuss the potential dangers of taking an
-empirical Bayesian approach to model choice (see the last paragraph of
-``Assessing prior sensitivity of \msb'' in \citet{Oaks2012}), and we expand on
-this here by exploring both the theoretical and practical implications of using
-empirically informed priors for Bayesian model choice.
+priors to avoid excluding the truth \textit{a priori}, resulting in too much
+density in improbable regions of parameter space.
+
+\citet{Hickerson2013} suggest that \citet{Oaks2012} should have used their
+data to inform their prior on divergence times (i.e., an empirical Bayesian
+approach).
+However, \citet{Oaks2012} did use empirically informed priors to mimic
+empirical situations in which a large amount of prior information is available;
+they did so to assess the prior sensitivity of the method and to determine if
+uniform priors in such situations are narrow enough to avoid spurious support
+of models with few divergence-time parameters.
+The results of \citet{Oaks2012} and \citet{Hickerson2013} show the method is
+highly sensitive to the prior on divergence times.
+Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
+divergences under the empirically informed priors.
+
+The main argument of \citet{Hickerson2013} is that the priors used by
+\citet{Oaks2012} were not informative enough.
+They suggest a very narrow, highly informed uniform prior on divergence times
+is necessary to avoid the method's preference for models with few
+divergence-time parameters.
+Such an empirical Bayesian approach to model selection raises some theoretical
+and practical concerns, some of which were discussed by \citet{Oaks2012} (see
+the last paragraph of ``Assessing prior sensitivity of \msb'' in
+\citet{Oaks2012}); we expand on this here.
 
@@
 We already have data from flipping a different coin 20 times that landed both
-heads and tails 10 times, and so we decide to use these data in specifying a
-beta prior on fairness of the new coin of $beta(a=10, b=10)$
+heads and tails 10 times each, and so we decide to use these data in specifying
+a beta prior on fairness of the new coin of $beta(a=10, b=10)$
 (Figure~\ref{figCoinFlip}).
-We collect data by flipping the coin $N=10$ times, $y=3$ of which land heads.
+We collect data by flipping the coin of interest $N=10$ times, $y=3$ of which
+land heads.
 Given the beta distribution is a conjugate prior for a binomial likelihood, the
@@
     Bayesian model choice}
-% \cwlnote{}{Should this be part of the empirical Bayes section, or something
-%     else?}
-
-% \cwlnote{}{It seems like this section can be boiled down to ``empirical Bayes
-%     approaches for model choice when using uniform priors can exclude the
-%     $truth$ from the prior, which is a violation of Bayes theorem" or something
-%     like that. Picking apart what they calculated and we calculated seems like
-%     too much detail. I would try to simplify how this is being presented. This
-%     can lead into how the use of more appropriate prior distributions should be
-%     used.}
 In addition to the theoretical concerns discussed above, there are practical
@@
 see Table~6 of \citet{Oaks2012}).
-Thus, the results of \citet{Hickerson2013} indicate that integration over
-parameter space with low likelihood is biasing the method toward models with
-smaller parameter space, and as a consequence is biasing the method toward
-models that exclude the truth (i.e., toward estimating model-averaged
-posteriors dominated by models that exclude true values of parameters of the
-model).
-
-%CWL:This should be in the section on model-averaging, not empirical Bayes
-%limitations
-We explored the propensity of the empirically informed model-averaging approach
-of \citet{Hickerson2013} to exclude the truth in two ways.
-First, we re-analyzed the Philippines dataset using the model-averaging approach
-of \citet{Hickerson2013}, but set one of prior models with a uniform prior on
-divergence times that is unrealistically narrow, and almost certainly excludes
-most, if not all, of the true divergence times of the 22 taxon pairs.
+
+The strong preference for the questionable priors on divergence times (their
+$M_1$ and $M_2$) in the model-averaged results of \citet{Hickerson2013} suggest
+their approach is biased toward models with less parameter space, and as a
+consequence, is biased toward estimating model-averaged posteriors dominated by
+models that exclude true values of parameters of the model.
+We explored this possibility in two ways.
+First, we re-analyzed the Philippines dataset using the model-averaging
+approach of \citet{Hickerson2013}, but set one of the prior models with a
+uniform prior on divergence times that is unrealistically narrow, and almost
+certainly excludes most, if not all, of the true divergence times of the 22
+taxon pairs.
 If small likelihoods of large models cause the method to prefer models with
 less parameter space, we expect \msb will preferentially sample from this
-incorrect model yielding a marginal posterior that is incorrect (i.e., the
+incorrect model yielding a posterior that is incorrect (i.e., the
 model-averaged posterior will be dominated by an incorrect model that excludes
@@
 
-For our re-analysis of the Philippines dataset we used the model-averaging
+\subsubsection*{Re-analyses of the Philippines dataset using empirical Bayesian
+model averaging}
+
+For our re-analyses of the Philippines dataset we followed the model-averaging
 approach of \citet{Hickerson2013}, but with a reduced set of prior models to
-avoid mixing time units.
+avoid their error of mixing units of time (see SI for details).
 We used five prior models, all of which had priors on population sizes of
@@
 Our results show that the model-averaging approach of \citet{Hickerson2013}
-does not reduce the method's bias toward models with less parameter space,
-but rather allows it to manifest in a rather dangerous way.
-The method strongly prefers the prior model with the narrowest distribution on
-divergence times across all three of our analyses, even when this model is
-almost certainly incorrect and excludes the true divergence times of the
-Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
-Unfortunately, ``checking'' the priors by plotting the summary statistics from
-1000 random samples of each prior model along the first two orthogonal axes of
-a principle component analysis, as recommended by \citet{Hickerson2013},
-provides no warning of a problem (Figure~\ref{figPCA}).
+does not reduce the method's support of models with less parameter space.
+Rather, the method strongly prefers the prior model with the narrowest
+distribution on divergence times across all three of our analyses, even when
+this model is almost certainly incorrect and excludes the true divergence times
+of the Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
+
+However, \citet{Hickerson2013} vetted the priors used in their model-averaging
+approach via ``graphical checks,'' in which the summary statistics from 1000
+random samples of each prior model are plotted along the first two orthogonal
+axes of a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
+To determine if such prior predictive analyses would indicate the $M_{1A}$ and
+$M_{1B}$ models are problematic, we performed these graphical checks on our
+prior models.
+Unfortunately, these prior predictive checks provide no warning that these
+priors are too narrow (Figure~S\ref{figPCA}).
+Rather, the graphs suggest these incorrect priors are ``better fit''
+(Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
+\citet{Oaks2012} (Figure~S\ref{figPCA}D--F).
+
 Given that the results of \citet{Hickerson2013} strongly prefer the models with
@@
 model-averaged results are dominated by models that exclude at least some of
-the 22 true divergence times, making their results difficult to interpret
-(in addition to the error of time units).
+the 22 true divergence times, making their results difficult to interpret.
+
+\subsubsection*{Simulation-based assessment of Hickerson et al.'s
+\citeyear{Hickerson2013} model averaging over empirical priors}
 
@@
 (\citeyear{Hickerson2013}) approach to exclude the truth, we simulated 1000
-datasets in which the divergence times for the 22 population pairs are drawn
+datasets in which the divergence times for the 22-population pairs are drawn
 randomly from an exponential distribution with a mean of 0.5 ($\divt{} \sim
@@
 excluding the true value of at least one parameter.
-Whenever the Bayes factor preferred a model excluding the truth we counted
-the number of 22 true divergence times that were excluded by the preferred
+Whenever the Bayes factor preferred a model excluding the truth we counted the
+number of the 22 true divergence times that were excluded by the preferred
 model.
 Our results show that the model-averaging approach of \citet{Hickerson2013}
-excludes the true values of parameters in 97\% of the replicates (90\% with
-GLM-regression adjustment), excluding up to 21 of the 22 true divergence times
-(Figure~\ref{figExclusionSimTau}).
+favors a model that excludes the true values of parameters in 97\% of the
+replicates (90\% with GLM-regression adjustment), excluding up to 21 of the 22
+true divergence times (Figure~\ref{figExclusionSimTau}).
 We also used the mode estimate of the preferred model for each replicate to
-estimate the number of true values excluded, which produced very similar
-results.
-Furthermore, the posterior probability of excluding at least one true parameter
+estimate the number of true parameter values excluded, which produced very
+similar results.
+Importantly, the posterior probability of excluding at least one true parameter
 value is very high in nearly all of the replicates
@@
 
-The results of our empirical and simulation-based analyses clearly demonstrate
-the danger of using narrow, empirically guided uniform priors in a Bayesian
-model-averaging framework.
-% This is especially concerning in \msb, which has already been demonstrated to
-% be biased away from models with more parameter space \citep{Oaks2012}.
-The consequence of this approach is a high risk of obtaining a model-averaged
-posterior estimate that is heavily weighted toward incorrect models that
-exclude true values of model parameters.
+The results of the preceding empirical and simulation-based analyses clearly
+demonstrate the risk of using narrow, empirically guided uniform priors in a
+Bayesian model-averaging framework.
+The consequence of this approach is obtaining a model-averaged posterior
+estimate that is heavily weighted toward incorrect models that exclude true
+values of model parameters.
 
 \subsection*{Additional thoughts on empirical priors in Bayesian model choice}
-Given all of the theoretical and practical issues with empirical Bayes
-approaches to Bayesian model choice discussed in the proceeding sections, it is
-quite clear why one should use caution before overly criticizing an
-investigator's choice of priors after having seen the resulting posterior.
-%In a strictly Bayesian world, this is cheating.
-As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a large
-amount of uncertainty regarding the divergence times of the 22 population pairs
-under study.
+The preceding sections are not a general critique of Bayesian model averaging.
+Rather, model averaging can provide an elegant way of incorporating
+model uncertainty in Bayesian inference.
+However, when averaging over models with narrow and broad uniform priors on a
+parameter with a likelihood density that is not expected to be uniformly
+distributed, the posterior can be dominated by models that exclude from
+consideration the true values of parameters due to the larger marginal
+likelihoods of the models that are integrated over less space with high prior
+weight yet low likelihood.
+
+Given the theoretical and practical issues with empirical Bayes approaches to
+Bayesian model choice, it is clear why one should use caution before overly
+criticizing an investigator's choice of priors \emph{after} having seen the
+resulting posterior.
+As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a
+large amount of uncertainty regarding the divergence times of the 22 population
+pairs under study.
 Two of these pairs represent distinct species, and the taxonomy of many groups
@@
 \citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
-    Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013}.
-\citet{Oaks2012} also discuss how the sole use of uniform priors in \msb makes
-it very difficult for investigators to express their prior uncertainty without
-putting high prior density in regions of improbable parameter space.
-The alternative, as shown above, is to risk excluding the truth before
-performing the analysis.
-
+Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013,Rafe2013AREES,Siler2014kikuchii}.
+When limited to uniformly distributed priors, the alternative to priors that
+reflect prior uncertainty, as shown above, is to risk excluding the true values
+one seeks to estimate.
+However continuous distributions perhaps more appropriate as priors for
+positive real-valued parameters have been shown to greatly reduce spurious
+support for clustered divergence models while allowing prior uncertainty to be
+accommodated \citep{Oaks2014dpp}.
+
+\section*{Assessing the power of the model-averaging approach of
+    \citet{Hickerson2013}}
 While our results above clearly demonstrate the risks inherent to the empirical
@@
 such risk if the method does indeed increase the power of the method and thus
-decrease bias toward clustered divergence models, as claimed by
-\citet{Hickerson2013}.
-In the next section, we use simulation-based analyses to evaluate the power of
-Hickerson et al.'s (\citeyear{Hickerson2013}) approach, which show the method is
-still biased toward highly clustered models of divergence.
-
-\section*{Assessing the power of the model-averaging approach of
-    \citet{Hickerson2013}}
-% \cwlnote{}{I think the empirical use of the model-averaging should go here,
-%     instead of above in the empirical Bayes section}
-
-As recommended by \citet{Oaks2012}, we perform simulation-based analyses
-to explore the power of the approximate Bayesian model-averaging approach
-proposed by \citet{Hickerson2013}.
+decrease bias toward clustered divergence models.
+We assess this possibility using simulation-based analyses.
 Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
@@
 We find that the approach of \citet{Hickerson2013} struggles to estimate the
-dispersion index of divergence times (\vmratio{}) across most of the \divt{max}
+variance of divergence times (\vmratio{}) across most of the \divt{max}
 we simulated, whether evaluating the unadjusted
-(Figure~\ref{figPowerAccOmegaMedian}) or GLM-adjusted
-(Figure~\ref{figPowerAccOmegaModeGLM}) posterior estimates.
+(Figure~S\ref{figPowerAccuracy}A--F) or GLM-adjusted
+(Figure~S\ref{figPowerAccuracy}G--L) posterior estimates.
 The method only estimates \vmratio{} relatively well when the simulated
 distribution of divergence times is identical to one of the prior models
-(Figures~\ref{figPowerAccOmegaMedian}E \& \ref{figPowerAccOmegaModeGLM}E).
-This is consistent with the conclusion of \citet{Oaks2012} that \msb lacks
-robustness and is highly sensitive to the prior distribution deviating from the
-true, underlying distribution of the data.
+(Figures~S\ref{figPowerAccuracy}E\&K).
+This is consistent with the findings of \citet{Oaks2012}  and
+\citet{Hickerson2013} that \msb is highly sensitive to the prior distribution
+deviating from the true, underlying distribution of the data.
 
@@
 consistently infers highly clustered divergences across all the \divt{max} we
-simulated (Figure~\ref{figPowerNumExcluded}).
-The method is most likely to infer the extreme case of a single divergence event
-when populations diverged randomly over the past \globalcoalunit generations.
-Even when divergences are random over the past $8\globalpopsize$ generations,
-the most likely inference is only two divergence events, and a single 
+simulated (Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
+The method is most likely to infer the extreme case of a single divergence
+event shared across 22 taxa when populations diverged randomly over the past
+\globalcoalunit generations.
+When divergences are random over the past $8\globalpopsize$ generations, the
+most likely inference is only two divergence events, and a single
 divergence is still estimated in more than 10\% of the replicates.
-It is very interesting to note that as \divt{max} increases, but before the
+It is interesting to note that as \divt{max} increases, but before the
 estimates are finally pulled away from $\numt{} = 1$, the distribution of
-\numt{} estimates closely mirror the odd U-shaped prior on divergence models
-used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
+\numt{} estimates closely mirror the U-shaped prior on divergence models
+used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
 (\citeyear{Oaks2012}) Figure 5B).
-This supports the conclusions of \citet{Oaks2012} that this U-shaped
-prior coupled with the poor marginal likelihoods of models with many
-\divt{} parameters, is a major cause of the method's bias toward
-clustered divergence models.
+This suggests this U-shaped prior coupled with the small marginal likelihoods
+of models with many divergence parameters, is a major cause of the method's
+bias toward clustered divergence models; this is consistent with
+\citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.
 
@@
 method strongly supports one divergence in greater than 27\% of the replicates
-across all the \divt{max} we simulated (Figure~\ref{figPowerOmegaProb}).
-Following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
+across all the \divt{max} we simulated (Figure~\ref{figPower4}E--H \&
+S\ref{figPower6}G--L);
+following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
 the criterion for incorrect inference of a single divergence event.
-There is strong support for a single divergence event in more than 90\% of the
-replicates when divergences are random over the past $2.4\globalpopsize$
-generations, and more than 60\% when over the past $3.2\globalpopsize$
-generations or less (Figure~\ref{figPowerOmegaProb}).
-
-Contrary to Hickerson et al.'s (\citeyear{Hickerson2013}) claim that their
-model-averaging approach ``can discriminate complex multispecies histories and
-correctly reject synchronous divergence, even when discrete divergence times
-differ by much less than \ldots millions of generations,'' our results show
-that method is biased toward supporting the extreme case of a single
-divergence event when populations diverged randomly over the last
-$8\globalpopsize$ generations.
+Furthermore, there is strong support for a single divergence event in
+more than 90\% of the replicates when divergences are random over the past
+$2.4\globalpopsize$ generations, and more than 60\% when over the past
+$3.2\globalpopsize$ generations (Figure~\ref{figPower4}E--H \&
+S\ref{figPower6}G--L).
+
+Our results show that the empirical Bayesian model-averaging approach leads to
+spuriously strong support for the extreme case of a single divergence event
+when populations diverged randomly over the last $8\globalpopsize$ generations.
 To put this on the scale roughly consistent with a vertebrate mitochondrial
@@
 million generations.
-
-The results of our power analyses further demonstrate the propensity of
+Given the empirically Bayesian model-averaging approach of \citet{Hickerson2013}
+lacks power to detect random variation in divergence times over such
+timescales, it is difficult to justify the risk of the approach to exclude
+regions of parameter space containing the truth.
+
+Also, the results of our power analyses further demonstrate the propensity of
 Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
 parameter values.
-Across all but one of the \divt{max} we simulated, the method excludes the
-truth in a large proportion of replicates, and across many of the \divt{max} it
-will exclude a large proportion of the true divergence time values
-(Figure~\ref{figPowerNumExcluded}).
-The posterior probability of excluding at least on true divergence value is
-also quite high across many of the \divt{max}
-(Figure~\ref{figPowerProbExclusion}).
+Across all but one of the \divt{max} we simulated, in a large proportion of
+replicates the method favors a model that excludes the truth, and across many
+of the \divt{max} the preferred model will exclude a large proportion of the
+true divergence-time values (Figure~\ref{figPowerExclusion4}A--D \&
+S\ref{figPowerExclusion6}A--F).
+Importantly, the posterior probability of excluding at least one true
+divergence value is also quite high across many of the \divt{max}
+(Figure~\ref{figPowerExclusion4}E--H \& S\ref{figPowerExclusion6}G--L).
 Only when the data are identically distributed as one of the prior models does
 the method avoid excluding the truth more than 5\% of the time
-(Figure~\ref{figPowerNumExcluded}E).
-Again, this demonstrates the method's lack of robustness.
-
-Given our results, we conclude that the approach of \citet{Hickerson2013}
-excludes regions of parameter space containing the truth, and lacks power to
-detect random variation in divergence times over a scale of $8\globalpopsize$
-generations or more.
-This roughly translates to millions of generations for mitochondrial loci, and
-hundreds of millions of generations for nuclear loci.
-
-%CWL: I would suggest removing this section
-\section*{The power analysis of \citet{Hickerson2013}}
-\citet{Hickerson2013} present a power analysis in which they find that
-\msb has the power to infer multiple divergence events when divergence times
-are random over hundreds of thousands of generations (rather than millions as
-demonstrated by \citet{Oaks2012}).
-It is not surprising that under limited parameter space the method has
-increased power to detect temporal variation in divergences over narrower time
-windows than millions of generations.
-However, what is important to consider is whether those conditions are relevant
-to real-world applications of the method.
-\citet{Oaks2012} explore the behavior of the method under three divergence time
-priors, as narrow as 0--5 coalescent units. This is quite narrow considering
+(Figure~\ref{figPowerExclusion4}C).
+Again, this demonstrates the method's sensitivity to priors.
+
+\section*{The importance of power analysis to guide applications of \msb}
+\citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
+divergence-time prior of 0--1 coalescent units ago.
+They found that under these prior conditions \msb can, assuming a
+per-site rate of $1.92\e{-8}$ mutations per generation, detect multiple
+divergence events among 18 taxa when the true divergences were random over
+hundreds of thousands of generations or more.
+\citet{Oaks2012} performed similar power analyses under three uniform
+divergence-time priors as narrow as 0--5 coalescent units, and found the
+method was able to detect multiple events among 22 when divergences were random
+over millions of generations.
+As recommended by \citet{Oaks2012}, it is important that investigators perform
+power analyses to determine the method's power for their dataset, and decide if
+\msb has sufficient temporal resolution to address their hypotheses; in the
+case of the Philippines dataset, it did not.
+When doing so, it is important to consider what prior conditions are
+relevant to their empirical system.
+A divergence-time prior of 0--5 coalescent units is quite narrow considering
 that this expresses the prior belief that all 22 taxon pairs diverged within
 this window.
-\citet{Hickerson2013} limit their power analysis to a single prior of 0--1
-coalescent unit.
-This prior setting assumes that there is enough a priori information to
-be 100\% certain that all taxa (18 for their simulations) diverged within
-the last 4$\globalpopsize$ generations.
-This does not seem applicable to most empirical systems.
-Furthermore, even when such extensive prior information is available, to only
-be able to detect multiple divergences on the scale of hundreds of thousands of
-generations does not seem very powerful.
-
-\citet{Hickerson2013} present information only on the extreme case of a
-single divergence event.
-As discussed in \citet{Oaks2012}, it seems odd to consider the estimation
-of two divergence events as a success when divergence is random over hundreds
-of thousands (or millions) of generations.
-In an empirical system such as the Philippines, where island fragmentation has
-occurred at least 10 times over the past several millions of years
-\citep{Haq1987,Rohling1998,Siddall2003,Miller2005}, an estimate where 22 taxa
-share even a handful of divergence events would be of biogeographic interest.
-
-Lastly, \citet{Hickerson2013} translate their results from units of
-\globalcoalunit generations to generations assuming a mutation rate of
-$1.92\e{-8}$ mutations per site per generation.
-If we translate their results to a scale more consistent with rates of mutation
-of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
-by \citet{Hickerson2013} the method can only reliably reject the extreme case
-of one divergence event when divergences are random over a period of more than
-3 million generations.
-Again, considering \citet{Hickerson2013} are assuming prior knowledge that
-all population pairs diverged within the last coalescent unit, the
-method has low power to detect random variation in divergence times.
-
-\section*{Model likelihoods or insufficient prior sampling as the cause of bias
-in \msb}
-One of the points of disagreement between \citet{Hickerson2013} and
-\citet{Oaks2012} is the underlying mechanism causing the bias toward models
-with clustered divergences (i.e., models with few divergence events).
-\citet{Oaks2012} present simulation-based analyses that suggest the broad
-uniform prior on divergence time parameters decrease the marginal likelihoods
-of models with more divergence time parameters, whereas \citet{Hickerson2013}
-argue that the bias is inherent to the inefficient rejection algorithm
-implemented in \msb.
-
-\citet{Hickerson2013} present an interesting probabilistic argument
- to show that insufficient prior sampling is to blame for the bias.
-They argue the widest of the three priors used by \citet{Oaks2012} was very
-unlikely to produce any prior samples with large numbers of divergence times
-that were consistent with the Philippine data.
-There are a few issues with their argument.
-First, the probabilities they present assume the gene divergence times
-estimated by \citet{Oaks2012} are correct.
-The sole purpose of the estimates of ultrametric gene trees presented by
-\citet{Oaks2012} was to provide a very rough comparison of the gene divergence
-times across the 22 taxa.  These analyses assumed an arbitrary strict clock of
-$2\e{-8}$ for all taxa, and are, of course, subject to estimation error.
+Certainly, it is  rare for there to be enough \emph{a priori} information to
+be certain that all taxa diverged within the last 4$\globalpopsize$ generations
+(i.e., 0--1 coalescent units).
+Also, it seems unlikely that when such prior information is available that
+being able to detect multiple divergences on the scale of hundreds of thousands
+of generations will add much insight about the evolutionary history of the
+taxa; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$),
+this translates to a temporal resolution of 3 million generations or more.
+
+As noted by \citet{Oaks2012}, inferring more than one divergence time
+shared across all taxa does not confirm the method is working well when
+analyzing data generated under random temporal variation in divergences (e.g.,
+an inference of two divergence events could be biogeographically interesting
+yet spurious).
+Thus, it is important that investigators not limit their assessment of the
+method's power to only differentiating inferences of one event or more (i.e.
+$\numt{} = 1$ versus $\numt{} > 1$).
+Rather, looking at the distribution of estimates, as in \citet{Oaks2012},
+provides much more information about the behavior of the method.
+
+\section*{The causes of support for models of co-divergence}
+To determine how best to improve the behavior of \msb, it is important to
+determine the mechanism by which broad uniform priors cause spurious support
+for clustered models of divergence.
+It is well established that vague priors can be problematic in Bayesian model
+selection.
+Models that integrate over more parameter space characterized by low
+probability of producing the data and relatively high prior density will have
+smaller marginal likelihoods \citep{Jeffreys1935,Lindley1957}.
+Given the uniformly distributed priors on divergence times (and other nuisance
+parameters) employed in \msb, models with more divergence parameters will be
+forced to integrate over \emph{much} greater parameter space, all with equal
+prior density, and much of it with low likelihood.
+In light of this fundamental statistical issue, it is not surprising that the
+method tends to support simple models.
+
+However, \citet{Hickerson2013} conclude that the bias is due to insufficient
+sampling.
+They argue the widest of the three priors used by \citet{Oaks2012} would
+infrequently produce samples with many independent population divergence times
+as recent as the estimated gene divergence times presented in \citet{Oaks2012}.
+However, this argument assumes the gene divergence times presented in
+\citet{Oaks2012} were estimated without error.
+These estimates were intended to only provide a rough comparison of the
+gene divergence times across the 22 taxa.
+These analyses assumed an arbitrary strict mutation rate of $2\e{-8}$ for all
+taxa, and are, of course, subject to estimation error.
 Furthermore, the branch-length units of the gene trees are in millions of
-years, whereas the divergence time prior of \msb is in generations, thus the
-logic of \citet{Hickerson2013} requires the additional assumption that all 22
-Philippine taxa have a generation time of one year.
-Thus, the argument of \citet{Hickerson2013} that divergence time estimates of
+years, whereas the divergence time prior of \msb is in generations, thus
+\citet{Hickerson2013} make the implicit assumption that all 22 Philippine taxa
+have a generation time of one year.
+The argument of \citet{Hickerson2013} that divergence time estimates of
 \citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
-questionable, especially given our findings on the behavior of empirically
-informed priors presented above.
-
-Finally, even \emph{if} we make all of these assumptions and assume the gene
-divergence times are estimated without error, the probabilistic argument only
-applies to one of the three different priors used by \citet{Oaks2012}.
-The narrowest prior on divergence times used by \citet{Oaks2012} closely mirrors
-the range of estimates of gene-divergence times, and applying Hickerson et al.'s
-(\citeyear{Hickerson2013}) probability equations demonstrates that the prior
-is densely populated with samples with large numbers of divergence parameters
-that are consistent with the gene divergence estimates.
-Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
-sampling is to blame for the bias, it should be much reduced under the narrow
-prior on \divt{}.
+questionable, especially given our findings presented above regarding the
+behavior of \msb when empirically informed uniform priors are employed.
+
+Even if we assume the gene divergence times are estimated without error,
+Hickerson et al.'s \citeyear{Hickerson2013} argument only applies to one of the
+three different priors used by \citet{Oaks2012}.
+The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5
+mya)$) closely mirrors the range of estimates of gene-divergence times.
+Applying Hickerson et al.'s (\citeyear{Hickerson2013}) sampling-probability
+argument demonstrates this prior is densely populated with samples with large
+numbers of divergence parameters with values younger than the estimated gene
+divergence estimates.
+Thus, if insufficient prior sampling is to blame for the bias, it should be
+much reduced under the narrow prior on \divt{}.
 However, the magnitude of the bias is very similar across all three priors
@@
 To disentangle whether model likelihoods or insufficient prior sampling is to
-blame for the biases revealed by \citet{Oaks2012}, we must look at the
+blame for the method's spurious support for simple models, we must look at the
 different predictions made by these two phenomena.
 One example, as discussed by \citet{Oaks2012}, is that insufficient prior
-sampling should create higher variance in posterior estimates, and thus it
-should cause analyses to be sensitive to the number of samples drawn from the
-prior.
-\citet{Oaks2012} do not see such sensitivity when they compare prior sample
+sampling should create large variance among posterior estimates, and thus it
+should cause analyses to be highly sensitive to the number of samples drawn
+from the prior.
+Furthermore, if sampling error is biased toward models with less
+parameter space, as suggested by \citet{Hickerson2013}, we expect to see
+support for these models decrease as sampling increases.
+\citet{Oaks2012} did not see such sensitivity when they compared prior sample
 sizes of $2\e6$, $5\e6$, and $10^7$.
@@
 To explore this prediction further, we repeat the analysis of the Philippines
-dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim U(0,
-10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
-(0.0005, 0.02)$), using a very large prior sample size of $10^8$.
+dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim
+U(0, 10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{}
+\sim (0.0005, 0.02)$), using a very large prior sample size of $10^8$.
 When we look at the trace of the estimates of the dispersion index of
 divergence times (\vmratio{}) as the prior samples accumulate
-(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or
-GLM-regression-adjusted estimates.
-This suggests that insufficient prior sampling did not play a large
-role in the bias found by \citet{Oaks2012}.
-
-Our results presented above that demonstrate the bias of the model-averaging
-approach of \citet{Hickerson2013} both toward models with narrower \divt{}
-priors (Table~\ref{tabModelChoiceEmpirical} and
-Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
-and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
-\ref{figPowerOmegaProb}) strongly suggest the primary cause is the uniform
-priors reducing the marginal likelihoods of large models.
-In all of our model-averaging analyses, all prior models have the same number
-of samples.
-% Thus, it is difficult to explain the method's propensity toward the model with
-% the smallest parameter space via insufficient prior sampling.
-While analyses that sample each model proportional to their relative parameter
-space could be explored, it seems much more likely that the broad uniform
-priors are simply inhibiting the marginal likelihoods of spacious models.
-
-As discussed by \citet{Oaks2012}, a straightforward prediction of reduced
-likelihoods of complex models is that they should disappear as the model
-generating the data converges to the prior.
-\citet{Oaks2012} test this prediction by performing 100,000 simulations to
+(Figure~\ref{figSamplingError}) we do not see the trend predicted by
+\citet{Hickerson2013} in either the unadjusted or GLM-regression-adjusted
+estimates.
+While sampling error is always a reality, it does not appear to be playing a
+large role in the results of \citet{Oaks2012} or presented above.
+
+As discussed by \citet{Oaks2012}, a straightforward prediction if marginal
+likelihoods are causing the preference for simple models is that the bias should
+disappear as the model generating the data converges to the prior.
+\citet{Oaks2012} tested this prediction by performing 100,000 simulations to
 assess the model-choice behavior of \msb when the prior model is correct.
-They find that the bias actually switches direction (at least for the
-regression-adjusted estimates) and the method tends to underestimate the
-probability of the model with one divergence event (see Figure 4 of
+The results confirm the prediction as \msb actually tends to underestimate
+the probability of the one-divergence model (see Figure 4 of
 \citet{Oaks2012}).
-The same prediction is not as straightforward for the insufficient-sampling
-hypothesis.
+We confirmed this same behavior for the model-averaging approach used by
+\citet{Hickerson2013} (see SI text and Figure~S\ref{figValidationMCBehavior}).
+These results are not clearly predicted if insufficient sampling was causing
+the bias.
 Even when the prior is correct, due to the discrete uniform prior on
@@
 Thus, the results of the simulations of \citet{Oaks2012} are more consistent
-with marginal model likelihoods causing the bias toward models with fewer
-divergence time parameters.
-
-Taken together, the results presented here and in \citet{Oaks2012} support
-reduced marginal likelihoods of large models as a primary mechanism by which
-broad uniform priors cause biases in \msb.
-Nonetheless, posterior sampling error will always be present in any numerical
-Bayesian approximation method.
-Thus, insufficient sampling of the prior will contribute to the error of all
-approximate Bayesian estimates.
-However, there is no strong evidence that it is playing a large role in
-the biases revealed herein and by \citet{Oaks2012}.
-
-\section*{Some general thoughts on the model of \msb}
-Our results demonstrating the poor behavior of the model-averaging approach
-proposed by \citet{Hickerson2013} are not too surprising.
-While this approach is trivial to implement, it is not a trivial change
-to the \msb model.
-To better understand the reason for this, it might help to step back and get a
-sense of the scale of the \msb model.
-To do this, we will use the dataset of 22 vertebrate taxon pairs of
-\citet{Oaks2012} as an example.
-
-Following the model description and notation of \citet{Oaks2012}, let us tally
-up all of the free parameters in the \msb model.  Under the simplest model in
-\msb (i.e., assuming no migration and
-no intra-locus recombination), the number of parameters for each
-taxon pair include:
+with the fundamental sensitivity of marginal likelihoods to priors.
+
+This is further demonstrated by the results presented herein that show the
+model-averaging approach of \citet{Hickerson2013} prefers models with narrower
+\divt{} priors (Table~\ref{tabModelChoiceEmpirical} and
+Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerExclusion4})
+and models with fewer \divt{} parameters (Figure~\ref{figPower4}).
+In all of these analyses, each of the prior models have the same number of
+samples.
+Thus, while sampling error will always be present in any numerical Bayesian
+approximation method, insufficient sampling is an unlikely explanation for
+support of prior models with less parameter space.
+While analyses that sample each model proportional to their parameter space
+could be explored, it is clear that the marginal likelihoods under broad
+uniform priors on divergence times will be greater for models with fewer
+divergence-time dimensions.
+
+\section*{General thoughts on the model of \msb}
+Our findings are not surprising in light of the difficult inference problem
+with which \msb is faced.
+To get a sense of the difficulty of this problem, we tally up all the free
+parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
+Under the simplest model in \msb (i.e., assuming no migration and no
+intra-locus recombination), the number of parameters for each taxon pair
+include:
 The population sizes of the ancestral and descendant populations
@@
 the magnitude of population contraction in each of the descendant
-populations (\bottleScalar{1}{} and \bottleScalar{2}{}) and the
+populations (\bottleScalar{1}{} and \bottleScalar{2}{}), the
 timing of these contractions (\bottleTime{}), and the $N-1$ node heights
 (coalescent times) of the gene tree that gave rise to the $N$ gene
-copies sampled from both populations of the pair.
-Because we only have a single locus for each taxon, there are no locus-specific
-\myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
-\locusRateHetShapeParameter shape parameter for the gamma prior distribution on
-\locusMutationRateScalar{}.
+copies sampled from both populations.
 Lastly, there are between one and 22 divergence time parameters \divt{} in
 the vector \divtvector.
-Overall, for our Philippines data under the simplest model in \msb, there are
-581--602 free parameters (depending on the cardinality of \divtvector).
-Furthermore, under this rich model, the method is sampling over 1002 divergence
-models \citep[i.e., the number of integer partitions of $Y=22$;][]{Oaks2012}.
-
-This is a very difficult inference problem, and the method only uses four
-summary statistics calculated from the sequence alignment of each taxon pair:
+Overall, when applying the simplest model in \msb to the dataset of
+\citet{Oaks2012} with 22 taxon pairs, there are 581--602 free parameters
+(depending on the number of divergence-time parameters in \divtvector).
+Furthermore, under this rich model, the method is estimating the probability of
+1002 divergence models \citep[i.e., the number of integer partitions of
+$Y=22$;][]{Oaks2012}.
+
+This is a very difficult inference problem, especially considering that all the
+information in the sequence alignment of each taxon pair is distilled into four
+summary statistics:
 $\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
@@
 More summary statistics can be used in \msb, but most are highly correlated
-with these four statistics (which are even highly correlated among themselves),
-and thus contribute little additional information about the data.
-
-The large number of parameters and divergence models relative to 
-summary statistics is undoubtedly a key reason the method is so sensitive
-to the prior distributions.
-It also likely contributes to the method's lack of robustness.
-% We use robustness here as a measure of a method's insensitivity to model
-% violations.
-Robustness is an extremely important characteristic of a method to gauge its
+with the four default statistics, and thus contribute little additional
+information about the parameters from the sequence data.
+The large number of parameters and divergence models relative to the amount of
+information in the data is undoubtedly a key reason the method lacks robustness
+to prior conditions.
+Robustness is an important characteristic of a method to gauge its
 applicability to real-world data, because we know the model and priors will be
@@
 
-The approach of \citet{Hickerson2013} adds an additional dimension of model
-choice to the model. They expand the model to sample over eight prior models.
-This extends the original model to having 582-603 free parameters and, more
-importantly, sampling over 8016 unique model states across both divergence
-models and prior models.
-This is a non-trivial extension of the model and, given the method's very large
-number of parameters and divergence models relative to the information used
-from the data, likely plays a major role in the poor behavior of this approach.
-% Certainly, such a large modification of the model demands simulation-based
-% assessments of the behavior of the new model, as we provide here.
-
-In theory, the approach of \citet{Hickerson2013} is very appealing.  It sums
-over multiple candidate prior models to produce a posterior estimate
-marginalized over the uncertainty in prior choice.
-In general, Bayesian model-averaging is a powerful approach that leverages a
-great strength of Bayesian statistical procedures, namely the ability to
-obtain marginalized estimates that incorporate uncertainty in nuisance
-parameters.
-However, given the basic \msb model is already struggling to estimate
-a huge number of parameters and model probabilities with scant information
-from the data, it is not surprising that adding another dimension of
-model choice to the method causes problems.
-
-The recommendations of \citet{Oaks2012} for mitigating the bias and lack of
-robustness of \msb are actually similar to those of \citet{Hickerson2013}, but
-avoid the need for imposing additional model choice.
-\citet{Oaks2012} suggest that uniform priors are inappropriate for many
+The model-averaging approach of \citet{Hickerson2013} adds an additional
+dimension of model choice to the model.
+They expand the model to sample over eight prior models.
+This adds an additional free parameter to the model and, more importantly,
+forces the model to sample over 8016 unique models.
+While this approach is trivial to implement, it is a non-trivial extension of
+the model.
+In theory, the model-averaging approach of \citet{Hickerson2013} is very
+appealing.
+It leverages a great strength of Bayesian statistical procedures, namely the
+ability to obtain marginalized estimates that incorporate uncertainty in
+nuisance parameters.
+However, when averaging over both narrow-empirical and diffuse uniform priors
+for a parameter that is expected to have a very non-uniform likelihood density,
+and in the context of a model-choice method that is highly sensitive to priors,
+it is not too surprising that the approach struggles.
+
+The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
+of \msb are similar to those of \citet{Hickerson2013}, but avoid the
+need for imposing an additional dimension of model choice.
+\citet{Oaks2012} suggest that uniform priors may not be appropriate for many
 parameters of the \msb model, and recommend the use of probability
 distributions from the exponential family.
-If we look at the prior distribution on the divergence time parameter \divt{}
-imposed by the model-averaging approach of \citet{Hickerson2013} we see
-it is a mixture of overlapping uniforms with lower limits of zero
-(Figure~\ref{figMCTauPrior}).
-This looks very much like an exponential distribution, except that each sample
-of divergence times are restricted to the hard bounds of one of the eight prior
-models.
-Thus, it seems more appropriate to place a gamma prior (the exponential being a
-special case of the gamma) on divergence times.
-This would provide investigators flexibility to represent their prior
-uncertainty in model parameters while avoiding broad uniform
-distributions and the need of costly model-averaging.
-This seems like a natural solution to the problem of skewed marginal
-likelihoods and/or sampling error, which are both likely exacerbated by the
-fact that uniform distributions are the only choice of prior for many of the
-parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
-\descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
-\bottleScalar{1}{}, \bottleScalar{2}{}, \locusRateHetShapeParameter,
-\migrationRate{}, \recombinationRate;][]{Oaks2012}.
-
-% Importantly, this strategy is easy to implement.
-Given that ABC methods only need to draw random values from prior distributions
-(hence there are no difficult prior and proposal ratios to calculate, etc.) it
-is easy to use the most appropriate distributions on parameters.
-Also, this approach reduces the temptation of empirically guided priors.
-The investigator can place the majority of the prior density in regions of
-parameter space they believe, \emph{a priori}, are most plausible, but still
-capture uncertainty in the tails of distributions with low density.
-As discussed in \citet{Oaks2012} the use of uniform priors necessitate
-broad priors to avoid excluding the truth \emph{a priori}.
-Thus, to avoid the behavior of the method under these uniform priors, using the
-data is a very tempting, albeit risky, option \citep{Hickerson2013}.
-More appropriate prior distributions would alleviate this issue.
-
-%CWL: This seems to nit-picky. Probably not worth including it
-\section*{Other clarifications}
-
-\subsection*{Graphical prior comparisons}
-\citet{Hickerson2013} advocate the use of what they call graphical checks of
-prior models.
-This entails generating a small number (1000) of random samples from the prior
-and plotting the resulting summary statistics in comparison to the observed
-statistics to see if they coincide (see Figure 1 of \citet{Hickerson2013}).
-As we show above, this strategy can be misleading, because the resulting plots
-of this approach have little correlation with the appropriateness of priors.
-Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
-dataset analyzed by \citet{Hickerson2013}), we do not expect that 1000
-\emph{random} draws from the vast prior parameter space will yield data and
-summary statistics consistent with the observed data.
-In fact, when such random draws are tightly clustered around the observed
-statistics, this can be an indication that the prior is over-fit, as we show
-above (Table~\ref{tabModelChoiceEmpirical} and Figure~\ref{figPCA}).
-Thus, using such plots to select priors should be avoided, and the use of
-posterior predictive analyses would be much more informative about the overall
-fit of models.
-
-\subsection*{Saturation of summary statistics}
-\citet{Hickerson2013} claim the priors used by \citet{Oaks2012} ``cause much of
-the explored parameter space to be beyond the threshold of saturation in most
-mtDNA genes.'' To explore this possibility, we simulated datasets under prior
-settings that match two of the three priors used by \citet{Oaks2012}:
-$\meanDescendantTheta{} \sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim
-U(0.0005, 0.02)$.
-Under this prior, we draw divergence time parameters from a uniform
-distribution of $U(0, 20)$, simulate datasets, and plot the \divt{} values
-against the summary statistics calculated from the resulting datasets
-(Figure~\ref{figSaturationPlot}).
-Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
-and 10 suffered from little to no effect from saturation.
-Even at divergence times of 20 coalescent units, there is still signal in the
-summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
-Thus, the assertion of \citet{Hickerson2013} does not apply to at least
-two of the priors used by \citet{Oaks2012} and, as a result, does not
-explain the bias they found.
+If we look at the prior distribution on divergence-time parameters imposed by
+the model-averaging approach of \citet{Hickerson2013} we see it is a mixture of
+overlapping uniforms with lower limits of zero (Figure~\ref{figMCTauPrior}).
+This looks very much like an exponential distribution, except that in any state
+of the model, all the divergence-time parameters are restricted to the hard
+bounds of one of the uniform distributions.
+Thus, it seems more appropriate to simply place a gamma prior (the exponential
+being a special case of the gamma) on divergence times.
+This would capture the prior uncertainty that \citet{Hickerson2013} are
+suggesting for divergence times (Figure~\ref{figMCTauPrior}) while avoiding
+costly model-averaging and the constraint that all divergence times must fall
+within the hard bounds of the current model state.
+It also would allow an investigator to place the majority of the prior density
+in regions of parameter space they believe, \emph{a priori}, are most
+plausible, but still capture uncertainty in the tails of distributions with low
+density.
+Most importantly, we have found the use of more flexible distributions in place
+of uniform priors improves the power of the method to detect temporal variation
+in divergences and reduces the bias toward models of clustered divergences
+\citep{Oaks2014dpp}.
 
 \section*{Conclusions}
-We demonstrate how the approximate Bayesian model choice method implemented in
+We demonstrate how the approximate Bayesian model-choice method implemented in
 \msb can be strongly biased away from models with greater parameter space.
@@
 truth \emph{a priori}.
-This likely reduces the marginal likelihoods of models with greater parameter
-space, either due to more divergence time parameters or broader prior
-distributions on those parameters.
+These broad priors reduce the marginal likelihoods of models with more
+divergence-time parameters or broader prior distributions on those parameters.
 We show that the empirical Bayesian model-averaging approach of
@@
 of the parameters.
-Exploring alternative prior probability distributions on most of the model's
-parameters, in addition to different priors over divergence models, should
-help mitigate the method's biases.
-
-Based on our results, we discourage the use of empirically informed uniform
-priors in Bayesian model choice, especially when combined with model-averaging
-\ldots
-
-We do not ``assumed that all previous \msb results are invalid,'' as suggested
-by \citet{Hickerson2013}.
-Rather, we conservatively recommend that the common inference of temporally
-clustered divergences \citep{Barber2010, Bell2012, Carnaval2009, Chan2011,
-Daza2010, Hickerson2006, Huang2011, Lawson2010, Leache2007, Plouviez2009,
-Stone2012, Voje2009}, when not accompanied with the necessary simulation-based
-analyses to guide the interpretation of such results, should be treated with
-caution, because the method has been shown to spuriously infer clustered
-divergences over a range of prior conditions.
+
+Whether or not one chooses to use empirically informed priors, our results
+suggest that it is difficult to choose an uniformly distributed prior on
+divergence times that is broad enough to confidently contain the true values of
+parameters while being narrow enough to avoid bias toward models with less
+parameter space.
+While we do not assume ``that all previous \msb results are invalid,'' as
+suggested by \citet{Hickerson2013}, we do conservatively recommend that the
+common inference of temporally clustered divergences \citep{Barber2010,
+    Bell2012, Carnaval2009, Chan2011, Daza2010, Hickerson2006, Huang2011,
+Lawson2010, Leache2007, Plouviez2009, Stone2012, Voje2009}, when not
+accompanied with the necessary analyses to assess the robustness and temporal
+resolution of such results, should be treated with caution, because the method
+has been shown to spuriously infer clustered divergences over a range of prior
+conditions.
+Fortunately, alternative prior probability distributions allow prior
+uncertainty to be accommodated while avoiding excessive prior density in
+regions of low likelihood, which greatly improves the power of the method
+\citet{Oaks2014dpp}.
 
 The work presented herein follows the principles of Open Notebook Science.
-All aspects of the work were recorded in real-time via version-control
-software and are publicly available at
+All aspects of the work were recorded in real-time via version-control software
+and are publicly available at
 \href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.

### 2013-12-13 manuscripts/reply/reply.tex
Updating reply ms.
Incorporating changes made in `short-repy.tex` into `reply.tex` and
`reply-main.tex`. Leaving `short-reply.tex` in repo, because Mark has
been working on this copy.

@@
     \noindent\textbf{ABSTRACT} \\
-    \noindent Recently, \citet{Oaks2012} presented a simulation-based
-    assessment of the approximate Bayesian phylogeographical model choice
-    method implemented in \msb, and found the method to lack robustness and
-    power to detect temporal variation in divergences that are randomly
-    distributed over relatively broad periods of evolutionary history.
-    They conclude that a likely mechanism behind the bias toward models with
-    few divergence events is the small marginal likelihoods of models with more
-    divergence time parameters due to the integration over vast parameter space
-    with low likeihood and high prior density that is imposed by broad uniform
-    priors.
-    \citet{Hickerson2013} have responded to this work, questioning the priors
-    used by \citet{Oaks2012} and presenting an approximate Bayesian
-    model-averaging approach that uses narrow, empirically informed uniform
-    priors.
-    \citet{Hickerson2013} use this method to re-analyze the empirical dataset
-    of \citet{Oaks2012} comprised of 22 pairs of vertebrate populations from
-    the Philippines, and conclude the model-averaging approach circumvents the
-    issues revealed by \citet{Oaks2012} and yields robust inference of
-    divergence models.
-    Furthermore, they argue that the biases revealed by \citet{Oaks2012} are
-    inherent to the inefficient rejection algorithm implemented in \msb, rather
-    than a result of low marginal likelihoods of parameter-rich models.
-    Here we discuss an error in the methodology of \citet{Hickerson2013} that
-    renders some of their results invalid and the remainder difficult to
-    interpret.
-    Furthermore, we demonstrate that the approach of \citet{Hickerson2013} is
-    biased toward models with less parameter space, which manifests in a strong
-    tendency to sample predominantly from models that exclude the true values
-    of the model's parameters.
-    Our results suggest that their model-averaged results are dominated by
-    models that likely exclude at least some of the true divergence times
-    across the 22 pairs of Philippine vertebrates, which adds to the
-    difficulty in interpreting their results.
-    We also show the approach of \citet{Hickerson2013} is still biased toward
-    inferring simultaneous divergences even when the true divergences are
-    randomly distributed over millions of generations.
-    Our results, coupled with those of \citet{Oaks2012}, strongly suggest that
-    low marginal likelihoods of complex models is a primary cause of the bias
-    of \msb toward models with less parameter space.
-    We discuss ways to mitigate this bias without resorting to adding an
-    additional dimension of model choice to the model or using questionable
-    empirically informed uniform priors.
+    \noindent Biogeographers frequently seek to explain population and species
+    differentiation on geographical phenomena.
+    Establishing that a set of population splitting events occurred
+    at the same time can be a potentially persuasive argument that a set of taxa
+    were affected by the same geographic events.
+    Unfortunately, estimating divergence times precisely when one lacks
+    precise information about the rate of molecular evolution is a notoriously
+    difficult problem in evolutionary biology.
+    \citep{Huang2011} introduced an approximate Bayesian approach which can be
+    used to estimate the probabilities of models in which multiple sets of taxa
+    diverge at the same time. 
+    This method (implemented in the software \msb) does not require a common
+    set of character data that links all of the taxa under investigation, but
+    does require the relative mutation rates among the taxa and loci to be
+    assumed \emph{a priori}.
+    Recently, \citet{Oaks2012} used this model-choice framework to 
+    study 22 pairs of vertebrate lineages which are distributed across
+    the Philippines; they also studied the behavior of the \msb approach
+    using computer simulations.
+    \citet{Oaks2012} found that the model was very sensitive to the prior and
+    that the inference method had low power to detect variation in divergences
+    times.
+    These results were not surprising, in light of a rich statistical literature
+    showing that the marginal likelihood of a model is sensitive to the
+    use of vague priors.
+    Because this sensitivity to prior assumptions affects the crucial insights 
+    that a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended
+    that users of the approach should carefully assess the robustness of their 
+    conclusions to different priors.
+    According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
+    was due to excessively broad priors on divergence times, leading to 
+    inadequate numbers of simulation replicates.
+    They proposed a new model-averaging approach that uses narrow, empirically
+    informed uniform priors.
+    Here we demonstrate that the approach of \citet{Hickerson2013} is
+    dangerous in the sense that the empirically-derived priors often
+    exclude from consideration the true values of the models' parameters.
+    On a more fundamental level, we question the value of adopting an empirical
+    Bayesian stance for this model-choice problem, because it can mislead model
+    posterior probabilities, which are inherently measures of belief in the
+    models after prior knowledge is updated by the data.
+    The robust Bayesian approach of conducting analyses under a variety
+    of priors can reveal prior sensitivity and communicate which assumptions
+    underlie model inference.
+    Furthermore, simulations provide insight into the temporal resolution of
+    the method, which in turn helps guide interpretation of results.
 
@@
 
-%CWL:
-% \highLight{The goal of this paper needs to be to highlight the major
-%     differences of opinion on how ABC can be used in comparative phylogeography
-%     and how these changes to msbayes can be made that will improve the
-%     inference in the future. We should avoid nit-picking the Hickerson paper
-%     and try to stay positive.}
 \input{reply-main.tex}
@@
 \section*{Acknowledgments}
-We thank the National Science Foundation for supporting this work (DEB
-1011423).
+We thank Jake Esselstyn for insightful comments that greatly improved this
+work.
+J.\ Oaks thanks the National Science Foundation for supporting this work (DEB
+1011423 and DBI 1308885).
 J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
@@
 \bibliography{../bib/references}
-% \bibliography{references}
 

### 2013-12-17 power-comparison/bin/plot-power-results-comparison.py
Updating power plotting script and plots.
Updating `plot-power-results-comparison.py` plotting script of
the `power-comparison` project, and the generated plots. Fixing
Bayes factor lines and frequencies of highs support. Also,
some font-size adjustments.

@@ -121,11 +121,26 @@ def create_plots(info_path,
         output_dir = project_util.IMAGE_DIR
     if not os.path.exists(output_dir):
         os.mkdir(output_dir)
+    prior_prob_psi_one = {
+            1: 0.0022,  # dpp
+            2: 0.0454,  # old
+            3: 0.0011,  # uniform
+            4: 0.0454,  # u-shaped
+            }
     prior_prob_omega_less_than = {
-            1: 0.0025575,
-            2: 0.049283,
-            3: 0.002171,
-            4: 0.049191}
+            1: 0.0028,  # dpp
+            2: 0.0496,  # old
+            3: 0.0026,  # uniform
+            4: 0.0491,  # u-shaped
+            }
+    bf_prob_psi = dict(zip(prior_prob_psi_one.iterkeys(),
+            [probability.get_probability_from_bayes_factor(bf = 10,
+                    prior_prob = prior_prob_psi_one[k]
+                    ) for k in prior_prob_psi_one.iterkeys()]))
+    bf_prob_omega = dict(zip(prior_prob_omega_less_than.iterkeys(),
+            [probability.get_probability_from_bayes_factor(bf = 10,
+                    prior_prob = prior_prob_omega_less_than[k]
+                    ) for k in prior_prob_omega_less_than.iterkeys()]))
     psi_res, omega_res, prior_index_to_name = parse_results(dmc_sim,
             observed_prefixes_to_include = observed_prefixes_to_include,
             observed_suffixes_to_include = observed_suffixes_to_include,
@@ -209,7 +224,8 @@ def create_plots(info_path,
                 row_labels.append('')
             row_label_size = 28.0
             row_label_offset = 0.08
-            draw_bayes_factor_line = False
+            # draw_bayes_factor_line = False
+            draw_bayes_factor_line = True
             # if num_columns > 4:
             #     draw_bayes_factor_line = True
 
@@ -251,6 +267,7 @@ def create_plots(info_path,
                     div_model_prior = div_model_prior,
                     dpp_concentration_mean = dpp_concentration_mean,
                     bayes_factor = 10,
+                    bayes_factor_prob = bf_prob_psi[prior_index],
                     draw_bayes_factor_line = draw_bayes_factor_line,
                     x_title = r'Posterior probability of one divergence',
                     width = width,
@@ -260,7 +277,7 @@ def create_plots(info_path,
             for cfg, sp in psi_prob_plot.cfg_to_subplot.iteritems():
[…]

### 2013-12-18 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Working on dpp ms.
In the process of integrating new figures into the text, and removing
old figures... A ways to go yet.

@@
-%% Validation accuracy
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelDPP}
-}{figValidationAccuracyDPPDPP}
+%% Validation model choice
+\mFigure{../../validation/images/validation-model-choice-psi.pdf}{
+    \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeNum = 1$}
+}{figValidationModelChoicePsi}
 
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelUniform}
-}{figValidationAccuracyDPPUniform}
+\mFigure{../../validation/images/validation-model-choice-omega.pdf}{
+    \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeDispersion < 0.01$}
+}{figValidationModelChoiceOmega}
 
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelOld}
-}{figValidationAccuracyDPPOld}
+%% Power results
+\mFigure{../../power-comparison/images/old-power-psi-4.pdf}{
+    \powerCommentSummary{\powerSeriesOld}
+    \powerPsiComment{\powerSeriesOld}
+    \timeConversionComment
+    Four of the six data-generating models of the \powerSeriesOld series
+    are shown;
+    please see Figure~S\ref{figPowerPsiOld} for all results.
+}{figPowerPsiOld4}
 
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPP}{\modelUshaped}
-}{figValidationAccuracyDPPUshaped}
+\mFigure{../../power-comparison/images/old-power-omega-4.pdf}{
+    \powerCommentSummary{\powerSeriesOld}
+    \powerDispersionComment{\powerSeriesOld}
+    \timeConversionComment
+    Four of the six data-generating models of the \powerSeriesOld series
+    are shown;
+    please see Figure~S\ref{figPowerPsiOld} for all results.
+}{figPowerPsiOld4}
 
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelDPP}
-}{figValidationAccuracyUniformDPP}
+\mFigure{../../power-comparison/images/old-power-psi-prob-4.pdf}{
+    \powerSupportCommentSummary{\powerSeriesOld}
+    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
+    \timeConversionComment
+    Four of the six data-generating models of the \powerSeriesOld series
+    are shown;
+    please see Figure~S\ref{figPowerPsiOld} for all results.
+}{figPowerPsiOld4}
 
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelUniform}
-}{figValidationAccuracyUniformUniform}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelOld}
-}{figValidationAccuracyUniformOld}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUniform}{\modelUshaped}
-}{figValidationAccuracyUniformUshaped}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelDPP}
-}{figValidationAccuracyOldDPP}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelUniform}
-}{figValidationAccuracyOldUniform}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelOld}
-}{figValidationAccuracyOldOld}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelOld}{\modelUshaped}
-}{figValidationAccuracyOldUshaped}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelDPP}
-}{figValidationAccuracyUshapedDPP}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelUniform}
-}{figValidationAccuracyUshapedUniform}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelOld}
-}{figValidationAccuracyUshapedOld}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_accuracy.pdf}{
-    \validationAccuracyCaption{\modelUshaped}{\modelUshaped}
-}{figValidationAccuracyUshapedUshaped}
-
-%% Validation model choice
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelDPP}
-}{figValidationModelChoiceDPPDPP}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelUniform}
-}{figValidationModelChoiceDPPUniform}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelOld}
-}{figValidationModelChoiceDPPOld}
-
-\mFigure{../../validation/results/dpp/pymsbayes-results/plots/prior-dpp_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPP}{\modelUshaped}
-}{figValidationModelChoiceDPPUshaped}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelDPP}
-}{figValidationModelChoiceUniformDPP}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelUniform}
-}{figValidationModelChoiceUniformUniform}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelOld}
-}{figValidationModelChoiceUniformOld}
-
-\mFigure{../../validation/results/uniform/pymsbayes-results/plots/prior-uniform_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUniform}{\modelUshaped}
-}{figValidationModelChoiceUniformUshaped}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelDPP}
-}{figValidationModelChoiceOldDPP}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelUniform}
-}{figValidationModelChoiceOldUniform}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelOld}
-}{figValidationModelChoiceOldOld}
-
-\mFigure{../../validation/results/old/pymsbayes-results/plots/prior-old_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelOld}{\modelUshaped}
-}{figValidationModelChoiceOldUshaped}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelDPP}
-}{figValidationModelChoiceUshapedDPP}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-uniform_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelUniform}
-}{figValidationModelChoiceUshapedUniform}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-old_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelOld}
-}{figValidationModelChoiceUshapedOld}
-
-\mFigure{../../validation/results/u-shaped/pymsbayes-results/plots/prior-u-shaped_prior-u-shaped_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelUshaped}{\modelUshaped}
-}{figValidationModelChoiceUshapedUshaped}
-
-%% Validation ordered models
-\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
-    \validationAccuracyCaption{\modelDPPOrdered}{\modelDPPOrdered}
-}{figValidationAccuracyDPPOrderedDPPOrdered}
-
-\mFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
-    \validationModelChoiceCaption{\modelDPPOrdered}{\modelDPPOrdered}
-}{figValidationModelChoiceDPPOrderedDPPOrdered}
-
-%% Power accuracy results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerAccuracyExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerAccuracyExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelOld}
-}{figPowerAccuracyExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerAccuracyExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerAccuracyUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerAccuracyUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerAccuracyUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerAccuracyUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerAccuracyOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerAccuracyOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelOld}
-}{figPowerAccuracyOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_accuracy_omega_median.pdf}{
-    \powerAccuracyCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerAccuracyOldUshaped}
-
-%% Power psi results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerPsiExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerPsiExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelOld}
-}{figPowerPsiExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerPsiExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerPsiUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerPsiUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerPsiUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerPsiUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerPsiOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerPsiOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelOld}
-}{figPowerPsiOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_mode.pdf}{
-    \powerPsiCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerPsiOldUshaped}
-
-%% Power dispersion results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerDispersionExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerDispersionExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelOld}
-}{figPowerDispersionExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerDispersionExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerDispersionUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerDispersionUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerDispersionUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerDispersionUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerDispersionOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerDispersionOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelOld}
-}{figPowerDispersionOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_median.pdf}{
-    \powerDispersionCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerDispersionOldUshaped}
-
-%% Power psi probability results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerPsiProbExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerPsiProbExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelOld}
-}{figPowerPsiProbExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerPsiProbExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerPsiProbUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerPsiProbUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerPsiProbUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerPsiProbUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerPsiProbOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerPsiProbOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelOld}
-}{figPowerPsiProbOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_psi_prob.pdf}{
-    \powerPsiProbCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerPsiProbOldUshaped}
-
-%% Power dispersion prob results
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelDPP}
-}{figPowerDispersionProbExpDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelUniform}
-}{figPowerDispersionProbExpUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelOld}
-}{figPowerDispersionProbExpOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/exp_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesExp}{\modelUshaped}
-}{figPowerDispersionProbExpUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelDPP}
-}{figPowerDispersionProbUniformDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUniform}
-}{figPowerDispersionProbUniformUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelOld}
-}{figPowerDispersionProbUniformOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/uniform_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesUniform}{\modelUshaped}
-}{figPowerDispersionProbUniformUshaped}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_dpp_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelDPP}
-}{figPowerDispersionProbOldDPP}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_uniform_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelUniform}
-}{figPowerDispersionProbOldUniform}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_old_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelOld}
-}{figPowerDispersionProbOldOld}
-
-\mFigure{../../power-comparison/results/pymsbayes-results/plots/old_u-shaped_power_omega_prob.pdf}{
-    \powerDispersionProbCaption{\powerSeriesOld}{\modelUshaped}
-}{figPowerDispersionProbOldUshaped}
+\mFigure{../../power-comparison/images/old-power-omega-prob-4.pdf}{
+    \powerSupportCommentSummary{\powerSeriesOld}
+    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesOld}
+    \timeConversionComment
+    Four of the six data-generating models of the \powerSeriesOld series
+    are shown;
+    please see Figure~S\ref{figPowerPsiOld} for all results.
+}{figPowerPsiOld4}

### 2013-12-18 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Updating dpp manuscript.
Correcting posterior distribution equations. Making the marginal
likelihood (denominator) of Bayes rule conditional on the parameters
that are fixed by the user.
Updating DPP manuscript.
Adding subsection placeholders for empirical analyses of the Philippines
data in both Methods and Results.
Working on dpp ms.
In the process of integrating new figures into the text, and removing
old figures... A ways to go yet.

@@
         \mutationRateScalarConstantVector)}{
-        p(\alignmentVector)},
+        p(\alignmentVector \given \hkyModelVector, \ploidyScalarVector,
+        \mutationRateScalarConstantVector)},
     \label{eq:fullModelCompactJoint}
@@
         p(\demographicParamVector)}{
-        p(\alignmentVector)},
+        p(\alignmentVector \given \hkyModelVector, \ploidyScalarVector,
+        \mutationRateScalarConstantVector)},
     \label{eq:fullModelCompact}
@@
         p(\demographicParamVector)}{
-        p(\ssSpace)},
+        p(\ssSpace \given \hkyModelVector, \ploidyScalarVector,
+        \mutationRateScalarConstantVector)},
     \label{eq:approxModelCompact}
@@
 
+\subsection*{An empirical application}
+\ldots
+
 \section*{Results}
 \subsection*{Validation analyses: Estimation accuracy}
-In terms of estimating summary statistics of model parameters (\divTimeNum,
-\divTimeDispersion, and \divTimeMean), the \modelDPP performs similarly when
-applied to datasets generated under all four models of Table \ref{tabPriors}
-(Figs.\
-\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP}).
-The same is true for models \modelUniform (Figs.\
-\labelcref{figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform})
-and \modelUshaped (Figs.\
-\labelcref{figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped}).
-All three of the models with exponential priors on nuisance parameters
-(\modelDPP, \modelUniform, and \modelUshaped) perform well when applied
-to the data generated under the original \msb model, \modelOld (Figs.\
-\labelcref{figValidationAccuracyOldDPP,figValidationAccuracyOldUniform,figValidationAccuracyOldUshaped}).
-The \modelOld model also performed well on its own datasets (Figure
-\ref{figValidationAccuracyOldOld}), but performed poorly when applied to the
-data generated under the other three models (Figs.\
-\labelcref{figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld}).
-Thus, the newly implemented models display greater robustness than
-the original model.
-
-Posterior adjustment with GLM regression does improve the estimation
-accuracy for the continuous variables \divTimeDispersion and \divTimeMean,
-however, this is not always true for estimating the discrete variable
-\divTimeNum (Figs.\
-\labelcref{figValidationAccuracyDPPDPP,figValidationAccuracyUniformDPP,figValidationAccuracyOldDPP,figValidationAccuracyUshapedDPP,figValidationAccuracyDPPUniform,figValidationAccuracyUniformUniform,figValidationAccuracyOldUniform,figValidationAccuracyUshapedUniform,figValidationAccuracyDPPUshaped,figValidationAccuracyUniformUshaped,figValidationAccuracyOldUshaped,figValidationAccuracyUshapedUshaped,figValidationAccuracyDPPOld,figValidationAccuracyUniformOld,figValidationAccuracyUshapedOld,figValidationAccuracyOldOld}).
+In terms of estimating the variance of divergence times (\divTimeDispersion),
+the models with exponentially distributed priors (\modelUshaped, \modelUniform,
+and \modelDPP) perform similarly when applied to datasets generated under all
+four of the models of Table \ref{tabPriors}
+(Figure~S\ref{figValidationAccuracyOmega}).
+The \modelOld model performs similarly to these models when applied to its
+own datasets, but is sensitive to model violations and is more biased when
+applied to data generated under the other three models
+(Figure~S\ref{figValidationAccuracyOmega}).
+This results are similar for the GLM-adjusted estimates of \divTimeDispersion,
+albeit the regression adjustment tends to improve estimates of this continuous
+statistic for all the models
+(Figure~S\ref{figValidationAccuracyPsiGlm}).
+
+The same general pattern is seen for estimates of \divTimeMean, with
+the all four models performing similarly when applied to the data generated
+under \modelOld, the models with exponentially distributed priors performing
+similarly when applied to data generated under other models, and the
+\modelOld is more biased when applied to data generated under other models
+(Figure~S\ref{figValidationAccuracyTime}).
+Also, the regression adjustment tends to slightly improve estimates of
+this continuous statistic for all of the models
+(Figure~S\ref{figValidationAccuracyTimeGlm}).
+
+In terms of estimating the number of divergence events (\divTimeNum), the
+\modelDPP model has the lowest root mean square error (RMSE) when applied
+to data generated under most of the models of Table \ref{tabPriors}
+(Figure~S\ref{figValidationAccuracyPsi}).
+The \modelOld model performs slightly better when applied to its own data,
+but is the worst performer when applied to data generated under other models
+(Figure~S\ref{figValidationAccuracyPsi}).
+There is a trend of $\modelDPP > \modelUniform > \modelUshaped > \modelOld$
+in terms of estimation accuracy as measured by RMSE when the models are applied
+to data generated under most of the models
+(Figure~S\ref{figValidationAccuracyPsi}).
+Unlike for the continuous statistics, regression adjustment of this discrete
+statistic tends to increase estimation bias; all of the models tend to
+underestimate \divTimeDispersion after the GLM-adjustment
+(Figure~S\ref{figValidationAccuracyPsiGlm}).
 
@@
 estimate the posterior probabilities of divergence models.
-Our results demonstrate that the unadjusted estimates of divergence models are
-generally more accurate than the regression-adjusted estimates, especially when
-using $\divTimeDispersion < 0.01$ to assess the probability of the one
-divergence model (Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
-Thus, we will focus our discussion of the results on the unadjusted estimates.
-
+Consistent with the findings of \citet{Oaks2012}, our results demonstrate that
+the unadjusted estimates of divergence-model posterior probabilities are
+generally more accurate than regression-adjusted estimates
+(Figure~\ref{figValidationModelChoicePsi} versus
+S\ref{figValidationModelChoicePsiGlm} and
+Figure~\ref{figValidationModelChoiceOmega} versus
+S\ref{figValidationModelChoiceOmegaGlm}).
+Regression adjustment results in biased estimates of the posterior probability
+of the one-divergence model when all model assumptions are satisfied, which is
+well illustrated in Figure S\ref{figValidationModelChoiceOmegaGlm}.
+As a result, we will focus our discussion of the results on the unadjusted
+estimates.
+
+We find that all four models accurately estimate the posterior probability of
+the one-divergence model when applied to their own datasets (i.e., when the
+prior is correct; see diagonal of Figures~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).
+The \modelUniform and \modelDPP show robustness to prior violations and
+perform well when applied to data generated under other models
+(Figures~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).
+However, both are less accurate and tends to underestimate the probability of the
+one-divergence model when applied to the data generated under \modelUshaped
+(Figures~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).
 The model-choice behavior of the \modelDPP is relatively accurate when applied
@@
 
+\subsection*{Empirical results}
+\ldots
+
 \section*{Discussion}

### 2013-12-18 manuscripts/dpp-msbayes/dpp-msbayes-si-figures.tex
Working on dpp ms.
In the process of integrating new figures into the text, and removing
old figures... A ways to go yet.

@@
 %% Validation accuracy
-\siFigure{../../validation/images/validation-accuracy-psi.pdf}{
-    \validationAccuracyComparisonCaption{unadjusted}{\divTimeNum}
-    \jitterComment
-}{figValidationAccuracyPsi}
-
-\siFigure{../../validation/images/validation-accuracy-psi-glm.pdf}{
-    \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeNum}
-    \jitterComment
-}{figValidationAccuracyPsiGlm}
-
 \siFigure{../../validation/images/validation-accuracy-omega.pdf}{
@@
 
+\siFigure{../../validation/images/validation-accuracy-psi.pdf}{
+    \validationAccuracyComparisonCaption{unadjusted}{\divTimeNum}
+    \jitterComment
+}{figValidationAccuracyPsi}
+
+\siFigure{../../validation/images/validation-accuracy-psi-glm.pdf}{
+    \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeNum}
+    \jitterComment
+}{figValidationAccuracyPsiGlm}
+
 %% Validation model choice
-\siFigure{../../validation/images/validation-model-choice-psi.pdf}{
-    \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeNum = 1$}
-}{figValidationModelChoicePsi}
-
 \siFigure{../../validation/images/validation-model-choice-psi-glm.pdf}{
@@
 }{figValidationModelChoicePsiGlm}
-
-\siFigure{../../validation/images/validation-model-choice-omega.pdf}{
-    \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeDispersion < 0.01$}
-}{figValidationModelChoiceOmega}
 
@@
 %% Power accuracy results
-\siFigure{../../power-comparison/images/old-power-omega-accuracy-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/old-power-omega-accuracy-6.pdf}{
     \powerAccuracyComparisonCaption{\powerSeriesOld}
@@
 
-\siFigure{../../power-comparison/images/uniform-power-omega-accuracy-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/uniform-power-omega-accuracy-6.pdf}{
     \powerAccuracyComparisonCaption{\powerSeriesUniform}
@@
 
-\siFigure{../../power-comparison/images/exp-power-omega-accuracy-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/exp-power-omega-accuracy-6.pdf}{
     \powerAccuracyComparisonCaption{\powerSeriesExp}
@@
 %% Power psi results
-\siFigure{../../power-comparison/images/old-power-psi-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/old-power-psi-6.pdf}{
     \powerComment{\powerSeriesOld}
@@
 
-\siFigure{../../power-comparison/images/uniform-power-psi-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/uniform-power-psi-6.pdf}{
     \powerComment{\powerSeriesUniform}
@@
 
-\siFigure{../../power-comparison/images/exp-power-psi-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/exp-power-psi-6.pdf}{
     \powerComment{\powerSeriesExp}
@@
 %% Power dispersion results
-\siFigure{../../power-comparison/images/old-power-omega-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/old-power-omega-6.pdf}{
     \powerComment{\powerSeriesOld}
@@
 
-\siFigure{../../power-comparison/images/uniform-power-omega-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/uniform-power-omega-6.pdf}{
     \powerComment{\powerSeriesUniform}
@@
 
-\siFigure{../../power-comparison/images/exp-power-omega-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/exp-power-omega-6.pdf}{
     \powerComment{\powerSeriesExp}
@@
 %% Power psi probability results
-\siFigure{../../power-comparison/images/old-power-psi-prob-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/old-power-psi-prob-6.pdf}{
     \powerSupportComment{\powerSeriesOld}
@@
 
-\siFigure{../../power-comparison/images/uniform-power-psi-prob-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/uniform-power-psi-prob-6.pdf}{
     \powerSupportComment{\powerSeriesUniform}
@@
 
-\siFigure{../../power-comparison/images/exp-power-psi-prob-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/exp-power-psi-prob-6.pdf}{
     \powerSupportComment{\powerSeriesExp}
@@
 %% Power dispersion prob results
-\siFigure{../../power-comparison/images/old-power-omega-prob-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/old-power-omega-prob-6.pdf}{
     \powerSupportComment{\powerSeriesOld}
@@
 
-\siFigure{../../power-comparison/images/uniform-power-omega-prob-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/uniform-power-omega-prob-6.pdf}{
     \powerSupportComment{\powerSeriesUniform}
@@
 
-\siFigure{../../power-comparison/images/exp-power-omega-prob-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/exp-power-omega-prob-6.pdf}{
     \powerSupportComment{\powerSeriesExp}

### 2013-12-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on DPP ms.
Finishing incorporation of the new figures. Also, updating some of the
new figures to add an extra digit for the exponential distributions of
divergence times.

@@
 \ref{figValidationModelChoiceOmega}).
-However, both are less accurate and tends to underestimate the probability of the
+However, both are less accurate and tend to underestimate the probability of the
 one-divergence model when applied to the data generated under \modelUshaped
@@
 \ref{figValidationModelChoiceOmega}).
-The model-choice behavior of the \modelDPP is relatively accurate when applied
-to datasets generated under models \modelDPP, \modelUniform, and \modelOld
-(Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP}).
-It is less accurate and tends to underestimate the probability of the
-one-divergence model when applied to the data generated under \modelUshaped
-(Figure \ref{figValidationModelChoiceUshapedDPP}).
-The \modelUniform model performs almost identically as \modelDPP in
-terms of model-choice accuracy (Figs.\
-\labelcref{figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}).
-
-The original \msb model \modelOld performs reasonably well when applied
-to its own datasets (Figure \ref{figValidationModelChoiceOldOld}), but
-it performs poorly when applied to data generated under any of the other
-models (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
-In these cases the \modelOld model is strongly biased towards over-estimating
-the posterior probability of the one-divergence model, especially when the
-divergence models are not distributed under its U-shaped prior (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld}).
+In contrast, the original is biased toward overestimating the posterior
+probability of the one-divergence model when applied to data generated under
+the other three models
+(Figures~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).
+This bias is particularly strong whenever divergence models are not
+distributed under its U-shaped prior
+(Figure~\ref{figValidationModelChoicePsi}C--D).
 The other model with the U-shaped prior on divergence models, but exponential
-priors on nuisance parameters (\modelUshaped), performs similarly to \modelOld
-in that it performs well when applied to its own data, but overestimates
-the probability of the one-divergence model when applied to data generated
-by the other models (Figs.\
-\labelcref{figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceUshapedUshaped}).
-However, the bias toward over-estimating the probability of the one-divergence
-model is stronger in the \modelOld model than \modelUshaped.
+priors on parameters (\modelUshaped), performs similarly to the \modelOld model
+in that it performs well when applied to its own data, but overestimates the
+probability of the one-divergence model when applied to data generated by the
+other models
+(Figures~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).
+However, the bias is stronger in the \modelOld model than \modelUshaped.
 
@@
 relatively robust in terms of model-choice accuracy, and when model violations
-do cause them to be biased, they tend to under-estimate the probability of
-the biogeographically interesting model of a single divergence event.
-In contrast, the \modelOld model is very sensitive to model violations,
-and strongly over-estimates the one-divergence model whenever the model
-is misspecified.
+do cause them to be biased, they tend to under-estimate the probability of the
+biogeographically interesting model of a single divergence event.
+In contrast, the \modelOld model is very sensitive to model violations, and
+strongly over-estimates the one-divergence model whenever the model is
+misspecified.
 Furthermore, the results suggest that using exponentially distributed priors on
@@
 \subsection*{Validation analyses: Ordered divergence models}
-Our results show that the ordered models of divergence, \modelDPPOrdered,
-perform similarly to sampling over unordered models (\modelDPP) when the prior
-is correct (Figs.\
-\labelcref{figValidationAccuracyDPPOrderedDPPOrdered,figValidationModelChoiceDPPOrderedDPPOrdered}).
+Our results show that the method performs similarily when sampling over ordered
+models of divergence
+(Figures S\ref{figValidationAccuracyDPPOrderedDPPOrdered} \&
+S\ref{figValidationModelChoiceDPPOrderedDPPOrdered}).
 This suggests that the method is not adversely affected by the large increase
@@
 All of the models we evaluated (Table \ref{tabPriors}) struggle to estimate the
-dispersion index of divergence times \divTimeDispersion regardless of which of
-the three series of models (Table \ref{tabPowerModels}) the data were generated
-under (Figs.\
-\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform,figPowerAccuracyExpOld,figPowerAccuracyExpUshaped,figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped,figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
+variance of divergence times \divTimeDispersion regardless of which of the
+three series of models (Table \ref{tabPowerModels}) the data were generated
+under
+(Figures
+S\labelcref{figPowerAccuracyOld,figPowerAccuracyUniform,figPowerAccuracyExp}).
+The models with the U-shaped prior on divergence models (\modelOld and
+\modelUshaped) tend to underestimate the variance in divergence times 
+(Plots A--L of Figures
+S\labelcref{figPowerAccuracyOld,figPowerAccuracyUniform,figPowerAccuracyExp}).
+whereas the models with Uniform or Dirichlet-process priors over divergence
+models tend to overestimate variance in divergence times
+(Plots M--X of Figures
+S\labelcref{figPowerAccuracyOld,figPowerAccuracyUniform,figPowerAccuracyExp}).
+
 When the divergence times of the 22 population pairs are randomly drawn from a
 series of exponential priors (\powerSeriesExp), the \modelDPP model is the
-best estimator of \divTimeDispersion, followed by \modelUniform (figs.\
-\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform}).
+best estimator of \divTimeDispersion, followed by \modelUniform
+(Figure~\ref{figPowerAccuracyExp}).
 The \modelOld model is strongly biased toward underestimating
 \divTimeDispersion, estimating values of zero for most of the replicates across
-all the data models of \powerSeriesExp (Figure \ref{figPowerAccuracyExpOld}).
+all the data models of \powerSeriesExp (Figure \ref{figPowerAccuracyExp}).
 The results of the \modelUshaped model are intermediate between those of
 \modelOld and the new models \modelDPP and \modelUniform
-(Figure \ref{figPowerAccuracyExpUshaped}).
-
-When the true divergence times are randomly drawn from a series of uniform
-priors (\powerSeriesUniform), all of the models poorly estimate
-\divTimeDispersion (Figs.\
-\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped}).
-The \modelDPP and \modelUniform models tend to over-estimate the variance in
-divergence times (Figs.\
-\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform}), whereas
-the \modelOld model, again, underestimates \divTimeDispersion, estimating
-values of zero for most replicates across all the data models of
-\powerSeriesUniform (Figure \ref{figPowerAccuracyUniformOld}).
+(Figure \ref{figPowerAccuracyExp}).
+
+Similarly, when the true divergence times are randomly drawn from a series of
+uniform priors (\powerSeriesUniform), the \modelDPP and \modelUniform models
+tend to over-estimate the variance in divergence times, whereas the \modelOld
+model, underestimates \divTimeDispersion, estimating values of zero for most
+replicates across all the data models of \powerSeriesUniform
+(Figure~\ref{figPowerAccuracyUniform}).
 Again, the performance of the \modelUshaped model is intermediate between the
 \modelOld and \modelDPP/\modelUniform models (Figure
-\ref{figPowerAccuracyUniformUshaped}).
-This pattern of results across the four models is very similar to their
-behavior when applied to data simulated under the series of models of
-\powerSeriesOld (Figs.\
-\labelcref{figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
+\ref{figPowerAccuracyUniform}).
+The results are very similar when the four models are applied to the data
+simulated under the \powerSeriesOld seires of models
+(Figure~\ref{figPowerAccuracyOld}).
 
 \subsection*{Power analyses: Model choice}
-There is a large amount of heterogeneity in estimating the number of divergence
-events (\divTimeNum) both among the prior models, and across the different
-series of data models (Table \ref{tabPowerModels}) (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped}).
-The \modelDPP model tends to infer highly clustered divergence models across
-all three series of data models when the divergences are recent, but performs
-the best of the four prior models  when divergences are random over a timescale
-of about one to two coalescent units (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiUniformDPP,figPowerPsiOldDPP,}).
-The \modelUniform model performs the best when divergences are random
-over the most recent timescales, but tends to only infer a tight
-cluster of \divTimeNum values even when divergences are random
-over longer periods (Figs.\
-\labelcref{figPowerPsiExpUniform,figPowerPsiUniformUniform,figPowerPsiOldUniform,}).
-The \modelOld model performs the worst of the four prior models across
-all three series of data models, inferring a single divergence event across
-most of the 18,000 simulations (Figs.\
-\labelcref{figPowerPsiExpOld,figPowerPsiUniformOld,figPowerPsiOldOld,}).
-Even when the simulated data are identically distributed as the \modelOld model
-(except for the divergence times) the \modelOld is still strongly biased
-towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
-model is the least likely to infer a single event when applied to these data
-(Figure \ref{figPowerPsiOldDPP}).
-Using exponential priors on nuisance parameters does increase the power
-of the \modelUshaped model compared to \modelOld across all three series
-of data models, but the U-shaped prior still prevents the model
-from performing as well as the \modelDPP and \modelUniform models (Figs.\
-\labelcref{figPowerPsiExpUshaped,figPowerPsiUniformUshaped,figPowerPsiOldUshaped,}).
+Our modifications of the \msb model decrease the method's bias toward clustered
+divergences when applied to data generated under random divergence times
+(Figure~\ref{figPowerPsiOld4}A--D \& plots A--F of Figures
+S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
+The \modelOld model performs the worst of the four models across
+all three series of data-generating models, inferring a single divergence event across
+most of the 18,000 simulations
+(Figure~\ref{figPowerPsiOld4}A--D \& plots A--F of Figures
+S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
+Importantly, the \modelOld tends to strongly support these estimates
+of one divergence across most of the simulations
+(Figure~\ref{figPowerPsiProbOld4}A--D \& plots A--F of Figures
+S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+The \modelDPP model also prefers the one-divergence model 
+when divergences are random over narrow windows of time, but
+performs much better when divergences are random over a timescale
+of 1--2 coalescent units
+(Figure~\ref{figPowerPsiOld4}M--P \& plots S--X of Figures
+S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
+However, even when \modelDPP infers the one-divergence model over narrow
+timescales, the posterior probability support is always low
+(Figure~\ref{figPowerPsiProbOld4}M--P \& plots S--X of Figures
+S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+% Even when the simulated data are identically distributed as the \modelOld model
+% (except for the divergence times) the \modelOld is still strongly biased
+% towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
+% model is the least likely to infer a single event when applied to these data
+% (Figure \ref{figPowerPsiOldDPP}).
+The \modelUniform model never infers the one-divergence model in any of the
+simulation replicates but still tends to infer relatively few (4--6) divergence
+events when divergences are random over longer periods 
+(Figure~\ref{figPowerPsiOld4} I--L \& plots M--R of Figures
+S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
+Using exponential priors on divergence-time and theta parameters does increase
+the power of the \modelUshaped model compared to \modelOld across all three
+series of data models, but the U-shaped prior still prevents the model from
+performing as well as the \modelDPP and \modelUniform models 
+(Figure~\ref{figPowerPsiOld4} \& 
+S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
 
 The improved power of the new models are even more pronounced when looking at
-estimates of the dispersion index of divergence times (\divTimeDispersion)
-across the simulations (Figs.\
-\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped}).
+estimates of the variance of divergence times (\divTimeDispersion) across the
+simulations
+(Figure~\ref{figPowerOmegaOld4} \& 
+S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).
+The performance among the models is so different, that the histograms of
+\divTimeDispersion estimates cannot be plotted along a shared x-axis.
 The \modelDPP and \modelUniform models perform similarly across all three
@@
 one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
-replicates across all the simulations (Figs.\
-\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionOldDPP,figPowerDispersionOldUniform}).
-The \modelOld model infers values consistent with a single divergence event in
-most of the replicates across all the simulations (Figs.\
-\labelcref{figPowerDispersionExpOld,figPowerDispersionUniformOld,figPowerDispersionOldOld}).
-Using exponential priors on nuisance parameters greatly increases the power
-of the \modelUshaped model relative to \modelOld, but it still has
-less power than the models with DPP or uniform priors across
-divergence models (Figs.\
-\labelcref{figPowerDispersionExpUshaped,figPowerDispersionUniformUshaped,figPowerDispersionOldUshaped}).
-
-When looking at the estimated posterior probability of the single-divergence
-model across the power analyses, we also see increased power of the new models
-(Figs.\
-\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped}).
-The \modelDPP model estimates low posterior probability of
-$\divTimeNum = 1$ across all of the simulations (Figs.\
-\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbUniformDPP,figPowerPsiProbOldDPP}).
-The \modelDPP does infer strong support for one divergence as gauged by a Bayes
-factor of greater than 10.
-However, this is not that unexpected given the low prior probability of a
-single divergence under this model.
-Also, the Bayes factor calculations for the \modelDPP replicates are only
-approximate; they are calculated based on the prior mean of the concentration
-parameter \concentrationParam.
-
-The \modelUniform model infers very low posterior probabilities for the
-one-divergence model across all the simulations (Figs.\
-\labelcref{figPowerPsiProbExpUniform,figPowerPsiProbUniformUniform,figPowerPsiProbOldUniform}),
-whereas the \modelOld model infers high posterior probabilities of a single
-divergence for most replicates across all simulations (Figs.\
-\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld}).
-Again, we see that the exponential priors on nuisance parameters greatly
-increase power and result in lower estimates of the probability of
-one divergence when comparing \modelUshaped to \modelOld (Figs.\
-\labelcref{figPowerPsiProbExpUshaped,figPowerPsiProbUniformUshaped,figPowerPsiProbOldUshaped}).
+replicates across all the simulations.
+In contrast, the \modelOld model infers values consistent with a single divergence event in
+most of the replicates across all the simulations.
+Using exponential priors on divergence-time and theta parameters greatly
+increases the power of the \modelUshaped model to detect variation in
+divergence times relative to \modelOld, but it still has less power than the
+models with Dirichlet-process or uniform priors across divergence models 
+(Figure~\ref{figPowerOmegaOld4} \& 
+S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).
+
+As mentioned above, we also see increased power of the new models when
+looking at the estimated posterior probability of the one-divergence model
+across the power analyses 
+(Figure~\ref{figPowerPsiProbOld4} \& 
+S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+The \modelDPP and \modelUniform models estimate low posterior probability of
+$\divTimeNum = 1$ across all of the simulations.
+This is in contrast to the \modelOld model, which infers high posterior probabilities of a single
+divergence for most replicates across all simulations
+(Figure~\ref{figPowerPsiProbOld4} \& 
+S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+The exponential priors on divergence-time and theta parameters (model
+\modelUshaped) result in lower estimates of the probability of one divergence
+when compared to \modelOld, but higher estimates when compared to \modelUniform
+and \modelDPP
+(Figure~\ref{figPowerPsiProbOld4} \& 
+S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+The \modelDPP and \modelUniform models do frequently support the one-model
+according to a Bayes factor criterion of greater than 10, however still less
+frequently than the \modelOld model.
+This result is not surprising given the extremely small prior probability of
+the one-divergence model under the \modelDPP and \modelUniform models (I.e.,
+very few posterior samples of the one-divergence model will result in a
+large Bayes factor under these models).
+However, the small posterior probability of the one-divergence model estimated
+under \modelDPP and \modelUniform should prevent an investigator from
+overinterpreting the Bayes factor as strong support for clustered divergences.
 
 Lastly, when we look at the estimated posterior probability of
-\divTimeDispersion being consistent with one divergence parameter
+\divTimeDispersion being consistent with one shared divergence
 ($p(\divTimeDispersion < 0.01 | \ssSpace)$), we see the same pattern of model
@@
 all simulations, \modelOld inferring high probabilities, and \modelUshaped
-inferring intermediate values (Figs.\
-\labelcref{figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
+inferring intermediate values
+(Figure~\ref{figPowerOmegaProbOld4} \& 
+S\labelcref{figPowerOmegaProbOld,figPowerOmegaUniform,figPowerOmegaExp}).
 
@@
 \section*{Discussion}
-Our results demonstrate that using alternative priors on nuisance parameters
-and divergence models improved the behavior of the \msb model.
+Our results demonstrate that using alternative priors on parameters and
+divergence models improved the behavior of the \msb model.
 In our implementation, model-choice estimation is more accurate and shows
-greater robustness to model violations (Figs.\
-\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
+greater robustness to model violations
+(Figure~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).
 The original model is very sensitive to model violations and, when present,
 strongly over-estimates the probability of one divergence event shared
-across all taxa (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
-When more appropriate priors are used for nuisance parameters, and either
-a Dirichlet-process or uniform prior applied across divergence models,
-the model is less sensitive to violations (Figs.\
-\labelcref{figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}),
-and, when they do cause bias, the method is conservative and tends to
-under-estimate the probability of biogeographically interesting models of
-temporally clustered divergences (Figs.\
-\labelcref{figValidationModelChoiceUshapedDPP,figValidationModelChoiceUshapedUniform}).
+across all taxa
+(Figure~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).
+When more appropriate priors are used for divergence-time and theta parameters,
+and either a Dirichlet-process or uniform prior applied across divergence
+models, the model is less sensitive to violations, and, when they do cause
+bias, the method is conservative and tends to under-estimate the probability of
+biogeographically interesting models of temporally clustered divergences
+(Figure~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).
 
@@
 detect random variation in divergence times, reducing the bias toward
-estimating clustered divergences (Figs.\
-\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped,figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped,figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped,figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
+estimating clustered divergences
+(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}.
 Our results are similar to those of \citet{Oaks2012} in that we also find
 the \msb model to infer strong support for clustered divergences when
-divergences are random over quite broad timescales (Figs.\
-\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld,figPowerDispersionProbExpOld,figPowerDispersionProbUniformOld,figPowerDispersionProbOldOld}).
+divergences are random over quite broad timescales
+(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}.
 Our results expand on this by showing that this behavior is consistent
-across a range conditions underlying the data.
+across a range of conditions underlying the data.
 When using the new priors, the method is less prone to spurious inference of
@@
 (\modelUshaped), we were able to determine the relative effects of the
-modifications to model.
+modifications to the model.
 Across all of our analyses, the results show that using better priors on
-nuisance parameters alone does improve the performance of the method.
+divergence-time and nuisance parameters alone does improve the performance of
+the method.
 The magnitude of the bias toward inferring support for the one-divergence
 model when there are model violations is reduced when the exponential
-priors are used in place of the uniform priors (Figs.\
-\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped}).
+priors are used in place of the uniform priors
+(Figure~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).
 Furthermore, better priors on nuisance parameters improves the method's
@@
 models of temporally clustered divergences is caused by a combination of
-(1) uniform priors on nuisance parameters hindering the marginal likelihoods of
-models with more \divTime{} parameters and
+(1) uniform priors on parameters hindering the marginal likelihoods of models
+with more \divTime{} parameters and
 (2) the U-shaped prior on divergence models giving low prior density to models
@@
 caution when using the method and interpreting its results.
-The method attempts to approximate the posterior of a very parameter rich model
+The method attempts to approximate the posterior of a very parameter-rich model
 using relatively little information from the data.

### 2013-12-19 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex
Working on DPP ms.
Finishing incorporation of the new figures. Also, updating some of the
new figures to add an extra digit for the exponential distributions of
divergence times.

@@
     \sffamily
-    % \footnotesize
+    \footnotesize
     % \addtolength{\tabcolsep}{-0.08cm}
@@
     \caption{The models we used to simulate pseudo-replicate datasets for
-        assessing the power of the models in Table \ref{tabPriors}.}
+        assessing the power of the models in Table \ref{tabPriors}.
+        Distributions of divergence times are given in units of \globalcoalunit
+        generations and millions of generations ago (MGA), the latter
+        assuming a per-site rate of 1\e{-8} mutations per generation.}
     \centering
-    \begin{tabular}{ l l l l l }
+    \begin{tabular}{ l l l l l l }
         \toprule
-        & \multicolumn{4}{c}{Priors} \\
-        \cmidrule(){2-5}
-        Model series & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
+        & \multicolumn{5}{c}{Priors} \\
+        \cmidrule(){2-6}
+        Model series & \divTimeIndexVector & \divTime{} (\globalcoalunit gens.) & \divTime{} (MGA) & \ancestralTheta{} & \descendantTheta{}{} \\
         \midrule
@@
                             & $\divTime{} \sim U(0,0.2)$
+                            & $\divTime{} \sim U(0,0.5)$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTime{} \sim U(0,0.4)$
+                            & $\divTime{} \sim U(0,1.0)$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTime{} \sim U(0,0.6)$
+                            & $\divTime{} \sim U(0,1.5)$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTime{} \sim U(0,0.8)$
+                            & $\divTime{} \sim U(0,2.0)$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTime{} \sim U(0,1.0)$
+                            & $\divTime{} \sim U(0,2.5)$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTime{} \sim U(0,2.0)$
+                            & $\divTime{} \sim U(0,5.0)$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTime{} \sim U(0,0.2)$
+                            & $\divTime{} \sim U(0,0.5)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTime{} \sim U(0,0.4)$
+                            & $\divTime{} \sim U(0,1.0)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTime{} \sim U(0,0.6)$
+                            & $\divTime{} \sim U(0,1.5)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTime{} \sim U(0,0.8)$
+                            & $\divTime{} \sim U(0,2.0)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTime{} \sim U(0,1.0)$
+                            & $\divTime{} \sim U(0,2.5)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTime{} \sim U(0,2.0)$
+                            & $\divTime{} \sim U(0,5.0)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
             \powerSeriesExp & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(17.3205)$
+                            & $\divTime{} \sim Exp(17.32)$
+                            & $\divTime{} \sim Exp(mean = 0.14)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(8.6603)$
+                            & $\divTime{} \sim Exp(8.66)$
+                            & $\divTime{} \sim Exp(mean = 0.29)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(5.7735)$
+                            & $\divTime{} \sim Exp(5.77)$
+                            & $\divTime{} \sim Exp(mean = 0.43)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(4.3301)$
+                            & $\divTime{} \sim Exp(4.33)$
+                            & $\divTime{} \sim Exp(mean = 0.58)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(3.4641)$
+                            & $\divTime{} \sim Exp(3.46)$
+                            & $\divTime{} \sim Exp(mean = 0.72)$
                             & $\ancestralTheta{} \sim Exp(40)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(1.7321)$
+                            & $\divTime{} \sim Exp(1.73)$
+                            & $\divTime{} \sim Exp(mean = 1.44)$
                             & $\ancestralTheta{} \sim Exp(40)$

### 2013-12-19 power-comparison/bin/plot-power-results-comparison.py

@@ -633,7 +633,7 @@ def tau_prior_in_generations(cfg, mu = 1e-8):
         return r'$\tau \sim U(0, \, {0:.1f} \, \mathsf{{MGA}})$'.format(upper_tau)
     elif isinstance(cfg.tau, probability.GammaDistribution):
         mean_tau = (cfg.tau.mean * (mean_theta / mu)) / 1000000.0
-        return r'$\tau \sim Exp(\mathsf{{mean}} = {0:.1f} \, \mathsf{{MGA}})$'.format(mean_tau)
+        return r'$\tau \sim Exp(\mathsf{{mean}} = {0:.2f} \, \mathsf{{MGA}})$'.format(mean_tau)
     else:
         raise Exception('unsupported tau distribution: {0}'.format(type(cfg.tau)))
 

### 2013-12-23 empirical-analyses/bin/plot-results-comparison.py
Updating plotting script for empirical results.
Updating formatting of column and row labels in plots for the
empirical results for the dpp ms.

@@ -214,8 +214,7 @@ def create_plots(
         os.mkdir(out_dir)
     dpp_num_div_values, dpp_div_model_values, dpp_ordered_model_values = get_dpp_prior_values(
                     config_path = dpp_config_path,
-                    # num_samples = 1000000)
-                    num_samples = 10000)
+                    num_samples = 1000000)
     dpp_hists = get_histograms(config_path = dpp_config_path,
             info_path = dpp_info_path,
             num_div_values = dpp_num_div_values,
@@ -247,7 +246,7 @@ def create_plots(
         hists.append(dpp_simple_hists[i])
 
     column_labels = [r'$\mathbf{M}_{msBayes}$', r'$\mathbf{M}_{Uniform}$', r'$\mathbf{M}_{DPP_{ }}$', r'$\mathbf{M}^{inform}_{DPP}$', r'$\mathbf{M}^{simple}_{DPP}$']
-    row_labels = ['Posterior', 'Prior', r'Prior $E(p(\mathbf{t}))$', r'Prior $E(p(\mathbf{t^{\circ}}))$']
+    row_labels = ['Posterior', 'Prior', r'Prior $\mathbb{E}(p(\mathbf{t}))$', r'Prior $\mathbb{E}(p(\mathbf{t^{\circ}}))$']
 
     pg = PlotGrid(subplots = hists,
             num_columns = 5,
@@ -303,7 +302,7 @@ def create_negros_panay_plots(
 
     hists = [unordered_hists[0]] + ordered_hists
 
-    row_labels = [r'Posterior $\mathbf{M}_{DPP}$', r'Posterior $\mathbf{M}^{\circ}_{DPP}$', 'Prior', r'Prior $E(p(\mathbf{t}))$', r'Prior $E(p(\mathbf{t^{\circ}}))$']
+    row_labels = [r'Posterior $\mathbb{M}_{DPP}$', r'Posterior $\mathbb{M}^{\circ}_{DPP}$', 'Prior', r'Prior $\mathbb{E}(p(\mathbf{t}))$', r'Prior $\mathbb{E}(p(\mathbf{t^{\circ}}))$']
 
     pg = PlotGrid(subplots = hists,
             num_columns = 1,

### 2013-12-25 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Adding empirical methods to DPP ms.

@@
 \subsection*{An empirical application}
-\ldots
+We also assessed the behavior of the newly implemented models when applied to
+the empirical dataset of \citet{Oaks2012} comprised of sequence data from 22
+pairs of taxa from the Philippine Islands (Drayad DOI: 10.5061/dryad.5s07m).
+We analyzed these data under five different models, which are detailed in
+Table~\ref{tabEmpiricalModels}.
+All of these models except one (\empModelDPPSimple) have six free demographic
+parameters per pair of taxa (
+\ancestralTheta{},
+\descendantTheta{1}{},
+\descendantTheta{2}{},
+\bottleTime{},
+\bottleScalar{1}{}, and
+\bottleScalar{2}{}),
+in addition to the $\popSampleSize{i}{} - 1$ coalescent times.
+Three of these models use a Dirichlet-process prior on divergence models:
+\empModelDPP, \empModelDPPInform, and \empModelDPPSimple.
+The \empModelDPP model is meant to represent the prior that \citet{Oaks2012}
+would have selected to represent their prior uncertainty about the parameters
+of the model if provided the more flexible distributions that are now
+implemented.
+To assess prior sensitivity, the \empModelDPPInform model uses a more
+informative exponentially distributed prior on divergence times, but otherwise
+is identical to \empModelDPP.
+To assess sensitivity to parameterization, we also applied the simplest
+possible model under the new implementation (\empModelDPPSimple) with only a
+single demographic parameter (\myTheta{}) per taxon pair, in addition to the
+$\popSampleSize{i}{} - 1$ coalescent times.
+We also applied the original \msBayes model (\empModelOld) with priors
+selected to make it comparable to the \empModelDPP model;
+the uniform prior on divergence times was selected to have the same variance as
+the exponential prior of the \empModelDPP, and the mean prior on population
+size was selected to have the same mean so that the models are on the same
+timescale.
+We also applied a model with a uniform distribution over divergence models
+(\empModelUniform).
+For each of these models, we simulated 2\e{7} samples from the prior, and
+retained an approximate posterior of the 10,000 samples with the smallest
+Euclidean distance from the summary statistics calculated from the empirical
+sequence alignments.
+
+To compare models that sample over ordered versus unordered models of
+divergence, we also analyzed the data from the subset of nine taxon pairs that
+are sampled from the Islands of Negros and Panay in the Philippines.
+The model we used for these analyses had a Dirichlet-process prior over
+divergence models and two demographic parameters (\ancestralTheta{} and
+\descendantTheta{}{}) for each pair of taxa, in addition to the
+$\popSampleSize{i}{} - 1$ coalescent times (see Table~\ref{tabEmpiricalModels}
+for details).
+One of the models, which we denote \npModelDPPOrdered, maintained the identity
+of the taxon pairs and sampled over ordered models of divergence, while the
+other (\npModelDPP) re-sorted the summary statistics of pair, losing the
+identity of the taxa and thus sampled over unordered models of divergence.
+For both analyses, we simulated 5\e{7} samples from the prior and retained an
+approximate posterior of 10,000 samples.
 

### 2013-12-25 empirical-analyses/bin/plot-results-comparison.py
Updating plotting script for empirical results.
Updating `plot-results-comparison.py` script that generates plots of the
empirical results for the dpp ms. Updating to create a plot of the
marginal divergence times for the Negros-Panay results under the ordered
model. Also, adding the resulting plot.

@@ -5,7 +5,7 @@ import sys
 
 import matplotlib
 
-from pymsbayes import teams
+from pymsbayes import teams, plotting
 from pymsbayes.utils.parsing import DMCSimulationResults, spreadsheet_iter
 from pymsbayes.config import MsBayesConfig
 from pymsbayes.plotting import (Ticks, HistData, ScatterPlot, PlotGrid)
@@ -330,6 +330,30 @@ def create_negros_panay_plots(
     pg.reset_figure()
     pg.savefig(os.path.join(out_dir, 'negros-panay-psi.pdf'))
 
+def create_time_plot(config_path,
+        info_path,
+        out_dir,
+        iteration_index = 249):
+    if not os.path.exists(out_dir):
+        os.mkdir(out_dir)
+    cfg = MsBayesConfig(config_path)
+    dmc = DMCSimulationResults(info_path)
+    sum_path = (dmc.get_result_path_prefix(1, 1, 1) + 
+            '{0}-posterior-summary.txt'.format(iteration_index))
+    labels = []
+    for t in cfg.taxa:
+        l = t.strip().split('.')
+        labels.append(r'\textit{{{0} {1}}}'.format(l[0], l[1]))
+    pg = plotting.get_marginal_divergence_time_plot(
+            config_path = config_path,
+            posterior_summary_path = sum_path,
+            labels = labels,
+            label_size = 12.0,
+            x_tick_label_size = 12.0,
+            x_label_size = 16.0,
+            y_label_size = 16.0)
+    pg.savefig(os.path.join(out_dir, 'negros-panay-marginal-times.pdf'))
+
 def main_cli():
     create_plots( 
             dpp_config_path = project_util.PHILIPPINES_DPP_CFG,
@@ -345,6 +369,10 @@ def main_cli():
             config_path = project_util.NEGROS_PANAY_CFG,
             ordered_info_path = project_util.NP_DPP_ORDERED_INFO,
             unordered_info_path = project_util.NP_DPP_UNORDERED_INFO,
+            out_dir = project_util.PLOT_DIR)
+    create_time_plot(
+            config_path = project_util.NEGROS_PANAY_CFG,
+            info_path = project_util.NP_DPP_ORDERED_INFO,
             out_dir = project_util.PLOT_DIR)
 
 if __name__ == '__main__':

### 2013-12-26 empirical-analyses/bin/plot-results-comparison.py
More work on plotting empirical results.
Working on script for plotting empirical results for DPP ms.
Updating marginal times plot.

@@ -348,10 +348,13 @@ def create_time_plot(config_path,
             config_path = config_path,
             posterior_summary_path = sum_path,
             labels = labels,
+            horizontal = True,
             label_size = 12.0,
-            x_tick_label_size = 12.0,
-            x_label_size = 16.0,
-            y_label_size = 16.0)
+            measure_tick_label_size = 12.0,
+            measure_axis_label = 'Divergence time',
+            measure_axis_label_size = 16.0,
+            label_axis_label = 'Taxon pair',
+            label_axis_label_size = 16.0)
     pg.savefig(os.path.join(out_dir, 'negros-panay-marginal-times.pdf'))
 
 def main_cli():

### 2013-12-26 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Working on DPP ms.
Finishing incorporation of the new figures. Also, updating some of the
new figures to add an extra digit for the exponential distributions of
divergence times.

@@
     please see Figure~S\ref{figPowerPsiOld} for all results.
-}{figPowerPsiOld4}
+}{figPowerOmegaOld4}
 
@@
     please see Figure~S\ref{figPowerPsiOld} for all results.
-}{figPowerPsiOld4}
+}{figPowerPsiProbOld4}
 
@@
     please see Figure~S\ref{figPowerPsiOld} for all results.
-}{figPowerPsiOld4}
+}{figPowerOmegaProbOld4}
+
+%% Empirical results
+\mFigure{../../empirical-analyses/plots/philippines-psi.pdf}{
+    The (A--E) posterior and (F--J) prior probabilities of the number of
+    divergence events (\divTimeNum) when the data of the 22 pairs of taxa from
+    the Philippines are analyzed under the five models indicated at the top of
+    column of plots (Table~\ref{tabEmpiricalModels}).
+    The average prior probability of a discrete (K--O) unordered and (P--T)
+    ordered model of divergence (\divTimeIndexVector) with \divTimeNum
+    divergence-time parameters is also shown.
+    The posterior median of the variance of divergence times
+    (\divTimeDispersion) is also given for each model, followed by the 95\%
+    highest posterior density interval in parantheses.
+}{figPhilippines}
+
+\widthFigure{0.4}{../../empirical-analyses/plots/negros-panay-psi.pdf}{listformat=figList}{
+    The posterior probabilities of the number of divergence events,
+    \divTimeNum, when the data of the 9 pairs of taxa from Negros and Panay
+    Islands are analyzed under the DPP model that samples over (A) unordered
+    and (B) ordered models of divergence (Table~\ref{tabEmpiricalModels}).
+    Both models share the same (C) prior probability of the number of
+    divergence events, and the average prior probability of a discrete (D)
+    unordered and (E) ordered model of divergence (\divTimeIndexVector) with
+    \divTimeNum divergence-time parameters.
+    The posterior median of the variance of divergence times
+    (\divTimeDispersion) is also given for each model, followed by the 95\%
+    highest posterior density interval in parantheses.
+}{figNegrosPanay}

### 2013-12-26 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex

@@
         assessing the power of the models in Table \ref{tabPriors}.
-        Distributions of divergence times are given in units of \globalcoalunit
-        generations and millions of generations ago (MGA), the latter
-        assuming a per-site rate of 1\e{-8} mutations per generation.}
+        The distributions of divergence times are given in units of
+        \globalcoalunit generations and millions of generations ago (MGA), the
+        latter assuming a per-site rate of 1\e{-8} mutations per generation.}
     \centering
@@
 \end{table}
+
+\begin{table}[htbp]
+    \sffamily
+    \footnotesize
+    \addtolength{\tabcolsep}{-0.08cm}
+    \rowcolors{2}{}{myGray}
+    %\captionsetup{font=footnotesize}
+    \caption{The models used to analyze the data from the 22 pairs of taxa from
+        the Philippines ($\mathbf{M}$), and a subset of nine of those pairs
+        from the Islands of Negros and Panay ($\mathbb{M}$).
+        In addition to the $\popSampleSize{}{} - 1$ coalescent times, the
+        \empModelDPPSimple has only a single \myTheta{} parameter for
+        each taxon pair.
+        The remaining $\mathbf{M}$ models have three \myTheta{}, two
+        \bottleScalar{}{}, and one \bottleTime{} parameters.
+        The distributions of divergence times are given in units of
+        \globalcoalunit generations and millions of generations ago (MGA), the
+        latter assuming a per-site rate of 1\e{-8} mutations per generation.
+        The \npModelDPP model (and its \npModelDPPOrdered counterpart
+        that samples over ordered divergence models) has only two
+        \myTheta{} parameters (the descendant populations of each
+        pair share the same \myTheta{} parameter, and there are no
+        bottleneck parameters).}
+    \centering
+    \begin{tabular}{ l l l l l l l }
+        \toprule
+        & \multicolumn{6}{c}{Priors} \\
+        \cmidrule(){2-7}
+        Model & \divTimeIndexVector & \divTime{} (\globalcoalunit gens.) & \divTime{} (MGA) & \ancestralTheta{} & \descendantTheta{}{} & \bottleScalar{}{} \\
+        \midrule
+            \empModelOld & $\divTimeIndexVector \sim \priorOld$
+                      & $\divTime{} \sim U(0,34.64)$
+                      & $\divTime{} \sim U(0,17.3)$
+                      & $\ancestralTheta{} \sim U(0, 0.01)$
+                      & $\descendantThetaMean{} \sim U(0, 0.01)$ 
+                      & $\bottleScalar{}{} \sim U(0, 1)$ \\
+            \empModelUniform & $\divTimeIndexVector \sim \priorUniform$
+                      & $\divTime{} \sim Exp(0.1)$
+                      & $\divTime{} \sim Exp(mean=5)$
+                      & $\ancestralTheta{} \sim Exp(200)$
+                      & $\descendantTheta{}{} \sim Exp(200)$
+                      & $\bottleScalar{}{} \sim Beta(5, 1)$ \\
+            \empModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
+                      & $\divTime{} \sim Exp(0.1)$
+                      & $\divTime{} \sim Exp(mean=5)$
+                      & $\ancestralTheta{} \sim Exp(200)$
+                      & $\descendantTheta{}{} \sim Exp(200)$
+                      & $\bottleScalar{}{} \sim Beta(5, 1)$ \\
+            \empModelDPPInform & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
+                      & $\divTime{} \sim Exp(1/6)$
+                      & $\divTime{} \sim Exp(mean=3)$
+                      & $\ancestralTheta{} \sim Exp(200)$
+                      & $\descendantTheta{}{} \sim Exp(200)$
+                      & $\bottleScalar{}{} \sim Beta(5, 1)$ \\
+            \empModelDPPSimple & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
+                      & $\divTime{} \sim Exp(0.1)$
+                      & $\divTime{} \sim Exp(mean=5)$
+                      & $\myTheta{} \sim Exp(200)$
+                      & -
+                      & - \\
+            \npModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,5.0)}$
+                      & $\divTime{} \sim Exp(0.1)$
+                      & $\divTime{} \sim Exp(mean=5)$
+                      & $\ancestralTheta{} \sim Exp(200)$
+                      & $\descendantTheta{}{} \sim Exp(200)$
+                      & - \\
+        \bottomrule
+    \end{tabular}
+    \label{tabEmpiricalModels}
+\end{table}

### 2013-12-26 manuscripts/reply/reply-figures.tex

@@
     \citet{Hickerson2013} when applied to simulated datasets in which divergence
-    times for 22 pairs of populations is drawn from an exponential
+    times of 22 pairs of populations are drawn from an exponential
     distribution, $\divt{} \sim Exp(2)$.
@@
     model-averaging approach of \citet{Hickerson2013} is applied to simulated
-    datasets in which divergence times for 22 pairs of populations is drawn
+    datasets in which divergence times of 22 pairs of populations are drawn
     from an exponential distribution, $\divt{} \sim Exp(2)$.
@@
     support the extreme model of one divergence when applied to simulated
-    datasets in which the divergence times are randomly drawn from the uniform
-    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
-    column of plots (divergence-time distributions are given in units of
-    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
-    mutations per generation).
+    datasets in which the divergence times of 22 pairs of populations are
+    randomly drawn from the uniform distributions $\divt{} \sim U(0,
+    \divt{max})$ indicated at the top of each column of plots (divergence-time
+    distributions are given in units of millions of generations ago (MGA)
+    assuming a per-site rate of 1\e{-8} mutations per generation).
     Four of the six \divt{max} we simulated are provided; please see
@@
     model-averaging approach of \citet{Hickerson2013} is applied to simlated
-    datasets in which divergence times are randomly drawn from the uniform
-    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
-    column of plots (divergence-time distributions are given in units of
-    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
-    mutations per generation).
+    datasets in which divergence times of 22 pairs of populations are randomly
+    drawn from the uniform distributions $\divt{} \sim U(0, \divt{max})$
+    indicated at the top of each column of plots (divergence-time distributions
+    are given in units of millions of generations ago (MGA) assuming a per-site
+    rate of 1\e{-8} mutations per generation).
     Four of the six \divt{max} we simulated are provided; please see

### 2013-12-26 manuscripts/reply/reply-main.tex

@@
 model selection
-\citep[e.g.,][]{Jeffreys1935,Lindley1957}.
+\citep[e.g.,][]{Jeffreys1939,Lindley1957}.
 Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
@@
 probability of producing the data and relatively high prior density will have
-smaller marginal likelihoods \citep{Jeffreys1935,Lindley1957}.
+smaller marginal likelihoods \citep{Jeffreys1939,Lindley1957}.
 Given the uniformly distributed priors on divergence times (and other nuisance

### 2013-12-26 manuscripts/reply/reply-si-figures.tex

@@
     informed model-averaging approach of \citet{Hickerson2013} is applied to
-    simlated datasets in which divergence times are randomly drawn from the
-    uniform distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top
-    of each column of plots (divergence-time distributions are given in units
-    of millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
-    mutations per generation).
+    simlated datasets in which divergence times of 22 pairs of populations are
+    randomly drawn from the uniform distributions $\divt{} \sim U(0,
+    \divt{max})$ indicated at the top of each column of plots (divergence-time
+    distributions are given in units of millions of generations ago (MGA)
+    assuming a per-site rate of 1\e{-8} mutations per generation).
 }{figPowerAccuracy}
@@
     support the extreme model of one divergence when applied to simulated
-    datasets in which the divergence times are randomly drawn from the uniform
-    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
-    column of plots (divergence-time distributions are given in units of
-    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
-    mutations per generation).
+    datasets in which the divergence times of 22 pairs of populations are
+    randomly drawn from the uniform distributions $\divt{} \sim U(0,
+    \divt{max})$ indicated at the top of each column of plots (divergence-time
+    distributions are given in units of millions of generations ago (MGA)
+    assuming a per-site rate of 1\e{-8} mutations per generation).
 }{figPower6}
@@
     model-averaging approach of \citet{Hickerson2013} is applied to simlated
-    datasets in which divergence times are randomly drawn from the uniform
-    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
-    column of plots (divergence-time distributions are given in units of
-    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
-    mutations per generation).
+    datasets in which divergence times of 22 pairs of populations are randomly
+    drawn from the uniform distributions $\divt{} \sim U(0, \divt{max})$
+    indicated at the top of each column of plots (divergence-time distributions
+    are given in units of millions of generations ago (MGA) assuming a per-site
+    rate of 1\e{-8} mutations per generation).
 }{figPowerExclusion6}

### 2013-12-26 manuscripts/reply/reply-si-main.tex

@@
 $\meanDescendantTheta{} \sim U(0.0005, 0.04)$.
-As described in \citet{Oaks2012}, the divergence-time parameters of the model
-implemented in \msb are scaled relative to a constant reference population
-size, \myTheta{C}.
-This reference population size is defined in terms of the upper limit of the
+As described in \citet{Oaks2012}, the divergence-time parameters in the model
+implemented in \msb are in generations scaled relative to a constant
+reference-population size, \myTheta{C}.
+This reference-population size is defined in terms of the upper limit of the
 uniform prior on the mean size of the descendant populations,
@@
 U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
-$\myTheta{C} = \uniformMax{\meanDescendantTheta{}}/2$.
+the size of the constant reference populations is $\myTheta{C} =
+\uniformMax{\meanDescendantTheta{}}/2$.
 Thus, the model used by \citet{Hickerson2013} mixes two different units of

### 2014-01-09 empirical-analyses/bin/plot-div-models.py
Updating plots of empirical results.
Adding time units (4N_C generations) to the time axis in the
divergence model and marginal time plots. This entailed changes
to the plotting script in the `empirical-analyses` project.

@@ -39,7 +39,7 @@ def get_div_model_plot_grid(
             plot_label_schema = 'uppercase',
             plot_label_offset = 0,
             plot_label_size = 12.0,
-            y_title = 'Divergence time',
+            y_title = 'Divergence time ($4N_C$ generations)',
             y_title_size = 14.0,
             y_tick_label_size = 10.0,
             right_text_size = 10.0,

### 2014-01-09 empirical-analyses/bin/plot-results-comparison.py

@@ -351,7 +351,7 @@ def create_time_plot(config_path,
             horizontal = True,
             label_size = 12.0,
             measure_tick_label_size = 12.0,
-            measure_axis_label = 'Divergence time',
+            measure_axis_label = 'Divergence time ($4N_C$ generations)',
             measure_axis_label_size = 16.0,
             label_axis_label = 'Taxon pair',
             label_axis_label_size = 16.0)

### 2014-01-09 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Adding divergence model plots to dpp ms.

@@
 $\popSampleSize{i}{} - 1$ coalescent times.
-We also applied the original \msBayes model (\empModelOld) with priors
+We also applied the original \msb model (\empModelOld) with priors
 selected to make it comparable to the \empModelDPP model;

### 2014-01-09 manuscripts/dpp-msbayes/dpp-msbayes-si-figures.tex
Working on DPP ms.
Finishing incorporation of the new figures. Also, updating some of the
new figures to add an extra digit for the exponential distributions of
divergence times.
Adding divergence model plots to dpp ms.

@@
     \timeConversionComment
-}{figPowerPsiOld}
+}{figPowerOmegaOld}
 
@@
     \timeConversionComment
-}{figPowerPsiUniform}
+}{figPowerOmegaUniform}
 
@@
     \timeConversionComment
-}{figPowerPsiExp}
+}{figPowerOmegaExp}
 
@@
     \timeConversionComment
-}{figPowerPsiOld}
+}{figPowerPsiProbOld}
 
@@
     \timeConversionComment
-}{figPowerPsiUniform}
+}{figPowerPsiProbUniform}
 
@@
     \timeConversionComment
-}{figPowerPsiExp}
+}{figPowerPsiProbExp}
 
@@
     \timeConversionComment
-}{figPowerPsiOld}
+}{figPowerOmegaProbOld}
 
@@
     \timeConversionComment
-}{figPowerPsiUniform}
+}{figPowerOmegaProbUniform}
 
@@
     \timeConversionComment
-}{figPowerPsiExp}
+}{figPowerOmegaProbExp}
+
+% Empirical results
+\siEightFigure{../../empirical-analyses/plots/philippines-dpp-div-models.pdf}{
+    The divergence-model results when the 22 pairs of taxa from the Philippines
+    are analyzed under the \empModelDPP model (Table~\ref{tabEmpiricalModels}).
+    The 10 unordered divergence models with highest posterior probability
+    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
+    indicate the inferred number of taxon pairs that diverged at each event.
+    The times indicate the posterior median and 95\% highest posterior density
+    (HPD) interval conditional on each divergence model.
+    For each model, times are summarized across posterior samples by the number
+    of taxon pairs associated with each divergence. For models in which there
+    are multiple divergence events with the same number of taxon pairs, the
+    events are sorted by time to summarize the divergence times in a consistant
+    way.
+}{figDivModelsDPP}
+
+\siEightFigure{../../empirical-analyses/plots/philippines-dpp-inform-div-models.pdf}{
+    The divergence-model results when the 22 pairs of taxa from the Philippines
+    are analyzed under the \empModelDPPInform model
+    (Table~\ref{tabEmpiricalModels}).
+    The 10 unordered divergence models with highest posterior probability
+    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
+    indicate the inferred number of taxon pairs that diverged at each event.
+    The times indicate the posterior median and 95\% highest posterior density
+    (HPD) interval conditional on each divergence model.
+    For each model, times are summarized across posterior samples by the number
+    of taxon pairs associated with each divergence. For models in which there
+    are multiple divergence events with the same number of taxon pairs, the
+    events are sorted by time to summarize the divergence times in a consistant
+    way.
+}{figDivModelsDPPInform}
+
+\siEightFigure{../../empirical-analyses/plots/philippines-dpp-simple-div-models.pdf}{
+    The divergence-model results when the 22 pairs of taxa from the Philippines
+    are analyzed under the \empModelDPPSimple model
+    (Table~\ref{tabEmpiricalModels}).
+    The 10 unordered divergence models with highest posterior probability
+    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
+    indicate the inferred number of taxon pairs that diverged at each event.
+    The times indicate the posterior median and 95\% highest posterior density
+    (HPD) interval conditional on each divergence model.
+    For each model, times are summarized across posterior samples by the number
+    of taxon pairs associated with each divergence. For models in which there
+    are multiple divergence events with the same number of taxon pairs, the
+    events are sorted by time to summarize the divergence times in a consistant
+    way.
+}{figDivModelsDPPSimple}
+
+\siEightFigure{../../empirical-analyses/plots/philippines-uniform-div-models.pdf}{
+    The divergence-model results when the 22 pairs of taxa from the Philippines
+    are analyzed under the \empModelUniform model
+    (Table~\ref{tabEmpiricalModels}).
+    The 10 unordered divergence models with highest posterior probability
+    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
+    indicate the inferred number of taxon pairs that diverged at each event.
+    The times indicate the posterior median and 95\% highest posterior density
+    (HPD) interval conditional on each divergence model.
+    For each model, times are summarized across posterior samples by the number
+    of taxon pairs associated with each divergence. For models in which there
+    are multiple divergence events with the same number of taxon pairs, the
+    events are sorted by time to summarize the divergence times in a consistant
+    way.
+}{figDivModelsUniform}
+
+\siEightFigure{../../empirical-analyses/plots/philippines-old-div-models.pdf}{
+    The divergence-model results when the 22 pairs of taxa from the Philippines
+    are analyzed under the \empModelOld model
+    (Table~\ref{tabEmpiricalModels}).
+    The 10 unordered divergence models with highest posterior probability
+    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
+    indicate the inferred number of taxon pairs that diverged at each event.
+    The times indicate the posterior median and 95\% highest posterior density
+    (HPD) interval conditional on each divergence model.
+    For each model, times are summarized across posterior samples by the number
+    of taxon pairs associated with each divergence. For models in which there
+    are multiple divergence events with the same number of taxon pairs, the
+    events are sorted by time to summarize the divergence times in a consistant
+    way.
+}{figDivModelsOld}
+
+\siEightFigure{../../empirical-analyses/plots/negros-panay-dpp-unordered-div-models.pdf}{
+    The divergence-model results when the 9 pairs of taxa from the Islands of
+    Negros and Panay are analyzed under the \npModelDPP model sampling over
+    unordered models of divergence.
+    (Table~\ref{tabEmpiricalModels}).
+    The 10 unordered divergence models with highest posterior probability
+    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
+    indicate the inferred number of taxon pairs that diverged at each event.
+    The times indicate the posterior median and 95\% highest posterior density
+    (HPD) interval conditional on each divergence model.
+    For each model, times are summarized across posterior samples by the number
+    of taxon pairs associated with each divergence. For models in which there
+    are multiple divergence events with the same number of taxon pairs, the
+    events are sorted by time to summarize the divergence times in a consistant
+    way.
+}{figDivModelsNP}
+
+\siEightFigure{../../empirical-analyses/plots/negros-panay-dpp-ordered-div-models.pdf}{
+    The divergence-model results when the 9 pairs of taxa from the Islands of
+    Negros and Panay are analyzed under the \npModelDPPOrdered model sampling
+    over ordered models of divergence.
+    (Table~\ref{tabEmpiricalModels}).
+    The posterior sample of divergence models were summarized while ignoring
+    the identity of the taxon pairs in order to compare the results of the
+    \npModelDPP model.
+    The 10 unordered divergence models with highest posterior probability
+    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
+    indicate the inferred number of taxon pairs that diverged at each event.
+    The times indicate the posterior median and 95\% highest posterior density
+    (HPD) interval conditional on each divergence model.
+    For each model, times are summarized across posterior samples by the number
+    of taxon pairs associated with each divergence. For models in which there
+    are multiple divergence events with the same number of taxon pairs, the
+    events are sorted by time to summarize the divergence times in a consistant
+    way.
+}{figDivModelsNPOrdered}
+
+\siFigure{../../empirical-analyses/plots/negros-panay-marginal-times.pdf}{
+    The marginal divergence-time results when the 9 pairs of taxa from the
+    Islands of Negros and Panay are analyzed under the \npModelDPPOrdered model
+    that samples over ordered models of divergence
+    (Table~\ref{tabEmpiricalModels}).
+}{figMarginalTimes}

### 2014-01-09 manuscripts/utils/macros.tex
Incorporating new figures into reply ms.
Incorporating updated figures (the same results presented in a
more concise manner). Also, some edits throughout the text.
Working on dpp ms.
In the process of integrating new figures into the text, and removing
old figures... A ways to go yet.
Adding plots of empirical results to dpp ms.
Also, adding table of the models used in the empirical analyses.
Adding divergence model plots to dpp ms.

@@
 \newcommand{\paramSampleMatrix}{\ensuremath{\boldsymbol{\paramSampleVector{}}}\xspace}
+\newcommand{\ordered}{\ensuremath{\circ}\xspace}
 \newcommand{\modelDPP}{\ensuremath{M_{DPP}}\xspace}
-\newcommand{\modelDPPOrdered}{\ensuremath{M^{\circ}_{DPP}}\xspace}
+\newcommand{\modelDPPOrdered}{\ensuremath{M^{\ordered}_{DPP}}\xspace}
 \newcommand{\modelUniform}{\ensuremath{M_{Uniform}}\xspace}
@@
 \newcommand{\powerSeriesExp}{\ensuremath{\mathcal{M}_{Exp}}\xspace}
+\newcommand{\empModelOld}{\ensuremath{\mathbf{M}_{msBayes}}\xspace}
+\newcommand{\empModelUniform}{\ensuremath{\mathbf{M}_{Uniform}}\xspace}
+\newcommand{\empModelDPP}{\ensuremath{\mathbf{M}_{DPP}}\xspace}
+\newcommand{\empModelDPPInform}{\ensuremath{\mathbf{M}^{inform}_{DPP}}\xspace}
+\newcommand{\empModelDPPSimple}{\ensuremath{\mathbf{M}^{simple}_{DPP}}\xspace}
+\newcommand{\npModelDPP}{\ensuremath{\mathbb{M}_{DPP}}\xspace}
+\newcommand{\npModelDPPOrdered}{\ensuremath{\mathbb{M}^{\ordered}_{DPP}}\xspace}
 
@@
 \newcommand{\bottleTimeVector}{\ensuremath{\boldsymbol{\bottleTime{}}}\xspace}
-\newcommand{\bottleTime}[1]{\ensuremath{\divt{B\ifArg{#1}{,}#1}}\xspace}
+\newcommand{\bottleTime}[1]{\ensuremath{\divt{B\protect\ifArg{#1}{,}#1}}\xspace}
 \newcommand{\bottleScalarVector}[1]{\ensuremath{\boldsymbol{\bottleScalar{#1}{}}}\xspace}
@@
     the #1 series of models.}
+\newcommand{\powerCommentSummary}[1]{The power of models
+    (A--D) \modelOld, (E--H) \modelUshaped, (I--L) \modelUniform, and (M--P)
+    \modelDPP to detect random variation in divergence times as simulated under
+    the #1 series of models.}
 \newcommand{\powerSupportComment}[1]{
@@
     \modelUniform, and (S--X) \modelDPP to support one divergence event when
+    there is random variation in divergence times as simulated under the #1
+    series of models.}
+\newcommand{\powerSupportCommentSummary}[1]{
+    The tendency of models (A--D) \modelOld, (E--H) \modelUshaped, (I--L)
+    \modelUniform, and (M--P) \modelDPP to support one divergence event when
     there is random variation in divergence times as simulated under the #1
@@
 
+\newcommand{\widthSidewaysFigure}[5]{\begin{sidewaysfigure}[htbp]
+\begin{center}
+    \includegraphics[width=#1\textwidth]{#2}
+    \captionsetup{#3}
+    \caption{#4}
+    \label{#5}
+    \end{center}
+    \end{sidewaysfigure}}
+
+\newcommand{\heightSidewaysFigure}[5]{\begin{sidewaysfigure}[htbp]
+\begin{center}
+    \includegraphics[height=#1]{#2}
+    \captionsetup{#3}
+    \caption{#4}
+    \label{#5}
+    \end{center}
+    \end{sidewaysfigure}}
+
 \newcommand{\mFigure}[3]{\widthFigure{1.0}{#1}{listformat=figList}{#2}{#3}\clearpage}
 \newcommand{\siFigure}[3]{\widthFigure{1.0}{#1}{name=Figure S, labelformat=noSpace, listformat=sFigList}{#2}{#3}\clearpage}
+\newcommand{\siEightFigure}[3]{\widthFigure{0.8}{#1}{name=Figure S, labelformat=noSpace, listformat=sFigList}{#2}{#3}\clearpage}
+\newcommand{\siSidewaysFigure}[3]{\widthSidewaysFigure{1.0}{#1}{name=Figure S, labelformat=noSpace, listformat=sFigList}{#2}{#3}\clearpage}

### 2014-01-10 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex
Playing with empirical model table of dpp ms.
Working on empirical model table for dpp ms.

--- 2013-12-19 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex
@@
 \end{table}
+
+\begin{table}[htbp]
+    \sffamily
+    \footnotesize
+    \addtolength{\tabcolsep}{-0.08cm}
+    \rowcolors{2}{myGray}{}
+    %\captionsetup{font=footnotesize}
+    \caption{The models used to analyze the data from the 22 pairs of taxa from
+        the Philippines ($\mathbf{M}$), and a subset of nine of those pairs
+        from the Islands of Negros and Panay ($\mathbb{M}$).
+        In addition to the $\popSampleSize{}{} - 1$ coalescent times, the
+        \empModelDPPSimple has only a single \myTheta{} parameter for
+        each taxon pair.
+        The remaining $\mathbf{M}$ models have three \myTheta{}, two
+        \bottleScalar{}{}, and one \bottleTime{} parameters.
+        The distributions of divergence times are given in units of
+        \globalcoalunit generations and millions of generations ago (MGA), the
+        latter assuming a per-site rate of 1\e{-8} mutations per generation.
+        The \npModelDPP model (and its \npModelDPPOrdered counterpart
+        that samples over ordered divergence models) has only two
+        \myTheta{} parameters (the descendant populations of each
+        pair share the same \myTheta{} parameter, and there are no
+        bottleneck parameters).}
+    \centering
+    \begin{tabulary}{\textwidth}{ l L }
+        \toprule
+        Model & Priors \\
+        \midrule
+            \empModelOld & $\divTimeIndexVector \sim \priorOld$ \tb
+                      $\divTime{} \sim U(0,34.64)$ \tb
+                      $\divTime{} \sim U(0,17.3 \; MGA)$ \tb
+                      $\ancestralTheta{} \sim U(0, 0.01)$ \tb
+                      $\descendantTheta{1}{}, \descendantTheta{2}{} \sim
+                            Beta(1, 1) \times 2 \times U(0, 0.01)$ \tb
+                      $\bottleScalar{1}{} \sim U(0, 1)$ \tb
+                      $\bottleScalar{2}{} \sim U(0, 1)$ \\[0.25em]
+            \empModelUniform & $\divTimeIndexVector \sim \priorUniform$ \tb
+                      $\divTime{} \sim Exp(0.1)$ \tb
+                      $\divTime{} \sim Exp(mean=5 \; MGA)$ \tb
+                      $\ancestralTheta{} \sim Exp(200)$ \tb
+                      $\descendantTheta{1}{} \sim Exp(200)$ \tb
+                      $\descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\bottleScalar{1}{} \sim Beta(5, 1)$ \tb
+                      $\bottleScalar{2}{} \sim Beta(5, 1)$ \\[0.25em]
+            \empModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ \tb
+                      $\divTime{} \sim Exp(0.1)$ \tb
+                      $\divTime{} \sim Exp(mean=5 \; MGA)$ \tb
+                      $\ancestralTheta{} \sim Exp(200)$ \tb
+                      $\descendantTheta{1}{} \sim Exp(200)$ \tb
+                      $\descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\bottleScalar{1}{} \sim Beta(5, 1)$ \tb
+                      $\bottleScalar{2}{} \sim Beta(5, 1)$ \tb \\
+            \empModelDPPInform & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ \tb
+                      $\divTime{} \sim Exp(1/6)$ \tb
+                      $\divTime{} \sim Exp(mean=3 \; MGA)$ \tb
+                      $\ancestralTheta{} \sim Exp(200)$ \tb
+                      $\descendantTheta{1}{} \sim Exp(200)$ \tb
+                      $\descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\bottleScalar{1}{} \sim Beta(5, 1)$ \tb
+                      $\bottleScalar{2}{} \sim Beta(5, 1)$ \tb \\
+            \empModelDPPSimple & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ \tb
+                      $\divTime{} \sim Exp(0.1)$ \tb
+                      $\divTime{} \sim Exp(mean=5 \; MGA)$ \tb
+                      $\ancestralTheta{} = \descendantTheta{1}{} = \descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\bottleScalar{1}{} = \bottleScalar{2}{} = 1.0$ \\[0.25em]
+            \npModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,5.0)}$ \tb
+                      $\divTime{} \sim Exp(0.1)$ \tb
+                      $\divTime{} \sim Exp(mean=5 \; MGA)$ \tb
+                      $\ancestralTheta{} \sim Exp(200)$ \tb
+                      $\descendantTheta{1}{} = \descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\bottleScalar{1}{} = \bottleScalar{2}{} = 1.0$ \\[0.25em]
+        \bottomrule
+    \end{tabulary}
+    \label{tabEmpiricalModels}
+\end{table}

### 2014-01-13 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex
Working on model tables for dpp manuscript.

@@
     \sffamily
-    % \footnotesize
+    \footnotesize
     \addtolength{\tabcolsep}{-0.08cm}
@@
         for the validation analyses and $Gamma(1.5,18.1)$ for the power
-        analyses.}
+        analyses. The distributions of divergence times are given in units of
+        \globalcoalunit generations followed in brackets by units of millions
+        of generations ago (MGA), with the former converted to the latter
+        assuming a per-site rate of 1\e{-8} mutations per generation. For model
+        \modelOld, the priors for theta parameters are $\ancestralTheta{} \sim
+        U(0, 0.05)$ and $\descendantTheta{1}{}, \descendantTheta{2}{} \sim
+        Beta(1, 1) \times 2 \times U(0, 0.05)$. The later is summarized as
+        $\descendantThetaMean{} \sim U(0, 0.05)$. For the \modelDPP and
+        \modelUniform, and \modelUshaped models, \ancestralTheta{},
+        \descendantTheta{1}{}, and \descendantTheta{2}{} are independently
+        and exponentially distributed with a mean of 0.025.}
     \centering
@@
         \cmidrule(){2-5}
-        Model & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
+        Model & \divTimeIndexVector & \divTime{} & \myTheta{} &  \\
         \midrule
             \modelOld & $\divTimeIndexVector \sim \priorOld$
-                      & $\divTime{} \sim U(0,10)$
+                      & $\divTime{} \sim U(0,10 \; [25 \; MGA])$
                       & $\ancestralTheta{} \sim U(0, 0.05)$
@@
             \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
-                      & $\divTime{} \sim Exp(0.3464)$
-                      & $\ancestralTheta{} \sim Exp(40)$
-                      & $\descendantTheta{}{} \sim Exp(40)$ \\
+                      & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
+                      & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
             \modelUniform & $\divTimeIndexVector \sim \priorUniform$
-                          & $\divTime{} \sim Exp(0.3464)$
-                          & $\ancestralTheta{} \sim Exp(40)$
-                          & $\descendantTheta{}{} \sim Exp(40)$ \\
+                          & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
+                          & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
             \modelUshaped & $\divTimeIndexVector \sim \priorOld$
-                          & $\divTime{} \sim Exp(0.3464)$
-                          & $\ancestralTheta{} \sim Exp(40)$
-                          & $\descendantTheta{}{} \sim Exp(40)$ \\
+                          & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
+                          & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
         \bottomrule
@@
     \footnotesize
+    % \scriptsize
     % \addtolength{\tabcolsep}{-0.08cm}
     \rowcolors{2}{}{myGray}
-    %\captionsetup{font=footnotesize}
+    % \captionsetup{font=footnotesize}
     \caption{The models we used to simulate pseudo-replicate datasets for
-        assessing the power of the models in Table \ref{tabPriors}.
-        The distributions of divergence times are given in units of
-        \globalcoalunit generations and millions of generations ago (MGA), the
-        latter assuming a per-site rate of 1\e{-8} mutations per generation.}
+        assessing the power of the models in Table \ref{tabPriors}.  The
+        distributions of divergence times are given in units of \globalcoalunit
+        generations followed in brackets by units of millions of generations
+        ago (MGA), with the former converted to the latter assuming a per-site
+        rate of 1\e{-8} mutations per generation. For all of the
+        \powerSeriesOld models, the priors for theta parameters are
+        $\ancestralTheta{} \sim U(0, 0.05)$ and $\descendantTheta{1}{},
+        \descendantTheta{2}{} \sim Beta(1, 1) \times 2 \times U(0, 0.05)$. The
+        later is summarized as $\descendantThetaMean{} \sim U(0, 0.05)$. For
+        the \powerSeriesUniform and \powerSeriesExp models, \ancestralTheta{},
+        \descendantTheta{1}{}, and \descendantTheta{2}{} are independently
+        and exponentially distributed with a mean of 0.025.}
     \centering
-    \begin{tabular}{ l l l l l l }
+    \begin{tabular}{ l l l l l }
         \toprule
-        & \multicolumn{5}{c}{Priors} \\
-        \cmidrule(){2-6}
-        Model series & \divTimeIndexVector & \divTime{} (\globalcoalunit gens.) & \divTime{} (MGA) & \ancestralTheta{} & \descendantTheta{}{} \\
+        & \multicolumn{4}{c}{Priors} \\
+        \cmidrule(){2-5}
+        Model series & \divTimeIndexVector & \divTime{} & \myTheta{} &  \\
         \midrule
             \powerSeriesOld & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.2)$
-                            & $\divTime{} \sim U(0,0.5)$
+                            & $\divTime{} \sim U(0,0.2 \; [0.5 \; MGA])$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.4)$
-                            & $\divTime{} \sim U(0,1.0)$
+                            & $\divTime{} \sim U(0,0.4 \; [1.0 \; MGA])$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.6)$
-                            & $\divTime{} \sim U(0,1.5)$
+                            & $\divTime{} \sim U(0,0.6 \; [1.5 \; MGA])$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.8)$
-                            & $\divTime{} \sim U(0,2.0)$
+                            & $\divTime{} \sim U(0,0.8 \; [2.0 \; MGA])$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,1.0)$
-                            & $\divTime{} \sim U(0,2.5)$
+                            & $\divTime{} \sim U(0,1.0 \; [2.5 \; MGA])$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,2.0)$
-                            & $\divTime{} \sim U(0,5.0)$
+                            & $\divTime{} \sim U(0,2.0 \; [5.0 \; MGA])$
                             & $\ancestralTheta{} \sim U(0, 0.05)$
@@
             \powerSeriesUniform & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.2)$
-                            & $\divTime{} \sim U(0,0.5)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim U(0,0.2 \; [0.5 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.4)$
-                            & $\divTime{} \sim U(0,1.0)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim U(0,0.4 \; [1.0 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.6)$
-                            & $\divTime{} \sim U(0,1.5)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim U(0,0.6 \; [1.5 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,0.8)$
-                            & $\divTime{} \sim U(0,2.0)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim U(0,0.8 \; [2.0 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,1.0)$
-                            & $\divTime{} \sim U(0,2.5)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim U(0,1.0 \; [2.5 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim U(0,2.0)$
-                            & $\divTime{} \sim U(0,5.0)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim U(0,2.0 \; [5.0 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
         \midrule
             \powerSeriesExp & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(17.32)$
-                            & $\divTime{} \sim Exp(mean = 0.14)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim Exp(mean = 0.058 \; [0.14 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(8.66)$
-                            & $\divTime{} \sim Exp(mean = 0.29)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim Exp(mean = 0.115 \; [0.29 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(5.77)$
-                            & $\divTime{} \sim Exp(mean = 0.43)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim Exp(mean = 0.173 \; [0.43 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(4.33)$
-                            & $\divTime{} \sim Exp(mean = 0.58)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim Exp(mean = 0.231 \; [0.58 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(3.46)$
-                            & $\divTime{} \sim Exp(mean = 0.72)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim Exp(mean = 0.289 \; [0.72 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                             & $\divTimeNum = 22$
-                            & $\divTime{} \sim Exp(1.73)$
-                            & $\divTime{} \sim Exp(mean = 1.44)$
-                            & $\ancestralTheta{} \sim Exp(40)$
-                            & $\descendantTheta{}{} \sim Exp(40)$ \\
+                            & $\divTime{} \sim Exp(mean = 0.577 \; [1.44 \; MGA])$
+                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
         \bottomrule
@@
             \empModelOld & $\divTimeIndexVector \sim \priorOld$ \tb
-                      $\divTime{} \sim U(0,34.64)$ \tb
-                      $\divTime{} \sim U(0,17.3 \; MGA)$ \tb
+        $\divTime{} \sim U(0,34.64 \, [17.3 \; MGA])$ \tb
                       $\ancestralTheta{} \sim U(0, 0.01)$ \tb
@@
             \empModelUniform & $\divTimeIndexVector \sim \priorUniform$ \tb
-                      $\divTime{} \sim Exp(0.1)$ \tb
-                      $\divTime{} \sim Exp(mean=5 \; MGA)$ \tb
-                      $\ancestralTheta{} \sim Exp(200)$ \tb
-                      $\descendantTheta{1}{} \sim Exp(200)$ \tb
-                      $\descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\divTime{} \sim Exp(mean=10 \; [5 \; MGA])$ \tb
+                      $\ancestralTheta{} \sim Exp(mean=0.005)$ \tb
+                      $\descendantTheta{1}{} \sim Exp(mean=0.005)$ \tb
+                      $\descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                       $\bottleScalar{1}{} \sim Beta(5, 1)$ \tb
@@
             \empModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ \tb
-                      $\divTime{} \sim Exp(0.1)$ \tb
-                      $\divTime{} \sim Exp(mean=5 \; MGA)$ \tb
-                      $\ancestralTheta{} \sim Exp(200)$ \tb
-                      $\descendantTheta{1}{} \sim Exp(200)$ \tb
-                      $\descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\divTime{} \sim Exp(mean=10 \; [5 \; MGA])$ \tb
+                      $\ancestralTheta{} \sim Exp(mean=0.005)$ \tb
+                      $\descendantTheta{1}{} \sim Exp(mean=0.005)$ \tb
+                      $\descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                       $\bottleScalar{1}{} \sim Beta(5, 1)$ \tb
@@
             \empModelDPPInform & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ \tb
-                      $\divTime{} \sim Exp(1/6)$ \tb
-                      $\divTime{} \sim Exp(mean=3 \; MGA)$ \tb
-                      $\ancestralTheta{} \sim Exp(200)$ \tb
-                      $\descendantTheta{1}{} \sim Exp(200)$ \tb
-                      $\descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\divTime{} \sim Exp(mean=6 \; [3 \; MGA])$ \tb
+                      $\ancestralTheta{} \sim Exp(mean=0.005)$ \tb
+                      $\descendantTheta{1}{} \sim Exp(mean=0.005)$ \tb
+                      $\descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                       $\bottleScalar{1}{} \sim Beta(5, 1)$ \tb
@@
             \empModelDPPSimple & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ \tb
-                      $\divTime{} \sim Exp(0.1)$ \tb
-                      $\divTime{} \sim Exp(mean=5 \; MGA)$ \tb
-                      $\ancestralTheta{} = \descendantTheta{1}{} = \descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\divTime{} \sim Exp(mean=10 \; [5 \; MGA])$ \tb
+                      $\ancestralTheta{} = \descendantTheta{1}{} = \descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                       $\bottleScalar{1}{} = \bottleScalar{2}{} = 1.0$ \\[0.25em]
             \npModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,5.0)}$ \tb
-                      $\divTime{} \sim Exp(0.1)$ \tb
-                      $\divTime{} \sim Exp(mean=5 \; MGA)$ \tb
-                      $\ancestralTheta{} \sim Exp(200)$ \tb
-                      $\descendantTheta{1}{} = \descendantTheta{2}{} \sim Exp(200)$ \tb
+                      $\divTime{} \sim Exp(mean=10 \; [5 \; MGA])$ \tb
+                      $\ancestralTheta{} \sim Exp(mean=0.005)$ \tb
+                      $\descendantTheta{1}{} = \descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                       $\bottleScalar{1}{} = \bottleScalar{2}{} = 1.0$ \\[0.25em]

### 2014-01-14 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex
More work on model tables for dpp manuscript.

@@
         The remaining $\mathbf{M}$ models have three \myTheta{}, two
-        \bottleScalar{}{}, and one \bottleTime{} parameters.
+        \bottleScalar{}{}, and one \bottleTime{} parameter.
         The distributions of divergence times are given in units of
-        \globalcoalunit generations and millions of generations ago (MGA), the
-        latter assuming a per-site rate of 1\e{-8} mutations per generation.
+        \globalcoalunit generations followed in brackets by units of millions
+        of generations ago (MGA), with the former converted to the latter
+        assuming a per-site rate of 1\e{-8} mutations per generation.
         The \npModelDPP model (and its \npModelDPPOrdered counterpart

### 2014-01-15 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on dpp manuscript.
Proofreading and incorporating edits throughout the manuscript.

@@
 Thus, being able to accurately estimate models of shared evolutionary
-histories across co-distributed taxa would provide an important
+history across co-distributed taxa would provide an important
 tool for better understanding how regional and global biogeographic
@@
 combination of broad, uniformly distributed priors on nuisance parameters and a
-U-shaped prior on divergence models are causing the bias.
+``U-shaped'' prior on divergence models are causing the bias.
 The model implemented in \msb uses a discrete uniform prior over the number of
@@
 divergence models that puts most of the prior mass on models with either very
-few or very many divergence times shared across the taxa.
-Furthermore, uniform priors used for most of the model's nuisance parameters
-likely hinder the marginal likelihoods of parameter-rich models, i.e., models
-with many divergence time parameters \citep{Lindley1957}.
-Thus, together, the priors on divergence models and nuisance parameters used in
-\msb could create a strong bias toward models with very few divergence time
-parameters \citep{Oaks2012}.
+few or very many divergence times shared across the taxa (see Figure 5 of
+\citet{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
+Furthermore, uniform priors used for most of the model's parameters
+result in small marginal likelihoods of parameter-rich models, i.e., models
+with many divergence-time parameters \citep{Jeffreys1935, Lindley1957}.
+Thus, together, the priors used in \msb will often create a bias toward
+models with few divergence-time parameters \citep{Oaks2012}.
 
@@
 of the approximate-Bayesian model-choice approach.
-We implement a Dirichlet-process prior over all possible models of divergence,
-and use gamma and beta probability distributions in place of uniform priors on
-many of the model's parameters.
+We implement a Dirichlet-process prior (DPP) over all possible models of
+divergence, and use gamma and beta probability distributions in place of
+uniform priors on many of the model's parameters.
 Our simulation-based analyses show that the new implementation has improved
 robustness, accuracy, and power compared to the original model.
-Also, our results show that both the uniform priors on nuisance parameters and
-the U-shaped prior on divergence models contribute to the biases of \msb.
+Also, our results show that both the uniform priors on demographic and
+divergence-time parameters and the U-shaped prior on divergence models
+contribute to the biases of \msb.
+When the Oaks et al.'s \citeyear{Oaks2012} sequence data of 22 pairs of taxa
+from the Philippines is reanalyzed under the new model, we find a large amount
+of posterior uncertainty in the number of divergence events shared among the
+taxa; a result that congruent with intuition given the richness of the model
+and the relatively small amount of information in the data.
 
@@
 prior on \locusMutationRateScalar{}, and
-$p(\alignmentVector)$ is the probability of the data (or the marginal
-likelihood of the model).
+$p(\alignmentVector \given \hkyModelVector, \ploidyScalarVector,
+\mutationRateScalarConstantVector)$
+is the probability of the data (or the marginal likelihood of the model) given
+the fixed constants provided by the investigator.
 \end{linenomath}
@@
 and gives rise to the two sampled descendant populations.
-Each \demographicParams{} contains the following demographic parameters:
-The mutation-scaled sizes of the ancestral, \ancestralTheta{}, and descendant
-populations, \descendantTheta{1}{} and \descendantTheta{2}{};
+Each \demographicParams{} contains the following demographic parameters: The
+mutation-scaled effective sizes ($\myTheta{} = 4N\mutationRate$) of the
+ancestral, \ancestralTheta{}, and descendant populations, \descendantTheta{1}{}
+and \descendantTheta{2}{};
 the proportion of the first, \bottleScalar{1}{}, and second population,
@@
 \end{equation}
-where \divTimeVector is an ordered set of unique divergence time parameters
+where \divTimeVector is an ordered set of unique divergence-time parameters
 $\{\divTime{1}, \ldots, \divTime{\divTimeNum}\}$ whose length
 \divTimeNum can range from 1 to \npairs{},
-and \divTimeIndexVector is the vector of indices of divergence time parameters 
+and \divTimeIndexVector is the vector of indices of divergence-time parameters 
 $(\divTimeIndex{1}, \ldots, \divTimeIndex{\npairs{}})$
@@
 Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
-Thus, given the number of unique divergence time classes in
+Thus, given the number of unique divergence-time classes in
 \divTimeIndexVector, this determines the probability of prior term
 $p(\divTimeVector \given \divTimeIndexVector)$.
-The divergence time parameters are in coalescent units relative to the size of
+The divergence times are in coalescent units relative to the size of
 a constant reference population, which we denote \myTheta{C}, that is equal to
@@
 \begin{equation}
-    \myTheta{C} = \mathbb{E}(\descendantTheta{}{}).
+    \myTheta{C} = \mathbb{E}(\dThetas),
     \label{eq:thetaC}
 \end{equation}
+where \popIndex is the descendant-population index which can be either 1 or 2.
 Given the size of the descendant populations are \iid as
-$\descendantTheta{}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+$\dThetas \sim Gamma(\gammaShape{\descendantTheta{}{}},
 \gammaScale{\descendantTheta{}{}})$,
@@
 \begin{equation}
-    \divTime{G_C} = \divTime \times \frac{
+    \divTime{G_C} = \divTime \times \frac{\myTheta{C}}{\mutationRate} = \divTime
+    \times \frac{
     \gammaShape{\descendantTheta{}{}}\gammaScale{\descendantTheta{}{}}}
@@
 where \gammaShape{\concentrationParam} and \gammaScale{\concentrationParam} are
-specified by the user.
+specified by the investigator.
 
@@
 \end{equation}
-and also the prior probability of the number of divergence time parameters
+and also the prior probability of the number of divergence-time parameters
 \begin{equation}
@@
 Equations \ref{eq:dppPriorSameClass} and \ref{eq:dppPriorNumClasses} show that
-smaller values of \concentrationParam will favor fewer divergence time
+smaller values of \concentrationParam will favor fewer divergence-time
 parameters, and thus more clustered models of divergence, whereas larger values
-will favor more divergence time parameters, and thus less clustered models of
+will favor more divergence-time parameters, and thus less clustered models of
 divergence.
@@
 (2) a Dirichlet-process prior on all ordered divergence models.
-The latter provides the user with a great deal of flexibility in
+The latter provides an investigator with a great deal of flexibility in
 expressing their prior beliefs about models of divergence.
@@
 \bottleScalar{2}{}).
-These priors provide the user with much greater flexibility in expressing prior
-uncertainty regarding the parameters of the model.
+These priors provide an investigator with much greater flexibility in
+expressing prior uncertainty regarding the parameters of the model.
 
@@
 This has the advantage of reducing the sample space of the number of discrete
-divergence models \divTimeMapVector, but there are at least two disadvantages.
+divergence models \divTimeIndexVector, but there are at least two disadvantages.
 First, additional information in the data is lost.
@@
 each pair and are thus exchangeable \citep{Hickerson2006,Huang2011}.
-This is actually incorrect for two of reasons.
+This is actually incorrect for two reasons.
 First, the expectation of $\pi_b$ is not independent of samples sizes.
@@
 are retained.
-The range of tolerance is determined by specifying the number of desired
+The range of tolerance is determined by specifying the desired number of
 posterior samples to be retained.
@@
 $\descendantThetaMean{} \sim U(0, 0.05)$;
-(2) a Dirichlet-process prior model, \modelDPP, with priors
-$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
+(2) a model with the \msb prior on unordered divergence models, but with
+exponential priors on nuisance parameters,
+\modelUshaped, with priors
+$\divTimeIndexVector \sim \priorOld$,
+$\divTime{} \sim Exp(mean=2.887)$,
+$\ancestralTheta{} \sim Exp(mean=0.025)$,
 and
-$\descendantTheta{}{} \sim Exp(40)$;
+$\dThetas \sim Exp(mean=0.025)$.
 (3) a model with uniform prior probability over unordered divergence models,
@@
 $\divTimeIndexVector \sim \priorUniform$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
+$\divTime{} \sim Exp(mean=2.887)$,
+$\ancestralTheta{} \sim Exp(mean=0.025)$,
 and
-$\descendantTheta{}{} \sim Exp(40)$;
+$\dThetas \sim Exp(mean=0.025)$;
 and
-(4) a model with the \msb prior on unordered divergence models, but with
-exponential priors on nuisance parameters,
-\modelUshaped, with priors
-$\divTimeIndexVector \sim \priorOld$,
-$\divTime{} \sim Exp(0.3464)$,
-$\ancestralTheta{} \sim Exp(40)$,
+(4) a Dirichlet-process prior model, \modelDPP, with priors
+$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
+$\divTime{} \sim Exp(mean=2.887)$,
+$\ancestralTheta{} \sim Exp(mean=0.025)$,
 and
-$\descendantTheta{}{} \sim Exp(40)$.
+$\dThetas \sim Exp(mean=0.025)$.
 We selected the exponential prior on divergence time used in models \modelDPP,
@@
 
-For each of the four models, we generated $1\e6$ samples from the prior, and
-simulated 50,000 pseudo-replicate datasets, also drawn from the prior.
-We then analyzed each of the replicate datasets, retaining a
+For each of the four models, we simulated $1\e6$ samples from the prior and
+50,000 datasets, also drawn from the prior.
+We then analyzed each of the simulated datasets, retaining a
 posterior of 1000 samples from the respective prior.
@@
 To assess the robustness of each of the four models, we also analyzed the
-pseudo-replicate datasets simulated under the other three models.
+datasets simulated under the other three models.
 Overall, for each model, we produced 200,000 posterior estimates,
-50,000 from the replicate datasets simulated under that model,
-and 150,000 estimated from the replicate datasets simulated under the
+50,000 from the datasets simulated under that model,
+and 150,000 estimated from the datasets simulated under the
 other three models.
@@
 We denote this model as \modelDPPOrdered.
-We generated $1\e6$ prior samples and 50,000 pseudo-replicate datasets, and
+We simulated $1\e6$ prior samples and 50,000 datasets, and
 analyzed them as above.
-We were not able to analyze the replicate datasets of the other models under
+We were not able to analyze the simulated datasets of the other models under
 the ordered model, because the identity of the population pairs is not
@@
 For each of the four models, we generated $2\e6$ samples from the prior.
-We then generated pseudo-observed datasets from three series of models.
+We then simulated datasets from three series of models
+(Table~\ref{tabPowerModels}).
 One series of models, which we denote \powerSeriesOld, is identically
@@
 2.0, in \globalcoalunit generations.
-A second series of models, \powerSeriesUniform, are identically distributed as
-\modelUniform except that the 22 divergence times are randomly drawn from the
-same series of uniform priors as above.
-The third series of models, \powerSeriesExp, is also identically
-distributed as \modelUniform except that the 22 divergence times are randomly
-drawn from a series of of exponential distributions: $Exp(17.3205)$,
-$Exp(8.6603)$, $Exp(5.7735)$, $Exp(4.3301)$, $Exp(3.4641)$, and $Exp(1.7321)$.
+A second series of models, \powerSeriesUniform, is identically distributed as
+\modelUniform and \modelDPP except that the 22 divergence times are randomly
+drawn from the same series of uniform priors as above.
+The third series of models, \powerSeriesExp, is also identically distributed as
+\modelUniform and \modelDPP except that the 22 divergence times are randomly
+drawn from a series of of exponential distributions:
+$Exp(mean=0.058)$, $Exp(mean=0.115)$, $Exp(mean=0.173)$, $Exp(mean=0.231)$,
+$Exp(mean=0.289)$, and $Exp(mean=0.577)$.
 These exponential distributions have the same variance as their uniform
@@
 For each of the six models in each of the three series of models, we simulated
-1000 pseudo-replicate datasets (18,000 datasets in total).
-We then analyzed each replicate dataset under all four prior models
+1000 datasets (18,000 datasets in total).
+We then analyzed each simulated dataset under all four prior models
 (Table~\ref{tabPriors}), producing 72,000 posterior estimates, each with 1000
@@
 the uniform prior on divergence times was selected to have the same variance as
-the exponential prior of the \empModelDPP, and the mean prior on population
+the exponential prior of the \empModelDPP, and the prior on population
 size was selected to have the same mean so that the models are on the same
@@
 To compare models that sample over ordered versus unordered models of
-divergence, we also analyzed the data from the subset of nine taxon pairs that
+divergence, we also analyzed the data from the subset of nine-taxon pairs that
 are sampled from the Islands of Negros and Panay in the Philippines.
@@
 of the taxon pairs and sampled over ordered models of divergence, while the
-other (\npModelDPP) re-sorted the summary statistics of pair, losing the
-identity of the taxa and thus sampled over unordered models of divergence.
+other (\npModelDPP) re-sorted the summary statistics of the pairs by $\pi_b$,
+losing the identity of the taxa and thus sampled over unordered models of
+divergence.
 For both analyses, we simulated 5\e{7} samples from the prior and retained an
@@
 (Figure~S\ref{figValidationAccuracyOmega}).
-This results are similar for the GLM-adjusted estimates of \divTimeDispersion,
+Results are similar for the GLM-adjusted estimates of \divTimeDispersion,
 albeit the regression adjustment tends to improve estimates of this continuous
@@
 
-The same general pattern is seen for estimates of \divTimeMean, with
-the all four models performing similarly when applied to the data generated
-under \modelOld, the models with exponentially distributed priors performing
-similarly when applied to data generated under other models, and the
-\modelOld is more biased when applied to data generated under other models
+The same general pattern is seen for estimates of \divTimeMean, with all four
+models performing similarly when applied to the data generated under the
+\modelOld model, the models with exponentially distributed priors performing
+similarly when applied to data generated under the other three models, and the
+\modelOld model is sensitive to model violations and is more biased whenever
+applied to data generated under other models
 (Figure~S\ref{figValidationAccuracyTime}).
@@
 generally more accurate than regression-adjusted estimates
-(Figure~\ref{figValidationModelChoicePsi} versus
+(compare the plots along the upper-left to lower-right diagonal for
+Figure~\ref{figValidationModelChoicePsi} versus
 S\ref{figValidationModelChoicePsiGlm} and
@@
 \ref{figValidationModelChoiceOmega}).
-In contrast, the original is biased toward overestimating the posterior
+In contrast, the \modelOld model is biased toward overestimating the posterior
 probability of the one-divergence model when applied to data generated under
@@
 best estimator of \divTimeDispersion, followed by \modelUniform
-(Figure~\ref{figPowerAccuracyExp}).
+(Figure~S\ref{figPowerAccuracyExp}).
 The \modelOld model is strongly biased toward underestimating
 \divTimeDispersion, estimating values of zero for most of the replicates across
-all the data models of \powerSeriesExp (Figure \ref{figPowerAccuracyExp}).
+all the data models of \powerSeriesExp (Figure~S\ref{figPowerAccuracyExp}).
 The results of the \modelUshaped model are intermediate between those of
 \modelOld and the new models \modelDPP and \modelUniform
-(Figure \ref{figPowerAccuracyExp}).
+(Figure~S\ref{figPowerAccuracyExp}).
 
@@
 tend to over-estimate the variance in divergence times, whereas the \modelOld
-model, underestimates \divTimeDispersion, estimating values of zero for most
+model underestimates \divTimeDispersion, estimating values of zero for most
 replicates across all the data models of \powerSeriesUniform
-(Figure~\ref{figPowerAccuracyUniform}).
+(Figure~S\ref{figPowerAccuracyUniform}).
 Again, the performance of the \modelUshaped model is intermediate between the
-\modelOld and \modelDPP/\modelUniform models (Figure
-\ref{figPowerAccuracyUniform}).
+\modelOld and \modelDPP/\modelUniform models (Figure~S\ref{figPowerAccuracyUniform}).
 The results are very similar when the four models are applied to the data
-simulated under the \powerSeriesOld seires of models
-(Figure~\ref{figPowerAccuracyOld}).
+simulated under the \powerSeriesOld series of models
+(Figure~S\ref{figPowerAccuracyOld}).
 
@@
 divergences when applied to data generated under random divergence times
-(Figure~\ref{figPowerPsiOld4}A--D \& plots A--F of Figures
+(Figure~\ref{figPowerPsiOld4} \& 
 S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
@@
 (Figure~\ref{figPowerPsiProbOld4}A--D \& plots A--F of Figures
-S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+S\labelcref{figPowerPsiProbOld,figPowerPsiProbUniform,figPowerPsiProbExp}).
 The \modelDPP model also prefers the one-divergence model 
@@
 (Figure~\ref{figPowerPsiProbOld4}M--P \& plots S--X of Figures
-S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+S\labelcref{figPowerPsiProbOld,figPowerPsiProbUniform,figPowerPsiProbExp}).
 % Even when the simulated data are identically distributed as the \modelOld model
@@
 
-The improved power of the new models are even more pronounced when looking at
+The improved power of the new models is even more pronounced when looking at
 estimates of the variance of divergence times (\divTimeDispersion) across the
@@
 \subsection*{Empirical results}
-\ldots
+As expected based on the results of \citet{Oaks2012}, when the Philippines data
+are analyzed under the \empModelOld model, there is strong support for very few
+divergence events shared among all 22 pairs of taxa
+(Figure~\ref{figPhilippines}A).
+When these data are analyzed under models allowed under the new implementation,
+there is much less support for highly clustered models and much greater
+uncertainty regarding the number of divergence events shared among the taxa,
+especially under the DPP models (Figure~\ref{figPhilippines}B--E).
+Figure \ref{figPhilippines} also shows the prior distribution across the number
+of divergence events (\divTimeNum) for each of the model, as well as the
+average prior probability of an unordered and ordered model of divergence
+(\divTimeIndexVector) across \divTimeNum for each model.
+Estimates under the new models tend to be similar to the prior, which is
+expected under such a parameter-rich model when there is limited information
+from the data (four summary statistics from a single locus for each pair of
+taxa).
+
+The disparity between the results of the \empModelOld model and the new models
+is even more pronounced when we look at the 10 divergence models
+(\divTimeMapVector) estimated to have the highest probability under each of the
+models
+(Figures~S\labelcref{figDivModelsDPP,figDivModelsDPPInform,figDivModelsDPPSimple,figDivModelsUniform,figDivModelsOld}).
+Again, the new models estimate more divergences, a large amount of posterior
+uncertainty, and an order of magnitude smaller probability for the top model
+when compared to the \empModelOld model.
+
+Figure \ref{figNegrosPanay} shows the estimated posterior probability
+distribution over the number of divergence events when the data from the
+nine-taxon pairs of the Islands of Negros and Panay are analyzed under the DPP
+models that sample over unordered (\npModelDPP) and ordered
+(\npModelDPPOrdered) models of divergence.
+The result are similar under both models and, again, yield a large amount of
+uncertainty about the number of divergence events that is similar to the prior
+uncertainty.
+Both models estimate a similar set of 10 unordered divergence models with the
+highest posterior probability
+(Figures~S\labelcref{figDivModelsNP,figDivModelsNPOrdered}).
+
+The main advantage of the \npModelDPPOrdered over the \npModelDPP is that the
+identity of the taxa is maintained, and thus a fully marginalized estimate of
+divergence times across the taxa can be obtained
+(Figure~S\ref{figMarginalTimes}).
+Furthermore, the probability of co-divergence among any set of taxa can be
+estimated from the posterior sample.
 
@@
 phylogeograpical models in a full-likelihood Bayesian framework is becoming
-computationally practical, especially when considering that the computational
-load of simulating millions of random datasets from the prior under the ABC
-approach is nontrivial.
+computationally practical, especially when considering that simulating millions
+of random datasets from the prior under the simple ABC rejection approach is
+inefficient and computationally nontrivial.

### 2014-01-15 manuscripts/dpp-msbayes/dpp-msbayes.tex
Working on dpp manuscript.
Proofreading and incorporating edits throughout the manuscript.

@@
     Such events predict temporally clustered divergences times, a pattern
-    that can be estimated from genetic data from co-distributed species.
+    that can be estimated using genetic data from co-distributed species.
     We introduce a new approximate-Bayesian model for comparative
@@
     divergences across taxa from multi-locus DNA sequence data.
-    Our model is an extension of the model implemented in \msb.
-    By reparameterizing the model, and using more flexible priors on divergence
-    models and nuisance parameters, we improve the robustness, accuracy, and power
-    of the method for estimating the posterior probabilities of models
-    of divergence across taxa.
-    Our results demonstrate that the bias of the \msb model toward inferring
-    models of clustered divergences is caused by a combination of (1) uniform
-    priors on nuisance parameters hindering the marginal likelihoods of models
-    with more divergence time parameters, and (2) a prior on divergence models
-    that disfavors models with intermediate numbers of divergence time
+    Our model is an extension of that implemented in \msb.
+    By reparameterizing the model, introducing more flexible priors on
+    demographic and divergence-time parameters, and implementing a
+    non-parametric Dirichlet-process prior over divergence models, we improve
+    the robustness, accuracy, and power of the method for estimating shared
+    evolutionary history across taxa.
+    Our results demonstrate that the tendency of the \msb model to spuriously
+    support models of clustered divergences is caused by a combination of (1)
+    uniform priors on nuisance parameters hindering the marginal likelihoods of
+    models with more divergence-time parameters, and (2) a prior on divergence
+    models that disfavors models with intermediate numbers of divergence-time
     parameters.
+    The new method addresses both of these issues and yields more robust
+    estimates of posterior uncertainty, and thus greatly reduces the tendency
+    of the model to incorrectly estimate biogeographically interesting models
+    with strong support.
 

### 2014-01-15 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More work on main text of dpp manuscript.

@@
 of posterior uncertainty in the number of divergence events shared among the
-taxa; a result that congruent with intuition given the richness of the model
+taxa; a result congruent with intuition given the richness of the model
 and the relatively small amount of information in the data.
@@
 
-Overall, the result suggest that the \modelDPP and \modelUniform models are
+Overall, the results suggest that the \modelDPP and \modelUniform models are
 relatively robust in terms of model-choice accuracy, and when model violations
@@
 S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
-Using exponential priors on divergence-time and theta parameters does increase
-the power of the \modelUshaped model compared to \modelOld across all three
-series of data models, but the U-shaped prior still prevents the model from
-performing as well as the \modelDPP and \modelUniform models 
+Using exponential priors on divergence-time and \myTheta{} parameters does
+increase the power of the \modelUshaped model compared to \modelOld across all
+three series of data models, but the U-shaped prior still prevents the model
+from performing as well as the \modelDPP and \modelUniform models 
 (Figure~\ref{figPowerPsiOld4} \& 
@@
 replicates across all the simulations.
-In contrast, the \modelOld model infers values consistent with a single divergence event in
-most of the replicates across all the simulations.
-Using exponential priors on divergence-time and theta parameters greatly
+In contrast, the \modelOld model infers values consistent with a single
+divergence event in most of the replicates across all the simulations.
+Using exponential priors on divergence-time and \myTheta{} parameters greatly
 increases the power of the \modelUshaped model to detect variation in
@@
 S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).
+An estimated value of \divTimeDispersion less than 0.01 is an arbitrary
+threshold for ``simultaneous'' divergence used throughout the \msb
+literature.
+Albeit arbitrary, \cite{Oaks2012} did show via simulation that the true value
+of \divTimeDispersion will almost always be greater than 0.01 when divergences
+are random over periods of 0.1 \globalcoalunit generations or more (see
+Figure~S4 of \citet{Oaks2012}).
 
@@
 S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
-The exponential priors on divergence-time and theta parameters (model
+The exponential priors on divergence-time and \myTheta{} parameters (model
 \modelUshaped) result in lower estimates of the probability of one divergence
@@
 S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
-The \modelDPP and \modelUniform models do frequently support the one-model
-according to a Bayes factor criterion of greater than 10, however still less
-frequently than the \modelOld model.
+The \modelDPP and \modelUniform models do frequently support the one-divergence
+model according to a Bayes factor criterion of greater than 10, however still
+less frequently than the \modelOld model.
 This result is not surprising given the extremely small prior probability of
@@
 are analyzed under the \empModelOld model, there is strong support for very few
-divergence events shared among all 22 pairs of taxa
+divergence events shared among all 22 pairs of taxa with a maximum \emph{a
+posteriori} (MAP) estimate of one-shared divergence
 (Figure~\ref{figPhilippines}A).
@@
 is even more pronounced when we look at the 10 divergence models
-(\divTimeMapVector) estimated to have the highest probability under each of the
+(\divTimeIndexVector) estimated to have the highest probability under each of the
 models
@@
 Again, the new models estimate more divergences, a large amount of posterior
-uncertainty, and an order of magnitude smaller probability for the top model
-when compared to the \empModelOld model.
+uncertainty, and an order of magnitude smaller probability for their respective
+MAP model when compared to the \empModelOld model.
 
@@
 distribution over the number of divergence events when the data from the
-nine-taxon pairs of the Islands of Negros and Panay are analyzed under the DPP
+nine-taxon pairs from the Islands of Negros and Panay are analyzed under the DPP
 models that sample over unordered (\npModelDPP) and ordered
@@
 The original model is very sensitive to model violations and, when present,
-strongly over-estimates the probability of one divergence event shared
+strongly over-estimates the probability of one-divergence event shared
 across all taxa
@@
 \ref{figValidationModelChoiceOmega}).
-When more appropriate priors are used for divergence-time and theta parameters,
-and either a Dirichlet-process or uniform prior applied across divergence
-models, the model is less sensitive to violations, and, when they do cause
-bias, the method is conservative and tends to under-estimate the probability of
-biogeographically interesting models of temporally clustered divergences
-(Figure~\ref{figValidationModelChoicePsi} \&
+When more appropriate priors are used for divergence-time and demographic
+parameters, and either a Dirichlet-process or uniform prior applied across
+divergence models, the model is less sensitive to violations, and, when they do
+cause bias, the method is conservative and tends to underestimate the
+probability of biogeographically interesting models of temporally clustered
+divergences (Figure~\ref{figValidationModelChoicePsi} \&
 \ref{figValidationModelChoiceOmega}).
@@
 Furthermore, the modifications to the model also improve the method's power to
-detect random variation in divergence times, reducing the bias toward
-estimating clustered divergences
-(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}.
+detect random variation in divergence times, reducing the tendency to estimate
+clustered divergences
+(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
 Our results are similar to those of \citet{Oaks2012} in that we also find
@@
 divergences are random over quite broad timescales
-(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}.
+(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
 Our results expand on this by showing that this behavior is consistent
 across a range of conditions underlying the data.
-When using the new priors, the method is less prone to spurious inference of
-clustered divergence models, and much less likely to incorrectly infer such
-models with strong support.
+The new method \dppmsbayes has greater power to detect random temporal variation in
+divergences, is less prone to spurious inference of clustered divergence
+models, and much less likely to incorrectly infer such models with strong
+support
+(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
 
 By evaluating a model intermediate between the old and new implementation
-(\modelUshaped), we were able to determine the relative effects of the
+(\modelUshaped), we were able to determine the relative effects of our
 modifications to the model.
 Across all of our analyses, the results show that using better priors on
-divergence-time and nuisance parameters alone does improve the performance of
-the method.
+divergence-time and demographic parameters alone does improve the performance
+of the method.
 The magnitude of the bias toward inferring support for the one-divergence
@@
 \ref{figValidationModelChoiceOmega}).
-Furthermore, better priors on nuisance parameters improves the method's
-power to detect temporally random divergences.
-Throughout our analyses, the intermediate model (\modelUshaped) performs
-better than the \msb model, but not as well as the models with alternative
-priors on divergence models.
+Furthermore, better priors on parameters improves the method's power to detect
+temporally random divergences
+(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
+Throughout our analyses, the intermediate model (\modelUshaped) performs better
+than the \msb model, but not as well as the models with alternative priors on
+divergence models.
 This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
 models of temporally clustered divergences is caused by a combination of
-(1) uniform priors on parameters hindering the marginal likelihoods of models
-with more \divTime{} parameters and
+(1) small marginal likelihoods of models with more \divTime{} parameters due to
+uniform priors on divergence-time and demographic parameters and
 (2) the U-shaped prior on divergence models giving low prior density to models
 with intermediate numbers of divergence parameters.
-
-While our modifications improve the behavior of the model, we urge
-caution when using the method and interpreting its results.
+The former essentially rules out models with many \divTime{} parameters, which
+causes the latter to act like an "L-shaped" prior with a spike of
+prior density on the one-divergence model.
+Given the parameter richness of the model and the relatively small amount of
+information in the summary statistics, it is not surprising that the
+combination of these two factors can create a strong tendency to infer
+clustered models of divergence.
+
+While our modifications improve the behavior of the model, we urge caution when
+using the method and interpreting its results.
 The method attempts to approximate the posterior of a very parameter-rich model
 using relatively little information from the data.
-For example, when applied to the dataset of Philippine vertebrates analyzed by
+For example, when applied to the dataset of 22 taxon pairs from the Philippines
 \citet{Oaks2012} the model can have as many as 604--625 free parameters
-(depending on \divTimeNum), and samples over 1002 divergence models.
-Our modifications allow this model to be simplified to 471--492 free
-parameters.
+(depending on \divTimeNum), and samples over 1002 unordered divergence models.
+Even under the simplest possible model allowed under our new implementation,
+the model still has 471--492 free parameters.
+Furthermore, the coalescent and mutational models specified by these parameters
+predict a large amount of variation in possible datasets even when the
+parameter values are known.
 Sampling over hundreds of parameters and more than a thousand models is a
@@
 power, while improved, may still be low.
-Furthermore, ABC methods in general have been shown to be biased
-\citep{Robert2011} for model choice.
+While we see much less prior sensitivity under the new model compared to those
+observed by \citet{Oaks2012}, we still some evidence of such sensitivity when
+comparing the results of the empirical data analyzed under a diffuse
+(\empModelDPP) and informative (\empModelDPPInform) divergence-time prior
+(Figure~\ref{figPhilippines} C versus D).
+The fact that the posterior shifts toward the prior under the informative prior
+suggests that the shift away from the prior toward fewer divergence events
+under the diffuse prior might be still be caused by small marginal likelihoods
+of models with more divergence-time parameters (Figure~\ref{figPhilippines}).
+
+Nonetheless, it is reassuring to see a large amount of posterior uncertainty
+when the new implementation is applied to the empirical datasets
+(Figure~\labelcref{figPhilippines,figNegrosPanay}).
+Applications of the \msb model often result in strong posterior support for
+estimated scenarios
+\citep{Barber2010,Bell2012,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012},
+as we found here (Figure~\ref{figPhilippines}).
+Given the richness of the model, the variance of the processes being modeled,
+and the relatively small amount of information in the summary statistics
+calculated from the sequence data, we do not expect to find strong posterior
+support for any scenario.
+Based on results of the empirical and power analyses
+(Figures~\labelcref{figPowerPsiProbOld4,figPowerOmegaProbOld4,figPhilippines,figNegrosPanay})
+the new implementation more accurately reflects posterior uncertainty and
+avoids spurious support for biogeographic scenarios.
+
+We also urge caution when applying \dppmsbayes due to the fact that ABC methods
+in general are known to be biased for model choice \citep{Robert2011}.
 As recommended by \citet{Oaks2012}, any results from the method should be
 accompanied by assessments of prior sensitivity and power.
-We encourage investigators to view this method and its predecessor as a means
-of exploring their data rather than a rigorous means of evaluating hypotheses.
+We encourage investigators to view this method as a means of exploring their
+data rather than a rigorous means of evaluating hypotheses.
 
@@
 full-likelihood methods that can leverage all of the information present in the
-sequence data will become increasingly important for robustly estimating the
-temporal history of divergences across taxa \citep{JeetDiss}.
+sequence data will become increasingly important for robustly estimating shared
+evolutionary history across taxa \citep{JeetDiss}.
 With improving numerical methods for sampling over models of differing
-dimensionality \citep{Green1995,Lemey2009} and increasing efficiency of
-likelihood calculations \citep{Ayres2012}, analyzing rich comparative
-phylogeograpical models in a full-likelihood Bayesian framework is becoming
-computationally practical, especially when considering that simulating millions
-of random datasets from the prior under the simple ABC rejection approach is
-inefficient and computationally nontrivial.
+dimensionality \citep{Green1995,Lemey2009}, advances in Monte Carlo techniques
+\citep{Jordan2012}, and increasing efficiency of likelihood calculations
+\citep{Ayres2012}, analyzing rich comparative phylogeograpical models in a
+full-likelihood Bayesian framework is becoming computationally practical,
+especially when considering that simulating millions of random datasets from
+the prior under the simple ABC rejection approach is inefficient and
+computationally nontrivial.

### 2014-01-15 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on reply and dpp ms.
More edits to dpp ms, and adding Efron (2008) citation to
reply.

@@
 
-Furthermore, the modifications to the model also improve the method's power to
-detect random variation in divergence times, reducing the tendency to estimate
-clustered divergences
+The modifications to the model also improve the method's power to detect random
+variation in divergence times, reducing the tendency to estimate clustered
+divergences
 (Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
@@
 For example, when applied to the dataset of 22 taxon pairs from the Philippines
-\citet{Oaks2012} the model can have as many as 604--625 free parameters
+\citep{Oaks2012} the model can have as many as 604--625 free parameters
 (depending on \divTimeNum), and samples over 1002 unordered divergence models.
@@
 the model still has 471--492 free parameters.
-Furthermore, the coalescent and mutational models specified by these parameters
+Furthermore, the stochastic coalescent and mutational processes being modeled
 predict a large amount of variation in possible datasets even when the
 parameter values are known.
-Sampling over hundreds of parameters and more than a thousand models is a
-difficult inference problem, especially when only using a small number of
-summary statistics calculated from the sequence alignments of each taxon pair.
+The richness and stochastic nature of the model makes for a difficult inference
+problem, especially when using a small number of summary statistics calculated
+from the sequence alignments of each taxon pair.
 The population-genetic summary statistics used by our method contain little
@@
 While we see much less prior sensitivity under the new model compared to those
-observed by \citet{Oaks2012}, we still some evidence of such sensitivity when
-comparing the results of the empirical data analyzed under a diffuse
-(\empModelDPP) and informative (\empModelDPPInform) divergence-time prior
-(Figure~\ref{figPhilippines} C versus D).
+observed by \citet{Oaks2012}, we still see an effect when comparing the results
+of the empirical data analyzed under a diffuse (\empModelDPP) and informative
+(\empModelDPPInform) divergence-time prior (Figure~\ref{figPhilippines} C
+versus D).
 The fact that the posterior shifts toward the prior under the informative prior
 suggests that the shift away from the prior toward fewer divergence events
-under the diffuse prior might be still be caused by small marginal likelihoods
+under the diffuse prior might still be caused by small marginal likelihoods
 of models with more divergence-time parameters (Figure~\ref{figPhilippines}).
@@
 Based on results of the empirical and power analyses
-(Figures~\labelcref{figPowerPsiProbOld4,figPowerOmegaProbOld4,figPhilippines,figNegrosPanay})
+(Figures~\labelcref{figPowerPsiProbOld4,figPowerOmegaProbOld4,figPhilippines,figNegrosPanay}),
 the new implementation more accurately reflects posterior uncertainty and
@@
 
-We also urge caution when applying \dppmsbayes due to the fact that ABC methods
+We also urge caution when using \dppmsbayes due to the fact that ABC methods
 in general are known to be biased for model choice \citep{Robert2011}.

### 2014-01-15 manuscripts/reply/reply-main.tex
Working on reply and dpp ms.
More edits to dpp ms, and adding Efron (2008) citation to
reply.

@@
 in some models and under confident in others.
+Nonetheless, empirical Bayesian model choice does perform well for some
+inference problems.
+Particularly, in cases where large aggregate data sets are used for many
+parallel model-choice problems simultaneously, pooling information to inform
+priors can lead to favorable group-wise frequentist coverage across tests
+\citep{Efron2008}.
 

### 2014-01-16 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Minor edits to dpp and reply manuscripts.

@@
 uncertainty, and an order of magnitude smaller probability for their respective
-MAP model when compared to the \empModelOld model.
+MAP-divergence model when compared to the \empModelOld model
+(Figures~S\labelcref{figDivModelsDPP,figDivModelsDPPInform,figDivModelsDPPSimple,figDivModelsUniform,figDivModelsOld}).
 
@@
 (\npModelDPPOrdered) models of divergence.
-The result are similar under both models and, again, yield a large amount of
+The results are similar under both models and, again, yield a large amount of
 uncertainty about the number of divergence events that is similar to the prior
@@
 
-The modifications to the model also improve the method's power to detect random
+Our modifications also improve the method's power to detect random
 variation in divergence times, reducing the tendency to estimate clustered
@@
 (Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
-Our results are similar to those of \citet{Oaks2012} in that we also find
-the \msb model to infer strong support for clustered divergences when
+Our results are similar to those of \citet{Oaks2012} in that we find
+\msb will often infer strong support for clustered divergences when
 divergences are random over quite broad timescales
@@
 \ref{figValidationModelChoiceOmega}).
-Furthermore, better priors on parameters improves the method's power to detect
+Furthermore, using exponential priors improves the method's power to detect
 temporally random divergences
@@
 divergence models.
-This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
-models of temporally clustered divergences is caused by a combination of
+This suggests, as predicted by \citet{Oaks2012}, that the tendency of \msb to
+erroneously support models of temporally clustered divergences is caused by a
+combination of
 (1) small marginal likelihoods of models with more \divTime{} parameters due to
@@
 For example, when applied to the dataset of 22 taxon pairs from the Philippines
-\citep{Oaks2012} the model can have as many as 604--625 free parameters
+\citep{Oaks2012} the model has as many as 604--625 free parameters
 (depending on \divTimeNum), and samples over 1002 unordered divergence models.

### 2014-01-16 manuscripts/reply/reply.tex
Updating reply ms.
Another working title.
Minor edits to dpp and reply manuscripts.

@@
 
-\newcommand{\msTitle}{Nuh uh: A reply to Hickerson et al.
+\newcommand{\msTitle}{Implications of uniformly distributed, empirically
+    informed priors for phylogeographical model selection: A reply to Hickerson
+    et al.
 \xspace}
@@
     difficult problem in evolutionary biology.
-    \citep{Huang2011} introduced an approximate Bayesian approach which can be
+    \citet{Huang2011} introduced an approximate Bayesian approach which can be
     used to estimate the probabilities of models in which multiple sets of taxa

### 2014-01-30 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Updating DPP manuscript.
Adding Adam's comments to abstract and intro.

@@
 variation within and among the affected lineages.
-Thus, being able to accurately estimate models of shared evolutionary
-history across co-distributed taxa would provide an important
-tool for better understanding how regional and global biogeographic
-processes affect biodiversity.
+Thus, accurately estimating models of shared evolutionary history across
+co-distributed taxa provides an important tool for better understanding how
+regional and global biogeographic processes affect biodiversity.
 
@@
 approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
-Recently, however, \citet{Oaks2012} found that \msb can be strongly biased
-toward spuriously estimating biogeographically interesting models of highly
-clustered divergence times across taxa.
-The results of their simulation-based assessment of the method suggest that a
-combination of broad, uniformly distributed priors on nuisance parameters and a
-``U-shaped'' prior on divergence models are causing the bias.
-The model implemented in \msb uses a discrete uniform prior over the number of
-possible divergence parameters, which \citet{Oaks2012} show imposes a prior on
-divergence models that puts most of the prior mass on models with either very
-few or very many divergence times shared across the taxa (see Figure 5 of
+Given aligned sequence data for \npairs{} pairs of populations, \msb 
+estimates the number of divergence events shared among the pairs, the timing of
+the events, and the assignment of pairs to the events, while integrating out
+uncertainty in demographic parameters and the genealogical history of the
+sequences.
+% samples over all the possible ways of assigning the divergences between the
+% populations of each pair to 1--\npairs{} divergence events, while estimating
+% the timing of the events and integrating out uncertainty in demographic
+% parameters and the genealogical history of the sequences.
+Thus, when the number of pairs is not trivial, the method samples over a large
+number of divergence models of differing dimensionality (i.e., all the possible
+partitions of \npairs{} pairs to 1--\npairs{} divergence-time parameters), and,
+in so doing, estimates the posterior probability of each possible divergence
+model.
+
+For priors on divergence-time and demographic parameters, \msb uses continuous
+uniform probability distributions
+This causes  divergence models with more divergence-time parameters to
+integrate over a \emph{much} greater parameter space with low likelihood yet
+high prior density, resulting in small marginal likelihoods relative to models
+with fewer divergence-time parameters.
+Furthermore, \msb uses a discrete uniform prior over the number of divergence
+events $1,2,\ldots,\npairs{}$.
+Because there are many more possible divergence models for intermediate numbers
+of divergence events, this imposes a prior on divergence models that strongly
+disfavors such models and puts most of the prior mass on models with either
+very few or very many divergence-time parameters (see Figure 5 of
 \citet{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
-Furthermore, uniform priors used for most of the model's parameters
-result in small marginal likelihoods of parameter-rich models, i.e., models
-with many divergence-time parameters \citep{Jeffreys1935, Lindley1957}.
-Thus, together, the priors used in \msb will often create a bias toward
-models with few divergence-time parameters \citep{Oaks2012}.
+Given that models with many divergence-time parameters have small marginal
+likelihoods due to the uniform priors on divergence-time and demographic
+parameters, the U-shaped prior will strongly favor models with very
+few divergence events.
+As a result, even when highly informative uniform priors are used for
+divergence-time and demographic parameters, \msb will often strongly support
+biogeographically interesting models with a small number of divergence events
+shared among the taxa when, in reality, divergences were random over broad
+timescales (\citet{Oaks2012}).
+
+% Recently, however, \citet{Oaks2012} found that \msb can be strongly biased
+% toward spuriously estimating biogeographically interesting models of highly
+% clustered divergence times across taxa.
+% The results of their simulation-based assessment of the method suggest that a
+% combination of broad, uniformly distributed priors on nuisance parameters and a
+% ``U-shaped'' prior on divergence models are causing the bias.
+% The model implemented in \msb uses a discrete uniform prior over the number of
+% possible divergence parameters, which \citet{Oaks2012} show imposes a prior on
+% divergence models that puts most of the prior mass on models with either very
+% few or very many divergence times shared across the taxa (see Figure 5 of
+% \citet{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
+% Furthermore, uniform priors used for most of the model's parameters
+% result in small marginal likelihoods of parameter-rich models, i.e., models
+% with many divergence-time parameters \citep{Jeffreys1935, Lindley1957}.
+% Thus, together, the priors used in \msb will often create a bias toward
+% models with few divergence-time parameters \citep{Oaks2012}.
 

### 2014-01-30 manuscripts/dpp-msbayes/dpp-msbayes.tex
Updating DPP manuscript.
Adding Adam's comments to abstract and intro.

@@
     evolutionary history across taxa.
-    Our results demonstrate that the tendency of the \msb model to spuriously
-    support models of clustered divergences is caused by a combination of (1)
-    uniform priors on nuisance parameters hindering the marginal likelihoods of
-    models with more divergence-time parameters, and (2) a prior on divergence
-    models that disfavors models with intermediate numbers of divergence-time
-    parameters.
-    The new method addresses both of these issues and yields more robust
-    estimates of posterior uncertainty, and thus greatly reduces the tendency
-    of the model to incorrectly estimate biogeographically interesting models
-    with strong support.
+    Our results demonstrate the improved performance of the new method is due
+    to (1) more appropriate priors on divergence-time and demographic
+    parameters that avoid prohibitively small marginal likelihoods for models
+    with more divergence events,
+    and (2) the Dirichlet-process providing a flexible prior on divergence
+    histories that does not strongly disfavor models with intermediate numbers
+    of divergence events.
+    % Our results demonstrate that the tendency of the \msb model to spuriously
+    % support models of clustered divergences is caused by a combination of (1)
+    % uniform priors on nuisance parameters hindering the marginal likelihoods of
+    % models with more divergence-time parameters, and (2) a prior on divergence
+    % models that disfavors models with intermediate numbers of divergence-time
+    % parameters.
+    The new method yields more robust estimates of posterior uncertainty, and
+    thus greatly reduces the tendency of the model to incorrectly estimate
+    biogeographically interesting models with strong support.
 

### 2014-01-30 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More work on intro of DPP ms.
Incorporating more of Adam's comments and making some other tweaks as
well.

@@
 For priors on divergence-time and demographic parameters, \msb uses continuous
-uniform probability distributions
+uniform probability distributions.
 This causes divergence models with more divergence-time parameters to
@@
 high prior density, resulting in small marginal likelihoods relative to models
-with fewer divergence-time parameters.
+with fewer divergence-time parameters \citep{Jeffreys1939,Lindley1957}.
+Given that the marginal likelihood of a model weighted by its prior is what
+determines its posterior probability, this causes \msb to support models with
+fewer divergence events.
 Furthermore, \msb uses a discrete uniform prior over the number of divergence
@@
 \citet{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
-Given that models with many divergence-time parameters have small marginal
-likelihoods due to the uniform priors on divergence-time and demographic
-parameters, the U-shaped prior will strongly favor models with very
-few divergence events.
-As a result, even when highly informative uniform priors are used for
-divergence-time and demographic parameters, \msb will often strongly support
-biogeographically interesting models with a small number of divergence events
-shared among the taxa when, in reality, divergences were random over broad
-timescales (\citet{Oaks2012}).
+Given that models with many divergence events have small marginal likelihoods
+due to the uniform priors on divergence-time parameters, the U-shaped prior
+will effectively create a strong prior preference for models with very few
+divergence events.
+Recently, \citet{Oaks2012} found via simulation that \msb will often strongly
+support biogeographically interesting models with a small number of divergence
+events shared among the taxa, even when divergences were random over broad
+timescales.
+They suggested this behavior was due to the combination of uniform priors on
+parameters and the U-shaped prior on divergence models.
 
@@
 
-Following the recommendations of \citet{Oaks2012}, we modify the \msb model to
-test whether alternative parameterizations and priors will improve the behavior
-of the approximate-Bayesian model-choice approach.
+We modify the \msb model to test whether alternative parameterizations and
+priors will improve the behavior of the approximate-Bayesian model-choice
+approach.
 We implement a Dirichlet-process prior (DPP) over all possible models of
@@
 robustness, accuracy, and power compared to the original model.
-Also, our results show that both the uniform priors on demographic and
+Also, our results confirm that both the uniform priors on demographic and
 divergence-time parameters and the U-shaped prior on divergence models
 contribute to the biases of \msb.
-When the Oaks et al.'s \citeyear{Oaks2012} sequence data of 22 pairs of taxa
-from the Philippines is reanalyzed under the new model, we find a large amount
-of posterior uncertainty in the number of divergence events shared among the
-taxa; a result congruent with intuition given the richness of the model
-and the relatively small amount of information in the data.
+After reanalyzing sequence data of 22 pairs of taxa from the Philippines
+\citep{Oaks2012} under the new model, we find a large amount of posterior
+uncertainty in the number of divergence events shared among the taxa; a result
+congruent with intuition given the richness of the model and the relatively
+small amount of information in the data.
 

### 2014-01-31 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More tweaks to intro of DPP manuscript.

@@
 Thus, accurately estimating models of shared evolutionary history across
-co-distributed taxa provides an important tool for better understanding how
-regional and global biogeographic processes affect biodiversity.
+co-distributed taxa is important for better understanding how regional and
+global biogeographic processes affect biodiversity.
 
@@
 % parameters and the genealogical history of the sequences.
-Thus, when the number of pairs is not trivial, the method samples over a large
-number of divergence models of differing dimensionality (i.e., all the possible
-partitions of \npairs{} pairs to 1--\npairs{} divergence-time parameters), and,
-in so doing, estimates the posterior probability of each possible divergence
-model.
+Thus, the method samples over all possible divergence models of differing
+dimensionality (i.e., all the possible partitions of \npairs{} pairs to
+1--\npairs{} divergence-time parameters), and, in so doing, estimates the
+posterior probability of each model.
 
@@
 uniform probability distributions.
-This causes divergence models with more divergence-time parameters to
-integrate over a \emph{much} greater parameter space with low likelihood yet
-high prior density, resulting in small marginal likelihoods relative to models
-with fewer divergence-time parameters \citep{Jeffreys1939,Lindley1957}.
+This causes divergence models with more divergence-time parameters to integrate
+over a \emph{much} greater parameter space with low likelihood yet high prior
+density, resulting in small marginal likelihoods relative to models with fewer
+divergence-time parameters \citep{Jeffreys1939,Lindley1957}.
 Given that the marginal likelihood of a model weighted by its prior is what
-determines its posterior probability, this causes \msb to support models with
-fewer divergence events.
+determines its posterior probability, this can cause support for models with
+fewer divergence events \citep{Oaks2012}.
 Furthermore, \msb uses a discrete uniform prior over the number of divergence
 events $1,2,\ldots,\npairs{}$.
-Because there are many more possible divergence models for intermediate numbers
-of divergence events, this imposes a prior on divergence models that strongly
-disfavors such models and puts most of the prior mass on models with either
-very few or very many divergence-time parameters (see Figure 5 of
-\citet{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
-Given that models with many divergence events have small marginal likelihoods
-due to the uniform priors on divergence-time parameters, the U-shaped prior
-will effectively create a strong prior preference for models with very few
-divergence events.
+Because there are many more possible assignments of populations pairs to
+intermediate numbers of divergence events, this imposes a prior on divergence
+models that puts most of the prior mass on models with either very few or very
+many divergence-time parameters (see Figure 5 of \citet{Oaks2012}; for brevity
+we will refer to this prior as ``U-shaped'').
+Given that models with many divergence events can have small marginal
+likelihoods due to the uniform priors on divergence-time parameters, the
+U-shaped prior will effectively create a strong prior preference for models
+with very few divergence events.
 Recently, \citet{Oaks2012} found via simulation that \msb will often strongly
@@
 uniform priors on many of the model's parameters.
-Our simulation-based analyses show that the new implementation has improved
-robustness, accuracy, and power compared to the original model.
-Also, our results confirm that both the uniform priors on demographic and
-divergence-time parameters and the U-shaped prior on divergence models
-contribute to the biases of \msb.
-After reanalyzing sequence data of 22 pairs of taxa from the Philippines
+Using simulations, we show that the new implementation has improved robustness,
+accuracy, and power compared to the original model.
+Also, our results confirm that the improved performance of the new model
+is due to a combination of
+(1) more flexible priors on divergence-time and demographic parameters that
+avoid placing high prior density in improbable regions of parameter space, and
+(2) a diffuse Dirichlet-process prior that does not strongly disfavor
+divergence models with intermediate numbers of divergence events.
+% Also, our results confirm that both the uniform priors on demographic and
+% divergence-time parameters and the U-shaped prior on divergence models
+% contribute to the tendency of \msb.
+After reanalyzing sequence data from 22 pairs of taxa from the Philippines
 \citep{Oaks2012} under the new model, we find a large amount of posterior
 uncertainty in the number of divergence events shared among the taxa; a result
-congruent with intuition given the richness of the model and the relatively
-small amount of information in the data.
+in contrast with the original \msb model \citep{Oaks2012} and congruent with
+intuition given the richness of the model and the relatively small amount of
+information in the data.
 
@@
 \subsubsection*{The priors on nuisance parameters}
-Following the recommendations of \citet{Oaks2012}, we have replaced the use of
-continuous uniform distributions for priors on many of the
-model's parameters (\divTime{}, \ancestralTheta{}, \descendantTheta{1}{},
-\descendantTheta{2}{}, \bottleScalar{1}{}, \bottleScalar{2}{},
-\recombinationRate, \migrationRate{}) with more flexible parametric
-distributions from the exponential family.
+We have replaced the use of continuous uniform distributions for priors on many
+of the model's parameters (\divTime{}, \ancestralTheta{},
+\descendantTheta{1}{}, \descendantTheta{2}{}, \bottleScalar{1}{},
+\bottleScalar{2}{}, \recombinationRate, \migrationRate{}) with more flexible
+parametric distributions from the exponential family.
 We introduce gamma-distributed priors for rate parameters that have a sample
@@
 
-Another distinction between our model and that of \msb, is the prior on the
-sizes of the descendant populations of each pair.
+In addition, we have modified the prior on the sizes of the descendant
+populations of each pair.
 As described by \citet{Oaks2012}, \msb uses the joint prior

### 2014-01-31 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Another tweak to the intro of DPP manuscript.

@@
 dimensionality (i.e., all the possible partitions of \npairs{} pairs to
-1--\npairs{} divergence-time parameters), and, in so doing, estimates the
-posterior probability of each model.
+$1,2,\ldots,\npairs{}$ divergence-time parameters), and, in so doing, estimates
+the posterior probability of each model.
 

### 2014-01-31 manuscripts/reply/reply-main.tex
Minor edits to dpp and reply manuscripts.
Updating reply manuscript.
Incorporating comments from Adam and adding some missing `linenomath`
environments to paragraphs with equations.

@@
    \item The use of vague priors on the divergence-time parameters can lead to
-       support models with few divergence events shared across taxa. Thus, the
-       primary inference enabled by the approach is very sensitive to the
+       support for models with few divergence events shared across taxa. Thus,
+       the primary inference enabled by the approach is very sensitive to the
        priors on divergence times.
@@
 This is not surprising given the rich statistical literature that shows that
-model marginal likelihoods are quite sensitive to the priors used in Bayesian
-model selection
+marginal likelihoods are very sensitive to the priors used in Bayesian model
+selection
 \citep[e.g.,][]{Jeffreys1939,Lindley1957}.
@@
 
+\begin{linenomath}
 Whereas empirical Bayes approaches can provide powerful methods for parameter
@@
 parameters.
+\end{linenomath}
 
@@
 Nonetheless, empirical Bayesian model choice does perform well for some
-inference problems.
+model-choice problems.
 Particularly, in cases where large aggregate data sets are used for many
@@
 
+\begin{linenomath}
 These distinctions between Bayesian parameter estimation and model choice can
@@
 and data.
+\end{linenomath}
 
@@
 the narrowest prior on divergence times, it seems quite likely that their
-model-averaged results are dominated by models that exclude at least some of
-the 22 true divergence times, making their results difficult to interpret.
+model-averaged results are dominated by models that exclude true divergence
+times, making their results difficult to interpret.
 
@@
 one seeks to estimate.
-However continuous distributions perhaps more appropriate as priors for
-positive real-valued parameters have been shown to greatly reduce spurious
-support for clustered divergence models while allowing prior uncertainty to be
-accommodated \citep{Oaks2014dpp}.
+However, continuous distributions more appropriate as priors for positive
+real-valued parameters have been shown to greatly reduce spurious support for
+clustered divergence models while allowing prior uncertainty to be accommodated
+\citep{Oaks2014dpp}.
 

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Updating model description in dpp manuscript.
Explicitly referring to Bayes' rule and the chain rule of probability in
the initial introduction of the model's posterior.

@@
 \mutationRateScalarConstantVector in hand, the joint posterior distribution
-of our model is given by
+of our model is given by Bayes' rule as
 \begin{equation}
@@
 \end{equation}
-which can be expanded out into components that are assumed to be independent
-to get
+which can be expanded using the chain rule of probability into components that
+are assumed to be independent to get
 \begin{equation}

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Reworking intro paragraphs in dpp manuscript.

@@
 \section*{Introduction}
+
 Understanding the processes that generate biodiversity and regulate
 community assembly is a major goal of evolutionary biology.
-Given the dynamic nature of our planet, a major component of this goal is
-understanding how geological and climatic processes can affect the
-evolutionary history of whole groups of co-distributed species and their
-associated microbiota, including pathogens.
-Such large-scale processes predict patterns of shared evolutionary history
-across affected populations of organisms, e.g., temporally clustered
-divergences across co-distributed species.
-At recent timescales, such events should leave a signature in the genetic
-variation within and among the affected lineages.
-Thus, accurately estimating models of shared evolutionary history across
-co-distributed taxa is important for better understanding how regional and
-global biogeographic processes affect biodiversity.
+Large-scale changes to the environment, including geological and climatic
+events, can affect the evolutionary history of entire communities of
+co-distributed species and their associated microbiota.
+For example, by partitioning communities, such an event can isolate groups of
+populations and cause a temporal cluster of speciation events across
+co-distributed taxa.
+Given the dynamic nature of our planet, such biogeographical processes likely
+play a significant role in determining diversification rates and patterns.
+At recent timescales, temporal clusters of diversification caused by
+biogeographical events can leave a signature in the genetic variation within
+and among the affected lineages.
+Thus, methods for accurately estimating models of shared evolutionary events
+across co-distributed taxa from genetic data are important for better
+understanding how regional and global biogeographical processes affect
+biodiversity.
 
@@
 approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
+The method models temporally clustered divergences across taxa caused by a
+biogeographical event (or a ``divergence event'') as a single, instantaneous
+occurrence.
+In other words, a divergence event causes a set of taxa to share the same
+moment of divergence along a continuous time scale (i.e., simultaneous
+divergence).
 Given aligned sequence data for \npairs{} pairs of populations, \msb 
@@
 the events, and the assignment of pairs to the events, while integrating out
-uncertainty in demographic parameters and the genealogical history of the
+uncertainty in demographic parameters and the genealogical histories of the
 sequences.
@@
 dimensionality (i.e., all the possible partitions of \npairs{} pairs to
-$1,2,\ldots,\npairs{}$ divergence-time parameters), and, in so doing, estimates
+$1,2,\ldots,\npairs{}$ divergence events), and, in so doing, estimates
 the posterior probability of each model.
@@
 uncertainty in the number of divergence events shared among the taxa; a result
-in contrast with the original \msb model \citep{Oaks2012} and congruent with
+in contrast with the original \msb model and congruent with
 intuition given the richness of the model and the relatively small amount of

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Updating formating of manuscripts.
Updating preamble to remove section title and caption formatting
options. Using `\maketitle` in dpp manuscript to create title page.
Changing all `section*` calls to `section` in dpp manuscript.

@@
-\section*{Introduction}
+\section{Introduction}
 
@@
 
-\section*{Methods}
-
-\subsection*{The model}
+\section{Methods}
+
+\subsection{The model}
 In this section, we describe our model, which is a modification of the model
@@
 
-\subsubsection*{Likelihood terms of Equation \ref{eq:fullModelCompact}}
+\subsubsection{Likelihood terms of Equation \ref{eq:fullModelCompact}}
 \begin{linenomath}
@@
 
-\subsubsection*{Prior terms of Equation \ref{eq:fullModelCompact}}
+\subsubsection{Prior terms of Equation \ref{eq:fullModelCompact}}
 \begin{linenomath}
@@
 
-\subsubsection*{Priors on divergence models}
+\subsubsection{Priors on divergence models}
 \begin{linenomath}
@@
 
-\subsection*{Differences between our model and the original \msb model}
-\subsubsection*{The prior on divergence models}
+\subsection{Differences between our model and the original \texttt{msBayes} model}
+\subsubsection{The prior on divergence models}
 One of the key differences between our model and that of \msb \citep{Huang2011}
@@
 
-\subsubsection*{Estimating ordered divergence models}
+\subsubsection{Estimating ordered divergence models}
 As mentioned above, \msb samples over unordered divergence models
@@
 
-\subsubsection*{The priors on nuisance parameters}
+\subsubsection{The priors on nuisance parameters}
 We have replaced the use of continuous uniform distributions for priors on many
@@
 
-\subsubsection*{Flexibility in parameterizing the model}
+\subsubsection{Flexibility in parameterizing the model}
 In our implementation, we provide the ability to control the richness of the
@@
 
-\subsubsection*{Time scale}
+\subsubsection{Time scale}
 As described above, in our model divergence times are in units of
@@
 
-\subsection*{ABC estimation of the posterior of the model}
-\subsubsection*{Sampling from the prior}
+\subsection{ABC estimation of the posterior of the model}
+\subsubsection{Sampling from the prior}
 To estimate the approximate posterior of Equation \ref{eq:approxModelCompact},
@@
 
-\subsubsection*{Ordering of taxon-specific summary statistics}
+\subsubsection{Ordering of taxon-specific summary statistics}
 As alluded to in the model description, \msb does not maintain the order of the
@@
 
-\subsubsection*{Obtaining an approximate posterior from the prior samples}
+\subsubsection{Obtaining an approximate posterior from the prior samples}
 We use a rejection algorithm to retain an approximate posterior sample of
@@
 
-\subsection*{Assessing model-choice behavior and robustness}
+\subsection{Assessing model-choice behavior and robustness}
 Following the simulation-based approach of \citet{Oaks2012}, we characterize
@@
 
-\subsection*{Assessing power}
+\subsection{Assessing power}
 We evaluated the power of the same four models (Table~\ref{tabPriors}) to
@@
 
-\subsection*{An empirical application}
+\subsection{An empirical application}
 We also assessed the behavior of the newly implemented models when applied to
@@
 
-\section*{Results}
-\subsection*{Validation analyses: Estimation accuracy}
+\section{Results}
+\subsection{Validation analyses: Estimation accuracy}
 In terms of estimating the variance of divergence times (\divTimeDispersion),
@@
 
-\subsection*{Validation analyses: Model-choice accuracy}
+\subsection{Validation analyses: Model-choice accuracy}
 The \msb model, and our modification of it, is a model-choice method
@@
 
-\subsection*{Validation analyses: Ordered divergence models}
+\subsection{Validation analyses: Ordered divergence models}
 Our results show that the method performs similarily when sampling over ordered
@@
 
-\subsection*{Power analyses: Estimation accuracy}
+\subsection{Power analyses: Estimation accuracy}
 All of the models we evaluated (Table \ref{tabPriors}) struggle to estimate the
@@
 
-\subsection*{Power analyses: Model choice}
+\subsection{Power analyses: Model choice}
 Our modifications of the \msb model decrease the method's bias toward clustered
@@
 
-\subsection*{Empirical results}
+\subsection{Empirical results}
 As expected based on the results of \citet{Oaks2012}, when the Philippines data
@@
 
-\section*{Discussion}
+\section{Discussion}
 Our results demonstrate that using alternative priors on parameters and

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes.tex
Updating formating of manuscripts.
Updating preamble to remove section title and caption formatting
options. Using `\maketitle` in dpp manuscript to create title page.
Changing all `section*` calls to `section` in dpp manuscript.

@@
     Estimating Shared Evolutionary History \xspace}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{document}
-\doublespacing
-\raggedright
-\setlength{\parindent}{0.5in}
-\begin{linenumbers}
+% \doublespacing
+% \raggedright
+% \setlength{\parindent}{0.5in}
+% \begin{linenumbers}
 
-\begin{titlepage}
-    \begin{flushleft}
-        \sffamily
+\title{\msTitle}
 
-        \MakeUppercase{\large\bfseries \msTitle}
+\author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
+\author[1]{Mark T.\ Holder}
+\affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
 
-        \vspace{12pt}
-        \textbf{Running head:} \MakeUppercase{Improved Estimation of Shared Evolutionary History}
+% \date{\today}
+\date{\parbox{\linewidth}{\centering%
+    \today\endgraf\bigskip
+    \textbf{Running head}: Improved Estimation of Shared Evolutionary History}}
 
-        \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2}$ and Mark T.\ Holder$^{1}$
+\maketitle
 
-        \bigskip
-        $^1$\emph{Department of Ecology and Evolutionary Biology,
-            % Biodiversity Institute,
-            University of Kansas,
-            Lawrence, Kansas 66045}\\[.1in]
-        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
-        joaks1@gmail.com})\\
-
-    \end{flushleft}
-\end{titlepage}
-
-{\sffamily
-    \noindent\textbf{ABSTRACT} \\
-    \noindent To understand the processes that generate biodiversity, it is
+\begin{abstract}
+    To understand the processes that generate biodiversity, it is
     important to account for large-scale processes that affect the evolutionary
@@
     computation; model choice; phylogeography; biogeography} 
-}
+\end{abstract}
 
@@
 
-\section*{Acknowledgments}
+\section{Acknowledgments}
 J.\ Oaks thanks the National Science Foundation for supporting this work (DEB
@@
 
-\end{linenumbers}
+% \end{linenumbers}
 
@@
 \setcounter{page}{1}
+\setcounter{section}{0}
 
@@
 
+\section*{Supporting Information}
 % PUT MAIN TEXT CITATION HERE
-{
-\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape supporting information}}
-\begin{thebibliography}{1}
-\providecommand{\natexlab}[1]{#1}
-\providecommand{\url}[1]{\texttt{#1}}
-\providecommand{\urlprefix}{URL }
-
-\bibitem[{Oaks and Holder (2014)Oaks and Holder}]{Oaks2014dpp}
-Oaks, J.~R.\ and M.~T.\ Holder,
-  % 2013.
-\newblock \msTitle
-% \newblock Evolution 67:991--1010.
-
-\end{thebibliography}
-}
+\hangindent=1cm
+Oaks, J.~R.\ and M.~T.\ Holder. \msTitle.
 

### 2014-02-19 manuscripts/reply/reply-main.tex
Some miscellaneous edits to reply ms.
Updating reply ms formating.
Changing all `section*` calls to `section`.

@@
 
-\section*{The potential implications of empirical Bayesian model choice}
+\section{The potential implications of empirical Bayesian model choice}
 \citet{Hickerson2013} repeatedly refer to the prior distributions used by
@@
 
-\subsection*{Theoretical implications of empirical priors for Bayesian model
+\subsection{Theoretical implications of empirical priors for Bayesian model
 choice}
@@
 Nonetheless, empirical Bayesian model choice does perform well for some
-model-choice problems.
+problems.
 Particularly, in cases where large aggregate data sets are used for many
@@
 \citep{Efron2008}.
+However, this is far removed from the single model-choice problem of \msb.
 
@@
 
-\subsection*{Practical concerns about empirically informed uniform priors for
+\subsection{Practical concerns about empirically informed uniform priors for
     Bayesian model choice}
@@
 
-\subsubsection*{Re-analyses of the Philippines dataset using empirical Bayesian
+\subsubsection{Re-analyses of the Philippines dataset using empirical Bayesian
 model averaging}
@@
 
-\subsubsection*{Simulation-based assessment of Hickerson et al.'s
+\subsubsection{Simulation-based assessment of Hickerson et al.'s
 \citeyear{Hickerson2013} model averaging over empirical priors}
@@
 
-\subsection*{Additional thoughts on empirical priors in Bayesian model choice}
+\subsection{Additional thoughts on empirical priors in Bayesian model choice}
 The preceding sections are not a general critique of Bayesian model averaging.
@@
 
-\section*{Assessing the power of the model-averaging approach of
+\section{Assessing the power of the model-averaging approach of
     \citet{Hickerson2013}}
@@
 
-\section*{The importance of power analysis to guide applications of \msb}
+\section{The importance of power analysis to guide applications of \msb}
 \citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
@@
 
-\section*{The causes of support for models of co-divergence}
+\section{The causes of support for models of co-divergence}
 To determine how best to improve the behavior of \msb, it is important to
@@
 
-\section*{General thoughts on the model of \msb}
+\section{General thoughts on the model of \msb}
 Our findings are not surprising in light of the difficult inference problem
@@
 
-\section*{Conclusions}
+\section{Conclusions}
 We demonstrate how the approximate Bayesian model-choice method implemented in

### 2014-02-19 manuscripts/reply/reply-si-main.tex
Updating reply ms formating.
Changing all `section*` calls to `section`.

@@
-\section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
+\section{An error in Hickerson et al.'s re-analysis of the Philippines data}
 \citet{Hickerson2013} re-analyzed the dataset of \citet{Oaks2012} using a
@@
 %CWL: This seems to specific, not sure about including
-\section*{Validation analyses}
+\section{Validation analyses}
 % \highLight{Not sure where to put this section}.
@@
 
-\section*{Additional clarifications from \citet{Hickerson2013}}
+\section{Additional clarifications from \citet{Hickerson2013}}
 
-\subsection*{Saturation of summary statistics}
+\subsection{Saturation of summary statistics}
 \citet{Hickerson2013} claim the priors used by \citet{Oaks2012} ``cause much of
@@
 
-\subsection*{Graphical prior comparisons}
+\subsection{Graphical prior comparisons}
 \citet{Hickerson2013} advocate the use of what they call graphical checks of
@@
 
-\subsection*{Differing utilities of \numt{} and \vmratio{} in \msb}
+\subsection{Differing utilities of \numt{} and \vmratio{} in \msb}
 The primary component of the \msb model is the vector of divergence

### 2014-02-19 manuscripts/reply/reply-main.tex
Adjusting formatting of reply ms.
Changing all `section*` calls to `section`. Switching to `\maketitle`
for formatting of the title page.

@@
 
-\section{The importance of power analysis to guide applications of \msb}
+\section{The importance of power analysis to guide applications of \texttt{msBayes}}
 \citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
@@
 
-\section{General thoughts on the model of \msb}
+\section{General thoughts on the model of \texttt{msBayes}}
 Our findings are not surprising in light of the difficult inference problem

### 2014-02-19 manuscripts/reply/reply-si-main.tex
Adjusting formatting of reply ms.
Changing all `section*` calls to `section`. Switching to `\maketitle`
for formatting of the title page.

@@
 
-\subsection{Differing utilities of \numt{} and \vmratio{} in \msb}
+\subsection{Differing utilities of \numt{} and \vmratio{} in \texttt{msBayes}}
 The primary component of the \msb model is the vector of divergence

### 2014-02-19 manuscripts/reply/reply.tex
Updating reply ms formating.
Changing all `section*` calls to `section`.
Adjusting formatting of reply ms.
Changing all `section*` calls to `section`. Switching to `\maketitle`
for formatting of the title page.

@@
 
-\newcommand{\msTitle}{Implications of uniformly distributed, empirically
-    informed priors for phylogeographical model selection: A reply to Hickerson
-    et al.
-\xspace}
+\title{Implications of uniformly distributed, empirically informed priors for
+phylogeographical model selection: A reply to Hickerson et al.}
+
+\author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
+\author[2]{Jeet Sukumaran}
+\author[3]{Charles W.\ Linkem}
+\author[4]{Cameron D.\ Siler}
+\author[1]{Rafe M.\ Brown}
+\affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
+\affil[2]{Department of Biology, Duke University, Durham, North Carolina 27708}
+\affil[3]{Department of Biology, University of Washington, Seattle, Washington 98195}
+\affil[4]{Sam Noble Museum, Department of Biology, University of Oklahoma, Norman, Oklahoma 73072}
+
+% \date{\today}
+\date{\parbox{\linewidth}{\centering%
+    \today\endgraf\bigskip
+    \textbf{Running head}: Approximate Bayesian model choice}}
+
+\makeatletter
+\let\msTitle\@title
+\let\msAuthor\@author
+\let\msDate\@date
+\makeatother
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{document}
-\doublespacing
-\raggedright
-\setlength{\parindent}{0.5in}
-\begin{linenumbers}
+% \doublespacing
+% \raggedright
+% \setlength{\parindent}{0.5in}
+% \begin{linenumbers}
 
-\begin{titlepage}
-    \begin{flushleft}
-        \sffamily
+\maketitle
 
-        \MakeUppercase{\large\bfseries \msTitle}
-
-        \vspace{12pt}
-        \textbf{Running head:} \MakeUppercase{Approximate Bayesian model
-        choice}
-
-        \vspace{12pt}
-        Jamie R.\ Oaks$^{1,2}$, Jeet Sukumaran$^{3}$, Charles W.\ Linkem$^{4}$,
-        Cameron D.\ Siler$^{5}$, and Rafe M.\ Brown$^{1}$
-
-        \bigskip
-        $^1$\emph{Department of Ecology and Evolutionary Biology,
-            Biodiversity Institute,
-            University of Kansas,
-            Lawrence, Kansas 66045}\\[.1in]
-        $^3$\emph{Department of Biology,
-            Duke University,
-            Durham, North Carolina 27708} \\[.1in]
-        $^4$\emph{Department of Biology,
-            University of Washington,
-            Seattle, Washington 98195}\\[.1in]
-        $^5$\emph{Sam Noble Museum,
-            Department of Biology,
-            University of Oklahoma,
-            Norman, Oklahoma 73072}\\[.1in]
-        $^2$\emph{Corresponding author} (\href{mailto:joaks1@gmail.com}{\tt
-        joaks1@gmail.com})\\
-
-    \end{flushleft}
-\end{titlepage}
-
-{\sffamily
-    \noindent\textbf{ABSTRACT} \\
-    \noindent Biogeographers frequently seek to explain population and species
+\begin{abstract}
+    Biogeographers frequently seek to explain population and species
     differentiation on geographical phenomena.
@@
         model choice; empirical Bayes} 
-}
+\end{abstract}
 
@@
 
-\section*{Acknowledgments}
+\section{Acknowledgments}
 We thank Jake Esselstyn for insightful comments that greatly improved this
@@
 
+\section*{Supporting Information}
 % PUT MAIN TEXT CITATION HERE
-{
-\renewcommand{\refname}{\noindent\MakeUppercase{\LARGE\sffamily\upshape supporting information}}
-\begin{thebibliography}{1}
-\providecommand{\natexlab}[1]{#1}
-\providecommand{\url}[1]{\texttt{#1}}
-\providecommand{\urlprefix}{URL }
+\hangindent=1cm
+Oaks, J.~R., J.~Sukumaran, C.~W.\ Linkem, C.~D.\ Siler, and R.~M.\ Brown. \msTitle.
 
-\bibitem[{Oaks et~al.(2014)Oaks, Sukumaran, Esselstyn, Linkem, Siler, Brown,
-  and Holder}]{Oaks2013reply}
-Oaks, J.~R., J.~Sukumaran, J.~A.\ Esselstyn, C.~W.\ Linkem, C.~D.\ Siler,
-    R.~M.\ Brown, and M.~T.\ Holder,
-  % 2013.
-\newblock \msTitle
-% \newblock Evolution 67:991--1010.
-
-\end{thebibliography}
-}
-
-\doublespacing
+% \doublespacing
 \input{reply-si-main.tex}

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Using `\url` for all URLs in manuscript.

@@
 software package \dppmsbayes
-(\href{https://github.com/joaks1/dpp-msbayes}{https://github.com/joaks1/dpp-msbayes}).
+(\href{https://github.com/joaks1/dpp-msbayes}{\url{https://github.com/joaks1/dpp-msbayes}}).
 Our analyses described below utilize the freely avaliable, open-source software
-package \pymsbayes (\href{https://github.com/joaks1/PyMsBayes}{https://github.com/joaks1/PyMsBayes}),
+package \pymsbayes (\href{https://github.com/joaks1/PyMsBayes}{\url{https://github.com/joaks1/PyMsBayes}}),
 which is a Python API that provides a multi-processing interface to \msb and
@@
 freely and publicly available in real-time at
-\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
+\href{https://github.com/joaks1/msbayes-experiments}{\url{https://github.com/joaks1/msbayes-experiments}}.
 All information necessary to reproduce our results is provided there.

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes.tex
Updating title page metadata for dpp manuscript.
Defining variables to preserve values for title page data, so they
can be used beyond the title page.
Updating acknowledgements in both manuscripts.

@@
 
-\newcommand{\msTitle}{An Improved Approximate-Bayesian Model-choice Method for
-    Estimating Shared Evolutionary History \xspace}
+\title{An Improved Approximate-Bayesian Model-choice Method for Estimating
+Shared Evolutionary History}
+
+\author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
+\author[1]{Mark T.\ Holder}
+\affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
+
+% \date{\today}
+\date{\parbox{\linewidth}{\centering%
+    \today\endgraf\bigskip
+    \textbf{Running head}: Improved Estimation of Shared Evolutionary History}}
+
+\makeatletter
+\let\msTitle\@title
+\let\msAuthor\@author
+\let\msDate\@date
+\makeatother
 
@@
 % \begin{linenumbers}
-
-\title{\msTitle}
-
-\author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
-\author[1]{Mark T.\ Holder}
-\affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
-
-% \date{\today}
-\date{\parbox{\linewidth}{\centering%
-    \today\endgraf\bigskip
-    \textbf{Running head}: Improved Estimation of Shared Evolutionary History}}
 
@@
 \section{Acknowledgments}
-J.\ Oaks thanks the National Science Foundation for supporting this work (DEB
+I am greatful to Melissa Callahan, Mark Holder, Emily McTavish, Daniel Money,
+Jordan Koch, and Adam Leach\'{e} for insightful comments that greatly improved
+this work.
+I thank the National Science Foundation for supporting this work (DEB
 1011423 and DBI 1308885).
-J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
+This work was also supported by the University of Kansas (KU) Office of Graduate
 Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
@@
 Biodiversity Institute.
+I also thank Mark Holder, the KU Information and Telecommunication Technology
+Center, KU Computing Center, and the iPlant Collaborative for the computational
+support necessary to conduct the analyses presented herein.
 

### 2014-02-19 manuscripts/reply/reply.tex
Updating acknowledgements in both manuscripts.

@@
 \section{Acknowledgments}
-We thank Jake Esselstyn for insightful comments that greatly improved this
-work.
+We thank Melissa Callahan, Jake Esselstyn, Mark Holder, Emily McTavish, Daniel
+Money, Jordan Koch, and Adam Leach\'{e} for insightful comments that greatly
+improved this work.
 J.\ Oaks thanks the National Science Foundation for supporting this work (DEB
@@
 Biodiversity Institute.
+We also thank Mark Holder, the KU Information and Telecommunication Technology
+Center, KU Computing Center, and the iPlant Collaborative for the computational
+support necessary to conduct the analyses presented herein.
 

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Changing first person to singular in dpp ms.

@@
 many divergence-time parameters (see Figure 5 of \citet{Oaks2012}; for brevity
-we will refer to this prior as ``U-shaped'').
+I will refer to this prior as ``U-shaped'').
 Given that models with many divergence events can have small marginal
@@
 
-We modify the \msb model to test whether alternative parameterizations and
+Here, I modify the \msb model to test whether alternative parameterizations and
 priors will improve the behavior of the approximate-Bayesian model-choice
 approach.
-We implement a Dirichlet-process prior (DPP) over all possible models of
+I implement a Dirichlet-process prior (DPP) over all possible models of
 divergence, and use gamma and beta probability distributions in place of
 uniform priors on many of the model's parameters.
-Using simulations, we show that the new implementation has improved robustness,
+Using simulations, I show that the new implementation has improved robustness,
 accuracy, and power compared to the original model.
-Also, our results confirm that the improved performance of the new model
+Also, my results confirm that the improved performance of the new model
 is due to a combination of
@@
 After reanalyzing sequence data from 22 pairs of taxa from the Philippines
-\citep{Oaks2012} under the new model, we find a large amount of posterior
+\citep{Oaks2012} under the new model, I find a large amount of posterior
 uncertainty in the number of divergence events shared among the taxa; a result
@@
 \subsection{The model}
-In this section, we describe our model, which is a modification of the model
+In this section, I describe the model, which is a modification of the model
 implemented in \msb \citep{Huang2011,Oaks2012}.
@@
 (\href{https://github.com/joaks1/dpp-msbayes}{\url{https://github.com/joaks1/dpp-msbayes}}).
-Our analyses described below utilize the freely avaliable, open-source software
-package \pymsbayes (\href{https://github.com/joaks1/PyMsBayes}{\url{https://github.com/joaks1/PyMsBayes}}),
+To perform the analyses described below, I used the freely avaliable,
+open-source software package \pymsbayes
+(\href{https://github.com/joaks1/PyMsBayes}{\url{https://github.com/joaks1/PyMsBayes}}),
 which is a Python API that provides a multi-processing interface to \msb and
 \dppmsbayes.
-We performed the work described below following the principles of
+I performed the work described below following the principles of
 Open Notebook Science.
-Using version-control software, we make progress in all aspects of the work
+Using version-control software, I make progress in all aspects of the work
 freely and publicly available in real-time at
 \href{https://github.com/joaks1/msbayes-experiments}{\url{https://github.com/joaks1/msbayes-experiments}}.
-All information necessary to reproduce our results is provided there.
-We follow much of the notation of \citet{Oaks2012}, but modify it to
-aid in our description of the new model.
-A summary of our notation can be found in Table \ref{tabNotation}.
-
-We assume an investigator is interested in inferring the distribution
+All information necessary to reproduce my results is provided there.
+I follow much of the notation of \citet{Oaks2012}, but modify it to
+aid in my description of the new model.
+A summary of my notation can be found in Table \ref{tabNotation}.
+
+I assume an investigator is interested in inferring the distribution
 of divergence times among \npairs{} pairs of populations.
@@
 sampled across the \npairs{} pairs of populations.
-We use \alignment{i}{j} to represent the multiple sequence alignment of
+I use \alignment{i}{j} to represent the multiple sequence alignment of
 locus $j$ for population pair $i$.
@@
 i.e., a vector of sequence alignments for all pairs and loci.
-We let \geneTree{i}{j} represent the gene tree upon which \alignment{i}{j}
+I let \geneTree{i}{j} represent the gene tree upon which \alignment{i}{j}
 evolved according to fixed \hky substitution model parameters \hkyModel{i}{j}.
@@
 \mutationRateScalarConstantVector in hand, the joint posterior distribution
-of our model is given by Bayes' rule as
+of the model is given by Bayes' rule as
 \begin{equation}
@@
 To avoid calculating the likelihood terms of Equation \ref{eq:fullModelCompact},
-we distill each sequence alignment \alignment{}{} into a vector of insufficient
+I distill each sequence alignment \alignment{}{} into a vector of insufficient
 summary statistics \alignmentSS{}{}, thus replacing the full dataset
@@
 \tol being greater than zero.
-We describe the full model in detail before delving into the numerical
+I describe the full model in detail before delving into the numerical
 method of estimating the approximate model.
@@
 The divergence times are in coalescent units relative to the size of
-a constant reference population, which we denote \myTheta{C}, that is equal to
+a constant reference population, which I denote \myTheta{C}, that is equal to
 the expectation of the prior on the size of the descendant populations
@@
 The \divTime{} parameters are in units of $\myTheta{C}/\mutationRate$
-generations, which we denote as \globalcoalunit generations.
+generations, which I denote as \globalcoalunit generations.
 Thus, each \divTime{} within \divTimeVector is proportional to time and can be
-converted to the number of generations of the reference population, which we
+converted to the number of generations of the reference population, which I
 denote \divTime{G_C}, by assuming a mutation rate and multiplying by the
@@
 As described by \citet{Oaks2012}, to get the divergence times in units
-proportional to the expected number of mutations, we must
+proportional to the expected number of mutations, I must
 scale them by the realized population size for locus $j$ of population-pair $i$ 
@@
     OeisPartitionTriangle,Malenfant2011}.
-We denote the set of all possible integer partitions of \npairs{} as
+I denote the set of all possible integer partitions of \npairs{} as
 \integerPartitionSet{\npairs{}} and the length of that set as
@@
 \begin{linenomath}
-We implement two prior probability distributions over the space
+I implement two prior probability distributions over the space
 of all possible discrete divergence models (\divTimeIndexVector).
@@
 (discrete divergence models).
-We denote this prior as
+I denote this prior as
 $\divTimeIndexVector \sim \priorUniform$.
@@
 
-The second prior we implement is based on the Dirichlet-process, which is a
+The second prior is based on the Dirichlet-process, which is a
 stochastic process that groups elements into an unknown number of discrete
@@
 which determines how clustered the process will be.
-We use the Dirichlet process to place a prior over all possible ordered
-partitions of \npairs{} elements, which we denote as $\divTimeIndexVector \sim
+I use the Dirichlet process to place a prior over all possible ordered
+partitions of \npairs{} elements, which I denote as $\divTimeIndexVector \sim
 \priorDPP{}$.
-We take a hierarchical approach and use a prior probability
+I take a hierarchical approach and use a prior probability
 distribution (i.e., hyperprior) for \concentrationParam.
-More specifically, we use a gamma-distributed prior $\concentrationParam \sim
+More specifically, I use a gamma-distributed prior $\concentrationParam \sim
 Gamma(\gammaShape{\concentrationParam}, \gammaScale{\concentrationParam})$,
@@
 
-\subsection{Differences between our model and the original \texttt{msBayes} model}
+\subsection{Differences between my model and the original \texttt{msBayes} model}
 \subsubsection{The prior on divergence models}
-One of the key differences between our model and that of \msb \citep{Huang2011}
+One of the key differences between my model and that of \msb \citep{Huang2011}
 is the prior distribution on divergence models.
@@
 parameters are represented at least once (see Equation 2 of \citet{Oaks2012}).
-We denote this prior used by \msb as $\divTimeIndexVector \sim \priorOld$.
+I denote this prior used by \msb as $\divTimeIndexVector \sim \priorOld$.
 \citet{Oaks2012} discuss how placing a uniform prior over the number of
@@
 Figure 5B of \citet{Oaks2012}).
-To avoid this, we place priors directly on the sample space of divergence
+To avoid this, I place priors directly on the sample space of divergence
 models, thus eliminating the parameter \numt{} from the model.
-We introduce two priors on divergence models:
+I introduce two priors on divergence models:
 (1) a prior that is uniform over all unordered divergence models, and
@@
 associated with it, is discarded.
-In our implementation, inference can be done on either unordered or ordered
+In my implementation, inference can be done on either unordered or ordered
 models of divergence.
@@
 \subsubsection{The priors on nuisance parameters}
-We have replaced the use of continuous uniform distributions for priors on many
+I have replaced the use of continuous uniform distributions for priors on many
 of the model's parameters (\divTime{}, \ancestralTheta{},
@@
 parametric distributions from the exponential family.
-We introduce gamma-distributed priors for rate parameters that have a sample
+I introduce gamma-distributed priors for rate parameters that have a sample
 space of all positive real numbers (\divTime{}, \ancestralTheta{},
@@
 
-In addition, we have modified the prior on the sizes of the descendant
+In addition, I have modified the prior on the sizes of the descendant
 populations of each pair.
@@
 populations of each pair.
-Under our model, the sizes of the descendant populations of each
+Under my model, the sizes of the descendant populations of each
 pair are \iid as
@@
 \subsubsection{Flexibility in parameterizing the model}
-In our implementation, we provide the ability to control the richness of the
+In my implementation, I provide the ability to control the richness of the
 model.
@@
 
-We also provide the option of eliminating the parameters associated with the
+I also provide the option of eliminating the parameters associated with the
 post-divergence bottlenecks in the descendant populations of each pair
@@
 Also, rather than eliminate the bottleneck parameters,
-we allow \bottleScalar{1}{} and \bottleScalar{2}{} to be constrained to be
+I allow \bottleScalar{1}{} and \bottleScalar{2}{} to be constrained to be
 equal, which removes one free parameter from the model for each of the
@@
 
-Overall, our implementation allows an investigator to specify a model that has
+Overall, my implementation allows an investigator to specify a model that has
 as many as seven parameters per population pair
@@
 \subsubsection{Time scale}
-As described above, in our model divergence times are in units of
+As described above, in my model divergence times are in units of
 $\myTheta{C}/\mutationRate$ generations, where \myTheta{C} is the expectation
@@
 To estimate the approximate posterior of Equation \ref{eq:approxModelCompact},
-we use an ABC rejection algorithm.
+I use an ABC rejection algorithm.
 The first step of this algorithm entails collecting a random sample of
@@
 Each sample is generated by
-(1) drawing values of all the model's parameters, which we denote \hpvector{},
+(1) drawing values of all the model's parameters, which I denote \hpvector{},
 from their respective prior distributions;
@@
 the same summary statistic vector estimated from the observed data \ssVectorObs.
-After repeating this procedure \numPriorSamples times, we obtain a random
+After repeating this procedure \numPriorSamples times, I obtain a random
 sample of parameter vectors
@@
 
-For all of our analyses below, we use four summary statistics for each of
+For all of the analyses below, I use four summary statistics for each of
 the pairs of populations:
@@
 respectively \citep{Hickerson2006,Huang2011,Oaks2012}.
-We use the new notation to avoid confusion that the values represent prior or
+I use the new notation to avoid confusion that the values represent prior or
 posterior expectations, and to indicate that \vmratio{} is not a parameter of
@@
 population pairs.
-However, we do not provide a formal proof that this re-ordering of the pairs
+However, I do not provide a formal proof that this re-ordering of the pairs
 does not introduce bias.
@@
 
-To maintain compatibility and comparability with \msb we maintain
+To maintain compatibility and comparability with \msb I maintain
 the re-ordering of taxon-specific summary statistics by $\pi_b$.
-However, we also allow the order to be maintained, and ordered
+However, I also allow the order to be maintained, and ordered
 divergence models to be estimated.
@@
 \subsubsection{Obtaining an approximate posterior from the prior samples}
-We use a rejection algorithm to retain an approximate posterior sample of
+I use a rejection algorithm to retain an approximate posterior sample of
 \paramSampleVector{} from the prior sample
@@
 of regression techniques \citep{Beaumont2002,Blum2009,Leuenberger2010}.
-For our analyses, we use the general linear model (GLM) regression adjustment
+For my analyses, I use the general linear model (GLM) regression adjustment
 \citep{Leuenberger2010} as implemented in \abctoolbox
@@
 \subsection{Assessing model-choice behavior and robustness}
-Following the simulation-based approach of \citet{Oaks2012}, we characterize
+Following the simulation-based approach of \citet{Oaks2012}, I characterize
 the behavior of several models under the ideal conditions where the
@@
 for analysis (i.e., the prior is correct).
-We selected four model priors for these analyses (Table~\ref{tabPriors}):
+I selected four model priors for these analyses (Table~\ref{tabPriors}):
 (1) A model to represent the original \msb model, \modelOld, with
@@
 $\dThetas \sim Exp(mean=0.025)$.
-We selected the exponential prior on divergence time used in models \modelDPP,
+I selected the exponential prior on divergence time used in models \modelDPP,
 \modelUniform, and \modelUshaped to have the same variance as the uniform prior
 in model \modelOld.
-We selected the exponential prior on population size used in models \modelDPP,
+I selected the exponential prior on population size used in models \modelDPP,
 \modelUniform, and \modelUniform to have the same mean as the uniform prior in
@@
 and unordered divergence models.
-For all of our simulations, we used a simulated data structure of eight
+For all of my simulations, I used a simulated data structure of eight
 population pairs, with a single 1000 base-pair locus sampled from 10
@@
 
-For each of the four models, we simulated $1\e6$ samples from the prior and
+For each of the four models, I simulated $1\e6$ samples from the prior and
 50,000 datasets, also drawn from the prior.
-We then analyzed each of the simulated datasets, retaining a
+I then analyzed each of the simulated datasets, retaining a
 posterior of 1000 samples from the respective prior.
@@
 posterior samples \citep{Leuenberger2010}.
-To assess the robustness of each of the four models, we also analyzed the
+To assess the robustness of each of the four models, I also analyzed the
 datasets simulated under the other three models.
-Overall, for each model, we produced 200,000 posterior estimates,
+Overall, for each model, I produced 200,000 posterior estimates,
 50,000 from the datasets simulated under that model,
@@
 
-For each set of 50,000 simulated datasets, we used the posterior estimates
+For each set of 50,000 simulated datasets, I used the posterior estimates
 to assess the model-choice behavior of each model.
-We did this by assigning the 50,000 estimates of the posterior probability
+I did this by assigning the 50,000 estimates of the posterior probability
 of one divergence event to 20 bins of width 0.05, and plotted
@@
 event \citep{Huelsenbeck2004,Oaks2012}.
-We did this using two criteria for the one divergence model:
+I did this using two criteria for the one divergence model:
 (1) the number of divergence time parameters ($\divTimeNum = 1$) and
@@
 0.01$][]{Hickerson2006,Huang2011}.
-We used the one-divergence model to assess model-choice behavior, because
+I used the one-divergence model to assess model-choice behavior, because
 it is often of biogeographic interest and is easily comparable among
@@
 
-In addition to the four models above, we also assessed the behavior of a model
+In addition to the four models above, I also assessed the behavior of a model
 with ordered divergence models (i.e., the order of the taxon-specific summary
 statistic vectors were maintained for the observed and simulated datasets).
-For this, we used a model with identical priors as \modelDPP, but that samples
+For this, I used a model with identical priors as \modelDPP, but that samples
 over ordered divergence models.
-We denote this model as \modelDPPOrdered.
-We simulated $1\e6$ prior samples and 50,000 datasets, and
+I denote this model as \modelDPPOrdered.
+I simulated $1\e6$ prior samples and 50,000 datasets, and
 analyzed them as above.
-We were not able to analyze the simulated datasets of the other models under
+I was not able to analyze the simulated datasets of the other models under
 the ordered model, because the identity of the population pairs is not
@@
 \subsection{Assessing power}
-We evaluated the power of the same four models (Table~\ref{tabPriors}) to
+I evaluated the power of the same four models (Table~\ref{tabPriors}) to
 detect random variation in divergence times using methods similar to
@@
 $\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
-we use the prior
+I use the prior
 $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
 over the discrete divergence models for the model \modelDPP.
-The reason for this difference is because for all of our power simulations, we
+The reason for this difference is because for all of the power simulations, I
 used a simulated data structure identical to that of the empirical dataset of
@@
 of populations.
-For each of the four models, we generated $2\e6$ samples from the prior.
-We then simulated datasets from three series of models
+For each of the four models, I generated $2\e6$ samples from the prior.
+I then simulated datasets from three series of models
 (Table~\ref{tabPowerModels}).
-One series of models, which we denote \powerSeriesOld, is identically
+One series of models, which I denote \powerSeriesOld, is identically
 distributed as \modelOld except that the divergence times for each of the 22
@@
 
-For each of the six models in each of the three series of models, we simulated
+For each of the six models in each of the three series of models, I simulated
 1000 datasets (18,000 datasets in total).
-We then analyzed each simulated dataset under all four prior models
+I then analyzed each simulated dataset under all four prior models
 (Table~\ref{tabPriors}), producing 72,000 posterior estimates, each with 1000
 samples.
-We also estimated a GLM-regression adjusted posterior from each of the
+I also estimated a GLM-regression adjusted posterior from each of the
 posterior samples \citep{Leuenberger2010}.
@@
 \subsection{An empirical application}
-We also assessed the behavior of the newly implemented models when applied to
+I also assessed the behavior of the newly implemented models when applied to
 the empirical dataset of \citet{Oaks2012} comprised of sequence data from 22
 pairs of taxa from the Philippine Islands (Drayad DOI: 10.5061/dryad.5s07m).
-We analyzed these data under five different models, which are detailed in
+I analyzed these data under five different models, which are detailed in
 Table~\ref{tabEmpiricalModels}.
@@
 is identical to \empModelDPP.
-To assess sensitivity to parameterization, we also applied the simplest
+To assess sensitivity to parameterization, I also applied the simplest
 possible model under the new implementation (\empModelDPPSimple) with only a
@@
 $\popSampleSize{i}{} - 1$ coalescent times.
-We also applied the original \msb model (\empModelOld) with priors
+I also applied the original \msb model (\empModelOld) with priors
 selected to make it comparable to the \empModelDPP model;
@@
 timescale.
-We also applied a model with a uniform distribution over divergence models
+I also applied a model with a uniform distribution over divergence models
 (\empModelUniform).
-For each of these models, we simulated 2\e{7} samples from the prior, and
+For each of these models, I simulated 2\e{7} samples from the prior, and
 retained an approximate posterior of the 10,000 samples with the smallest
@@
 To compare models that sample over ordered versus unordered models of
-divergence, we also analyzed the data from the subset of nine-taxon pairs that
+divergence, I also analyzed the data from the subset of nine-taxon pairs that
 are sampled from the Islands of Negros and Panay in the Philippines.
-The model we used for these analyses had a Dirichlet-process prior over
+The model I used for these analyses had a Dirichlet-process prior over
 divergence models and two demographic parameters (\ancestralTheta{} and
@@
 for details).
-One of the models, which we denote \npModelDPPOrdered, maintained the identity
+One of the models, which I denote \npModelDPPOrdered, maintained the identity
 of the taxon pairs and sampled over ordered models of divergence, while the
@@
 divergence.
-For both analyses, we simulated 5\e{7} samples from the prior and retained an
+For both analyses, I simulated 5\e{7} samples from the prior and retained an
 approximate posterior of 10,000 samples.
@@
 \subsection{Validation analyses: Model-choice accuracy}
-The \msb model, and our modification of it, is a model-choice method
+The \msb model, and my modification of it, is a model-choice method
 with the primary purpose of estimating the probabilities of models
@@
 estimate the posterior probabilities of divergence models.
-Consistent with the findings of \citet{Oaks2012}, our results demonstrate that
+Consistent with the findings of \citet{Oaks2012}, my results demonstrate that
 the unadjusted estimates of divergence-model posterior probabilities are
@@
 well illustrated in Figure S\ref{figValidationModelChoiceOmegaGlm}.
-As a result, we will focus our discussion of the results on the unadjusted
+As a result, I will focus my discussion of the results on the unadjusted
 estimates.
 
-We find that all four models accurately estimate the posterior probability of
+I find that all four models accurately estimate the posterior probability of
 the one-divergence model when applied to their own datasets (i.e., when the
@@
 \subsection{Validation analyses: Ordered divergence models}
-Our results show that the method performs similarily when sampling over ordered
+The results show that the method performs similarily when sampling over ordered
 models of divergence
@@
 \subsection{Power analyses: Estimation accuracy}
-All of the models we evaluated (Table \ref{tabPriors}) struggle to estimate the
+All of the models I evaluated (Table \ref{tabPriors}) struggle to estimate the
 variance of divergence times \divTimeDispersion regardless of which of the
@@
 \subsection{Power analyses: Model choice}
-Our modifications of the \msb model decrease the method's bias toward clustered
+The modifications of the \msb model decrease the method's bias toward clustered
 divergences when applied to data generated under random divergence times
@@
 
-As mentioned above, we also see increased power of the new models when
+As mentioned above, I also see increased power of the new models when
 looking at the estimated posterior probability of the one-divergence model
@@
 
-Lastly, when we look at the estimated posterior probability of
+Lastly, when I look at the estimated posterior probability of
 \divTimeDispersion being consistent with one shared divergence
-($p(\divTimeDispersion < 0.01 | \ssSpace)$), we see the same pattern of model
+($p(\divTimeDispersion < 0.01 | \ssSpace)$), I see the same pattern of model
 behavior, with \modelDPP and \modelUniform inferring low probabilities across
@@
 The disparity between the results of the \empModelOld model and the new models
-is even more pronounced when we look at the 10 divergence models
+is even more pronounced when I look at the 10 divergence models
 (\divTimeIndexVector) estimated to have the highest probability under each of the
@@
 \section{Discussion}
-Our results demonstrate that using alternative priors on parameters and
+My results demonstrate that using alternative priors on parameters and
 divergence models improved the behavior of the \msb model.
-In our implementation, model-choice estimation is more accurate and shows
+In my implementation, model-choice estimation is more accurate and shows
 greater robustness to model violations
@@
 
-Our modifications also improve the method's power to detect random
+The modifications also improve the method's power to detect random
 variation in divergence times, reducing the tendency to estimate clustered
@@
 (Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
-Our results are similar to those of \citet{Oaks2012} in that we find
+My results are similar to those of \citet{Oaks2012} in that I find
 \msb will often infer strong support for clustered divergences when
@@
 (Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
-Our results expand on this by showing that this behavior is consistent
+My results expand on this by showing that this behavior is consistent
 across a range of conditions underlying the data.
@@
 By evaluating a model intermediate between the old and new implementation
-(\modelUshaped), we were able to determine the relative effects of our
+(\modelUshaped), I was able to determine the relative effects of my 
 modifications to the model.
-Across all of our analyses, the results show that using better priors on
+Across all of the analyses, the results show that using better priors on
 divergence-time and demographic parameters alone does improve the performance
@@
 (Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
-Throughout our analyses, the intermediate model (\modelUshaped) performs better
+Throughout the analyses, the intermediate model (\modelUshaped) performs better
 than the \msb model, but not as well as the models with alternative priors on
@@
 
-While our modifications improve the behavior of the model, we urge caution when
+While the modifications improve the behavior of the model, I urge caution when
 using the method and interpreting its results.
@@
 (depending on \divTimeNum), and samples over 1002 unordered divergence models.
-Even under the simplest possible model allowed under our new implementation,
+Even under the simplest possible model allowed under the new implementation,
 the model still has 471--492 free parameters.
@@
 from the sequence alignments of each taxon pair.
-The population-genetic summary statistics used by our method contain little
+The population-genetic summary statistics used by the method contain little
 information about many of the free parameters in the model.
-Thus, we expect the improved method will still be sensitive to priors, and the
+Thus, I expect the improved method will still be sensitive to priors, and the
 power, while improved, may still be low.
-While we see much less prior sensitivity under the new model compared to those
-observed by \citet{Oaks2012}, we still see an effect when comparing the results
+While I see much less prior sensitivity under the new model compared to those
+observed by \citet{Oaks2012}, I still see an effect when comparing the results
 of the empirical data analyzed under a diffuse (\empModelDPP) and informative
@@
 \citep{Barber2010,Bell2012,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012},
-as we found here (Figure~\ref{figPhilippines}).
+as I found here (Figure~\ref{figPhilippines}).
 Given the richness of the model, the variance of the processes being modeled,
 and the relatively small amount of information in the summary statistics
-calculated from the sequence data, we do not expect to find strong posterior
+calculated from the sequence data, I do not expect to find strong posterior
 support for any scenario.
@@
 
-We also urge caution when using \dppmsbayes due to the fact that ABC methods
+I also urge caution when using \dppmsbayes due to the fact that ABC methods
 in general are known to be biased for model choice \citep{Robert2011}.
@@
 accompanied by assessments of prior sensitivity and power.
-We encourage investigators to view this method as a means of exploring their
+I encourage investigators to view this method as a means of exploring their
 data rather than a rigorous means of evaluating hypotheses.
 
-Given the difficulty of this estimation problem, we anticipate that
+Given the difficulty of this estimation problem, I anticipate that
 full-likelihood methods that can leverage all of the information present in the

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes.tex
Changing first person to singular in dpp ms.

@@
     that can be estimated using genetic data from co-distributed species.
-    We introduce a new approximate-Bayesian model for comparative
+    I introduce a new approximate-Bayesian model for comparative
     phylogeographical model-choice that estimates the temporal distribution of
     divergences across taxa from multi-locus DNA sequence data.
-    Our model is an extension of that implemented in \msb.
+    The model is an extension of that implemented in \msb.
     By reparameterizing the model, introducing more flexible priors on
     demographic and divergence-time parameters, and implementing a
-    non-parametric Dirichlet-process prior over divergence models, we improve
+    non-parametric Dirichlet-process prior over divergence models, I improve
     the robustness, accuracy, and power of the method for estimating shared
     evolutionary history across taxa.
-    Our results demonstrate the improved performance of the new method is due
+    My results demonstrate the improved performance of the new method is due
     to (1) more appropriate priors on divergence-time and demographic
@@
     of divergence events.
-    % Our results demonstrate that the tendency of the \msb model to spuriously
+    % The results demonstrate that the tendency of the \msb model to spuriously
     % support models of clustered divergences is caused by a combination of (1)

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Making some edits to the DPP manuscript.

@@
 I follow much of the notation of \citet{Oaks2012}, but modify it to
-aid in my description of the new model.
+aid in the description of the new model.
 A summary of my notation can be found in Table \ref{tabNotation}.
@@
 i.e., a vector of sequence alignments for all pairs and loci.
-I let \geneTree{i}{j} represent the gene tree upon which \alignment{i}{j}
+Let \geneTree{i}{j} represent the gene tree upon which \alignment{i}{j}
 evolved according to fixed \hky substitution model parameters \hkyModel{i}{j}.
@@
 \begin{linenomath}
-The prior term for the vector of divergence times for each of
-the \npairs{} pairs of populations can be expanded as
+The prior term for the vector of divergence times for each of the \npairs{}
+pairs of populations, \divTimeMapVector, can be expanded as
 \begin{equation}
@@
 \end{equation}
-where \divTimeVector is an ordered set of unique divergence-time parameters
+where \divTimeVector is an ordered set of divergence-time parameters
 $\{\divTime{1}, \ldots, \divTime{\divTimeNum}\}$ whose length
 \divTimeNum can range from 1 to \npairs{},
-and \divTimeIndexVector is the vector of indices of divergence-time parameters 
-$(\divTimeIndex{1}, \ldots, \divTimeIndex{\npairs{}})$
-that map the divergence times in \divTimeVector to each of the \npairs{}
-pairs of populations.
+and \divTimeIndexVector is a vector of indices %of divergence-time parameters 
+$(\divTimeIndex{1}, \ldots, \divTimeIndex{\npairs{}})$,
+where
+$\divTimeIndex{i} \in \{1, \ldots, \divTimeNum\}$.
+These indices map each of the \npairs{} pairs of populations to a
+divergence-time parameter in \divTimeVector.
 Thus, \divTimeMapVector is the result of applying the mapping function
@@
 the \npairs{} pairs of populations.
-For example, if \divTimeVector contains a single divergence time parameter
+For example, if \divTimeVector contains a single divergence-time parameter
 \divTime{1}, all \npairs{} pairs of populations are constrained to diverge at
@@
 \npairs{} times, and \divTimeMapVector would contain the value \divTime{1}
-repeated \npairs{} times), whereas if it contains \npairs{} divergence time
+repeated \npairs{} times), whereas if it contains \npairs{} divergence-time
 parameters, the model is fully generalized to allow all of the pairs to diverge

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes.tex
Updating author line of dpp manuscript.
Making some edits to the DPP manuscript.

@@
 \author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
-\author[1]{Mark T.\ Holder}
+% \author[1]{Mark T.\ Holder}
 \affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
@@
     demographic and divergence-time parameters, and implementing a
-    non-parametric Dirichlet-process prior over divergence models, I improve
+    non-parametric Dirichlet-process prior over divergence models, I improved
     the robustness, accuracy, and power of the method for estimating shared
     evolutionary history across taxa.
-    My results demonstrate the improved performance of the new method is due
+    The results demonstrate the improved performance of the new method is due
     to (1) more appropriate priors on divergence-time and demographic
@@
 \hangindent=1cm
-Oaks, J.~R.\ and M.~T.\ Holder. \msTitle.
+Oaks, J.\ R. \msTitle.
 

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More edits to dpp manuscript.

@@
 
+Unlike the model implemented in \msb, here I place priors on
+\divTimeIndexVector and \divTime{}, rather than \divTimeNum and
+\divTime{}.
+As a result, \divTimeIndexVector determines the number of divergence-time
+parameters (\divTimeNum) in the model.
+Below, I first describe the prior used for \divTime{} and the timescale it
+imposes on the model before discussing the priors implemented for
+\divTimeIndexVector.
+
 \begin{linenomath}
@@
 \begin{equation}
-    \myTheta{C} = \mathbb{E}(\dThetas),
+    % \myTheta{C} = \mathbb{E}(\dThetas),
+    \myTheta{C} = \mathbb{E}(\descendantTheta{}{}),
     \label{eq:thetaC}
 \end{equation}
-where \popIndex is the descendant-population index which can be either 1 or 2.
+% where \popIndex is the descendant-population index which can be either 1 or 2.
 Given the size of the descendant populations are \iid as
-$\dThetas \sim Gamma(\gammaShape{\descendantTheta{}{}},
+$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+\gammaScale{\descendantTheta{}{}})$ and
+$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
 \gammaScale{\descendantTheta{}{}})$,
@@
 scale, the relative mutation rates among the pairs of populations are assumed
-to be fixed and known according to the user-provided values in
+to be known and fixed according to the user-provided values in
 \mutationRateScalarConstantVector.
@@
 As described by \citet{Oaks2012}, to get the divergence times in units
-proportional to the expected number of mutations, I must
-scale them by the realized population size for locus $j$ of population-pair $i$ 
+proportional to the expected number of mutations, they must be
+scaled by the realized population size for locus $j$ of population-pair $i$ 
 \begin{equation}
@@
 $\ancestralTheta{} \sim Exp(mean=0.025)$,
+$\descendantTheta{1}{} \sim Exp(mean=0.025)$,
 and
-$\dThetas \sim Exp(mean=0.025)$.
+$\descendantTheta{2}{} \sim Exp(mean=0.025)$.
 (3) a model with uniform prior probability over unordered divergence models,
@@
 $\ancestralTheta{} \sim Exp(mean=0.025)$,
+$\descendantTheta{1}{} \sim Exp(mean=0.025)$,
 and
-$\dThetas \sim Exp(mean=0.025)$;
+$\descendantTheta{2}{} \sim Exp(mean=0.025)$.
 and
@@
 $\ancestralTheta{} \sim Exp(mean=0.025)$,
+$\descendantTheta{1}{} \sim Exp(mean=0.025)$,
 and
-$\dThetas \sim Exp(mean=0.025)$.
+$\descendantTheta{2}{} \sim Exp(mean=0.025)$.
 I selected the exponential prior on divergence time used in models \modelDPP,

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Edit to DPP section of DPP ms.

@@
 regarding divergence models.
-The \concentrationParam determines the prior probability that any two
-pairs of populations $i$ and $j$ will be assigned to the same divergence time
-parameter
+The concentration parameter \concentrationParam determines the prior
+probability that any two pairs of populations $i$ and $j$ will be assigned to
+the same divergence-time parameter
 \begin{equation}

### 2014-02-19 prior-tests/bin/generate-priors-no-sort.sh
Adding `prior-tests`.

@@ -1,0 +1,51 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=16G
+#$ -l vf=16G
+#$ -q all.q
+#$ -pe orte 8
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
+
+nprocs=8
+nprior=10000
+batch_size=100
+nsums=10000
+seed=37851841
+
+output_dir="../priors"
+if [ ! -d "$output_dir" ]
+then
+    mkdir -p $output_dir
+fi
+
+dmc.py --np $nprocs \
+    -r 1 \
+    -o ../configs/equal.cfg \
+    -p ../configs/*.cfg \
+    -n $nprior \
+    --prior-batch-size $batch_size \
+    --num-standardizing-samples $nsums \
+    --sort-index 0 \
+    --output-dir $output_dir \
+    --staging-dir $staging_dir \
+    --temp-dir $staging_dir \
+    --seed $seed \
+    --compress \
+    --generate-samples-only
+
+echo "Here are the contents of the local temp directory '${staging_dir}':"
+ls -Fla $staging_dir
+echo 'Removing the local temp directory...'
+rm -r $staging_dir
+echo 'Done!'
+

### 2014-02-19 prior-tests/bin/generate-priors-no-sort.sh
Updating prior-generating script for prior-tests.

@@ -40,7 +40,6 @@ dmc.py --np $nprocs \
     --staging-dir $staging_dir \
     --temp-dir $staging_dir \
     --seed $seed \
-    --compress \
     --generate-samples-only
 
 echo "Here are the contents of the local temp directory '${staging_dir}':"

### 2014-02-20 prior-tests/bin/generate-priors-no-sort.sh
Updating `test-prior` shell command script.
It now uses the new `dmc_plot_stats.py` script of the `PyMsBayes`
package.

@@ -1,50 +1,0 @@
-#! /bin/sh
-#$ -S /bin/bash
-#$ -cwd
-#$ -V
-#$ -l h_vmem=16G
-#$ -l vf=16G
-#$ -q all.q
-#$ -pe orte 8
-
-if [ -n "$SGE_O_WORKDIR" ]
-then
-    source ~/.bash_profile
-    cd /share/work1
-    cd $SGE_O_WORKDIR
-fi
-
-staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
-
-nprocs=8
-nprior=10000
-batch_size=100
-nsums=10000
-seed=37851841
-
-output_dir="../priors"
-if [ ! -d "$output_dir" ]
-then
-    mkdir -p $output_dir
-fi
-
-dmc.py --np $nprocs \
-    -r 1 \
-    -o ../configs/equal.cfg \
-    -p ../configs/*.cfg \
-    -n $nprior \
-    --prior-batch-size $batch_size \
-    --num-standardizing-samples $nsums \
-    --sort-index 0 \
-    --output-dir $output_dir \
-    --staging-dir $staging_dir \
-    --temp-dir $staging_dir \
-    --seed $seed \
-    --generate-samples-only
-
-echo "Here are the contents of the local temp directory '${staging_dir}':"
-ls -Fla $staging_dir
-echo 'Removing the local temp directory...'
-rm -r $staging_dir
-echo 'Done!'
-

### 2014-02-20 prior-tests/bin/plot.sh

@@ -1,0 +1,22 @@
+#! /bin/bash
+
+cfg_dir="../configs"
+plot_dir="../plots"
+nproc=8
+if [ ! -d "$plot_dir" ]
+then
+    mkdir -p $plot_dir
+fi
+
+for config in ${cfg_dir}/*.cfg
+do
+    conf="$(basename $config)"
+    prefix=${conf/\.cfg/}
+    out_dir=${plot_dir}/${prefix}
+    if [ ! -d "$out_dir" ]
+    then
+        mkdir $out_dir
+    fi
+    dmc_plot_stats.py -c "$config" -n 1000 --np $nproc --compress -s pi -o "$out_dir" --seed 199065646
+done
+

### 2014-02-21 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More editing of the DPP ms.

@@
 \end{equation}
-The \divTime{} parameters are in units of $\myTheta{C}/\mutationRate$
-generations, which I denote as \globalcoalunit generations.
+More specifically, the \divTime{} parameters are in units of
+$\myTheta{C}/\mutationRate$ generations, which I denote as \globalcoalunit
+generations.
 Thus, each \divTime{} within \divTimeVector is proportional to time and can be
@@
 
-The second prior is based on the Dirichlet-process, which is a
-stochastic process that groups elements into an unknown number of discrete
+The second prior is based on the Dirichlet process, which is a stochastic
+process that groups random variables into an unknown number of discrete
 parameter classes \citep{Ferguson1973,Antoniak1974}.
-The process is controlled by the concentration parameter \concentrationParam,
-which determines how clustered the process will be.
-I use the Dirichlet process to place a prior over all possible ordered
-partitions of \npairs{} elements, which I denote as $\divTimeIndexVector \sim
-\priorDPP{}$.
+The Dirichlet process has been used as a nonparametric Bayesian approach to
+many inference problems in evolutionary biology
+\citep{Lartillot2004,Huelsenbeck2007,HuelsenbeckSuchard2007,Ane2007,Heath2011,Heath2012}.
+Here, I use the Dirichlet process to place a prior over all possible ordered
+partitions of \npairs{} population pairs into divergence-time parameter classes
+(i.e., ``divergence events'').
+As discussed above, the time of each divergence-time parameter is drawn from
+the base distribution $\divTime{} \sim Gamma(\gammaShape{\divTime{}},
+\gammaScale{\divTime{}})$.
+The partitioning of the population pairs to divergence-time classes is
+controlled by the concentration parameter \concentrationParam, which determines
+how clustered the process will be.
 I take a hierarchical approach and use a prior probability
-distribution (i.e., hyperprior) for \concentrationParam.
+distribution (i.e., hyperprior) for \concentrationParam \citep{Escobar1995}.
 More specifically, I use a gamma-distributed prior $\concentrationParam \sim
@@
 specified by the investigator.
+I use $\divTimeIndexVector \sim \priorDPP{}$ to denote this Dirichlet-process
+prior.
 
@@
 
-\subsection{Differences between my model and the original \texttt{msBayes} model}
+\subsection{Differences between this model and the original \texttt{msBayes} model}
 \subsubsection{The prior on divergence models}
@@
 \subsubsection{Flexibility in parameterizing the model}
-In my implementation, I provide the ability to control the richness of the
+In the new implementation, I provide the ability to control the richness of the
 model.
@@
 (\myTheta{}),
-in addition to the $\popSampleSize{i}{} - 1$ coalescence-time parameters.
+in addition to the $\popSampleSize{i}{} - 1$ coalescence-time parameters
+(i.e., the node heights of the gene tree).
 
 \subsubsection{Time scale}
-As described above, in my model divergence times are in units of
+As described above, divergence times are in units of
 $\myTheta{C}/\mutationRate$ generations, where \myTheta{C} is the expectation
@@
 sizes.
+This is only equal to the expectation of the prior if the lower limit is zero.
 
@@
 of divergence time parameters (\divTimeNum).
-These have been denoted as \meant{}{}, \vart{}{}, \vmratio{}, and \numt{},
-respectively \citep{Hickerson2006,Huang2011,Oaks2012}.
-I use the new notation to avoid confusion that the values represent prior or
-posterior expectations, and to indicate that \vmratio{} is not a parameter of
-the model and \numt{} is no longer a parameter in the new implementation.
+Previously, these have been denoted as \meant{}{}, \vart{}{}, \vmratio{}, and
+\numt{}, respectively \citep{Hickerson2006,Huang2011,Oaks2012}.
+I use \divTimeMean and \divTimeVar in place \meant{}{} and \vart{}{} to make
+clear that these values do not represent the prior or posterior
+expectation/variance of divergence times.
+I use \divTimeDispersion in place of \vmratio{} to clarify that this is a
+statistic rather than a parameter of the model.
+Lastly, I use \divTimeNum in place of \numt{}, because the number of
+divergence-time parameters is no longer a parameter in the new implementation.
 
@@
 The original descriptions of the \msb model claim that this re-ordering is
-justified by the fact that of $\pi_b$ (and other summary statistics) are
-unrelated to the sample sizes \popSampleSize{1}{} and \popSampleSize{2}{} of
-each pair and are thus exchangeable \citep{Hickerson2006,Huang2011}.
-This is actually incorrect for two reasons.
-First, the expectation of $\pi_b$ is not independent of samples sizes.
-If there are more than one coalescent events in the ancestor, which is expected
-to be common on phylogeographic timescales, more samples will increase the
-probability of capturing these deeper events, and thus affect the average
-pairwise differences between the descendant populations.
-Also, other statistics that estimate gross diversity (e.g., $\pi$ and
-$\theta_W$) are clearly not independent of sample size.
-Second, for variables to be exchangeable, they do not need to be independent,
+justified by the fact that the expected value of $\pi_b$ is
+unrelated to sample size \popSampleSize{1}{} and \popSampleSize{2}{} and
+thus exchangeable among pairs \citep{Hickerson2006,Huang2011}.
+This is actually incorrect for three reasons.
+First, the independence of the expected value of $\pi_b$ and sample size
+breaks down as $\pi_b$ approaches its lower limit of zero (i.e., very recent or
+no divergence between populations), which might be common given the timescales
+of most phylogeographical studies.
+Second, the entire vector of summary statistics \alignmentSS{}{} for each pair
+of populations is re-ordered across pairs, which implies that the justification
+for re-ordering $\pi_b$ applies to all the statistics within each
+\alignmentSS{}{}.
+% First, the expectation of $\pi_b$ is not independent of samples sizes.
+% If there are more than one coalescent events in the ancestor, which is expected
+% to be common on phylogeographic timescales, more samples will increase the
+% probability of capturing these deeper events, and thus affect the average
+% pairwise differences between the descendant populations.
+However, the expectations for statistics that estimate gross diversity (e.g.,
+$\pi$ and $\theta_W$) are not independent of sample size for structured
+populations, and other statistics are not independent of sample size in general
+(e.g., $SD(\pi-\theta_W)$).
+Third, for variables to be exchangeable, they do not need to be independent,
 but their marginal distributions must be the same (i.e., they must be
@@
 is to minimize Euclidean distances between simulated datasets in which the true
-model has a single divergence event, and (2) the reordering increases the
-model's tendency to infer a single divergence event \citep{Huang2011}, it seems
-possible that this approach could be biasing the method \citep{Oaks2012}.
-
-To maintain compatibility and comparability with \msb I maintain
-the re-ordering of taxon-specific summary statistics by $\pi_b$.
-However, I also allow the order to be maintained, and ordered
-divergence models to be estimated.
+model has a single divergence event \citep{Huang2011}, and (2) the reordering
+increases the model's tendency to infer a single divergence event
+\citep{Huang2011}, it seems possible that this approach could be biasing the
+method \citep{Oaks2012}.
+
+For comparability with \msb, I maintain the re-ordering of taxon-specific
+summary statistics by $\pi_b$.  However, I also allow the order to be
+maintained, and ordered divergence models to be estimated.
 
@@
 \citep{Leuenberger2010} as implemented in \abctoolbox
-\citep[v1.1;][]{ABCtoolbox}, which \citet{Oaks2012} showed performed very
+\citep[v1.1;][]{ABCtoolbox}, which \citet{Oaks2012} showed performs very
 similarly to weighted local-linear regression and multinomial logistic
@@
 I selected four model priors for these analyses (Table~\ref{tabPriors}):
-(1) A model to represent the original \msb model, \modelOld, with
-priors
-$\divTimeIndexVector \sim \priorOld$,
-$\divTime{} \sim U(0,10)$,
-$\ancestralTheta{} \sim U(0, 0.05)$,
-and
-$\descendantThetaMean{} \sim U(0, 0.05)$;
-(2) a model with the \msb prior on unordered divergence models, but with
-exponential priors on nuisance parameters,
-\modelUshaped, with priors
-$\divTimeIndexVector \sim \priorOld$,
-$\divTime{} \sim Exp(mean=2.887)$,
-$\ancestralTheta{} \sim Exp(mean=0.025)$,
-$\descendantTheta{1}{} \sim Exp(mean=0.025)$,
-and
-$\descendantTheta{2}{} \sim Exp(mean=0.025)$.
-(3) a model with uniform prior probability over unordered divergence models,
-\modelUniform, with priors
-$\divTimeIndexVector \sim \priorUniform$,
-$\divTime{} \sim Exp(mean=2.887)$,
-$\ancestralTheta{} \sim Exp(mean=0.025)$,
-$\descendantTheta{1}{} \sim Exp(mean=0.025)$,
-and
-$\descendantTheta{2}{} \sim Exp(mean=0.025)$.
-and
-(4) a Dirichlet-process prior model, \modelDPP, with priors
-$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
-$\divTime{} \sim Exp(mean=2.887)$,
-$\ancestralTheta{} \sim Exp(mean=0.025)$,
-$\descendantTheta{1}{} \sim Exp(mean=0.025)$,
-and
-$\descendantTheta{2}{} \sim Exp(mean=0.025)$.
+\begin{description}
+    \item[\modelOld] a model to represent the original \msb implementation with
+        the U-shaped prior on unordered divergence models and uniform priors on
+        divergence-time and demographic parameters; $\divTimeIndexVector \sim
+        \priorOld$, $\divTime{} \sim U(0,10)$, $\ancestralTheta{} \sim U(0,
+        0.05)$, and $\descendantThetaMean{} \sim U(0, 0.05)$
+    \item[\modelUshaped] a model with the U-shaped prior of \msb on unordered
+        divergence models, but with exponential priors on divergence-time and
+        demographic parameters; $\divTimeIndexVector \sim \priorOld$,
+        $\divTime{} \sim Exp(mean=2.887)$, $\ancestralTheta{} \sim
+        Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$, and
+        $\descendantTheta{2}{} \sim Exp(mean=0.025)$
+    \item[\modelUniform] a model with a uniform prior over unordered divergence
+        models and exponential priors on divergence-time and demographic
+        parameters; $\divTimeIndexVector \sim \priorUniform$, $\divTime{} \sim
+        Exp(mean=2.887)$, $\ancestralTheta{} \sim Exp(mean=0.025)$,
+        $\descendantTheta{1}{} \sim Exp(mean=0.025)$, and
+        $\descendantTheta{2}{} \sim Exp(mean=0.025)$
+    \item[\modelDPP] a model with a Dirichlet-process prior on ordered
+        divergence models and exponential priors on divergence-time and
+        demographic parameters; $\divTimeIndexVector \sim \priorDPP{\sim
+        Gamma(2,2)}$, $\divTime{} \sim Exp(mean=2.887)$, $\ancestralTheta{}
+        \sim Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$,
+        and $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
+\end{description}
 I selected the exponential prior on divergence time used in models \modelDPP,
@@
 parameters for each population pair, two uniformly distributed ($beta(1,1)$)
-\bottleScalar{}{} parameters per pair, no migration, no recombination,
-and unordered divergence models.
-For all of my simulations, I used a simulated data structure of eight
-population pairs, with a single 1000 base-pair locus sampled from 10
-individuals from each population.
+\bottleScalar{}{} parameters per pair, no migration, no recombination, and
+resorting of taxon-specific summary statistics by $\pi_b$ (i.e., sampling
+unordered divergence models).
+For all simulations, I used a data structure of eight population pairs, with a
+single 1000 base-pair locus sampled from 10 individuals from each population.
 
@@
 50,000 from the datasets simulated under that model,
-and 150,000 estimated from the datasets simulated under the
+and 150,000 from the datasets simulated under the
 other three models.
@@
 (2) the dispersion index of divergence times \citep[$\divTimeDispersion <
-0.01$][]{Hickerson2006,Huang2011}.
+0.01$;][]{Hickerson2006,Huang2011}.
 I used the one-divergence model to assess model-choice behavior, because
@@
 In addition to the four models above, I also assessed the behavior of a model
-with ordered divergence models (i.e., the order of the taxon-specific summary
-statistic vectors were maintained for the observed and simulated datasets).
-For this, I used a model with identical priors as \modelDPP, but that samples
-over ordered divergence models.
+that samples over ordered divergence models (i.e., the order of the
+taxon-specific summary statistic vectors were maintained for the observed and
+simulated datasets); all other settings were identical to the \modelDPP model.
 I denote this model as \modelDPPOrdered.

### 2014-02-21 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More editing of dpp ms.

@@
 I implement two prior probability distributions over the space
-of all possible discrete divergence models (\divTimeIndexVector).
+of all possible divergence models (\divTimeIndexVector).
 The first simply gives all possible unordered partitions of \npairs{} elements
@@
 i.e., a discrete uniform prior over all the integer partitions of \npairs{}
-(discrete divergence models).
+(unordered divergence models).
 I denote this prior as
@@
 divergence parameters (denoted \divTimeNum here, and as \numt{} in
-\citet{Huang2011}) imposes an U-shaped prior over discrete divergence
+\citet{Huang2011}) imposes an U-shaped prior over divergence
 models (\divTimeIndexVector; see
@@
 \citep[$\pi_b$;][]{NeiLi1979,Huang2011}.
-This has the advantage of reducing the sample space of the number of discrete
+This has the advantage of reducing the sample space of the number of
 divergence models \divTimeIndexVector, but there are at least two disadvantages.
@@
 \citet{Oaks2012}.
-The only difference from the prior models used in the validation analyses above
-is the prior on the concentration parameter \concentrationParam for the
-\modelDPP model.
-Rather than the prior
-$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
-I use the prior
-$\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
-over the discrete divergence models for the model \modelDPP.
-The reason for this difference is because for all of the power simulations, I
-used a simulated data structure identical to that of the empirical dataset of
-Philippine vertebrates analyzed by \citet{Oaks2012}, which consists of 22 pairs
-of populations.
+For all power simulations, I used a data structure identical to that of the
+empirical dataset of Philippine vertebrates analyzed by \citet{Oaks2012}, which
+consists of 22 pairs of populations.
+Due to the larger number of pairs, I used a different hyperprior on the
+concentration parameter for the \modelDPP model; I used a prior of
+$\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ over divergence
+models for the model \modelDPP.
+All other aspects of the four models in Table~\ref{tabPriors} were identical to
+those used in the validation analyses described above.
+
 For each of the four models, I generated $2\e6$ samples from the prior.

### 2014-02-21 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More editing of DPP ms.

@@
 for analysis (i.e., the prior is correct).
-I selected four model priors for these analyses (Table~\ref{tabPriors}):
-\begin{description}
-    \item[\modelOld] a model to represent the original \msb implementation with
+I selected the following four model priors for these analyses
+(Table~\ref{tabPriors}).
+\begin{enumerate}
+    \item The \modelOld model represents the original \msb implementation with
         the U-shaped prior on unordered divergence models and uniform priors on
@@
         \priorOld$, $\divTime{} \sim U(0,10)$, $\ancestralTheta{} \sim U(0,
-        0.05)$, and $\descendantThetaMean{} \sim U(0, 0.05)$
-    \item[\modelUshaped] a model with the U-shaped prior of \msb on unordered
+        0.05)$, and $\descendantThetaMean{} \sim U(0, 0.05)$.
+    \item The \modelUshaped model with the U-shaped prior of \msb on unordered
         divergence models, but with exponential priors on divergence-time and
@@
         Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$, and
-        $\descendantTheta{2}{} \sim Exp(mean=0.025)$
-    \item[\modelUniform] a model with a uniform prior over unordered divergence
-        models and exponential priors on divergence-time and demographic
-        parameters; $\divTimeIndexVector \sim \priorUniform$, $\divTime{} \sim
-        Exp(mean=2.887)$, $\ancestralTheta{} \sim Exp(mean=0.025)$,
-        $\descendantTheta{1}{} \sim Exp(mean=0.025)$, and
-        $\descendantTheta{2}{} \sim Exp(mean=0.025)$
-    \item[\modelDPP] a model with a Dirichlet-process prior on ordered
+        $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
+    \item The \modelUniform model with a uniform prior over unordered
+        divergence models and exponential priors on divergence-time and
+        demographic parameters; $\divTimeIndexVector \sim \priorUniform$,
+        $\divTime{} \sim Exp(mean=2.887)$, $\ancestralTheta{} \sim
+        Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$, and
+        $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
+    \item The \modelDPP model with a Dirichlet-process prior on ordered
         divergence models and exponential priors on divergence-time and
@@
         and $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
-\end{description}
+\end{enumerate}
 I selected the exponential prior on divergence time used in models \modelDPP,
@@
 those used in the validation analyses described above.
-
 For each of the four models, I generated $2\e6$ samples from the prior.
-I then simulated datasets from three series of models
-(Table~\ref{tabPowerModels}).
-One series of models, which I denote \powerSeriesOld, is identically
-distributed as \modelOld except that the divergence times for each of the 22
-pairs of populations are randomly drawn from a series of uniform distributions,
-$U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and
-2.0, in \globalcoalunit generations.
-A second series of models, \powerSeriesUniform, is identically distributed as
-\modelUniform and \modelDPP except that the 22 divergence times are randomly
-drawn from the same series of uniform priors as above.
-The third series of models, \powerSeriesExp, is also identically distributed as
-\modelUniform and \modelDPP except that the 22 divergence times are randomly
-drawn from a series of of exponential distributions:
-$Exp(mean=0.058)$, $Exp(mean=0.115)$, $Exp(mean=0.173)$, $Exp(mean=0.231)$,
-$Exp(mean=0.289)$, and $Exp(mean=0.577)$.
-These exponential distributions have the same variance as their uniform
+
+Next, I simulated datasets from three series of models in which the divergence
+times of the 22 pairs were random (i.e., no clustering; $\divTimeNum = 22$).
+The models comprising each series differ in the variance of the distribution
+from which the divergence times are randomly drawn.
+When the variance of random divergence times is small, all of the models in
+Table~\ref{tabPriors} are expected to struggle to detect this variation and
+will often incorrectly estimate highly clustered models of divergence (i.e.,
+few divergence events).
+The goal is to assess how much temporal variation in random divergence times is
+necessary before the behavior of the models of Table~\ref{tabPriors} begins to
+improve.
+This will determine the timescales over which the models can reliably detect
+random variation in divergence times and avoid spurious inference of
+biogeographically interesting clustered divergence models.
+
+Specifically, I simulated datasets from the following three series of six
+models (Table~\ref{tabPowerModels}).
+\begin{enumerate}
+    \item The \powerSeriesOld models are identically distributed as \modelOld
+        except the divergence times for each of the 22 pairs of
+        populations are randomly drawn from a series of uniform distributions,
+        $U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8,
+        1.0, and 2.0, in \globalcoalunit generations.
+    \item The \powerSeriesUniform models are identically distributed as
+        \modelUniform and \modelDPP except the 22 divergence times are
+        randomly drawn from the same series of uniform priors as above.
+    \item The \powerSeriesExp models are also identically distributed as
+        \modelUniform and \modelDPP except the 22 divergence times are
+        randomly drawn from a series of of exponential distributions:
+        $Exp(mean=0.058)$, $Exp(mean=0.115)$, $Exp(mean=0.173)$,
+        $Exp(mean=0.231)$, $Exp(mean=0.289)$, and $Exp(mean=0.577)$. These
+        exponential distributions have the same variance as their uniform
         counterparts in the first two series of models.
-
+\end{enumerate}
 For each of the six models in each of the three series of models, I simulated
@@
 I also assessed the behavior of the newly implemented models when applied to
-the empirical dataset of \citet{Oaks2012} comprised of sequence data from 22
-pairs of taxa from the Philippine Islands (Drayad DOI: 10.5061/dryad.5s07m).
+the empirical dataset of \citet{Oaks2012}, which is comprised of sequence data
+from 22 pairs of taxa from the Philippine Islands (Drayad DOI:
+10.5061/dryad.5s07m).
 I analyzed these data under five different models, which are detailed in
@@
 \empModelDPP, \empModelDPPInform, and \empModelDPPSimple.
-The \empModelDPP model is meant to represent the prior that \citet{Oaks2012}
-would have selected to represent their prior uncertainty about the parameters
-of the model if provided the more flexible distributions that are now
-implemented.
+The \empModelDPP model represents the priors that \citet{Oaks2012} would have
+selected to reflect their prior uncertainty about the parameters of the model
+if provided the more flexible distributions that are now implemented.
 To assess prior sensitivity, the \empModelDPPInform model uses a more
@@
 $\popSampleSize{i}{} - 1$ coalescent times.
-I also applied the original \msb model (\empModelOld) with priors
-selected to make it comparable to the \empModelDPP model;
+I also applied the original \msb model (\empModelOld) with priors selected to
+make the results directly comparable to those of the \empModelDPP model;
 the uniform prior on divergence times was selected to have the same variance as
-the exponential prior of the \empModelDPP, and the prior on population
+the exponential prior of the \empModelDPP model, and the prior on population
 size was selected to have the same mean so that the models are on the same
@@
 divergence models improved the behavior of the \msb model.
-In my implementation, model-choice estimation is more accurate and shows
+In the new implementation, model-choice estimation is more accurate and shows
 greater robustness to model violations

### 2014-02-22 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More editing of the DPP ms.

@@
 
-The original descriptions of the \msb model claim that this re-ordering is
-justified by the fact that the expected value of $\pi_b$ is
-unrelated to sample size \popSampleSize{1}{} and \popSampleSize{2}{} and
-thus exchangeable among pairs \citep{Hickerson2006,Huang2011}.
-This is actually incorrect for three reasons.
-First, the independence of the expected value of $\pi_b$ and sample size
-breaks down as $\pi_b$ approaches its lower limit of zero (i.e., very recent or
-no divergence between populations), which might be common given the timescales
-of most phylogeographical studies.
-Second, the entire vector of summary statistics \alignmentSS{}{} for each pair
+In the original descriptions of the \msb model \citep{Hickerson2006,Huang2011},
+this re-ordering is justified by the fact that the expected value of $\pi_b$ is
+unrelated to sample size \popSampleSize{1}{} and \popSampleSize{2}{} and thus
+exchangeable among pairs.
+However, this is incorrect for two reasons.
+First, the entire vector of summary statistics \alignmentSS{}{} for each pair
 of populations is re-ordered across pairs, which implies that the justification
@@
 \alignmentSS{}{}.
-% First, the expectation of $\pi_b$ is not independent of samples sizes.
-% If there are more than one coalescent events in the ancestor, which is expected
-% to be common on phylogeographic timescales, more samples will increase the
-% probability of capturing these deeper events, and thus affect the average
-% pairwise differences between the descendant populations.
 However, the expectations for statistics that estimate gross diversity (e.g.,
@@
 (e.g., $SD(\pi-\theta_W)$).
-Third, for variables to be exchangeable, they do not need to be independent,
-but their marginal distributions must be the same (i.e., they must be
-identically distributed).
-The simulated alignments and their summary statistics are not identically
-distributed because of differing sample sizes \emph{and} taxon- and
+Second, and more importantly, having the same expectation does not ensure
+random variables are exchangeable.
+Rather, for variables to be exchangeable their marginal distributions must be
+the same (i.e., they must be identically distributed).
+None of the summary statistics used by \msb, including $\pi_b$, are identically
+distributed across different sample sizes.
+In addition to differences in sample size, the simulated alignments and their
+summary statistics are not identically distributed because of the taxon- and
 locus-specific parameters \hkyModel{}{}, \mutationRateScalarConstant{}{}, and
@@
 I did this by assigning the 50,000 estimates of the posterior probability
-of one divergence event to 20 bins of width 0.05, and plotted
+of one-divergence event to 20 bins of width 0.05, and plotted
 the estimated posterior probability of each bin against the proportion of
@@
 event \citep{Huelsenbeck2004,Oaks2012}.
-I did this using two criteria for the one divergence model:
-(1) the number of divergence time parameters ($\divTimeNum = 1$) and
-(2) the dispersion index of divergence times \citep[$\divTimeDispersion <
-0.01$;][]{Hickerson2006,Huang2011}.
-I used the one-divergence model to assess model-choice behavior, because
+If the method is unbiased, the points should fall near the identity line,
+whereas any trend away from the identity line indicates bias.
+I did this using two criteria for the one-divergence model:
+(1) the number of divergence-time parameters ($\divTimeNum = 1$) and
+(2) the dispersion index of divergence times ($\divTimeDispersion < 0.01$).
+For the latter, $\divTimeDispersion < 0.01$ has been commonly used as an
+arbitrary criterion for a single ``simultaneous'' divergence event
+\citep[e.g.,][]{Hickerson2006,Leache2007,Stone2012}.
+I focused on the one-divergence model to assess model-choice behavior, because
 it is often of biogeographic interest and is easily comparable among
@@
 and \modelDPP) perform similarly when applied to datasets generated under all
-four of the models of Table \ref{tabPriors}
+four of the models in Table \ref{tabPriors}
 (Figure~S\ref{figValidationAccuracyOmega}).
@@
 statistic for all the models
-(Figure~S\ref{figValidationAccuracyPsiGlm}).
-
-The same general pattern is seen for estimates of \divTimeMean, with all four
-models performing similarly when applied to the data generated under the
-\modelOld model, the models with exponentially distributed priors performing
-similarly when applied to data generated under the other three models, and the
-\modelOld model is sensitive to model violations and is more biased whenever
-applied to data generated under other models
+(Figure~S\ref{figValidationAccuracyOmegaGlm}).
+
+The same general pattern is seen for estimates of \divTimeMean, with
+(1) all four models performing similarly when applied to the data generated
+under the \modelOld model,
+(2) the models with exponentially distributed priors performing similarly when
+applied to data generated under the other three models, and
+(3) the \modelOld model is sensitive to model violations and is more biased
+whenever applied to data generated under other models
 (Figure~S\ref{figValidationAccuracyTime}).
@@
 statistic tends to increase estimation bias; all of the models tend to
-underestimate \divTimeDispersion after the GLM-adjustment
+underestimate \divTimeNum after the GLM-adjustment
 (Figure~S\ref{figValidationAccuracyPsiGlm}).
@@
 \ref{figValidationModelChoiceOmega}).
-The \modelUniform and \modelDPP show robustness to prior violations and
+The \modelUniform and \modelDPP models show robustness to prior violations and
 perform well when applied to data generated under other models
@@
 In contrast, the \modelOld model is very sensitive to model violations, and
-strongly over-estimates the one-divergence model whenever the model is
-misspecified.
+strongly over-estimates the probability of the one-divergence model whenever
+the model is misspecified.
 Furthermore, the results suggest that using exponentially distributed priors on
@@
 S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
-Using exponential priors on divergence-time and \myTheta{} parameters does
+Using exponential priors on divergence-time and demographic parameters does
 increase the power of the \modelUshaped model compared to \modelOld across all
@@
 divergence event in most of the replicates across all the simulations.
-Using exponential priors on divergence-time and \myTheta{} parameters greatly
+Using exponential priors on divergence-time and demographic parameters greatly
 increases the power of the \modelUshaped model to detect variation in
@@
 S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).
-An estimated value of \divTimeDispersion less than 0.01 is an arbitrary
-threshold for ``simultaneous'' divergence used throughout the \msb
-literature.
-Albeit arbitrary, \cite{Oaks2012} did show via simulation that the true value
-of \divTimeDispersion will almost always be greater than 0.01 when divergences
-are random over periods of 0.1 \globalcoalunit generations or more (see
-Figure~S4 of \citet{Oaks2012}).
-
-As mentioned above, I also see increased power of the new models when
+% An estimated value of \divTimeDispersion less than 0.01 is an arbitrary
+% threshold for ``simultaneous'' divergence used throughout the \msb
+% literature.
+Although the \divTimeDispersion threshold of 0.01 is arbitrary, \cite{Oaks2012}
+did show via simulation that the true value of \divTimeDispersion will almost
+always be greater than 0.01 when divergences are random over periods of 0.1
+coalescent units or more (see Figure~S4 of \citet{Oaks2012}).
+
+As mentioned above, the increased power of the new models is also evident when
 looking at the estimated posterior probability of the one-divergence model
@@
 S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
-The exponential priors on divergence-time and \myTheta{} parameters (model
+The exponential priors on divergence-time and demographic parameters (model
 \modelUshaped) result in lower estimates of the probability of one divergence
@@
 The \modelDPP and \modelUniform models do frequently support the one-divergence
-model according to a Bayes factor criterion of greater than 10, however still
+model according to a Bayes factor criterion of greater than 10, but still
 less frequently than the \modelOld model.
 This result is not surprising given the extremely small prior probability of
-the one-divergence model under the \modelDPP and \modelUniform models (I.e.,
+the one-divergence model under the \modelDPP and \modelUniform models (i.e.,
 very few posterior samples of the one-divergence model will result in a
@@
 
-Lastly, when I look at the estimated posterior probability of
+Lastly, when looking at the estimated posterior probability of
 \divTimeDispersion being consistent with one shared divergence
-($p(\divTimeDispersion < 0.01 | \ssSpace)$), I see the same pattern of model
+($p(\divTimeDispersion < 0.01 | \ssSpace)$), I find the same pattern of model
 behavior, with \modelDPP and \modelUniform inferring low probabilities across
@@
 are analyzed under the \empModelOld model, there is strong support for very few
-divergence events shared among all 22 pairs of taxa with a maximum \emph{a
+divergence events shared among all 22 pairs of taxa, with a maximum \emph{a
 posteriori} (MAP) estimate of one-shared divergence
 (Figure~\ref{figPhilippines}A).
-When these data are analyzed under models allowed under the new implementation,
+When these data are analyzed using models allowed by the new implementation,
 there is much less support for highly clustered models and much greater
@@
 Figure \ref{figPhilippines} also shows the prior distribution across the number
-of divergence events (\divTimeNum) for each of the model, as well as the
+of divergence events (\divTimeNum) for each model, as well as the
 average prior probability of an unordered and ordered model of divergence
-(\divTimeIndexVector) across \divTimeNum for each model.
+(\divTimeIndexVector) across \divTimeNum.
 Estimates under the new models tend to be similar to the prior, which is
@@
 The disparity between the results of the \empModelOld model and the new models
-is even more pronounced when I look at the 10 divergence models
+is even more pronounced when looking at the 10 divergence models
 (\divTimeIndexVector) estimated to have the highest probability under each of the
@@
 distribution over the number of divergence events when the data from the
-nine-taxon pairs from the Islands of Negros and Panay are analyzed under the DPP
+nine-taxon pairs from the Islands of Negros and Panay are analyzed under DPP
 models that sample over unordered (\npModelDPP) and ordered
@@
 \ref{figValidationModelChoiceOmega}).
-The original model is very sensitive to model violations and, when present,
+The original model is very sensitive to violations and, when present,
 strongly over-estimates the probability of one-divergence event shared
@@
 across a range of conditions underlying the data.
-The new method \dppmsbayes has greater power to detect random temporal variation in
-divergences, is less prone to spurious inference of clustered divergence
-models, and much less likely to incorrectly infer such models with strong
-support
+The new method, \dppmsbayes, has greater power to detect random temporal
+variation in divergences, is less prone to spurious inference of clustered
+divergence models, and much less likely to incorrectly infer such models with
+strong support
 (Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
@@
 By evaluating a model intermediate between the old and new implementation
-(\modelUshaped), I was able to determine the relative effects of my 
+(\modelUshaped), I was able to determine the relative affects of my 
 modifications to the model.
@@
 power, while improved, may still be low.
-While I see much less prior sensitivity under the new model compared to those
-observed by \citet{Oaks2012}, I still see an effect when comparing the results
-of the empirical data analyzed under a diffuse (\empModelDPP) and informative
-(\empModelDPPInform) divergence-time prior (Figure~\ref{figPhilippines} C
-versus D).
+While there is much less prior sensitivity under the new model compared to
+those observed by \citet{Oaks2012}, there is still an effect when comparing the
+results of the empirical data analyzed under a diffuse (\empModelDPP) and
+informative (\empModelDPPInform) divergence-time prior
+(Figure~\ref{figPhilippines} C versus D).
 The fact that the posterior shifts toward the prior under the informative prior
@@
 estimated scenarios
-\citep{Barber2010,Bell2012,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012},
+\citep{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012},
 as I found here (Figure~\ref{figPhilippines}).
@@
 and the relatively small amount of information in the summary statistics
-calculated from the sequence data, I do not expect to find strong posterior
-support for any scenario.
+calculated from the sequence data, finding strong posterior support for any
+scenario is unexpected.
 Based on results of the empirical and power analyses
@@
 the new implementation more accurately reflects posterior uncertainty and
-avoids spurious support for biogeographic scenarios.
+avoids spurious support for biogeographical scenarios.
 

### 2014-02-22 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Adding divergence model plots to dpp ms.
More editing of DPP ms.

--- 2013-12-18 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
@@
     please see Figure~S\ref{figPowerPsiOld} for all results.
-}{figPowerPsiOld4}
+}{figPowerOmegaProbOld4}
+
+%% Empirical results
+\mFigure{../../empirical-analyses/plots/philippines-psi.pdf}{
+    The (A--E) posterior and (F--J) prior probabilities of the number of
+    divergence events (\divTimeNum) when the data of the 22 pairs of taxa from
+    the Philippines are analyzed under the five models indicated at the top of
+    each column of plots (Table~\ref{tabEmpiricalModels}).
+    The average prior probability of an (K--O) unordered and (P--T)
+    ordered model of divergence (\divTimeIndexVector) with \divTimeNum
+    divergence-time parameters is also shown.
+    The posterior median of the dispersion index of divergence times
+    (\divTimeDispersion) is also given for each model, followed by the 95\%
+    highest posterior density interval in parantheses.
+}{figPhilippines}
+
+\widthFigure{0.4}{../../empirical-analyses/plots/negros-panay-psi.pdf}{listformat=figList}{
+    The posterior probabilities of the number of divergence events,
+    \divTimeNum, when the data of the 9 pairs of taxa from Negros and Panay
+    Islands are analyzed under the DPP model that samples over (A) unordered
+    and (B) ordered models of divergence (Table~\ref{tabEmpiricalModels}).
+    Both models share the same (C) prior probability of the number of
+    divergence events, and the average prior probability of an (D)
+    unordered and (E) ordered model of divergence (\divTimeIndexVector) with
+    \divTimeNum divergence-time parameters.
+    The posterior median of the dispersion index of divergence times
+    (\divTimeDispersion) is also given for each model, followed by the 95\%
+    highest posterior density interval in parantheses.
+}{figNegrosPanay}
+\clearpage

### 2014-02-22 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More editing of DPP ms.

@@
 dimensionality (i.e., all the possible partitions of \npairs{} pairs to
-$1,2,\ldots,\npairs{}$ divergence events), and, in so doing, estimates
+$1,2,\ldots,\npairs{}$ divergence-time parameters), and, in so doing, estimates
 the posterior probability of each model.
@@
 \locusMutationRateScalar{\nlociTotal})$
-is a vector of locus-specific mutation-rate scaling parameters for each of the
+is a vector of locus-specific mutation-rate multipliers for each of the
 \nlociTotal loci,
@@
 Each \demographicParams{} contains the following demographic parameters: The
-mutation-scaled effective sizes ($\myTheta{} = 4N\mutationRate$) of the
+mutation-rate-scaled effective sizes ($\myTheta{} = 4N\mutationRate$) of the
 ancestral, \ancestralTheta{}, and descendant populations, \descendantTheta{1}{}
@@
 The term $p(\locusRateHetShapeParameter)$ is the prior density
-function for the shape-parameter of the gamma-distributed prior on
+function for the shape parameter of the gamma-distributed prior on
 rate heterogeneity among loci.
 This prior is $\locusRateHetShapeParameter \sim U(1, 20)$.
-The prior probability of the vector of locus-specific mutation-rate scaling parameters
+The prior probability of the vector of locus-specific mutation-rate multipliers
 given \locusRateHetShapeParameter then expands out as a product over

### 2014-02-22 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex
Working on dpp manuscript.
Proofreading and incorporating edits throughout the manuscript.
Changing first person to singular in dpp ms.
More editing of DPP ms.

@@
         \geneTreeVector & Vector containing the gene trees of each locus from each population pair: $(\geneTree{1}{1},\ldots,\geneTree{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \divTimeNum & Number of unique population divergence times among the \npairs{} population pairs. \tn
-        \divTime{} & One of the \divTimeNum \iid divergence time paramters in \globalcoalunit generations. \tn
-        \divTimeVector & Ordered set of divergence time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
-        \divTimeIndex{i} & The index of the divergence time parameter in \divTimeVector that is mapped to population pair $i$. \tn
-        \divTimeIndexVector & Vector of divergence time parameter indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn
+        \divTimeNum & Number of population divergence-time parameters shared among the \npairs{} population pairs. \tn
+        \divTime{} & Time of population divergence in \globalcoalunit generations. \tn
+        \divTimeVector & Set of divergence-time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
+        \divTimeIndex{i} & The index of the divergence-time in \divTimeVector to which population pair $i$ is mapped. \tn
+        \divTimeIndexVector & Vector of divergence-time indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn
         \divTimeMap{i} & Time of divergence in \globalcoalunit generations between the populations of pair $i$. \tn
@@
         \divTimeScaledVector & Vector containing the scaled divergence times of each locus from each population pair: $(\divTimeScaled{1}{1},\ldots,\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \descendantTheta{1}{i}, \descendantTheta{2}{i} & \myTheta{} of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
-        \ancestralTheta{i} & \myTheta{} of the population ancestral to pair $i$. \tn
+        \descendantTheta{1}{i}, \descendantTheta{2}{i} & Mutation-rate-scaled effective population size of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
+        \ancestralTheta{i} & Mutation-rate-scaled effective population size of the population ancestral to pair $i$. \tn
         \descendantThetaVector{1}, \descendantThetaVector{2} & Vectors $(\descendantTheta{1}{1},\ldots,\descendantTheta{1}{\npairs{}})$ and $(\descendantTheta{2}{1},\ldots,\descendantTheta{2}{\npairs{}})$, respectively.
@@
         \ancestralThetaVector & Vector containing the \ancestralTheta{} parameters for each population pair: $(\ancestralTheta{1},\ldots,\ancestralTheta{\npairs{}})$. \tn
-        \locusMutationRateScalar{j} & mutation-rate scaling parameter for locus $j$. \tn
-        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate parameters: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
-        \locusRateHetShapeParameter & Hyper-parameter for the shape of the gamma prior distribution on each \locusMutationRateScalar{}. \tn
+        \locusMutationRateScalar{j} & Mutation-rate multiplier of locus $j$. \tn
+        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate multipliers: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
+        \locusRateHetShapeParameter & The shape parameter of the gamma prior distribution on \locusMutationRateScalar{}. \tn
         \bottleScalar{1}{i}, \bottleScalar{2}{i} & \myTheta{}-scaling parameters that determine the magnitude of the population bottleneck in the $1^{st}$ and $2^{nd}$ descendant population of pair $i$, respectively.
@@
 \tn
-        \bottleTime{i} & Proportion of time between present and \divt{i} when the bottleneck ends for the descendant populations of pair $i$. \tn
+        \bottleTime{i} & Proportion of time between present and \divTimeMap{i} when the bottleneck ends for the descendant populations of pair $i$. \tn
         \bottleTimeVector & Vector containing the \bottleTime{} parameters for each population pair: $(\bottleTime{1},\ldots,\bottleTime{\npairs{}})$. \tn
@@
         \migrationRateVector & Vector containing the migration rates for each population pair: $(\migrationRate{1},\ldots,\migrationRate{\npairs{}})$. \tn
-        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
-This constant is intended to allow the user to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa.
+        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the investigator for locus $j$ of pair $i$.
+This constant is required to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa.
 \tn
-        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$.
-This constant is intended to allow the user to scale \myTheta{} for differences in mutation rates among loci or among taxa.
+        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the investigator for locus $j$ of pair $i$.
+This constant is required to scale \myTheta{} for differences in mutation rates among loci or among taxa.
 \tn
@@
         \mutationRateScalarConstantVector & Vector of mutation-rate scaling constants: $(\mutationRateScalarConstant{1}{1},\ldots,\mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$. \tn
-        \divTimeMean  & Mean of \divTimeMapVector. \tn
-        \divTimeVar  & Variance of \divTimeMapVector. \tn
-        \divTimeDispersion  & Dispersion index of \divTimeMapVector ($\divTimeVar/\divTimeMean$). \tn
+        \divTimeMean  & Mean of divergence times across the \npairs{} population pairs. \tn
+        \divTimeVar  & Variance of divergence times across the \npairs{} population pairs. \tn
+        \divTimeDispersion  & Dispersion index of divergence times across the \npairs{} population pairs ($\divTimeVar/\divTimeMean$). \tn
         \numPriorSamples & Number of samples from the joint prior. \tn
@@
     %\captionsetup{font=footnotesize}
-    \caption{The model priors evaluated in our simulation-based analyses.  For
+    \caption{The models evaluated in the simulation-based analyses.  For
         the \modelDPP model, the prior on the concentration parameter,
@@
                       & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
-            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
+            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
                           & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
@@
                           & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
-            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
+            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
                       & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
@@
     % \captionsetup{font=footnotesize}
-    \caption{The models we used to simulate pseudo-replicate datasets for
+    \caption{The models used to simulate pseudo-replicate datasets for
         assessing the power of the models in Table \ref{tabPriors}.  The

### 2014-02-22 manuscripts/reply/reply-main.tex
Adding Intro section title to reply manuscript.
Using `\url` for all URLs in manuscript.
A few edits to reply ms.

@@
+\section{Introduction}
+
 \msb \citep{Huang2011} uses approximate Bayesian computation (ABC) to estimate
@@
 three different priors used by \citet{Oaks2012}.
-The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5
-mya)$) closely mirrors the range of estimates of gene-divergence times.
+The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5)$)
+closely mirrors the range of estimates of gene-divergence times if we assume
+the same mutation rate and one generation per year (0--5 million years ago).
 Applying Hickerson et al.'s (\citeyear{Hickerson2013}) sampling-probability
@@
 While sampling error is always a reality, it does not appear to be playing a
-large role in the results of \citet{Oaks2012} or presented above.
+large role in the biases revealed by the results of \citet{Oaks2012} or
+presented above.
 
@@
 Thus, while sampling error will always be present in any numerical Bayesian
-approximation method, insufficient sampling is an unlikely explanation for
+approximation method, insufficient sampling is an unlikely explanation for the
 support of prior models with less parameter space.
@@
 and are publicly available at
-\href{https://github.com/joaks1/msbayes-experiments}{https://github.com/joaks1/msbayes-experiments}.
+\href{https://github.com/joaks1/msbayes-experiments}{\url{https://github.com/joaks1/msbayes-experiments}}.
 All information necessary to reproduce our results is provided there.

### 2014-02-22 manuscripts/reply/reply.tex
A few edits to reply ms.

@@
     difficult problem in evolutionary biology.
-    \citet{Huang2011} introduced an approximate Bayesian approach which can be
-    used to estimate the probabilities of models in which multiple sets of taxa
-    diverge at the same time. 
-    This method (implemented in the software \msb) does not require a common
-    set of character data that links all of the taxa under investigation, but
-    does require the relative mutation rates among the taxa and loci to be
-    assumed \emph{a priori}.
+    \citet{Huang2011} introduced an approximate Bayesian approach (implemented
+    in the software \msb) which can be used to estimate the probabilities of
+    models in which multiple sets of taxa diverge at the same time. 
+    % This method  does not require a common set of character data that links all
+    % of the taxa under investigation, but does require the relative mutation
+    % rates among the taxa and loci to be assumed \emph{a priori}.
     Recently, \citet{Oaks2012} used this model-choice framework to 

### 2014-02-23 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More editing of DPP ms.
Incorporating Hickerson's and our response papers into the intro.

@@
 determines its posterior probability, this can cause support for models with
-fewer divergence events \citep{Oaks2012}.
+fewer divergence events \citep{Oaks2012,Oaks2014reply}.
 Furthermore, \msb uses a discrete uniform prior over the number of divergence
@@
 with very few divergence events.
-Recently, \citet{Oaks2012} found via simulation that \msb will often strongly
-support biogeographically interesting models with a small number of divergence
-events shared among the taxa, even when divergences were random over broad
-timescales.
+
+Recently, \citet{Oaks2012} and \citet{Oaks2014reply} found via simulation that
+\msb will often strongly support biogeographically interesting models with a
+small number of divergence events shared among taxa, even when divergences were
+random over broad timescales.
 They suggested this behavior was due to the combination of uniform priors on
-parameters and the U-shaped prior on divergence models.
+parameters causing small marginal likelihoods of richer models and the
+U-shaped prior on divergence models.
+\citet{Hickerson2013} suggested the problem is caused by sampling error, and
+proposed as a solution an approximate-Bayesian model averaging approach that
+samples over empirically informed uniform priors.
+However, \citet{Oaks2014reply} evaluated the approach proposed by
+\citet{Hickerson2013} using simulations and found that it did not mitigate the
+method's propensity to incorrectly infer clustered divergences, and often
+preferred priors that excluded the true values of the model's parameters.
+Here, I describe a new method that successfully mitigates spurious inference
+of biogeographically interesting scenarios while avoiding negative side effects
+of empirically informed uniform priors.
 
@@
 
-Here, I modify the \msb model to test whether alternative parameterizations and
-priors will improve the behavior of the approximate-Bayesian model-choice
-approach.
-I implement a Dirichlet-process prior (DPP) over all possible models of
-divergence, and use gamma and beta probability distributions in place of
-uniform priors on many of the model's parameters.
+In this study, I introduce a new method, implemented in the software
+\dppmsbayes, that extends the model of \msb.
+I use this method to test whether alternative parameterizations and priors
+improve the behavior of the approximate-Bayesian model-choice approach to
+estimating shared divergence events.
+The new approach uses a Dirichlet-process prior (DPP) over all possible models
+of divergence, and gamma and beta probability distributions in place of uniform
+priors on many of the model's parameters.
 Using simulations, I show that the new implementation has improved robustness,

### 2014-02-23 manuscripts/dpp-msbayes/dpp-msbayes.tex
More editing of DPP ms.
Incorporating Hickerson's and our response papers into the intro.

@@
     that can be estimated using genetic data from co-distributed species.
-    I introduce a new approximate-Bayesian model for comparative
+    I introduce a new approximate-Bayesian method for comparative
     phylogeographical model-choice that estimates the temporal distribution of

### 2014-02-24 manuscripts/dpp-msbayes/dpp-msbayes-si-figures.tex
Working on dpp manuscript.
Proofreading and incorporating edits throughout the manuscript.
More editing of DPP ms.
Updating figures for dpp ms.
Using compressed and bitmapped versions of some of the more
vector-graphic-dense figures for the DPP ms.

@@
 %% Validation accuracy
-\siFigure{../../validation/images/validation-accuracy-omega.pdf}{
+% \siFigure{../../validation/images/validation-accuracy-omega.pdf}{
+\siFigure{../../validation/images/validation-accuracy-omega-compress.pdf}{
     \validationAccuracyComparisonCaption{unadjusted}{\divTimeDispersion}
@@
 
-\siFigure{../../validation/images/validation-accuracy-omega-glm.pdf}{
+% \siFigure{../../validation/images/validation-accuracy-omega-glm.pdf}{
+\siFigure{../../validation/images/validation-accuracy-omega-glm-compress.pdf}{
     \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeDispersion}
@@
 
-\siFigure{../../validation/images/validation-accuracy-time.pdf}{
+% \siFigure{../../validation/images/validation-accuracy-time.pdf}{
+\siFigure{../../validation/images/validation-accuracy-time-compress.pdf}{
     \validationAccuracyComparisonCaption{unadjusted}{\divTimeMean}
@@
 
-\siFigure{../../validation/images/validation-accuracy-time-glm.pdf}{
+% \siFigure{../../validation/images/validation-accuracy-time-glm.pdf}{
+\siFigure{../../validation/images/validation-accuracy-time-glm-compress.pdf}{
     \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeMean}
@@
 
-\siFigure{../../validation/images/validation-accuracy-psi.pdf}{
+% \siFigure{../../validation/images/validation-accuracy-psi.pdf}{
+\siFigure{../../validation/images/validation-accuracy-psi-compress.pdf}{
     \validationAccuracyComparisonCaption{unadjusted}{\divTimeNum}
@@
 
-\siFigure{../../validation/images/validation-accuracy-psi-glm.pdf}{
+% \siFigure{../../validation/images/validation-accuracy-psi-glm.pdf}{
+\siFigure{../../validation/images/validation-accuracy-psi-glm-compress.pdf}{
     \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeNum}
@@
 %% Validation ordered models
-\siFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
+% \siFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
+\siFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy-compress.pdf}{
     \validationAccuracyCaption{\modelDPPOrdered}{\modelDPPOrdered}
@@
 %% Power accuracy results
-\siSidewaysFigure{../../power-comparison/images/old-power-omega-accuracy-6.pdf}{
-    \powerAccuracyComparisonCaption{\powerSeriesOld}
-}{figPowerAccuracyOld}
+% \siSidewaysFigure{../../power-comparison/images/exp-power-omega-accuracy-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/exp-power-omega-accuracy-6-compress.pdf}{
+    \powerAccuracyComparisonCaption{\powerSeriesExp}
+}{figPowerAccuracyExp}
 
-\siSidewaysFigure{../../power-comparison/images/uniform-power-omega-accuracy-6.pdf}{
+% \siSidewaysFigure{../../power-comparison/images/uniform-power-omega-accuracy-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/uniform-power-omega-accuracy-6-compress.pdf}{
     \powerAccuracyComparisonCaption{\powerSeriesUniform}
@@
 
-\siSidewaysFigure{../../power-comparison/images/exp-power-omega-accuracy-6.pdf}{
-    \powerAccuracyComparisonCaption{\powerSeriesExp}
-}{figPowerAccuracyExp}
+% \siSidewaysFigure{../../power-comparison/images/old-power-omega-accuracy-6.pdf}{
+\siSidewaysFigure{../../power-comparison/images/old-power-omega-accuracy-6-compress.pdf}{
+    \powerAccuracyComparisonCaption{\powerSeriesOld}
+}{figPowerAccuracyOld}
 
@@
 
+%% Power psi probability results
+\siSidewaysFigure{../../power-comparison/images/old-power-psi-prob-6.pdf}{
+    \powerSupportComment{\powerSeriesOld}
+    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
+    \timeConversionComment
+}{figPowerPsiProbOld}
+
+\siSidewaysFigure{../../power-comparison/images/uniform-power-psi-prob-6.pdf}{
+    \powerSupportComment{\powerSeriesUniform}
+    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesUniform}
+    \timeConversionComment
+}{figPowerPsiProbUniform}
+
+\siSidewaysFigure{../../power-comparison/images/exp-power-psi-prob-6.pdf}{
+    \powerSupportComment{\powerSeriesExp}
+    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesExp}
+    \timeConversionComment
+}{figPowerPsiProbExp}
+
 %% Power dispersion results
@@
 }{figPowerOmegaExp}
-
-%% Power psi probability results
-\siSidewaysFigure{../../power-comparison/images/old-power-psi-prob-6.pdf}{
-    \powerSupportComment{\powerSeriesOld}
-    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
-    \timeConversionComment
-}{figPowerPsiProbOld}
-
-\siSidewaysFigure{../../power-comparison/images/uniform-power-psi-prob-6.pdf}{
-    \powerSupportComment{\powerSeriesUniform}
-    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesUniform}
-    \timeConversionComment
-}{figPowerPsiProbUniform}
-
-\siSidewaysFigure{../../power-comparison/images/exp-power-psi-prob-6.pdf}{
-    \powerSupportComment{\powerSeriesExp}
-    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesExp}
-    \timeConversionComment
-}{figPowerPsiProbExp}
 
@@
     (Table~\ref{tabEmpiricalModels}).
+    The times indicate the posterior median and 95\% highest posterior density
+    (HPD) interval.
 }{figMarginalTimes}

### 2014-02-24 manuscripts/bin/compress-images.sh
Adding `compress-images.sh` script.

@@ -1,0 +1,20 @@
+#! /bin/bash
+
+paths="../../validation/images/validation-accuracy-omega-glm.pdf
+../../validation/images/validation-accuracy-omega.pdf
+../../validation/images/validation-accuracy-psi-glm.pdf
+../../validation/images/validation-accuracy-psi.pdf
+../../validation/images/validation-accuracy-time-glm.pdf
+../../validation/images/validation-accuracy-time.pdf
+../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf
+../../power-comparison/images/exp-power-omega-accuracy-6.pdf
+../../power-comparison/images/old-power-omega-accuracy-6.pdf
+../../power-comparison/images/uniform-power-omega-accuracy-6.pdf"
+
+
+for f in $paths
+do
+    n=${f/\.pdf/-compress\.pdf}
+    convert -density 300 -compress jpeg -quality 60 $f $n
+done
+

### 2014-02-26 manuscripts/reply/reply.tex
Updating Jeet's address on reply ms.
Updating acknowledgements of reply ms.
Updating authors and acknowledgements for reply ms.

@@
 \author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
-\author[2]{Jeet Sukumaran}
-\author[3]{Charles W.\ Linkem}
-\author[4]{Cameron D.\ Siler}
-\author[1]{Rafe M.\ Brown}
+\author[2]{Charles W.\ Linkem}
+\author[3]{Jeet Sukumaran}
 \affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
-\affil[2]{Department of Biology, Duke University, Durham, North Carolina 27708}
-\affil[3]{Department of Biology, University of Washington, Seattle, Washington 98195}
-\affil[4]{Sam Noble Museum, Department of Biology, University of Oklahoma, Norman, Oklahoma 73072}
+\affil[2]{Department of Biology, University of Washington, Seattle, Washington 98195}
+\affil[3]{Department of Ecology and Evolutionary Biology, University of Michigan, Ann Arbor, Michigan 48109}
 
@@
 \section{Acknowledgments}
-We thank Melissa Callahan, Jake Esselstyn, Mark Holder, Emily McTavish, Daniel
-Money, Jordan Koch, and Adam Leach\'{e} for insightful comments that greatly
-improved this work.
-J.\ Oaks thanks the National Science Foundation for supporting this work (DEB
-1011423 and DBI 1308885).
+We thank Melissa Callahan, Jake Esselstyn, Cameron Siler, Mark Holder, Rafe
+Brown, Emily McTavish, Daniel Money, Jordan Koch, and Adam Leach\'{e} for
+insightful comments that greatly improved this work.
+We thank Michael Hickerson and co-authors for generously providing their data.
+J.\ Oaks and C.\ Linkem thank the National Science Foundation for supporting
+this work (DEB 1011423, DBI 1308885 and BIO-1202754).
 J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
@@
 \hangindent=1cm
-Oaks, J.~R., J.~Sukumaran, C.~W.\ Linkem, C.~D.\ Siler, and R.~M.\ Brown. \msTitle.
+Oaks, J.~R., C.~W.\ Linkem, and J.~Sukumaran. \msTitle
 

### 2014-02-28 manuscripts/dpp-msbayes/cover-letter/cover-letter.tex
Adding initial draft of cover letter for dpp ms.

\documentclass[letterpaper]{letter}
\usepackage{anysize}
\marginsize{1.25in}{1.25in}{0.5in}{0.5in}
\usepackage{setspace}
\usepackage{url}
\usepackage{verbatim}
\usepackage{parskip}
\usepackage{xspace}
\usepackage{hyperref}
\hypersetup{pdfborder={0 0 0}, colorlinks=true, urlcolor=blue}

\newcommand{\ignore}[1]{}
\newcommand{\super}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\sub}[1]{\ensuremath{_{\textrm{#1}}}}
\newcommand{\dC}{\ensuremath{^\circ{\textrm{C}}}}
\newcommand{\highLight}[1]{\textcolor{magenta}{\MakeUppercase{#1}}\xspace}

\makeatletter
\let\@texttop\relax
\makeatother

\newenvironment{myEnumerate}{
  \begin{enumerate}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}}
  {\end{enumerate}}

\newenvironment{myItemize}{
  \begin{itemize}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}}
  {\end{itemize}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\signature{Jamie Oaks \\ 
                 Postdoctoral Fellow \\ 
                 Department of Biology \\
                 University of Washington \\ 
                 Box 351800 \\ 
                 Seattle, WA 98195 \\ 
                 \href{mailto:joaks1@uw.edu}{\texttt{joaks1@uw.edu}} \\ 
                 \href{http://www.phyletica.com}{\texttt{www.phyletica.com}}}
\begin{document}
\begin{letter}{Journal/Society \\
                address \\
                here}
\opening{Dear Editor:}
% \raggedright
I am submitting a manuscript entitled ``An Improved Approximate-Bayesian
Model-choice Method for Estimating Shared Evolutionary History'' to be
considered for publication in \highLight{journal here}.
The field of research is comparative biogeography and Bayesian model choice.
All of the computer code for the newly implemented model I present in the
the manuscript is freely available in the open-source
software packages \texttt{dpp-msbayes} and \texttt{PyMsBayes}
(\href{https://github.com/joaks1/dpp-msbayes}{\url{https://github.com/joaks1/dpp-msbayes}}
and
\href{https://github.com/joaks1/PyMsBayes}{\url{https://github.com/joaks1/PyMsBayes}},
respectively).
Also, I performed the work described in the manuscript following the principles
of Open Notebook Science; using version-control software, I make progress in
all aspects of the work freely and publicly available in real-time at
\href{https://github.com/joaks1/msbayes-experiments}{\url{https://github.com/joaks1/msbayes-experiments}}.
All information necessary to reproduce my results is provided there.

In this paper, I introduce a novel Bayesian method for estimating the
probability of shared divergence histories across a set of taxa from
multi-locus DNA sequence data.
More specifically, given a set of species pairs, the method estimates
the probabilities of all possible models in which multiple sets of taxa
can diverge at the same time.
These models range from a single divergence-time parameter (i.e., simultaneous
divergence of all pairs of taxa) to the fully generalized model in which each
pair of taxa diverged at a unique time.
The current state-of-the-art method for this estimation problem is implemented
in the software package \texttt{msBayes}.
By introducing more flexible priors on the parameters the model and a
non-parametric Dirichlet-process prior over divergence models, the new method
has improved robustness, accuracy, and power for estimating shared evolutionary
history across taxa when compared to \texttt{msBayes}.
Also, simulations have shown that \texttt{msBayes} will often to incorrectly
support biogeographically interesting scenarios of shared divergences when
divergences were random over broad time periods.
My results show that the new method is \emph{much} less likely to make such
misleading and costly errors.

The new method provides an important tool for evolutionary biologists to
investigate how large-scale changes to the environment, including geological
and climatic events, have affected the evolutionary history of entire
communities of co-distributed species and their associated microbiota.
Given the dynamic nature of our planet, such biogeographical processes likely
play a significant role regulating biodiversification and community assembly.

Thank you for your consideration, and I look forward to your correspondence
regarding this manuscript.

\addtolength{\medskipamount}{-5pt}
\closing{Sincerely,}
\end{letter}

\end{document}

### 2014-02-28 manuscripts/dpp-msbayes/dpp-msbayes-abstract.tex
Breaking up dpp ms into more files.
Moving abstract and acknowledgments to separate files, and also making
the title a macro.

--- 2014-02-23 manuscripts/dpp-msbayes/dpp-msbayes.tex
@@
-%&<latex>
-\documentclass[letterpaper,12pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{../utils/preamble.tex}
-\input{../utils/macros.tex}
-
-\title{An Improved Approximate-Bayesian Model-choice Method for Estimating
-Shared Evolutionary History}
-
-\author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
-% \author[1]{Mark T.\ Holder}
-\affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
-
-% \date{\today}
-\date{\parbox{\linewidth}{\centering%
-    \today\endgraf\bigskip
-    \textbf{Running head}: Improved Estimation of Shared Evolutionary History}}
-
-\makeatletter
-\let\msTitle\@title
-\let\msAuthor\@author
-\let\msDate\@date
-\makeatother
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-% \doublespacing
-% \raggedright
-% \setlength{\parindent}{0.5in}
-% \begin{linenumbers}
-
-\maketitle
-
-\begin{abstract}
 To understand the processes that generate biodiversity, it is
@@
 biogeographically interesting models with strong support.
-
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Dirichlet-process prior; approximate Bayesian
-    computation; model choice; phylogeography; biogeography} 
-\end{abstract}
-
-\newpage
-
-\input{dpp-msbayes-main.tex}
-
-\section{Acknowledgments}
-I am greatful to Melissa Callahan, Mark Holder, Emily McTavish, Daniel Money,
-Jordan Koch, and Adam Leach\'{e} for insightful comments that greatly improved
-this work.
-I thank the National Science Foundation for supporting this work (DEB
-1011423 and DBI 1308885).
-This work was also supported by the University of Kansas (KU) Office of Graduate
-Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
-Society, KU Department of Ecology and Evolutionary Biology, and the KU
-Biodiversity Institute.
-I also thank Mark Holder, the KU Information and Telecommunication Technology
-Center, KU Computing Center, and the iPlant Collaborative for the computational
-support necessary to conduct the analyses presented herein.
-
-\bibliography{../bib/references}
-
-%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\renewcommand\listfigurename{Figure Captions}
-\cftsetindents{fig}{0cm}{2.2cm}
-\renewcommand\cftdotsep{\cftnodots}
-\setlength\cftbeforefigskip{10pt}
-\cftpagenumbersoff{fig}
-\listoffigures
-
-% \end{linenumbers}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\input{dpp-msbayes-tables.tex}
-
-\clearpage
-
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
-\input{dpp-msbayes-figures.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-\setcounter{section}{0}
-
-\singlespacing
-
-\section*{Supporting Information}
-% PUT MAIN TEXT CITATION HERE
-\hangindent=1cm
-Oaks, J.\ R. \msTitle.
-
-\newpage
-\singlespacing
-
-\input{dpp-msbayes-si-figures.tex}
-
-\end{document}

### 2014-02-28 manuscripts/dpp-msbayes/dpp-msbayes-acknowledgments.tex
Breaking up dpp ms into more files.
Moving abstract and acknowledgments to separate files, and also making
the title a macro.

--- 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes.tex
@@
-%&<latex>
-\documentclass[letterpaper,12pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{../utils/preamble.tex}
-\input{../utils/macros.tex}
-
-\title{An Improved Approximate-Bayesian Model-choice Method for Estimating
-Shared Evolutionary History}
-
-\author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
-% \author[1]{Mark T.\ Holder}
-\affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
-
-% \date{\today}
-\date{\parbox{\linewidth}{\centering%
-    \today\endgraf\bigskip
-    \textbf{Running head}: Improved Estimation of Shared Evolutionary History}}
-
-\makeatletter
-\let\msTitle\@title
-\let\msAuthor\@author
-\let\msDate\@date
-\makeatother
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-% \doublespacing
-% \raggedright
-% \setlength{\parindent}{0.5in}
-% \begin{linenumbers}
-
-\maketitle
-
-\begin{abstract}
-    To understand the processes that generate biodiversity, it is
-    important to account for large-scale processes that affect the evolutionary
-    history of groups of co-distributed populations of organisms.
-    Such events predict temporally clustered divergences times, a pattern
-    that can be estimated using genetic data from co-distributed species.
-    I introduce a new approximate-Bayesian model for comparative
-    phylogeographical model-choice that estimates the temporal distribution of
-    divergences across taxa from multi-locus DNA sequence data.
-    The model is an extension of that implemented in \msb.
-    By reparameterizing the model, introducing more flexible priors on
-    demographic and divergence-time parameters, and implementing a
-    non-parametric Dirichlet-process prior over divergence models, I improved
-    the robustness, accuracy, and power of the method for estimating shared
-    evolutionary history across taxa.
-    The results demonstrate the improved performance of the new method is due
-    to (1) more appropriate priors on divergence-time and demographic
-    parameters that avoid prohibitively small marginal likelihoods for models
-    with more divergence events,
-    and (2) the Dirichlet-process providing a flexible prior on divergence
-    histories that does not strongly disfavor models with intermediate numbers
-    of divergence events.
-    % The results demonstrate that the tendency of the \msb model to spuriously
-    % support models of clustered divergences is caused by a combination of (1)
-    % uniform priors on nuisance parameters hindering the marginal likelihoods of
-    % models with more divergence-time parameters, and (2) a prior on divergence
-    % models that disfavors models with intermediate numbers of divergence-time
-    % parameters.
-    The new method yields more robust estimates of posterior uncertainty, and
-    thus greatly reduces the tendency of the model to incorrectly estimate
-    biogeographically interesting models with strong support.
-
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Dirichlet-process prior; approximate Bayesian
-    computation; model choice; phylogeography; biogeography} 
-\end{abstract}
-
-\newpage
-
-\input{dpp-msbayes-main.tex}
-
-\section{Acknowledgments}
 I am greatful to Melissa Callahan, Mark Holder, Emily McTavish, Daniel Money,
@@
 support necessary to conduct the analyses presented herein.
-
-\bibliography{../bib/references}
-
-%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\renewcommand\listfigurename{Figure Captions}
-\cftsetindents{fig}{0cm}{2.2cm}
-\renewcommand\cftdotsep{\cftnodots}
-\setlength\cftbeforefigskip{10pt}
-\cftpagenumbersoff{fig}
-\listoffigures
-
-% \end{linenumbers}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\input{dpp-msbayes-tables.tex}
-
-\clearpage
-
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
-\input{dpp-msbayes-figures.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-\setcounter{section}{0}
-
-\singlespacing
-
-\section*{Supporting Information}
-% PUT MAIN TEXT CITATION HERE
-\hangindent=1cm
-Oaks, J.\ R. \msTitle.
-
-\newpage
-\singlespacing
-
-\input{dpp-msbayes-si-figures.tex}
-
-\end{document}

### 2014-02-28 manuscripts/dpp-msbayes/dpp-msbayes.tex
Breaking up dpp ms into more files.
Moving abstract and acknowledgments to separate files, and also making
the title a macro.

--- 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes.tex
@@
 
-\title{An Improved Approximate-Bayesian Model-choice Method for Estimating
-Shared Evolutionary History}
+\title{\dppTitle}
 
@@
 \begin{abstract}
-    To understand the processes that generate biodiversity, it is
-    important to account for large-scale processes that affect the evolutionary
-    history of groups of co-distributed populations of organisms.
-    Such events predict temporally clustered divergences times, a pattern
-    that can be estimated using genetic data from co-distributed species.
-    I introduce a new approximate-Bayesian model for comparative
-    phylogeographical model-choice that estimates the temporal distribution of
-    divergences across taxa from multi-locus DNA sequence data.
-    The model is an extension of that implemented in \msb.
-    By reparameterizing the model, introducing more flexible priors on
-    demographic and divergence-time parameters, and implementing a
-    non-parametric Dirichlet-process prior over divergence models, I improved
-    the robustness, accuracy, and power of the method for estimating shared
-    evolutionary history across taxa.
-    The results demonstrate the improved performance of the new method is due
-    to (1) more appropriate priors on divergence-time and demographic
-    parameters that avoid prohibitively small marginal likelihoods for models
-    with more divergence events,
-    and (2) the Dirichlet-process providing a flexible prior on divergence
-    histories that does not strongly disfavor models with intermediate numbers
-    of divergence events.
-    % The results demonstrate that the tendency of the \msb model to spuriously
-    % support models of clustered divergences is caused by a combination of (1)
-    % uniform priors on nuisance parameters hindering the marginal likelihoods of
-    % models with more divergence-time parameters, and (2) a prior on divergence
-    % models that disfavors models with intermediate numbers of divergence-time
-    % parameters.
-    The new method yields more robust estimates of posterior uncertainty, and
-    thus greatly reduces the tendency of the model to incorrectly estimate
-    biogeographically interesting models with strong support.
+    \input{dpp-msbayes-abstract.tex}
 
@@
 \section{Acknowledgments}
-I am greatful to Melissa Callahan, Mark Holder, Emily McTavish, Daniel Money,
-Jordan Koch, and Adam Leach\'{e} for insightful comments that greatly improved
-this work.
-I thank the National Science Foundation for supporting this work (DEB
-1011423 and DBI 1308885).
-This work was also supported by the University of Kansas (KU) Office of Graduate
-Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
-Society, KU Department of Ecology and Evolutionary Biology, and the KU
-Biodiversity Institute.
-I also thank Mark Holder, the KU Information and Telecommunication Technology
-Center, KU Computing Center, and the iPlant Collaborative for the computational
-support necessary to conduct the analyses presented herein.
+\input{dpp-msbayes-acknowledgments.tex}
 

### 2014-02-28 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Updating bib management and formating in dpp ms.
Removing us of natbib, and adding use of `plos2009.bst` for plos-style
citations and bibliography.

@@
 mutations occur in populations and how they are inherited over generations
-\citep{Hudson1990,WakeleyCoalescent}.
+\cite{Hudson1990,WakeleyCoalescent}.
 Thus, a method for estimating historical patterns of divergences across taxa
@@
 The software package \msb implements such an approach in an
-approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
+approximate-Bayesian model-choice framework \cite{Hickerson2006,Huang2011}.
 The method models temporally clustered divergences across taxa caused by a
@@
 density, resulting in small marginal likelihoods relative to models with fewer
-divergence-time parameters \citep{Jeffreys1939,Lindley1957}.
+divergence-time parameters \cite{Jeffreys1939,Lindley1957}.
 Given that the marginal likelihood of a model weighted by its prior is what
 determines its posterior probability, this can cause support for models with
-fewer divergence events \citep{Oaks2012,Oaks2014reply}.
+fewer divergence events \cite{Oaks2012,Oaks2014reply}.
 Furthermore, \msb uses a discrete uniform prior over the number of divergence
@@
 models that puts most of the prior mass on models with either very few or very
-many divergence-time parameters (see Figure 5 of \citet{Oaks2012}; for brevity
+many divergence-time parameters (see Figure 5 of \cite{Oaks2012}; for brevity
 I will refer to this prior as ``U-shaped'').
@@
 
-Recently, \citet{Oaks2012} and \citet{Oaks2014reply} found via simulation that
+Recently, \cite{Oaks2012} and \cite{Oaks2014reply} found via simulation that
 \msb will often strongly support biogeographically interesting models with a
@@
 U-shaped prior on divergence models.
-\citet{Hickerson2013} suggested the problem is caused by sampling error, and
+\cite{Hickerson2013} suggested the problem is caused by sampling error, and
 proposed as a solution an approximate-Bayesian model averaging approach that
 samples over empirically informed uniform priors.
-However, \citet{Oaks2014reply} evaluated the approach proposed by
-\citet{Hickerson2013} using simulations and found that it did not mitigate the
+However, \cite{Oaks2014reply} evaluated the approach proposed by
+\cite{Hickerson2013} using simulations and found that it did not mitigate the
 method's propensity to incorrectly infer clustered divergences, and often
@@
 
-% Recently, however, \citet{Oaks2012} found that \msb can be strongly biased
+% Recently, however, \cite{Oaks2012} found that \msb can be strongly biased
 % toward spuriously estimating biogeographically interesting models of highly
@@
 % The model implemented in \msb uses a discrete uniform prior over the number of
-% possible divergence parameters, which \citet{Oaks2012} show imposes a prior on
+% possible divergence parameters, which \cite{Oaks2012} show imposes a prior on
 % divergence models that puts most of the prior mass on models with either very
 % few or very many divergence times shared across the taxa (see Figure 5 of
-% \citet{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
+% \cite{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
 % Furthermore, uniform priors used for most of the model's parameters
 % result in small marginal likelihoods of parameter-rich models, i.e., models
-% with many divergence-time parameters \citep{Jeffreys1935, Lindley1957}.
+% with many divergence-time parameters \cite{Jeffreys1935, Lindley1957}.
 % Thus, together, the priors used in \msb will often create a bias toward
-% models with few divergence-time parameters \citep{Oaks2012}.
+% models with few divergence-time parameters \cite{Oaks2012}.
 
@@
 After reanalyzing sequence data from 22 pairs of taxa from the Philippines
-\citep{Oaks2012} under the new model, I find a large amount of posterior
+\cite{Oaks2012} under the new model, I find a large amount of posterior
 uncertainty in the number of divergence events shared among the taxa; a result
@@
 In this section, I describe the model, which is a modification of the model
-implemented in \msb \citep{Huang2011,Oaks2012}.
+implemented in \msb \cite{Huang2011,Oaks2012}.
 The code implementing the new model is freely available in the open-source
@@
 All information necessary to reproduce my results is provided there.
-I follow much of the notation of \citet{Oaks2012}, but modify it to
+I follow much of the notation of \cite{Oaks2012}, but modify it to
 aid in the description of the new model.
@@
 is the probability of the sequence alignment of locus $j$ for population pair
-$i$ given the gene tree and \hky substitution model parameters \citep[i.e., the
-``Felsenstein likelihood'';][]{Felsenstein1981}.
+$i$ given the gene tree and \hky substitution model parameters \cite[i.e., the
+``Felsenstein likelihood'';]{Felsenstein1981}.
 The model allows for an intra-locus recombination rate \recombinationRate,
@@
 heights) of gene tree \geneTree{i}{j} is given by a multi-population
-Kingman-coalescent model \citep{Kingman1982} where the ancestral population of
+Kingman-coalescent model \cite{Kingman1982} where the ancestral population of
 size
@@
 \begin{linenomath}
-As described by \citet{Oaks2012}, to get the divergence times in units
+As described by \cite{Oaks2012}, to get the divergence times in units
 proportional to the expected number of mutations, they must be
@@
 \end{equation}
-which is the Bell Number \citep{Bell1934}.
+which is the Bell Number \cite{Bell1934}.
 The original \msb model only samples over the unordered realizations of
 \divTimeIndexVector, such that the sample space is reduced to all the possible
-integer partitions of \npairs{} \citep{Oaks2012,Huang2011,OeisPartitionNumber,
+integer partitions of \npairs{} \cite{Oaks2012,Huang2011,OeisPartitionNumber,
     OeisPartitionTriangle,Malenfant2011}.
@@
 process that groups random variables into an unknown number of discrete
-parameter classes \citep{Ferguson1973,Antoniak1974}.
+parameter classes \cite{Ferguson1973,Antoniak1974}.
 The Dirichlet process has been used as a nonparametric Bayesian approach to
 many inference problems in evolutionary biology
-\citep{Lartillot2004,Huelsenbeck2007,HuelsenbeckSuchard2007,Ane2007,Heath2011,Heath2012}.
+\cite{Lartillot2004,Huelsenbeck2007,HuelsenbeckSuchard2007,Ane2007,Heath2011,Heath2012}.
 Here, I use the Dirichlet process to place a prior over all possible ordered
@@
 I take a hierarchical approach and use a prior probability
-distribution (i.e., hyperprior) for \concentrationParam \citep{Escobar1995}.
+distribution (i.e., hyperprior) for \concentrationParam \cite{Escobar1995}.
 More specifically, I use a gamma-distributed prior $\concentrationParam \sim
@@
 \subsubsection{The prior on divergence models}
-One of the key differences between my model and that of \msb \citep{Huang2011}
+One of the key differences between my model and that of \msb \cite{Huang2011}
 is the prior distribution on divergence models.
-As discussed in \citet{Oaks2012}, in \msb the prior used for
+As discussed in \cite{Oaks2012}, in \msb the prior used for
 \divTimeIndexVector is a combination of a discrete uniform prior over the
@@
 appears in \divTimeIndexVector, with the constraint that all \divTime{}
-parameters are represented at least once (see Equation 2 of \citet{Oaks2012}).
+parameters are represented at least once (see Equation 2 of \cite{Oaks2012}).
 I denote this prior used by \msb as $\divTimeIndexVector \sim \priorOld$.
-\citet{Oaks2012} discuss how placing a uniform prior over the number of
+\cite{Oaks2012} discuss how placing a uniform prior over the number of
 divergence parameters (denoted \divTimeNum here, and as \numt{} in
-\citet{Huang2011}) imposes an U-shaped prior over divergence
+\cite{Huang2011}) imposes an U-shaped prior over divergence
 models (\divTimeIndexVector; see
-Figure 5B of \citet{Oaks2012}).
+Figure 5B of \cite{Oaks2012}).
 To avoid this, I place priors directly on the sample space of divergence
@@
 populations of each pair.
-As described by \citet{Oaks2012}, \msb uses the joint prior
+As described by \cite{Oaks2012}, \msb uses the joint prior
 \begin{equation}
@@
 of the prior on descendant-population size.
-As described by \citet{Oaks2012}, in \msb, \myTheta{C} is half of the upper
+As described by \cite{Oaks2012}, in \msb, \myTheta{C} is half of the upper
 limit of the continuous uniform prior on the mean of the descendant population
@@
 the pairs of populations:
-$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
-\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
+$\pi$ \cite{Tajima1983}, $\theta_W$ \cite{Watterson1975}, $\pi_{net}$
+\cite{Takahata1985}, and $SD(\pi-\theta_W)$ \cite{Tajima1989}.
 Furthermore, in addition to model parameters, each sample \hpvector{}
@@
 Previously, these have been denoted as \meant{}{}, \vart{}{}, \vmratio{}, and
-\numt{}, respectively \citep{Hickerson2006,Huang2011,Oaks2012}.
+\numt{}, respectively \cite{Hickerson2006,Huang2011,Oaks2012}.
 I use \divTimeMean and \divTimeVar in place \meant{}{} and \vart{}{} to make
@@
 pairwise differences between the descendant populations
-\citep[$\pi_b$;][]{NeiLi1979,Huang2011}.
+\cite[$\pi_b$;]{NeiLi1979,Huang2011}.
 This has the advantage of reducing the sample space of the number of
@@
 
-In the original descriptions of the \msb model \citep{Hickerson2006,Huang2011},
+In the original descriptions of the \msb model \cite{Hickerson2006,Huang2011},
 this re-ordering is justified by the fact that the expected value of $\pi_b$ is
@@
 is to minimize Euclidean distances between simulated datasets in which the true
-model has a single divergence event \citep{Huang2011}, and (2) the reordering
+model has a single divergence event \cite{Huang2011}, and (2) the reordering
 increases the model's tendency to infer a single divergence event
-\citep{Huang2011}, it seems possible that this approach could be biasing the
-method \citep{Oaks2012}.
+\cite{Huang2011}, it seems possible that this approach could be biasing the
+method \cite{Oaks2012}.
 
@@
 Post-hoc adjustment of the posterior sample can also be performed with a number
-of regression techniques \citep{Beaumont2002,Blum2009,Leuenberger2010}.
+of regression techniques \cite{Beaumont2002,Blum2009,Leuenberger2010}.
 For my analyses, I use the general linear model (GLM) regression adjustment
-\citep{Leuenberger2010} as implemented in \abctoolbox
-\citep[v1.1;][]{ABCtoolbox}, which \citet{Oaks2012} showed performs very
+\cite{Leuenberger2010} as implemented in \abctoolbox
+\cite[v1.1;]{ABCtoolbox}, which \cite{Oaks2012} showed performs very
 similarly to weighted local-linear regression and multinomial logistic
-regression adjustments \citep{Beaumont2002} for \msb posteriors.
+regression adjustments \cite{Beaumont2002} for \msb posteriors.
 
 \subsection{Assessing model-choice behavior and robustness}
-Following the simulation-based approach of \citet{Oaks2012}, I characterize
+Following the simulation-based approach of \cite{Oaks2012}, I characterize
 the behavior of several models under the ideal conditions where the
@@
 A GLM-regression adjusted posterior was also estimated from each of the
-posterior samples \citep{Leuenberger2010}.
+posterior samples \cite{Leuenberger2010}.
 To assess the robustness of each of the four models, I also analyzed the
@@
 replicates in that bin with a true value consistent with one divergence
-event \citep{Huelsenbeck2004,Oaks2012}.
+event \cite{Huelsenbeck2004,Oaks2012}.
 If the method is unbiased, the points should fall near the identity line,
@@
 arbitrary criterion for a single ``simultaneous'' divergence event
-\citep[e.g.,][]{Hickerson2006,Leache2007,Stone2012}.
+\cite[e.g.,]{Hickerson2006,Leache2007,Stone2012}.
 I focused on the one-divergence model to assess model-choice behavior, because
@@
 detect random variation in divergence times using methods similar to
-\citet{Oaks2012}.
+\cite{Oaks2012}.
 For all power simulations, I used a data structure identical to that of the
-empirical dataset of Philippine vertebrates analyzed by \citet{Oaks2012}, which
+empirical dataset of Philippine vertebrates analyzed by \cite{Oaks2012}, which
 consists of 22 pairs of populations.
@@
 I also estimated a GLM-regression adjusted posterior from each of the
-posterior samples \citep{Leuenberger2010}.
+posterior samples \cite{Leuenberger2010}.
 
@@
 I also assessed the behavior of the newly implemented models when applied to
-the empirical dataset of \citet{Oaks2012}, which is comprised of sequence data
+the empirical dataset of \cite{Oaks2012}, which is comprised of sequence data
 from 22 pairs of taxa from the Philippine Islands (Drayad DOI:
@@
 All of these models except one (\empModelDPPSimple) have six free demographic
-parameters per pair of taxa (
-\ancestralTheta{},
+parameters per pair of taxa (\ancestralTheta{},
 \descendantTheta{1}{},
@@
 \empModelDPP, \empModelDPPInform, and \empModelDPPSimple.
-The \empModelDPP model represents the priors that \citet{Oaks2012} would have
+The \empModelDPP model represents the priors that \cite{Oaks2012} would have
 selected to reflect their prior uncertainty about the parameters of the model
@@
 estimate the posterior probabilities of divergence models.
-Consistent with the findings of \citet{Oaks2012}, my results demonstrate that
+Consistent with the findings of \cite{Oaks2012}, my results demonstrate that
 the unadjusted estimates of divergence-model posterior probabilities are
@@
 always be greater than 0.01 when divergences are random over periods of 0.1
-coalescent units or more (see Figure~S4 of \citet{Oaks2012}).
+coalescent units or more (see Figure~S4 of \cite{Oaks2012}).
 
@@
 \subsection{Empirical results}
-As expected based on the results of \citet{Oaks2012}, when the Philippines data
+As expected based on the results of \cite{Oaks2012}, when the Philippines data
 are analyzed under the \empModelOld model, there is strong support for very few
@@
 (Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
-My results are similar to those of \citet{Oaks2012} in that I find
+My results are similar to those of \cite{Oaks2012} in that I find
 \msb will often infer strong support for clustered divergences when
@@
 divergence models.
-This suggests, as predicted by \citet{Oaks2012}, that the tendency of \msb to
+This suggests, as predicted by \cite{Oaks2012}, that the tendency of \msb to
 erroneously support models of temporally clustered divergences is caused by a
@@
 For example, when applied to the dataset of 22 taxon pairs from the Philippines
-\citep{Oaks2012} the model has as many as 604--625 free parameters
+\cite{Oaks2012} the model has as many as 604--625 free parameters
 (depending on \divTimeNum), and samples over 1002 unordered divergence models.
@@
 While there is much less prior sensitivity under the new model compared to
-those observed by \citet{Oaks2012}, there is still an effect when comparing the
+those observed by \cite{Oaks2012}, there is still an effect when comparing the
 results of the empirical data analyzed under a diffuse (\empModelDPP) and
@@
 estimated scenarios
-\citep{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012},
+\cite{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012},
 as I found here (Figure~\ref{figPhilippines}).
@@
 I also urge caution when using \dppmsbayes due to the fact that ABC methods
-in general are known to be biased for model choice \citep{Robert2011}.
-As recommended by \citet{Oaks2012}, any results from the method should be
+in general are known to be biased for model choice \cite{Robert2011}.
+As recommended by \cite{Oaks2012}, any results from the method should be
 accompanied by assessments of prior sensitivity and power.
@@
 sequence data will become increasingly important for robustly estimating shared
-evolutionary history across taxa \citep{JeetDiss}.
+evolutionary history across taxa \cite{JeetDiss}.
 With improving numerical methods for sampling over models of differing
-dimensionality \citep{Green1995,Lemey2009}, advances in Monte Carlo techniques
-\citep{Jordan2012}, and increasing efficiency of likelihood calculations
-\citep{Ayres2012}, analyzing rich comparative phylogeograpical models in a
+dimensionality \cite{Green1995,Lemey2009}, advances in Monte Carlo techniques
+\cite{Jordan2012}, and increasing efficiency of likelihood calculations
+\cite{Ayres2012}, analyzing rich comparative phylogeograpical models in a
 full-likelihood Bayesian framework is becoming computationally practical,

### 2014-02-28 manuscripts/dpp-msbayes/cover-letter/cover-letter.tex
Some minor reformatting to dpp ms.
Also, updating cover letter.

@@
-\documentclass[letterpaper]{letter}
+\documentclass[letterpaper,11pt]{letter}
 \usepackage{anysize}
-\marginsize{1.25in}{1.25in}{0.5in}{0.5in}
+\marginsize{0.8in}{0.8in}{-0.2in}{0.3in}
 \usepackage{setspace}
@@
 
-\signature{Jamie Oaks \\ 
-                 Postdoctoral Fellow \\ 
+\signature{Jamie Oaks, Postdoctoral Fellow \\ 
                  Department of Biology \\
                  University of Washington \\ 
-                 Box 351800 \\ 
+                 % Box 351800 \\ 
                  Seattle, WA 98195 \\ 
@@
 \begin{document}
-\begin{letter}{Journal/Society \\
-                address \\
-                here}
+\begin{letter}{\emph{PLOS Computational Biology} Editorial Board \\
+                        1160 Battery Street \\
+                        Koshland Building East, Suite 100 \\
+                    San Francisco, CA 94111}
 \opening{Dear Editor:}
 % \raggedright
-I am submitting a manuscript entitled ``An Improved Approximate-Bayesian
-Model-choice Method for Estimating Shared Evolutionary History'' to be
-considered for publication in \highLight{journal here}.
+% I am submitting a manuscript entitled ``An Improved Approximate-Bayesian
+% Model-choice Method for Estimating Shared Evolutionary History'' to be
+% considered for publication in \highLight{journal here}.
+I am submitting a presubmission inquiry regarding a manuscript entitled ``An
+Improved Approximate-Bayesian Model-choice Method for Estimating Shared
+Evolutionary History'' that I wish to be considered for publication in
+\emph{PLOS Computational Biology}.
 The field of research is comparative biogeography and Bayesian model choice.
-All of the computer code for the newly implemented model I present in the
-the manuscript is freely available in the open-source
-software packages \texttt{dpp-msbayes} and \texttt{PyMsBayes}
+
+In the paper, I introduce a novel Bayesian method for estimating the
+probability of shared divergence histories across a set of taxa from
+multi-locus DNA sequence data.
+The new method provides an important tool for evolutionary biologists to
+investigate how large-scale changes to the environment, including geological
+and climatic events, have affected the evolutionary history of entire
+communities of co-distributed species and their associated microbiota.
+Given the dynamic nature of our planet, such biogeographical processes likely
+play a significant role in regulating biodiversification and community assembly.
+
+More specifically, given a set of species pairs, the method estimates
+the probabilities of all possible models in which multiple sets of taxa
+can diverge at the same time.
+These models range from a single divergence event shared across all of the
+taxa, to the fully generalized model in which each pair of taxa diverged at a
+unique time.
+The current state-of-the-art method for this estimation problem is implemented
+in the popular software package \texttt{msBayes}, which has a broad community
+of users.
+I introduce a non-parametric Bayesian approach to this inference problem that
+uses a Dirichlet-process prior over all possible divergence models.
+Using extensive simulations and applications to empirical data, I demonstrate
+that the new method has improved robustness, accuracy, and power for estimating
+shared evolutionary history across taxa when compared to \texttt{msBayes}.
+Also, previous simulation work has shown that \texttt{msBayes} often
+incorrectly supports biogeographically compelling scenarios of simultaneous
+divergence across taxa, even when divergences were random over broad time
+periods.
+Importantly, my results show that the new method is \emph{much} less likely to
+make such misleading and costly errors.
+
+All of the computer code for the new method presented in the the manuscript is
+freely available in the open-source software packages \texttt{dpp-msbayes} and
+\texttt{PyMsBayes}
 (\href{https://github.com/joaks1/dpp-msbayes}{\url{https://github.com/joaks1/dpp-msbayes}}
@@
 
-In this paper, I introduce a novel Bayesian method for estimating the
-probability of shared divergence histories across a set of taxa from
-multi-locus DNA sequence data.
-More specifically, given a set of species pairs, the method estimates
-the probabilities of all possible models in which multiple sets of taxa
-can diverge at the same time.
-These models range from a single divergence-time parameter (i.e., simultaneous
-divergence of all pairs of taxa) to the fully generalized model in which each
-pair of taxa diverged at a unique time.
-The current state-of-the-art method for this estimation problem is implemented
-in the software package \texttt{msBayes}.
-By introducing more flexible priors on the parameters the model and a
-non-parametric Dirichlet-process prior over divergence models, the new method
-has improved robustness, accuracy, and power for estimating shared evolutionary
-history across taxa when compared to \texttt{msBayes}.
-Also, simulations have shown that \texttt{msBayes} will often to incorrectly
-support biogeographically interesting scenarios of shared divergences when
-divergences were random over broad time periods.
-My results show that the new method is \emph{much} less likely to make such
-misleading and costly errors.
-
-The new method provides an important tool for evolutionary biologists to
-investigate how large-scale changes to the environment, including geological
-and climatic events, have affected the evolutionary history of entire
-communities of co-distributed species and their associated microbiota.
-Given the dynamic nature of our planet, such biogeographical processes likely
-play a significant role regulating biodiversification and community assembly.
-
 Thank you for your consideration, and I look forward to your correspondence
@@
 
-\addtolength{\medskipamount}{-5pt}
+\addtolength{\medskipamount}{-6pt}
 \closing{Sincerely,}

### 2014-02-28 manuscripts/reply/cover-letter/cover-letter.tex
Adding cover letter for reply ms.

\documentclass[letterpaper]{letter}
\usepackage{anysize}
\marginsize{1in}{1in}{0.5in}{0.5in}
\usepackage{setspace}
\usepackage{url}
\usepackage{verbatim}
\usepackage{parskip}
\usepackage{hyperref}
\hypersetup{pdfborder={0 0 0}, colorlinks=true, urlcolor=blue}

\newcommand{\ignore}[1]{}
\newcommand{\super}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\sub}[1]{\ensuremath{_{\textrm{#1}}}}
\newcommand{\dC}{\ensuremath{^\circ{\textrm{C}}}}

\makeatletter
\let\@texttop\relax
\makeatother

\newenvironment{myEnumerate}{
  \begin{enumerate}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}}
  {\end{enumerate}}

\newenvironment{myItemize}{
  \begin{itemize}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}}
  {\end{itemize}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\signature{Jamie Oaks \\
                 Postdoctoral Fellow \\ 
                 Department of Biology \\
                 University of Washington \\ 
                 Box 351800 \\ 
                 Seattle, WA 98195 \\ 
                 USA \\
                 \href{mailto:joaks1@uw.edu}{\texttt{joaks1@uw.edu}} \\ 
                 \href{http://www.phyletica.com}{\texttt{www.phyletica.com}}}
%\address{Biodiversity Institute \\ University of Kansas \\ Dyche Hall, 1345 Jayhawk Blvd \\ Lawrence, KS 66045}
\begin{document}
\begin{letter}{The Society for the Study of Evolution \\
                     PO Box 299 \\
                     St. Louis, MO 63166-0299}
\opening{Dear Editor:}
% \raggedright
I am submitting a manuscript entitled ``Implications of uniformly distributed,
empirically informed priors for phylogeographical model selection: A reply to
Hickerson et al.'' to be considered for publication in the \emph{International
Journal of Organic Evolution}.
This manuscript is a response to a paper by Hickerson et al.\ (2014,
\emph{Evolution} 68:284--294), which was a response to a paper published last
year by my co-authors and me (Oaks et al.\ 2013, \emph{Evolution}
67:991--1010).
Former Associate Editor Dr.\ Laura Kubatko invited us to write this response,
with the intention of handling its review.
The following is an excerpt from Dr.\ Kubatko's comments in the decision letter
for Hickerson et al.\ (2014):

\begin{quote}
Thus, I propose the following: Hickerson et al. should prepare and
submit the final version of their response.  I urge them to fully
consider all of the points raised by Oaks in his most recent review of their
paper. However, since the point of view of Hickerson et al. seems to
be at odds with Oaks et al., I would like to give Oaks et al. the opportunity
to respond, if they so choose. I have therefore recommended acceptance
for this version of the manuscript, and I ask that Hickerson et al. provide a
copy of the final version to Jamie Oaks once they have submitted it, so
that he and his co-authors can consider whether they wish to respond.
\end{quote}

In our response, we use simulated and biological data to evaluate the empirical
Bayesian method proposed by Hickerson et al.\ (2014) as a solution to the
problems revealed by Oaks et al.\ (2013) regarding the power and accuracy of the
comparative phylogeographical method implemented in the software
\texttt{msBayes}.
We demonstrate their approach does not mitigate the method's bias toward
supporting models with less parameter space, but rather allows it to manifest
in strong preference for models that often exclude from consideration the true
values of the model's parameters.
Our results also strongly suggest the underlying statistical cause of the
method's tendency to support simple models.
More generally, we discuss the theoretical and practical issues of empirical
Bayesian approaches to model choice; these issues are important for empiricists
applying methods of Bayesian model selection to carefully consider.
Lastly, we discuss ways to mitigate the biases of the \texttt{msBayes} method
while avoiding negative side effects of empirically informed uniform priors.

We performed the work described in the manuscript following the principles of
Open Notebook Science; using version-control software, we make progress in all
aspects of the work freely and publicly available in real-time at
\href{https://github.com/joaks1/msbayes-experiments}{\url{https://github.com/joaks1/msbayes-experiments}}.
All information necessary to reproduce my results is provided there.

Thank you for your consideration, and I look forward to your correspondence
regarding this manuscript.

\addtolength{\medskipamount}{-5pt}
\closing{Sincerely,}
\end{letter}

\end{document}

### 2014-03-01 manuscripts/reply/reply-abstract.tex
Reformating reply ms for Evolution.
Also moving the abstract and acknowledgments to separate files, and
resizing the sampling error figure to fit within the margins.

--- 2014-02-26 manuscripts/reply/reply.tex
@@
-%&<latex>
-\documentclass[letterpaper,12pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{../utils/preamble.tex}
-\input{../utils/macros.tex}
-
-\title{Implications of uniformly distributed, empirically informed priors for
-phylogeographical model selection: A reply to Hickerson et al.}
-
-\author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
-\author[2]{Charles W.\ Linkem}
-\author[3]{Jeet Sukumaran}
-\affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
-\affil[2]{Department of Biology, University of Washington, Seattle, Washington 98195}
-\affil[3]{Department of Ecology and Evolutionary Biology, University of Michigan, Ann Arbor, Michigan 48109}
-
-% \date{\today}
-\date{\parbox{\linewidth}{\centering%
-    \today\endgraf\bigskip
-    \textbf{Running head}: Approximate Bayesian model choice}}
-
-\makeatletter
-\let\msTitle\@title
-\let\msAuthor\@author
-\let\msDate\@date
-\makeatother
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-% \doublespacing
-% \raggedright
-% \setlength{\parindent}{0.5in}
-% \begin{linenumbers}
-
-\maketitle
-
-\begin{abstract}
 Biogeographers frequently seek to explain population and species
@@
 the method, which in turn helps guide interpretation of results.
-
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Approximate Bayesian computation; Bayesian
-        model choice; empirical Bayes} 
-\end{abstract}
-
-\newpage
-
-\input{reply-main.tex}
-
-\section{Acknowledgments}
-We thank Melissa Callahan, Jake Esselstyn, Cameron Siler, Mark Holder, Rafe
-Brown, Emily McTavish, Daniel Money, Jordan Koch, and Adam Leach\'{e} for
-insightful comments that greatly improved this work.
-We thank Michael Hickerson and co-authors for generously providing their data.
-J.\ Oaks and C.\ Linkem thank the National Science Foundation for supporting
-this work (DEB 1011423, DBI 1308885 and BIO-1202754).
-J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
-Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
-Society, KU Department of Ecology and Evolutionary Biology, and the KU
-Biodiversity Institute.
-We also thank Mark Holder, the KU Information and Telecommunication Technology
-Center, KU Computing Center, and the iPlant Collaborative for the computational
-support necessary to conduct the analyses presented herein.
-
-\bibliography{../bib/references}
-
-%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\renewcommand\listfigurename{Figure Captions}
-\cftsetindents{fig}{0cm}{2.2cm}
-\renewcommand\cftdotsep{\cftnodots}
-\setlength\cftbeforefigskip{10pt}
-\cftpagenumbersoff{fig}
-\listoffigures
-
-% \end{linenumbers}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\input{reply-tables.tex}
-
-\clearpage
-
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
-\input{reply-figures.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-\setcounter{section}{0}
-
-\singlespacing
-
-\section*{Supporting Information}
-% PUT MAIN TEXT CITATION HERE
-\hangindent=1cm
-Oaks, J.~R., C.~W.\ Linkem, and J.~Sukumaran. \msTitle
-
-% \doublespacing
-\input{reply-si-main.tex}
-
-\newpage
-\singlespacing
-
-\input{reply-si-figures.tex}
-
-\end{document}

### 2014-03-01 manuscripts/reply/reply-acknowledgments.tex
Reformating reply ms for Evolution.
Also moving the abstract and acknowledgments to separate files, and
resizing the sampling error figure to fit within the margins.

--- 2014-02-26 manuscripts/reply/reply.tex
@@
-%&<latex>
-\documentclass[letterpaper,12pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{../utils/preamble.tex}
-\input{../utils/macros.tex}
-
-\title{Implications of uniformly distributed, empirically informed priors for
-phylogeographical model selection: A reply to Hickerson et al.}
-
-\author[1]{Jamie R.\ Oaks\thanks{Corresponding author: \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}}\thanks{Current address: Department of Biology, University of Washington, Seattle, Washington 98195}}
-\author[2]{Charles W.\ Linkem}
-\author[3]{Jeet Sukumaran}
-\affil[1]{Department of Ecology and Evolutionary Biology, University of Kansas, Lawrence, Kansas 66045}
-\affil[2]{Department of Biology, University of Washington, Seattle, Washington 98195}
-\affil[3]{Department of Ecology and Evolutionary Biology, University of Michigan, Ann Arbor, Michigan 48109}
-
-% \date{\today}
-\date{\parbox{\linewidth}{\centering%
-    \today\endgraf\bigskip
-    \textbf{Running head}: Approximate Bayesian model choice}}
-
-\makeatletter
-\let\msTitle\@title
-\let\msAuthor\@author
-\let\msDate\@date
-\makeatother
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-% \doublespacing
-% \raggedright
-% \setlength{\parindent}{0.5in}
-% \begin{linenumbers}
-
-\maketitle
-
-\begin{abstract}
-    Biogeographers frequently seek to explain population and species
-    differentiation on geographical phenomena.
-    Establishing that a set of population splitting events occurred
-    at the same time can be a potentially persuasive argument that a set of taxa
-    were affected by the same geographic events.
-    Unfortunately, estimating divergence times precisely when one lacks
-    precise information about the rate of molecular evolution is a notoriously
-    difficult problem in evolutionary biology.
-    \citet{Huang2011} introduced an approximate Bayesian approach (implemented
-    in the software \msb) which can be used to estimate the probabilities of
-    models in which multiple sets of taxa diverge at the same time. 
-    % This method  does not require a common set of character data that links all
-    % of the taxa under investigation, but does require the relative mutation
-    % rates among the taxa and loci to be assumed \emph{a priori}.
-    Recently, \citet{Oaks2012} used this model-choice framework to 
-    study 22 pairs of vertebrate lineages which are distributed across
-    the Philippines; they also studied the behavior of the \msb approach
-    using computer simulations.
-    \citet{Oaks2012} found that the model was very sensitive to the prior and
-    that the inference method had low power to detect variation in divergences
-    times.
-    These results were not surprising, in light of a rich statistical literature
-    showing that the marginal likelihood of a model is sensitive to the
-    use of vague priors.
-    Because this sensitivity to prior assumptions affects the crucial insights 
-    that a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended
-    that users of the approach should carefully assess the robustness of their 
-    conclusions to different priors.
-    According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
-    was due to excessively broad priors on divergence times, leading to 
-    inadequate numbers of simulation replicates.
-    They proposed a new model-averaging approach that uses narrow, empirically
-    informed uniform priors.
-    Here we demonstrate that the approach of \citet{Hickerson2013} is
-    dangerous in the sense that the empirically-derived priors often
-    exclude from consideration the true values of the models' parameters.
-    On a more fundamental level, we question the value of adopting an empirical
-    Bayesian stance for this model-choice problem, because it can mislead model
-    posterior probabilities, which are inherently measures of belief in the
-    models after prior knowledge is updated by the data.
-    The robust Bayesian approach of conducting analyses under a variety
-    of priors can reveal prior sensitivity and communicate which assumptions
-    underlie model inference.
-    Furthermore, simulations provide insight into the temporal resolution of
-    the method, which in turn helps guide interpretation of results.
-
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Approximate Bayesian computation; Bayesian
-        model choice; empirical Bayes} 
-\end{abstract}
-
-\newpage
-
-\input{reply-main.tex}
-
-\section{Acknowledgments}
 We thank Melissa Callahan, Jake Esselstyn, Cameron Siler, Mark Holder, Rafe
@@
 support necessary to conduct the analyses presented herein.
-
-\bibliography{../bib/references}
-
-%% LIST OF FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\renewcommand\listfigurename{Figure Captions}
-\cftsetindents{fig}{0cm}{2.2cm}
-\renewcommand\cftdotsep{\cftnodots}
-\setlength\cftbeforefigskip{10pt}
-\cftpagenumbersoff{fig}
-\listoffigures
-
-% \end{linenumbers}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\singlespacing
-
-\input{reply-tables.tex}
-
-\clearpage
-
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
-\input{reply-figures.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-\setcounter{section}{0}
-
-\singlespacing
-
-\section*{Supporting Information}
-% PUT MAIN TEXT CITATION HERE
-\hangindent=1cm
-Oaks, J.~R., C.~W.\ Linkem, and J.~Sukumaran. \msTitle
-
-% \doublespacing
-\input{reply-si-main.tex}
-
-\newpage
-\singlespacing
-
-\input{reply-si-figures.tex}
-
-\end{document}

### 2014-03-01 manuscripts/reply/reply-figures.tex
A few edits to reply ms.
Reformating reply ms for Evolution.
Also moving the abstract and acknowledgments to separate files, and
resizing the sampling error figure to fit within the margins.

@@
 %% Sampling error
-\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
+\widthFigure{0.9}{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{listformat=figList}{
     Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
@@
     samples are accumulated. Each pair of points is based on 1000 posterior samples retained
-    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
+    from the prior. Both (A) unadjusted and (B) GLM-regression-adjusted estimates are shown.
+    The data analyzed were the 22 pairs of Philippine taxa from \citet{Oaks2012}.
     Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
@@
 }{figSamplingError}
+\clearpage
 

### 2014-03-01 manuscripts/reply/reply.tex
Moving `bibliographystyle` call in manuscripts.
Moving from preamble to main tex documents for each of the manuscripts.
Moving use of natbib to reply ms only.
Reformating reply ms for Evolution.
Also moving the abstract and acknowledgments to separate files, and
resizing the sampling error figure to fit within the margins.

@@
 \input{../utils/preamble.tex}
+\usepackage[round]{natbib}
 \input{../utils/macros.tex}
+
+\newcommand{\keywords}{Approximate Bayesian computation; Bayesian model choice;
+empirical Bayes; phylogeography; biogeography\xspace}
 
@@
     \today\endgraf\bigskip
-    \textbf{Running head}: Approximate Bayesian model choice}}
+    \textbf{Running head}: Approximate Bayesian model choice\endgraf\bigskip
+    \textbf{Key words}: \keywords\endgraf\bigskip
+    \textbf{Counts}: 7219 words, 7 figures, 1 table, 7 supporting figures\endgraf\bigskip
+    \textbf{Archival location}: \href{https://github.com/joaks1/msbayes-experiments}{\url{https://github.com/joaks1/msbayes-experiments}}}}
 
@@
 \begin{document}
-% \doublespacing
-% \raggedright
-% \setlength{\parindent}{0.5in}
-% \begin{linenumbers}
 
@@
 
+\newpage
+\doublespacing
+\raggedright
+% \setlength{\parindent}{0.5in}
+\begin{linenumbers}
+
 \begin{abstract}
-    Biogeographers frequently seek to explain population and species
-    differentiation on geographical phenomena.
-    Establishing that a set of population splitting events occurred
-    at the same time can be a potentially persuasive argument that a set of taxa
-    were affected by the same geographic events.
-    Unfortunately, estimating divergence times precisely when one lacks
-    precise information about the rate of molecular evolution is a notoriously
-    difficult problem in evolutionary biology.
-    \citet{Huang2011} introduced an approximate Bayesian approach (implemented
-    in the software \msb) which can be used to estimate the probabilities of
-    models in which multiple sets of taxa diverge at the same time. 
-    % This method  does not require a common set of character data that links all
-    % of the taxa under investigation, but does require the relative mutation
-    % rates among the taxa and loci to be assumed \emph{a priori}.
-    Recently, \citet{Oaks2012} used this model-choice framework to 
-    study 22 pairs of vertebrate lineages which are distributed across
-    the Philippines; they also studied the behavior of the \msb approach
-    using computer simulations.
-    \citet{Oaks2012} found that the model was very sensitive to the prior and
-    that the inference method had low power to detect variation in divergences
-    times.
-    These results were not surprising, in light of a rich statistical literature
-    showing that the marginal likelihood of a model is sensitive to the
-    use of vague priors.
-    Because this sensitivity to prior assumptions affects the crucial insights 
-    that a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended
-    that users of the approach should carefully assess the robustness of their 
-    conclusions to different priors.
-    According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
-    was due to excessively broad priors on divergence times, leading to 
-    inadequate numbers of simulation replicates.
-    They proposed a new model-averaging approach that uses narrow, empirically
-    informed uniform priors.
-    Here we demonstrate that the approach of \citet{Hickerson2013} is
-    dangerous in the sense that the empirically-derived priors often
-    exclude from consideration the true values of the models' parameters.
-    On a more fundamental level, we question the value of adopting an empirical
-    Bayesian stance for this model-choice problem, because it can mislead model
-    posterior probabilities, which are inherently measures of belief in the
-    models after prior knowledge is updated by the data.
-    The robust Bayesian approach of conducting analyses under a variety
-    of priors can reveal prior sensitivity and communicate which assumptions
-    underlie model inference.
-    Furthermore, simulations provide insight into the temporal resolution of
-    the method, which in turn helps guide interpretation of results.
+    \input{reply-abstract.tex}
 
-    \vspace{12pt}
-    \noindent\textbf{KEY WORDS: Approximate Bayesian computation; Bayesian
-        model choice; empirical Bayes} 
+    % \vspace{12pt}
+    % \noindent\textbf{KEY WORDS: \keywords} 
 \end{abstract}
@@
 \section{Acknowledgments}
-We thank Melissa Callahan, Jake Esselstyn, Cameron Siler, Mark Holder, Rafe
-Brown, Emily McTavish, Daniel Money, Jordan Koch, and Adam Leach\'{e} for
-insightful comments that greatly improved this work.
-We thank Michael Hickerson and co-authors for generously providing their data.
-J.\ Oaks and C.\ Linkem thank the National Science Foundation for supporting
-this work (DEB 1011423, DBI 1308885 and BIO-1202754).
-J.\ Oaks was also supported by the University of Kansas (KU) Office of Graduate
-Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
-Society, KU Department of Ecology and Evolutionary Biology, and the KU
-Biodiversity Institute.
-We also thank Mark Holder, the KU Information and Telecommunication Technology
-Center, KU Computing Center, and the iPlant Collaborative for the computational
-support necessary to conduct the analyses presented herein.
+\input{reply-acknowledgments.tex}
 
+\bibliographystyle{../bib/evolution.bst}
 \bibliography{../bib/references}
@@
 \newpage
-\singlespacing
+% \singlespacing
 
@@
 
-% \end{linenumbers}
+\end{linenumbers}
 

### 2014-03-01 manuscripts/reply/reply-main.tex
Updating reply ms.
Moving coin flip figure to SI, changing to right justified, minor edit
to section title.

@@
 a beta prior on fairness of the new coin of $beta(a=10, b=10)$
-(Figure~\ref{figCoinFlip}).
+(Figure~S\ref{figCoinFlip}).
 We collect data by flipping the coin of interest $N=10$ times, $y=3$ of which
@@
 beta(a + y, b + N - y)$, which for the new dataset is simply $beta(13, 17)$
-(Figure~\ref{figCoinFlip}).
+(Figure~S\ref{figCoinFlip}).
 The maximum a posteriori (MAP) estimate of the probability of heads is 0.429,
@@
 results in a ``posterior'' distribution of $beta(16, 24)$
-(Figure~\ref{figCoinFlip}), with a MAP estimate of 0.395, and $p(M_1 \given
+(Figure~S\ref{figCoinFlip}), with a MAP estimate of 0.395, and $p(M_1 \given
 y=3, N=10) = 0.10$.
@@
 By using the empirically informed prior, we ignored prior uncertainty, leading
-to an underestimate of our posterior uncertainty (Figure~\ref{figCoinFlip}).
+to an underestimate of our posterior uncertainty (Figure~S\ref{figCoinFlip}).
 While this did not greatly affect our estimate of \myTheta{}, it misled us
@@
 
-\section{The importance of power analysis to guide applications of \texttt{msBayes}}
+\section{The importance of power analyses to guide applications of \texttt{msBayes}}
 \citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform

### 2014-03-01 model-choice/bin/plot-power-results-pretty.py
Updating plots for reply ms.
Manuscript central cannot handle any Type 3 fonts in PDFs.  matplotlib
uses Type 3 fonts by default. I had to update plotting code and re-plot
most of the figures for the reply manuscript so that they now use
TrueType fonts.

@@ -426,7 +426,7 @@ def create_plots(info_path):
             margin_left = 0.025,
             margin_bottom = 0.1,
             margin_right = 1,
-            margin_top = 0.85,
+            margin_top = 0.84,
             padding_between_horizontal = 0.5)
 
     fig = ex_plot.create_column_grid(

### 2014-03-05 manuscripts/utils/macros.tex
Working on empirical model table for dpp ms.
Working on model tables for dpp manuscript.
Working on dpp manuscript.
Proofreading and incorporating edits throughout the manuscript.
More editing of DPP ms.
Breaking up dpp ms into more files.
Moving abstract and acknowledgments to separate files, and also making
the title a macro.
Updating manuscripts' preamble and maco files.
Moving some definitions from preamble to macro file.
Reverting back to numerical citations for dpp ms.
Reverting back to the use of `cite` in place of `citep` and `citep` for
dpp ms.

Changes were made in:
    `dpp-msbayes-

@@
 \newcommand{\change}[1]{{\color{black} #1}\xspace}
+
+\definecolor{myGray}{gray}{0.9}
+
+%% redefine multicolumn to not override rowcolors
+\let\oldmulticolumn\multicolumn
+\makeatletter
+\newcommand{\mcrowcolors}{%
+  \renewcommand{\multicolumn}[3]{%
+    \oldmulticolumn{##1}{##2}{\ifodd\rownum \@oddrowcolor\else\@evenrowcolor\fi ##3}%
+  }}
+\makeatother
+
+\newcommand{\dppTitle}{An Improved Approximate-Bayesian Model-choice Method for
+Estimating Shared Evolutionary History\xspace}
 
@@
 \newcommand{\dC}{\ensuremath{^\circ{\textrm{C}}}}
+\newcommand{\tb}{\hspace{2em}}
 
@@
 \newcommand{\descendantTheta}[2]{\ensuremath{\theta_{D#1\protect\ifArg{#2}{,}#2}}\xspace}
+\newcommand{\popIndex}{\ensuremath{z}\xspace}
+\newcommand{\dThetas}{\ensuremath{\descendantTheta{\popIndex}{}}\xspace}
 \newcommand{\meanDescendantTheta}[1]{\ensuremath{\descendantTheta{}{#1}}\xspace}
@@
     plot symbols.}
-\newcommand{\timeConversionComment}{The divergence times were randomly drawn as
+\newcommand{\timeConversionComment}{The 22 divergence times were randomly drawn as
     indicated above each column of plots, where time is respresented as
     millions of generations ago (MGA) according to a per-site rate of 1\e{-8}
-    mutations per generation.}
+    mutations per generation.\xspace}
 \newcommand{\validationModelChoiceComparisonCaption}[2]{Model-choice accuracy
@@
     of divergence times ($\hat{\vmratio{}}$) from \simulationScheme.
-    The threshold for one divergence event \citep{Hickerson2006} is indicated
+    The threshold for one divergence event \cite{Hickerson2006} is indicated
     by the dashed line, and the estimated probability of inferring one

### 2014-03-06 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Adding bmc-style figure titles to dpp ms.
Adding bmc-style figure titles to `dpp-msbayes-figure.tex` enclosed
withing `ifbmc` statements.

@@
 %% Validation model choice
-\mFigure{../../validation/images/validation-model-choice-psi.pdf}{
+\mFigure{../../validation/images/validation-model-choice-psi.pdf}{%
+    \ifbmc{\csentence{Comparison of model-choice behavior.}}{}
     \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeNum = 1$}
@@
 \mFigure{../../validation/images/validation-model-choice-omega.pdf}{
+    \ifbmc{\csentence{Comparison of model-choice behavior.}}{}
     \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeDispersion < 0.01$}
@@
 \mFigure{../../power-comparison/images/old-power-psi-4.pdf}{
+    \ifbmc{\csentence{Power comparison.}}{}
     \powerCommentSummary{\powerSeriesOld}
@@
 \mFigure{../../power-comparison/images/old-power-omega-4.pdf}{
+    \ifbmc{\csentence{Power comparison.}}{}
     \powerCommentSummary{\powerSeriesOld}
@@
 \mFigure{../../power-comparison/images/old-power-psi-prob-4.pdf}{
+    \ifbmc{\csentence{Power comparison.}}{}
     \powerSupportCommentSummary{\powerSeriesOld}
@@
 \mFigure{../../power-comparison/images/old-power-omega-prob-4.pdf}{
+    \ifbmc{\csentence{Power comparison.}}{}
     \powerSupportCommentSummary{\powerSeriesOld}
@@
 \mFigure{../../empirical-analyses/plots/philippines-psi.pdf}{
+    \ifbmc{\csentence{Empirical results for 22 Philippine taxa.}}{}
     The (A--E) posterior and (F--J) prior probabilities of the number of
@@
 \widthFigure{0.4}{../../empirical-analyses/plots/negros-panay-psi.pdf}{listformat=figList}{
+    \ifbmc{\csentence{Empirical results for 9 Philippine taxa.}}{}
     The posterior probabilities of the number of divergence events,

### 2014-03-06 manuscripts/dpp-msbayes/dpp-msbayes-abstract.tex
Updating abstract for dpp ms.
Adding section titles to the abstract per bmc guidelines. The
bmc-specific section title commands are wrapped in `ifbmc` statements.

@@
+\ifbmc{\parttitle{Background}}{}
 To understand the processes that generate biodiversity, it is
@@
 The model is an extension of that implemented in \msb.
+\ifbmc{\parttitle{Results}}{}
 By reparameterizing the model, introducing more flexible priors on
@@
 evolutionary history across taxa.
+\ifbmc{\parttitle{Conclusions}}{}
 The results demonstrate the improved performance of the new method is due

### 2014-03-06 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Reverting bib management in dpp ms.
Reverting back to natbib bib management for dpp ms.
Reverting back to numerical citations for dpp ms.
Reverting back to the use of `cite` in place of `citep` and `citep` for
dpp ms.

Changes were made in:
    `dpp-msbayes-main.tex`
    `dpp-msbayes-tables.tex`
    `../utils/macros.tex`
Updating main text of dpp ms.
Updating the section titles and adding a conclusions section to
`dpp-msbayes-main.tex`.

@@
-\section{Introduction}
+% \section{Introduction}
+\section{Background}
 
@@
 parameter classes \cite{Ferguson1973,Antoniak1974}.
-The Dirichlet process has been used as a nonparametric Bayesian approach to
+The Dirichlet process has been used as a non-parametric Bayesian approach to
 many inference problems in evolutionary biology
@@
 computationally nontrivial.
+
+\section{Conclusions}
+I introduced a new approximate-Bayesian model-choice method for estimating
+shared divergence histories across taxa using DNA sequence data.
+The new method, \dppmsbayes, takes a non-parametric approach to the problem by
+using a Dirichlet-process prior on the temporal distribution of divergences
+across taxa.
+The new method shows improved robustness, accuracy, and power compared to the
+existing method, \msb.
+Compared to \msb, the new approach better estimates posterior uncertainty,
+which greatly reduces the chances of incorrectly estimating biogeographically
+interesting scenarios of shared divergence events.
+This new tool will allow evolutionary biologists to better leverage comparative
+genetic data to assess the affects of regional and global biogeographical
+processes on biodiversity.

### 2014-03-06 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Fixing figure references in dpp figures.
In `dpp-msbayes-figures.tex` some of the figure captions reference other
figures in the SI. A copy-and-paste error had all of them referencing
the same figure in the SI. This commit corrects these references.

@@
     are shown;
-    please see Figure~S\ref{figPowerPsiOld} for all results.
+    please see Figure~S\ref{figPowerOmegaOld} for all results.
 }{figPowerOmegaOld4}
@@
     are shown;
-    please see Figure~S\ref{figPowerPsiOld} for all results.
+    please see Figure~S\ref{figPowerPsiProbOld} for all results.
 }{figPowerPsiProbOld4}
@@
     are shown;
-    please see Figure~S\ref{figPowerPsiOld} for all results.
+    please see Figure~S\ref{figPowerOmegaProbOld} for all results.
 }{figPowerOmegaProbOld4}

### 2014-03-07 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Updating references in dpp ms.
Editing sentences where `citet` was previously used. Making these
sentences compatible with the current use of `cite`.

Also, adding HKY85 reference, and updating natbib to sort and compress
numerical citations.

@@
 
-Recently, \cite{Oaks2012} and \cite{Oaks2014reply} found via simulation that
+Recently, Oaks et al.\ \cite{Oaks2012,Oaks2014reply} found via simulation that
 \msb will often strongly support biogeographically interesting models with a
@@
 U-shaped prior on divergence models.
-\cite{Hickerson2013} suggested the problem is caused by sampling error, and
-proposed as a solution an approximate-Bayesian model averaging approach that
-samples over empirically informed uniform priors.
-However, \cite{Oaks2014reply} evaluated the approach proposed by
-\cite{Hickerson2013} using simulations and found that it did not mitigate the
-method's propensity to incorrectly infer clustered divergences, and often
-preferred priors that excluded the true values of the model's parameters.
+Hickerson et al.\ \cite{Hickerson2013} suggested the problem is caused by
+sampling error, and proposed as a solution an approximate-Bayesian model
+averaging approach that samples over empirically informed uniform priors.
+However, Oaks et al.\ \cite{Oaks2014reply} evaluated the approach proposed by
+Hickerson et al.\ \cite{Hickerson2013} using simulations and found that it did
+not mitigate the method's propensity to incorrectly infer clustered
+divergences, and often preferred priors that excluded the true values of the
+model's parameters.
 Here, I describe a new method that successfully mitigates spurious inference
@@
 All information necessary to reproduce my results is provided there.
-I follow much of the notation of \cite{Oaks2012}, but modify it to
+I follow much of the notation of Oaks et al.\ \cite{Oaks2012}, but modify it to
 aid in the description of the new model.
@@
 The first term,
-$p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})$,
-is the probability of the sequence alignment of locus $j$ for population pair
-$i$ given the gene tree and \hky substitution model parameters \cite[i.e., the
-``Felsenstein likelihood'';]{Felsenstein1981}.
+$p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})$, is the
+probability of the sequence alignment of locus $j$ for population pair $i$
+given the gene tree and \hky \cite{HKY} substitution model parameters
+\cite[i.e., the ``Felsenstein likelihood'']{Felsenstein1981}.
 The model allows for an intra-locus recombination rate \recombinationRate,
@@
 \begin{linenomath}
-As described by \cite{Oaks2012}, to get the divergence times in units
-proportional to the expected number of mutations, they must be
-scaled by the realized population size for locus $j$ of population-pair $i$ 
+As described by Oaks et al.\ \cite{Oaks2012}, to get the divergence times in
+units proportional to the expected number of mutations, they must be scaled by
+the realized population size for locus $j$ of population-pair $i$ 
 \begin{equation}
@@
 is the prior distribution on divergence models.
-As discussed in \cite{Oaks2012}, in \msb the prior used for
+As discussed in Oaks et al.\ \cite{Oaks2012}, in \msb the prior used for
 \divTimeIndexVector is a combination of a discrete uniform prior over the
@@
 I denote this prior used by \msb as $\divTimeIndexVector \sim \priorOld$.
-\cite{Oaks2012} discuss how placing a uniform prior over the number of
-divergence parameters (denoted \divTimeNum here, and as \numt{} in
+Oaks et al.\ \cite{Oaks2012} discuss how placing a uniform prior over the
+number of divergence parameters (denoted \divTimeNum here, and as \numt{} in
 \cite{Huang2011}) imposes an U-shaped prior over divergence
@@
 populations of each pair.
-As described by \cite{Oaks2012}, \msb uses the joint prior
+As described by Oaks et al.\ \cite{Oaks2012}, \msb uses the joint prior
 \begin{equation}
@@
 of the prior on descendant-population size.
-As described by \cite{Oaks2012}, in \msb, \myTheta{C} is half of the upper
-limit of the continuous uniform prior on the mean of the descendant population
-sizes.
+As described by Oaks et al.\ \cite{Oaks2012}, in \msb, \myTheta{C} is half of
+the upper limit of the continuous uniform prior on the mean of the descendant
+population sizes.
 This is only equal to the expectation of the prior if the lower limit is zero.
@@
 pairwise differences between the descendant populations
-\cite[$\pi_b$;]{NeiLi1979,Huang2011}.
+($\pi_b$) \cite{NeiLi1979,Huang2011}.
 This has the advantage of reducing the sample space of the number of
@@
 \cite{Leuenberger2010} as implemented in \abctoolbox
-\cite[v1.1;]{ABCtoolbox}, which \cite{Oaks2012} showed performs very
-similarly to weighted local-linear regression and multinomial logistic
+\cite[v1.1;]{ABCtoolbox}, which Oaks et al.\ \cite{Oaks2012} showed performs
+very similarly to weighted local-linear regression and multinomial logistic
 regression adjustments \cite{Beaumont2002} for \msb posteriors.
@@
 \subsection{Assessing model-choice behavior and robustness}
-Following the simulation-based approach of \cite{Oaks2012}, I characterize
-the behavior of several models under the ideal conditions where the
-data are generated from parameters drawn from the same prior distributions used
-for analysis (i.e., the prior is correct).
+Following the simulation-based approach of Oaks et al.\ \cite{Oaks2012}, I
+characterize the behavior of several models under the ideal conditions where
+the data are generated from parameters drawn from the same prior distributions
+used for analysis (i.e., the prior is correct).
 I selected the following four model priors for these analyses
@@
 arbitrary criterion for a single ``simultaneous'' divergence event
-\cite[e.g.,]{Hickerson2006,Leache2007,Stone2012}.
+(e.g., \cite{Hickerson2006,Leache2007,Stone2012}).
 I focused on the one-divergence model to assess model-choice behavior, because
@@
 detect random variation in divergence times using methods similar to
-\cite{Oaks2012}.
+Oaks et al.\ \cite{Oaks2012}.
 For all power simulations, I used a data structure identical to that of the
-empirical dataset of Philippine vertebrates analyzed by \cite{Oaks2012}, which
-consists of 22 pairs of populations.
+empirical dataset of Philippine vertebrates analyzed by Oaks et al.\
+\cite{Oaks2012}, which consists of 22 pairs of populations.
 Due to the larger number of pairs, I used a different hyperprior on the
@@
 I also assessed the behavior of the newly implemented models when applied to
-the empirical dataset of \cite{Oaks2012}, which is comprised of sequence data
-from 22 pairs of taxa from the Philippine Islands (Drayad DOI:
+the empirical dataset of Oaks et al.\ \cite{Oaks2012}, which is comprised of
+sequence data from 22 pairs of taxa from the Philippine Islands (Drayad DOI:
 10.5061/dryad.5s07m).
@@
 \empModelDPP, \empModelDPPInform, and \empModelDPPSimple.
-The \empModelDPP model represents the priors that \cite{Oaks2012} would have
-selected to reflect their prior uncertainty about the parameters of the model
-if provided the more flexible distributions that are now implemented.
+The \empModelDPP model represents the priors that Oaks et al.\ \cite{Oaks2012}
+would have selected to reflect their prior uncertainty about the parameters of
+the model if provided the more flexible distributions that are now implemented.
 To assess prior sensitivity, the \empModelDPPInform model uses a more
@@
 estimate the posterior probabilities of divergence models.
-Consistent with the findings of \cite{Oaks2012}, my results demonstrate that
-the unadjusted estimates of divergence-model posterior probabilities are
-generally more accurate than regression-adjusted estimates
+Consistent with the findings of Oaks et al.\ \cite{Oaks2012}, my results
+demonstrate that the unadjusted estimates of divergence-model posterior
+probabilities are generally more accurate than regression-adjusted estimates
 (compare the plots along the upper-left to lower-right diagonal for
@@
 % literature.
-Although the \divTimeDispersion threshold of 0.01 is arbitrary, \cite{Oaks2012}
-did show via simulation that the true value of \divTimeDispersion will almost
-always be greater than 0.01 when divergences are random over periods of 0.1
-coalescent units or more (see Figure~S4 of \cite{Oaks2012}).
+Although the \divTimeDispersion threshold of 0.01 is arbitrary, Oaks et al.\
+\cite{Oaks2012} did show via simulation that the true value of
+\divTimeDispersion will almost always be greater than 0.01 when divergences are
+random over periods of 0.1 coalescent units or more (see Figure~S4 of
+\cite{Oaks2012}).
 
@@
 \subsection{Empirical results}
-As expected based on the results of \cite{Oaks2012}, when the Philippines data
-are analyzed under the \empModelOld model, there is strong support for very few
-divergence events shared among all 22 pairs of taxa, with a maximum \emph{a
-posteriori} (MAP) estimate of one-shared divergence
+As expected based on the results of Oaks et al.\ \cite{Oaks2012}, when the
+Philippines data are analyzed under the \empModelOld model, there is strong
+support for very few divergence events shared among all 22 pairs of taxa, with
+a maximum \emph{a posteriori} (MAP) estimate of one-shared divergence
 (Figure~\ref{figPhilippines}A).
@@
 (Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
-My results are similar to those of \cite{Oaks2012} in that I find
-\msb will often infer strong support for clustered divergences when
-divergences are random over quite broad timescales
+My results are similar to those of Oaks et al.\ \cite{Oaks2012} in that I find
+\msb will often infer strong support for clustered divergences when divergences
+are random over quite broad timescales
 (Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
@@
 divergence models.
-This suggests, as predicted by \cite{Oaks2012}, that the tendency of \msb to
-erroneously support models of temporally clustered divergences is caused by a
-combination of
+This suggests, as predicted by Oaks et al.\ \cite{Oaks2012,Oaks2014reply}, that
+the tendency of \msb to erroneously support models of temporally clustered
+divergences is caused by a combination of
 (1) small marginal likelihoods of models with more \divTime{} parameters due to
@@
 For example, when applied to the dataset of 22 taxon pairs from the Philippines
-\cite{Oaks2012} the model has as many as 604--625 free parameters
+\cite{Oaks2012}, the model has as many as 604--625 free parameters
 (depending on \divTimeNum), and samples over 1002 unordered divergence models.
@@
 While there is much less prior sensitivity under the new model compared to
-those observed by \cite{Oaks2012}, there is still an effect when comparing the
-results of the empirical data analyzed under a diffuse (\empModelDPP) and
-informative (\empModelDPPInform) divergence-time prior
+those observed by Oaks et al.\ \cite{Oaks2012}, there is still an effect when
+comparing the results of the empirical data analyzed under a diffuse
+(\empModelDPP) and informative (\empModelDPPInform) divergence-time prior
 (Figure~\ref{figPhilippines} C versus D).
@@
 estimated scenarios
-\cite{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012},
+(e.g.,
+\cite{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012}),
 as I found here (Figure~\ref{figPhilippines}).
@@
 in general are known to be biased for model choice \cite{Robert2011}.
-As recommended by \cite{Oaks2012}, any results from the method should be
-accompanied by assessments of prior sensitivity and power.
+As recommended by Oaks et al.\ \cite{Oaks2012}, any results from the method
+should be accompanied by assessments of prior sensitivity and power.
 I encourage investigators to view this method as a means of exploring their

### 2014-03-07 manuscripts/dpp-msbayes/cover-letter/cover-letter.tex
Updating cover letter for dpp ms.

@@
                  % Box 351800 \\ 
-                 Seattle, WA 98195 \\ 
+                 Seattle, WA 98195, USA \\
                  \href{mailto:joaks1@uw.edu}{\texttt{joaks1@uw.edu}} \\ 
@@
 \begin{document}
-\begin{letter}{\emph{PLOS Computational Biology} Editorial Board \\
-                        1160 Battery Street \\
-                        Koshland Building East, Suite 100 \\
-                    San Francisco, CA 94111}
+\begin{letter}{\emph{BMC Evolutionary Biology} \\
+                        BioMed Central \\
+                        236 Gray's Inn Road \\
+                        London WC1X 8HB \\
+                        United Kingdom}
 \opening{Dear Editor:}
 % \raggedright
-% I am submitting a manuscript entitled ``An Improved Approximate-Bayesian
-% Model-choice Method for Estimating Shared Evolutionary History'' to be
-% considered for publication in \highLight{journal here}.
-I am submitting a presubmission inquiry regarding a manuscript entitled ``An
-Improved Approximate-Bayesian Model-choice Method for Estimating Shared
-Evolutionary History'' that I wish to be considered for publication in
-\emph{PLOS Computational Biology}.
-The field of research is comparative biogeography and Bayesian model choice.
+I am submitting a manuscript entitled ``An Improved Approximate-Bayesian
+Model-choice Method for Estimating Shared Evolutionary History'' to be
+considered for publication in \emph{BMC Evolutionary Biology}.
+This is a methodological article, and the field of research is comparative
+biogeography and Bayesian model choice.
 

### 2014-03-07 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Updating figure titles for dpp ms.

@@
 \mFigure{../../validation/images/validation-model-choice-psi.pdf}{%
-    \ifbmc{\csentence{Comparison of model-choice behavior.}}{}
+    \ifbmc{\csentence{Comparison of model-choice accuracy.}}{}
     \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeNum = 1$}
@@
 \mFigure{../../validation/images/validation-model-choice-omega.pdf}{
-    \ifbmc{\csentence{Comparison of model-choice behavior.}}{}
+    \ifbmc{\csentence{Comparison of model-choice accuracy using
+    \divTimeDispersion threshold.}}{}
     \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeDispersion < 0.01$}
@@
 \mFigure{../../power-comparison/images/old-power-psi-4.pdf}{
-    \ifbmc{\csentence{Power comparison.}}{}
+    \ifbmc{\csentence{Power to avoid spurious estimation of clustered
+    divergences when divergence times are random.}}{}
     \powerCommentSummary{\powerSeriesOld}
@@
 \mFigure{../../power-comparison/images/old-power-omega-4.pdf}{
-    \ifbmc{\csentence{Power comparison.}}{}
+    \ifbmc{\csentence{Power to avoid spurious estimation of small temporal
+    variance in divergences when divergence times are random.}}{}
     \powerCommentSummary{\powerSeriesOld}
@@
 \mFigure{../../power-comparison/images/old-power-psi-prob-4.pdf}{
-    \ifbmc{\csentence{Power comparison.}}{}
+    \ifbmc{\csentence{Power to avoid spurious support for one divergence event
+    when divergence times are random.}}{}
     \powerSupportCommentSummary{\powerSeriesOld}
@@
 \mFigure{../../power-comparison/images/old-power-omega-prob-4.pdf}{
-    \ifbmc{\csentence{Power comparison.}}{}
+    \ifbmc{\csentence{Power to avoid spurious support for no temporal variance
+    in divergences (i.e., $\divTimeDispersion < 0.01$) when divergence times
+    are random.}}{}
     \powerSupportCommentSummary{\powerSeriesOld}
@@
 \mFigure{../../empirical-analyses/plots/philippines-psi.pdf}{
-    \ifbmc{\csentence{Empirical results for 22 Philippine taxa.}}{}
+    \ifbmc{\csentence{Estimated number of divergence events for 22 taxa from
+    the Philippines.}}{}
     The (A--E) posterior and (F--J) prior probabilities of the number of
@@
 \widthFigure{0.4}{../../empirical-analyses/plots/negros-panay-psi.pdf}{listformat=figList}{
-    \ifbmc{\csentence{Empirical results for 9 Philippine taxa.}}{}
+    \ifbmc{\csentence{Estimated number of divergence events for 9 taxa from the
+    Philippines.}}{}
     The posterior probabilities of the number of divergence events,

### 2014-04-26 slides/bin/plot-power-results.py
More work on evo-wibo slides.
Also, updating several plots for slides.

@@ -210,6 +210,9 @@ def create_plots(info_path,
             pg.row_label_size = row_label_size
             pg.row_label_offset = row_label_offset
             pg.reset_figure()
+            pg.set_shared_x_limits()
+            pg.set_shared_y_limits()
+            pg.reset_figure()
             pg.savefig(os.path.join(output_dir,
                     prefix + '_power_psi_mode.pdf'))
             pg.column_labels = ['' for x in column_labels]

### 2014-05-02 slides/bin/reject-sampling-plot.py
Adding plotting script for slides.
Adding plotting script to generate plots for cartoon example of ABC
rejection algorithm.

@@ -1,0 +1,160 @@
+#! /usr/bin/env python
+
+import os
+import sys
+import random
+import math
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib
+import numpy
+from matplotlib import pyplot as plt
+from matplotlib import cm
+from scipy.stats import gamma
+import numpy as np
+
+def get_rejection_plot_3d(maximum = 1.0,
+        observed = (0.5, 0.5, 0.5),
+        nsamples = 0,
+        show_rejected = True,
+        rng = None):
+    if not rng:
+        rng = random.Random()
+    fig = plt.figure()
+    ax = fig.add_subplot(111, projection = '3d')
+    ax.plot([observed[0]], [observed[1]], [observed[2]],
+            marker = "o",
+            zorder=100
+            )
+    ax.set_xlim((0.0, maximum))
+    ax.set_ylim((0.0, maximum))
+    ax.set_zlim((0.0, maximum))
+    rect = [-0.12, 0, 1, 1.07]
+    fig.tight_layout(pad = 0.25, rect = rect)
+    u = np.linspace(0, 2 * np.pi, 100)
+    v = np.linspace(0, np.pi, 100)
+    
+    x = (0.1 * np.outer(np.cos(u), np.sin(v))) + observed[0]
+    y = (0.1 * np.outer(np.sin(u), np.sin(v))) + observed[1]
+    z = (0.1 * np.outer(np.ones(np.size(u)), np.cos(v))) + observed[2]
+    if show_rejected:
+        ax.plot_surface(x, y, z,  rstride=10, cstride=15, linewidth = 0.1,
+                color='none',
+                edgecolor='0.5',
+                zorder=1000
+                )
+    if nsamples > 0:
+        draws = [[], [], []]
+        for i in range(nsamples):
+            draws[0].append(rng.random())
+            draws[1].append(rng.random())
+            draws[2].append(rng.random())
+        if not show_rejected:
+            s = [[], [], []]
+            for i in range(nsamples):
+                e = math.sqrt(math.pow(draws[0][i] - observed[0], 2) +
+                        math.pow(draws[1][i] - observed[1], 2) +
+                        math.pow(draws[2][i] - observed[2], 2))
+                if e < 0.1:
+                    s[0].append(draws[0][i])
+                    s[1].append(draws[1][i])
+                    s[2].append(draws[2][i])
+            draws = s
+        ax.plot(draws[0], draws[1], draws[2],
[…]

### 2014-05-27 exchangeability/bin/run-reject-ml.sh
Adding reject script for mult-locus exchange.

@@ -1,0 +1,56 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=16G
+#$ -l vf=16G
+#$ -q all.q
+#$ -pe orte 8
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+staging_dir=$(mktemp -d /tmp/output.XXXXXXXXX)
+
+reps=50000
+nprocs=8
+nprior=1000000
+batch_size=25000
+nsums=100000
+npost=1000
+nquantiles=1000
+seed=8174634137
+
+output_dir="../results/multi-locus-sort"
+if [ ! -d "$output_dir" ]
+then
+    mkdir -p $output_dir
+fi
+
+dmc.py --np $nprocs \
+    -r $reps \
+    -o ../configs/prior-multi-locus.cfg \
+    -p ../priors/multi-locus-sort/pymsbayes-results/pymsbayes-output/prior-stats-summaries \
+    -n $nprior \
+    --prior-batch-size $batch_size \
+    --num-posterior-samples $npost \
+    --num-standardizing-samples $nsums \
+    -q $nquantiles \
+    --sort-index 7 \
+    --output-dir $output_dir \
+    --staging-dir $staging_dir \
+    --temp-dir $staging_dir \
+    --seed $seed \
+    --no-global-estimate \
+    --compress
+
+echo "Here are the contents of the local temp directory '${staging_dir}':"
+ls -Fla $staging_dir
+echo 'Removing the local temp directory...'
+rm -r $staging_dir
+echo 'Done!'
+

### 2014-05-28 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on reviews of dpp-msbayes ms.
Incorporating comments of reviewer 1.

@@
 substitution and lineage coalescence.
+\change{
+The goal would be to estimate the posterior probability of models ranging from
+a single divergence-time parameter (i.e., simultaneous divergence of all
+co-distributed taxa) to the fully generalized model in which each taxon
+diverged at a unique time.}
 
@@
 
+\change{
+\msb has been used to address biogeographical questions in variety of empirical
+systems. Some examples include
+(1) whether rising of the Isthmus of Panama caused co-divergence across species
+of echinoids co-distributed across the Pacific and Atlantic sides of the ishmus
+\cite{Hickerson2006}, 
+(2) if an historical seaway across the Baja Peninsula caused co-divergence
+across species of squamates and mammals co-distributed both north and south of
+the putative seaway \cite{Leache2007},
+(3) if species of gall-wasps and their associated parasitoids share divergences
+across putative glacial refugia \cite{Stone2012}, and
+(4) whether repeated coalescence and fragmentation of the oceanic Islands of
+the Philippines caused by Pleistocene sea-level fluctuations caused
+diversification of vertebrate taxa distributed across the islands
+\cite{Oaks2012}.
+Such applications of the method often result in strong posterior support for
+co-divergence among some of the taxa investigated (e.g.,
+\cite{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012}).
+}
+
 For priors on divergence-time and demographic parameters, \msb uses continuous
@@
 
+\change{
 Recently, Oaks et al.\ \cite{Oaks2012,Oaks2014reply} found via simulation that
-\msb will often strongly support biogeographically interesting models with a
-small number of divergence events shared among taxa, even when divergences were
-random over broad timescales.
+\msb will often strongly support models with a small number of divergence
+events shared among taxa, even when divergences were random over broad
+timescales.}
 They suggested this behavior was due to the combination of uniform priors on
@@
 model's parameters.
-Here, I describe a new method that successfully mitigates spurious inference
-of biogeographically interesting scenarios while avoiding negative side effects
-of empirically informed uniform priors.
+\change{
+Here, I describe a new method that successfully mitigates spurious inference of
+co-divergence while avoiding negative side effects of empirically informed
+uniform priors.}
 
@@
 model.
-For the \myTheta{} parameters, the model can be fully generalized to
+\change{
+For the \myTheta{} parameters, by default, the model is fully generalized to
 allow each population pair to have three parameters:
 \ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}.
-Furthermore, any model of \myTheta{} parameters nested within this
+Furthermore, if an investigator prefers to reduce the number of parameters, any
+model of \myTheta{} parameters nested within this
 general model can also be specified, including the most restricted model
 where the ancestral and descendant populations of each pair share
-a single \myTheta{} parameter.
+a single \myTheta{} parameter.}
 
@@
 exchangeable among pairs.
-However, this is incorrect for two reasons.
+\change{
+This is incorrect for two reasons.}
 First, the entire vector of summary statistics \alignmentSS{}{} for each pair
@@
 \alignmentSS{}{}.
+\change{
 However, the expectations for statistics that estimate gross diversity (e.g.,
 $\pi$ and $\theta_W$) are not independent of sample size for structured
-populations, and other statistics are not independent of sample size in general
-(e.g., $SD(\pi-\theta_W)$).
+populations (e.g., the divergent pairs of populations modeled by \msb), and
+other statistics are not independent of sample size in general (e.g.,
+$SD(\pi-\theta_W)$).}
 Second, and more importantly, having the same expectation does not ensure
@@
 improve.
+\change{
 This will determine the timescales over which the models can reliably detect
-random variation in divergence times and avoid spurious inference of
-biogeographically interesting clustered divergence models.
+random variation in divergence times and avoid spurious inference of clustered
+divergence models.}
 
@@
 
+\change{
 Overall, the results suggest that the \modelDPP and \modelUniform models are
@@
 do cause them to be biased, they tend to under-estimate the probability of the
-biogeographically interesting model of a single divergence event.
+model with a single, shared divergence event.}
 In contrast, the \modelOld model is very sensitive to model violations, and
@@
 S\ref{figValidationModelChoiceDPPOrderedDPPOrdered}).
-This suggests that the method is not adversely affected by the large increase
+This suggests that the method is not adversely affected by the increase
 in the number of possible discrete models (from 22 unordered to 4140 ordered
 models) when there are eight pairs of populations.
+\change{
+This is encouraging, because, as discussed above, the sorting of the summary
+statistics calculated from sequence alignments is not valid for empirical
+datasets in which there is variation in the number of individuals sampled per
+population, generation times across taxa, nucleotide substitution parameters,
+or the number, length, mutation rate, or ploidy of the loci sampled.
+Thus, this sorting should not be performed for any empirical application
+of the method.}
 
@@
 \ref{figValidationModelChoiceOmega}).
+\change{
 When more appropriate priors are used for divergence-time and demographic
@@
 cause bias, the method is conservative and tends to underestimate the
-probability of biogeographically interesting models of temporally clustered
-divergences (Figure~\ref{figValidationModelChoicePsi} \&
-\ref{figValidationModelChoiceOmega}).
+probability of models with temporally clustered divergences
+(Figure~\ref{figValidationModelChoicePsi} \&
+\ref{figValidationModelChoiceOmega}).}
 
@@
 
-I also urge caution when using \dppmsbayes due to the fact that ABC methods
-in general are known to be biased for model choice \cite{Robert2011}.
+\change{
+I also urge caution when using \dppmsbayes due to the lack of theoretical
+validation of Bayesian model choice when the full data are replaced by
+insufficient summary statistics, and the recent demonstration that ABC
+estimates of model posterior probabilities can be biased when insufficient
+summary statistics are employed \cite{Rober2011}.}
 As recommended by Oaks et al.\ \cite{Oaks2012}, any results from the method
@@
 existing method, \msb.
+\change{
 Compared to \msb, the new approach better estimates posterior uncertainty,
-which greatly reduces the chances of incorrectly estimating biogeographically
-interesting scenarios of shared divergence events.
+which greatly reduces the chances of incorrectly estimating biogeographical
+scenarios of shared divergence events.
+This is important, because models of shared divergence events are often of
+particular interest to researchers who employ the method.
+}
 This new tool will allow evolutionary biologists to better leverage comparative

### 2014-05-29 manuscripts/dpp-msbayes/reviews/response.tex
Working on reviews of dpp-msbayes ms.
Adding initial draft of the response to reviews. Also, making some minor
revisions to the main text of the ms regarding Reviewer 1's comments.

%&<latex>
\documentclass[12pt]{article}

\usepackage{parskip}
\input{../../utils/preamble.tex}
\usepackage[numbers,sort&compress]{natbib}
\input{../../utils/macros.tex}

\newcounter{commentCounter}
\newcommand{\revcomment}[1]{{\addtocounter{commentCounter}{1}}
    \medskip \hrule \noindent
\textbf{\arabic{section}.\arabic{commentCounter}}: {\sl #1}\par\xspace}
\newcommand{\response}[1]{{\addtolength{\leftskip}{0.25in} #1\par}\xspace}

\let\quoteOld\quote
\let\endquoteOld\endquote
% \renewenvironment{quote}{\sffamily\slshape\small\quoteOld}{\endquoteOld}
\renewenvironment{quote}{\sffamily\small\quoteOld}{\endquoteOld}

% \makeatletter
% \renewcommand{\section}{\@startsection{section}{1}{0mm}%
%     {-8pt}%
%     {4pt}%
% {\sffamily\large\itshape}}
% \makeatother

\begin{document}
% \raggedright
\setlength{\parindent}{0in}

We greatly appreciate the helpful comments of the editors and reviewers. In
this document, we provide responses to these comments, including descriptions
of the changes we have made.

\section{Editor's Comments}
\setcounter{commentCounter}{0}
\revcomment{
    Looking at Reviewer 2's list of issues, they appear to relate only to the
    presentation, and Reviewer 2 is not asking for more analysis.  These
    requests for revision are requests for comments, clarifications, and
    statements of caveats to be made by the author.  I think that if these
    requests are attended to then the ms will be improved.
}
\response{
    ADD!!
}

\revcomment{
    (Concerning Reviewer 2's comment 1.  I did not get the impression that the
    author was criticizing Bayesian foundations, but this Reviewer did, and
    that in itself shows that there is room for misunderstanding.  I suspect
    that the author may feel that the intention of the paragraph is clear
    enough, but perhaps the wording here can be adjusted in the light of this
    reviewer's comment to more clearly say that you are talking about options
    for prior choice, that msBayes has opted for uniform priors, and your
    comments on that.)
}
\response{
    ADD!!
}

\revcomment{
    I would like to ask that in replying to the requests for revision that a
    point-by-point be made to make it easy for us to see what you have done."
}
\response{
    ADD!!
}

\section{Reviewer 1's Comments}
\setcounter{commentCounter}{0}
\revcomment{
    In reference to ``Bayesian model-choice method'' in the second paragraph
    of page 2, Reviewer 1 requested a clearer description of the models being
    sampled by such a method.
}
\response{
    I have added the following sentence to this paragraph that provides a
    specific possibility for the model sample space and, thanks to Reviewer 1's
    suggestion, provides a smoother transition into the description of the \msb
    model in the following paragraph:
    \begin{quote}
        More specifically, the sample space of such a model-choice procedure
        could include all models ranging from a single divergence-time
        parameter (i.e., simultaneous divergence of all co-distributed taxa) to
        the fully generalized model in which each taxon diverged at a unique
        time.
    \end{quote}
}

\revcomment{
    Reviewer 1 felt the equating of ``biogeographically interesting models''
    with models of shared divergences was unnecessarily subjective.
}
\response{
    I agree. I have removed the use of the model descriptor ``biogeographically
    interesting'' from the manuscript, and have added the following to the
    Conclusions:
    \begin{quote}
        Compared to \msb, the new approach better estimates posterior
        uncertainty, which greatly reduces the chances of incorrectly
        estimating biogeographical scenarios of shared divergence events.  This
        is important, because models of shared divergence events are often of
        particular interest to researchers who employ these methods.
    \end{quote}
    Thanks to Reviewer 1's comment, this is a much clearer description of what
    I was intending to convey with ``biogeographically interesting.''
}

\revcomment{
    Many readers may be generally unaware of several important factors of how
    phylogeographers use msBayes. I'd suggest a review paragraph somewhere in
    the intro that describes your empirical paper and others such that the
    readers have a better concept of why these technical details matter.
}
\response{
    Good point. I have added the following paragraph to the introduction to
    familiarize readers with how \msb has been applied, and the general results
    of such applications.
    \begin{quote}
        \msb has been used to address biogeographical questions in variety of
        empirical systems. Some examples include (1) whether rising of the
        Isthmus of Panama caused co-divergence across species of echinoids
        co-distributed across the Pacific and Atlantic sides of the ishmus
        \cite{Hickerson2006}, (2) if an historical seaway across the Baja
        Peninsula caused co-divergence across species of squamates and mammals
        co-distributed both north and south of the putative seaway
        \cite{Leache2007}, (3) if species of gall-wasps and their associated
        parasitoids share divergences across putative glacial refugia
        \cite{Stone2012}, and (4) whether repeated coalescence and
        fragmentation of the oceanic Islands of the Philippines during by
        Pleistocene sea-level fluctuations caused diversification of vertebrate
        taxa distributed across the islands \cite{Oaks2012}.  Such applications
        of the method often result in strong posterior support for
        co-divergence among some of the taxa investigated (e.g.,
        \cite{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012}).
    \end{quote}
}

\revcomment{
    In reference to ``I provide the ability to control the richness of the model'' on Page 9,
    Reviewer 1 commented:

    While I appreciate how this flexibility is a selling point to someone who
    is very familiar with this model (such as the author), it might also be
    viewed as an increased difficulty for the average user. Is this to say that
    the 'richness' of the model has to be specified by the user, or can be
    specified? Please clarify.
}
\response{
    To clarify that the user is not required to control the richness of the
    model, I have changed the subsequent sentences to:
    \begin{quote}
        For the \myTheta{} parameters, by default, the model is fully
        generalized to allow each population pair to have three parameters:
        \ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}.
        Furthermore, if an investigator prefers to reduce the number of
        parameters, any model of \myTheta{} parameters nested within this
        general model can also be specified, including the most restricted
        model where the ancestral and descendant populations of each pair share
        a single \myTheta{} parameter.
    \end{quote}
}

\revcomment{
    Reviewer 1 commented that I awkwardly started two consecutive sentences
    with ``However'' on Page 11.
}
\response{
    Thank you for catching this. I have removed ``However'' from the beginning
    of the first sentence, because it was unnecessary.
}

\revcomment{
    Reviewer 1 commented that I was underselling the results in section
    ``Validation analyses: Ordered divergence models'' on Page 16.
}
\response{
    I have expanded this section by including the following sentences:
    \begin{quote}
        ADD!!
    \end{quote}
}

\revcomment{
    Given these caveats, what (in your opinion) is the best usage of this
    approach to phylogeographic analysis?
}
\response{
    ADD!!!
}

\revcomment{
    Reviewer 1 requested that we elaborate about how Robert et al's
    \cite{Robert2011} critique of ABC model choice relates to \dppmsbayes.
}
\response{
    I have expanded:
    \begin{quote}
        I also urge caution when using dpp-msbayes due to the fact that ABC
        methods in general are known to be biased for model choice
        \cite{Robert2011}.
    \end{quote}
    to:
    \begin{quote}
        I also urge caution when using \dppmsbayes due to the lack of
        theoretical validation of Bayesian model choice when the full data are
        replaced by summary statistics that are insufficient for discriminating
        across models under comparision \cite{Robert2011}, which is certainly
        the case here.  Robert et al.\ \cite{Robert2011} demonstrated that ABC
        estimates of model posterior probabilities can be biased when such
        across-model insufficient statistics are used \cite{Robert2011}.
    \end{quote}
    I think this makes it more clear how the findings of \cite{Robert2011} are
    potentially relevant to the posterior probabilities estimated by
    \dppmsbayes and \msb.
}

\section{Reviewer 2's Comments}
\setcounter{commentCounter}{0}

\revcomment{
    The complaint that ``This causes divergence models with more divergence-time
    parameters to integrate over a much greater parameter space with low
    likelihood yet high prior density, resulting in small marginal likelihoods
    relative to models with fewer divergence-time parameters.'' (p.2) is
    essentially stressing the point with Occam's razor argument. Which I deem a
    rather positive feature of Bayesian model choice. A reflection on the
    determination of the prior distribution away from uniform priors thus
    sounds most timely, rather than a criticism of the Bayesian foundations.
}
\response{
    ADD!!
}

\revcomment{
    The current paper claims to have reached a satisfactory prior modelling
    with ``improved robustness, accuracy, and power'' (p.3). If I understand
    correctly, the changes are in replacing a uniform distribution with a Gamma
    or a Dirichlet prior.  Which means introducing a seriously large number of
    hyperparameters into the picture. Having a lot of flexibility in the prior
    also means a lot of variability in the resulting inference. This is a
    general feature of all Bayesian modellings and I am rather surprised by the
    conclusion that things look better by taking this step away from the
    uniform prior. Depending on the problem they could also look worse.
}
\response{
    ADD!!
}

\revcomment{
    The ABC part is rather standard, except for the rather strange feature that
    the divergence times are used to construct summary statistics (p.10). I
    find it strange because these times are not observed for the actual data.
    So I must be missing something, which should be spelled out more clearly.
    (And I object to the sentence ``I have introduced a new approximate-Bayesian
    model choice method.'' (p.21) since there is no new ABC technique, simply a
    different form of prior.)
}
\response{
    ADD!!
}

\revcomment{
    I do not understand much the part about (reshuffling) introducing bias as
    detailed on p.11: every approximate method gives a ``biased'' answer in the
    sense this answer is not the true and proper posterior distribution. Using
    a different (re-ordered) vector of statistics provides a different ABC
    outcome, hence a different approximate posterior, for which it seems truly
    impossible to check whether or not it increases the discrepancy from the
    true posterior, compared with the other version. I always find it annoying
    to see the word `bias' used in a vague meaning and esp. within a Bayesian
    setting. All Bayesian methods are biased.
}
\response{
    ADD!!
}

\revcomment{
    Quoting [43] as concluding that ABC model choice is biased is
    somewhatmisleading: the warning therein was that Bayes factors and
    posterior probabilities resulting from those ABC steps could be quite
    unrelated with those based on the whole dataset. That the same authors
    concluded that the proper choice of summary statistics leads to a
    consistent model choice shows ABC model choice is not necessarily
    ``biased''.
}
\response{
    ADD!!
}

\revcomment{
    I also fail to understand why the posterior probability of model i should
    be distributed as a uniform (``If the method is unbiased, the points should
    fall near the identity line'') when the data is from model i: this is not a
    p-value but a posterior probability and the posterior probability is not
    the frequentist coverage.
}
\response{
    ADD!!
}

\revcomment{
    My overall problem with the paper is that, all in all, this is a single
    Monte Carlo study about the performances of some classes of priors and, as
    such, it does not carry enough weight to validate an approach that remains
    highly subjective in the selection of its hyperparameters. I think this
    remains a controlled experiment with simulated data where the true
    parameters are know and the prior is ``true''. This obviously helps in
    getting better performances. In a different and less controlled experiment,
    a new experimenter may (should?) be at a loss as to which hyperparameters
    to chose.
}
\response{
    ADD!!
}

\revcomment{
    The conclusion of the paper that ``With improving numerical methods (...),
    advances in Monte Carlo techniques and increasing efficiency of likelihood
    calculations, analyzing rich comparative phylo-geographical models in a
    full-likelihood Bayesian framework is becoming computationally feasible.''
    (p.21) sounds overly optimistic and rather premature. As to this day, I do
    not know of any significant advance in computing the observed likelihood
    for the population genetics models ABC is currently handling. (The SMC
    algorithm of Bouchard-Côté, Sankaraman and Jordan 2012 does not apply to
    Kingman's coalescent, as far as I can judge.) This is certainly a goal
    worth pursuing and borrowing strength from multiple techniques cannot hurt,
    but it remains so far a lofty goal, still beyond our reach... I thus think
    the major message of the paper is to reinforce our earlier calls for
    caution when interpreting the output of an ABC model choice (p.20), or even
    of a regular Bayesian analysis, agreeing that we should aim at seeing ``a
    large amount of posterior uncertainty'' rather than posterior probability
    values close to 0 and 1.
}
\response{
    ADD!!
}

\bibliographystyle{../../bib/bmc-mathphys.bst}
\bibliography{../../bib/references}

\end{document}

### 2014-05-29 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on reviews of dpp-msbayes ms.
Adding initial draft of the response to reviews. Also, making some minor
revisions to the main text of the ms regarding Reviewer 1's comments.

@@
 \change{
-The goal would be to estimate the posterior probability of models ranging from
-a single divergence-time parameter (i.e., simultaneous divergence of all
-co-distributed taxa) to the fully generalized model in which each taxon
-diverged at a unique time.}
+More specifically, the sample space of such a model-choice procedure could
+include all models ranging from a single divergence-time parameter (i.e.,
+simultaneous divergence of all co-distributed taxa) to the fully generalized
+model in which each taxon diverged at a unique time.}
 
@@
 (4) whether repeated coalescence and fragmentation of the oceanic Islands of
-the Philippines caused by Pleistocene sea-level fluctuations caused
+the Philippines during by Pleistocene sea-level fluctuations caused
 diversification of vertebrate taxa distributed across the islands
@@
 In addition to differences in sample size, the simulated alignments and their
-summary statistics are not identically distributed because of the taxon- and
+summary statistics are not \change{exchangeable} because of the taxon- and
 locus-specific parameters \hkyModel{}{}, \mutationRateScalarConstant{}{}, and
@@
 I also urge caution when using \dppmsbayes due to the lack of theoretical
-validation of Bayesian model choice when the full data are replaced by
-insufficient summary statistics, and the recent demonstration that ABC
-estimates of model posterior probabilities can be biased when insufficient
-summary statistics are employed \cite{Rober2011}.}
+validation of Bayesian model choice when the full data are replaced by summary
+statistics that are insufficient for discriminating across models under
+comparision \cite{Robert2011}, which is certainly the case here.
+Robert et al.\ \cite{Robert2011} demonstrated that ABC estimates of model
+posterior probabilities can be biased when such across-model insufficient
+statistics are used \cite{Robert2011}.
 As recommended by Oaks et al.\ \cite{Oaks2012}, any results from the method
 should be accompanied by assessments of prior sensitivity and power.
-I encourage investigators to view this method as a means of exploring their
-data rather than a rigorous means of evaluating hypotheses.
+Until full-likelihood methods are available to assess the accuracy of ABC
+estimates of divergence-model posterior probabilities, I encourage
+investigators to view this method as a means of exploring their data rather
+than a rigorous means of evaluating hypotheses.}
 
@@
 This is important, because models of shared divergence events are often of
-particular interest to researchers who employ the method.
+particular interest to researchers who employ these methods.
 }

### 2014-05-29 manuscripts/dpp-msbayes/reviews/response.tex
Working on revisions and review response for dpp ms.

@@
         replaced by summary statistics that are insufficient for discriminating
-        across models under comparision \cite{Robert2011}, which is certainly
+        across models under comparison \cite{Robert2011}, which is certainly
         the case here.  Robert et al.\ \cite{Robert2011} demonstrated that ABC
@@
     Quoting [43] as concluding that ABC model choice is biased is
-    somewhatmisleading: the warning therein was that Bayes factors and
+    somewhat misleading: the warning therein was that Bayes factors and
     posterior probabilities resulting from those ABC steps could be quite
@@
 \response{
-    ADD!!
+    This is a very good point, thank you. I have expanded this section
+    to be clearer about the findings of \cite{Robert2011} and how they
+    relate to the methods discussed in the paper:
+    \begin{quote}
+        I also urge caution when using \dppmsbayes due to the lack of
+        theoretical validation of Bayesian model choice when the full data are
+        replaced by summary statistics that are insufficient for discriminating
+        across models under comparison \cite{Robert2011}, which is certainly
+        the case here.  Robert et al.\ \cite{Robert2011} demonstrated that ABC
+        estimates of model posterior probabilities can be biased when such
+        across-model insufficient statistics are used \cite{Robert2011}.
+    \end{quote}
 }
@@
 \response{
-    ADD!!
+    I agree that the justification of why the identity line should be considered
+    ``unbiased'' was too vague and relied entirely on the reference to
+    \cite{Huelsenbeck2004}. I have expanded this section out to be very
+    explicit what is meant by ``bias'' and why it is desirable for the method
+    to produce estimates that fall along the identity line:
+    \begin{quote}
+        Ideally, the estimated posterior probability of the one-divergence model
+        should estimate the probability that the one-divergence model is
+        correct given the data.  For large numbers of simulation replicates,
+        the proportion of the replicates in each bin for which the
+        one-divergence model is true will approximate the probability that the
+        one-divergence model is correct.  Thus, if the method has the desirable
+        behavior such that the estimated posterior probability of
+        the one-divergence model is an unbiased estimate of the probability
+        that the one-divergence model is correct given the data, the points
+        should fall near the identity line.  For example, let us say the method
+        estimates a posterior probability of 0.90 for 1000 simulated datasets.
+        If the method is accurately estimating the probability that the
+        one-divergence model is correct given the data, then the one-divergence
+        model should be the true model in approximately 900 of the 1000
+        replicates.  Any trend away from the identity line indicates the method
+        is biased in the sense that it is not correctly estimating the
+        probability that the one-divergence model is correct given the data.
+    \end{quote}
 }
@@
 \response{
-    ADD!!
+    I agree. However, the goal of the paper is not to demonstrate, for example,
+    that gamma probability distributions are generally superior to uniform
+    probability distributions for Bayesian statistics. Nor is the goal to
+    solve the subjective nature of prior selection in Bayesian statistics.
+    My primary goal was to develop an improved comparative method that
+    biogeographers can use to investigate the affect of historical events on
+    diversification.
+
+    I used extensive simulations and biological data to compare, as fairly as
+    possible, the new approach to the current state-of-the-art method for this
+    inference problem. I agree the new approach has many weaknesses, including
+    the subjective nature of choosing priors when the truth is unknown. However,
+    these weaknesses apply to both \dppmsbayes and \msbayes (and to most
+    empirical applications of Bayesian statistical methods where extensive
+    prior information is lacking). However, the results strongly support
+    the overall conclusion that the primary goal was accomplished, and the
+    new approach out performs \msb under a range of conditions.
 }
@@
 \response{
+    It's already been done \cite{JeetDiss}.
+    I can cite skyline/skyride plots, species trees, snapp, IM, etc. (all
+    full-likelihood coalescent methods).
+    Bryant et al.: analytical integration of gene trees for biallelic markers
+    under the Kingman coalescent.
     ADD!!

### 2014-05-29 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on revisions and review response for dpp ms.

@@
 event \cite{Huelsenbeck2004,Oaks2012}.
-If the method is unbiased, the points should fall near the identity line,
-whereas any trend away from the identity line indicates bias.
-I did this using two criteria for the one-divergence model:
+\change{
+Ideally, the estimated posterior probability of the one-divergence model
+should estimate the probability that the one-divergence model is correct
+given the data.
+For large numbers of simulation replicates, the proportion of the replicates in
+each bin for which the one-divergence model is true will approximate the
+probability that the one-divergence model is correct.
+Thus, if the method has the desirable behavior such that the
+estimated posterior probability of the one-divergence model is an unbiased
+estimate of the probability that the one-divergence model is correct given the
+data, the points should fall near the identity line.
+For example, let us say the method estimates a posterior probability of 0.90
+for 1000 simulated datasets.
+If the method is accurately estimating the probability that the one-divergence
+model is correct given the data, then the one-divergence model should be the
+true model in approximately 900 of the 1000 replicates.
+Any trend away from the identity line indicates the method is biased in the
+sense that it is not correctly estimating the probability that the
+one-divergence model is correct given the data.
+
+I constructed these plots using two criteria for the one-divergence model:
+}
 (1) the number of divergence-time parameters ($\divTimeNum = 1$) and
@@
 statistics that are insufficient for discriminating across models under
-comparision \cite{Robert2011}, which is certainly the case here.
+comparison \cite{Robert2011}, which is certainly the case here.
 Robert et al.\ \cite{Robert2011} demonstrated that ABC estimates of model

### 2014-05-30 manuscripts/dpp-msbayes/reviews/response.tex
Work on revisions and review response for dpp ms.

@@
 \response{
-    ADD!!
+    Sorry for the confusion here. This statement was not intened as
+    a general criticism of Bayesian foundations. Rather, the intent was to
+    highlight why prior choice is very important in Bayesian model choice.
+    I have added the following sentences to clarify:
+    \begin{quote}
+        This serves as a ``natural'' penalty for over-parameterization, which
+        is a strength of Bayesian model choice.  However, given the sensitivity
+        of marginal likelihoods to the prior, care is needed when selecting
+        prior distributions \cite{Lindley1957}.  Selecting distributions that
+        are broad or very differently shaped than the likelihood density (e.g.,
+        a uniform prior for divergence times) can lead to small marginal
+        likelihoods of parameter-rich models even if they are correct.
+    \end{quote}
 }
@@
 \response{
-    ADD!!
+    Both uniform and gamma distributions have two parameters. Thus, the new
+    model only has one additional hyperparameter compared to the original
+    msBayes model: the concentration parameter on the Dirichlet process prior
+    on divergence models. Also, as described in the manuscript, the new method
+    offers a uniform prior over divergence models, which would have the same
+    number of hyper-parameters as the original method.
+
+    Also, in both msBayes and dpp-msbayes the hyper-parameters are fixed, and
+    thus do not contribute variability to the model. While a hiearchical
+    approach that places hyper-priors on the various hyper-parameters would be
+    interesting to explore for either the msBayes or dpp-msbayes model, it is
+    beyond the scope of the current study.
+
+    I have added some text to various places in the manuscript to clarify that
+    the hyper-parameters are fixed:
+    \begin{quote}
+        If the recombination rate \recombinationRate is allowed to be non-zero,
+        the prior term $p(\recombinationRate)$ would be added to Equation
+        \ref{eq:fullModelCompact}, and the prior would be $\recombinationRate
+        \sim Gamma(\gammaShape{\recombinationRate},
+        \gammaScale{\recombinationRate})$, \change{where
+        \gammaShape{\recombinationRate} and \gammaScale{\recombinationRate} are
+        specified by the investigator.}
+    \end{quote}
+
+    \begin{quote}
+        The priors for the demographic parameters are:
+        $\ancestralTheta{} \sim Gamma(\gammaShape{\ancestralTheta{}},
+        \gammaScale{\ancestralTheta{}})$,
+        $\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+        \gammaScale{\descendantTheta{}{}})$,
+        $\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+        \gammaScale{\descendantTheta{}{}})$,
+        $\bottleScalar{1}{} \sim Beta(\betaA{\bottleScalar{}{}},
+        \betaB{\bottleScalar{}{}})$,
+        $\bottleScalar{2}{} \sim Beta(\betaA{\bottleScalar{}{}},
+        \betaB{\bottleScalar{}{}})$,
+        $\bottleTime{} \sim U(0, 1)$,
+        and
+        $\migrationRate{} \sim Gamma(\gammaShape{\migrationRate{}},
+        \gammaScale{\migrationRate{}})$.
+        \change{
+        By default, \ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}
+        share the same prior (i.e., $\gammaShape{\ancestralTheta{}} =
+        \gammaShape{\descendantTheta{}{}}$ and $\gammaScale{\ancestralTheta{}} =
+        \gammaScale{\descendantTheta{}{}}$), but a separate gamma-distributed prior can
+        be assigned to \ancestralTheta{}.
+        Also, the \bottleScalar{1}{}, \bottleScalar{2}{}, and \migrationRate{}
+        parameters are optional (i.e., it can be assumed that there is no migration
+        between populations of each pair, and that the population size of each
+        descendant population has been constant through time).
+        For the parameters included in the model, the hyper-parameters of the prior
+        distributions are specified by the investigator.}
+    \end{quote}
+
+    \begin{quote}
+        Each \divTime{} within \divTimeVector is \iid as $\divTime{} \sim
+        Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$, \change{
+        where \gammaShape{\divTime{}} and \gammaScale{\divTime{}} are specified
+        by the investigator}.
+    \end{quote}
 }
@@
 \response{
+    Regarding the reshuffling of the summary statistics calculated from
+    different alignments of sequences, those statistics are simply not
+    exchangeable. So, reshuffling them in a manner that is not conistent across
+    all simulations and the observed data (as is done in msBayes) is not
+    mathematically valid. Also, if elements are exchangeable, their order will
+    not affect the likelihood (or the posterior, barring sampling error). Thus,
+    if the goal is to approximate the likelihood, I would hope the reshuffling
+    would also have little affect on the approximate posterior. The fact that
+    \cite{Huang2011} demonstrate that it does strongly affect the posterior
+    suggests this reshuffling is not only mathematically incorrect, but also is
+    misleading the results.
+
     ADD!!
@@
     ``unbiased'' was too vague and relied entirely on the reference to
-    \cite{Huelsenbeck2004}. I have expanded this section out to be very
+    \cite{Huelsenbeck2004}. I have expanded this section to be very
     explicit what is meant by ``bias'' and why it is desirable for the method
@@
 \response{
-    It's already been done \cite{JeetDiss}.
-    I can cite skyline/skyride plots, species trees, snapp, IM, etc. (all
-    full-likelihood coalescent methods).
-    Bryant et al.: analytical integration of gene trees for biallelic markers
-    under the Kingman coalescent.
-    ADD!!
+    Full-likelihood methods in this domain are not as lofty of a goal as you
+    might think. The likelihood function of the msBayes model is tractable,
+    and, in fact, has already been derived and implemented via reversible-jump
+    MCMC (see \cite{JeetDiss}; albeit, not readily available yet).
+    Also, there are plenty of examples of rich, Kingman-coalescent models
+    implemented in full-likelihood Bayesian frameworks.
+    A small sample includes \cite{Drummond2005,Drummond2008,Drummond2002,Kuhner1995,Hey2007,Hey2010,Bryant2012,Shapiro2011,Vaughan2014,Heled2010,Lemey2009,Lemey2010,Pybus2012,Baele2012,Minin2008beast,Bloomquist2010}.
 }

### 2014-05-30 exchangeability/bin/parse-results-ml.py
Adding multi-locus exchange results.

@@ -1,0 +1,26 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes.utils.parsing import DMCSimulationResults
+from pymsbayes.utils.messaging import get_logger
+
+_LOG = get_logger(__name__)
+
+def main_cli():
+    bin_dir = os.path.abspath(os.path.dirname(__file__))
+    project_dir = os.path.abspath(os.path.dirname(bin_dir))
+    result_dir = os.path.abspath(os.path.join(project_dir, 'results'))
+    info_path = os.path.join(result_dir, 'multi-locus-sort', 'pymsbayes-results',
+            'pymsbayes-info.txt')
+    _LOG.info('Parsing and writing results...')
+    results = DMCSimulationResults(info_path)
+    prior_indices = results.prior_index_to_config.keys()
+    results.write_result_summaries(
+            prior_indices = prior_indices,
+            include_tau_exclusion_info = False)
+
+if __name__ == '__main__':
+    main_cli()
+

### 2014-05-30 exchangeability/bin/parse-results-ml.sh

@@ -1,0 +1,17 @@
+#! /bin/sh
+#$ -S /bin/bash
+#$ -cwd
+#$ -V
+#$ -l h_vmem=2G
+#$ -l vf=2G
+#$ -q all.q
+
+if [ -n "$SGE_O_WORKDIR" ]
+then
+    source ~/.bash_profile
+    cd /share/work1
+    cd $SGE_O_WORKDIR
+fi
+
+which python
+python parse-results-ml.py

### 2014-05-30 manuscripts/dpp-msbayes/dpp-msbayes-main.tex

@@
 uniform probability distributions.
+\change{
 This causes divergence models with more divergence-time parameters to integrate
 over a \emph{much} greater parameter space with low likelihood yet high prior
-density, resulting in small marginal likelihoods relative to models with fewer
-divergence-time parameters \cite{Jeffreys1939,Lindley1957}.
+density, which can result in small marginal likelihoods relative to models with
+fewer divergence-time parameters \cite{Jeffreys1939,Lindley1957}.
 Given that the marginal likelihood of a model weighted by its prior is what
@@
 fewer divergence events \cite{Oaks2012,Oaks2014reply}.
+This serves as a ``natural'' penalty for over-parameterization, which is a
+strength of Bayesian model choice.
+However, given the sensitivity of marginal likelihoods to the prior, care is
+needed when selecting prior distributions \cite{Lindley1957}.
+Selecting distributions that are broad or very differently shaped than the
+likelihood density (e.g., a uniform prior for divergence times) can lead to
+small marginal likelihoods of parameter-rich models even if they are correct.
+}
+
 Furthermore, \msb uses a discrete uniform prior over the number of divergence
@@
     \alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$.
-For each population pair, the means of the summary statistics are calculated
-across the \nloci{} loci, reducing the vector to
-$\ssVectorObs = (\alignmentSSObs{1}{}, \ldots,
+\change{
+Optionally, for each population pair, the means of the summary statistics can
+be calculated across the \nloci{} loci, and the vector can be further reduced
+to $\ssVectorObs = (\alignmentSSObs{1}{}, \ldots,
 \alignmentSSObs{\npairs{}}{})$.
-This allows us to estimate the approximate joint posterior distribution
+Either way, with \ssVectorObs in hand, we can estimate the approximate joint
+posterior distribution}
 {\small
@@
 1/\locusRateHetShapeParameter)$.
-If the recombination rate \recombinationRate is allowed to be
-non-zero, the prior term $p(\recombinationRate)$ would be added
-to Equation \ref{eq:fullModelCompact}, and the prior would be
-$\recombinationRate \sim Gamma(\gammaShape{\recombinationRate},
-\gammaScale{\recombinationRate})$.
+If the recombination rate \recombinationRate is allowed to be non-zero, the
+prior term $p(\recombinationRate)$ would be added to Equation
+\ref{eq:fullModelCompact}, and the prior would be $\recombinationRate \sim
+Gamma(\gammaShape{\recombinationRate}, \gammaScale{\recombinationRate})$,
+\change{where \gammaShape{\recombinationRate} and
+\gammaScale{\recombinationRate} are specified by the investigator.}
 \end{linenomath}
@@
 \gammaScale{\migrationRate{}})$.
+\change{
+By default, \ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}
+share the same prior (i.e., $\gammaShape{\ancestralTheta{}} =
+\gammaShape{\descendantTheta{}{}}$ and $\gammaScale{\ancestralTheta{}} =
+\gammaScale{\descendantTheta{}{}}$), but a separate gamma-distributed prior can
+be assigned to \ancestralTheta{}.
+Also, the \bottleScalar{1}{}, \bottleScalar{2}{}, and \migrationRate{}
+parameters are optional (i.e., it can be assumed that there is no migration
+between populations of each pair, and that the population size of each
+descendant population has been constant through time).
+For the parameters included in the model, the hyper-parameters of the prior
+distributions are specified by the investigator.}
 \end{linenomath}
@@
 Each \divTime{} within \divTimeVector is \iid as $\divTime{} \sim
-Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$.
+Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$,
+\change{
+where \gammaShape{\divTime{}} and \gammaScale{\divTime{}} are specified by the
+investigator}.
 Thus, given the number of unique divergence-time classes in
@@
 Each sample is generated by
+\change{
 (1) drawing values of all the model's parameters, which I denote \hpvector{},
 from their respective prior distributions;
-(2) rescaling the divergence times
-$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
-from units proportional to time to units proportional to the expected number
-of mutations via Equation \ref{eq:divTimeScaled} to get 
-$\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
-\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$;
-(3) simulating gene trees $\geneTreeVector = (\geneTree{1}{1}, \ldots, 
+% (2) rescaling the divergence times
+% $\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
+% from units proportional to time to units proportional to the expected number
+% of mutations via Equation \ref{eq:divTimeScaled} to get 
+% $\divTimeScaledVector = (\divTimeScaled{1}{1}, \ldots,
+% \divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$;
+(2) simulating gene trees $\geneTreeVector = (\geneTree{1}{1}, \ldots, 
 \geneTree{\npairs{}}{\nloci{\npairs{}}})$
@@
 a multi-population Kingman-coalescent model given the demographic parameters;
-(4) simulating sequence alignments 
+(3) simulating sequence alignments 
 $\alignmentVector = (\alignment{1}{1}, \ldots, \alignment{\npairs{}}{\nloci{\npairs{}}})$
@@
 dataset;
-(5) calculating population genetic summary statistics
+and
+(4) calculating population genetic summary statistics
 $\ssVector{} = (\alignmentSS{1}{1}, \ldots, \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$
 from the simulated sequence alignments;
-% (these are the same statistics
-% calculated from the observed alignments, \ssVectorObs{});
-and (6) reducing the summary statistics to the means across loci for each
-population pair to get
-$\ssVector{} = (\alignmentSS{1}{}, \ldots, \alignmentSS{\npairs{}}{})$, which is
-the same summary statistic vector estimated from the observed data \ssVectorObs.
+Optionally, an additional step can be performed to reduce the summary
+statistics to the means across loci for each population pair to get
+$\ssVector{} = (\alignmentSS{1}{}, \ldots, \alignmentSS{\npairs{}}{})$.
+Either way, \ssVector{} contains the same summary statistics as those estimated
+from the observed data \ssVectorObs.
+}
 After repeating this procedure \numPriorSamples times, I obtain a random
@@
 ($\pi_b$) \cite{NeiLi1979,Huang2011}.
+\change{
+Thus, the summary statistics being compared when calculating the Euclidean
+distance between the observed data and each simulation are often from sequence
+alignments of \emph{different} taxon pairs and loci.
+In other words, the compared statistics are often calulated from datasets
+simulated with \emph{different} numbers and length of sequences, and
+potentially different numbers of loci, and different fixed \hky model
+parameters (\hkyModel{}{}) and mutation-rate (\mutationRateScalarConstant{}{})
+and ploidy (\ploidyScalar{}{}) scaling parameters.}
+
 This has the advantage of reducing the sample space of the number of

### 2014-05-30 exchangeability/bin/plot-results-ml.py
Adding plotting script to exchange experiment.

@@ -1,0 +1,22 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from pymsbayes import plotting
+
+def main_cli():
+    bin_dir = os.path.abspath(os.path.dirname(__file__))
+    project_dir = os.path.abspath(os.path.dirname(bin_dir))
+    result_dir = os.path.abspath(os.path.join(project_dir, 'results'))
+    info_path = os.path.join(result_dir, 'multi-locus-sort', 'pymsbayes-results',
+            'pymsbayes-info.txt')
+    plotting.plot_validation_results(info_path,
+            omega_symbol = r'D_T',
+            psi_symbol = r'|\mathbf{\tau}|',
+            mean_time_symbol = r'\bar{T}',
+            math_font = None)
+
+if __name__ == '__main__':
+    main_cli()
+

### 2014-05-31 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on exchangeability section of dpp ms.

@@
 \change{
-Thus, the summary statistics being compared when calculating the Euclidean
+As a result, the summary statistics being compared when calculating the Euclidean
 distance between the observed data and each simulation are often from sequence
-alignments of \emph{different} taxon pairs and loci.
-In other words, the compared statistics are often calulated from datasets
-simulated with \emph{different} numbers and length of sequences, and
-potentially different numbers of loci, and different fixed \hky model
-parameters (\hkyModel{}{}) and mutation-rate (\mutationRateScalarConstant{}{})
-and ploidy (\ploidyScalar{}{}) scaling parameters.}
+alignments of \emph{different} taxon pairs and/or loci.
+More specifically, the summary statistics calculated from the observed sequence
+alignments are being compared to summary statistics calulated from datasets
+simulated with potentially \emph{different} 
+(1) numbers of sequences,
+(2) length of sequences,
+(3) numbers of loci,
+(4) \hky model parameters (\hkyModel{}{}),
+(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
+and
+(6) ploidy multipliers (\ploidyScalar{}{}).}
 
@@
 By discarding the identity of the \npairs{} pairs of populations, all
-pair-specific information about the amount of data (e.g., the number gene
+pair-specific information about the amount of data (e.g., the number of gene
 copies collected from each of the populations [\popSampleSize{1}{} and
@@
 \locusMutationRateScalar{}) is lost.
-Second, the results are more difficult to interpret, because they can be no
+Second, the results are more difficult to interpret, because they can no
 longer be directly associated to the taxa under study.
@@
 
-Thus, the theoretical basis for this reshuffling is questionable.
-It may be justifiable as an additional coarsening of the data,
-by removing the information associated with the identity of the
-population pairs.
-However, I do not provide a formal proof that this re-ordering of the pairs
-does not introduce bias.
-One can imagine situations in which the sampling intensity (i.e., the
-number of gene copies, loci, and locus length) is highly skewed across
-pairs and/or there is large heterogeneity in mutational processes
-(e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
-In such cases, it seems possible that discarding this information could
-cause bias.
-Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
-is to minimize Euclidean distances between simulated datasets in which the true
-model has a single divergence event \cite{Huang2011}, and (2) the reordering
-increases the model's tendency to infer a single divergence event
-\cite{Huang2011}, it seems possible that this approach could be biasing the
-method \cite{Oaks2012}.
+\change{
+Thus, whenever the six parameters listed above vary across any of the taxa or
+loci (i.e., nearly all empirical applications), the taxon-specific summary
+statistics \alignmentSS{}{} are not exchangeable.
+Hence, the reshuffling of the summary statistic vectors in not mathematically
+valid.
+The magnitude of the affect of this violaton of exchangeability is currently
+unknown.
+Huang et al.\ \cite{Huang2011} demonstrated that the reordering of the summary
+statistic vectors can greatly increase the method's tendency to infer a single
+divergence event.
+By definition, if the summary statistic vectors were exhangeable, the
+reordering would not change the likelihood or posterior (barring sampling
+error).
+Thus, the results of Huang et al.\ \cite{Huang2011} suggest the
+reordering of the statistics could be introducing a bias.
+% Thus, the theoretical basis for this reshuffling is questionable.
+% It may be justifiable as an additional coarsening of the data,
+% by removing the information associated with the identity of the
+% population pairs.
+% However, I do not provide a formal proof that this re-ordering of the pairs
+% does not introduce bias.
+% One can imagine situations in which the sampling intensity (i.e., the
+% number of gene copies, loci, and locus length) is highly skewed across
+% pairs and/or there is large heterogeneity in mutational processes
+% (e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
+% In such cases, it seems possible that discarding this information could
+% cause bias.
+% Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
+% is to minimize Euclidean distances between simulated datasets in which the true
+% model has a single divergence event \cite{Huang2011}, and (2) the reordering
+% increases the model's tendency to infer a single divergence event
+% \cite{Huang2011}, it seems possible that this approach could be biasing the
+% method \cite{Oaks2012}.
+}
 
@@
 maintained, and ordered divergence models to be estimated.
+\change{
+Also, in all of the simulation-based analyses described below, the
+summary statistic vectors \emph{are} exchangeable, because all of the
+simulated datasets have the same 
+(1) numbers of sequences,
+(2) length of sequences,
+(3) numbers of loci,
+(4) \hky model parameters (\hkyModel{}{}),
+(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
+and
+(6) ploidy multipliers (\ploidyScalar{}{}).}
 

### 2014-05-31 manuscripts/dpp-msbayes/dpp-msbayes-abstract.tex
More editing of main text and abstract of dpp ms.

@@
 thus greatly reduces the tendency of the model to incorrectly estimate
-biogeographically interesting models with strong support.
+\change{models of shared evolutionary history} with strong support.
+% biogeographically interesting models with strong support.

### 2014-05-31 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More editing of main text and abstract of dpp ms.

@@
 \change{
-More specifically, the sample space of such a model-choice procedure could
-include all models ranging from a single divergence-time parameter (i.e.,
-simultaneous divergence of all co-distributed taxa) to the fully generalized
-model in which each taxon diverged at a unique time.}
+The sample space of such a model-choice procedure would include all models
+ranging from a single divergence-time parameter (i.e., simultaneous divergence
+of all co-distributed taxa) to the fully generalized model in which each taxon
+diverged at a unique time.}
 
@@
 systems. Some examples include
-(1) whether rising of the Isthmus of Panama caused co-divergence across species
-of echinoids co-distributed across the Pacific and Atlantic sides of the ishmus
-\cite{Hickerson2006}, 
+(1) whether the rise of the Isthmus of Panama caused co-divergence among
+species of echinoids co-distributed across the Pacific and Atlantic sides of
+the isthmus \cite{Hickerson2006}, 
 (2) if an historical seaway across the Baja Peninsula caused co-divergence
@@
 (4) whether repeated coalescence and fragmentation of the oceanic Islands of
-the Philippines during by Pleistocene sea-level fluctuations caused
+the Philippines during Pleistocene sea-level fluctuations caused
 diversification of vertebrate taxa distributed across the islands
@@
 fewer divergence events \cite{Oaks2012,Oaks2014reply}.
-This serves as a ``natural'' penalty for over-parameterization, which is a
-strength of Bayesian model choice.
+% The use of marginal likelihoods to estimat model probabilities provides a
+% ``natural'' penalty for overparameterization, and can be a great strength of
+% the Bayesian approach.
+This is not a critique of Bayesian model choice in general; comparing models by
+their marginal likelihoods provides a ``natural'' penalty for
+overparameterization and can be a great strength of the Bayesian approach.
 However, given the sensitivity of marginal likelihoods to the prior, care is
 needed when selecting prior distributions \cite{Lindley1957}.
-Selecting distributions that are broad or very differently shaped than the
-likelihood density (e.g., a uniform prior for divergence times) can lead to
-small marginal likelihoods of parameter-rich models even if they are correct.
+Selecting distributions that will often place high prior density in large
+regions of parameter space with low likelihood can lead to small marginal
+likelihoods of parameter-rich models even if they are correct.
 }
@@
 accuracy, and power compared to the original model.
-Also, my results confirm that the improved performance of the new model
+The results confirm that the improved performance of the new model
 is due to a combination of
@@
 (\href{https://github.com/joaks1/PyMsBayes}{\url{https://github.com/joaks1/PyMsBayes}}),
-which is a Python API that provides a multi-processing interface to \msb and
+which provides a multi-processing interface to \msb and
 \dppmsbayes.
@@
 \alignmentSSObs{\npairs{}}{})$.
-Either way, with \ssVectorObs in hand, we can estimate the approximate joint
+With \ssVectorObs in hand, we can estimate the approximate joint
 posterior distribution}
@@
 
-\subsubsection{Likelihood terms of Equation \ref{eq:fullModelCompact}}
+% \subsubsection{Likelihood terms of Equation \ref{eq:fullModelCompact}}
+\change{
+\subsubsection{Likelihood and gene-tree prior terms of Equation \ref{eq:fullModelCompact}}
+}
 \begin{linenomath}
-The likelihood terms of Equation \ref{eq:fullModelCompact} can
-be expanded out as a product over population pairs and loci
+% The likelihood terms of Equation \ref{eq:fullModelCompact} can
+\change{
+The likelihood and gene-tree prior terms of Equation \ref{eq:fullModelCompact}
+}
+can be expanded out as a product over population pairs and loci
 \begin{equation}
@@
 
-\subsubsection{Prior terms of Equation \ref{eq:fullModelCompact}}
+% \subsubsection{Prior terms of Equation \ref{eq:fullModelCompact}}
+\change{
+\subsubsection{Additional prior terms of Equation \ref{eq:fullModelCompact}}
+}
 \begin{linenomath}
@@
 \end{equation}
-The priors for the demographic parameters are:
+\change{
+The priors for the demographic parameters are
 $\ancestralTheta{} \sim Gamma(\gammaShape{\ancestralTheta{}},
@@
 $\migrationRate{} \sim Gamma(\gammaShape{\migrationRate{}},
-\gammaScale{\migrationRate{}})$.
-\change{
+\gammaScale{\migrationRate{}})$,
+where the hyper-parameters of each prior distribution can be specified by
+the investigator.
 By default, \ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}
@@
 Also, the \bottleScalar{1}{}, \bottleScalar{2}{}, and \migrationRate{}
-parameters are optional (i.e., it can be assumed that there is no migration
-between populations of each pair, and that the population size of each
-descendant population has been constant through time).
-For the parameters included in the model, the hyper-parameters of the prior
-distributions are specified by the investigator.}
+parameters are optional (i.e., the investigator can assume that there has been
+no migration between populations of each pair and/or the population size of
+each descendant population has been constant through time).}
 \end{linenomath}
@@
 population sizes.
-This is only equal to the expectation of the prior if the lower limit is zero.
+This is only equal to the expectation of the prior if the lower limit of the
+prior is zero.
 
@@
 $\ssVector{} = (\alignmentSS{1}{1}, \ldots, \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$
-from the simulated sequence alignments;
+from the simulated sequence alignments.
 Optionally, an additional step can be performed to reduce the summary
@@
 }
-After repeating this procedure \numPriorSamples times, I obtain a random
+After repeating this procedure \numPriorSamples times, we have a random
 sample of parameter vectors
@@
 
-For all of the analyses below, I use four summary statistics for each of
-the pairs of populations:
+For all of the analyses below, I use four summary statistics for each pair of
+populations:
 $\pi$ \cite{Tajima1983}, $\theta_W$ \cite{Watterson1975}, $\pi_{net}$
@@
 
+\subsubsection{Obtaining an approximate posterior from the prior samples}
+I use a rejection algorithm to retain an approximate posterior sample of
+\paramSampleVector{} from the prior sample
+$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$.
+First, the observed summary statistics \ssVectorObs, and the summary statistics
+of the prior samples 
+$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$,
+are standardized using the means and standard deviations of the statistics from
+the prior sample (i.e., the prior mean is subtracted from each statistic, and the
+difference is divided by the prior standard deviation).
+After all statistics are standardized, the Euclidean distance between
+\ssVectorObs and each \ssVector{} within \ssMatrix is calculated.
+The samples that fall within a range of tolerance \tol around \ssVectorObs
+are retained.
+The range of tolerance is determined by specifying the desired number of
+posterior samples to be retained.
+Post-hoc adjustment of the posterior sample can also be performed with a number
+of regression techniques \cite{Beaumont2002,Blum2009,Leuenberger2010}.
+For analyses below, I use the general linear model (GLM) regression adjustment
+\cite{Leuenberger2010} as implemented in \abctoolbox v1.1
+\cite{ABCtoolbox}, which Oaks et al.\ \cite{Oaks2012} showed performs
+very similarly to weighted local-linear regression and multinomial logistic
+regression adjustments \cite{Beaumont2002} for \msb posteriors.
+
 \subsubsection{Ordering of taxon-specific summary statistics}
@@
 \change{
-As a result, the summary statistics being compared when calculating the Euclidean
-distance between the observed data and each simulation are often from sequence
-alignments of \emph{different} taxon pairs and/or loci.
+As a result, the summary statistics being compared when calculating the
+Euclidean distance between the observed data and each simulated dataset often
+represent sequence alignments of \emph{different} taxon pairs and/or loci.
 More specifically, the summary statistics calculated from the observed sequence
-alignments are being compared to summary statistics calulated from datasets
+alignments are being compared to summary statistics calculated from datasets
 simulated with potentially \emph{different} 
-(1) numbers of sequences,
-(2) length of sequences,
-(3) numbers of loci,
+(1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
+(2) length of alignments,
+(3) numbers of loci (\nloci{}),
 (4) \hky model parameters (\hkyModel{}{}),
@@
 distributed across different sample sizes.
-In addition to differences in sample size, the simulated alignments and their
-summary statistics are not \change{exchangeable} because of the taxon- and
-locus-specific parameters \hkyModel{}{}, \mutationRateScalarConstant{}{}, and
-\ploidyScalar{}{}.
-
-\change{
-Thus, whenever the six parameters listed above vary across any of the taxa or
-loci (i.e., nearly all empirical applications), the taxon-specific summary
-statistics \alignmentSS{}{} are not exchangeable.
-Hence, the reshuffling of the summary statistic vectors in not mathematically
+\change{
+More generally, the simulated alignments and their summary statistics are not
+identically distributed if there is any variation among taxa or loci in the
+(1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
+(2) length of alignments,
+(3) numbers of loci (\nloci{}),
+(4) \hky model parameters (\hkyModel{}{}),
+(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
+or
+(6) ploidy multipliers (\ploidyScalar{}{}).}
+\change{
+Whenever such variation is present (i.e., nearly all empirical applications),
+the taxon-specific summary statistics \alignmentSS{}{} are not exchangeable,
+and the reshuffling of the summary statistic vectors in not mathematically
 valid.
-The magnitude of the affect of this violaton of exchangeability is currently
-unknown.
+
+The magnitude of the affect of this violation of exchangeability is not known.
 Huang et al.\ \cite{Huang2011} demonstrated that the reordering of the summary
@@
 divergence event.
-By definition, if the summary statistic vectors were exhangeable, the
+By definition, if the summary statistic vectors were exchangeable, the
 reordering would not change the likelihood or posterior (barring sampling
 error).
-Thus, the results of Huang et al.\ \cite{Huang2011} suggest the
-reordering of the statistics could be introducing a bias.
+Thus, the results of Huang et al.\ \cite{Huang2011} suggest the reordering of
+the statistics can introduce bias.
 % Thus, the theoretical basis for this reshuffling is questionable.
@@
 
-For comparability with \msb, I maintain the re-ordering of taxon-specific
-summary statistics by $\pi_b$.  However, I also allow the order to be
-maintained, and ordered divergence models to be estimated.
-\change{
-Also, in all of the simulation-based analyses described below, the
-summary statistic vectors \emph{are} exchangeable, because all of the
-simulated datasets have the same 
+\change{
+For comparability with \msb, I maintain the option for re-ordering
+taxon-specific summary statistics by $\pi_b$.
+However, by default, the order is preserved, and ordered divergence models are
+estimated.
+In all of the simulation-based analyses described below, the summary statistic
+vectors \emph{are} exchangeable, because the simulated datasets have the same 
 (1) numbers of sequences,
@@
 (3) numbers of loci,
-(4) \hky model parameters (\hkyModel{}{}),
-(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
+(4) \hky model parameters,
+(5) mutation-rate multipliers,
 and
-(6) ploidy multipliers (\ploidyScalar{}{}).}
-
-\subsubsection{Obtaining an approximate posterior from the prior samples}
-I use a rejection algorithm to retain an approximate posterior sample of
-\paramSampleVector{} from the prior sample
-$\paramSampleMatrix = (\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples})$.
-First, the observed summary statistics \ssVectorObs, and the summary statistics
-of the prior samples 
-$\ssMatrix = (\ssVector{1}, \ldots, \ssVector{\numPriorSamples})$,
-are standardized using the means and standard deviations of the statistics from
-the prior sample (i.e., the prior mean is subtracted from each statistic, and the
-difference is divided by the prior standard deviation).
-After all statistics are standardized, the Euclidean distance between
-\ssVectorObs and each \ssVector{} within \ssMatrix is calculated.
-The samples that fall within a range of tolerance \tol around \ssVectorObs
-are retained.
-The range of tolerance is determined by specifying the desired number of
-posterior samples to be retained.
-Post-hoc adjustment of the posterior sample can also be performed with a number
-of regression techniques \cite{Beaumont2002,Blum2009,Leuenberger2010}.
-For my analyses, I use the general linear model (GLM) regression adjustment
-\cite{Leuenberger2010} as implemented in \abctoolbox
-\cite[v1.1;]{ABCtoolbox}, which Oaks et al.\ \cite{Oaks2012} showed performs
-very similarly to weighted local-linear regression and multinomial logistic
-regression adjustments \cite{Beaumont2002} for \msb posteriors.
+(6) ploidy multipliers.}
 
@@
 Ideally, the estimated posterior probability of the one-divergence model
-should estimate the probability that the one-divergence model is correct
-given the data.
+should estimate the probability that the one-divergence model is correct.
 For large numbers of simulation replicates, the proportion of the replicates in
 each bin for which the one-divergence model is true will approximate the
-probability that the one-divergence model is correct.
-Thus, if the method has the desirable behavior such that the
-estimated posterior probability of the one-divergence model is an unbiased
-estimate of the probability that the one-divergence model is correct given the
-data, the points should fall near the identity line.
+probability that the one-divergence model is the correct model.
+Thus, if the method has the desirable behavior such that the estimated
+posterior probability of the one-divergence model is an unbiased estimate of
+the probability that the one-divergence model is correct, the points should
+fall near the identity line.
 For example, let us say the method estimates a posterior probability of 0.90
@@
 Any trend away from the identity line indicates the method is biased in the
-sense that it is not correctly estimating the probability that the
-one-divergence model is correct given the data.
+sense that it is not accurately estimating the probability that the
+one-divergence model is the correct model.
 

### 2014-06-01 manuscripts/dpp-msbayes/reviews/response.tex
Working on dpp ms revisions.

--- 2014-05-29 manuscripts/dpp-msbayes/reviews/response.tex
@@
 
-We greatly appreciate the helpful comments of the editors and reviewers. In
-this document, we provide responses to these comments, including descriptions
-of the changes we have made.
+I greatly appreciate the helpful comments of the editor and reviewers. In this
+document, I provide responses to these comments, including descriptions of the
+changes I have made.
 
@@
 \response{
-    ADD!!
+    Reviewer 2's comments were very helpful, and have helped to improve the
+    clarity of the manuscript. Below, I address each of Reviewer 2's concerns
+    along with the corresponding changes I have made in light of the comments.
 }
@@
 \response{
-    ADD!!
+    I agree that there was room for misunderstanding given that the context of
+    the statements were not made clear. Please see the below for the specific
+    edits I have made in response to this helpful comment.
 }
@@
     I would like to ask that in replying to the requests for revision that a
-    point-by-point be made to make it easy for us to see what you have done."
+    point-by-point be made to make it easy for us to see what you have done.
 }
 \response{
-    ADD!!
+    Please see below for my point-by-point revisions. Also, I am attaching a
+    PDF of the manuscript with all of the revisions highlighted in color.
 }
@@
     I have added the following sentence to this paragraph that provides a
-    specific possibility for the model sample space and, thanks to Reviewer 1's
+    clearer description of the model sample space and, thanks to Reviewer 1's
     suggestion, provides a smoother transition into the description of the \msb
@@
     \begin{quote}
-        More specifically, the sample space of such a model-choice procedure
-        could include all models ranging from a single divergence-time
-        parameter (i.e., simultaneous divergence of all co-distributed taxa) to
-        the fully generalized model in which each taxon diverged at a unique
-        time.
+        The sample space of such a model-choice procedure would include all
+        models ranging from a single divergence-time parameter (i.e.,
+        simultaneous divergence of all co-distributed taxa) to the fully
+        generalized model in which each taxon diverged at a unique time.
     \end{quote}
@@
     \begin{quote}
-        \msb has been used to address biogeographical questions in variety of
-        empirical systems. Some examples include (1) whether rising of the
-        Isthmus of Panama caused co-divergence across species of echinoids
-        co-distributed across the Pacific and Atlantic sides of the ishmus
+        \msb has been used to address biogeographical questions in a variety of
+        empirical systems. Some examples include (1) whether the rise of the
+        Isthmus of Panama caused co-divergence among species of echinoids
+        co-distributed across the Pacific and Atlantic sides of the isthmus
         \cite{Hickerson2006}, (2) if an historical seaway across the Baja
@@
         parasitoids share divergences across putative glacial refugia
-        \cite{Stone2012}, and (4) whether repeated coalescence and
-        fragmentation of the oceanic Islands of the Philippines during by
-        Pleistocene sea-level fluctuations caused diversification of vertebrate
-        taxa distributed across the islands \cite{Oaks2012}.  Such applications
-        of the method often result in strong posterior support for
-        co-divergence among some of the taxa investigated (e.g.,
+        \cite{Stone2012}, and (4) whether repeated fragmentation of the oceanic
+        Islands of the Philippines during Pleistocene sea-level fluctuations
+        caused diversification of vertebrate taxa distributed across the
+        islands \cite{Oaks2012}.  Such applications of the method often result
+        in strong posterior support for co-divergence among all or subsets of
+        the taxa investigated (e.g.,
         \cite{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012}).
@@
     \begin{quote}
-        ADD!!
+        This is encouraging, because, as discussed above, estimating unordered
+        models of divergence by shuffling the summary statistic vectors
+        calculated from the sequence alignments is not valid for most empirical
+        datasets.  Given these results, estimation of unordered divergence
+        models should be avoided for empirical applications of the method.
     \end{quote}
@@
 \response{
-    ADD!!!
+    This was an important observation by Reviewer 1 that I was listing a lot of
+    potential problems with the new method without providing any advice or
+    guidance on how to navigate such potential pitfalls and how to best use the
+    method. I have added the following paragraph following the discussion of
+    caveats:
+    \begin{quote}
+        Given all of these caveats, I encourage investigators to view this
+        method as a means of exploring their data for general temporal patterns
+        of divergences across taxa, rather than a rigorous means of evaluating
+        hypotheses.  As recommended by Oaks et al.\ \cite{Oaks2012}, any
+        results from the method should be accompanied by (1) analyses under a
+        variety of priors to assess the assumptions underlying model inference
+        and the prior sensitivity of the results, and (2) simulation-based
+        power analyses to provide insight into the temporal resolution of the
+        method.  Both approaches are important to help guide the interpretation
+        of results.
+    \end{quote}
 }
@@
         replaced by summary statistics that are insufficient for discriminating
-        across models under comparision \cite{Robert2011}, which is certainly
+        across models under comparison \cite{Robert2011}, which is certainly
         the case here.  Robert et al.\ \cite{Robert2011} demonstrated that ABC
-        estimates of model posterior probabilities can be biased when such
-        across-model insufficient statistics are used \cite{Robert2011}.
+        estimates of model posterior probabilities can be inaccurate when such
+        across-model insufficient statistics are used.
     \end{quote}
@@
 \response{
-    ADD!!
+    Sorry for the confusion here. This statement was not intended as
+    a general criticism of Bayesian foundations. Rather, the intent was to
+    highlight why prior choice is very important in Bayesian model choice.
+    I have added the following sentences to clarify:
+    \begin{quote}
+        This is not a critique of Bayesian model choice in general; comparing
+        models by their marginal likelihoods provides a ``natural'' penalty for
+        over-parameterization and can be a great strength of the Bayesian
+        approach.  However, given the sensitivity of marginal likelihoods to
+        the prior, care is needed when selecting prior distributions
+        \cite{Lindley1957}.  Selecting distributions that will often place high
+        prior density in large regions of parameter space with low likelihood
+        can lead to small marginal likelihoods of parameter-rich models even if
+        they are correct.
+    \end{quote}
 }
@@
 \response{
-    ADD!!
+    Both uniform and gamma distributions have two parameters. Thus, the new
+    model only has one additional hyperparameter compared to the original
+    msBayes model: the concentration parameter on the Dirichlet process prior
+    on divergence models. Also, as described in the manuscript, the new method
+    offers a uniform prior over divergence models, which would have the same
+    number of hyper-parameters as the original method.
+
+    Also, in both msBayes and dpp-msbayes the hyper-parameters are fixed, and
+    thus do not contribute variability to the model. While a hiearchical
+    approach that places hyper-priors on the various hyper-parameters would be
+    interesting to explore for either the msBayes or dpp-msbayes model, it is
+    beyond the scope of the current study.
+
+    I have added some text to various places in the manuscript to clarify that
+    the hyper-parameters are fixed rather than estimated:
+    \begin{quote}
+        If the recombination rate \recombinationRate is allowed to be non-zero,
+        the prior term $p(\recombinationRate)$ would be added to Equation
+        2, and the prior would be $\recombinationRate
+        \sim Gamma(\gammaShape{\recombinationRate},
+        \gammaScale{\recombinationRate})$, \change{where
+        \gammaShape{\recombinationRate} and \gammaScale{\recombinationRate} are
+        specified by the investigator.}
+    \end{quote}
+
+    \begin{quote}
+        The priors for the demographic parameters are $\ancestralTheta{} \sim
+        Gamma(\gammaShape{\ancestralTheta{}}, \gammaScale{\ancestralTheta{}})$,
+        $\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
+        \gammaScale{\descendantTheta{}{}})$, $\descendantTheta{2}{} \sim
+        Gamma(\gammaShape{\descendantTheta{}{}},
+        \gammaScale{\descendantTheta{}{}})$, $\bottleScalar{1}{} \sim
+        Beta(\betaA{\bottleScalar{}{}}, \betaB{\bottleScalar{}{}})$,
+        $\bottleScalar{2}{} \sim Beta(\betaA{\bottleScalar{}{}},
+        \betaB{\bottleScalar{}{}})$, $\bottleTime{} \sim U(0, 1)$, and
+        $\migrationRate{} \sim Gamma(\gammaShape{\migrationRate{}},
+        \gammaScale{\migrationRate{}})$, \change{where the hyper-parameters of
+        each prior distribution can be specified by the investigator.  By
+        default, \ancestralTheta{}, \descendantTheta{1}{}, and
+        \descendantTheta{2}{} share the same prior (i.e.,
+        $\gammaShape{\ancestralTheta{}} = \gammaShape{\descendantTheta{}{}}$
+        and $\gammaScale{\ancestralTheta{}} =
+        \gammaScale{\descendantTheta{}{}}$), but a separate gamma-distributed
+        prior can be assigned to \ancestralTheta{}.  Also, the
+        \bottleScalar{1}{}, \bottleScalar{2}{}, and \migrationRate{} parameters
+        are optional (i.e., the investigator can assume that there has been no
+        migration between populations of each pair and/or the population size
+        of each descendant population has been constant through time).}
+    \end{quote}
+
+    \begin{quote}
+        Each \divTime{} within \divTimeVector is \iid as $\divTime{} \sim
+        Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$, \change{
+        where \gammaShape{\divTime{}} and \gammaScale{\divTime{}} are specified
+        by the investigator}.
+    \end{quote}
 }
@@
 \response{
-    ADD!!
+    Reviewer 2 is correct that there is no new ABC numerical machinery
+    presented in the manuscript, only a new model. To make this clearer,
+    I have changed the sentence to:
+    \begin{quote}
+        I introduced a new model for estimating shared divergence histories
+        across taxa from DNA sequence data within an approximate-Bayesian
+        model-choice framework.
+    \end{quote}
+
+    Also, the intuition of Reviewer 2 is correct, I do not use the divergence
+    times to calculate summary statistics. I mention the divergence times in
+    the description of the ABC algorithm with the hope of making it clear that
+    the times are scaled (see Equation (12)) prior to the simulation of the
+    data (from which the summary statistics are calculated). This scaling is
+    simply to go from units proportional to time, to units that are
+    proportional to the expected number of mutations. Because, this is already
+    described earlier in the description of the model, I have removed the
+    re-description of this scaling from the description of the ABC algorithm.
+
+    This section now reads as:
+
+    \begin{quote}
+        Each sample is generated by \change{(1) drawing values of all the
+            model's parameters, which I denote \hpvector{}, from their
+            respective prior distributions; (2) simulating gene trees
+            $\geneTreeVector = (\geneTree{1}{1}, \ldots,
+            \geneTree{\npairs{}}{\nloci{\npairs{}}})$ for each locus of each
+            population pair by drawing coalescent times from a multi-population
+            Kingman-coalescent model given the demographic parameters; (3)
+            simulating sequence alignments $\alignmentVector =
+            (\alignment{1}{1}, \ldots,
+            \alignment{\npairs{}}{\nloci{\npairs{}}})$ along the gene trees
+            under the \hky substitution parameters $\hkyModelVector =
+            (\hkyModel{1}{1}, \ldots, \hkyModel{\npairs{}}{\nloci{\npairs{}}})$
+            that have the same number of sequences and sequence lengths as the
+            observed dataset; and (4) calculating population genetic summary
+            statistics $\ssVector{} = (\alignmentSS{1}{1}, \ldots,
+            \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$ from the simulated
+        sequence alignments.}
+    \end{quote}
+
+    Many thanks to Reviewer 2 for catching this potential source of confusion.
 }
@@
 \response{
-    ADD!!
+    Regarding the reshuffling of the summary statistics calculated from
+    different alignments of sequences, those statistics are not exchangeable.
+    So, reshuffling them in a manner that is not consistent across all
+    simulations and the observed data (as is done in msBayes) is not
+    mathematically valid. Also, if elements are exchangeable, their order will
+    not affect the likelihood (or the posterior). Thus, if the goal is to
+    approximate the likelihood, the reshuffling should also have little affect
+    on the approximate posterior. The fact that \cite{Huang2011} demonstrate
+    that it does strongly affect the posterior suggests this reshuffling is not
+    only mathematically incorrect, but also is misleading the results.
+
+    I have revised this section to make it more precise and clear:
+    \begin{quote}
+        \change{
+        The re-ordering of the summary statistic vectors also has an important
+        implication for the ABC algorithm.  When calculating the Euclidean
+        distance between the observed data and each simulated dataset, the
+        summary statistics being compared often represent sequence alignments
+        of \emph{different} taxon pairs and/or loci.  More specifically, the
+        summary statistics calculated from the observed sequence alignments are
+        being compared to summary statistics calculated from datasets simulated
+        with potentially \emph{different} (1) numbers of sequences
+        (\popSampleSize{1}{} and \popSampleSize{2}{}), (2) length of
+        alignments, (3) numbers of loci (\nloci{}), (4) \hky model parameters
+        (\hkyModel{}{}), (5) mutation-rate multipliers
+        (\mutationRateScalarConstant{}{}), and (6) ploidy multipliers
+        (\ploidyScalar{}{}).}
+
+        In the original descriptions of the \msb method
+        \cite{Hickerson2006,Huang2011}, this re-ordering is justified by the
+        fact that the expected value of $\pi_b$ is unrelated to sample size
+        \popSampleSize{1}{} and \popSampleSize{2}{} and thus exchangeable among
+        pairs.  \change{ This is incorrect for two reasons.} First, the entire
+        vector of summary statistics \alignmentSS{}{} for each pair of
+        populations is re-ordered across pairs, which implies that the
+        justification for re-ordering $\pi_b$ applies to all the statistics
+        within each \alignmentSS{}{}.  \change{However, the expectations for
+        statistics that estimate gross diversity (e.g., $\pi$ and $\theta_W$)
+        are not independent of sample size for structured populations (e.g.,
+        the divergent pairs of populations modeled by \msb), and other
+        statistics are not independent of sample size in general (e.g.,
+        $SD(\pi-\theta_W)$).  Second, and more importantly, having the same
+        expectation does not ensure random variables are exchangeable.  Rather,
+        for variables to be exchangeable their marginal distributions must be
+        the same (i.e., they must be identically distributed).  None of the
+        summary statistics used by \msb, including $\pi_b$, are identically
+        distributed when there is any variation among taxa or loci in the (1)
+        numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}), (2)
+        length of alignments, (3) numbers of loci (\nloci{}), (4) \hky model
+        parameters (\hkyModel{}{}), (5) mutation-rate multipliers
+        (\mutationRateScalarConstant{}{}), or (6) ploidy multipliers
+        (\ploidyScalar{}{}).  Whenever such variation is present (i.e., nearly
+        all empirical applications), the taxon-specific summary statistics
+        \alignmentSS{}{} are not exchangeable, and the reshuffling of the
+        summary statistic vectors is not mathematically valid.
+
+        The magnitude of the affect of this violation of exchangeability is not
+        known.  Huang et al.\ \cite{Huang2011} demonstrated that the reordering
+        of the summary statistic vectors can greatly increase the method's
+        tendency to infer a single divergence event.  By definition, if the
+        summary statistic vectors were exchangeable, the reordering would not
+        change the likelihood or posterior (barring sampling error).  Thus, the
+        results of Huang et al.\ \cite{Huang2011} suggest the reordering of the
+        statistics is potentially introducing sizeable error to the analysis.}
+    \end{quote}
 }
@@
 \response{
-    ADD!!
+    This is a very good point, thank you. I have expanded this section to be
+    clearer about the findings of \cite{Robert2011} and how they relate to the
+    methods discussed in the paper:
+    \begin{quote}
+        I also urge caution when using \dppmsbayes due to the lack of
+        theoretical validation of Bayesian model choice when the full data are
+        replaced by summary statistics that are insufficient for discriminating
+        across models under comparison \cite{Robert2011}, which is certainly
+        the case here.  Robert et al.\ \cite{Robert2011} demonstrated that ABC
+        estimates of model posterior probabilities can be inaccurate when such
+        across-model insufficient statistics are used.
+    \end{quote}
 }
@@
 \response{
-    ADD!!
+    I agree that the justification of why the identity line should be
+    considered ``unbiased'' was too vague and relied entirely on the reference
+    to \cite{Huelsenbeck2004}. I have expanded this section to be very explicit
+    what is meant by ``bias'' and why it is desirable for the method to produce
+    estimates that fall along the identity line:
+    \begin{quote}
+        Ideally, the estimated posterior probability of the one-divergence
+        model should estimate the probability that the one-divergence model is
+        correct.  For large numbers of simulation replicates, the proportion of
+        the replicates in each bin for which the one-divergence model is true
+        will approximate the probability that the one-divergence model is the
+        correct model.  Thus, if the method has the desirable behavior such
+        that the estimated posterior probability of the one-divergence model is
+        an unbiased estimate of the probability that the one-divergence model
+        is correct, the points should fall near the identity line.  For
+        example, let us say the method estimates a posterior probability of
+        0.90 for 1000 datasets simulated from the prior.  If the method is
+        accurately estimating the probability that the one-divergence model is
+        correct given the data, then the one-divergence model should be the
+        true model in approximately 900 of the 1000 replicates.  Any trend away
+        from the identity line indicates the method is biased in the sense that
+        it is not accurately estimating the probability that the one-divergence
+        model is the correct model.
+    \end{quote}
 }
@@
 \response{
-    ADD!!
+    I agree. However, the goal of the paper is not to demonstrate, for example,
+    that gamma probability distributions are generally superior to uniform
+    probability distributions for Bayesian model-choice. Nor is the goal to
+    solve the subjective nature of prior selection in Bayesian statistics.
+    My primary goal was to develop an improved comparative method that
+    biogeographers can use to investigate the affect of historical events on
+    diversification.
+
+    I used extensive simulations and biological data to compare, as fairly as
+    possible, the new approach to the current state-of-the-art method for this
+    inference problem. I agree the new approach has many weaknesses, including
+    the subjective nature of choosing priors when the truth is unknown.
+    However, these weaknesses apply to \emph{both} \dppmsbayes and \msb
+    (and to all empirical applications of Bayesian statistical methods where
+    extensive prior information is lacking). However, the results strongly
+    support the overall conclusion that the primary goal was accomplished, and
+    the new approach outperforms \msb under a range of conditions.
 }
@@
     for the population genetics models ABC is currently handling. (The SMC
-    algorithm of Bouchard-Côté, Sankaraman and Jordan 2012 does not apply to
+    algorithm of Bouchard-C\^{o}t\'{e}, Sankaraman and Jordan 2012 does not apply to
     Kingman's coalescent, as far as I can judge.) This is certainly a goal
@@
 \response{
-    ADD!!
+    Full-likelihood methods in this domain are not as lofty of a goal as you
+    might think. The likelihood function of the msBayes model is tractable,
+    and, in fact, has already been derived and implemented via reversible-jump
+    MCMC (see \cite{JeetDiss}; albeit, not readily available yet).
+    Also, there are plenty of examples of rich, Kingman-coalescent models
+    implemented in a full-likelihood Bayesian framework.
+    A small sample includes \cite{Drummond2005,Drummond2008,Drummond2002,Kuhner1995,Hey2007,Hey2010,Bryant2012,Shapiro2011,Vaughan2014,Heled2010,Lemey2009,Lemey2010,Pybus2012,Baele2012,Minin2008beast,Bloomquist2010}.
+    Also, the SMC algorithm of \cite{Jordan2012} can be used to estimate
+    phylogenetic trees using an arbitrary prior on trees (e.g., Birth-death
+    process, coalescent, etc.).
 }

### 2014-06-01 manuscripts/dpp-msbayes/dpp-msbayes-acknowledgments.tex
Working on dpp ms revisions.

@@
+\change{
 I am greatful to Melissa Callahan, Mark Holder, Emily McTavish, Daniel Money,
-Jordan Koch, and Adam Leach\'{e} for insightful comments that greatly improved
-this work.
+Jordan Koch, Adam Leach\'{e}, Peter Foster, two anonymous reviewers, and
+Christian Robert and his blog
+(\href{http://xianblog.wordpress.com/}{\url{http://xianblog.wordpress.com/}})
+for insightful comments that greatly improved this work.
+}
 I thank the National Science Foundation for supporting this work (DEB

### 2014-06-01 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on dpp ms revisions.

@@
 \change{
-\msb has been used to address biogeographical questions in variety of empirical
-systems. Some examples include
+\msb has been used to address biogeographical questions in a variety of
+empirical systems. Some examples include
 (1) whether the rise of the Isthmus of Panama caused co-divergence among
@@
 across putative glacial refugia \cite{Stone2012}, and
-(4) whether repeated coalescence and fragmentation of the oceanic Islands of
-the Philippines during Pleistocene sea-level fluctuations caused
-diversification of vertebrate taxa distributed across the islands
-\cite{Oaks2012}.
+(4) whether repeated fragmentation of the oceanic Islands of the Philippines
+during Pleistocene sea-level fluctuations caused diversification of vertebrate
+taxa distributed across the islands \cite{Oaks2012}.
 Such applications of the method often result in strong posterior support for
-co-divergence among some of the taxa investigated (e.g.,
+co-divergence among all or subsets of the taxa investigated (e.g.,
 \cite{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012}).
@@
 their marginal likelihoods provides a ``natural'' penalty for
-overparameterization and can be a great strength of the Bayesian approach.
+over-parameterization and can be a great strength of the Bayesian approach.
 However, given the sensitivity of marginal likelihoods to the prior, care is
@@
 events $1,2,\ldots,\npairs{}$.
-Because there are many more possible assignments of populations pairs to
+Because there are many more possible assignments of population pairs to
 intermediate numbers of divergence events, this imposes a prior on divergence
@@
 U-shaped prior on divergence models.
-Hickerson et al.\ \cite{Hickerson2013} suggested the problem is caused by
+Hickerson et al.\ \cite{Hickerson2013} suggested the problem was caused by
 sampling error, and proposed as a solution an approximate-Bayesian model
@@
 \end{equation}
-\change{
 The priors for the demographic parameters are
@@
 \gammaScale{\migrationRate{}})$,
+\change{
 where the hyper-parameters of each prior distribution can be specified by
@@
 ($\pi_b$) \cite{NeiLi1979,Huang2011}.
-\change{
-As a result, the summary statistics being compared when calculating the
-Euclidean distance between the observed data and each simulated dataset often
-represent sequence alignments of \emph{different} taxon pairs and/or loci.
+This has the advantage of reducing the sample space of possible divergence
+models \divTimeIndexVector, but there are at least two disadvantages.
+First, additional information in the data is lost.
+By discarding the identity of the \npairs{} pairs of populations, all
+pair-specific information about the amount of data (e.g., the number of gene
+copies collected from each of the populations [\popSampleSize{1}{} and
+\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
+taxon- and locus-specific parameters (\hkyModel{}{},
+\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
+\locusMutationRateScalar{}) is lost.
+\change{
+Second, the results are more difficult to interpret, because divergence models
+and parameter estimates cannot be directly associated to the taxa under study.
+}
+
+\change{
+The re-ordering of the summary statistic vectors also has an important
+implication for the ABC algorithm.
+When calculating the Euclidean distance between the observed data and each
+simulated dataset, the summary statistics being compared often represent
+sequence alignments of \emph{different} taxon pairs and/or loci.
 More specifically, the summary statistics calculated from the observed sequence
@@
 
-This has the advantage of reducing the sample space of the number of
-divergence models \divTimeIndexVector, but there are at least two disadvantages.
-First, additional information in the data is lost.
-By discarding the identity of the \npairs{} pairs of populations, all
-pair-specific information about the amount of data (e.g., the number of gene
-copies collected from each of the populations [\popSampleSize{1}{} and
-\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
-taxon- and locus-specific parameters (\hkyModel{}{},
-\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
-\locusMutationRateScalar{}) is lost.
-Second, the results are more difficult to interpret, because they can no
-longer be directly associated to the taxa under study.
-
-In the original descriptions of the \msb model \cite{Hickerson2006,Huang2011},
+In the original descriptions of the \msb method \cite{Hickerson2006,Huang2011},
 this re-ordering is justified by the fact that the expected value of $\pi_b$ is
@@
 other statistics are not independent of sample size in general (e.g.,
-$SD(\pi-\theta_W)$).}
+$SD(\pi-\theta_W)$).
 Second, and more importantly, having the same expectation does not ensure
@@
 None of the summary statistics used by \msb, including $\pi_b$, are identically
-distributed across different sample sizes.
-\change{
-More generally, the simulated alignments and their summary statistics are not
-identically distributed if there is any variation among taxa or loci in the
+distributed when there is any variation among taxa or loci in the
 (1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
@@
 or
-(6) ploidy multipliers (\ploidyScalar{}{}).}
-\change{
+(6) ploidy multipliers (\ploidyScalar{}{}).
 Whenever such variation is present (i.e., nearly all empirical applications),
 the taxon-specific summary statistics \alignmentSS{}{} are not exchangeable,
-and the reshuffling of the summary statistic vectors in not mathematically
+and the reshuffling of the summary statistic vectors is not mathematically
 valid.
@@
 Thus, the results of Huang et al.\ \cite{Huang2011} suggest the reordering of
-the statistics can introduce bias.
+the statistics is potentially introducing sizeable error to the analysis.
 % Thus, the theoretical basis for this reshuffling is questionable.
@@
 For example, let us say the method estimates a posterior probability of 0.90
-for 1000 simulated datasets.
+for 1000 datasets simulated from the prior.
 If the method is accurately estimating the probability that the one-divergence
@@
 \subsection{Validation analyses: Ordered divergence models}
-The results show that the method performs similarily when sampling over ordered
+The results show that the method performs similarly when sampling over ordered
 models of divergence
@@
 \change{
-This is encouraging, because, as discussed above, the sorting of the summary
-statistics calculated from sequence alignments is not valid for empirical
-datasets in which there is variation in the number of individuals sampled per
-population, generation times across taxa, nucleotide substitution parameters,
-or the number, length, mutation rate, or ploidy of the loci sampled.
-Thus, this sorting should not be performed for any empirical application
-of the method.}
+This is encouraging, because, as discussed above, estimating unordered models
+of divergence by shuffling the summary statistic vectors calculated from the
+sequence alignments is not valid for most empirical datasets.
+Given these results, estimation of unordered divergence models should be
+avoided for empirical applications of the method.}
 
@@
 parameters, and either a Dirichlet-process or uniform prior applied across
-divergence models, the model is less sensitive to violations, and, when they do
-cause bias, the method is conservative and tends to underestimate the
-probability of models with temporally clustered divergences
+divergence models, the model is less sensitive to violations, and, when
+violations do cause bias, the method is conservative and tends to underestimate
+the probability of models with temporally clustered divergences
 (Figure~\ref{figValidationModelChoicePsi} \&
@@
 Robert et al.\ \cite{Robert2011} demonstrated that ABC estimates of model
-posterior probabilities can be biased when such across-model insufficient
-statistics are used \cite{Robert2011}.
+posterior probabilities can be inaccurate when such across-model insufficient
+statistics are used.
+
+Given all of these caveats, I encourage investigators to view this method as a
+means of exploring their data for general temporal patterns of divergences
+across taxa, rather than a rigorous means of evaluating hypotheses.
 As recommended by Oaks et al.\ \cite{Oaks2012}, any results from the method
-should be accompanied by assessments of prior sensitivity and power.
-Until full-likelihood methods are available to assess the accuracy of ABC
-estimates of divergence-model posterior probabilities, I encourage
-investigators to view this method as a means of exploring their data rather
-than a rigorous means of evaluating hypotheses.}
+should be accompanied by
+(1) analyses under a variety of priors to assess the assumptions underlying
+model inference and the prior sensitivity of the results, and
+(2) simulation-based power analyses to provide insight into the temporal
+resolution of the method.
+Both approaches are important to help guide the interpretation of results.
+}
 
@@
 \section{Conclusions}
-I introduced a new approximate-Bayesian model-choice method for estimating
-shared divergence histories across taxa using DNA sequence data.
+\change{
+I introduced a new model for estimating shared divergence histories across taxa
+from DNA sequence data within an approximate-Bayesian model-choice framework.
+}
 The new method, \dppmsbayes, takes a non-parametric approach to the problem by

### 2014-06-02 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
dpp ms ready for resubmission.

@@
 \change{
-Here, I describe a new method that successfully mitigates spurious inference of
-co-divergence while avoiding negative side effects of empirically informed
+Here, I describe a new approach that successfully mitigates spurious inference
+of co-divergence while avoiding negative side effects of empirically informed
 uniform priors.}
@@
 the same (i.e., they must be identically distributed).
-None of the summary statistics used by \msb, including $\pi_b$, are identically
-distributed when there is any variation among taxa or loci in the
+\emph{None} of the summary statistics used by \msb, including $\pi_b$, have
+this property when there is any variation among taxa or loci in the
 (1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
@@
 divergence models, the model is less sensitive to violations, and, when
-violations do cause bias, the method is conservative and tends to underestimate
+violations do cause bias, the method tends to underestimate
 the probability of models with temporally clustered divergences
 (Figure~\ref{figValidationModelChoicePsi} \&
-\ref{figValidationModelChoiceOmega}).}
+\ref{figValidationModelChoiceOmega}).
+Given that clustered models are often of particular interest to biogeographers,
+this behavior of the new method can be considered conservative.
+}
 

### 2014-06-02 manuscripts/dpp-msbayes/reviews/response.tex
Another tweak to the dpp ms results section.
*Now*, it should be ready for re-submission.

@@
 
-\section{Editor's Comments}
+\section{Editor's Comments}\label{ed}
 \setcounter{commentCounter}{0}
@@
 \response{
-    Reviewer 2's comments were very helpful, and have helped to improve the
-    clarity of the manuscript. Below, I address each of Reviewer 2's concerns
-    along with the corresponding changes I have made in light of the comments.
+    Reviewer 2's comments were very helpful, and have improved the clarity of
+    the manuscript. Below, in Section \ref{rev2}, I address each of Reviewer
+    2's concerns along with the corresponding changes I have made in light of
+    the comments.
 }
@@
     I agree that there was room for misunderstanding given that the context of
-    the statements were not made clear. Please see the below for the specific
-    edits I have made in response to this helpful comment.
+    the statements were not made clear. Please see Section \ref{rev2} below for
+    the specific edits I have made in response to this helpful comment.
 }
@@
 
-\section{Reviewer 1's Comments}
+\section{Reviewer 1's Comments}\label{rev1}
 \setcounter{commentCounter}{0}
@@
 
-\section{Reviewer 2's Comments}
+\section{Reviewer 2's Comments}\label{rev2}
 \setcounter{commentCounter}{0}

### 2014-06-02 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Another tweak to the dpp ms results section.
*Now*, it should be ready for re-submission.

@@
 uncertainty.
+\change{
+The small difference between the results of the \npModelDPP and
+\npModelDPPOrdered models is consistent across multiple analyses, and thus
+could be due to error introduced to the \npModelDPP model by the invalid
+shuffling of the summary statitic vectors.
+}
 Both models estimate a similar set of 10 unordered divergence models with the
@@
 
-The main advantage of the \npModelDPPOrdered over the \npModelDPP is that the
-identity of the taxa is maintained, and thus a fully marginalized estimate of
-divergence times across the taxa can be obtained
-(Figure~S\ref{figMarginalTimes}).
-Furthermore, the probability of co-divergence among any set of taxa can be
+\change{
+The main advantages of the \npModelDPPOrdered model over the \npModelDPP are
+that
+(1) the incorrect shuffling of the summary statistic vectors is avoided,
+(2) the identity of the taxa is maintained, and thus a fully marginalized
+estimate of divergence times across the taxa can be obtained
+(Figure~S\ref{figMarginalTimes}), and
+(3) the probability of co-divergence among any set of taxa can be
 estimated from the posterior sample.
+}
 

### 2014-06-02 manuscripts/dpp-msbayes/reviews/response.tex
Review response ready for resubmission with dpp ms.

@@
     Also, the intuition of Reviewer 2 is correct, I do not use the divergence
-    times to calculate summary statistics. I mention the divergence times in
+    times to calculate summary statistics. I mentioned the divergence times in
     the description of the ABC algorithm with the hope of making it clear that
@@
 \revcomment{
-    Quoting [43] as concluding that ABC model choice is biased is
-    somewhat misleading: the warning therein was that Bayes factors and
-    posterior probabilities resulting from those ABC steps could be quite
-    unrelated with those based on the whole dataset. That the same authors
-    concluded that the proper choice of summary statistics leads to a
-    consistent model choice shows ABC model choice is not necessarily
-    ``biased''.
+    Quoting [43] [[citation 43 from the manuscript is now 10 in the current
+    document]] as concluding that ABC model choice is biased is somewhat
+    misleading: the warning therein was that Bayes factors and posterior
+    probabilities resulting from those ABC steps could be quite unrelated with
+    those based on the whole dataset. That the same authors concluded that the
+    proper choice of summary statistics leads to a consistent model choice
+    shows ABC model choice is not necessarily ``biased''.
 }

### 2014-06-02 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Adding Dryad citation to dpp ms.

@@
 the empirical dataset of Oaks et al.\ \cite{Oaks2012}, which is comprised of
-sequence data from 22 pairs of taxa from the Philippine Islands (Drayad DOI:
-10.5061/dryad.5s07m).
+sequence data from 22 pairs of taxa from the Philippine Islands
+(\cite{Oaks2012dryad}; Dryad DOI: 10.5061/dryad.5s07m).
 I analyzed these data under five different models, which are detailed in

### 2014-06-02 manuscripts/dpp-msbayes/cover-letter/cover-letter.tex
Updating dpp ms cover letter for re-submission.

@@
 \usepackage{anysize}
-\marginsize{0.8in}{0.8in}{-0.2in}{0.3in}
+\marginsize{0.6in}{0.6in}{-0.4in}{0.3in}
 \usepackage{setspace}
@@
 % \raggedright
-I am submitting a manuscript entitled ``An Improved Approximate-Bayesian
+I am re-submitting a manuscript entitled ``An Improved Approximate-Bayesian
 Model-choice Method for Estimating Shared Evolutionary History'' to be
 considered for publication in \emph{BMC Evolutionary Biology}.
+The manuscript ID of the previous submission was 9419461361233299.
 This is a methodological article, and the field of research is comparative
@@
 
+The previous submission of this manuscript was handled by Editor Dr.\  Peter
+G.\ Foster, who obtained two anonymous reviews.
+I have incorporated the revisions recommended by Dr.\ Foster and the reviewers.
+My point-by-point response to reviewer comments follows this letter.
 Thank you for your consideration, and I look forward to your correspondence

### 2014-06-16 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Making some edits to dpp ms in proofing stage.
Adding Table S1 to supporting information that give an example of the
divergence model sample space when Y=4 taxa.

Also, making some edits to the method section where the sample space of
ordered/unordered divergence models is introduced. Making the language
in this section more exact and introducing my loose use of "ordered" and
"unordered" subsequent to this section.

@@
 \begin{linenomath}
+\change{
 As for the prior term $p(\divTimeIndexVector)$, the total sample space of
-ordered realizations of \divTimeIndexVector is all the possible partitions of
-\npairs{} elements into 1 to \npairs{} categories.
+\divTimeIndexVector is all the possible partitionings of the \npairs{} pairs of
+populations into 1 to \npairs{} divergence-time classes, where each
+partitioning consists of non-overlapping and non-empty subsets whose union is
+the \npairs{} pairs;
+Hereinafter, I refer to these partitionings as ``ordered'' divergence models or
+partitions.
+}
 The total number of possible partitions is a sum of the Stirling numbers of
@@
 \end{equation}
-which is the Bell Number \cite{Bell1934}.
-The original \msb model only samples over the unordered realizations of
+which is the Bell number \cite{Bell1934}.
+The original \msb model samples over the unordered realizations of
 \divTimeIndexVector, such that the sample space is reduced to all the possible
 integer partitions of \npairs{} \cite{Oaks2012,Huang2011,OeisPartitionNumber,
-    OeisPartitionTriangle,Malenfant2011}.
-I denote the set of all possible integer partitions of \npairs{} as
-\integerPartitionSet{\npairs{}} and the length of that set as
-\integerPartitionNum{\npairs{}}.
+OeisPartitionTriangle,Malenfant2011} (Table~S\ref{tabSampleSpace}).
+\change{
+I denote the set of all possible integer partitions of the \npairs{} pairs of
+populations as \integerPartitionSet{\npairs{}} and the length of that set as
+\integerPartitionNum{\npairs{}}, and I hereinafter refer to these integer
+partitions as ``unordered'' divergence models or partitions.}
 The advantages, disadvantages, and justification of ignoring the order
@@
 \bottleScalar{}{} parameters per pair, no migration, no recombination, and
-resorting of taxon-specific summary statistics by $\pi_b$ (i.e., sampling
+re-sorting of taxon-specific summary statistics by $\pi_b$ (i.e., sampling
 unordered divergence models).

### 2014-06-16 manuscripts/dpp-msbayes/dpp-msbayes-si-tables.tex
Making some edits to dpp ms in proofing stage.
Adding Table S1 to supporting information that give an example of the
divergence model sample space when Y=4 taxa.

Also, making some edits to the method section where the sample space of
ordered/unordered divergence models is introduced. Making the language
in this section more exact and introducing my loose use of "ordered" and
"unordered" subsequent to this section.

%:TABLE-model-sample-spaces
\begin{table}%[htbp]
    % \small
    % \addtolength{\tabcolsep}{-0.9mm}
    \centering
    \rowcolors{2}{}{myGray} \mcrowcolors
    \captionsetup{name=Table S, labelformat=noSpace}
    \caption{An example showing the number of divergence events (\divTimeNum)
        and the associated sample space of the unordered divergence models
        (integer partitions of \npairs{} pairs) and ordered divergence models
        (partitions of \npairs{} pairs) for $\npairs{} = 4$ pairs of
        populations.}
    \begin{tabular}{ c | l | l }
        \divTimeNum & Unordered divergence models & Ordered divergence models \\
        \hline
        1 & $4$          & 1111                                     \\
        2 & $3+1$; $2+2$ & 1112, 1121, 1211, 2111, 1122, 1212, 1221 \\
        3 & $2+1+1$      & 1123, 1213, 1231, 1223, 1232, 1233       \\
        4 & $1+1+1+1$    & 1234                                     \\
    \end{tabular}
    \label{tabSampleSpace}
\end{table}

### 2014-06-16 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Final edits to dpp ms in proofing stage.

@@
     (\divTimeDispersion) is also given for each model, followed by the 95\%
-    highest posterior density interval in parantheses.
+    highest posterior density interval in parentheses.
 }{figPhilippines}
@@
     (\divTimeDispersion) is also given for each model, followed by the 95\%
-    highest posterior density interval in parantheses.
+    highest posterior density interval in parentheses.
 }{figNegrosPanay}

### 2014-06-16 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Final edits to dpp ms in proofing stage.

@@
 number of divergence parameters (denoted \divTimeNum here, and as \numt{} in
-\cite{Huang2011}) imposes an U-shaped prior over divergence
+\cite{Huang2011}) imposes an ``U-shaped'' prior over divergence
 models (\divTimeIndexVector; see
@@
 S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
-Importantly, the \modelOld tends to strongly support these estimates
-of one divergence across most of the simulations
+Importantly, the \modelOld \change{model} tends to strongly support these
+estimates of one divergence across most of the simulations
 (Figure~\ref{figPowerPsiProbOld4}A--D \& plots A--F of Figures
@@
 (Figure~\ref{figPowerPsiProbOld4} \& 
-S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+S\labelcref{figPowerPsiProbOld,figPowerPsiProbUniform,figPowerPsiProbExp}).
 The \modelDPP and \modelUniform models estimate low posterior probability of
@@
 (Figure~\ref{figPowerPsiProbOld4} \& 
-S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+S\labelcref{figPowerPsiProbOld,figPowerPsiProbUniform,figPowerPsiProbExp}).
 The exponential priors on divergence-time and demographic parameters (model
@@
 (Figure~\ref{figPowerPsiProbOld4} \& 
-S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
+S\labelcref{figPowerPsiProbOld,figPowerPsiProbUniform,figPowerPsiProbExp}).
 The \modelDPP and \modelUniform models do frequently support the one-divergence
@@
 (Figure~\ref{figPowerOmegaProbOld4} \& 
-S\labelcref{figPowerOmegaProbOld,figPowerOmegaUniform,figPowerOmegaExp}).
+S\labelcref{figPowerOmegaProbOld,figPowerOmegaProbUniform,figPowerOmegaProbExp}).
 
@@
 could be due to error introduced to the \npModelDPP model by the invalid
-shuffling of the summary statitic vectors.
+shuffling of the summary statistic vectors.
 }

### 2014-06-17 empirical-analyses/bin/plot-results.py
Updating plots for empirical Philippine results.
Modifying `empirical-analyses/bin/plot-results.py` to plot dpp-msbayes
vs msbayes psi results with the priors included. Also committing updated
PDFs of other plots produced by this script, but the code to produce
these plots was not changed (slight differences in fonts due to
different versions of PyMsBayes and/or matplotlib).

@@ -22,7 +22,7 @@ def get_dpp_psi_values(num_elements, shape, scale, num_sims = 100000):
     for i in range(num_sims):
         a = conc.draw()
         x = p.dirichlet_process_draw(a)
-        psis.append(len(set(x)))
+        psis.append(len(set(x.partition)))
     return psis
 
 def create_plots(dpp_info_path, old_info_path, out_dir):
@@ -149,6 +149,63 @@ def create_plots(dpp_info_path, old_info_path, out_dir):
     pg.reset_figure()
     pg.savefig(os.path.join(out_dir, 'philippines-dpp-psi-posterior-prior-lablels.pdf'))
 
+    prior_psis_old = []
+    for i in range(22):
+        prior_psis_old.extend([i + 1] * 100)
+    prior_hd_old = HistData(x = prior_psis_old,
+            normed = True,
+            bins = bins,
+            histtype = 'bar',
+            align = 'mid',
+            orientation = 'vertical',
+            zorder = 0)
+    prior_hist_old = ScatterPlot(hist_data_list = [prior_hd_old],
+            x_label = 'Number of divergence events',
+            y_label = 'Prior probability',
+            xticks_obj = xticks_obj)
+    prior_hist.set_xlim(left = bins[0], right = bins[-1])
+    prior_hist.set_ylim(bottom = 0.0, top = 0.5)
+
+    hist.set_ylim(bottom = 0.0, top = 0.5)
+    prior_hist.set_ylim(bottom = 0.0, top = 0.5)
+
+    for h in [hist_old, hist, prior_hist_old, prior_hist]:
+        h.set_ylabel(ylabel = '')
+        h.set_xlabel(xlabel = '')
+        h.set_title_text('')
+        h.set_extra_y_label('')
+
+    pg = PlotGrid(subplots = [hist_old, hist, prior_hist_old, prior_hist],
+            num_columns = 2,
+            height = 6.0,
+            width = 8.0,
+            share_x = True,
+            share_y = False,
+            label_schema = None,
+            auto_height = False,
+            title = r'Number of divergence events',
+            title_top = False,
+            title_size = 16.0,
+            y_title = 'Probability',
+            y_title_size = 16.0,
+            column_labels = [r'msBayes', r'dpp-msbayes'],
+            row_labels = ['Posterior', 'Prior'],
+            column_label_offset = 0.07,
[…]

### 2014-06-17 slides/bin/plot-power-results.py
Updating plots for slides.
Updating plotting scripts that plot the validation and power results for
use in slides. Also, updating the plot PDFs generated by these scripts.
Making fonts larger in the plots.

@@ -164,11 +164,13 @@ def create_plots(info_path,
             height = 3.5
             column_label_offset = 0.14
             margin_top = 0.83
-            margin_bottom = 0.085
-            margin_left = 0.017
+            margin_bottom = 0.1
+            margin_left = 0.025
             margin_right = 0.95
-            column_label_size = 22.0
-            title_size = 18.0
+            column_label_size = 26.0
+            title_size = 22.0
+            y_title_size = 20.0
+            right_text_size = 14.0
             row_labels = []
             if prior_name == 'old':
                 row_labels.append(r'$M_{msBayes}$')
@@ -180,7 +182,7 @@ def create_plots(info_path,
                 row_labels.append(r'$M_{DPP}$')
             else:
                 row_labels.append('')
-            row_label_size = 28.0
+            row_label_size = 32.0
             row_label_offset = 0.08
 
 
@@ -191,7 +193,11 @@ def create_plots(info_path,
                     num_columns = num_columns,
                     width = width,
                     height = height,
-                    x_title = r'Estimated number of divergence events (mode)')
+                    x_title = r'Estimated number of divergence events (mode)',
+                    y_title_size = y_title_size,
+                    include_right_text = False,
+                    text_size = right_text_size,
+                    xtick_label_size = 14.0)
             pg = psi_plot.create_grid()
             pg.label_schema = None
             column_labels = []
@@ -203,6 +209,7 @@ def create_plots(info_path,
             pg.margin_top = margin_top
             pg.margin_left = margin_left
             pg.margin_right = margin_right
+            pg.padding_between_horizontal = 2.0
             pg.column_label_offset = column_label_offset
             pg.column_label_size = column_label_size
             pg.title_size = title_size
@@ -239,9 +246,13 @@ def create_plots(info_path,
                     bayes_factor = 10,
                     draw_bayes_factor_line = False,
                     x_title = r'Posterior probability of one divergence',
+                    y_title_size = y_title_size,
                     width = width,
                     height = height,
-                    num_columns = num_columns)
+                    num_columns = num_columns,
+                    text_size = right_text_size,
+                    xtick_label_size = 14.0,
+                    pretty_xtick_labels = True)
             pg = psi_prob_plot.create_grid()
             pg.label_schema = None
             column_labels = []
@@ -256,10 +267,14 @@ def create_plots(info_path,
             pg.margin_right = margin_right
             pg.column_label_offset = column_label_offset
[…]

### 2014-06-17 slides/bin/plot-validation-results.py

@@ -80,17 +80,17 @@ def main_cli():
             share_x = True,
             share_y = True,
             title = 'Posterior probability of one divergence',
-            title_size = 12.0,
+            title_size = 16.0,
             title_top = False,
             y_title = 'True probability of one divergence',
             y_title_position = 0.03,
-            y_title_size = 12.0,
+            y_title_size = 16.0,
             height = 6.0,
             auto_height = False,
             column_labels = [r'$M_{msBayes}$', r'$M_{DPP}$'],
             row_labels = [r'$M_{msBayes}$', r'$M_{DPP}$'],
-            column_label_size = 20.0,
-            row_label_size = 20.0,
+            column_label_size = 22.0,
+            row_label_size = 22.0,
             column_label_offset = 0.05,
             row_label_offset = 0.05,
             super_title = r'\textbf{Data model}',
@@ -100,8 +100,8 @@ def main_cli():
             super_y_title_right = True)
     pg.label_schema = None
     pg.auto_adjust_margins = False
-    pg.margin_bottom = 0.04
-    pg.margin_left = 0.06
+    pg.margin_bottom = 0.05
+    pg.margin_left = 0.07
     pg.margin_top = 0.90
     pg.margin_right = 0.92
     pg.reset_figure()
@@ -126,18 +126,18 @@ def main_cli():
             share_x = True,
             share_y = True,
             title = 'Posterior probability of one divergence',
-            title_size = 12.0,
+            title_size = 16.0,
             title_top = False,
             y_title = 'True probability of one divergence',
             y_title_position = 0.015,
-            y_title_size = 12.0,
+            y_title_size = 16.0,
             height = 4.8,
             width = 11.0,
             auto_height = False,
             row_labels = [r'$M_{msBayes}$', r'$M_{DPP}$'],
             column_labels = [r'$M_{msBayes}$', r'$M_{DPP}$', r'$M_{Uniform}$', r'$M_{Ushaped}$'],
-            column_label_size = 20.0,
-            row_label_size = 20.0,
+            column_label_size = 22.0,
+            row_label_size = 22.0,
             column_label_offset = 0.05,
             row_label_offset = 0.05,
             super_title = r'\textbf{Data model}',
@@ -147,8 +147,8 @@ def main_cli():
             super_y_title_right = True)
     pg.label_schema = None
     pg.auto_adjust_margins = False
-    pg.margin_bottom = 0.05
[…]

### 2014-06-17 slides/bin/plot-power-results.py
Updating plots of power results for slides.
Updating plotting script that creates plots of power results for use in
slides. Adding code to create a single plot of the psi results for the
msBayes model under the broadest range of random divergence times.
Updating PDFs of other plots created by the script (these haven't
changed).

@@ -227,15 +227,33 @@ def create_plots(info_path,
             pg.savefig(os.path.join(output_dir,
                     prefix + '_power_psi_mode_headless.pdf'))
 
-            # psi_plot_glm = PowerPlotGrid(
-            #         observed_config_to_estimates = cfg_to_psi_glm,
-            #         variable = 'psi',
-            #         variable_symbol = r'|\mathbf{\tau}|',
-            #         num_columns = 2,
-            #         margin_top = 0.975)
-            # fig = psi_plot_glm.create_grid()
-            # fig.savefig(os.path.join(output_dir,
-            #         prefix + '_power_psi_mode_glm.pdf'))
+            psi_plot.width = 5.0
+            psi_plot.height = 4.0
+            pg = psi_plot.create_column_grid(
+                subplot_indices_to_exclude = [0,1,2],
+                x_title_size = 16.0,
+                y_title_size = 15.0,
+                add_column_labels = True,
+                column_label_size = 18.0,
+                column_label_offset = 0.14,
+                plot_label_size = 12.0,
+                right_text_size = 10.0,
+                x_tick_label_size = 14.0,
+                share_x = False,
+                share_y = False)
+            pg.label_schema = None
+            pg.column_labels = [[t for (c, t) in sorted(column_labels, key = lambda x : x[0])][-1]]
+            pg.margin_bottom = 0.07
+            pg.margin_top = 0.88
+            pg.margin_left = 0.07
+            pg.margin_right = 0.99
+            pg.column_label_offset = 0.04
+            pg.column_label_size = column_label_size
+            # pg.title_size = title_size
+            pg.row_labels = None
+            pg.reset_figure()
+            pg.savefig(os.path.join(output_dir,
+                    prefix + '_power_psi_mode_last.pdf'))
 
             psi_prob_plot = ProbabilityPowerPlotGrid(
                     observed_config_to_estimates = cfg_to_psi_prob,

### 2014-07-02 manuscripts/reply/reply-figures.tex
Updating reply ms.
Moving coin flip figure to SI, changing to right justified, minor edit
to section title.
Working on reply ms.
Moving sampling error figure and the plot of Hickerson et al.'s
mixture-model prior on divergence times to the Supporting Information.

--- 2013-12-26 manuscripts/reply/reply-figures.tex
@@
-\mFigure{../images/coin_flip.pdf}{
-    A plot of three beta probability density functions that represent a prior
-    (black; $beta(10, 10)$), posterior (blue; $beta(13, 17)$), and empirical
-    Bayes density (red; $beta(16, 24)$) for a dataset of 10 coin flips, three
-    of which are successes.
-}{figCoinFlip}
-
 %% Exclusion simulation results
@@
 }{figPowerExclusion4}
-
-%% Sampling error
-\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
-    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
-    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
-    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
-    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
-    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
-    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
-}{figSamplingError}
-
-%% Model-choice tau prior
-\mFigure{../../model-choice/tau-prior/tau_prior.pdf}{
-    The prior distribution on divergence times imposed by the model-averaging prior
-    comprised of five models with different uniform priors on \divt{}:
-    $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_2$ ($\divt{} \sim U(0, 1)$), $M_3$
-    ($\divt{} \sim U(0, 5)$), $M_4$ ($\divt{} \sim U(0, 10)$), $M_5$ ($\divt{}
-    \sim U(0, 20)$).
-}{figMCTauPrior}

### 2014-07-02 manuscripts/reply/reply-main.tex
Working on reply ms.
Moving sampling error figure and the plot of Hickerson et al.'s
mixture-model prior on divergence times to the Supporting Information.

@@
 divergence times (\vmratio{}) as the prior samples accumulate
-(Figure~\ref{figSamplingError}) we do not see the trend predicted by
+(Figure~S\ref{figSamplingError}) we do not see the trend predicted by
 \citet{Hickerson2013} in either the unadjusted or GLM-regression-adjusted
@@
 the model-averaging approach of \citet{Hickerson2013} we see it is a mixture of
-overlapping uniforms with lower limits of zero (Figure~\ref{figMCTauPrior}).
+overlapping uniforms with lower limits of zero (Figure~S\ref{figMCTauPrior}).
 This looks very much like an exponential distribution, except that in any state
@@
 This would capture the prior uncertainty that \citet{Hickerson2013} are
-suggesting for divergence times (Figure~\ref{figMCTauPrior}) while avoiding
+suggesting for divergence times (Figure~S\ref{figMCTauPrior}) while avoiding
 costly model-averaging and the constraint that all divergence times must fall

### 2014-07-02 manuscripts/reply/reply-si-figures.tex
Updating reply ms.
Moving coin flip figure to SI, changing to right justified, minor edit
to section title.
Working on reply ms.
Moving sampling error figure and the plot of Hickerson et al.'s
mixture-model prior on divergence times to the Supporting Information.

@@
+\siFigure{../images/coin_flip.pdf}{
+    A plot of three beta probability density functions that represent a prior
+    (black; $beta(10, 10)$), posterior (blue; $beta(13, 17)$), and empirical
+    Bayes density (red; $beta(16, 24)$) for a dataset of 10 coin flips, three
+    of which are successes.
+}{figCoinFlip}
+
 \siFigure{../../response-redux/hickerson-et-al-posterior/hickerson-posterior-1k/mean_by_dispersion.pdf}{
@@
 
+%% Sampling error
+\widthFigure{0.9}{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{listformat=sFigList}{
+    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
+    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
+    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
+    from the prior. Both (A) unadjusted and (B) GLM-regression-adjusted estimates are shown.
+    The data analyzed were the 22 pairs of Philippine taxa from \citet{Oaks2012}.
+    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
+    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
+}{figSamplingError}
+\clearpage
+
 %% Validation results
@@
 }{figSaturationPlot}
+
+%% Model-choice tau prior
+\siFigure{../../model-choice/tau-prior/tau_prior.pdf}{
+    The prior distribution on divergence times imposed by the model-averaging prior
+    comprised of five models with different uniform priors on \divt{}:
+    $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_2$ ($\divt{} \sim U(0, 1)$), $M_3$
+    ($\divt{} \sim U(0, 5)$), $M_4$ ($\divt{} \sim U(0, 10)$), $M_5$ ($\divt{}
+    \sim U(0, 20)$).
+}{figMCTauPrior}

### 2014-07-02 manuscripts/reply/reply-main.tex
Working on reply ms.
Moving coin-flipping example to supporting information.

@@
 However, this is far removed from the single model-choice problem of \msb.
-
-\begin{linenomath}
-These distinctions between Bayesian parameter estimation and model choice can
-be illustrated with a simple example.
-Let us say we are interested in the fairness of a particular coin, and we
-denote the unknown probability of it landing heads as \myTheta{}.
-More specifically, we are interested in the probability of two models, $M_1$
-and $M_2$.
-In both models the outcomes of flipping the coin are assumed to be binomially
-distributed, but under $M_1$ the coin is weighted toward landing heads (i.e.,
-$\myTheta{} > 0.5)$), whereas under $M_2$, the coin is weighted toward landing
-tails (i.e., $\myTheta{} < 0.5$).
-We already have data from flipping a different coin 20 times that landed both
-heads and tails 10 times each, and so we decide to use these data in specifying
-a beta prior on fairness of the new coin of $beta(a=10, b=10)$
-(Figure~S\ref{figCoinFlip}).
-We collect data by flipping the coin of interest $N=10$ times, $y=3$ of which
-land heads.
-Given the beta distribution is a conjugate prior for a binomial likelihood, the
-posterior distribution has the nice analytical form $\theta \given y,N \sim
-beta(a + y, b + N - y)$, which for the new dataset is simply $beta(13, 17)$
-(Figure~S\ref{figCoinFlip}).
-The maximum a posteriori (MAP) estimate of the probability of heads is 0.429,
-and following Equation~\ref{eq:marginallikelhiood} the marginal likelihoods of
-our models of interest are
-\begin{equation}
-    p(y=3, N=10 \given M_1) = \int_{0.5}^{1} p(y=3, N=10 \given
-    \myTheta{}, M_1)p(\myTheta{} \given M_1) d\myTheta{} \approx 0.029,
-\end{equation}
-and
-\begin{equation}
-    p(y=3, N=10 \given M_2) = \int_{0}^{0.5} p(y=3, N=10 \given
-    \myTheta{}, M_2)p(\myTheta{} \given M_2) d\myTheta{} \approx 0.097.
-\end{equation}
-Given the models have equal probability under our prior, we can calculate the
-posterior probability of Model 1 as
-\begin{equation}
-    p(M_1 \given y=3, N=10) = \frac{p(y=3, N=10 \given M_1)}{p(y=3, N=10 \given
-    M_1) + p(y=3, N=10 \given M_2)} \approx 0.23.
-\end{equation}
-This is the correct posterior probability of Model 1 given our prior
-and data.
-\end{linenomath}
-
-To give the data more weight relative to the prior, we could use it twice, and
-calculate an empirical Bayes estimate using a prior of $beta(13,17)$.  This
-results in a ``posterior'' distribution of $beta(16, 24)$
-(Figure~S\ref{figCoinFlip}), with a MAP estimate of 0.395, and $p(M_1 \given
-y=3, N=10) = 0.10$.
-The estimated posterior distribution of the parameter, and resulting MAP
-estimate, is similar whether or not an empirically informed prior is used.
-However, the posterior probability of Model 1 is very sensitive to the
-empirical prior, decreasing by 56\%.
-By using the empirically informed prior, we ignored prior uncertainty, leading
-to an underestimate of our posterior uncertainty (Figure~S\ref{figCoinFlip}).
-While this did not greatly affect our estimate of \myTheta{}, it misled us
-to be overconfident in Model 2.
-% Our mode estimate of \myTheta{} under the empirical prior might be closer
-% to the true value than our posterior mode.
-% Let us assume the truth of the matter is that the coin under study has a broad
-% flat edge, and as a result lands on its edge at a certain frequency.
-% Hence, both of our models are incorrect (i.e., the outcomes of flipping the
-% coin are not binomially distributed).
-
-% This can be demonstrated with a simple, albeit contrived, example.
-% Let us say that principal investigator Mary and her new postdoc Will are
-% interested in the hypothesis that the mass of George Washington's periwig
-% renders the quarter dollar of the United States unfair.
-% That is to say their null hypothesis is that the probability of a US
-% quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
-% They can certainly evaluate this hypothesis, they can have undergraduate
-% worker, Joe, flip the coin for them while they tabulate the results.
-% But being Bayesians, before they call Joe into the lab, they agree on a
-% prior probability to place on the set of all possible probabilities that
-% the quarter will land heads when it is flipped.
-% Given that neither of them have a quarter, and their prior knowledge that Joe
-% moonlights as a magician, and is notorious for performing coin and card tricks
-% in the lab, they suspect there is a good chance that the quarter Joe uses will
-% be either two-headed or two-tailed.
-% So, knowing that the beta distribution is the conjugate prior for a binomial
-% likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
-% (Figure~\ref{figCoinFlip}).
-
-% Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
-% begin flipping it.
-% After five tosses, four of which was heads, Joe decides that academics are
-% crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
-% being computational biologists, are satisfied with their
-% empirical dataset of $y = 4$ heads out of $N = 5$ trials.
-% They know from the conjugacy of the beta prior, that the posterior distribution
-% has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
-% in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
-% of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
-% This allows them to plug these values into the beta cumulative distribution
-% function to determine that the posterior probability of their hypothesis is
-% $p(\theta < 0.5 | y=1, N=5) = 0.088$.
-% Given their prior belief and dataset, this indeed is the correct posterior
-% probability of the hypothesis, and Mary and Will should now update their
-% posterior belief accordingly.
-
-% However, reflecting upon the results of their experiment
-% (Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
-% Their prior looks very ``poorly selected,'' and if they had only known that the
-% coin was fair before the data were collected, they would have selected a much
-% better prior.
-% Clearly, from their results, they should have used a prior centered
-% around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
-% much ``better.''
-% Rather than resort to flipping the coin five more times, Mary and Will decide
-% to redo their analysis using the much better ``prior'' of $p(\theta) \sim
-% beta(4.5, 1.5)$.
-% This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
-% probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
-% Now convinced that the US quarter is unfair, albeit not due to the mass of
-% President Washington's head as they hypothesized, Will begins composing an
-% e-mail of complaint addressed to the U.S.\ Mint.
-
-% If Joe's flips were a representative sample, Mary and Will's empirical Bayes
-% estimate might very well be a better point estimate for the parameter
-% \myTheta{}.
-% However, their empirical Bayes estimate of the probability of their hypothesis
-% is incorrect and biased.
-% This simple example shows how parameter estimation is fundamentally different
-% from estimating the probability of a model.
-% While empirically informed priors can be used to obtain well-behaved parameter
-% estimators, using them for model choice is much less certain.
+In the Supporting Information we use a simple example to help highlight the
+distinctions between Bayesian parameter estimation and model choice.
 

### 2014-07-02 manuscripts/reply/reply-si-main.tex
Working on reply ms.
Moving coin-flipping example to supporting information.

@@
 time units indicated (Figure~\ref{figJointPosteriorHickerson}).
+
+\section{Theoretical implications of empirical priors for Bayesian model
+choice---A simple example}
+\begin{linenomath}
+The distinctions between Bayesian parameter estimation and model choice
+discussed in the main text can be illustrated with a simple example.
+Let us say we are interested in the fairness of a particular coin, and we
+denote the unknown probability of it landing heads as \myTheta{}.
+More specifically, we are interested in the probability of two models, $M_1$
+and $M_2$.
+In both models the outcomes of flipping the coin are assumed to be binomially
+distributed, but under $M_1$ the coin is weighted toward landing heads (i.e.,
+$\myTheta{} > 0.5)$), whereas under $M_2$, the coin is weighted toward landing
+tails (i.e., $\myTheta{} < 0.5$).
+We already have data from flipping a different coin 20 times that landed both
+heads and tails 10 times each, and so we decide to use these data in specifying
+a beta prior on fairness of the new coin of $beta(a=10, b=10)$
+(Figure~S\ref{figCoinFlip}).
+We collect data by flipping the coin of interest $N=10$ times, $y=3$ of which
+land heads.
+Given the beta distribution is a conjugate prior for a binomial likelihood, the
+posterior distribution has the nice analytical form $\theta \given y,N \sim
+beta(a + y, b + N - y)$, which for the new dataset is simply $beta(13, 17)$
+(Figure~S\ref{figCoinFlip}).
+The maximum a posteriori (MAP) estimate of the probability of heads is 0.429,
+and following Equation~\ref{eq:marginallikelhiood} in the main text the
+marginal likelihoods of our models of interest are
+\begin{equation}
+    p(y=3, N=10 \given M_1) = \int_{0.5}^{1} p(y=3, N=10 \given
+    \myTheta{}, M_1)p(\myTheta{} \given M_1) d\myTheta{} \approx 0.029,
+\end{equation}
+and
+\begin{equation}
+    p(y=3, N=10 \given M_2) = \int_{0}^{0.5} p(y=3, N=10 \given
+    \myTheta{}, M_2)p(\myTheta{} \given M_2) d\myTheta{} \approx 0.097.
+\end{equation}
+Given the models have equal probability under our prior, we can calculate the
+posterior probability of Model 1 as
+\begin{equation}
+    p(M_1 \given y=3, N=10) = \frac{p(y=3, N=10 \given M_1)}{p(y=3, N=10 \given
+    M_1) + p(y=3, N=10 \given M_2)} \approx 0.23.
+\end{equation}
+This is the correct posterior probability of Model 1 given our prior
+and data.
+\end{linenomath}
+
+To give the data more weight relative to the prior, we could use it twice, and
+calculate an empirical Bayes estimate using a prior of $beta(13,17)$.  This
+results in a ``posterior'' distribution of $beta(16, 24)$
+(Figure~S\ref{figCoinFlip}), with a MAP estimate of 0.395, and $p(M_1 \given
+y=3, N=10) = 0.10$.
+The estimated posterior distribution of the parameter, and resulting MAP
+estimate, is similar whether or not an empirically informed prior is used.
+However, the posterior probability of Model 1 is very sensitive to the
+empirical prior, decreasing by 56\%.
+By using the empirically informed prior, we ignored prior uncertainty, leading
+to an underestimate of our posterior uncertainty (Figure~S\ref{figCoinFlip}).
+While this did not greatly affect our estimate of \myTheta{}, it misled us
+to be overconfident in Model 2.
+% Our mode estimate of \myTheta{} under the empirical prior might be closer
+% to the true value than our posterior mode.
+% Let us assume the truth of the matter is that the coin under study has a broad
+% flat edge, and as a result lands on its edge at a certain frequency.
+% Hence, both of our models are incorrect (i.e., the outcomes of flipping the
+% coin are not binomially distributed).
+
+% This can be demonstrated with a simple, albeit contrived, example.
+% Let us say that principal investigator Mary and her new postdoc Will are
+% interested in the hypothesis that the mass of George Washington's periwig
+% renders the quarter dollar of the United States unfair.
+% That is to say their null hypothesis is that the probability of a US
+% quarter landing heads when tossed is less than 0.5 ($\theta < 0.5$).
+% They can certainly evaluate this hypothesis, they can have undergraduate
+% worker, Joe, flip the coin for them while they tabulate the results.
+% But being Bayesians, before they call Joe into the lab, they agree on a
+% prior probability to place on the set of all possible probabilities that
+% the quarter will land heads when it is flipped.
+% Given that neither of them have a quarter, and their prior knowledge that Joe
+% moonlights as a magician, and is notorious for performing coin and card tricks
+% in the lab, they suspect there is a good chance that the quarter Joe uses will
+% be either two-headed or two-tailed.
+% So, knowing that the beta distribution is the conjugate prior for a binomial
+% likelihood, they decide to use a $beta(a=0.5, b=0.5)$ prior distribution
+% (Figure~\ref{figCoinFlip}).
+
+% Mary calls Joe into the lab, confirms that he has a quarter, and tells him to
+% begin flipping it.
+% After five tosses, four of which was heads, Joe decides that academics are
+% crazy and leaves the lab to pursue a major in theatre.  Mary and Will, both
+% being computational biologists, are satisfied with their
+% empirical dataset of $y = 4$ heads out of $N = 5$ trials.
+% They know from the conjugacy of the beta prior, that the posterior distribution
+% has the nice analytical form $\theta|y,N \sim beta(a + y, b + N - y)$, which
+% in this case is simply $beta(4.5, 1.5)$; this is the true posterior distribution
+% of $\theta$ given their prior belief and data (Figure~\ref{figCoinFlip}).
+% This allows them to plug these values into the beta cumulative distribution
+% function to determine that the posterior probability of their hypothesis is
+% $p(\theta < 0.5 | y=1, N=5) = 0.088$.
+% Given their prior belief and dataset, this indeed is the correct posterior
+% probability of the hypothesis, and Mary and Will should now update their
+% posterior belief accordingly.
+
+% However, reflecting upon the results of their experiment
+% (Figure~\ref{figCoinFlip}), Mary and Will regret their choice of prior.
+% Their prior looks very ``poorly selected,'' and if they had only known that the
+% coin was fair before the data were collected, they would have selected a much
+% better prior.
+% Clearly, from their results, they should have used a prior centered
+% around 0.2; their data suggest a prior of $beta(4.5, 1.5)$ would have been
+% much ``better.''
+% Rather than resort to flipping the coin five more times, Mary and Will decide
+% to redo their analysis using the much better ``prior'' of $p(\theta) \sim
+% beta(4.5, 1.5)$.
+% This gives a ``posterior'' of $\theta|y,N \sim beta(8.5, 2.5)$, and a
+% probability of their hypothesis of $p(\theta < 0.5 | y=1, N=5) = 0.026$.
+% Now convinced that the US quarter is unfair, albeit not due to the mass of
+% President Washington's head as they hypothesized, Will begins composing an
+% e-mail of complaint addressed to the U.S.\ Mint.
+
+% If Joe's flips were a representative sample, Mary and Will's empirical Bayes
+% estimate might very well be a better point estimate for the parameter
+% \myTheta{}.
+% However, their empirical Bayes estimate of the probability of their hypothesis
+% is incorrect and biased.
+% This simple example shows how parameter estimation is fundamentally different
+% from estimating the probability of a model.
+% While empirically informed priors can be used to obtain well-behaved parameter
+% estimators, using them for model choice is much less certain.
 

### 2014-07-02 manuscripts/reply/reply-abstract.tex
More editing of reply ms.

@@
 using computer simulations.
-\citet{Oaks2012} found that the model was very sensitive to the prior and
-that the inference method had low power to detect variation in divergences
-times.
+\citet{Oaks2012} found the method was very sensitive to prior assumptions and
+had low power to detect variation in divergence times.
 These results were not surprising, in light of a rich statistical literature
@@
 underlie model inference.
-Furthermore, simulations provide insight into the temporal resolution of
-the method, which in turn helps guide interpretation of results.
+Furthermore, simulations provide insight into the temporal resolution of the
+method, which helps guide interpretation of results.

### 2014-07-02 manuscripts/reply/reply-main.tex
Working on reply ms.
Fixing two citations of `Oaks2014dpp`; changing `citet` to `citep`.
More editing of reply ms.

@@
 unique time.
-A full description of the model is given in \citet{Oaks2012}.
-Frequently, the exact values of the divergence times are hard 
-to determine because researchers typically lack precise knowledge 
-of the rate at which substitutions occur.
-The exact divergence time is therefore often viewed as a nuisance parameter,
-and researchers focus on the number of divergence events.
+% A full description of the model is given in \citet{Oaks2012}.
+% Frequently, the exact values of the divergence times are hard 
+% to determine because researchers typically lack precise knowledge 
+% of the rate at which substitutions occur.
+% The exact divergence time is therefore often viewed as a nuisance parameter,
+% and researchers focus on the number of divergence events.
 The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
@@
 prior density imposed by the uniformly distributed prior on divergence times.
+% By using such priors on divergence times, we are essentially telling the
+% method to favor models with few divergence times even if they do not
+% explain the data very well.
 Note, this is a statistical issue, extrinsic to \msb, and does not question the
@@
 
-\citet{Hickerson2013} suggest the bias is the result of a numerical issue
+\citet{Hickerson2013} suggest the problem is the result of a numerical issue
 intrinsic to \msb, concluding the method's rejection algorithm is inefficient
@@
 detect temporal variation in divergences
-\citet{Oaks2014dpp}.
+\citep{Oaks2014dpp}.
 
@@
 their approach is biased toward models with less parameter space, and as a
-consequence, is biased toward estimating model-averaged posteriors dominated by
-models that exclude true values of parameters of the model.
+consequence, will estimate model-averaged posteriors dominated by models that
+exclude true values of parameters of the model.
 We explored this possibility in two ways.
@@
 regions of low likelihood, which greatly improves the power of the method
-\citet{Oaks2014dpp}.
+\citep{Oaks2014dpp}.
 

### 2014-07-09 manuscripts/reply/reply-main.tex
Working on revisions of reply ms.

@@
 prior density imposed by the uniformly distributed prior on divergence times.
+\change{
 % By using such priors on divergence times, we are essentially telling the
@@
 % explain the data very well.
-Note, this is a statistical issue, extrinsic to \msb, and does not question the
-soundness of the numerical approximation machinery of \msb.
+This affect of the priors on marginal likelihoods creates a penalty for
+over-parameterization in Bayesian model choice.
+However, for this inference problem, the uniform priors on divergence times can
+create a strong preference for simple models that are often of interest to
+biogeographers even if these models do not explain the data very well.
+This phenomenon is extrinsic to the numerical approximation machinery of \msb;
+with such prior assumptions, the true posterior will favor models with
+few divergence events.
+However, such results, even if correct from the perspective of Bayesian model
+choice, do not provide the biogeographical insights that a researcher who
+employs \msb seeks to gain.
+For example, a biogeographer would like to interpret the posterior probability
+of a simultaneous divergence model as an estimate of the probability that this
+model is correct.
+However, when such prior assumptions are made about divergence times, the
+estimated posterior probability is better interpretted as the probability that
+you cannot completely rule out that the simultaneous-divergence model can
+explain the data.
+% Note, this is a statistical issue, extrinsic to \msb, and does not question the
+% soundness of the numerical approximation machinery of \msb.
+}
 
@@
 \citet{Oaks2012} suggest alternative prior probability distributions on
-divergence-times and other nuisance parameters could increase the marginal
-likelihoods of models with more divergence-time parameters and thus reduce
-spurious support for models of temporally clustered divergences.
+divergence-times could increase the marginal likelihoods of models with more
+divergence-time parameters and thus reduce spurious support for models of
+temporally clustered divergences.
 Alternatively, \citet{Hickerson2013} present an approach that uses empirically
@@
 
-In this paper, we discuss the potential theoretical and practical
-considerations of using empirically informed priors for Bayesian model choice.
+\change{
+In this paper, we discuss theoretical and practical considerations of using
+empirically informed priors for Bayesian model choice.
+}
 Furthermore, we evaluate the empirical Bayesian model-averaging approach of
@@
 parameters.
-Our results, when combined with those of \citet{Oaks2012}, suggest that it is
-difficult to choose a uniformly distributed prior on divergence times that is
-broad enough to confidently contain the true values of parameters while being
-narrow enough to avoid spurious support of models with less parameter space.
+\change{
+Our results, show that it is difficult to choose a uniformly distributed prior
+on divergence times that is broad enough to confidently contain the true values
+of parameters while being narrow enough to avoid spurious support of models
+with less parameter space.
+}
 However, recent work shows more flexible probability distributions without a
@@
 \section{The potential implications of empirical Bayesian model choice}
-\citet{Hickerson2013} repeatedly refer to the prior distributions used by
-\citet{Oaks2012} for divergence-time parameters as ``poorly selected.''
-However, this is misleading, as \citet{Oaks2012} discuss in detail how they
-selected their prior to reflect the large amount of uncertainty about the
-timing of divergences across all 22 of the taxa in their study (see section
-``Specifying and simulating the joint prior'' of \citet{Oaks2012}).
-\citet{Oaks2012} further discuss how the use of uniform prior distributions in
-\msb for many of the model's parameters requires investigators to select broad
-priors to avoid excluding the truth \textit{a priori}, resulting in too much
-density in improbable regions of parameter space.
-
-\citet{Hickerson2013} suggest that \citet{Oaks2012} should have used their
-data to inform their prior on divergence times (i.e., an empirical Bayesian
-approach).
-However, \citet{Oaks2012} did use empirically informed priors to mimic
-empirical situations in which a large amount of prior information is available;
-they did so to assess the prior sensitivity of the method and to determine if
-uniform priors in such situations are narrow enough to avoid spurious support
-of models with few divergence-time parameters.
+\change{
 The results of \citet{Oaks2012} and \citet{Hickerson2013} show the method is
@@
 Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
-divergences under the empirically informed priors.
-
+divergences when using a prior on divergence times that is strongly
+informed by the data being analyzed.
 The main argument of \citet{Hickerson2013} is that the priors used by
 \citet{Oaks2012} were not informative enough.
+}
+%
+% \citet{Hickerson2013} repeatedly refer to the prior distributions used by
+% \citet{Oaks2012} for divergence-time parameters as ``poorly selected.''
+% However, this is misleading, as \citet{Oaks2012} discuss in detail how they
+% selected their prior to reflect the large amount of uncertainty about the
+% timing of divergences across all 22 of the taxa in their study (see section
+% ``Specifying and simulating the joint prior'' of \citet{Oaks2012}).
+% \citet{Oaks2012} further discuss how the use of uniform prior distributions in
+% \msb for many of the model's parameters requires investigators to select broad
+% priors to avoid excluding the truth \textit{a priori}, resulting in too much
+% density in improbable regions of parameter space.
+
+% \citet{Hickerson2013} suggest that \citet{Oaks2012} should have used their
+% data to inform their prior on divergence times (i.e., an empirical Bayesian
+% approach).
+% However, \citet{Oaks2012} did use empirically informed priors to mimic
+% empirical situations in which a large amount of prior information is available;
+% they did so to assess the prior sensitivity of the method and to determine if
+% uniform priors in such situations are narrow enough to avoid spurious support
+% of models with few divergence-time parameters.
+% The results of \citet{Oaks2012} and \citet{Hickerson2013} show the method is
+% highly sensitive to the prior on divergence times.
+% Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
+% divergences under the empirically informed priors.
+%
+% The main argument of \citet{Hickerson2013} is that the priors used by
+% \citet{Oaks2012} were not informative enough.
 They suggest a very narrow, highly informed uniform prior on divergence times
@@
 often exhibit favorable frequentist properties.
-Furthermore, many post-hoc correction methods have been developed for
-estimating confidence-intervals from empirical Bayes estimates of
-distributions that often exhibit well-behaved frequentist coverage
-probabilities
+% Furthermore, many post-hoc correction methods have been developed for
+% estimating confidence-intervals from empirical Bayes estimates of
+% distributions that often exhibit well-behaved frequentist coverage
+% probabilities
 \citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.
@@
 due to using the data to inform the priors.
-This is why empirical Bayes can work well for estimating the values of model
-parameters.
+By giving the data more weight relative to the prior, the goal is that the
+posterior distribution will be peaked near the true value(s) of the model's
+parameter(s).
+% This is why empirical Bayes can work well for estimating the values of model
+% parameters.
 \end{linenomath}
@@
 than parameter estimation.
+Also, unlike model parameters, the posterior probabilities of candidate
+models have no clear true values.
+% There is no such justification for empirical Bayesian model choice, because
+% there are no true values for the model probabilities being estimated.
+Model posterior probabilities are inherently measures of our belief in the
+models after our prior beliefs are updated by the data being analyzed.
+This complicates the meaning of model posterior probabilities when Bayes' rule
+is violated by informing priors with the same data to be analyzed.
+By using the data twice, we fail to account for prior uncertainty and mislead
+our posterior beliefs in the models being compared; we will be over confident
+in some models and under confident in others.
 
@@
 posterior uncertainty.
-Also, unlike model parameters, the posterior probabilities of candidate
-models have no clear true values.
 % In some simple situations we might suspect that one model is correct and the
@@
 % consideration are oversimplifications of reality.
-Model posterior probabilities are inherently measures of our belief in the
-models after our prior beliefs are updated by the data being analyzed.
 
-This complicates the meaning of model posterior probabilities when Bayes' rule
-is violated by informing priors with the same data to be analyzed.
-For empirical Bayesian parameter estimation, the objective is that by giving the
-data more weight relative to the prior, the posterior distribution will be
-peaked near the true value(s) of the model's parameter(s).
-There is no such justification for empirical Bayesian model choice, because
-there are no true values for the model probabilities being estimated.
-By using the data twice, we fail to account for prior uncertainty and mislead
-our posterior beliefs in the models being compared; we will be over confident
-in some models and under confident in others.
 Nonetheless, empirical Bayesian model choice does perform well for some
@@
 In addition to the theoretical concerns discussed above, there are practical
-problems with using narrow, empirically informed, uniform priors for a method
-that has already been shown to be biased toward models with less parameter
-space.
+problems with using narrow, empirically informed, uniform priors.
+% for a method that has already been shown to be biased toward models with less
+% parameter space.
 The results of Hickerson et al.'s \citeyear{Hickerson2013} reanalysis of the
@@
 \citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
-divergence times for at least some of the Philippine taxa, a major problem when
-using uniform priors.
+divergence times for at least some of the Philippine taxa.
+%, a major problem when using uniform priors.
 \citet{Hickerson2013} set this prior to match the 95\% highest posterior
@@
 
-The strong preference for the questionable priors on divergence times (their
-$M_1$ and $M_2$) in the model-averaged results of \citet{Hickerson2013} suggest
-their approach is biased toward models with less parameter space, and as a
+\change{
+Hickerson et al.'s \citeyear{Hickerson2013} results that strongly prefer the
+narrowest priors on divergence times (their $M_1$ and $M_2$) suggest their
+approach is biased toward models with less parameter space, and as a
 consequence, will estimate model-averaged posteriors dominated by models that
-exclude true values of parameters of the model.
+exclude true values of model's parameters.
+}
 We explored this possibility in two ways.
@@
 distribution on divergence times across all three of our analyses, even when
-this model is almost certainly incorrect and excludes the true divergence times
+this model is incorrect and excludes the true divergence times
 of the Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
@@
 
-Given that the results of \citet{Hickerson2013} strongly prefer the models with
-the narrowest prior on divergence times, it seems quite likely that their
-model-averaged results are dominated by models that exclude true divergence
-times, making their results difficult to interpret.
+% Given that the results of \citet{Hickerson2013} strongly prefer the models with
+% the narrowest prior on divergence times, it seems quite likely that their
+% model-averaged results are dominated by models that exclude true divergence
+% times, making their results difficult to interpret.
 
@@
 We then repeated the analysis described above using $1\e6$ random samples from
-prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, retaining 1000 posterior
-samples for each of the 1000 simulated datasets.
+each of the prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, retaining 1000
+posterior samples for each of the 1000 simulated datasets.
 
@@
 We also used the mode estimate of the preferred model for each replicate to
-estimate the number of true parameter values excluded, which produced very
-similar results.
+estimate the number of true parameter values excluded, which produced nearly
+identical results.
 Importantly, the posterior probability of excluding at least one true parameter
@@
 weight yet low likelihood.
-
-Given the theoretical and practical issues with empirical Bayes approaches to
-Bayesian model choice, it is clear why one should use caution before overly
-criticizing an investigator's choice of priors \emph{after} having seen the
-resulting posterior.
-As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a
-large amount of uncertainty regarding the divergence times of the 22 population
-pairs under study.
-Two of these pairs represent distinct species, and the taxonomy of many groups
-in the Philippines has repeatedly been shown to mask deeply divergent lineages
-\citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
-Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013,Rafe2013AREES,Siler2014kikuchii}.
+\change{
 When limited to uniformly distributed priors, the alternative to priors that
-reflect prior uncertainty, as shown above, is to risk excluding the true values
-one seeks to estimate.
+reflect prior uncertainty is to risk excluding the true values one seeks to
+estimate, as shown above.
 However, continuous distributions more appropriate as priors for positive
@@
 \citep{Oaks2014dpp}.
+}
+
+% Given the theoretical and practical issues with empirical Bayes approaches to
+% Bayesian model choice, it is clear why one should use caution before overly
+% criticizing an investigator's choice of priors \emph{after} having seen the
+% resulting posterior.
+% As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a
+% large amount of uncertainty regarding the divergence times of the 22 population
+% pairs under study.
+% Two of these pairs represent distinct species, and the taxonomy of many groups
+% in the Philippines has repeatedly been shown to mask deeply divergent lineages
+% \citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
+% Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013,Rafe2013AREES,Siler2014kikuchii}.
 
@@
 The method is most likely to infer the extreme case of a single divergence
-event shared across 22 taxa when populations diverged randomly over the past
-\globalcoalunit generations.
+event shared across all 22 taxa when populations diverged randomly over the
+past \globalcoalunit generations.
 When divergences are random over the past $8\globalpopsize$ generations, the
@@
 million generations.
-Given the empirically Bayesian model-averaging approach of \citet{Hickerson2013}
-lacks power to detect random variation in divergence times over such
-timescales, it is difficult to justify the risk of the approach to exclude
-regions of parameter space containing the truth.
+% Given the empirically Bayesian model-averaging approach of \citet{Hickerson2013}
+% lacks power to detect random variation in divergence times over such
+% timescales, it is difficult to justify the risk of the approach to exclude
+% regions of parameter space containing the truth.
 
@@
 (Figure~\ref{figPowerExclusion4}C).
-Again, this demonstrates the method's sensitivity to priors.
+% Again, this demonstrates the method's sensitivity to priors.
 
@@
 of generations will add much insight about the evolutionary history of the
-taxa; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$),
-this translates to a temporal resolution of 3 million generations or more.
+taxa.
+% ; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), this
+% translates to a temporal resolution of 3 million generations or more.
 
@@
 with which \msb is faced.
-To get a sense of the difficulty of this problem, we tally up all the free
-parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
+% To get a sense of the difficulty of this problem, we tally up all the free
+% parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
 Under the simplest model in \msb (i.e., assuming no migration and no
@@
 \citet{Oaks2012} with 22 taxon pairs, there are 581--602 free parameters
-(depending on the number of divergence-time parameters in \divtvector).
-Furthermore, under this rich model, the method is estimating the probability of
-1002 divergence models \citep[i.e., the number of integer partitions of
-$Y=22$;][]{Oaks2012}.
+that model highly stochastic coalescent and mutational processes.
+Furthermore, under this rich stochastic model, the method is estimating the
+probability of 1002 divergence models \citep[i.e., the number of integer
+partitions of $Y=22$;][]{Oaks2012}.
 
@@
 \section{Conclusions}
+% We demonstrate how the approximate Bayesian model-choice method implemented in
+% \msb can be strongly biased away from models with greater parameter space.
 We demonstrate how the approximate Bayesian model-choice method implemented in
-\msb can be strongly biased away from models with greater parameter space.
+\msb can be spuriously support models with less parameter space.
 As suggested by \citet{Oaks2012}, this is likely caused by the use of uniform
 priors on most of the model's parameters.
-Uniform distributions necessitate the use of broad priors that place high prior
-density in unlikely regions of parameter space, less the risk of excluding the
-truth \emph{a priori}.
+Uniform distributions necessitate the use of priors that place high density in
+unlikely regions of parameter space, less the risk of excluding the truth
+\emph{a priori}.
 These broad priors reduce the marginal likelihoods of models with more
@@
 of the parameters.
+Also, our results show that it is difficult to choose an uniformly distributed
+prior on divergence times that is broad enough to confidently contain the true
+values of parameters while being narrow enough to avoid bias toward models with
+less parameter space.
 
-Whether or not one chooses to use empirically informed priors, our results
-suggest that it is difficult to choose an uniformly distributed prior on
-divergence times that is broad enough to confidently contain the true values of
-parameters while being narrow enough to avoid bias toward models with less
-parameter space.
-While we do not assume ``that all previous \msb results are invalid,'' as
-suggested by \citet{Hickerson2013}, we do conservatively recommend that the
-common inference of temporally clustered divergences \citep{Barber2010,
+\change{
+% Whether or not one chooses to use empirically informed priors,
+% We do not assume results of all previous applications of \msb results are invalid, as
+% suggested by \citet{Hickerson2013}.
+% We do conservatively recommend that
+The common inference of temporally clustered divergences \citep{Barber2010,
     Bell2012, Carnaval2009, Chan2011, Daza2010, Hickerson2006, Huang2011,
@@
 accompanied with the necessary analyses to assess the robustness and temporal
-resolution of such results, should be treated with caution, because the method
-has been shown to spuriously infer clustered divergences over a range of prior
+resolution of such results, should be treated with caution, because \msb has
+been shown to spuriously infer clustered divergences over a range of prior
 conditions.
+}
 Fortunately, alternative prior probability distributions allow prior

### 2014-07-10 manuscripts/reply/reply-abstract.tex
Working on overhaul of intro for reply ms.

@@
 According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
-was due to excessively broad priors on divergence times, leading to 
+was due to excessively broad priors on divergence times leading to 
 inadequate numbers of simulation replicates.

### 2014-07-10 manuscripts/reply/reply-main.tex
Working on overhaul of intro for reply ms.

@@
 \section{Introduction}
+Biogeographers frequently seek to explain population and species
+differentiation on geographical phenomena.
+Establishing that a set of population splitting events occurred
+at the same time can be a potentially persuasive argument that a set of taxa
+were affected by the same geographic events.
+% Unfortunately, estimating divergence times precisely when one lacks
+% precise information about the rate of molecular evolution is a notoriously
+% difficult problem in evolutionary biology.
+\citet{Huang2011} introduced an approximate Bayesian approach (implemented
+in the software \msb) which can be used to estimate the probabilities of
+models in which multiple sets of taxa diverge at the same time. 
+% This method  does not require a common set of character data that links all
+% of the taxa under investigation, but does require the relative mutation
+% rates among the taxa and loci to be assumed \emph{a priori}.
 
-\msb \citep{Huang2011} uses approximate Bayesian computation (ABC) to estimate
-the distribution of divergence times among co-distributed pairs of taxa.
-It approximates a posterior probability over models that range from a single
-divergence-time parameter (i.e., simultaneous divergence of all pairs of taxa)
-to the fully generalized model in which each pair of taxa diverged at a
-unique time.
-% A full description of the model is given in \citet{Oaks2012}.
-% Frequently, the exact values of the divergence times are hard 
-% to determine because researchers typically lack precise knowledge 
-% of the rate at which substitutions occur.
-% The exact divergence time is therefore often viewed as a nuisance parameter,
-% and researchers focus on the number of divergence events.
-The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
-on the fundamental methodological point about the model selection performed in \msb:
+Recently, \citet{Oaks2012} used this model-choice framework to 
+study 22 pairs of vertebrate lineages which are distributed across
+the Philippines; they also studied the behavior of the \msb approach
+using computer simulations.
+They found the method was very sensitive to prior assumptions and suggested
+this was due to the fact that the marginal likelihoods of models are sensitive
+to the use of uniform priors on divergence times.
+In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
+analyses was caused by sampling error due to the inefficiency of the method's
+rejection algorithm.
+They proposed a model-averaging approach that uses narrow, empirically
+informed uniform priors.
+
+Whereas the papers of \citet{Oaks2012} and \citet{Hickerson2013} differ about
+the cause and solution, they agree on the fundamental methodological point
+about the model selection performed in \msb:
+% The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
+% on the fundamental methodological point about the model selection performed in \msb:
 \begin{itemize}
@@
 \end{itemize}
-
-This is not surprising given the rich statistical literature that shows that
+This is not surprising given the rich statistical literature showing that
 marginal likelihoods are very sensitive to the priors used in Bayesian model
@@
 \citep[e.g.,][]{Jeffreys1939,Lindley1957}.
-Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
-for models with few divergence parameters is the greater marginal likelihoods
-of these models under vague uniform priors, relative to more parameter-rich
-models.
-Models with more divergence-time parameters integrate over \emph{much} greater
-parameter space with low probability of producing the data, yet relatively high
-prior density imposed by the uniformly distributed prior on divergence times.
-\change{
+% We know that under broad uniform priors on divergence times, models with few
+% divergence parameters will have greater marginal likelihoods relative to more
+% parameter-rich models.
+% Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
+% for models with few divergence parameters is the greater marginal likelihoods
+% of these models under vague uniform priors, relative to more parameter-rich
+% models.
+Under vague uniform priors on divergence times, models with more
+divergence-time parameters integrate over \emph{much} greater parameter space
+with low probability of producing the data, yet relatively high prior density
+imposed by the uniformly distributed prior.
 % By using such priors on divergence times, we are essentially telling the
@@
 % explain the data very well.
-This affect of the priors on marginal likelihoods creates a penalty for
-over-parameterization in Bayesian model choice.
-However, for this inference problem, the uniform priors on divergence times can
-create a strong preference for simple models that are often of interest to
-biogeographers even if these models do not explain the data very well.
-This phenomenon is extrinsic to the numerical approximation machinery of \msb;
-with such prior assumptions, the true posterior will favor models with
-few divergence events.
-However, such results, even if correct from the perspective of Bayesian model
-choice, do not provide the biogeographical insights that a researcher who
-employs \msb seeks to gain.
-For example, a biogeographer would like to interpret the posterior probability
-of a simultaneous divergence model as an estimate of the probability that this
-model is correct.
-However, when such prior assumptions are made about divergence times, the
-estimated posterior probability is better interpretted as the probability that
-you cannot completely rule out that the simultaneous-divergence model can
-explain the data.
+Thus, uniform priors on divergence times can impose a strong ``penalty'' for
+divergence-time parameters, creating a strong preference for simple
+models that are often of interest to biogeographers, even if these models do
+not explain the data very well.
+% It is important to note that this is not due to estimation error
+% (i.e., this phenomenon is extrinsic to the numerical approximation machinery of
+% \msb).
+% Rather, the \emph{true} posterior will tend to support models with few
+% divergence-time parameters under broad uniform priors on divergence times.
+Such posterior support for simultaneous divergence, even if correct from the
+perspective of Bayesian model choice, does not provide the biogeographical
+insights that a researcher who employs \msb seeks to gain.
+% Ideally, the estimated posterior probability of a simultaneous-divergence model
+% can be interpreted as an estimate of the probability that this model is correct.
+% However, when using uniform priors on divergence times, it might be better
+% interpreted as an estimate of the probability that the simultaneous-divergence
+% model cannot be completely ruled out as a possible explanation of the data.
 % Note, this is a statistical issue, extrinsic to \msb, and does not question the
 % soundness of the numerical approximation machinery of \msb.
-}
 
-\citet{Hickerson2013} suggest the problem is the result of a numerical issue
-intrinsic to \msb, concluding the method's rejection algorithm is inefficient
-and will be biased toward models with fewer divergence-time parameters, because
-the parameter space of these models are more densely sampled relative to models
-with more divergence-time parameters.
-We agree that sampling error is present in all numerical Bayesian estimation
-methods, however, we show here that it is not a major contributer to the
-biases found by \citet{Oaks2012}.
-We present results here of analyses that, when combined with the results
-of \citet{Oaks2012}, strongly suggest that \msb prefers models with few
-divergence-time parameters, because they have greater likelihoods when
-averaged over broad uniform priors on divergences times.
+% Alternatively, \citet{Hickerson2013} suggest the problem is the result of a
+% numerical issue intrinsic to \msb, concluding the method's rejection algorithm
+% is inefficient and will be biased toward models with fewer divergence-time
+% parameters, because the parameter space of these models are more densely
+% sampled relative to models with more divergence-time parameters.
+% We agree that sampling error is present in all numerical Bayesian estimation
+% methods, however, we show here that it is not a major contributer to the
+% biases found by \citet{Oaks2012}.
+% We present results here of analyses that, when combined with the results
+% of \citet{Oaks2012}, strongly suggest that \msb prefers models with few
+% divergence-time parameters, because they have greater likelihoods when
+% averaged over broad uniform priors on divergences times.
 
-\citet{Oaks2012} suggest alternative prior probability distributions on
-divergence-times could increase the marginal likelihoods of models with more
-divergence-time parameters and thus reduce spurious support for models of
-temporally clustered divergences.
-Alternatively, \citet{Hickerson2013} present an approach that uses empirically
-informed uniform priors and Bayesian model-averaging in an attempt to accommodate
-uncertainty in selecting priors.
-In general, we agree with the use of Bayesian model averaging to obtain a
-posterior estimate marginalized over uncertainty in prior choice.
-However, we question whether adding an additional dimension of model choice
-that samples over empirically informed uniform priors is a fruitful solution
-for a model-choice method that is highly sensitive to priors.
+% \citet{Oaks2012} suggest alternative prior probability distributions on
+% divergence-times could increase the marginal likelihoods of models with more
+% divergence-time parameters and thus reduce spurious support for models of
+% temporally clustered divergences.
+% Alternatively, \citet{Hickerson2013} present an approach that uses empirically
+% informed uniform priors and Bayesian model-averaging in an attempt to accommodate
+% uncertainty in selecting priors.
+% In general, we agree with the use of Bayesian model averaging to obtain a
+% posterior estimate marginalized over uncertainty in prior choice.
+% However, we question whether adding an additional dimension of model choice
+% that samples over empirically informed uniform priors is a fruitful solution
+% for a model-choice method that is highly sensitive to priors.
 
 \change{
-In this paper, we discuss theoretical and practical considerations of using
-empirically informed priors for Bayesian model choice.
+Here, we discuss theoretical considerations for using empirically informed
+priors for Bayesian model choice.
 }
-Furthermore, we evaluate the empirical Bayesian model-averaging approach of
+We then evaluate the empirical Bayesian model-averaging approach of
 \citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
@@
 for details).
-Here, we avoid this error, but still find their approach to be biased toward
+We avoid this error, but still find their approach to be biased toward
 clustered divergence models.
@@
 }
-However, recent work shows more flexible probability distributions without a
-hard upper bound (e.g., gamma) can accommodate prior uncertainty in divergence
-times without inhibiting the marginal likelihoods of models with more
-divergence-time parameters, and as a result, increase the method's power to
-detect temporal variation in divergences
-\citep{Oaks2014dpp}.
+We also explore the predictions of the alternative causes of this problem
+proposed by \citet{Oaks2012} and \citet{Hickerson2013}.
+We find the prior sensitivity of the marginal likelihoods is a likely
+culprit,
+and show that sampling error is not a major contributer to the biases found
+by \citet{Oaks2012}.
+Fortunately, \citet{Oaks2014dpp} has introduced a new method that uses more
+flexible probability distributions to accommodate prior uncertainty in
+divergence times without inhibiting the marginal likelihoods of models with
+more divergence-time parameters, which increases the method's robustness and
+power to detect temporal variation in divergences \citep{Oaks2014dpp}.
+
+% \msb \citep{Huang2011} uses approximate Bayesian computation (ABC) to estimate
+% the distribution of divergence times among co-distributed pairs of taxa.
+% It approximates a posterior probability over models that range from a single
+% divergence-time parameter (i.e., simultaneous divergence of all pairs of taxa)
+% to the fully generalized model in which each pair of taxa diverged at a
+% unique time.
+% % A full description of the model is given in \citet{Oaks2012}.
+% % Frequently, the exact values of the divergence times are hard 
+% % to determine because researchers typically lack precise knowledge 
+% % of the rate at which substitutions occur.
+% % The exact divergence time is therefore often viewed as a nuisance parameter,
+% % and researchers focus on the number of divergence events.
+% The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
+% on the fundamental methodological point about the model selection performed in \msb:
+% \begin{itemize}
+%    \item The use of vague priors on the divergence-time parameters can lead to
+%        support for models with few divergence events shared across taxa. Thus,
+%        the primary inference enabled by the approach is very sensitive to the
+%        priors on divergence times.
+% \end{itemize}
+% This is not surprising given the rich statistical literature that shows that
+% marginal likelihoods are very sensitive to the priors used in Bayesian model
+% selection
+% \citep[e.g.,][]{Jeffreys1939,Lindley1957}.
+% Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
+% for models with few divergence parameters is the greater marginal likelihoods
+% of these models under vague uniform priors, relative to more parameter-rich
+% models.
+
+% \change{
+% Models with more divergence-time parameters integrate over \emph{much} greater
+% parameter space with low probability of producing the data, yet relatively high
+% prior density imposed by the uniformly distributed prior.
+% % By using such priors on divergence times, we are essentially telling the
+% % method to favor models with few divergence times even if they do not
+% % explain the data very well.
+% Thus, uniform priors on divergence times can impose a strong ``penalty'' for
+% divergence-time parameters, creating a strong preference for simple
+% models that are often of interest to biogeographers, even if these models do
+% not explain the data very well.
+% It is important to note that this is not due to estimation error
+% (i.e., this phenomenon is extrinsic to the numerical approximation machinery of
+% \msb).
+% Rather, the \emph{true} posterior will tend to support models with few
+% divergence-time parameters under broad uniform priors on divergence times.
+% Such posterior support for simultaneous divergence, even if correct from the
+% perspective of Bayesian model choice, does not provide the biogeographical
+% insights that a researcher who employs \msb seeks to gain.
+% % Ideally, the estimated posterior probability of a simultaneous-divergence model
+% % can be interpreted as an estimate of the probability that this model is correct.
+% % However, when using uniform priors on divergence times, it might be better
+% % interpreted as an estimate of the probability that the simultaneous-divergence
+% % model cannot be completely ruled out as a possible explanation of the data.
+% % Note, this is a statistical issue, extrinsic to \msb, and does not question the
+% % soundness of the numerical approximation machinery of \msb.
+% }
+
+% \citet{Hickerson2013} suggest the problem is the result of a numerical issue
+% intrinsic to \msb, concluding the method's rejection algorithm is inefficient
+% and will be biased toward models with fewer divergence-time parameters, because
+% the parameter space of these models are more densely sampled relative to models
+% with more divergence-time parameters.
+% We agree that sampling error is present in all numerical Bayesian estimation
+% methods, however, we show here that it is not a major contributer to the
+% biases found by \citet{Oaks2012}.
+% % We present results here of analyses that, when combined with the results
+% % of \citet{Oaks2012}, strongly suggest that \msb prefers models with few
+% % divergence-time parameters, because they have greater likelihoods when
+% % averaged over broad uniform priors on divergences times.
+
+% \citet{Oaks2012} suggest alternative prior probability distributions on
+% divergence-times could increase the marginal likelihoods of models with more
+% divergence-time parameters and thus reduce spurious support for models of
+% temporally clustered divergences.
+% Alternatively, \citet{Hickerson2013} present an approach that uses empirically
+% informed uniform priors and Bayesian model-averaging in an attempt to accommodate
+% uncertainty in selecting priors.
+% In general, we agree with the use of Bayesian model averaging to obtain a
+% posterior estimate marginalized over uncertainty in prior choice.
+% However, we question whether adding an additional dimension of model choice
+% that samples over empirically informed uniform priors is a fruitful solution
+% for a model-choice method that is highly sensitive to priors.
+
+% \change{
+% In this paper, we discuss theoretical and practical considerations of using
+% empirically informed priors for Bayesian model choice.
+% }
+% Furthermore, we evaluate the empirical Bayesian model-averaging approach of
+% \citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
+% \citet{Oaks2012}.
+% In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
+% made an error by mixing different units of time, which makes the results
+% presented in their response difficult to interpret (see Supporting Information
+% for details).
+% Here, we avoid this error, but still find their approach to be biased toward
+% clustered divergence models.
+% Furthermore, the approach provides another means by which the method can
+% ``escape'' models with large parameter space, which manifests in the preference
+% of models that exclude from consideration the true values of the model's
+% parameters.
+% \change{
+% Our results, show that it is difficult to choose a uniformly distributed prior
+% on divergence times that is broad enough to confidently contain the true values
+% of parameters while being narrow enough to avoid spurious support of models
+% with less parameter space.
+% }
+% However, recent work shows more flexible probability distributions without a
+% hard upper bound (e.g., gamma) can accommodate prior uncertainty in divergence
+% times without inhibiting the marginal likelihoods of models with more
+% divergence-time parameters, and as a result, increase the method's power to
+% detect temporal variation in divergences
+% \citep{Oaks2014dpp}.
 
@@
 \change{
-The results of \citet{Oaks2012} and \citet{Hickerson2013} show the method is
-highly sensitive to the prior on divergence times.
-Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
-divergences when using a prior on divergence times that is strongly
+% The results of \citet{Oaks2012} and \citet{Hickerson2013} show the method is
+% highly sensitive to the prior on divergence times.
+\citet{Oaks2012} found that \msb would incorrectly support models of temporally
+clustered divergences even when using a prior on divergence times that was
 informed by the data being analyzed.
@@
 % density in improbable regions of parameter space.
-
+%
 % \citet{Hickerson2013} suggest that \citet{Oaks2012} should have used their
@@
 % parameter space.
-The results of Hickerson et al.'s \citeyear{Hickerson2013} reanalysis of the
+The results of Hickerson et al.'s (\citeyear{Hickerson2013}) reanalysis of the
 Philippine dataset strongly favored models with the narrowest, empirically
@@
 \change{
-Hickerson et al.'s \citeyear{Hickerson2013} results that strongly prefer the
+Hickerson et al.'s (\citeyear{Hickerson2013}) results that strongly prefer the
 narrowest priors on divergence times (their $M_1$ and $M_2$) suggest their
@@
 \subsubsection{Simulation-based assessment of Hickerson et al.'s
-\citeyear{Hickerson2013} model averaging over empirical priors}
+    (\citeyear{Hickerson2013}) model averaging over empirical priors}
 
@@
 Even if we assume the gene divergence times are estimated without error,
-Hickerson et al.'s \citeyear{Hickerson2013} argument only applies to one of the
-three different priors used by \citet{Oaks2012}.
+Hickerson et al.'s (\citeyear{Hickerson2013}) argument only applies to one of
+the three different priors used by \citet{Oaks2012}.
 The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5)$)

### 2014-07-18 manuscripts/reply/reply-main.tex
More work on intro of reply ms.

@@
 They found the method was very sensitive to prior assumptions and suggested
-this was due to the fact that the marginal likelihoods of models are sensitive
-to the use of uniform priors on divergence times.
+this was due to the fact that the marginal likelihoods of divergence models are
+sensitive to the use of uniform priors on divergence times.
+\citet{Oaks2012} suggested that alternative probability distributions (e.g.,
+gamma) could better represent prior uncertainty about divergence times while
+mitigating the disparity of marginal likelihoods across divergence models.
 In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
@@
 rejection algorithm.
-They proposed a model-averaging approach that uses narrow, empirically
-informed uniform priors.
+As a potential solution, they proposed a model-averaging approach that uses
+narrow, empirically informed uniform priors.
 
@@
 proposed by \citet{Oaks2012} and \citet{Hickerson2013}.
-We find the prior sensitivity of the marginal likelihoods is a likely
-culprit,
-and show that sampling error is not a major contributer to the biases found
-by \citet{Oaks2012}.
-Fortunately, \citet{Oaks2014dpp} has introduced a new method that uses more
+Our finings suggest the primary cause is the disparity in marginal likelihoods
+across possible divergence models imposed by the uniform priors on divergence
+times.
+Furthermore, we show that sampling error is not a major contributer to the
+biases found by \citet{Oaks2012}.
+Fortunately, \citet{Oaks2014dpp} has introduced a method that uses more
 flexible probability distributions to accommodate prior uncertainty in

### 2014-07-19 manuscripts/reply/reply-main.tex
More editing of main text of reply ms.

@@
 % difficult problem in evolutionary biology.
-\citet{Huang2011} introduced an approximate Bayesian approach (implemented
-in the software \msb) which can be used to estimate the probabilities of
-models in which multiple sets of taxa diverge at the same time. 
+\citet{Huang2011} introduced an approximate Bayesian approach (implemented in
+the software \msb) to estimating the probabilities of models in which multiple
+sets of taxa diverge at the same time. 
 % This method  does not require a common set of character data that links all
@@
 
-Recently, \citet{Oaks2012} used this model-choice framework to 
-study 22 pairs of vertebrate lineages which are distributed across
-the Philippines; they also studied the behavior of the \msb approach
-using computer simulations.
+Recently, \citet{Oaks2012} used this model-choice framework to study 22 pairs
+of vertebrate lineages distributed across the Philippines; they also studied
+the behavior of the \msb approach using computer simulations.
 They found the method was very sensitive to prior assumptions and suggested
@@
 sensitive to the use of uniform priors on divergence times.
-\citet{Oaks2012} suggested that alternative probability distributions (e.g.,
-gamma) could better represent prior uncertainty about divergence times while
-mitigating the disparity of marginal likelihoods across divergence models.
+They also suggested alternative probability distributions (e.g., gamma) could
+better represent prior uncertainty about divergence times while mitigating the
+disparity of marginal likelihoods across divergence models.
 In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
@@
 for details).
-We avoid this error, but still find their approach to be biased toward
-clustered divergence models.
+We avoid this error, but still find their approach will often spuriously
+support clustered divergence models.
 Furthermore, the approach provides another means by which the method can
@@
 }
-We also explore the predictions of the alternative causes of this problem
-proposed by \citet{Oaks2012} and \citet{Hickerson2013}.
-Our finings suggest the primary cause is the disparity in marginal likelihoods
-across possible divergence models imposed by the uniform priors on divergence
-times.
-Furthermore, we show that sampling error is not a major contributer to the
-biases found by \citet{Oaks2012}.
+We also explore the predictions of the causes of the bias proposed by
+\citet{Oaks2012} and \citet{Hickerson2013}.
+We show the primary cause of the method's tendency to incorrectly infer shared
+divergences is the disparity in marginal likelihoods across divergence models
+imposed by the uniform priors on divergence times.
+We also show that sampling error is not a major contributer to the problem.
 Fortunately, \citet{Oaks2014dpp} has introduced a method that uses more
@@
 more divergence-time parameters, which increases the method's robustness and
-power to detect temporal variation in divergences \citep{Oaks2014dpp}.
+power to detect temporal variation in divergences.
 
@@
 % highly sensitive to the prior on divergence times.
-\citet{Oaks2012} found that \msb would incorrectly support models of temporally
+\citet{Oaks2012} found \msb incorrectly supports models of temporally
 clustered divergences even when using a prior on divergence times that was
@@
 Empirical Bayes is a well-studied branch of Bayesian statistics that has given
-rise to many inference methods that provide means of parameter estimation that
-often exhibit favorable frequentist properties.
+rise to many inference methods for obtaining parameter estimates that often
+exhibit favorable frequentist properties.
 % Furthermore, many post-hoc correction methods have been developed for
@@
 \begin{linenomath}
-Whereas empirical Bayes approaches can provide powerful methods for parameter
-estimation, a theoretical justification for empirical Bayes approaches to model
-choice is questionable.
+Whereas empirical Bayesian approaches can provide powerful methods for
+parameter estimation, a theoretical justification for empirical Bayesian
+approaches to model choice is questionable.
 In Bayesian model choice, the primary goal is not to estimate parameters, but
-to estimate the relative probabilities of candidate models.
+to estimate the probabilities of candidate models.
 In a simple example where two candidate models, $M_1$ and $M_2$, are being
@@
 due to using the data to inform the priors.
-By giving the data more weight relative to the prior, the goal is that the
-posterior distribution will be peaked near the true value(s) of the model's
-parameter(s).
+By giving the data more weight relative to the prior, the goal of empirical
+Bayesian parameter estimation is to estimate a distribution that will be peaked
+near the true value(s) of the model's parameter(s).
 % This is why empirical Bayes can work well for estimating the values of model
@@
 updates the prior to yield the model's posterior probability.
-Thus the integral over the entire parameter space of the likelihood weighted by
-the prior probability density is no longer a normalizing constant, but rather
-is what informs the posterior probability of that model from the data.
+The integral over the entire parameter space of the likelihood weighted by the
+prior probability density is no longer a normalizing constant, but rather is
+what informs the posterior probability of that model from the data.
 As a result, Bayesian model choice tends to be much more sensitive to priors
@@
 Philippine dataset strongly favored models with the narrowest, empirically
-informed prior on divergence time, and thus their model-averaged posterior
+informed prior on divergence times, and thus their model-averaged posterior
 estimates are dominated by models $M_1$ and $M_2$ (see Table 1 of
@@
 \change{
-Hickerson et al.'s (\citeyear{Hickerson2013}) results that strongly prefer the
-narrowest priors on divergence times (their $M_1$ and $M_2$) suggest their
-approach is biased toward models with less parameter space, and as a
-consequence, will estimate model-averaged posteriors dominated by models that
-exclude true values of model's parameters.
+The strong preference for the narrowest priors on divergence times used by
+\citet{Hickerson2013} suggest their approach is biased toward models with less
+parameter space and, as a consequence, will estimate model-averaged posteriors
+dominated by models that exclude true values of the model's parameters.
 }
@@
 approach of \citet{Hickerson2013}, but set one of the prior models with a
-uniform prior on divergence times that is unrealistically narrow, and almost
+uniform prior on divergence times that is unrealistically narrow and almost
 certainly excludes most, if not all, of the true divergence times of the 22
@@
 
-The results of the preceding empirical and simulation-based analyses clearly
-demonstrate the risk of using narrow, empirically guided uniform priors in a
-Bayesian model-averaging framework.
+The results of the above empirical and simulation analyses clearly demonstrate
+the risk of using narrow, empirically guided uniform priors in a Bayesian
+model-averaging framework.
 The consequence of this approach is obtaining a model-averaged posterior
@@
 values of model parameters.
-
-\subsection{Additional thoughts on empirical priors in Bayesian model choice}
-The preceding sections are not a general critique of Bayesian model averaging.
+% \subsection{Additional thoughts on empirical priors in Bayesian model choice}
+This is not a general critique of Bayesian model averaging.
 Rather, model averaging can provide an elegant way of incorporating
@@
 consideration the true values of parameters due to the larger marginal
-likelihoods of the models that are integrated over less space with high prior
+likelihoods of the models that integrate over less space with high prior
 weight yet low likelihood.
+
 \change{
-When limited to uniformly distributed priors, the alternative to priors that
-reflect prior uncertainty is to risk excluding the true values one seeks to
-estimate, as shown above.
-However, continuous distributions more appropriate as priors for positive
+When using uniformly distributed priors, the alternative to priors that reflect
+prior uncertainty is to risk excluding the true values one seeks to estimate,
+as shown above.
+Fortunately, continuous distributions more appropriate as priors for positive
 real-valued parameters have been shown to greatly reduce spurious support for
@@
 Bayesian model-choice approach used by \citet{Hickerson2013}, one could justify
-such risk if the method does indeed increase the power of the method and thus
-decrease bias toward clustered divergence models.
+such risk if the approach does indeed increase power to detect temporal
+variation in divergences and thus decrease bias toward clustered divergence
+models.
 We assess this possibility using simulation-based analyses.
@@
 Our results show that the empirical Bayesian model-averaging approach leads to
-spuriously strong support for the extreme case of a single divergence event
-when populations diverged randomly over the last $8\globalpopsize$ generations.
+erroneous support for highly clustered divergences when populations diverge
+randomly over the last $8\globalpopsize$ generations.
 To put this on the scale roughly consistent with a vertebrate mitochondrial
-locus, assuming a mutation rate of $2\e{-8}$ per site per generation, this
-translates to 5 million generations.
+locus, assuming a mutation rate of $1\e{-8}$ per site per generation, this
+translates to 10 million generations.
 Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
@@
 parameter values.
-Across all but one of the \divt{max} we simulated, in a large proportion of
-replicates the method favors a model that excludes the truth, and across many
-of the \divt{max} the preferred model will exclude a large proportion of the
-true divergence-time values (Figure~\ref{figPowerExclusion4}A--D \&
+Across all but one of the \divt{max} we simulated, the method favors a model
+that excludes the truth in a large proportion of replicates, and across many of
+the \divt{max} the preferred model will exclude a large proportion of the true
+divergence-time values (Figure~\ref{figPowerExclusion4}A--D \&
 S\ref{figPowerExclusion6}A--F).
@@
 \citet{Oaks2012} performed similar power analyses under three uniform
-divergence-time priors as narrow as 0--5 coalescent units, and found the
-method was able to detect multiple events among 22 when divergences were random
+divergence-time priors as narrow as 0--5 coalescent units, and found the method
+was able to detect multiple events among 22 taxa when divergences were random
 over millions of generations.
@@
 To determine how best to improve the behavior of \msb, it is important to
-determine the mechanism by which broad uniform priors cause spurious support
+determine the mechanism by which broad uniform priors cause incorrect support
 for clustered models of divergence.
@@
 Hickerson et al.'s (\citeyear{Hickerson2013}) argument only applies to one of
-the three different priors used by \citet{Oaks2012}.
+the three priors used by \citet{Oaks2012}.
 The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5)$)
@@
 To disentangle whether model likelihoods or insufficient prior sampling is to
-blame for the method's spurious support for simple models, we must look at the
+blame for the method's erroneous support for simple models, we must look at the
 different predictions made by these two phenomena.
@@
 approximation method, insufficient sampling is an unlikely explanation for the
-support of prior models with less parameter space.
+erroneous support of models with less parameter space.
 While analyses that sample each model proportional to their parameter space
@@
 % parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
-Under the simplest model in \msb (i.e., assuming no migration and no
-intra-locus recombination), the number of parameters for each taxon pair
-include:
-The population sizes of the ancestral and descendant populations
-(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
-the magnitude of population contraction in each of the descendant
-populations (\bottleScalar{1}{} and \bottleScalar{2}{}), the
-timing of these contractions (\bottleTime{}), and the $N-1$ node heights
-(coalescent times) of the gene tree that gave rise to the $N$ gene
-copies sampled from both populations.
-Lastly, there are between one and 22 divergence time parameters \divt{} in
-the vector \divtvector.
-Overall, when applying the simplest model in \msb to the dataset of
-\citet{Oaks2012} with 22 taxon pairs, there are 581--602 free parameters
-that model highly stochastic coalescent and mutational processes.
+% Under the simplest model in \msb (i.e., assuming no migration and no
+% intra-locus recombination), the number of parameters for each taxon pair
+% include:
+% The population sizes of the ancestral and descendant populations
+% (\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
+% the magnitude of population contraction in each of the descendant
+% populations (\bottleScalar{1}{} and \bottleScalar{2}{}), the
+% timing of these contractions (\bottleTime{}), and the $N-1$ node heights
+% (coalescent times) of the gene tree that gave rise to the $N$ gene
+% copies sampled from both populations.
+% Lastly, there are between one and 22 divergence time parameters \divt{} in
+% the vector \divtvector.
+When applying the simplest model in \msb (i.e., assuming no migration and no
+intra-locus recombination) to the dataset of \citet{Oaks2012} with 22 taxon
+pairs, there are 581--602 free parameters that model highly stochastic
+coalescent and mutational processes.
 Furthermore, under this rich stochastic model, the method is estimating the
@@
 partitions of $Y=22$;][]{Oaks2012}.
-
-This is a very difficult inference problem, especially considering that all the
-information in the sequence alignment of each taxon pair is distilled into four
-summary statistics:
+Furthermore, all the information in the sequence alignment of each taxon pair
+is distilled into four summary statistics:
 $\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
@@
 The model-averaging approach of \citet{Hickerson2013} adds an additional
-dimension of model choice to the model.
-They expand the model to sample over eight prior models.
+dimension of model choice to sample over eight prior models.
 This adds an additional free parameter to the model and, more importantly,
-forces the model to sample over 8016 unique models.
+forces the model to sample over 8016 unique models for 22 taxa.
 While this approach is trivial to implement, it is a non-trivial extension of
@@
 density.
-Most importantly, we have found the use of more flexible distributions in place
-of uniform priors improves the power of the method to detect temporal variation
-in divergences and reduces the bias toward models of clustered divergences
-\citep{Oaks2014dpp}.
+Indeed, \citet{Oaks2014dpp} has shown that the use of more flexible
+distributions in place of uniform priors improves the power of the method to
+detect temporal variation in divergences and reduces the bias toward models of
+clustered divergences.
 
@@
 We demonstrate how the approximate Bayesian model-choice method implemented in
-\msb can be spuriously support models with less parameter space.
+\msb can spuriously support models with less parameter space.
 As suggested by \citet{Oaks2012}, this is likely caused by the use of uniform
-priors on most of the model's parameters.
+priors on divergence times.
 Uniform distributions necessitate the use of priors that place high density in
-unlikely regions of parameter space, less the risk of excluding the truth
-\emph{a priori}.
-These broad priors reduce the marginal likelihoods of models with more
-divergence-time parameters or broader prior distributions on those parameters.
+unlikely regions of parameter space, less the risk of excluding the true
+divergence times \emph{a priori}.
+These broad uniform priors reduce the marginal likelihoods of models with more
+divergence-time parameters.
 We show that the empirical Bayesian model-averaging approach of
@@
 resolution of such results, should be treated with caution, because \msb has
-been shown to spuriously infer clustered divergences over a range of prior
+been shown to erroneously infer clustered divergences over a range of prior
 conditions.
 }
-Fortunately, alternative prior probability distributions allow prior
-uncertainty to be accommodated while avoiding excessive prior density in
-regions of low likelihood, which greatly improves the power of the method
-\citep{Oaks2014dpp}.
+Fortunately, \citet{Oaks2014dpp} has shown that alternative probability
+distributions allow prior uncertainty to be accommodated while avoiding
+excessive prior density in regions of low likelihood, which greatly improves
+the ability to infer shared divergence histories.
 

### 2014-07-19 manuscripts/reply/reply-abstract.tex
Shortening abstract of reply ms.

@@
-Biogeographers frequently seek to explain population and species
-differentiation on geographical phenomena.
-Establishing that a set of population splitting events occurred
-at the same time can be a potentially persuasive argument that a set of taxa
-were affected by the same geographic events.
-Unfortunately, estimating divergence times precisely when one lacks
-precise information about the rate of molecular evolution is a notoriously
-difficult problem in evolutionary biology.
-\citet{Huang2011} introduced an approximate Bayesian approach (implemented
-in the software \msb) which can be used to estimate the probabilities of
-models in which multiple sets of taxa diverge at the same time. 
+% Biogeographers frequently seek to explain population and species
+% differentiation on geographical phenomena.
+Establishing that a set of population splitting events occurred at the same
+time can be a potentially persuasive argument that the populations were
+affected by the same geographic event.
+% Unfortunately, estimating divergence times precisely when one lacks
+% precise information about the rate of molecular evolution is a notoriously
+% difficult problem in evolutionary biology.
+% \citet{Huang2011} introduced an approximate-Bayesian approach, \msb, for
+% estimating the probabilities of models in which multiple sets of taxa diverge
+% at the same time.
 % This method  does not require a common set of character data that links all
@@
 % rates among the taxa and loci to be assumed \emph{a priori}.
-Recently, \citet{Oaks2012} used this model-choice framework to 
-study 22 pairs of vertebrate lineages which are distributed across
-the Philippines; they also studied the behavior of the \msb approach
-using computer simulations.
+Recently, \citet{Oaks2012} used an approximate-Bayesian model choice method
+(\msb) to estimate the support for such a pattern of divergences across 22
+pairs of taxa from the Philippines.
+They also studied the behavior of \msb using computer simulations.
 \citet{Oaks2012} found the method was very sensitive to prior assumptions and
 had low power to detect variation in divergence times.
-These results were not surprising, in light of a rich statistical literature
+These results were not surprising in light of a rich statistical literature
 showing that the marginal likelihood of a model is sensitive to the
@@
 inadequate numbers of simulation replicates.
-They proposed a new model-averaging approach that uses narrow, empirically
+They proposed a model-averaging approach that uses narrow, empirically
 informed uniform priors.

### 2014-07-19 manuscripts/reply/reply-acknowledgments.tex
Updating acknowledgments for reply ms.
Adding editor and reviewers.

@@
 We thank Melissa Callahan, Jake Esselstyn, Cameron Siler, Mark Holder, Rafe
-Brown, Emily McTavish, Daniel Money, Jordan Koch, and Adam Leach\'{e} for
-insightful comments that greatly improved this work.
+Brown, Emily McTavish, Daniel Money, Jordan Koch, Adam Leach\'{e}, Luke Harmon,
+and three anonymous reviewers for insightful comments that greatly improved
+this work.
 We thank Michael Hickerson and co-authors for generously providing their data.

### 2014-07-22 manuscripts/reply/reply-abstract.tex
More editing of reply ms.

@@
 informed uniform priors.
-Here we demonstrate that the approach of \citet{Hickerson2013} is
-dangerous in the sense that the empirically-derived priors often
-exclude from consideration the true values of the models' parameters.
+Here we demonstrate that the approach of \citet{Hickerson2013} will often
+incorrectly support shared divergence events and is dangerous in the sense that
+the empirically-derived priors often exclude from consideration the true values
+of the models' parameters.
 On a more fundamental level, we question the value of adopting an empirical
-Bayesian stance for this model-choice problem, because it can mislead model
-posterior probabilities, which are inherently measures of belief in the
-models after prior knowledge is updated by the data.
-The robust Bayesian approach of conducting analyses under a variety
-of priors can reveal prior sensitivity and communicate which assumptions
-underlie model inference.
+Bayesian stance for this model-choice problem.
+% , because it can mislead model
+% posterior probabilities, which are inherently measures of belief in the models
+% after prior knowledge is updated by the data.
+The robust Bayesian approach of conducting analyses under a variety of priors
+can reveal prior sensitivity and communicate which assumptions underlie model
+inference.
 Furthermore, simulations provide insight into the temporal resolution of the
 method, which helps guide interpretation of results.
+Fortunately, as predicted by fundamental principles of Bayesian model choice,
+more appropriate priors on the model's parameters increase the method's
+robustness and power to detect temporal variation in divergences
+\citep{Oaks2014dpp}.

### 2014-07-22 manuscripts/reply/reply-main.tex
More editing of reply ms.

@@
 }
-We also explore the predictions of the causes of the bias proposed by
+We also explore the predictions of the potential causes of the bias proposed by
 \citet{Oaks2012} and \citet{Hickerson2013}.
@@
 model choice.
-In Equation \ref{eq:bayesrule}, we see that the posterior density of any
-state $\myTheta{}$ of the model, is the prior density updated by the
-probability of the data given the state $\myTheta{}$ (the likelihood of
-$\myTheta{}$).
-The integral over the entire parameter space of the model, which is defined by
-the priors, only appears as a normalizing constant in the denominator.
-Thus, as long as the prior distribution contains the values of $\myTheta{}$
-that maximize the probability of the data (i.e., the maximum likelihood) and
-the data are strongly informative relative to the prior, the values of the
-parameters that maximize the posterior distribution will be relatively robust
-to prior choice, even if the posterior distribution is technically incorrect
-due to using the data to inform the priors.
-By giving the data more weight relative to the prior, the goal of empirical
-Bayesian parameter estimation is to estimate a distribution that will be peaked
-near the true value(s) of the model's parameter(s).
-% This is why empirical Bayes can work well for estimating the values of model
-% parameters.
 \end{linenomath}
 
-% Unlike parameter estimates, posterior probabilities represent a summary of the
-% entire posterior distribution.
-% Thus, given that empirical Bayesian posterior distributions are not accurate,
-% there is no guarantee regarding the accuracy of probabilities summarized from
-% them.
-
+In Equation \ref{eq:bayesrule}, we see that the posterior density of any state
+$\myTheta{}$ of the model, is the prior density updated by the probability of
+the data given $\myTheta{}$ (the likelihood of $\myTheta{}$).
+The integral over the entire parameter space of the model only appears as a
+normalizing constant in the denominator.
+Thus, as long as the prior distribution contains the values of $\myTheta{}$
+that maximize the probability of the data and the data are strongly informative
+relative to the prior, the values of the parameters that maximize the posterior
+distribution will be relatively robust to prior choice, even if the posterior
+is technically incorrect due to using the data to inform the priors.
 However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
@@
 prior probability density is no longer a normalizing constant, but rather is
-what informs the posterior probability of that model from the data.
+what informs the posterior probability of the model from the data.
 As a result, Bayesian model choice tends to be much more sensitive to priors
 than parameter estimation.
-Also, unlike model parameters, the posterior probabilities of candidate
-models have no clear true values.
+Another important difference of Bayesian model choice illustrated by
+Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the
+posterior probability of a model, is not a function of \myTheta{}, because
+it is integrated out of the marginal likelihoods of the candidate models.
+Thus, unlike parameter estimates, the estimated posterior probability of a
+model is a single value (rather than a distribution) lacking a measure of
+posterior uncertainty.
+
+The justification for an empirical Bayesian approach to parameter estimation is
+that giving the data more weight relative to the prior (i.e., using the data
+twice), will often shift the peak of the estimated distribution nearer to the
+true value(s) of the model's parameter(s).
+However, there is often no such justification for model selection, because
+unlike model parameters, the posterior probabilities of candidate models often
+have no clear true values.
 % There is no such justification for empirical Bayesian model choice, because
@@
 in some models and under confident in others.
-
-Another important difference of Bayesian model choice illustrated by
-Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the
-posterior probability of a model, is not a function of \myTheta{}, because
-it is integrated out of the marginal likelihoods of the candidate models.
-Thus, unlike parameter estimates, the estimated posterior probability of a
-model is a single value (rather than a distribution) lacking a measure of
-posterior uncertainty.
-% In some simple situations we might suspect that one model is correct and the
-% rest are wrong, but in most applications, we know that all the models under
-% consideration are oversimplifications of reality.
 
@@
 For each model, we retained the 10,000 samples with the smallest Euclidean
-distance from the observed summary statistics, standardizing the statistics
-using the prior means and standard deviations of the given model.
+distance from the observed summary statistics after standardizing the
+statistics using the prior means and standard deviations of the given model.
 From the remaining 50,000 samples, we then retained the 10,000 samples with the
@@
 \citet{Hickerson2013} likely excludes the true divergence times of at least
-some of the 22 taxa, we are nearly certain that these narrower priors are
-incorrect and exclude most, if not all, of the divergence times of the
-Philippine taxa.
+some of the 22 taxa, we are nearly certain that these narrower priors exclude
+most, if not all, of the divergence times of the Philippine taxa.
 
 Our results show that the model-averaging approach of \citet{Hickerson2013}
-does not reduce the method's support of models with less parameter space.
+does not reduce the method's tendency to support models with less parameter
+space.
 Rather, the method strongly prefers the prior model with the narrowest
@@
 When using uniformly distributed priors, the alternative to priors that reflect
-prior uncertainty is to risk excluding the true values one seeks to estimate,
-as shown above.
+prior uncertainty is to risk excluding the true values one seeks to estimate.
 Fortunately, continuous distributions more appropriate as priors for positive
@@
 
-\section{General thoughts on the model of \texttt{msBayes}}
+% \section{General thoughts on the model of \texttt{msBayes}}
+\section{A difficult inference problem}
 Our findings are not surprising in light of the difficult inference problem
@@
 to prior conditions.
-Robustness is an important characteristic of a method to gauge its
-applicability to real-world data, because we know the model and priors will be
-wrong to some degree.
+% Robustness is an important characteristic of a method to gauge its
+% applicability to real-world data, because we know the model and priors will be
+% wrong to some degree.
 
@@
 distributions in place of uniform priors improves the power of the method to
-detect temporal variation in divergences and reduces the bias toward models of
+detect temporal variation in divergences and reduces incorrect support for
 clustered divergences.

### 2014-07-23 manuscripts/reply/reply-main.tex
More editing of reply ms.

@@
 The strong preference for the narrowest priors on divergence times used by
-\citet{Hickerson2013} suggest their approach is biased toward models with less
+\citet{Hickerson2013} suggests their approach is biased toward models with less
 parameter space and, as a consequence, will estimate model-averaged posteriors
@@
 simulated (Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
-The method is most likely to infer the extreme case of a single divergence
-event shared across all 22 taxa when populations diverged randomly over the
-past \globalcoalunit generations.
-When divergences are random over the past $8\globalpopsize$ generations, the
-most likely inference is only two divergence events, and a single
-divergence is still estimated in more than 10\% of the replicates.
-It is interesting to note that as \divt{max} increases, but before the
-estimates are finally pulled away from $\numt{} = 1$, the distribution of
-\numt{} estimates closely mirror the U-shaped prior on divergence models
-used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
-(\citeyear{Oaks2012}) Figure 5B).
-This suggests this U-shaped prior coupled with the small marginal likelihoods
-of models with many divergence parameters, is a major cause of the method's
-bias toward clustered divergence models; this is consistent with
-\citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.
-
-Looking at our simulation results in terms of the posterior probability of the
-dispersion index of divergence times supporting the extreme case of one
-divergence event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the
-method strongly supports one divergence in greater than 27\% of the replicates
-across all the \divt{max} we simulated (Figure~\ref{figPower4}E--H \&
-S\ref{figPower6}G--L);
-following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
-the criterion for incorrect inference of a single divergence event.
-Furthermore, there is strong support for a single divergence event in
-more than 90\% of the replicates when divergences are random over the past
-$2.4\globalpopsize$ generations, and more than 60\% when over the past
-$3.2\globalpopsize$ generations (Figure~\ref{figPower4}E--H \&
+If we follow \citet{Hickerson2013} and use a Bayes factor of greater than 10 as
+the criterion for incorrect inference of a single divergence event, we find
+that the approach will often strongly support the extreme case of one
+divergence event across all our simulation conditions (Figure~\ref{figPower4}E--H \&
 S\ref{figPower6}G--L).
-
-Our results show that the empirical Bayesian model-averaging approach leads to
-erroneous support for highly clustered divergences when populations diverge
-randomly over the last $8\globalpopsize$ generations.
-To put this on the scale roughly consistent with a vertebrate mitochondrial
+Overall, the empirical Bayesian model-averaging approach leads to erroneous
+support for highly clustered divergences when populations diverge randomly over
+the last $8\globalpopsize$ generations.
+To put this on a timescale roughly consistent with a vertebrate mitochondrial
 locus, assuming a mutation rate of $1\e{-8}$ per site per generation, this
@@
 million generations.
+
+% The method is most likely to infer the extreme case of a single divergence
+% event shared across all 22 taxa when populations diverged randomly over the
+% past \globalcoalunit generations.
+% When divergences are random over the past $8\globalpopsize$ generations, the
+% most likely inference is only two divergence events, and a single
+% divergence is still estimated in more than 10\% of the replicates.
+It is interesting to note that as \divt{max} increases, but before the
+estimates are finally pulled away from $\numt{} = 1$, the distribution of
+\numt{} estimates closely mirror the ``U-shaped'' prior on divergence models
+used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
+(\citeyear{Oaks2012}) Figure 5B).
+This suggests the U-shaped prior coupled with the small marginal likelihoods
+of models with many divergence parameters is a major cause of the method's
+bias toward clustered divergence models; this is consistent with
+\citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.
+
+% Looking at our simulation results in terms of the posterior probability of the
+% dispersion index of divergence times supporting the extreme case of one
+% divergence event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the
+% method strongly supports one divergence in greater than 27\% of the replicates
+% across all the \divt{max} we simulated (Figure~\ref{figPower4}E--H \&
+% S\ref{figPower6}G--L);
+% following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
+% the criterion for incorrect inference of a single divergence event.
+% Furthermore, there is strong support for a single divergence event in
+% more than 90\% of the replicates when divergences are random over the past
+% $2.4\globalpopsize$ generations, and more than 60\% when over the past
+% $3.2\globalpopsize$ generations (Figure~\ref{figPower4}E--H \&
+% S\ref{figPower6}G--L).
+
 % Given the empirically Bayesian model-averaging approach of \citet{Hickerson2013}

### 2014-07-25 manuscripts/reply/reply-main.tex
More editing of reply ms.

@@
 Our results, show that it is difficult to choose a uniformly distributed prior
-on divergence times that is broad enough to confidently contain the true values
-of parameters while being narrow enough to avoid spurious support of models
+on divergence times that is broad enough to confidently contain the true
+parameter values while being narrow enough to avoid spurious support for models
 with less parameter space.
@@
 the data given $\myTheta{}$ (the likelihood of $\myTheta{}$).
-The integral over the entire parameter space of the model only appears as a
-normalizing constant in the denominator.
+The marginal likelihood of the model only appears as a normalizing constant in
+the denominator.
 Thus, as long as the prior distribution contains the values of $\myTheta{}$
-that maximize the probability of the data and the data are strongly informative
+that maximize the likelihood and the data are strongly informative
 relative to the prior, the values of the parameters that maximize the posterior
@@
 The integral over the entire parameter space of the likelihood weighted by the
-prior probability density is no longer a normalizing constant, but rather is
-what informs the posterior probability of the model from the data.
-As a result, Bayesian model choice tends to be much more sensitive to priors
-than parameter estimation.
+prior density is no longer a normalizing constant, rather it is how the data
+inform the posterior probability of the model.
+Because the prior probability distributions placed on the model's parameters
+have a strong affect on the integrated, or ``average'', likelihood of a model,
+Bayesian model choice tends to be much more sensitive to priors than parameter
+estimation \citep{Jeffreys1939,Lindley1957}.
 Another important difference of Bayesian model choice illustrated by
@@
 that giving the data more weight relative to the prior (i.e., using the data
-twice), will often shift the peak of the estimated distribution nearer to the
+twice) will often shift the peak of the estimated distribution nearer to the
 true value(s) of the model's parameter(s).

### 2014-07-27 manuscripts/reply/reply-abstract.tex
More editing of reply ms.

@@
 use of vague priors.
-Because this sensitivity to prior assumptions affects the crucial insights 
-that a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended
-that users of the approach should carefully assess the robustness of their 
+Because this sensitivity to prior assumptions affects the crucial insights that
+a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended that
+users of the approach should carefully assess the robustness of their
 conclusions to different priors.
-According to \citet{Hickerson2013}, the lack of robustness in \msb analyses
-was due to excessively broad priors on divergence times leading to 
-inadequate numbers of simulation replicates.
+In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
+analyses was due to the method's inefficient rejection algorithm and broad
+priors on divergence times leading to inadequate numbers of simulation
+replicates.
 They proposed a model-averaging approach that uses narrow, empirically
@@
 Fortunately, as predicted by fundamental principles of Bayesian model choice,
-more appropriate priors on the model's parameters increase the method's
-robustness and power to detect temporal variation in divergences
+more appropriate priors on the models' parameters can accommodate prior
+uncertainty while increasing the method's robustness and power to detect
+temporal variation in divergences
 \citep{Oaks2014dpp}.

### 2014-07-27 manuscripts/reply/reply-main.tex
More editing of reply ms.

@@
 % difficult problem in evolutionary biology.
-\citet{Huang2011} introduced an approximate Bayesian approach (implemented in
-the software \msb) to estimating the probabilities of models in which multiple
-sets of taxa diverge at the same time. 
+% \citet{Huang2011} introduced an approximate Bayesian approach (implemented in
+% the software \msb) to estimating the probabilities of models in which multiple
+% sets of taxa diverge at the same time. 
+The approximate-Bayesian method, \msb, allows biogeographers to estimate the
+probabilities of models in which multiple sets of taxa diverge at the same
+time \citep{Hickerson2006,Huang2011}.
 % This method  does not require a common set of character data that links all
@@
 We show the primary cause of the method's tendency to incorrectly infer shared
-divergences is the disparity in marginal likelihoods across divergence models
-imposed by the uniform priors on divergence times.
+divergences is the uniform priors on divergence times leading to small marginal
+likelihoods of models with more divergence-time parameters.
 We also show that sampling error is not a major contributer to the problem.
@@
 flexible probability distributions to accommodate prior uncertainty in
-divergence times without inhibiting the marginal likelihoods of models with
-more divergence-time parameters, which increases the method's robustness and
-power to detect temporal variation in divergences.
+divergence times without overly inhibiting the marginal likelihoods of models
+with more divergence-time parameters, which increases the method's robustness
+and power to detect temporal variation in divergences.
 
@@
 Empirical Bayes is a well-studied branch of Bayesian statistics that has given
-rise to many inference methods for obtaining parameter estimates that often
+rise to many methods for obtaining parameter estimates that often
 exhibit favorable frequentist properties.
@@
 Thus, as long as the prior distribution contains the values of $\myTheta{}$
-that maximize the likelihood and the data are strongly informative
+under which the data are probable and the data are strongly informative
 relative to the prior, the values of the parameters that maximize the posterior
@@
 Another important difference of Bayesian model choice illustrated by
-Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the
-posterior probability of a model, is not a function of \myTheta{}, because
-it is integrated out of the marginal likelihoods of the candidate models.
+Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the posterior
+probability of a model, is not a function of \myTheta{}, because the parameters
+are integrated out of the marginal likelihoods of the candidate models.
 Thus, unlike parameter estimates, the estimated posterior probability of a
@@
 true value(s) of the model's parameter(s).
-However, there is often no such justification for model selection, because
-unlike model parameters, the posterior probabilities of candidate models often
-have no clear true values.
+However, there is no such justification for model selection, because unlike
+model parameters, the posterior probabilities of candidate models often have no
+clear true values.
 % There is no such justification for empirical Bayesian model choice, because
@@
 \change{
-The strong preference for the narrowest priors on divergence times used by
-\citet{Hickerson2013} suggests their approach is biased toward models with less
-parameter space and, as a consequence, will estimate model-averaged posteriors
-dominated by models that exclude true values of the model's parameters.
+The strong preference for the narrowest prior on divergence times suggests the
+approach of \citet{Hickerson2013} is biased toward models with less parameter
+space and, as a consequence, will estimate model-averaged posteriors dominated
+by models that exclude true values of the model's parameters.
 }
@@
 Our results show that the model-averaging approach of \citet{Hickerson2013}
-does not reduce the method's tendency to support models with less parameter
-space.
-Rather, the method strongly prefers the prior model with the narrowest
-distribution on divergence times across all three of our analyses, even when
-this model is incorrect and excludes the true divergence times
-of the Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
-
+strongly prefers the prior model with the narrowest distribution on divergence
+times across all three of our analyses, even when this model is incorrect and
+excludes the true divergence times of the Philippine taxa
+(Table~\ref{tabModelChoiceEmpirical}).
 However, \citet{Hickerson2013} vetted the priors used in their model-averaging
@@
 models.
-We assess this possibility using simulation-based analyses.
+We assess this possibility using simulations.
 Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
@@
 
-We find that the approach of \citet{Hickerson2013} struggles to estimate the
-variance of divergence times (\vmratio{}) across most of the \divt{max}
-we simulated, whether evaluating the unadjusted
-(Figure~S\ref{figPowerAccuracy}A--F) or GLM-adjusted
-(Figure~S\ref{figPowerAccuracy}G--L) posterior estimates.
-The method only estimates \vmratio{} relatively well when the simulated
-distribution of divergence times is identical to one of the prior models
-(Figures~S\ref{figPowerAccuracy}E\&K).
-This is consistent with the findings of \citet{Oaks2012}  and
-\citet{Hickerson2013} that \msb is highly sensitive to the prior distribution
-deviating from the true, underlying distribution of the data.
-
-Furthermore, our results demonstrate that the approach of \citet{Hickerson2013}
-consistently infers highly clustered divergences across all the \divt{max} we
-simulated (Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
+Our results demonstrate that the approach of \citet{Hickerson2013} consistently
+infers highly clustered divergences across all the \divt{max} we simulated
+(Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
 If we follow \citet{Hickerson2013} and use a Bayes factor of greater than 10 as
@@
 S\ref{figPower6}G--L).
+The method also struggles to estimate the variance of divergence times
+(\vmratio{}), whether evaluating the unadjusted
+(Figure~S\ref{figPowerAccuracy}A--F) or GLM-adjusted
+(Figure~S\ref{figPowerAccuracy}G--L) posterior estimates.
 Overall, the empirical Bayesian model-averaging approach leads to erroneous
@@
 To put this on a timescale roughly consistent with a vertebrate mitochondrial
-locus, assuming a mutation rate of $1\e{-8}$ per site per generation, this
+locus (assuming a per-site mutation rate of $1\e{-8}$ per generation) this
 translates to 10 million generations.
@@
 % divergence is still estimated in more than 10\% of the replicates.
-It is interesting to note that as \divt{max} increases, but before the
-estimates are finally pulled away from $\numt{} = 1$, the distribution of
-\numt{} estimates closely mirror the ``U-shaped'' prior on divergence models
-used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
-(\citeyear{Oaks2012}) Figure 5B).
-This suggests the U-shaped prior coupled with the small marginal likelihoods
-of models with many divergence parameters is a major cause of the method's
-bias toward clustered divergence models; this is consistent with
-\citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.
+% It is interesting to note that as \divt{max} increases, but before the
+% estimates are finally pulled away from $\numt{} = 1$, the distribution of
+% \numt{} estimates closely mirror the ``U-shaped'' prior on divergence models
+% used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
+% (\citeyear{Oaks2012}) Figure 5B).
+% This suggests the U-shaped prior coupled with the small marginal likelihoods
+% of models with many divergence parameters is a major cause of the method's
+% bias toward clustered divergence models; this is consistent with
+% \citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.
 
@@
 (Figure~\ref{figPowerExclusion4}E--H \& S\ref{figPowerExclusion6}G--L).
-Only when the data are identically distributed as one of the prior models does
-the method avoid excluding the truth more than 5\% of the time
-(Figure~\ref{figPowerExclusion4}C).
+% Only when the data are identically distributed as one of the prior models does
+% the method avoid excluding the truth more than 5\% of the time
+% (Figure~\ref{figPowerExclusion4}C).
 % Again, this demonstrates the method's sensitivity to priors.
@@
 over millions of generations.
-As recommended by \citet{Oaks2012}, it is important that investigators perform
-power analyses to determine the method's power for their dataset, and decide if
-\msb has sufficient temporal resolution to address their hypotheses; in the
-case of the Philippines dataset, it did not.
+It is important that investigators perform power analyses to determine the
+method's power for their dataset, and decide if \msb has sufficient temporal
+resolution to address their hypotheses; in the case of the Philippines dataset,
+it did not.
 When doing so, it is important to consider what prior conditions are
@@
 Also, it seems unlikely that when such prior information is available that
-being able to detect multiple divergences on the scale of hundreds of thousands
-of generations will add much insight about the evolutionary history of the
-taxa.
+being able to detect more than one divergence event in the face of 18 random
+divergences over hundreds of thousands of generations will provide much insight
+into the evolutionary history of the taxa.
 % ; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), this
@@
 
-As noted by \citet{Oaks2012}, inferring more than one divergence time
-shared across all taxa does not confirm the method is working well when
-analyzing data generated under random temporal variation in divergences (e.g.,
-an inference of two divergence events could be biogeographically interesting
-yet spurious).
+Inferring more than one divergence time shared across all taxa does not confirm
+the method is working well when analyzing data generated under random temporal
+variation in divergences (e.g., an inference of two divergence events could be
+biogeographically interesting yet spurious).
 Thus, it is important that investigators not limit their assessment of the
@@
 $\numt{} = 1$ versus $\numt{} > 1$).
-Rather, looking at the distribution of estimates, as in \citet{Oaks2012},
-provides much more information about the behavior of the method.
+Rather, looking at the distribution of estimates, as in Figure~\ref{figPower4}
+and \citet{Oaks2012}, provides much more information about the behavior of the
+method.
 
@@
 smaller marginal likelihoods \citep{Jeffreys1939,Lindley1957}.
-Given the uniformly distributed priors on divergence times (and other nuisance
-parameters) employed in \msb, models with more divergence parameters will be
-forced to integrate over \emph{much} greater parameter space, all with equal
-prior density, and much of it with low likelihood.
+Given the uniformly distributed priors on divergence times employed in \msb,
+models with more divergence parameters will be ``averaged'' over much greater
+parameter space, all with equal prior weight, and much of it with low
+likelihood.
 In light of this fundamental statistical issue, it is not surprising that the
@@
 gene divergence times across the 22 taxa.
-These analyses assumed an arbitrary strict mutation rate of $2\e{-8}$ for all
-taxa, and are, of course, subject to estimation error.
+These analyses assumed an arbitrary strict per-site mutation rate of $2\e{-8}$
+per generation for all taxa, and are, of course, subject to estimation error.
 Furthermore, the branch-length units of the gene trees are in millions of
@@
 have a generation time of one year.
-The argument of \citet{Hickerson2013} that divergence time estimates of
-\citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
-questionable, especially given our findings presented above regarding the
+The argument of \citet{Hickerson2013} that divergence-time estimates of
+\citet{Oaks2012} should set an upper bound on the prior for divergence times
+seems questionable, especially given our findings presented above regarding the
 behavior of \msb when empirically informed uniform priors are employed.
@@
 approximation method, insufficient sampling is an unlikely explanation for the
-erroneous support of models with less parameter space.
+erroneous support for models with less parameter space.
 While analyses that sample each model proportional to their parameter space
@@
 \msb can spuriously support models with less parameter space.
-As suggested by \citet{Oaks2012}, this is likely caused by the use of uniform
-priors on divergence times.
+This is caused by the use of uniform priors on divergence times.
 Uniform distributions necessitate the use of priors that place high density in
@@
 prior on divergence times that is broad enough to confidently contain the true
-values of parameters while being narrow enough to avoid bias toward models with
-less parameter space.
+values of parameters while being narrow enough to avoid erroneous support for
+models with less parameter space.
 

### 2014-07-27 manuscripts/reply/reply-si-main.tex
More editing of reply ms.

@@
 U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
-the size of the constant reference populations is $\myTheta{C} =
+the size of the constant reference population is $\myTheta{C} =
 \uniformMax{\meanDescendantTheta{}}/2$.
@@
 The error is easily illustrated by re-plotting their results with the different
-time units indicated (Figure~\ref{figJointPosteriorHickerson}).
+time units indicated (Figure~S\ref{figJointPosteriorHickerson}).
 
@@
 \subsection{Graphical prior comparisons}
-\citet{Hickerson2013} advocate the use of what they call graphical checks of
-prior models.
+\citet{Hickerson2013} advocate the use graphical checks of prior models.
 This prior-predictive approach entails generating a small number (1000) of
@@
 is contained within the divergence-time vector that \vmratio{} is summarizing.
-Clearly, the number of divergence time parameters within the vector and their
-values is more informative than its variance (i.e., the dispersion index is not
+Clearly, the number of divergence-time parameters within the vector and their
+values is more informative than the variance (i.e., the dispersion index is not
 a sufficient statistic for \divtvector).
-\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
-better than \numt{}.''
+\citet{Hickerson2013} also argue that \msb can estimate \vmratio{} much better
+than \numt{}.
 However, \citet{Oaks2012} demonstrate that even when all assumptions of the

### 2014-07-28 manuscripts/dpp-msbayes/dpp-msbayes-abstract.tex
Adding proof corrections and minor edits for dpp ms.

@@
 \ifbmc{\parttitle{Background}}{}
-To understand the processes that generate biodiversity, it is
-important to account for large-scale processes that affect the evolutionary
-history of groups of co-distributed populations of organisms.
+To understand biological diversification, it is important to account for
+large-scale processes that affect the evolutionary history of groups of
+co-distributed populations of organisms.
 Such events predict temporally clustered divergences times, a pattern
@@
 The new method yields more robust estimates of posterior uncertainty, and
-thus greatly reduces the tendency of the model to incorrectly estimate
+thus greatly reduces the tendency to incorrectly estimate
 \change{models of shared evolutionary history} with strong support.

### 2014-07-28 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Adding proof corrections and minor edits for dpp ms.

@@
 partitioning consists of non-overlapping and non-empty subsets whose union is
-the \npairs{} pairs;
+the \npairs{} pairs.
 Hereinafter, I refer to these partitionings as ``ordered'' divergence models or

### 2014-07-29 manuscripts/reply/reply-abstract.tex
More editing of reply ms.

@@
 incorrectly support shared divergence events and is dangerous in the sense that
-the empirically-derived priors often exclude from consideration the true values
-of the models' parameters.
+the empirically-derived uniform priors often exclude from consideration the
+true values of the models' parameters.
 On a more fundamental level, we question the value of adopting an empirical
@@
 method, which helps guide interpretation of results.
-Fortunately, as predicted by fundamental principles of Bayesian model choice,
-more appropriate priors on the models' parameters can accommodate prior
-uncertainty while increasing the method's robustness and power to detect
-temporal variation in divergences
+As predicted by fundamental principles of Bayesian model choice, more flexible
+distributions that are better suited as priors for the models' parameters can
+accommodate prior uncertainty while increasing the method's robustness and
+power to detect temporal variation in divergences
 \citep{Oaks2014dpp}.

### 2014-07-29 manuscripts/reply/reply-figures.tex
More editing of reply ms.

@@
 \mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-4.pdf}{
-    Histograms of the (A--D) number of true divergence-time parameters excluded
+    Histograms of the (A--D) number of true divergence times excluded
     from the preferred model and the (E--H) posterior probability of excluding
-    at least one divergence-time parameter when the empirically informed
+    at least one true divergence time when the empirically informed
     model-averaging approach of \citet{Hickerson2013} is applied to simlated

### 2014-07-29 manuscripts/reply/reply-main.tex
More editing of reply ms.

@@
 ``escape'' models with large parameter space, which manifests in the preference
-of models that exclude from consideration the true values of the model's
+of models that exclude from consideration the true values of the models'
 parameters.
@@
 less parameter space, we expect \msb will preferentially sample from this
-incorrect model yielding a posterior that is incorrect (i.e., the
+erroneous model yielding a posterior that is incorrect (i.e., the
 model-averaged posterior will be dominated by an incorrect model that excludes
@@
 Following \citet{Hickerson2013}, each of these models had the following
-priors on divergence time parameters:
+priors on divergence-time parameters:
 $M_1$, $\divt{} \sim U(0, 0.1)$;
@@
 axes of a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
-To determine if such prior predictive analyses would indicate the $M_{1A}$ and
+To determine if such prior-predictive analyses would indicate the $M_{1A}$ and
 $M_{1B}$ models are problematic, we performed these graphical checks on our
 prior models.
-Unfortunately, these prior predictive checks provide no warning that these
+Unfortunately, these prior-predictive checks provide no warning that these
 priors are too narrow (Figure~S\ref{figPCA}).
@@
 Importantly, the posterior probability of excluding at least one true parameter
-value is very high in nearly all of the replicates
+value is very high in most replicates
 (Figure~\ref{figExclusionSimProb}).
@@
 \change{
-When using uniformly distributed priors, the alternative to priors that reflect
-prior uncertainty is to risk excluding the true values one seeks to estimate.
-Fortunately, continuous distributions more appropriate as priors for positive
-real-valued parameters have been shown to greatly reduce spurious support for
-clustered divergence models while allowing prior uncertainty to be accommodated
+When using uniformly distributed priors, the alternative to capturing prior
+uncertainty is to risk excluding the true values one seeks to estimate.
+Fortunately, more flexible continuous distributions that are better suited as
+priors for the positive real-valued parameters of the \msb model have been
+shown to greatly reduce spurious support for clustered divergence models while
+allowing prior uncertainty to be accommodated
 \citep{Oaks2014dpp}.
@@
 Overall, the empirical Bayesian model-averaging approach leads to erroneous
-support for highly clustered divergences when populations diverge randomly over
-the last $8\globalpopsize$ generations.
-To put this on a timescale roughly consistent with a vertebrate mitochondrial
-locus (assuming a per-site mutation rate of $1\e{-8}$ per generation) this
-translates to 10 million generations.
-Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
-million generations.
+support for highly clustered divergences when populations diverged randomly
+over the last $8\globalpopsize$ generations.
+For loci with per-site rates of mutation on the order of $1\e{-8}$ and
+$1\e{-9}$ per generation, this translates to 10 million and 100 million
+generations, respectively.
 
@@
 over millions of generations.
-It is important that investigators perform power analyses to determine the
+It is important that investigators perform simulations to determine the
 method's power for their dataset, and decide if \msb has sufficient temporal
@@
 Given the uniformly distributed priors on divergence times employed in \msb,
-models with more divergence parameters will be ``averaged'' over much greater
-parameter space, all with equal prior weight, and much of it with low
-likelihood.
+the likelihood of models with more divergence parameters will be ``averaged''
+over much greater parameter space, all with equal prior weight, and much of it
+with low likelihood.
 In light of this fundamental statistical issue, it is not surprising that the
@@
 \citet{Oaks2012} were estimated without error.
-These estimates were intended to only provide a rough comparison of the
+These estimates were intended to provide only a rough comparison of the
 gene divergence times across the 22 taxa.
-These analyses assumed an arbitrary strict per-site mutation rate of $2\e{-8}$
+These analyses assumed an arbitrary strict per-site rate of $2\e{-8}$ mutations
 per generation for all taxa, and are, of course, subject to estimation error.
 Furthermore, the branch-length units of the gene trees are in millions of
-years, whereas the divergence time prior of \msb is in generations, thus
+years, whereas the divergence-time prior of \msb is in generations, thus
 \citet{Hickerson2013} make the implicit assumption that all 22 Philippine taxa
@@
 
-Even if we assume the gene divergence times are estimated without error,
-Hickerson et al.'s (\citeyear{Hickerson2013}) argument only applies to one of
-the three priors used by \citet{Oaks2012}.
-The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5)$)
-closely mirrors the range of estimates of gene-divergence times if we assume
-the same mutation rate and one generation per year (0--5 million years ago).
+Even if we assume the arbitrary strict clock is correct, gene divergence times
+were estimated without error, and all 22 taxa have one-year generation times,
+Hickerson et al.'s
+(\citeyear{Hickerson2013}) argument only applies to one of the three priors
+used by \citet{Oaks2012}.
+Under these assumptions, the narrowest prior on divergence times used by
+\citet{Oaks2012} ($U(0, 5)$) closely mirrors the range of estimates of
+gene-divergence times (0--5 million years ago).
 Applying Hickerson et al.'s (\citeyear{Hickerson2013}) sampling-probability
@@
 different predictions made by these two phenomena.
-One example, as discussed by \citet{Oaks2012}, is that insufficient prior
-sampling should create large variance among posterior estimates, and thus it
-should cause analyses to be highly sensitive to the number of samples drawn
-from the prior.
+For example, insufficient prior sampling should create large variance among
+posterior estimates and cause analyses to be highly sensitive to the number of
+samples drawn from the prior.
 Furthermore, if sampling error is biased toward models with less
@@
 
-As discussed by \citet{Oaks2012}, a straightforward prediction if marginal
-likelihoods are causing the preference for simple models is that the bias should
-disappear as the model generating the data converges to the prior.
+A straightforward prediction if marginal likelihoods are causing the preference
+for simple models is that the bias should disappear as the model generating the
+data converges to the prior.
 \citet{Oaks2012} tested this prediction by performing 100,000 simulations to
@@
 the bias.
-Even when the prior is correct, due to the discrete uniform prior on
-\numt{} implemented in \msb, models with larger numbers of divergence
-events (and thus greater parameter space) will still be less densely
-sampled than those with fewer divergence events \citep{Oaks2012}.
+Even when the prior is correct, due to the discrete uniform prior on the number
+of divergence events (\numt{}) implemented in \msb, models with larger numbers
+of divergence-time parameters (and thus greater parameter space) will still be
+far less densely sampled than those with fewer divergence events
+\citep{Oaks2012}.
 Thus, the results of the simulations of \citet{Oaks2012} are more consistent
@@
 Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerExclusion4})
-and models with fewer \divt{} parameters (Figure~\ref{figPower4}).
+and fewer \divt{} parameters (Figure~\ref{figPower4}).
 In all of these analyses, each of the prior models have the same number of
@@
 % the vector \divtvector.
-When applying the simplest model in \msb (i.e., assuming no migration and no
-intra-locus recombination) to the dataset of \citet{Oaks2012} with 22 taxon
+When applying \msb to the dataset of \citet{Oaks2012} with 22 taxon
 pairs, there are 581--602 free parameters that model highly stochastic
 coalescent and mutational processes.
-Furthermore, under this rich stochastic model, the method is estimating the
+Under this rich stochastic model, the method is estimating the
 probability of 1002 divergence models \citep[i.e., the number of integer
@@
 Furthermore, all the information in the sequence alignment of each taxon pair
-is distilled into four summary statistics:
-$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
-\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
-That gives us a total of 88 summary statistics (four from each of the 22 taxon
-pairs), which contain minimal information about many of the $\approx 600$
+is distilled into four summary statistics.
+This gives us a total of 88 summary statistics (four from each of the 22 taxon
+pairs) that contain minimal information about many of the $\approx 600$
 parameters in the model.
@@
 and in the context of a model-choice method that is highly sensitive to priors,
-it is not too surprising that the approach struggles.
+it is not surprising that the approach struggles.
 
@@
 need for imposing an additional dimension of model choice.
-\citet{Oaks2012} suggest that uniform priors may not be appropriate for many
+\citet{Oaks2012} suggest that uniform priors may not be ideal for many
 parameters of the \msb model, and recommend the use of probability
@@
 Thus, it seems more appropriate to simply place a gamma prior (the exponential
-being a special case of the gamma) on divergence times.
+being a special case) on divergence times.
 This would capture the prior uncertainty that \citet{Hickerson2013} are

### 2014-07-29 manuscripts/reply/reply-si-figures.tex
More edits to reply ms.

@@
     The three models that likely exclude true values of some divergence times
-    of the 22 pairs of Philippine vertebrate taxa (A--C) appear to have a
-    better ``fit'' than the priors that likely cover the true divergence times
+    of the 22 pairs of Philippine taxa (A--C) appear to have a ``better fit''
+    than the valid priors that likely cover the true divergence times
     (D--F).
@@
     The accuracy of (A--F) unadjusted and (G--L) GLM-adjusted estimates of
-    dispersion index of divergence times (\vmratio{}) when the empirically
+    the dispersion index of divergence times (\vmratio{}) when the empirically
     informed model-averaging approach of \citet{Hickerson2013} is applied to
@@
 \siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-6.pdf}{
-    Histograms of the (A--F) number of true divergence-time parameters excluded
+    Histograms of the (A--F) number of true divergence times excluded
     from the preferred model and the (G--L) posterior probability of excluding
-    at least one divergence-time parameter when the empirically informed
+    at least one true divergence time when the empirically informed
     model-averaging approach of \citet{Hickerson2013} is applied to simlated

### 2014-07-29 manuscripts/reply/reply-si-main.tex
More edits to reply ms.

@@
 
-The fact that their posterior samples are in different units makes the results
-of \citet{Hickerson2013} difficult to interpret, and renders their
-regression-adjusted results invalid.
-A fundamental assumption of regression is that all of the values of the
-response variable are in the same units.
+A fundamental assumption of the \msb model and post hoc regression adjustment
+is that all possible values of the parameter of interest (divergence times) are
+in the same units.
 Thus, the results in sections ``Using ABC Model Comparison to Weight
@@
 the Philippines Vertebrate Data'' and presented in Figure 2 of
-\citet{Hickerson2013} should be disregarded.
+\citet{Hickerson2013} are invalid and should be disregarded.
 The error is easily illustrated by re-plotting their results with the different
@@
 summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
-Thus, the assertion of \citet{Hickerson2013} does not apply to at least
-two of the priors used by \citet{Oaks2012} and, as a result, does not
-explain the bias they found.
+Thus, the assertion of \citet{Hickerson2013} that the priors used by
+\citet{Oaks2012} sample parameter space in which the mtDNA alignments are
+saturated by substitutions is incorrect and, as a result, does not explain the
+bias they found.
 
@@
 variance.
-This plot shows, not surprisingly, that \vmratio{} has very little information
+This plot shows that \vmratio{} has very little information
 about the number of divergences among taxa.
@@
 % We struggle to follow this logic.
-However, all of the information about the temporal distribution of divergences
-is contained within the divergence-time vector that \vmratio{} is summarizing.
-Clearly, the number of divergence-time parameters within the vector and their
-values is more informative than the variance (i.e., the dispersion index is not
-a sufficient statistic for \divtvector).
+However, the number of divergence-time parameters within the vector and their
+values contains all of the information about the temporal distribution of
+divergences, and is much more informative than the variance (i.e., the
+dispersion index is not a sufficient statistic for \divtvector).
 \citet{Hickerson2013} also argue that \msb can estimate \vmratio{} much better
@@
 It is easier, less subjective, and more interpretable to estimate the
-probability of the discrete parameter of the model, \numt{}, is at its lower
-limit of one.
+probability of the model with one divergence-time parameter (i.e., $\numt{} =
+1$).
 Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better

### 2014-07-29 manuscripts/reply/reply-main.tex
Working on overhaul of intro for reply ms.

@@
 were affected by the same geographic events.
-% Unfortunately, estimating divergence times precisely when one lacks
-% precise information about the rate of molecular evolution is a notoriously
-% difficult problem in evolutionary biology.
-% \citet{Huang2011} introduced an approximate Bayesian approach (implemented in
-% the software \msb) to estimating the probabilities of models in which multiple
-% sets of taxa diverge at the same time. 
 The approximate-Bayesian method, \msb, allows biogeographers to estimate the
@@
 time \citep{Hickerson2006,Huang2011}.
-% This method  does not require a common set of character data that links all
-% of the taxa under investigation, but does require the relative mutation
-% rates among the taxa and loci to be assumed \emph{a priori}.
 
@@
 the behavior of the \msb approach using computer simulations.
-They found the method was very sensitive to prior assumptions and suggested
-this was due to the fact that the marginal likelihoods of divergence models are
-sensitive to the use of uniform priors on divergence times.
-They also suggested alternative probability distributions (e.g., gamma) could
-better represent prior uncertainty about divergence times while mitigating the
-disparity of marginal likelihoods across divergence models.
-In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
-analyses was caused by sampling error due to the inefficiency of the method's
-rejection algorithm.
-As a potential solution, they proposed a model-averaging approach that uses
-narrow, empirically informed uniform priors.
-
-Whereas the papers of \citet{Oaks2012} and \citet{Hickerson2013} differ about
-the cause and solution, they agree on the fundamental methodological point
-about the model selection performed in \msb:
+They found the method was very sensitive to prior assumptions and often
+incorrectly supported shared divergences across taxa, to which
+\citet{Hickerson2013} responded.
+Both papers agree on the fundamental methodological point about the model
+selection performed in \msb:
 % The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
@@
 \end{itemize}
+However, the two papers suggest alternative causes for this behavior:
+\begin{itemize}
+    \item \citep{Hickerson2013} suggest broad uniform priors cause numerical
+        problems in which the ABC rejection algorithm is unable to sufficiently
+        sample the space of the models in reasonable computational time. The
+        method will incorrectly support models with fewer divergence-time
+        parameters because they are better sampled.
+    \item \citet{Oaks2012} suggest the cause is more fundamental; the broad
+        uniform priors simply lead to very small marginal likelihoods of models
+        with many divergence-time parameters.
+\end{itemize}
+In the former case, the true posterior would support heterogeneity in 
+divergences among taxa in the face of temporally random divergences; the
+problem only arises from our inability to accurately estimate the posterior.
+In the latter case, the true posterior would support simultaneous divergences
+among taxa in the face of random divergences. I.e., under the prior
+assumptions about divergence times, this is the correct answer from 
+the perspective of Bayesian model choice.
+Such posterior support for simultaneous divergence, even if correct, does not
+provide the biogeographical insights that a researcher who employs \msb seeks
+to gain.
+
+The distinction between these potential causes is important for improving our
+ability to estimate shared divergence histories.
+If the cause is numerical, then the model is sound and we simply need to
+improve our estimation procedures.
+For example, \citet{Hickerson2013} proposed a model-averaging approach that
+uses narrow, empirically informed uniform priors in the hope that with less
+parameter space to sample, the rejection algorithm will produce better
+estimates of the posterior.
+If the cause is marginal likelihoods, we need to correct the model, because
+improving the estimation machinery will only increase our ability to estimate
+biogeographically misleading posteriors.
+Accordingly, \citet{Oaks2014dpp} has introduced a method that uses more
+flexible probability distributions to accommodate prior uncertainty in
+divergence times without overly inhibiting the marginal likelihoods of models
+with more divergence-time parameters, which increases the method's robustness
+and power to detect temporal variation in divergences.
+
 This is not surprising given the rich statistical literature showing that
 marginal likelihoods are very sensitive to the priors used in Bayesian model
-selection
-\citep[e.g.,][]{Jeffreys1939,Lindley1957}.
-% We know that under broad uniform priors on divergence times, models with few
-% divergence parameters will have greater marginal likelihoods relative to more
-% parameter-rich models.
-% Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
-% for models with few divergence parameters is the greater marginal likelihoods
-% of these models under vague uniform priors, relative to more parameter-rich
-% models.
-Under vague uniform priors on divergence times, models with more
-divergence-time parameters integrate over \emph{much} greater parameter space
-with low probability of producing the data, yet relatively high prior density
-imposed by the uniformly distributed prior.
-% By using such priors on divergence times, we are essentially telling the
-% method to favor models with few divergence times even if they do not
-% explain the data very well.
+selection \citep[e.g.,][]{Jeffreys1939,Lindley1957}.
 Thus, uniform priors on divergence times can impose a strong ``penalty'' for
@@
 not explain the data very well.
-% It is important to note that this is not due to estimation error
-% (i.e., this phenomenon is extrinsic to the numerical approximation machinery of
-% \msb).
-% Rather, the \emph{true} posterior will tend to support models with few
-% divergence-time parameters under broad uniform priors on divergence times.
-Such posterior support for simultaneous divergence, even if correct from the
-perspective of Bayesian model choice, does not provide the biogeographical
-insights that a researcher who employs \msb seeks to gain.
-% Ideally, the estimated posterior probability of a simultaneous-divergence model
-% can be interpreted as an estimate of the probability that this model is correct.
-% However, when using uniform priors on divergence times, it might be better
-% interpreted as an estimate of the probability that the simultaneous-divergence
-% model cannot be completely ruled out as a possible explanation of the data.
-% Note, this is a statistical issue, extrinsic to \msb, and does not question the
-% soundness of the numerical approximation machinery of \msb.
-
-% Alternatively, \citet{Hickerson2013} suggest the problem is the result of a
-% numerical issue intrinsic to \msb, concluding the method's rejection algorithm
-% is inefficient and will be biased toward models with fewer divergence-time
-% parameters, because the parameter space of these models are more densely
-% sampled relative to models with more divergence-time parameters.
-% We agree that sampling error is present in all numerical Bayesian estimation
-% methods, however, we show here that it is not a major contributer to the
-% biases found by \citet{Oaks2012}.
-% We present results here of analyses that, when combined with the results
-% of \citet{Oaks2012}, strongly suggest that \msb prefers models with few
-% divergence-time parameters, because they have greater likelihoods when
-% averaged over broad uniform priors on divergences times.
-
-% \citet{Oaks2012} suggest alternative prior probability distributions on
-% divergence-times could increase the marginal likelihoods of models with more
-% divergence-time parameters and thus reduce spurious support for models of
-% temporally clustered divergences.
-% Alternatively, \citet{Hickerson2013} present an approach that uses empirically
-% informed uniform priors and Bayesian model-averaging in an attempt to accommodate
-% uncertainty in selecting priors.
-% In general, we agree with the use of Bayesian model averaging to obtain a
-% posterior estimate marginalized over uncertainty in prior choice.
-% However, we question whether adding an additional dimension of model choice
-% that samples over empirically informed uniform priors is a fruitful solution
-% for a model-choice method that is highly sensitive to priors.
 

### 2014-07-30 manuscripts/reply/reply-main.tex
More work on overhaul of intro to reply ms.

@@
 They found the method was very sensitive to prior assumptions and often
-incorrectly supported shared divergences across taxa, to which
-\citet{Hickerson2013} responded.
-Both papers agree on the fundamental methodological point about the model
-selection performed in \msb:
+incorrectly supported shared divergences across taxa that diverged randomly
+over broad time periods, to which \citet{Hickerson2013} responded.
+\citet{Oaks2012} and \citet{Hickerson2013} agree on the fundamental
+methodological point about the model selection performed in \msb:
 % The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
@@
    \item The use of vague priors on the divergence-time parameters can lead to
-       support for models with few divergence events shared across taxa. Thus,
-       the primary inference enabled by the approach is very sensitive to the
-       priors on divergence times.
+       spurious support for models with few divergence events shared across
+       taxa. Thus, the primary inference enabled by the approach is very
+       sensitive to the priors on divergence times.
 \end{itemize}
-However, the two papers suggest alternative causes for this behavior:
+However, the two papers suggest alternative mechanisms by which the priors on
+divergence times cause this behavior:
 \begin{itemize}
-    \item \citep{Hickerson2013} suggest broad uniform priors cause numerical
-        problems in which the ABC rejection algorithm is unable to sufficiently
-        sample the space of the models in reasonable computational time. The
-        method will incorrectly support models with fewer divergence-time
-        parameters because they are better sampled.
+    \item \citet{Hickerson2013} suggest that under broad uniform priors, the
+        rejection algorithm implemented in \msb is unable to adequately sample
+        the space of the models within reasonable computational time.
+        They propose the method incorrectly supports models with fewer
+        divergence-time parameters because they are better sampled.
     \item \citet{Oaks2012} suggest the cause is more fundamental; the broad
@@
         with many divergence-time parameters.
+        Models with more divergence-time parameters integrate over \emph{much}
+        greater parameter space with low probability of producing the data, yet
+        relatively high prior density, resulting in small marginal likelihoods.
 \end{itemize}
-In the former case, the true posterior would support heterogeneity in 
-divergences among taxa in the face of temporally random divergences; the
-problem only arises from our inability to accurately estimate the posterior.
-In the latter case, the true posterior would support simultaneous divergences
-among taxa in the face of random divergences. I.e., under the prior
-assumptions about divergence times, this is the correct answer from 
-the perspective of Bayesian model choice.
-Such posterior support for simultaneous divergence, even if correct, does not
-provide the biogeographical insights that a researcher who employs \msb seeks
-to gain.
+In the former case, the true (correct) posterior supports models with many
+divergence-time parameters given data generated under temporally random
+divergences, but the problem arises from our inability to accurately estimate
+the posterior.
+In the latter case, estimation error is not the issue, because the true
+posterior supports simultaneous divergences given data from randomly diverged
+taxa; this is the correct (albeit biogeographically misleading) answer given
+the prior assumptions about divergence times imposed by broad uniform priors.
+Such posterior support for simultaneous divergence, even if correct from the
+perspective of Bayesian model choice, does not provide the biogeographical
+insights that a researcher who employs \msb seeks to gain.
 
-The distinction between these potential causes is important for improving our
-ability to estimate shared divergence histories.
-If the cause is numerical, then the model is sound and we simply need to
-improve our estimation procedures.
+While these phenomena are not mutually exclusive, it is important to
+distinguish between them in order to determine how best to improve our ability
+to estimate shared divergence histories.
+If the cause is estimation error, then the model is sound and we need to
+improve our numerical estimation procedures.
 For example, \citet{Hickerson2013} proposed a model-averaging approach that
@@
 estimates of the posterior.
-If the cause is marginal likelihoods, we need to correct the model, because
-improving the estimation machinery will only increase our ability to estimate
+If the cause is marginal likelihoods, we need to correct the model; improving
+the estimation machinery will only increase our ability to estimate
 biogeographically misleading posteriors.
@@
 
-This is not surprising given the rich statistical literature showing that
-marginal likelihoods are very sensitive to the priors used in Bayesian model
-selection \citep[e.g.,][]{Jeffreys1939,Lindley1957}.
-Thus, uniform priors on divergence times can impose a strong ``penalty'' for
-divergence-time parameters, creating a strong preference for simple
-models that are often of interest to biogeographers, even if these models do
-not explain the data very well.
-
-\change{
 Here, we discuss theoretical considerations for using empirically informed
-priors for Bayesian model choice.
-}
-We then evaluate the empirical Bayesian model-averaging approach of
-\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
-\citet{Oaks2012}.
+priors for Bayesian model choice and evaluate the empirical Bayesian
+model-averaging approach of \citet{Hickerson2013} as a potential solution to
+the biases of \msb revealed by \citet{Oaks2012}.
 In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
-made an error by mixing different units of time, which makes the results
-presented in their response difficult to interpret (see Supporting Information
-for details).
-We avoid this error, but still find their approach will often spuriously
-support clustered divergence models.
-Furthermore, the approach provides another means by which the method can
-``escape'' models with large parameter space, which manifests in the preference
-of models that exclude from consideration the true values of the models'
+made an error by mixing different units of time, which invalidates the results
+presented in their response (see Supporting Information for details).
+We avoid this error, but find their approach will often incorrectly support
+(1) clustered divergence models and
+(2) models that exclude from consideration the true values of the models'
 parameters.
-\change{
-Our results, show that it is difficult to choose a uniformly distributed prior
-on divergence times that is broad enough to confidently contain the true
-parameter values while being narrow enough to avoid spurious support for models
-with less parameter space.
-}
-We also explore the predictions of the potential causes of the bias proposed by
-\citet{Oaks2012} and \citet{Hickerson2013}.
+% Our results show that it is difficult to choose a uniformly distributed prior
+% on divergence times that is broad enough to confidently contain the true
+% parameter values while being narrow enough to avoid spurious support for models
+% with less parameter space.
+We also explore the distinct predictions made by the alternative causes of the
+bias proposed by \citet{Oaks2012} and \citet{Hickerson2013}.
 We show the primary cause of the method's tendency to incorrectly infer shared
@@
 We also show that sampling error is not a major contributer to the problem.
-Fortunately, \citet{Oaks2014dpp} has introduced a method that uses more
-flexible probability distributions to accommodate prior uncertainty in
-divergence times without overly inhibiting the marginal likelihoods of models
-with more divergence-time parameters, which increases the method's robustness
-and power to detect temporal variation in divergences.
-
-% \msb \citep{Huang2011} uses approximate Bayesian computation (ABC) to estimate
-% the distribution of divergence times among co-distributed pairs of taxa.
-% It approximates a posterior probability over models that range from a single
-% divergence-time parameter (i.e., simultaneous divergence of all pairs of taxa)
-% to the fully generalized model in which each pair of taxa diverged at a
-% unique time.
-% % A full description of the model is given in \citet{Oaks2012}.
-% % Frequently, the exact values of the divergence times are hard 
-% % to determine because researchers typically lack precise knowledge 
-% % of the rate at which substitutions occur.
-% % The exact divergence time is therefore often viewed as a nuisance parameter,
-% % and researchers focus on the number of divergence events.
-% The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
-% on the fundamental methodological point about the model selection performed in \msb:
-% \begin{itemize}
-%    \item The use of vague priors on the divergence-time parameters can lead to
-%        support for models with few divergence events shared across taxa. Thus,
-%        the primary inference enabled by the approach is very sensitive to the
-%        priors on divergence times.
-% \end{itemize}
-% This is not surprising given the rich statistical literature that shows that
-% marginal likelihoods are very sensitive to the priors used in Bayesian model
-% selection
-% \citep[e.g.,][]{Jeffreys1939,Lindley1957}.
-% Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
-% for models with few divergence parameters is the greater marginal likelihoods
-% of these models under vague uniform priors, relative to more parameter-rich
-% models.
-
-% \change{
-% Models with more divergence-time parameters integrate over \emph{much} greater
-% parameter space with low probability of producing the data, yet relatively high
-% prior density imposed by the uniformly distributed prior.
-% % By using such priors on divergence times, we are essentially telling the
-% % method to favor models with few divergence times even if they do not
-% % explain the data very well.
-% Thus, uniform priors on divergence times can impose a strong ``penalty'' for
-% divergence-time parameters, creating a strong preference for simple
-% models that are often of interest to biogeographers, even if these models do
-% not explain the data very well.
-% It is important to note that this is not due to estimation error
-% (i.e., this phenomenon is extrinsic to the numerical approximation machinery of
-% \msb).
-% Rather, the \emph{true} posterior will tend to support models with few
-% divergence-time parameters under broad uniform priors on divergence times.
-% Such posterior support for simultaneous divergence, even if correct from the
-% perspective of Bayesian model choice, does not provide the biogeographical
-% insights that a researcher who employs \msb seeks to gain.
-% % Ideally, the estimated posterior probability of a simultaneous-divergence model
-% % can be interpreted as an estimate of the probability that this model is correct.
-% % However, when using uniform priors on divergence times, it might be better
-% % interpreted as an estimate of the probability that the simultaneous-divergence
-% % model cannot be completely ruled out as a possible explanation of the data.
-% % Note, this is a statistical issue, extrinsic to \msb, and does not question the
-% % soundness of the numerical approximation machinery of \msb.
-% }
-
-% \citet{Hickerson2013} suggest the problem is the result of a numerical issue
-% intrinsic to \msb, concluding the method's rejection algorithm is inefficient
-% and will be biased toward models with fewer divergence-time parameters, because
-% the parameter space of these models are more densely sampled relative to models
-% with more divergence-time parameters.
-% We agree that sampling error is present in all numerical Bayesian estimation
-% methods, however, we show here that it is not a major contributer to the
-% biases found by \citet{Oaks2012}.
-% % We present results here of analyses that, when combined with the results
-% % of \citet{Oaks2012}, strongly suggest that \msb prefers models with few
-% % divergence-time parameters, because they have greater likelihoods when
-% % averaged over broad uniform priors on divergences times.
-
-% \citet{Oaks2012} suggest alternative prior probability distributions on
-% divergence-times could increase the marginal likelihoods of models with more
-% divergence-time parameters and thus reduce spurious support for models of
-% temporally clustered divergences.
-% Alternatively, \citet{Hickerson2013} present an approach that uses empirically
-% informed uniform priors and Bayesian model-averaging in an attempt to accommodate
-% uncertainty in selecting priors.
-% In general, we agree with the use of Bayesian model averaging to obtain a
-% posterior estimate marginalized over uncertainty in prior choice.
-% However, we question whether adding an additional dimension of model choice
-% that samples over empirically informed uniform priors is a fruitful solution
-% for a model-choice method that is highly sensitive to priors.
-
-% \change{
-% In this paper, we discuss theoretical and practical considerations of using
-% empirically informed priors for Bayesian model choice.
-% }
-% Furthermore, we evaluate the empirical Bayesian model-averaging approach of
-% \citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
-% \citet{Oaks2012}.
-% In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
-% made an error by mixing different units of time, which makes the results
-% presented in their response difficult to interpret (see Supporting Information
-% for details).
-% Here, we avoid this error, but still find their approach to be biased toward
-% clustered divergence models.
-% Furthermore, the approach provides another means by which the method can
-% ``escape'' models with large parameter space, which manifests in the preference
-% of models that exclude from consideration the true values of the model's
-% parameters.
-% \change{
-% Our results, show that it is difficult to choose a uniformly distributed prior
-% on divergence times that is broad enough to confidently contain the true values
-% of parameters while being narrow enough to avoid spurious support of models
-% with less parameter space.
-% }
-% However, recent work shows more flexible probability distributions without a
-% hard upper bound (e.g., gamma) can accommodate prior uncertainty in divergence
-% times without inhibiting the marginal likelihoods of models with more
-% divergence-time parameters, and as a result, increase the method's power to
-% detect temporal variation in divergences
-% \citep{Oaks2014dpp}.
+This is not surprising given the rich statistical literature showing that
+marginal likelihoods are very sensitive to the priors used in Bayesian model
+selection
+\citep[e.g.,][]{Jeffreys1939,Lindley1957}.
+As predicted by this fundamental principle of Bayesian model selection,
+alternative prior assumptions can improve robustness of the method and greatly
+mitigate misleading support for models of shared divergences
+\citet{Oaks2014dpp}.
 
@@
 \change{
-% The results of \citet{Oaks2012} and \citet{Hickerson2013} show the method is
-% highly sensitive to the prior on divergence times.
 \citet{Oaks2012} found \msb incorrectly supports models of temporally
@@
 %
-% \citet{Hickerson2013} repeatedly refer to the prior distributions used by
-% \citet{Oaks2012} for divergence-time parameters as ``poorly selected.''
-% However, this is misleading, as \citet{Oaks2012} discuss in detail how they
-% selected their prior to reflect the large amount of uncertainty about the
-% timing of divergences across all 22 of the taxa in their study (see section
-% ``Specifying and simulating the joint prior'' of \citet{Oaks2012}).
-% \citet{Oaks2012} further discuss how the use of uniform prior distributions in
-% \msb for many of the model's parameters requires investigators to select broad
-% priors to avoid excluding the truth \textit{a priori}, resulting in too much
-% density in improbable regions of parameter space.
-%
-% \citet{Hickerson2013} suggest that \citet{Oaks2012} should have used their
-% data to inform their prior on divergence times (i.e., an empirical Bayesian
-% approach).
-% However, \citet{Oaks2012} did use empirically informed priors to mimic
-% empirical situations in which a large amount of prior information is available;
-% they did so to assess the prior sensitivity of the method and to determine if
-% uniform priors in such situations are narrow enough to avoid spurious support
-% of models with few divergence-time parameters.
-% The results of \citet{Oaks2012} and \citet{Hickerson2013} show the method is
-% highly sensitive to the prior on divergence times.
-% Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
-% divergences under the empirically informed priors.
-%
-% The main argument of \citet{Hickerson2013} is that the priors used by
-% \citet{Oaks2012} were not informative enough.
 They suggest a very narrow, highly informed uniform prior on divergence times

### 2014-07-30 manuscripts/reply/reply-main.tex
More work on intro for reply ms.

@@
 \end{itemize}
-In the former case, the true (correct) posterior supports models with many
-divergence-time parameters given data generated under temporally random
-divergences, but the problem arises from our inability to accurately estimate
-the posterior.
+In the former case, the problem is estimation error; the true (correct)
+posterior supports models with many divergence-time parameters given data
+generated under temporally random divergences, but we are unable to accurately
+estimate the posterior.
 In the latter case, estimation error is not the issue, because the true

### 2014-07-30 manuscripts/reply/reviews/response.tex

%&<latex>
\documentclass[12pt]{article}

\usepackage{parskip}
\input{../../utils/preamble.tex}
\usepackage[numbers,sort&compress]{natbib}
\input{../../utils/macros.tex}

\newcounter{commentCounter}
\newcommand{\revcomment}[1]{{\addtocounter{commentCounter}{1}}
    \medskip \hrule \noindent
\textbf{\arabic{section}.\arabic{commentCounter}}: {\sl #1}\par\xspace}
\newcommand{\response}[1]{{\addtolength{\leftskip}{0.25in} #1\par}\xspace}

\let\quoteOld\quote
\let\endquoteOld\endquote
% \renewenvironment{quote}{\sffamily\slshape\small\quoteOld}{\endquoteOld}
\renewenvironment{quote}{\sffamily\small\quoteOld}{\endquoteOld}

% \makeatletter
% \renewcommand{\section}{\@startsection{section}{1}{0mm}%
%     {-8pt}%
%     {4pt}%
% {\sffamily\large\itshape}}
% \makeatother

\begin{document}
% \raggedright
\setlength{\parindent}{0in}

I greatly appreciate the helpful comments of the editor and reviewers. In this
document, I provide responses to these comments, including descriptions of the
changes I have made.

\section{Editor's Comments}\label{ed}
\setcounter{commentCounter}{0}
\revcomment{
    The most negative review suggests that the authors (and Hickerson et al)
    might have reached a point of diminishing returns, where further exchange
    in the pages of Evolution might not be informative for our readership. My
    own view is somewhat in line with this reviewer. I do think there are some
    new ideas in the present manuscript, and I also want to honor the previous
    AE who solicited an additional response from Oaks and colleagues to the
    last Hickerson et al. paper. But I would also say that after this paper,
    enough is enough, in my opinion \ldots
}
\response{
    We agree. We are very appreciative of this opportunity to clarify some
    aspects of Hickerson et al.'s response, and feel our reply will make a nice
    end to an interesting discussion about Bayesian model selection.
}

\revcomment{
     I also echo the suggestion from reviewer 2 to shorten this paper as much
     as is practical, especially by moving things to the supplement. One can
     safely assume that readers still following the thread have read the
     preceding papers in the series, I think.
}
\response{
    We are greatful for this insight by the editor and reviewers that our
    paper was too long and lacked focus.
    We have greatly reduced the length of the main text and moved some items
    (including 2 figures) to the supplement.
}

\revcomment{
    One small point: there is a typo on line 19: Oaks (2014) should be (Oaks
    2014).
}
\response{
    Corrected. Thank you.
}

\section{Reviewer 1's Comments}\label{rev1}
\setcounter{commentCounter}{0}
\revcomment{
    At this point, I think at least for the readers of Evolution, we are
    reaching the point of diminishing returns and perhaps the discussion has
    burned out. The arguments of how broad priors or mixes of specific and
    diffuse priors yield poor posteriors is indeed important but perhaps
    another venue would be appropriate for this discussion---also with a
    greater emphasis on solutions, if they exist for non-model organisms, to
    common problems affecting most molecular ecologists (particularly regarding
    substitution rate uncertainty).

}
\response{
    We agree that Bayesian statisticians are aware of the issues raised in our
    paper. However, I think among biologists there is common misunderstanding
    about some of the subtle but fundamental differences between Bayesian
    approaches to parameter estimation and model selection. Mainly, the
    implicit and strong statements that priors on the models' parameters can
    make about the probability of the models. Thus, we feel the readership of
    \emph{Evolution} will benefit from issues discussed in the paper.

    We have added emphasis of solutions to the problem in several places in the
    manuscript (abstract, introduction, and conclusions). The solution was the
    focus of a different paper \citep{Oaks2014dpp} that introduces an improved
    method for estimating shared divergence histories. We now draw attention to
    this positive outcome from this work.
}

\section{Reviewer 2's Comments}\label{rev2}
\setcounter{commentCounter}{0}

\revcomment{
    The paper is rather long for simply arguing against a model averaging
    approach. I recommend to move the coin example to the supp info, as this is
    a simple text book example that does require to be published in a journal
    such as evolution, but is still a nice illustration so that it should be
    published as supp info to the paper. Also, I think the paper sometimes
    reads like a direct accusation of the findings in Hickerson et al, to agree
    to which they might be more emotional than required for the presented
    arguments. Paragraphs such as L96-100, L104-108, L222-237 or L287-289
    (among others later in the text and, especially, in the supp info) for
    instance, could simply be omitted or shortened, which would make reading
    the paper much more enjoyable, as it would focus on the valid arguments,
    instead of highlighting subtleties only personally important to people
    included in the debate.

}
\response{
    We thank Reviewer 2 for this insight. We agree that the paper was too long
    and lacked focus. Reviewer 2 also points out several sections that might be
    interpreted as emotionally charged or polemical, which we appreciate. We
    did not intend for our response to read as an ``attack'' of the work of
    Hickerson et al., but Reviewer 2 helped us to realize that wording in a
    response paper needs to be carefully chosen to avoid this interpretation.

    We have moved the coin-flipping example to the supplement.

    In regard to the sections specifically highlighted by Reviewer 2:
    \begin{description}
        \item[L96-100 and L104-108]
            We have reduced this section to only three sentences that summarize
            only the information necessary to introduce our subsequent
            discussion.
        \item[L222-237]
            We see how this paragraph can be interpreted as polemical. However,
            this brief summary of the methods and results of Hickerson et al.'s
            model-averaging approach is needed to set up the motivation for our
            subsequent analyses.
        \item[L287-289]
            We agree this section was superfluous. We have removed it.
    \end{description}

    We have also removed and edited many other sections of the paper to try to
    eliminate sections that might seem overly negative. Specifically, we
    removed all quotations of sections from Hickerson et al. While we only
    intended to accurately represent their positions, we felt this is too
    easily interpreted as contentious.
}

\revcomment{
    While criticizing the approach proposed by Hickerson et al. and admitting
    the power problem of msBayes, the paper falls short of providing much of an
    alternative. However, I do think that there could be other ways of solving
    the problem of too many parameters for few summary statistics (section 6).
    This is especially true since ABC is known to be particularly sensitive to
    the choice of summary statistics for model choice (see Roberts et al. PNAS
    2011). I think the paper would be much stronger, if it could also discuss
    (but not necessarily test or implement) potential solutions to the apparent
    problem.
}
\response{
    This is a great point, and we have added emphasis to solutions to the
    problem, as recommended by Reviewer 2. As mentioned above, solutions to
    this problem were the focus of a separate paper, and were too extensive to
    discuss in detail in the response paper. However, we highlight the main
    points of this paper to make the response more positive, and to to refer
    readers to an improved methodology.
}

\revcomment{
    One alternative suggestion could be to formulate the problem as a
    hierarchical model, in which one seeks only to infer mean and variance in
    divergence times. Such a model would be much easier to fit, as the summary
    statistics ordering problem discussed (and solved) in the original msBayes
    paper could be circumvented easily, as the model would not seek to infer
    taxon specific divergence times. Also, it could then be compared to a model
    of zero variance using either Bayes factors, or by assessing the posterior
    probability of the variance to be zero using a simulation / bootstrap
    approach. If a model of co-divergence is preferred, the inferred posterior
    on the mean divergence would directly serve as an estimate of the joint
    divergence time. If independent divergences are inferred, then a simpler
    model of a single taxon pair could be used to infer the divergence times of
    each taxon pair individually, as they were found to be independent. For
    such a model, the same ABC simulations could be reused for all pairs,
    making this step computationally very cheap. But I'm sure there would also
    be other alternatives.
}
\response{
    These are great points raised by Reviewer 2. Many of these are
    discussed in detail in a separate paper \citep{Oaks2014dpp}.
    We highlight the improved methodology presented there to add a positive
    note to the response.
}

\revcomment{
    A minor issue on L53: while it is true that all numerical methods suffer
    from large prior ranges, it is also true that effect is pronounced in ABC,
    as the sampling occurs from the prior, and not the posteriors, as is done
    in techniques such as MCMC. Hence, it is not a problem that is completely
    extrinsic to the ABC machinery.
}
\response{
    We thank Reviewer 2 for pointing out this section, which was admitedly very
    confusing. We have re-written this section (and almost all of the
    introduction) to try to make this distinction clearer. Our goal is to
    distinguish two different ways that broad uniform priors could cause the
    poor behavior of msBayes that was shown in the origin Oaks et al. paper.
    \begin{enumerate}
        \item As pointed out by reviewer 2, these priors could cause numerical
            issues, where the simple ABC rejection machinery is unable to
            sufficiently sample the space of the model in manageable
            computational time.
        \item The priors can cause rich models to have very small (but correct)
            marginal likelihoods.
    \end{enumerate}

    These issues are not mutually exclusive, but in the former the true
    posterior would support heterogeneity in divergence times in the face of
    random divergences (i.e., the problem only arises from our inability to
    estimate that posterior).
    Whereas in the later, the true posterior (i.e., an analytical solution)
    would correctly tell us that the most probable model is simultaneous
    divergence despite random variation in divergence times. I.e., this is
    actually the correct answer under the prior assumptions from the
    perspective of Bayesian model choice. Unfortunately, this correct answer
    does not provide the biogeographical insights we are using the method to
    gain.

    This is an important distinction, because if the prior conditions are such
    that we are in situation 2, any arguments about the estimation procedure
    are moot, because even the true posterior is biogeographcially misleading.
}

\bibliographystyle{../../bib/bmc-mathphys.bst}
\bibliography{../../bib/references}

\end{document}

### 2014-07-30 manuscripts/reply/reply-abstract.tex
More updating of reply ms.
Finishing draft of reply ms overhaul.

--- 2014-07-27 manuscripts/reply/reply-abstract.tex
@@
-% Biogeographers frequently seek to explain population and species
-% differentiation on geographical phenomena.
 Establishing that a set of population splitting events occurred at the same
@@
 affected by the same geographic event.
-% Unfortunately, estimating divergence times precisely when one lacks
-% precise information about the rate of molecular evolution is a notoriously
-% difficult problem in evolutionary biology.
-% \citet{Huang2011} introduced an approximate-Bayesian approach, \msb, for
-% estimating the probabilities of models in which multiple sets of taxa diverge
-% at the same time.
-% This method  does not require a common set of character data that links all
-% of the taxa under investigation, but does require the relative mutation
-% rates among the taxa and loci to be assumed \emph{a priori}.
 Recently, \citet{Oaks2012} used an approximate-Bayesian model choice method
@@
 Here we demonstrate that the approach of \citet{Hickerson2013} will often
-incorrectly support shared divergence events and is dangerous in the sense that
-the empirically-derived priors often exclude from consideration the true values
-of the models' parameters.
+erroneously support shared divergence events and is dangerous in the sense that
+the empirically-derived uniform priors often exclude from consideration the
+true values of the models' parameters.
 On a more fundamental level, we question the value of adopting an empirical
 Bayesian stance for this model-choice problem.
-% , because it can mislead model
-% posterior probabilities, which are inherently measures of belief in the models
-% after prior knowledge is updated by the data.
 The robust Bayesian approach of conducting analyses under a variety of priors
@@
 inference.
-Furthermore, simulations provide insight into the temporal resolution of the
-method, which helps guide interpretation of results.
-Fortunately, as predicted by fundamental principles of Bayesian model choice,
-more appropriate priors on the models' parameters can accommodate prior
-uncertainty while increasing the method's robustness and power to detect
-temporal variation in divergences
-\citep{Oaks2014dpp}.
+% Furthermore, simulations provide insight into the temporal resolution of the
+% method, which helps guide interpretation of results.
+As predicted by fundamental principles of Bayesian model choice, more flexible
+distributions can accommodate prior uncertainty about the models' parameters
+while increasing the method's robustness and power to detect temporal variation
+in divergences \citep{Oaks2014dpp}.

### 2014-07-30 manuscripts/reply/reply-main.tex
Finishing draft of reply ms overhaul.

@@
 They found the method was very sensitive to prior assumptions and often
-incorrectly supported shared divergences across taxa that diverged randomly
-over broad time periods, to which \citet{Hickerson2013} responded.
+supports shared divergences across taxa that diverged randomly over broad time
+periods, to which \citet{Hickerson2013} responded.
 \citet{Oaks2012} and \citet{Hickerson2013} agree on the fundamental
 methodological point about the model selection performed in \msb:
-% The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
-% on the fundamental methodological point about the model selection performed in \msb:
 \begin{itemize}
@@
         rejection algorithm implemented in \msb is unable to adequately sample
-        the space of the models within reasonable computational time.
-        They propose the method incorrectly supports models with fewer
-        divergence-time parameters because they are better sampled.
-    \item \citet{Oaks2012} suggest the cause is more fundamental; the broad
-        uniform priors simply lead to very small marginal likelihoods of models
-        with many divergence-time parameters.
-        Models with more divergence-time parameters integrate over \emph{much}
-        greater parameter space with low probability of producing the data, yet
-        relatively high prior density, resulting in small marginal likelihoods.
+        the space of the models within reasonable computational time, which
+        leads to support for models with fewer divergence-time parameters
+        because they are better sampled.
+    \item \citet{Oaks2012} suggest the cause is more fundamental; the uniform
+        priors on divergence times simply lead to very small marginal
+        likelihoods of models with many divergence-time parameters. These
+        models are ``averaged'' over much greater parameter space with large
+        prior weight and small likelihood \citep{Jeffreys1939,Lindley1957}.
 \end{itemize}
@@
 posterior supports simultaneous divergences given data from randomly diverged
-taxa; this is the correct (albeit biogeographically misleading) answer given
-the prior assumptions about divergence times imposed by broad uniform priors.
+taxa.
+I.e., this is the correct (albeit biogeographically misleading) answer given
+the prior assumption that divergence times are broadly and uniformly
+distributed.
 Such posterior support for simultaneous divergence, even if correct from the
@@
 improve our numerical estimation procedures.
-For example, \citet{Hickerson2013} proposed a model-averaging approach that
-uses narrow, empirically informed uniform priors in the hope that with less
-parameter space to sample, the rejection algorithm will produce better
-estimates of the posterior.
-If the cause is marginal likelihoods, we need to correct the model; improving
-the estimation machinery will only increase our ability to estimate
-biogeographically misleading posteriors.
+For example, Markov chain or sequential Monte Carlo algorithms might improve
+the efficiency of sampling from the posterior.
+Rather than alter the sampling algorithm, \citet{Hickerson2013} tried a
+model-averaging approach that uses narrow, empirically informed uniform priors
+in the hope that with less parameter space to sample, the rejection algorithm
+will produce better estimates of the posterior.
+Here, we discuss theoretical considerations for using empirically informed
+priors for Bayesian model choice and evaluate the approach of
+\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
+\citet{Oaks2012}.
+In their analyses, \citet{Hickerson2013} made an error by mixing different
+units of time, which invalidates the results presented in their response (see
+Supporting Information for details).
+We correct this error, but still find their approach will often support
+(1) clustered divergence models when divergences are random, and
+(2) models that exclude from consideration the true values of the
+parameters.
+The behavior of their approach also suggests that estimation error is not the
+primary cause of the issue.
+
+If support for models with few divergence events is caused by their larger
+marginal likelihoods, we need to correct the model; improving the estimation
+machinery will only increase our ability to estimate true posteriors that are
+difficult to interpret.
 Accordingly, \citet{Oaks2014dpp} has introduced a method that uses more
@@
 divergence times without overly inhibiting the marginal likelihoods of models
-with more divergence-time parameters, which increases the method's robustness
-and power to detect temporal variation in divergences.
-
-Here, we discuss theoretical considerations for using empirically informed
-priors for Bayesian model choice and evaluate the empirical Bayesian
-model-averaging approach of \citet{Hickerson2013} as a potential solution to
-the biases of \msb revealed by \citet{Oaks2012}.
-In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
-made an error by mixing different units of time, which invalidates the results
-presented in their response (see Supporting Information for details).
-We avoid this error, but find their approach will often incorrectly support
-(1) clustered divergence models and
-(2) models that exclude from consideration the true values of the models'
-parameters.
-% Our results show that it is difficult to choose a uniformly distributed prior
-% on divergence times that is broad enough to confidently contain the true
-% parameter values while being narrow enough to avoid spurious support for models
-% with less parameter space.
-We also explore the distinct predictions made by the alternative causes of the
-bias proposed by \citet{Oaks2012} and \citet{Hickerson2013}.
-We show the primary cause of the method's tendency to incorrectly infer shared
-divergences is the uniform priors on divergence times leading to small marginal
-likelihoods of models with more divergence-time parameters.
-We also show that sampling error is not a major contributer to the problem.
+with more divergence-time parameters.
+This greatly increases the method's robustness and power to detect temporal
+variation in divergences \citep{Oaks2014dpp}.
 This is not surprising given the rich statistical literature showing that
@@
 \citep[e.g.,][]{Jeffreys1939,Lindley1957}.
-As predicted by this fundamental principle of Bayesian model selection,
-alternative prior assumptions can improve robustness of the method and greatly
-mitigate misleading support for models of shared divergences
-\citet{Oaks2014dpp}.
+
+We also explore the distinct predictions made by the alternative causes of the
+bias proposed by \citet{Oaks2012} and \citet{Hickerson2013}.
+We show the primary cause of the method's tendency to support models of shared
+divergences is the large marginal likelihoods of these models under the prior
+assumption of uniformly distributed divergence times.
+We also show that sampling error is not a major contributer to the problem.
 
 \section{The potential implications of empirical Bayesian model choice}
-\change{
-\citet{Oaks2012} found \msb incorrectly supports models of temporally
+\citet{Oaks2012} found \msb spuriously supports models of temporally
 clustered divergences even when using a prior on divergence times that was
@@
 \citet{Oaks2012} were not informative enough.
-}
-%
 They suggest a very narrow, highly informed uniform prior on divergence times
@@
 exhibit favorable frequentist properties.
-% Furthermore, many post-hoc correction methods have been developed for
-% estimating confidence-intervals from empirical Bayes estimates of
-% distributions that often exhibit well-behaved frequentist coverage
-% probabilities
 \citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.
@@
 Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the posterior
-probability of a model, is not a function of \myTheta{}, because the parameters
+probability of a model, is not a function of \myTheta{} because the parameters
 are integrated out of the marginal likelihoods of the candidate models.
@@
 clear true values.
-% There is no such justification for empirical Bayesian model choice, because
-% there are no true values for the model probabilities being estimated.
 Model posterior probabilities are inherently measures of our belief in the
@@
 problems with using narrow, empirically informed, uniform priors.
-% for a method that has already been shown to be biased toward models with less
-% parameter space.
 The results of Hickerson et al.'s (\citeyear{Hickerson2013}) reanalysis of the
@@
 divergence times for at least some of the Philippine taxa.
-%, a major problem when using uniform priors.
 \citet{Hickerson2013} set this prior to match the 95\% highest posterior
@@
 
-\change{
 The strong preference for the narrowest prior on divergence times suggests the
@@
 space and, as a consequence, will estimate model-averaged posteriors dominated
-by models that exclude true values of the model's parameters.
-}
+by models that exclude true values of the parameters.
 We explored this possibility in two ways.
@@
 less parameter space, we expect \msb will preferentially sample from this
-erroneous model yielding a posterior that is incorrect (i.e., the
-model-averaged posterior will be dominated by an incorrect model that excludes
+erroneous model yielding a posterior that is misleading (i.e., the
+model-averaged posterior will be dominated by a model that excludes
 the truth).
@@
 Following \citet{Hickerson2013}, each of these models had the following
-priors on divergence-time parameters:
+priors on divergence times:
 $M_1$, $\divt{} \sim U(0, 0.1)$;
@@
 strongly prefers the prior model with the narrowest distribution on divergence
-times across all three of our analyses, even when this model is incorrect and
-excludes the true divergence times of the Philippine taxa
+times across all three of our analyses, even when this model excludes the true
+divergence times of the Philippine taxa
 (Table~\ref{tabModelChoiceEmpirical}).
@@
 priors are too narrow (Figure~S\ref{figPCA}).
-Rather, the graphs suggest these incorrect priors are ``better fit''
+Rather, the graphs suggest these invalid priors are ``better fit''
 (Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
 \citet{Oaks2012} (Figure~S\ref{figPCA}D--F).
-
-% Given that the results of \citet{Hickerson2013} strongly prefer the models with
-% the narrowest prior on divergence times, it seems quite likely that their
-% model-averaged results are dominated by models that exclude true divergence
-% times, making their results difficult to interpret.
 
@@
 (Table~\ref{tabModelChoiceEmpirical}).
-We then repeated the analysis described above using $1\e6$ random samples from
-each of the prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, retaining 1000
+We then repeated the model-averaging  analysis described above, retaining 1000
 posterior samples for each of the 1000 simulated datasets.
-
 For each simulation replicate, we estimated the Bayes factor in favor
@@
 model.
+
 Our results show that the model-averaging approach of \citet{Hickerson2013}
@@
 true divergence times (Figure~\ref{figExclusionSimTau}).
-We also used the mode estimate of the preferred model for each replicate to
-estimate the number of true parameter values excluded, which produced nearly
-identical results.
 Importantly, the posterior probability of excluding at least one true parameter
@@
 The consequence of this approach is obtaining a model-averaged posterior
-estimate that is heavily weighted toward incorrect models that exclude true
-values of model parameters.
-% \subsection{Additional thoughts on empirical priors in Bayesian model choice}
+estimate that is heavily weighted toward models that exclude true
+values of the parameters.
 This is not a general critique of Bayesian model averaging.
@@
 
-\change{
 When using uniformly distributed priors, the alternative to capturing prior
@@
 \citep{Oaks2014dpp}.
-}
-
-% Given the theoretical and practical issues with empirical Bayes approaches to
-% Bayesian model choice, it is clear why one should use caution before overly
-% criticizing an investigator's choice of priors \emph{after} having seen the
-% resulting posterior.
-% As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a
-% large amount of uncertainty regarding the divergence times of the 22 population
-% pairs under study.
-% Two of these pairs represent distinct species, and the taxonomy of many groups
-% in the Philippines has repeatedly been shown to mask deeply divergent lineages
-% \citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
-% Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013,Rafe2013AREES,Siler2014kikuchii}.
 
@@
 such risk if the approach does indeed increase power to detect temporal
-variation in divergences and thus decrease bias toward clustered divergence
-models.
+variation in divergences.
 We assess this possibility using simulations.
@@
 All other parameters were identically distributed as the prior models.
-As above, we generated a prior sample of $5\e6$ total samples from the five
-prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$
-(Table~\ref{tabModelChoiceEmpirical}).
-For each simulated dataset, we approximated the posterior
-by retaining 1000 samples from the prior with the smallest Euclidean distance
-from the true summary statistics, as described above.
-In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
-posterior samples for each of them.
+As above, we generated $5\e6$ samples from the prior models in
+Table~\ref{tabModelChoiceEmpirical}.
+For each of the 6000 simulated datasets, we approximated the posterior
+by retaining 1000 samples from the prior.
 
@@
 
-% The method is most likely to infer the extreme case of a single divergence
-% event shared across all 22 taxa when populations diverged randomly over the
-% past \globalcoalunit generations.
-% When divergences are random over the past $8\globalpopsize$ generations, the
-% most likely inference is only two divergence events, and a single
-% divergence is still estimated in more than 10\% of the replicates.
-% It is interesting to note that as \divt{max} increases, but before the
-% estimates are finally pulled away from $\numt{} = 1$, the distribution of
-% \numt{} estimates closely mirror the ``U-shaped'' prior on divergence models
-% used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
-% (\citeyear{Oaks2012}) Figure 5B).
-% This suggests the U-shaped prior coupled with the small marginal likelihoods
-% of models with many divergence parameters is a major cause of the method's
-% bias toward clustered divergence models; this is consistent with
-% \citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.
-
-% Looking at our simulation results in terms of the posterior probability of the
-% dispersion index of divergence times supporting the extreme case of one
-% divergence event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the
-% method strongly supports one divergence in greater than 27\% of the replicates
-% across all the \divt{max} we simulated (Figure~\ref{figPower4}E--H \&
-% S\ref{figPower6}G--L);
-% following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
-% the criterion for incorrect inference of a single divergence event.
-% Furthermore, there is strong support for a single divergence event in
-% more than 90\% of the replicates when divergences are random over the past
-% $2.4\globalpopsize$ generations, and more than 60\% when over the past
-% $3.2\globalpopsize$ generations (Figure~\ref{figPower4}E--H \&
-% S\ref{figPower6}G--L).
-
-% Given the empirically Bayesian model-averaging approach of \citet{Hickerson2013}
-% lacks power to detect random variation in divergence times over such
-% timescales, it is difficult to justify the risk of the approach to exclude
-% regions of parameter space containing the truth.
-
 Also, the results of our power analyses further demonstrate the propensity of
@@
 the \divt{max} the preferred model will exclude a large proportion of the true
-divergence-time values (Figure~\ref{figPowerExclusion4}A--D \&
+divergence times (Figure~\ref{figPowerExclusion4}A--D \&
 S\ref{figPowerExclusion6}A--F).
@@
 (Figure~\ref{figPowerExclusion4}E--H \& S\ref{figPowerExclusion6}G--L).
-% Only when the data are identically distributed as one of the prior models does
-% the method avoid excluding the truth more than 5\% of the time
-% (Figure~\ref{figPowerExclusion4}C).
-% Again, this demonstrates the method's sensitivity to priors.
 
@@
 hundreds of thousands of generations or more.
-\citet{Oaks2012} performed similar power analyses under three uniform
-divergence-time priors as narrow as 0--5 coalescent units, and found the method
-was able to detect multiple events among 22 taxa when divergences were random
-over millions of generations.
-It is important that investigators perform simulations to determine the
+It is important that investigators perform such simulations to determine the
 method's power for their dataset, and decide if \msb has sufficient temporal
@@
 it did not.
-When doing so, it is important to consider what prior conditions are
-relevant to their empirical system.
-A divergence-time prior of 0--5 coalescent units is quite narrow considering
-that this expresses the prior belief that all 22 taxon pairs diverged within
-this window.
-Certainly, it is  rare for there to be enough \emph{a priori} information to
-be certain that all taxa diverged within the last 4$\globalpopsize$ generations
-(i.e., 0--1 coalescent units).
+When doing so, it is important to consider what prior conditions are relevant
+to the empirical system.
+It is rare for there to be enough \emph{a priori} information to be certain
+that all taxa diverged within the last 4$\globalpopsize$ generations (i.e.,
+0--1 coalescent units).
 Also, it seems unlikely that when such prior information is available that
@@
 into the evolutionary history of the taxa.
-% ; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), this
-% translates to a temporal resolution of 3 million generations or more.
 
@@
 To determine how best to improve the behavior of \msb, it is important to
-determine the mechanism by which broad uniform priors cause incorrect support
-for clustered models of divergence.
+determine the mechanism by which broad uniform priors cause support for
+clustered models of divergence.
 It is well established that vague priors can be problematic in Bayesian model
@@
 over much greater parameter space, all with equal prior weight, and much of it
-with low likelihood.
+with small likelihood.
 In light of this fundamental statistical issue, it is not surprising that the
@@
 
-However, \citet{Hickerson2013} conclude that the bias is due to insufficient
-sampling.
+However, \citet{Hickerson2013} conclude that the bias is caused by estimation
+error due to insufficient sampling.
 They argue the widest of the three priors used by \citet{Oaks2012} would
@@
 Hickerson et al.'s
-(\citeyear{Hickerson2013}) argument only applies to one of the three priors
-used by \citet{Oaks2012}.
+(\citeyear{Hickerson2013}) argument only applies to one of the \emph{three}
+priors used by \citet{Oaks2012}.
 Under these assumptions, the narrowest prior on divergence times used by
@@
 explored by \citet{Oaks2012}.
-\citet{Hickerson2013} point out a case where the narrow prior performs
+\citet{Hickerson2013} point out a case where the narrowest prior performs
 slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
@@
 in \msb, and there are many cases after \citet{Oaks2012} corrected the 
-bug where the narrow prior performs slightly worse (see panels D--J of
+bug where the narrowest prior performs slightly worse (see panels D--J of
 Figures 3 and S12).
@@
 samples drawn from the prior.
-Furthermore, if sampling error is biased toward models with less
-parameter space, as suggested by \citet{Hickerson2013}, we expect to see
+Furthermore, if sampling error is \emph{biasing} estimates toward models with
+less parameter space, as suggested by \citet{Hickerson2013}, we expect to see
 support for these models decrease as sampling increases.
@@
 (Figure~S\ref{figSamplingError}) we do not see the trend predicted by
-\citet{Hickerson2013} in either the unadjusted or GLM-regression-adjusted
-estimates.
+\citet{Hickerson2013}.
 While sampling error is always a reality, it does not appear to be playing a
@@
 assess the model-choice behavior of \msb when the prior model is correct.
-The results confirm the prediction as \msb actually tends to underestimate
-the probability of the one-divergence model (see Figure 4 of
-\citet{Oaks2012}).
+The results confirm the prediction:
+\msb estimates the probability of the one-divergence model quite well when the
+prior is correct (see Figure 4 of \citet{Oaks2012}).
 We confirmed this same behavior for the model-averaging approach used by
@@
 These results are not clearly predicted if insufficient sampling was causing
-the bias.
+estimation bias.
 Even when the prior is correct, due to the discrete uniform prior on the number
@@
 
-% \section{General thoughts on the model of \texttt{msBayes}}
-\section{A difficult inference problem}
-Our findings are not surprising in light of the difficult inference problem
-with which \msb is faced.
-% To get a sense of the difficulty of this problem, we tally up all the free
-% parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
-% Under the simplest model in \msb (i.e., assuming no migration and no
-% intra-locus recombination), the number of parameters for each taxon pair
-% include:
-% The population sizes of the ancestral and descendant populations
-% (\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
-% the magnitude of population contraction in each of the descendant
-% populations (\bottleScalar{1}{} and \bottleScalar{2}{}), the
-% timing of these contractions (\bottleTime{}), and the $N-1$ node heights
-% (coalescent times) of the gene tree that gave rise to the $N$ gene
-% copies sampled from both populations.
-% Lastly, there are between one and 22 divergence time parameters \divt{} in
-% the vector \divtvector.
-When applying \msb to the dataset of \citet{Oaks2012} with 22 taxon
-pairs, there are 581--602 free parameters that model highly stochastic
-coalescent and mutational processes.
-Under this rich stochastic model, the method is estimating the
-probability of 1002 divergence models \citep[i.e., the number of integer
-partitions of $Y=22$;][]{Oaks2012}.
-Furthermore, all the information in the sequence alignment of each taxon pair
-is distilled into four summary statistics.
-This gives us a total of 88 summary statistics (four from each of the 22 taxon
-pairs) that contain minimal information about many of the $\approx 600$
-parameters in the model.
-More summary statistics can be used in \msb, but most are highly correlated
-with the four default statistics, and thus contribute little additional
-information about the parameters from the sequence data.
-The large number of parameters and divergence models relative to the amount of
-information in the data is undoubtedly a key reason the method lacks robustness
-to prior conditions.
-% Robustness is an important characteristic of a method to gauge its
-% applicability to real-world data, because we know the model and priors will be
-% wrong to some degree.
-
-The model-averaging approach of \citet{Hickerson2013} adds an additional
-dimension of model choice to sample over eight prior models.
-This adds an additional free parameter to the model and, more importantly,
-forces the model to sample over 8016 unique models for 22 taxa.
-While this approach is trivial to implement, it is a non-trivial extension of
-the model.
-In theory, the model-averaging approach of \citet{Hickerson2013} is very
-appealing.
+\section{Improving inference of shared divergences}
+In theory, the model-averaging approach of \citet{Hickerson2013} is appealing.
 It leverages a great strength of Bayesian statistical procedures, namely the
@@
 nuisance parameters.
-However, when averaging over both narrow-empirical and diffuse uniform priors
-for a parameter that is expected to have a very non-uniform likelihood density,
-and in the context of a model-choice method that is highly sensitive to priors,
-it is not surprising that the approach struggles.
+However, when sampling over models with narrow-empirical and diffuse uniform
+priors for a parameter that is expected to have a very non-uniform likelihood
+density, models that exclude the true values of the parameters we aim to
+estimate will often have the largest marginal likelihood.
 
@@
 distributions from the exponential family.
-If we look at the prior distribution on divergence-time parameters imposed by
+If we look at the prior distribution on divergence times imposed by
 the model-averaging approach of \citet{Hickerson2013} we see it is a mixture of
@@
 This looks very much like an exponential distribution, except that in any state
-of the model, all the divergence-time parameters are restricted to the hard
+of the model, all the divergence times are restricted to the hard
 bounds of one of the uniform distributions.
@@
 density.
-Indeed, \citet{Oaks2014dpp} has shown that the use of more flexible
-distributions in place of uniform priors improves the power of the method to
-detect temporal variation in divergences and reduces incorrect support for
-clustered divergences.
+Indeed, \citet{Oaks2014dpp} has shown that the use of gamma distributions in
+place of uniform priors improves the power of the method to detect temporal
+variation in divergences and reduces erroneous support for clustered
+divergences.
 
 \section{Conclusions}
-% We demonstrate how the approximate Bayesian model-choice method implemented in
-% \msb can be strongly biased away from models with greater parameter space.
 We demonstrate how the approximate Bayesian model-choice method implemented in
@@
 of the parameters.
-Also, our results show that it is difficult to choose an uniformly distributed
+Our results show that it is difficult to choose an uniformly distributed
 prior on divergence times that is broad enough to confidently contain the true
@@
 models with less parameter space.
+More generally, it is important to carefully choose prior assumptions about
+parameters in Bayesian model selection; these prior assumptions strongly
+influence the posterior probabilities we seek to estimate.
 
-\change{
-% Whether or not one chooses to use empirically informed priors,
-% We do not assume results of all previous applications of \msb results are invalid, as
-% suggested by \citet{Hickerson2013}.
-% We do conservatively recommend that
 The common inference of temporally clustered divergences \citep{Barber2010,
@@
 conditions.
-}
 Fortunately, \citet{Oaks2014dpp} has shown that alternative probability

### 2014-07-30 manuscripts/reply/reply-si-main.tex
Finishing draft of reply ms overhaul.

@@
 
-%CWL: This seems to specific, not sure about including
 \section{Validation analyses}
-% \highLight{Not sure where to put this section}.
-% \highLight{This justifies presenting the unadjusted results in this paper.}
-% \highLight{Is it worth including just for this purpose?}
 Following \citet{Oaks2012}, we characterize the model-choice behavior of the
@@
 the model probabilities are biased.
+
+\section{A difficult inference problem}
+In the main text, we discuss how the prior assumption of uniformly distributed
+divergence times in \msb leads to posteriors that are difficult to interpret.
+However, it is also important to consider the difficult inference problem with
+which \msb is faced.
+When applying \msb to the dataset of \citet{Oaks2012} with 22 taxon
+pairs, there are 581--602 free parameters that model highly stochastic
+coalescent and mutational processes.
+Under this rich stochastic model, the method is estimating the
+probability of 1002 divergence models \citep[i.e., the number of integer
+partitions of $Y=22$;][]{Oaks2012}.
+Furthermore, all the information in the sequence alignment of each taxon pair
+is distilled into four summary statistics.
+This gives us a total of 88 summary statistics (four from each of the 22 taxon
+pairs) that contain minimal information about many of the $\approx 600$
+parameters in the model.
+More summary statistics can be used in \msb, but most are highly correlated
+with the four default statistics, and thus contribute little additional
+information about the parameters from the sequence data.
+The large number of parameters and divergence models relative to the amount of
+information in the data is undoubtedly another reason the method lacks
+robustness to prior conditions.
 

### 2014-07-30 manuscripts/reply/reply-main.tex
More minor tweaks to intro of reply ms.

@@
 marginal likelihoods, we need to correct the model; improving the estimation
-machinery will only increase our ability to estimate true posteriors that are
-difficult to interpret.
+machinery will only increase our ability to estimate misleading posteriors.
 Accordingly, \citet{Oaks2014dpp} has introduced a method that uses more

### 2014-07-31 manuscripts/reply/reply-acknowledgments.tex
More editing of reply ms.
Changing discussions of "estimation error" to "numerical
approximation error" as per Vladimir's comments. Adding Vlad
to the acknowledgments as well.

--- 2014-03-01 manuscripts/reply/reply-acknowledgments.tex
@@
 We thank Melissa Callahan, Jake Esselstyn, Cameron Siler, Mark Holder, Rafe
-Brown, Emily McTavish, Daniel Money, Jordan Koch, and Adam Leach\'{e} for
-insightful comments that greatly improved this work.
+Brown, Emily McTavish, Daniel Money, Jordan Koch, Adam Leach\'{e}, Vladimir
+Minin, Luke Harmon, and three anonymous reviewers for insightful comments that
+greatly improved this work.
 We thank Michael Hickerson and co-authors for generously providing their data.

### 2014-07-31 manuscripts/reply/reply-main.tex
More editing of reply ms.
Changing discussions of "estimation error" to "numerical
approximation error" as per Vladimir's comments. Adding Vlad
to the acknowledgments as well.

@@
 the behavior of the \msb approach using computer simulations.
-They found the method was very sensitive to prior assumptions and often
+They found the method is very sensitive to prior assumptions and often
 supports shared divergences across taxa that diverged randomly over broad time
@@
         the space of the models within reasonable computational time, which
-        leads to support for models with fewer divergence-time parameters
+        leads to bias toward models with fewer divergence-time parameters
         because they are better sampled.
@@
 \end{itemize}
-In the former case, the problem is estimation error; the true (correct)
-posterior supports models with many divergence-time parameters given data
-generated under temporally random divergences, but we are unable to accurately
-estimate the posterior.
-In the latter case, estimation error is not the issue, because the true
+In the former case, the problem is numerical approximation error due to
+insufficient computation;
+the true (correct) posterior supports models with many divergence-time
+parameters given data generated under temporally random divergences, but we are
+unable to accurately approximate the posterior.
+In the latter case, numerical error is not the issue, because the true
 posterior supports simultaneous divergences given data from randomly diverged
@@
 to estimate shared divergence histories.
-If the cause is estimation error, then the model is sound and we need to
-improve our numerical estimation procedures.
-For example, Markov chain or sequential Monte Carlo algorithms might improve
-the efficiency of sampling from the posterior.
+If the cause is numerical approximation error, then the model is sound and we
+need to increase our computational effort or improve our Monte Carlo
+integration procedures.
+For example, Markov chain or sequential Monte Carlo algorithms might sample the
+posterior more efficiently than the simple Monte Carlo rejection sampler
+currently implemented in \msb.
 Rather than alter the sampling algorithm, \citet{Hickerson2013} tried a
@@
 parameters.
-The behavior of their approach also suggests that estimation error is not the
-primary cause of the issue.
+The behavior of their approach also suggests that insufficient computation is
+not the primary cause of the issue.
 
 If support for models with few divergence events is caused by their larger
-marginal likelihoods, we need to correct the model; improving the estimation
-machinery will only increase our ability to estimate misleading posteriors.
+marginal likelihoods, we need to correct the model; improving the numerical
+approximation machinery will only increase our ability to estimate the true
+posterior that leads us to the wrong interpretation about evolutionary history.
 Accordingly, \citet{Oaks2014dpp} has introduced a method that uses more
@@
 assumption of uniformly distributed divergence times.
-We also show that sampling error is not a major contributer to the problem.
+We also show that numerical approximation error is not a major contributer to
+the problem.
 
@@
 
-However, \citet{Hickerson2013} conclude that the bias is caused by estimation
-error due to insufficient sampling.
+However, \citet{Hickerson2013} conclude that the bias is caused by
+numerical approximation error due to insufficient computation.
 They argue the widest of the three priors used by \citet{Oaks2012} would
-infrequently produce samples with many independent population divergence times
-as recent as the estimated gene divergence times presented in \citet{Oaks2012}.
+infrequently produce random samples with many independent population divergence
+times as recent as the estimated gene divergence times presented in
+\citet{Oaks2012}.
 However, this argument assumes the gene divergence times presented in
@@
 
-To disentangle whether model likelihoods or insufficient prior sampling is to
+To disentangle whether model likelihoods or numerical approximation error is to
 blame for the method's erroneous support for simple models, we must look at the
 different predictions made by these two phenomena.
-For example, insufficient prior sampling should create large variance among
-posterior estimates and cause analyses to be highly sensitive to the number of
-samples drawn from the prior.
-Furthermore, if sampling error is \emph{biasing} estimates toward models with
-less parameter space, as suggested by \citet{Hickerson2013}, we expect to see
-support for these models decrease as sampling increases.
+For example, numerical error due to insufficient prior sampling should create
+large variance among posterior estimates and cause analyses to be highly
+sensitive to the number of samples drawn from the prior.
+Furthermore, if insufficient prior sampling is \emph{biasing} estimates toward
+models with less parameter space, as suggested by \citet{Hickerson2013}, we
+expect to see support for these models decrease as sampling from the prior
+increases.
 \citet{Oaks2012} did not see such sensitivity when they compared prior sample
@@
 \citet{Hickerson2013}.
-While sampling error is always a reality, it does not appear to be playing a
-large role in the biases revealed by the results of \citet{Oaks2012} or
-presented above.
+While approximation error is always present for any numerical analysis, it does
+not appear to be playing a large role in the biases revealed by the results of
+\citet{Oaks2012} or presented above.
 
@@
 \citet{Hickerson2013} (see SI text and Figure~S\ref{figValidationMCBehavior}).
-These results are not clearly predicted if insufficient sampling was causing
-estimation bias.
+These results are not clearly predicted if insufficient computation was causing
+numerical error.
 Even when the prior is correct, due to the discrete uniform prior on the number
@@
 and fewer \divt{} parameters (Figure~\ref{figPower4}).
-In all of these analyses, each of the prior models have the same number of
-samples.
-Thus, while sampling error will always be present in any numerical Bayesian
-approximation method, insufficient sampling is an unlikely explanation for the
-erroneous support for models with less parameter space.
+In all of these analyses, the same number of random samples were drawn from
+each of the prior models.
+Thus, while approximation error will always be present in any numerical
+approximation method, insufficient prior sampling is an unlikely explanation
+for the erroneous support for models with less parameter space.
 While analyses that sample each model proportional to their parameter space
@@
 More generally, it is important to carefully choose prior assumptions about
-parameters in Bayesian model selection; these prior assumptions strongly
-influence the posterior probabilities we seek to estimate.
+parameters in Bayesian model selection; these prior assumptions about
+parameters strongly influence the posterior probabilities we seek to estimate.
 

### 2014-08-01 manuscripts/reply/reply-main.tex
More work on intro of reply ms.

@@
 divergence times cause this behavior:
-\begin{itemize}
-    \item \citet{Hickerson2013} suggest that under broad uniform priors, the
-        rejection algorithm implemented in \msb is unable to adequately sample
-        the space of the models within reasonable computational time, which
-        leads to bias toward models with fewer divergence-time parameters
-        because they are better sampled.
-    \item \citet{Oaks2012} suggest the cause is more fundamental; the uniform
-        priors on divergence times simply lead to very small marginal
-        likelihoods of models with many divergence-time parameters. These
-        models are ``averaged'' over much greater parameter space with large
-        prior weight and small likelihood \citep{Jeffreys1939,Lindley1957}.
-\end{itemize}
-In the former case, the problem is numerical approximation error due to
+\begin{description}
+    \item[Hypothesis 1: Numerical approximation error \citep{Hickerson2013}]
+        Under broad uniform priors, the rejection algorithm implemented in \msb
+        is unable to adequately sample the space of the models within
+        reasonable computational time, which leads to bias toward models with
+        fewer divergence-time parameters because they are better sampled.
+    % \item \citet{Hickerson2013} suggest that under broad uniform priors, the
+    %     rejection algorithm implemented in \msb is unable to adequately sample
+    %     the space of the models within reasonable computational time, which
+    %     leads to bias toward models with fewer divergence-time parameters
+    %     because they are better sampled.
+    \item[Hypothesis 2: Strongly weighted marginal likelihoods
+        \citep{Oaks2012}]
+        The uniform priors on divergence times simply lead to very small
+        marginal likelihoods of models with many divergence-time parameters.
+        These models are ``averaged'' over much greater parameter space with
+        large prior weight and small likelihood
+        \citep{Jeffreys1939,Lindley1957}.
+    % \item \citet{Oaks2012} suggest the cause is more fundamental; the uniform
+    %     priors on divergence times simply lead to very small marginal
+    %     likelihoods of models with many divergence-time parameters. These
+    %     models are ``averaged'' over much greater parameter space with large
+    %     prior weight and small likelihood \citep{Jeffreys1939,Lindley1957}.
+\end{description}
+In Hypothesis 1, the problem is numerical approximation error due to
 insufficient computation;
-the true (correct) posterior supports models with many divergence-time
-parameters given data generated under temporally random divergences, but we are
-unable to accurately approximate the posterior.
-In the latter case, numerical error is not the issue, because the true
-posterior supports simultaneous divergences given data from randomly diverged
-taxa.
-I.e., this is the correct (albeit biogeographically misleading) answer given
-the prior assumption that divergence times are broadly and uniformly
-distributed.
-Such posterior support for simultaneous divergence, even if correct from the
-perspective of Bayesian model choice, does not provide the biogeographical
+% In the former case, the problem is numerical approximation error due to
+% insufficient computation;
+the exact posterior supports models with many divergence-time parameters
+given data generated under temporally random divergences, but we are unable to
+accurately approximate the posterior.
+In Hypothesis 2, the problem is more fundamental, because the exact posterior
+supports simultaneous divergences given data from randomly diverged taxa.
+% In the latter case, numerical error is not the issue, because the exact
+% posterior supports simultaneous divergences given data from randomly diverged
+% taxa.
+I.e., given the assumption that divergence times are broadly and uniformly
+distributed, the exact posterior from Bayes' rule leads us to the wrong
+conclusions about evolutonary history.
+% I.e., this is the correct (albeit biogeographically misleading) answer given
+% the prior assumption that divergence times are broadly and uniformly
+% distributed.
+Such posterior support for simultaneous divergence, even if ``correct'' from
+the perspective of Bayesian model choice, does not provide the biogeographical
 insights that a researcher who employs \msb seeks to gain.
@@
 While these phenomena are not mutually exclusive, it is important to
-distinguish between them in order to determine how best to improve our ability
-to estimate shared divergence histories.
-If the cause is numerical approximation error, then the model is sound and we
-need to increase our computational effort or improve our Monte Carlo
-integration procedures.
+distinguish between them in order to determine how to improve our ability to
+estimate shared divergence histories.
+If Hypothesis 1 is correct, then the model is sound and we need to increase our
+computational effort or improve our Monte Carlo integration procedures.
 For example, Markov chain or sequential Monte Carlo algorithms might sample the
 posterior more efficiently than the simple Monte Carlo rejection sampler
-currently implemented in \msb.
-Rather than alter the sampling algorithm, \citet{Hickerson2013} tried a
-model-averaging approach that uses narrow, empirically informed uniform priors
-in the hope that with less parameter space to sample, the rejection algorithm
-will produce better estimates of the posterior.
+implemented in \msb.
+Rather than alter the sampling algorithm, \citet{Hickerson2013} tried using
+narrow, empirically informed uniform priors in the hope that with less
+parameter space to sample, the rejection algorithm will produce better
+estimates of the posterior.
 Here, we discuss theoretical considerations for using empirically informed
@@
 parameters.
-The behavior of their approach also suggests that insufficient computation is
-not the primary cause of the issue.
+The behavior of their approach also suggests that insufficient computation
+(Hypothesis 1) is not the primary cause of the issue.
 
-If support for models with few divergence events is caused by their larger
-marginal likelihoods, we need to correct the model; improving the numerical
-approximation machinery will only increase our ability to estimate the true
-posterior that leads us to the wrong interpretation about evolutionary history.
+If Hypothesis 2 is correct, we need to correct the model, because no amount of
+computation will help; even if we could calculate the exact posterior, we would
+still reach the wrong interpretation about evolutionary history.
 Accordingly, \citet{Oaks2014dpp} has introduced a method that uses more
@@
 
-We also explore the distinct predictions made by the alternative causes of the
-bias proposed by \citet{Oaks2012} and \citet{Hickerson2013}.
-We show the primary cause of the method's tendency to support models of shared
-divergences is the large marginal likelihoods of these models under the prior
-assumption of uniformly distributed divergence times.
-We also show that numerical approximation error is not a major contributer to
-the problem.
+We also explore the distinct predictions made by Hypotheses 1 and 2.
+We show the behavior of \msb matches the predictions of Hypothesis 2,
+but not Hypothesis 1.
+This strongly suggests the method tends to support models of shared divergences
+not due to insufficient computation, but rather due to the larger marginal
+likelihoods of these models under the prior assumption of uniformly distributed
+divergence times.
+% We also show that numerical approximation error is not a major contributer to
+% the problem.
 

### 2014-08-01 manuscripts/reply/reply-main.tex
More tweaking of intro for reply ms.

@@
         These models are ``averaged'' over much greater parameter space with
-        large prior weight and small likelihood
+        small likelihood, but large prior weight.
         \citep{Jeffreys1939,Lindley1957}.
@@
 distributed, the exact posterior from Bayes' rule leads us to the wrong
-conclusions about evolutonary history.
+conclusions about evolutionary history.
 % I.e., this is the correct (albeit biogeographically misleading) answer given
@@
 parameters.
-The behavior of their approach also suggests that insufficient computation
-(Hypothesis 1) is not the primary cause of the issue.
+% The behavior of their approach also suggests that insufficient computation
+% (Hypothesis 1) is not the primary cause of the issue.
 

### 2014-08-02 manuscripts/reply/reply-main.tex
More editing of reply ms.

--- 2014-08-01 manuscripts/reply/reply-main.tex
@@
 divergence times cause this behavior:
-\begin{description}
-    \item[Hypothesis 1: Numerical approximation error \citep{Hickerson2013}]
+\begin{enumerate}[label=\textsl{\textbf{Hypothesis \arabic*)}},ref=\arabic*,align=left]
+    \item \textsl{\textbf{Numerical approximation error}} \citep{Hickerson2013}:
         Under broad uniform priors, the rejection algorithm implemented in \msb
@@
         fewer divergence-time parameters because they are better sampled.
+        \label{hypError}
     % \item \citet{Hickerson2013} suggest that under broad uniform priors, the
@@
     %     because they are better sampled.
-    \item[Hypothesis 2: Strongly weighted marginal likelihoods
-        \citep{Oaks2012}]
-        The uniform priors on divergence times simply lead to very small
-        marginal likelihoods of models with many divergence-time parameters.
-        These models are ``averaged'' over much greater parameter space with
-        large prior weight and small likelihood
-        \citep{Jeffreys1939,Lindley1957}.
+    \item \textsl{\textbf{Strongly weighted marginal likelihoods}}
+        \citep{Oaks2012}:
+        The uniform priors on divergence times lead to very small marginal
+        likelihoods (and thus posterior probabilities) of models with many
+        divergence-time parameters.  The likelihood of these models is
+        ``averaged'' over a much greater parameter space in which there is a
+        large amount of prior weight and small probability of producing the
+        data.  \citep{Jeffreys1939,Lindley1957}. \label{hypML}
     % \item \citet{Oaks2012} suggest the cause is more fundamental; the uniform
@@
     %     prior weight and small likelihood \citep{Jeffreys1939,Lindley1957}.
-\end{description}
-In Hypothesis 1, the problem is numerical approximation error due to
+\end{enumerate}
+In Hypothesis~\ref{hypError}, the problem is numerical approximation error due to
 insufficient computation;
-% In the former case, the problem is numerical approximation error due to
-% insufficient computation;
-the exact posterior supports models with many divergence-time parameters
-given data generated under temporally random divergences, but we are unable to
-accurately approximate the posterior.
-In Hypothesis 2, the problem is more fundamental, because the exact posterior
-supports simultaneous divergences given data from randomly diverged taxa.
+given data from randomly diverged taxa, the exact posterior supports a model
+with many divergence-time parameters, but we are unable to accurately
+approximate this posterior.
+In Hypothesis~\ref{hypML}, the problem is more fundamental; given data from
+randomly diverged taxa, the exact posterior supports a model with a small
+number of simultaneous divergences.
 % In the latter case, numerical error is not the issue, because the exact
@@
 distributed, the exact posterior from Bayes' rule leads us to the wrong
-conclusions about evolutonary history.
+conclusion about evolutionary history.
 % I.e., this is the correct (albeit biogeographically misleading) answer given
@@
 estimate shared divergence histories.
-If Hypothesis 1 is correct, then the model is sound and we need to increase our
-computational effort or improve our Monte Carlo integration procedures.
+If Hypothesis~\ref{hypError} is correct, then the model is sound and we need to
+increase our computational effort or improve our Monte Carlo integration
+procedures.
 For example, Markov chain or sequential Monte Carlo algorithms might sample the
@@
 narrow, empirically informed uniform priors in the hope that with less
-parameter space to sample, the rejection algorithm will produce better
+parameter space to sample, the rejection algorithm would produce better
 estimates of the posterior.
@@
 parameters.
-The behavior of their approach also suggests that insufficient computation
-(Hypothesis 1) is not the primary cause of the issue.
 
-If Hypothesis 2 is correct, we need to correct the model, because no amount of
-computation will help; even if we could calculate the exact posterior, we would
-still reach the wrong interpretation about evolutionary history.
+If Hypothesis~\ref{hypML} is correct, we need to correct the model, because no
+amount of computation will help; even if we could calculate the exact
+posterior, we would still reach the wrong interpretation about evolutionary
+history.
 Accordingly, \citet{Oaks2014dpp} has introduced a method that uses more
@@
 We also explore the distinct predictions made by Hypotheses 1 and 2.
-We show the behavior of \msb matches the predictions of Hypothesis 2,
-but not Hypothesis 1.
+We show the behavior of \msb matches the predictions of Hypothesis~\ref{hypML},
+but not Hypothesis~\ref{hypError}.
 This strongly suggests the method tends to support models of shared divergences
-not due to insufficient computation, but rather due to the larger marginal
+not because of insufficient computation, but rather due to the larger marginal
 likelihoods of these models under the prior assumption of uniformly distributed
 divergence times.
-% We also show that numerical approximation error is not a major contributer to
-% the problem.
 
 \section{The potential implications of empirical Bayesian model choice}
-\citet{Oaks2012} found \msb spuriously supports models of temporally
-clustered divergences even when using a prior on divergence times that was
-informed by the data being analyzed.
-The main argument of \citet{Hickerson2013} is that the priors used by
-\citet{Oaks2012} were not informative enough.
-They suggest a very narrow, highly informed uniform prior on divergence times
-is necessary to avoid the method's preference for models with few
-divergence-time parameters.
+% \citet{Oaks2012} found \msb spuriously supports models of temporally
+% clustered divergences even when using a prior on divergence times that was
+% informed by the data being analyzed.
+% The main argument of \citet{Hickerson2013} is that the priors used by
+% \citet{Oaks2012} were not informative enough.
+\citet{Hickerson2013} suggest a very narrow, highly informed uniform prior on
+divergence times is necessary to avoid the method's preference for models with
+few divergence-time parameters.
 Such an empirical Bayesian approach to model selection raises some theoretical
@@
 Particularly, in cases where large aggregate data sets are used for many
-parallel model-choice problems simultaneously, pooling information to inform
+parallel model-choice problems, pooling information to inform
 priors can lead to favorable group-wise frequentist coverage across tests
@@
 If small likelihoods of large models cause the method to prefer models with
-less parameter space, we expect \msb will preferentially sample from this
-erroneous model yielding a posterior that is misleading (i.e., the
-model-averaged posterior will be dominated by a model that excludes
-the truth).
+less parameter space (Hypothesis~\ref{hypML}), we expect \msb will
+preferentially sample from this erroneous model yielding a posterior that is
+misleading (i.e., the model-averaged posterior will be dominated by a model
+that excludes the truth).
 Second, we generated simulated datasets for which the divergence times are
@@
 (Table~\ref{tabModelChoiceEmpirical}).
-However, \citet{Hickerson2013} vetted the priors used in their model-averaging
+Given that the same number of simuations were sampled from each prior model,
+this behavior is not clearly predicted by insufficient computation
+(Hypothesis~\ref{hypError}), but is a straightforward prediction of
+Hypothesis~\ref{hypML}.
+
+\citet{Hickerson2013} vetted the priors used in their model-averaging
 approach via ``graphical checks,'' in which the summary statistics from 1000
@@
 Rather, the graphs suggest these invalid priors are ``better fit''
-(Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
-\citet{Oaks2012} (Figure~S\ref{figPCA}D--F).
+(Figure~S\ref{figPCA}A--C) than the valid priors used by \citet{Oaks2012}
+(Figure~S\ref{figPCA}D--F).
 
@@
 excluding the true value of at least one parameter.
-Whenever the Bayes factor preferred a model excluding the truth we counted the
+Whenever the Bayes factor preferred a model excluding the truth, we counted the
 number of the 22 true divergence times that were excluded by the preferred
@@
 model uncertainty in Bayesian inference.
-However, when averaging over models with narrow and broad uniform priors on a
-parameter with a likelihood density that is not expected to be uniformly
-distributed, the posterior can be dominated by models that exclude from
-consideration the true values of parameters due to the larger marginal
-likelihoods of the models that integrate over less space with high prior
-weight yet low likelihood.
+However, as predicted by Hypothesis~\ref{hypML}, when averaging over models
+with narrow and broad uniform priors on a parameter that is not expected to
+have a uniformly distributed likelihood density, the posterior can be dominated
+by models that exclude from consideration the true values of parameters due to
+their larger marginal likelihoods (these models integrate over less space with
+high prior weight and low likelihood).
 
@@
 All other parameters were identically distributed as the prior models.
-As above, we generated $5\e6$ samples from the prior models in
-Table~\ref{tabModelChoiceEmpirical}.
+As above, we generated $5\e6$ samples from prior models $M_1$--$M_5$
+(Table~\ref{tabModelChoiceEmpirical}).
 For each of the 6000 simulated datasets, we approximated the posterior
@@
 (Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
-If we follow \citet{Hickerson2013} and use a Bayes factor of greater than 10 as
-the criterion for incorrect inference of a single divergence event, we find
-that the approach will often strongly support the extreme case of one
-divergence event across all our simulation conditions (Figure~\ref{figPower4}E--H \&
-S\ref{figPower6}G--L).
+The approach often strongly supports (Bayes factor of greater than 10) the
+extreme case of one divergence event across all our simulation conditions
+(Figure~\ref{figPower4}E--H \& S\ref{figPower6}G--L).
 The method also struggles to estimate the variance of divergence times
@@
 
+%%
 \section{The causes of support for models of co-divergence}

### 2014-08-02 manuscripts/reply/reply-main.tex
More editing of reply ms.

@@
         \label{hypError}
-    % \item \citet{Hickerson2013} suggest that under broad uniform priors, the
-    %     rejection algorithm implemented in \msb is unable to adequately sample
-    %     the space of the models within reasonable computational time, which
-    %     leads to bias toward models with fewer divergence-time parameters
-    %     because they are better sampled.
     \item \textsl{\textbf{Strongly weighted marginal likelihoods}}
@@
         data.  \citep{Jeffreys1939,Lindley1957}. \label{hypML}
-    % \item \citet{Oaks2012} suggest the cause is more fundamental; the uniform
-    %     priors on divergence times simply lead to very small marginal
-    %     likelihoods of models with many divergence-time parameters. These
-    %     models are ``averaged'' over much greater parameter space with large
-    %     prior weight and small likelihood \citep{Jeffreys1939,Lindley1957}.
 \end{enumerate}
@@
 number of simultaneous divergences.
-% In the latter case, numerical error is not the issue, because the exact
-% posterior supports simultaneous divergences given data from randomly diverged
-% taxa.
 I.e., given the assumption that divergence times are broadly and uniformly
@@
 conclusion about evolutionary history.
-% I.e., this is the correct (albeit biogeographically misleading) answer given
-% the prior assumption that divergence times are broadly and uniformly
-% distributed.
 Such posterior support for simultaneous divergence, even if ``correct'' from
@@
 over much greater parameter space, all with equal prior weight, and much of it
-with small likelihood.
+with small likelihood (Hypothesis~\ref{hypML}).
 In light of this fundamental statistical issue, it is not surprising that the
@@
 
-However, \citet{Hickerson2013} conclude that the bias is caused by
-numerical approximation error due to insufficient computation.
+However, \citet{Hickerson2013} conclude that the bias is caused by numerical
+approximation error due to insufficient computation
+(Hypothesis~\ref{hypError}).
 They argue the widest of the three priors used by \citet{Oaks2012} would
@@
 have a generation time of one year.
-The argument of \citet{Hickerson2013} that divergence-time estimates of
-\citet{Oaks2012} should set an upper bound on the prior for divergence times
-seems questionable, especially given our findings presented above regarding the
-behavior of \msb when empirically informed uniform priors are employed.
+The argument of \citet{Hickerson2013} that these estimated gene divergence
+times should be used to set the upper bound of the uniform prior on divergence
+times for the \msb analysis of the same sequence data seems questionable,
+especially given our findings presented above regarding the behavior of \msb
+when empirically informed uniform priors are employed.
 
-Even if we assume the arbitrary strict clock is correct, gene divergence times
-were estimated without error, and all 22 taxa have one-year generation times,
-Hickerson et al.'s
+Even if we assume (1) the arbitrary strict clock is correct, (2) gene
+divergence times were estimated without error, and (3) all 22 taxa have
+one-year generation times, Hickerson et al.'s
 (\citeyear{Hickerson2013}) argument only applies to one of the \emph{three}
@@
 \citet{Oaks2012} ($U(0, 5)$) closely mirrors the range of estimates of
-gene-divergence times (0--5 million years ago).
+gene divergence times (0--5 million years ago).
 Applying Hickerson et al.'s (\citeyear{Hickerson2013}) sampling-probability
@@
 divergence estimates.
-Thus, if insufficient prior sampling is to blame for the bias, it should be
-much reduced under the narrow prior on \divt{}.
+Thus, if insufficient prior sampling (Hypothesis~\ref{hypError}) is to blame
+for the bias, it should be much reduced under the narrow prior on \divt{}.
 However, the magnitude of the bias is very similar across all three priors
@@
 slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
-However, it is important to note that these results suffered from a bug
-in \msb, and there are many cases after \citet{Oaks2012} corrected the 
-bug where the narrowest prior performs slightly worse (see panels D--J of
-Figures 3 and S12).
+However, it is important to note that these results suffered from a bug in
+\msb, and after \citet{Oaks2012} corrected the bug, there are many cases where
+the narrowest prior performs slightly worse (see panels D--J of Figures 3 and
+S12).
 
-To disentangle whether model likelihoods or numerical approximation error is to
-blame for the method's erroneous support for simple models, we must look at the
-different predictions made by these two phenomena.
-For example, numerical error due to insufficient prior sampling should create
-large variance among posterior estimates and cause analyses to be highly
-sensitive to the number of samples drawn from the prior.
+To disentangle whether Hypothesis 1 or 2 is the primary cause the method's
+erroneous support for simple models, we must look at the different predictions
+made by these two phenomena.
+For example, numerical error due to insufficient prior sampling
+(Hypothesis~\ref{hypError}) should create large variance among posterior
+estimates and cause analyses to be highly sensitive to the number of samples
+drawn from the prior.
 Furthermore, if insufficient prior sampling is \emph{biasing} estimates toward
@@
 (Figure~S\ref{figSamplingError}) we do not see the trend predicted by
-\citet{Hickerson2013}.
+Hypothesis~\ref{hypError}.
 While approximation error is always present for any numerical analysis, it does
@@
 
-A straightforward prediction if marginal likelihoods are causing the preference
-for simple models is that the bias should disappear as the model generating the
-data converges to the prior.
+A straightforward prediction if strongly weighted marginal likelihoods are
+causing the preference for simple models (Hypothesis~\ref{hypML}) is that the
+bias should disappear as the model generating the data converges to the prior.
 \citet{Oaks2012} tested this prediction by performing 100,000 simulations to
 assess the model-choice behavior of \msb when the prior model is correct.
-The results confirm the prediction:
-\msb estimates the probability of the one-divergence model quite well when the
-prior is correct (see Figure 4 of \citet{Oaks2012}).
+The results confirm the prediction of Hypothesis~\ref{hypML}:
+\msb estimates the probability of the one-divergence model quite well (or even
+\emph{under}estimates it) when the prior is correct (see Figure 4 of
+\citet{Oaks2012}).
 We confirmed this same behavior for the model-averaging approach used by
@@
 Thus, the results of the simulations of \citet{Oaks2012} are more consistent
-with the fundamental sensitivity of marginal likelihoods to priors.
+with the fundamental sensitivity of marginal likelihoods to priors
+(Hypothesis~\ref{hypML}).
 
@@
 density, models that exclude the true values of the parameters we aim to
-estimate will often have the largest marginal likelihood.
+estimate will often have the largest marginal likelihoods.
 
 The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
-of \msb are similar to those of \citet{Hickerson2013}, but avoid the
-need for imposing an additional dimension of model choice.
+of \msb are similar to those of \citet{Hickerson2013}, but avoid the need for
+imposing an additional dimension of model choice and using priors that often
+exclude the truth.
 \citet{Oaks2012} suggest that uniform priors may not be ideal for many
@@
 \citet{Hickerson2013} does not mitigate this bias, but rather causes it to
-manifest by sampling predominantly from models that may exclude the true values
-of the parameters.
-Our results show that it is difficult to choose an uniformly distributed
-prior on divergence times that is broad enough to confidently contain the true
-values of parameters while being narrow enough to avoid erroneous support for
-models with less parameter space.
+manifest by sampling predominantly from models that often exclude the true
+values of the divergence times the method is trying to estimate.
+Our results show that it is difficult to choose an uniformly distributed prior
+on divergence times that is broad enough to confidently contain the true values
+of parameters while being narrow enough to avoid strongly weighted and
+misleading posterior support for models with less parameter space.
 More generally, it is important to carefully choose prior assumptions about
-parameters in Bayesian model selection; these prior assumptions about
-parameters strongly influence the posterior probabilities we seek to estimate.
+parameters in Bayesian model selection, because they can strongly influence the
+\emph{true} posterior probabilities of the models we seek to estimate.
 
-The common inference of temporally clustered divergences \citep{Barber2010,
-    Bell2012, Carnaval2009, Chan2011, Daza2010, Hickerson2006, Huang2011,
-    Lawson2010, Leache2007, Plouviez2009, Stone2012, Voje2009}, when not
-accompanied with the necessary analyses to assess the robustness and temporal
-resolution of such results, should be treated with caution, because \msb has
-been shown to erroneously infer clustered divergences over a range of prior
+The common inference of temporally clustered historical events
+\citep{Barber2010, Bell2012, Carnaval2009, Chan2011, Chan2014, Daza2010,
+    Hickerson2006, Huang2011, Lawson2010, Leache2007, Plouviez2009, Stone2012,
+    Voje2009},
+when not accompanied with the necessary analyses to assess the robustness and
+temporal resolution of such results, should be treated with caution, because
+\msb has been shown to erroneously infer clustered events over a range of prior
 conditions.
@@
 excessive prior density in regions of low likelihood, which greatly improves
-the ability to infer shared divergence histories.
+inference of shared divergence histories.
 

### 2014-08-02 manuscripts/reply/reply-abstract.tex
Starting to re-work abstract of reply ms.

@@
 (\msb) to estimate the support for such a pattern of divergences across 22
-pairs of taxa from the Philippines.
-They also studied the behavior of \msb using computer simulations.
-\citet{Oaks2012} found the method was very sensitive to prior assumptions and
-had low power to detect variation in divergence times.
-These results were not surprising in light of a rich statistical literature
-showing that the marginal likelihood of a model is sensitive to the
-use of vague priors.
-Because this sensitivity to prior assumptions affects the crucial insights that
-a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended that
-users of the approach should carefully assess the robustness of their
-conclusions to different priors.
-In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
-analyses was due to the method's inefficient rejection algorithm and broad
-priors on divergence times leading to inadequate numbers of simulation
-replicates.
-They proposed a model-averaging approach that uses narrow, empirically
-informed uniform priors.
+pairs of taxa from the Philippines; they also studied the behavior of \msb
+using computer simulations.
+\citet{Hickerson2013} responded to this work.
+Both papers agree that the primary inference enabled by the method is very
+sensitive to prior assumptions, which can lead the approach to erroneously
+support shared divergences across taxa.
+% \citet{Oaks2012} found the method was very sensitive to prior assumptions and
+% often incorrectly supported models with few divergence events shared
+% across taxa.
+\citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
+marginal likelihoods of models to the uniformly distributed priors on
+divergence times.  They suggested alternative distributions for prior
+uncertainty about divergence times, and recommended that users of the method
+should conduct analyses under a variety of priors to reveal prior sensitivity
+and communicate which assumptions underlie model inference.
+% These results were not surprising in light of a rich statistical literature
+% showing that the marginal likelihood of a model is sensitive to the
+% use of vague priors.
+\citet{Hickerson2013} suggested the lack of robustness in \msb analyses was due
+to the inefficiency of the method's rejection algorithm under broad uniform
+priors on divergence times leading to numerical approximation error.
+They proposed a model-averaging approach that uses narrow, empirically informed
+uniform priors to overcome this.
 Here we demonstrate that the approach of \citet{Hickerson2013} will often
@@
 true values of the models' parameters.
-On a more fundamental level, we question the value of adopting an empirical
-Bayesian stance for this model-choice problem.
-The robust Bayesian approach of conducting analyses under a variety of priors
-can reveal prior sensitivity and communicate which assumptions underlie model
-inference.
+We also show this behavior of the method is primarily due to marginal
+likelihoods strongly weighted by the prior assumption of uniformly distributed
+divergence times, and not numerical approximation error.
+% On a more fundamental level, we question the value of adopting an empirical
+% Bayesian stance for this model-choice problem.
+% The robust Bayesian approach of conducting analyses under a variety of priors
+% can reveal prior sensitivity and communicate which assumptions underlie model
+% inference.
 % Furthermore, simulations provide insight into the temporal resolution of the

### 2014-08-03 manuscripts/reply/reply-abstract.tex
Working on abstract of reply ms.

@@
 using computer simulations.
-\citet{Hickerson2013} responded to this work.
-Both papers agree that the primary inference enabled by the method is very
-sensitive to prior assumptions, which can lead the approach to erroneously
-support shared divergences across taxa.
+% \citet{Hickerson2013} responded to this work.
+They found the primary inference enabled by the method is very sensitive to
+prior assumptions and often erroneously supports shared divergences across
+taxa.
 % \citet{Oaks2012} found the method was very sensitive to prior assumptions and
@@
 % across taxa.
-\citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
-marginal likelihoods of models to the uniformly distributed priors on
-divergence times.  They suggested alternative distributions for prior
-uncertainty about divergence times, and recommended that users of the method
-should conduct analyses under a variety of priors to reveal prior sensitivity
-and communicate which assumptions underlie model inference.
+\citet{Oaks2012} suggested this was due to the sensitivity of marginal
+likelihoods of models to the uniformly distributed priors on divergence times.
+To mitigate this sensitivity they suggested alternative distributions for
+priors on divergence times.
+They also recommended that users of the method should conduct analyses under a
+variety of priors to reveal prior sensitivity and communicate which assumptions
+underlie model inference.
 % These results were not surprising in light of a rich statistical literature
@@
 % use of vague priors.
-\citet{Hickerson2013} suggested the lack of robustness in \msb analyses was due
-to the inefficiency of the method's rejection algorithm under broad uniform
-priors on divergence times leading to numerical approximation error.
+In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
+analyses was caused by numerical approximation error stemming from the
+inefficiency of the method's rejection algorithm under broad uniform priors on
+divergence times.
 They proposed a model-averaging approach that uses narrow, empirically informed
@@
 true values of the models' parameters.
-We also show this behavior of the method is primarily due to marginal
-likelihoods strongly weighted by the prior assumption of uniformly distributed
-divergence times, and not numerical approximation error.
+% On a more fundamental level, we question the value of adopting an empirical
+% Bayesian stance for this model-choice problem.
+We also show the tendency of \msb analyses to support models of shared
+divergences is primarily due to the larger marginal likelihoods of these models
+when they are strongly weighted by the prior assumption of uniformly
+distributed divergence times; numerical approximation error is not a tenable
+explanation for the bias.
 % On a more fundamental level, we question the value of adopting an empirical
@@
 distributions can accommodate prior uncertainty about the models' parameters
-while increasing the method's robustness and power to detect temporal variation
+without placing excessive prior weight in vast regions of parameter space with
+low likelihood;
+this increases the method's robustness and power to detect temporal variation
 in divergences \citep{Oaks2014dpp}.

### 2014-08-04 manuscripts/reply/reply-abstract.tex
More editing of abstract and text of reply ms.

--- 2014-07-30 manuscripts/reply/reply-abstract.tex
@@
 affected by the same geographic event.
-Recently, \citet{Oaks2012} used an approximate-Bayesian model choice method
-(\msb) to estimate the support for such a pattern of divergences across 22
-pairs of taxa from the Philippines.
-They also studied the behavior of \msb using computer simulations.
-\citet{Oaks2012} found the method was very sensitive to prior assumptions and
-had low power to detect variation in divergence times.
-These results were not surprising in light of a rich statistical literature
-showing that the marginal likelihood of a model is sensitive to the
-use of vague priors.
-Because this sensitivity to prior assumptions affects the crucial insights that
-a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended that
-users of the approach should carefully assess the robustness of their
-conclusions to different priors.
-In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
-analyses was due to the method's inefficient rejection algorithm and broad
-priors on divergence times leading to inadequate numbers of simulation
-replicates.
-They proposed a model-averaging approach that uses narrow, empirically
-informed uniform priors.
-Here we demonstrate that the approach of \citet{Hickerson2013} will often
+Recently, \citet{Oaks2012} assessed the ability of an approximate-Bayesian
+model choice method (\msb) to estimate such patterns of divergences across
+taxa, to which \citet{Hickerson2013} responded.
+Both papers agree that the primary inference enabled by the method is very
+sensitive to prior assumptions and often erroneously supports shared
+divergences across taxa.
+However, the papers differ in the best explanation and solution for this
+problem.
+\citet{Hickerson2013} suggested the lack of robustness in \msb analyses is
+caused by numerical approximation error due to the inefficiency of the method's
+rejection algorithm under broad uniform priors on divergence times
+(Hypothesis~\ref{hypError}).
+To overcome this, they proposed a model-averaging approach that uses narrow,
+empirically informed uniform priors.
+Here we demonstrate the approach of \citet{Hickerson2013} will often
 erroneously support shared divergence events and is dangerous in the sense that
 the empirically-derived uniform priors often exclude from consideration the
-true values of the models' parameters.
-On a more fundamental level, we question the value of adopting an empirical
-Bayesian stance for this model-choice problem.
-The robust Bayesian approach of conducting analyses under a variety of priors
-can reveal prior sensitivity and communicate which assumptions underlie model
-inference.
-% Furthermore, simulations provide insight into the temporal resolution of the
-% method, which helps guide interpretation of results.
+true values of the divergence-time parameters.
+\citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
+marginal likelihoods to the uniformly distributed priors on divergence times
+(Hypothesis~\ref{hypML}).
+To mitigate this sensitivity, they suggested alternative probability
+distributions to accommodate prior uncertainty about divergence times.
 As predicted by fundamental principles of Bayesian model choice, more flexible
-distributions can accommodate prior uncertainty about the models' parameters
-while increasing the method's robustness and power to detect temporal variation
+distributions can accommodate prior uncertainty about parameters without
+placing excessive prior weight in vast regions of parameter space with low
+likelihood;
+this increases the method's robustness and power to detect temporal variation
 in divergences \citep{Oaks2014dpp}.
+We also show that the tendency of \msb analyses to support models of shared
+divergences is primarily due to Hypothesis~\ref{hypML};
+Hypothesis~\ref{hypError} is not a tenable explanation for the bias.
+All of these papers have highlighted the importance of choosing prior
+assumptions about parameters in Bayesian model selection, because they can
+strongly influence the posterior probabilities of the models we seek to
+estimate.
+No amount of computation can rescue our inference if we choose priors on
+parameters such that the exact posterior supports the wrong model of
+evolutionary history.

### 2014-08-04 manuscripts/reply/reply-main.tex
More editing of abstract and text of reply ms.

@@
 to estimate the probabilities of candidate models.
-In a simple example where two candidate models, $M_1$ and $M_2$, are being
-compared, the goal is to estimate the posterior probabilities of these two
-models.
-Again, we can use Bayes' rule to calculate this as
+In a simple example with two candidate models, $M_1$ and $M_2$, we can use
+Bayes' rule to calculate the posterior probability of $M_1$ as
 \begin{equation}
@@
 less parameter space (Hypothesis~\ref{hypML}), we expect \msb will
-preferentially sample from this erroneous model yielding a posterior that is
+preferentially sample from this erroneous prior yielding a posterior that is
 misleading (i.e., the model-averaged posterior will be dominated by a model
@@
 divergence events among 18 taxa when the true divergences were random over
-hundreds of thousands of generations or more.
+150,000 generations or more.
 It is important that investigators perform such simulations to determine the
@@
 Also, it seems unlikely that when such prior information is available that
-being able to detect more than one divergence event in the face of 18 random
-divergences over hundreds of thousands of generations will provide much insight
-into the evolutionary history of the taxa.
+being able to detect more than one divergence event in the face of 18
+divergences that were random over 150,000+ generations will
+provide much insight into the evolutionary history of the taxa.
 

### 2014-08-05 manuscripts/reply/reply-main.tex
More editing of reply ms.

@@
 (Hypothesis~\ref{hypError}).
-They argue the widest of the three priors used by \citet{Oaks2012} would
-infrequently produce random samples with many independent population divergence
-times as recent as the estimated gene divergence times presented in
-\citet{Oaks2012}.
-However, this argument assumes the gene divergence times presented in
-\citet{Oaks2012} were estimated without error.
-These estimates were intended to provide only a rough comparison of the
-gene divergence times across the 22 taxa.
-These analyses assumed an arbitrary strict per-site rate of $2\e{-8}$ mutations
-per generation for all taxa, and are, of course, subject to estimation error.
-Furthermore, the branch-length units of the gene trees are in millions of
-years, whereas the divergence-time prior of \msb is in generations, thus
+They argue the widest of the three priors on divergence times used by
+\citet{Oaks2012} would infrequently produce random samples with many
+independent population divergence times as recent as the estimated gene
+divergence times presented in \citet{Oaks2012}.
+% They also also suggest that Oaks et al.'s \citeyear{Oaks2012} estimated gene
+% divergence times should be used to set the upper bound of the uniform prior on
+% divergence times for the \msb analysis of the same sequence data.
+% Given the gene-tree estimates were intended to provide only a rough comparison
+% of the gene divergence times across the 22 taxa \citep{Oaks2012}, and our
+% findings presented above regarding the behavior of \msb when empirically
+% informed uniform priors are employed, such an approach should not be used.
+However, this sampling-probability argument is based on some questionable
+assumptions.
+Oaks et al.'s (\citeyear{Oaks2012}) gene-tree estimates were intended to
+provide only a rough comparison of the gene divergence times across the 22 taxa
+and assumed an arbitrary strict per-site rate of $2\e{-8}$ mutations per
+generation for all taxa.
+Furthermore, because the branch-length units of the gene trees are in millions
+of years whereas the divergence-time prior of \msb is in generations,
 \citet{Hickerson2013} make the implicit assumption that all 22 Philippine taxa
 have a generation time of one year.
-The argument of \citet{Hickerson2013} that these estimated gene divergence
-times should be used to set the upper bound of the uniform prior on divergence
-times for the \msb analysis of the same sequence data seems questionable,
-especially given our findings presented above regarding the behavior of \msb
-when empirically informed uniform priors are employed.
-
-Even if we assume (1) the arbitrary strict clock is correct, (2) gene
-divergence times were estimated without error, and (3) all 22 taxa have
-one-year generation times, Hickerson et al.'s
-(\citeyear{Hickerson2013}) argument only applies to one of the \emph{three}
-priors used by \citet{Oaks2012}.
-Under these assumptions, the narrowest prior on divergence times used by
-\citet{Oaks2012} ($U(0, 5)$) closely mirrors the range of estimates of
-gene divergence times (0--5 million years ago).
-Applying Hickerson et al.'s (\citeyear{Hickerson2013}) sampling-probability
-argument demonstrates this prior is densely populated with samples with large
-numbers of divergence parameters with values younger than the estimated gene
-divergence estimates.
-Thus, if insufficient prior sampling (Hypothesis~\ref{hypError}) is to blame
-for the bias, it should be much reduced under the narrow prior on \divt{}.
+More importantly, even if we assume (1) the arbitrary strict clock is correct,
+(2) gene divergence times were estimated without error, and (3) all 22 taxa
+have one-year generation times, Hickerson et al.'s (\citeyear{Hickerson2013})
+argument actually demonstrates that the models used by \citet{Oaks2012} with
+narrower priors on divergence times are densely populated with samples with
+large numbers of divergence parameters with values younger than the estimated
+gene divergence estimates.
+Thus, if \citet{Hickerson2013} are correct, analyses under these narrow priors
+should be much less biased toward clustered models of divergence.
 However, the magnitude of the bias is very similar across all three priors
@@
 Furthermore, if insufficient prior sampling is \emph{biasing} estimates toward
-models with less parameter space, as suggested by \citet{Hickerson2013}, we
-expect to see support for these models decrease as sampling from the prior
-increases.
+models with less parameter space we expect to see support for these models
+decrease as sampling from the prior increases.
 \citet{Oaks2012} did not see such sensitivity when they compared prior sample
@@
 Hypothesis~\ref{hypError}.
-While approximation error is always present for any numerical analysis, it does
+While approximation error is always present in any numerical analysis, it does
 not appear to be playing a large role in the biases revealed by the results of
@@
 These results are not clearly predicted if insufficient computation was causing
-numerical error.
+numerical error (Hypothesis~\ref{hypError}).
 Even when the prior is correct, due to the discrete uniform prior on the number
@@
 and fewer \divt{} parameters (Figure~\ref{figPower4}).
-In all of these analyses, the same number of random samples were drawn from
-each of the prior models.
-Thus, while approximation error will always be present in any numerical
-approximation method, insufficient prior sampling is an unlikely explanation
-for the erroneous support for models with less parameter space.
-While analyses that sample each model proportional to their parameter space
-could be explored, it is clear that the marginal likelihoods under broad
-uniform priors on divergence times will be greater for models with fewer
-divergence-time dimensions.
+For these model-averaging analyses, insufficient prior sampling
+(Hypothesis~\ref{hypError}) is an unlikely explanation for the erroneous support
+for models with less parameter space, because (1) all of the prior models share
+the same dimensionality, and (2) the same number of random samples were drawn
+from each of the prior models.
+These results are predicted by Hypothesis~\ref{hypML}, because the marginal
+likelihoods will be higher for models with narrower priors on divergence times
+and fewer divergence-time dimensions (these models integrate over less space
+with large prior weight and small likelihood).
 
@@
 manifest by sampling predominantly from models that often exclude the true
-values of the divergence times the method is trying to estimate.
+values of the divergence times.
 Our results show that it is difficult to choose an uniformly distributed prior

### 2014-08-05 manuscripts/reply/reply-abstract.tex
More tweaking of reply ms.

@@
 sensitive to prior assumptions and often erroneously supports shared
-divergences across taxa.
+divergences across taxa when prior uncertainty about divergence times
+is represented by a uniform distribution.
 However, the papers differ in the best explanation and solution for this
@@
 We also show that the tendency of \msb analyses to support models of shared
-divergences is primarily due to Hypothesis~\ref{hypML};
+divergences is primarily due to Hypothesis~\ref{hypML}, whereas
 Hypothesis~\ref{hypError} is not a tenable explanation for the bias.

### 2014-08-05 manuscripts/reply/reply-main.tex
More tweaking of reply ms.

@@
 \begin{itemize}
-   \item The use of vague priors on the divergence-time parameters can lead to
-       spurious support for models with few divergence events shared across
-       taxa. Thus, the primary inference enabled by the approach is very
-       sensitive to the priors on divergence times.
+   \item Representing prior uncertainty about divergence-time parameters with a
+       uniform distribution can lead to spurious support for models with few
+       divergence events shared across taxa. Thus, the primary inference
+       enabled by the approach is very sensitive to the priors on divergence
+       times.
 \end{itemize}
@@
 \end{enumerate}
-In Hypothesis~\ref{hypError}, the problem is numerical approximation error due to
-insufficient computation;
-given data from randomly diverged taxa, the exact posterior supports a model
-with many divergence-time parameters, but we are unable to accurately
-approximate this posterior.
+In Hypothesis~\ref{hypError}, the problem is numerical approximation error due
+to insufficient computation.
+In this scenario, given data from taxa that diverged randomly through time, the
+exact posterior supports a model with many divergence-time parameters, but we
+are unable to accurately approximate this posterior.
 In Hypothesis~\ref{hypML}, the problem is more fundamental; given data from
-randomly diverged taxa, the exact posterior supports a model with a small
-number of simultaneous divergences.
-I.e., given the assumption that divergence times are broadly and uniformly
-distributed, the exact posterior from Bayes' rule leads us to the wrong
-conclusion about evolutionary history.
+taxa that diverged randomly through time, the exact posterior supports a model
+with simultaneous divergences across taxa.
+I.e., when accommodating prior uncertainty about divergence times with a
+uniform distribution, the exact posterior from Bayes' rule leads us to the
+wrong conclusion about evolutionary history.
 Such posterior support for simultaneous divergence, even if ``correct'' from
@@
 priors for Bayesian model choice and evaluate the approach of
-\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
-\citet{Oaks2012}.
+\citet{Hickerson2013} as a potential solution to the biases of \msb.
 In their analyses, \citet{Hickerson2013} made an error by mixing different
@@
 but not Hypothesis~\ref{hypError}.
-This strongly suggests the method tends to support models of shared divergences
-not because of insufficient computation, but rather due to the larger marginal
-likelihoods of these models under the prior assumption of uniformly distributed
-divergence times.
+This strongly suggests that the method tends to support models of shared
+divergences not because of insufficient computation, but rather due to the
+larger marginal likelihoods of these models under the prior assumption of
+uniformly distributed divergence times.
 
@@
 They argue the widest of the three priors on divergence times used by
-\citet{Oaks2012} would infrequently produce random samples with many
-independent population divergence times as recent as the estimated gene
-divergence times presented in \citet{Oaks2012}.
+\citet{Oaks2012} would infrequently produce random samples of parameter values
+with many independent population divergence times as recent as the estimated
+gene divergence times presented in \citet{Oaks2012}.
 % They also also suggest that Oaks et al.'s \citeyear{Oaks2012} estimated gene
@@
 
-To disentangle whether Hypothesis 1 or 2 is the primary cause the method's
+To disentangle whether Hypothesis 1 or 2 is the primary cause of the method's
 erroneous support for simple models, we must look at the different predictions
@@
 For these model-averaging analyses, insufficient prior sampling
-(Hypothesis~\ref{hypError}) is an unlikely explanation for the erroneous support
+(Hypothesis~\ref{hypError}) is an untenable explanation for the erroneous support
 for models with less parameter space, because (1) all of the prior models share
@@
 from each of the prior models.
-These results are predicted by Hypothesis~\ref{hypML}, because the marginal
-likelihoods will be higher for models with narrower priors on divergence times
-and fewer divergence-time dimensions (these models integrate over less space
-with large prior weight and small likelihood).
+However, these results are predicted by Hypothesis~\ref{hypML}, because the
+marginal likelihoods will be higher for models with narrower priors on
+divergence times and fewer divergence-time dimensions (these models integrate
+over less space with large prior weight and small likelihood).
 

### 2014-08-06 manuscripts/reply/reviews/response.tex
More work on reply ms and review response.

@@
     end to an interesting discussion about Bayesian model selection.
+
+    We have re-written a lot of the manuscript and feel the new version will be
+    of general interest to readers interested in Bayesian selection of
+    evolutionary models.
 }
@@
     paper was too long and lacked focus.
-    We have greatly reduced the length of the main text and moved some items
-    (including 2 figures) to the supplement.
+    We greatly reduced the length of the main text, moved some items (including
+    2 figures) to the supplement, and re-wrote a lot of the manuscript to focus
+    on the most interesting aspects of Bayesian model choice brought to
+    light by this series of papers.
 }
@@
         \item[L96-100 and L104-108]
-            We have reduced this section to only three sentences that summarize
-            only the information necessary to introduce our subsequent
-            discussion.
+            We have removed this section.
         \item[L222-237]
@@
     We thank Reviewer 2 for pointing out this section, which was admitedly very
-    confusing. We have re-written this section (and almost all of the
-    introduction) to try to make this distinction clearer. Our goal is to
-    distinguish two different ways that broad uniform priors could cause the
-    poor behavior of msBayes that was shown in the origin Oaks et al. paper.
-    \begin{enumerate}
-        \item As pointed out by reviewer 2, these priors could cause numerical
-            issues, where the simple ABC rejection machinery is unable to
-            sufficiently sample the space of the model in manageable
-            computational time.
-        \item The priors can cause rich models to have very small (but correct)
-            marginal likelihoods.
-    \end{enumerate}
+    confusing. We have re-written this section (and a lot of the manuscript) to
+    try to make this distinction clearer. Our goal is to distinguish two
+    different ways that broad uniform priors could cause the poor behavior of
+    msBayes that was shown in the original Oaks et al. paper.
 
-    These issues are not mutually exclusive, but in the former the true
-    posterior would support heterogeneity in divergence times in the face of
-    random divergences (i.e., the problem only arises from our inability to
-    estimate that posterior).
-    Whereas in the later, the true posterior (i.e., an analytical solution)
-    would correctly tell us that the most probable model is simultaneous
-    divergence despite random variation in divergence times. I.e., this is
-    actually the correct answer under the prior assumptions from the
-    perspective of Bayesian model choice. Unfortunately, this correct answer
-    does not provide the biogeographical insights we are using the method to
-    gain.
+    We now pose these as alternative hypotheses to explain the methods tendency
+    to support clustered divergence times. We also discuss the distinct
+    predictions made by each hypothesis, and how best to improve the method
+    under each of them.
 
-    This is an important distinction, because if the prior conditions are such
-    that we are in situation 2, any arguments about the estimation procedure
-    are moot, because even the true posterior is biogeographcially misleading.
+    This has greatly improved the clarity, focus, and general interest of the
+    paper, and we are very thankful that Reviewer 2 steered us in this
+    direction.
 }

### 2014-08-06 manuscripts/reply/reply-abstract.tex
More work on reply ms and review response.

@@
 Recently, \citet{Oaks2012} assessed the ability of an approximate-Bayesian
-model choice method (\msb) to estimate such patterns of divergences across
-taxa, to which \citet{Hickerson2013} responded.
+model choice method (\msb) to estimate such a pattern of simultaneous
+divergence across taxa, to which \citet{Hickerson2013} responded.
 Both papers agree that the primary inference enabled by the method is very
@@
 empirically informed uniform priors.
-Here we demonstrate the approach of \citet{Hickerson2013} will often
-erroneously support shared divergence events and is dangerous in the sense that
-the empirically-derived uniform priors often exclude from consideration the
-true values of the divergence-time parameters.
 \citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
@@
 distributions to accommodate prior uncertainty about divergence times.
-As predicted by fundamental principles of Bayesian model choice, more flexible
-distributions can accommodate prior uncertainty about parameters without
-placing excessive prior weight in vast regions of parameter space with low
-likelihood;
-this increases the method's robustness and power to detect temporal variation
-in divergences \citep{Oaks2014dpp}.
-We also show that the tendency of \msb analyses to support models of shared
-divergences is primarily due to Hypothesis~\ref{hypML}, whereas
-Hypothesis~\ref{hypError} is not a tenable explanation for the bias.
+Here we use analyses of simulated and empirical data to show that the tendency
+of \msb analyses to support models of shared divergences is primarily due to
+Hypothesis~\ref{hypML}, whereas Hypothesis~\ref{hypError} is an untenable
+explanation for the bias.
+We also demonstrate the approach of \citet{Hickerson2013} will often
+erroneously support shared divergence events and is dangerous in the sense that
+the empirically-derived uniform priors often exclude from consideration the
+true values of the divergence-time parameters.
+\citep{Oaks2014dpp} showed, as predicted by fundamental principles of Bayesian
+model choice, more flexible distributions can accommodate prior uncertainty
+about parameters without placing excessive prior weight in vast regions of
+parameter space with low likelihood; this increases the method's robustness and
+power to detect temporal variation in divergences.
 All of these papers have highlighted the importance of choosing prior
-assumptions about parameters in Bayesian model selection, because they can
-strongly influence the posterior probabilities of the models we seek to
-estimate.
-No amount of computation can rescue our inference if we choose priors on
-parameters such that the exact posterior supports the wrong model of
-evolutionary history.
+assumptions about parameters in Bayesian model selection, because they
+can strongly influence the posterior probabilities of the models we seek
+to estimate.
+If we choose priors placing too much weight in unlikely regions of parameter
+space such that the true posterior supports the wrong model of evolutionary
+history, no amount of computation can rescue our inference.

### 2014-08-06 manuscripts/reply/reviews/response.tex
More work on review response for reply ms.

@@
 \usepackage[numbers,sort&compress]{natbib}
+\usepackage{enumitem}
 \input{../../utils/macros.tex}
@@
 
-I greatly appreciate the helpful comments of the editor and reviewers. In this
-document, I provide responses to these comments, including descriptions of the
-changes I have made.
+My co-authors and I greatly appreciate the helpful comments of the editor and
+reviewers.
+In this document, I provide our responses to these comments, including
+descriptions of the changes we have made.
 
 \section{Editor's Comments}\label{ed}
+\setcounter{commentCounter}{0}
+\revcomment{
+    In particular, the paper needs to be more focused and would benefit from
+    avoiding unnecessary polemical considerations. Importantly, the core
+    arguments could be made in much fewer words, given previous publications. A
+    resubmission should be cut by at least 25\% in the main text. Similarly
+    consider removing 1 or 2 figures (perhaps fig. 5 and 6).
+}
+\response{
+    We have reduced the number of words in the main text be 26\%, and
+    moved Figures 5 and 6 to the Supporting Information.
+}
+
+\section{Associate Editor's Comments}\label{ed}
 \setcounter{commentCounter}{0}
@@
     In regard to the sections specifically highlighted by Reviewer 2:
-    \begin{description}
+    \begin{description}[align=left,labelindent=0.5in,leftmargin=0.75in]
         \item[L96-100 and L104-108]
@@
 
+\section{Reviewer 3's Comments}\label{rev2}
+\setcounter{commentCounter}{0}
+
+\revcomment{
+     I feel this is an excellent paper and is ready for publication in Evolution.
+}
+\response{
+    Thank you.
+}
+
 \bibliographystyle{../../bib/bmc-mathphys.bst}

### 2014-08-06 manuscripts/reply/reply-abstract.tex
More editing of reply ms.

@@
 Establishing that a set of population splitting events occurred at the same
-time can be a potentially persuasive argument that the populations were
-affected by the same geographic event.
+time can be a potentially persuasive argument that a common process affected
+the populations.
 Recently, \citet{Oaks2012} assessed the ability of an approximate-Bayesian
@@
 is represented by a uniform distribution.
-However, the papers differ in the best explanation and solution for this
+However, the papers differ about the best explanation and solution for this
 problem.
@@
 marginal likelihoods to the uniformly distributed priors on divergence times
-(Hypothesis~\ref{hypML}).
-To mitigate this sensitivity, they suggested alternative probability
-distributions to accommodate prior uncertainty about divergence times.
-Here we use analyses of simulated and empirical data to show that the tendency
+(Hypothesis~\ref{hypML}), and proposed alternative prior probability
+distributions to mitigate this sensitivity.
+As predicted by fundamental principles of Bayesian model choice, more flexible
+distributions that accommodate prior uncertainty about parameters without
+placing excessive weight in vast regions of parameter space with low
+likelihood increase the method's robustness and power to detect temporal
+variation in divergences \citep{Oaks2014dpp}.
+Here, we use analyses of simulated and empirical data to show that the tendency
 of \msb analyses to support models of shared divergences is primarily due to
@@
 explanation for the bias.
-We also demonstrate the approach of \citet{Hickerson2013} will often
-erroneously support shared divergence events and is dangerous in the sense that
-the empirically-derived uniform priors often exclude from consideration the
-true values of the divergence-time parameters.
-\citep{Oaks2014dpp} showed, as predicted by fundamental principles of Bayesian
-model choice, more flexible distributions can accommodate prior uncertainty
-about parameters without placing excessive prior weight in vast regions of
-parameter space with low likelihood; this increases the method's robustness and
-power to detect temporal variation in divergences.
+We also demonstrate the empirical Bayesian model-averaging approach of
+\citet{Hickerson2013} will often erroneously support shared divergence events
+and is dangerous in the sense that the empirically-derived uniform priors often
+exclude from consideration the true values of the divergence-time parameters.
 All of these papers have highlighted the importance of choosing prior
-assumptions about parameters in Bayesian model selection, because they
-can strongly influence the posterior probabilities of the models we seek
-to estimate.
-If we choose priors placing too much weight in unlikely regions of parameter
+assumptions about parameters in Bayesian model selection.
+If our prior assumptions place too much weight in unlikely regions of parameter
 space such that the true posterior supports the wrong model of evolutionary

### 2014-08-06 manuscripts/reply/reply-main.tex
More editing of reply ms.

@@
         The uniform priors on divergence times lead to very small marginal
-        likelihoods (and thus posterior probabilities) of models with many
-        divergence-time parameters.  The likelihood of these models is
+        likelihoods (and thus smaller posterior probabilities) of models with
+        many divergence-time parameters.  The likelihood of these models is
         ``averaged'' over a much greater parameter space in which there is a
@@
 In this scenario, given data from taxa that diverged randomly through time, the
-exact posterior supports a model with many divergence-time parameters, but we
-are unable to accurately approximate this posterior.
+exact (true) posterior supports a model with many divergence-time parameters,
+but we are unable to accurately approximate this posterior.
 In Hypothesis~\ref{hypML}, the problem is more fundamental; given data from
@@
 Accordingly, \citet{Oaks2014dpp} has introduced a method that uses more
-flexible probability distributions to accommodate prior uncertainty in
-divergence times without overly inhibiting the marginal likelihoods of models
-with more divergence-time parameters.
+flexible probability distributions (e.g., gamma) to accommodate prior
+uncertainty in divergence times without overly inhibiting the marginal
+likelihoods of models with more divergence-time parameters.
 This greatly increases the method's robustness and power to detect temporal
@@
 
-We also explore the distinct predictions made by Hypotheses 1 and 2.
+We also use analyses of simulated and empirical data to explore the distinct
+predictions made by Hypotheses 1 and 2.
 We show the behavior of \msb matches the predictions of Hypothesis~\ref{hypML},
@@
 \emph{true} posterior probabilities of the models we seek to estimate.
+No amount of computation can rescue our inference if our prior assumptions
+place too much weight in unlikely regions of parameter space such that the
+exact posterior supports the wrong model of evolutionary history.
 

### 2014-08-06 manuscripts/reply/reply-si-figures.tex
More editing of reply ms.

@@
 %% Sampling error
-\widthFigure{0.9}{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{listformat=sFigList}{
+\widthFigure{0.9}{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{name=Figure S, labelformat=noSpace, listformat=sFigList}{
     Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)

### 2014-08-07 manuscripts/reply/reviews/response.tex
More work on reply ms and response to reviews.

@@
 \let\endquoteOld\endquote
-% \renewenvironment{quote}{\sffamily\slshape\small\quoteOld}{\endquoteOld}
 \renewenvironment{quote}{\sffamily\small\quoteOld}{\endquoteOld}
 
-% \makeatletter
-% \renewcommand{\section}{\@startsection{section}{1}{0mm}%
-%     {-8pt}%
-%     {4pt}%
-% {\sffamily\large\itshape}}
-% \makeatother
-
 \begin{document}
-% \raggedright
 \setlength{\parindent}{0in}
 
-My co-authors and I greatly appreciate the helpful comments of the editor and
-reviewers.
-In this document, I provide our responses to these comments, including
-descriptions of the changes we have made.
+My co-authors and I greatly appreciate the helpful comments of the editors and
+reviewers.  In this document, I provide our responses to these comments,
+including descriptions of the changes we have made.
 
@@
 \response{
-    We have reduced the number of words in the main text be 26\%, and
-    moved Figures 5 and 6 to the Supporting Information.
+    We have reduced the number of words in the main text by a little more than
+    25\%, and moved Figures 5 and 6 to the Supporting Information.
+    We have re-written many sections to make the paper less polemical and more
+    focused.
 }
@@
     We have re-written a lot of the manuscript and feel the new version will be
-    of general interest to readers interested in Bayesian selection of
+    of greater interest to readers interested in Bayesian selection of
     evolutionary models.
@@
 \response{
-    We agree that Bayesian statisticians are aware of the issues raised in our
-    paper. However, I think among biologists there is common misunderstanding
-    about some of the subtle but fundamental differences between Bayesian
-    approaches to parameter estimation and model selection. Mainly, the
-    implicit and strong statements that priors on the models' parameters can
-    make about the probability of the models. Thus, we feel the readership of
-    \emph{Evolution} will benefit from issues discussed in the paper.
-
+    We have re-written much of the manuscript in order to make it of greater
+    interest to the broad readership of \emph{Evolution}.
     We have added emphasis of solutions to the problem in several places in the
@@
     method for estimating shared divergence histories. We now draw attention to
-    this positive outcome from this work.
+    this positive outcome.
 }
@@
         \item[L287-289]
-            We agree this section was superfluous. We have removed it.
+            We agree this section was superfluous and removed it.
     \end{description}
@@
 \response{
-    This is a great point, and we have added emphasis to solutions to the
+    This is a great point, and we have added emphasis on solutions to the
     problem, as recommended by Reviewer 2. As mentioned above, solutions to
     this problem were the focus of a separate paper, and were too extensive to
-    discuss in detail in the response paper. However, we highlight the main
+    discuss in detail in the response paper. However, we now highlight the main
     points of this paper to make the response more positive, and to to refer
@@
     different ways that broad uniform priors could cause the poor behavior of
-    msBayes that was shown in the original Oaks et al. paper.
-
-    We now pose these as alternative hypotheses to explain the methods tendency
-    to support clustered divergence times. We also discuss the distinct
-    predictions made by each hypothesis, and how best to improve the method
-    under each of them.
+    msBayes that was shown in the original Oaks et al. paper.  We now pose
+    these as alternative hypotheses to explain the method's tendency to support
+    clustered divergence times. We also discuss the distinct predictions made
+    by each hypothesis, and how best to improve the method under each of them.
 

### 2014-08-07 manuscripts/reply/cover-letter/cover-letter.tex
Updating reply cover letter.
More work on reply ms and response to reviews.

@@
 \usepackage{anysize}
-\marginsize{1in}{1in}{0.5in}{0.5in}
+\marginsize{1in}{1in}{0in}{0.5in}
 \usepackage{setspace}
@@
 % \raggedright
-I am submitting a manuscript entitled ``Implications of uniformly distributed,
-empirically informed priors for phylogeographical model selection: A reply to
-Hickerson et al.'' to be considered for publication in the \emph{International
-Journal of Organic Evolution}.
+I am re-submitting a manuscript (previous submission ID 14-0137) entitled
+``Implications of uniformly distributed, empirically informed priors for
+phylogeographical model selection:
+A reply to Hickerson et al.'' to be considered for publication in the
+\emph{International Journal of Organic Evolution}.
 This manuscript is a response to a paper by Hickerson et al.\ (2014,
@@
 67:991--1010).
-Former Associate Editor Dr.\ Laura Kubatko invited us to write this response,
-with the intention of handling its review.
-The following is an excerpt from Dr.\ Kubatko's comments in the decision letter
-for Hickerson et al.\ (2014):
+Former Associate Editor Dr.\ Laura Kubatko invited us to write this response.
 
-\begin{quote}
-Thus, I propose the following: Hickerson et al. should prepare and
-submit the final version of their response.  I urge them to fully
-consider all of the points raised by Oaks in his most recent review of their
-paper. However, since the point of view of Hickerson et al. seems to
-be at odds with Oaks et al., I would like to give Oaks et al. the opportunity
-to respond, if they so choose. I have therefore recommended acceptance
-for this version of the manuscript, and I ask that Hickerson et al. provide a
-copy of the final version to Jamie Oaks once they have submitted it, so
-that he and his co-authors can consider whether they wish to respond.
-\end{quote}
-
-In our response, we use simulated and biological data to evaluate the empirical
-Bayesian method proposed by Hickerson et al.\ (2014) as a solution to the
-problems revealed by Oaks et al.\ (2013) regarding the power and accuracy of the
-comparative phylogeographical method implemented in the software
-\texttt{msBayes}.
+In the manuscript, we highlight alternative hypotheses for why the popular
+comparative phylogeographcial method, \texttt{msBayes}, often incorrectly
+infers simultaneous divergence across co-distributed species.
+Using analyses of simulated and empirical data, we test the distinct
+predictions of these hypotheses and find the cause of the bias is not
+approximation error, but rather strongly weighted marginal likelihoods of
+divergence models under uniform priors on divergence times.
+We also evaluate the empirical Bayesian method proposed by Hickerson et al.\
+(2014) as a solution to the problems revealed by Oaks et al.\ (2013) regarding
+the power and accuracy of the method.
 We demonstrate their approach does not mitigate the method's bias toward
@@
 values of the model's parameters.
-Our results also strongly suggest the underlying statistical cause of the
-method's tendency to support simple models.
 More generally, we discuss the theoretical and practical issues of empirical
@@
 applying methods of Bayesian model selection to carefully consider.
-Lastly, we discuss ways to mitigate the biases of the \texttt{msBayes} method
-while avoiding negative side effects of empirically informed uniform priors.
+We highlight ways to mitigate the biases of the \texttt{msBayes} method while
+avoiding negative side effects of empirically informed uniform priors.
+Overall, this paper highlights the importance of choosing prior assumptions
+about parameters in Bayesian selection of models of evolution; if our prior
+assumptions place too much weight in unlikely regions of parameter space, the
+true posterior will often support the wrong model of evolutionary history.
 
@@
 \href{https://github.com/joaks1/msbayes-experiments}{\url{https://github.com/joaks1/msbayes-experiments}}.
-All information necessary to reproduce my results is provided there.
+All information necessary to reproduce our results is provided there.
 
+The previous submission of this manuscript was handled by Associate Editor Dr.\
+Luke Harmon who obtained three anonymous reviews.
+We have incorporated the helpful revisions recommended by Dr.\ Harmon and the
+reviewers.
+My point-by-point response to their comments follows this letter.
 Thank you for your consideration, and I look forward to your correspondence

### 2014-08-07 manuscripts/reply/reply-abstract.tex
More work on reply ms and response to reviews.

@@
 distributions to mitigate this sensitivity.
+Here, we use analyses of simulated and empirical data to show that the tendency
+of \msb analyses to support models of shared divergences is primarily due to
+Hypothesis~\ref{hypML}, whereas Hypothesis~\ref{hypError} is an untenable
+explanation for the bias.
+If our prior assumptions place too much weight in unlikely regions of parameter
+space such that the true posterior supports the wrong model of evolutionary
+history, no amount of computation can rescue our inference. 
+We also demonstrate the empirical Bayesian model-averaging approach of
+\citet{Hickerson2013} is dangerous in the sense that the empirically-derived
+uniform priors often exclude from consideration the true values of the
+divergence-time parameters.
+Furthermore, this approach will often strongly support models of highly
+clustered divergence times when analyzing data from taxa that diverged randomly
+over millions of generations.
 As predicted by fundamental principles of Bayesian model choice, more flexible
@@
 variation in divergences \citep{Oaks2014dpp}.
-Here, we use analyses of simulated and empirical data to show that the tendency
-of \msb analyses to support models of shared divergences is primarily due to
-Hypothesis~\ref{hypML}, whereas Hypothesis~\ref{hypError} is an untenable
-explanation for the bias.
-We also demonstrate the empirical Bayesian model-averaging approach of
-\citet{Hickerson2013} will often erroneously support shared divergence events
-and is dangerous in the sense that the empirically-derived uniform priors often
-exclude from consideration the true values of the divergence-time parameters.
-All of these papers have highlighted the importance of choosing prior
-assumptions about parameters in Bayesian model selection.
-If our prior assumptions place too much weight in unlikely regions of parameter
-space such that the true posterior supports the wrong model of evolutionary
-history, no amount of computation can rescue our inference.

### 2014-08-07 manuscripts/reply/reply-abstract.tex
More work on abstract of reply ms.

@@
 explanation for the bias.
-If our prior assumptions place too much weight in unlikely regions of parameter
-space such that the true posterior supports the wrong model of evolutionary
-history, no amount of computation can rescue our inference. 
+I.e., if our prior assumptions place too much weight in unlikely regions of
+parameter space such that the true posterior supports the wrong model of
+evolutionary history, no amount of computation can rescue our inference. 
 We also demonstrate the empirical Bayesian model-averaging approach of
-\citet{Hickerson2013} is dangerous in the sense that the empirically-derived
-uniform priors often exclude from consideration the true values of the
-divergence-time parameters.
-Furthermore, this approach will often strongly support models of highly
-clustered divergence times when analyzing data from taxa that diverged randomly
-over millions of generations.
-As predicted by fundamental principles of Bayesian model choice, more flexible
-distributions that accommodate prior uncertainty about parameters without
-placing excessive weight in vast regions of parameter space with low
+\citet{Hickerson2013} does not mitigate the method's tendency to strongly
+support models of highly clustered divergence times when analyzing data from
+taxa that diverged randomly over millions of generations.
+Furthermore, we show this approach is dangerous in the sense that the
+empirically-derived uniform priors often exclude from consideration the true
+values of the divergence-time parameters.
+Fortunately, as predicted by fundamental principles of Bayesian model choice,
+more flexible distributions that accommodate prior uncertainty about parameters
+without placing excessive weight in vast regions of parameter space with low
 likelihood increase the method's robustness and power to detect temporal

### 2014-08-07 manuscripts/reply/reply-abstract.tex
More work on the abstract of reply ms.

@@
 \citet{Hickerson2013} suggested the lack of robustness in \msb analyses is
-caused by numerical approximation error due to the inefficiency of the method's
-rejection algorithm under broad uniform priors on divergence times
+caused by numerical approximation error due to the inability of the method's
+rejection algorithm to adequately sample parameter space within reasonable
+computational limits when using broad uniform priors on divergence times
 (Hypothesis~\ref{hypError}).
@@
 empirically informed uniform priors.
-\citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
-marginal likelihoods to the uniformly distributed priors on divergence times
-(Hypothesis~\ref{hypML}), and proposed alternative prior probability
-distributions to mitigate this sensitivity.
+\citet{Oaks2012} suggested the method's behavior was caused by the strong
+weight of uniformly distributed priors on divergence times leading to smaller
+marginal likelihoods (and posterior probabilities) of models with more
+divergence time parameters
+(Hypothesis~\ref{hypML}); they proposed alternative prior probability
+distributions to avoid such strongly weighted marginal likelihoods.
+% \citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
+% marginal likelihoods to the uniformly distributed priors on divergence times
+% (Hypothesis~\ref{hypML}), and proposed alternative prior probability
+% distributions to mitigate this sensitivity.
 Here, we use analyses of simulated and empirical data to show that the tendency

### 2014-08-07 manuscripts/reply/reply-abstract.tex
More tweaking of abstract for reply ms.

@@
 problem.
-\citet{Hickerson2013} suggested the lack of robustness in \msb analyses is
-caused by numerical approximation error due to the inability of the method's
-rejection algorithm to adequately sample parameter space within reasonable
-computational limits when using broad uniform priors on divergence times
-(Hypothesis~\ref{hypError}).
-To overcome this, they proposed a model-averaging approach that uses narrow,
-empirically informed uniform priors.
+\citet{Hickerson2013} suggested numerical approximation error causes \msb
+analyses to be biased toward models of clustered divergences because the the
+method's rejection algorithm is unable to adequately sample the parameter space
+of richer models within reasonable computational limits when using broad
+uniform priors on divergence times (Hypothesis~\ref{hypError}).
+As a potential solution, they proposed a model-averaging approach that uses
+narrow, empirically informed uniform priors.
 \citet{Oaks2012} suggested the method's behavior was caused by the strong
 weight of uniformly distributed priors on divergence times leading to smaller
-marginal likelihoods (and posterior probabilities) of models with more
-divergence time parameters
-(Hypothesis~\ref{hypML}); they proposed alternative prior probability
-distributions to avoid such strongly weighted marginal likelihoods.
+marginal likelihoods (and thus smaller posterior probabilities) of models with
+more divergence time parameters (Hypothesis~\ref{hypML}); they proposed
+alternative prior probability distributions to avoid such strongly weighted
+posteriors.
 % \citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
@@
 We also demonstrate the empirical Bayesian model-averaging approach of
-\citet{Hickerson2013} does not mitigate the method's tendency to strongly
-support models of highly clustered divergence times when analyzing data from
-taxa that diverged randomly over millions of generations.
+\citet{Hickerson2013} does not mitigate the method's tendency to erroneously
+support models of highly clustered divergences.
 Furthermore, we show this approach is dangerous in the sense that the

### 2014-08-07 manuscripts/reply/reply-abstract.tex
Fixing "the the" in abstract of reply ms.

@@
 \citet{Hickerson2013} suggested numerical approximation error causes \msb
-analyses to be biased toward models of clustered divergences because the the
+analyses to be biased toward models of clustered divergences because the
 method's rejection algorithm is unable to adequately sample the parameter space

### 2014-08-07 manuscripts/reply/reply-abstract.tex
More editing of abstract and main text of reply ms.

@@
 problem.
+\citet{Oaks2012} suggested the method's behavior was caused by the strong
+weight of uniformly distributed priors on divergence times leading to smaller
+marginal likelihoods (and thus smaller posterior probabilities) of models with
+more divergence time parameters (Hypothesis~\ref{hypML}); they proposed
+alternative prior probability distributions to avoid such strongly weighted
+posteriors.
 \citet{Hickerson2013} suggested numerical approximation error causes \msb
@@
 narrow, empirically informed uniform priors.
-\citet{Oaks2012} suggested the method's behavior was caused by the strong
-weight of uniformly distributed priors on divergence times leading to smaller
-marginal likelihoods (and thus smaller posterior probabilities) of models with
-more divergence time parameters (Hypothesis~\ref{hypML}); they proposed
-alternative prior probability distributions to avoid such strongly weighted
-posteriors.
-% \citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
-% marginal likelihoods to the uniformly distributed priors on divergence times
-% (Hypothesis~\ref{hypML}), and proposed alternative prior probability
-% distributions to mitigate this sensitivity.
-Here, we use analyses of simulated and empirical data to show that the tendency
-of \msb analyses to support models of shared divergences is primarily due to
-Hypothesis~\ref{hypML}, whereas Hypothesis~\ref{hypError} is an untenable
-explanation for the bias.
-I.e., if our prior assumptions place too much weight in unlikely regions of
-parameter space such that the true posterior supports the wrong model of
-evolutionary history, no amount of computation can rescue our inference. 
-We also demonstrate the empirical Bayesian model-averaging approach of
-\citet{Hickerson2013} does not mitigate the method's tendency to erroneously
-support models of highly clustered divergences.
-Furthermore, we show this approach is dangerous in the sense that the
-empirically-derived uniform priors often exclude from consideration the true
-values of the divergence-time parameters.
+Here, we use analyses of simulated and empirical data to demonstrate that the
+approach of \citet{Hickerson2013} does not mitigate the method's tendency to
+erroneously support models of highly clustered divergences, and is dangerous in
+the sense that the empirically-derived uniform priors often exclude from
+consideration the true values of the divergence-time parameters.
+Our results also show that the tendency of \msb analyses to support models of
+shared divergences is primarily due to Hypothesis~\ref{hypML}, whereas
+Hypothesis~\ref{hypError} is an untenable explanation for the bias.
+Overall, this series of papers demonstrate that if our prior assumptions place
+too much weight in unlikely regions of parameter space such that the exact
+posterior supports the wrong model of evolutionary history, no amount of
+computation can rescue our inference. 
 Fortunately, as predicted by fundamental principles of Bayesian model choice,
@@
 likelihood increase the method's robustness and power to detect temporal
-variation in divergences \citep{Oaks2014dpp}.
+variation in divergences.

### 2014-08-07 manuscripts/reply/reply-main.tex
More editing of abstract and main text of reply ms.

@@
 \begin{enumerate}[label=\textsl{\textbf{Hypothesis \arabic*)}},ref=\arabic*,align=left]
-    \item \textsl{\textbf{Numerical approximation error}} \citep{Hickerson2013}:
-        Under broad uniform priors, the rejection algorithm implemented in \msb
-        is unable to adequately sample the space of the models within
-        reasonable computational time, which leads to bias toward models with
-        fewer divergence-time parameters because they are better sampled.
-        \label{hypError}
     \item \textsl{\textbf{Strongly weighted marginal likelihoods}}
@@
         data.  \citep{Jeffreys1939,Lindley1957}. \label{hypML}
+    \item \textsl{\textbf{Numerical approximation error}} \citep{Hickerson2013}:
+        Under broad uniform priors, the rejection algorithm implemented in \msb
+        is unable to adequately sample the space of the models within
+        reasonable computational time, which leads to bias toward models with
+        fewer divergence-time parameters because they are better sampled.
+        \label{hypError}
 \end{enumerate}
@@
 parameters in Bayesian model selection, because they can strongly influence the
-\emph{true} posterior probabilities of the models we seek to estimate.
+posterior probabilities of the models we seek to compare.
 No amount of computation can rescue our inference if our prior assumptions

### 2014-09-25 manuscripts/reply/reply-abstract.tex
Incorporating some edits to proofs.

@@
-Establishing that a set of population splitting events occurred at the same
+Establishing that a set of population-splitting events occurred at the same
 time can be a potentially persuasive argument that a common process affected
@@
 Recently, \citet{Oaks2012} assessed the ability of an approximate-Bayesian
-model choice method (\msb) to estimate such a pattern of simultaneous
+model-choice method (\msb) to estimate such a pattern of simultaneous
 divergence across taxa, to which \citet{Hickerson2013} responded.
@@
 marginal likelihoods (and thus smaller posterior probabilities) of models with
-more divergence time parameters (Hypothesis~\ref{hypML}); they proposed
+more divergence-time parameters (Hypothesis~\ref{hypML}); they proposed
 alternative prior probability distributions to avoid such strongly weighted

### 2014-09-25 manuscripts/reply/reply-main.tex
Incorporating some edits to proofs.

@@
 differentiation on geographical phenomena.
-Establishing that a set of population splitting events occurred
+Establishing that a set of population-splitting events occurred
 at the same time can be a potentially persuasive argument that a set of taxa
@@
         data.  \citep{Jeffreys1939,Lindley1957}. \label{hypML}
-    \item \textsl{\textbf{Numerical approximation error}} \citep{Hickerson2013}:
+    \item \textsl{\textbf{Numerical-approximation error}} \citep{Hickerson2013}:
         Under broad uniform priors, the rejection algorithm implemented in \msb
@@
 \end{enumerate}
-In Hypothesis~\ref{hypError}, the problem is numerical approximation error due
+In Hypothesis~\ref{hypError}, the problem is numerical-approximation error due
 to insufficient computation.
@@
 \begin{linenomath}
-Whereas empirical Bayesian approaches can provide powerful methods for
+Although empirical Bayesian approaches can provide powerful methods for
 parameter estimation, a theoretical justification for empirical Bayesian
@@
 
-However, \citet{Hickerson2013} conclude that the bias is caused by numerical
+However, \citet{Hickerson2013} conclude that the bias is caused by numerical-
 approximation error due to insufficient computation
@@
 \section{Conclusions}
-We demonstrate how the approximate Bayesian model-choice method implemented in
+We demonstrate how the approximate-Bayesian model-choice method implemented in
 \msb can spuriously support models with less parameter space.

### 2014-09-26 manuscripts/reply/reply-main.tex
Incorporating edits from proofs.

@@
 the narrowest prior performs slightly worse (see panels D--J of Figures 3 and
-S12).
+S12 of \citet{Oaks2012}).
 

### 2014-10-30 slides/osu/slides.tex
Adding slides for osu.

\input{../utils/preamble.tex}
\input{../../manuscripts/utils/macros.tex}
\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
\bibliography{../../manuscripts/bib/references}

\title[Improving inference of evolutionary history]{Improving inference of evolutionary history}
\subtitle{Developing comparative methods for genomic data}

\author[J.\ Oaks]{
    Jamie R.\ Oaks\inst{1}
}
\institute[University of Washington]{
    \inst{1}%
        Department of Biology, University of Washington
}

% \date{\today}
\date{November 3, 2014}

\begin{document}
\maketitle

% I. What I do
% II. What I've been focusing on recently (started in PhD and continuing in postdoc)
%   a. Dynamic planet---need to account for large-scale processes
%       i. E.g., SE Asia---dynamic landscape
\begin{frame}
    \frametitle{Processes of diversification}
        \begin{itemize}
            \item<1-> Large-scale geological and climatic processes are
                important in biodiversification and community assembly
            \item<2-> Accounting for such processes will better our
                understanding of biodiversity
            \item<3-> We need methods for inferring evolutionary patterns
                predicted by historical events from contemporary populations
        \end{itemize}
\end{frame}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
\begin{frame}
    \frametitle{Southeast Asia}    
\end{frame}
}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
\begin{frame}
    \frametitle{Southeast Asia}    
\end{frame}
}

%   b. Predictions of climate-driven diversification
\begin{frame}
    \frametitle<1-4>{Community scale processes}
    \frametitle<5->{Divergence model choice}
    \begin{columns}[c]
        \column{.4\textwidth}
        \begin{onlyenv}<1-4>
            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
            \end{minipage}
        \end{onlyenv}
        \begin{onlyenv}<5>
            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
                \begin{displaybox}[0.95\linewidth]
                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
                        \[
                            \divTimeMapVector = (\divTimeMap{1},\divTimeMap{2},\divTimeMap{3})
                        \]\vspace{0mm}
                        \[
                            model = 111
                        \]\vspace{0mm}
                        \[
                            \divTimeVector = \{\divTime{1}\}
                        \]\vspace{0mm}
                    \end{minipage}
                \end{displaybox}
            \end{minipage}
        \end{onlyenv}
        \begin{onlyenv}<6>
            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
                \begin{displaybox}[0.95\linewidth]
                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
                        \[
                            \divTimeMapVector = (260,260,260)
                        \]\vspace{0mm}
                        \[
                            model = 111
                        \]\vspace{0mm}
                        \[
                            \divTimeVector = \{260\}
                        \]\vspace{0mm}
                    \end{minipage}
                \end{displaybox}
            \end{minipage}
        \end{onlyenv}
        \begin{onlyenv}<7>
            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
                \begin{displaybox}[0.95\linewidth]
                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
                        \[
                            \divTimeMapVector = (397,260,260)
                        \]\vspace{0mm}
                        \[
                            model = 211
                        \]\vspace{0mm}
                        \[
                            \divTimeVector = \{260,397\}
                        \]\vspace{0mm}
                    \end{minipage}
                \end{displaybox}
            \end{minipage}
        \end{onlyenv}
        \begin{onlyenv}<8>
            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
                \begin{displaybox}[0.95\linewidth]
                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
                        \[
                            \divTimeMapVector = (260,397,260)
                        \]\vspace{0mm}
                        \[
                            model = 121
                        \]\vspace{0mm}
                        \[
                            \divTimeVector = \{260,397\}
                        \]\vspace{0mm}
                    \end{minipage}
                \end{displaybox}
            \end{minipage}
        \end{onlyenv}
        \begin{onlyenv}<9>
            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
                \begin{displaybox}[0.95\linewidth]
                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
                        \[
                            \divTimeMapVector = (260,260,397)
                        \]\vspace{0mm}
                        \[
                            model = 112
                        \]\vspace{0mm}
                        \[
                            \divTimeVector = \{260,397\}
                        \]\vspace{0mm}
                    \end{minipage}
                \end{displaybox}
            \end{minipage}
        \end{onlyenv}
        \begin{onlyenv}<10>
            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
                \begin{displaybox}[0.95\linewidth]
                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
                        \[
                            \divTimeMapVector = (260,95,397)
                        \]\vspace{0mm}
                        \[
                            model = 123
                        \]\vspace{0mm}
                        \[
                            \divTimeVector = \{260,95,397\}
                        \]\vspace{0mm}
                    \end{minipage}
                \end{displaybox}
            \end{minipage}
        \end{onlyenv}
        % \begin{onlyenv}<11>
        %     \begin{minipage}[c][0.5\textheight][c]{\linewidth}
        %         \begin{displaybox}[0.95\linewidth]
        %             \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
        %                 \[
        %                     \divTimeMapVector = (260,260,260)
        %                 \]\vspace{0mm}
        %                 \[
        %                     model = 111
        %                 \]\vspace{0mm}
        %                 \[
        %                     \divTimeVector = \{260\}
        %                 \]\vspace{0mm}
        %             \end{minipage}
        %         \end{displaybox}
        %     \end{minipage}
        % \end{onlyenv}
        \begin{onlyenv}<11-13>
            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
                \begin{displaybox}[0.95\linewidth]
                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
                        \[
                            \divTimeMapVector = (\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})
                        \]\vspace{0mm}
                        \[
                            model = \divModel{i}
                        \]\vspace{0mm}
                        \[
                            \divTimeVector = \{\divTime{1},\ldots,\divTime{\divTimeNum}\}
                        \]\vspace{0mm}
                    \end{minipage}
                \end{displaybox}
            \end{minipage}
        \end{onlyenv}
        \begin{onlyenv}<14>
            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
                \begin{mydescription}
                    \item[\alignmentVector] Sequence alignments
                    \item[\divTimeMapVector] Divergence times
                    \item[\divModel{}] Divergence model
                    \item[\geneTreeVector] Gene trees
                    \item[\hkyModelVector] Substitution parameters
                    \item[\demographicParamVector] Demographic parameters
                    % \item[\allParameters] $(\divTimeMapVector, \geneTreeVector,
                    %     \hkyModelVector, \demographicParamVector)$
                \end{mydescription}
            \end{minipage}
        \end{onlyenv}
        \begin{uncoverenv}<12->
            \begin{minipage}[c][0.18\textheight][c]{\linewidth}
            \begin{flushleft}
                \small We want to infer \textcolor{blue}{\divModel{}} and
                \textcolor{blue}{\divTimeMapVector} given DNA sequence
                alignments
                \textcolor{blue}{\alignmentVector}
            \end{flushleft}
            \end{minipage}
        \end{uncoverenv}
        \column{.6\textwidth}
        \begin{minipage}[t][0.8\textheight][c]{\linewidth}
        \includegraphics<1>[height=6cm]{../images/community-ellipse-small.pdf}
        % \includegraphics<2>[height=6cm]{../images/community-ellipse-barrier-small.pdf}
        \includegraphics<2>[height=6cm]{../images/community-ellipse-split-small.pdf}
        \includegraphics<3>[height=6cm]{../images/div-model-cartoon-taxa.pdf}
        \includegraphics<4>[height=6cm]{../images/div-model-cartoon-shared-no-labels.pdf}
        \includegraphics<5-6>[height=6cm]{../images/div-model-cartoon-shared.pdf}
        \includegraphics<7>[height=6cm]{../images/div-model-cartoon-2-1.pdf}
        \includegraphics<8>[height=6cm]{../images/div-model-cartoon-2-2.pdf}
        \includegraphics<9>[height=6cm]{../images/div-model-cartoon-2-3.pdf}
        \includegraphics<10>[height=6cm]{../images/div-model-cartoon-general.pdf}
        \includegraphics<11-12>[height=6cm]{../images/div-model-cartoon-shared.pdf}
        \includegraphics<13->[height=6cm]{../images/div-model-cartoon-sp-tree-shared.pdf}
        \end{minipage}
    \end{columns}
\end{frame}

%   c. Testing for predicted patterns via model choice.
%       i. Bayesian model choice

\begin{frame}[t]
    \frametitle{Bayesian model choice}
    \begin{block}{\it Full model:}
        \begin{minipage}[c][3.8cm][c]{\linewidth}
            \begin{uncoverenv}<1->
            \[
                p(
                  \divTimeMapVector,
                  \geneTreeVector,
                  \hkyModelVector,
                  \demographicParamVector
                  \given \alignmentVector, \divModel{i})
                  =
                \frac{
                    p(\alignmentVector \given
                      \divTimeMapVector,
                      \geneTreeVector,
                      \hkyModelVector,
                      \demographicParamVector,
                      \divModel{i}
                      )
                      p(
                        \divTimeMapVector,
                        \geneTreeVector,
                        \hkyModelVector,
                        \demographicParamVector
                        \given \divModel{i}
                        )
                    }{p(\alignmentVector \given \divModel{i})}
            \]\vspace{-1mm}
            \end{uncoverenv}
            \begin{uncoverenv}<2->
            \[
                p(\alignmentVector \given \divModel{i}) =
                \int_{\allParameters{i}}
                p(\alignmentVector \given \allParameters{i}, \divModel{i})
                p(\allParameters{i} \given \divModel{i})
                d\allParameters{i}
            \]\vspace{-1mm}
            \end{uncoverenv}
            \begin{uncoverenv}<3->
            \[
                p(\divModel{i} \given \alignmentVector) =
                \frac{
                    p(\alignmentVector \given \divModel{i})
                    p(\divModel{i})
                }{
                    \sum_{i} p(\alignmentVector \given \divModel{i})
                    p(\divModel{i})
                }
            \]\vspace{-1mm}
            \end{uncoverenv}
        \end{minipage}
    \end{block}
    \vspace{1mm}
    \begin{uncoverenv}<4>
        \begin{center}
            \msb: Approximate Bayesian computation (ABC)

            % \vspace{-5mm}
            % \[ \alignmentVector \, \to \, \ssVectorObs \, \to \, \ssSpace\]
        \end{center}
    \end{uncoverenv}
    \barefootnote{\tiny\shortfullcite{Huang2011}. \hspace{2mm}\shortfullcite{Oaks2012}.}
\end{frame}

%   d. Method exists---applied it to comparative dataset from Philippines
%       i. Surprising results!
%       ii. Rather than publish, did some detective work

\begin{frame}
\begin{columns}[c]
    \column{.5\textwidth}
        \begin{table}%[htbp]
            \scriptsize
            \addtolength{\tabcolsep}{-0.09cm}
            \centering
            \begin{tabular}{ l c c }
                % \toprule
                \textbf{Species} & {\boldmath $n_1$} & {\boldmath $n_2$} \\
                \hline
                \textbf{\textsc{Mammals}} & & \\
                \emph{Crocidura beatus}             & 12 & 11 \\
                \emph{Crocidura negrina-panayensis} & 12 & 6  \\
                \emph{Hipposideros obscurus}        & 19 & 9  \\
                \emph{Hipposideros pygmaeus}        & 3  & 12 \\
                \emph{Cynopterus brachyotis}        & 20 & 8  \\
                \emph{Cynopterus brachyotis}        & 8  & 14 \\
                \emph{Haplonycteris fischeri}       & 29 & 8  \\
                \emph{Haplonycteris fischeri}       & 9  & 21 \\
                \emph{Macroglossus minimus}         & 19 & 4  \\
                \emph{Macroglossus minimus}         & 8  & 10 \\
                \emph{Ptenochirus jagori}           & 4  & 7  \\
                \emph{Ptenochirus jagori}           & 8  & 8  \\
                \emph{Ptenochirus minor}            & 30 & 9  \\
                \textbf{\textsc{Squamates}} & & \\
                \emph{Cyrtodactylus gubaot-sumuroi} & 29 & 6  \\
                \emph{Cyrtodactylus annulatus}      & 14 & 3  \\
                \emph{Cyrtodactylus philippinicus}  & 6  & 14 \\
                \emph{Gekko mindorensis}            & 8  & 11 \\
                \emph{Insulasaurus arborens}        & 22 & 10 \\
                \emph{Pinoyscincus jagori}          & 8  & 8  \\
                \emph{Dendrelaphis marenae}         & 6  & 6  \\
                \textbf{\textsc{Anurans}}  & & \\
                \emph{Limnonectes leytensis}        & 4  & 2  \\
                \emph{Limnonectes magnus}           & 2  & 3  \\
                \hline
            \end{tabular}
        \end{table}
    \column{.5\textwidth}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=1.5cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/gekko-mindorensis.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=1.5cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
\end{columns}
\end{frame}

\begin{frame}
\begin{columns}[c]
    \column{.5\textwidth}
        \begin{table}%[htbp]
            \scriptsize
            \addtolength{\tabcolsep}{-0.09cm}
            \def\firstrowcolor{}
            \def\secondrowcolor{}
            \def\thirdrowcolor{}
            \def\fourthrowcolor{}
            \def\fifthrowcolor{}
            \only<2>{\def\firstrowcolor{\rowcolor{rowred}}}
            \only<3>{\def\secondrowcolor{\rowcolor{rowred}}}
            \only<4>{\def\thirdrowcolor{\rowcolor{rowred}}}
            \only<5>{\def\fourthrowcolor{\rowcolor{rowred}}}
            \only<6>{\def\fifthrowcolor{\rowcolor{rowred}}}
            \centering
            \begin{tabular}{ l c c }
                % \toprule
                \textbf{Species} & {\boldmath $n_1$} & {\boldmath $n_2$} \\
                \hline
                \textbf{\textsc{Mammals}} & & \\
                \fifthrowcolor  \emph{Crocidura beatus}             & 12 & 11 \\
                \firstrowcolor  \emph{Crocidura negrina-panayensis} & 12 & 6  \\
                \thirdrowcolor  \emph{Hipposideros obscurus}        & 19 & 9  \\
                \secondrowcolor \emph{Hipposideros pygmaeus}        & 3  & 12 \\
                \thirdrowcolor  \emph{Cynopterus brachyotis}        & 20 & 8  \\
                \firstrowcolor  \emph{Cynopterus brachyotis}        & 8  & 14 \\
                \thirdrowcolor  \emph{Haplonycteris fischeri}       & 29 & 8  \\
                \firstrowcolor  \emph{Haplonycteris fischeri}       & 9  & 21 \\
                \thirdrowcolor  \emph{Macroglossus minimus}         & 19 & 4  \\
                \firstrowcolor  \emph{Macroglossus minimus}         & 8  & 10 \\
                \thirdrowcolor  \emph{Ptenochirus jagori}           & 4  & 7  \\
                \firstrowcolor  \emph{Ptenochirus jagori}           & 8  & 8  \\
                \thirdrowcolor  \emph{Ptenochirus minor}            & 30 & 9  \\
                \textbf{\textsc{Squamates}} & & \\
                \fifthrowcolor  \emph{Cyrtodactylus gubaot-sumuroi} & 29 & 6  \\
                \secondrowcolor \emph{Cyrtodactylus annulatus}      & 14 & 3  \\
                \firstrowcolor  \emph{Cyrtodactylus philippinicus}  & 6  & 14 \\
                \firstrowcolor  \emph{Gekko mindorensis}            & 8  & 11 \\
                \firstrowcolor  \emph{Insulasaurus arborens}        & 22 & 10 \\
                \fourthrowcolor \emph{Pinoyscincus jagori}          & 8  & 8  \\
                \firstrowcolor  \emph{Dendrelaphis marenae}         & 6  & 6  \\
                \textbf{\textsc{Anurans}}  & & \\
                \secondrowcolor \emph{Limnonectes leytensis}        & 4  & 2  \\
                \secondrowcolor \emph{Limnonectes magnus}           & 2  & 3  \\
                \hline
            \end{tabular}
        \end{table}
    \column{.5\textwidth}
        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
        \includegraphics<2>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-bohol_mindanao.png}
        \includegraphics<4>[width=\textwidth]{../images/maps/Philippines-leyte_mindanao.png}
        \includegraphics<5>[width=\textwidth]{../images/maps/Philippines-mindanao_samar.png}
        \includegraphics<6>[width=\textwidth]{../images/maps/Philippines-leyte_samar.png}
\end{columns}
\end{frame}

\section{Empirical results}
\subsection{Too good to be true?}

% \begin{frame}
% \frametitle{Outline}
% \tableofcontents[currentsection,currentsubsection]
% \end{frame}

\begin{frame}
    \frametitle{Empirical results}
    \begin{columns}[c]
        \column{.4\textwidth}
            {\small
            Strong support for simultaneous divergence of all 22 taxon pairs \\
            \vspace{1cm}
            pp $>$ 0.96 \\
            \vspace{1cm}
            $\sim$100,000--250,000 years ago}
        \column{.6\textwidth}
        \includegraphics[width=\textwidth]{../images/jointDensityPlotsSlide.jpg}
    \end{columns}
\end{frame}

%   e. Evaluated method via simulation
%       i. biased in alarming direction

\section{Simulation-based assessment of method}

\begin{frame}
    \frametitle{Simulation-based power analyses}
    What is ``simultaneous''?
    \begin{myitemize}
        \item<1-> Simulate datasets in which all 22 divergence times are random
        \uncover<1->{
            \smallskip
            \begin{myitemize}
                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
                \smallskip
                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
                \smallskip
                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
                \smallskip
                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
                \smallskip
            \end{myitemize}
        \item<1-> $MGA$ = Millions of Generations Ago
        }
        \item<2-> Simulate 1000 datasets for each \divTime{} distribution
        \item<2-> Analyze all 4000 datasets as we did the empirical data
    \end{myitemize}
\end{frame}

\begin{frame}[t]
    \frametitle{Simulation-based power analyses: Results}
    \vspace{1cm}
        \centerline{
        \includegraphics[width=1.13\textwidth]{../images/old_old_power_psi_mode.pdf}}
        % \vspace{0mm}
        % \centerline{
        % \includegraphics<2>[width=1.13\textwidth]{../images/old-sims_power_psi_prob_headless.pdf}}
    \barefootnote{\shortfullcite{Oaks2012}}
\end{frame}

\begin{frame}
    \frametitle{Simulation-based power analyses: Results}
        \msb will often infer clustered divergences when divergences are random
        over millions of generations. \\

    \bigskip
    The empirical results are likely spurious \\

    \begin{onlyenv}<2->
        \begin{block}{Objective:}
            Use principles of probability to extend \msb framework for improved
            estimation of shared evolutionary history
        \end{block}
    \end{onlyenv}
    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
\end{frame}

%   f. Rather than abandon approach---look for theoretical reasons for behavior
%       i. Strongly weighted posterior (marginal likelihood)
%       ii. Combinatorics
%   g. Developed new method---dpp-msbayes
%       i. Used DPP---common in Bayeisn nonparametrics
%           1. Flexible and mathematically convenient way of assigning prior
%           probability to all possible divergence models (partitions) directly
%           (avoid combinatoric issue)
%       ii. More flexible priors on parameters of model
%       iii. Implemented method in much more efficient multi-processing
%       framework.
%   e. Results---improved!
%   f. Empirical results...
%       i. Reducing bias and increasing power provides more "honest" estimate
%       of uncertainty
%       ii. A lot of uncertainty (Good! and Bad!)
%       iii. What next? More data and more power!

\begin{frame}[label=improvements]
    \frametitle{An improved method}
    Potential improvements:
    \begin{enumerate}
        \item Alternative priors on parameters that increase marginal
            likelihoods of rich models
        \item Alternative approach to modeling the temporal distribution of
            divergences
    \end{enumerate}
    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
\end{frame}

\begin{frame}[t]
    \vspace{-2mm}
    \begin{displaybox}[5.5cm]
        \small
        \[
            p(X) = \int_{\theta} p(X \given \theta) p(\theta) \mathrm{d}\theta
        \]%\vspace{0mm}
    \end{displaybox}

    \vspace{-1mm}
    \begin{center}
        % \includegraphics<1>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
        \includegraphics<2>[height=7.8cm]{../images/marginal-plot-3d.png}
        \includegraphics<3>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
        \includegraphics<4>[height=7.8cm]{../images/marginal-plot-2d.pdf}
    \end{center}
\end{frame}

\againframe{improvements}

\begin{frame}
    \frametitle{Prior on divergence models}
    \begin{itemize}
        \item \msb uses a discrete uniform prior on the \emph{number} of
            divergence events
    \end{itemize}
    % \smallskip
    \centerline{
        \includegraphics<1->[width=\textwidth]{../images/partition_numbers.pdf}}
    \begin{block}<2->{\it Potential solution:}
        Place flexible prior directly on the sample space of divergence models
    \end{block}
\end{frame}

\begin{frame}
    \frametitle{New method: \dppmsbayes}
    \begin{itemize}
        % \item<1-> Reparameterized the model implemented in \msb
        \item<1-> Replaced uniform priors on continuous parameters with gamma and
            beta distributions
        \item<1-> Dirichlet process prior (DPP) over all possible divergence
            models
        \item<1-> Develop multi-processing interface to improve performance for
            accommodating genomic datasets
        % \item<4-> Uniform prior over divergence models
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{\dppmsbayes: Simulation-based assessment}
        Simulate 50,000 datasets under three models\\
        \smallskip
        \begin{description}
            \item[$M_{msBayes}$]
                \begin{itemize}
                    \item U-shaped prior on divergence models
                    \item Uniform priors on continuous parameters
                \end{itemize}
            \smallskip
            \item[$M_{Ushaped}$]
                \begin{itemize}
                    \item U-shaped prior on divergence models
                    \item Gamma priors on continuous parameters
                \end{itemize}
            \smallskip
            % \item[$M_{Uniform}$]
            %     \begin{itemize}
            %         \item Uniform prior on divergence models
            %         \item Gamma priors on continuous parameters
            %     \end{itemize}
            % \smallskip
            \item[$M_{DPP}$]
                \begin{itemize}
                    \item DPP prior on divergence models
                    \item Gamma priors on continuous parameters
                \end{itemize}
        \end{description}
        \smallskip
        Analyze all datasets under each of the models
\end{frame}

\begin{frame}
    \frametitle{\dppmsbayes: Simulation results}
    \centerline{
        \includegraphics<1>[width=0.85\textwidth]{../images/validation-model-choice-old-dpp.pdf}
        \includegraphics<2>[width=1.13\textwidth]{../images/validation-model-choice-old-dpp-full.pdf}}
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\begin{frame}
    \frametitle{\dppmsbayes: Simulation-based power analyses}
    \begin{myitemize}
        \item Simulate datasets in which all 22 divergence times are random
        \uncover{
            \smallskip
            \begin{myitemize}
                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
                \smallskip
                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
                \smallskip
                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
                \smallskip
                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
                \smallskip
            \end{myitemize}
        \item $MGA$ = Millions of Generations Ago
        }
        \item Simulate 1000 datasets for each \divTime{} distribution
        % \item Analyze all 4000 datasets as we did the empirical data
        \item Analyze all 4000 datasets under models $M_{msBayes}$, $M_{Ushaped}$, and $M_{DPP}$
    \end{myitemize}
\end{frame}

\begin{frame}[t]
    \frametitle{\dppmsbayes: Power results}
    \vspace{1cm}
        \centerline{
        \includegraphics[width=1.13\textwidth]<1->{../images/old_old_power_psi_mode.pdf}}
        \vspace{0mm}
        \centerline{
        \includegraphics[width=1.13\textwidth]<2->{../images/old_dpp_power_psi_mode_headless.pdf}}
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\begin{frame}[t]
    \frametitle{\dppmsbayes: Power results}
    \vspace{1cm}
        \centerline{
        \includegraphics[width=1.13\textwidth]{../images/old_old_power_psi_prob.pdf}}
        \vspace{0mm}
        \centerline{
        \includegraphics[width=1.13\textwidth]{../images/old_dpp_power_psi_prob_headless.pdf}}
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\begin{frame}
    \frametitle{\dppmsbayes: Power results}
    % \vspace{1cm}
        \centerline{
        \includegraphics[width=\textwidth]{../images/old_old_power_psi_prob.pdf}}
        \vspace{0mm}
        \centerline{
        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_psi_prob_headless.pdf}}
        \vspace{0mm}
        \centerline{
        \includegraphics[width=\textwidth]{../images/old_dpp_power_psi_prob_headless.pdf}}
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\begin{frame}
    \frametitle{Empirical results: Philippine diversification}
    \centerline{
    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp.pdf}}
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\begin{frame}
    \frametitle{Conclusions}
    \begin{itemize}
        \item<1-> New method for estimating shared evolutionary history shows
            improved:
        \begin{enumerate}
            \item Estimation of posterior uncertainty
            \item Model-choice accuracy
            \item Power to detect temporal variation across divergences
            \item Robustness to model violations
            \item Speed!
        \end{enumerate}

    \end{itemize}

    \ignore{
    \bigskip
    \begin{uncoverenv}<2->
    Caveats:\\
    \begin{itemize}
            \item<1-> Estimating a very rich (600+ parameters for 22 taxa) model
                using limited information from the data

            \begin{itemize}
                \item Likely sensitive to prior assumptions
                \item Be skeptical of strongly supported results
            \end{itemize}

    \end{itemize}
    \end{uncoverenv}
    }
\end{frame}

\begin{frame}
    \frametitle{Recommendations}
    For Bayesian model choice, choose priors carefully\\
    \bigskip
    ABC model choice estimates should be accompanied by:
    \begin{enumerate}
        \item Simulation-based power analyses
        \item Assessment of prior sensitivity
    \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{Everything is on GitHub\ldots}
    Software:\\
    \begin{itemize}
        \item \texttt{dpp-msbayes}:
            \href{https://github.com/joaks1/dpp-msbayes}{\tt
            https://github.com/joaks1/dpp-msbayes}

        \item \texttt{PyMsBayes}:
            \href{https://github.com/joaks1/PyMsBayes}{\tt
            https://github.com/joaks1/PyMsBayes}

        \item ABACUS: Approximate BAyesian C UtilitieS.
            \href{https://github.com/joaks1/abacus}{\tt
            https://github.com/joaks1/abacus}
    \end{itemize}

    \medskip
    Open-Science Notebook:\\
    \begin{itemize}
        \item \texttt{msbayes-experiments}:
            \href{https://github.com/joaks1/msbayes-experiments}{\tt
            https://github.com/joaks1/msbayes-experiments}
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Testing limate-driven diversification}
    \textbf{Did repeated fragmentation of islands during inter-glacial rises in sea level
    promote diversification?}\\

    \bigskip
    \begin{uncoverenv}<2->
    \ldots left with uncertainty regarding climate-driven diversification \\
    \end{uncoverenv}

    \begin{uncoverenv}<3->
    \begin{itemize}
            \item More data

            \item More power 
    \end{itemize}
    \end{uncoverenv}
\end{frame}

%   g. Collected thousands of loci from across the genome of almost 300
%   individuals of two genera of geckos from across the Philippine Islands
%       i. Prelim results with dpp-msbayes (can accommodate NGS data!)

\begin{frame}
    \frametitle{Genomic data}
    \begin{columns}[c]
        \column{.5\textwidth}
        \begin{myitemize}
            \item Collecting genomic data from two gekkonid radiations
                co-distributed Southeast Asia
            \begin{myitemize}
                \item \emph{Cyrtodactylus}---12+ species
                \item \emph{Gekko}---13+ species
            \end{myitemize}
            \item Whole genome assemblies for each genus
            \item RAD-seq libraries constructed for $\approx150$ individuals
                from each genus
        \end{myitemize}
        \column{.5\textwidth}
        \includegraphics[width=\textwidth]{../images/photos/cyrt-agusanensis.jpg} \quad
        \includegraphics[width=\textwidth]{../images/photos/gekko-mindorensis.jpg}
    \end{columns}
\end{frame}

\begin{frame}
    Prelim results from 5 population pairs/1000 loci \ldots
\end{frame}

%   h. Currently working on full-likelihood implementation of the model.
%       i. Working with Jeet---has full-likelihood up and running---inefficient
%       ii. Take advantage of recent developments to analytically integrate
%       gene trees---might even be faster than simulations?!
%       iii. Uses ALL information in genetic data

\begin{frame}
    \frametitle{Full-likelihood approach}
    \begin{itemize}[<+->]
        \item Full-likelihood Bayesian implementation \footnote{\tiny\shortfullcite{JeetDiss}}
            \begin{itemize}
                \item \ldots inefficient
            \end{itemize}
        \item Currently working on incorporating new developments on analytical
            integration of gene trees
            \footnote{\tiny\shortfullcite{Bryant2012}}
            \begin{itemize}
                \item Very efficient numerical approximation of posterior 
                \item Applicable to NGS datasets
                \item Uses all the information in the sequence data
            \end{itemize}
    \end{itemize}
\end{frame}

\section{Future directions}

% III. Future work
%   a. Theory/computation
%       i. Extend full-likelihood approach to phylogenetic framework
%           1. Can infer relative mutation rates/pop sizes from data (fewer
%           assumptions)
%           2. Much better representation of diversification than pairwise
%           approach.
%       ii. Parameter estimation approach (Lindley's paradox).

\begin{frame}
    \frametitle{Future directions---Computational}
    \begin{itemize}
        \item<1-> Full-phylogenetic framework
    \end{itemize}

    \begin{center}
        % \includegraphics<1->[height=5.5cm]{../images/div-model-cartoon-sp-tree-shared-wide.pdf}
        \includegraphics<1->[height=5.5cm]{../images/div-model-cartoon-sp-phylogeny.pdf}
    \end{center}
\end{frame}

%   b. Empirical
%       i. Gobi (NGS --> NSF)
%       ii. West-Central Africa
%           1. unfunded collaborator
%       iii. Compare diversification of mainland SE Asia to Oceanic Islands
%           1. Planning pre-proposal
%       iv. Long-term study of small island of sunda shelf
%           1. Tiny islands with pops of Eutropis multifasciata and
%           Lepidodactylus lugubris
%           2. Inundated during last interglacial ~120,000 years ago, spent the
%           better part of next 100k years as part of exposed sunda shelf, then
%           fragmented during current interglacial
%           3. Both species vary in reproductive mode (oviparity/ovoviviparity,
%           parthenogenesis), skinks in color patterns
%           4. Affects of recent, severe fragmentation on diversification and
%           demography
%           5. Comparative genomics to understand evolution of color and
%           reproductive mode
%           6. Evolutionary ecology---what ecological context influence
%           evolution asexuality/live-bearing/color.
\begin{frame}
    \frametitle{Future directions---Empirical}
    \begin{itemize}
        \item<1-> Many exciting empirical systems for applying new methods!
    \end{itemize}
\end{frame}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../../../../Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}}
\begin{frame}
    % \frametitle{Gobi Desert}
\end{frame}
}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../../../../Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}}
\begin{frame}
    % \frametitle{Gobi Desert}
\end{frame}
}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../../../../Dropbox/field-photos/africa/nigeria-landscape.jpg}}
\begin{frame}
    % \frametitle{Gobi Desert}
\end{frame}
}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../../../../Dropbox/field-photos/malaysia/malaysia-burungs.jpg}}
\begin{frame}
    % \frametitle{Gobi Desert}
\end{frame}
}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
\begin{frame}
    \frametitle{Southeast Asia}    
\end{frame}
}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
\begin{frame}
    \frametitle{Southeast Asia}    
\end{frame}
}

\begin{frame}
    \frametitle{Acknowledgments}
    \begin{columns}[t]
        \column{.5\textwidth}
            {\bf Ideas and feedback:}
            \begin{myitemize}
                \item Holder Lab
                \item KU Herpetology
                \item Melissa Callahan
            \end{myitemize}
            \smallskip
            {\bf Computation:}
            \begin{myitemize}
                \item KU ITTC
                \item KU Computing Center
                \item iPlant
            \end{myitemize}
        \column{.5\textwidth}
            {\bf Funding:}
            \begin{myitemize}
                \item NSF
                \item KU Grad Studies, EEB \& BI
                \item SSB
                \item Sigma Xi
            \end{myitemize}
            \smallskip
            {\bf Photo credits:}
            \begin{myitemize}
                \item Rafe Brown, Cam Siler, \& Jake Esselstyn
                \item FMNH Philippine Mammal Website:
                    \begin{myitemize}
                        \item D.S.\ Balete, M.R.M.\ Duya, \& J.\ Holden
                    \end{myitemize}
                \item PhyloPic!
            \end{myitemize}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Questions?}    
    \begin{center}
        {
        \Large
        \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
        }
    \end{center}
\end{frame}

% Extra slides

\begin{frame}[noframenumbering]
    \frametitle{Easy as ABC}
    \centerline{
    \includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}
    \includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}
    \includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}
    \includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}
    \includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}
    \includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}
    \includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}
    \includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}
    \includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}
    \includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}
    }
\end{frame}

\begin{frame}[noframenumbering]
    \frametitle{Causes of bias: Insufficient sampling}
    \begin{itemize}
        \item Models with more parameter space are less densely sampled
        \item Could explain bias toward small models in extreme cases
        \item {\bf Predicts large variance in posterior estimates}
        \begin{itemize}
            \item We explored empirical and simulation-based analyses with
                2, 5, and 10 million prior samples, and estimates were
                very similar
        \end{itemize}
    \end{itemize}
    \smallskip
    \centerline{
    \includegraphics[width=\textwidth]{../images/omega_over_sampling.pdf}}
\end{frame}

\begin{frame}[noframenumbering]
    \frametitle{\dppmsbayes: Simulation results}
    % \vspace{1cm}
        \centerline{
        \includegraphics[width=\textwidth]{../images/old_old_power_psi_mode.pdf}}
        \vspace{0mm}
        \centerline{
        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_psi_mode_headless.pdf}}
        \vspace{0mm}
        \centerline{
        \includegraphics[width=\textwidth]{../images/old_dpp_power_psi_mode_headless.pdf}}
    % \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\begin{frame}[noframenumbering]
    \frametitle{\dppmsbayes: Simulation results}
    % \vspace{1cm}
        \centerline{
        \includegraphics[width=\textwidth]{../images/old_old_power_omega_median.pdf}}
        \vspace{0mm}
        \centerline{
        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_omega_median_headless.pdf}}
        \vspace{0mm}
        \centerline{
        \includegraphics[width=\textwidth]{../images/old_dpp_power_omega_median_headless.pdf}}
    % \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\begin{frame}[noframenumbering]
    \frametitle{Empirical results: Philippine diversification}
    \centerline{
    \includegraphics[width=0.8\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp-with-prior.pdf}}
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\end{document}

### 2014-11-01 slides/bin/plot-validation-results.py
Updating validation plotting script for slides.
The script now generates plots of validation results that are
easier to use in slides.

@@ -57,23 +57,52 @@ def main_cli():
             share_x = True,
             share_y = True,
             title = 'Posterior probability of one divergence',
-            title_size = 12.0,
+            title_size = 14.0,
             title_top = False,
             y_title = 'True probability of one divergence',
             y_title_position = 0.001,
-            y_title_size = 12.0,
-            width = 5.2,
-            height = 4.0,
+            y_title_size = 14.0,
+            column_labels = [r'$M_{msBayes}$'],
+            column_label_size = 22.0,
+            column_label_offset = 0.04,
+            width = 4.6,
+            height = 3.7,
             auto_height = False)
     pg.label_schema = None
     pg.auto_adjust_margins = False
-    pg.margin_bottom = 0.05
-    pg.margin_left = 0.04
-    pg.margin_top = 1
-    pg.margin_right = 1
+    pg.margin_bottom = 0.06
+    pg.margin_left = 0.05
+    pg.margin_top = 0.91
+    pg.margin_right = 0.94
     pg.reset_figure()
 
     pg.savefig('../images/validation-model-choice-old.pdf')
+
+    pg = plotting.PlotGrid(subplots = sub_plot_list[-1:],
+            num_columns = 1,
+            share_x = True,
+            share_y = True,
+            title = 'Posterior probability of one divergence',
+            title_size = 14.0,
+            title_top = False,
+            y_title = 'True probability of one divergence',
+            y_title_position = 0.001,
+            y_title_size = 14.0,
+            column_labels = [r'$M_{DPP}$'],
+            column_label_size = 22.0,
+            column_label_offset = 0.04,
+            width = 4.6,
+            height = 3.7,
+            auto_height = False)
+    pg.label_schema = None
+    pg.auto_adjust_margins = False
+    pg.margin_bottom = 0.06
+    pg.margin_left = 0.05
+    pg.margin_top = 0.91
+    pg.margin_right = 0.94
+    pg.reset_figure()
+
+    pg.savefig('../images/validation-model-choice-dpp.pdf')
 
     pg = plotting.PlotGrid(subplots = sub_plot_list,
             num_columns = 2,
@@ -154,6 +183,75 @@ def main_cli():
     pg.reset_figure()
 
     pg.savefig('../images/validation-model-choice-old-dpp-full.pdf')
+
+    for p in sub_plot_list:
+        p.set_extra_y_label('')
+        p.set_title_text('')
+
+    old_violations = sub_plot_list[1:3]
[…]

### 2014-11-01 slides/bin/partition-number-plots.py
More updating of plots for slides.

@@ -76,7 +76,7 @@ def create_plots(n = 22,
                     y_label_size = y_label_size,
                     ylim = (0, ymax),
                     )
-    ymax = 1.1 * max(prob_div_models)
+    ymax = 1.05 * max(prob_div_models)
     prob_plot = plotting.ScatterPlot(
                     bar_data_list = [prob_bar_data],
                     x_label = '# of divergence events',

### 2014-11-01 slides/bin/plot-validation-results.py

@@ -62,7 +62,7 @@ def main_cli():
             y_title = 'True probability of one divergence',
             y_title_position = 0.001,
             y_title_size = 14.0,
-            column_labels = [r'$M_{msBayes}$'],
+            column_labels = [r'msBayes'],
             column_label_size = 22.0,
             column_label_offset = 0.04,
             width = 4.6,
@@ -88,7 +88,7 @@ def main_cli():
             y_title = 'True probability of one divergence',
             y_title_position = 0.001,
             y_title_size = 14.0,
-            column_labels = [r'$M_{DPP}$'],
+            column_labels = [r'dpp-msbayes'],
             column_label_size = 22.0,
             column_label_offset = 0.04,
             width = 4.6,
@@ -204,7 +204,7 @@ def main_cli():
             height = 6.5,
             width = 4.5,
             auto_height = False,
-            column_labels = [r'$M_{msBayes}$'],
+            column_labels = [r'msBayes'],
             column_label_size = 22.0,
             column_label_offset = 0.05,
             # super_title = r'\textbf{Data model}',
@@ -235,7 +235,7 @@ def main_cli():
             height = 6.5,
             width = 4.5,
             auto_height = False,
-            column_labels = [r'$M_{DPP}$'],
+            column_labels = [r'dpp-msbayes'],
             column_label_size = 22.0,
             column_label_offset = 0.05,
             # super_title = r'\textbf{Data model}',

### 2014-11-01 slides/osu/slides.tex
Working on osu slides.
Adding photos and cartoons.

@@
     \begin{columns}[c]
-        \column{.4\textwidth}
+        \column{.5\textwidth}
         \begin{onlyenv}<1-4>
@@
                         \[
+                            \divTimeVector = \{\divTime{1}\}
+                        \]\vspace{0mm}
+                        \[
+                            model = 111
+                        \]\vspace{0mm}
+                        \[
                             \divTimeMapVector = (\divTimeMap{1},\divTimeMap{2},\divTimeMap{3})
-                        \]\vspace{0mm}
-                        \[
-                            model = 111
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeVector = \{\divTime{1}\}
                         \]\vspace{0mm}
@@
                         \[
-                            \divTimeMapVector = (260,260,260)
+                            \divTimeVector = \{130\}
                         \]\vspace{0mm}
@@
                         \[
-                            \divTimeVector = \{260\}
+                            \divTimeMapVector = (130,130,130)
                         \]\vspace{0mm}
@@
                         \[
-                            \divTimeMapVector = (397,260,260)
+                            \divTimeVector = \{130,397\}
                         \]\vspace{0mm}
@@
                         \[
-                            \divTimeVector = \{260,397\}
+                            \divTimeMapVector = (397,130,130)
                         \]\vspace{0mm}
@@
                         \[
-                            \divTimeMapVector = (260,397,260)
+                            \divTimeVector = \{130,397\}
                         \]\vspace{0mm}
@@
                         \[
-                            \divTimeVector = \{260,397\}
+                            \divTimeMapVector = (130,397,130)
                         \]\vspace{0mm}
@@
                         \[
-                            \divTimeMapVector = (260,260,397)
+                            \divTimeVector = \{130,397\}
                         \]\vspace{0mm}
@@
                         \[
-                            \divTimeVector = \{260,397\}
+                            \divTimeVector = \{130,397\}
                         \]\vspace{0mm}
@@
         \end{onlyenv}
-        \begin{onlyenv}<10>
+        \begin{onlyenv}<10-12>
             \begin{minipage}[c][0.5\textheight][c]{\linewidth}
@@
                         \[
-                            \divTimeMapVector = (260,95,397)
-                        \]\vspace{0mm}
-                        \[
-                            model = 123
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeVector = \{260,95,397\}
+                            \divTimeVector = \{130,397,260\}
+                        \]\vspace{0mm}
+                        \[
+                            model = 312
+                        \]\vspace{0mm}
+                        \[
+                            \divTimeMapVector = (260,130,397)
                         \]\vspace{0mm}
@@
         %                 \[
-        %                     \divTimeMapVector = (260,260,260)
+        %                     \divTimeVector = \{\divTime{1},\ldots,\divTime{\divTimeNum}\}
         %                 \]\vspace{0mm}
         %                 \[
-        %                     model = 111
+        %                     model = \divModel{i}
         %                 \]\vspace{0mm}
         %                 \[
-        %                     \divTimeVector = \{260\}
+        %                     \divTimeMapVector = (\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})
         %                 \]\vspace{0mm}
@@
         % \end{onlyenv}
-        \begin{onlyenv}<11-13>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeMapVector = (\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})
-                        \]\vspace{0mm}
-                        \[
-                            model = \divModel{i}
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeVector = \{\divTime{1},\ldots,\divTime{\divTimeNum}\}
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<14>
+        \begin{onlyenv}<13>
             \begin{minipage}[c][0.5\textheight][c]{\linewidth}
@@
                     \item[\divModel{}] Divergence model
-                    \item[\geneTreeVector] Gene trees
-                    \item[\hkyModelVector] Substitution parameters
-                    \item[\demographicParamVector] Demographic parameters
+                    \item[\allParameters{}] Gene trees
+                    \item[\ ] Substitution parameters
+                    \item[\ ] Demographic parameters
                     % \item[\allParameters] $(\divTimeMapVector, \geneTreeVector,
@@
         \end{onlyenv}
-        \begin{uncoverenv}<12->
+        \begin{uncoverenv}<11->
             \begin{minipage}[c][0.18\textheight][c]{\linewidth}
@@
         \begin{minipage}[t][0.8\textheight][c]{\linewidth}
-        \includegraphics<1>[height=6cm]{../images/community-ellipse-small.pdf}
-        % \includegraphics<2>[height=6cm]{../images/community-ellipse-barrier-small.pdf}
-        \includegraphics<2>[height=6cm]{../images/community-ellipse-split-small.pdf}
-        \includegraphics<3>[height=6cm]{../images/div-model-cartoon-taxa.pdf}
-        \includegraphics<4>[height=6cm]{../images/div-model-cartoon-shared-no-labels.pdf}
-        \includegraphics<5-6>[height=6cm]{../images/div-model-cartoon-shared.pdf}
-        \includegraphics<7>[height=6cm]{../images/div-model-cartoon-2-1.pdf}
-        \includegraphics<8>[height=6cm]{../images/div-model-cartoon-2-2.pdf}
-        \includegraphics<9>[height=6cm]{../images/div-model-cartoon-2-3.pdf}
-        \includegraphics<10>[height=6cm]{../images/div-model-cartoon-general.pdf}
-        \includegraphics<11-12>[height=6cm]{../images/div-model-cartoon-shared.pdf}
-        \includegraphics<13->[height=6cm]{../images/div-model-cartoon-sp-tree-shared.pdf}
+        \includegraphics<1>[height=6cm]{../images/dmc-cartoon-islands.pdf}
+        \includegraphics<2>[height=6cm]{../images/dmc-cartoon-taxa.pdf}
+        \includegraphics<3>[height=6cm]{../images/dmc-cartoon-sea-level.pdf}
+        \includegraphics<4>[height=6cm]{../images/dmc-cartoon-shared-no-labels.pdf}
+        \includegraphics<5-6>[height=6cm]{../images/dmc-cartoon-shared.pdf}
+        \includegraphics<7>[height=6cm]{../images/dmc-cartoon-2-1.pdf}
+        \includegraphics<8>[height=6cm]{../images/dmc-cartoon-2-2.pdf}
+        \includegraphics<9>[height=6cm]{../images/dmc-cartoon-2-3.pdf}
+        \includegraphics<10-11>[height=6cm]{../images/dmc-cartoon-general.pdf}
+        \includegraphics<12->[height=6cm]{../images/dmc-cartoon-pop-general.pdf}
         \end{minipage}
@@
                   \divTimeMapVector,
-                  \geneTreeVector,
-                  \hkyModelVector,
-                  \demographicParamVector
+                  \allParameters{}
                   \given \alignmentVector, \divModel{i})
@@
                       \divTimeMapVector,
-                      \geneTreeVector,
-                      \hkyModelVector,
-                      \demographicParamVector,
+                      \allParameters{},
                       \divModel{i}
@@
                         \divTimeMapVector,
-                        \geneTreeVector,
-                        \hkyModelVector,
-                        \demographicParamVector
+                        \allParameters{}
                         \given \divModel{i}
@@
                 p(\alignmentVector \given \divModel{i}) =
-                \int_{\allParameters{i}}
-                p(\alignmentVector \given \allParameters{i}, \divModel{i})
-                p(\allParameters{i} \given \divModel{i})
-                d\allParameters{i}
+                \int_{\divTimeMapVector} \int_{\allParameters{}}
+                p(\alignmentVector \given \divTimeMapVector, \allParameters{}, \divModel{i})
+                p(\divTimeMapVector, \allParameters{} \given \divModel{i})
+                d\divTimeMapVector d\allParameters{}
             \]\vspace{-1mm}
@@
 
+\ignore{
 \begin{frame}
@@
             \def\firstrowcolor{}
-            \def\secondrowcolor{}
-            \def\thirdrowcolor{}
-            \def\fourthrowcolor{}
-            \def\fifthrowcolor{}
+            % \def\secondrowcolor{}
+            % \def\thirdrowcolor{}
+            % \def\fourthrowcolor{}
+            % \def\fifthrowcolor{}
             \only<2>{\def\firstrowcolor{\rowcolor{rowred}}}
-            \only<3>{\def\secondrowcolor{\rowcolor{rowred}}}
-            \only<4>{\def\thirdrowcolor{\rowcolor{rowred}}}
-            \only<5>{\def\fourthrowcolor{\rowcolor{rowred}}}
-            \only<6>{\def\fifthrowcolor{\rowcolor{rowred}}}
+            % \only<3>{\def\secondrowcolor{\rowcolor{rowred}}}
+            % \only<4>{\def\thirdrowcolor{\rowcolor{rowred}}}
+            % \only<5>{\def\fourthrowcolor{\rowcolor{rowred}}}
+            % \only<6>{\def\fifthrowcolor{\rowcolor{rowred}}}
             \centering
@@
                 \textbf{\textsc{Mammals}} & & \\
-                \fifthrowcolor  \emph{Crocidura beatus}             & 12 & 11 \\
+                  \emph{Crocidura beatus}             & 12 & 11 \\
                 \firstrowcolor  \emph{Crocidura negrina-panayensis} & 12 & 6  \\
-                \thirdrowcolor  \emph{Hipposideros obscurus}        & 19 & 9  \\
-                \secondrowcolor \emph{Hipposideros pygmaeus}        & 3  & 12 \\
-                \thirdrowcolor  \emph{Cynopterus brachyotis}        & 20 & 8  \\
+                  \emph{Hipposideros obscurus}        & 19 & 9  \\
+                  \emph{Hipposideros pygmaeus}        & 3  & 12 \\
+                  \emph{Cynopterus brachyotis}        & 20 & 8  \\
                 \firstrowcolor  \emph{Cynopterus brachyotis}        & 8  & 14 \\
-                \thirdrowcolor  \emph{Haplonycteris fischeri}       & 29 & 8  \\
+                  \emph{Haplonycteris fischeri}       & 29 & 8  \\
                 \firstrowcolor  \emph{Haplonycteris fischeri}       & 9  & 21 \\
-                \thirdrowcolor  \emph{Macroglossus minimus}         & 19 & 4  \\
+                  \emph{Macroglossus minimus}         & 19 & 4  \\
                 \firstrowcolor  \emph{Macroglossus minimus}         & 8  & 10 \\
-                \thirdrowcolor  \emph{Ptenochirus jagori}           & 4  & 7  \\
+                  \emph{Ptenochirus jagori}           & 4  & 7  \\
                 \firstrowcolor  \emph{Ptenochirus jagori}           & 8  & 8  \\
-                \thirdrowcolor  \emph{Ptenochirus minor}            & 30 & 9  \\
+                  \emph{Ptenochirus minor}            & 30 & 9  \\
                 \textbf{\textsc{Squamates}} & & \\
-                \fifthrowcolor  \emph{Cyrtodactylus gubaot-sumuroi} & 29 & 6  \\
-                \secondrowcolor \emph{Cyrtodactylus annulatus}      & 14 & 3  \\
+                  \emph{Cyrtodactylus gubaot-sumuroi} & 29 & 6  \\
+                  \emph{Cyrtodactylus annulatus}      & 14 & 3  \\
                 \firstrowcolor  \emph{Cyrtodactylus philippinicus}  & 6  & 14 \\
@@
                 \firstrowcolor  \emph{Insulasaurus arborens}        & 22 & 10 \\
-                \fourthrowcolor \emph{Pinoyscincus jagori}          & 8  & 8  \\
+                  \emph{Pinoyscincus jagori}          & 8  & 8  \\
                 \firstrowcolor  \emph{Dendrelaphis marenae}         & 6  & 6  \\
                 \textbf{\textsc{Anurans}}  & & \\
-                \secondrowcolor \emph{Limnonectes leytensis}        & 4  & 2  \\
-                \secondrowcolor \emph{Limnonectes magnus}           & 2  & 3  \\
+                  \emph{Limnonectes leytensis}        & 4  & 2  \\
+                  \emph{Limnonectes magnus}           & 2  & 3  \\
                 \hline
@@
         \includegraphics<2>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-bohol_mindanao.png}
-        \includegraphics<4>[width=\textwidth]{../images/maps/Philippines-leyte_mindanao.png}
-        \includegraphics<5>[width=\textwidth]{../images/maps/Philippines-mindanao_samar.png}
-        \includegraphics<6>[width=\textwidth]{../images/maps/Philippines-leyte_samar.png}
+        % \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-bohol_mindanao.png}
+        % \includegraphics<4>[width=\textwidth]{../images/maps/Philippines-leyte_mindanao.png}
+        % \includegraphics<5>[width=\textwidth]{../images/maps/Philippines-mindanao_samar.png}
+        % \includegraphics<6>[width=\textwidth]{../images/maps/Philippines-leyte_samar.png}
+\end{columns}
+\end{frame}
+}
+
+\begin{frame}
+\begin{columns}[c]
+    \column{.5\textwidth}
+        \centerline{
+        \includegraphics<1>[height=1.5cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1>[height=1.5cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=1.5cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1>[height=1.5cm]{../images/photos/gekko-mindorensis.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=1.5cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1>[height=1.5cm]{../images/photos/cyrt-agusanensis.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=1.5cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1>[height=1.5cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
+    \column{.5\textwidth}
+        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
+        \includegraphics<2>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
 \end{columns}

### 2014-11-02 slides/osu/slides.tex
More work on osu slides.

@@
     \frametitle{Southeast Asia}    
-\end{frame}
-}
+    \begin{columns}
+        \column{0.6\textwidth}
+        \ \\
+
+        \column{0.45\textwidth}
+
+        \vspace{-2.4cm}
+
+        \begin{uncoverenv}<2->
+        \colorbox{white}{
+            \begin{minipage}[t][3.6cm][t]{1.1\textwidth}
+                \begin{onlyenv}<2-3>
+                \begin{itemize}
+                    \item<2-> Repeated coalescence and fragmentation of island complexes
+                    \item<3-> Prominent paradigm for explaining Philippine biodiversity
+                \end{itemize}
+                \end{onlyenv}
+
+                \begin{onlyenv}<4>
+                \begin{itemize}
+                    \item<4> \textbf{Did repeated fragmentation of islands
+                        during inter-glacial rises in sea level promote
+                        diversification?}
+                \end{itemize}
+                \end{onlyenv}
+
+                \begin{onlyenv}<5->
+                \begin{itemize}
+                    \item<5-> Climate-driven model of diversification
+                    \item<6-> Testable prediction:
+                    \begin{itemize}
+                        \item Temporally clustered divergences among taxa
+                            co-distributed across fragmented islands
+                    \end{itemize}
+                \end{itemize}
+                \end{onlyenv}
+            \end{minipage}
+        }
+    \end{uncoverenv}
+    \end{columns}
+\end{frame}
+}
+
+% \begin{frame}
+%     \frametitle{Philippine Archipelago}
+%     \begin{center}
+%         \includegraphics<1>[width=.5\textwidth]{../images/maps/Philippines-present.png}
+%         \includegraphics<2>[width=.5\textwidth]{../images/maps/Philippines.png}
+%     \end{center}
+% \end{frame}
+
+% \begin{frame}
+%     \frametitle{Climate-driven diversification model}
+%     \begin{columns}[c]
+%         \column{.5\textwidth}
+%             \begin{myitemize}
+%                 \item<1-> Repeated coalescence and fragmentation of island complexes
+%                 \item<2-> Prominent paradigm for explaining Philippine biodiversity
+%                 \item<3-> Proposed as model of diversification
+%             \end{myitemize}
+%         \column{.5\textwidth}
+%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
+%     \end{columns}
+% \end{frame}
+
+% \begin{frame}
+%     \frametitle{Testing climate-driven diversification}
+%     \begin{columns}[c]
+%         \column{.5\textwidth}
+%     \textbf{Did repeated fragmentation of islands during inter-glacial rises in sea level
+%     promote diversification?}\\
+%     \bigskip
+%     \uncover<2->{
+%     Model has testable prediction:\\
+%     \begin{itemize}
+%         \item Temporally clustered divergences among taxa
+%             co-distributed across fragmented islands
+%     \end{itemize}
+%     }
+%         \column{.5\textwidth}
+%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
+%     \end{columns}
+% \end{frame}
 
@@
 %       ii. Rather than publish, did some detective work
+
+{
+\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
+\begin{frame}
+    \frametitle{Climate-driven diversification}    
+    \begin{columns}
+        \column{0.6\textwidth}
+        \ \\
+
+        \column{0.4\textwidth}
+
+        \vspace{-2cm}
+
+        \colorbox{white}{
+            \begin{minipage}[t]{1.0\textwidth}
+                \raggedright
+                \textbf{Did repeated fragmentation of islands during
+                    inter-glacial rises in sea level promote diversification?}
+            \end{minipage}
+        }
+    \end{columns}
+\end{frame}
+}
 
@@
 \begin{frame}
+    \frametitle{Climate-driven diversification}
 \begin{columns}[c]
     \column{.5\textwidth}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=1.5cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=1.5cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=1.5cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
+        \vspace{-1cm}
+        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
+        \end{minipage}
+
     \column{.5\textwidth}
+        \vspace{-1cm}
+        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
         \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
         \includegraphics<2>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-\end{columns}
-\end{frame}
-
-\section{Empirical results}
-\subsection{Too good to be true?}
-
-% \begin{frame}
-% \frametitle{Outline}
-% \tableofcontents[currentsection,currentsubsection]
-% \end{frame}
-
-\begin{frame}
-    \frametitle{Empirical results}
-    \begin{columns}[c]
-        \column{.4\textwidth}
-            {\small
-            Strong support for simultaneous divergence of all 22 taxon pairs \\
-            \vspace{1cm}
-            pp $>$ 0.96 \\
-            \vspace{1cm}
-            $\sim$100,000--250,000 years ago}
-        \column{.6\textwidth}
-        \includegraphics[width=\textwidth]{../images/jointDensityPlotsSlide.jpg}
+        \begin{onlyenv}<3->
+            \begin{itemize}
+                \item Strong support for simultaneous divergence of all 22 taxon pairs
+
+                \item Posterior probability $>$ 0.96 \\
+
+                \item $\sim$100,000--250,000 years ago
+            \end{itemize}
+        \end{onlyenv}
+        \end{minipage}
 \end{columns}
@@
     \frametitle{Simulation-based power analyses}
-    What is ``simultaneous''?
+    % What is ``simultaneous''?
     \begin{myitemize}
-        \item<1-> Simulate datasets in which all 22 divergence times are random
-        \uncover<1->{
+        \item<1-> What if divergences were random?
+        \item<2-> Simulate datasets in which all 22 divergence times are random
+        \uncover<3->{
             \smallskip
@@
             \end{myitemize}
-        \item<1-> $MGA$ = Millions of Generations Ago
-        }
-        \item<2-> Simulate 1000 datasets for each \divTime{} distribution
-        \item<2-> Analyze all 4000 datasets as we did the empirical data
+        \item<3-> $MGA$ = Millions of Generations Ago
+        }
+        \item<4-> Simulate 1000 datasets for each \divTime{} distribution
+        \item<4-> Analyze all 4000 datasets as we did the empirical data
     \end{myitemize}
@@
 
-\begin{frame}[t]
+\begin{frame}
     \frametitle{Simulation-based power analyses: Results}
-    \vspace{1cm}
         \centerline{
         \includegraphics[width=1.13\textwidth]{../images/old_old_power_psi_mode.pdf}}
-        % \vspace{0mm}
-        % \centerline{
-        % \includegraphics<2>[width=1.13\textwidth]{../images/old-sims_power_psi_prob_headless.pdf}}
-    \barefootnote{\shortfullcite{Oaks2012}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Simulation-based power analyses: Results}
-        \msb will often infer clustered divergences when divergences are random
-        over millions of generations. \\
-
-    \bigskip
-    The empirical results are likely spurious \\
-
-    \begin{onlyenv}<2->
+
+    \begin{uncoverenv}<2->
+        \begin{itemize}[<+->]
+            \item \msb will often infer clustered divergences when divergences
+                are random over millions of generations.
+
+            \item The empirical results are likely spurious
+        \end{itemize}
+    \end{uncoverenv}
+
+    \begin{uncoverenv}<3->
         \begin{block}{Objective:}
@@
         \end{block}
-    \end{onlyenv}
+    \end{uncoverenv}
     \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
@@
 
-\begin{frame}
+\begin{frame}[t]
     \frametitle{Prior on divergence models}
@@
     \end{itemize}
-    % \smallskip
-    \centerline{
-        \includegraphics<1->[width=\textwidth]{../images/partition_numbers.pdf}}
-    \begin{block}<2->{\it Potential solution:}
+
+    \begin{uncoverenv}<2->
+    \begin{columns}
+        \column{0.5\textwidth}
+            \centerline{
+                \includegraphics<2->[width=1.0\textwidth]{../images/number-of-div-models-22-unordered.pdf}}
+        \column{0.5\textwidth}
+            \centerline{
+                \includegraphics<3->[width=1.0\textwidth]{../images/prob-of-div-models-22-unordered.pdf}}
+    \end{columns}
+    \end{uncoverenv}
+
+    \begin{uncoverenv}<4->
+    \begin{block}{\it Potential solution:}
         Place flexible prior directly on the sample space of divergence models
     \end{block}
+    \end{uncoverenv}
 \end{frame}
@@
     \frametitle{New method: \dppmsbayes}
-    \begin{itemize}
+    \begin{itemize}[<+->]
         % \item<1-> Reparameterized the model implemented in \msb
@@
     \frametitle{\dppmsbayes: Simulation-based assessment}
-        Simulate 50,000 datasets under three models\\
-        \smallskip
-        \begin{description}
-            \item[$M_{msBayes}$]
-                \begin{itemize}
-                    \item U-shaped prior on divergence models
-                    \item Uniform priors on continuous parameters
-                \end{itemize}
-            \smallskip
-            \item[$M_{Ushaped}$]
-                \begin{itemize}
-                    \item U-shaped prior on divergence models
-                    \item Gamma priors on continuous parameters
-                \end{itemize}
-            \smallskip
-            % \item[$M_{Uniform}$]
-            %     \begin{itemize}
-            %         \item Uniform prior on divergence models
-            %         \item Gamma priors on continuous parameters
-            %     \end{itemize}
-            % \smallskip
-            \item[$M_{DPP}$]
-                \begin{itemize}
-                    \item DPP prior on divergence models
-                    \item Gamma priors on continuous parameters
-                \end{itemize}
-        \end{description}
-        \smallskip
-        Analyze all datasets under each of the models
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Simulation results}
-    \centerline{
-        \includegraphics<1>[width=0.85\textwidth]{../images/validation-model-choice-old-dpp.pdf}
-        \includegraphics<2>[width=1.13\textwidth]{../images/validation-model-choice-old-dpp-full.pdf}}
+        \uncover<1->{
+        Validation:\\
+
+        \smallskip
+        \begin{itemize}
+            \item Simulate 50,000 datasets and analyze each under the same
+                model
+        \end{itemize}
+        }
+
+        \uncover<2->{
+        \bigskip
+        Robustness:\\
+        \begin{itemize}
+            \item Simulate datasets that violate model assumptions and analyze
+                each of them
+        \end{itemize}
+        }
+
+        \uncover<3->{
+        \bigskip
+        Done for \msb and \dppmsbayes
+        }
+\end{frame}
+
+\begin{frame}
+    \frametitle{\dppmsbayes: Validation results}
+    \begin{columns}
+        \column{0.5\textwidth}
+            \includegraphics<1->[width=1.0\textwidth]{../images/validation-model-choice-old.pdf}
+        \column{0.5\textwidth}
+            \includegraphics<2->[width=1.0\textwidth]{../images/validation-model-choice-dpp.pdf}
+    \end{columns}
     \barefootnote{\shortfullcite{Oaks2014dpp}}
@@
 \begin{frame}
+    \frametitle{\dppmsbayes: Robustness results}
+    \begin{columns}
+        \column{0.5\textwidth}
+            \centering{
+            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-old-violations.pdf}
+            }
+        \column{0.5\textwidth}
+            \centering{
+            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-dpp-violations.pdf}
+            }
+    \end{columns}
+
+    \begin{uncoverenv}<2->
+    \begin{itemize}
+        \item Improved model-choice accuracy and robustness
+    \end{itemize}
+    \end{uncoverenv}
+    \vspace{-0.5cm}
+    \barefootnote{\shortfullcite{Oaks2014dpp}}
+\end{frame}
+
+\begin{frame}
     \frametitle{\dppmsbayes: Simulation-based power analyses}
+    Does the new method have power to detect variation in divergence times?\\
+
+    \bigskip
     \begin{myitemize}
-        \item Simulate datasets in which all 22 divergence times are random
-        \uncover{
+        \item<2-> Simulate datasets in which all 22 divergence times are random
             \smallskip
@@
             \end{myitemize}
-        \item $MGA$ = Millions of Generations Ago
-        }
-        \item Simulate 1000 datasets for each \divTime{} distribution
-        % \item Analyze all 4000 datasets as we did the empirical data
-        \item Analyze all 4000 datasets under models $M_{msBayes}$, $M_{Ushaped}$, and $M_{DPP}$
+        \item<2-> $MGA$ = Millions of Generations Ago
+
+        \item<3-> Simulate 1000 datasets for each \divTime{} distribution
+        \item<3-> Analyze all 4000 datasets with \msb and \dppmsbayes
     \end{myitemize}
@@
     \frametitle{\dppmsbayes: Power results}
-    \vspace{1cm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<1->{../images/old_old_power_psi_mode.pdf}}
+        \begin{onlyenv}<1-2>
+        \centerline{
+        \includegraphics[width=1.13\textwidth]<1-2>{../images/power-psi-old-annotate-no-pp.pdf}}
         \vspace{0mm}
         \centerline{
-        \includegraphics[width=1.13\textwidth]<2->{../images/old_dpp_power_psi_mode_headless.pdf}}
+        \includegraphics[width=1.13\textwidth]<2>{../images/power-psi-dpp-annotate-no-pp.pdf}}
+        \end{onlyenv}
+
+        \begin{onlyenv}<3->
+        \centerline{
+        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-old-annotate-pp.pdf}}
+        \vspace{0mm}
+        \centerline{
+        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-dpp-annotate-pp.pdf}}
+        \end{onlyenv}
+
+    \begin{uncoverenv}<4->
+    \begin{itemize}
+        \item Improved power to detect temporal variation across divergences
+    \end{itemize}
+    \end{uncoverenv}
     \barefootnote{\shortfullcite{Oaks2014dpp}}
@@
 
-\begin{frame}[t]
-    \frametitle{\dppmsbayes: Power results}
-    \vspace{1cm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]{../images/old_old_power_psi_prob.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]{../images/old_dpp_power_psi_prob_headless.pdf}}
+{
+\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
+\begin{frame}
+    \frametitle{Climate-driven diversification}    
+    \begin{columns}
+        \column{0.6\textwidth}
+        \ \\
+
+        \column{0.4\textwidth}
+
+        \vspace{-2cm}
+
+        \colorbox{white}{
+            \begin{minipage}[t]{1.0\textwidth}
+                \raggedright
+                \textbf{Did repeated fragmentation of islands during
+                    inter-glacial rises in sea level promote diversification?}
+            \end{minipage}
+        }
+    \end{columns}
+\end{frame}
+}
+
+\begin{frame}
+    \frametitle{Climate-driven diversification}
+\begin{columns}[c]
+    \column{.5\textwidth}
+        \vspace{-1cm}
+        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
+        % \vspace{0.5mm}
+        \centerline{
+        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
+        \end{minipage}
+
+    \column{.5\textwidth}
+        \vspace{-1cm}
+        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
+        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
+        \end{minipage}
+\end{columns}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Empirical results}
+    \centerline{
+    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp.pdf}}
+
+    \begin{uncoverenv}<2->
+    \begin{itemize}
+        \item More ``honest'' estimate of posterior uncertainty
+    \end{itemize}
+    \end{uncoverenv}
     \barefootnote{\shortfullcite{Oaks2014dpp}}
@@
 \begin{frame}
-    \frametitle{\dppmsbayes: Power results}
-    % \vspace{1cm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_old_power_psi_prob.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_psi_prob_headless.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_dpp_power_psi_prob_headless.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Empirical results: Philippine diversification}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Conclusions}
+    \frametitle{Empirical results}
+    \centerline{
+    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
+    \smallskip
+    \centerline{
+    \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Summary}
     \begin{itemize}
@@
             \begin{enumerate}
-            \item Estimation of posterior uncertainty
-            \item Model-choice accuracy
-            \item Power to detect temporal variation across divergences
-            \item Robustness to model violations
-            \item Speed!
+                \item<2-> Estimation of posterior uncertainty
+                \item<2-> Model-choice accuracy
+                \item<2-> Power to detect temporal variation across divergences
+                \item<2-> Robustness to model violations
+                \item<2-> Speed!
             \end{enumerate}
 
-    \end{itemize}
-
-    \ignore{
-    \bigskip
-    \begin{uncoverenv}<2->
-    Caveats:\\
-    \begin{itemize}
-            \item<1-> Estimating a very rich (600+ parameters for 22 taxa) model
-                using limited information from the data
-
-            \begin{itemize}
-                \item Likely sensitive to prior assumptions
-                \item Be skeptical of strongly supported results
-            \end{itemize}
-
-    \end{itemize}
-    \end{uncoverenv}
-    }
-\end{frame}
-
-\begin{frame}
-    \frametitle{Recommendations}
-    For Bayesian model choice, choose priors carefully\\
-    \bigskip
-    ABC model choice estimates should be accompanied by:
-    \begin{enumerate}
-        \item Simulation-based power analyses
-        \item Assessment of prior sensitivity
-    \end{enumerate}
+        \smallskip
+        \item<3-> Climate-driven diversification model?
+            \begin{itemize}
+                \item<4-> Results consistent with prediction of clustered
+                    divergences
+                \item<4-> Results suggest multiple co-divergences
+                \item<4-> However, there is a lot of uncertainty!
+            \end{itemize}
+
+        \smallskip
+        \item<5-> What to do?
+            \begin{itemize}
+                \item<6-> More data!
+                \item<6-> More power! 
+            \end{itemize}
+    \end{itemize}
 \end{frame}
@@
 
-\begin{frame}
-    \frametitle{Testing limate-driven diversification}
-    \textbf{Did repeated fragmentation of islands during inter-glacial rises in sea level
-    promote diversification?}\\
-
-    \bigskip
-    \begin{uncoverenv}<2->
-    \ldots left with uncertainty regarding climate-driven diversification \\
-    \end{uncoverenv}
-
-    \begin{uncoverenv}<3->
-    \begin{itemize}
-            \item More data
-
-            \item More power 
-    \end{itemize}
-    \end{uncoverenv}
-\end{frame}
-
 %   g. Collected thousands of loci from across the genome of almost 300
@@
 \begin{frame}
-    \frametitle{Genomic data}
+    \frametitle{More data}
     \begin{columns}[c]
@@
 \begin{frame}
-    Prelim results from 5 population pairs/1000 loci \ldots
+    \frametitle{More data---preliminary results}
+    Results for 1000 loci from 5 pairs of \emph{Gekko mindorensis} populations\\
+
+    \begin{center}
+        \includegraphics<1->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
+    \end{center}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Full-likelihood approach}
+    \frametitle{More power}
     \begin{itemize}[<+->]
@@
             \begin{itemize}
-                \item \ldots inefficient
-            \end{itemize}
-        \item Currently working on incorporating new developments on analytical
-            integration of gene trees
-            \footnote{\tiny\shortfullcite{Bryant2012}}
+                \item Uses all the information in the sequence data
+                \item Applicable to much deeper timescales
+                \item \ldots Currently inefficient
+            \end{itemize}
+        \item Working to incorporate new developments on analytical integration
+            of gene trees \footnote{\tiny\shortfullcite{Bryant2012}}
             \begin{itemize}
@@
                 \item Applicable to NGS datasets
-                \item Uses all the information in the sequence data
             \end{itemize}
@@
     \begin{center}
-        % \includegraphics<1->[height=5.5cm]{../images/div-model-cartoon-sp-tree-shared-wide.pdf}
-        \includegraphics<1->[height=5.5cm]{../images/div-model-cartoon-sp-phylogeny.pdf}
+        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
+        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
     \end{center}
@@
 
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../../../../Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}}
-\begin{frame}
-    % \frametitle{Gobi Desert}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../../../../Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}}
-\begin{frame}
-    % \frametitle{Gobi Desert}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../../../../Dropbox/field-photos/africa/nigeria-landscape.jpg}}
-\begin{frame}
-    % \frametitle{Gobi Desert}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../../../../Dropbox/field-photos/malaysia/malaysia-burungs.jpg}}
-\begin{frame}
-    % \frametitle{Gobi Desert}
-\end{frame}
-}
+\begin{frame}
+    \frametitle{Gobi Desert}
+    \vspace{-2mm}
+    \begin{columns}
+        % \column{\dimexpr\paperwidth-10pt}
+        \column{\dimexpr\paperwidth}
+        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
+    \end{columns}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Gobi Desert}
+    \vspace{-2mm}
+    \begin{columns}
+        % \column{\dimexpr\paperwidth-10pt}
+        \column{\dimexpr\paperwidth}
+        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
+    \end{columns}
+\end{frame}
+
+\begin{frame}[t]
+    \frametitle{West-Central Africa}
+    \begin{center}
+        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
+        \hspace{0.3mm}
+        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure4_herps.jpeg}
+    \end{center}
+
+    \uncover<2->{
+    \begin{center}
+        \includegraphics[height=4.5cm]{../images/number-of-divs-africa.pdf}
+    \end{center}
+    }
+\end{frame}
+
+\begin{frame}
+    \frametitle{``Micro''-islands of the Sunda Shelf}
+    \vspace{-2mm}
+    \begin{columns}
+        % \column{\dimexpr\paperwidth-10pt}
+        \column{\dimexpr\paperwidth}
+        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/malaysia/malaysia-burungs.jpg}
+    \end{columns}
+\end{frame}
 
@@
 \begin{frame}
-    \frametitle{Southeast Asia}    
+    \frametitle{``Micro''-islands of the Sunda Shelf}
 \end{frame}
@@
             \begin{myitemize}
+                \item Leach\'{e} Lab
+                \item Minin Lab
                 \item Holder Lab
-                \item KU Herpetology
+                \item Brown Lab/KU Herpetology
                 \item Melissa Callahan
@@
             \smallskip
-            {\bf Computation:}
-            \begin{myitemize}
-                \item KU ITTC
-                \item KU Computing Center
-                \item iPlant
-            \end{myitemize}
+            {\bf Computation:}\\
+            \includegraphics<1->[height={8mm}]{../images/iplant.jpg}
+            \hspace{0.5mm}
+            \includegraphics<1->[height={8mm}]{../images/kuittc.png}
+            \hspace{0.5mm}
+            \includegraphics<1->[height={8mm}]{../images/uw.png}\\
+
         \column{.5\textwidth}
-            {\bf Funding:}
-            \begin{myitemize}
-                \item NSF
-                \item KU Grad Studies, EEB \& BI
-                \item SSB
-                \item Sigma Xi
-            \end{myitemize}
+            {\bf Funding:}\\
+            \includegraphics<1->[height={8mm}]{../images/nsf.jpg}
+            \hspace{0.5mm}
+            \includegraphics<1->[height={8mm}]{../images/ngs.jpg}
+            \hspace{0.5mm}
+            \includegraphics<1->[height={8mm}]{../images/ssb.png}\\
+
             \smallskip
@@
             \begin{myitemize}
-                \item Rafe Brown, Cam Siler, \& Jake Esselstyn
+                \item Rafe Brown, Cam Siler, Jesse Grismer, \& Jake Esselstyn
                 \item FMNH Philippine Mammal Website:

### 2014-11-03 slides/osu/slides.tex
Updating osu slides.

@@
 \begin{document}
+% \maketitle
+\begin{frame}
+    \begin{columns}[c]
+        \column{.5\textwidth}
             \maketitle
+        \column{.5\textwidth}
+            \begin{figure}
+                \begin{center}
+                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
+                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
+                \end{center}
+            \end{figure}
+    \end{columns}
+\end{frame}
 
@@
 %       i. E.g., SE Asia---dynamic landscape
+{
+\setbeamercolor{background canvas}{bg=black}
+\setbeamercolor{whitetext}{fg=white}
 \begin{frame}
     \frametitle{Processes of diversification}
-        \begin{itemize}
-            \item<1-> Large-scale geological and climatic processes are
+    \begin{columns}[c]
+        \column{.5\textwidth}
+        {\usebeamercolor[fg]{whitetext}
+        \begin{itemize}
+            \item<2-> Large-scale geological and climatic processes are
                 important in biodiversification and community assembly
-            \item<2-> Accounting for such processes will better our
+            \item<3-> Accounting for such processes will better our
                 understanding of biodiversity
-            \item<3-> We need methods for inferring evolutionary patterns
+            \item<4-> We need methods for inferring evolutionary patterns
                 predicted by historical events from contemporary populations
         \end{itemize}
-\end{frame}
+        }
+        \column{.5\textwidth}
+        \begin{figure}
+            \begin{center}
+            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
+            \end{center}
+        \end{figure}
+    \end{columns}
+\end{frame}
+}
 
@@
                     \item<2-> Repeated coalescence and fragmentation of island complexes
-                    \item<3-> Prominent paradigm for explaining Philippine biodiversity
+                    \item<3-> Prominent paradigm for explaining biodiversity
                 \end{itemize}
@@
 
-\ignore{
-\begin{frame}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \begin{table}%[htbp]
-            \scriptsize
-            \addtolength{\tabcolsep}{-0.09cm}
-            \centering
-            \begin{tabular}{ l c c }
-                % \toprule
-                \textbf{Species} & {\boldmath $n_1$} & {\boldmath $n_2$} \\
-                \hline
-                \textbf{\textsc{Mammals}} & & \\
-                \emph{Crocidura beatus}             & 12 & 11 \\
-                \emph{Crocidura negrina-panayensis} & 12 & 6  \\
-                \emph{Hipposideros obscurus}        & 19 & 9  \\
-                \emph{Hipposideros pygmaeus}        & 3  & 12 \\
-                \emph{Cynopterus brachyotis}        & 20 & 8  \\
-                \emph{Cynopterus brachyotis}        & 8  & 14 \\
-                \emph{Haplonycteris fischeri}       & 29 & 8  \\
-                \emph{Haplonycteris fischeri}       & 9  & 21 \\
-                \emph{Macroglossus minimus}         & 19 & 4  \\
-                \emph{Macroglossus minimus}         & 8  & 10 \\
-                \emph{Ptenochirus jagori}           & 4  & 7  \\
-                \emph{Ptenochirus jagori}           & 8  & 8  \\
-                \emph{Ptenochirus minor}            & 30 & 9  \\
-                \textbf{\textsc{Squamates}} & & \\
-                \emph{Cyrtodactylus gubaot-sumuroi} & 29 & 6  \\
-                \emph{Cyrtodactylus annulatus}      & 14 & 3  \\
-                \emph{Cyrtodactylus philippinicus}  & 6  & 14 \\
-                \emph{Gekko mindorensis}            & 8  & 11 \\
-                \emph{Insulasaurus arborens}        & 22 & 10 \\
-                \emph{Pinoyscincus jagori}          & 8  & 8  \\
-                \emph{Dendrelaphis marenae}         & 6  & 6  \\
-                \textbf{\textsc{Anurans}}  & & \\
-                \emph{Limnonectes leytensis}        & 4  & 2  \\
-                \emph{Limnonectes magnus}           & 2  & 3  \\
-                \hline
-            \end{tabular}
-        \end{table}
-    \column{.5\textwidth}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=1.5cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=1.5cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1>[height=1.5cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-\end{columns}
-\end{frame}
-
-\begin{frame}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \begin{table}%[htbp]
-            \scriptsize
-            \addtolength{\tabcolsep}{-0.09cm}
-            \def\firstrowcolor{}
-            % \def\secondrowcolor{}
-            % \def\thirdrowcolor{}
-            % \def\fourthrowcolor{}
-            % \def\fifthrowcolor{}
-            \only<2>{\def\firstrowcolor{\rowcolor{rowred}}}
-            % \only<3>{\def\secondrowcolor{\rowcolor{rowred}}}
-            % \only<4>{\def\thirdrowcolor{\rowcolor{rowred}}}
-            % \only<5>{\def\fourthrowcolor{\rowcolor{rowred}}}
-            % \only<6>{\def\fifthrowcolor{\rowcolor{rowred}}}
-            \centering
-            \begin{tabular}{ l c c }
-                % \toprule
-                \textbf{Species} & {\boldmath $n_1$} & {\boldmath $n_2$} \\
-                \hline
-                \textbf{\textsc{Mammals}} & & \\
-                  \emph{Crocidura beatus}             & 12 & 11 \\
-                \firstrowcolor  \emph{Crocidura negrina-panayensis} & 12 & 6  \\
-                  \emph{Hipposideros obscurus}        & 19 & 9  \\
-                  \emph{Hipposideros pygmaeus}        & 3  & 12 \\
-                  \emph{Cynopterus brachyotis}        & 20 & 8  \\
-                \firstrowcolor  \emph{Cynopterus brachyotis}        & 8  & 14 \\
-                  \emph{Haplonycteris fischeri}       & 29 & 8  \\
-                \firstrowcolor  \emph{Haplonycteris fischeri}       & 9  & 21 \\
-                  \emph{Macroglossus minimus}         & 19 & 4  \\
-                \firstrowcolor  \emph{Macroglossus minimus}         & 8  & 10 \\
-                  \emph{Ptenochirus jagori}           & 4  & 7  \\
-                \firstrowcolor  \emph{Ptenochirus jagori}           & 8  & 8  \\
-                  \emph{Ptenochirus minor}            & 30 & 9  \\
-                \textbf{\textsc{Squamates}} & & \\
-                  \emph{Cyrtodactylus gubaot-sumuroi} & 29 & 6  \\
-                  \emph{Cyrtodactylus annulatus}      & 14 & 3  \\
-                \firstrowcolor  \emph{Cyrtodactylus philippinicus}  & 6  & 14 \\
-                \firstrowcolor  \emph{Gekko mindorensis}            & 8  & 11 \\
-                \firstrowcolor  \emph{Insulasaurus arborens}        & 22 & 10 \\
-                  \emph{Pinoyscincus jagori}          & 8  & 8  \\
-                \firstrowcolor  \emph{Dendrelaphis marenae}         & 6  & 6  \\
-                \textbf{\textsc{Anurans}}  & & \\
-                  \emph{Limnonectes leytensis}        & 4  & 2  \\
-                  \emph{Limnonectes magnus}           & 2  & 3  \\
-                \hline
-            \end{tabular}
-        \end{table}
-    \column{.5\textwidth}
-        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
-        \includegraphics<2>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        % \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-bohol_mindanao.png}
-        % \includegraphics<4>[width=\textwidth]{../images/maps/Philippines-leyte_mindanao.png}
-        % \includegraphics<5>[width=\textwidth]{../images/maps/Philippines-mindanao_samar.png}
-        % \includegraphics<6>[width=\textwidth]{../images/maps/Philippines-leyte_samar.png}
-\end{columns}
-\end{frame}
-}
-
 \begin{frame}
@@
         \vspace{-1cm}
+        \begin{onlyenv}<1>
         \begin{minipage}[t][1.0\textheight][c]{\linewidth}
         \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
+        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/rafe.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/cam.jpg}}
+        \centerline{
+        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/charles.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/jake.jpg}}
+        \end{minipage}
+        \end{onlyenv}
+
+        \begin{onlyenv}<2->
+        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
+        \centerline{
+        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
+        \hspace{0.3mm}
+        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
         % \vspace{0.5mm}
         \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
+        \includegraphics<2->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
+        \hspace{0.3mm}
+        \includegraphics<2->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
         % \vspace{0.5mm}
         \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
+        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
+        \hspace{0.3mm}
+        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
         % \vspace{0.5mm}
         \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
+        \includegraphics<2->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
+        \hspace{0.3mm}
+        \includegraphics<2->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
         % \vspace{0.5mm}
         \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
+        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
+        \hspace{0.3mm}
+        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
         % \vspace{0.5mm}
         \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
+        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
+        \hspace{0.3mm}
+        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
+        \end{minipage}
+        \end{onlyenv}
 
@@
         \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
-        \includegraphics<2>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        \begin{onlyenv}<3->
+        \includegraphics<1-2>[width=\textwidth]{../images/maps/Philippines.png}
+        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
+        \begin{onlyenv}<4->
             \begin{itemize}
@@
             beta distributions
-        \item<1-> Dirichlet process prior (DPP) over all possible divergence
+        \item<2-> Dirichlet process prior (DPP) over all possible divergence
             models
-        \item<1-> Develop multi-processing interface to improve performance for
+        \item<3-> Develop multi-processing interface to improve performance for
             accommodating genomic datasets
@@
             \end{itemize}
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Everything is on GitHub\ldots}
-    Software:\\
-    \begin{itemize}
-        \item \texttt{dpp-msbayes}:
-            \href{https://github.com/joaks1/dpp-msbayes}{\tt
-            https://github.com/joaks1/dpp-msbayes}
-
-        \item \texttt{PyMsBayes}:
-            \href{https://github.com/joaks1/PyMsBayes}{\tt
-            https://github.com/joaks1/PyMsBayes}
-
-        \item ABACUS: Approximate BAyesian C UtilitieS.
-            \href{https://github.com/joaks1/abacus}{\tt
-            https://github.com/joaks1/abacus}
-    \end{itemize}
-
-    \medskip
-    Open-Science Notebook:\\
-    \begin{itemize}
-        \item \texttt{msbayes-experiments}:
-            \href{https://github.com/joaks1/msbayes-experiments}{\tt
-            https://github.com/joaks1/msbayes-experiments}
     \end{itemize}
@@
         \end{myitemize}
+        \centerline{
+        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/misc/lee-and-me.jpg}
+        \hspace{2mm}
+        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/people/rafe.jpg}}
         \column{.5\textwidth}
@@
     \frametitle{More power}
+    \centerline{
+    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/mtholder.jpg}
+    \hspace{0.6mm}
+    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/jeet2.jpg}
+    \hspace{0.6mm}
+    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/vladimir.jpg}
+    \hspace{0.6mm}
+    \includegraphics<1->[height=2cm]{../images/revbayes.png}}
+
     \begin{itemize}[<+->]
@@
 %           evolution asexuality/live-bearing/color.
-\begin{frame}
-    \frametitle{Future directions---Empirical}
-    \begin{itemize}
-        \item<1-> Many exciting empirical systems for applying new methods!
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Gobi Desert}
+{
+\setbeamercolor{background canvas}{bg=black}
+\setbeamercolor{whitetext}{fg=white}
+\begin{frame}
+    \frametitle{Processes of diversification}
+    \begin{columns}[c]
+        \column{.5\textwidth}
+        {\usebeamercolor[fg]{whitetext}
+        \begin{itemize}
+            \item<1-> These comparative tools will allow evolutionary
+                biologists to leverage genomic data to infer shared
+                evolutionary history
+            \item<2-> This will better our understanding of how community scale
+                processes affect the generation and assembly of biodiversity
+            \item<3-> Many exciting empirical systems for applying new methods!
+        \end{itemize}
+        }
+        \column{.5\textwidth}
+        \begin{figure}
+            \begin{center}
+            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
+            \end{center}
+        \end{figure}
+    \end{columns}
+\end{frame}
+}
+
+\begin{frame}
+    \frametitle{Future directions: Gobi Desert}
     \vspace{-2mm}
@@
 \begin{frame}
-    \frametitle{Gobi Desert}
+    \frametitle{Future directions: Gobi Desert}
     \vspace{-2mm}
@@
 \begin{frame}[t]
-    \frametitle{West-Central Africa}
+    \frametitle{Future directions: West-Central Africa}
     \begin{center}
@@
     \end{center}
+
+    \only<1>{
+    \begin{center}
+        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/adam.png}
+        \hspace{0.3mm}
+        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/fujita.jpg}
+    \end{center}
+    }
 
@@
 \begin{frame}
+    \frametitle{Everything is on GitHub\ldots}
+    Software:\\
+    \begin{itemize}
+        \item \texttt{dpp-msbayes}:
+            \href{https://github.com/joaks1/dpp-msbayes}{\tt
+            https://github.com/joaks1/dpp-msbayes}
+
+        \item \texttt{PyMsBayes}:
+            \href{https://github.com/joaks1/PyMsBayes}{\tt
+            https://github.com/joaks1/PyMsBayes}
+
+        \item ABACUS: Approximate BAyesian C UtilitieS.
+            \href{https://github.com/joaks1/abacus}{\tt
+            https://github.com/joaks1/abacus}
+    \end{itemize}
+
+    \medskip
+    Open-Science Notebook:\\
+    \begin{itemize}
+        \item \texttt{msbayes-experiments}:
+            \href{https://github.com/joaks1/msbayes-experiments}{\tt
+            https://github.com/joaks1/msbayes-experiments}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
     \frametitle{Questions?}    
+    \begin{columns}[c]
+        \column{.5\textwidth}
         \begin{center}
@@
         \end{center}
+        \column{.5\textwidth}
+            \begin{figure}
+                \begin{center}
+                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
+                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
+                \end{center}
+            \end{figure}
+    \end{columns}
 \end{frame}

### 2014-11-03 slides/bin/marginal_plot.py
Adding marginal likelihood plot for slides.

@@ -227,6 +227,15 @@ def main_cli():
             include_constrained_density = False)
     fig.savefig('../images/marginal-plot-3d-prior.png', dpi=300)
 
+    ax, fig = get_marginal_plot_3d(maximum = maximum,
+            mean = (0.15, 0.247),
+            variance = (0.039, 0.026),
+            covariance=0.0,
+            npoints = 100,
+            include_prior = False,
+            include_constrained_density = True)
+    fig.savefig('../images/marginal-plot-3d-constrained.png', dpi=300)
+
     ax, fig = get_marginal_plot_2d(maximum = maximum,
         likelihood_shape = 50.0,
         likelihood_scale = 0.002,

### 2014-11-03 slides/osu/slides.tex
More work on slides for osu.

@@
 \begin{frame}
-    \frametitle{Southeast Asia}    
+    \frametitle{Climate-driven diversification}    
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Southeast Asia}    
+    \frametitle{Climate-driven diversification}    
     \begin{columns}
@@
             \begin{minipage}[t][3.6cm][t]{1.1\textwidth}
-                \begin{onlyenv}<2-3>
-                \begin{itemize}
-                    \item<2-> Repeated coalescence and fragmentation of island complexes
-                    \item<3-> Prominent paradigm for explaining biodiversity
-                \end{itemize}
-                \end{onlyenv}
-
-                \begin{onlyenv}<4>
-                \begin{itemize}
-                    \item<4> \textbf{Did repeated fragmentation of islands
+                % \begin{onlyenv}<2-3>
+                % \begin{itemize}
+                %     \item<2-> Repeated coalescence and fragmentation of island complexes
+                %     \item<3-> Prominent paradigm for explaining biodiversity
+                % \end{itemize}
+                % \end{onlyenv}
+
+                \begin{onlyenv}<2>
+                \begin{itemize}
+                    \item \textbf{Did repeated fragmentation of islands
                         during inter-glacial rises in sea level promote
@@
 
-                \begin{onlyenv}<5->
-                \begin{itemize}
-                    \item<5-> Climate-driven model of diversification
-                    \item<6-> Testable prediction:
+                \begin{onlyenv}<3->
+                \begin{itemize}
+                    \item<3-> Climate-driven model of diversification
+                    \item<4-> Testable prediction:
                     \begin{itemize}
@@
 \begin{frame}
-    \frametitle<1-4>{Community scale processes}
+    \frametitle<1-4>{Climate-driven model: Predictions}
     \frametitle<5->{Divergence model choice}
@@
         % \includegraphics<1>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<2>[height=7.8cm]{../images/marginal-plot-3d.png}
-        \includegraphics<3>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<4>[height=7.8cm]{../images/marginal-plot-2d.pdf}
+        \includegraphics<2>[height=7.8cm]{../images/marginal-plot-3d-bare.png}
+        \includegraphics<3>[height=7.8cm]{../images/marginal-plot-3d-constrained.png}
+        \includegraphics<4>[height=7.8cm]{../images/marginal-plot-3d.png}
+        \includegraphics<5>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
+        \includegraphics<6>[height=7.8cm]{../images/marginal-plot-2d.pdf}
     \end{center}
@@
 \begin{frame}
+    \frametitle{Everything is on GitHub\ldots}
+    Software:\\
+    \begin{itemize}
+        \item \texttt{dpp-msbayes}:
+            \href{https://github.com/joaks1/dpp-msbayes}{\tt
+            https://github.com/joaks1/dpp-msbayes}
+
+        \item \texttt{PyMsBayes}:
+            \href{https://github.com/joaks1/PyMsBayes}{\tt
+            https://github.com/joaks1/PyMsBayes}
+
+        \item ABACUS: Approximate BAyesian C UtilitieS.
+            \href{https://github.com/joaks1/abacus}{\tt
+            https://github.com/joaks1/abacus}
+    \end{itemize}
+
+    \medskip
+    Open-Science Notebook:\\
+    \begin{itemize}
+        \item \texttt{msbayes-experiments}:
+            \href{https://github.com/joaks1/msbayes-experiments}{\tt
+            https://github.com/joaks1/msbayes-experiments}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
     \frametitle{Acknowledgments}
@@
     \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Everything is on GitHub\ldots}
-    Software:\\
-    \begin{itemize}
-        \item \texttt{dpp-msbayes}:
-            \href{https://github.com/joaks1/dpp-msbayes}{\tt
-            https://github.com/joaks1/dpp-msbayes}
-
-        \item \texttt{PyMsBayes}:
-            \href{https://github.com/joaks1/PyMsBayes}{\tt
-            https://github.com/joaks1/PyMsBayes}
-
-        \item ABACUS: Approximate BAyesian C UtilitieS.
-            \href{https://github.com/joaks1/abacus}{\tt
-            https://github.com/joaks1/abacus}
-    \end{itemize}
-
-    \medskip
-    Open-Science Notebook:\\
-    \begin{itemize}
-        \item \texttt{msbayes-experiments}:
-            \href{https://github.com/joaks1/msbayes-experiments}{\tt
-            https://github.com/joaks1/msbayes-experiments}
-    \end{itemize}
 \end{frame}

### 2015-01-29 slides/bin/partition-number-plots.py
Adding divergence model plots for slides.
Updating script to plot the number and prior probability of divergence
models for 10 taxa, and adding the new plots.

@@ -130,3 +130,10 @@ if __name__ == '__main__':
 
     num_plot.savefig(os.path.join(project_util.IMAGE_DIR, 'number-of-div-models-22-unordered.pdf'))
     prob_plot.savefig(os.path.join(project_util.IMAGE_DIR, 'prob-of-div-models-22-unordered.pdf'))
+
+
+    num_plot, prob_plot = create_plots(n = 10,
+            ordered = True)
+
+    num_plot.savefig(os.path.join(project_util.IMAGE_DIR, 'number-of-div-models-10.pdf'))
+    prob_plot.savefig(os.path.join(project_util.IMAGE_DIR, 'prob-of-div-models-10.pdf'))

### 2015-01-29 slides/osu/slides.tex
Updating osu slides.
Making paths to some of the photos full rather than relative.

@@
         \centerline{
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/rafe.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/cam.jpg}}
-        \centerline{
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/charles.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/jake.jpg}}
+        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/cam.jpg}}
+        \centerline{
+        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/charles.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/jake.jpg}}
         \end{minipage}
@@
         \centerline{
-        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/misc/lee-and-me.jpg}
+        \includegraphics<1->[height=2.5cm]{/home/jamie/Dropbox/field-photos/misc/lee-and-me.jpg}
         \hspace{2mm}
-        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/people/rafe.jpg}}
+        \includegraphics<1->[height=2.5cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}}
         \column{.5\textwidth}
@@
     \centerline{
-    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/mtholder.jpg}
+    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/mtholder.jpg}
     \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/jeet2.jpg}
+    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
     \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/vladimir.jpg}
+    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/vladimir.jpg}
     \hspace{0.6mm}
@@
         \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
+        \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
     \end{columns}
@@
         \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
+        \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
     \end{columns}
@@
     \begin{center}
-        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
-        \hspace{0.3mm}
-        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure4_herps.jpeg}
+        \includegraphics<1->[height={3cm}]{/home/jamie/Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
+        \hspace{0.3mm}
+        \includegraphics<1->[height={3cm}]{/home/jamie/Dropbox/field-photos/africa/Figure4_herps.jpeg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/adam.png}
-        \hspace{0.3mm}
-        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/fujita.jpg}
+        \includegraphics[height={3.5cm}]{/home/jamie/Dropbox/field-photos/people/adam.png}
+        \hspace{0.3mm}
+        \includegraphics[height={3.5cm}]{/home/jamie/Dropbox/field-photos/people/fujita.jpg}
     \end{center}
@@
         \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/malaysia/malaysia-burungs.jpg}
+        \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/malaysia/malaysia-burungs.jpg}
     \end{columns}

### 2015-01-30 slides/bin/marginal_plot.py
Adding plots for slides.
Updating plotting script for marginal likelihoods, and adding plots
of 2D marginal plots of normal likelihoods to match the 3D plots.

@@ -7,7 +7,7 @@ import matplotlib
 import numpy
 from matplotlib import pyplot as plt
 from matplotlib import cm
-from scipy.stats import gamma
+from scipy.stats import gamma, norm
 
 def get_bivariate_normal_and_uniform_densities(maximum = 1.0,
         mean = (0.15, 0.247),
@@ -79,6 +79,82 @@ def get_marginal_plot_2d(maximum = 1.0,
         prior_label_x = 0.5):
     x = numpy.linspace(0.0000001, maximum, npoints)
     likelihood = gamma(likelihood_shape, scale = likelihood_scale)
+    y = [likelihood.pdf(i) for i in x]
+    fig = plt.figure()
+    ax = fig.add_subplot(1,1,1)
+    likelihood_line = ax.plot(x, y)
+    plt.setp(likelihood_line,
+            color = '0.3',
+            linestyle = '-',
+            linewidth = linewidth,
+            marker = '',
+            zorder = 200)
+    max_idx = y.index(max(y)) + int(round(0.01 * npoints))
+    label_target = (x[max_idx], y[max_idx])
+    label_position = (label_target[0] + (0.08 * maximum), label_target[1])
+    plt.annotate(r'$p(X \mid \, \theta)$',
+            xy = label_target,
+            arrowprops = dict(arrowstyle = '->'),
+            xytext = label_position,
+            size = 18.0)
+    prior_x = prior_label_x * maximum
+    u_density = 1.0 / maximum
+    ymax = max(ax.get_ylim())
+    prior_label_position = (prior_x, u_density + (0.1 * ymax))
+    if include_uniform_prior:
+        u = [u_density for i in range(len(x))]
+        u_line = ax.plot(x, u)
+        plt.setp(u_line,
+                color = 'r',
+                linestyle = '-',
+                linewidth = linewidth,
+                marker = '',
+                zorder = 0)
+        u_label_target = (prior_x + (0.04 * maximum), u_density)
+        plt.annotate(r'$p(\theta)$',
+                xy = u_label_target,
+                arrowprops = dict(arrowstyle = '->'),
+                xytext = prior_label_position,
+                size = 18.0)
+                # verticalalignment = 'bottom',
+                # horizontalalignment = 'center')
+    if include_gamma_prior:
+        g_prior = gamma(prior_shape, scale = prior_scale)
+        g = [g_prior.pdf(i) for i in x]
[…]

### 2015-01-30 slides/bin/marginal_plot.py
Updating marginal likelihood plots for slides.
Changing x-axis labels of 2D plots to match 3D plots.

@@ -152,6 +152,7 @@ def get_marginal_plot_2d_normal(maximum = 1.0,
         include_uniform_prior = True,
         include_gamma_prior = True,
         linewidth = 2.0,
+        x_axis_label = r'$\theta$',
         prior_label_x = 0.5):
     x = numpy.linspace(0.0000001, maximum, npoints)
     likelihood = norm(mean, variance)
@@ -213,7 +214,7 @@ def get_marginal_plot_2d_normal(maximum = 1.0,
                 size = 18.0)
                 # verticalalignment = 'center',
                 # horizontalalignment = 'center')
-    ax.set_xlabel(r'$\theta$', size=18.0)
+    ax.set_xlabel(x_axis_label, size=18.0)
     ax.set_ylabel(r'Density', size=18.0)
     rect = [0, 0, 1, 1]
     fig.tight_layout(pad = 0.25, rect = rect)
@@ -331,6 +332,7 @@ def main_cli():
         npoints = 500,
         include_uniform_prior = True,
         include_gamma_prior = True,
+        x_axis_label = r'$T_1$',
         linewidth = 2.0)
     fig.savefig('../images/normal-marginal-plot-2d.pdf')
 
@@ -364,6 +366,7 @@ def main_cli():
         npoints = 500,
         include_uniform_prior = True,
         include_gamma_prior = False,
+        x_axis_label = r'$T_1$',
         linewidth = 2.0)
     fig.savefig('../images/normal-marginal-plot-2d-uniform-prior.pdf')
 

### 2015-02-18 slides/bordeaux/slides.tex
Adding Bordeaux slides.

--- 2015-01-29 slides/osu/slides.tex
@@
 
-\title[Improving inference of evolutionary history]{Improving inference of evolutionary history}
-\subtitle{Developing comparative methods for genomic data}
+% \title[Estimating shared history]{An Improved Approximate-Bayesian Method for Estimating Shared
+% Evolutionary History}
+
+\title[Inferring community history]{Inferring Community-scale Diversification
+    from Genomic Data}
 
@@
 % \date{\today}
-\date{November 3, 2014}
+\date{February 18, 2015}
 
 \begin{document}
-% \maketitle
-\begin{frame}
-    \begin{columns}[c]
-        \column{.5\textwidth}
 \maketitle
-        \column{.5\textwidth}
-            \begin{figure}
-                \begin{center}
-                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
-                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
-                \end{center}
-            \end{figure}
-    \end{columns}
-\end{frame}
 
-% I. What I do
-% II. What I've been focusing on recently (started in PhD and continuing in postdoc)
-%   a. Dynamic planet---need to account for large-scale processes
-%       i. E.g., SE Asia---dynamic landscape
 {
@@
         \begin{itemize}
-            \item<2-> Large-scale geological and climatic processes are
-                important in biodiversification and community assembly
+            \item<2-> Community-scale processes are important in
+                biodiversification and community assembly
             \item<3-> Accounting for such processes will better our
-                understanding of biodiversity
+                understanding of evolutionary history and
+                community structure
             \item<4-> We need methods for inferring evolutionary patterns
@@
 \begin{frame}
-    \frametitle{Climate-driven diversification}    
+    % \frametitle{Climate-driven diversification}
+
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Climate-driven diversification}    
+    % \frametitle{Climate-driven diversification}    
     \begin{columns}
         \column{0.6\textwidth}
-        \ \\
 
-        \column{0.45\textwidth}
+        \vspace{6.5cm}
 
-        \vspace{-2.4cm}
-
-        \begin{uncoverenv}<2->
-        \colorbox{white}{
-            \begin{minipage}[t][3.6cm][t]{1.1\textwidth}
-                % \begin{onlyenv}<2-3>
-                % \begin{itemize}
-                %     \item<2-> Repeated coalescence and fragmentation of island complexes
-                %     \item<3-> Prominent paradigm for explaining biodiversity
-                % \end{itemize}
-                % \end{onlyenv}
-
-                \begin{onlyenv}<2>
-                \begin{itemize}
-                    \item \textbf{Did repeated fragmentation of islands
-                        during inter-glacial rises in sea level promote
-                        diversification?}
-                \end{itemize}
-                \end{onlyenv}
-
-                \begin{onlyenv}<3->
-                \begin{itemize}
-                    \item<3-> Climate-driven model of diversification
-                    \item<4-> Testable prediction:
-                    \begin{itemize}
-                        \item Temporally clustered divergences among taxa
-                            co-distributed across fragmented islands
-                    \end{itemize}
-                \end{itemize}
-                \end{onlyenv}
-            \end{minipage}
-        }
-    \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-% \begin{frame}
-%     \frametitle{Philippine Archipelago}
-%     \begin{center}
-%         \includegraphics<1>[width=.5\textwidth]{../images/maps/Philippines-present.png}
-%         \includegraphics<2>[width=.5\textwidth]{../images/maps/Philippines.png}
-%     \end{center}
-% \end{frame}
-
-% \begin{frame}
-%     \frametitle{Climate-driven diversification model}
-%     \begin{columns}[c]
-%         \column{.5\textwidth}
-%             \begin{myitemize}
-%                 \item<1-> Repeated coalescence and fragmentation of island complexes
-%                 \item<2-> Prominent paradigm for explaining Philippine biodiversity
-%                 \item<3-> Proposed as model of diversification
-%             \end{myitemize}
-%         \column{.5\textwidth}
-%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
-%     \end{columns}
-% \end{frame}
-
-% \begin{frame}
-%     \frametitle{Testing climate-driven diversification}
-%     \begin{columns}[c]
-%         \column{.5\textwidth}
-%     \textbf{Did repeated fragmentation of islands during inter-glacial rises in sea level
-%     promote diversification?}\\
-%     \bigskip
-%     \uncover<2->{
-%     Model has testable prediction:\\
-%     \begin{itemize}
-%         \item Temporally clustered divergences among taxa
-%             co-distributed across fragmented islands
-%     \end{itemize}
-%     }
-%         \column{.5\textwidth}
-%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
-%     \end{columns}
-% \end{frame}
-
-%   b. Predictions of climate-driven diversification
-\begin{frame}
-    \frametitle<1-4>{Climate-driven model: Predictions}
-    \frametitle<5->{Divergence model choice}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{onlyenv}<1-4>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<5>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{\divTime{1}\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 111
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (\divTimeMap{1},\divTimeMap{2},\divTimeMap{3})
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<6>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 111
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (130,130,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<7>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 211
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (397,130,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<8>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 121
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (130,397,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<9>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 112
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<10-12>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397,260\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 312
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (260,130,397)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        % \begin{onlyenv}<11>
-        %     \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-        %         \begin{displaybox}[0.95\linewidth]
-        %             \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-        %                 \[
-        %                     \divTimeVector = \{\divTime{1},\ldots,\divTime{\divTimeNum}\}
-        %                 \]\vspace{0mm}
-        %                 \[
-        %                     model = \divModel{i}
-        %                 \]\vspace{0mm}
-        %                 \[
-        %                     \divTimeMapVector = (\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})
-        %                 \]\vspace{0mm}
-        %             \end{minipage}
-        %         \end{displaybox}
-        %     \end{minipage}
-        % \end{onlyenv}
-        \begin{onlyenv}<13>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{mydescription}
-                    \item[\alignmentVector] Sequence alignments
-                    \item[\divTimeMapVector] Divergence times
-                    \item[\divModel{}] Divergence model
-                    \item[\allParameters{}] Gene trees
-                    \item[\ ] Substitution parameters
-                    \item[\ ] Demographic parameters
-                    % \item[\allParameters] $(\divTimeMapVector, \geneTreeVector,
-                    %     \hkyModelVector, \demographicParamVector)$
-                \end{mydescription}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{uncoverenv}<11->
-            \begin{minipage}[c][0.18\textheight][c]{\linewidth}
-            \begin{flushleft}
-                \small We want to infer \textcolor{blue}{\divModel{}} and
-                \textcolor{blue}{\divTimeMapVector} given DNA sequence
-                alignments
-                \textcolor{blue}{\alignmentVector}
-            \end{flushleft}
-            \end{minipage}
-        \end{uncoverenv}
-        \column{.6\textwidth}
-        \begin{minipage}[t][0.8\textheight][c]{\linewidth}
-        \includegraphics<1>[height=6cm]{../images/dmc-cartoon-islands.pdf}
-        \includegraphics<2>[height=6cm]{../images/dmc-cartoon-taxa.pdf}
-        \includegraphics<3>[height=6cm]{../images/dmc-cartoon-sea-level.pdf}
-        \includegraphics<4>[height=6cm]{../images/dmc-cartoon-shared-no-labels.pdf}
-        \includegraphics<5-6>[height=6cm]{../images/dmc-cartoon-shared.pdf}
-        \includegraphics<7>[height=6cm]{../images/dmc-cartoon-2-1.pdf}
-        \includegraphics<8>[height=6cm]{../images/dmc-cartoon-2-2.pdf}
-        \includegraphics<9>[height=6cm]{../images/dmc-cartoon-2-3.pdf}
-        \includegraphics<10-11>[height=6cm]{../images/dmc-cartoon-general.pdf}
-        \includegraphics<12->[height=6cm]{../images/dmc-cartoon-pop-general.pdf}
-        \end{minipage}
-    \end{columns}
-\end{frame}
-
-%   c. Testing for predicted patterns via model choice.
-%       i. Bayesian model choice
-
-\begin{frame}[t]
-    \frametitle{Bayesian model choice}
-    \begin{block}{\it Full model:}
-        \begin{minipage}[c][3.8cm][c]{\linewidth}
-            \begin{uncoverenv}<1->
-            \[
-                p(
-                  \divTimeMapVector,
-                  \allParameters{}
-                  \given \alignmentVector, \divModel{i})
-                  =
-                \frac{
-                    p(\alignmentVector \given
-                      \divTimeMapVector,
-                      \allParameters{},
-                      \divModel{i}
-                      )
-                      p(
-                        \divTimeMapVector,
-                        \allParameters{}
-                        \given \divModel{i}
-                        )
-                    }{p(\alignmentVector \given \divModel{i})}
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-            \begin{uncoverenv}<2->
-            \[
-                p(\alignmentVector \given \divModel{i}) =
-                \int_{\divTimeMapVector} \int_{\allParameters{}}
-                p(\alignmentVector \given \divTimeMapVector, \allParameters{}, \divModel{i})
-                p(\divTimeMapVector, \allParameters{} \given \divModel{i})
-                d\divTimeMapVector d\allParameters{}
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-            \begin{uncoverenv}<3->
-            \[
-                p(\divModel{i} \given \alignmentVector) =
-                \frac{
-                    p(\alignmentVector \given \divModel{i})
-                    p(\divModel{i})
-                }{
-                    \sum_{i} p(\alignmentVector \given \divModel{i})
-                    p(\divModel{i})
-                }
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-        \end{minipage}
-    \end{block}
-    \vspace{1mm}
-    \begin{uncoverenv}<4>
-        \begin{center}
-            \msb: Approximate Bayesian computation (ABC)
-
-            % \vspace{-5mm}
-            % \[ \alignmentVector \, \to \, \ssVectorObs \, \to \, \ssSpace\]
-        \end{center}
-    \end{uncoverenv}
-    \barefootnote{\tiny\shortfullcite{Huang2011}. \hspace{2mm}\shortfullcite{Oaks2012}.}
-\end{frame}
-
-%   d. Method exists---applied it to comparative dataset from Philippines
-%       i. Surprising results!
-%       ii. Rather than publish, did some detective work
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
         \ \\
@@
 
+        \begin{uncoverenv}<2->
         \colorbox{white}{
@@
         }
+        \end{uncoverenv}
     \end{columns}
@@
         \centerline{
-        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}
+        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}
         \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/cam.jpg}}
+        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rob.jpg}}
         \centerline{
-        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/charles.jpg}
+        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/charles.jpg}
         \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/jake.jpg}}
+        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/cam.jpg}}
+        \centerline{
+        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jake.jpg}}
+        \centerline{
+        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/allie.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/Luke.jpg}
+        \hspace{0.3mm}
+        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/anthony.png}}
         \end{minipage}
@@
         \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        \begin{onlyenv}<4->
-            \begin{itemize}
-                \item Strong support for simultaneous divergence of all 22 taxon pairs
+        % \begin{onlyenv}<4->
+        %     \begin{itemize}
+        %         \item Strong support for simultaneous divergence of all 22 taxon pairs
 
-                \item Posterior probability $>$ 0.96 \\
+        %         \item Posterior probability $>$ 0.96 \\
 
-                \item $\sim$100,000--250,000 years ago
-            \end{itemize}
-        \end{onlyenv}
+        %         \item $\sim$100,000--250,000 years ago
+        %     \end{itemize}
+        % \end{onlyenv}
         \end{minipage}
@@
 
-%   e. Evaluated method via simulation
-%       i. biased in alarming direction
+\begin{frame}[t]
+    \frametitle<1-4>{Climate-driven diversification: Predictions}
+    \frametitle<5->{Divergence-model choice}
 
-\section{Simulation-based assessment of method}
-
-\begin{frame}
-    \frametitle{Simulation-based power analyses}
-    % What is ``simultaneous''?
-    \begin{myitemize}
-        \item<1-> What if divergences were random?
-        \item<2-> Simulate datasets in which all 22 divergence times are random
-        \uncover<3->{
-            \smallskip
-            \begin{myitemize}
-                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
-                \smallskip
-            \end{myitemize}
-        \item<3-> $MGA$ = Millions of Generations Ago
+    \centering{
+    \includegraphics<1>[height=7.5cm]{../images/dmc-cartoon-islands.pdf}
+    \includegraphics<2>[height=7.5cm]{../images/dmc-cartoon-taxa.pdf}
+    \includegraphics<3>[height=7.5cm]{../images/dmc-cartoon-sea-level.pdf}
+    \includegraphics<4>[height=7.5cm]{../images/dmc-cartoon-shared-no-labels.pdf}
+    \includegraphics<5>[height=7.5cm]{../images/dmc-cartoon-shared.pdf}
+    \includegraphics<6>[height=7.5cm]{../images/dmc-cartoon-2-1.pdf}
+    \includegraphics<7>[height=7.5cm]{../images/dmc-cartoon-2-2.pdf}
+    \includegraphics<8>[height=7.5cm]{../images/dmc-cartoon-2-3.pdf}
+    \includegraphics<9>[height=7.5cm]{../images/dmc-cartoon-general.pdf}
     }
-        \item<4-> Simulate 1000 datasets for each \divTime{} distribution
-        \item<4-> Analyze all 4000 datasets as we did the empirical data
-    \end{myitemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Simulation-based power analyses: Results}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]{../images/old_old_power_psi_mode.pdf}}
-
-    \begin{uncoverenv}<2->
-        \begin{itemize}[<+->]
-            \item \msb will often infer clustered divergences when divergences
-                are random over millions of generations.
-
-            \item The empirical results are likely spurious
-        \end{itemize}
-    \end{uncoverenv}
-
-    \begin{uncoverenv}<3->
-        \begin{block}{Objective:}
-            Use principles of probability to extend \msb framework for improved
-            estimation of shared evolutionary history
-        \end{block}
-    \end{uncoverenv}
-    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
-\end{frame}
-
-%   f. Rather than abandon approach---look for theoretical reasons for behavior
-%       i. Strongly weighted posterior (marginal likelihood)
-%       ii. Combinatorics
-%   g. Developed new method---dpp-msbayes
-%       i. Used DPP---common in Bayeisn nonparametrics
-%           1. Flexible and mathematically convenient way of assigning prior
-%           probability to all possible divergence models (partitions) directly
-%           (avoid combinatoric issue)
-%       ii. More flexible priors on parameters of model
-%       iii. Implemented method in much more efficient multi-processing
-%       framework.
-%   e. Results---improved!
-%   f. Empirical results...
-%       i. Reducing bias and increasing power provides more "honest" estimate
-%       of uncertainty
-%       ii. A lot of uncertainty (Good! and Bad!)
-%       iii. What next? More data and more power!
-
-\begin{frame}[label=improvements]
-    \frametitle{An improved method}
-    Potential improvements:
-    \begin{enumerate}
-        \item Alternative priors on parameters that increase marginal
-            likelihoods of rich models
-        \item Alternative approach to modeling the temporal distribution of
-            divergences
-    \end{enumerate}
-    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
 \end{frame}
@@
 \begin{frame}[t]
+    \frametitle{Divergence-model choice}
+
+    \vspace{-7mm}
+
+    \begin{minipage}[t][0.1\textheight][c]{1.1\linewidth}
+        \begin{adjustwidth}{-0.5em}{}
+            \begin{tabular}{ p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} }
+                $m_1$ & $m_2$ & $m_3$ & $m_4$ & $m_5$ \\
+            \end{tabular}
+        \end{adjustwidth}
+    \end{minipage}
+
     \vspace{-2mm}
-    \begin{displaybox}[5.5cm]
-        \small
-        \[
-            p(X) = \int_{\theta} p(X \given \theta) p(\theta) \mathrm{d}\theta
-        \]%\vspace{0mm}
-    \end{displaybox}
 
-    \vspace{-1mm}
+    \begin{minipage}[t][0.35\textheight][c]{1.1\linewidth}
+        \begin{adjustwidth}{-2.3em}{}
+        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-shared.pdf}
+        \hspace{0.2mm}
+        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-1.pdf}
+        \hspace{0.2mm}
+        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-2.pdf}
+        \hspace{0.2mm}
+        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-3.pdf}
+        \hspace{0.2mm}
+        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-general.pdf}
+        \end{adjustwidth}
+    \end{minipage}
+
+    \vspace{3.5mm}
+
+    \begin{minipage}[c][0.35\textheight][c]{\linewidth}
+        \begin{uncoverenv}<2->
             \begin{center}
-        % \includegraphics<1>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<2>[height=7.8cm]{../images/marginal-plot-3d-bare.png}
-        \includegraphics<3>[height=7.8cm]{../images/marginal-plot-3d-constrained.png}
-        \includegraphics<4>[height=7.8cm]{../images/marginal-plot-3d.png}
-        \includegraphics<5>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<6>[height=7.8cm]{../images/marginal-plot-2d.pdf}
+            We want to infer \textcolor{blue}{\divModel{}} and
+            \textcolor{blue}{\divTimeMapVector} given DNA sequence
+            alignments
+            \textcolor{blue}{\alignmentVector}
         \end{center}
-\end{frame}
-
-\againframe{improvements}
-
-\begin{frame}[t]
-    \frametitle{Prior on divergence models}
-    \begin{itemize}
-        \item \msb uses a discrete uniform prior on the \emph{number} of
-            divergence events
-    \end{itemize}
-
-    \begin{uncoverenv}<2->
-    \begin{columns}
-        \column{0.5\textwidth}
-            \centerline{
-                \includegraphics<2->[width=1.0\textwidth]{../images/number-of-div-models-22-unordered.pdf}}
-        \column{0.5\textwidth}
-            \centerline{
-                \includegraphics<3->[width=1.0\textwidth]{../images/prob-of-div-models-22-unordered.pdf}}
-    \end{columns}
         \end{uncoverenv}
 
-    \begin{uncoverenv}<4->
-    \begin{block}{\it Potential solution:}
-        Place flexible prior directly on the sample space of divergence models
-    \end{block}
+        \vspace{4mm}
+
+        \begin{uncoverenv}<3->
+            \begin{displaybox}[0.85\linewidth]
+                \begin{minipage}[c][0.12\textheight][c]{\linewidth}
+
+                \[
+                    p(\divTimeMapVector,
+                      \divModel{i},
+                      \allParameters{}
+                      \given \alignmentVector)
+                      =
+                    \frac{
+                        p(\alignmentVector \given
+                          \divTimeMapVector,
+                          \allParameters{},
+                          \divModel{i})
+                        p(\divTimeMapVector,
+                          \allParameters{}
+                          \given \divModel{i})
+                        p(\divModel{i})
+                        }{p(\alignmentVector)}
+                \]
+                \vspace{-1mm}
+                \end{minipage}
+            \end{displaybox}
         \end{uncoverenv}
+    \end{minipage}
 \end{frame}
@@
     \frametitle{New method: \dppmsbayes}
-    \begin{itemize}[<+->]
+    \begin{itemize}
         % \item<1-> Reparameterized the model implemented in \msb
-        \item<1-> Replaced uniform priors on continuous parameters with gamma and
-            beta distributions
-        \item<2-> Dirichlet process prior (DPP) over all possible divergence
+        \item<1-> Dirichlet-process prior (DPP) over all possible divergence
             models
-        \item<3-> Develop multi-processing interface to improve performance for
+        \item<2-> Flexible priors on continuous parameters to avoid strongly
+            weighted posteriors
+        \item<3-> Multi-processing interface to improve performance for
             accommodating genomic datasets
@@
     \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Simulation-based assessment}
-        \uncover<1->{
-        Validation:\\
-
-        \smallskip
-        \begin{itemize}
-            \item Simulate 50,000 datasets and analyze each under the same
-                model
-        \end{itemize}
-        }
-
-        \uncover<2->{
-        \bigskip
-        Robustness:\\
-        \begin{itemize}
-            \item Simulate datasets that violate model assumptions and analyze
-                each of them
-        \end{itemize}
-        }
-
-        \uncover<3->{
-        \bigskip
-        Done for \msb and \dppmsbayes
-        }
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Validation results}
-    \begin{columns}
-        \column{0.5\textwidth}
-            \includegraphics<1->[width=1.0\textwidth]{../images/validation-model-choice-old.pdf}
-        \column{0.5\textwidth}
-            \includegraphics<2->[width=1.0\textwidth]{../images/validation-model-choice-dpp.pdf}
-    \end{columns}
     \barefootnote{\shortfullcite{Oaks2014dpp}}
@@
 \begin{frame}
-    \frametitle{\dppmsbayes: Robustness results}
-    \begin{columns}
-        \column{0.5\textwidth}
-            \centering{
-            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-old-violations.pdf}
+    \frametitle{\dppmsbayes: Performance}
+    \begin{itemize}
+        \item<1-> Extensive computer simulations to compare performance to previous
+            state-of-the-art methods (\msb).
+        \item<2-> New method for estimating shared evolutionary history shows
+            improved:
+            \begin{enumerate}
+                \item<2-> Improved model-choice accuracy 
+                \item<2-> More robustness to model violations
+                \item<2-> Greater power to detect variation in divergence times
+                \item<2-> Faster!
+            \end{enumerate}
+    \end{itemize}
+
+    \uncover<3->{
+    \textbf{\dppmsbayes allows biologists to leverage comparative genomic data
+        to assess the affects of community-scale processes on biodiversity}
     }
-        \column{0.5\textwidth}
-            \centering{
-            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-dpp-violations.pdf}
-            }
-    \end{columns}
-
-    \begin{uncoverenv}<2->
-    \begin{itemize}
-        \item Improved model-choice accuracy and robustness
-    \end{itemize}
-    \end{uncoverenv}
-    \vspace{-0.5cm}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Simulation-based power analyses}
-    Does the new method have power to detect variation in divergence times?\\
-
-    \bigskip
-    \begin{myitemize}
-        \item<2-> Simulate datasets in which all 22 divergence times are random
-            \smallskip
-            \begin{myitemize}
-                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
-                \smallskip
-            \end{myitemize}
-        \item<2-> $MGA$ = Millions of Generations Ago
-
-        \item<3-> Simulate 1000 datasets for each \divTime{} distribution
-        \item<3-> Analyze all 4000 datasets with \msb and \dppmsbayes
-    \end{myitemize}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{\dppmsbayes: Power results}
-        \begin{onlyenv}<1-2>
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<1-2>{../images/power-psi-old-annotate-no-pp.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<2>{../images/power-psi-dpp-annotate-no-pp.pdf}}
-        \end{onlyenv}
-
-        \begin{onlyenv}<3->
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-old-annotate-pp.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-dpp-annotate-pp.pdf}}
-        \end{onlyenv}
-
-    \begin{uncoverenv}<4->
-    \begin{itemize}
-        \item Improved power to detect temporal variation across divergences
-    \end{itemize}
-    \end{uncoverenv}
     \barefootnote{\shortfullcite{Oaks2014dpp}}
@@
 \begin{frame}
-    \frametitle{Climate-driven diversification}    
+    % \frametitle{Climate-driven diversification}    
     \begin{columns}
         \column{0.6\textwidth}
+
+        \vspace{6.5cm}
+
         \ \\
@@
 
+        \begin{uncoverenv}<1->
         \colorbox{white}{
@@
         \vspace{-1cm}
+        \begin{onlyenv}<1->
         \begin{minipage}[t][1.0\textheight][c]{\linewidth}
@@
         \end{minipage}
+        \end{onlyenv}
 
@@
         \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
+        \includegraphics<1->[width=\textwidth]{../images/maps/Philippines.png}
         \end{minipage}
@@
 \begin{frame}
-    \frametitle{Empirical results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp.pdf}}
-
-    \begin{uncoverenv}<2->
-    \begin{itemize}
-        \item More ``honest'' estimate of posterior uncertainty
-    \end{itemize}
-    \end{uncoverenv}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Empirical results}
+    \frametitle{Results}
     \centerline{
@@
     \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
+    \barefootnote{\shortfullcite{Oaks2014dpp}}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Summary}
+    \frametitle{Current work: More data}
     \begin{itemize}
-        \item<1-> New method for estimating shared evolutionary history shows
-            improved:
-            \begin{enumerate}
-                \item<2-> Estimation of posterior uncertainty
-                \item<2-> Model-choice accuracy
-                \item<2-> Power to detect temporal variation across divergences
-                \item<2-> Robustness to model violations
-                \item<2-> Speed!
-            \end{enumerate}
-
-        \smallskip
-        \item<3-> Climate-driven diversification model?
-            \begin{itemize}
-                \item<4-> Results consistent with prediction of clustered
-                    divergences
-                \item<4-> Results suggest multiple co-divergences
-                \item<4-> However, there is a lot of uncertainty!
+        \item<1-> Collecting genomic data from taxa co-distributed across Southeast
+            Asia
+        \item<2-> Preliminary results for 1000 loci from 5 pairs of \emph{Gekko
+                mindorensis} populations
     \end{itemize}
 
-        \smallskip
-        \item<5-> What to do?
-            \begin{itemize}
-                \item<6-> More data!
-                \item<6-> More power! 
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-%   g. Collected thousands of loci from across the genome of almost 300
-%   individuals of two genera of geckos from across the Philippine Islands
-%       i. Prelim results with dpp-msbayes (can accommodate NGS data!)
-
-\begin{frame}
-    \frametitle{More data}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{myitemize}
-            \item Collecting genomic data from two gekkonid radiations
-                co-distributed Southeast Asia
-            \begin{myitemize}
-                \item \emph{Cyrtodactylus}---12+ species
-                \item \emph{Gekko}---13+ species
-            \end{myitemize}
-            \item Whole genome assemblies for each genus
-            \item RAD-seq libraries constructed for $\approx150$ individuals
-                from each genus
-        \end{myitemize}
-        \centerline{
-        \includegraphics<1->[height=2.5cm]{/home/jamie/Dropbox/field-photos/misc/lee-and-me.jpg}
-        \hspace{2mm}
-        \includegraphics<1->[height=2.5cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}}
-        \column{.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/photos/cyrt-agusanensis.jpg} \quad
-        \includegraphics[width=\textwidth]{../images/photos/gekko-mindorensis.jpg}
-    \end{columns}
+    \begin{center}
+        \includegraphics<2->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
+    \end{center}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{More data---preliminary results}
-    Results for 1000 loci from 5 pairs of \emph{Gekko mindorensis} populations\\
-
-    \begin{center}
-        \includegraphics<1->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
-    \end{center}
-\end{frame}
-
-%   h. Currently working on full-likelihood implementation of the model.
-%       i. Working with Jeet---has full-likelihood up and running---inefficient
-%       ii. Take advantage of recent developments to analytically integrate
-%       gene trees---might even be faster than simulations?!
-%       iii. Uses ALL information in genetic data
-
-\begin{frame}
-    \frametitle{More power}
+    \frametitle{Current work: More power}
     \centerline{
@@
     \begin{itemize}[<+->]
-        \item Full-likelihood Bayesian implementation \footnote{\tiny\shortfullcite{JeetDiss}}
+        \item Full-likelihood Bayesian implementation
             \begin{itemize}
-                \item Uses all the information in the sequence data
-                \item Applicable to much deeper timescales
-                \item \ldots Currently inefficient
+                \item Uses all the information in the data
+                \item Applicable to deeper timescales
             \end{itemize}
-        \item Working to incorporate new developments on analytical integration
-            of gene trees \footnote{\tiny\shortfullcite{Bryant2012}}
+        \item Analytically integrate over gene trees
+            \footnote{\tiny\shortfullcite{Bryant2012}}
             \begin{itemize}
@@
 
-\section{Future directions}
+\section{Conclusions}
 
-% III. Future work
-%   a. Theory/computation
-%       i. Extend full-likelihood approach to phylogenetic framework
-%           1. Can infer relative mutation rates/pop sizes from data (fewer
-%           assumptions)
-%           2. Much better representation of diversification than pairwise
-%           approach.
-%       ii. Parameter estimation approach (Lindley's paradox).
+\begin{frame}
+    \frametitle{Climate-driven diversification model?}
+    \begin{itemize}
+        \item<1-> Results consistent with prediction of clustered
+            divergences
+        \item<2-> Results suggest multiple co-divergences
+        \item<3-> However, there is a lot of uncertainty!
+    \end{itemize}
+    \barefootnote{\shortfullcite{Oaks2014dpp}}
+\end{frame}
+
+\section{Teaching}
+
+\begin{frame}
+    BIO 180 stuff
+\end{frame}
+
+\begin{frame}
+    Software carpentry \\
+
+    Workshops
+\end{frame}
+
+\begin{frame}
+    \frametitle{Questions?}    
+    \begin{center}
+        {
+        \Large
+        \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
+        }
+    \end{center}
+\end{frame}
 
@@
 
-%   b. Empirical
-%       i. Gobi (NGS --> NSF)
-%       ii. West-Central Africa
-%           1. unfunded collaborator
-%       iii. Compare diversification of mainland SE Asia to Oceanic Islands
-%           1. Planning pre-proposal
-%       iv. Long-term study of small island of sunda shelf
-%           1. Tiny islands with pops of Eutropis multifasciata and
-%           Lepidodactylus lugubris
-%           2. Inundated during last interglacial ~120,000 years ago, spent the
-%           better part of next 100k years as part of exposed sunda shelf, then
-%           fragmented during current interglacial
-%           3. Both species vary in reproductive mode (oviparity/ovoviviparity,
-%           parthenogenesis), skinks in color patterns
-%           4. Affects of recent, severe fragmentation on diversification and
-%           demography
-%           5. Comparative genomics to understand evolution of color and
-%           reproductive mode
-%           6. Evolutionary ecology---what ecological context influence
-%           evolution asexuality/live-bearing/color.
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
 \begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
-        \begin{itemize}
-            \item<1-> These comparative tools will allow evolutionary
-                biologists to leverage genomic data to infer shared
-                evolutionary history
-            \item<2-> This will better our understanding of how community scale
-                processes affect the generation and assembly of biodiversity
-            \item<3-> Many exciting empirical systems for applying new methods!
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Future directions: Gobi Desert}
-    \vspace{-2mm}
-    \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
-    \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Future directions: Gobi Desert}
-    \vspace{-2mm}
-    \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
-    \end{columns}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Future directions: West-Central Africa}
-    \begin{center}
-        \includegraphics<1->[height={3cm}]{/home/jamie/Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
-        \hspace{0.3mm}
-        \includegraphics<1->[height={3cm}]{/home/jamie/Dropbox/field-photos/africa/Figure4_herps.jpeg}
-    \end{center}
-
-    \only<1>{
-    \begin{center}
-        \includegraphics[height={3.5cm}]{/home/jamie/Dropbox/field-photos/people/adam.png}
-        \hspace{0.3mm}
-        \includegraphics[height={3.5cm}]{/home/jamie/Dropbox/field-photos/people/fujita.jpg}
-    \end{center}
-    }
-
-    \uncover<2->{
-    \begin{center}
-        \includegraphics[height=4.5cm]{../images/number-of-divs-africa.pdf}
-    \end{center}
+    \centerline{
+    \includegraphics<1>[height=8cm]{../images/bd-cartoon-no-lines.pdf}
+    \includegraphics<2>[height=8cm]{../images/bd-cartoon-shared.pdf}
     }
@@
 \begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-    \vspace{-2mm}
 \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/malaysia/malaysia-burungs.jpg}
+    \column{.5\textwidth}
+    \begin{minipage}[c]{\columnwidth}
+    \begin{itemize}
+        \item Blah
+    \end{itemize}
+    \end{minipage}
+
+    \column{.6\textwidth}
+    \begin{minipage}[t][\textheight][c]{\linewidth}
+    \includegraphics<1>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
+    \includegraphics<2>[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
+    \end{minipage}
 \end{columns}
@@
 
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Everything is on GitHub\ldots}
-    Software:\\
-    \begin{itemize}
-        \item \texttt{dpp-msbayes}:
-            \href{https://github.com/joaks1/dpp-msbayes}{\tt
-            https://github.com/joaks1/dpp-msbayes}
-
-        \item \texttt{PyMsBayes}:
-            \href{https://github.com/joaks1/PyMsBayes}{\tt
-            https://github.com/joaks1/PyMsBayes}
-
-        \item ABACUS: Approximate BAyesian C UtilitieS.
-            \href{https://github.com/joaks1/abacus}{\tt
-            https://github.com/joaks1/abacus}
-    \end{itemize}
-
-    \medskip
-    Open-Science Notebook:\\
-    \begin{itemize}
-        \item \texttt{msbayes-experiments}:
-            \href{https://github.com/joaks1/msbayes-experiments}{\tt
-            https://github.com/joaks1/msbayes-experiments}
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Acknowledgments}
-    \begin{columns}[t]
-        \column{.5\textwidth}
-            {\bf Ideas and feedback:}
-            \begin{myitemize}
-                \item Leach\'{e} Lab
-                \item Minin Lab
-                \item Holder Lab
-                \item Brown Lab/KU Herpetology
-                \item Melissa Callahan
-            \end{myitemize}
-            \smallskip
-            {\bf Computation:}\\
-            \includegraphics<1->[height={8mm}]{../images/iplant.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/kuittc.png}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/uw.png}\\
-
-        \column{.5\textwidth}
-            {\bf Funding:}\\
-            \includegraphics<1->[height={8mm}]{../images/nsf.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/ngs.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/ssb.png}\\
-
-            \smallskip
-            {\bf Photo credits:}
-            \begin{myitemize}
-                \item Rafe Brown, Cam Siler, Jesse Grismer, \& Jake Esselstyn
-                \item FMNH Philippine Mammal Website:
-                    \begin{myitemize}
-                        \item D.S.\ Balete, M.R.M.\ Duya, \& J.\ Holden
-                    \end{myitemize}
-                \item PhyloPic!
-            \end{myitemize}
-    \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Questions?}    
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{center}
-            {
-            \Large
-            \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
-            }
-        \end{center}
-        \column{.5\textwidth}
-            \begin{figure}
-                \begin{center}
-                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
-                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
-                \end{center}
-            \end{figure}
-    \end{columns}
-\end{frame}
-
-% Extra slides
-
-\begin{frame}[noframenumbering]
-    \frametitle{Easy as ABC}
-    \centerline{
-    \includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}
-    \includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}
-    \includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}
-    \includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}
-    \includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}
-    \includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}
-    \includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}
-    \includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}
-    \includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}
-    \includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}
-    }
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{Causes of bias: Insufficient sampling}
-    \begin{itemize}
-        \item Models with more parameter space are less densely sampled
-        \item Could explain bias toward small models in extreme cases
-        \item {\bf Predicts large variance in posterior estimates}
-        \begin{itemize}
-            \item We explored empirical and simulation-based analyses with
-                2, 5, and 10 million prior samples, and estimates were
-                very similar
-        \end{itemize}
-    \end{itemize}
-    \smallskip
-    \centerline{
-    \includegraphics[width=\textwidth]{../images/omega_over_sampling.pdf}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{\dppmsbayes: Simulation results}
-    % \vspace{1cm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_old_power_psi_mode.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_psi_mode_headless.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_dpp_power_psi_mode_headless.pdf}}
-    % \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{\dppmsbayes: Simulation results}
-    % \vspace{1cm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_old_power_omega_median.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_omega_median_headless.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_dpp_power_omega_median_headless.pdf}}
-    % \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{Empirical results: Philippine diversification}
-    \centerline{
-    \includegraphics[width=0.8\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp-with-prior.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
 \end{document}

### 2015-02-18 slides/bordeaux/slides-1.tex
Working on bordeaux slides.

--- 2015-02-18 slides/bordeaux/slides.tex
@@
 \begin{document}
-\maketitle
+
+% \maketitle
+\begin{frame}[noframenumbering]
+    \titlepage
+\end{frame}
 
@@
 \begin{frame}
-    \frametitle{Processes of diversification}
+    % \frametitle{Processes of diversification}
     \begin{columns}[c]
@@
                 community structure
-            \item<4-> We need methods for inferring evolutionary patterns
-                predicted by historical events from contemporary populations
+            \item<4-> We need methods for inferring patterns predicted by
+                community-scale events from contemporary populations
         \end{itemize}
@@
 \begin{frame}[t]
-    \frametitle<1-4>{Climate-driven diversification: Predictions}
-    \frametitle<5->{Divergence-model choice}
+    % \frametitle<1-4>{Climate-driven diversification: Predictions}
+    % \frametitle<5->{Divergence-model choice}
+    \frametitle{Inferring community history}
 
     \centering{
-    \includegraphics<1>[height=7.5cm]{../images/dmc-cartoon-islands.pdf}
-    \includegraphics<2>[height=7.5cm]{../images/dmc-cartoon-taxa.pdf}
-    \includegraphics<3>[height=7.5cm]{../images/dmc-cartoon-sea-level.pdf}
-    \includegraphics<4>[height=7.5cm]{../images/dmc-cartoon-shared-no-labels.pdf}
-    \includegraphics<5>[height=7.5cm]{../images/dmc-cartoon-shared.pdf}
-    \includegraphics<6>[height=7.5cm]{../images/dmc-cartoon-2-1.pdf}
-    \includegraphics<7>[height=7.5cm]{../images/dmc-cartoon-2-2.pdf}
-    \includegraphics<8>[height=7.5cm]{../images/dmc-cartoon-2-3.pdf}
-    \includegraphics<9>[height=7.5cm]{../images/dmc-cartoon-general.pdf}
+    % \includegraphics<1>[height=7.5cm]{../images/dmc-cartoon-islands.pdf}
+    % \includegraphics<2>[height=7.5cm]{../images/dmc-cartoon-taxa.pdf}
+    \includegraphics<1>[height=7.5cm]{../images/dmc-cartoon-sea-level.pdf}
+    % \includegraphics<4>[height=7.5cm]{../images/dmc-cartoon-shared-no-labels.pdf}
+    \includegraphics<2>[height=7.5cm]{../images/dmc-cartoon-shared.pdf}
+    \includegraphics<3>[height=7.5cm]{../images/dmc-cartoon-2-1.pdf}
+    \includegraphics<4>[height=7.5cm]{../images/dmc-cartoon-2-2.pdf}
+    \includegraphics<5>[height=7.5cm]{../images/dmc-cartoon-2-3.pdf}
+    \includegraphics<6>[height=7.5cm]{../images/dmc-cartoon-general.pdf}
     }
@@
 \begin{frame}[t]
-    \frametitle{Divergence-model choice}
+    % \frametitle{Divergence-model choice}
+    \frametitle{Inferring community history}
 
@@
             weighted posteriors
-        \item<3-> Multi-processing interface to improve performance for
-            accommodating genomic datasets
+        \item<3-> Multi-processing interface to accommodate genomic datasets
         % \item<4-> Uniform prior over divergence models
@@
 
+\section{Conclusions}
+
 \begin{frame}
+    \frametitle{Results}
+    \begin{itemize}
+        \item<1-> Results consistent with prediction of clustered
+            divergences
+        \item<2-> Results suggest multiple interglacial periods were important
+            drivers of diversification and community assembly across Islands of
+            Southeast Asia
+        \item<3-> However, there is a lot of uncertainty!
+    \end{itemize}
+    \barefootnote{\shortfullcite{Oaks2014dpp}}
+\end{frame}
+
+\begin{frame}[t]
     \frametitle{Current work: More data}
+    \begin{minipage}[t][0.15\textheight][c]{\linewidth}
     \begin{itemize}
-        \item<1-> Collecting genomic data from taxa co-distributed across Southeast
-            Asia
+        \item<1-> Collecting genomic data from taxa co-distributed across
+            Southeast Asian Islands and Mainland
         \item<2-> Preliminary results for 1000 loci from 5 pairs of \emph{Gekko
@@
     \end{itemize}
+    \end{minipage}
 
@@
 
-    \begin{itemize}[<+->]
-        \item Full-likelihood Bayesian implementation
     \begin{itemize}
-                \item Uses all the information in the data
-                \item Applicable to deeper timescales
+        \item<1-> Full-likelihood Bayesian implementation
+            \begin{itemize}
+                \item<2-> Uses all the information in the data
+                \item<2-> Applicable to deeper timescales
             \end{itemize}
-        \item Analytically integrate over gene trees
+        \item<3-> Analytically integrate over gene trees
             \footnote{\tiny\shortfullcite{Bryant2012}}
             \begin{itemize}
-                \item Very efficient numerical approximation of posterior 
-                \item Applicable to NGS datasets
+                \item<4-> Very efficient numerical approximation of posterior 
+                \item<4-> Applicable to NGS datasets
             \end{itemize}
@@
 
-\section{Conclusions}
+\section{Teaching}
+
+{
+\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/kane.jpg}}
+\begin{frame}[t]
+    \vspace{4.5cm}
+
+    \uncover<2->{
+    Working with the UW Biology Education Research Group (BERG) \\
+    }
+
+    \begin{itemize}
+        \item<3-> Designing active-learning modules to teach fundamental concepts
+            in evolution and statistics
+        \item<4-> Taking an evidence-based approach to the classroom as a lecturer
+            for Introductory Biology
+    \end{itemize}
+\end{frame}
+}
 
 \begin{frame}
-    \frametitle{Climate-driven diversification model?}
+    \frametitle{Collaborative teaching}
+    \centerline{
+    \includegraphics[height=1.5cm]{../images/software-carpentry.pdf}
+    }
+    Software Carpentry Bootcamp\\
+
+    \begin{uncoverenv}<2->
     \begin{itemize}
-        \item<1-> Results consistent with prediction of clustered
-            divergences
-        \item<2-> Results suggest multiple co-divergences
-        \item<3-> However, there is a lot of uncertainty!
+        \item Two-day crash course in basic command line and software skills
+        \item Taught an international group of biologists how to increase their
+            research productivity.
     \end{itemize}
+    \end{uncoverenv}
+
+    Molecular Evolution Workshops\\
+
+    \begin{uncoverenv}<3->
+    \begin{itemize}
+        \item Introduce international groups of graduate students, postdocs,
+            and faculty to the fundamentals of molecular evolution and
+            state-of-the-art software
+    \end{itemize}
+    \end{uncoverenv}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Acknowledgments}
+    \begin{columns}[t]
+        \column{.5\textwidth}
+            {\bf Ideas and feedback:}
+            \begin{myitemize}
+                \item Leach\'{e} Lab
+                \item Minin Lab
+                \item Holder Lab
+                \item Brown Lab/KU Herpetology
+                \item Melissa Callahan
+            \end{myitemize}
+            \smallskip
+            {\bf Computation:}\\
+            \includegraphics<1->[height={8mm}]{../images/iplant.jpg}
+            \hspace{0.5mm}
+            \includegraphics<1->[height={8mm}]{../images/kuittc.png}
+            \hspace{0.5mm}
+            \includegraphics<1->[height={8mm}]{../images/uw.png}\\
+
+        \column{.5\textwidth}
+            {\bf Funding:}\\
+            \includegraphics<1->[height={8mm}]{../images/nsf.jpg}
+            \hspace{0.5mm}
+            \includegraphics<1->[height={8mm}]{../images/ngs.jpg}
+            \hspace{0.5mm}
+            \includegraphics<1->[height={8mm}]{../images/ssb.png}\\
+
+            \smallskip
+            {\bf Photo credits:}
+            \begin{myitemize}
+                \item Rafe Brown, Cam Siler, Jesse Grismer, \& Jake Esselstyn
+                \item FMNH Philippine Mammal Website:
+                    \begin{myitemize}
+                        \item D.S.\ Balete, M.R.M.\ Duya, \& J.\ Holden
+                    \end{myitemize}
+                \item PhyloPic!
+            \end{myitemize}
+    \end{columns}
+\end{frame}
+
+% \begin{frame}
+%     \frametitle{Questions?}    
+%     \begin{columns}[c]
+%         \column{.5\textwidth}
+%         \begin{center}
+%             {
+%             \Large
+%             \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
+%             }
+%         \end{center}
+%         \column{.5\textwidth}
+%             \begin{figure}
+%                 \begin{center}
+%                 \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
+%                 \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
+%                 \end{center}
+%             \end{figure}
+%     \end{columns}
+% \end{frame}
+
+% Extra slides
+
+\begin{frame}[noframenumbering]
+    \frametitle{Easy as ABC}
+    \centerline{
+    \includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}
+    \includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}
+    \includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}
+    \includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}
+    \includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}
+    \includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}
+    \includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}
+    \includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}
+    \includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}
+    \includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}
+    }
+\end{frame}
+
+\begin{frame}[noframenumbering]
+    \frametitle{Causes of bias: Insufficient sampling}
+    \begin{itemize}
+        \item Models with more parameter space are less densely sampled
+        \item Could explain bias toward small models in extreme cases
+        \item {\bf Predicts large variance in posterior estimates}
+        \begin{itemize}
+            \item We explored empirical and simulation-based analyses with
+                2, 5, and 10 million prior samples, and estimates were
+                very similar
+        \end{itemize}
+    \end{itemize}
+    \smallskip
+    \centerline{
+    \includegraphics[width=\textwidth]{../images/omega_over_sampling.pdf}}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+    \frametitle{\dppmsbayes: Simulation results}
+    % \vspace{1cm}
+        \centerline{
+        \includegraphics[width=\textwidth]{../images/old_old_power_psi_mode.pdf}}
+        \vspace{0mm}
+        \centerline{
+        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_psi_mode_headless.pdf}}
+        \vspace{0mm}
+        \centerline{
+        \includegraphics[width=\textwidth]{../images/old_dpp_power_psi_mode_headless.pdf}}
+    % \barefootnote{\shortfullcite{Oaks2014dpp}}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+    \frametitle{\dppmsbayes: Simulation results}
+    % \vspace{1cm}
+        \centerline{
+        \includegraphics[width=\textwidth]{../images/old_old_power_omega_median.pdf}}
+        \vspace{0mm}
+        \centerline{
+        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_omega_median_headless.pdf}}
+        \vspace{0mm}
+        \centerline{
+        \includegraphics[width=\textwidth]{../images/old_dpp_power_omega_median_headless.pdf}}
+    % \barefootnote{\shortfullcite{Oaks2014dpp}}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+    \frametitle{Empirical results: Philippine diversification}
+    \centerline{
+    \includegraphics[width=0.8\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp-with-prior.pdf}}
     \barefootnote{\shortfullcite{Oaks2014dpp}}
@@
 
-\section{Teaching}
-
-\begin{frame}
-    BIO 180 stuff
-\end{frame}
-
-\begin{frame}
-    Software carpentry \\
-
-    Workshops
-\end{frame}
-
-\begin{frame}
-    \frametitle{Questions?}    
-    \begin{center}
-        {
-        \Large
-        \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
-        }
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Future directions---Computational}
-    \begin{itemize}
-        \item<1-> Full-phylogenetic framework
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
-        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \centerline{
-    \includegraphics<1>[height=8cm]{../images/bd-cartoon-no-lines.pdf}
-    \includegraphics<2>[height=8cm]{../images/bd-cartoon-shared.pdf}
-    }
-\end{frame}
-
-\begin{frame}
-\begin{columns}
-    \column{.5\textwidth}
-    \begin{minipage}[c]{\columnwidth}
-    \begin{itemize}
-        \item Blah
-    \end{itemize}
-    \end{minipage}
-
-    \column{.6\textwidth}
-    \begin{minipage}[t][\textheight][c]{\linewidth}
-    \includegraphics<1>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
-    \includegraphics<2>[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
-    \end{minipage}
-\end{columns}
-\end{frame}
-
 \end{document}

### 2015-02-18 slides/bordeaux/slides-2.tex
Working on bordeaux slides.

\input{../utils/preamble.tex}
\input{../../manuscripts/utils/macros.tex}
\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
\bibliography{../../manuscripts/bib/references}

% \title[Estimating shared history]{An Improved Approximate-Bayesian Method for Estimating Shared
% Evolutionary History}

\title[Future directions]{Future Directions}

\author[J.\ Oaks]{
    Jamie R.\ Oaks\inst{1}
}
\institute[University of Washington]{
    \inst{1}%
        Department of Biology, University of Washington
}

% \date{\today}
\date{February 18, 2015}

\begin{document}

\begin{frame}[t]
    \frametitle{New methods for inferring community history}
    \vspace{-5mm}
    \begin{minipage}[t][0.1\textheight][c]{1\linewidth}
    \begin{itemize}
        \item<2-> Incorporate comparative models of community-scale
            co-diversification into a full-phylogenetic framework
    \end{itemize}
    \end{minipage}

    \vspace{-5mm}
    \begin{minipage}[t][0.9\textheight][c]{1\linewidth}
    \begin{center}
        \includegraphics<3>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
        \includegraphics<4>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
    \end{center}
    \end{minipage}
\end{frame}

\begin{frame}
    \begin{columns}
        \column{0.4\textwidth}
        \begin{minipage}[c]{\columnwidth}
        \begin{adjustwidth}{-2.3em}{}
        \begin{onlyenv}<1-6>
        \begin{itemize}
            \item<2-> Current phylogenetic methods assume speciation and extinction
                occur independently of one another
            \item<3-> We know this assumption is frequently violated
            \item<4-> I will generalize a birth-death model to accommodate
                ``shared'' events
            \item<6> Infer events jointly across multiple trees
        \end{itemize}
        \end{onlyenv}
        \begin{onlyenv}<7->
        \only<7->{\hspace{0.5cm}\textbf{\large Benefits:}\\}
        \begin{itemize}
            \item<8-> Improve phylogenetic inference and divergence-time estimation
            \item<9-> Provide a general framework for inferring the evolutionary
                history of whole communities
            \item<10-> Possible epidemiological applications
        \end{itemize}
        \end{onlyenv}
        \end{adjustwidth}
        \end{minipage}

        \column{0.6\textwidth}
        \begin{minipage}[t][\textheight][c]{\linewidth}
        \centerline{
        \includegraphics<1-3>[width=\columnwidth]{../images/bd-cartoon-no-lines.pdf}
        \includegraphics<4>[width=\columnwidth]{../images/bd-cartoon-shared.pdf}
        \includegraphics<5>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
        \includegraphics<6->[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
        }
        \end{minipage}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Inferring the evolutionary context of ecosystems}
    \begin{itemize}[<+->]
        \item Use comparative framework to estimate the history of floral and
            faunal communities across Europe
        \item A better understanding of the diversification and assembly
            of communities will provide a framework for understanding
            the affect of past climate change
        \item This will provide context for predicting future change
    \end{itemize}
\end{frame}

\begin{frame}
    \begin{itemize}[<+->]
        \item Generalized birth-death process might better model
            epidemiological dynamics
        \item Could be relevant to the evolution of grapevine pathogens
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Strategies for integration}
    \begin{itemize}[<+->]
        \item In order to predict the response of ecosystems to environmental
            change, we need a general model of ecosystem processes
        \item Organize a series ``hackathons'' to stimulate synthesis among the
            biologists, engineers, and theorists across LabEx 
        \item Each hackathon would focus on integrating processes and existing
            models between particular specialties and/or ecosystems
    \end{itemize}
\end{frame}

\end{document}

### 2015-02-18 slides/images/pull-images.sh

@@ -11,3 +11,4 @@ else
 fi
 
 curl -o darwin-tol-copyright-boris-kulikov-2007.jpg http://www.boriskulikov.com/images/DarwinTreeOfLife-%20L.jpg
+curl -o kane.jpg http://uwberg.com/wp-content/uploads/2014/05/berg-hero.jpg

### 2015-02-18 slides/bordeaux/slides-1.tex
Working on bordeaux slides

@@
 
-        \vspace{4mm}
+        % \vspace{1mm}
 
@@
         \end{uncoverenv}
+
+        \vspace{-1mm}
+
+        \begin{uncoverenv}<4->
+            \begin{center}
+                \begin{itemize}
+                    \small
+                    \item Dirichlet-process prior (DPP) over all possible divergence
+                        models
+                    % \item<2-> Flexible priors on continuous parameters to avoid strongly
+                    %     weighted posteriors
+                    \item Multi-processing interface to accommodate genomic datasets
+                \end{itemize}
+            \end{center}
+        \end{uncoverenv}
     \end{minipage}
-\end{frame}
-
-\begin{frame}
-    \frametitle{New method: \dppmsbayes}
-    \begin{itemize}
-        % \item<1-> Reparameterized the model implemented in \msb
-        \item<1-> Dirichlet-process prior (DPP) over all possible divergence
-            models
-        \item<2-> Flexible priors on continuous parameters to avoid strongly
-            weighted posteriors
-        \item<3-> Multi-processing interface to accommodate genomic datasets
-        % \item<4-> Uniform prior over divergence models
-    \end{itemize}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
 \end{frame}
@@
     \begin{itemize}
-        \item<1-> Extensive computer simulations to compare performance to previous
+        \item<1-> Used computer simulations to compare performance to previous
             state-of-the-art methods (\msb).
@@
 
-    \uncover<3->{
-    \textbf{\dppmsbayes allows biologists to leverage comparative genomic data
-        to assess the affects of community-scale processes on biodiversity}
-    }
+    % \uncover<3->{
+    % \textbf{\dppmsbayes allows biologists to leverage comparative genomic data
+    %     to assess the affects of community-scale processes on biodiversity}
+    % }
     \barefootnote{\shortfullcite{Oaks2014dpp}}
@@
 {
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
+% \usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
+% \begin{frame}
+%     % \frametitle{Climate-driven diversification}    
+%     \begin{columns}
+%         \column{0.6\textwidth}
 
-        \vspace{6.5cm}
+%         \vspace{6.5cm}
 
-        \ \\
+%         \ \\
 
-        \column{0.4\textwidth}
+%         \column{0.4\textwidth}
 
-        \vspace{-2cm}
+%         \vspace{-2cm}
 
-        \begin{uncoverenv}<1->
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-        \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
+%         \begin{uncoverenv}<1->
+%         \colorbox{white}{
+%             \begin{minipage}[t]{1.0\textwidth}
+%                 \raggedright
+%                 \textbf{Did repeated fragmentation of islands during
+%                     inter-glacial rises in sea level promote diversification?}
+%             \end{minipage}
+%         }
+%         \end{uncoverenv}
+%     \end{columns}
+% \end{frame}
+% }
 
@@
     \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
+    \includegraphics[height=5cm]{../../empirical-analyses/plots/philippines-dpp-psi-posterior.pdf}}
     \smallskip
@@
     \begin{itemize}
-        \item<1-> Results consistent with prediction of clustered
-            divergences
+        \item \textbf{dpp-msbayes allows biologists to leverage comparative
+                genomic data to assess the affects of community-scale processes
+                on biodiversity}
+        % \item<1-> Results consistent with prediction of clustered
+        %     divergences
         \item<2-> Results suggest multiple interglacial periods were important

### 2015-02-18 slides/bordeaux/slides-2.tex
Working on bordeaux slides

@@
 \begin{frame}
-    \frametitle{Inferring the evolutionary context of ecosystems}
-    \begin{itemize}[<+->]
-        \item Use comparative framework to estimate the history of floral and
-            faunal communities across Europe
-        \item A better understanding of the diversification and assembly
-            of communities will provide a framework for understanding
-            the affect of past climate change
-        \item This will provide context for predicting future change
+    \frametitle{Relevance to LabEx COTE}
+    \begin{itemize}
+        \item<2-> Use novel comparative method to estimate the history of floral
+            and faunal communities across Europe
+        \item<3-> Elucidate the history of communities and their response to past
+            environmental climate change
+        \item<4-> This will provide context for predicting future change
+        \item<5-> Collaboration with BIOGECO partners
     \end{itemize}
@@
 \begin{frame}
+    \frametitle{Relevance to LabEx COTE}
     \begin{itemize}[<+->]
-        \item Generalized birth-death process might better model
-            epidemiological dynamics
-        \item Could be relevant to the evolution of grapevine pathogens
+        \item Generalized birth-death process will be relevant to
+            modeling epidemiological dynamics
+        \item Relevant to the evolution of grapevine pathogens
+        \item Collaboration with SV partners to investigate disease dynamics
     \end{itemize}
@@
     \frametitle{Strategies for integration}
-    \begin{itemize}[<+->]
-        \item In order to predict the response of ecosystems to environmental
+    \begin{itemize}
+        \item<2-> In order to predict the response of ecosystems to environmental
             change, we need a general model of ecosystem processes
-        \item Organize a series ``hackathons'' to stimulate synthesis among the
-            biologists, engineers, and theorists across LabEx 
-        \item Each hackathon would focus on integrating processes and existing
-            models between particular specialties and/or ecosystems
+        \item<3-> This requires integration across LabEx COTE partners
+        \item<3-> Organize a series ``hackathons'' to stimulate synthesis among the
+            biologists, engineers, theorists, and modelers across LabEx COTE
+        \item<4-> Each hackathon would focus on integrating concepts and existing
+            models between particular ecosystems and/or ecosystem processes
     \end{itemize}
@@
 
+\begin{frame}
+    \frametitle{Teaching prospects}
+    \begin{itemize}[<+->]
+        \item Teach courses at the nexus of evolution, ecology, statistics,
+            and computer science
+        \begin{itemize}
+            \item Computational biology
+            \item Advance statistics in biology (likelihood and Bayesian
+                approaches)
+            \item Advanced methods in evolutionary genetics and
+                molecular ecology
+        \end{itemize}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Questions?}    
+    \begin{columns}[c]
+        \column{.5\textwidth}
+        \begin{center}
+            {
+            \Large
+            \href{mailto:joaks1@uw.edu}{\tt joaks1@uw.edu}
+            }
+        \end{center}
+        \column{.5\textwidth}
+            \begin{figure}
+                \begin{center}
+                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
+                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
+                \end{center}
+            \end{figure}
+    \end{columns}
+\end{frame}
+
 \end{document}

### 2015-02-28 slides/bin/create-tree-gif.sh
More work on script to create tree gif for slides.

@@ -1,7 +1,16 @@
 #!/bin/bash
 
+# Run tree plotting script
 python plot-sec-tree.py
 
+# frame_dir="../images/mascot-tree-animation-frames"
+# if [ ! -d "$frame_dir" ]
+# then
+#     mkdir -p "$frame_dir"
+# fi
+
+
+# Get the dimensions of the tree png
 dimensions="$(identify sec-tree-bare.png | egrep -o ' [0-9]+x[0-9]+ ')"
 dimensions=${dimensions/ /}
 width=${dimensions/x*/}
@@ -9,17 +18,28 @@ height=${dimensions/*x/}
 
 slice=$(expr $width / 100)
 
+# Create series of masked tree pngs
+# cnt=0
 for i in $(seq -w 0 $slice $width)
 do
     current_width=$(expr $width - $i)
     echo "$i of $width"
     convert sec-tree-bare.png -alpha set -region "${current_width}x${height}+${i}+0" -alpha transparent mt-${i}.png
+    # cp "mt-${i}.png" "${frame_dir}/mascot-tree-frame-${cnt}.png"
+    # cnt=$(expr $cnt + 1)
 done
 
+# cp sec-tree-bare.png "${frame_dir}/mascot-tree-frame-${cnt}.png"
+# cnt=$(expr $cnt + 1)
+# cp sec-tree.png "${frame_dir}/mascot-tree-frame-${cnt}.png"
+
+# Create gif from png files
 convert -layers OptimizePlus -delay 5 mt-*.png sec-tree-bare.png -delay 250 sec-tree.png -loop 0 ../images/mascot-tree.gif
 
-convert -background white -alpha remove -layers OptimizePlus -delay 5 mt-*.png sec-tree-bare.png -delay 250 sec-tree.png -loop 0 ../images/mascot-tree2.gif
+# convert -background white -alpha remove -layers OptimizePlus -delay 5 mt-*.png sec-tree-bare.png -delay 250 sec-tree.png -loop 0 ../images/mascot-tree2.gif
 
 mv sec-tree.png ../images/mascot-tree.png
+mv sec-tree-bare.png ../images/mascot-tree-bare.png
 
-rm sec-tree-bare.png mt-*.png
+# Remove intermediate files
+rm mt-*.png

### 2015-03-03 slides/auburn/intro-mascot-tree.tex
Working on slides for auburn.

\begin{frame}
    \begin{center}
        % \transwipe<2>[direction=0, duration=3]
        % \includegraphics<2>[width=\textwidth]{../images/mascot-tree.png}
        \includegraphics[width=\textwidth]{../images/mascot-tree.png}

    \end{center}
\end{frame}

### 2015-03-03 slides/auburn/intro-se-asia.tex
Working on slides for auburn.

--- 2015-02-18 slides/bordeaux/slides.tex
@@
-\input{../utils/preamble.tex}
-\input{../../manuscripts/utils/macros.tex}
-\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
-\bibliography{../../manuscripts/bib/references}
-
-% \title[Estimating shared history]{An Improved Approximate-Bayesian Method for Estimating Shared
-% Evolutionary History}
-
-\title[Inferring community history]{Inferring Community-scale Diversification
-    from Genomic Data}
-
-\author[J.\ Oaks]{
-    Jamie R.\ Oaks\inst{1}
-}
-\institute[University of Washington]{
-    \inst{1}%
-        Department of Biology, University of Washington
-}
-
-% \date{\today}
-\date{February 18, 2015}
-
-\begin{document}
-\maketitle
-
 {
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
-        \begin{itemize}
-            \item<2-> Community-scale processes are important in
-                biodiversification and community assembly
-            \item<3-> Accounting for such processes will better our
-                understanding of evolutionary history and
-                community structure
-            \item<4-> We need methods for inferring evolutionary patterns
-                predicted by historical events from contemporary populations
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
+\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120-solid.png}}
 \begin{frame}
@@
 {
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
+\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
 \begin{frame}
@@
 }
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1>
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rob.jpg}}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/charles.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/cam.jpg}}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jake.jpg}}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/allie.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/Luke.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/anthony.png}}
-        \end{minipage}
-        \end{onlyenv}
-
-        \begin{onlyenv}<2->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1-2>[width=\textwidth]{../images/maps/Philippines.png}
-        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        % \begin{onlyenv}<4->
-        %     \begin{itemize}
-        %         \item Strong support for simultaneous divergence of all 22 taxon pairs
-
-        %         \item Posterior probability $>$ 0.96 \\
-
-        %         \item $\sim$100,000--250,000 years ago
-        %     \end{itemize}
-        % \end{onlyenv}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle<1-4>{Climate-driven diversification: Predictions}
-    \frametitle<5->{Divergence-model choice}
-
-    \centering{
-    \includegraphics<1>[height=7.5cm]{../images/dmc-cartoon-islands.pdf}
-    \includegraphics<2>[height=7.5cm]{../images/dmc-cartoon-taxa.pdf}
-    \includegraphics<3>[height=7.5cm]{../images/dmc-cartoon-sea-level.pdf}
-    \includegraphics<4>[height=7.5cm]{../images/dmc-cartoon-shared-no-labels.pdf}
-    \includegraphics<5>[height=7.5cm]{../images/dmc-cartoon-shared.pdf}
-    \includegraphics<6>[height=7.5cm]{../images/dmc-cartoon-2-1.pdf}
-    \includegraphics<7>[height=7.5cm]{../images/dmc-cartoon-2-2.pdf}
-    \includegraphics<8>[height=7.5cm]{../images/dmc-cartoon-2-3.pdf}
-    \includegraphics<9>[height=7.5cm]{../images/dmc-cartoon-general.pdf}
-    }
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Divergence-model choice}
-
-    \vspace{-7mm}
-
-    \begin{minipage}[t][0.1\textheight][c]{1.1\linewidth}
-        \begin{adjustwidth}{-0.5em}{}
-            \begin{tabular}{ p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} }
-                $m_1$ & $m_2$ & $m_3$ & $m_4$ & $m_5$ \\
-            \end{tabular}
-        \end{adjustwidth}
-    \end{minipage}
-
-    \vspace{-2mm}
-
-    \begin{minipage}[t][0.35\textheight][c]{1.1\linewidth}
-        \begin{adjustwidth}{-2.3em}{}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-shared.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-1.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-2.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-3.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-general.pdf}
-        \end{adjustwidth}
-    \end{minipage}
-
-    \vspace{3.5mm}
-
-    \begin{minipage}[c][0.35\textheight][c]{\linewidth}
-        \begin{uncoverenv}<2->
-            \begin{center}
-            We want to infer \textcolor{blue}{\divModel{}} and
-            \textcolor{blue}{\divTimeMapVector} given DNA sequence
-            alignments
-            \textcolor{blue}{\alignmentVector}
-        \end{center}
-        \end{uncoverenv}
-
-        \vspace{4mm}
-
-        \begin{uncoverenv}<3->
-            \begin{displaybox}[0.85\linewidth]
-                \begin{minipage}[c][0.12\textheight][c]{\linewidth}
-
-                \[
-                    p(\divTimeMapVector,
-                      \divModel{i},
-                      \allParameters{}
-                      \given \alignmentVector)
-                      =
-                    \frac{
-                        p(\alignmentVector \given
-                          \divTimeMapVector,
-                          \allParameters{},
-                          \divModel{i})
-                        p(\divTimeMapVector,
-                          \allParameters{}
-                          \given \divModel{i})
-                        p(\divModel{i})
-                        }{p(\alignmentVector)}
-                \]
-                \vspace{-1mm}
-                \end{minipage}
-            \end{displaybox}
-        \end{uncoverenv}
-    \end{minipage}
-\end{frame}
-
-\begin{frame}
-    \frametitle{New method: \dppmsbayes}
-    \begin{itemize}
-        % \item<1-> Reparameterized the model implemented in \msb
-        \item<1-> Dirichlet-process prior (DPP) over all possible divergence
-            models
-        \item<2-> Flexible priors on continuous parameters to avoid strongly
-            weighted posteriors
-        \item<3-> Multi-processing interface to improve performance for
-            accommodating genomic datasets
-        % \item<4-> Uniform prior over divergence models
-    \end{itemize}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Performance}
-    \begin{itemize}
-        \item<1-> Extensive computer simulations to compare performance to previous
-            state-of-the-art methods (\msb).
-        \item<2-> New method for estimating shared evolutionary history shows
-            improved:
-            \begin{enumerate}
-                \item<2-> Improved model-choice accuracy 
-                \item<2-> More robustness to model violations
-                \item<2-> Greater power to detect variation in divergence times
-                \item<2-> Faster!
-            \end{enumerate}
-    \end{itemize}
-
-    \uncover<3->{
-    \textbf{\dppmsbayes allows biologists to leverage comparative genomic data
-        to assess the affects of community-scale processes on biodiversity}
-    }
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-
-        \vspace{6.5cm}
-
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \begin{uncoverenv}<1->
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-        \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1->[width=\textwidth]{../images/maps/Philippines.png}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
-    \smallskip
-    \centerline{
-    \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Current work: More data}
-    \begin{itemize}
-        \item<1-> Collecting genomic data from taxa co-distributed across Southeast
-            Asia
-        \item<2-> Preliminary results for 1000 loci from 5 pairs of \emph{Gekko
-                mindorensis} populations
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<2->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Current work: More power}
-    \centerline{
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/mtholder.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/vladimir.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../images/revbayes.png}}
-
-    \begin{itemize}[<+->]
-        \item Full-likelihood Bayesian implementation
-            \begin{itemize}
-                \item Uses all the information in the data
-                \item Applicable to deeper timescales
-            \end{itemize}
-        \item Analytically integrate over gene trees
-            \footnote{\tiny\shortfullcite{Bryant2012}}
-            \begin{itemize}
-                \item Very efficient numerical approximation of posterior 
-                \item Applicable to NGS datasets
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-\section{Conclusions}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification model?}
-    \begin{itemize}
-        \item<1-> Results consistent with prediction of clustered
-            divergences
-        \item<2-> Results suggest multiple co-divergences
-        \item<3-> However, there is a lot of uncertainty!
-    \end{itemize}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\section{Teaching}
-
-\begin{frame}
-    BIO 180 stuff
-\end{frame}
-
-\begin{frame}
-    Software carpentry \\
-
-    Workshops
-\end{frame}
-
-\begin{frame}
-    \frametitle{Questions?}    
-    \begin{center}
-        {
-        \Large
-        \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
-        }
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Future directions---Computational}
-    \begin{itemize}
-        \item<1-> Full-phylogenetic framework
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
-        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \centerline{
-    \includegraphics<1>[height=8cm]{../images/bd-cartoon-no-lines.pdf}
-    \includegraphics<2>[height=8cm]{../images/bd-cartoon-shared.pdf}
-    }
-\end{frame}
-
-\begin{frame}
-\begin{columns}
-    \column{.5\textwidth}
-    \begin{minipage}[c]{\columnwidth}
-    \begin{itemize}
-        \item Blah
-    \end{itemize}
-    \end{minipage}
-
-    \column{.6\textwidth}
-    \begin{minipage}[t][\textheight][c]{\linewidth}
-    \includegraphics<1>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
-    \includegraphics<2>[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
-    \end{minipage}
-\end{columns}
-\end{frame}
-
-\end{document}

### 2015-03-03 slides/bin/plot-sec-tree.py

@@ -7,30 +7,45 @@ from PIL import Image
 import ete2
 
 _extinct_lengths = {
-        "mascot-sewanee-tiger": 0.02,
-        "mascot-tulane-green-wave": 0.5,
-        "mascot-georgia-tech-yellow-jacket": 0.5,
+        "mascot-sewanee-tiger": 0.14,
+        "mascot-tulane-green-wave": 0.29,
+        "mascot-georgia-tech-yellow-jacket": 0.43,
+        }
+_extinct_lengths_shared = {
+        "mascot-sewanee-tiger": 0.15,
+        "mascot-tulane-green-wave": 0.26,
+        "mascot-georgia-tech-yellow-jacket": 0.43,
         }
 _sec_tree_ladder = """
-(((((((((((mascot-auburn-tiger.png:0.01,mascot-lsu-tiger.jpg:0.01):0.01,mascot-missouri-tiger.png:0.02):0.04,mascot-sewanee-tiger.jpg:{mascot-sewanee-tiger}):0.05,mascot-kentucky-wild-cat.jpg:0.11):0.11,
-((((mascot-georgia-bulldog.jpg:0.01,mascot-mississippi-state-bulldog.gif:0.01):0.03,mascot-texas-am-reveille.gif:0.04):0.06,mascot-tennessee-smokey.png:0.1):0.03,mascot-ole-miss-rebel-black-bear.jpg:0.13):0.09):0.005,
-mascot-arkansas-razorback.png:0.225):0.095,
-mascot-vanderbilt-commodore-cartoon.gif:0.32):0.025,
-mascot-alabama-big-al.gif:0.345):0.145,
-(mascot-sc-gamecock.png:0.12,mascot-florida-gator.png:0.12):0.37):0.08,
+(((((((((((mascot-auburn-tiger.png:0.03,mascot-lsu-tiger.jpg:0.03):0.04,mascot-missouri-tiger.png:0.07):0.12,mascot-sewanee-tiger.jpg:{mascot-sewanee-tiger}):0.13,mascot-kentucky-wild-cat.jpg:0.32):0.08,
+((((mascot-georgia-bulldog.jpg:0.01,mascot-mississippi-state-bulldog.gif:0.01):0.03,mascot-texas-am-reveille.gif:0.04):0.08,mascot-tennessee-smokey.png:0.12):0.22,mascot-ole-miss-rebel-black-bear.jpg:0.34):0.06):0.005,
[…]

### 2015-03-03 slides/bin/plot-sec-tree-gram.py
Working on slides.
Adding another script for plotting mascot tree, this time using Gram
rather than ete2.

@@ -1,0 +1,130 @@
+#! /usr/bin/env python
+
+import os
+import sys
+
+from PIL import Image
+import ete2
+
+from Gram import TreeGram
+var.punctuation = var.phylip_punctuation
+
+_wide_images = ['mascot-missouri-tiger.png',
+                'mascot-florida-gator.png',
+                'mascot-arkansas-razorback.png',
+                ]
+_extinct_images = [
+        "mascot-sewanee-tiger.jpg",
+        "mascot-tulane-green-wave.png",
+        "mascot-georgia-tech-yellow-jacket.png",
+        ]
+_extinct_lengths = {
+        "mascot-sewanee-tiger": 0.14,
+        "mascot-tulane-green-wave": 0.29,
+        "mascot-georgia-tech-yellow-jacket": 0.43,
+        }
+_extinct_lengths_shared = {
+        "mascot-sewanee-tiger": 0.15,
+        "mascot-tulane-green-wave": 0.26,
+        "mascot-georgia-tech-yellow-jacket": 0.43,
+        }
+_sec_tree_ladder = """
+(((((((((((mascot-auburn-tiger.png:0.03,mascot-lsu-tiger.jpg:0.03):0.04,mascot-missouri-tiger.png:0.07):0.12,mascot-sewanee-tiger.jpg:{mascot-sewanee-tiger}):0.13,mascot-kentucky-wild-cat.jpg:0.32):0.08,
+((((mascot-georgia-bulldog.jpg:0.01,mascot-mississippi-state-bulldog.png:0.01):0.03,mascot-texas-am-reveille.png:0.04):0.08,mascot-tennessee-smokey.png:0.12):0.22,mascot-ole-miss-rebel-black-bear.jpg:0.34):0.06):0.005,
+mascot-arkansas-razorback.png:0.405):0.045,
+mascot-vanderbilt-commodore-cartoon.png:0.45):0.025,
+mascot-alabama-big-al.png:0.475):0.015,
+(mascot-sc-gamecock.png:0.33,mascot-florida-gator.png:0.33):0.16):0.08,
+mascot-georgia-tech-yellow-jacket.png:{mascot-georgia-tech-yellow-jacket}):0.02,
+mascot-tulane-green-wave.png:{mascot-tulane-green-wave}):0.01;
+""".format(**_extinct_lengths)
+_sec_tree = """
[…]

### 2015-03-03 slides/images/pull-images.sh

@@ -16,16 +16,20 @@ curl -o mascot-tulane-green-wave.png http://upload.wikimedia.org/wikipedia/en/9/
 curl -o mascot-sewanee-tiger.jpg http://www.sewaneetigers.com/sports/wbkb/2010-11/releases/MBB%200115110836.html%20h35g7.jpeg
 curl -o mascot-georgia-tech-yellow-jacket.png http://upload.wikimedia.org/wikipedia/en/8/8f/GeorgiaTechYellowJackets.png
 curl -o mascot-alabama-big-al.gif http://www.survivingcollege.com/wp-content/uploads/2013/07/University-of-Alabama-Crimson-Tide-Big-Al-Elephant-Mascot-Logo.gif
+convert mascot-alabama-big-al.gif mascot-alabama-big-al.png
 curl -o mascot-florida-gator.png http://upload.wikimedia.org/wikipedia/en/thumb/1/12/Florida_Gators_logo.svg/470px-Florida_Gators_logo.svg.png
 curl -o mascot-sc-gamecock.png http://upload.wikimedia.org/wikipedia/commons/thumb/6/6a/South_Carolina_Gamecocks_Block_C_logo.svg/182px-South_Carolina_Gamecocks_Block_C_logo.svg.png
 curl -o mascot-vanderbilt-commodore.png http://redanglespanish.files.wordpress.com/2011/12/commodore.png
 curl -o mascot-vanderbilt-commodore-cartoon.gif http://www.nashvillescene.com/binary/0511/vanderbilt_mascot.gif
+convert mascot-vanderbilt-commodore-cartoon.gif mascot-vanderbilt-commodore-cartoon.png
 curl -o mascot-texas-am-reveille.gif http://content.sportslogos.net/logos/34/866/thumbs/86645362001.gif
+convert mascot-texas-am-reveille.gif mascot-texas-am-reveille.png
 curl -o mascot-tennessee-smokey.png http://content.sportslogos.net/logos/34/861/full/2892_tennessee_volunteers-mascot-2005.png
 curl -o mascot-arkansas-razorback.png http://upload.wikimedia.org/wikipedia/commons/e/ef/Arkansas-Razorback-Logo-2001.png
 curl -o mascot-ole-miss-rebel-black-bear.jpg http://uwire.com/wp-content/uploads/2010/10/news_bear_contrib.jpg
[…]

### 2015-03-04 slides/mascot-tree/mascot-tree-shared-events.tikz.tex
Working on slides.

%% This is a tikz file
\tikzset{node lower left/.style={font=\scriptsize,anchor=north east,text height=0.192cm,text depth=0.055cm,inner sep=0.03cm},
leaf/.style={font=\small,anchor=west,text height=0.240cm,text depth=0.068cm},
node upper left/.style={font=\scriptsize,anchor=south east,text height=0.192cm,text depth=0.055cm,inner sep=0.03cm},
bracket label/.style={font=\small,anchor=west,text height=0.240cm,text depth=0.068cm,inner sep=0.1cm},
node upper right/.style={font=\scriptsize,anchor=south west,text height=0.192cm,text depth=0.055cm,inner sep=0.03cm},
node right/.style={font=\scriptsize,anchor=west,text height=0.192cm,text depth=0.055cm,inner sep=0.03cm},
branch/.style={font=\tiny,text height=0.144cm,text depth=0.041cm,inner sep=0.025cm},
root/.style={font=\small,anchor=east,text height=0.240cm,text depth=0.068cm},
node lower right/.style={font=\scriptsize,anchor=north west,text height=0.192cm,text depth=0.055cm,inner sep=0.03cm}}
\begin{tikzpicture}[very thick,inner sep=0.1cm]
% 1:\hspace{11mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-tulane-green-wave.png}
% |
% |3:\hspace{11mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-georgia-tech-yellow-jacket.png}
% |
% | 13:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-auburn-tiger.png}
% |+12
% |114:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-lsu-tiger.jpg}
% ||
% 0105:\hspace{9mm}\includegraphics[width=8mm,resolution=150]{../images/mascot-missouri-tiger.png}
% ||
% |96:\hspace{11mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-sewanee-tiger.jpg}
% ||
% |17:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-kentucky-wild-cat.jpg}
% ||
% ||22:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-georgia-bulldog.jpg}
% 2821
% |203:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-mississippi-state-bulldog.png}
% ||
% |194:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-texas-am-reveille.png}
% ||
% |18:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-tennessee-smokey.png}
% ||
% |26:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-ole-miss-rebel-black-bear.jpg}
% |6
% |27:\hspace{9mm}\includegraphics[width=8mm,resolution=150]{../images/mascot-arkansas-razorback.png}
% 5|
% |28:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-vanderbilt-commodore-cartoon.png}
% |
% 429:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-alabama-big-al.png}
% |
% |31:\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-sc-gamecock.png}
% +30
%  32:\hspace{9mm}\includegraphics[width=8mm,resolution=150]{../images/mascot-florida-gator.png}

% The scale is 24.347458, and the yScale is 0.650000

%% Coordinates of nodes.
\draw [black!40,line width=10.0pt] (6.340,-0.500) -- (6.340,11.000);
\draw [black!40,line width=10.0pt] (13.400,-0.500) -- (13.400,11.000);
\draw [draw=none] (10.97, 9.75) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};
\draw [draw=none] (6.34, 10.4) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};
\draw [draw=none] (13.4, 7.15) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};
\coordinate (n0) at (0.000,7.947);
\coordinate (n1) at (6.330,10.400);
\coordinate (n1p) at (0.000,10.400);
\coordinate (n2) at (0.487,5.493);
\coordinate (n2p) at (0.000,5.493);
\coordinate (n3) at (10.956,9.750);
\coordinate (n3p) at (0.487,9.750);
\coordinate (n4) at (2.435,1.237);
\coordinate (n4p) at (0.487,1.237);
\coordinate (n5) at (2.800,2.148);
\coordinate (n5p) at (2.435,2.148);
\coordinate (n6) at (3.409,2.996);
\coordinate (n6p) at (2.800,2.996);
\coordinate (n7) at (4.504,4.042);
\coordinate (n7p) at (3.409,4.042);
\coordinate (n8) at (4.626,5.484);
\coordinate (n8p) at (4.504,5.484);
\coordinate (n9) at (6.330,7.109);
\coordinate (n9p) at (4.626,7.109);
\coordinate (n10) at (9.739,7.719);
\coordinate (n10p) at (6.330,7.719);
\coordinate (n11) at (12.661,8.287);
\coordinate (n11p) at (9.739,8.287);
\coordinate (n12) at (13.391,8.775);
\coordinate (n12p) at (12.661,8.775);
\coordinate (n13) at (14.365,9.100);
\coordinate (n13p) at (13.391,9.100);
\coordinate (n14) at (14.365,8.450);
\coordinate (n14p) at (13.391,8.450);
\coordinate (n15) at (14.365,7.800);
\coordinate (n15p) at (12.661,7.800);
\coordinate (n16) at (13.391,7.150);
\coordinate (n16p) at (9.739,7.150);
\coordinate (n17) at (14.365,6.500);
\coordinate (n17p) at (6.330,6.500);
\coordinate (n18) at (6.330,3.859);
\coordinate (n18p) at (4.626,3.859);
\coordinate (n19) at (11.443,4.469);
\coordinate (n19p) at (6.330,4.469);
\coordinate (n20) at (13.391,5.038);
\coordinate (n20p) at (11.443,5.038);
\coordinate (n21) at (14.122,5.525);
\coordinate (n21p) at (13.391,5.525);
\coordinate (n22) at (14.365,5.850);
\coordinate (n22p) at (14.122,5.850);
\coordinate (n23) at (14.365,5.200);
\coordinate (n23p) at (14.122,5.200);
\coordinate (n24) at (14.365,4.550);
\coordinate (n24p) at (13.391,4.550);
\coordinate (n25) at (14.365,3.900);
\coordinate (n25p) at (11.443,3.900);
\coordinate (n26) at (14.365,3.250);
\coordinate (n26p) at (6.330,3.250);
\coordinate (n27) at (14.365,2.600);
\coordinate (n27p) at (4.504,2.600);
\coordinate (n28) at (14.365,1.950);
\coordinate (n28p) at (3.409,1.950);
\coordinate (n29) at (14.365,1.300);
\coordinate (n29p) at (2.800,1.300);
\coordinate (n30) at (6.330,0.325);
\coordinate (n30p) at (2.435,0.325);
\coordinate (n31) at (14.365,0.650);
\coordinate (n31p) at (6.330,0.650);
\coordinate (n32) at (14.365,0.000);
\coordinate (n32p) at (6.330,0.000);

%% horizontal lines
\draw (n1p) -- (n1);
\draw (n2p) -- (n2);
\draw (n3p) -- (n3);
\draw (n4p) -- (n4);
\draw (n5p) -- (n5);
\draw (n6p) -- (n6);
\draw (n7p) -- (n7);
\draw (n8p) -- (n8);
\draw (n9p) -- (n9);
\draw (n10p) -- (n10);
\draw (n11p) -- (n11);
\draw (n12p) -- (n12);
\draw (n13p) -- (n13);
\draw (n14p) -- (n14);
\draw (n15p) -- (n15);
\draw (n16p) -- (n16);
\draw (n17p) -- (n17);
\draw (n18p) -- (n18);
\draw (n19p) -- (n19);
\draw (n20p) -- (n20);
\draw (n21p) -- (n21);
\draw (n22p) -- (n22);
\draw (n23p) -- (n23);
\draw (n24p) -- (n24);
\draw (n25p) -- (n25);
\draw (n26p) -- (n26);
\draw (n27p) -- (n27);
\draw (n28p) -- (n28);
\draw (n29p) -- (n29);
\draw (n30p) -- (n30);
\draw (n31p) -- (n31);
\draw (n32p) -- (n32);

%% vertical lines
\draw [line cap=rect] (n1p) -- (n2p);
\draw [line cap=rect] (n3p) -- (n4p);
\draw [line cap=rect] (n5p) -- (n30p);
\draw [line cap=rect] (n6p) -- (n29p);
\draw [line cap=rect] (n7p) -- (n28p);
\draw [line cap=rect] (n8p) -- (n27p);
\draw [line cap=rect] (n9p) -- (n18p);
\draw [line cap=rect] (n10p) -- (n17p);
\draw [line cap=rect] (n11p) -- (n16p);
\draw [line cap=rect] (n12p) -- (n15p);
\draw [line cap=rect] (n13p) -- (n14p);
\draw [line cap=rect] (n19p) -- (n26p);
\draw [line cap=rect] (n20p) -- (n25p);
\draw [line cap=rect] (n21p) -- (n24p);
\draw [line cap=rect] (n22p) -- (n23p);
\draw [line cap=rect] (n31p) -- (n32p);

%% leaf labels
\node at (n1) {\hspace{11mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-tulane-green-wave.png}};
\node at (n3) {\hspace{11mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-georgia-tech-yellow-jacket.png}};
\node at (n13) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-auburn-tiger.png}};
\node at (n14) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-lsu-tiger.jpg}};
\node at (n15) {\hspace{9mm}\includegraphics[width=8mm,resolution=150]{../images/mascot-missouri-tiger.png}};
\node at (n16) {\hspace{11mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-sewanee-tiger.jpg}};
\node at (n17) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-kentucky-wild-cat.jpg}};
\node at (n22) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-georgia-bulldog.jpg}};
\node at (n23) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-mississippi-state-bulldog.png}};
\node at (n24) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-texas-am-reveille.png}};
\node at (n25) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-tennessee-smokey.png}};
\node at (n26) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-ole-miss-rebel-black-bear.jpg}};
\node at (n27) {\hspace{9mm}\includegraphics[width=8mm,resolution=150]{../images/mascot-arkansas-razorback.png}};
\node at (n28) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-vanderbilt-commodore-cartoon.png}};
\node at (n29) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-alabama-big-al.png}};
\node at (n31) {\hspace{9mm}\includegraphics[height=6.5mm,resolution=150]{../images/mascot-sc-gamecock.png}};
\node at (n32) {\hspace{9mm}\includegraphics[width=8mm,resolution=150]{../images/mascot-florida-gator.png}};

\end{tikzpicture}

### 2015-03-04 slides/mascot-tree/mascot-tree.tikz.tex
Working on slides.

--- 2015-03-04 slides/mascot-tree/mascot-tree-shared-events.tikz.tex
@@
 %% Coordinates of nodes.
-\draw [black!40,line width=10.0pt] (6.340,-0.500) -- (6.340,11.000);
-\draw [black!40,line width=10.0pt] (13.400,-0.500) -- (13.400,11.000);
+\draw [black!00,line width=10.0pt] (6.340,-0.500) -- (6.340,11.000);
+\draw [black!00,line width=10.0pt] (13.400,-0.500) -- (13.400,11.000);
 \draw [draw=none] (10.97, 9.75) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};
-\draw [draw=none] (6.34, 10.4) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};
-\draw [draw=none] (13.4, 7.15) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};
+\draw [draw=none] (7.08, 10.4) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};
+\draw [draw=none] (13.17, 7.15) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};
 \coordinate (n0) at (0.000,7.947);
-\coordinate (n1) at (6.330,10.400);
+\coordinate (n1) at (7.061,10.400);
 \coordinate (n1p) at (0.000,10.400);
@@
 \coordinate (n8p) at (4.504,5.484);
-\coordinate (n9) at (6.330,7.109);
+\coordinate (n9) at (6.574,7.109);
 \coordinate (n9p) at (4.626,7.109);
 \coordinate (n10) at (9.739,7.719);
-\coordinate (n10p) at (6.330,7.719);
+\coordinate (n10p) at (6.574,7.719);
 \coordinate (n11) at (12.661,8.287);
 \coordinate (n11p) at (9.739,8.287);
-\coordinate (n12) at (13.391,8.775);
+\coordinate (n12) at (13.635,8.775);
 \coordinate (n12p) at (12.661,8.775);
 \coordinate (n13) at (14.365,9.100);
-\coordinate (n13p) at (13.391,9.100);
+\coordinate (n13p) at (13.635,9.100);
 \coordinate (n14) at (14.365,8.450);
-\coordinate (n14p) at (13.391,8.450);
+\coordinate (n14p) at (13.635,8.450);
 \coordinate (n15) at (14.365,7.800);
 \coordinate (n15p) at (12.661,7.800);
-\coordinate (n16) at (13.391,7.150);
+\coordinate (n16) at (13.148,7.150);
 \coordinate (n16p) at (9.739,7.150);
 \coordinate (n17) at (14.365,6.500);
-\coordinate (n17p) at (6.330,6.500);
-\coordinate (n18) at (6.330,3.859);
+\coordinate (n17p) at (6.574,6.500);
+\coordinate (n18) at (6.087,3.859);
 \coordinate (n18p) at (4.626,3.859);
 \coordinate (n19) at (11.443,4.469);
-\coordinate (n19p) at (6.330,4.469);
+\coordinate (n19p) at (6.087,4.469);
 \coordinate (n20) at (13.391,5.038);
@@
 \coordinate (n26) at (14.365,3.250);
-\coordinate (n26p) at (6.330,3.250);
+\coordinate (n26p) at (6.087,3.250);
 \coordinate (n27) at (14.365,2.600);

### 2015-03-04 slides/bin/plot-sec-tree-gram.py

@@ -63,7 +63,7 @@ def get_image_path(leaf_name):
     return os.path.join(os.path.pardir, 'images', leaf_name)
 
 def plot_tree(tree, base_name = 'tree', dir_name = 'gram', show_events = False):
-    read(_sec_tree_shared)
+    read(tree)
     print var.trees
     t = var.trees[-1]
     for n in t.iterNodes():
@@ -87,7 +87,7 @@ def plot_tree(tree, base_name = 'tree', dir_name = 'gram', show_events = False):
             node_name = n.name
             if node_name in _extinct_images:
                 lmargin += 2
-            n.name = r"\hspace{{{0}mm}}\includegraphics[{1}={2}mm,resolution=150]{{../{3}}}".format(
+            n.name = r"\hspace{{{0}mm}}\includegraphics[{1}={2}mm,resolution=150]{{{3}}}".format(
                     lmargin,
                     dimension_name,
                     dimension,
@@ -97,25 +97,36 @@ def plot_tree(tree, base_name = 'tree', dir_name = 'gram', show_events = False):
     tg.tgDefaultLineThickness = 'very thick'
     tg.baseName = base_name
     tg.dirName = dir_name
-    s = tg.code(r"""\draw [draw=none] (6.34, 10.4) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};""")
-    s = tg.code(r"""\draw [draw=none] (13.4, 7.15) node[circle, minimum width=10pt, draw=none, inner sep=0pt, path picture={\draw[red] (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);}] {};""")
+    l1 = tg.gramLine(6.34,-0.5,6.34,11)
+    # l1.lineStyle = 'loosely dashed'
+    # l1.lineThickness = 'ultra thick'
+    l1.lineThickness = 10.0
+    l2 = tg.gramLine(13.4,-0.5,13.4,11)
+    # l2.lineStyle = 'loosely dashed'
+    # l2.lineThickness = 'ultra thick'
+    l2.lineThickness = 10.0
+    l1.colour = 'black!40'
+    l2.colour = 'black!40'
+    if not show_events:
+        # print dir(l1)
+        l1.colour = 'black!00'
+        l2.colour = 'black!00'
[…]

### 2015-03-04 slides/auburn/abc.tex
Working on slides.

\begin{frame}
\newcommand{\matrixonebottom}{6.0}
\newcommand{\matrixtwobottom}{3.0}
\newcommand{\matrixthreebottom}{0.0}
\newcommand{\matrixspace}{0.285}
% \begin{picture}(300,200)(0,0)
    % \put(0,-34){
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2.2cm,on grid,auto]
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*9) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*8) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*7) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*6) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*5) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*4) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*3) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*2) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*1) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*0) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};

	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*9) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*8) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*7) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*6) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*5) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*4) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*3) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*2) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*1) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*0) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};

	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*9) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*8) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*7) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*6) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*5) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*4) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*3) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*2) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*1) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*0) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<2->] at (2.6,1.3) (m3) {\pgftext{\includegraphics[height=28mm]{../images/colormat1.pdf}}};
	\node [visible on=<2->] at (2.6,4.3) (m2) {\pgftext{\includegraphics[height=28mm]{../images/colormat2.pdf}}};
	\node [visible on=<2->] at (2.6,7.3) (m1) {\pgftext{\includegraphics[height=28mm]{../images/colormat3.pdf}}};

    \node[visible on=<3->] at (8, 7.3) (s1) {$S_1$};
    \node[visible on=<3->] at (8, 4.3) (s2) {$S_2$};
    \node[visible on=<3->] at (8, 1.3) (s3) {$S_3$};

    \path[->] (5.2,7.3) edge [visible on=<3->,very thick, bend left] (s1);
    \path[->] (5.2,4.3) edge [visible on=<3->,very thick, bend left] (s2);
    \path[->] (5.2,1.3) edge [visible on=<3->,very thick, bend left] (s3);

\end{tikzpicture}
\end{center}
    % }
% \end{picture}
\end{frame}

\begin{frame}[t]
    % \frametitle{Easy as ABC}
    \begin{center}
    \begin{tikzpicture}[shorten >=1pt,node distance=2.2cm,on grid,auto]
        \node[] at (0,0) {\includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}};
        \node[] at (0,0) {\includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}};
        \node[] at (0,0) {\includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}};
        \node[] at (0,0) {\includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}};
        \node[] at (0,0) {\includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}};
        \node[] at (0,0) {\includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}};
        \node[] at (0,0) {\includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}};
        \node[] at (0,0) {\includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}};
        \node[] at (0,0) {\includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}};
        \node[] at (0,0) {\includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}};
        \node[visible on=<1->] at (-2.9, -3.7) (s1) {$S_1$};
        \node[visible on=<1->] at (3.4, -2.9) (s2) {$S_2$};
        \node[visible on=<1->] at (5.3, 0.7) (s3) {$S_3$};
    \end{tikzpicture}
    \end{center}
    % \centerline{
    % \includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}
    % \includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}
    % \includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}
    % \includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}
    % \includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}
    % \includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}
    % \includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}
    % \includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}
    % \includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}
    % \includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}
    % }
\end{frame}

### 2015-03-04 slides/bin/reject-sampling-plot.py

@@ -16,6 +16,7 @@ def get_rejection_plot_3d(maximum = 1.0,
         observed = (0.5, 0.5, 0.5),
         nsamples = 0,
         show_rejected = True,
+        axis_labels = ('$S_1$', '$S_2$', '$S_3$'),
         rng = None):
     if not rng:
         rng = random.Random()
@@ -69,6 +70,9 @@ def get_rejection_plot_3d(maximum = 1.0,
                 zorder = 500,
                 )
     
+    ax.set_xlabel(axis_labels[0])
+    ax.set_ylabel(axis_labels[1])
+    ax.set_zlabel(axis_labels[2])
     return ax, fig
 
 def main_cli():

### 2015-03-05 slides/auburn/abc.tex
Working on slides.

@@
-\begin{frame}
-\newcommand{\matrixonebottom}{6.0}
-\newcommand{\matrixtwobottom}{3.0}
-\newcommand{\matrixthreebottom}{0.0}
-\newcommand{\matrixspace}{0.285}
-% \begin{picture}(300,200)(0,0)
-    % \put(0,-34){
+\tikzstyle{nAfill} = [fill=purple!25]
+\tikzstyle{nCfill} = [fill=green!25]
+\tikzstyle{nGfill} = [fill=orange!25]
+\tikzstyle{nTfill} = [fill=blue!25]
+\tikzstyle{nucA} = [draw=black, nAfill,
+    minimum width=2.3ex, minimum height=2.3ex, node distance=2.3ex]
+\tikzstyle{space} = [
+    minimum width=2.3ex, minimum height=2.3ex, node distance=2.3ex]
+\tikzstyle{lizard} = [
+    minimum width=0ex, minimum height=0ex, node distance=0ex]
+\tikzstyle{nucC} = [nucA, nCfill]
+\tikzstyle{nucG} = [nucA, nGfill]
+\tikzstyle{nucT} = [nucA, nTfill]
+
+\begin{frame}[t]
+    % \frametitle{Easy as ABC}
+    \begin{columns}
+        \column{0.2\textwidth}
+        \includegraphics<1->[width=1.0\columnwidth]{../images/dmc-cartoon-islands-crop.pdf}
+        \column{0.8\textwidth}
     \begin{center}
-\begin{tikzpicture}[shorten >=1pt,node distance=2.2cm,on grid,auto]
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*9) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*8) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*7) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*6) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*5) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*4) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*3) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*2) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*1) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*0) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
+    \begin{tikzpicture}[font=\sffamily]
 
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*9) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*8) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*7) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*6) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*5) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*4) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*3) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*2) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*1) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*0) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
+        \node[visible on=<2->, name=m1r01c01, nucA] {\nA};
+        \node[visible on=<2->, name=m1r02c01, nucA, below of=m1r01c01] {\nA};
+        \node[visible on=<2->, name=m1r03c01, nucA, below of=m1r02c01] {\nA};
+        \node[visible on=<2->, name=m1r04c01, nucG, below of=m1r03c01] {\nG};
+        \node[visible on=<2->, name=m1r05c01, nucG, below of=m1r04c01] {\nG};
+        \node[visible on=<2->, name=m1r06c01, nucG, below of=m1r05c01] {\nG};
 
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*9) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*8) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*7) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*6) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*5) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*4) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*3) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*2) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*1) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
-	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*0) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
-	\node [visible on=<2->] at (2.6,1.3) (m3) {\pgftext{\includegraphics[height=28mm]{../images/colormat1.pdf}}};
-	\node [visible on=<2->] at (2.6,4.3) (m2) {\pgftext{\includegraphics[height=28mm]{../images/colormat2.pdf}}};
-	\node [visible on=<2->] at (2.6,7.3) (m1) {\pgftext{\includegraphics[height=28mm]{../images/colormat3.pdf}}};
+        \node[visible on=<2->, name=m1r01c02, nucC, right of=m1r01c01] {\nC};
+        \node[visible on=<2->, name=m1r02c02, nucC, below of=m1r01c02] {\nC};
+        \node[visible on=<2->, name=m1r03c02, nucC, below of=m1r02c02] {\nC};
+        \node[visible on=<2->, name=m1r04c02, nucC, below of=m1r03c02] {\nC};
+        \node[visible on=<2->, name=m1r05c02, nucC, below of=m1r04c02] {\nC};
+        \node[visible on=<2->, name=m1r06c02, nucC, below of=m1r05c02] {\nC};
 
-    \node[visible on=<3->] at (8, 7.3) (s1) {$S_1$};
-    \node[visible on=<3->] at (8, 4.3) (s2) {$S_2$};
-    \node[visible on=<3->] at (8, 1.3) (s3) {$S_3$};
+        \node[visible on=<2->, name=m1r01c03, nucG, right of=m1r01c02] {\nG};
+        \node[visible on=<2->, name=m1r02c03, nucG, below of=m1r01c03] {\nG};
+        \node[visible on=<2->, name=m1r03c03, nucG, below of=m1r02c03] {\nG};
+        \node[visible on=<2->, name=m1r04c03, nucG, below of=m1r03c03] {\nG};
+        \node[visible on=<2->, name=m1r05c03, nucG, below of=m1r04c03] {\nG};
+        \node[visible on=<2->, name=m1r06c03, nucG, below of=m1r05c03] {\nG};
 
-    \path[->] (5.2,7.3) edge [visible on=<3->,very thick, bend left] (s1);
-    \path[->] (5.2,4.3) edge [visible on=<3->,very thick, bend left] (s2);
-    \path[->] (5.2,1.3) edge [visible on=<3->,very thick, bend left] (s3);
+        \node[visible on=<2->, name=m1r01c04, nucA, right of=m1r01c03] {\nA};
+        \node[visible on=<2->, name=m1r02c04, nucA, below of=m1r01c04] {\nA};
+        \node[visible on=<2->, name=m1r03c04, nucA, below of=m1r02c04] {\nA};
+        \node[visible on=<2->, name=m1r04c04, nucA, below of=m1r03c04] {\nA};
+        \node[visible on=<2->, name=m1r05c04, nucA, below of=m1r04c04] {\nA};
+        \node[visible on=<2->, name=m1r06c04, nucT, below of=m1r05c04] {\nT};
 
+        \node[visible on=<2->, name=m1r01c05, nucA, right of=m1r01c04] {\nA};
+        \node[visible on=<2->, name=m1r02c05, nucA, below of=m1r01c05] {\nA};
+        \node[visible on=<2->, name=m1r03c05, nucA, below of=m1r02c05] {\nA};
+        \node[visible on=<2->, name=m1r04c05, nucA, below of=m1r03c05] {\nA};
+        \node[visible on=<2->, name=m1r05c05, nucA, below of=m1r04c05] {\nA};
+        \node[visible on=<2->, name=m1r06c05, nucA, below of=m1r05c05] {\nA};
+
+        \node[visible on=<2->, name=m1r01c06, nucT, right of=m1r01c05] {\nT};
+        \node[visible on=<2->, name=m1r02c06, nucT, below of=m1r01c06] {\nT};
+        \node[visible on=<2->, name=m1r03c06, nucC, below of=m1r02c06] {\nC};
+        \node[visible on=<2->, name=m1r04c06, nucC, below of=m1r03c06] {\nC};
+        \node[visible on=<2->, name=m1r05c06, nucC, below of=m1r04c06] {\nC};
+        \node[visible on=<2->, name=m1r06c06, nucC, below of=m1r05c06] {\nC};
+
+        \node[visible on=<2->, name=m1r01c07, nucG, right of=m1r01c06] {\nG};
+        \node[visible on=<2->, name=m1r02c07, nucG, below of=m1r01c07] {\nG};
+        \node[visible on=<2->, name=m1r03c07, nucG, below of=m1r02c07] {\nG};
+        \node[visible on=<2->, name=m1r04c07, nucG, below of=m1r03c07] {\nG};
+        \node[visible on=<2->, name=m1r05c07, nucG, below of=m1r04c07] {\nG};
+        \node[visible on=<2->, name=m1r06c07, nucG, below of=m1r05c07] {\nG};
+
+        \node[visible on=<2->, name=m1r01c08, nucT, right of=m1r01c07] {\nT};
+        \node[visible on=<2->, name=m1r02c08, nucT, below of=m1r01c08] {\nT};
+        \node[visible on=<2->, name=m1r03c08, nucT, below of=m1r02c08] {\nT};
+        \node[visible on=<2->, name=m1r04c08, nucT, below of=m1r03c08] {\nT};
+        \node[visible on=<2->, name=m1r05c08, nucT, below of=m1r04c08] {\nT};
+        \node[visible on=<2->, name=m1r06c08, nucT, below of=m1r05c08] {\nT};
+
+        \node[visible on=<2->, name=m1r01c09, nucG, right of=m1r01c08] {\nG};
+        \node[visible on=<2->, name=m1r02c09, nucG, below of=m1r01c09] {\nG};
+        \node[visible on=<2->, name=m1r03c09, nucG, below of=m1r02c09] {\nG};
+        \node[visible on=<2->, name=m1r04c09, nucG, below of=m1r03c09] {\nG};
+        \node[visible on=<2->, name=m1r05c09, nucG, below of=m1r04c09] {\nG};
+        \node[visible on=<2->, name=m1r06c09, nucG, below of=m1r05c09] {\nG};
+
+        \node[visible on=<2->, name=m1r01c10, nucC, right of=m1r01c09] {\nC};
+        \node[visible on=<2->, name=m1r02c10, nucC, below of=m1r01c10] {\nC};
+        \node[visible on=<2->, name=m1r03c10, nucC, below of=m1r02c10] {\nC};
+        \node[visible on=<2->, name=m1r04c10, nucT, below of=m1r03c10] {\nT};
+        \node[visible on=<2->, name=m1r05c10, nucT, below of=m1r04c10] {\nT};
+        \node[visible on=<2->, name=m1r06c10, nucT, below of=m1r05c10] {\nT};
+
+        \node[visible on=<2->, name=vspace1, space, below of=m1r06c01] {};
+
+        \node[visible on=<2->, name=m2r01c01, nucT, below of=vspace1] {\nT};
+        \node[visible on=<2->, name=m2r02c01, nucT, below of=m2r01c01] {\nT};
+        \node[visible on=<2->, name=m2r03c01, nucT, below of=m2r02c01] {\nT};
+        \node[visible on=<2->, name=m2r04c01, nucC, below of=m2r03c01] {\nC};
+        \node[visible on=<2->, name=m2r05c01, nucC, below of=m2r04c01] {\nC};
+        \node[visible on=<2->, name=m2r06c01, nucC, below of=m2r05c01] {\nC};
+
+        \node[visible on=<2->, name=m2r01c02, nucC, right of=m2r01c01] {\nC};
+        \node[visible on=<2->, name=m2r02c02, nucC, below of=m2r01c02] {\nC};
+        \node[visible on=<2->, name=m2r03c02, nucC, below of=m2r02c02] {\nC};
+        \node[visible on=<2->, name=m2r04c02, nucC, below of=m2r03c02] {\nC};
+        \node[visible on=<2->, name=m2r05c02, nucC, below of=m2r04c02] {\nC};
+        \node[visible on=<2->, name=m2r06c02, nucC, below of=m2r05c02] {\nC};
+
+        \node[visible on=<2->, name=m2r01c03, nucG, right of=m2r01c02] {\nG};
+        \node[visible on=<2->, name=m2r02c03, nucG, below of=m2r01c03] {\nG};
+        \node[visible on=<2->, name=m2r03c03, nucG, below of=m2r02c03] {\nG};
+        \node[visible on=<2->, name=m2r04c03, nucG, below of=m2r03c03] {\nG};
+        \node[visible on=<2->, name=m2r05c03, nucG, below of=m2r04c03] {\nG};
+        \node[visible on=<2->, name=m2r06c03, nucG, below of=m2r05c03] {\nG};
+
+        \node[visible on=<2->, name=m2r01c04, nucC, right of=m2r01c03] {\nC};
+        \node[visible on=<2->, name=m2r02c04, nucC, below of=m2r01c04] {\nC};
+        \node[visible on=<2->, name=m2r03c04, nucT, below of=m2r02c04] {\nT};
+        \node[visible on=<2->, name=m2r04c04, nucT, below of=m2r03c04] {\nT};
+        \node[visible on=<2->, name=m2r05c04, nucT, below of=m2r04c04] {\nT};
+        \node[visible on=<2->, name=m2r06c04, nucT, below of=m2r05c04] {\nT};
+
+        \node[visible on=<2->, name=m2r01c05, nucA, right of=m2r01c04] {\nA};
+        \node[visible on=<2->, name=m2r02c05, nucA, below of=m2r01c05] {\nA};
+        \node[visible on=<2->, name=m2r03c05, nucA, below of=m2r02c05] {\nA};
+        \node[visible on=<2->, name=m2r04c05, nucA, below of=m2r03c05] {\nA};
+        \node[visible on=<2->, name=m2r05c05, nucA, below of=m2r04c05] {\nA};
+        \node[visible on=<2->, name=m2r06c05, nucA, below of=m2r05c05] {\nA};
+
+        \node[visible on=<2->, name=m2r01c06, nucC, right of=m2r01c05] {\nC};
+        \node[visible on=<2->, name=m2r02c06, nucC, below of=m2r01c06] {\nC};
+        \node[visible on=<2->, name=m2r03c06, nucC, below of=m2r02c06] {\nC};
+        \node[visible on=<2->, name=m2r04c06, nucC, below of=m2r03c06] {\nC};
+        \node[visible on=<2->, name=m2r05c06, nucC, below of=m2r04c06] {\nC};
+        \node[visible on=<2->, name=m2r06c06, nucC, below of=m2r05c06] {\nC};
+
+        \node[visible on=<2->, name=m2r01c07, nucG, right of=m2r01c06] {\nG};
+        \node[visible on=<2->, name=m2r02c07, nucG, below of=m2r01c07] {\nG};
+        \node[visible on=<2->, name=m2r03c07, nucG, below of=m2r02c07] {\nG};
+        \node[visible on=<2->, name=m2r04c07, nucG, below of=m2r03c07] {\nG};
+        \node[visible on=<2->, name=m2r05c07, nucG, below of=m2r04c07] {\nG};
+        \node[visible on=<2->, name=m2r06c07, nucG, below of=m2r05c07] {\nG};
+
+        \node[visible on=<2->, name=m2r01c08, nucT, right of=m2r01c07] {\nT};
+        \node[visible on=<2->, name=m2r02c08, nucT, below of=m2r01c08] {\nT};
+        \node[visible on=<2->, name=m2r03c08, nucT, below of=m2r02c08] {\nT};
+        \node[visible on=<2->, name=m2r04c08, nucT, below of=m2r03c08] {\nT};
+        \node[visible on=<2->, name=m2r05c08, nucT, below of=m2r04c08] {\nT};
+        \node[visible on=<2->, name=m2r06c08, nucT, below of=m2r05c08] {\nT};
+
+        \node[visible on=<2->, name=m2r01c09, nucA, right of=m2r01c08] {\nA};
+        \node[visible on=<2->, name=m2r02c09, nucA, below of=m2r01c09] {\nA};
+        \node[visible on=<2->, name=m2r03c09, nucA, below of=m2r02c09] {\nA};
+        \node[visible on=<2->, name=m2r04c09, nucG, below of=m2r03c09] {\nG};
+        \node[visible on=<2->, name=m2r05c09, nucG, below of=m2r04c09] {\nG};
+        \node[visible on=<2->, name=m2r06c09, nucG, below of=m2r05c09] {\nG};
+
+        \node[visible on=<2->, name=m2r01c10, nucC, right of=m2r01c09] {\nC};
+        \node[visible on=<2->, name=m2r02c10, nucC, below of=m2r01c10] {\nC};
+        \node[visible on=<2->, name=m2r03c10, nucC, below of=m2r02c10] {\nC};
+        \node[visible on=<2->, name=m2r04c10, nucC, below of=m2r03c10] {\nC};
+        \node[visible on=<2->, name=m2r05c10, nucC, below of=m2r04c10] {\nC};
+        \node[visible on=<2->, name=m2r06c10, nucC, below of=m2r05c10] {\nC};
+
+        \node[visible on=<2->, name=vspace2, space, below of=m2r06c01] {};
+
+        \node[visible on=<2->, name=m3r01c01, nucC, below of=vspace2] {\nC};
+        \node[visible on=<2->, name=m3r02c01, nucC, below of=m3r01c01] {\nC};
+        \node[visible on=<2->, name=m3r03c01, nucC, below of=m3r02c01] {\nC};
+        \node[visible on=<2->, name=m3r04c01, nucC, below of=m3r03c01] {\nC};
+        \node[visible on=<2->, name=m3r05c01, nucC, below of=m3r04c01] {\nC};
+        \node[visible on=<2->, name=m3r06c01, nucC, below of=m3r05c01] {\nC};
+
+        \node[visible on=<2->, name=m3r01c02, nucA, right of=m3r01c01] {\nA};
+        \node[visible on=<2->, name=m3r02c02, nucA, below of=m3r01c02] {\nA};
+        \node[visible on=<2->, name=m3r03c02, nucA, below of=m3r02c02] {\nA};
+        \node[visible on=<2->, name=m3r04c02, nucT, below of=m3r03c02] {\nT};
+        \node[visible on=<2->, name=m3r05c02, nucT, below of=m3r04c02] {\nT};
+        \node[visible on=<2->, name=m3r06c02, nucT, below of=m3r05c02] {\nT};
+
+        \node[visible on=<2->, name=m3r01c03, nucG, right of=m3r01c02] {\nG};
+        \node[visible on=<2->, name=m3r02c03, nucG, below of=m3r01c03] {\nG};
+        \node[visible on=<2->, name=m3r03c03, nucG, below of=m3r02c03] {\nG};
+        \node[visible on=<2->, name=m3r04c03, nucG, below of=m3r03c03] {\nG};
+        \node[visible on=<2->, name=m3r05c03, nucG, below of=m3r04c03] {\nG};
+        \node[visible on=<2->, name=m3r06c03, nucG, below of=m3r05c03] {\nG};
+
+        \node[visible on=<2->, name=m3r01c04, nucT, right of=m3r01c03] {\nT};
+        \node[visible on=<2->, name=m3r02c04, nucT, below of=m3r01c04] {\nT};
+        \node[visible on=<2->, name=m3r03c04, nucT, below of=m3r02c04] {\nT};
+        \node[visible on=<2->, name=m3r04c04, nucT, below of=m3r03c04] {\nT};
+        \node[visible on=<2->, name=m3r05c04, nucC, below of=m3r04c04] {\nC};
+        \node[visible on=<2->, name=m3r06c04, nucC, below of=m3r05c04] {\nC};
+
+        \node[visible on=<2->, name=m3r01c05, nucA, right of=m3r01c04] {\nA};
+        \node[visible on=<2->, name=m3r02c05, nucA, below of=m3r01c05] {\nA};
+        \node[visible on=<2->, name=m3r03c05, nucA, below of=m3r02c05] {\nA};
+        \node[visible on=<2->, name=m3r04c05, nucA, below of=m3r03c05] {\nA};
+        \node[visible on=<2->, name=m3r05c05, nucA, below of=m3r04c05] {\nA};
+        \node[visible on=<2->, name=m3r06c05, nucA, below of=m3r05c05] {\nA};
+
+        \node[visible on=<2->, name=m3r01c06, nucC, right of=m3r01c05] {\nC};
+        \node[visible on=<2->, name=m3r02c06, nucC, below of=m3r01c06] {\nC};
+        \node[visible on=<2->, name=m3r03c06, nucC, below of=m3r02c06] {\nC};
+        \node[visible on=<2->, name=m3r04c06, nucC, below of=m3r03c06] {\nC};
+        \node[visible on=<2->, name=m3r05c06, nucC, below of=m3r04c06] {\nC};
+        \node[visible on=<2->, name=m3r06c06, nucC, below of=m3r05c06] {\nC};
+
+        \node[visible on=<2->, name=m3r01c07, nucC, right of=m3r01c06] {\nC};
+        \node[visible on=<2->, name=m3r02c07, nucC, below of=m3r01c07] {\nC};
+        \node[visible on=<2->, name=m3r03c07, nucC, below of=m3r02c07] {\nC};
+        \node[visible on=<2->, name=m3r04c07, nucT, below of=m3r03c07] {\nT};
+        \node[visible on=<2->, name=m3r05c07, nucT, below of=m3r04c07] {\nT};
+        \node[visible on=<2->, name=m3r06c07, nucT, below of=m3r05c07] {\nT};
+
+        \node[visible on=<2->, name=m3r01c08, nucG, right of=m3r01c07] {\nG};
+        \node[visible on=<2->, name=m3r02c08, nucG, below of=m3r01c08] {\nG};
+        \node[visible on=<2->, name=m3r03c08, nucG, below of=m3r02c08] {\nG};
+        \node[visible on=<2->, name=m3r04c08, nucG, below of=m3r03c08] {\nG};
+        \node[visible on=<2->, name=m3r05c08, nucG, below of=m3r04c08] {\nG};
+        \node[visible on=<2->, name=m3r06c08, nucG, below of=m3r05c08] {\nG};
+
+        \node[visible on=<2->, name=m3r01c09, nucG, right of=m3r01c08] {\nG};
+        \node[visible on=<2->, name=m3r02c09, nucG, below of=m3r01c09] {\nG};
+        \node[visible on=<2->, name=m3r03c09, nucG, below of=m3r02c09] {\nG};
+        \node[visible on=<2->, name=m3r04c09, nucG, below of=m3r03c09] {\nG};
+        \node[visible on=<2->, name=m3r05c09, nucG, below of=m3r04c09] {\nG};
+        \node[visible on=<2->, name=m3r06c09, nucG, below of=m3r05c09] {\nG};
+
+        \node[visible on=<2->, name=m3r01c10, nucT, right of=m3r01c09] {\nT};
+        \node[visible on=<2->, name=m3r02c10, nucT, below of=m3r01c10] {\nT};
+        \node[visible on=<2->, name=m3r03c10, nucT, below of=m3r02c10] {\nT};
+        \node[visible on=<2->, name=m3r04c10, nucT, below of=m3r03c10] {\nT};
+        \node[visible on=<2->, name=m3r05c10, nucT, below of=m3r04c10] {\nT};
+        \node[visible on=<2->, name=m3r06c10, nucT, below of=m3r05c10] {\nT};
+
+	    \node [visible on=<1->, left of=m1r01c01] (s1l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
+	    \node [visible on=<1->, left of=m1r02c01] (s1l2a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
+	    \node [visible on=<1->, left of=m1r03c01] (s1l3a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
+	    \node [visible on=<1->, left of=m1r04c01] (s1l1b) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
+	    \node [visible on=<1->, left of=m1r05c01] (s1l2b) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
+	    \node [visible on=<1->, left of=m1r06c01] (s1l3b) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
+
+	    \node [visible on=<1->, left of=m2r01c01] (s2l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
+	    \node [visible on=<1->, left of=m2r02c01] (s2l2a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
+	    \node [visible on=<1->, left of=m2r03c01] (s2l3a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
+	    \node [visible on=<1->, left of=m2r04c01] (s2l1b) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
+	    \node [visible on=<1->, left of=m2r05c01] (s2l2b) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
+	    \node [visible on=<1->, left of=m2r06c01] (s2l3b) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
+
+	    \node [visible on=<1->, left of=m3r01c01] (s3l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
+	    \node [visible on=<1->, left of=m3r02c01] (s3l2a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
+	    \node [visible on=<1->, left of=m3r03c01] (s3l3a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
+	    \node [visible on=<1->, left of=m3r04c01] (s3l1b) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
+	    \node [visible on=<1->, left of=m3r05c01] (s3l2b) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
+	    \node [visible on=<1->, left of=m3r06c01] (s3l3b) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
+
+        \node[visible on=<3->, name=m1hspace1, space, right of=m1r03c10] {};
+        \node[visible on=<3->, name=m1hspace2, space, right of=m1hspace1] {};
+        \node[visible on=<3->, name=m1hspace3, space, right of=m1hspace2] {};
+        \node[visible on=<3->, name=m1hspace4, space, right of=m1hspace3] {};
+        \node[visible on=<3->, name=m1hspace5, space, right of=m1hspace4] {};
+        \node[visible on=<3->, name=s1, right of=m1hspace5, visible on=<1->] {$S_1$};
+
+        \node[visible on=<3->, name=m2hspace1, space, right of=m2r03c10] {};
+        \node[visible on=<3->, name=m2hspace2, space, right of=m2hspace1] {};
+        \node[visible on=<3->, name=m2hspace3, space, right of=m2hspace2] {};
+        \node[visible on=<3->, name=m2hspace4, space, right of=m2hspace3] {};
+        \node[visible on=<3->, name=m2hspace5, space, right of=m2hspace4] {};
+        \node[visible on=<3->, name=s2, right of=m2hspace5, visible on=<1->] {$S_2$};
+
+        \node[visible on=<3->, name=m3hspace1, space, right of=m3r03c10] {};
+        \node[visible on=<3->, name=m3hspace2, space, right of=m3hspace1] {};
+        \node[visible on=<3->, name=m3hspace3, space, right of=m3hspace2] {};
+        \node[visible on=<3->, name=m3hspace4, space, right of=m3hspace3] {};
+        \node[visible on=<3->, name=m3hspace5, space, right of=m3hspace4] {};
+        \node[visible on=<3->, name=s3, right of=m3hspace5, visible on=<1->] {$S_3$};
+
+        \path[->] (m1hspace1) edge [visible on=<3->,very thick, bend left] (s1);
+        \path[->] (m2hspace1) edge [visible on=<3->,very thick, bend left] (s2);
+        \path[->] (m3hspace1) edge [visible on=<3->,very thick, bend left] (s3);
     \end{tikzpicture}
     \end{center}
-    % }
-% \end{picture}
+    \end{columns}
 \end{frame}
@@
         \node[] at (0,0) {\includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}};
-        \node[] at (0,0) {\includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}};
-        \node[] at (0,0) {\includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}};
+        % \node[] at (0,0) {\includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}};
+        \node[] at (0,0) {\includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}};
         \node[visible on=<1->] at (-2.9, -3.7) (s1) {$S_1$};

### 2015-03-05 slides/auburn/dpp-intro.tex
Working on slides.

\begin{frame}[t]
    \frametitle{Sampling divergence models---a novel approach}
    \begin{itemize}[<+->]
        \item The divergence models are ways of assigning our taxa to events
        \item A Dirichlet process prior (DPP) model is a ``natural'' fit
            \begin{itemize}[<+->]
                \item Common nonparametric Bayesian approach to assigning
                    variables to an unknown number of categories
            \end{itemize}
    \end{itemize}

    \begin{uncoverenv}<2->
    \begin{center}
        \includegraphics[height=0.4\textheight]{../images/peter-dirichlet.jpg}
    \end{center}
    \end{uncoverenv}
\end{frame}

### 2015-03-05 slides/auburn/full-likelihood.tex
Working on slides.

--- 2015-02-18 slides/bordeaux/slides.tex
@@
-\input{../utils/preamble.tex}
-\input{../../manuscripts/utils/macros.tex}
-\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
-\bibliography{../../manuscripts/bib/references}
-
-% \title[Estimating shared history]{An Improved Approximate-Bayesian Method for Estimating Shared
-% Evolutionary History}
-
-\title[Inferring community history]{Inferring Community-scale Diversification
-    from Genomic Data}
-
-\author[J.\ Oaks]{
-    Jamie R.\ Oaks\inst{1}
-}
-\institute[University of Washington]{
-    \inst{1}%
-        Department of Biology, University of Washington
-}
-
-% \date{\today}
-\date{February 18, 2015}
-
-\begin{document}
-\maketitle
-
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
-        \begin{itemize}
-            \item<2-> Community-scale processes are important in
-                biodiversification and community assembly
-            \item<3-> Accounting for such processes will better our
-                understanding of evolutionary history and
-                community structure
-            \item<4-> We need methods for inferring evolutionary patterns
-                predicted by historical events from contemporary populations
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}
-
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-
-        \vspace{6.5cm}
-
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \begin{uncoverenv}<2->
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-        \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1>
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rob.jpg}}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/charles.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/cam.jpg}}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jake.jpg}}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/allie.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/Luke.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/anthony.png}}
-        \end{minipage}
-        \end{onlyenv}
-
-        \begin{onlyenv}<2->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1-2>[width=\textwidth]{../images/maps/Philippines.png}
-        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        % \begin{onlyenv}<4->
-        %     \begin{itemize}
-        %         \item Strong support for simultaneous divergence of all 22 taxon pairs
-
-        %         \item Posterior probability $>$ 0.96 \\
-
-        %         \item $\sim$100,000--250,000 years ago
-        %     \end{itemize}
-        % \end{onlyenv}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle<1-4>{Climate-driven diversification: Predictions}
-    \frametitle<5->{Divergence-model choice}
-
-    \centering{
-    \includegraphics<1>[height=7.5cm]{../images/dmc-cartoon-islands.pdf}
-    \includegraphics<2>[height=7.5cm]{../images/dmc-cartoon-taxa.pdf}
-    \includegraphics<3>[height=7.5cm]{../images/dmc-cartoon-sea-level.pdf}
-    \includegraphics<4>[height=7.5cm]{../images/dmc-cartoon-shared-no-labels.pdf}
-    \includegraphics<5>[height=7.5cm]{../images/dmc-cartoon-shared.pdf}
-    \includegraphics<6>[height=7.5cm]{../images/dmc-cartoon-2-1.pdf}
-    \includegraphics<7>[height=7.5cm]{../images/dmc-cartoon-2-2.pdf}
-    \includegraphics<8>[height=7.5cm]{../images/dmc-cartoon-2-3.pdf}
-    \includegraphics<9>[height=7.5cm]{../images/dmc-cartoon-general.pdf}
-    }
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Divergence-model choice}
-
-    \vspace{-7mm}
-
-    \begin{minipage}[t][0.1\textheight][c]{1.1\linewidth}
-        \begin{adjustwidth}{-0.5em}{}
-            \begin{tabular}{ p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} }
-                $m_1$ & $m_2$ & $m_3$ & $m_4$ & $m_5$ \\
-            \end{tabular}
-        \end{adjustwidth}
-    \end{minipage}
-
-    \vspace{-2mm}
-
-    \begin{minipage}[t][0.35\textheight][c]{1.1\linewidth}
-        \begin{adjustwidth}{-2.3em}{}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-shared.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-1.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-2.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-3.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-general.pdf}
-        \end{adjustwidth}
-    \end{minipage}
-
-    \vspace{3.5mm}
-
-    \begin{minipage}[c][0.35\textheight][c]{\linewidth}
-        \begin{uncoverenv}<2->
-            \begin{center}
-            We want to infer \textcolor{blue}{\divModel{}} and
-            \textcolor{blue}{\divTimeMapVector} given DNA sequence
-            alignments
-            \textcolor{blue}{\alignmentVector}
-        \end{center}
-        \end{uncoverenv}
-
-        \vspace{4mm}
-
-        \begin{uncoverenv}<3->
-            \begin{displaybox}[0.85\linewidth]
-                \begin{minipage}[c][0.12\textheight][c]{\linewidth}
-
-                \[
-                    p(\divTimeMapVector,
-                      \divModel{i},
-                      \allParameters{}
-                      \given \alignmentVector)
-                      =
-                    \frac{
-                        p(\alignmentVector \given
-                          \divTimeMapVector,
-                          \allParameters{},
-                          \divModel{i})
-                        p(\divTimeMapVector,
-                          \allParameters{}
-                          \given \divModel{i})
-                        p(\divModel{i})
-                        }{p(\alignmentVector)}
-                \]
-                \vspace{-1mm}
-                \end{minipage}
-            \end{displaybox}
-        \end{uncoverenv}
-    \end{minipage}
-\end{frame}
-
-\begin{frame}
-    \frametitle{New method: \dppmsbayes}
-    \begin{itemize}
-        % \item<1-> Reparameterized the model implemented in \msb
-        \item<1-> Dirichlet-process prior (DPP) over all possible divergence
-            models
-        \item<2-> Flexible priors on continuous parameters to avoid strongly
-            weighted posteriors
-        \item<3-> Multi-processing interface to improve performance for
-            accommodating genomic datasets
-        % \item<4-> Uniform prior over divergence models
-    \end{itemize}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Performance}
-    \begin{itemize}
-        \item<1-> Extensive computer simulations to compare performance to previous
-            state-of-the-art methods (\msb).
-        \item<2-> New method for estimating shared evolutionary history shows
-            improved:
-            \begin{enumerate}
-                \item<2-> Improved model-choice accuracy 
-                \item<2-> More robustness to model violations
-                \item<2-> Greater power to detect variation in divergence times
-                \item<2-> Faster!
-            \end{enumerate}
-    \end{itemize}
-
-    \uncover<3->{
-    \textbf{\dppmsbayes allows biologists to leverage comparative genomic data
-        to assess the affects of community-scale processes on biodiversity}
-    }
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-
-        \vspace{6.5cm}
-
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \begin{uncoverenv}<1->
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-        \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1->[width=\textwidth]{../images/maps/Philippines.png}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
-    \smallskip
-    \centerline{
-    \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Current work: More data}
-    \begin{itemize}
-        \item<1-> Collecting genomic data from taxa co-distributed across Southeast
-            Asia
-        \item<2-> Preliminary results for 1000 loci from 5 pairs of \emph{Gekko
-                mindorensis} populations
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<2->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
-    \end{center}
-\end{frame}
-
 \begin{frame}
@@
 \end{frame}
-
-\section{Conclusions}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification model?}
-    \begin{itemize}
-        \item<1-> Results consistent with prediction of clustered
-            divergences
-        \item<2-> Results suggest multiple co-divergences
-        \item<3-> However, there is a lot of uncertainty!
-    \end{itemize}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\section{Teaching}
-
-\begin{frame}
-    BIO 180 stuff
-\end{frame}
-
-\begin{frame}
-    Software carpentry \\
-
-    Workshops
-\end{frame}
-
-\begin{frame}
-    \frametitle{Questions?}    
-    \begin{center}
-        {
-        \Large
-        \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
-        }
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Future directions---Computational}
-    \begin{itemize}
-        \item<1-> Full-phylogenetic framework
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
-        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \centerline{
-    \includegraphics<1>[height=8cm]{../images/bd-cartoon-no-lines.pdf}
-    \includegraphics<2>[height=8cm]{../images/bd-cartoon-shared.pdf}
-    }
-\end{frame}
-
-\begin{frame}
-\begin{columns}
-    \column{.5\textwidth}
-    \begin{minipage}[c]{\columnwidth}
-    \begin{itemize}
-        \item Blah
-    \end{itemize}
-    \end{minipage}
-
-    \column{.6\textwidth}
-    \begin{minipage}[t][\textheight][c]{\linewidth}
-    \includegraphics<1>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
-    \includegraphics<2>[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
-    \end{minipage}
-\end{columns}
-\end{frame}
-
-\end{document}

### 2015-03-05 slides/auburn/full-phylo.tex
Working on slides.

--- 2015-02-18 slides/bordeaux/slides-2.tex
@@
-\input{../utils/preamble.tex}
-\input{../../manuscripts/utils/macros.tex}
-\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
-\bibliography{../../manuscripts/bib/references}
-
-% \title[Estimating shared history]{An Improved Approximate-Bayesian Method for Estimating Shared
-% Evolutionary History}
-
-\title[Future directions]{Future Directions}
-
-\author[J.\ Oaks]{
-    Jamie R.\ Oaks\inst{1}
-}
-\institute[University of Washington]{
-    \inst{1}%
-        Department of Biology, University of Washington
-}
-
-% \date{\today}
-\date{February 18, 2015}
-
-\begin{document}
-
 \begin{frame}[t]
@@
 \end{frame}
-
-\begin{frame}
-    \frametitle{Inferring the evolutionary context of ecosystems}
-    \begin{itemize}[<+->]
-        \item Use comparative framework to estimate the history of floral and
-            faunal communities across Europe
-        \item A better understanding of the diversification and assembly
-            of communities will provide a framework for understanding
-            the affect of past climate change
-        \item This will provide context for predicting future change
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \begin{itemize}[<+->]
-        \item Generalized birth-death process might better model
-            epidemiological dynamics
-        \item Could be relevant to the evolution of grapevine pathogens
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Strategies for integration}
-    \begin{itemize}[<+->]
-        \item In order to predict the response of ecosystems to environmental
-            change, we need a general model of ecosystem processes
-        \item Organize a series ``hackathons'' to stimulate synthesis among the
-            biologists, engineers, and theorists across LabEx 
-        \item Each hackathon would focus on integrating processes and existing
-            models between particular specialties and/or ecosystems
-    \end{itemize}
-\end{frame}
-
-\end{document}

### 2015-03-05 slides/auburn/future-empirical.tex
Working on slides.

--- 2015-01-29 slides/osu/slides.tex
@@
-\input{../utils/preamble.tex}
-\input{../../manuscripts/utils/macros.tex}
-\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
-\bibliography{../../manuscripts/bib/references}
-
-\title[Improving inference of evolutionary history]{Improving inference of evolutionary history}
-\subtitle{Developing comparative methods for genomic data}
-
-\author[J.\ Oaks]{
-    Jamie R.\ Oaks\inst{1}
-}
-\institute[University of Washington]{
-    \inst{1}%
-        Department of Biology, University of Washington
-}
-
-% \date{\today}
-\date{November 3, 2014}
-
-\begin{document}
-% \maketitle
-\begin{frame}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-            \maketitle
-        \column{.5\textwidth}
-            \begin{figure}
-                \begin{center}
-                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
-                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
-                \end{center}
-            \end{figure}
-    \end{columns}
-\end{frame}
-
-% I. What I do
-% II. What I've been focusing on recently (started in PhD and continuing in postdoc)
-%   a. Dynamic planet---need to account for large-scale processes
-%       i. E.g., SE Asia---dynamic landscape
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
-        \begin{itemize}
-            \item<2-> Large-scale geological and climatic processes are
-                important in biodiversification and community assembly
-            \item<3-> Accounting for such processes will better our
-                understanding of biodiversity
-            \item<4-> We need methods for inferring evolutionary patterns
-                predicted by historical events from contemporary populations
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    \frametitle{Climate-driven diversification}    
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-        \ \\
-
-        \column{0.45\textwidth}
-
-        \vspace{-2.4cm}
-
-        \begin{uncoverenv}<2->
-        \colorbox{white}{
-            \begin{minipage}[t][3.6cm][t]{1.1\textwidth}
-                % \begin{onlyenv}<2-3>
-                % \begin{itemize}
-                %     \item<2-> Repeated coalescence and fragmentation of island complexes
-                %     \item<3-> Prominent paradigm for explaining biodiversity
-                % \end{itemize}
-                % \end{onlyenv}
-
-                \begin{onlyenv}<2>
-                \begin{itemize}
-                    \item \textbf{Did repeated fragmentation of islands
-                        during inter-glacial rises in sea level promote
-                        diversification?}
-                \end{itemize}
-                \end{onlyenv}
-
-                \begin{onlyenv}<3->
-                \begin{itemize}
-                    \item<3-> Climate-driven model of diversification
-                    \item<4-> Testable prediction:
-                    \begin{itemize}
-                        \item Temporally clustered divergences among taxa
-                            co-distributed across fragmented islands
-                    \end{itemize}
-                \end{itemize}
-                \end{onlyenv}
-            \end{minipage}
-        }
-    \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-% \begin{frame}
-%     \frametitle{Philippine Archipelago}
-%     \begin{center}
-%         \includegraphics<1>[width=.5\textwidth]{../images/maps/Philippines-present.png}
-%         \includegraphics<2>[width=.5\textwidth]{../images/maps/Philippines.png}
-%     \end{center}
-% \end{frame}
-
-% \begin{frame}
-%     \frametitle{Climate-driven diversification model}
-%     \begin{columns}[c]
-%         \column{.5\textwidth}
-%             \begin{myitemize}
-%                 \item<1-> Repeated coalescence and fragmentation of island complexes
-%                 \item<2-> Prominent paradigm for explaining Philippine biodiversity
-%                 \item<3-> Proposed as model of diversification
-%             \end{myitemize}
-%         \column{.5\textwidth}
-%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
-%     \end{columns}
-% \end{frame}
-
-% \begin{frame}
-%     \frametitle{Testing climate-driven diversification}
-%     \begin{columns}[c]
-%         \column{.5\textwidth}
-%     \textbf{Did repeated fragmentation of islands during inter-glacial rises in sea level
-%     promote diversification?}\\
-%     \bigskip
-%     \uncover<2->{
-%     Model has testable prediction:\\
-%     \begin{itemize}
-%         \item Temporally clustered divergences among taxa
-%             co-distributed across fragmented islands
-%     \end{itemize}
-%     }
-%         \column{.5\textwidth}
-%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
-%     \end{columns}
-% \end{frame}
-
-%   b. Predictions of climate-driven diversification
-\begin{frame}
-    \frametitle<1-4>{Climate-driven model: Predictions}
-    \frametitle<5->{Divergence model choice}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{onlyenv}<1-4>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<5>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{\divTime{1}\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 111
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (\divTimeMap{1},\divTimeMap{2},\divTimeMap{3})
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<6>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 111
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (130,130,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<7>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 211
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (397,130,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<8>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 121
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (130,397,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<9>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 112
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<10-12>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397,260\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 312
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (260,130,397)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        % \begin{onlyenv}<11>
-        %     \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-        %         \begin{displaybox}[0.95\linewidth]
-        %             \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-        %                 \[
-        %                     \divTimeVector = \{\divTime{1},\ldots,\divTime{\divTimeNum}\}
-        %                 \]\vspace{0mm}
-        %                 \[
-        %                     model = \divModel{i}
-        %                 \]\vspace{0mm}
-        %                 \[
-        %                     \divTimeMapVector = (\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})
-        %                 \]\vspace{0mm}
-        %             \end{minipage}
-        %         \end{displaybox}
-        %     \end{minipage}
-        % \end{onlyenv}
-        \begin{onlyenv}<13>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{mydescription}
-                    \item[\alignmentVector] Sequence alignments
-                    \item[\divTimeMapVector] Divergence times
-                    \item[\divModel{}] Divergence model
-                    \item[\allParameters{}] Gene trees
-                    \item[\ ] Substitution parameters
-                    \item[\ ] Demographic parameters
-                    % \item[\allParameters] $(\divTimeMapVector, \geneTreeVector,
-                    %     \hkyModelVector, \demographicParamVector)$
-                \end{mydescription}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{uncoverenv}<11->
-            \begin{minipage}[c][0.18\textheight][c]{\linewidth}
-            \begin{flushleft}
-                \small We want to infer \textcolor{blue}{\divModel{}} and
-                \textcolor{blue}{\divTimeMapVector} given DNA sequence
-                alignments
-                \textcolor{blue}{\alignmentVector}
-            \end{flushleft}
-            \end{minipage}
-        \end{uncoverenv}
-        \column{.6\textwidth}
-        \begin{minipage}[t][0.8\textheight][c]{\linewidth}
-        \includegraphics<1>[height=6cm]{../images/dmc-cartoon-islands.pdf}
-        \includegraphics<2>[height=6cm]{../images/dmc-cartoon-taxa.pdf}
-        \includegraphics<3>[height=6cm]{../images/dmc-cartoon-sea-level.pdf}
-        \includegraphics<4>[height=6cm]{../images/dmc-cartoon-shared-no-labels.pdf}
-        \includegraphics<5-6>[height=6cm]{../images/dmc-cartoon-shared.pdf}
-        \includegraphics<7>[height=6cm]{../images/dmc-cartoon-2-1.pdf}
-        \includegraphics<8>[height=6cm]{../images/dmc-cartoon-2-2.pdf}
-        \includegraphics<9>[height=6cm]{../images/dmc-cartoon-2-3.pdf}
-        \includegraphics<10-11>[height=6cm]{../images/dmc-cartoon-general.pdf}
-        \includegraphics<12->[height=6cm]{../images/dmc-cartoon-pop-general.pdf}
-        \end{minipage}
-    \end{columns}
-\end{frame}
-
-%   c. Testing for predicted patterns via model choice.
-%       i. Bayesian model choice
-
-\begin{frame}[t]
-    \frametitle{Bayesian model choice}
-    \begin{block}{\it Full model:}
-        \begin{minipage}[c][3.8cm][c]{\linewidth}
-            \begin{uncoverenv}<1->
-            \[
-                p(
-                  \divTimeMapVector,
-                  \allParameters{}
-                  \given \alignmentVector, \divModel{i})
-                  =
-                \frac{
-                    p(\alignmentVector \given
-                      \divTimeMapVector,
-                      \allParameters{},
-                      \divModel{i}
-                      )
-                      p(
-                        \divTimeMapVector,
-                        \allParameters{}
-                        \given \divModel{i}
-                        )
-                    }{p(\alignmentVector \given \divModel{i})}
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-            \begin{uncoverenv}<2->
-            \[
-                p(\alignmentVector \given \divModel{i}) =
-                \int_{\divTimeMapVector} \int_{\allParameters{}}
-                p(\alignmentVector \given \divTimeMapVector, \allParameters{}, \divModel{i})
-                p(\divTimeMapVector, \allParameters{} \given \divModel{i})
-                d\divTimeMapVector d\allParameters{}
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-            \begin{uncoverenv}<3->
-            \[
-                p(\divModel{i} \given \alignmentVector) =
-                \frac{
-                    p(\alignmentVector \given \divModel{i})
-                    p(\divModel{i})
-                }{
-                    \sum_{i} p(\alignmentVector \given \divModel{i})
-                    p(\divModel{i})
-                }
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-        \end{minipage}
-    \end{block}
-    \vspace{1mm}
-    \begin{uncoverenv}<4>
-        \begin{center}
-            \msb: Approximate Bayesian computation (ABC)
-
-            % \vspace{-5mm}
-            % \[ \alignmentVector \, \to \, \ssVectorObs \, \to \, \ssSpace\]
-        \end{center}
-    \end{uncoverenv}
-    \barefootnote{\tiny\shortfullcite{Huang2011}. \hspace{2mm}\shortfullcite{Oaks2012}.}
-\end{frame}
-
-%   d. Method exists---applied it to comparative dataset from Philippines
-%       i. Surprising results!
-%       ii. Rather than publish, did some detective work
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1>
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/cam.jpg}}
-        \centerline{
-        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/charles.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{/home/jamie/Dropbox/field-photos/people/jake.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-        \begin{onlyenv}<2->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1-2>[width=\textwidth]{../images/maps/Philippines.png}
-        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        \begin{onlyenv}<4->
-            \begin{itemize}
-                \item Strong support for simultaneous divergence of all 22 taxon pairs
-
-                \item Posterior probability $>$ 0.96 \\
-
-                \item $\sim$100,000--250,000 years ago
-            \end{itemize}
-        \end{onlyenv}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-%   e. Evaluated method via simulation
-%       i. biased in alarming direction
-
-\section{Simulation-based assessment of method}
-
-\begin{frame}
-    \frametitle{Simulation-based power analyses}
-    % What is ``simultaneous''?
-    \begin{myitemize}
-        \item<1-> What if divergences were random?
-        \item<2-> Simulate datasets in which all 22 divergence times are random
-        \uncover<3->{
-            \smallskip
-            \begin{myitemize}
-                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
-                \smallskip
-            \end{myitemize}
-        \item<3-> $MGA$ = Millions of Generations Ago
-        }
-        \item<4-> Simulate 1000 datasets for each \divTime{} distribution
-        \item<4-> Analyze all 4000 datasets as we did the empirical data
-    \end{myitemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Simulation-based power analyses: Results}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]{../images/old_old_power_psi_mode.pdf}}
-
-    \begin{uncoverenv}<2->
-        \begin{itemize}[<+->]
-            \item \msb will often infer clustered divergences when divergences
-                are random over millions of generations.
-
-            \item The empirical results are likely spurious
-        \end{itemize}
-    \end{uncoverenv}
-
-    \begin{uncoverenv}<3->
-        \begin{block}{Objective:}
-            Use principles of probability to extend \msb framework for improved
-            estimation of shared evolutionary history
-        \end{block}
-    \end{uncoverenv}
-    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
-\end{frame}
-
-%   f. Rather than abandon approach---look for theoretical reasons for behavior
-%       i. Strongly weighted posterior (marginal likelihood)
-%       ii. Combinatorics
-%   g. Developed new method---dpp-msbayes
-%       i. Used DPP---common in Bayeisn nonparametrics
-%           1. Flexible and mathematically convenient way of assigning prior
-%           probability to all possible divergence models (partitions) directly
-%           (avoid combinatoric issue)
-%       ii. More flexible priors on parameters of model
-%       iii. Implemented method in much more efficient multi-processing
-%       framework.
-%   e. Results---improved!
-%   f. Empirical results...
-%       i. Reducing bias and increasing power provides more "honest" estimate
-%       of uncertainty
-%       ii. A lot of uncertainty (Good! and Bad!)
-%       iii. What next? More data and more power!
-
-\begin{frame}[label=improvements]
-    \frametitle{An improved method}
-    Potential improvements:
-    \begin{enumerate}
-        \item Alternative priors on parameters that increase marginal
-            likelihoods of rich models
-        \item Alternative approach to modeling the temporal distribution of
-            divergences
-    \end{enumerate}
-    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
-\end{frame}
-
-\begin{frame}[t]
-    \vspace{-2mm}
-    \begin{displaybox}[5.5cm]
-        \small
-        \[
-            p(X) = \int_{\theta} p(X \given \theta) p(\theta) \mathrm{d}\theta
-        \]%\vspace{0mm}
-    \end{displaybox}
-
-    \vspace{-1mm}
-    \begin{center}
-        % \includegraphics<1>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<2>[height=7.8cm]{../images/marginal-plot-3d-bare.png}
-        \includegraphics<3>[height=7.8cm]{../images/marginal-plot-3d-constrained.png}
-        \includegraphics<4>[height=7.8cm]{../images/marginal-plot-3d.png}
-        \includegraphics<5>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<6>[height=7.8cm]{../images/marginal-plot-2d.pdf}
-    \end{center}
-\end{frame}
-
-\againframe{improvements}
-
-\begin{frame}[t]
-    \frametitle{Prior on divergence models}
-    \begin{itemize}
-        \item \msb uses a discrete uniform prior on the \emph{number} of
-            divergence events
-    \end{itemize}
-
-    \begin{uncoverenv}<2->
-    \begin{columns}
-        \column{0.5\textwidth}
-            \centerline{
-                \includegraphics<2->[width=1.0\textwidth]{../images/number-of-div-models-22-unordered.pdf}}
-        \column{0.5\textwidth}
-            \centerline{
-                \includegraphics<3->[width=1.0\textwidth]{../images/prob-of-div-models-22-unordered.pdf}}
-    \end{columns}
-    \end{uncoverenv}
-
-    \begin{uncoverenv}<4->
-    \begin{block}{\it Potential solution:}
-        Place flexible prior directly on the sample space of divergence models
-    \end{block}
-    \end{uncoverenv}
-\end{frame}
-
-\begin{frame}
-    \frametitle{New method: \dppmsbayes}
-    \begin{itemize}[<+->]
-        % \item<1-> Reparameterized the model implemented in \msb
-        \item<1-> Replaced uniform priors on continuous parameters with gamma and
-            beta distributions
-        \item<2-> Dirichlet process prior (DPP) over all possible divergence
-            models
-        \item<3-> Develop multi-processing interface to improve performance for
-            accommodating genomic datasets
-        % \item<4-> Uniform prior over divergence models
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Simulation-based assessment}
-        \uncover<1->{
-        Validation:\\
-
-        \smallskip
-        \begin{itemize}
-            \item Simulate 50,000 datasets and analyze each under the same
-                model
-        \end{itemize}
-        }
-
-        \uncover<2->{
-        \bigskip
-        Robustness:\\
-        \begin{itemize}
-            \item Simulate datasets that violate model assumptions and analyze
-                each of them
-        \end{itemize}
-        }
-
-        \uncover<3->{
-        \bigskip
-        Done for \msb and \dppmsbayes
-        }
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Validation results}
-    \begin{columns}
-        \column{0.5\textwidth}
-            \includegraphics<1->[width=1.0\textwidth]{../images/validation-model-choice-old.pdf}
-        \column{0.5\textwidth}
-            \includegraphics<2->[width=1.0\textwidth]{../images/validation-model-choice-dpp.pdf}
-    \end{columns}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Robustness results}
-    \begin{columns}
-        \column{0.5\textwidth}
-            \centering{
-            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-old-violations.pdf}
-            }
-        \column{0.5\textwidth}
-            \centering{
-            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-dpp-violations.pdf}
-            }
-    \end{columns}
-
-    \begin{uncoverenv}<2->
-    \begin{itemize}
-        \item Improved model-choice accuracy and robustness
-    \end{itemize}
-    \end{uncoverenv}
-    \vspace{-0.5cm}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Simulation-based power analyses}
-    Does the new method have power to detect variation in divergence times?\\
-
-    \bigskip
-    \begin{myitemize}
-        \item<2-> Simulate datasets in which all 22 divergence times are random
-            \smallskip
-            \begin{myitemize}
-                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
-                \smallskip
-            \end{myitemize}
-        \item<2-> $MGA$ = Millions of Generations Ago
-
-        \item<3-> Simulate 1000 datasets for each \divTime{} distribution
-        \item<3-> Analyze all 4000 datasets with \msb and \dppmsbayes
-    \end{myitemize}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{\dppmsbayes: Power results}
-        \begin{onlyenv}<1-2>
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<1-2>{../images/power-psi-old-annotate-no-pp.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<2>{../images/power-psi-dpp-annotate-no-pp.pdf}}
-        \end{onlyenv}
-
-        \begin{onlyenv}<3->
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-old-annotate-pp.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-dpp-annotate-pp.pdf}}
-        \end{onlyenv}
-
-    \begin{uncoverenv}<4->
-    \begin{itemize}
-        \item Improved power to detect temporal variation across divergences
-    \end{itemize}
-    \end{uncoverenv}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Empirical results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp.pdf}}
-
-    \begin{uncoverenv}<2->
-    \begin{itemize}
-        \item More ``honest'' estimate of posterior uncertainty
-    \end{itemize}
-    \end{uncoverenv}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Empirical results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
-    \smallskip
-    \centerline{
-    \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Summary}
-    \begin{itemize}
-        \item<1-> New method for estimating shared evolutionary history shows
-            improved:
-            \begin{enumerate}
-                \item<2-> Estimation of posterior uncertainty
-                \item<2-> Model-choice accuracy
-                \item<2-> Power to detect temporal variation across divergences
-                \item<2-> Robustness to model violations
-                \item<2-> Speed!
-            \end{enumerate}
-
-        \smallskip
-        \item<3-> Climate-driven diversification model?
-            \begin{itemize}
-                \item<4-> Results consistent with prediction of clustered
-                    divergences
-                \item<4-> Results suggest multiple co-divergences
-                \item<4-> However, there is a lot of uncertainty!
-            \end{itemize}
-
-        \smallskip
-        \item<5-> What to do?
-            \begin{itemize}
-                \item<6-> More data!
-                \item<6-> More power! 
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-%   g. Collected thousands of loci from across the genome of almost 300
-%   individuals of two genera of geckos from across the Philippine Islands
-%       i. Prelim results with dpp-msbayes (can accommodate NGS data!)
-
-\begin{frame}
-    \frametitle{More data}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{myitemize}
-            \item Collecting genomic data from two gekkonid radiations
-                co-distributed Southeast Asia
-            \begin{myitemize}
-                \item \emph{Cyrtodactylus}---12+ species
-                \item \emph{Gekko}---13+ species
-            \end{myitemize}
-            \item Whole genome assemblies for each genus
-            \item RAD-seq libraries constructed for $\approx150$ individuals
-                from each genus
-        \end{myitemize}
-        \centerline{
-        \includegraphics<1->[height=2.5cm]{/home/jamie/Dropbox/field-photos/misc/lee-and-me.jpg}
-        \hspace{2mm}
-        \includegraphics<1->[height=2.5cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}}
-        \column{.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/photos/cyrt-agusanensis.jpg} \quad
-        \includegraphics[width=\textwidth]{../images/photos/gekko-mindorensis.jpg}
-    \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{More data---preliminary results}
-    Results for 1000 loci from 5 pairs of \emph{Gekko mindorensis} populations\\
-
-    \begin{center}
-        \includegraphics<1->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
-    \end{center}
-\end{frame}
-
-%   h. Currently working on full-likelihood implementation of the model.
-%       i. Working with Jeet---has full-likelihood up and running---inefficient
-%       ii. Take advantage of recent developments to analytically integrate
-%       gene trees---might even be faster than simulations?!
-%       iii. Uses ALL information in genetic data
-
-\begin{frame}
-    \frametitle{More power}
-    \centerline{
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/mtholder.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/vladimir.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../images/revbayes.png}}
-
-    \begin{itemize}[<+->]
-        \item Full-likelihood Bayesian implementation \footnote{\tiny\shortfullcite{JeetDiss}}
-            \begin{itemize}
-                \item Uses all the information in the sequence data
-                \item Applicable to much deeper timescales
-                \item \ldots Currently inefficient
-            \end{itemize}
-        \item Working to incorporate new developments on analytical integration
-            of gene trees \footnote{\tiny\shortfullcite{Bryant2012}}
-            \begin{itemize}
-                \item Very efficient numerical approximation of posterior 
-                \item Applicable to NGS datasets
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-\section{Future directions}
-
-% III. Future work
-%   a. Theory/computation
-%       i. Extend full-likelihood approach to phylogenetic framework
-%           1. Can infer relative mutation rates/pop sizes from data (fewer
-%           assumptions)
-%           2. Much better representation of diversification than pairwise
-%           approach.
-%       ii. Parameter estimation approach (Lindley's paradox).
-
-\begin{frame}
-    \frametitle{Future directions---Computational}
-    \begin{itemize}
-        \item<1-> Full-phylogenetic framework
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
-        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
-    \end{center}
-\end{frame}
-
-%   b. Empirical
-%       i. Gobi (NGS --> NSF)
-%       ii. West-Central Africa
-%           1. unfunded collaborator
-%       iii. Compare diversification of mainland SE Asia to Oceanic Islands
-%           1. Planning pre-proposal
-%       iv. Long-term study of small island of sunda shelf
-%           1. Tiny islands with pops of Eutropis multifasciata and
-%           Lepidodactylus lugubris
-%           2. Inundated during last interglacial ~120,000 years ago, spent the
-%           better part of next 100k years as part of exposed sunda shelf, then
-%           fragmented during current interglacial
-%           3. Both species vary in reproductive mode (oviparity/ovoviviparity,
-%           parthenogenesis), skinks in color patterns
-%           4. Affects of recent, severe fragmentation on diversification and
-%           demography
-%           5. Comparative genomics to understand evolution of color and
-%           reproductive mode
-%           6. Evolutionary ecology---what ecological context influence
-%           evolution asexuality/live-bearing/color.
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
-        \begin{itemize}
-            \item<1-> These comparative tools will allow evolutionary
-                biologists to leverage genomic data to infer shared
-                evolutionary history
-            \item<2-> This will better our understanding of how community scale
-                processes affect the generation and assembly of biodiversity
-            \item<3-> Many exciting empirical systems for applying new methods!
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
 \begin{frame}
@@
 }
-
-\begin{frame}
-    \frametitle{Everything is on GitHub\ldots}
-    Software:\\
-    \begin{itemize}
-        \item \texttt{dpp-msbayes}:
-            \href{https://github.com/joaks1/dpp-msbayes}{\tt
-            https://github.com/joaks1/dpp-msbayes}
-
-        \item \texttt{PyMsBayes}:
-            \href{https://github.com/joaks1/PyMsBayes}{\tt
-            https://github.com/joaks1/PyMsBayes}
-
-        \item ABACUS: Approximate BAyesian C UtilitieS.
-            \href{https://github.com/joaks1/abacus}{\tt
-            https://github.com/joaks1/abacus}
-    \end{itemize}
-
-    \medskip
-    Open-Science Notebook:\\
-    \begin{itemize}
-        \item \texttt{msbayes-experiments}:
-            \href{https://github.com/joaks1/msbayes-experiments}{\tt
-            https://github.com/joaks1/msbayes-experiments}
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Acknowledgments}
-    \begin{columns}[t]
-        \column{.5\textwidth}
-            {\bf Ideas and feedback:}
-            \begin{myitemize}
-                \item Leach\'{e} Lab
-                \item Minin Lab
-                \item Holder Lab
-                \item Brown Lab/KU Herpetology
-                \item Melissa Callahan
-            \end{myitemize}
-            \smallskip
-            {\bf Computation:}\\
-            \includegraphics<1->[height={8mm}]{../images/iplant.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/kuittc.png}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/uw.png}\\
-
-        \column{.5\textwidth}
-            {\bf Funding:}\\
-            \includegraphics<1->[height={8mm}]{../images/nsf.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/ngs.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/ssb.png}\\
-
-            \smallskip
-            {\bf Photo credits:}
-            \begin{myitemize}
-                \item Rafe Brown, Cam Siler, Jesse Grismer, \& Jake Esselstyn
-                \item FMNH Philippine Mammal Website:
-                    \begin{myitemize}
-                        \item D.S.\ Balete, M.R.M.\ Duya, \& J.\ Holden
-                    \end{myitemize}
-                \item PhyloPic!
-            \end{myitemize}
-    \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Questions?}    
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{center}
-            {
-            \Large
-            \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
-            }
-        \end{center}
-        \column{.5\textwidth}
-            \begin{figure}
-                \begin{center}
-                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
-                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
-                \end{center}
-            \end{figure}
-    \end{columns}
-\end{frame}
-
-% Extra slides
-
-\begin{frame}[noframenumbering]
-    \frametitle{Easy as ABC}
-    \centerline{
-    \includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}
-    \includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}
-    \includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}
-    \includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}
-    \includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}
-    \includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}
-    \includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}
-    \includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}
-    \includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}
-    \includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}
-    }
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{Causes of bias: Insufficient sampling}
-    \begin{itemize}
-        \item Models with more parameter space are less densely sampled
-        \item Could explain bias toward small models in extreme cases
-        \item {\bf Predicts large variance in posterior estimates}
-        \begin{itemize}
-            \item We explored empirical and simulation-based analyses with
-                2, 5, and 10 million prior samples, and estimates were
-                very similar
-        \end{itemize}
-    \end{itemize}
-    \smallskip
-    \centerline{
-    \includegraphics[width=\textwidth]{../images/omega_over_sampling.pdf}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{\dppmsbayes: Simulation results}
-    % \vspace{1cm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_old_power_psi_mode.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_psi_mode_headless.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_dpp_power_psi_mode_headless.pdf}}
-    % \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{\dppmsbayes: Simulation results}
-    % \vspace{1cm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_old_power_omega_median.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_omega_median_headless.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_dpp_power_omega_median_headless.pdf}}
-    % \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{Empirical results: Philippine diversification}
-    \centerline{
-    \includegraphics[width=0.8\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp-with-prior.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\end{document}

### 2015-03-05 slides/auburn/intro-mascot-tree.tex
Working on slides.

@@
         % \includegraphics<2>[width=\textwidth]{../images/mascot-tree.png}
-        \includegraphics[width=\textwidth]{../images/mascot-tree.png}
+        \includegraphics<1>[width=\textwidth]{../mascot-tree/mascot-tree-shared.pdf}
+        \includegraphics<2->[width=\textwidth]{../mascot-tree/mascot-tree-shared-events.pdf}
 
@@
 \end{frame}
+
+\begin{frame}
+    \begin{columns}
+        \column{0.4\textwidth}
+        \begin{minipage}[c]{\columnwidth}
+        \begin{adjustwidth}{-2.3em}{}
+        \begin{onlyenv}<1-6>
+        \begin{itemize}
+            \item<2-> Current phylogenetic methods assume events are independent 
+            \item<3-> We know this assumption is frequently violated
+            \item<4-> Why account for this non-independence?
+            \begin{itemize}
+                \item<5-> Improve phylogenetic inference
+                \item<6-> Provide a general framework for inferring the affect
+                    of community-scale processes of diversification
+            \end{itemize}
+        \end{itemize}
+        \end{onlyenv}
+        \end{adjustwidth}
+        \end{minipage}
+
+        \column{0.6\textwidth}
+        \begin{minipage}[t][\textheight][c]{\linewidth}
+        \centerline{
+        \includegraphics<1->[width=\columnwidth]{../mascot-tree/mascot-tree-shared-events.pdf}
+        }
+        \end{minipage}
+    \end{columns}
+\end{frame}

### 2015-03-05 slides/auburn/more-data.tex
Working on slides.

--- 2015-02-18 slides/bordeaux/slides.tex
@@
-\input{../utils/preamble.tex}
-\input{../../manuscripts/utils/macros.tex}
-\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
-\bibliography{../../manuscripts/bib/references}
-
-% \title[Estimating shared history]{An Improved Approximate-Bayesian Method for Estimating Shared
-% Evolutionary History}
-
-\title[Inferring community history]{Inferring Community-scale Diversification
-    from Genomic Data}
-
-\author[J.\ Oaks]{
-    Jamie R.\ Oaks\inst{1}
-}
-\institute[University of Washington]{
-    \inst{1}%
-        Department of Biology, University of Washington
-}
-
-% \date{\today}
-\date{February 18, 2015}
-
-\begin{document}
-\maketitle
-
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
+\begin{frame}[t]
+    \frametitle{Current work: More data}
+    \begin{minipage}[t][0.15\textheight][c]{\linewidth}
     \begin{itemize}
-            \item<2-> Community-scale processes are important in
-                biodiversification and community assembly
-            \item<3-> Accounting for such processes will better our
-                understanding of evolutionary history and
-                community structure
-            \item<4-> We need methods for inferring evolutionary patterns
-                predicted by historical events from contemporary populations
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}
-
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-
-        \vspace{6.5cm}
-
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \begin{uncoverenv}<2->
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-        \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1>
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rob.jpg}}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/charles.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/cam.jpg}}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jake.jpg}}
-        \centerline{
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/allie.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/Luke.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/anthony.png}}
-        \end{minipage}
-        \end{onlyenv}
-
-        \begin{onlyenv}<2->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1-2>[width=\textwidth]{../images/maps/Philippines.png}
-        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        % \begin{onlyenv}<4->
-        %     \begin{itemize}
-        %         \item Strong support for simultaneous divergence of all 22 taxon pairs
-
-        %         \item Posterior probability $>$ 0.96 \\
-
-        %         \item $\sim$100,000--250,000 years ago
-        %     \end{itemize}
-        % \end{onlyenv}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle<1-4>{Climate-driven diversification: Predictions}
-    \frametitle<5->{Divergence-model choice}
-
-    \centering{
-    \includegraphics<1>[height=7.5cm]{../images/dmc-cartoon-islands.pdf}
-    \includegraphics<2>[height=7.5cm]{../images/dmc-cartoon-taxa.pdf}
-    \includegraphics<3>[height=7.5cm]{../images/dmc-cartoon-sea-level.pdf}
-    \includegraphics<4>[height=7.5cm]{../images/dmc-cartoon-shared-no-labels.pdf}
-    \includegraphics<5>[height=7.5cm]{../images/dmc-cartoon-shared.pdf}
-    \includegraphics<6>[height=7.5cm]{../images/dmc-cartoon-2-1.pdf}
-    \includegraphics<7>[height=7.5cm]{../images/dmc-cartoon-2-2.pdf}
-    \includegraphics<8>[height=7.5cm]{../images/dmc-cartoon-2-3.pdf}
-    \includegraphics<9>[height=7.5cm]{../images/dmc-cartoon-general.pdf}
-    }
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Divergence-model choice}
-
-    \vspace{-7mm}
-
-    \begin{minipage}[t][0.1\textheight][c]{1.1\linewidth}
-        \begin{adjustwidth}{-0.5em}{}
-            \begin{tabular}{ p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} }
-                $m_1$ & $m_2$ & $m_3$ & $m_4$ & $m_5$ \\
-            \end{tabular}
-        \end{adjustwidth}
-    \end{minipage}
-
-    \vspace{-2mm}
-
-    \begin{minipage}[t][0.35\textheight][c]{1.1\linewidth}
-        \begin{adjustwidth}{-2.3em}{}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-shared.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-1.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-2.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-3.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-general.pdf}
-        \end{adjustwidth}
-    \end{minipage}
-
-    \vspace{3.5mm}
-
-    \begin{minipage}[c][0.35\textheight][c]{\linewidth}
-        \begin{uncoverenv}<2->
-            \begin{center}
-            We want to infer \textcolor{blue}{\divModel{}} and
-            \textcolor{blue}{\divTimeMapVector} given DNA sequence
-            alignments
-            \textcolor{blue}{\alignmentVector}
-        \end{center}
-        \end{uncoverenv}
-
-        \vspace{4mm}
-
-        \begin{uncoverenv}<3->
-            \begin{displaybox}[0.85\linewidth]
-                \begin{minipage}[c][0.12\textheight][c]{\linewidth}
-
-                \[
-                    p(\divTimeMapVector,
-                      \divModel{i},
-                      \allParameters{}
-                      \given \alignmentVector)
-                      =
-                    \frac{
-                        p(\alignmentVector \given
-                          \divTimeMapVector,
-                          \allParameters{},
-                          \divModel{i})
-                        p(\divTimeMapVector,
-                          \allParameters{}
-                          \given \divModel{i})
-                        p(\divModel{i})
-                        }{p(\alignmentVector)}
-                \]
-                \vspace{-1mm}
-                \end{minipage}
-            \end{displaybox}
-        \end{uncoverenv}
-    \end{minipage}
-\end{frame}
-
-\begin{frame}
-    \frametitle{New method: \dppmsbayes}
-    \begin{itemize}
-        % \item<1-> Reparameterized the model implemented in \msb
-        \item<1-> Dirichlet-process prior (DPP) over all possible divergence
-            models
-        \item<2-> Flexible priors on continuous parameters to avoid strongly
-            weighted posteriors
-        \item<3-> Multi-processing interface to improve performance for
-            accommodating genomic datasets
-        % \item<4-> Uniform prior over divergence models
-    \end{itemize}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Performance}
-    \begin{itemize}
-        \item<1-> Extensive computer simulations to compare performance to previous
-            state-of-the-art methods (\msb).
-        \item<2-> New method for estimating shared evolutionary history shows
-            improved:
-            \begin{enumerate}
-                \item<2-> Improved model-choice accuracy 
-                \item<2-> More robustness to model violations
-                \item<2-> Greater power to detect variation in divergence times
-                \item<2-> Faster!
-            \end{enumerate}
-    \end{itemize}
-
-    \uncover<3->{
-    \textbf{\dppmsbayes allows biologists to leverage comparative genomic data
-        to assess the affects of community-scale processes on biodiversity}
-    }
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-
-        \vspace{6.5cm}
-
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \begin{uncoverenv}<1->
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-        \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1->[width=\textwidth]{../images/maps/Philippines.png}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
-    \smallskip
-    \centerline{
-    \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Current work: More data}
-    \begin{itemize}
-        \item<1-> Collecting genomic data from taxa co-distributed across Southeast
-            Asia
+        \item<1-> Collecting genomic data from taxa co-distributed across
+            Southeast Asian Islands and Mainland
         \item<2-> Preliminary results for 1000 loci from 5 pairs of \emph{Gekko
@@
 \end{frame}
-
-\begin{frame}
-    \frametitle{Current work: More power}
-    \centerline{
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/mtholder.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/vladimir.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../images/revbayes.png}}
-
-    \begin{itemize}[<+->]
-        \item Full-likelihood Bayesian implementation
-            \begin{itemize}
-                \item Uses all the information in the data
-                \item Applicable to deeper timescales
-            \end{itemize}
-        \item Analytically integrate over gene trees
-            \footnote{\tiny\shortfullcite{Bryant2012}}
-            \begin{itemize}
-                \item Very efficient numerical approximation of posterior 
-                \item Applicable to NGS datasets
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-\section{Conclusions}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification model?}
-    \begin{itemize}
-        \item<1-> Results consistent with prediction of clustered
-            divergences
-        \item<2-> Results suggest multiple co-divergences
-        \item<3-> However, there is a lot of uncertainty!
-    \end{itemize}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\section{Teaching}
-
-\begin{frame}
-    BIO 180 stuff
-\end{frame}
-
-\begin{frame}
-    Software carpentry \\
-
-    Workshops
-\end{frame}
-
-\begin{frame}
-    \frametitle{Questions?}    
-    \begin{center}
-        {
-        \Large
-        \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
-        }
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Future directions---Computational}
-    \begin{itemize}
-        \item<1-> Full-phylogenetic framework
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
-        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \centerline{
-    \includegraphics<1>[height=8cm]{../images/bd-cartoon-no-lines.pdf}
-    \includegraphics<2>[height=8cm]{../images/bd-cartoon-shared.pdf}
-    }
-\end{frame}
-
-\begin{frame}
-\begin{columns}
-    \column{.5\textwidth}
-    \begin{minipage}[c]{\columnwidth}
-    \begin{itemize}
-        \item Blah
-    \end{itemize}
-    \end{minipage}
-
-    \column{.6\textwidth}
-    \begin{minipage}[t][\textheight][c]{\linewidth}
-    \includegraphics<1>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
-    \includegraphics<2>[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
-    \end{minipage}
-\end{columns}
-\end{frame}
-
-\end{document}

### 2015-03-05 slides/auburn/open-science.tex
Working on slides.

--- 2014-11-03 slides/osu/slides.tex
@@
-\input{../utils/preamble.tex}
-\input{../../manuscripts/utils/macros.tex}
-\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
-\bibliography{../../manuscripts/bib/references}
-
-\title[Improving inference of evolutionary history]{Improving inference of evolutionary history}
-\subtitle{Developing comparative methods for genomic data}
-
-\author[J.\ Oaks]{
-    Jamie R.\ Oaks\inst{1}
-}
-\institute[University of Washington]{
-    \inst{1}%
-        Department of Biology, University of Washington
-}
-
-% \date{\today}
-\date{November 3, 2014}
-
-\begin{document}
-% \maketitle
-\begin{frame}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-            \maketitle
-        \column{.5\textwidth}
-            \begin{figure}
-                \begin{center}
-                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
-                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
-                \end{center}
-            \end{figure}
-    \end{columns}
-\end{frame}
-
-% I. What I do
-% II. What I've been focusing on recently (started in PhD and continuing in postdoc)
-%   a. Dynamic planet---need to account for large-scale processes
-%       i. E.g., SE Asia---dynamic landscape
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
-        \begin{itemize}
-            \item<2-> Large-scale geological and climatic processes are
-                important in biodiversification and community assembly
-            \item<3-> Accounting for such processes will better our
-                understanding of biodiversity
-            \item<4-> We need methods for inferring evolutionary patterns
-                predicted by historical events from contemporary populations
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    \frametitle{Southeast Asia}    
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Southeast Asia}    
-    \begin{columns}
-        \column{0.6\textwidth}
-        \ \\
-
-        \column{0.45\textwidth}
-
-        \vspace{-2.4cm}
-
-        \begin{uncoverenv}<2->
-        \colorbox{white}{
-            \begin{minipage}[t][3.6cm][t]{1.1\textwidth}
-                \begin{onlyenv}<2-3>
-                \begin{itemize}
-                    \item<2-> Repeated coalescence and fragmentation of island complexes
-                    \item<3-> Prominent paradigm for explaining biodiversity
-                \end{itemize}
-                \end{onlyenv}
-
-                \begin{onlyenv}<4>
-                \begin{itemize}
-                    \item<4> \textbf{Did repeated fragmentation of islands
-                        during inter-glacial rises in sea level promote
-                        diversification?}
-                \end{itemize}
-                \end{onlyenv}
-
-                \begin{onlyenv}<5->
-                \begin{itemize}
-                    \item<5-> Climate-driven model of diversification
-                    \item<6-> Testable prediction:
-                    \begin{itemize}
-                        \item Temporally clustered divergences among taxa
-                            co-distributed across fragmented islands
-                    \end{itemize}
-                \end{itemize}
-                \end{onlyenv}
-            \end{minipage}
-        }
-    \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-% \begin{frame}
-%     \frametitle{Philippine Archipelago}
-%     \begin{center}
-%         \includegraphics<1>[width=.5\textwidth]{../images/maps/Philippines-present.png}
-%         \includegraphics<2>[width=.5\textwidth]{../images/maps/Philippines.png}
-%     \end{center}
-% \end{frame}
-
-% \begin{frame}
-%     \frametitle{Climate-driven diversification model}
-%     \begin{columns}[c]
-%         \column{.5\textwidth}
-%             \begin{myitemize}
-%                 \item<1-> Repeated coalescence and fragmentation of island complexes
-%                 \item<2-> Prominent paradigm for explaining Philippine biodiversity
-%                 \item<3-> Proposed as model of diversification
-%             \end{myitemize}
-%         \column{.5\textwidth}
-%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
-%     \end{columns}
-% \end{frame}
-
-% \begin{frame}
-%     \frametitle{Testing climate-driven diversification}
-%     \begin{columns}[c]
-%         \column{.5\textwidth}
-%     \textbf{Did repeated fragmentation of islands during inter-glacial rises in sea level
-%     promote diversification?}\\
-%     \bigskip
-%     \uncover<2->{
-%     Model has testable prediction:\\
-%     \begin{itemize}
-%         \item Temporally clustered divergences among taxa
-%             co-distributed across fragmented islands
-%     \end{itemize}
-%     }
-%         \column{.5\textwidth}
-%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
-%     \end{columns}
-% \end{frame}
-
-%   b. Predictions of climate-driven diversification
-\begin{frame}
-    \frametitle<1-4>{Community scale processes}
-    \frametitle<5->{Divergence model choice}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{onlyenv}<1-4>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<5>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{\divTime{1}\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 111
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (\divTimeMap{1},\divTimeMap{2},\divTimeMap{3})
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<6>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 111
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (130,130,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<7>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 211
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (397,130,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<8>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 121
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (130,397,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<9>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 112
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<10-12>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397,260\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 312
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (260,130,397)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        % \begin{onlyenv}<11>
-        %     \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-        %         \begin{displaybox}[0.95\linewidth]
-        %             \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-        %                 \[
-        %                     \divTimeVector = \{\divTime{1},\ldots,\divTime{\divTimeNum}\}
-        %                 \]\vspace{0mm}
-        %                 \[
-        %                     model = \divModel{i}
-        %                 \]\vspace{0mm}
-        %                 \[
-        %                     \divTimeMapVector = (\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})
-        %                 \]\vspace{0mm}
-        %             \end{minipage}
-        %         \end{displaybox}
-        %     \end{minipage}
-        % \end{onlyenv}
-        \begin{onlyenv}<13>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{mydescription}
-                    \item[\alignmentVector] Sequence alignments
-                    \item[\divTimeMapVector] Divergence times
-                    \item[\divModel{}] Divergence model
-                    \item[\allParameters{}] Gene trees
-                    \item[\ ] Substitution parameters
-                    \item[\ ] Demographic parameters
-                    % \item[\allParameters] $(\divTimeMapVector, \geneTreeVector,
-                    %     \hkyModelVector, \demographicParamVector)$
-                \end{mydescription}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{uncoverenv}<11->
-            \begin{minipage}[c][0.18\textheight][c]{\linewidth}
-            \begin{flushleft}
-                \small We want to infer \textcolor{blue}{\divModel{}} and
-                \textcolor{blue}{\divTimeMapVector} given DNA sequence
-                alignments
-                \textcolor{blue}{\alignmentVector}
-            \end{flushleft}
-            \end{minipage}
-        \end{uncoverenv}
-        \column{.6\textwidth}
-        \begin{minipage}[t][0.8\textheight][c]{\linewidth}
-        \includegraphics<1>[height=6cm]{../images/dmc-cartoon-islands.pdf}
-        \includegraphics<2>[height=6cm]{../images/dmc-cartoon-taxa.pdf}
-        \includegraphics<3>[height=6cm]{../images/dmc-cartoon-sea-level.pdf}
-        \includegraphics<4>[height=6cm]{../images/dmc-cartoon-shared-no-labels.pdf}
-        \includegraphics<5-6>[height=6cm]{../images/dmc-cartoon-shared.pdf}
-        \includegraphics<7>[height=6cm]{../images/dmc-cartoon-2-1.pdf}
-        \includegraphics<8>[height=6cm]{../images/dmc-cartoon-2-2.pdf}
-        \includegraphics<9>[height=6cm]{../images/dmc-cartoon-2-3.pdf}
-        \includegraphics<10-11>[height=6cm]{../images/dmc-cartoon-general.pdf}
-        \includegraphics<12->[height=6cm]{../images/dmc-cartoon-pop-general.pdf}
-        \end{minipage}
-    \end{columns}
-\end{frame}
-
-%   c. Testing for predicted patterns via model choice.
-%       i. Bayesian model choice
-
-\begin{frame}[t]
-    \frametitle{Bayesian model choice}
-    \begin{block}{\it Full model:}
-        \begin{minipage}[c][3.8cm][c]{\linewidth}
-            \begin{uncoverenv}<1->
-            \[
-                p(
-                  \divTimeMapVector,
-                  \allParameters{}
-                  \given \alignmentVector, \divModel{i})
-                  =
-                \frac{
-                    p(\alignmentVector \given
-                      \divTimeMapVector,
-                      \allParameters{},
-                      \divModel{i}
-                      )
-                      p(
-                        \divTimeMapVector,
-                        \allParameters{}
-                        \given \divModel{i}
-                        )
-                    }{p(\alignmentVector \given \divModel{i})}
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-            \begin{uncoverenv}<2->
-            \[
-                p(\alignmentVector \given \divModel{i}) =
-                \int_{\divTimeMapVector} \int_{\allParameters{}}
-                p(\alignmentVector \given \divTimeMapVector, \allParameters{}, \divModel{i})
-                p(\divTimeMapVector, \allParameters{} \given \divModel{i})
-                d\divTimeMapVector d\allParameters{}
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-            \begin{uncoverenv}<3->
-            \[
-                p(\divModel{i} \given \alignmentVector) =
-                \frac{
-                    p(\alignmentVector \given \divModel{i})
-                    p(\divModel{i})
-                }{
-                    \sum_{i} p(\alignmentVector \given \divModel{i})
-                    p(\divModel{i})
-                }
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-        \end{minipage}
-    \end{block}
-    \vspace{1mm}
-    \begin{uncoverenv}<4>
-        \begin{center}
-            \msb: Approximate Bayesian computation (ABC)
-
-            % \vspace{-5mm}
-            % \[ \alignmentVector \, \to \, \ssVectorObs \, \to \, \ssSpace\]
-        \end{center}
-    \end{uncoverenv}
-    \barefootnote{\tiny\shortfullcite{Huang2011}. \hspace{2mm}\shortfullcite{Oaks2012}.}
-\end{frame}
-
-%   d. Method exists---applied it to comparative dataset from Philippines
-%       i. Surprising results!
-%       ii. Rather than publish, did some detective work
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1>
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/rafe.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/cam.jpg}}
-        \centerline{
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/charles.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/jake.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-        \begin{onlyenv}<2->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1-2>[width=\textwidth]{../images/maps/Philippines.png}
-        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        \begin{onlyenv}<4->
-            \begin{itemize}
-                \item Strong support for simultaneous divergence of all 22 taxon pairs
-
-                \item Posterior probability $>$ 0.96 \\
-
-                \item $\sim$100,000--250,000 years ago
-            \end{itemize}
-        \end{onlyenv}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-%   e. Evaluated method via simulation
-%       i. biased in alarming direction
-
-\section{Simulation-based assessment of method}
-
-\begin{frame}
-    \frametitle{Simulation-based power analyses}
-    % What is ``simultaneous''?
-    \begin{myitemize}
-        \item<1-> What if divergences were random?
-        \item<2-> Simulate datasets in which all 22 divergence times are random
-        \uncover<3->{
-            \smallskip
-            \begin{myitemize}
-                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
-                \smallskip
-            \end{myitemize}
-        \item<3-> $MGA$ = Millions of Generations Ago
-        }
-        \item<4-> Simulate 1000 datasets for each \divTime{} distribution
-        \item<4-> Analyze all 4000 datasets as we did the empirical data
-    \end{myitemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Simulation-based power analyses: Results}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]{../images/old_old_power_psi_mode.pdf}}
-
-    \begin{uncoverenv}<2->
-        \begin{itemize}[<+->]
-            \item \msb will often infer clustered divergences when divergences
-                are random over millions of generations.
-
-            \item The empirical results are likely spurious
-        \end{itemize}
-    \end{uncoverenv}
-
-    \begin{uncoverenv}<3->
-        \begin{block}{Objective:}
-            Use principles of probability to extend \msb framework for improved
-            estimation of shared evolutionary history
-        \end{block}
-    \end{uncoverenv}
-    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
-\end{frame}
-
-%   f. Rather than abandon approach---look for theoretical reasons for behavior
-%       i. Strongly weighted posterior (marginal likelihood)
-%       ii. Combinatorics
-%   g. Developed new method---dpp-msbayes
-%       i. Used DPP---common in Bayeisn nonparametrics
-%           1. Flexible and mathematically convenient way of assigning prior
-%           probability to all possible divergence models (partitions) directly
-%           (avoid combinatoric issue)
-%       ii. More flexible priors on parameters of model
-%       iii. Implemented method in much more efficient multi-processing
-%       framework.
-%   e. Results---improved!
-%   f. Empirical results...
-%       i. Reducing bias and increasing power provides more "honest" estimate
-%       of uncertainty
-%       ii. A lot of uncertainty (Good! and Bad!)
-%       iii. What next? More data and more power!
-
-\begin{frame}[label=improvements]
-    \frametitle{An improved method}
-    Potential improvements:
-    \begin{enumerate}
-        \item Alternative priors on parameters that increase marginal
-            likelihoods of rich models
-        \item Alternative approach to modeling the temporal distribution of
-            divergences
-    \end{enumerate}
-    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
-\end{frame}
-
-\begin{frame}[t]
-    \vspace{-2mm}
-    \begin{displaybox}[5.5cm]
-        \small
-        \[
-            p(X) = \int_{\theta} p(X \given \theta) p(\theta) \mathrm{d}\theta
-        \]%\vspace{0mm}
-    \end{displaybox}
-
-    \vspace{-1mm}
-    \begin{center}
-        % \includegraphics<1>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<2>[height=7.8cm]{../images/marginal-plot-3d.png}
-        \includegraphics<3>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<4>[height=7.8cm]{../images/marginal-plot-2d.pdf}
-    \end{center}
-\end{frame}
-
-\againframe{improvements}
-
-\begin{frame}[t]
-    \frametitle{Prior on divergence models}
-    \begin{itemize}
-        \item \msb uses a discrete uniform prior on the \emph{number} of
-            divergence events
-    \end{itemize}
-
-    \begin{uncoverenv}<2->
-    \begin{columns}
-        \column{0.5\textwidth}
-            \centerline{
-                \includegraphics<2->[width=1.0\textwidth]{../images/number-of-div-models-22-unordered.pdf}}
-        \column{0.5\textwidth}
-            \centerline{
-                \includegraphics<3->[width=1.0\textwidth]{../images/prob-of-div-models-22-unordered.pdf}}
-    \end{columns}
-    \end{uncoverenv}
-
-    \begin{uncoverenv}<4->
-    \begin{block}{\it Potential solution:}
-        Place flexible prior directly on the sample space of divergence models
-    \end{block}
-    \end{uncoverenv}
-\end{frame}
-
-\begin{frame}
-    \frametitle{New method: \dppmsbayes}
-    \begin{itemize}[<+->]
-        % \item<1-> Reparameterized the model implemented in \msb
-        \item<1-> Replaced uniform priors on continuous parameters with gamma and
-            beta distributions
-        \item<2-> Dirichlet process prior (DPP) over all possible divergence
-            models
-        \item<3-> Develop multi-processing interface to improve performance for
-            accommodating genomic datasets
-        % \item<4-> Uniform prior over divergence models
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Simulation-based assessment}
-        \uncover<1->{
-        Validation:\\
-
-        \smallskip
-        \begin{itemize}
-            \item Simulate 50,000 datasets and analyze each under the same
-                model
-        \end{itemize}
-        }
-
-        \uncover<2->{
-        \bigskip
-        Robustness:\\
-        \begin{itemize}
-            \item Simulate datasets that violate model assumptions and analyze
-                each of them
-        \end{itemize}
-        }
-
-        \uncover<3->{
-        \bigskip
-        Done for \msb and \dppmsbayes
-        }
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Validation results}
-    \begin{columns}
-        \column{0.5\textwidth}
-            \includegraphics<1->[width=1.0\textwidth]{../images/validation-model-choice-old.pdf}
-        \column{0.5\textwidth}
-            \includegraphics<2->[width=1.0\textwidth]{../images/validation-model-choice-dpp.pdf}
-    \end{columns}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Robustness results}
-    \begin{columns}
-        \column{0.5\textwidth}
-            \centering{
-            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-old-violations.pdf}
-            }
-        \column{0.5\textwidth}
-            \centering{
-            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-dpp-violations.pdf}
-            }
-    \end{columns}
-
-    \begin{uncoverenv}<2->
-    \begin{itemize}
-        \item Improved model-choice accuracy and robustness
-    \end{itemize}
-    \end{uncoverenv}
-    \vspace{-0.5cm}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Simulation-based power analyses}
-    Does the new method have power to detect variation in divergence times?\\
-
-    \bigskip
-    \begin{myitemize}
-        \item<2-> Simulate datasets in which all 22 divergence times are random
-            \smallskip
-            \begin{myitemize}
-                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
-                \smallskip
-            \end{myitemize}
-        \item<2-> $MGA$ = Millions of Generations Ago
-
-        \item<3-> Simulate 1000 datasets for each \divTime{} distribution
-        \item<3-> Analyze all 4000 datasets with \msb and \dppmsbayes
-    \end{myitemize}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{\dppmsbayes: Power results}
-        \begin{onlyenv}<1-2>
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<1-2>{../images/power-psi-old-annotate-no-pp.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<2>{../images/power-psi-dpp-annotate-no-pp.pdf}}
-        \end{onlyenv}
-
-        \begin{onlyenv}<3->
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-old-annotate-pp.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-dpp-annotate-pp.pdf}}
-        \end{onlyenv}
-
-    \begin{uncoverenv}<4->
-    \begin{itemize}
-        \item Improved power to detect temporal variation across divergences
-    \end{itemize}
-    \end{uncoverenv}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Empirical results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp.pdf}}
-
-    \begin{uncoverenv}<2->
-    \begin{itemize}
-        \item More ``honest'' estimate of posterior uncertainty
-    \end{itemize}
-    \end{uncoverenv}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Empirical results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
-    \smallskip
-    \centerline{
-    \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Summary}
-    \begin{itemize}
-        \item<1-> New method for estimating shared evolutionary history shows
-            improved:
-            \begin{enumerate}
-                \item<2-> Estimation of posterior uncertainty
-                \item<2-> Model-choice accuracy
-                \item<2-> Power to detect temporal variation across divergences
-                \item<2-> Robustness to model violations
-                \item<2-> Speed!
-            \end{enumerate}
-
-        \smallskip
-        \item<3-> Climate-driven diversification model?
-            \begin{itemize}
-                \item<4-> Results consistent with prediction of clustered
-                    divergences
-                \item<4-> Results suggest multiple co-divergences
-                \item<4-> However, there is a lot of uncertainty!
-            \end{itemize}
-
-        \smallskip
-        \item<5-> What to do?
-            \begin{itemize}
-                \item<6-> More data!
-                \item<6-> More power! 
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-%   g. Collected thousands of loci from across the genome of almost 300
-%   individuals of two genera of geckos from across the Philippine Islands
-%       i. Prelim results with dpp-msbayes (can accommodate NGS data!)
-
-\begin{frame}
-    \frametitle{More data}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{myitemize}
-            \item Collecting genomic data from two gekkonid radiations
-                co-distributed Southeast Asia
-            \begin{myitemize}
-                \item \emph{Cyrtodactylus}---12+ species
-                \item \emph{Gekko}---13+ species
-            \end{myitemize}
-            \item Whole genome assemblies for each genus
-            \item RAD-seq libraries constructed for $\approx150$ individuals
-                from each genus
-        \end{myitemize}
-        \centerline{
-        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/misc/lee-and-me.jpg}
-        \hspace{2mm}
-        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/people/rafe.jpg}}
-        \column{.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/photos/cyrt-agusanensis.jpg} \quad
-        \includegraphics[width=\textwidth]{../images/photos/gekko-mindorensis.jpg}
-    \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{More data---preliminary results}
-    Results for 1000 loci from 5 pairs of \emph{Gekko mindorensis} populations\\
-
-    \begin{center}
-        \includegraphics<1->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
-    \end{center}
-\end{frame}
-
-%   h. Currently working on full-likelihood implementation of the model.
-%       i. Working with Jeet---has full-likelihood up and running---inefficient
-%       ii. Take advantage of recent developments to analytically integrate
-%       gene trees---might even be faster than simulations?!
-%       iii. Uses ALL information in genetic data
-
-\begin{frame}
-    \frametitle{More power}
-    \centerline{
-    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/mtholder.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/jeet2.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/vladimir.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../images/revbayes.png}}
-
-    \begin{itemize}[<+->]
-        \item Full-likelihood Bayesian implementation \footnote{\tiny\shortfullcite{JeetDiss}}
-            \begin{itemize}
-                \item Uses all the information in the sequence data
-                \item Applicable to much deeper timescales
-                \item \ldots Currently inefficient
-            \end{itemize}
-        \item Working to incorporate new developments on analytical integration
-            of gene trees \footnote{\tiny\shortfullcite{Bryant2012}}
-            \begin{itemize}
-                \item Very efficient numerical approximation of posterior 
-                \item Applicable to NGS datasets
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-\section{Future directions}
-
-% III. Future work
-%   a. Theory/computation
-%       i. Extend full-likelihood approach to phylogenetic framework
-%           1. Can infer relative mutation rates/pop sizes from data (fewer
-%           assumptions)
-%           2. Much better representation of diversification than pairwise
-%           approach.
-%       ii. Parameter estimation approach (Lindley's paradox).
-
-\begin{frame}
-    \frametitle{Future directions---Computational}
-    \begin{itemize}
-        \item<1-> Full-phylogenetic framework
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
-        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
-    \end{center}
-\end{frame}
-
-%   b. Empirical
-%       i. Gobi (NGS --> NSF)
-%       ii. West-Central Africa
-%           1. unfunded collaborator
-%       iii. Compare diversification of mainland SE Asia to Oceanic Islands
-%           1. Planning pre-proposal
-%       iv. Long-term study of small island of sunda shelf
-%           1. Tiny islands with pops of Eutropis multifasciata and
-%           Lepidodactylus lugubris
-%           2. Inundated during last interglacial ~120,000 years ago, spent the
-%           better part of next 100k years as part of exposed sunda shelf, then
-%           fragmented during current interglacial
-%           3. Both species vary in reproductive mode (oviparity/ovoviviparity,
-%           parthenogenesis), skinks in color patterns
-%           4. Affects of recent, severe fragmentation on diversification and
-%           demography
-%           5. Comparative genomics to understand evolution of color and
-%           reproductive mode
-%           6. Evolutionary ecology---what ecological context influence
-%           evolution asexuality/live-bearing/color.
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
-        \begin{itemize}
-            \item<1-> These comparative tools will allow evolutionary
-                biologists to leverage genomic data to infer shared
-                evolutionary history
-            \item<2-> This will better our understanding of how community scale
-                processes affect the generation and assembly of biodiversity
-            \item<3-> Many exciting empirical systems for applying new methods!
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Future directions: Gobi Desert}
-    \vspace{-2mm}
-    \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
-    \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Future directions: Gobi Desert}
-    \vspace{-2mm}
-    \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
-    \end{columns}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Future directions: West-Central Africa}
-    \begin{center}
-        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
-        \hspace{0.3mm}
-        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure4_herps.jpeg}
-    \end{center}
-
-    \only<1>{
-    \begin{center}
-        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/adam.png}
-        \hspace{0.3mm}
-        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/fujita.jpg}
-    \end{center}
-    }
-
-    \uncover<2->{
-    \begin{center}
-        \includegraphics[height=4.5cm]{../images/number-of-divs-africa.pdf}
-    \end{center}
-    }
-\end{frame}
-
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-    \vspace{-2mm}
-    \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/malaysia/malaysia-burungs.jpg}
-    \end{columns}
-\end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Acknowledgments}
-    \begin{columns}[t]
-        \column{.5\textwidth}
-            {\bf Ideas and feedback:}
-            \begin{myitemize}
-                \item Leach\'{e} Lab
-                \item Minin Lab
-                \item Holder Lab
-                \item Brown Lab/KU Herpetology
-                \item Melissa Callahan
-            \end{myitemize}
-            \smallskip
-            {\bf Computation:}\\
-            \includegraphics<1->[height={8mm}]{../images/iplant.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/kuittc.png}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/uw.png}\\
-
-        \column{.5\textwidth}
-            {\bf Funding:}\\
-            \includegraphics<1->[height={8mm}]{../images/nsf.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/ngs.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/ssb.png}\\
-
-            \smallskip
-            {\bf Photo credits:}
-            \begin{myitemize}
-                \item Rafe Brown, Cam Siler, Jesse Grismer, \& Jake Esselstyn
-                \item FMNH Philippine Mammal Website:
-                    \begin{myitemize}
-                        \item D.S.\ Balete, M.R.M.\ Duya, \& J.\ Holden
-                    \end{myitemize}
-                \item PhyloPic!
-            \end{myitemize}
-    \end{columns}
-\end{frame}
-
 \begin{frame}
@@
 \end{frame}
-
-\begin{frame}
-    \frametitle{Questions?}    
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{center}
-            {
-            \Large
-            \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
-            }
-        \end{center}
-        \column{.5\textwidth}
-            \begin{figure}
-                \begin{center}
-                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
-                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
-                \end{center}
-            \end{figure}
-    \end{columns}
-\end{frame}
-
-% Extra slides
-
-\begin{frame}[noframenumbering]
-    \frametitle{Easy as ABC}
-    \centerline{
-    \includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}
-    \includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}
-    \includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}
-    \includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}
-    \includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}
-    \includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}
-    \includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}
-    \includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}
-    \includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}
-    \includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}
-    }
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{Causes of bias: Insufficient sampling}
-    \begin{itemize}
-        \item Models with more parameter space are less densely sampled
-        \item Could explain bias toward small models in extreme cases
-        \item {\bf Predicts large variance in posterior estimates}
-        \begin{itemize}
-            \item We explored empirical and simulation-based analyses with
-                2, 5, and 10 million prior samples, and estimates were
-                very similar
-        \end{itemize}
-    \end{itemize}
-    \smallskip
-    \centerline{
-    \includegraphics[width=\textwidth]{../images/omega_over_sampling.pdf}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{\dppmsbayes: Simulation results}
-    % \vspace{1cm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_old_power_psi_mode.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_psi_mode_headless.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_dpp_power_psi_mode_headless.pdf}}
-    % \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{\dppmsbayes: Simulation results}
-    % \vspace{1cm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_old_power_omega_median.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_omega_median_headless.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_dpp_power_omega_median_headless.pdf}}
-    % \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{Empirical results: Philippine diversification}
-    \centerline{
-    \includegraphics[width=0.8\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp-with-prior.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\end{document}

### 2015-03-05 slides/auburn/paic-study.tex
Working on slides.

--- 2015-02-18 slides/bordeaux/slides.tex
@@
-\input{../utils/preamble.tex}
-\input{../../manuscripts/utils/macros.tex}
-\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
-\bibliography{../../manuscripts/bib/references}
+% {
+% \usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
+% \begin{frame}
+%     % \frametitle{Climate-driven diversification}
 
-% \title[Estimating shared history]{An Improved Approximate-Bayesian Method for Estimating Shared
-% Evolutionary History}
-
-\title[Inferring community history]{Inferring Community-scale Diversification
-    from Genomic Data}
-
-\author[J.\ Oaks]{
-    Jamie R.\ Oaks\inst{1}
-}
-\institute[University of Washington]{
-    \inst{1}%
-        Department of Biology, University of Washington
-}
-
-% \date{\today}
-\date{February 18, 2015}
-
-\begin{document}
-\maketitle
-
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
-        \begin{itemize}
-            \item<2-> Community-scale processes are important in
-                biodiversification and community assembly
-            \item<3-> Accounting for such processes will better our
-                understanding of evolutionary history and
-                community structure
-            \item<4-> We need methods for inferring evolutionary patterns
-                predicted by historical events from contemporary populations
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}
-
-\end{frame}
-}
+% \end{frame}
+% }
 
@@
 
-\begin{frame}[t]
-    \frametitle<1-4>{Climate-driven diversification: Predictions}
-    \frametitle<5->{Divergence-model choice}
-
-    \centering{
-    \includegraphics<1>[height=7.5cm]{../images/dmc-cartoon-islands.pdf}
-    \includegraphics<2>[height=7.5cm]{../images/dmc-cartoon-taxa.pdf}
-    \includegraphics<3>[height=7.5cm]{../images/dmc-cartoon-sea-level.pdf}
-    \includegraphics<4>[height=7.5cm]{../images/dmc-cartoon-shared-no-labels.pdf}
-    \includegraphics<5>[height=7.5cm]{../images/dmc-cartoon-shared.pdf}
-    \includegraphics<6>[height=7.5cm]{../images/dmc-cartoon-2-1.pdf}
-    \includegraphics<7>[height=7.5cm]{../images/dmc-cartoon-2-2.pdf}
-    \includegraphics<8>[height=7.5cm]{../images/dmc-cartoon-2-3.pdf}
-    \includegraphics<9>[height=7.5cm]{../images/dmc-cartoon-general.pdf}
-    }
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Divergence-model choice}
-
-    \vspace{-7mm}
-
-    \begin{minipage}[t][0.1\textheight][c]{1.1\linewidth}
-        \begin{adjustwidth}{-0.5em}{}
-            \begin{tabular}{ p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} }
-                $m_1$ & $m_2$ & $m_3$ & $m_4$ & $m_5$ \\
-            \end{tabular}
+\begin{frame}
+    \frametitle{Results}
+    \begin{adjustwidth}{-4.5em}{-2.5em}
+    \begin{columns}
+        \column{0.3\linewidth}
+    \begin{itemize}
+            \small
+        \item<2-> Results consistent with prediction of clustered
+            divergences
+        \item<3-> Results suggest multiple interglacial periods were important
+            drivers of diversification and community assembly across Islands of
+            Southeast Asia
+        \item<4-> However, there is a lot of uncertainty!
+    \end{itemize}
+        \column{0.6\linewidth}
+    \centerline{
+    \includegraphics[width=0.85\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior.pdf}}
+    \smallskip
+    \centerline{
+    \includegraphics[width=0.8\textwidth]{../images/sea-level-only.pdf}}
+    \end{columns}
     \end{adjustwidth}
-    \end{minipage}
-
-    \vspace{-2mm}
-
-    \begin{minipage}[t][0.35\textheight][c]{1.1\linewidth}
-        \begin{adjustwidth}{-2.3em}{}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-shared.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-1.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-2.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-3.pdf}
-        \hspace{0.2mm}
-        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-general.pdf}
-        \end{adjustwidth}
-    \end{minipage}
-
-    \vspace{3.5mm}
-
-    \begin{minipage}[c][0.35\textheight][c]{\linewidth}
-        \begin{uncoverenv}<2->
-            \begin{center}
-            We want to infer \textcolor{blue}{\divModel{}} and
-            \textcolor{blue}{\divTimeMapVector} given DNA sequence
-            alignments
-            \textcolor{blue}{\alignmentVector}
-        \end{center}
-        \end{uncoverenv}
-
-        \vspace{4mm}
-
-        \begin{uncoverenv}<3->
-            \begin{displaybox}[0.85\linewidth]
-                \begin{minipage}[c][0.12\textheight][c]{\linewidth}
-
-                \[
-                    p(\divTimeMapVector,
-                      \divModel{i},
-                      \allParameters{}
-                      \given \alignmentVector)
-                      =
-                    \frac{
-                        p(\alignmentVector \given
-                          \divTimeMapVector,
-                          \allParameters{},
-                          \divModel{i})
-                        p(\divTimeMapVector,
-                          \allParameters{}
-                          \given \divModel{i})
-                        p(\divModel{i})
-                        }{p(\alignmentVector)}
-                \]
-                \vspace{-1mm}
-                \end{minipage}
-            \end{displaybox}
-        \end{uncoverenv}
-    \end{minipage}
-\end{frame}
-
-\begin{frame}
-    \frametitle{New method: \dppmsbayes}
-    \begin{itemize}
-        % \item<1-> Reparameterized the model implemented in \msb
-        \item<1-> Dirichlet-process prior (DPP) over all possible divergence
-            models
-        \item<2-> Flexible priors on continuous parameters to avoid strongly
-            weighted posteriors
-        \item<3-> Multi-processing interface to improve performance for
-            accommodating genomic datasets
-        % \item<4-> Uniform prior over divergence models
-    \end{itemize}
     \barefootnote{\shortfullcite{Oaks2014dpp}}
@@
 
-\begin{frame}
-    \frametitle{\dppmsbayes: Performance}
-    \begin{itemize}
-        \item<1-> Extensive computer simulations to compare performance to previous
-            state-of-the-art methods (\msb).
-        \item<2-> New method for estimating shared evolutionary history shows
-            improved:
-            \begin{enumerate}
-                \item<2-> Improved model-choice accuracy 
-                \item<2-> More robustness to model violations
-                \item<2-> Greater power to detect variation in divergence times
-                \item<2-> Faster!
-            \end{enumerate}
-    \end{itemize}
-
-    \uncover<3->{
-    \textbf{\dppmsbayes allows biologists to leverage comparative genomic data
-        to assess the affects of community-scale processes on biodiversity}
-    }
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    % \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-
-        \vspace{6.5cm}
-
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \begin{uncoverenv}<1->
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-        \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1->[width=\textwidth]{../images/maps/Philippines.png}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
-    \smallskip
-    \centerline{
-    \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Current work: More data}
-    \begin{itemize}
-        \item<1-> Collecting genomic data from taxa co-distributed across Southeast
-            Asia
-        \item<2-> Preliminary results for 1000 loci from 5 pairs of \emph{Gekko
-                mindorensis} populations
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<2->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Current work: More power}
-    \centerline{
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/mtholder.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/vladimir.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../images/revbayes.png}}
-
-    \begin{itemize}[<+->]
-        \item Full-likelihood Bayesian implementation
-            \begin{itemize}
-                \item Uses all the information in the data
-                \item Applicable to deeper timescales
-            \end{itemize}
-        \item Analytically integrate over gene trees
-            \footnote{\tiny\shortfullcite{Bryant2012}}
-            \begin{itemize}
-                \item Very efficient numerical approximation of posterior 
-                \item Applicable to NGS datasets
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-\section{Conclusions}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification model?}
-    \begin{itemize}
-        \item<1-> Results consistent with prediction of clustered
-            divergences
-        \item<2-> Results suggest multiple co-divergences
-        \item<3-> However, there is a lot of uncertainty!
-    \end{itemize}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\section{Teaching}
-
-\begin{frame}
-    BIO 180 stuff
-\end{frame}
-
-\begin{frame}
-    Software carpentry \\
-
-    Workshops
-\end{frame}
-
-\begin{frame}
-    \frametitle{Questions?}    
-    \begin{center}
-        {
-        \Large
-        \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
-        }
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Future directions---Computational}
-    \begin{itemize}
-        \item<1-> Full-phylogenetic framework
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
-        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
-    \end{center}
-\end{frame}
-
-\begin{frame}
-    \centerline{
-    \includegraphics<1>[height=8cm]{../images/bd-cartoon-no-lines.pdf}
-    \includegraphics<2>[height=8cm]{../images/bd-cartoon-shared.pdf}
-    }
-\end{frame}
-
-\begin{frame}
-\begin{columns}
-    \column{.5\textwidth}
-    \begin{minipage}[c]{\columnwidth}
-    \begin{itemize}
-        \item Blah
-    \end{itemize}
-    \end{minipage}
-
-    \column{.6\textwidth}
-    \begin{minipage}[t][\textheight][c]{\linewidth}
-    \includegraphics<1>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
-    \includegraphics<2>[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
-    \end{minipage}
-\end{columns}
-\end{frame}
-
-\end{document}
+% \begin{frame}
+%     \frametitle{Results}
+%     \begin{itemize}
+%         % \item \textbf{dpp-msbayes allows biologists to leverage comparative
+%         %         genomic data to assess the affects of community-scale processes
+%         %         on biodiversity}
+%         \item<1-> Results consistent with prediction of clustered
+%             divergences
+%         \item<2-> Results suggest multiple interglacial periods were important
+%             drivers of diversification and community assembly across Islands of
+%             Southeast Asia
+%         \item<3-> However, there is a lot of uncertainty!
+%     \end{itemize}
+%     \barefootnote{\shortfullcite{Oaks2014dpp}}
+% \end{frame}

### 2015-03-05 slides/auburn/sampling-models.tex
Working on slides.

--- 2014-11-03 slides/osu/slides.tex
@@
-\input{../utils/preamble.tex}
-\input{../../manuscripts/utils/macros.tex}
-\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
-\bibliography{../../manuscripts/bib/references}
-
-\title[Improving inference of evolutionary history]{Improving inference of evolutionary history}
-\subtitle{Developing comparative methods for genomic data}
-
-\author[J.\ Oaks]{
-    Jamie R.\ Oaks\inst{1}
-}
-\institute[University of Washington]{
-    \inst{1}%
-        Department of Biology, University of Washington
-}
-
-% \date{\today}
-\date{November 3, 2014}
-
-\begin{document}
-% \maketitle
-\begin{frame}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-            \maketitle
-        \column{.5\textwidth}
-            \begin{figure}
-                \begin{center}
-                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
-                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
-                \end{center}
-            \end{figure}
-    \end{columns}
-\end{frame}
-
-% I. What I do
-% II. What I've been focusing on recently (started in PhD and continuing in postdoc)
-%   a. Dynamic planet---need to account for large-scale processes
-%       i. E.g., SE Asia---dynamic landscape
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
+\begin{frame}[t]
+    \frametitle{Sampling divergence models---previous approach}
     \begin{itemize}
-            \item<2-> Large-scale geological and climatic processes are
-                important in biodiversification and community assembly
-            \item<3-> Accounting for such processes will better our
-                understanding of biodiversity
-            \item<4-> We need methods for inferring evolutionary patterns
-                predicted by historical events from contemporary populations
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    \frametitle{Southeast Asia}    
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Southeast Asia}    
-    \begin{columns}
-        \column{0.6\textwidth}
-        \ \\
-
-        \column{0.45\textwidth}
-
-        \vspace{-2.4cm}
-
-        \begin{uncoverenv}<2->
-        \colorbox{white}{
-            \begin{minipage}[t][3.6cm][t]{1.1\textwidth}
-                \begin{onlyenv}<2-3>
-                \begin{itemize}
-                    \item<2-> Repeated coalescence and fragmentation of island complexes
-                    \item<3-> Prominent paradigm for explaining biodiversity
-                \end{itemize}
-                \end{onlyenv}
-
-                \begin{onlyenv}<4>
-                \begin{itemize}
-                    \item<4> \textbf{Did repeated fragmentation of islands
-                        during inter-glacial rises in sea level promote
-                        diversification?}
-                \end{itemize}
-                \end{onlyenv}
-
-                \begin{onlyenv}<5->
-                \begin{itemize}
-                    \item<5-> Climate-driven model of diversification
-                    \item<6-> Testable prediction:
-                    \begin{itemize}
-                        \item Temporally clustered divergences among taxa
-                            co-distributed across fragmented islands
-                    \end{itemize}
-                \end{itemize}
-                \end{onlyenv}
-            \end{minipage}
-        }
-    \end{uncoverenv}
-    \end{columns}
-\end{frame}
-}
-
-% \begin{frame}
-%     \frametitle{Philippine Archipelago}
-%     \begin{center}
-%         \includegraphics<1>[width=.5\textwidth]{../images/maps/Philippines-present.png}
-%         \includegraphics<2>[width=.5\textwidth]{../images/maps/Philippines.png}
-%     \end{center}
-% \end{frame}
-
-% \begin{frame}
-%     \frametitle{Climate-driven diversification model}
-%     \begin{columns}[c]
-%         \column{.5\textwidth}
-%             \begin{myitemize}
-%                 \item<1-> Repeated coalescence and fragmentation of island complexes
-%                 \item<2-> Prominent paradigm for explaining Philippine biodiversity
-%                 \item<3-> Proposed as model of diversification
-%             \end{myitemize}
-%         \column{.5\textwidth}
-%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
-%     \end{columns}
-% \end{frame}
-
-% \begin{frame}
-%     \frametitle{Testing climate-driven diversification}
-%     \begin{columns}[c]
-%         \column{.5\textwidth}
-%     \textbf{Did repeated fragmentation of islands during inter-glacial rises in sea level
-%     promote diversification?}\\
-%     \bigskip
-%     \uncover<2->{
-%     Model has testable prediction:\\
-%     \begin{itemize}
-%         \item Temporally clustered divergences among taxa
-%             co-distributed across fragmented islands
-%     \end{itemize}
-%     }
-%         \column{.5\textwidth}
-%             \includegraphics[width=\textwidth]{../images/maps/Philippines.png}
-%     \end{columns}
-% \end{frame}
-
-%   b. Predictions of climate-driven diversification
-\begin{frame}
-    \frametitle<1-4>{Community scale processes}
-    \frametitle<5->{Divergence model choice}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{onlyenv}<1-4>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<5>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{\divTime{1}\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 111
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (\divTimeMap{1},\divTimeMap{2},\divTimeMap{3})
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<6>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 111
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (130,130,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<7>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 211
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (397,130,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<8>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 121
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (130,397,130)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<9>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 112
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeVector = \{130,397\}
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{onlyenv}<10-12>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{displaybox}[0.95\linewidth]
-                    \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-                        \[
-                            \divTimeVector = \{130,397,260\}
-                        \]\vspace{0mm}
-                        \[
-                            model = 312
-                        \]\vspace{0mm}
-                        \[
-                            \divTimeMapVector = (260,130,397)
-                        \]\vspace{0mm}
-                    \end{minipage}
-                \end{displaybox}
-            \end{minipage}
-        \end{onlyenv}
-        % \begin{onlyenv}<11>
-        %     \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-        %         \begin{displaybox}[0.95\linewidth]
-        %             \begin{minipage}[c][0.45\textheight][c]{0.95\linewidth}
-        %                 \[
-        %                     \divTimeVector = \{\divTime{1},\ldots,\divTime{\divTimeNum}\}
-        %                 \]\vspace{0mm}
-        %                 \[
-        %                     model = \divModel{i}
-        %                 \]\vspace{0mm}
-        %                 \[
-        %                     \divTimeMapVector = (\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})
-        %                 \]\vspace{0mm}
-        %             \end{minipage}
-        %         \end{displaybox}
-        %     \end{minipage}
-        % \end{onlyenv}
-        \begin{onlyenv}<13>
-            \begin{minipage}[c][0.5\textheight][c]{\linewidth}
-                \begin{mydescription}
-                    \item[\alignmentVector] Sequence alignments
-                    \item[\divTimeMapVector] Divergence times
-                    \item[\divModel{}] Divergence model
-                    \item[\allParameters{}] Gene trees
-                    \item[\ ] Substitution parameters
-                    \item[\ ] Demographic parameters
-                    % \item[\allParameters] $(\divTimeMapVector, \geneTreeVector,
-                    %     \hkyModelVector, \demographicParamVector)$
-                \end{mydescription}
-            \end{minipage}
-        \end{onlyenv}
-        \begin{uncoverenv}<11->
-            \begin{minipage}[c][0.18\textheight][c]{\linewidth}
-            \begin{flushleft}
-                \small We want to infer \textcolor{blue}{\divModel{}} and
-                \textcolor{blue}{\divTimeMapVector} given DNA sequence
-                alignments
-                \textcolor{blue}{\alignmentVector}
-            \end{flushleft}
-            \end{minipage}
-        \end{uncoverenv}
-        \column{.6\textwidth}
-        \begin{minipage}[t][0.8\textheight][c]{\linewidth}
-        \includegraphics<1>[height=6cm]{../images/dmc-cartoon-islands.pdf}
-        \includegraphics<2>[height=6cm]{../images/dmc-cartoon-taxa.pdf}
-        \includegraphics<3>[height=6cm]{../images/dmc-cartoon-sea-level.pdf}
-        \includegraphics<4>[height=6cm]{../images/dmc-cartoon-shared-no-labels.pdf}
-        \includegraphics<5-6>[height=6cm]{../images/dmc-cartoon-shared.pdf}
-        \includegraphics<7>[height=6cm]{../images/dmc-cartoon-2-1.pdf}
-        \includegraphics<8>[height=6cm]{../images/dmc-cartoon-2-2.pdf}
-        \includegraphics<9>[height=6cm]{../images/dmc-cartoon-2-3.pdf}
-        \includegraphics<10-11>[height=6cm]{../images/dmc-cartoon-general.pdf}
-        \includegraphics<12->[height=6cm]{../images/dmc-cartoon-pop-general.pdf}
-        \end{minipage}
-    \end{columns}
-\end{frame}
-
-%   c. Testing for predicted patterns via model choice.
-%       i. Bayesian model choice
-
-\begin{frame}[t]
-    \frametitle{Bayesian model choice}
-    \begin{block}{\it Full model:}
-        \begin{minipage}[c][3.8cm][c]{\linewidth}
-            \begin{uncoverenv}<1->
-            \[
-                p(
-                  \divTimeMapVector,
-                  \allParameters{}
-                  \given \alignmentVector, \divModel{i})
-                  =
-                \frac{
-                    p(\alignmentVector \given
-                      \divTimeMapVector,
-                      \allParameters{},
-                      \divModel{i}
-                      )
-                      p(
-                        \divTimeMapVector,
-                        \allParameters{}
-                        \given \divModel{i}
-                        )
-                    }{p(\alignmentVector \given \divModel{i})}
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-            \begin{uncoverenv}<2->
-            \[
-                p(\alignmentVector \given \divModel{i}) =
-                \int_{\divTimeMapVector} \int_{\allParameters{}}
-                p(\alignmentVector \given \divTimeMapVector, \allParameters{}, \divModel{i})
-                p(\divTimeMapVector, \allParameters{} \given \divModel{i})
-                d\divTimeMapVector d\allParameters{}
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-            \begin{uncoverenv}<3->
-            \[
-                p(\divModel{i} \given \alignmentVector) =
-                \frac{
-                    p(\alignmentVector \given \divModel{i})
-                    p(\divModel{i})
-                }{
-                    \sum_{i} p(\alignmentVector \given \divModel{i})
-                    p(\divModel{i})
-                }
-            \]\vspace{-1mm}
-            \end{uncoverenv}
-        \end{minipage}
-    \end{block}
-    \vspace{1mm}
-    \begin{uncoverenv}<4>
-        \begin{center}
-            \msb: Approximate Bayesian computation (ABC)
-
-            % \vspace{-5mm}
-            % \[ \alignmentVector \, \to \, \ssVectorObs \, \to \, \ssSpace\]
-        \end{center}
-    \end{uncoverenv}
-    \barefootnote{\tiny\shortfullcite{Huang2011}. \hspace{2mm}\shortfullcite{Oaks2012}.}
-\end{frame}
-
-%   d. Method exists---applied it to comparative dataset from Philippines
-%       i. Surprising results!
-%       ii. Rather than publish, did some detective work
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{onlyenv}<1>
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/rafe.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/cam.jpg}}
-        \centerline{
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/charles.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/jake.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-        \begin{onlyenv}<2->
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-        \end{onlyenv}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1-2>[width=\textwidth]{../images/maps/Philippines.png}
-        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
-        \begin{onlyenv}<4->
-            \begin{itemize}
-                \item Strong support for simultaneous divergence of all 22 taxon pairs
-
-                \item Posterior probability $>$ 0.96 \\
-
-                \item $\sim$100,000--250,000 years ago
-            \end{itemize}
-        \end{onlyenv}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-%   e. Evaluated method via simulation
-%       i. biased in alarming direction
-
-\section{Simulation-based assessment of method}
-
-\begin{frame}
-    \frametitle{Simulation-based power analyses}
-    % What is ``simultaneous''?
-    \begin{myitemize}
-        \item<1-> What if divergences were random?
-        \item<2-> Simulate datasets in which all 22 divergence times are random
-        \uncover<3->{
-            \smallskip
-            \begin{myitemize}
-                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
-                \smallskip
-            \end{myitemize}
-        \item<3-> $MGA$ = Millions of Generations Ago
-        }
-        \item<4-> Simulate 1000 datasets for each \divTime{} distribution
-        \item<4-> Analyze all 4000 datasets as we did the empirical data
-    \end{myitemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Simulation-based power analyses: Results}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]{../images/old_old_power_psi_mode.pdf}}
-
-    \begin{uncoverenv}<2->
-        \begin{itemize}[<+->]
-            \item \msb will often infer clustered divergences when divergences
-                are random over millions of generations.
-
-            \item The empirical results are likely spurious
-        \end{itemize}
-    \end{uncoverenv}
-
-    \begin{uncoverenv}<3->
-        \begin{block}{Objective:}
-            Use principles of probability to extend \msb framework for improved
-            estimation of shared evolutionary history
-        \end{block}
-    \end{uncoverenv}
-    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
-\end{frame}
-
-%   f. Rather than abandon approach---look for theoretical reasons for behavior
-%       i. Strongly weighted posterior (marginal likelihood)
-%       ii. Combinatorics
-%   g. Developed new method---dpp-msbayes
-%       i. Used DPP---common in Bayeisn nonparametrics
-%           1. Flexible and mathematically convenient way of assigning prior
-%           probability to all possible divergence models (partitions) directly
-%           (avoid combinatoric issue)
-%       ii. More flexible priors on parameters of model
-%       iii. Implemented method in much more efficient multi-processing
-%       framework.
-%   e. Results---improved!
-%   f. Empirical results...
-%       i. Reducing bias and increasing power provides more "honest" estimate
-%       of uncertainty
-%       ii. A lot of uncertainty (Good! and Bad!)
-%       iii. What next? More data and more power!
-
-\begin{frame}[label=improvements]
-    \frametitle{An improved method}
-    Potential improvements:
-    \begin{enumerate}
-        \item Alternative priors on parameters that increase marginal
-            likelihoods of rich models
-        \item Alternative approach to modeling the temporal distribution of
-            divergences
-    \end{enumerate}
-    \barefootnote{\tiny\shortfullcite{Oaks2012}. \hspace{2mm}\shortfullcite{Oaks2014reply}}
-\end{frame}
-
-\begin{frame}[t]
-    \vspace{-2mm}
-    \begin{displaybox}[5.5cm]
-        \small
-        \[
-            p(X) = \int_{\theta} p(X \given \theta) p(\theta) \mathrm{d}\theta
-        \]%\vspace{0mm}
-    \end{displaybox}
-
-    \vspace{-1mm}
-    \begin{center}
-        % \includegraphics<1>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<2>[height=7.8cm]{../images/marginal-plot-3d.png}
-        \includegraphics<3>[height=7.8cm]{../images/marginal-plot-2d-uniform-prior.pdf}
-        \includegraphics<4>[height=7.8cm]{../images/marginal-plot-2d.pdf}
-    \end{center}
-\end{frame}
-
-\againframe{improvements}
-
-\begin{frame}[t]
-    \frametitle{Prior on divergence models}
-    \begin{itemize}
-        \item \msb uses a discrete uniform prior on the \emph{number} of
-            divergence events
+        \item<2-> Place a discrete uniform prior on the
+            \textit{\textbf{number}} of events
     \end{itemize}
@@
             \centerline{
-                \includegraphics<2->[width=1.0\textwidth]{../images/number-of-div-models-22-unordered.pdf}}
+                \includegraphics<2->[width=1.0\textwidth]{../images/number-of-div-models-10.pdf}}
         \column{0.5\textwidth}
             \centerline{
-                \includegraphics<3->[width=1.0\textwidth]{../images/prob-of-div-models-22-unordered.pdf}}
+                \includegraphics<3->[width=1.0\textwidth]{../images/prob-of-div-models-10.pdf}}
     \end{columns}
@@
 
-    \begin{uncoverenv}<4->
+    \begin{itemize}
+        \item<4-> This can create a strong bias toward models of co-divergence
+    \end{itemize}
+
+    \begin{uncoverenv}<5->
     \begin{block}{\it Potential solution:}
@@
 \end{frame}
-
-\begin{frame}
-    \frametitle{New method: \dppmsbayes}
-    \begin{itemize}[<+->]
-        % \item<1-> Reparameterized the model implemented in \msb
-        \item<1-> Replaced uniform priors on continuous parameters with gamma and
-            beta distributions
-        \item<2-> Dirichlet process prior (DPP) over all possible divergence
-            models
-        \item<3-> Develop multi-processing interface to improve performance for
-            accommodating genomic datasets
-        % \item<4-> Uniform prior over divergence models
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Simulation-based assessment}
-        \uncover<1->{
-        Validation:\\
-
-        \smallskip
-        \begin{itemize}
-            \item Simulate 50,000 datasets and analyze each under the same
-                model
-        \end{itemize}
-        }
-
-        \uncover<2->{
-        \bigskip
-        Robustness:\\
-        \begin{itemize}
-            \item Simulate datasets that violate model assumptions and analyze
-                each of them
-        \end{itemize}
-        }
-
-        \uncover<3->{
-        \bigskip
-        Done for \msb and \dppmsbayes
-        }
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Validation results}
-    \begin{columns}
-        \column{0.5\textwidth}
-            \includegraphics<1->[width=1.0\textwidth]{../images/validation-model-choice-old.pdf}
-        \column{0.5\textwidth}
-            \includegraphics<2->[width=1.0\textwidth]{../images/validation-model-choice-dpp.pdf}
-    \end{columns}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Robustness results}
-    \begin{columns}
-        \column{0.5\textwidth}
-            \centering{
-            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-old-violations.pdf}
-            }
-        \column{0.5\textwidth}
-            \centering{
-            \includegraphics<1->[width=0.9\textwidth]{../images/validation-model-choice-dpp-violations.pdf}
-            }
-    \end{columns}
-
-    \begin{uncoverenv}<2->
-    \begin{itemize}
-        \item Improved model-choice accuracy and robustness
-    \end{itemize}
-    \end{uncoverenv}
-    \vspace{-0.5cm}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{\dppmsbayes: Simulation-based power analyses}
-    Does the new method have power to detect variation in divergence times?\\
-
-    \bigskip
-    \begin{myitemize}
-        \item<2-> Simulate datasets in which all 22 divergence times are random
-            \smallskip
-            \begin{myitemize}
-                \item $\divTime{} \sim U(0, \,0.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,1.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,2.5 \,MGA)$
-                \smallskip
-                \item $\divTime{} \sim U(0, \,5.0 \,MGA)$
-                \smallskip
-            \end{myitemize}
-        \item<2-> $MGA$ = Millions of Generations Ago
-
-        \item<3-> Simulate 1000 datasets for each \divTime{} distribution
-        \item<3-> Analyze all 4000 datasets with \msb and \dppmsbayes
-    \end{myitemize}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{\dppmsbayes: Power results}
-        \begin{onlyenv}<1-2>
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<1-2>{../images/power-psi-old-annotate-no-pp.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<2>{../images/power-psi-dpp-annotate-no-pp.pdf}}
-        \end{onlyenv}
-
-        \begin{onlyenv}<3->
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-old-annotate-pp.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=1.13\textwidth]<3->{../images/power-psi-dpp-annotate-pp.pdf}}
-        \end{onlyenv}
-
-    \begin{uncoverenv}<4->
-    \begin{itemize}
-        \item Improved power to detect temporal variation across divergences
-    \end{itemize}
-    \end{uncoverenv}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{Climate-driven diversification}    
-    \begin{columns}
-        \column{0.6\textwidth}
-        \ \\
-
-        \column{0.4\textwidth}
-
-        \vspace{-2cm}
-
-        \colorbox{white}{
-            \begin{minipage}[t]{1.0\textwidth}
-                \raggedright
-                \textbf{Did repeated fragmentation of islands during
-                    inter-glacial rises in sea level promote diversification?}
-            \end{minipage}
-        }
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Climate-driven diversification}
-\begin{columns}[c]
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
-        % \vspace{0.5mm}
-        \centerline{
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
-        \hspace{0.3mm}
-        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
-        \end{minipage}
-
-    \column{.5\textwidth}
-        \vspace{-1cm}
-        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
-        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
-        \end{minipage}
-\end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Empirical results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp.pdf}}
-
-    \begin{uncoverenv}<2->
-    \begin{itemize}
-        \item More ``honest'' estimate of posterior uncertainty
-    \end{itemize}
-    \end{uncoverenv}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Empirical results}
-    \centerline{
-    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
-    \smallskip
-    \centerline{
-    \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Summary}
-    \begin{itemize}
-        \item<1-> New method for estimating shared evolutionary history shows
-            improved:
-            \begin{enumerate}
-                \item<2-> Estimation of posterior uncertainty
-                \item<2-> Model-choice accuracy
-                \item<2-> Power to detect temporal variation across divergences
-                \item<2-> Robustness to model violations
-                \item<2-> Speed!
-            \end{enumerate}
-
-        \smallskip
-        \item<3-> Climate-driven diversification model?
-            \begin{itemize}
-                \item<4-> Results consistent with prediction of clustered
-                    divergences
-                \item<4-> Results suggest multiple co-divergences
-                \item<4-> However, there is a lot of uncertainty!
-            \end{itemize}
-
-        \smallskip
-        \item<5-> What to do?
-            \begin{itemize}
-                \item<6-> More data!
-                \item<6-> More power! 
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-%   g. Collected thousands of loci from across the genome of almost 300
-%   individuals of two genera of geckos from across the Philippine Islands
-%       i. Prelim results with dpp-msbayes (can accommodate NGS data!)
-
-\begin{frame}
-    \frametitle{More data}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{myitemize}
-            \item Collecting genomic data from two gekkonid radiations
-                co-distributed Southeast Asia
-            \begin{myitemize}
-                \item \emph{Cyrtodactylus}---12+ species
-                \item \emph{Gekko}---13+ species
-            \end{myitemize}
-            \item Whole genome assemblies for each genus
-            \item RAD-seq libraries constructed for $\approx150$ individuals
-                from each genus
-        \end{myitemize}
-        \centerline{
-        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/misc/lee-and-me.jpg}
-        \hspace{2mm}
-        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/people/rafe.jpg}}
-        \column{.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/photos/cyrt-agusanensis.jpg} \quad
-        \includegraphics[width=\textwidth]{../images/photos/gekko-mindorensis.jpg}
-    \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{More data---preliminary results}
-    Results for 1000 loci from 5 pairs of \emph{Gekko mindorensis} populations\\
-
-    \begin{center}
-        \includegraphics<1->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
-    \end{center}
-\end{frame}
-
-%   h. Currently working on full-likelihood implementation of the model.
-%       i. Working with Jeet---has full-likelihood up and running---inefficient
-%       ii. Take advantage of recent developments to analytically integrate
-%       gene trees---might even be faster than simulations?!
-%       iii. Uses ALL information in genetic data
-
-\begin{frame}
-    \frametitle{More power}
-    \centerline{
-    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/mtholder.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/jeet2.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/vladimir.jpg}
-    \hspace{0.6mm}
-    \includegraphics<1->[height=2cm]{../images/revbayes.png}}
-
-    \begin{itemize}[<+->]
-        \item Full-likelihood Bayesian implementation \footnote{\tiny\shortfullcite{JeetDiss}}
-            \begin{itemize}
-                \item Uses all the information in the sequence data
-                \item Applicable to much deeper timescales
-                \item \ldots Currently inefficient
-            \end{itemize}
-        \item Working to incorporate new developments on analytical integration
-            of gene trees \footnote{\tiny\shortfullcite{Bryant2012}}
-            \begin{itemize}
-                \item Very efficient numerical approximation of posterior 
-                \item Applicable to NGS datasets
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
-\section{Future directions}
-
-% III. Future work
-%   a. Theory/computation
-%       i. Extend full-likelihood approach to phylogenetic framework
-%           1. Can infer relative mutation rates/pop sizes from data (fewer
-%           assumptions)
-%           2. Much better representation of diversification than pairwise
-%           approach.
-%       ii. Parameter estimation approach (Lindley's paradox).
-
-\begin{frame}
-    \frametitle{Future directions---Computational}
-    \begin{itemize}
-        \item<1-> Full-phylogenetic framework
-    \end{itemize}
-
-    \begin{center}
-        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
-        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
-    \end{center}
-\end{frame}
-
-%   b. Empirical
-%       i. Gobi (NGS --> NSF)
-%       ii. West-Central Africa
-%           1. unfunded collaborator
-%       iii. Compare diversification of mainland SE Asia to Oceanic Islands
-%           1. Planning pre-proposal
-%       iv. Long-term study of small island of sunda shelf
-%           1. Tiny islands with pops of Eutropis multifasciata and
-%           Lepidodactylus lugubris
-%           2. Inundated during last interglacial ~120,000 years ago, spent the
-%           better part of next 100k years as part of exposed sunda shelf, then
-%           fragmented during current interglacial
-%           3. Both species vary in reproductive mode (oviparity/ovoviviparity,
-%           parthenogenesis), skinks in color patterns
-%           4. Affects of recent, severe fragmentation on diversification and
-%           demography
-%           5. Comparative genomics to understand evolution of color and
-%           reproductive mode
-%           6. Evolutionary ecology---what ecological context influence
-%           evolution asexuality/live-bearing/color.
-{
-\setbeamercolor{background canvas}{bg=black}
-\setbeamercolor{whitetext}{fg=white}
-\begin{frame}
-    \frametitle{Processes of diversification}
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        {\usebeamercolor[fg]{whitetext}
-        \begin{itemize}
-            \item<1-> These comparative tools will allow evolutionary
-                biologists to leverage genomic data to infer shared
-                evolutionary history
-            \item<2-> This will better our understanding of how community scale
-                processes affect the generation and assembly of biodiversity
-            \item<3-> Many exciting empirical systems for applying new methods!
-        \end{itemize}
-        }
-        \column{.5\textwidth}
-        \begin{figure}
-            \begin{center}
-            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
-            \end{center}
-        \end{figure}
-    \end{columns}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Future directions: Gobi Desert}
-    \vspace{-2mm}
-    \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
-    \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Future directions: Gobi Desert}
-    \vspace{-2mm}
-    \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
-    \end{columns}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Future directions: West-Central Africa}
-    \begin{center}
-        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
-        \hspace{0.3mm}
-        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure4_herps.jpeg}
-    \end{center}
-
-    \only<1>{
-    \begin{center}
-        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/adam.png}
-        \hspace{0.3mm}
-        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/fujita.jpg}
-    \end{center}
-    }
-
-    \uncover<2->{
-    \begin{center}
-        \includegraphics[height=4.5cm]{../images/number-of-divs-africa.pdf}
-    \end{center}
-    }
-\end{frame}
-
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-    \vspace{-2mm}
-    \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/malaysia/malaysia-burungs.jpg}
-    \end{columns}
-\end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-\end{frame}
-}
-
-\begin{frame}
-    \frametitle{Acknowledgments}
-    \begin{columns}[t]
-        \column{.5\textwidth}
-            {\bf Ideas and feedback:}
-            \begin{myitemize}
-                \item Leach\'{e} Lab
-                \item Minin Lab
-                \item Holder Lab
-                \item Brown Lab/KU Herpetology
-                \item Melissa Callahan
-            \end{myitemize}
-            \smallskip
-            {\bf Computation:}\\
-            \includegraphics<1->[height={8mm}]{../images/iplant.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/kuittc.png}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/uw.png}\\
-
-        \column{.5\textwidth}
-            {\bf Funding:}\\
-            \includegraphics<1->[height={8mm}]{../images/nsf.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/ngs.jpg}
-            \hspace{0.5mm}
-            \includegraphics<1->[height={8mm}]{../images/ssb.png}\\
-
-            \smallskip
-            {\bf Photo credits:}
-            \begin{myitemize}
-                \item Rafe Brown, Cam Siler, Jesse Grismer, \& Jake Esselstyn
-                \item FMNH Philippine Mammal Website:
-                    \begin{myitemize}
-                        \item D.S.\ Balete, M.R.M.\ Duya, \& J.\ Holden
-                    \end{myitemize}
-                \item PhyloPic!
-            \end{myitemize}
-    \end{columns}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Everything is on GitHub\ldots}
-    Software:\\
-    \begin{itemize}
-        \item \texttt{dpp-msbayes}:
-            \href{https://github.com/joaks1/dpp-msbayes}{\tt
-            https://github.com/joaks1/dpp-msbayes}
-
-        \item \texttt{PyMsBayes}:
-            \href{https://github.com/joaks1/PyMsBayes}{\tt
-            https://github.com/joaks1/PyMsBayes}
-
-        \item ABACUS: Approximate BAyesian C UtilitieS.
-            \href{https://github.com/joaks1/abacus}{\tt
-            https://github.com/joaks1/abacus}
-    \end{itemize}
-
-    \medskip
-    Open-Science Notebook:\\
-    \begin{itemize}
-        \item \texttt{msbayes-experiments}:
-            \href{https://github.com/joaks1/msbayes-experiments}{\tt
-            https://github.com/joaks1/msbayes-experiments}
-    \end{itemize}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Questions?}    
-    \begin{columns}[c]
-        \column{.5\textwidth}
-        \begin{center}
-            {
-            \Large
-            \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
-            }
-        \end{center}
-        \column{.5\textwidth}
-            \begin{figure}
-                \begin{center}
-                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
-                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
-                \end{center}
-            \end{figure}
-    \end{columns}
-\end{frame}
-
-% Extra slides
-
-\begin{frame}[noframenumbering]
-    \frametitle{Easy as ABC}
-    \centerline{
-    \includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}
-    \includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}
-    \includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}
-    \includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}
-    \includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}
-    \includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}
-    \includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}
-    \includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}
-    \includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}
-    \includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}
-    }
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{Causes of bias: Insufficient sampling}
-    \begin{itemize}
-        \item Models with more parameter space are less densely sampled
-        \item Could explain bias toward small models in extreme cases
-        \item {\bf Predicts large variance in posterior estimates}
-        \begin{itemize}
-            \item We explored empirical and simulation-based analyses with
-                2, 5, and 10 million prior samples, and estimates were
-                very similar
-        \end{itemize}
-    \end{itemize}
-    \smallskip
-    \centerline{
-    \includegraphics[width=\textwidth]{../images/omega_over_sampling.pdf}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{\dppmsbayes: Simulation results}
-    % \vspace{1cm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_old_power_psi_mode.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_psi_mode_headless.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_dpp_power_psi_mode_headless.pdf}}
-    % \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{\dppmsbayes: Simulation results}
-    % \vspace{1cm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_old_power_omega_median.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_u-shaped_power_omega_median_headless.pdf}}
-        \vspace{0mm}
-        \centerline{
-        \includegraphics[width=\textwidth]{../images/old_dpp_power_omega_median_headless.pdf}}
-    % \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\begin{frame}[noframenumbering]
-    \frametitle{Empirical results: Philippine diversification}
-    \centerline{
-    \includegraphics[width=0.8\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-old-vs-dpp-with-prior.pdf}}
-    \barefootnote{\shortfullcite{Oaks2014dpp}}
-\end{frame}
-
-\end{document}

### 2015-03-05 slides/images/pull-images.sh

@@ -10,6 +10,7 @@ else
     base_dir="`pwd`"
 fi
 
+curl -o peter-dirichlet.jpg http://upload.wikimedia.org/wikipedia/commons/3/32/Peter_Gustav_Lejeune_Dirichlet.jpg
 curl -o darwin-tol-copyright-boris-kulikov-2007.jpg http://www.boriskulikov.com/images/DarwinTreeOfLife-%20L.jpg
 curl -o kane.jpg http://uwberg.com/wp-content/uploads/2014/05/berg-hero.jpg
 curl -o mascot-tulane-green-wave.png http://upload.wikimedia.org/wikipedia/en/9/9e/TulaneGreenWave.png

### 2015-03-06 slides/auburn/intro-mascot-tree.tex
Working on slides.

--- 2015-03-03 slides/auburn/intro-mascot-tree.tex
@@
-\begin{frame}
+\begin{frame}[t]
+\frametitle{Current state of phylogenetics}
+    \vspace{-3mm}
+    \uncover<2->{
+    \begin{adjustwidth}{-3.1em}{-3.1em}
+        \begin{itemize}
+            \item<2-> \textbf{Limiting assumption:} Events are independent across the tree
+            % \item<3-> We know this assumption is frequently violated
+        \end{itemize}
+    \end{adjustwidth}
+    }
+
+    \vspace{-5mm}
+
+    \begin{onlyenv}<1-2>
     \begin{center}
@@
         % \includegraphics<2>[width=\textwidth]{../images/mascot-tree.png}
-        \includegraphics[width=\textwidth]{../images/mascot-tree.png}
+        \includegraphics[width=\textwidth]{../mascot-tree/mascot-tree-shared.pdf}
 
     \end{center}
+    \end{onlyenv}
+
+    % \vspace{-1cm}
+    \begin{onlyenv}<3->
+    \begin{columns}
+        \column{0.4\textwidth}
+        \begin{minipage}[c]{\columnwidth}
+        \begin{adjustwidth}{-2.1em}{}
+        % \begin{onlyenv}<1-6>
+        \begin{itemize}
+            % \item<2-> Current phylogenetic methods assume events are independent 
+            \item<3-> We know this assumption is frequently violated
+            \item<5-> Why account for this non-independence?
+            \begin{enumerate}
+                \item<6-> Improve phylogenetic inference
+                \item<7-> \textbf{Provide a general framework for inferring the
+                        effect of community-scale processes on diversification}
+            \end{enumerate}
+        \end{itemize}
+        % \end{onlyenv}
+        \end{adjustwidth}
+        \end{minipage}
+
+        \column{0.6\textwidth}
+        \begin{minipage}[t][\textheight][c]{\linewidth}
+        \centerline{
+        \includegraphics<3>[width=1.1\columnwidth]{../mascot-tree/mascot-tree-shared.pdf}
+        \includegraphics<4->[width=1.1\columnwidth]{../mascot-tree/mascot-tree-shared-events.pdf}
+        }
+        \end{minipage}
+    \end{columns}
+    \end{onlyenv}
 \end{frame}

### 2015-03-07 slides/auburn/dpp-intro.tex
Working on slides.

@@
         \item The divergence models are ways of assigning our taxa to events
-        \item A Dirichlet process prior (DPP) model is a ``natural'' fit
+        \item A Dirichlet process prior (DPP) model is convenient and flexible
+            solution
             \begin{itemize}[<+->]
-                \item Common nonparametric Bayesian approach to assigning
-                    variables to an unknown number of categories
+                \item Common Bayesian approach to assigning variables to an
+                    unknown number of categories
+                \item Controlled by ``concentration'' parameter: $\alpha$
             \end{itemize}
@@
     \begin{uncoverenv}<2->
+    \begin{figure}
     \begin{center}
         \includegraphics[height=0.4\textheight]{../images/peter-dirichlet.jpg}
+        \caption{Peter Dirichlet}
     \end{center}
+    \end{figure}
     \end{uncoverenv}

### 2015-03-07 slides/auburn/full-phylo.tex
Working on slides.

@@
 \begin{frame}[t]
-    \frametitle{New methods for inferring community history}
+    \frametitle{Next step: A general framework}
     \vspace{-5mm}
@@
 \begin{frame}
+    \frametitle{Next step: A general framework}
+
+    \vspace{-5mm}
     \begin{columns}
@@
         \begin{adjustwidth}{-2.3em}{}
-        \begin{onlyenv}<1-6>
+        \begin{onlyenv}<1-4>
         \begin{itemize}
-            \item<2-> Current phylogenetic methods assume speciation and extinction
-                occur independently of one another
-            \item<3-> We know this assumption is frequently violated
-            \item<4-> I will generalize a birth-death model to accommodate
-                ``shared'' events
-            \item<6> Infer events jointly across multiple trees
+            \item<1-> Relax the assumption of independence of events across the
+                tree
+            \item<2-> I will generalize a birth-death model to accommodate
+                shared events
+            \item<4-> Jointly infer co-diversification across multiple trees
         \end{itemize}
         \end{onlyenv}
-        \begin{onlyenv}<7->
-        \only<7->{\hspace{0.5cm}\textbf{\large Benefits:}\\}
+        \begin{onlyenv}<5->
+        \only<5->{\hspace{0.5cm}\textbf{\large Benefits:}\\}
         \begin{itemize}
-            \item<8-> Improve phylogenetic inference and divergence-time estimation
-            \item<9-> Provide a general framework for inferring the evolutionary
-                history of whole communities
-            \item<10-> Possible epidemiological applications
+            \item<6-> Improve phylogenetic inference and divergence-time estimation
+            \item<7-> Provide a general framework for inferring shared
+                evolutionary history across communities of species
+            \item<8-> Epidemiological applications
         \end{itemize}
@@
         \centerline{
-        \includegraphics<1-3>[width=\columnwidth]{../images/bd-cartoon-no-lines.pdf}
-        \includegraphics<4>[width=\columnwidth]{../images/bd-cartoon-shared.pdf}
-        \includegraphics<5>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
-        \includegraphics<6->[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
+        \includegraphics<1>[width=\columnwidth]{../images/bd-cartoon-no-lines.pdf}
+        \includegraphics<2>[width=\columnwidth]{../images/bd-cartoon-shared.pdf}
+        \includegraphics<3>[width=0.7\columnwidth]{../images/bd-cartoon-shared-space.pdf}
+        \includegraphics<4->[width=0.7\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
         }

### 2015-03-07 slides/auburn/future-empirical.tex
Working on slides.

@@
+% \begin{frame}
+%     \frametitle{Empirical applications: Gobi Desert}
+%     \vspace{-2mm}
+%     \begin{columns}
+%         % \column{\dimexpr\paperwidth-10pt}
+%         \column{\dimexpr\paperwidth}
+%         \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
+%     \end{columns}
+% \end{frame}
+
+% \begin{frame}
+%     \frametitle{Empirical applications: Gobi Desert}
+%     \vspace{-2mm}
+%     \begin{columns}
+%         % \column{\dimexpr\paperwidth-10pt}
+%         \column{\dimexpr\paperwidth}
+%         \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
+%     \end{columns}
+% \end{frame}
+
 \begin{frame}
-    \frametitle{Future directions: Gobi Desert}
-    \vspace{-2mm}
+    \frametitle{Empirical applications: West-Central Africa}
     \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
+        \column{0.5\linewidth}
+        % \begin{minipage}[c][\textheight][c]{\columnwidth}
+        \begin{itemize}
+            \item Did climate cycles drive diversification and community
+                assembly across rainforest taxa?
+            \item<2-> Funded by NSF
+            \item<3-> Preliminary results with 300 loci from 3 taxa
+        \end{itemize}
+
+        \begin{minipage}[c][0.45\textheight][c]{\columnwidth}
+        \begin{center}
+            \only<1-2>{
+            \includegraphics[height={2.5cm}]{/home/jamie/Dropbox/field-photos/people/adam.png}
+            \hspace{0.3mm}
+            \includegraphics[height={2.5cm}]{/home/jamie/Dropbox/field-photos/people/fujita.jpg}
+            }
+
+            \includegraphics<3->[width=0.9\columnwidth]{../images/number-of-divs-africa.pdf}
+        \end{center}
+        \end{minipage}
+
+        \column{0.5\linewidth}
+        % \begin{minipage}[c][\textheight][c]{\columnwidth}
+        {\centering
+            \includegraphics[width=\columnwidth]{/home/jamie/Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
+
+            \vspace{5mm}
+
+            \includegraphics[width=\columnwidth]{/home/jamie/Dropbox/field-photos/africa/Figure4_herps.jpeg}
+
+        }
+        % \end{minipage}
     \end{columns}
@@
 \begin{frame}
-    \frametitle{Future directions: Gobi Desert}
-    \vspace{-2mm}
+    \frametitle{Empirical applications: Gobi Desert}
     \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
+        \column{0.5\linewidth}
+        % \begin{minipage}[c][\textheight][c]{\columnwidth}
+        \begin{itemize}
+            \item Did the formation of the Gobi cause co-diversification
+                across rock and sand-adapted communities of lizards?
+            \item<2-> Funding from National Geographic Society to
+                survey lizard biodiversity and collect preliminary
+                genetic data
+        \end{itemize}
+
+        \column{0.5\linewidth}
+        % \begin{minipage}[c][\textheight][c]{\columnwidth}
+        {\centering
+            \includegraphics[width=\columnwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
+
+            \vspace{2mm}
+            \includegraphics[width=\columnwidth]{/home/jamie/Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
+        }
+        % \end{minipage}
     \end{columns}
@@
 
-\begin{frame}[t]
-    \frametitle{Future directions: West-Central Africa}
-    \begin{center}
-        \includegraphics<1->[height={3cm}]{/home/jamie/Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
-        \hspace{0.3mm}
-        \includegraphics<1->[height={3cm}]{/home/jamie/Dropbox/field-photos/africa/Figure4_herps.jpeg}
-    \end{center}
+\begin{frame}
+    \frametitle{Empirical applications: Southeast Asia}
+    \begin{columns}
+        \column{0.5\linewidth}
+        \begin{itemize}[<+->]
+            \item Did glacial cycles promote co-diversification across SE Asia?
+            \item Compare patterns of diversification between mainland and
+                continental and oceanic island systems
+            \item Islands of the Sunda Shelf are an exciting system for
+                comparative evolutionary biology
+            \item NSF pre-proposal pending
+        \end{itemize}
 
-    \only<1>{
-    \begin{center}
-        \includegraphics[height={3.5cm}]{/home/jamie/Dropbox/field-photos/people/adam.png}
-        \hspace{0.3mm}
-        \includegraphics[height={3.5cm}]{/home/jamie/Dropbox/field-photos/people/fujita.jpg}
-    \end{center}
+        \column{0.5\linewidth}
+        {\centering
+            {\setlength{\fboxsep}{0pt}
+            \fbox{
+                \hspace{-1.2mm}\includegraphics[width=\columnwidth]{../images/maps/se-asia-120-crop.png}
+            }}
         }
-
-    \uncover<2->{
-    \begin{center}
-        \includegraphics[height=4.5cm]{../images/number-of-divs-africa.pdf}
-    \end{center}
-    }
-\end{frame}
-
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-    \vspace{-2mm}
-    \begin{columns}
-        % \column{\dimexpr\paperwidth-10pt}
-        \column{\dimexpr\paperwidth}
-        \includegraphics[width=\textwidth]{/home/jamie/Dropbox/field-photos/malaysia/malaysia-burungs.jpg}
     \end{columns}
 \end{frame}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-\end{frame}
-}
-
-{
-\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
-\begin{frame}
-    \frametitle{``Micro''-islands of the Sunda Shelf}
-\end{frame}
-}

### 2015-03-08 slides/auburn/intro-mascot-tree.tex
More work on slides.

@@
     \vspace{-3mm}
-    \uncover<2->{
+    \uncover<3->{
     \begin{adjustwidth}{-3.1em}{-3.1em}
         \begin{itemize}
-            \item<2-> \textbf{Limiting assumption:} Events are independent across the tree
+            \item<3-> \textbf{Limiting assumption:} Events are independent across the tree
             % \item<3-> We know this assumption is frequently violated
@@
 
-    \begin{onlyenv}<1-2>
+    \begin{onlyenv}<2-3>
     \begin{center}
@@
     % \vspace{-1cm}
-    \begin{onlyenv}<3->
+    \begin{onlyenv}<4->
     \begin{columns}
@@
             % \item<2-> Current phylogenetic methods assume events are independent 
-            \item<3-> We know this assumption is frequently violated
-            \item<5-> Why account for this non-independence?
+            \item<4-> We know this assumption is frequently violated
+            \item<6-> Why account for this non-independence?
             \begin{enumerate}
-                \item<6-> Improve phylogenetic inference
-                \item<7-> \textbf{Provide a general framework for inferring the
+                \item<7-> Improve phylogenetic inference
+                \item<8-> \textbf{Provide a general framework for inferring the
                         effect of community-scale processes on diversification}
@@
         \centerline{
-        \includegraphics<3>[width=1.1\columnwidth]{../mascot-tree/mascot-tree-shared.pdf}
-        \includegraphics<4->[width=1.1\columnwidth]{../mascot-tree/mascot-tree-shared-events.pdf}
+        \includegraphics<4>[width=1.1\columnwidth]{../mascot-tree/mascot-tree-shared.pdf}
+        \includegraphics<5->[width=1.1\columnwidth]{../mascot-tree/mascot-tree-shared-events.pdf}
         }

### 2015-03-09 slides/bin/plot-validation-results.py
Working on slides.

@@ -137,6 +137,36 @@ def main_cli():
 
     pg.savefig('../images/validation-model-choice-old-dpp.pdf')
 
+    for p in sub_plot_list:
+        p.set_extra_y_label('')
+        p.set_title_text('')
+
+    pg = plotting.PlotGrid(subplots = sub_plot_list[-1:],
+            num_columns = 1,
+            share_x = True,
+            share_y = True,
+            title = 'Posterior probability of one divergence',
+            title_size = 14.0,
+            title_top = False,
+            y_title = 'True probability of one divergence',
+            y_title_position = 0.001,
+            y_title_size = 14.0,
+            # column_labels = [r'dpp-msbayes'],
+            column_label_size = 22.0,
+            column_label_offset = 0.04,
+            width = 4.8,
+            height = 3.6,
+            auto_height = False)
+    pg.label_schema = None
+    pg.auto_adjust_margins = False
+    pg.margin_bottom = 0.06
+    pg.margin_left = 0.06
+    pg.margin_top = 0.99
+    pg.margin_right = 0.94
+    pg.reset_figure()
+
+    pg.savefig('../images/validation-model-choice-dpp-no-label.pdf')
+
     sub_plot_list = [
             sub_plots['old']['old'],
             sub_plots['dpp']['old'],
@@ -252,6 +282,41 @@ def main_cli():
     pg.reset_figure()
 
     pg.savefig('../images/validation-model-choice-dpp-violations.pdf')
+    
+    for p in sub_plot_list:
+        p.set_extra_y_label('')
+        p.set_title_text('')
+
+    pg = plotting.PlotGrid(subplots = dpp_violations,
+            num_columns = 2,
+            share_x = True,
+            share_y = True,
+            title = 'Posterior probability of one divergence',
+            title_size = 16.0,
+            title_top = False,
+            y_title = 'True probability of one divergence',
+            y_title_position = 0.015,
+            y_title_size = 16.0,
+            height = 4,
+            width = 10,
+            auto_height = False,
+            # column_labels = ['',''],
+            column_label_size = 22.0,
+            column_label_offset = 0.05,
+            # super_title = r'\textbf{Data model}',
+            # super_y_title = r'\textbf{Analysis model}',
+            super_title_size = 20.0,
+            super_y_title_size = 20.0,
+            super_y_title_right = True)
+    pg.label_schema = None
[…]
